### 2022-05-12 body.tex
Initial commit

\begin{abstract}
\input{abstract}
\end{abstract}

\maketitle

\section{Introduction}
\label{sec:intro}

\section{Conclusion}
\label{sec:conclusion}

\setlength{\bibsep}{.8ex}
\bibliographystyle{tex-macros/ACM-Reference-Format.bst}
\bibliography{refs.bib}

### 2022-05-12 make/makefile.tex
Initial commit

# This is a makefile template to be copied & modified. Copy it to
# "Makefile" in your directory and do "chmod u+w Makefile".
# -- Andrew Myers

DOC = paper
#
# change this line to contain the name of the document without extension

TARGETS = $(DOC).pdf
#
# TARGETS could be $(DOC).dvi instead, if there are no figures

TEXS = $(DOC).tex *.tex *.sty
#
# Add any included TeX files to TEXS

FIGS =
#
# List of postscript figures to be included

PAPERSROOT=..
MAKEDIR=$(PAPERSROOT)/make

include $(MAKEDIR)/commondefs
#
# Change MAKEDIR to . if you have commondefs installed in the current
# directory.

default: $(TARGETS)

BIBFILE = $(DOC).bib
# BIBFILE = $(PAPERSROOT)/bibtex/master.bib
#
# Ignore this if you're not using bibtex. Change the $(DOC) part if your bib
# file has a different name from your document

$(DOC).dvi: $(TEXS) $(FIGS) $(DOC).stamp
$(DOC).pdf: $(TEXS) $(FIGS) $(DOC).stamp
#
# To use bibtex, add these to dependency list:
# $(DOC).bbl $(BIBFILE)
# 
# To use makeindex, add this to dependency list:
# $(DOC).ind
#
# To use glossaries, add this to dependency list:
# $(DOC).glo

$(DOC).bbl: $(BIBFILE) $(DOC).stamp
$(DOC).glo $(DOC).ind: $(DOC).stamp
$(DOC).ps: $(DOC).dvi $(FIGS)
#
# These are standard dependencies. Shouldn't need to modify these.

VPATH = $(PAPERSROOT)/tex-macros

include $(COMMONRULES)

### 2022-05-12 tex-macros/aplcomments-test.tex
Initial commit

\documentclass[12pt]{article}
\usepackage{times,graphicx}
\IfFileExists{.comments}{
\typeout{** Showing comments **}
\usepackage{hyperref}
\usepackage{aplcomments}
}{
\usepackage[disabled]{aplcomments}
}
%\usepackage[inline]{aplcomments}
\textwidth 7in
\advance\oddsidemargin -1.25in
\usepackage{aplcomments}
\newcommenter{Andrew}{0.0,0.5,1.0}
\begin{document}

This song was written for the Cornell Computer Science Department
holiday party in December 2009. It was performed by Andrew Myers,
with Dexter Kozen accompanying\Andrew{awesomely}.

\Andrew{Lyrics by Andrew Myers.}

This is another paragraph.
\Andrew{and another comment}

And here is a third paragraph, which should be indented even if
comments are being used.

\Andrew{A comment before a section heading might get eaten}

\section{Lyrics follow this point}

\Andrew{A comment after a section heading should be okay}

{
\obeylines
\setlength\parskip{0.7em}
\Andrew{The beginning}
A new student came by just the other day
to start some research right away
But there were papers to review and grants to raise
and he took courses while I was away
And he gave talks 'fore I knew it, and good ones too
He'd say, ``I'm gonna be like you, yeah"
"You know I'm gonna be like you"
\hrule
Refrain:
\fbox{
\begin{minipage}{\linewidth}
The committee meetings and the blind reviews,
Writing homeworks and exams, and grading them too.
``When can we meet, Professor?"
``Don't know when, but we'll do some research then"
``You know we'll do some nice work then"
\end{minipage}
}
My student started third year just the other day
He said, ``My first paper's going well now that I've got an RA"
``Can you help me write the intro today?"
I said, ``Not right now, I have exams to grade"
And he walked away but his smile never dimmed
He said, ``I'm gonna be like him, yeah"
``You know I'm gonna be like him"
\hrule
(Refrain)
\hrule
Well, he finished his thesis just the other day
Such a brilliant researcher I just had to say,
``Hey, I'm proud of you, want to talk for a while?"
He shook his head and then said with a smile
``What I really need now is a very strong letter"
``See you later, can you write it soon, please?
See you later, can you write it soon?"
\hrule
And as I sent off the email it occurred to me
He'd ended up just like me.
Yeah, my student was just like me.
}

\begin{figure}

\begin{center}
\Andrew{comment in a figure}
\fbox{Figure}
\end{center}
\caption{This is a caption\Andrew{and a comment in a caption}}
\label{fig}
\end{figure}

\ref{fig}
final paragraph

\end{document}

### 2022-05-12 title.tex
Initial commit

\def\thetitle{%
Title
}

\def\thesubtitle{%
No Subtitle
}

\ifreport
\title{\textls[-15]{\thetitle \mbox{(Technical Report)}}}
\else
\title{\textls[-15]{\thetitle}}
\fi

%\subtitle{\thesubtitle}

### 2022-05-12 body.tex
Adding gitignore.
Save.

@@
 
+\section{Metatheory}
+\label{sec:metatheory}
+\subsection{Syntax}
+We start with the formulation of the syntax, which is based on the formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT) given by \cite{coquand2018canonicity}.
+Thus we will omit most of the discussion about MLTT here but focus on the new facility we introduced.
+Though we formulate in Latex, our thought process is mainly following \cite{kaposi2017type} where we use a type theoretical framework (like Agda) as our meta-logic instead of a the conventional constructive set theory as \cite{coquand2018canonicity}.
+
+More concretely, we consider the syntax of our type theory as a \textit{Quotient Inductive Inductive Type} (QIIT), and most of the reasoning will be formulated using the \textit{algebra} of QIIT (mapping out function from this quotiented data-type).
+Each judgement will actually be represented by a QIIT type, which are mutual-recursively defined because those judgements are inter-dependently defined.
+Quotient is used to represent judgemental(definitional) equality, thanks to which we can have a concise representation.
+
+Compared to the conventional formulation of type theory, we highlight some details of difference in this style of formluation of dependent: 1.
+this modern formulation of type theory does't rely on any notion of operational semantic but only equality
+%We will recover canonicity afterwards, which implies the computational ability of our theory. 
+;
+2.
+instead of using meta-level substitution, we use explicit substitution and debruijn indices.
+This is also called substitution calculus in the literature and is favoured due to its clear categorical semantic.
+
+\subsection{Standard Model, for Consistency}
+Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
+We prove the consistency by extending the standard model from \citep{kaposi2017type} to interpret our theory.
+
+Notice that we interpret the bottom type using empty set, and thus we won't have syntactically derived
+
+\subsection{Canonicity}
+
 \section{Conclusion}

### 2022-05-13 body.tex
Save.

--- 2022-05-12 body.tex
@@
 
+\section{Metatheory}
+\label{sec:metatheory}
+
+We start with the formulation of the syntax, which is based on the formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT) given by \citep{coquand2018canonicity}.
+Thus we will omit most of the discussion about MLTT here but focus on the new facility we introduced.
+
+Though formulated in Latex, we are following \citep{kaposi2017type} to use a type theoretical framework (like Agda) as our meta-logic, 
+and using \textit{Quotient Inductive Inductive Type} (QIIT) to represent judgemental equality.
+Each judgement will actually be represented by a QIIT type, which are mutual-recursively defined together.
+Quotient is used to represent judgemental(definitional) equality, thanks to which we can have a concise representation.
+This way, for all four kinds of judgement we are using, there are always equality among these data.
+Doing so, we can omit some obvious equality judgement.
+
+Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).
+
+We also highlight some special feature in this style of formulation of dependent type theory: 1. this modern formulation of type theory does't have operational semantic but only equality 
+%We will recover canonicity afterwards, which implies the computational ability of our theory. 
+;
+2.
+instead of using meta-level substitution, we use explicit substitution and debruijn indices in our intrinsic formulation.
+This is also called substitution calculus in the literature and is favoured due to its clear categorical semantic;
+3. we are dealing with quotient data, so we are actually manipulating equivalence classes of terms.
+
+We will write out the syntax of the whole type theory for completeness.
+
+\newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}
+\newcommand{\goodCtx}[2]{{ {#1} \ ok }}
+\newcommand{\goodType}[3]{{ {#1} \vdash {#2} }}
+\newcommand{\goodTerm}[3]{{ {#1} \vdash {#2} : {#3} }}
+\newcommand{\goodSub}[3]{{ {#1} \vdash {#2} : {#3} }}
+\newcommand{\goodSig}[3]{{ {#1} \vdash {#2} \ \  Sig^{#3} }}
+\newcommand{\goodWSig}[3]{{ {#1} \vdash {#2} \ \ WSig^{#3} }}
+\newcommand{\goodSeal}[3]{{ {#1} \vdash {#2} \  |\  {#3} }}
+\newcommand{\goodInh}[3]{{ {#1} \vdash {#2} \twoheadrightarrow {#3}}}
+\newcommand{\nat}{\mathbf{N}}
+
+\newcommand{\cU}{{\mathcal{U}}}
+\newcommand{\cB}{{\mathbb{B}}}
+
+\subsection{Syntax}
+
+\input{syntax}
+
+\subsection{Standard Model for Consistency}
+Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
+We prove the consistency by extending the standard model from \citep{kaposi2017type} to interpret our theory.
+
+Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we interpret this derivation in our standard model and we can conclude $\denotes{t} \in \emptyset$, which is a contradiction.
+
+\subsection{Proof Relevant Logical Relation for Canonicity}
+We follow \citep{sterling2019algebraic,coquand2018canonicity} to construct the canonicity model.
+
 \section{Conclusion}

### 2022-05-13 syntax.tex
Save.

Firstly, there are four standard kinds of judgement in MLTT, well-typed context $\goodCtx{\Gamma}{l}$, well-typed type $\goodType{\Gamma}{T}{l}$, well-typed term $\goodTerm{\Gamma}{t}{T}$ and well-typed substitution $\goodSub{\Gamma}{\gamma}{\Delta}$. We are following \citep{kaposi2019gluing} but we will omit the universe level attached to the judgement, to make it more readable. After these, we will have four new kinds of judgement.

Because we consider each judgement as a (dependent) QIIT type, thus each judgement has a \textit{type} as well (otherwise how can they be encoded in Agda?). Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements. This \textit{well-formed-ness} might still be confusing thus we spell out their Agda counterpart to make things more understandable. Of course, a sequent is well-formed doesn't mean it is derivable. Every sequent, judgement and data we discuss here will be well-formed because we are working in a type-theoretical framework. Due to our intrinsic setting, we can omit a lot of presumption, for example we don't spell out that "Type-Universe" require a well-typed context, even though it did implicitly.

\begin{figure}[H]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}$$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma, \tau}{n}}
      {\goodSeal{\Gamma}{\sigma}{\tau} \text{ is well-formed}}
      $$
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Sig  : Con → ℕ → Set
      data WSig : Con i → ℕ → Set 
      data Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
    \end{minted}
  \end{minipage}
\end{figure}

Now we spell out some of the typing rules, not meant to be comprehensive 
$$ 
\Rule[name=Empty Context]{}{\goodCtx{\cdot}{0}} 
\quad
\Rule[name=Context Extension]
{\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{A}{j}}
{\goodCtx{\Gamma, A}{i \cup j}}  
$$

$$
\Rule[name=Type Universe]
{}
{\goodType{\Gamma}{\cU_j}{j + 1}}
\quad 
\Rule[name=Boolean]
{}
{\goodType{\Gamma}{\cB}{0}}
\quad 
\Rule[name=Bottom]
{}
{\goodType{\Gamma}{\bot}{0}}
\quad 
\Rule[name=Function]
{\goodType{\Gamma}{A}{j} 
  \quad \goodType{\Gamma, A}{B}{k}}
{\goodType{\Gamma}{\Pi A B}{j \cup k}}
$$

$$
\Rule[name=Type Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}}
\quad 
\Rule[name=Function Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma \uparrow] }{j \cup k}}
$$
$$
\Rule[name=Base Type Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{\cU_j[\gamma] \equiv \cU_j }{j + 1} \quad
 \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
 \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
}
$$

### 2022-05-13 syntax.tex
Save.

@@
 Every sequent, judgement and data we discuss here will be well-formed because we are working in a type-theoretical framework.
-Due to our intrinsic setting, we can omit a lot of presumption, for example we don't spell out that "Type-Universe" require a well-typed context, even though it did implicitly.
+Due to our intrinsic setting, we can omit a lot of presumption, for example we don't spell out that the rule "Type-Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
+
+Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{\tau}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{\cL\tau[\pi_1]}$.
+Even though we will introduce rules for sealing judgement, we will later show that all these rules are derivable by term judgements.
 
@@
       {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma, \tau}{n}}
-      {\goodSeal{\Gamma}{\sigma}{\tau} \text{ is well-formed}}
+      {\goodSeal{\Gamma}{\_}{\sigma}{\tau} \text{ is well-formed}}
       $$
@@
       data WSig : Con i → ℕ → Set 
-      data Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
       data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
+      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
     \end{minted}
@@
 
-Now we spell out some of the typing rules, not meant to be comprehensive 
+Now we show case some of the typing rules, and demonstrate their typed encoding in our (fake) Agda, to help the reader  grasp the rough idea of our type system and understand the Agda-formulation of the typing rules easier.
 $$ 
@@
 \quad 
-\Rule[name=Function Subst]
+\Rule[name=Func Subst]
 {\goodSub{\Gamma}{\gamma}{\Delta}}
-{\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma \uparrow] }{j \cup k}}
+{\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}}
 $$
@@
 $$
+$$
+\Rule[name=Univ-1]
+{\goodType{\Gamma}{T}{j}}
+{\goodTerm{\Gamma}{c \ T}{\cU_j}
+}\quad
+\Rule[name=Univ-2]
+{\goodTerm{\Gamma}{T}{\cU_j}}
+{\goodType{\Gamma}{El \ T}{j}}
+\quad
+\Rule[name=True]
+{}
+{\goodTerm{\Gamma}{tt}{\cB}}
+\quad
+\Rule[name=False]
+{}
+{\goodTerm{\Gamma}{ff}{\cB}}
+$$
+$$
+\Rule[name=Term Subst]
+{\goodTerm{\Delta}{t}{T}
+  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
+\quad 
+\Rule[name=Func intro]
+{\goodTerm{\Gamma, A}{t}{B}}
+{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
+\quad 
+\Rule[name=Lam Subst]
+{}
+{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}}
+$$
+$$
+\Rule[name=Base Type/Term Subst]
+{\goodSub{\Gamma}{\gamma}{\Delta}}
+{\goodType{\Gamma}{(El \ T)[\gamma] \equiv El \ (T[\gamma]) }{j} \quad
+ \goodTerm{\Gamma}{tt[\gamma] \equiv tt}{\cB} \quad 
+ \goodTerm{\Gamma}{ff[\gamma] \equiv ff}{\cB} 
+}
+$$
+
+$$
+\Rule[name=Empty Subst]
+{}{\goodSub{\Gamma}{\epsilon}{\cdot}}
+\quad
+\Rule[name=Id Subst]
+{}{\goodSub{\Gamma}{id}{\Gamma}}
+\quad
+\Rule[name=Subst Extension]
+{\goodSub{\Gamma}{\sigma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
+{\goodSub{\Gamma}{(\sigma, t)}{(\Delta, A)}}
+$$
+
+$$
+\Rule[name=Proj Subst-1]
+{\goodSub{\Gamma}{\sigma}{(\Delta, A)}}
+{\goodSub{\Gamma}{\pi_1 \sigma}{\Delta}}
+\quad
+\Rule[name=Proj Subst-2]
+{\goodSub{\Gamma}{\sigma}{(\Delta, A)}}
+{\goodSub{\Gamma}{\pi_2 \sigma}{A[\pi_1 \sigma]}}
+\quad
+\Rule[name=Proj-Ext]
+{}
+{\goodSub{\Gamma}{(\pi_1 \sigma, \pi_2 \sigma) \equiv \sigma}{\Delta}}
+$$
+
+We have shown case some exemplar rules for all four standard type judgements, now we focus on the new equipments.

### 2022-05-14 metatheory.tex
Save.

We start with the formulation of the syntax, which is based on the formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT) given by \citep{coquand2018canonicity}. Thus we will omit most of the discussion about MLTT here but focus on the new facility we introduced.

Though formulated in Latex, we are following \citep{kaposi2017type} to use a type theoretical framework (like Agda) as our meta-logic, 
and using \textit{Quotient Inductive Inductive Type} (QIIT) to represent judgemental equality. Each judgement will actually be represented by a QIIT type, which are mutual-recursively defined together. Quotient is used to represent judgemental(definitional) equality, thanks to which we can have a concise representation. This way, for all four kinds of judgement we are using, there are always equality among these data. Doing so, we can omit some obvious equality judgement.

Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).

We also highlight some special feature in this style of formulation of dependent type theory: 1. this modern formulation of type theory does't have operational semantic but only equality 
%We will recover canonicity afterwards, which implies the computational ability of our theory. 
;
2. instead of using meta-level substitution, we use explicit substitution and debruijn indices in our intrinsic formulation. This is also called substitution calculus in the literature and is favoured due to its clear categorical semantic; 
3. we are dealing with quotient data, so we are actually manipulating equivalence classes of terms.

We will write out the syntax of the whole type theory for completeness.

\newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}
\newcommand{\goodCtx}[2]{{ {#1} \ ok }}
\newcommand{\goodType}[3]{{ {#1} \vdash {#2} }}
\newcommand{\goodTerm}[3]{{ {#1} \vdash {#2} : {#3} }}
\newcommand{\goodSub}[3]{{ {#1} \vdash {#2} : {#3} }}
\newcommand{\goodSig}[3]{{ {#1} \vdash {#2} \ \  Sig^{#3} }}
\newcommand{\goodWSig}[3]{{ {#1} \vdash {#2} \ \ WSig^{#3} }}
\newcommand{\goodSeal}[4]{{ {#1} \vdash {#2} : {#3} \  |\  {#4} }}
\newcommand{\goodInh}[3]{{ {#1} \vdash {#2} \twoheadrightarrow {#3}}}
\newcommand{\nat}{\mathbf{N}}

\newcommand{\cU}{{\mathcal{U}}}
\newcommand{\cB}{{\mathbb{B}}}
\newcommand{\cL}{{\mathcal{L}}}

\subsection{Syntax}

\input{syntax}

\subsection{Standard Model for Consistency}
Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory. We prove the consistency by extending the standard model from \citep{kaposi2017type} to interpret our theory.

Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we interpret this derivation in our standard model and we can conclude $\denotes{t} \in \emptyset$, which is a contradiction.

\subsection{Proof Relevant Logical Relation for Canonicity}
We follow \citep{sterling2019algebraic,coquand2018canonicity} to construct the canonicity model.

### 2022-05-14 metatheory.tex
Save.

@@
 We start with the formulation of the syntax, which is based on the formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT) given by \citep{coquand2018canonicity}.
-Thus we will omit most of the discussion about MLTT here but focus on the new facility we introduced.
+Thus we will omit a great deal of details about MLTT here but focus on the new facility we introduced.
 
-Though formulated in Latex, we are following \citep{kaposi2017type} to use a type theoretical framework (like Agda) as our meta-logic, 
-and using \textit{Quotient Inductive Inductive Type} (QIIT) to represent judgemental equality.
+Though phrased in Latex, we are following \citep{kaposi2017type} to use as our meta-logic a type theoretical framework (like Agda) , 
+together with \textit{Quotient Inductive Inductive Type} (QIIT).
 Each judgement will actually be represented by a QIIT type, which are mutual-recursively defined together.
@@
 Doing so, we can omit some obvious equality judgement.
-
 Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).
@@
 2.
-instead of using meta-level substitution, we use explicit substitution and debruijn indices in our intrinsic formulation.
+instead of using meta-level substitution, we use explicit substitution and de Bruijn indices in our intrinsic formulation.
 This is also called substitution calculus in the literature and is favoured due to its clear categorical semantic;
@@
 \newcommand{\cL}{{\mathcal{L}}}
+\newcommand{\bW}{{\mathbb{W}}}
 

### 2022-05-14 syntax.tex
Save.

@@
 Firstly, there are four standard kinds of judgement in MLTT, well-typed context $\goodCtx{\Gamma}{l}$, well-typed type $\goodType{\Gamma}{T}{l}$, well-typed term $\goodTerm{\Gamma}{t}{T}$ and well-typed substitution $\goodSub{\Gamma}{\gamma}{\Delta}$.
-We are following \citep{kaposi2019gluing} but we will omit the universe level attached to the judgement, to make it more readable.
+We are following \citep{kaposi2019gluing} but we will omit the universe level attached to the judgement here, to make it more readable.
 After these, we will have four new kinds of judgement.
 
-Because we consider each judgement as a (dependent) QIIT type, thus each judgement has a \textit{type} as well (otherwise how can they be encoded in Agda?).
-Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements.
-This \textit{well-formed-ness} might still be confusing thus we spell out their Agda counterpart to make things more understandable.
-Of course, a sequent is well-formed doesn't mean it is derivable.
-Every sequent, judgement and data we discuss here will be well-formed because we are working in a type-theoretical framework.
+Because we consider each judgement as a (dependent) QIIT type, each judgement has a \textit{type} and each sequent is well-typed as well (otherwise how can they be encoded in Agda?).
+Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements and their Agda counterpart to make things more understandable.
+Of course, a well-formed sequent  is not necessarily derivable.
 Due to our intrinsic setting, we can omit a lot of presumption, for example we don't spell out that the rule "Type-Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
@@
       data Sig  : Con → ℕ → Set
-      data WSig : Con i → ℕ → Set 
+      data WSig : Con → ℕ → Set 
       data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
@@
   \end{minipage}
+  \caption{Well-formed-ness of Four New Judgements}
 \end{figure}
@@
 Now we show case some of the typing rules, and demonstrate their typed encoding in our (fake) Agda, to help the reader  grasp the rough idea of our type system and understand the Agda-formulation of the typing rules easier.
+
+  \label{fig:rules:well-typed-ctx}
+\judgebox{\goodCtx{\Gamma}{i}}
 $$ 
@@
 
+\judgebox{ \goodType{\Gamma}{T}{i} }
 $$
@@
 $$
+\judgebox{ \goodTerm{\Gamma}{t}{T} }
 $$
@@
 $$
-
+\judgebox{\goodSub{\Gamma}{\sigma}{\Delta}}
 $$
@@
 
-We have shown case some exemplar rules for all four standard type judgements, now we focus on the new equipments.
+We have shown case some exemplar rules for all four standard type judgments, now we focus on the newly introduced facility.
+
+We have two kinds of signatures, one for extensible inductive type, the other for linkages.
+
+Recall in W-type, a pair of type $x : A \vdash B(x)$ decides a W-type.
+For simplicity,  we use a list of pairs of such $A, B$ as the signature for a given inductive type.
+Doing so, all the constructors will have a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
+We only formulate non-dependent eliminator, \textit{recursor}.
+Each recursor is decided by a given linkage.
+
+$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields.
+We omit the naming/label information for fields in the meta-theory.
+We can use $\cL$ to retrieve the corresponding types of linkage of a given signature.
+Our signature corresponds to Coq's Module Type and our linkage corresponds to Coq's Module.
+Of course we need to put an eye on the substitution lemma.
+
+Signatures and Linkages are both constructed from the empty stuff, and respect substitution lemma for sure.
+Intuitively, both of their rules should be motivated by that of dependent pairs, so an obvious question is why is that sealing judgement there?
+
+The sealing judgement is like a syntactic sugar helpful for future inheritance.
+Consider the scenario we have constructed a linkage meant to be recursor $  ..
+(T : \bW(\sigma)) \vdash L : \cL(\tau)$.
+Now we extend $\sigma$ with more constructors into $\sigma'$, but we don't have any rules to transfer $L$ to the context of $\bW(\sigma')$.
+Thus the correct way is to construct $L$ in an \textit{sealed} context, $  ..
+(T : \cU) \vdash L : \cL(\tau)$, and single out the \textit{coverage checking} (because we don't have to inherits coverage checking -- we should redo it everytime we program)
+
+$$
+\Rule[name=Linkage Type]
+{\goodSig{\Gamma}{\sigma}{n}}
+{\goodType{\Gamma}{\cL \sigma}{i}}
+\quad
+\Rule[name=Sig/Linkage Subst]
+{\goodSig{\Delta}{\sigma}{n}
+  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+{\goodSig{\Gamma}{\sigma[\gamma]}{n}
+  \quad \goodType{\Gamma}{\cL (\sigma[\gamma]) \equiv (\cL \sigma)[\gamma]}{i}
+}
+$$

### 2022-05-14 metatheory.tex
Save.

@@
 \newcommand{\goodSeal}[4]{{ {#1} \vdash {#2} : {#3} \  |\  {#4} }}
-\newcommand{\goodInh}[3]{{ {#1} \vdash {#2} \twoheadrightarrow {#3}}}
+\newcommand{\goodInh}[4]{{ {#1} \vdash {#2} : {#3} \twoheadrightarrow {#4}}}
 \newcommand{\nat}{\mathbf{N}}
@@
 \subsection{Standard Model for Consistency}
-Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{kaposi2017type} to interpret our theory.
 
-Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we interpret this derivation in our standard model and we can conclude $\denotes{t} \in \emptyset$, which is a contradiction.
+\input{standardmodel.tex}
 

### 2022-05-14 standardmodel.tex
Save.

Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory. We prove the consistency by extending the standard model from \citep{kaposi2017type}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into Agda components that is also respecting judgemental equalities.\footnote{Of course, it is also possible to interpret to constructive set theory.} 
Here we will only show some interpretation, the complete version please refer to appendix.

\begin{align*}
  ss
\end{align*}

Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we interpret this derivation in our standard model and we can conclude $\denotes{t} : Data.Empty$, 
% which is basically $\denotes{t} \in \emptyset$ set theoretically, 
and that is a contradiction.

### 2022-05-14 syntax.tex
Save.

@@
       $$
+      \Rule[name=Tm]
+      {\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{T}{j}}
+      {\goodTerm{\Gamma}{\_}{T}}
+      $$
+      $$
+      \Rule[name=Tms]
+      {\goodCtx{\Gamma}{i} \quad \goodCtx{\Delta}{i}}
+      {\goodSub{\Gamma}{\_}{\Delta} \text{ is well-formed}}
+      $$
+      $$
       \Rule[name=Sig]
       {\goodCtx{\Gamma}{i} \quad n \in \nat}
-      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}$$
+      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
+      $$
       $$
@@
     \begin{minted}[]{agda}
+      data Con   : Set 
+      data Ty    : Con → Set   
+      data Tms   : Con → Con → Set 
+      data Tm    : (Γ : Con) → Ty Γ  → Set 
       data Sig  : Con → ℕ → Set
@@
 
-Recall in W-type, a pair of type $x : A \vdash B(x)$ decides a W-type.
+\judgebox{\goodWSig{\Gamma}{\sigma}{n}}
+$$
+\Rule[name=Inductive Type]
+{\goodWSig{\Gamma}{\sigma}{n}}
+{\goodType{\Gamma}{W\sigma}{i}}
+\quad
+\Rule[name=Inductive Sig]
+{\goodWSig{\Delta}{\sigma}{n}
+  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+{\goodWSig{\Gamma}{\sigma[\gamma]}{n}
+  \quad \goodType{\Gamma}{W (\sigma[\gamma]) \equiv (W \sigma)[\gamma]}{i}}
+$$
+
+Recall in W-type \citep{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type.
 For simplicity,  we use a list of pairs of such $A, B$ as the signature for a given inductive type.
 Doing so, all the constructors will have a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
-We only formulate non-dependent eliminator, \textit{recursor}.
-Each recursor is decided by a given linkage.
+Of course we need to put an eye on the substitution law for each piece of the syntax.
+In our meta-theory, we only formulate non-dependent eliminator, i.e.
+the \textit{recursor}.
+Each recursor is constructed by a linkage contained with functions dealing with each case.
+Thus code-reuse for recursor is delegated by the inheritance for linkage.
 
-$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields.
-We omit the naming/label information for fields in the meta-theory.
-We can use $\cL$ to retrieve the corresponding types of linkage of a given signature.
-Our signature corresponds to Coq's Module Type and our linkage corresponds to Coq's Module.
-Of course we need to put an eye on the substitution lemma.
-
-Signatures and Linkages are both constructed from the empty stuff, and respect substitution lemma for sure.
-Intuitively, both of their rules should be motivated by that of dependent pairs, so an obvious question is why is that sealing judgement there?
-
-The sealing judgement is like a syntactic sugar helpful for future inheritance.
-Consider the scenario we have constructed a linkage meant to be recursor $  ..
-(T : \bW(\sigma)) \vdash L : \cL(\tau)$.
-Now we extend $\sigma$ with more constructors into $\sigma'$, but we don't have any rules to transfer $L$ to the context of $\bW(\sigma')$.
-Thus the correct way is to construct $L$ in an \textit{sealed} context, $  ..
-(T : \cU) \vdash L : \cL(\tau)$, and single out the \textit{coverage checking} (because we don't have to inherits coverage checking -- we should redo it everytime we program)
-
+\judgebox{\goodSig{\Gamma}{\sigma}{n} }
 $$
@@
 {\goodSig{\Gamma}{\sigma[\gamma]}{n}
-  \quad \goodType{\Gamma}{\cL (\sigma[\gamma]) \equiv (\cL \sigma)[\gamma]}{i}
+  \quad \goodType{\Gamma}{\cL (\sigma[\gamma]) \equiv (\cL \sigma)[\gamma]}{i}}
+$$
+
+$$
+\Rule[name=Empty Sig]
+{}
+{\goodSig{\Gamma}{\nu\cdot}{0}}
+\quad
+\Rule[name=Sig Add]
+{\goodSig{\Gamma}{\sigma}{n} 
+ \quad \goodSeal{\Gamma}{f}{\sigma}{\sigma'}
+ \quad \goodType{\Gamma, \sigma'}{T}{i}}
+{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
+$$
+$$
+\Rule[name=Empty Lnkg]
+{}
+{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
+\quad
+\Rule[name=Lnkg Add]
+{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+\quad  \goodSeal{\Gamma}{f}{\sigma}{\sigma'} 
+ \quad \goodTerm{\Gamma, \cL \sigma'}{t}{T}
 }
+{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
 $$
+
+$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields.
+We omit the naming/label information for fields in the meta-theory.
+We can use $\cL$ to retrieve the corresponding types of linkage of a given signature.
+Our signature corresponds to Coq's Module Type and our linkage corresponds to Coq's Module.
+
+The signature for linkage closely follows the signature for inductive type.
+They are both constructed from the empty stuff, and respect substitution lemma for sure.
+We have projection for each componenet due to harmony \citep{pfenning2009lecture} (local soundness and completeness), but we omit them and the substitution laws for either concatenation and projection.
+For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.
+
+Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by that of dependent pairs, so an obvious question is why is that sealing judgement there?
+
+\judgebox{[\goodSeal{\Gamma}{s}{\sigma}{\sigma'}] = [\goodTerm{\Gamma, \cL\sigma}{s}{\cL \sigma'[\pi_1]}] }
+$$
+\Rule[name=Seal Subst]
+{\goodSeal{\Delta}{s}{\sigma}{\sigma'}
+  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+{\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{\sigma'[\gamma]}}
+=
+\frac
+{\goodTerm{\Delta, \cL\sigma}{s}{\cL\sigma'[\pi_1]}
+  \quad  \goodSub{\Gamma}{\gamma}{\Delta}  }
+{\goodTerm{\Gamma,\cL\sigma[\gamma]}{s[\gamma^\uparrow]}{\cL\sigma'[\pi_1][\gamma^\uparrow]\equiv\cL\sigma'[\gamma][\pi_1]}}
+$$
+$$
+\Rule[name=Seal-Id]
+{}
+{\goodSeal{\Gamma}{id_s}{\sigma}{\sigma}}
+= \goodTerm{\Gamma, \cL\sigma}{\pi_2}{\cL\sigma[\pi_1]}
+$$
+\begin{figure}[H]
+
+\centering
+\captionsetup{justification=centering}
+
+\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a short hand of the other)}
+
+\begin{minted}[]{agda}
+Seal : (Γ : Con) → (σ : Sig Γ n) → (τ : Sig Γ n) → Set 
+Seal Γ σ τ = Tm (Γ, 𝓛σ) 𝓛τ[p₁]
+
+-[-]S : (f : Seal Γ σ₁ σ₂) → (τ : Tms Θ Γ) → Seal Θ σ₁[τ] σ₂[τ]
+f[τ]S = t[τ ↑ 𝓛σ₁]
+
+idₛ : Seal Γ σ σ
+idₛ = π₁
+\end{minted}
+
+\end{figure}
+
+As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{\sigma'}$ is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{\cL\tau[\pi_1]}$.
+The substitution law should understand as a lemma, and easily derivable by using $\gamma^\uparrow$.
+
+The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance.
+Consider the scenario we have constructed a linkage meant to be a recursor $  ..
+(T : \bW(\sigma)) \vdash L : \cL(\tau)$.
+Now we extend $\sigma$ with more constructors into $\sigma'$, but we don't have any rules to transfer $L$ to the context of $\bW(\sigma')$.
+Thus the correct way is to construct $L$ in an \textit{sealed} context, $  ..
+(T : \cU) \vdash L : \cL(\tau)$, and single out the \textit{recursor term}  $..
+(T : \bW(\sigma)), L : \cL(\tau) \vdash \ \text{ Recursor } \ T \ L ..
+: R $ also functioning as \textit{coverage checking}.
+(Because we don't need to inherit coverage checking -- we should redo it in every linkage) Here, the recursor should also be \textit{seal-ed} so that every field is relying on an abstract, overridable recursor.
+Thus the sealing judgement will be helpful to shift the context from $..
+(T : \bW(\sigma)) ..
+$ to $..
+(T : \cU) ..$
+
+However, in most of the cases, we don't need to seal anything, and thus the "Seal-Id" is used.
+For notation clearance, we will sometime use $sf'$ to emphasize that we consider that seal $s$ as a term $(sf' \ s)$.
+
+\judgebox{\goodInh{\Gamma}{h}{\sigma}{\tau}}
+
+Finally we introduce the inheritance judgement.
+Inheritance is represented by judgement so naturally first order data (i.e.
+not something function can return), but it itself looks very much alike a function that transform a given linkage.

### 2022-05-15 standardmodel.tex
Save.

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{kaposi2017type}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into Agda components that is also respecting judgemental equalities.\footnote{Of course, it is also possible to interpret to constructive set theory.}
-Here we will only show some interpretation, the complete version please refer to appendix.
+We prove the consistency by extending the standard model from \citep{kaposi2017type}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities for accessibility\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.}
+
+Here we will only show some interpretation and also omit the universe level to sketch out the idea, please refer to appendix for the complete version.
+
+\newcommand{\goodTypeS}[3]{{ {#1} \models_S {#2} }}
+\newcommand{\goodTermS}[3]{{ {#1} \models_S {#2} : {#3} }}
+\newcommand{\goodSubS}[3]{{ {#1} \models_S {#2} : {#3} }}
+\newcommand{\goodSigS}[3]{{ {#1} \models_S {#2} \ \  Sig^{#3} }}
+\newcommand{\goodWSigS}[3]{{ {#1} \models_S {#2} \ \ WSig^{#3} }}
+\newcommand{\goodSealS}[4]{{ {#1} \models_S {#2} : {#3} \  |\  {#4} }}
+\newcommand{\goodInhS}[4]{{ {#1} \models_S {#2} : {#3} \twoheadrightarrow {#4}}}
 
 \begin{align*}
-  ss
+  \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set} \\
+  \denotesS{\goodType{\Gamma}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gamma}{i}} \\
+  & \text{ and we define } \goodTypeS{\Gamma}{T}{i} : \iff \denotesS{\goodType{\Gamma}{T}{i}} \text{ is defined } \\
+  \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gamma}{\gamma}{\Delta}} : \iff \denotesS{\gamma} \in \denotesS{\goodSub{\Gamma}{\_}{\Delta}} \\
+  \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gamma}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
+  \denotesS{\goodTerm{\Gamma}{\_}{T}} & \text{ is a Set (of dependent function) where }   {\goodTermS{\Gamma}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gamma}{\_}{T}} \\
+  \denotesS{\goodTerm{\Gamma}{\_}{T}} & = \prod_{\gamma \in \denotesS{\goodCtx{\Gamma}{i}}}\denotesS{T}(\gamma) \\
+  % 
+  \denotesS{\goodSig{\Gamma}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gamma}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gamma}{\_}{n}} \\
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} & \text{ is a family of sets, dependent on } \denotesS{\sigma}  \\
+  \denotesS{\goodSig{\Gamma}{\_}{n+1}} &= \{
+    (\denotesS{\sigma}, \denotesS{\sigma'}, \denotesS{f}, \denotesS{T}) :
+      {\goodSigS{\Gamma}{\sigma}{n}}
+      \land  {\goodSigS{\Gamma}{\sigma'}{n}} \\  
+      & \quad \quad \quad \land  {\goodSealS{\Gamma}{f}{\sigma}{\sigma'}}
+      \land  {\goodTypeS{\Gamma, \cL \sigma'}{T}{i}} 
+  \} \\ 
+  \denotesS{\goodSig{\Gamma}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{(m, f) :  m \in \denotesS{\cL \tau}(\gamma) \land f \in \prod_{m' \in \denotesS{\cL \tau'}(\gamma)}\denotesS{T}(\gamma, m')  \}  \\
+  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{\tau'}, \_ , \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
 \end{align*}
 
-Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we interpret this derivation in our standard model and we can conclude $\denotes{t} : Data.Empty$,
-% which is basically $\denotes{t} \in \emptyset$ set theoretically, 
+Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have  $\denotesS{t}() \in \denotesS{\bot}() = \emptyset$ 
 and that is a contradiction.

### 2022-05-15 standardmodel.tex
Save.

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{kaposi2017type}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities for accessibility\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.}
+We prove the consistency by extending the standard model from \citep{kaposi2017type}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities for accessibility\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
+we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.
+we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma.
+We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$, and also we omit the verification of equality (i.e.
+the judgementally-equal syntax needs to be mapped to the same data for well-defined-ness).
 
@@
 
+\newcommand{\goodCtxS}[2]{{ {#1} \ \models_S }}
 \newcommand{\goodTypeS}[3]{{ {#1} \models_S {#2} }}
@@
 \begin{align*}
-  \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set} \\
+  \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set and we define } \goodCtxS{\Gamma}{i} :\iff \denotesS{\goodCtx{\Gamma}{i}} \text{ is defined} \\
+  \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
   \denotesS{\goodType{\Gamma}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gamma}{i}} \\
@@
   \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gamma}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
-  \denotesS{\goodTerm{\Gamma}{\_}{T}} & \text{ is a Set (of dependent function) where }   {\goodTermS{\Gamma}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gamma}{\_}{T}} \\
+  \denotesS{\goodTerm{\Gamma}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gamma}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gamma}{\_}{T}} \\
   \denotesS{\goodTerm{\Gamma}{\_}{T}} & = \prod_{\gamma \in \denotesS{\goodCtx{\Gamma}{i}}}\denotesS{T}(\gamma) \\
@@
   & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{\tau'}, \_ , \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} \ (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
+  \denotesS{\goodSig{\Gamma}{\nu\cdot}{0}} &= \star \\
+  \denotesS{\goodSig{\Gamma}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, \sigma', f, T) \\
+  \denotesS{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot} } \ (\gamma) &= \star \\
+  \denotesS{\goodTerm{\Gamma}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\gamma)&= (o(\gamma), (\lambda \ m' . t'(\gamma, m'))) \\
+  \denotesS{\goodType{\Gamma}{\bot}{i}} \ (\gamma) &= \emptyset \\
+  \denotesS{\goodType{\Gamma}{\cB}{i}} \ (\gamma) &= \{0, 1\} \quad \text{a specific two element set} \\
 \end{align*}
 
-Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have  $\denotesS{t}() \in \denotesS{\bot}() = \emptyset$ 
+We will take some examples on the verification of validity of this model: 1. to verify compatibility lemma ; 2. to verify the well-defined-ness of this model,
+
+Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cL\sigma}{i}}$ is mutual recursively defined together.
+What's more, the substitution law around these two are also proved together by induction.
+
+We omit the model here for extensible inductive type and inheritance judgement because they require inductive facility in meta-logic.
+In set theory, that would be using \textit{rule sets}\citep{timany2017consistency,aczel1998relating}, which would be more confusing for the reader.
+In appendix we will use the native inductive notion from Agda to construct models.
+
+But we can sketch out their construction -- 1.
+for extensible W type, we simply construct a similar inductive type reflect the same idea, and the challenge mainly lies in satisfying the substitution law; 2.
+for inheritance judgement, we inductively define an almost-identical syntactic structure as model, but not a QIIT, and we inductively define meta-level substitution to implement the explicit substitution.
+We also define $inh$ operation by induction strictly following $\beta$-law.
+
+Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot \Rightarrow \denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
 and that is a contradiction.

### 2022-05-15 standardmodel.tex
Save.

@@
 We omit the model here for extensible inductive type and inheritance judgement because they require inductive facility in meta-logic.
-In set theory, that would be using \textit{rule sets}\citep{timany2017consistency,aczel1998relating}, which would be more confusing for the reader.
+In set theory, that would be using \textit{rule sets}~\citep{timany2017consistency,aczel1998relating}, which would be more confusing for the reader.
 In appendix we will use the native inductive notion from Agda to construct models.
@@
 for extensible W type, we simply construct a similar inductive type reflect the same idea, and the challenge mainly lies in satisfying the substitution law; 2.
-for inheritance judgement, we inductively define an almost-identical syntactic structure as model, but not a QIIT, and we inductively define meta-level substitution to implement the explicit substitution.
+for inheritance judgement, we inductively define an almost-identical syntactic structure as model, (but not a QIIT), and we inductively define meta-level substitution to implement the explicit substitution.
 We also define $inh$ operation by induction strictly following $\beta$-law.

### 2022-05-15 standardmodel.tex
TODO: 1. finish inh-rules type setting 2. examples for verification 3. formalize argument for canonicity

@@
 the judgementally-equal syntax needs to be mapped to the same data for well-defined-ness).
+
+Note that, since every piece is intrinsic, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gamma}{t}{T}}$.
 
@@
 We omit the model here for extensible inductive type and inheritance judgement because they require inductive facility in meta-logic.
-In set theory, that would be using \textit{rule sets}~\citep{timany2017consistency,aczel1998relating}, which would be more confusing for the reader.
+In set theory, that would be using \textit{Rule sets}~\citep{timany2017consistency,aczel1998relating}, which would be more confusing for the reader.
 In appendix we will use the native inductive notion from Agda to construct models.

### 2022-05-18 introduction.tex
Save. Introduction.

% mention expression problem and family polymorphism solving it
% proof engineering, not being focused, fampoly on proof engineering
% contribution

### 2022-05-18 coqimpl.tex
Save.

After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each command into a bunch of Coq commands (the surface Vernacular command) instead of modify the code-base of Coq.

Doing so, Despite of the possible difficulty of the future maintenance, this approach has various advantages: 1. it is the most easy and accessible way to prototype; 2. we have a clear definition of trusted-base -- the whole Coq; 3. it is easy to debug -- we just need to check the translated commands; 4. it can be well-incorporated with the VSCoq (and proof general I believe); 5. it is more accessible for the the interested audience who can easily adapt our plugin and give a try -- otherwise they have to download and re-build the whole customized Coq; 6. it is more stable because the surface syntax of Coq should be stable across different versions; 7. still, this plugin can capture the key mechanism of implementing family polymorphism inside Coq and act as a reference for guiding an appropriate implementation of Family Polymorphism inside all sorts of proof assistants.

### 2022-05-30 coqimpl.tex
Save.

@@
 it is more stable because the surface syntax of Coq should be stable across different versions; 7.
-still, this plugin can capture the key mechanism of implementing family polymorphism inside Coq and act as a reference for guiding an appropriate implementation of Family Polymorphism inside all sorts of proof assistants.
+still, this plugin can capture the key ingredients of implementing family polymorphism inside Coq and act as a reference for guiding an appropriate implementation of Family Polymorphism inside all sorts of proof assistants.
+
+Propositional Partial Recursor can prove injection and discrimination of the constructors.
+Propositional Partial Recursor is good enough because it and its computational axiom can prove that the vanilla inductive type can "embed" into the extensible inductive type. 
+(i.e. there will be an left inversion of that injection, which is also witnessing the fact that those "constructors" can really act as constructors) 
+Thus 1.
+every future extension can support this partial recursion; 2.
+every type support this partial recursor with its computational axiom at least "support these constructors" (because of the embedding).

### 2022-06-04 coqimpl.tex
Save.

@@
-After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each command into a bunch of Coq commands (the surface Vernacular command) instead of modify the code-base of Coq.
+After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
 
@@
 
+\textbf{Family compiled into module, Family type into module type and Context into module parameter} This is the main idea of compiling family components into Coq's primitives.
+To achieve incremental type checking, we notice the nature of late binding: for each defining field $ ..
+\cL\sigma \vdash t : T $, we can compile $t$ using ``universal-quantifier-wrapped term'' : $..
+\vdash \lambda \cL \sigma.
+t : \forall \cL \sigma.
+T$.
+However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping -- each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each term of the field will compile into a parametrized module (functor).
+Doing so we can also get rapid feedback from type checking when the users are defining each field.
+Overriding is achieved in the same manner -- we just need to make sure the Overriding term has the same type as the original term.
+
+% Concrete details of compilation of a family term data structure
+Now that we know a family data structure is actually a list of compiled modules, we need to aggregate them into a complete compiled module.
+Thankfully, Coq's \textit{Include} command makes things easy and reasonably fast.
+
+% Compilation into Coq can give us computational interpretation for free.
+
+\textbf{Only deal with inheritance judgement}
+
+\textbf{Reasoning Requires Non-Overridable}
+However, unlike general-purposed programming, Coq should be able to  \textit{reason} about each program.
+However, sometimes reasoning requires a specific program and overriding on them is unacceptable.
+For example, we may define a field $add : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to proving its commutativity in the following field $comm$.
+However, to prove so, we need to know more than just a type interface of $add$ but also its \textbf{concrete definition}.
+Otherwise, in the future, once $add$ is overridden into a subtraction function and $comm$ will have problem to be  inherited.
+To achieve such exposure of definition in meta-theory we only need to expose the information as propositional equality as part of the signature, but in our Coq plugin, we can achieve that even easier by exposing the whole definition directly into module type.
+Doing so we can have a better judgemental equality and we can avoid overriding because subtraction doesn't have the same definition as $add$.
+We require the users to decorate field as `Final' so that the plugin will proceed with this special treatment.
+
+\textbf{Special Fields: Overridable Family and Sealing as Interface}
+
+\textbf{Special Fields : Extensible Inductive Type and Recursor}
+
+We also provide the tool of `FTheorem' to wrap the proof-irrelevant recursor.
+This tool can avoid most boiler-plate code when writing a standard recursor, and is also open to extension like `FRecursor'.
+
+\textbf{Propositional Partial Recursor}
 Propositional Partial Recursor can prove injection and discrimination of the constructors.

### 2022-06-05 coqimpl.tex
Save.

@@
-After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
+After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
 
-Doing so, Despite of the possible difficulty of the future maintenance, this approach has various advantages: 1.
-it is the most easy and accessible way to prototype; 2.
+Despite of the possible difficulty of the future maintenance, this approach has various advantages: 1.
+it is the easiest and the most accessible way to prototype as it relieves us the necessity of familarity of Coq base, especially for the implementation of the module and functor; 2.
 we have a clear definition of trusted-base -- the whole Coq; 3.
 it is easy to debug -- we just need to check the translated commands; 4.
-it can be well-incorporated with the VSCoq (and proof general I believe); 5.
-it is more accessible for the the interested audience who can easily adapt our plugin and give a try -- otherwise they have to download and re-build the whole customized Coq; 6.
+it can be well-incorporated with the existent tools like VSCoq; 5.
+it is more accessible for the the interested audience who can then easily adapt our plugin and give a try -- otherwise they have to download and re-build the whole customized Coq; 6.
 it is more stable because the surface syntax of Coq should be stable across different versions; 7.
@@
 
-\textbf{Family compiled into module, Family type into module type and Context into module parameter} This is the main idea of compiling family components into Coq's primitives.
+\textbf{Family compiled into module.} This is the main idea of compiling family components into Coq's primitives: we compile family into coq module, family type into module type and context of a given term into parameters of the module.
 To achieve incremental type checking, we notice the nature of late binding: for each defining field $ ..
@@
 
-% Concrete details of compilation of a family term data structure
 Now that we know a family data structure is actually a list of compiled modules, we need to aggregate them into a complete compiled module.
 Thankfully, Coq's \textit{Include} command makes things easy and reasonably fast.
+% Concrete details of compilation of a family term data structure
 
@@
 
-\textbf{Only deal with inheritance judgement}
+In meta-theory, we consider each family (including children family) as an inheritance judgement -- thus we directly construct the data structure representing inheritance judgement during interactive proving.
+As before, an empty family would be an inheritance judgement with empty family as input.
+Consider inheritance judgement as data gives us the ability to consider mix-in -- even though we didn't provide a formal definition here, we will illustrate its usage in the example.
+% {Only deal with inheritance judgement}
 
-\textbf{Reasoning Requires Non-Overridable}
+\textbf{Reasoning Requires Non-Overridable.}
 However, unlike general-purposed programming, Coq should be able to  \textit{reason} about each program.
@@
 
-\textbf{Special Fields: Overridable Family and Sealing as Interface}
+\textbf{Special Fields: Overridable and Sealed Family as Interface.}
 
-\textbf{Special Fields : Extensible Inductive Type and Recursor}
+\textbf{Special Fields: Extensible Inductive Type and Defining Recursor.}
 
@@
 
-\textbf{Propositional Partial Recursor}
+\textbf{Propositional Partial Recursor.}
 Propositional Partial Recursor can prove injection and discrimination of the constructors.
@@
 every type support this partial recursor with its computational axiom at least "support these constructors" (because of the embedding).
+
+\textbf{Hack: Delayed Verifying Fact.}
+
+\section{Example}
+We provide two examples here for illustration and evaluation.
+
+\subsection{Type Safety for STLC}
+We start with the very basic type safety proof of STLC, largely adapted from Software Foundation, which is also one of the primal motivating exampele of this project.
+
+The key highlight of our example here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately.
+And we use a  \textit{not-yet-completely-defined} \textbf{mix-in} to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
+This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/family}.
+This implies, we can really think of programming language feature as plugin, not only in compilers or interpreters, but also in meta-theoretic reasoning.
+
+\subsection{Abstract Interpreters for Imp}
+The second example is adapted and modified from the Familia paper -- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
+Then we extend the language and we instantiate the postulation for both concrete interpreters and abstract interpreters.
+Thanks to the compilation, we can directly run the resulting abstract interpreters.

### 2022-06-05 coqimpl.tex
Save.

@@
 \subsection{Type Safety for STLC}
-We start with the very basic type safety proof of STLC, largely adapted from Software Foundation, which is also one of the primal motivating exampele of this project.
+We start with the very basic type safety proof of STLC, largely adapted from Software Foundation, which is also one of the primal motivating example of this project.
 
-The key highlight of our example here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately.
-And we use a  \textit{not-yet-completely-defined} \textbf{mix-in} to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
-This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/family}.
-This implies, we can really think of programming language feature as plugin, not only in compilers or interpreters, but also in meta-theoretic reasoning.
+Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the examples of MTC.
+And we use a \textit{not-yet-completely-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
+This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/inherited family}.
+Compared to MTC, our example uses small-step operational semantic by exploiting the extensible inductive type, and most of the proof is directly adapted from the one in Software Foundation, resulting a more accessible proof.
+
+% We have problem on inversion lemma. Check if it is the same problem as MTC
+We use `Closing Fact' to state and prove the inversion lemma instead of using the extensible proving mechanism `FTheorem'.
+The reason is that 1.
+it introduce much less boilerplate code because the proof for these inversion lemma should be just simple case analysis and we should rely on Coq to auto-generate them; 2.
+it shouldn't bother us in the future because any extension on the syntax should still satisfy these inversion lemmas; 3.
+most importantly, we believe this inversion ``lemma'' should be part of the definition of the syntax instead of considering the syntax as a mere concrete inductive type.
+We should postulate this inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify that our inductive definition did satisfy the constraint, which is exactly what we expect from `Closing Fact'.
 
 \subsection{Abstract Interpreters for Imp}
-The second example is adapted and modified from the Familia paper -- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
-Then we extend the language and we instantiate the postulation for both concrete interpreters and abstract interpreters.
+The second example is adapted and modified from the Familia paper -- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
+Then we extend the language feature and we instantiate the postulation on computation for both concrete interpreters and abstract interpreters.
 Thanks to the compilation, we can directly run the resulting abstract interpreters.

### 2022-06-05 canonicitymodel.tex
Save.

We follow \citep{sterling2019algebraic,coquand2018canonicity} to construct the canonicity model. We will first specify the formulation of the basic canonicity theorem.

We will resort to both conventional math notation and Agda syntax to explain the main idea of canonicity model.

% After gluing, we figure out the following "canonicity" theorem as well
% the canonical form of extensible inductive type, linkage and boolean

### 2022-06-05 coqimpl.tex
Save.

@@
 
-Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the examples of MTC.
+Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the examples of MTC\citep{delaware2013,forsta2020}.
 And we use a \textit{not-yet-completely-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
@@
 Compared to MTC, our example uses small-step operational semantic by exploiting the extensible inductive type, and most of the proof is directly adapted from the one in Software Foundation, resulting a more accessible proof.
+% In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach.
+Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users -- which can be reflected from the distance of their approach and text-book Software Foundation proof.
 
@@
 \subsection{Abstract Interpreters for Imp}
-The second example is adapted and modified from the Familia paper -- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
+The second example is adapted and modified from the Familia paper
+% citation needed here
+-- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
 Then we extend the language feature and we instantiate the postulation on computation for both concrete interpreters and abstract interpreters.

### 2022-06-05 standardmodel.tex
Save.

--- 2022-05-15 standardmodel.tex
@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{kaposi2017type}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities for accessibility\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
+We prove the consistency by extending the standard model from \citep{kaposi2017type, kaposi2019gluing}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities for accessibility\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
 we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.
 we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma.
-We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$, and also we omit the verification of equality (i.e.
+We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is doing exactly an interpretation function, and in our agda formulation we only do 1 because 2 is directly implied by 1; we make 2 explicit here to make our text more accessible and uniform}, and also we omit the verification of equality (i.e.
 the judgementally-equal syntax needs to be mapped to the same data for well-defined-ness).
+
+Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gamma}{t}{T}}$ (i.e. every term is tracking its context and type)
 
@@
 
-We omit the model here for extensible inductive type and inheritance judgement because they require inductive facility in meta-logic.
-In set theory, that would be using \textit{rule sets}\citep{timany2017consistency,aczel1998relating}, which would be more confusing for the reader.
-In appendix we will use the native inductive notion from Agda to construct models.
+% We omit the model here for extensible inductive type and inheritance judgement because they require inductive facility in meta-logic. 
+We spell-out the model for extensible type and inheritance judgement here, and both of them require inductive facility in meta-logic.
+In set theory, that would be using \textit{Rule sets}~\citep{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead.
+% In appendix we will use the native inductive notion from Agda to construct models. 
+As expected, we use an inductive type with the constructor from W-type to model our extensible inductive type, and the recursor for the inductive type is directly modelling the recursor for our extensible inductive type.
+For inheritance judgement, we inductively define an almost-identical syntactic structure as model, (but not a QIIT), and we inductively define meta-level substitution to implement the explicit substitution.
+We also define $inh$ operation by induction strictly following $\beta$-law.
+Basically both ideas are using the syntax of themselves in meta-logic to model, because its definition is pretty "well-founded".
+Most importantly, this doesn't affect our aim to model bottom type using empty set to derive consistency.
 
-But we can sketch out their construction -- 1.
-for extensible W type, we simply construct a similar inductive type reflect the same idea, and the challenge mainly lies in satisfying the substitution law; 2.
-for inheritance judgement, we inductively define an almost-identical syntactic structure as model, but not a QIIT, and we inductively define meta-level substitution to implement the explicit substitution.
-We also define $inh$ operation by induction strictly following $\beta$-law.
-
-Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot \Rightarrow \denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
+Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
 and that is a contradiction.

### 2022-06-05 syntax.tex
Save.

@@
 $$
-\Rule[name=Inductive Type]
+\Rule[name=Emp WSig]
+{}
+{\goodWSig{\Gamma}{w\cdot}{0}}
+\quad
+\Rule[name=WSig Add]
+{\goodWSig{\Gamma}{s}{n}
+  \quad \goodType{\Gamma}{A}{i}
+  \quad \goodType{\Gamma, A}{B}{i}}
+{\goodWSig{\Gamma}{w^+ \  s \  A \  B}{n+1}}
+\quad
+\Rule[name=Ind Univ]
 {\goodWSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{W\sigma}{i}}
+{\goodType{\Gamma}{\bW \sigma}{i}}
+$$
+
+$$
+\Rule[name=WSig Proj]
+{\goodWSig{\Gamma}{s}{n} \quad j < n}
+{\goodType{\Gamma}{\pi^j_1 s}{i} \quad \goodType{\Gamma, \pi^j_1 s}{\pi^j_2  s}{i}}
 \quad
-\Rule[name=Inductive Sig]
+\Rule[name=Ind Sig]
 {\goodWSig{\Delta}{\sigma}{n}
@@
   \quad \goodType{\Gamma}{W (\sigma[\gamma]) \equiv (W \sigma)[\gamma]}{i}}
+$$
+
+$$
+\Rule[name=Ind Type]
+{\goodTerm{\Gamma}{T}{\bW \sigma}}
+{\goodTerm{\Gamma}{W \ T}{\cU}}
+\quad
+\Rule[name=Ind Term]
+{\goodTerm{\Gamma}{T}{\bW \sigma}
+  \quad \goodTerm{\Gamma}{a}{\pi^j_1\sigma}
+  \quad \goodTerm{\Gamma, \pi^j_1\sigma}{b}{\pi^j_2\sigma}}
+{\goodTerm{\Gamma}{Wsup \ T \ a \ b}{El\ (W\ T)} }
 $$
@@
 Finally we introduce the inheritance judgement.
-Inheritance is represented by judgement so naturally first order data (i.e.
-not something function can return), but it itself looks very much alike a function that transform a given linkage.
+Inheritance are judgement so naturally second class citizen data (i.e.
+not something function can return), and it itself looks very much alike a function that transform a given linkage.
+
+The syntax of the inheritance judgement is very close to how Family is defined in the surface syntax -- a Family with no parent is an inheritance with empty input; an extended Family is exactly an inheritance -- because the parent of an extended Family can be overridden thus inheritance should only define upon the ``interface'' instead of a concrete family.
+However, since our current formulation doesn't support further-binding (i.e.
+we cannot do inheritance inside a family), our formulation still have some distance to the real family polymorphism.

### 2022-06-05 syntax.tex
Save.

@@
 Of course we need to put an eye on the substitution law for each piece of the syntax.
+
 In our meta-theory, we only formulate non-dependent eliminator, i.e.
 the \textit{recursor}.
-Each recursor is constructed by a linkage contained with functions dealing with each case.
-Thus code-reuse for recursor is delegated by the inheritance for linkage.
+Each recursor is constructed by a linkage contained with functions/handlers/pattern-matching cases dealing with each constructor.
+Thus code-reuse for recursor is delegated by the inheritance of linkage -- to reuse one particular pattern-matching case, we just inherit the function/handler.
+This greatly simplifies the metatheory development.
 
@@
 
+Note that, there will be a lot of $\sigma_1, \sigma_1'$ and $\sigma_2, \sigma_2'$, in these cases, $\sigma_1'$ is the result of the sealing of $\sigma_1$, i.e.
+we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{\sigma_1'}, \goodSeal{\Gamma}{f_2}{\sigma_2}{\sigma_2'}$ when needed.
+
 \judgebox{\goodInh{\Gamma}{h}{\sigma}{\tau}}
+
+$$
+\Rule[name=Inh-Id]
+{}
+{\goodInh{\Gamma}{"inhid"}{\sigma}{\sigma}}
+\quad
+\Rule[name=Inh-Override]
+{
+\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}  
+\quad \goodType{\Gamma, \cL\sigma_1'}{T_1}{}
+\quad \goodType{\Gamma, \cL\sigma_2'}{T_2}{}
+  \quad \goodTerm{\Gamma, \cL \sigma_2'}{t}{T_2}}
+{\goodInh{\Gamma}{"inhov" \ h \ t}{(\nu^+ \  \sigma_1 \  T_1)}{(\nu^+ \  \sigma_2\  T_2)}}
+$$
+
+$$
+\Rule[name=Inh-Ext]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+  \quad \goodTerm{\Gamma, \cL \sigma_2'}{t}{T}}
+{\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}
+\quad
+\Rule[name=Inh-Inh]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+\quad \goodType{\Gamma, \cL\sigma_1'}{T}{}
+\quad \goodTerm{\Gamma, \cL\sigma_2'}{\uparrow^s}{\cL\sigma_1'[\pi_1]}
+}
+{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}
+$$
+$$
+\Rule[name=Inh+Inh]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+\quad \goodTerm{\Gamma, \cL\sigma_2'}{\uparrow^s}{\cL\sigma_1'[\pi_1]}
+\quad 
+\goodInh{\Gamma, \cL\sigma_2'}{i}{\tau_1[(\pi_1, \uparrow^s)]}{\tau_2}}
+{\goodInh{\Gamma}{"inhnest" \ h \ i}{(\nu^+ \  \sigma_1\  \cL\tau_1)}{(\nu^+ \  \sigma_2\  \cL\tau_2)}}
+\quad
+$$
+
+$$
+\Rule[name=Inh]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+\quad \goodTerm{\Gamma}{l}{\cL \sigma_1}
+}
+{\goodTerm{\Gamma}{("inh" \ h \ l)}{\cL \sigma_2}} 
+\quad 
+\Rule[name=Inh-Inh-beta]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}}
+{\goodTerm{\Gamma}{"inh" ("inhinh" \ h) (\mu^+ \ m \ t) \equiv \mu^+ \ ("inh" \ h \ m) \ t[(\pi_1, \uparrow^s)]}{}} 
+$$
 
@@
 Inheritance are judgement so naturally second class citizen data (i.e.
-not something function can return), and it itself looks very much alike a function that transform a given linkage.
+not something function can return and cannot be assigned to variables), and it looks very much alike a function that transform a given linkage.
 
@@
 we cannot do inheritance inside a family), our formulation still have some distance to the real family polymorphism.
+
+"inhid" and "inhext" are simple as expected -- the former corresponds to empty inheritance, and the latter gives the programmer the ability to add new field.
+
+"inhinh" is special because it requires user to provide a proof of ``upcasting'' from the context of the children to that of the parent.
+In other words, if I want to inherit a specific field from the parent, I have to convince the type checker that my current context can ``accommodate'' the inherited field.
+
+Override "inhov" is even more special because in mundane OO programming language, we usually require that the overriding term has the same type/interface as the overriden one.
+Here we don't since the very reason we want the same type/interface is that we hope other inherited fields can use this overridden/late-binding field without breaking abstraction, and this reason is already managed by the assumption of "inhinh", the upcasting proof $\uparrow^s$.
+Thus "inhov" is quite like "inhext", where overriding is just throwing away one parent field.
+This difference also show one of the distinct feature of our vanilla family inheritance compared to the mainstream OO inheritance -- family inheritance don't need to introduce the concept of ``subclassing'' between families and thus no bounded polymorphism yet\footnote{Of course once trait and interface comes in, we can have bounded polymorphism back}, and code reuse is mainly achieved by late binding.
+Without ``subclassing'', there is no way of confusing inheritance and subtyping.
+
+Finally we need to introduce nested inheritance "inhnest" for dealing with nested families.
+The rule is mostly direct and we need "upcasting" again when dealing with nested family in different but inheriting context.
+We also show how to use an inheritance judgement to derive children family and how the beta rule is defined.
+We omit most of the other substitution laws here.

### 2022-06-05 syntax.tex
Save.
YZ comments in syntax.tex

--- 2022-06-05 syntax.tex
@@
 
-Because we consider each judgement as a (dependent) QIIT type, each judgement has a \textit{type} and each sequent is well-typed as well (otherwise how can they be encoded in Agda?).
+Because we consider each judgement as a (dependent) QIIT type, each judgement has a \textit{type} and each sequent is well-typed as well (otherwise how can they be encoded in Agda?)\YZ{What's the distinction of judgments vs.
+sequents?}.
 Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements and their Agda counterpart to make things more understandable.
@@
 \Rule[name=Func Subst]
-{\goodSub{\Gamma}{\gamma}{\Delta}}
+{\goodSub{\Gamma}{\gamma}{\Delta}\YZ{What is the requirement on Delta?}}
 {\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}}
@@
 {\goodWSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{\bW \sigma}{i}}
+{\goodType{\Gamma}{\bW \sigma}{i}\YZ{Is there an introduction rule for terms of type 'bW sigma'?}}
 $$
@@
 : R $ also functioning as \textit{coverage checking}.
+\YZ{Does coverage checking mean exhaustiveness checking?}
 (Because we don't need to inherit coverage checking -- we should redo it in every linkage) Here, the recursor should also be \textit{seal-ed} so that every field is relying on an abstract, overridable recursor.
-Thus the sealing judgement will be helpful to shift the context from $..
-(T : \bW(\sigma)) ..
-$ to $..
-(T : \cU) ..$
+\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
+ML, sealing has the connotation of hiding things behind an existential
+type. So using the same word may cause the reader to have false
+presumptions about what is going on here.}
+Thus the sealing judgement will be helpful to shift the context from $.. (T : \bW(\sigma)) .. $ to $.. (T : \cU) ..$
+\YZ{Can we have a more concrete example of how the judgment is useful?}
 

### 2022-06-05 syntax.tex

--- 2022-06-05 syntax.tex
@@
 \Rule[name=Inh-Inh-beta]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}}
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+  \quad \goodTerm{\Gamma}{m}{\cL \sigma_1}
+  \quad \goodTerm{\Gamma, \cL\sigma_1'}{t}{T}
+}
 {\goodTerm{\Gamma}{"inh" ("inhinh" \ h) (\mu^+ \ m \ t) \equiv \mu^+ \ ("inh" \ h \ m) \ t[(\pi_1, \uparrow^s)]}{}} 

### 2022-06-06 standardmodel.tex
Reply to some of the comments. Still need time to resolve all the problems

@@
 we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma.
-We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is doing exactly an interpretation function, and in our agda formulation we only do 1 because 2 is directly implied by 1; we make 2 explicit here to make our text more accessible and uniform}, and also we omit the verification of equality (i.e.
+We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e.
 the judgementally-equal syntax needs to be mapped to the same data for well-defined-ness).

### 2022-06-06 syntax.tex
Reply to some of the comments. Still need time to resolve all the problems

@@
 Because we consider each judgement as a (dependent) QIIT type, each judgement has a \textit{type} and each sequent is well-typed as well (otherwise how can they be encoded in Agda?)\YZ{What's the distinction of judgments vs.
-sequents?}.
+sequents?}.\EDJreply{I am not sure about the correct terminology here.
+Here I want to say "even underivable judgement has to be well-typed".
+That's the main reason I spell out the fake Agda code there.
+I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}
 Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements and their Agda counterpart to make things more understandable.
@@
 \Rule[name=Func Subst]
-{\goodSub{\Gamma}{\gamma}{\Delta}\YZ{What is the requirement on Delta?}}
+{\goodSub{\Gamma}{\gamma}{\Delta}\YZ{What is the requirement on Delta?}
+\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context.
+This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well).
+For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing.
+What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda.
+For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer fake agda formulation) }
+}
 {\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}}
@@
 {\goodWSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{\bW \sigma}{i}\YZ{Is there an introduction rule for terms of type 'bW sigma'?}}
+{\goodType{\Gamma}{\bW \sigma}{i}\YZ{Is there an introduction rule for terms of type 'bW sigma'?}\EDJreply{Yes there is.
+I omit it because it is a singleton type, please see my newly added explanation at below.}}
 $$
@@
 Of course we need to put an eye on the substitution law for each piece of the syntax.
+Here $\bW \sigma$ is a large singleton type, and it only has one "element", the inductive type itself, inside.
+We formulate in this cumbersome way because we want to have inductive type to be a field element at the same time exposing the constructors for later \textit{pattern-matching coverage checking} (i.e.
+if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed)\EDJ{Here}.
 
@@
 We have projection for each componenet due to harmony \citep{pfenning2009lecture} (local soundness and completeness), but we omit them and the substitution laws for either concatenation and projection.
-For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.
+For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using }.
 
@@
 type. So using the same word may cause the reader to have false
-presumptions about what is going on here.}
+presumptions about what is going on here.}\EDJreply{You are right.
+Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type".
+Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}
 Thus the sealing judgement will be helpful to shift the context from $.. (T : \bW(\sigma)) .. $ to $.. (T : \cU) ..$
-\YZ{Can we have a more concrete example of how the judgment is useful?}
+\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{}
 

### 2022-06-06 syntax.tex
address comments

--- 2022-06-05 syntax.tex
@@
 
-Because we consider each judgement as a (dependent) QIIT type, each judgement has a \textit{type} and each sequent is well-typed as well (otherwise how can they be encoded in Agda?)\YZ{What's the distinction of judgments vs.
-sequents?}.
+Because we consider each judgement as a (dependent) QIIT type, each judgement has a \textit{type} and each sequent is well-typed as well (otherwise how can they be encoded in Agda?)
+%\YZ{What's the distinction of judgments vs.
+sequents?}.\EDJreply{I am not sure about the correct terminology here.
+Here I want to say "even underivable judgement has to be well-typed".
+That's the main reason I spell out the fake Agda code there.
+I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}
 Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements and their Agda counterpart to make things more understandable.
@@
 \Rule[name=Func Subst]
-{\goodSub{\Gamma}{\gamma}{\Delta}\YZ{What is the requirement on Delta?}}
+{\goodSub{\Gamma}{\gamma}{\Delta}\YZ{What is the requirement on Delta?}
+\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context.
+This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well).
+For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing.
+What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda.
+For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer fake agda formulation) }\YZreply{I was thinking about 'Delta |- A'.
+I suppose it is also implicitly enforced via intrinsically typed syntax.
+Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}
+}
 {\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}}
@@
 {\goodWSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{\bW \sigma}{i}\YZ{Is there an introduction rule for terms of type 'bW sigma'?}}
+{\goodType{\Gamma}{\bW \sigma}{i}}
 $$
+%\YZ{Is there an introduction rule for terms of type 'bW sigma'?}
+%\EDJreply{Yes there is. I omit it because it is a singleton type, please see my newly added explanation below.}
 
@@
 Of course we need to put an eye on the substitution law for each piece of the syntax.
+Here $\bW \sigma$ is a large singleton type \cite{stone2000}, and it only has one "element", the inductive type itself, inside.
+We formulate in this cumbersome way because we want to have inductive type to be an abstract field element at the same time exposing the constructors for later \textit{pattern-matching coverage checking} (i.e.
+if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}
 
@@
 Thus code-reuse for recursor is delegated by the inheritance of linkage -- to reuse one particular pattern-matching case, we just inherit the function/handler.
-This greatly simplifies the metatheory development.
+This greatly simplifies the meta\-theory development.
 
@@
 (Because we don't need to inherit coverage checking -- we should redo it in every linkage) Here, the recursor should also be \textit{seal-ed} so that every field is relying on an abstract, overridable recursor.
-\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
-ML, sealing has the connotation of hiding things behind an existential
-type. So using the same word may cause the reader to have false
-presumptions about what is going on here.}
+%\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
+%ML, sealing has the connotation of hiding things behind an existential
+%type. So using the same word may cause the reader to have false
+%presumptions about what is going on here.}\EDJreply{You are right.
+Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type".
+Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}
 Thus the sealing judgement will be helpful to shift the context from $.. (T : \bW(\sigma)) .. $ to $.. (T : \cU) ..$
-\YZ{Can we have a more concrete example of how the judgment is useful?}
+\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{}
 

### 2022-06-06 syntax.tex
Unresolve some comments asking for futher checking Some Further feedback. DM me if you think some my un-resolution is unnecessary.

@@
 
-Because we consider each judgement as a (dependent) QIIT type, each judgement has a \textit{type} and each sequent is well-typed as well (otherwise how can they be encoded in Agda?)
-%\YZ{What's the distinction of judgments vs.
+Because we consider each kind of judgement as a (dependent) QIIT type, thus each judgement has a \textit{type} and all the derivation of judgements are well-typed (QIIT terms) as well (otherwise how can they be encoded in Agda?)
+\YZ{What's the distinction of judgments vs.
 sequents?}.\EDJreply{I am not sure about the correct terminology here.
@@
 That's the main reason I spell out the fake Agda code there.
-I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}
-Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements and their Agda counterpart to make things more understandable.
-Of course, a well-formed sequent  is not necessarily derivable.
-Due to our intrinsic setting, we can omit a lot of presumption, for example we don't spell out that the rule "Type-Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
+I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}\EDJreply{I rephrase that so there won't be confusion for the future readers.
+Please check}
+Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements and their Agda counterpart to make things more accessible.
+Of course, a well-formed judgement is not necessarily derivable, just like a well-typed QIIT type is not necessarily inhabited.
+
+Due to our intrinsic setting, \textbf{we can and will omit a lot of presumption of the typing rules}, for example we don't spell out that the rule "Type-Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
 
@@
 I suppose it is also implicitly enforced via intrinsically typed syntax.
-Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}
+Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}\EDJreply{You are absolutely right!
+I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
+\quad \goodType{\Delta}{A}{} 
+\quad \goodType{\Delta, A}{B}{}
 }
@@
 We formulate in this cumbersome way because we want to have inductive type to be an abstract field element at the same time exposing the constructors for later \textit{pattern-matching coverage checking} (i.e.
-if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}
+if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type.
+They are basically the same thing but (W T) cannot be at the type position in a judgement, i.e.
+k |- q : (W T) is not allowed, we only have k |- q : El (W T).
+This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}
 
@@
 : R $ also functioning as \textit{coverage checking}.
-\YZ{Does coverage checking mean exhaustiveness checking?}
+\YZ{Does coverage checking mean exhaustiveness checking?}\EDJreply{Yes.
+The one make sure every case is handled in pattern-matching.
+Do you want me to replace the terminology using exhaustiveness checking?
+}
 (Because we don't need to inherit coverage checking -- we should redo it in every linkage) Here, the recursor should also be \textit{seal-ed} so that every field is relying on an abstract, overridable recursor.
-%\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
-%ML, sealing has the connotation of hiding things behind an existential
-%type. So using the same word may cause the reader to have false
-%presumptions about what is going on here.}\EDJreply{You are right.
+\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
+ML, sealing has the connotation of hiding things behind an existential
+type. So using the same word may cause the reader to have false
+presumptions about what is going on here.}\EDJreply{You are right.
 Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type".
-Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}
+Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear -- because I haven't changed any text}
 Thus the sealing judgement will be helpful to shift the context from $.. (T : \bW(\sigma)) .. $ to $.. (T : \cU) ..$
-\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{}
+\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{Working on it, don't resolve}
 

### 2022-06-06 syntax.tex
comments

@@
 
-Because we consider each kind of judgement as a (dependent) QIIT type, thus each judgement has a \textit{type} and all the derivation of judgements are well-typed (QIIT terms) as well (otherwise how can they be encoded in Agda?)
-\YZ{What's the distinction of judgments vs.
+Because we formalize our language as a shallow embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms (otherwise how can they be encoded in Agda?)
+%\YZ{What's the distinction of judgments vs.
 sequents?}.\EDJreply{I am not sure about the correct terminology here.
@@
 \Rule[name=Func Subst]
-{\goodSub{\Gamma}{\gamma}{\Delta}\YZ{What is the requirement on Delta?}
-\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context.
+{\goodSub{\Gamma}{\gamma}{\Delta}
+\quad \goodType{\Delta}{A}{} 
+\quad \goodType{\Delta, A}{B}{}
+}
+{\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}}
+$$
+%\YZ{What is the requirement on Delta?}
+%\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context.
 This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well).
@@
 I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
-\quad \goodType{\Delta}{A}{} 
-\quad \goodType{\Delta, A}{B}{}
-}
-{\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}}
-$$
 $$
@@
 k |- q : (W T) is not allowed, we only have k |- q : El (W T).
-This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}
+This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}\YZreply{I think explanation is warranted; readers unfamiliar with that line of work may find it confusing.
+BTW, is the elimination rule for "El (W T)" omitted?}
 
@@
 Do you want me to replace the terminology using exhaustiveness checking?
-}
+}\YZreply{Yes, I think exhaustiveness is standard parlance.}
 (Because we don't need to inherit coverage checking -- we should redo it in every linkage) Here, the recursor should also be \textit{seal-ed} so that every field is relying on an abstract, overridable recursor.

### 2022-06-06 coqimpl.tex
Reconsiderate phrase of "Extensible"

@@
 
-\textbf{Special Fields: Extensible Inductive Type and Defining Recursor.}
+\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
+
+Even though in our metatheory there is no notion of extensible inductive type, but in our implementation we have the syntax of extensible inductive type.
 

### 2022-06-06 standardmodel.tex
Reconsiderate phrase of "Extensible"

@@
 
-% We omit the model here for extensible inductive type and inheritance judgement because they require inductive facility in meta-logic. 
-We spell-out the model for extensible type and inheritance judgement here, and both of them require inductive facility in meta-logic.
+% We omit the model here for inductive type and inheritance judgement because they require inductive facility in meta-logic. 
+We spell-out the model for the inductive type and inheritance judgement here, and both of them require inductive facility in meta-logic.
 In set theory, that would be using \textit{Rule sets}~\citep{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead.
 % In appendix we will use the native inductive notion from Agda to construct models. 
-As expected, we use an inductive type with the constructor from W-type to model our extensible inductive type, and the recursor for the inductive type is directly modelling the recursor for our extensible inductive type.
+As expected, we use an inductive type in the meta-logic with the constructor from W-type to model our inductive type, and the recursor for the inductive type in the meta-logic is directly modelling the recursor for our inductive type in our theory.
 For inheritance judgement, we inductively define an almost-identical syntactic structure as model, (but not a QIIT), and we inductively define meta-level substitution to implement the explicit substitution.

### 2022-06-06 syntax.tex
Reconsiderate phrase of "Extensible"

@@
 
-We have two kinds of signatures, one for extensible inductive type, the other for linkages.
+We have two kinds of signatures, one for inductive type, the other for linkages.
 
@@
 Recall in W-type \citep{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type.
-For simplicity,  we use a list of pairs of such $A, B$ as the signature for a given inductive type.
-Doing so, all the constructors will have a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
+To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
+Thus we use a list of pairs of such $A_i, B_i$ as the signature for a given inductive type.
+
 Of course we need to put an eye on the substitution law for each piece of the syntax.

### 2022-06-11 syntax.tex
Save. Working on the example

@@
 We are following \citep{kaposi2019gluing} but we will omit the universe level attached to the judgement here, to make it more readable.
-After these, we will have four new kinds of judgement.
+After these, we will have four new kinds of judgement to model our linkage.
 
-Because we formalize our language as a shallow embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms (otherwise how can they be encoded in Agda?)
+Because we formalize our language as a deep embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms (otherwise how can they be encoded in Agda?)
 %\YZ{What's the distinction of judgments vs.
@@
 
-Due to our intrinsic setting, \textbf{we can and will omit a lot of presumption of the typing rules}, for example we don't spell out that the rule "Type-Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
+Due to our intrinsic setting, \textbf{we can and will omit a lot of presumption of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
 
-Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{\tau}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{\cL\tau[\pi_1]}$.
-Even though we will introduce rules for sealing judgement, we will later show that all these rules are derivable by term judgements.
+Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{\cL\tau[\pi_1]}$.
+Though we will still introduce rules for sealing judgement, and show that all these rules are derivable by term judgements.
 
@@
       \Rule[name=Seal]
-      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma, \tau}{n}}
-      {\goodSeal{\Gamma}{\_}{\sigma}{\tau} \text{ is well-formed}}
+      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
+      \quad \goodType{\Gamma}{A}{} }
+      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
       $$
@@
 
-Now we show case some of the typing rules, and demonstrate their typed encoding in our (fake) Agda, to help the reader  grasp the rough idea of our type system and understand the Agda-formulation of the typing rules easier.
+Now we show case some of the typing rules, we will have dependent function type and dependent pair type.
+Dependent pair type is used to model Coq module, which is useful when we need to model the ``compilation'' from linkages to modules.
 
@@
 $$
-\Rule[name=Type Subst]
-{\goodType{\Delta}{T}{j} 
-  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodType{\Gamma}{T[\gamma]}{j}}
 \quad 
-\Rule[name=Func Subst]
+\Rule[name=Func/DPair Subst]
 {\goodSub{\Gamma}{\gamma}{\Delta}
@@
 }
-{\goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}}
+{
+  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
+  \quad 
+  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
+}
 $$
@@
 $$
+\Rule[name=Type Subst]
+{\goodType{\Delta}{T}{j} 
+  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+{\goodType{\Gamma}{T[\gamma]}{j}}
+\quad
 \Rule[name=Base Type Subst]
@@
 {\goodType{\Gamma}{T}{j}}
-{\goodTerm{\Gamma}{c \ T}{\cU_j}
+{\goodTerm{\Gamma}{"c" \ T}{\cU_j}
 }\quad
@@
 {}
-{\goodTerm{\Gamma}{tt}{\cB}}
+{\goodTerm{\Gamma}{"tt"}{\cB}}
 \quad
@@
 {}
-{\goodTerm{\Gamma}{ff}{\cB}}
+{\goodTerm{\Gamma}{"ff"}{\cB}}
 $$
@@
 {\goodType{\Gamma}{(El \ T)[\gamma] \equiv El \ (T[\gamma]) }{j} \quad
- \goodTerm{\Gamma}{tt[\gamma] \equiv tt}{\cB} \quad 
- \goodTerm{\Gamma}{ff[\gamma] \equiv ff}{\cB} 
+ \goodTerm{\Gamma}{"tt"[\gamma] \equiv "tt"}{\cB} \quad 
+ \goodTerm{\Gamma}{"ff"[\gamma] \equiv "ff"}{\cB} 
 }
@@
 \Rule[name=Id Subst]
-{}{\goodSub{\Gamma}{id}{\Gamma}}
+{}{\goodSub{\Gamma}{"id"}{\Gamma}}
 \quad
@@
 {\goodTerm{\Gamma}{T}{\bW \sigma}}
-{\goodTerm{\Gamma}{W \ T}{\cU}}
+{\goodTerm{\Gamma}{"W" \ T}{\cU}}
 \quad
@@
   \quad \goodTerm{\Gamma, \pi^j_1\sigma}{b}{\pi^j_2\sigma}}
-{\goodTerm{\Gamma}{Wsup \ T \ a \ b}{El\ (W\ T)} }
+{\goodTerm{\Gamma}{"Wsup" \ T \ a \ b}{El\ ("W" \ T)} }
 $$
@@
 To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
-Thus we use a list of pairs of such $A_i, B_i$ as the signature for a given inductive type.
+Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type.
 
@@
 Here $\bW \sigma$ is a large singleton type \cite{stone2000}, and it only has one "element", the inductive type itself, inside.
-We formulate in this cumbersome way because we want to have inductive type to be an abstract field element at the same time exposing the constructors for later \textit{pattern-matching coverage checking} (i.e.
+We formulate in this cumbersome way because we want to have inductive type to be a field element and exposing the constructors at the type level for later \textit{pattern-matching coverage checking} (i.e.
 if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type.
@@
 This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}\YZreply{I think explanation is warranted; readers unfamiliar with that line of work may find it confusing.
-BTW, is the elimination rule for "El (W T)" omitted?}
+BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general.
+I just didn't write out the two equations c (El T) = T, El (c T) = T}.
+We follow the style of \textit{type universe à la Tarski} and distinguish types from their \textit{codes or names}, and thus we use $El\ T$ as the type given type name $T : \cU$ -- in these cases, $T$ is not allowed to locate at type position (i.e.
+at the right-most position of a term judgement) but $El \ T$ is.
 
@@
 {\goodSig{\Gamma}{\sigma}{n} 
- \quad \goodSeal{\Gamma}{f}{\sigma}{\sigma'}
- \quad \goodType{\Gamma, \sigma'}{T}{i}}
+ \quad \goodSeal{\Gamma}{f}{\sigma}{A}
+ \quad \goodType{\Gamma, A}{T}{i}}
 {\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
@@
 { \goodTerm{\Gamma}{o}{\cL \sigma} 
-\quad  \goodSeal{\Gamma}{f}{\sigma}{\sigma'} 
- \quad \goodTerm{\Gamma, \cL \sigma'}{t}{T}
+\quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
+ \quad \goodTerm{\Gamma, A}{t}{T}
 }
@@
 We have projection for each componenet due to harmony \citep{pfenning2009lecture} (local soundness and completeness), but we omit them and the substitution laws for either concatenation and projection.
-For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using }.
+For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using Implicit Arguments}.
 
-Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by that of dependent pairs, so an obvious question is why is that sealing judgement there?
+Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by the design of our linkage, so an obvious question is why is that sealing judgement there?
 
-\judgebox{[\goodSeal{\Gamma}{s}{\sigma}{\sigma'}] = [\goodTerm{\Gamma, \cL\sigma}{s}{\cL \sigma'[\pi_1]}] }
+\judgebox{[\goodSeal{\Gamma}{s}{\sigma}{A}] 
+= [\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}] }
 $$
 \Rule[name=Seal Subst]
-{\goodSeal{\Delta}{s}{\sigma}{\sigma'}
+{\goodSeal{\Delta}{s}{\sigma}{A}
   \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{\sigma'[\gamma]}}
+{\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{A[\gamma]}}
 =
 \frac
-{\goodTerm{\Delta, \cL\sigma}{s}{\cL\sigma'[\pi_1]}
+{\goodTerm{\Delta, \cC\sigma}{s}{A[\pi_1]}
   \quad  \goodSub{\Gamma}{\gamma}{\Delta}  }
-{\goodTerm{\Gamma,\cL\sigma[\gamma]}{s[\gamma^\uparrow]}{\cL\sigma'[\pi_1][\gamma^\uparrow]\equiv\cL\sigma'[\gamma][\pi_1]}}
+{\goodTerm{\Gamma,\cC\sigma[\gamma]}{s[\gamma^\uparrow]}{A[\pi_1][\gamma^\uparrow]\equiv A[\gamma][\pi_1]}}
 $$
@@
 {}
-{\goodSeal{\Gamma}{id_s}{\sigma}{\sigma}}
-= \goodTerm{\Gamma, \cL\sigma}{\pi_2}{\cL\sigma[\pi_1]}
+{\goodSeal{\Gamma}{id_s}{\sigma}{\cC \sigma}}
+= \goodTerm{\Gamma, \cC\sigma}{\pi_2}{\cC\sigma[\pi_1]}
 $$
@@
 
-As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{\sigma'}$ is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{\cL\tau[\pi_1]}$.
+As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just a short hand of term judgement $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$.
 The substitution law should understand as a lemma, and easily derivable by using $\gamma^\uparrow$.
@@
 Note that, there will be a lot of $\sigma_1, \sigma_1'$ and $\sigma_2, \sigma_2'$, in these cases, $\sigma_1'$ is the result of the sealing of $\sigma_1$, i.e.
-we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{\sigma_1'}, \goodSeal{\Gamma}{f_2}{\sigma_2}{\sigma_2'}$ when needed.
+we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}, \goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption when needed.
 
@@
 \goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}  
-\quad \goodType{\Gamma, \cL\sigma_1'}{T_1}{}
-\quad \goodType{\Gamma, \cL\sigma_2'}{T_2}{}
+\quad \goodType{\Gamma, A_1}{T_1}{}
+\quad \goodType{\Gamma, A_2}{T_2}{}
   \quad \goodTerm{\Gamma, \cL \sigma_2'}{t}{T_2}}
@@
 {\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodType{\Gamma, \cL\sigma_1'}{T}{}
-\quad \goodTerm{\Gamma, \cL\sigma_2'}{\uparrow^s}{\cL\sigma_1'[\pi_1]}
+\quad \goodType{\Gamma, A_1}{T}{}
+\quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}
 }
@@
 {\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodTerm{\Gamma, \cL\sigma_2'}{\uparrow^s}{\cL\sigma_1'[\pi_1]}
+\quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}
 \quad 
-\goodInh{\Gamma, \cL\sigma_2'}{i}{\tau_1[(\pi_1, \uparrow^s)]}{\tau_2}}
+\goodInh{\Gamma, A_2}{i}{\tau_1[(\pi_1, \uparrow^s)]}{\tau_2}}
 {\goodInh{\Gamma}{"inhnest" \ h \ i}{(\nu^+ \  \sigma_1\  \cL\tau_1)}{(\nu^+ \  \sigma_2\  \cL\tau_2)}}
@@
   \quad \goodTerm{\Gamma}{m}{\cL \sigma_1}
-  \quad \goodTerm{\Gamma, \cL\sigma_1'}{t}{T}
+  \quad \goodTerm{\Gamma, A_1}{t}{T}
 }

### 2022-06-11 syntactic-example.tex
Save. Add new section.

We use an example to better demonstrate different components of the syntax of our formulation. Our example will be the meta-theory encoding of the following pseudo-code:

### 2022-06-11 syntactic-example.tex
Save, Working on example.

@@
 We use an example to better demonstrate different components of the syntax of our formulation. Our example will be the meta-theory encoding of the following pseudo-code:
+
+\begin{minted}{Coq}
+  (* First Note that how inductive type is written in our format
+     we use 1, 0 to indicate unit type and bottom type *)
+  Inductive N :=
+    | O : 1 -> (0 -> N) -> N 
+    | S : 1 -> (1 -> N) -> N
+  (* For example, Number 1 is encoded using
+      1 = S () (fun _ -> O () (fun x -> x is ⊥)) *)
+
+Family A.
+  Inductive N := O : N. 
+  Family handler.
+    O : self[A].N * self[A].N 
+      := (self[A].O, self[A].O).
+  EndFamily.
+  pred : self[A].N -> self[A].N * self[A].N 
+    := (* handler as recursor *) .
+  k := self[A].pred self[A].O
+EndFamily.
+
+Family A2 extends A.
+  Extend Inductive N := S : N -> N.
+  Extend Family handler.
+    Inherit O.
+    S : N*N -> N*N 
+      := fun (x,y) => (y, self[A2].S y).
+  EndFamily.
+  pred : self[A].N -> self[A].N * self[A].N 
+    := (* .. Override ..  *) .
+  Inherits k
+EndFamily.
+\end{minted}
+
+\begin{align*}
+  "Osig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  w\cdot}{1} \\
+  "sig"_1 &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Osig"[\pi_1])}{1}  \\
+  "obj"_1 &:= \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ *}{\cL"sig"_1}\\
+  \cC"sig"_1 &= \goodType{\cdot}{\Sigma \ \top \ (\bW \ "Osig"[\pi_1]) }{} \\
+  "sig"_2 &:=  \goodSig{\cdot}{\nu^+ \ "sig"_1 \ {id_s} \ El ("W" ("pjr" \ \pi_2))}{2} \\ 
+  "O" &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  & \text{ and thus }  \goodTerm{\cdot, \cC"sig"_1}{"O"}{El ("W" ("pjr" \pi_2))} \\
+  "obj"_2 &:= \goodTerm{\cdot}{\mu^+ \ "obj"_1 \ "O"}{\cL"sig"_2} \\
+  "sig"_2' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU)El ("pjr" \pi_2)}{} \\ 
+  \cC"sig"_2 &= \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El \ W("pjr" \pi_2) }{} \\ 
+  "sl"_1 &:= ((*, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
+  & \text{ and thus }  \goodSeal{\cdot}{"sl"_1}{"sig"_2}{"sig"_2'} \\
+  "pN"_1 &:= \goodType{\cdot, "sig"_2'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "handler"_1 & := \goodTerm{\cdot, "sig"_2'}{\_}{\cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)} \\
+  "sig"_3 &:= \goodSig{\cdot}{\nu^+ "sig"_2 \{"sl"_1\} \ \cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)}{3} \\ 
+  "predT" := (\Pi & \ El ("W"("pjr" ("pjl" ("pjl" \pi_2)))) \\  
+  & (El ("W" ("pjr" ("pjl" ("pjl" \pi_2))))[\pi_1])) \\
+  "pred"_1 & := \goodTerm{\cdot, \cC "sig"_3}{\lambda ("Wrec" \_)}{predT} \\ 
+  "sig"_4 &:= \nu^+ \ "sig"_3 \ \{id_s\} \ "predT" \\ 
+  "sl"_4 &:= \goodSeal{\cdot}{\_}{"sig"_4}{"sig"_4'} \\ 
+  "pN"_2 &:= \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "sig"_5 &:= \nu^+ \ "sig"_4 \ \{"sl"_4\} \  "pN"_2 \\ 
+  "Nsig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  "Osig"}{1} \\ 
+  "sig"_{11} &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
+  "inh"_1 &:= \goodInh{\cdot}{"inhid"}{"sig"_1}{"sig"_{11}}\\
+  "O"' &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  & \text{ but }  \goodTerm{\cdot, \cC"sig"_{11}}{"O"'}{El ("W" ("pjr" \pi_2))} \\
+  "sig"_{21} &:= \nu^+ \ "sig"_{11} \ \{id_s\} (El (W (pjr p₂))) \\ 
+  "inh"_2 &:= \goodInh{\cdot}{"inhov" \ "inh"_1 \ "O"'}{"sig"_2}{"sig"_{21}}\\
+  "predT"' := (\Pi & \ El ("W"("pjr"  ("pjl" \pi_2))) \\  
+  & (El ("W" ("pjr" ("pjl"  \pi_2)))[\pi_1])) \\
+  "sig"_{22} &:= \nu^+ \ "sig"_{21} \ \{id_s\} \ "predT"' \\ 
+  "sig"_{21}' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU) (El ("pjr" \pi_2))}{} \\
+  "sig"_{22}' &:= \goodType{\cdot}{\Sigma "sig"_{21}' (\Pi \ El ("pjr"  ("pjl" \pi_2)) (El  ("pjr" ("pjl"  \pi_2))[\pi_1]))}{} \\
+  "sl"_{22} &:= \goodSeal{\cdot}{\_}{"sig"_{22}}{"sig"_{22}'} \\ 
+  \uparrow^s &:= \goodTerm{\cdot, "sig"_{22}'}{\_}{"sig"_2'[\pi_1]} \\ 
+  "inh"_3 &:= \goodInh{\cdot}{\_}{"sig"_2}{"sig"_{22}} \\ 
+  "pN"_3 &:= "pN"_1[(\pi_1, \uparrow^s)] \\ 
+  "inh"_{in0} &:= \goodInh{\cdot, "sig"_{22}'}{"inhid"}{(\nu^+ \ \nu\cdot \ ("pN"_1[\pi_1]))[(\pi_1, \uparrow^s)]}{(\nu^+ \cdot ("pN"_3[\pi_1]))} \\
+  "inh"_{in} &:= \goodInh{\cdot , "sig"_{22}'}{\_}{(\nu^+ \ \nu\cdot \ "pN"_1[\pi_1])[(\pi_1, \uparrow^s)]}{(\nu^+ \ \cdot \ ("pN"_3[\pi_1]) \ \{id_s\} (\Pi "pN"_3 ("pN"_3[\pi_1])) )[\pi_1]}\\
+  "sig"_{31} &:=  \nu^+ \ "sig"_{22} \ {"sl"_{22}}\  \cL(\nu^+ \cdot ("pN"_3[\pi_1])  (\Pi "pN"_3 ("pN"_3[\pi_1]))[\pi_1]) \\ 
+  "inh"_4 &:= \goodInh{\cdot}{"inh+inh" \ "inh"_3 \uparrow^s "inh"_{in}}{"sig"_3}{"sig"_{31}} 
+\end{align*}

### 2022-06-11 syntax.tex
Save. Add new section.
Save. Working on the example

@@
 
-The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance.
-Consider the scenario we have constructed a linkage meant to be a recursor $  ..
-(T : \bW(\sigma)) \vdash L : \cL(\tau)$.
-Now we extend $\sigma$ with more constructors into $\sigma'$, but we don't have any rules to transfer $L$ to the context of $\bW(\sigma')$.
-Thus the correct way is to construct $L$ in an \textit{sealed} context, $  ..
-(T : \cU) \vdash L : \cL(\tau)$, and single out the \textit{recursor term}  $..
-(T : \bW(\sigma)), L : \cL(\tau) \vdash \ \text{ Recursor } \ T \ L ..
-: R $ also functioning as \textit{coverage checking}.
+The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance -- due to the presence of inductive type in the context, not everything is inheritable.
+For example, $ X_1 : \bW \sigma_1 \vdash t : T$ in the parent family will generally not be able to be inherited to the children once we override the inductive type.
+In other words, there is no way ``transplanting'' $t$ into the context $X_2 : \bW \sigma_2$.
+This makes all the fields after this inductive type non-inheritable -- which shouldn't be the case because a lot of fields might not be directly relying on this concrete $\bW \sigma_1$.
+
+Sealing is used to exploit this fact -- with sealing, we can add well-typed terms $X_1 : \cU \vdash t : T$, into the parent family, in the context of $ X_1 : \bW \sigma_1$.
+Later when we want to inherit $t$, because $t$ only consider $X_1$ as an arbitrary type, inheritance is plausible.
+
+Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ and thus non-inheritable -- in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}.
+These fields together with inductive types are the one sealing responsible for hiding.
+After compilation of linkage type into existential type, we can easily hide the textit{concrete} definition of the inductive type while exposing all the fields using "Wsup" and "Wrec" like an abstract interface.
+Apparently, future fields relies on this abstract interface will have no problem on inheritance.
 \YZ{Does coverage checking mean exhaustiveness checking?}\EDJreply{Yes.
@@
 }\YZreply{Yes, I think exhaustiveness is standard parlance.}
-(Because we don't need to inherit coverage checking -- we should redo it in every linkage) Here, the recursor should also be \textit{seal-ed} so that every field is relying on an abstract, overridable recursor.
+Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden.
+For "Wsup" we only need to override each constructor and for "Wrec" it is just doing exhaustiveness checking inside each children family.
+The recursor handlers however, can be safely inherited.
 \YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
@@
 Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear -- because I haven't changed any text}
-Thus the sealing judgement will be helpful to shift the context from $.. (T : \bW(\sigma)) .. $ to $.. (T : \cU) ..$
+
 \YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{Working on it, don't resolve}
 
-However, in most of the cases, we don't need to seal anything, and thus the "Seal-Id" is used.
+However, there are cases we don't need to seal anything (i.e.
+doing exhaustiveness checking when using recursor), and thus the "Seal-Id" is used.
 For notation clearance, we will sometime use $sf'$ to emphasize that we consider that seal $s$ as a term $(sf' \ s)$.
@@
 We omit most of the other substitution laws here.
+
+\subsection{Exemplar Usage of Meta Theory}
+\input{syntactic-example.tex}

### 2022-06-11 syntax.tex
Save. Rephrasing.

@@
 These fields together with inductive types are the one sealing responsible for hiding.
-After compilation of linkage type into existential type, we can easily hide the textit{concrete} definition of the inductive type while exposing all the fields using "Wsup" and "Wrec" like an abstract interface.
+After compilation of linkage type into existential type, we can easily hide the \textit{concrete} definition of the inductive type while exposing all the fields using "Wsup" and "Wrec" like an abstract interface.
 Apparently, future fields relies on this abstract interface will have no problem on inheritance.
@@
 
-However, there are cases we don't need to seal anything (i.e.
+And these are also the cases we don't need to seal anything (i.e.
 doing exhaustiveness checking when using recursor), and thus the "Seal-Id" is used.
@@
 
-Note that, there will be a lot of $\sigma_1, \sigma_1'$ and $\sigma_2, \sigma_2'$, in these cases, $\sigma_1'$ is the result of the sealing of $\sigma_1$, i.e.
+Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e.
 we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}, \goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption when needed.

### 2022-06-12 standardmodel.tex
Save.

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{kaposi2017type, kaposi2019gluing}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities for accessibility\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
+We prove the consistency by extending the standard model from \citep{kaposi2017type, kaposi2019gluing}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
 we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.
@@
   \denotesS{\goodSig{\Gamma}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gamma}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gamma}{\_}{n}} \\
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} & \text{ is a family of sets, dependent on } \denotesS{\sigma}  \\
+  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} & \text{ is a family of sets, dependent on and defined inductively on } \denotesS{\sigma} \\   
   \denotesS{\goodSig{\Gamma}{\_}{n+1}} &= \{
-    (\denotesS{\sigma}, \denotesS{\sigma'}, \denotesS{f}, \denotesS{T}) :
+    (\denotesS{\sigma}, \denotesS{A}, \denotesS{f}, \denotesS{T}) :
       {\goodSigS{\Gamma}{\sigma}{n}}
-      \land  {\goodSigS{\Gamma}{\sigma'}{n}} \\  
-      & \quad \quad \quad \land  {\goodSealS{\Gamma}{f}{\sigma}{\sigma'}}
-      \land  {\goodTypeS{\Gamma, \cL \sigma'}{T}{i}} 
+      \land  {\goodTypeS{\Gamma}{A}{}} \\  
+      & \quad \quad \quad \land  {\goodSealS{\Gamma}{f}{\sigma}{A}}
+      \land  {\goodTypeS{\Gamma, A}{T}{i}} 
   \} \\ 
   \denotesS{\goodSig{\Gamma}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{(m, f) :  m \in \denotesS{\cL \tau}(\gamma) \land f \in \prod_{m' \in \denotesS{\cL \tau'}(\gamma)}\denotesS{T}(\gamma, m')  \}  \\
-  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{\tau'}, \_ , \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
+  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gamma}{\cC \tau}{}} \denotesS{T[sf \ f]} \\
+  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
+  \denotesS{\goodType{\Gamma}{\cC \sigma}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} & \text{ is a family of sets, dependent on } \denotesS{\sigma}  \\
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\gamma) \land t \in \prod_{m' \in \denotesS{A}(\gamma)}\denotesS{T}(\gamma, m')  \}  \\
+  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
   \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} \ (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
   \denotesS{\goodSig{\Gamma}{\nu\cdot}{0}} &= \star \\
-  \denotesS{\goodSig{\Gamma}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, \sigma', f, T) \\
+  \denotesS{\goodSig{\Gamma}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, A, f, T) \\
   \denotesS{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot} } \ (\gamma) &= \star \\
   \denotesS{\goodTerm{\Gamma}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\gamma)&= (o(\gamma), (\lambda \ m' . t'(\gamma, m'))) \\
+  \denotesS{\goodTerm{\Gamma}{\cCt \{\sigma\} o}{\cC\sigma}} & \text{ is a dependent function, defined inductively on } \denotesS{\sigma} \\
+  \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
+  &= \denotesS{\goodTerm{\Gamma}{\cCt (p_1\mu \ o)}{...}}, \denotesS{(p_2\mu \ o)[sf \ f][(id, \cCt (p_1\mu \ o))]} \\
+  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
+  \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
+  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
   \denotesS{\goodType{\Gamma}{\bot}{i}} \ (\gamma) &= \emptyset \\
@@
 
-Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cL\sigma}{i}}$ is mutual recursively defined together.
+Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together -- because $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
 What's more, the substitution law around these two are also proved together by induction.

### 2022-06-12 syntactic-example.tex
Save.

--- 2022-06-11 syntactic-example.tex
@@
-We use an example to better demonstrate different components of the syntax of our formulation. Our example will be the meta-theory encoding of the following pseudo-code:
+We use an example to better demonstrate different components of the syntax of our formulation. Our example will be the meta-theory encoding of the following pseudo-code.
+
+First Note that how inductive type is written in our format. We use 1, 0 here to indicate unit type and bottom type:
+\begin{minted}{Coq}
+Inductive N :=
+ | O : 1 -> (0 -> N) -> N 
+ | S : 1 -> (1 -> N) -> N
+(* For example, Number 1 is encoded using
+   1 = S () (fun _ -> O () (fun x -> elim-⊥ x)) *)
+\end{minted}
+
+\begin{figure}[H]
+\begin{minipage}[t]{0.5\linewidth}
+\begin{minted}{Coq}
+Family A.
+ Inductive N := O : N. 
+ (* Osig, sig₁, obj₁, sig₂,
+    O₁, obj₂, sig₂', sl₁ *)
+ Family handler.
+   O :self[A].N * self[A].N 
+     := (self[A].O, self[A].O).
+
+ EndFamily. (* pN₁, handler₁, sig₃ *)
+ pred 
+  : self[A].N → self[A].N×self[A].N 
+  := ... 
+  (* predT, pred₁, sig₄, sig₄', sl₄ *) 
+ k := self[A].pred self[A].O 
+(* pN₂, sig₅ *)
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \begin{minipage}[t]{0.45\linewidth}
+\begin{minted}{Coq}
+Family A₂ extends A.
+ Extend Inductive N := S : N -> N.
+ (* Nsig, sig₁₁, inh₁, O₂, sig₂₁, inh₂, 
+    sig₂₂, sig₂₁', sig₂₂', sl₂₂, inh₃ *)
+ Extend Family handler.
+   Inherit O. (* inhᵢₙ₀, pN₃ *)
+   S : N*N -> N*N 
+     := fun (x,y) => (y, self[A₂].S y).
+ EndFamily. (* inhᵢₙ, sig₃₁, inh₄*)
+ pred 
+  : self[A₂].N → self[A₂].N×self[A₂].N 
+  := ... 
+  (* sig₄₁, inh₅, sl₄₁ *)
+ Inherits k. 
+ (* inh₆ *)
+EndFamily.
+    \end{minted}
+  \end{minipage}
+  \caption{}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{minipage}[b]{0.3\linewidth}
+
+  \end{minipage}
+  \begin{minipage}[b]{0.6\linewidth}
+
+  \end{minipage}
+  \caption{}
+\end{figure}
+
+\begin{align*}
+  "Osig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  w\cdot}{1} \\
+  "sig"_1 &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Osig"[\pi_1])}{1}  \\
+  "obj"_1 &:= \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ *}{\cL"sig"_1}\\
+  \cC"sig"_1 &= \goodType{\cdot}{\Sigma \ \top \ (\bW \ "Osig"[\pi_1]) }{} \\
+  "sig"_2 &:=  \goodSig{\cdot}{\nu^+ \ "sig"_1 \ {id_s} \ El ("W" ("pjr" \ \pi_2))}{2} \\ 
+  "O"_1 &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  & \text{ and thus }  \goodTerm{\cdot, \cC"sig"_1}{"O"_1}{El ("W" ("pjr" \pi_2))} \\
+  "obj"_2 &:= \goodTerm{\cdot}{\mu^+ \ "obj"_1 \ "O"_1}{\cL"sig"_2} \\
+  "sig"_2' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU)El ("pjr" \pi_2)}{} \\ 
+  \cC"sig"_2 &= \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El \ W("pjr" \pi_2) }{} \\ 
+  "sl"_1 &:= ((*, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
+  & \text{ and thus }  \goodSeal{\cdot}{"sl"_1}{"sig"_2}{"sig"_2'} \\
+  "pN"_1 &:= \goodType{\cdot, "sig"_2'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "handler"_1 & := \goodTerm{\cdot, "sig"_2'}{\_}{\cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)} \\
+  "sig"_3 &:= \goodSig{\cdot}{\nu^+ "sig"_2 \{"sl"_1\} \ \cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)}{3} \\ 
+  "predT" := (\Pi & \ El ("W"("pjr" ("pjl"^2 \pi_2))) \\  
+  & (El ("W" ("pjr" ("pjl"^2  \pi_2)))[\pi_1])) \\
+  "pred"_1 & := \goodTerm{\cdot, \cC "sig"_3}{\lambda ("Wrec" \_)}{predT} \\ 
+  "sig"_4 &:= \nu^+ \ "sig"_3 \ \{id_s\} \ "predT" \\ 
+  "sl"_4 &:= \goodSeal{\cdot}{\_}{"sig"_4}{"sig"_4'} \\ 
+  "pN"_2 &:= \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "sig"_5 &:= \nu^+ \ "sig"_4 \ \{"sl"_4\} \  "pN"_2 \\ 
+  "Nsig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  "Osig"}{1} \\ 
+  "sig"_{11} &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
+  "inh"_1 &:= \goodInh{\cdot}{"inhid"}{"sig"_1}{"sig"_{11}}\\
+  "O"_2 &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  & \text{ but }  \goodTerm{\cdot, \cC"sig"_{11}}{"O"_2}{El ("W" ("pjr" \pi_2))} \\
+  "sig"_{21} &:= \nu^+ \ "sig"_{11} \ \{id_s\} (El (W (pjr p₂))) \\ 
+  "inh"_2 &:= \goodInh{\cdot}{"inhov" \ "inh"_1 \ "O"_2}{"sig"_2}{"sig"_{21}}\\
+  "S"_T := (\Pi & \ El ("W"("pjr"  ("pjl" \pi_2))) \\  
+  & (El ("W" ("pjr" ("pjl"  \pi_2)))[\pi_1])) \\
+  "sig"_{22} &:= \nu^+ \ "sig"_{21} \ \{id_s\} \ "S"_T \\ 
+  "sig"_{21}' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU) (El ("pjr" \pi_2))}{} \\
+  "sig"_{22}' &:= \goodType{\cdot}{\Sigma "sig"_{21}' (\Pi \ El ("pjr"  ("pjl" \pi_2)) (El  ("pjr" ("pjl"  \pi_2))[\pi_1]))}{} \\
+  "sl"_{22} &:= \goodSeal{\cdot}{\_}{"sig"_{22}}{"sig"_{22}'} \\ 
+  \uparrow^s &:= \goodTerm{\cdot, "sig"_{22}'}{\_}{"sig"_2'[\pi_1]} \\ 
+  "inh"_3 &:= \goodInh{\cdot}{\_}{"sig"_2}{"sig"_{22}} \\ 
+  "pN"_3 &:= "pN"_1[(\pi_1, \uparrow^s)] \\ 
+  "inh"_{in0} &:= \goodInh{\cdot, "sig"_{22}'}{"inhid"}{(\nu^+ \ \nu\cdot \ ("pN"_1[\pi_1]))[(\pi_1, \uparrow^s)]}{(\nu^+ \cdot ("pN"_3[\pi_1]))} \\
+  "inh"_{in} &:= \goodInh{\cdot , "sig"_{22}'}{\_}{(\nu^+ \ \nu\cdot \ "pN"_1[\pi_1])[(\pi_1, \uparrow^s)]}{(\nu^+ \ \cdot \ ("pN"_3[\pi_1]) \ \{id_s\} (\Pi "pN"_3 ("pN"_3[\pi_1])) )[\pi_1]}\\
+  "sig"_{31} &:=  \nu^+ \ "sig"_{22} \ {"sl"_{22}}\  \cL(\nu^+ \cdot ("pN"_3[\pi_1])  (\Pi "pN"_3 ("pN"_3[\pi_1]))[\pi_1]) \\ 
+  "inh"_4 &:= \goodInh{\cdot}{"inh+inh" \ "inh"_3 \uparrow^s "inh"_{in}}{"sig"_3}{"sig"_{31}} \\
+  "sig"_{41} &:= {\nu^+ \ "sig"_{31} \ (\Pi (El ("W" ("pjr" ("pjl"^4 \pi_2)))) (El ("W" ("pjr" ("pjl"^4 \pi_2))))[\pi_1])} \\ 
+  "inh"_5 &:= \goodInh{\cdot}{\_}{"sig"_4}{"sig"_{41}} \\ 
+  "sl"_{41} &:= \goodSeal{\cdot}{\_}{"sig"_{41}}{"sig"_{41}'} \\ 
+  \uparrow^s_2 &:= \goodTerm{\cdot, "sig"_{41}}{\_}{"sig"_4'[\pi_1]} \\ 
+  "inh"_6 &:= \goodInh{\cdot}{\_}{"inh"_5}{\uparrow^s_2}
+\end{align*}

### 2022-06-12 syntactic-example.tex
Save. Adding example.

--- 2022-06-11 syntactic-example.tex
@@
-We use an example to better demonstrate different components of the syntax of our formulation. Our example will be the meta-theory encoding of the following pseudo-code:
+We use an example to better demonstrate different components of the syntax of our formulation. Our example will be the meta-theory encoding of the following pseudo-code.
+
+First Note that how inductive type is written in our format. We use 1, 0 here to indicate unit type and bottom type:
+\begin{minted}{Coq}
+Inductive N :=
+ | O : 1 -> (0 -> N) -> N 
+ | S : 1 -> (1 -> N) -> N
+(* For example, Number 1 is encoded using
+   1 = S () (fun _ -> O () (fun x -> elim-⊥ x)) *)
+\end{minted}
+
+\begin{figure}[H]
+\begin{minipage}[t]{0.5\linewidth}
+\begin{minted}{Coq}
+Family A.
+ Inductive N := O : N. 
+ (* Osig, sig₁, obj₁, sig₂,
+    O₁, obj₂, sig₂', sl₁ *)
+ Family handler.
+   O :self[A].N * self[A].N 
+     := (self[A].O, self[A].O).
+
+ EndFamily. (* pN₁, handler₁, sig₃ *)
+ pred 
+  : self[A].N → self[A].N×self[A].N 
+  := ... 
+  (* predT, pred₁, sig₄, sig₄', sl₄ *) 
+ k := self[A].pred self[A].O 
+(* pN₂, sig₅ *)
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \begin{minipage}[t]{0.45\linewidth}
+\begin{minted}{Coq}
+Family A₂ extends A.
+ Extend Inductive N := S : N -> N.
+ (* Nsig, sig₁₁, inh₁, O₂, sig₂₁, inh₂, 
+    sig₂₂, sig₂₁', sig₂₂', sl₂₂, inh₃ *)
+ Extend Family handler.
+   Inherit O. (* inhᵢₙ₀, pN₃ *)
+   S : N*N -> N*N 
+     := fun (x,y) => (y, self[A₂].S y).
+ EndFamily. (* inhᵢₙ, sig₃₁, inh₄*)
+ pred 
+  : self[A₂].N → self[A₂].N×self[A₂].N 
+  := ... 
+  (* sig₄₁, inh₅, sl₄₁ *)
+ Inherits k. 
+ (* inh₆ *)
+EndFamily.
+    \end{minted}
+  \end{minipage}
+  \caption{}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{minipage}[t]{0.4\linewidth}
+\begin{align*}
+  "Osig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  w\cdot}{1} \\
+  "sig"_1 &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Osig"[\pi_1])}{1}  \\
+  "obj"_1 &:= \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ *}{\cL"sig"_1}\\
+  \cC"sig"_1 &= \goodType{\cdot}{\Sigma \ \top \ (\bW \ "Osig"[\pi_1]) }{} \\
+  "sig"_2 &:=  \goodSig{\cdot}{\nu^+ \ "sig"_1 \ {id_s} \ El ("W" ("pjr" \ \pi_2))}{2} \\ 
+  "O"_1 &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  & \text{ and thus }  \goodTerm{\cdot, \cC"sig"_1}{"O"_1}{El ("W" ("pjr" \pi_2))} \\
+  "obj"_2 &:= \goodTerm{\cdot}{\mu^+ \ "obj"_1 \ "O"_1}{\cL"sig"_2} \\
+  "sig"_2' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU)El ("pjr" \pi_2)}{} \\ 
+  \cC"sig"_2 &= \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El \ W("pjr" \pi_2) }{} \\ 
+  "sl"_1 &:= ((*, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
+  & \text{ and thus }  \goodSeal{\cdot}{"sl"_1}{"sig"_2}{"sig"_2'} \\
+  "pN"_1 &:= \goodType{\cdot, "sig"_2'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "handler"_1 & := \goodTerm{\cdot, "sig"_2'}{\_}{\cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)} \\
+  "sig"_3 &:= \goodSig{\cdot}{\nu^+ "sig"_2 \{"sl"_1\} \ \cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)}{3} \\ 
+  "predT" := (\Pi & \ El ("W"("pjr" ("pjl"^2 \pi_2))) \\  
+  & (El ("W" ("pjr" ("pjl"^2  \pi_2)))[\pi_1])) \\
+  "pred"_1 & := \goodTerm{\cdot, \cC "sig"_3}{\lambda ("Wrec" \_)}{predT} \\ 
+  "sig"_4 &:= \nu^+ \ "sig"_3 \ \{id_s\} \ "predT" \\ 
+  "sl"_4 &:= \goodSeal{\cdot}{\_}{"sig"_4}{"sig"_4'} \\ 
+  "pN"_2 &:= \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "sig"_5 &:= \nu^+ \ "sig"_4 \ \{"sl"_4\} \  "pN"_2 \\ 
+\end{align*}
+\end{minipage}%
+\begin{minipage}[t]{0.4\linewidth}
+\begin{align*}
+  "Nsig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  "Osig"}{1} \\ 
+  "sig"_{11} &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
+  "inh"_1 &:= \goodInh{\cdot}{"inhid"}{"sig"_1}{"sig"_{11}}\\
+  "O"_2 &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  & \text{ but }  \goodTerm{\cdot, \cC"sig"_{11}}{"O"_2}{El ("W" ("pjr" \pi_2))} \\
+  "sig"_{21} &:= \nu^+ \ "sig"_{11} \ \{id_s\} (El (W (pjr p₂))) \\ 
+  "inh"_2 &:= \goodInh{\cdot}{"inhov" \ "inh"_1 \ "O"_2}{"sig"_2}{"sig"_{21}}\\
+  "S"_T := (\Pi & \ El ("W"("pjr"  ("pjl" \pi_2))) \\  
+  & (El ("W" ("pjr" ("pjl"  \pi_2)))[\pi_1])) \\
+  "sig"_{22} &:= \nu^+ \ "sig"_{21} \ \{id_s\} \ "S"_T \\ 
+  "sig"_{21}' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU) (El ("pjr" \pi_2))}{} \\
+  "sig"_{22}' &:= \Sigma "sig"_{21}' (\Pi \ El ("pjr"  ("pjl" \pi_2)) \  ...) \\
+  "sl"_{22} &:= \goodSeal{\cdot}{\_}{"sig"_{22}}{"sig"_{22}'} \\ 
+  \uparrow^s &:= \goodTerm{\cdot, "sig"_{22}'}{\_}{"sig"_2'[\pi_1]} \\ 
+  "inh"_3 &:= \goodInh{\cdot}{\_}{"sig"_2}{"sig"_{22}} \\ 
+  "pN"_3 &:= "pN"_1[(\pi_1, \uparrow^s)] \\ 
+\end{align*}
+  \end{minipage}
+
+  \begin{minipage}{\linewidth}
+    \begin{align*}
+      "inh"_{in0} &:= \goodInh{\cdot, "sig"_{22}'}{"inhid"}{(\nu^+ \ \nu\cdot \ ("pN"_1[\pi_1]))[(\pi_1, \uparrow^s)]}{(\nu^+ \cdot ("pN"_3[\pi_1]))} \\
+      "inh"_{in} &:= \goodInh{\cdot , "sig"_{22}'}{\_}{(\nu^+ \ \nu\cdot \ "pN"_1[\pi_1])[(\pi_1, \uparrow^s)]}{(\nu^+ \ \cdot \ ("pN"_3[\pi_1]) \ \{id_s\} (\Pi "pN"_3 ("pN"_3[\pi_1])) )[\pi_1]}\\
+      "sig"_{31} &:=  \nu^+ \ "sig"_{22} \ {"sl"_{22}}\  \cL(\nu^+ \cdot ("pN"_3[\pi_1])  (\Pi "pN"_3 ("pN"_3[\pi_1]))[\pi_1]) \\ 
+      "inh"_4 &:= \goodInh{\cdot}{"inh+inh" \ "inh"_3 \uparrow^s "inh"_{in}}{"sig"_3}{"sig"_{31}} \\
+      "sig"_{41} &:= {\nu^+ \ "sig"_{31} \ (\Pi (El ("W" ("pjr" ("pjl"^4 \pi_2)))) (El ("W" ("pjr" ("pjl"^4 \pi_2))))[\pi_1])} \\ 
+      "inh"_5 &:= \goodInh{\cdot}{\_}{"sig"_4}{"sig"_{41}} \\ 
+      "sl"_{41} &:= \goodSeal{\cdot}{\_}{"sig"_{41}}{"sig"_{41}'} \\ 
+      \uparrow^s_2 &:= \goodTerm{\cdot, "sig"_{41}}{\_}{"sig"_4'[\pi_1]} \\ 
+      "inh"_6 &:= \goodInh{\cdot}{\_}{"inh"_5}{\uparrow^s_2}
+    \end{align*}  
+  \end{minipage}
+  \caption{Detailed Construction}
+\end{figure}

### 2022-06-12 syntactic-example.tex
Save. Example finished.

@@
 
-First Note that how inductive type is written in our format. We use 1, 0 here to indicate unit type and bottom type:
-\begin{minted}{Coq}
-Inductive N :=
- | O : 1 -> (0 -> N) -> N 
- | S : 1 -> (1 -> N) -> N
-(* For example, Number 1 is encoded using
-   1 = S () (fun _ -> O () (fun x -> elim-⊥ x)) *)
-\end{minted}
+We provide one example in its surface syntax as pseudocode, and their meta-theory encoding.
+This example is computing the predecessor of a mundane natural number -- however we split the inductive definition into two parts and making "O" and "S" standing alone.
+"pred" is defined as $n \mapsto ("pred" n, n)$, and thus a predecessor.
 
-\begin{figure}[H]
+\begin{figure}[H]\label{fig:example-pseudocode}
 \begin{minipage}[t]{0.5\linewidth}
@@
   \end{minipage}
-  \caption{}
+  \caption{Demonstrating Pseudo-Code}
 \end{figure}
 
-\begin{figure}[H]
+\begin{figure}
+  \begin{minipage}{\linewidth}
+    \begin{minted}{Coq}
+      Inductive N :=
+       | O : 1 -> (0 -> N) -> N 
+       | S : 1 -> (1 -> N) -> N
+      (* For example, Number 1 is encoded using
+         1 = S () (fun _ -> O () (fun x -> elim-⊥ x)) *)
+    \end{minted}
+  \end{minipage}
+
   \begin{minipage}[t]{0.4\linewidth}
@@
   "sig"_{11} &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
-  "inh"_1 &:= \goodInh{\cdot}{"inhid"}{"sig"_1}{"sig"_{11}}\\
+  "inh"_1 &:= \goodInh{\cdot}{"inhov" \  "inhid" \ *}{"sig"_1}{"sig"_{11}}\\
   "O"_2 &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
@@
   "sl"_{22} &:= \goodSeal{\cdot}{\_}{"sig"_{22}}{"sig"_{22}'} \\ 
+  "inh"_3 &:= \goodInh{\cdot}{\_}{"sig"_2}{"sig"_{22}} \\ 
   \uparrow^s &:= \goodTerm{\cdot, "sig"_{22}'}{\_}{"sig"_2'[\pi_1]} \\ 
-  "inh"_3 &:= \goodInh{\cdot}{\_}{"sig"_2}{"sig"_{22}} \\ 
   "pN"_3 &:= "pN"_1[(\pi_1, \uparrow^s)] \\ 
@@
 
-  \begin{minipage}{\linewidth}
+  \begin{minipage}{0.8\linewidth}
+    \centering
     \begin{align*}
@@
       "sig"_{31} &:=  \nu^+ \ "sig"_{22} \ {"sl"_{22}}\  \cL(\nu^+ \cdot ("pN"_3[\pi_1])  (\Pi "pN"_3 ("pN"_3[\pi_1]))[\pi_1]) \\ 
-      "inh"_4 &:= \goodInh{\cdot}{"inh+inh" \ "inh"_3 \uparrow^s "inh"_{in}}{"sig"_3}{"sig"_{31}} \\
+      "inh"_4 &:= \goodInh{\cdot}{"inhnest" \ "inh"_3 \uparrow^s "inh"_{in}}{"sig"_3}{"sig"_{31}} \\
       "sig"_{41} &:= {\nu^+ \ "sig"_{31} \ (\Pi (El ("W" ("pjr" ("pjl"^4 \pi_2)))) (El ("W" ("pjr" ("pjl"^4 \pi_2))))[\pi_1])} \\ 
-      "inh"_5 &:= \goodInh{\cdot}{\_}{"sig"_4}{"sig"_{41}} \\ 
+      "inh"_5 &:= \goodInh{\cdot}{inhov \ \_}{"sig"_4}{"sig"_{41}} \\ 
       "sl"_{41} &:= \goodSeal{\cdot}{\_}{"sig"_{41}}{"sig"_{41}'} \\ 
       \uparrow^s_2 &:= \goodTerm{\cdot, "sig"_{41}}{\_}{"sig"_4'[\pi_1]} \\ 
-      "inh"_6 &:= \goodInh{\cdot}{\_}{"inh"_5}{\uparrow^s_2}
+      "inh"_6 &:= \goodInh{\cdot}{"inhinh" \ "inh"_5 \ \uparrow^s_2}{"sig"_5}{(\nu^+ \ "sig"_{41} \{"sl"_{41}\} \ "pN"_2[(\pi_1, \uparrow^s_2)])}
     \end{align*}  
   \end{minipage}
-  \caption{Detailed Construction}
+  \caption{Detailed Construction}\label{fig:example-construction}
 \end{figure}
+
+Here we explain how  our detailed construction in Figure \ref{fig:example-construction} is constructed.
+We first elaborate that how our inductive type "N" is written in the specific format (the style of W-type).
+We use 1, 0 here to indicate unit type and bottom type.
+
+Next, we alias each part of the derivation with a name. Every derivation here is well-typed term. We also annotate in the original pseudocode the name of the related derivation.
+
+We start with the construction of "Osig" which is the signature of a standalone "O" constructor.
+With this we will have $"sig"_1, "obj"_1, "sig"_2, "obj"_2$ as the signature and object of the linkage with that inductive type, and the signature and the linkage once we export the concrete constructor $"O"_1$.
+
+Now sealing comes in to seal these concrete component into an abstract interface $"sig"_2'$ for decoupling the following fields.
+To make sure the sealing exists, we take a look at the resulting $\cC "sig"_2$, and doing abstraction on it to get $"sl"_1$.
+
+Now we construct the recursor, with the help of the handler module $"handler"_1$, with which we can construct the real recursor $"pred"_1$.
+Both detailed constructions are omitted.
+Then we again abstract the concrete recursor $"pred"_1$ away using $"sl"_4$ so that the field $"k"$ can applys it without resorting to the concrete recursor.
+Finally we have $"sig"_5$ as final signature.
+
+Now we construct the inheritance judgement for $"A"_2$.
+Still, we start with signature for inductive type, "Nsig".
+But to ``extend'' the inductive type, we actually \textit{override} the inductive type with the enriched inductive type in $"inh"_1$.
+Because inductive type is overridden, inheritance on the constructors are not pausible and thus we need to override the old $"O"_1$ constructor using $"O"_2$ and extend with new constructor for "S", resulting $"inh"_3$.
+These newly overridden constructors are sealed by $"sl"_{22}$.
+However, look closely, our sealed abstract interface in the children still has more fields than the sealed abstract interface in the parent.
+We still need $\uparrow^s$ to indicate the ``compatibility'' between two abstract interfaces so that the inheritance from the parents can work.
+It can be understood as ``upcasting'', but this ``upcasting'' is used to upcast the context of the children.
+We immeidately see its usage in $"pN"_3$ and $"inh"_{in0}$.
+Upon this, we construct the complete inner-inheritance $"inh"_{in}$ that is responsible for extending "handler".
+Then we continue the construction for $"inh"_4$ that ``nest'' the inner-inheritance into the whole $"inh"_3$, and we re-do exhaustiveness checking when constructing $"inh"_5$ using $"inhov"$ and "Wrec".
+Since it is another recursor requiring concrete inductive definition, we do another abstraction on it using $"sl"_41$.
+After that, $\uparrow^s_2$ again witnesses the compatibility of the abstract interface, and the two are both used for constructing $"inh"_6$, which is responsible for inheriting field "k".

### 2022-06-12 syntax.tex
Save.
Save. Example is added, please check.

@@
 
-Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{\cL\tau[\pi_1]}$.
-Though we will still introduce rules for sealing judgement, and show that all these rules are derivable by term judgements.
+Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.
 
@@
 
-We have shown case some exemplar rules for all four standard type judgments, now we focus on the newly introduced facility.
+We have shown case some exemplar rules for all four standard type judgments, and now we focus on the newly introduced facility.
 
@@
 To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
-Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type.
+Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig Proj} to extract and corresponding $A_i, B_i$.
 
@@
 $$
-\Rule[name=Linkage Type]
+\Rule[name=Lnkg Type/Compile]
 {\goodSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{\cL \sigma}{i}}
+{\goodType{\Gamma}{\cL \sigma}{i}
+\quad \goodType{\Gamma}{\cC \sigma}{}}
 \quad
-\Rule[name=Sig/Linkage Subst]
+\Rule[name=Sig/Lnkg Subst]
 {\goodSig{\Delta}{\sigma}{n}
@@
 $$
-\Rule[name=Empty Sig]
+\Rule[name=Ept Sig]
 {}
@@
 $$
+
 $$ 
-\Rule[name=Empty Lnkg]
+\Rule[name=Sig Proj]
+{\goodSig{\Gamma}{\sigma}{n+1}}
+{\goodSig{\Gamma}{p_1\nu \ \sigma}{n}
+\\ \goodType{\Gamma}{p_1\nu' \ \sigma}{}
+\\ \goodSeal{\Gamma}{p_f\nu \  \sigma}{p_1\nu \  \sigma}{p_1\nu' \sigma}
+\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}
+}
+$$
+
+$$
+\Rule[name=Ept Lnkg]
 {}
@@
 
+$$
+\Rule[name=Lnkg Proj]
+{\goodTerm{\Gamma}{o}{\cL\sigma}}
+{\goodTerm{\Gamma}{p_1\mu \ o}{\cL (p_1\nu \ \sigma)}
+\\ \goodTerm{\Gamma, p_1\nu' \ \sigma}{p_2\mu \ o}{p_2\nu \ \sigma}
+}
+\quad 
+\Rule[name=Lnkg Compile]
+{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+}
+{
+  \goodTerm{\Gamma}{\cCt o}{\cC \sigma}
+}
+$$
+
+$$
+\Rule[name=Compile]
+{}
+{\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
+\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}
+\\
+\goodType{\Gamma}{\cC (\nu^+ \ \sigma \ \{f\} \ T) \equiv 
+    \Sigma (\cC \sigma) T[sf \ f]}{}
+\\\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t[sf \ f][(id, \cCt \ o)]) }{}
+}
+$$
+% $$
+% \Rule
+% { \goodSeal{\Gamma}{f}{\sigma}{A}
+% \\ \goodType{\Gamma, A}{T}{}
+% \\ \goodTerm{\Gamma}{t}{T}
+% }{}
+% $$
+
 $\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields.
 We omit the naming/label information for fields in the meta-theory.
-We can use $\cL$ to retrieve the corresponding types of linkage of a given signature.
-Our signature corresponds to Coq's Module Type and our linkage corresponds to Coq's Module.
+We can use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).
 
-The signature for linkage closely follows the signature for inductive type.
+The signature (for the linkage) and the linkage rules are quite similar the signature for inductive type.
 They are both constructed from the empty stuff, and respect substitution lemma for sure.
-We have projection for each componenet due to harmony \citep{pfenning2009lecture} (local soundness and completeness), but we omit them and the substitution laws for either concatenation and projection.
+We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony \citep{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules).
+We omit them and the substitution laws for either field addition and projection.
 For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using Implicit Arguments}.
 
-Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by the design of our linkage, so an obvious question is why is that sealing judgement there?
+Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by the implementation of our linkage, so an obvious question is what is this sealing judgement for?
 
@@
 \begin{minted}[]{agda}
-Seal : (Γ : Con) → (σ : Sig Γ n) → (τ : Sig Γ n) → Set 
-Seal Γ σ τ = Tm (Γ, 𝓛σ) 𝓛τ[p₁]
+Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
+Seal Γ A τ = Tm (Γ, 𝓛σ) A[p₁]
 
--[-]S : (f : Seal Γ σ₁ σ₂) → (τ : Tms Θ Γ) → Seal Θ σ₁[τ] σ₂[τ]
-f[τ]S = t[τ ↑ 𝓛σ₁]
+-[-]S : (f : Seal Γ σ A) → (τ : Tms Θ Γ) → Seal Θ σ[τ] A[τ]
+f[τ]S = t[τ ↑ A]
 
-idₛ : Seal Γ σ σ
+idₛ : Seal Γ σ 𝓒σ
 idₛ = π₁
@@
 
-Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e.
-we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}, \goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption when needed.
+With the definition of sealing judgement in mind, we can see \ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the implementation.
+In the implementation, we will have "self__" all around when defining new fields in our plugin, and this "self__" is exactly the $A$ in the context of the two fields -- especially when we don't do any sealing, this $A = \cC \sigma$, which is exactly how the implementation expose "self__" as a module parameter.
+
+Next we talk about inheritance judgement.
+Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, and in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e.
+we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}, \goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption of judgements when needed.
 

### 2022-06-13 standardmodel.tex
Save.

@@
   \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set and we define } \goodCtxS{\Gamma}{i} :\iff \denotesS{\goodCtx{\Gamma}{i}} \text{ is defined} \\
-  \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
   \denotesS{\goodType{\Gamma}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gamma}{i}} \\
   & \text{ and we define } \goodTypeS{\Gamma}{T}{i} : \iff \denotesS{\goodType{\Gamma}{T}{i}} \text{ is defined } \\
+  \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
+  \denotesS{\goodCtx{\Gamma, A}{i}} &= \{(\gamma, a) : \gamma \in \denotesS{\goodCtx{\Gamma}{}}, a \in \denotesS{\goodType{\Gamma}{A}{}}(\gamma)\}
+  \\
   \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gamma}{\gamma}{\Delta}} : \iff \denotesS{\gamma} \in \denotesS{\goodSub{\Gamma}{\_}{\Delta}} \\
@@
   \denotesS{\goodTerm{\Gamma}{\_}{T}} & = \prod_{\gamma \in \denotesS{\goodCtx{\Gamma}{i}}}\denotesS{T}(\gamma) \\
-  % 
+  \denotesS{\goodType{\Gamma}{\Sigma A B}{}} &\text{ is a family of sets, defined dependent on } \denotesS{A}, \denotesS{B} \\
+  \denotesS{\goodType{\Gamma}{\Sigma A B}{}}(\gamma) &:= \sum_{a \in \denotesS{A}\gamma} \denotesS{B}(\gamma, a) \\
   \denotesS{\goodSig{\Gamma}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gamma}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gamma}{\_}{n}} \\
@@
 
-We will take some examples on the verification of validity of this model: 1. to verify compatibility lemma ; 2. to verify the well-defined-ness of this model,
+We still encourage the reader to read the (fake) Agda version because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation.
+Here, we need to clarify some ambiguity in our conventional math formulation.
+For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on $\denotesS{\Sigma}$'', the best way to understand is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ also defined inductively on the length of the signature.
+The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
+This is actually the formulation in Agda -- QIIT syntax consider $\cC$ as a constructor and our model needs to interpret this constructor to a function.
 
@@
 
-Notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
+We will take some examples on the verification of validity of this model: 1. to verify the well-defined-ness of this model; 2. to verify compatibility lemma.
+
+Finally, notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
 and that is a contradiction.

### 2022-06-13 syntax.tex
Save.

@@
 
-Due to our intrinsic setting, \textbf{we can and will omit a lot of presumption of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
+Due to our intrinsic setting, \textbf{we can and will omit a lot of presumptions of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
 
 Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.
+
+\newcommand\mathboxtext[1]{
+  \fcolorbox{black}{faint-gray}{\ensuremath{#1}}
+}
 
@@
   \end{minipage}
-  \caption{Well-formed-ness of Four New Judgements}
+\begin{align*}
+  &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
+  &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
+\end{align*}
+\caption{Well-formed-ness of Four New Judgements, and Agda-encoding-Judgements correspondence}
 \end{figure}
+
+% \begin{minted}[texcomments]{Coq}
+%   Γ : Con  (* $\goodCtx{\Gamma}{}$ *) 
+% \end{minted}
+% T : Ty Γ (* $\goodType{\Gamma}{T}{}$ *)       σ : Tms Γ Δ {- $\goodSub{\Gamma}{\sigma}{\Delta}$ -}        t : Tm Γ T {- $\goodTerm{\Gamma}{t}{T}$ -}
 
@@
 Dependent pair type is used to model Coq module, which is useful when we need to model the ``compilation'' from linkages to modules.
+We also have $\beta,\eta$-rule for both dependent function type and dependent pair type.
 
@@
 \quad
-\Rule[name=True]
+\Rule
 {}
-{\goodTerm{\Gamma}{"tt"}{\cB}}
+{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
 \quad 
-\Rule[name=False]
-{}
-{\goodTerm{\Gamma}{"ff"}{\cB}}
-$$
-$$
 \Rule[name=Term Subst]
@@
 {\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
-\quad 
-\Rule[name=Func intro]
+$$
+$$
+\Rule[]
 {\goodTerm{\Gamma, A}{t}{B}}
@@
 \quad 
-\Rule[name=Lam Subst]
+\Rule[]
+{\goodTerm{\Gamma}{u}{A} 
+\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
+{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
+\quad 
+\Rule[]
+{\goodTerm{\Gamma}{t}{\Sigma A B}}
+{\goodTerm{\Gamma}{"pjl" \ t}{A}
+\quad \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
+}
+$$
+$$
+\Rule
 {}
-{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}}
+{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
+\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
+\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
+}
 $$
+
 $$
@@
 Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, and in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e.
-we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}, \goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption of judgements when needed.
+we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}$, $\goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption of judgements when needed.
 

### 2022-06-13 standardmodel.tex
Save.

@@
   \denotesS{\goodType{\Gamma}{\cB}{i}} \ (\gamma) &= \{0, 1\} \quad \text{a specific two element set} \\
+  \denotesS{\goodWSig{\Gamma}{\_}{n}} &\text{ is a Set of list of pairs of type interpretation of length } n \\
+  & \text{ and we define } \goodWSigS{\Gamma}{w}{n} :\iff \denotesS{\goodWSig{\Gamma}{w}{n}} \in \denotesS{\goodWSig{\Gamma}{\_}{n}}\\
+  \denotesS{\goodWSig{\Gamma}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 
+  \denotesS{\goodWSig{\Gamma}{(w^+ \ A \ B \ w)}{n+1}}
+  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gamma}{w}{n}} \\ 
+  \denotesS{\goodType{\Gamma}{\pi^j_1~\sigma}{}} &= "fst"~({\denotesS{\sigma}}_{j})\\
+  \denotesS{\goodType{\Gamma, \pi^j_1~\sigma}{\pi^j_2~\sigma}{}} &= "snd"~({\denotesS{\sigma}}_j)\\
 \end{align*}
@@
 The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
-This is actually the formulation in Agda -- QIIT syntax consider $\cC$ as a constructor and our model needs to interpret this constructor to a function.
+
+\textit{This is actually the formulation in Agda} -- QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^M : "Sig"^M \Gamma^M~n \to "Ty"^M~\Gamma^M$ with each judgement type also interpreted.
+Basically we are simply constructing a function that mapping our QIIT data-type to some other data -- $"Con"$ to $"Con"^M$, "Ty : Con → Set" to "Ty$^M$ : Con$^M$ → Set" and etc.
+What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-defined-ness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
 
@@
 
-% We omit the model here for inductive type and inheritance judgement because they require inductive facility in meta-logic. 
-We spell-out the model for the inductive type and inheritance judgement here, and both of them require inductive facility in meta-logic.
+\begin{figure}
+  \centering
+\begin{minipage}{0.8\linewidth}
+  \begin{minted}[mathescape, escapeinside=||]{Agda}
+    data MW {i}{j} (σ : |$\denotesS{\texttt{WSig i }\cdot\texttt{ S}}$|) : Set (lsuc i) where
+      MWsup : |$n$| < S → (a : |$\pi^n_1$| σ) → ((|$\pi^n_2$| σ) a → MW σ) 
+              → MW σ
+      \end{minted}
+\end{minipage}
+\caption{Modeling Inductive Type}\label{fig:model-ind-type}
+\end{figure}
+
+We sketch the model for the inductive type and inheritance judgement here, and both of them require inductive facility in meta-logic.
 In set theory, that would be using \textit{Rule sets}~\citep{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead.
-% In appendix we will use the native inductive notion from Agda to construct models. 
-As expected, we use an inductive type in the meta-logic with the constructor from W-type to model our inductive type, and the recursor for the inductive type in the meta-logic is directly modelling the recursor for our inductive type in our theory.
+Both of them are somewhat ``interpreted by themselves'' in Agda, because their syntactic definition is pretty ``well-founded''.
+For inductive type, for each grounded signature, we have an inductive type using introduction rule of W-type as constructor, shown in Figure \ref*{fig:model-ind-type}.
+This "MWsup" when applying to $n$ will use the $n$-th constructor specify in the signature.
+This grounded-ness is required to verify the substitution laws.
+For inheritance judgement, we simply construct a term model of inheritance judgement using vanilla inductive type (instead of QIIT).
+Then we use meta-level substitution to model the explicit substitution, and we can define $\denotesS{"inh"~\_ ~\_}$ by induction on this term model.
+
+% As expected, we use an inductive type in the meta-logic with the constructor from W-type to model our inductive type, and the recursor for the inductive type in the meta-logic is directly modelling the recursor for our inductive type in our theory.
 For inheritance judgement, we inductively define an almost-identical syntactic structure as model, (but not a QIIT), and we inductively define meta-level substitution to implement the explicit substitution.
@@
 
-We will take some examples on the verification of validity of this model: 1. to verify the well-defined-ness of this model; 2. to verify compatibility lemma.
+We will take some examples on the verification of validity of this model: 1.
+to verify the well-defined-ness of this model (i.e.
+judgemental equality is modeled by equality inside the model); 2.
+to verify compatibility lemma.
 

### 2022-06-14 canonicitymodel.tex
Reconsiderate phrase of "Extensible"
Save.

@@
-We follow \citep{sterling2019algebraic,coquand2018canonicity} to construct the canonicity model. We will first specify the formulation of the basic canonicity theorem.
+We follow the reducibility argument from \citep{coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
+We will first specify the formulation of the most basic canonicity theorem.
 
-We will resort to both conventional math notation and Agda syntax to explain the main idea of canonicity model.
+\begin{theorem}[Canonicity]
+  For a closed boolean term $\goodTerm{\cdot}{t}{\mathbb{B}}$, we have $\goodTerm{\cdot}{t \equiv tt}{\mathbb{B}}$ or $\goodTerm{\cdot}{t \equiv ff}{\mathbb{B}}$ hold.
+\end{theorem}
 
-% After gluing, we figure out the following "canonicity" theorem as well
-% the canonical form of extensible inductive type, linkage and boolean
+Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
+% cite https://ncatlab.org/nlab/show/canonical+form here
+We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content}, 
+% citation needed
+then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.
+
+Now we sketch out the big picture of the proof.
+The canonicity model $"Tm"^C~\Gamma^C~"T"^C$ will map each syntax piece $("t : Tm Γ T") \mapsto "(t, tₚ) :" "Tm"^C~\Gamma^C~"T"^C = \sum~("t : Tm Γ T")."Tm"_p~\Gamma^C~"T"^C~"t"$ to a dependent pair\footnote{This $"Tm"_p$ is usually called dependent model.}, of which the first part will be the same as the input syntax.
+Using this mapping on closed boolean term $"t : Tm ⋅ "\mathbb{B}$ will give us $"Tm"_p~\cdot^C~\mathbb{B}^C~"t"$, which will unfolds to our final goal $"t" \equiv "tt" + "t" \equiv "ff"$.
+Roughly, this can read as ``every closed boolean term is reducible''.
+Thus once we can ``propagate'' $(\cdot)^C$ into every syntactic piece, and construct this model, we are done.
+In our (fake) Agda formulation, we use $(\cdot)_2$ to replace the notation of $(\cdot)^C$ because we consider most parts of this model are actually a dependent pair.
+
+Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
+The latter is more conventional and the former can be understood as type theoretic encoding of the latter\footnote{We usually use predicate to encode "belong to" relation in type theory}.
+From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out --
+% cite https://arxiv.org/abs/1907.11133
+using the terminology of the conventional formulation, $\Gamma_p$ will be the set of substitution into closed reducible terms, and we will prove $t[\gamma] \in T_p$, a set of closed reducible term of type $T$ when $\gamma \in \Gamma_p$.
+
+The proof can be found in appendix.
+
+With the logical relation and the help of eta rule, we figure out the following canonical form
+\begin{theorem}[Canonical Forms].
+  \begin{itemize}
+    \item if $\goodTerm{\cdot}{t}{El~("W"~*)}$ with $\goodTerm{\cdot}{*}{\bW~\sigma}$ and $\goodWSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv "Wsup"~j~a~b}{El~("W"~*)}$ for some $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot, B[(id, a)]}{b}{El~("W"~*)}$ and $j < n$
+    \item if $\goodTerm{\cdot}{t}{\mathbb{B}}$ then $\goodTerm{\cdot}{t \equiv "tt"}{\mathbb{B}}$ or $\goodTerm{\cdot}{t \equiv "ff"}{\mathbb{B}}$ 
+    \item if $\goodTerm{\cdot}{t}{\cL\sigma}$ with $\goodSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv \mu^+~o~t}{\cL\sigma}$
+
+      $\quad$ for some $\goodTerm{\cdot}{o}{\cL(p_1\nu~\sigma)}$ and $\goodTerm{\cdot, \cC(p_1\nu'~\sigma)}{t}{p_2\nu~\sigma}$
+    \item if $\goodTerm{\cdot}{t}{\Sigma A B}$ then $\goodTerm{\cdot}{t = (a, b)}{\Sigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[(id, a)]}$
+    \footnote{We emphasize the last one because $\cCt t$ is a dependent pair}
+  \end{itemize}
+\end{theorem}
+
+% the canonical form of inductive type, linkage and boolean

### 2022-06-14 standardmodel.tex
Save.

@@
 
-We still encourage the reader to read the (fake) Agda version because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation.
-Here, we need to clarify some ambiguity in our conventional math formulation.
+We encourage the reader to read the (fake) Agda-style proof in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation.
+Here, we need to clarify some ambiguity in our conventional formulation.
 For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on $\denotesS{\Sigma}$'', the best way to understand is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ also defined inductively on the length of the signature.
 The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
+\textbf{This is actually the formulation in Agda} -- QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.
 
-\textit{This is actually the formulation in Agda} -- QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^M : "Sig"^M \Gamma^M~n \to "Ty"^M~\Gamma^M$ with each judgement type also interpreted.
-Basically we are simply constructing a function that mapping our QIIT data-type to some other data -- $"Con"$ to $"Con"^M$, "Ty : Con → Set" to "Ty$^M$ : Con$^M$ → Set" and etc.
+Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data -- it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
 What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-defined-ness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
@@
 
-We will take some examples on the verification of validity of this model: 1.
-to verify the well-defined-ness of this model (i.e.
-judgemental equality is modeled by equality inside the model); 2.
-to verify compatibility lemma.
+The verification of validity of this model is of course required but largely omitted and we refer the reader to the appendix.
+For example, to verify the well-defined-ness of this model (i.e.
+judgemental equality is modeled by equality inside the model).
+If we want to verify compatibility lemma.
 
@@
 and that is a contradiction.
+
+The full model formulated in (fake) Agda syntax can be found in the appendix.

### 2022-06-14 canonicitymodel.tex
Save.

@@
-We follow the reducibility argument from \citep{coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
+We follow the reducibility argument from \citep{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model.
 We will first specify the formulation of the most basic canonicity theorem.
@@
 Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
-% cite https://ncatlab.org/nlab/show/canonical+form here
-We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content}, 
-% citation needed
+We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content\citep{altenkirch2016type}}, 
 then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.
@@
 The latter is more conventional and the former can be understood as type theoretic encoding of the latter\footnote{We usually use predicate to encode "belong to" relation in type theory}.
-From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out --
+From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out\citep{skorstengaard2019introduction} --
 % cite https://arxiv.org/abs/1907.11133
@@
 
-      $\quad$ for some $\goodTerm{\cdot}{o}{\cL(p_1\nu~\sigma)}$ and $\goodTerm{\cdot, \cC(p_1\nu'~\sigma)}{t}{p_2\nu~\sigma}$
+      $\quad$ for some $\goodTerm{\cdot}{o}{\cL(p_1\nu~\sigma)}$ and $\goodTerm{\cdot, p_1\nu'~\sigma}{t}{p_2\nu~\sigma}$
     \item if $\goodTerm{\cdot}{t}{\Sigma A B}$ then $\goodTerm{\cdot}{t = (a, b)}{\Sigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[(id, a)]}$

### 2022-06-14 standardmodel.tex
Save.

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{kaposi2017type, kaposi2019gluing}, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
+We prove the consistency by extending the standard model from \citep{altenkirch2016type,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
 we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.
@@
 Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together -- because $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
-What's more, the substitution law around these two are also proved together by induction.
 
@@
 
-The verification of validity of this model is of course required but largely omitted and we refer the reader to the appendix.
-For example, to verify the well-defined-ness of this model (i.e.
-judgemental equality is modeled by equality inside the model).
-If we want to verify compatibility lemma.
+The verification of validity of this model is largely omitted here and we refer the reader to the appendix.
+To verify our model preserve the judgemental equality, usually unpacking the definition is enough -- for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$.
+One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together.
+The compatibility lemmas can be verified directly -- for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{p_1\mu~o}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
 
+Since all the compatibility lemmas hold (by simply changing $\vdash$ into $\models$ in the syntactic rules then we would have all the compatibility lemmas\footnote{Apparently, in our (fake) Agda formulation, we don't need this extra step.}), we have constructed a model.
 Finally, notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$

### 2022-06-14 coqimpl.tex
Save.

--- 2022-06-05 coqimpl.tex
@@
 \cL\sigma \vdash t : T $, we can compile $t$ using ``universal-quantifier-wrapped term'' : $..
-\vdash \lambda \cL \sigma.
-t : \forall \cL \sigma.
+\vdash \lambda t : \forall \cL \sigma.
 T$.
 However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping -- each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each term of the field will compile into a parametrized module (functor).
-Doing so we can also get rapid feedback from type checking when the users are defining each field.
-Overriding is achieved in the same manner -- we just need to make sure the Overriding term has the same type as the original term.
+Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
+Overriding is achieved in the same manner -- we just need to make sure the overriding term has the same type as the original term.
 
+% Insert one pseudo-code example for explanation of the mechanism
+
+% Concrete details of compilation of a family term data structure
+
+% Explain the whole procedure using this pseudo-code example
 Now that we know a family data structure is actually a list of compiled modules, we need to aggregate them into a complete compiled module.
 Thankfully, Coq's \textit{Include} command makes things easy and reasonably fast.
-% Concrete details of compilation of a family term data structure
+One key feature of Coq's \textit{Include} is that, it can include not-completely instantiated functor (i.e.
+module with non-instantiated parameters) and it will try to find appropriate fields in the current surrounding module to instantiate those parameters.
+This feature can solve the fact that each defining field requires a ``self'' parameter.
 
@@
 
-In meta-theory, we consider each family (including children family) as an inheritance judgement -- thus we directly construct the data structure representing inheritance judgement during interactive proving.
-As before, an empty family would be an inheritance judgement with empty family as input.
-Consider inheritance judgement as data gives us the ability to consider mix-in -- even though we didn't provide a formal definition here, we will illustrate its usage in the example.
-% {Only deal with inheritance judgement}
+% explaining inheritance
+To simplify the implementation, we can consider each family as an inheritance and the standalone family are extending empty family.
+Thus we directly construct the data structure representing inheritance during interactive proving.
+An inheritance will fix the type for the input family and output family, and generally has three kinds of inheritance in OO language -- extend, override and inherit.
+Extending corresponds to adding new fields, overriding corresponds to modify existent fields and inheriting corresponds to directly taking fields from the parent.
+It is possible to consider inheritance as solely a data rather than a functor, and this actually provides us the ability to \text{mix-in} inheritance data.
 
 \textbf{Reasoning Requires Non-Overridable.}
-However, unlike general-purposed programming, Coq should be able to  \textit{reason} about each program.
+Unlike general-purposed programming, Coq should be able to  \textit{reason} about each program.
 However, sometimes reasoning requires a specific program and overriding on them is unacceptable.
-For example, we may define a field $add : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to proving its commutativity in the following field $comm$.
-However, to prove so, we need to know more than just a type interface of $add$ but also its \textbf{concrete definition}.
-Otherwise, in the future, once $add$ is overridden into a subtraction function and $comm$ will have problem to be  inherited.
-To achieve such exposure of definition in meta-theory we only need to expose the information as propositional equality as part of the signature, but in our Coq plugin, we can achieve that even easier by exposing the whole definition directly into module type.
+For example, we may define a field $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to proving its commutativity in a following field $"comm"$.
+But to prove so, we need to know more than just a type interface of $"add"$ but also its \textbf{concrete definition}.
+Otherwise, in the future, once $"add"$ is overridden into a subtraction function then $"comm"$ will have problem to be  inherited.
+To achieve such exposure of definition, in meta-theory we only need to expose the information as propositional equality as part of the signature, and in our Coq plugin, we can achieve that even easier by exposing the whole definition directly into module type.
 Doing so we can have a better judgemental equality and we can avoid overriding because subtraction doesn't have the same definition as $add$.
@@
 \textbf{Special Fields: Overridable and Sealed Family as Interface.}
+But there are cases reasoning and overriding are both required, especially when we need to use Coq to do computation. 
+% Basically, {a : int -> int; property : a is monotonic}, in this case
+%   we may want to override "a" using different computation, but we still wnat
+%   this refinement/property
 
-\textbf{Special Fields: Extensible Inductive Type and Defining Recursor.}
+\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
+We don't actually support extensible inductive type in Coq (because that require non-trivial research effort on theory side\footnote{What is the semantic of extensible inductive type?} and engineering), but we simulate extensible inductive type using overriding -- when extending an inductive type with new constructors $c$ to type $A$, our plugin will generate a new inductive type data structure $A'$ with this new $c$ on the fly and feed to Coq.
+But still, we must be careful on the exposing data in the family type (module type) for the following field.
 
+% Insert one pseudo-code example for explanation of the mechanism
+
+% Explain why the module type need extra care
+
+% explain how recursor is constructed
+% explain we have the incremental-checking for any recursor
+
+% explain ftheorem wrapping this complicated recursor
 We also provide the tool of `FTheorem' to wrap the proof-irrelevant recursor.
@@
 
-\textbf{Hack: Delayed Verifying Fact.}
+% Mention still needing total recursor, for exhaustiveness checking
+
+\textbf{Hack: Global Reasoning.}
+% explain the places using ``Closing Fact''
+% 1. No need to export partial recursor
+% 2. Computational rule for the total recursor
 
@@
 
-Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the examples of MTC\citep{delaware2013,forsta2020}.
+Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the style of the examples in MTC\citep{delaware2013,forsta2020}.
 And we use a \textit{not-yet-completely-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
@@
 \subsection{Abstract Interpreters for Imp}
-The second example is adapted and modified from the Familia paper
+The second example is adapted and modified from the Familia\citep{zhang2017familia}
 % citation needed here

### 2022-06-14 metatheory.tex
Save.
Save. Working on the example
light edits

@@
 Though phrased in Latex, we are following \citep{kaposi2017type} to use as our meta-logic a type theoretical framework (like Agda) , 
-together with \textit{Quotient Inductive Inductive Type} (QIIT).
-Each judgement will actually be represented by a QIIT type, which are mutual-recursively defined together.
-Quotient is used to represent judgemental(definitional) equality, thanks to which we can have a concise representation.
-This way, for all four kinds of judgement we are using, there are always equality among these data.
-Doing so, we can omit some obvious equality judgement.
-Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).
+together with \textit{Quotient Inductive Inductive Type} (QIIT). Each
+judgement will actually be represented by a QIIT type, which are
+mutually recursively defined together. Quotient is used to represent
+judgemental(definitional) equality, thanks to which we can have a
+concise representation. This way, for all four kinds of judgement we are
+using, there are always equality among these data. Doing so, we can omit
+some obvious equality judgement. Most of the reasoning can be formulated
+using the \textit{algebra} of QIIT (mapping out function from this
+quotient data-type).
 
@@
 \newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}
-\newcommand{\goodCtx}[2]{{ {#1} \ ok }}
+\newcommand{\denotesS}[1]{{{\llbracket {#1} \rrbracket}_S}}
+\newcommand{\goodCtx}[2]{{ {#1} \ \vdash }}
 \newcommand{\goodType}[3]{{ {#1} \vdash {#2} }}
@@
 \newcommand{\cL}{{\mathcal{L}}}
+\newcommand{\cC}{{\mathcal{C}}}
+\newcommand{\cCt}{{\mathcal{C}_t}}
 \newcommand{\bW}{{\mathbb{W}}}
@@
 \subsection{Proof Relevant Logical Relation for Canonicity}
-We follow \citep{sterling2019algebraic,coquand2018canonicity} to construct the canonicity model.
+\input{canonicitymodel.tex}

### 2022-06-14 syntax.tex
Save.
light edits

@@
-Firstly, there are four standard kinds of judgement in MLTT, well-typed context $\goodCtx{\Gamma}{l}$, well-typed type $\goodType{\Gamma}{T}{l}$, well-typed term $\goodTerm{\Gamma}{t}{T}$ and well-typed substitution $\goodSub{\Gamma}{\gamma}{\Delta}$.
-We are following \citep{kaposi2019gluing} but we will omit the universe level attached to the judgement here, to make it more readable.
-After these, we will have four new kinds of judgement to model our linkage.
+Firstly, there are four standard judgement forms in MLTT,
+well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
+$\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
+well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. Our presentation
+follows the style of \citep{kaposi2019gluing}, but for readability we
+omit universe levels attached to judgements.
+After these,
+we will have four new kinds of judgement to model our linkage.
 
@@
 Please check}
-Latex-wise, we spell out these \textit{well-formed-ness rule} of type judgements and their Agda counterpart to make things more accessible.
-Of course, a well-formed judgement is not necessarily derivable, just like a well-typed QIIT type is not necessarily inhabited.
+Latex-wise, we spell out these \textit{well-formedness rule} of type
+judgements and their Agda counterpart to make things more accessible. Of
+course, a well-formed judgement is not necessarily derivable, just like
+a well-typed QIIT type is not necessarily inhabited.
 
@@
 
-We have shown case some exemplar rules for all four standard type judgments, and now we focus on the newly introduced facility.
+We have shown case some exemplar rules for all four standard type
+judgments, and now we focus on the newly introduced facility.
 

### 2022-06-14 coqimpl.tex
Save.

@@
 Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
-Overriding is achieved in the same manner -- we just need to make sure the overriding term has the same type as the original term.
+% Insert one pseudo-code example for explanation of the mechanism
+\begin{figure}
+  \begin{minipage}[t]{0.25\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family A.
+Field a : nat 
+  := 1.
+Field b 
+  : self_A.a = self_A.a
+  := eq_refl.
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \begin{minipage}[t]{0.4\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Module a_4 (self_A: EmptySig).
+Definition a : nat := 1. End a_4.
+Module Type a_5 (self_A: EmptySig).
+Parameter (a : nat). End a_5.
+Module Type a_6 := a_5 EmptySig.
+\end{minted}
+  \end{minipage}
+  \begin{minipage}[t]{0.3\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Module b_7 (self_A: a_6).
+Definition b 
+  : self_A.a = self_A.a 
+  := eq_refl. End b_7.
 
-% Insert one pseudo-code example for explanation of the mechanism
+Module A. (* Final Aggregation *)
+Include a_4. Include b_7. End A.
+\end{minted}
+  \end{minipage}
+  \caption{Example Code and Plugin Translation}\label{fig:plugin-example1}
+\end{figure}
 
 % Concrete details of compilation of a family term data structure
+Taking Figure \ref*{fig:plugin-example1} as an example: when the user interactively input the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
+{The "EmptySig" is just empty module type.} These modules and module types will be type-checked by Coq on the fly.
+The computed "Module a_10" is then part of the context of "Field b" (as the module type of "self__A") for "b" to refer to.
+The internal representation of family "A" is simply the list of these ``compiled'' modules and module types.
+We can simply achieve incremental checking and overriding -- for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
 
-% Explain the whole procedure using this pseudo-code example
-Now that we know a family data structure is actually a list of compiled modules, we need to aggregate them into a complete compiled module.
-Thankfully, Coq's \textit{Include} command makes things easy and reasonably fast.
-One key feature of Coq's \textit{Include} is that, it can include not-completely instantiated functor (i.e.
-module with non-instantiated parameters) and it will try to find appropriate fields in the current surrounding module to instantiate those parameters.
-This feature can solve the fact that each defining field requires a ``self'' parameter.
-
-% Compilation into Coq can give us computational interpretation for free.
+Now that we know a family data structure is actually a list of compiled modules, we may want to aggregate them into one complete compiled module.
+We achieve that by repeatedly using Coq's "Include" command as the example illustrated.
+Note that, in our example, we include two functors without arguments -- this is due to one of the feature of "Include": non-instantiated parameters will be automatically instantiated by the appropriate fields in the surrounding defining module.
+The resulting module is just like mundane Coq's module and thus enjoy all the computational mechanism from Coq for free.
 
 % explaining inheritance
+Besides standalone family, we also need to implement family inheritance.
 To simplify the implementation, we can consider each family as an inheritance and the standalone family are extending empty family.
 Thus we directly construct the data structure representing inheritance during interactive proving.
-An inheritance will fix the type for the input family and output family, and generally has three kinds of inheritance in OO language -- extend, override and inherit.
+
+An inheritance data will be annotated with the type for the input family and output family, and generally there are three kinds of inheritance -- extend, override and inherit.
 Extending corresponds to adding new fields, overriding corresponds to modify existent fields and inheriting corresponds to directly taking fields from the parent.
-It is possible to consider inheritance as solely a data rather than a functor, and this actually provides us the ability to \text{mix-in} inheritance data.
+Thus it is possible to consider inheritance as solely a data rather than a functor, and this actually provides us the ability to \text{mix-in} inheritance data.
 
@@
 Otherwise, in the future, once $"add"$ is overridden into a subtraction function then $"comm"$ will have problem to be  inherited.
+A more concrete example is in Figure \ref*{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more.
+
 To achieve such exposure of definition, in meta-theory we only need to expose the information as propositional equality as part of the signature, and in our Coq plugin, we can achieve that even easier by exposing the whole definition directly into module type.

### 2022-06-14 standardmodel.tex
Save.

@@
   & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} \ (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
   \denotesS{\goodSig{\Gamma}{\nu\cdot}{0}} &= \star \\
   \denotesS{\goodSig{\Gamma}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, A, f, T) \\
-  \denotesS{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot} } \ (\gamma) &= \star \\
+  \denotesS{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot} } (\gamma) &= \star \\
   \denotesS{\goodTerm{\Gamma}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\gamma)&= (o(\gamma), (\lambda \ m' . t'(\gamma, m'))) \\
+  \denotesS{\goodTerm{\Gamma}{p_1\mu~t}{\cL(p_1\nu~\sigma)}}(\gamma) &=  "fst"~(t\gamma) \\ 
   \denotesS{\goodTerm{\Gamma}{\cCt \{\sigma\} o}{\cC\sigma}} & \text{ is a dependent function, defined inductively on } \denotesS{\sigma} \\
@@
 The verification of validity of this model is largely omitted here and we refer the reader to the appendix.
-To verify our model preserve the judgemental equality, usually unpacking the definition is enough -- for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$.
+To verify our model preserve the judgemental equality, usually unpacking the definition is enough -- for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$ by function extensionality.
 One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together.

### 2022-06-15 coqimpl.tex
Save.

@@
 T$.
-However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping -- each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each term of the field will compile into a parametrized module (functor).
+However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping -- each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a parametrized module (functor).
 Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
@@
 Parameter (a : nat). End a_5.
-Module Type a_6 := a_5 EmptySig.
+Module Type a_6 := a_5 EmptyMod.
 \end{minted}
@@
 Taking Figure \ref*{fig:plugin-example1} as an example: when the user interactively input the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
-{The "EmptySig" is just empty module type.} These modules and module types will be type-checked by Coq on the fly.
+{The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately.
 The computed "Module a_10" is then part of the context of "Field b" (as the module type of "self__A") for "b" to refer to.
@@
 Besides standalone family, we also need to implement family inheritance.
-To simplify the implementation, we can consider each family as an inheritance and the standalone family are extending empty family.
-Thus we directly construct the data structure representing inheritance during interactive proving.
+To simplify the implementation, we can consider all families as inheritance and the standalone families are extending empty family.
+Thus we only need to deal with inheritance during interactive proving.
 
@@
 A more concrete example is in Figure \ref*{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more.
+To fix that, we need more information on "a" than its type -- we need to expose its concrete definition "a = 1" to appear in the context of "b".
 
@@
 But there are cases reasoning and overriding are both required, especially when we need to use Coq to do computation.
+For example, we may want a field to be an arbitrary monotonic function.
+Again, we cannot split into two fields \mintinline{Coq}{(f : nat → nat, p : monotonic f)} because to provide "p" we need to have concrete definition of "f" in the family type, which makes "f" non-overridable at all.
+One option is to use sigma type \mintinline{Coq}{{f | monotonic f}} but that is cumbersome when we have a bundle of stuff with multiple refinements.
+
+To resolve this, we provide family sealing mechanism -- basically we allow the users to specify the family/module type, and thus we can have a ``local'' transparent, non-overridable field located in a overridable family.
+Using the above example, we simply make \mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an overridable family, but "f" to be non-overridable and transparent on its definition, and thus we can prove "p".
+Once we want to swap the implementation, we just override the whole ``sealed'' family.
+
+This ``sealed'' family actually simulate interface to a certain level -- we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}  , and later instantiate them with concrete implementation by overriding this family.
 % Basically, {a : int -> int; property : a is monotonic}, in this case
@@
 \textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
-We don't actually support extensible inductive type in Coq (because that require non-trivial research effort on theory side\footnote{What is the semantic of extensible inductive type?} and engineering), but we simulate extensible inductive type using overriding -- when extending an inductive type with new constructors $c$ to type $A$, our plugin will generate a new inductive type data structure $A'$ with this new $c$ on the fly and feed to Coq.
+We don't actually support extensible inductive type in Coq (because that require non-trivial research effort on theory side -- i.e.
+{What is the semantic of extensible inductive type?} -- and engineering), but we simulate extensible inductive type using overriding -- when extending an inductive type $A$ with new constructors $c$, our plugin will generate a new inductive definition $A'$ with this new $c$ on the fly and feed to Coq.
 But still, we must be careful on the exposing data in the family type (module type) for the following field.
@@
 % Insert one pseudo-code example for explanation of the mechanism
+% Use the natural number example
+\begin{figure}
+  \begin{minipage}[t]{0.3\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family B.
+  FInductive b : Set 
+    := tt : b | ff : b.
+  Family neg_handler. 
+    Field tt : self_B.b 
+      := self_B.ff.
+    Field ff : self_B.b 
+      := self_B.tt.
+  EndFamily.
+FRecursor neg 
+  about self_B.b 
+  motive (fun _ => self_B.b)
+  using self_B.neg_handler
+  by _rec.
+EndFamily.
+
+Family B2 extends B.
+  Extend FInductive b : Set 
+    := uu : b.
+  Extend Family neg_handler. 
+    Field uu : self_B2.b 
+      := self_B2.uu.
+  EndFamily. 
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \begin{minipage}[t]{0.65\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+  (* Abstraction for Inductive Type *)
+Module Type b_3 (self__B: EmptySig_4).
+Parameter (b : Set). Parameter (tt ff : b). End b_3.      
+Module Type b_6 := b_3 EmptyMod.
+  (* Recursor to Set for b *)
+Module b_rec_12 (self__B: b_6).
+Definition __recursor_type_b_rec :=
+  forall P : self__B.b -> Set,
+  P self__B.tt -> P self__B.ff 
+  -> forall __i : self__B.b, P __i. End b_rec_12. 
+  (* Field B.neg_handler.ff *)
+Module ff_24 (self__B: b_6) 
+             (self__neg_handler: tt_23 self__B).
+Definition ff : self__B.b := self__B.tt. End ff_24.
+  (* Intermediate Module solely for type checking *)
+Module v_33_34 (self__B: neg_handler_26).
+Include b_rec_12 self__B.
+Parameter (recursor_for_type_checking : __recursor_type_b_rec).
+Definition term_for_type_checking :=
+  recursor_for_type_checking (fun _ : self__B.b => self__B.b.)
+    self__B.neg_handler.tt
+    self__B.neg_handler.ff.  
+End v_33_34.
+\end{minted}
+  \end{minipage}
+\caption{Example Code for Inductive Type}\label{fig:plugin-example2}
+\end{figure}
+
+Taking \cref{fig:plugin-example2} as an example, when we define inductive type "W", the following "Field b" can only know "W" is a type (of "Set") and there are at least two constructors for it, just like what we do in module type "W_3".
+We cannot export the eliminator of "W" inside "W_3" because doing so "b" (and "b_19") will not be able to be inherited to the context where "W" is extended with a third constructor.
+Since the inhabited recursor is generally asserting there are only two constructors in "W".
+
+But we do have to provide a way to construct a recursor, and do type-checking (or exhaustiveness checking).
+Thus we need to extract a elimination principle for future usage.
+Here we show one exemplar result when extracting the inductive principle.
+
+To construct a recursor, we make each handler
+
+The key insight here is that, to make each recursor handler inheritable, we need to seal an abstraction around the inductive type; but to construct a recursor (and carry out the exhaustiveness checking), we need to break this abstraction and see the concrete definition of the inductive type.
+Our meta-theory need to handler these two seemingly contradicting ideas simultaneously.
+
+Since the recursor is constructed knowing a concrete inductive type, thus not inheritable, for all the children, we have to reconstruct and override with a new recursor (acting like an extra exhaustiveness checking).
+This is affordable since each recursor handler doesn't need to be rechecked.
 
@@
 % explain ftheorem wrapping this complicated recursor
-We also provide the tool of `FTheorem' to wrap the proof-irrelevant recursor.
+We also provide the tool of `FTheorem' to wrap the proof-irrelevant recursor (i.e.
+when reasoning stuff in "Prop").
 This tool can avoid most boiler-plate code when writing a standard recursor, and is also open to extension like `FRecursor'.

### 2022-06-15 coqimpl.tex
Save.

@@
   (* Abstraction for Inductive Type *)
-Module Type b_3 (self__B: EmptySig_4).
+Module Type b_3 (self_B: EmptySig_4).
 Parameter (b : Set). Parameter (tt ff : b). End b_3.      
@@
   (* Recursor to Set for b *)
-Module b_rec_12 (self__B: b_6).
+Module b_rec_12 (self_B: b_6).
 Definition __recursor_type_b_rec :=
-  forall P : self__B.b -> Set,
-  P self__B.tt -> P self__B.ff 
-  -> forall __i : self__B.b, P __i. End b_rec_12. 
+  forall P : self_B.b -> Set,
+  P self_B.tt -> P self_B.ff 
+  -> forall __i : self_B.b, P __i. End b_rec_12. 
   (* Field B.neg_handler.ff *)
-Module ff_24 (self__B: b_6) 
-             (self__neg_handler: tt_23 self__B).
-Definition ff : self__B.b := self__B.tt. End ff_24.
+Module ff_24 (self_B: b_6) 
+             (self__neg_handler: tt_23 self_B).
+Definition ff : self_B.b := self_B.tt. End ff_24.
   (* Intermediate Module solely for type checking *)
-Module v_33_34 (self__B: neg_handler_26).
-Include b_rec_12 self__B.
+Module v_33_34 (self_B: neg_handler_26).
+Include b_rec_12 self_B.
 Parameter (recursor_for_type_checking : __recursor_type_b_rec).
 Definition term_for_type_checking :=
-  recursor_for_type_checking (fun _ : self__B.b => self__B.b.)
-    self__B.neg_handler.tt
-    self__B.neg_handler.ff.  
+  recursor_for_type_checking (fun _ : self_B.b => self_B.b)
+    self_B.neg_handler.tt
+    self_B.neg_handler.ff.  
 End v_33_34.
@@
 
-Taking \cref{fig:plugin-example2} as an example, when we define inductive type "W", the following "Field b" can only know "W" is a type (of "Set") and there are at least two constructors for it, just like what we do in module type "W_3".
-We cannot export the eliminator of "W" inside "W_3" because doing so "b" (and "b_19") will not be able to be inherited to the context where "W" is extended with a third constructor.
-Since the inhabited recursor is generally asserting there are only two constructors in "W".
+Let's look at \cref{fig:plugin-example2}, the example of boolean and negation.
+When we define inductive type "b", the following "Field neg_handler.ff" can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface'' module type "b_3".
+We cannot export the eliminator of "b" inside "b_3" because doing so "neg_handler.ff" (and "ff_24") will not be able to be inherited to the context where "b" is extended with a third constructor, since the existence of recursor alone will assert that there are only two constructors in "b", not ``extensible'' at all!
 
-But we do have to provide a way to construct a recursor, and do type-checking (or exhaustiveness checking).
+But we do have to provide a way to construct a recursor, and execute type checking (or exhaustiveness checking).
 Thus we need to extract a elimination principle for future usage.
-Here we show one exemplar result when extracting the inductive principle.
+Here we show one exemplar extraction of the recursor in \mintinline{Coq}{Set} as module "b_rec_12".
 
-To construct a recursor, we make each handler
-
-The key insight here is that, to make each recursor handler inheritable, we need to seal an abstraction around the inductive type; but to construct a recursor (and carry out the exhaustiveness checking), we need to break this abstraction and see the concrete definition of the inductive type.
-Our meta-theory need to handler these two seemingly contradicting ideas simultaneously.
+To construct a recursor, we make for each constructor a handler and aggregate them inside a family.
+The whole family "neg_handler" can be considered as a large matching case.
+Doing so, family inheritance can function as matching-case reuse.
+Then we use "FRecursor" to specify the motive and inductive type.
+Here "FRecursor" will feed Coq
+a module "v_33_34" for doing type-checking -- what it does is basically assert the existence of the recursor and apply to see if handlers can pass the type checking by Coq.
 
 Since the recursor is constructed knowing a concrete inductive type, thus not inheritable, for all the children, we have to reconstruct and override with a new recursor (acting like an extra exhaustiveness checking).
-This is affordable since each recursor handler doesn't need to be rechecked.
+This is affordable since each recursor handler doesn't need to be rechecked -- look at "B2.neg_handler", there is only one new handler for "uu" and the handlers from the parent "B.neg_handler.tt,B.neg_handler.ff" are already type-checked and inherited(thanks to the fact that they only rely on ``the abstracted interface'' module "b_3").
+Of course, our plugin will ``inherits'' the recursor "neg" but what the plugin does is doing exhaustiveness checking, again.
 
@@
 % explain ftheorem wrapping this complicated recursor
-We also provide the tool of `FTheorem' to wrap the proof-irrelevant recursor (i.e.
-when reasoning stuff in "Prop").
-This tool can avoid most boiler-plate code when writing a standard recursor, and is also open to extension like `FRecursor'.
+We also provide a command "FTheorem" when the users want to use tactic programming instead.
+This tool can avoid most boiler-plate code (e.g.
+specify the handler family) when writing a recursor like above, invoke proof interaction mode, allow tactic usage, and is also open to extension like "FRecursor".
+
+To summarize, the key insight here is that, to make each recursor handler inheritable, we need to seal an abstraction around the inductive type (e.g.
+module "b_3"); but to construct a recursor (and carry out the exhaustiveness checking), we need to break this abstraction and see the concrete definition of the inductive type (e.g.
+module "b_rec_12").
+Our meta-theory need to handler these two seemingly contradicting ideas simultaneously.
 
 \textbf{Propositional Partial Recursor.}
+Apparently, the exposed module "b_3" in \cref*{fig:plugin-example2} is definitely not expressive enough.
+For example, we cannot prove the constructors are injective, i.e.
+\mintinline{Coq}{self_B.tt <> self_B.ff}.
+
 Propositional Partial Recursor can prove injection and discrimination of the constructors.

### 2022-06-15 coqimpl.tex
Save.

@@
 \textbf{Propositional Partial Recursor.}
-Apparently, the exposed module "b_3" in \cref*{fig:plugin-example2} is definitely not expressive enough.
+Apparently, the exposed module "b_3" in \cref*{fig:plugin-example2} is not expressive enough.
 For example, we cannot prove the constructors are injective, i.e.
-\mintinline{Coq}{self_B.tt <> self_B.ff}.
+\mintinline{Coq}{self_B.tt ≠ self_B.ff}.
+In the original setting, to prove injectiveness of constructor we only require a recursor and its computational rules.
+However, as we pointed out earlier, the recursor of an inductive type cannot be exposed into the context for the sake of extensibility.
+We may have to alter the formulation of the recursor.
 
-Propositional Partial Recursor can prove injection and discrimination of the constructors.
-Propositional Partial Recursor is good enough because it and its computational axiom can prove that the vanilla inductive type can "embed" into the extensible inductive type. 
-(i.e. there will be an left inversion of that injection, which is also witnessing the fact that those "constructors" can really act as constructors) 
-Thus 1.
-every future extension can support this partial recursion; 2.
-every type support this partial recursor with its computational axiom at least "support these constructors" (because of the embedding).
+The simplest idea is to allow partiality -- i.e.
+returning \mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
+We call this \textit{propositional partial recursor}.
+For example, the propositional partial recursor for \mintinline{Coq}{B.b} is
+\begin{minted}{Coq}
+b_prec : forall P : b -> Set, P tt -> P ff 
+    -> forall x : b, option (P x). 
+\end{minted}
+with appropriate computational axioms. The partiality doesn't break extensibility (i.e. all the future extension of "b" can support this "b_prec").
+
+What's more, partial recursor is theoretically enough for the "injection" and "discriminate" tactics.
+The rough idea to achieve this is to use "b_prec" to reflect our special inductive type into vanilla inductive type, and use the orignal "injection" and "discrimination" respectively.
+For example, we want to prove \mintinline{Coq}{self_B.tt ≠ self_B.ff}, we simply
+\begin{minted}{Coq}
+Definition reflect : self_B.b → option bool 
+                   := b_prec (fun _ => bool) true false.
+Definition tt_neq_ff : self_B.tt = self_B.ff → False.
+(* Because self_B.tt = self_B.ff → reflect self_B.tt = reflect self_B.ff 
+                → true = false    (by computational axiom)
+                → False           (by discriminate of bool) *)
+\end{minted}
+
+In fact, the above "reflect" function implies that propositional partial recursor is good enough to some extent, at least for non-indexed inductive type, because we can ``embed'' vanilla inductive type into types supporting partial recursor.
+Because "reflect" is actually a left inverse of an injection "bool → self_B.b".
+A bit more formally, and restrict our focus to non-indexed inductve types,
+
+\begin{theorem}\label{thm:prec-complete} Given a list of $n$ pairs of types $\{ x : "A"_i \vdash "B"_i(x) \}_{i}$
+  \begin{itemize}
+    \item Define $"C"$ as the inductive type using this list, i.e. 
+    \begin{minted}{Coq}
+Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
+    \end{minted}
+    \item Assume an arbitrary type \mintinline{Coq}{D : Set} that has ...
+    \begin{itemize}
+      \item following $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
+      \item a partial recursor 
+      \begin{minted}{Coq}
+prec : forall (R : Set), ..,  
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> R) -> R)), .., D -> R
+      \end{minted}
+      \item all the following $n$ (propositional) computational axioms: for all $i$, 
+      \begin{minted}{Coq}
+prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
+    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
+      \end{minted}
+    \end{itemize}
+  \end{itemize}
+  We will have an \textbf{embedding} from "C" to "D", more concretely, we can have \mintinline{Coq}{inj : C -> D} and
+
+  \mintinline{Coq}{linv : D -> option C} defined using eliminator and "prec", (both act similar to identity) such that
+  \begin{minted}{Coq}
+    forall c : C, linv (inj c) = some c
+  \end{minted}
+\end{theorem}
+
+We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can actually be considered as a constructor, because they can be reflected to real constructors of "C" using "linv".
+Thus, \cref{thm:prec-complete} implies that (1) every future extension of the inductive type can support this partial recursion (trivially); (2) every type support this partial recursor with its computational axiom at least support these constructors because of the embedding.
+In other words, \textit{a type (at least) supports these constructors if and only if this type supports the corresponding partial recursor (with the computational axiom)}.
+Thus we can argue that the partial recursor gives a sound and complete characterization of all the extension of a given (non-indexed) inductive type.
 

### 2022-06-15 metatheory.tex

@@
 
-Though phrased in Latex, we are following \citep{kaposi2017type} to use as our meta-logic a type theoretical framework (like Agda) , 
+Though phrased in Latex, we are following \citet{altkap2016} to use as
+our meta-logic a type theoretical framework (like Agda), 
 together with \textit{Quotient Inductive Inductive Type} (QIIT). Each

### 2022-06-15 syntax.tex

@@
 \end{align*}
-\caption{Well-formed-ness of Four New Judgements, and Agda-encoding-Judgements correspondence}
+\caption{Well-formedness of Four New Judgements, and Agda-encoding-Judgements correspondence}
 \end{figure}
@@
 
-Now we show case some of the typing rules, we will have dependent function type and dependent pair type.
-Dependent pair type is used to model Coq module, which is useful when we need to model the ``compilation'' from linkages to modules.
-We also have $\beta,\eta$-rule for both dependent function type and dependent pair type.
+Now we show selected typing rules: we will have dependent function types
+and dependent pair types. Dependent pair type is used to model Coq
+module, which is useful when we need to model the ``compilation'' from
+linkages to modules. We also have $\beta,\eta$-rule for both dependent
+function type and dependent pair type.
 
@@
   \quad \goodTerm{\Gamma}{a}{\pi^j_1\sigma}
-  \quad \goodTerm{\Gamma, \pi^j_1\sigma}{b}{\pi^j_2\sigma}}
-{\goodTerm{\Gamma}{"Wsup" \ T \ a \ b}{El\ ("W" \ T)} }
+  \quad \goodTerm{\Gamma, \pi^j_2\sigma[(id, a)]}{b}{El~("W"~T)}}
+{\goodTerm{\Gamma}{"Wsup" \ T \ a \ b}{El~("W"~T)} }
 $$
@@
 
-In our meta-theory, we only formulate non-dependent eliminator, i.e.
-the \textit{recursor}.
-Each recursor is constructed by a linkage contained with functions/handlers/pattern-matching cases dealing with each constructor.
-Thus code-reuse for recursor is delegated by the inheritance of linkage -- to reuse one particular pattern-matching case, we just inherit the function/handler.
-This greatly simplifies the meta\-theory development.
+In our meta-theory, we only formulate non-dependent eliminator, i.e. the
+\textit{recursor}. Each recursor is constructed by a linkage containing
+functions/handlers/pattern-matching cases dealing with each
+constructor. Thus code reuse for recursor is delegated to linkage
+inheritance---to reuse one particular pattern-matching case,
+we just inherit the function/handler. This greatly simplifies the
+meta\-theory development.
 
@@
 
-As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just a short hand of term judgement $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$.
-The substitution law should understand as a lemma, and easily derivable by using $\gamma^\uparrow$.
+As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
+a short hand of term judgement $\goodTerm{\Gamma , \cC
+\sigma}{\_}{A[\pi_1]}$. The substitution law should be understood as a
+lemma, and it is easily derivable using $\gamma^\uparrow$.
 
@@
 
-With the definition of sealing judgement in mind, we can see \ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the implementation.
-In the implementation, we will have "self__" all around when defining new fields in our plugin, and this "self__" is exactly the $A$ in the context of the two fields -- especially when we don't do any sealing, this $A = \cC \sigma$, which is exactly how the implementation expose "self__" as a module parameter.
+With the definition of sealing judgement in mind, we can see
+\ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the
+implementation. In the implementation, we will have "self__" all around
+when defining new fields in our plugin, and this "self__" is exactly the
+$A$ in the context of the two fields -- especially when we don't do any
+sealing, this $A = \cC \sigma$, which is exactly how the implementation
+exposes "self__" as a module parameter.
 

### 2022-06-15 coqimpl.tex
Save.

@@
 % 2. Computational rule for the total recursor
+However, to verify and prove the recursor and partial recursor (and their computational rules), our plugin has to compile the extensible family into a non-extensible Coq Module and prove the recursors by instantiating the content of Coq module with concrete inductive type.
+
+This style of ``Global Reasoning'' is that we can only prove the corresponding theorem when we know the ``full pictures'' of the inductive type.
+We provide "Closing Fact" as a command to carry out this global reasoning -- we can attach "Closing Fact" with either (proof) terms or Coq's proof script, then only when compilation to module happens, the term/proof script will put into type-check to see if a "Closing Fact" statement can be proved.
+Note that, "Closing Fact" cannot immeidately check the correctness of the term/proof script because this verification only happens when we finished defining a family, and thus not easy to handle.
+
+This "Closing Fact" can ease us from the necessity to let plugin generate proof for recursors and partial recursors, and thus making plugin development easier.
+For example, the computation rule for a concrete recursor like "B.neg" and partial recursors in \cref*{fig:plugin-example2} can be directly asserted by using "Closing Fact" and verified later during compilation.
+This ``Global Reasoning'' will be again useful in the later examples.
 

### 2022-06-15 coq-example.tex
Save. Working on example section

--- 2022-06-15 coqimpl.tex
@@
-After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
-
-Despite of the possible difficulty of the future maintenance, this approach has various advantages: 1.
-it is the easiest and the most accessible way to prototype as it relieves us the necessity of familarity of Coq base, especially for the implementation of the module and functor; 2.
-we have a clear definition of trusted-base -- the whole Coq; 3.
-it is easy to debug -- we just need to check the translated commands; 4.
-it can be well-incorporated with the existent tools like VSCoq; 5.
-it is more accessible for the the interested audience who can then easily adapt our plugin and give a try -- otherwise they have to download and re-build the whole customized Coq; 6.
-it is more stable because the surface syntax of Coq should be stable across different versions; 7.
-still, this plugin can capture the key ingredients of implementing family polymorphism inside Coq and act as a reference for guiding an appropriate implementation of Family Polymorphism inside all sorts of proof assistants.
-
-\textbf{Family compiled into module.} This is the main idea of compiling family components into Coq's primitives: we compile family into coq module, family type into module type and context of a given term into parameters of the module.
-To achieve incremental type checking, we notice the nature of late binding: for each defining field $ ..
-\cL\sigma \vdash t : T $, we can compile $t$ using ``universal-quantifier-wrapped term'' : $..
-\vdash \lambda t : \forall \cL \sigma.
-T$.
-However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping -- each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a parametrized module (functor).
-Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
-% Insert one pseudo-code example for explanation of the mechanism
-\begin{figure}
-  \begin{minipage}[t]{0.25\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Family A.
-Field a : nat 
-  := 1.
-Field b 
-  : self_A.a = self_A.a
-  := eq_refl.
-EndFamily.
-\end{minted}
-  \end{minipage}
-  \begin{minipage}[t]{0.4\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Module a_4 (self_A: EmptySig).
-Definition a : nat := 1. End a_4.
-Module Type a_5 (self_A: EmptySig).
-Parameter (a : nat). End a_5.
-Module Type a_6 := a_5 EmptyMod.
-\end{minted}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Module b_7 (self_A: a_6).
-Definition b 
-  : self_A.a = self_A.a 
-  := eq_refl. End b_7.
-
-Module A. (* Final Aggregation *)
-Include a_4. Include b_7. End A.
-\end{minted}
-  \end{minipage}
-  \caption{Example Code and Plugin Translation}\label{fig:plugin-example1}
-\end{figure}
-
-% Concrete details of compilation of a family term data structure
-Taking Figure \ref*{fig:plugin-example1} as an example: when the user interactively input the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
-{The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately.
-The computed "Module a_10" is then part of the context of "Field b" (as the module type of "self__A") for "b" to refer to.
-The internal representation of family "A" is simply the list of these ``compiled'' modules and module types.
-We can simply achieve incremental checking and overriding -- for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
-
-Now that we know a family data structure is actually a list of compiled modules, we may want to aggregate them into one complete compiled module.
-We achieve that by repeatedly using Coq's "Include" command as the example illustrated.
-Note that, in our example, we include two functors without arguments -- this is due to one of the feature of "Include": non-instantiated parameters will be automatically instantiated by the appropriate fields in the surrounding defining module.
-The resulting module is just like mundane Coq's module and thus enjoy all the computational mechanism from Coq for free.
-
-% explaining inheritance
-Besides standalone family, we also need to implement family inheritance.
-To simplify the implementation, we can consider all families as inheritance and the standalone families are extending empty family.
-Thus we only need to deal with inheritance during interactive proving.
-
-An inheritance data will be annotated with the type for the input family and output family, and generally there are three kinds of inheritance -- extend, override and inherit.
-Extending corresponds to adding new fields, overriding corresponds to modify existent fields and inheriting corresponds to directly taking fields from the parent.
-Thus it is possible to consider inheritance as solely a data rather than a functor, and this actually provides us the ability to \text{mix-in} inheritance data.
-
-\textbf{Reasoning Requires Non-Overridable.}
-Unlike general-purposed programming, Coq should be able to  \textit{reason} about each program.
-However, sometimes reasoning requires a specific program and overriding on them is unacceptable.
-For example, we may define a field $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to proving its commutativity in a following field $"comm"$.
-But to prove so, we need to know more than just a type interface of $"add"$ but also its \textbf{concrete definition}.
-Otherwise, in the future, once $"add"$ is overridden into a subtraction function then $"comm"$ will have problem to be  inherited.
-A more concrete example is in Figure \ref*{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more.
-To fix that, we need more information on "a" than its type -- we need to expose its concrete definition "a = 1" to appear in the context of "b".
-
-To achieve such exposure of definition, in meta-theory we only need to expose the information as propositional equality as part of the signature, and in our Coq plugin, we can achieve that even easier by exposing the whole definition directly into module type.
-Doing so we can have a better judgemental equality and we can avoid overriding because subtraction doesn't have the same definition as $add$.
-We require the users to decorate field as `Final' so that the plugin will proceed with this special treatment.
-
-\textbf{Special Fields: Overridable and Sealed Family as Interface.}
-But there are cases reasoning and overriding are both required, especially when we need to use Coq to do computation.
-For example, we may want a field to be an arbitrary monotonic function.
-Again, we cannot split into two fields \mintinline{Coq}{(f : nat → nat, p : monotonic f)} because to provide "p" we need to have concrete definition of "f" in the family type, which makes "f" non-overridable at all.
-One option is to use sigma type \mintinline{Coq}{{f | monotonic f}} but that is cumbersome when we have a bundle of stuff with multiple refinements.
-
-To resolve this, we provide family sealing mechanism -- basically we allow the users to specify the family/module type, and thus we can have a ``local'' transparent, non-overridable field located in a overridable family.
-Using the above example, we simply make \mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an overridable family, but "f" to be non-overridable and transparent on its definition, and thus we can prove "p".
-Once we want to swap the implementation, we just override the whole ``sealed'' family.
-
-This ``sealed'' family actually simulate interface to a certain level -- we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}  , and later instantiate them with concrete implementation by overriding this family.
-% Basically, {a : int -> int; property : a is monotonic}, in this case
-%   we may want to override "a" using different computation, but we still wnat
-%   this refinement/property
-
-\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
-We don't actually support extensible inductive type in Coq (because that require non-trivial research effort on theory side -- i.e.
-{What is the semantic of extensible inductive type?} -- and engineering), but we simulate extensible inductive type using overriding -- when extending an inductive type $A$ with new constructors $c$, our plugin will generate a new inductive definition $A'$ with this new $c$ on the fly and feed to Coq.
-But still, we must be careful on the exposing data in the family type (module type) for the following field.
-
-% Insert one pseudo-code example for explanation of the mechanism
-% Use the natural number example
-\begin{figure}
-  \begin{minipage}[t]{0.3\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Family B.
-  FInductive b : Set 
-    := tt : b | ff : b.
-  Family neg_handler. 
-    Field tt : self_B.b 
-      := self_B.ff.
-    Field ff : self_B.b 
-      := self_B.tt.
-  EndFamily.
-FRecursor neg 
-  about self_B.b 
-  motive (fun _ => self_B.b)
-  using self_B.neg_handler
-  by _rec.
-EndFamily.
-
-Family B2 extends B.
-  Extend FInductive b : Set 
-    := uu : b.
-  Extend Family neg_handler. 
-    Field uu : self_B2.b 
-      := self_B2.uu.
-  EndFamily. 
-EndFamily.
-\end{minted}
-  \end{minipage}
-  \begin{minipage}[t]{0.65\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-  (* Abstraction for Inductive Type *)
-Module Type b_3 (self_B: EmptySig_4).
-Parameter (b : Set). Parameter (tt ff : b). End b_3.      
-Module Type b_6 := b_3 EmptyMod.
-  (* Recursor to Set for b *)
-Module b_rec_12 (self_B: b_6).
-Definition __recursor_type_b_rec :=
-  forall P : self_B.b -> Set,
-  P self_B.tt -> P self_B.ff 
-  -> forall __i : self_B.b, P __i. End b_rec_12. 
-  (* Field B.neg_handler.ff *)
-Module ff_24 (self_B: b_6) 
-             (self__neg_handler: tt_23 self_B).
-Definition ff : self_B.b := self_B.tt. End ff_24.
-  (* Intermediate Module solely for type checking *)
-Module v_33_34 (self_B: neg_handler_26).
-Include b_rec_12 self_B.
-Parameter (recursor_for_type_checking : __recursor_type_b_rec).
-Definition term_for_type_checking :=
-  recursor_for_type_checking (fun _ : self_B.b => self_B.b)
-    self_B.neg_handler.tt
-    self_B.neg_handler.ff.  
-End v_33_34.
-\end{minted}
-  \end{minipage}
-\caption{Example Code for Inductive Type}\label{fig:plugin-example2}
-\end{figure}
-
-Let's look at \cref{fig:plugin-example2}, the example of boolean and negation.
-When we define inductive type "b", the following "Field neg_handler.ff" can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface'' module type "b_3".
-We cannot export the eliminator of "b" inside "b_3" because doing so "neg_handler.ff" (and "ff_24") will not be able to be inherited to the context where "b" is extended with a third constructor, since the existence of recursor alone will assert that there are only two constructors in "b", not ``extensible'' at all!
-
-But we do have to provide a way to construct a recursor, and execute type checking (or exhaustiveness checking).
-Thus we need to extract a elimination principle for future usage.
-Here we show one exemplar extraction of the recursor in \mintinline{Coq}{Set} as module "b_rec_12".
-
-To construct a recursor, we make for each constructor a handler and aggregate them inside a family.
-The whole family "neg_handler" can be considered as a large matching case.
-Doing so, family inheritance can function as matching-case reuse.
-Then we use "FRecursor" to specify the motive and inductive type.
-Here "FRecursor" will feed Coq
-a module "v_33_34" for doing type-checking -- what it does is basically assert the existence of the recursor and apply to see if handlers can pass the type checking by Coq.
-
-Since the recursor is constructed knowing a concrete inductive type, thus not inheritable, for all the children, we have to reconstruct and override with a new recursor (acting like an extra exhaustiveness checking).
-This is affordable since each recursor handler doesn't need to be rechecked -- look at "B2.neg_handler", there is only one new handler for "uu" and the handlers from the parent "B.neg_handler.tt,B.neg_handler.ff" are already type-checked and inherited(thanks to the fact that they only rely on ``the abstracted interface'' module "b_3").
-Of course, our plugin will ``inherits'' the recursor "neg" but what the plugin does is doing exhaustiveness checking, again.
-
-% Explain why the module type need extra care
-
-% explain how recursor is constructed
-% explain we have the incremental-checking for any recursor
-
-% explain ftheorem wrapping this complicated recursor
-We also provide a command "FTheorem" when the users want to use tactic programming instead.
-This tool can avoid most boiler-plate code (e.g.
-specify the handler family) when writing a recursor like above, invoke proof interaction mode, allow tactic usage, and is also open to extension like "FRecursor".
-
-To summarize, the key insight here is that, to make each recursor handler inheritable, we need to seal an abstraction around the inductive type (e.g.
-module "b_3"); but to construct a recursor (and carry out the exhaustiveness checking), we need to break this abstraction and see the concrete definition of the inductive type (e.g.
-module "b_rec_12").
-Our meta-theory need to handler these two seemingly contradicting ideas simultaneously.
-
-\textbf{Propositional Partial Recursor.}
-Apparently, the exposed module "b_3" in \cref*{fig:plugin-example2} is definitely not expressive enough.
-For example, we cannot prove the constructors are injective, i.e.
-\mintinline{Coq}{self_B.tt <> self_B.ff}.
-
-Propositional Partial Recursor can prove injection and discrimination of the constructors.
-Propositional Partial Recursor is good enough because it and its computational axiom can prove that the vanilla inductive type can "embed" into the extensible inductive type. 
-(i.e. there will be an left inversion of that injection, which is also witnessing the fact that those "constructors" can really act as constructors) 
-Thus 1.
-every future extension can support this partial recursion; 2.
-every type support this partial recursor with its computational axiom at least "support these constructors" (because of the embedding).
-
-% Mention still needing total recursor, for exhaustiveness checking
-
-\textbf{Hack: Global Reasoning.}
-% explain the places using ``Closing Fact''
-% 1. No need to export partial recursor
-% 2. Computational rule for the total recursor
-
-\section{Example}
 We provide two examples here for illustration and evaluation.

### 2022-06-15 coqimpl.tex
Save. Working on example section

@@
 
-This style of ``Global Reasoning'' is that we can only prove the corresponding theorem when we know the ``full pictures'' of the inductive type.
+This style of ``Global Reasoning'' is that we can only prove the corresponding theorem when we know the ``full pictures'', including the inductive type.
 We provide "Closing Fact" as a command to carry out this global reasoning -- we can attach "Closing Fact" with either (proof) terms or Coq's proof script, then only when compilation to module happens, the term/proof script will put into type-check to see if a "Closing Fact" statement can be proved.
@@
 This ``Global Reasoning'' will be again useful in the later examples.
-
-\section{Example}
-We provide two examples here for illustration and evaluation.
-
-\subsection{Type Safety for STLC}
-We start with the very basic type safety proof of STLC, largely adapted from Software Foundation, which is also one of the primal motivating example of this project.
-
-Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the style of the examples in MTC\citep{delaware2013,forsta2020}.
-And we use a \textit{not-yet-completely-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
-This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/inherited family}.
-Compared to MTC, our example uses small-step operational semantic by exploiting the extensible inductive type, and most of the proof is directly adapted from the one in Software Foundation, resulting a more accessible proof.
-% In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach.
-Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users -- which can be reflected from the distance of their approach and text-book Software Foundation proof.
-
-% We have problem on inversion lemma. Check if it is the same problem as MTC
-We use `Closing Fact' to state and prove the inversion lemma instead of using the extensible proving mechanism `FTheorem'.
-The reason is that 1.
-it introduce much less boilerplate code because the proof for these inversion lemma should be just simple case analysis and we should rely on Coq to auto-generate them; 2.
-it shouldn't bother us in the future because any extension on the syntax should still satisfy these inversion lemmas; 3.
-most importantly, we believe this inversion ``lemma'' should be part of the definition of the syntax instead of considering the syntax as a mere concrete inductive type.
-We should postulate this inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify that our inductive definition did satisfy the constraint, which is exactly what we expect from `Closing Fact'.
-
-\subsection{Abstract Interpreters for Imp}
-The second example is adapted and modified from the Familia\citep{zhang2017familia}
-% citation needed here
--- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
-Then we extend the language feature and we instantiate the postulation on computation for both concrete interpreters and abstract interpreters.
-Thanks to the compilation, we can directly run the resulting abstract interpreters.

### 2022-06-16 coq-example.tex
Save. Add Recursor Rules in metatheory.

@@
-We provide two examples here for illustration and evaluation.
+We provide two examples here for illustration and evaluation of our Coq plugin.
 
 \subsection{Type Safety for STLC}
-We start with the very basic type safety proof of STLC, largely adapted from Software Foundation, which is also one of the primal motivating example of this project.
+We start with the very basic type safety proof of STLC, largely adapted from Software Foundation\cite{pierce2014software}, which is also one of the primal motivating example of this project.
 
 Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the style of the examples in MTC\citep{delaware2013,forsta2020}.
-And we use a \textit{not-yet-completely-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
+And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
 This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/inherited family}.
@@
 % We have problem on inversion lemma. Check if it is the same problem as MTC
-We use `Closing Fact' to state and prove the inversion lemma instead of using the extensible proving mechanism `FTheorem'.
-The reason is that 1.
-it introduce much less boilerplate code because the proof for these inversion lemma should be just simple case analysis and we should rely on Coq to auto-generate them; 2.
-it shouldn't bother us in the future because any extension on the syntax should still satisfy these inversion lemmas; 3.
-most importantly, we believe this inversion ``lemma'' should be part of the definition of the syntax instead of considering the syntax as a mere concrete inductive type.
-We should postulate this inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify that our inductive definition did satisfy the constraint, which is exactly what we expect from `Closing Fact'.
+We use "Closing Fact" to state and prove the inversion lemma instead of using the extensible proving mechanism "FTheorem".
+The reasons are that (1) it introduce much less boilerplate code because the proof for these inversion lemma should be just simple case analysis and we should rely on Coq to auto-generate them; (2) it shouldn't bother us in the future because any extension on the syntax should still satisfy these inversion lemmas; (3) most importantly, we believe this inversion ``lemma'' should be part of the definition of the syntax instead of considering the syntax as a mere concrete inductive type.
+We should postulate this inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify that our inductive definition did satisfy the constraint, which is exactly what we expect from "Closing Fact".
+
+Our formulation for bare STLC is around 400+ LOCs; the two families implementing product and boolean features both take around 300+ LOCs. 
+% need comparison with the original implementation
+
+The biggest difference in the proof script comes from the fact that we are handling ``extensible'' inductive type instead of real inductive type, and thus inversion tactic is not working and we have to manually lift these tactic into inversion lemma.
+This cost a mild code bloat.
+
+Another difference comes from the fact that we need to use "Closing Fact" to manually verify the computational rules for each recursor and use tactic to ``run'' the recursor by repeatedly "rewrite" using those verified computational rules.
+This part is possible to be automatically handled by the plugin, but still it will result propositional computational rules.
 
@@
 Thanks to the compilation, we can directly run the resulting abstract interpreters.
+
+We construct four families, in one inheritance chain.
+
+The first family Lang define the big step operator (using fuel) of a while language, with expression. But it postulates the computational information of State used by Lang. It takes about 200+ LOCs.
+
+The second family LangwAbs define the abstract interpreter for it, together with a bunch of postulates, to prove the correctness of the abstract interpreter. It takes about 300+ LOCs.
+
+The third Family LangMore extend LangwAbs with nat constant and addition, and if-then-else control flow, of course retaining all the soundness theorem. It takes about 200+ LOCs.
+
+The fourth family LangRun instantiate some of the postulates of the LangMore and recover computational information of Lang.
+The concrete interpreter is doing as expected; the abstract interpreter is expected to act as a type-analyzer.
+It takes about 200+ LOCs.
+
+And at the end, we compile LangRun into module and use Coq to run it on some simple queries.

### 2022-06-16 syntax.tex
Save. Add Recursor Rules in metatheory.

@@
 $$
-\Rule[name=WSig Proj]
+\Rule[name=WSig-Proj]
 {\goodWSig{\Gamma}{s}{n} \quad j < n}
@@
 \quad
-\Rule[name=Ind Sig]
+\Rule[name=Ind-Sig]
 {\goodWSig{\Delta}{\sigma}{n}
@@
 $$
-\Rule[name=Ind Type]
+\Rule[name=Ind-Type]
 {\goodTerm{\Gamma}{T}{\bW \sigma}}
@@
 \quad
-\Rule[name=Ind Term]
+\Rule[name=Ind-Term]
 {\goodTerm{\Gamma}{T}{\bW \sigma}
@@
 Of course we need to put an eye on the substitution law for each piece of the syntax.
-Here $\bW \sigma$ is a large singleton type \cite{stone2000}, and it only has one "element", the inductive type itself, inside.
+Here $\bW \sigma$ is a large singleton type \cite{stone2000}, and it only has one ``element'', the inductive type itself, inside.
 We formulate in this cumbersome way because we want to have inductive type to be a field element and exposing the constructors at the type level for later \textit{pattern-matching coverage checking} (i.e.
@@
 
-In our meta-theory, we only formulate non-dependent eliminator, i.e. the
-\textit{recursor}. Each recursor is constructed by a linkage containing
-functions/handlers/pattern-matching cases dealing with each
-constructor. Thus code reuse for recursor is delegated to linkage
-inheritance---to reuse one particular pattern-matching case,
-we just inherit the function/handler. This greatly simplifies the
-meta\-theory development.
+The formulation of our recursor is closely related to the linkage, so we will postpone their description.
 
@@
 
-Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by the implementation of our linkage, so an obvious question is what is this sealing judgement for?
+Since our recursor handlers are closely related to linkages, we can finally talk about the rules of recursor now.
+
+$$
+\Rule[name=Hdler]
+{\goodType{\Gamma}{A}{}
+\quad \goodType{\Gamma, A}{B}{}
+\quad \goodType{\Gamma}{R}{}
+}
+{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
+\quad 
+\Rule[name=Hdlers]
+{\goodWSig{\Gamma}{\sigma}{n}
+\quad \goodType{\Gamma}{R}{}
+}
+{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}}
+$$
+$$
+\Rule[name=Hdler-Proj]
+{ \goodWSig{\Gamma}{\sigma}{N}
+\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\sigma~R)}
+\quad n < N
+}
+{\goodTerm{\Gamma}{"prjR"~n~o}{("RecSig1"~(\pi_1^n \sigma)~(\pi_2^n \sigma)~R)[\pi_1]}}
+$$
+$$
+\Rule[name=Rec-Constr]
+{ \goodWSig{\Gamma}{\sigma}{N}
+\quad \goodType{\Gamma}{T}{\bW \sigma}
+\quad \goodType{\Gamma}{R}{}
+\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\sigma~R)}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
+}
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
+$$
+
+In our meta-theory, we only formulate non-dependent eliminator, i.e. the
+\textit{recursor}. Each recursor is constructed by a linkage containing
+functions/handlers/pattern-matching cases dealing with each
+constructor. Thus code reuse for recursor is delegated to linkage
+inheritance---to reuse one particular pattern-matching case,
+we just inherit the function/handler. This greatly simplifies the
+metatheory development.
+
+Intuitively \ruleref{Hdler} is indicating the type of recursor handler for one specific constructor, and \ruleref{Hdlers} is the linkage type of the bundle of all recursor handlers of one given inductive type.
+We omit the definition of "RecSig" but \ruleref{Hdler-Proj} indicates its (apparent) internal definition.
+"prjR" is also the handler selection from the handler linkage, and is helpful for the formulation of $\beta$-rule of "Wrec" (omitted here).
+Finally \ruleref{Rec-Constr} is the used to construct a concrete recursor given a specific linkage of handlers.
+All these rules satisfy substitution laws and omitted here.
+
+Now, one missing piece for linkage is the sealing judgement, and both of the $\mu^+, \nu^+$ rules relies on that -- then what is this sealing judgement for?
 
@@
 
-Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ and thus non-inheritable -- in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}.
+Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and \ruleref{Rec-Constr}) and thus non-inheritable -- in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}.
 These fields together with inductive types are the one sealing responsible for hiding.

### 2022-06-16 canonicitymodel.tex
Save. Fix comments.

@@
 From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out\citep{skorstengaard2019introduction} --
-% cite https://arxiv.org/abs/1907.11133
 using the terminology of the conventional formulation, $\Gamma_p$ will be the set of substitution into closed reducible terms, and we will prove $t[\gamma] \in T_p$, a set of closed reducible term of type $T$ when $\gamma \in \Gamma_p$.

### 2022-06-16 coq-example.tex
Save. Fix comments.

@@
 The second example is adapted and modified from the Familia\citep{zhang2017familia}
-% citation needed here
 -- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.

### 2022-06-16 canonicitymodel.tex
Fixing biblography.

@@
 Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
-We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content\citep{altenkirch2016type}}, 
+We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content\citep{altkap2016}}, 
 then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.

### 2022-06-16 coq-example.tex
Fixing biblography.

@@
 \subsection{Abstract Interpreters for Imp}
-The second example is adapted and modified from the Familia\citep{zhang2017familia}
+The second example is adapted and modified from the Familia\citep{zm2017}
 -- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.

### 2022-06-16 standardmodel.tex
Fixing biblography.

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{altenkirch2016type,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
+We prove the consistency by extending the standard model from \citep{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
 we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.

### 2022-06-18 canonicitymodel.tex
Adapt style guide.

@@
-We follow the reducibility argument from \citep{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model.
+We follow the reducibility argument from~\citet{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model.
 We will first specify the formulation of the most basic canonicity theorem.
@@
 
-Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
-We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content\citep{altkap2016}}, 
+Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
+We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}}, 
 then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.
@@
 The latter is more conventional and the former can be understood as type theoretic encoding of the latter\footnote{We usually use predicate to encode "belong to" relation in type theory}.
-From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out\citep{skorstengaard2019introduction} --
-using the terminology of the conventional formulation, $\Gamma_p$ will be the set of substitution into closed reducible terms, and we will prove $t[\gamma] \in T_p$, a set of closed reducible term of type $T$ when $\gamma \in \Gamma_p$.
+From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out~\cite{skorstengaard2019introduction}---using the terminology of the conventional formulation, $\Gamma_p$ will be the set of substitution into closed reducible terms, and we will prove $t[\gamma] \in T_p$, a set of closed reducible term of type $T$ when $\gamma \in \Gamma_p$.
 

### 2022-06-18 coq-example.tex
Adapt style guide.

@@
 \subsection{Type Safety for STLC}
-We start with the very basic type safety proof of STLC, largely adapted from Software Foundation\cite{pierce2014software}, which is also one of the primal motivating example of this project.
+We start with the very basic type safety proof of STLC, largely adapted from Software Foundation~\cite{pierce2014software}, which is also one of the primal motivating example of this project.
 
-Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the style of the examples in MTC\citep{delaware2013,forsta2020}.
-And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature -- product and boolean -- with vanilla STLC.
+Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the style of the examples in MTC~\cite{delaware2013,forsta2020}.
+And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature---product and boolean---with vanilla STLC.
 This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/inherited family}.
@@
 % In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach.
-Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users -- which can be reflected from the distance of their approach and text-book Software Foundation proof.
+Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users---which can be reflected from the distance of their approach and text-book Software Foundation proof.
 
@@
 
-\subsection{Abstract Interpreters for Imp}
-The second example is adapted and modified from the Familia\citep{zm2017}
--- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
+\subsection{Abstract Interpreters for \texttt{Imp}}
+The second example is adapted and modified from \citet{zm2017}
+--- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
 Then we extend the language feature and we instantiate the postulation on computation for both concrete interpreters and abstract interpreters.

### 2022-06-18 coqimpl.tex
Adapt style guide.

@@
 it is the easiest and the most accessible way to prototype as it relieves us the necessity of familarity of Coq base, especially for the implementation of the module and functor; 2.
-we have a clear definition of trusted-base -- the whole Coq; 3.
-it is easy to debug -- we just need to check the translated commands; 4.
+we have a clear definition of trusted-base---the whole Coq; 3.
+it is easy to debug---we just need to check the translated commands; 4.
 it can be well-incorporated with the existent tools like VSCoq; 5.
-it is more accessible for the the interested audience who can then easily adapt our plugin and give a try -- otherwise they have to download and re-build the whole customized Coq; 6.
+it is more accessible for the the interested audience who can then easily adapt our plugin and give a try---otherwise they have to download and re-build the whole customized Coq; 6.
 it is more stable because the surface syntax of Coq should be stable across different versions; 7.
@@
 T$.
-However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping -- each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a parametrized module (functor).
+However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping---each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a parametrized module (functor).
 Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
@@
 % Concrete details of compilation of a family term data structure
-Taking Figure \ref*{fig:plugin-example1} as an example: when the user interactively input the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
+Taking \cref{fig:plugin-example1} as an example: when the user interactively input the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
 {The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately.
@@
 The internal representation of family "A" is simply the list of these ``compiled'' modules and module types.
-We can simply achieve incremental checking and overriding -- for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
+We can simply achieve incremental checking and overriding---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
 
@@
 We achieve that by repeatedly using Coq's "Include" command as the example illustrated.
-Note that, in our example, we include two functors without arguments -- this is due to one of the feature of "Include": non-instantiated parameters will be automatically instantiated by the appropriate fields in the surrounding defining module.
+Note that, in our example, we include two functors without arguments---this is due to one of the feature of "Include": non-instantiated parameters will be automatically instantiated by the appropriate fields in the surrounding defining module.
 The resulting module is just like mundane Coq's module and thus enjoy all the computational mechanism from Coq for free.
@@
 
-An inheritance data will be annotated with the type for the input family and output family, and generally there are three kinds of inheritance -- extend, override and inherit.
+An inheritance data will be annotated with the type for the input family and output family, and generally there are three kinds of inheritance---extend, override and inherit.
 Extending corresponds to adding new fields, overriding corresponds to modify existent fields and inheriting corresponds to directly taking fields from the parent.
@@
 Otherwise, in the future, once $"add"$ is overridden into a subtraction function then $"comm"$ will have problem to be  inherited.
-A more concrete example is in Figure \ref*{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more.
-To fix that, we need more information on "a" than its type -- we need to expose its concrete definition "a = 1" to appear in the context of "b".
+A more concrete example is in \cref{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more.
+To fix that, we need more information on "a" than its type---we need to expose its concrete definition "a = 1" to appear in the context of "b".
 
@@
 
-To resolve this, we provide family sealing mechanism -- basically we allow the users to specify the family/module type, and thus we can have a ``local'' transparent, non-overridable field located in a overridable family.
+To resolve this, we provide family sealing mechanism---basically we allow the users to specify the family/module type, and thus we can have a ``local'' transparent, non-overridable field located in a overridable family.
 Using the above example, we simply make \mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an overridable family, but "f" to be non-overridable and transparent on its definition, and thus we can prove "p".
@@
 
-This ``sealed'' family actually simulate interface to a certain level -- we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}  , and later instantiate them with concrete implementation by overriding this family.
+This ``sealed'' family actually simulate interface to a certain level---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}  , and later instantiate them with concrete implementation by overriding this family.
 % Basically, {a : int -> int; property : a is monotonic}, in this case
@@
 \textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
-We don't actually support extensible inductive type in Coq (because that require non-trivial research effort on theory side -- i.e.
-{What is the semantic of extensible inductive type?} -- and engineering), but we simulate extensible inductive type using overriding -- when extending an inductive type $A$ with new constructors $c$, our plugin will generate a new inductive definition $A'$ with this new $c$ on the fly and feed to Coq.
+We don't actually support extensible inductive type in Coq (because that require non-trivial research effort on theory side---i.e.
+{What is the semantic of extensible inductive type?}---and engineering), but we simulate extensible inductive type using overriding---when extending an inductive type $A$ with new constructors $c$, our plugin will generate a new inductive definition $A'$ with this new $c$ on the fly and feed to Coq.
 But still, we must be careful on the exposing data in the family type (module type) for the following field.
@@
 Here "FRecursor" will feed Coq
-a module "v_33_34" for doing type-checking -- what it does is basically assert the existence of the recursor and apply to see if handlers can pass the type checking by Coq.
+a module "v_33_34" for doing type-checking---what it does is basically assert the existence of the recursor and apply to see if handlers can pass the type checking by Coq.
 
 Since the recursor is constructed knowing a concrete inductive type, thus not inheritable, for all the children, we have to reconstruct and override with a new recursor (acting like an extra exhaustiveness checking).
-This is affordable since each recursor handler doesn't need to be rechecked -- look at "B2.neg_handler", there is only one new handler for "uu" and the handlers from the parent "B.neg_handler.tt,B.neg_handler.ff" are already type-checked and inherited(thanks to the fact that they only rely on ``the abstracted interface'' module "b_3").
+This is affordable since each recursor handler doesn't need to be rechecked---look at "B2.neg_handler", there is only one new handler for "uu" and the handlers from the parent "B.neg_handler.tt,B.neg_handler.ff" are already type-checked and inherited(thanks to the fact that they only rely on ``the abstracted interface'' module "b_3").
 Of course, our plugin will ``inherits'' the recursor "neg" but what the plugin does is doing exhaustiveness checking, again.
@@
 \textbf{Propositional Partial Recursor.}
-Apparently, the exposed module "b_3" in \cref*{fig:plugin-example2} is not expressive enough.
+Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is not expressive enough.
 For example, we cannot prove the constructors are injective, i.e.
@@
 
-The simplest idea is to allow partiality -- i.e.
+The simplest idea is to allow partiality---i.e.
 returning \mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
@@
 This style of ``Global Reasoning'' is that we can only prove the corresponding theorem when we know the ``full pictures'', including the inductive type.
-We provide "Closing Fact" as a command to carry out this global reasoning -- we can attach "Closing Fact" with either (proof) terms or Coq's proof script, then only when compilation to module happens, the term/proof script will put into type-check to see if a "Closing Fact" statement can be proved.
+We provide "Closing Fact" as a command to carry out this global reasoning---we can attach "Closing Fact" with either (proof) terms or Coq's proof script, then only when compilation to module happens, the term/proof script will put into type-check to see if a "Closing Fact" statement can be proved.
 Note that, "Closing Fact" cannot immeidately check the correctness of the term/proof script because this verification only happens when we finished defining a family, and thus not easy to handle.
@@
 This "Closing Fact" can ease us from the necessity to let plugin generate proof for recursors and partial recursors, and thus making plugin development easier.
-For example, the computation rule for a concrete recursor like "B.neg" and partial recursors in \cref*{fig:plugin-example2} can be directly asserted by using "Closing Fact" and verified later during compilation.
+For example, the computation rule for a concrete recursor like "B.neg" and partial recursors in \cref{fig:plugin-example2} can be directly asserted by using "Closing Fact" and verified later during compilation.
 This ``Global Reasoning'' will be again useful in the later examples.

### 2022-06-18 metatheory.tex
Swap section meta-theory and impl. Please check Thm 3.1, justify that partial recursor is good enough.
Adapt style guide.

@@
-We start with the formulation of the syntax, which is based on the formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT) given by \citep{coquand2018canonicity}.
+We start with the formulation of the syntax, which is based on the formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT) given by~\citet{coquand2018canonicity}.
 Thus we will omit a great deal of details about MLTT here but focus on the new facility we introduced.
 
-Though phrased in Latex, we are following \citet{altkap2016} to use as
+Though phrased in Latex, we are following~\citet{altkap2016} to use as
 our meta-logic a type theoretical framework (like Agda), 
@@
 
-\newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}
-\newcommand{\denotesS}[1]{{{\llbracket {#1} \rrbracket}_S}}
-\newcommand{\goodCtx}[2]{{ {#1} \ \vdash }}
-\newcommand{\goodType}[3]{{ {#1} \vdash {#2} }}
-\newcommand{\goodTerm}[3]{{ {#1} \vdash {#2} : {#3} }}
-\newcommand{\goodSub}[3]{{ {#1} \vdash {#2} : {#3} }}
-\newcommand{\goodSig}[3]{{ {#1} \vdash {#2} \ \  Sig^{#3} }}
-\newcommand{\goodWSig}[3]{{ {#1} \vdash {#2} \ \ WSig^{#3} }}
-\newcommand{\goodSeal}[4]{{ {#1} \vdash {#2} : {#3} \  |\  {#4} }}
-\newcommand{\goodInh}[4]{{ {#1} \vdash {#2} : {#3} \twoheadrightarrow {#4}}}
-\newcommand{\nat}{\mathbf{N}}
-
-\newcommand{\cU}{{\mathcal{U}}}
-\newcommand{\cB}{{\mathbb{B}}}
-\newcommand{\cL}{{\mathcal{L}}}
-\newcommand{\cC}{{\mathcal{C}}}
-\newcommand{\cCt}{{\mathcal{C}_t}}
-\newcommand{\bW}{{\mathbb{W}}}
-
 \subsection{Syntax}

### 2022-06-18 standardmodel.tex
Adapt style guide.

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from \citep{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
+We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
 we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.
@@
 We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e.
-the judgementally-equal syntax needs to be mapped to the same data for well-defined-ness).
+the judgementally-equal syntax needs to be mapped to the same data for well-definedness).
 
@@
 The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
-\textbf{This is actually the formulation in Agda} -- QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.
+\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.
 
-Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data -- it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
-What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-defined-ness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
+Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
+What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-definedness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
 
-Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together -- because $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
+Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
 
@@
 We sketch the model for the inductive type and inheritance judgement here, and both of them require inductive facility in meta-logic.
-In set theory, that would be using \textit{Rule sets}~\citep{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead.
+In set theory, that would be using \textit{Rule sets}~\cite{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead.
 Both of them are somewhat ``interpreted by themselves'' in Agda, because their syntactic definition is pretty ``well-founded''.
-For inductive type, for each grounded signature, we have an inductive type using introduction rule of W-type as constructor, shown in Figure \ref*{fig:model-ind-type}.
+For inductive type, for each grounded signature, we have an inductive type using introduction rule of W-type as constructor, shown in \cref{fig:model-ind-type}.
 This "MWsup" when applying to $n$ will use the $n$-th constructor specify in the signature.
@@
 The verification of validity of this model is largely omitted here and we refer the reader to the appendix.
-To verify our model preserve the judgemental equality, usually unpacking the definition is enough -- for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$ by function extensionality.
+To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$ by function extensionality.
 One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together.
-The compatibility lemmas can be verified directly -- for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{p_1\mu~o}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
+The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{p_1\mu~o}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
 

### 2022-06-18 syntactic-example.tex
Save.
Adapt style guide.

@@
 We provide one example in its surface syntax as pseudocode, and their meta-theory encoding.
-This example is computing the predecessor of a mundane natural number -- however we split the inductive definition into two parts and making "O" and "S" standing alone.
+This example is computing the predecessor of a mundane natural number, but we split the inductive definition into two parts and making "O" and "S" standing alone.
 "pred" is defined as $n \mapsto ("pred" n, n)$, and thus a predecessor.
@@
 Family A.
- Inductive N := O : N. 
+ FInductive N := O : N. 
  (* Osig, sig₁, obj₁, sig₂,
@@
  Family handler.
-   O :self[A].N * self[A].N 
+   O :self[A].N × self[A].N 
      := (self[A].O, self[A].O).
@@
  pred 
-  : self[A].N → self[A].N×self[A].N 
+  : self[A].N → self[A].N × self[A].N 
   := ... 
@@
 Family A₂ extends A.
- Extend Inductive N := S : N -> N.
+ Extend FInductive N := S : N → N.
  (* Nsig, sig₁₁, inh₁, O₂, sig₂₁, inh₂, 
@@
    Inherit O. (* inhᵢₙ₀, pN₃ *)
-   S : N*N -> N*N 
+   S : self[A].N × self[A].N 
+       → self[A].N × self[A].N 
      := fun (x,y) => (y, self[A₂].S y).
@@
 
-Here we explain how  our detailed construction in Figure \ref{fig:example-construction} is constructed.
+Here we explain how  our detailed construction in Figure \cref{fig:example-construction} is constructed.
 We first elaborate that how our inductive type "N" is written in the specific format (the style of W-type).

### 2022-06-18 syntax.tex
1. Recursor rules are added into metatheory 2. Coq Implementation section is done 3. The two Example is basically done, I only need to add statistics. Please Check.
Adapt style guide.

@@
 well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. Our presentation
-follows the style of \citep{kaposi2019gluing}, but for readability we
+follows the style of~\citet{kaposi2019gluing}, but for readability we
 omit universe levels attached to judgements.
@@
 
-Due to our intrinsic setting, \textbf{we can and will omit a lot of presumptions of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.
+Due to our intrinsic setting, \textbf{we can and will omit a lot of presumptions of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly---because it is not possible to have "not-well-formed" context in our setting.
 
@@
 $$
-\Rule[name=Emp WSig]
+\Rule[name=Emp-WSig]
 {}
@@
 \quad
-\Rule[name=WSig Add]
+\Rule[name=WSig-Add]
 {\goodWSig{\Gamma}{s}{n}
@@
 \quad
-\Rule[name=Ind Univ]
+\Rule[name=Ind-Univ]
 {\goodWSig{\Gamma}{\sigma}{n}}
@@
 
-Recall in W-type \citep{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type.
+Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type.
 To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
@@
 Of course we need to put an eye on the substitution law for each piece of the syntax.
-Here $\bW \sigma$ is a large singleton type \cite{stone2000}, and it only has one ``element'', the inductive type itself, inside.
+Here $\bW \sigma$ is a large singleton type~\cite{stone2000}, and it only has one ``element'', the inductive type itself, inside.
 We formulate in this cumbersome way because we want to have inductive type to be a field element and exposing the constructors at the type level for later \textit{pattern-matching coverage checking} (i.e.
@@
 I just didn't write out the two equations c (El T) = T, El (c T) = T}.
-We follow the style of \textit{type universe à la Tarski} and distinguish types from their \textit{codes or names}, and thus we use $El\ T$ as the type given type name $T : \cU$ -- in these cases, $T$ is not allowed to locate at type position (i.e.
+We follow the style of \textit{type universe à la Tarski} and distinguish types from their \textit{codes or names}, and thus we use $El\ T$ as the type given type name $T : \cU$---in these cases, $T$ is not allowed to locate at type position (i.e.
 at the right-most position of a term judgement) but $El \ T$ is.
@@
 They are both constructed from the empty stuff, and respect substitution lemma for sure.
-We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony \citep{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules).
+We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony~\cite{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules).
 We omit them and the substitution laws for either field addition and projection.
-For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using Implicit Arguments}.
+For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.\footnote{This omission can be supported by Agda using Implicit Arguments}
 
@@
 constructor. Thus code reuse for recursor is delegated to linkage
-inheritance---to reuse one particular pattern-matching case,
+inheritance----to reuse one particular pattern-matching case,
 we just inherit the function/handler. This greatly simplifies the
@@
 
-Now, one missing piece for linkage is the sealing judgement, and both of the $\mu^+, \nu^+$ rules relies on that -- then what is this sealing judgement for?
+Now, one missing piece for linkage is the sealing judgement, and both of the $\mu^+, \nu^+$ rules relies on that---then what is this sealing judgement for?
 
@@
 a short hand of term judgement $\goodTerm{\Gamma , \cC
-\sigma}{\_}{A[\pi_1]}$. The substitution law should be understood as a
+\sigma}{\_}{A[\pi_1]}$, a special form of \textit{local morphism}~\cite{abbott2003category}. The substitution law should be understood as a
 lemma, and it is easily derivable using $\gamma^\uparrow$.
 
-The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance -- due to the presence of inductive type in the context, not everything is inheritable.
+The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance---due to the presence of inductive type in the context, not everything is inheritable.
 For example, $ X_1 : \bW \sigma_1 \vdash t : T$ in the parent family will generally not be able to be inherited to the children once we override the inductive type.
 In other words, there is no way ``transplanting'' $t$ into the context $X_2 : \bW \sigma_2$.
-This makes all the fields after this inductive type non-inheritable -- which shouldn't be the case because a lot of fields might not be directly relying on this concrete $\bW \sigma_1$.
+This makes all the fields after this inductive type non-inheritable---which shouldn't be the case because a lot of fields might not be directly relying on this concrete $\bW \sigma_1$.
 
-Sealing is used to exploit this fact -- with sealing, we can add well-typed terms $X_1 : \cU \vdash t : T$, into the parent family, in the context of $ X_1 : \bW \sigma_1$.
+Sealing is used to exploit this fact---with sealing, we can add well-typed terms $X_1 : \cU \vdash t : T$, into the parent family, in the context of $ X_1 : \bW \sigma_1$.
 Later when we want to inherit $t$, because $t$ only consider $X_1$ as an arbitrary type, inheritance is plausible.
 
-Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and \ruleref{Rec-Constr}) and thus non-inheritable -- in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}.
+Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and \ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}.
 These fields together with inductive types are the one sealing responsible for hiding.
@@
 Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type".
-Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear -- because I haven't changed any text}
+Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear---because I haven't changed any text}
 
@@
 when defining new fields in our plugin, and this "self__" is exactly the
-$A$ in the context of the two fields -- especially when we don't do any
+$A$ in the context of the two fields---especially when we don't do any
 sealing, this $A = \cC \sigma$, which is exactly how the implementation
@@
 
-The syntax of the inheritance judgement is very close to how Family is defined in the surface syntax -- a Family with no parent is an inheritance with empty input; an extended Family is exactly an inheritance -- because the parent of an extended Family can be overridden thus inheritance should only define upon the ``interface'' instead of a concrete family.
+The syntax of the inheritance judgement is very close to how Family is defined in the surface syntax---a Family with no parent is an inheritance with empty input; an extended Family is exactly an inheritance---because the parent of an extended Family can be overridden thus inheritance should only define upon the ``interface'' instead of a concrete family.
 However, since our current formulation doesn't support further-binding (i.e.
@@
 
-"inhid" and "inhext" are simple as expected -- the former corresponds to empty inheritance, and the latter gives the programmer the ability to add new field.
+"inhid" and "inhext" are simple as expected---the former corresponds to empty inheritance, and the latter gives the programmer the ability to add new field.
 
@@
 Thus "inhov" is quite like "inhext", where overriding is just throwing away one parent field.
-This difference also show one of the distinct feature of our vanilla family inheritance compared to the mainstream OO inheritance -- family inheritance don't need to introduce the concept of ``subclassing'' between families and thus no bounded polymorphism yet\footnote{Of course once trait and interface comes in, we can have bounded polymorphism back}, and code reuse is mainly achieved by late binding.
+This difference also show one of the distinct feature of our vanilla family inheritance compared to the mainstream OO inheritance---family inheritance don't need to introduce the concept of ``subclassing'' between families and thus no bounded polymorphism yet\footnote{Of course once trait and interface comes in, we can have bounded polymorphism back}, and code reuse is mainly achieved by late binding.
 Without ``subclassing'', there is no way of confusing inheritance and subtyping.

### 2022-06-19 lang-design.tex
Add Lang Design Section

When adapting Family Polymorphism into dependent type, we choose to shrink and focus on only the essence of family (and inheritance) structure in \citet{zm2017}, and thus a lot of unrelated feature, like Interface, will be tailored. In this case, it will look like module with late-binding and inheritance as code reuse for module.

We also need it to have a good compatibility with Inductive Type, because we don't want to lose too much reasoning power of Coq.

### 2022-06-19 standardmodel.tex
Add Lang Design Section

@@
 One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together.
-The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{p_1\mu~o}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
+The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{(p_1\mu~o)}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
 

### 2022-06-19 syntax.tex
Add Lang Design Section

@@
 To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
-Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig Proj} to extract and corresponding $A_i, B_i$.
+Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig-Proj} to extract and corresponding $A_i, B_i$.
 

### 2022-06-19 coqimpl.tex
Adding Language Design Section.

@@
 \textbf{Family compiled into module.} This is the main idea of compiling family components into Coq's primitives: we compile family into coq module, family type into module type and context of a given term into parameters of the module.
-To achieve incremental type checking, we notice the nature of late binding: for each defining field $ ..
-\cL\sigma \vdash t : T $, we can compile $t$ using ``universal-quantifier-wrapped term'' : $..
+Unlike module, we need to faithfully reflect the \textit{late binding} nature of family : for each defining field $ ..
+\cL\sigma \vdash t : T $, we need to compile $t$ using ``universal-quantifier-wrapped term'' : $..
 \vdash \lambda t : \forall \cL \sigma.
 T$.
-However, instead of using Coq's universal quantifier, we use module and module parameters to achieve this wrapping---each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a parametrized module (functor).
+However, instead of using Coq's universal quantifier, we choose module and module parameters to achieve this wrapping---each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a functor parametrized by its context.
 Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
@@
 We can simply achieve incremental checking and overriding---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
+This compiled module type serve as an abstraction to let overriding possible.
 
@@
 Extending corresponds to adding new fields, overriding corresponds to modify existent fields and inheriting corresponds to directly taking fields from the parent.
-Thus it is possible to consider inheritance as solely a data rather than a functor, and this actually provides us the ability to \text{mix-in} inheritance data.
+Thus it is possible to consider inheritance as solely a data rather than a functor, and this actually provides us the possibility to \text{mix-in} inheritance data.
 
-\textbf{Reasoning Requires Non-Overridable.}
-Unlike general-purposed programming, Coq should be able to  \textit{reason} about each program.
-However, sometimes reasoning requires a specific program and overriding on them is unacceptable.
+\textbf{Reasoning Requires Non-Overridable.} This compilation strategy can allow fields to have type dependent on former fields thus fulfilling \ref{langdesign-req1}.
+However, there are some caveats, especially when the form fields are actually ``proof-relevant''.
 For example, we may define a field $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to proving its commutativity in a following field $"comm"$.
@@
 Otherwise, in the future, once $"add"$ is overridden into a subtraction function then $"comm"$ will have problem to be  inherited.
-A more concrete example is in \cref{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more.
+A more concrete example is in \cref{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more (and this is necessary so that "a" can be overridden in the future).
 To fix that, we need more information on "a" than its type---we need to expose its concrete definition "a = 1" to appear in the context of "b".
@@
 To achieve such exposure of definition, in meta-theory we only need to expose the information as propositional equality as part of the signature, and in our Coq plugin, we can achieve that even easier by exposing the whole definition directly into module type.
-Doing so we can have a better judgemental equality and we can avoid overriding because subtraction doesn't have the same definition as $add$.
+Doing so we can have a better judgemental equality and we prohibit overriding because subtraction doesn't have the same definition as $add$.
 We require the users to decorate field as `Final' so that the plugin will proceed with this special treatment.
@@
 \textbf{Special Fields: Overridable and Sealed Family as Interface.} 
-But there are cases reasoning and overriding are both required, especially when we need to use Coq to do computation.
+Non-overridable field is still not expressive enough to achieve the full extent of the dependent type programming.
+There are cases reasoning and overriding are both required, especially when we need to use Coq to do computation.
 For example, we may want a field to be an arbitrary monotonic function.
-Again, we cannot split into two fields \mintinline{Coq}{(f : nat → nat, p : monotonic f)} because to provide "p" we need to have concrete definition of "f" in the family type, which makes "f" non-overridable at all.
+Again, we cannot split into two fields \mintinline{Coq}{(f : nat → nat; p : monotonic f)} because to provide "p" we need to have concrete definition of "f" in the family type, which makes "f" non-overridable at all.
 One option is to use sigma type \mintinline{Coq}{{f | monotonic f}} but that is cumbersome when we have a bundle of stuff with multiple refinements.
 
-To resolve this, we provide family sealing mechanism---basically we allow the users to specify the family/module type, and thus we can have a ``local'' transparent, non-overridable field located in a overridable family.
-Using the above example, we simply make \mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an overridable family, but "f" to be non-overridable and transparent on its definition, and thus we can prove "p".
+To resolve this, we provide family sealing mechanism---basically we allow the users to specify the family/module type, and thus we can have a \textit{locally transparent}, \textit{non-overridable} field located in a \textit{overridable} family.
+Using the above example, we simply make \mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an \textit{overridable} family, but "f" to be non-overridable and transparent on its definition, and thus we can prove "p".
 Once we want to swap the implementation, we just override the whole ``sealed'' family.
 
-This ``sealed'' family actually simulate interface to a certain level---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}  , and later instantiate them with concrete implementation by overriding this family.
+This ``sealed'' family actually simulate interface to a certain level---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.
 % Basically, {a : int -> int; property : a is monotonic}, in this case
@@
 \textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
-We don't actually support extensible inductive type in Coq (because that require non-trivial research effort on theory side---i.e.
+We don't actually support extensible inductive type from \ref{langdesign-req2} in Coq (because that require non-trivial research effort on theory side---i.e.
 {What is the semantic of extensible inductive type?}---and engineering), but we simulate extensible inductive type using overriding---when extending an inductive type $A$ with new constructors $c$, our plugin will generate a new inductive definition $A'$ with this new $c$ on the fly and feed to Coq.
@@
 
-But we do have to provide a way to construct a recursor, and execute type checking (or exhaustiveness checking).
+But we do have to provide a way to construct an (extensible) recursor (as in \ref{langdesign-req3}), and execute type checking (and exhaustiveness checking).
 Thus we need to extract a elimination principle for future usage.
@@
 
-To construct a recursor, we make for each constructor a handler and aggregate them inside a family.
+To construct a recursor as in \ref{langdesign-req3}, we make for each constructor a handler and aggregate them inside a family.
 The whole family "neg_handler" can be considered as a large matching case.
@@
 \textbf{Propositional Partial Recursor.}
-Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is not expressive enough.
+Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is not expressive enough to really fulfill \ref{langdesign-req2}.
 For example, we cannot prove the constructors are injective, i.e.
@@
 However, as we pointed out earlier, the recursor of an inductive type cannot be exposed into the context for the sake of extensibility.
-We may have to alter the formulation of the recursor.
+Thus, apart from the (total) recursor, we need to have another formulation of the recursor i.e., we need to have a better characterization of ``extensible inductive type'' exported in the module type as interface.
 
@@
 % 2. Computational rule for the total recursor
-However, to verify and prove the recursor and partial recursor (and their computational rules), our plugin has to compile the extensible family into a non-extensible Coq Module and prove the recursors by instantiating the content of Coq module with concrete inductive type.
+However, to verify and prove the recursor and partial recursor and their computational rules (as in \ref{langdesign-req4}), our plugin has to compile the extensible family into a non-extensible Coq Module and prove the recursors by instantiating the content of Coq module with concrete inductive type.
 
-This style of ``Global Reasoning'' is that we can only prove the corresponding theorem when we know the ``full pictures'', including the inductive type.
+This style of ``Global Reasoning'' is that we can only prove the corresponding theorem when we know the ``full pictures'', especially the inductive type.
 We provide "Closing Fact" as a command to carry out this global reasoning---we can attach "Closing Fact" with either (proof) terms or Coq's proof script, then only when compilation to module happens, the term/proof script will put into type-check to see if a "Closing Fact" statement can be proved.
-Note that, "Closing Fact" cannot immeidately check the correctness of the term/proof script because this verification only happens when we finished defining a family, and thus not easy to handle.
+Note that, "Closing Fact" cannot immeidately check the correctness of the term/proof script because this verification only happens when we finished defining a family, and thus a bit harder to handle.
 
-This "Closing Fact" can ease us from the necessity to let plugin generate proof for recursors and partial recursors, and thus making plugin development easier.
+This "Closing Fact" can ease us from the necessity to let plugin generate proof for recursors and partial recursors and their computational rules (as in \ref{langdesign-req4}), and thus making plugin development easier.
 For example, the computation rule for a concrete recursor like "B.neg" and partial recursors in \cref{fig:plugin-example2} can be directly asserted by using "Closing Fact" and verified later during compilation.
 This ``Global Reasoning'' will be again useful in the later examples.
+
+\textbf{Hack: ``Meta-data''.}

### 2022-06-19 lang-design.tex
Adding Language Design Section.

@@
-When adapting Family Polymorphism into dependent type, we choose to shrink and focus on only the essence of family (and inheritance) structure in \citet{zm2017}, and thus a lot of unrelated feature, like Interface, will be tailored.
-In this case, it will look like module with late-binding and inheritance as code reuse for module.
+When adapting Family Polymorphism into dependent type, we choose to tailor and focus on only the essence of family (and inheritance) structure in \citet{zm2017}, and thus a lot of unrelated feature, like Interface, will be removed.
+In this case, it will look like module with late-binding and inheritance as code and proof reuse for module.
+We also need it to have a good compatibility with Inductive Type, because we don't want to lose too much reasoning power of Coq (and its tactic programming).
 
-We also need it to have a good compatibility with Inductive Type, because we don't want to lose too much reasoning power of Coq.
+Let's start with the basic example---STLC and its extension---to consider what kind of features are required and how much of them can be supported by family polymorphism.
+% add the example code here
+
+% We want to reason about fields in a family.
+\textlabel{Req (1)}{langdesign-req1}:\textbf{ we want to be able to reason about fields in a family}, just like in a module, where a field can reason about its former fields.
+More generally, we want it possible for a later field to have type dependent on the former fields.
+
+% we want extensible inductive type
+\textlabel{Req (2)}{langdesign-req2}:\textbf{ we want extensible inductive type}.
+The extension of inductive type is not supported by \citet{zm2017} thus we need further consideration about it on both implementation and metatheory.
+Additionally, \textlabel{Req (3)}{langdesign-req3}:\textbf{we also want extensible pattern matching and induction reasoning}.
+There are actually two kinds of ``pattern matching'', one for data and the other for induction reasoning, i.e.
+one use the eliminator to \mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the eliminator to \mintinline{Coq}{Prop}.
+Luckily, in this setting, extensible pattern matching can be easily expressed with family extension -- we just aggregate all case handlers of pattern matching into a family (as a bunch of functions), and then one family can encode one pattern matching, and family inheritance can express adding case handlers.
+Then we just need to introduce a primitive that will ``wrap'' that family into a recursive function.
+Induction reasoning can be handled in an almost identical way.
+
+% we want to be able to reason pattern matching as well
+However, there is still difference between data recursion and induction reasoning, because the later one is \textit{proof-irrelevant}.
+This difference leads to another issue: \textlabel{Req (4)}{langdesign-req4}:\textbf{we want to reason aboue extensible pattern matching}.
+Our substitution lemma is one example,
+
+% we want tactic programming and certain level of automation
+And finally, \textlabel{Req (5)}{langdesign-req5}:\textbf{we want tactic programming}.

### 2022-06-19 coqimpl.tex
Save language design

@@
 \textbf{Reasoning Requires Non-Overridable.} This compilation strategy can allow fields to have type dependent on former fields thus fulfilling \ref{langdesign-req1}.
-However, there are some caveats, especially when the form fields are actually ``proof-relevant''.
-For example, we may define a field $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to proving its commutativity in a following field $"comm"$.
+However, there are some caveats, especially when the former fields are actually ``proof-relevant''.
+For example, we may define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to prove its commutativity in a following field $"comm"$.
 But to prove so, we need to know more than just a type interface of $"add"$ but also its \textbf{concrete definition}.
 Otherwise, in the future, once $"add"$ is overridden into a subtraction function then $"comm"$ will have problem to be  inherited.
-A more concrete example is in \cref{fig:plugin-example1}, consider we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more (and this is necessary so that "a" can be overridden in the future).
+A more concrete example is in \cref{fig:plugin-example1}, consider the case we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more (and this is necessary so that "a" can be overridden in the future).
 To fix that, we need more information on "a" than its type---we need to expose its concrete definition "a = 1" to appear in the context of "b".
@@
 \mintinline{Coq}{self_B.tt ≠ self_B.ff}.
-In the original setting, to prove injectiveness of constructor we only require a recursor and its computational rules.
+In the original setting, to prove injectiveness of constructor of (vanilla) inductive type, we only require a recursor and its computational rules.
 However, as we pointed out earlier, the recursor of an inductive type cannot be exposed into the context for the sake of extensibility.
-Thus, apart from the (total) recursor, we need to have another formulation of the recursor i.e., we need to have a better characterization of ``extensible inductive type'' exported in the module type as interface.
+Thus we need to alter the formulation of the recursor i.e., we need to have a better characterization of ``extensible inductive type'' exported in the module type as interface.
 
@@
 
-\textbf{Hack: ``Meta-data''.}
+\textbf{Hack: ``Meta-data''.} Due to our implementation of the family, the family data is invisible for Coq internal and thus cause some inconvenience.
+For example, currently it is not possible to define a vanilla inductive type dependent on our ``extensible'' inductive type due to the invisiblity, unless we compile the whole family.
+Similarly, customized tactic expression requested by \ref{langdesign-req5} cannot use lemmas or theorems in the fields of the current defining family as well.
+
+To remedy this, we facilitate a "MetaData" command.
+This command will allow to bundle any original Coq primitive (in the current family context) into the module type, and thus visible to the following defining field.
+Doing so, we can define customized tactic as in \ref{langdesign-req5} that uses lemmas proved in the family.

### 2022-06-19 lang-design.tex
Save language design

@@
 
-Let's start with the basic example---STLC and its extension---to consider what kind of features are required and how much of them can be supported by family polymorphism.
+Let's start with the basic example---STLC and its extension in \cref{fig:STLC-example}---to consider what kind of features are required and how much of them can be supported by family polymorphism.
 % add the example code here
+
+\begin{figure}
+  \begin{minipage}[t]{0.47\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family STLC.
+Inductive ty : Set :=
+  | unit : ty
+  | arrow : ty -> ty -> ty.
+Inducitive term : Set := 
+  | var : id -> term 
+  | lam : id -> term -> term ...
+Fixpoint subst 
+  : id -> term -> term := ...
+Inductive has_type 
+  : context -> term -> type := ...
+Proposition subst_lemma :
+  has_type (Γ; x ↦ A) t T ->
+  has_type Γ u A ->
+  has_type Γ (subst x u t) T.
+Inducitve step : term -> term -> Prop 
+  := ...
+(* ... and more, end with type safety *)
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \begin{minipage}[t]{0.47\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family STLC_bool extends STLC.
+Extend Inductive ty : Set :=
+  | bool : ty.
+
+Extend Inducitive term : Set := 
+  | tt : term | ff : term 
+  | tif : term -> term -> term -> term
+Extend Fixpoint subst (* Need to handle new term *)
+  : id -> term -> term := ...
+Extend Inductive has_type (* .. and new ty too *)
+  : context -> term -> type := ...
+Extend Proposition subst_lemma :
+  ... (* Need to prove extra cases *)
+
+Extend Inducitve step : term -> term -> Prop 
+  := ... (* Need to expand this binary relation *)
+(* ... and more extension *)
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \caption{Example STLC and its extension}\label{fig:STLC-example}
+\end{figure}
 
@@
 More generally, we want it possible for a later field to have type dependent on the former fields.
+In the example of \cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require this feature.
 
 % we want extensible inductive type
-\textlabel{Req (2)}{langdesign-req2}:\textbf{ we want extensible inductive type}.
+\textlabel{Req (2)}{langdesign-req2}:\textbf{ we need extensible inductive type}, to extend "term" and "ty" in \cref{fig:STLC-example}.
 The extension of inductive type is not supported by \citet{zm2017} thus we need further consideration about it on both implementation and metatheory.
-Additionally, \textlabel{Req (3)}{langdesign-req3}:\textbf{we also want extensible pattern matching and induction reasoning}.
+Additionally, \textlabel{Req (3)}{langdesign-req3}:\textbf{we also want extensible pattern matching and induction reasoning}, to extend "subst" in \cref{fig:STLC-example}.
 There are actually two kinds of ``pattern matching'', one for data and the other for induction reasoning, i.e.
 one use the eliminator to \mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the eliminator to \mintinline{Coq}{Prop}.
-Luckily, in this setting, extensible pattern matching can be easily expressed with family extension -- we just aggregate all case handlers of pattern matching into a family (as a bunch of functions), and then one family can encode one pattern matching, and family inheritance can express adding case handlers.
+Luckily, in this setting, extensible pattern matching can be easily expressed with family extension -- we just aggregate all case handlers of pattern matching into one family (as a bunch of functions), and then one family can encode one pattern matching, and family inheritance can express adding case handlers.
 Then we just need to introduce a primitive that will ``wrap'' that family into a recursive function.
@@
 % we want to be able to reason pattern matching as well
-However, there is still difference between data recursion and induction reasoning, because the later one is \textit{proof-irrelevant}.
-This difference leads to another issue: \textlabel{Req (4)}{langdesign-req4}:\textbf{we want to reason aboue extensible pattern matching}.
-Our substitution lemma is one example,
-
+However, there is still difference between data recursion and induction reasoning, because the former one is \textit{proof-relevant}.
+This difference leads to another issue: \textlabel{Req (4)}{langdesign-req4}:\textbf{we need to reason about (the computation about) extensible pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
+When proving "subst_lemma", we have to know "subst" is invariant on "tt", i.e.
+\mintinline{Coq}{(subst i x tt) = tt}.
+This kind of information requires \textit{computational rules} from the recursors.
 % we want tactic programming and certain level of automation
-And finally, \textlabel{Req (5)}{langdesign-req5}:\textbf{we want tactic programming}.
+Finally, \textlabel{Req (5)}{langdesign-req5}:\textbf{we desire tactic programming} to relieve us from the hassle of direct manipulation of proof terms.

### 2022-06-20 body.tex
Save.
Save. Introduction.
Save. Add new section.
Save. Working on example section
Swap section meta-theory and impl. Please check Thm 3.1, justify that partial recursor is good enough.
Add Lang Design Section
Add appendix.

--- 2022-05-12 body.tex
@@
 \label{sec:intro}
+\input{introduction.tex}
+
+% Concept 
+% Contribution
+
+\section{Language Design}
+% what feature we want
+% what problem/formalization we want to achieve
+% Before/After Comparison of examples (with/without Fampoly)
+\label{sec:lang-design}
+\input{lang-design.tex}
+
+% Macro for all use
+
+\newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}
+\newcommand{\denotesS}[1]{{{\llbracket {#1} \rrbracket}_S}}
+\newcommand{\goodCtx}[2]{{ {#1} \ \vdash }}
+\newcommand{\goodType}[3]{{ {#1} \vdash {#2} }}
+\newcommand{\goodTerm}[3]{{ {#1} \vdash {#2} : {#3} }}
+\newcommand{\goodSub}[3]{{ {#1} \vdash {#2} : {#3} }}
+\newcommand{\goodSig}[3]{{ {#1} \vdash {#2} \ \  Sig^{#3} }}
+\newcommand{\goodWSig}[3]{{ {#1} \vdash {#2} \ \ WSig^{#3} }}
+\newcommand{\goodSeal}[4]{{ {#1} \vdash {#2} : {#3} \  |\  {#4} }}
+\newcommand{\goodInh}[4]{{ {#1} \vdash {#2} : {#3} \twoheadrightarrow {#4}}}
+\newcommand{\nat}{\mathbf{N}}
+
+\newcommand{\cU}{{\mathcal{U}}}
+\newcommand{\cB}{{\mathbb{B}}}
+\newcommand{\cL}{{\mathcal{L}}}
+\newcommand{\cC}{{\mathcal{C}}}
+\newcommand{\cCt}{{\mathcal{C}_t}}
+\newcommand{\bW}{{\mathbb{W}}}
+
+\section{Prototypical Coq Implementation}
+% we need to make this section before Metatheory
+% because a lot of our formulation in the metatheory
+%   is inspired by the implementation
+\label{sec:coqimpl}
+\input{coqimpl.tex}
+
+\section{Metatheory}
+
+\label{sec:metatheory}
+\input{metatheory.tex}
+
+\section{Example}\label{sec:coqexample}
+\input{coq-example.tex}
 
@@
 \bibliography{refs.bib}
+
+\appendix
+
+\section{Metatheory in (pseudo-)Agda}
+We make three agda source code files (publicly and anonymously) openly avaliable that include the corresponding QIIT syntax definition and models.
+
+For syntax, please check \href{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}{here}.
+
+For consistency model, please check \href{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}{here}.
+
+For canonicity model, please check \href{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}{here}.
+
+We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins.

### 2022-06-20 coq-example.tex
Add appendix.

@@
 Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the style of the examples in MTC~\cite{delaware2013,forsta2020}.
-And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature---product and boolean---with vanilla STLC.
+And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature---product and boolean (in \cref{fig:STLC-example})---with vanilla STLC.
 This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/inherited family}.

### 2022-06-20 body.tex
Adding pseudo-Agda into appendix. I actually have no better idea to include them but only attach a link. Please check.

@@
 \section{Metatheory in (pseudo-)Agda}
-We make three agda source code files (publicly and anonymously) openly avaliable that include the corresponding QIIT syntax definition and models.
+We make three agda source code files (publicly and anonymously) openly avaliable, and they include the corresponding QIIT syntax definition and models.
 

### 2022-06-20 lang-design.tex
Finishing Language Design Section. Try to fix graph embedding and format.
pass on 'lang design' section

--- 2022-06-19 lang-design.tex
@@
-When adapting Family Polymorphism into dependent type, we choose to shrink and focus on only the essence of family (and inheritance) structure in \citet{zm2017}, and thus a lot of unrelated feature, like Interface, will be tailored.
-In this case, it will look like module with late-binding and inheritance as code reuse for module.
+\YZ{looks more like a 'design space' rather than 'language design' section}
+When adapting Family Polymorphism into dependent types, we choose to
+focus only on the essence of family (and inheritance)
+structure in \citet{zm2017}, and thus a lot of unrelated features, like
+Interface, will be removed. In this case, it will look like module with
+late-binding and inheritance as code and proof reuse for module. We also
+need it to have good compatibility with inductive types, because we
+don't want to lose too much reasoning power of Coq (and its tactic
+programming).
 
-We also need it to have a good compatibility with Inductive Type, because we don't want to lose too much reasoning power of Coq.
+Let's start with a basic example---STLC and its extension in
+\cref{fig:STLC-example}---to consider what kind of features are required
+and how much of them can be supported by family polymorphism.
+% add the example code here
+
+\begin{figure}[!htb]
+  \begin{minipage}[t]{0.47\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family STLC.
+Inductive ty : Set :=
+  | unit : ty
+  | arrow : ty -> ty -> ty.
+Inductive term : Set := 
+  | var : id -> term 
+  | lam : id -> term -> term ...
+Fixpoint subst 
+  : id -> term -> term := ...
+Inductive has_type 
+  : context -> term -> type := ...
+Proposition subst_lemma :
+  has_type (Γ; x ↦ A) t T ->
+  has_type Γ u A ->
+  has_type Γ (subst x u t) T.
+Inductive step : term -> term -> Prop 
+  := ...
+(* ... and more, end with type safety *)
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \begin{minipage}[t]{0.47\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family STLC_bool extends STLC.
+Extend Inductive ty : Set :=
+  | bool : ty.
+
+Extend Inductive term : Set := 
+  | tt : term | ff : term 
+  | tif : term -> term -> term -> term
+Extend Fixpoint subst (* Need to handle new term *)
+  : id -> term -> term := ...
+Extend Inductive has_type (* .. and new ty too *)
+  : context -> term -> type := ...
+Extend Proposition subst_lemma :
+  ... (* Need to prove extra cases *)
+
+Extend Inductive step : term -> term -> Prop 
+  := ... (* Need to expand this binary relation *)
+(* ... and more extension *)
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \caption{Example STLC and its extension}\label{fig:STLC-example}
+\end{figure}
+
+% We want to reason about fields in a family.
+\textlabel{Req (1)}{langdesign-req1}:\textbf{ we want to be able to
+reason about fields in a family}, just like in a module, where a field
+can reason about its former fields. More generally, we want to allow a
+later field to be type-dependent on the former fields. In the example of
+\cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
+this feature.\YZ{rephrase Req to Challenges}
+
+% we want extensible inductive type
+\textlabel{Req (2)}{langdesign-req2}:\textbf{ we need extensible
+inductive types}, to extend "term" and "ty" in \cref{fig:STLC-example}.
+Extensible inductive types are not supported by \citet{zm2017}, so
+we need further consideration about it on both implementation and
+metatheory.
+
+\textlabel{Req (3)}{langdesign-req3}: \textbf{we also want extensible
+pattern matching and induction reasoning}, to extend "subst" in
+\cref{fig:STLC-example}.
+There are actually two kinds of ``pattern matching'', one for data and
+the other for induction reasoning, i.e., one uses the eliminator to
+\mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the
+eliminator to \mintinline{Coq}{Prop}. Luckily, in this setting,
+extensible pattern matching can be easily expressed with family
+extension---we just aggregate all case handlers of pattern matching
+into one family (as a bunch of functions), and then one family can
+encode one pattern matching, and family inheritance can express adding
+case handlers. Then we just need to introduce a primitive that will
+``wrap'' that family into a recursive function. Induction reasoning can
+be handled in an almost identical way.
+
+% we want to be able to reason pattern matching as well
+However, there is still difference between data recursion and induction
+reasoning, because the former one is \textit{proof-relevant}. This
+difference leads to another issue: \textlabel{Req (4)}{langdesign-req4}:
+\textbf{we need to reason about (the computation about) extensible
+pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
+When proving "subst_lemma", we have to know "subst" is invariant on
+"tt", i.e., \mintinline{Coq}{(subst i x tt) = tt}. This kind of
+information requires exposing \textit{computational rules} from the
+recursors.\YZ{What goes wrong if subst is not invariant or you do not
+know whether subst is invariant?}
+
+% we want tactic programming and certain level of automation
+Finally, \textlabel{Req (5)}{langdesign-req5}: \textbf{we desire tactic
+programming} to relieve us from the hassle of direct manipulation of
+proof terms.\YZ{What about modular type checking?}

### 2022-06-20 coqimpl.tex
Finishing Language Design Section. Try to fix graph embedding and format.
Comments Reply.

@@
 % Insert one pseudo-code example for explanation of the mechanism
-\begin{figure}
+\begin{figure}[!htb]
   \begin{minipage}[t]{0.25\linewidth}
@@
 The internal representation of family "A" is simply the list of these ``compiled'' modules and module types.
-We can simply achieve incremental checking and overriding---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
+We can simply achieve incremental checking and overriding as in \ref{langdesign-req0}---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
 This compiled module type serve as an abstraction to let overriding possible.
@@
 % Use the natural number example
-\begin{figure}
+\begin{figure}[!htb]
   \begin{minipage}[t]{0.3\linewidth}

### 2022-06-20 lang-design.tex
Comments Reply.

@@
-\YZ{looks more like a 'design space' rather than 'language design' section}
+\YZ{looks more like a 'design space' rather than 'language design' section}\EDJreply{Good point. Do I just change the section name?s}
 When adapting Family Polymorphism into dependent types, we choose to
@@
 
+First and foremost, \textlabel{Req (0)}{langdesign-req0}\textbf{we desire incremental(modular) type checking}.
+Notice that, the modularity here in Coq is a bit different from other language, because Coq is \textit{interactive} theorem proving so we actually need \textit{statement-wise} incremental type checking, not only for saving the compilation time, but also for the immediate feedback and type-checking for the Coq developer.
+
 % We want to reason about fields in a family.
@@
 \cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
-this feature.\YZ{rephrase Req to Challenges}
+this feature.\YZ{rephrase Req to Challenges}\EDJreply{But it is not really challenge because I didn't show their difficulty. It is more like requirements.}
 
@@
 recursors.\YZ{What goes wrong if subst is not invariant or you do not
-know whether subst is invariant?}
+know whether subst is invariant?}\EDJreply{Then it is unprovable because substitution lemma need to know this. You can just instantiate the whole lemma -- we want  
+has_type (Γ; x ↦ A) tt T ->
+has_type Γ u A ->
+has_type Γ (subst i x tt) T. But this final goal requires to know the computation of subst.}
 
@@
 programming} to relieve us from the hassle of direct manipulation of
-proof terms.\YZ{What about modular type checking?}
+proof terms.\YZ{What about modular type checking?}\EDJreply{Good point! Let me add that.}

### 2022-06-20 related-work.tex
Adding related-work section.

This section discusses related work.

\textbf{Modular Mechanized Metatheory: Syntactic Approach} We group the work that achieves modular mechanized metatheory by extending surface syntax of proof assistants as \textit{syntactic approach}.

The closest work to ours might be \citet{boite2004proo}--they try to support \textit{extended} inductive type and related proof reuse in Coq. This biggest difference lies in our (incomplete) adaption of Family Polymorphism -- we support proof reuse and extension by inheritance, but also support implementation swapping using overriding, and we also argue that Family Polymorphism can provide a better and cleaner way of organizing mechanized metatheory development. The other difference lies in the handling of the extension of the inductive type. In our paradigm, the discussion of \textit{extensible} inductive type and partial recursor is unavoidable, which characterizes all possible extension of a given inductive type. In their work, without using Family to organized the code, the extended inductive type can be considered as a different inductive type than the original one, possibly leading to more expression problem~\cite{wadler-ep}.

\textbf{Modular Mechanized Metatheory: Semantic Approach} We group the work using certain algebraic encoding (e.g. Church Encoding, datatype-generic programming) without modify the surface syntax of the proof assistants as \textit{semantic approach}, that includes \citet{delaware2013,forsta2020,liwei2022,schwaab2013modular, keuchel2013generic} and so on. Their advantages are various: this technique is more general and thus can be applied across different (dependent-typed) proof assistants and this technique can be easily distributed like a library and thus doesn't need to alter the core of the proof assistants. The downsides are mainly about the threshold: they are not friendly for the junior Coq development and these indirect encoding of datatype might lead to worse readability and accessibility compared to the direct style like ours.

\textbf{Solution to The Expression Problem}
% Have no idea how to write this :(

### 2022-06-22 coqimpl.tex
minimal edits in coqimpl.tex

--- 2022-06-19 coqimpl.tex
@@
-After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
+After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
 
-Despite of the possible difficulty of the future maintenance, this approach has various advantages: 1.
-it is the easiest and the most accessible way to prototype as it relieves us the necessity of familarity of Coq base, especially for the implementation of the module and functor; 2.
-we have a clear definition of trusted-base---the whole Coq; 3.
-it is easy to debug---we just need to check the translated commands; 4.
-it can be well-incorporated with the existent tools like VSCoq; 5.
-it is more accessible for the the interested audience who can then easily adapt our plugin and give a try---otherwise they have to download and re-build the whole customized Coq; 6.
-it is more stable because the surface syntax of Coq should be stable across different versions; 7.
-still, this plugin can capture the key ingredients of implementing family polymorphism inside Coq and act as a reference for guiding an appropriate implementation of Family Polymorphism inside all sorts of proof assistants.
+Despite possible difficulties for future maintenance, this approach has
+various advantages: 1. it is the easiest and the most accessible way to
+prototype as it relieves us the necessity of familarity of Coq base,
+especially for the implementation of the module and functor; 2. we have
+a clear definition of trusted-base---the whole Coq; 3. it is easy to
+debug---we just need to check the translated commands; 4. it can be
+well-incorporated with the existent tools like VSCoq; 5. it is more
+accessible for the interested audience who can then easily adapt our
+plugin and give a try---otherwise they have to download and re-build the
+whole customized Coq; 6. it is more stable because the surface syntax of
+Coq should be stable across different versions; 7. still, this plugin
+can capture the key ingredients of implementing family polymorphism
+inside Coq and act as a reference for guiding an appropriate
+implementation of Family Polymorphism inside all sorts of proof
+assistants.\YZ{Does the Coq implementation perform type checking to
+prevent illegal uses fam poly?}
 
-\textbf{Family compiled into module.} This is the main idea of compiling family components into Coq's primitives: we compile family into coq module, family type into module type and context of a given term into parameters of the module.
-Unlike module, we need to faithfully reflect the \textit{late binding} nature of family : for each defining field $ ..
-\cL\sigma \vdash t : T $, we need to compile $t$ using ``universal-quantifier-wrapped term'' : $..
-\vdash \lambda t : \forall \cL \sigma.
-T$.
-However, instead of using Coq's universal quantifier, we choose module and module parameters to achieve this wrapping---each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a functor parametrized by its context.
-Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
+\textbf{Family compiled into module.} This is the main idea of compiling
+family components into Coq's primitives: we compile family into Coq
+module, family type into module type, and the context of a given term
+into parameters of the module. Unlike module, we need to faithfully
+reflect the \textit{late binding} nature of family: for each defining
+field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
+``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
+\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
+choose module and module parameters to achieve this wrapping---each
+family type in the context of the judgement will be one module parameter
+with the compiled module type, and thus each field of the family will
+compile into a functor parametrized by its context. Doing so we can also
+get rapid feedback from type checking when the users are defining each
+field interactively with Coq.
 % Insert one pseudo-code example for explanation of the mechanism
-\begin{figure}
+\begin{figure}[!htb]
   \begin{minipage}[t]{0.25\linewidth}
@@
 \end{minted}
-  \end{minipage}
+  \end{minipage}\YZ{Can the self's be omitted?}
   \begin{minipage}[t]{0.4\linewidth}
@@
 % Concrete details of compilation of a family term data structure
-Taking \cref{fig:plugin-example1} as an example: when the user interactively input the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
+Taking \cref{fig:plugin-example1} as an example: when the user interactively inputs the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
 {The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately.

### 2022-06-22 coqimpl.tex

@@
 % Concrete details of compilation of a family term data structure
-Taking \cref{fig:plugin-example1} as an example: when the user interactively inputs the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
-{The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately.
-The computed "Module a_10" is then part of the context of "Field b" (as the module type of "self__A") for "b" to refer to.
-The internal representation of family "A" is simply the list of these ``compiled'' modules and module types.
-We can simply achieve incremental checking and overriding---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
-This compiled module type serve as an abstraction to let overriding possible.
+Taking \cref{fig:plugin-example1} as an example: when the user
+interactively inputs the line "Field a : nat := ..", our plugin will
+translate this statement into the "Module a_9", "Module Type a_10, a_11"
+three components. {The "EmptySig" is just empty module type.} Then,
+these modules and module types will be generated and type-checked by Coq
+immediately. The computed "Module a_10" is then part of the context of
+"Field b" (as the module type of "self__A") for "b" to refer to. The
+internal representation of family "A" is simply the list of these
+``compiled'' modules and module types. We can simply achieve incremental
+checking and overriding as in \ref{langdesign-req0}---for example, we
+want to override "Field a", we simply replace that compiled module with
+the new module inside the list. This compiled module type serve as an
+abstraction to let overriding possible.
 

### 2022-06-22 coqimpl.tex
typo?

--- 2022-06-22 coqimpl.tex
@@
 % Concrete details of compilation of a family term data structure
-Taking \cref{fig:plugin-example1} as an example: when the user interactively inputs the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components.
-{The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately.
-The computed "Module a_10" is then part of the context of "Field b" (as the module type of "self__A") for "b" to refer to.
-The internal representation of family "A" is simply the list of these ``compiled'' modules and module types.
-We can simply achieve incremental checking and overriding---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list.
-This compiled module type serve as an abstraction to let overriding possible.
+Taking \cref{fig:plugin-example1} as an example: when the user
+interactively inputs the line "Field a : nat := ..", our plugin will
+translate this statement into the "Module a_9", "Module Type a_10, a_11"
+three components.\YZ{Typo? Are they a_4, a_5, a_6 in Figure 2?}
+{The "EmptySig" is just empty module type.} Then,
+these modules and module types will be generated and type-checked by Coq
+immediately. The computed "Module a_10" is then part of the context of
+"Field b" (as the module type of "self__A") for "b" to refer to. The
+internal representation of family "A" is simply the list of these
+``compiled'' modules and module types. We can simply achieve incremental
+checking and overriding as in \ref{langdesign-req0}---for example, if we
+want to override "Field a", we simply replace that compiled module with
+the new module inside the list. This compiled module type serves as an
+abstraction that enables overriding.
 
-Now that we know a family data structure is actually a list of compiled modules, we may want to aggregate them into one complete compiled module.
-We achieve that by repeatedly using Coq's "Include" command as the example illustrated.
-Note that, in our example, we include two functors without arguments---this is due to one of the feature of "Include": non-instantiated parameters will be automatically instantiated by the appropriate fields in the surrounding defining module.
-The resulting module is just like mundane Coq's module and thus enjoy all the computational mechanism from Coq for free.
+Now that we know a family data structure is actually a list of compiled
+modules, we may want to aggregate them into one complete compiled
+module. We achieve that by repeatedly using Coq's "Include" command as
+the example illustrated. Note that, in our example, we include two
+functors without arguments---this is due to one of the feature of
+"Include": non-instantiated parameters will be automatically
+instantiated by the appropriate fields in the surrounding defining
+module. The resulting module is just like mundane Coq's module and thus
+enjoy all the computational mechanism from Coq for free.
 

### 2022-06-22 lang-design.tex
typo?

@@
 programming} to relieve us from the hassle of direct manipulation of
-proof terms.\YZ{What about modular type checking?}\EDJreply{Good point! Let me add that.}
+proof terms.

### 2022-06-23 related-work.tex
Save

@@
 The closest work to ours might be \citet{boite2004proo}--they try to support \textit{extended} inductive type and related proof reuse in Coq.
-This biggest difference lies in our (incomplete) adaption of Family Polymorphism -- we support proof reuse and extension by inheritance, but also support implementation swapping using overriding, and we also argue that Family Polymorphism can provide a better and cleaner way of organizing mechanized metatheory development.
+This biggest difference lies in our (incomplete) adaption of Family Polymorphism -- apart from supporting proof reuse and extension by inheritance, we also support implementation swapping using overriding, and we also argue that Family Polymorphism can provide a better and cleaner way of organizing mechanized metatheory development.
 The other difference lies in the handling of the extension of the inductive type.
-In our paradigm, the discussion of \textit{extensible} inductive type and partial recursor is unavoidable, which characterizes all possible extension of a given inductive type.
+In our paradigm, the discussion of \textit{extensible} inductive type and \textit{partial recursor} is unavoidable, which characterizes all possible extension of a given inductive type.
 In their work, without using Family to organized the code, the extended inductive type can be considered as a different inductive type than the original one, possibly leading to more expression problem~\cite{wadler-ep}.
+% Another similar work is proof weaving \citep{mulhern2006proof} but I don't see a evaluation section in that paper
+
+The other work
 

### 2022-06-23 coqimpl.tex

@@
 the example illustrated. Note that, in our example, we include two
-functors without arguments---this is due to one of the feature of
+functors without arguments---this is due to a perculiarity of Coq's
 "Include": non-instantiated parameters will be automatically
 instantiated by the appropriate fields in the surrounding defining
-module. The resulting module is just like mundane Coq's module and thus
+module.\YZ{'For example, ...'}
+The resulting module is just like mundane Coq's module and thus
 enjoy all the computational mechanism from Coq for free.
@@
 
-An inheritance data will be annotated with the type for the input family and output family, and generally there are three kinds of inheritance---extend, override and inherit.
-Extending corresponds to adding new fields, overriding corresponds to modify existent fields and inheriting corresponds to directly taking fields from the parent.
-Thus it is possible to consider inheritance as solely a data rather than a functor, and this actually provides us the possibility to \text{mix-in} inheritance data.
+An inheritance data will be annotated with the type for the input family
+and output family, and generally there are three kinds of
+inheritance---extend, override and inherit. Extending corresponds to
+adding new fields, overriding corresponds to modifying existent fields,
+and inheriting corresponds to directly taking fields from the parent.
+Thus, it is possible to consider inheritance as solely a data rather than
+a functor, and this actually provides us the possibility to
+\text{mix-in} inheritance data.
 
-\textbf{Reasoning Requires Non-Overridable.} This compilation strategy can allow fields to have type dependent on former fields thus fulfilling \ref{langdesign-req1}.
-However, there are some caveats, especially when the former fields are actually ``proof-relevant''.
-For example, we may define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field and intends to prove its commutativity in a following field $"comm"$.
-But to prove so, we need to know more than just a type interface of $"add"$ but also its \textbf{concrete definition}.
-Otherwise, in the future, once $"add"$ is overridden into a subtraction function then $"comm"$ will have problem to be  inherited.
-A more concrete example is in \cref{fig:plugin-example1}, consider the case we have "Field b : self_A.a = 1 := .." instead, then our compilation and Coq's type check will both fail: in the context of "b" (module type "a_6") we only know "a" is of type nat but nothing more (and this is necessary so that "a" can be overridden in the future).
-To fix that, we need more information on "a" than its type---we need to expose its concrete definition "a = 1" to appear in the context of "b".
+\textbf{Reasoning Requires Non-Overridable.} This compilation strategy
+can allow fields to be type-dependent on former fields thus fulfilling
+\ref{langdesign-req1}. However, there are some caveats, especially when
+the former fields are actually ``proof-relevant''. For example, we may
+define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
+and intend to prove its commutativity in a following field $"comm"$.
+But to prove so, we need to know more than just a type interface of
+"add" but also its \textbf{concrete definition}. Otherwise, in the
+future, once "add" is overridden into a subtraction function then
+"comm" will have problem to be inherited.
+A more concrete example is in \cref{fig:plugin-example1}, consider the
+case we have "Field b : self_A.a = 1 := .." instead. Then our
+compilation and Coq's type checking will both fail: in the context of
+"b" (module type "a_6") we only know "a" is of type "nat" but nothing
+more (and this is necessary so that "a" can be overridden in the
+future). To fix that, we need more information on "a" than its type---we
+need to expose the concrete definition "a = 1" to the context of "b".
+\YZ{I had a hard time working through these two paras. I guess they are
+saying that there is a need to expose definitional equalities, and
+therefore we need non-overridables. This causal flow should be made
+more explicit.}
 
-To achieve such exposure of definition, in meta-theory we only need to expose the information as propositional equality as part of the signature, and in our Coq plugin, we can achieve that even easier by exposing the whole definition directly into module type.
-Doing so we can have a better judgemental equality and we prohibit overriding because subtraction doesn't have the same definition as $add$.
-We require the users to decorate field as `Final' so that the plugin will proceed with this special treatment.
+To achieve such exposure of definition, in meta-theory we only need to
+expose the information as propositional equality as part of the
+signature, and in our Coq plugin, we can achieve that even easier by
+exposing the whole definition directly into module type. Doing so we can
+have a better judgemental equality and we prohibit overriding because
+subtraction doesn't have the same definition as $add$. We require the
+users to decorate field as `Final' so that the plugin will proceed with
+this special treatment.\YZ{This para is confusing: which kind of equality is used, propositional or judgmental? Where does 'Final' show up in the example?}
 

### 2022-06-23 coqimpl.tex
Reply part of the comments.

@@
 assistants.\YZ{Does the Coq implementation perform type checking to
-prevent illegal uses fam poly?}
+prevent illegal uses fam poly?}\EDJreply{Yes. Because I define feature of fampoly as a direct translation into Coq feature.  The fampoly is like a shorthand into multiple Coq commands. \\
+Do you find translation from my current metatheory into MLTT (without linkage) helpful? \\
+Actually I am also not sure what this question is aiming towards. \\
+Are you complaining there is no formal specification of my fampoly feature in paper and thus there is no proof my Coq surface syntax is related to metatheory?
+If so, I thought our Coq plugin is only advertising how powerful Fampoly is to solve expression problem in mechanized proving.
+I can decouple the relationship between our plugin and our metatheory if you find the gap is really there...
+And I can only claim metatheory is about a partial result on incorporating fampoly with dependent type, Coq plugin is using two example to illustrate how mechanized proving can get benefit and our Coq plugin is only inspired by our metatheory (and not claim any relationship).
+Doing so would require rearrange the metatheory section before this section.
+\\
+But I don't think the gap is there at all because if you want all possible gaps to disappear, then the action of using debruijn is wrong because we need to prove the one using debruijn and the surface syntax that is not using debruijn are ``equivalent''.
+What's more, module is not sigma type because Module is not a term in Coq and thus there is no formalism of Module at all.
+Let alone I don't see the plausibility of mechanizing dependent type without debruijn.
+\\
+Please comment back so that I know what corresponding modification is needed on the paper. 
+}
 
@@
 \end{minted}
-  \end{minipage}\YZ{Can the self's be omitted?}
+  \end{minipage}\YZ{Can the self's be omitted?}\EDJreply{Yes.
+But in our current plugin implementation it is not omitted.
+I think altering example is fine.
+If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}
   \begin{minipage}[t]{0.4\linewidth}
@@
 interactively inputs the line "Field a : nat := ..", our plugin will
-translate this statement into the "Module a_9", "Module Type a_10, a_11"
-three components.\YZ{Typo? Are they a_4, a_5, a_6 in Figure 2?}
+translate this statement into the "Module a_4", "Module Type a_5, a_6"
+three components.\YZ{Typo? Are they a_4, a_5, a_6 in Figure 2?}\EDJreply{Thanks, fixed.}
 {The "EmptySig" is just empty module type.} Then,
@@
 instantiated by the appropriate fields in the surrounding defining
-module.\YZ{'For example, ...'}
+module.\YZ{'For example, ...'}\EDJreply{I don't think explaining a Coq feature inside our paper is a good idea?
+But I have added this in the following line, please check and retain it if you find this ok.
+}
+For example, "b_7" requires a module of type "a_6".
+Inside module "A", when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat} inside the current surrounding "A".
+Then when we \mintinline{Coq}{Include b_7} without specifying the module parameters, Coq will try to find fields in the surrounding module, (i.e.
+module "A") to instantiate \mintinline{Coq}{self_A: a_6}.
+This is satisfied by the \mintinline{Coq}{a : nat} that is included earlier.
 The resulting module is just like mundane Coq's module and thus
@@
 can allow fields to be type-dependent on former fields thus fulfilling
-\ref{langdesign-req1}. However, there are some caveats, especially when
-the former fields are actually ``proof-relevant''. For example, we may
+\ref{langdesign-req1}. However, there are some cases we need to expose the concrete definition of a given fields (especially when we are proving properties about it).
+
+For example, we may
 define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
@@
 But to prove so, we need to know more than just a type interface of
-"add" but also its \textbf{concrete definition}. Otherwise, in the
+"add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
 future, once "add" is overridden into a subtraction function then
@@
 more (and this is necessary so that "a" can be overridden in the
-future). To fix that, we need more information on "a" than its type---we
+future).
+
+To fix that, we need more information on "a" than its type---we
 need to expose the concrete definition "a = 1" to the context of "b".
@@
 therefore we need non-overridables. This causal flow should be made
-more explicit.}
+more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph.}
 
-To achieve such exposure of definition, in meta-theory we only need to
-expose the information as propositional equality as part of the
-signature, and in our Coq plugin, we can achieve that even easier by
-exposing the whole definition directly into module type. Doing so we can
-have a better judgemental equality and we prohibit overriding because
+In our Coq plugin, we can achieve such exposure of definition by
+exposing the whole definition directly into module type. Doing so we also prohibit overriding because
 subtraction doesn't have the same definition as $add$. We require the
 users to decorate field as `Final' so that the plugin will proceed with
-this special treatment.\YZ{This para is confusing: which kind of equality is used, propositional or judgmental? Where does 'Final' show up in the example?}
+this special treatment.\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
+Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on taking about meta-theory.
+Now everything is about judgemental so I don't have to emphasize judgemental.
+I wanted to make the example small but I guess I can add one `Final' statement into our example.
+}
 

### 2022-06-23 coqimpl.tex
Language Design Section, second attempt.

@@
 In our Coq plugin, we can achieve such exposure of definition by
-exposing the whole definition directly into module type. Doing so we also prohibit overriding because
+exposing the whole definition directly into the compiled module type. Doing so we also prohibit overriding because
 subtraction doesn't have the same definition as $add$. We require the
@@
 
-To resolve this, we provide family sealing mechanism---basically we allow the users to specify the family/module type, and thus we can have a \textit{locally transparent}, \textit{non-overridable} field located in a \textit{overridable} family.
+To resolve this, we provide family sealing mechanism---basically we allow the users to specify the family/compiled module type, and thus we can have a \textit{locally transparent}, \textit{non-overridable} field located in a \textit{overridable} family.
 Using the above example, we simply make \mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an \textit{overridable} family, but "f" to be non-overridable and transparent on its definition, and thus we can prove "p".
@@
 {What is the semantic of extensible inductive type?}---and engineering), but we simulate extensible inductive type using overriding---when extending an inductive type $A$ with new constructors $c$, our plugin will generate a new inductive definition $A'$ with this new $c$ on the fly and feed to Coq.
-But still, we must be careful on the exposing data in the family type (module type) for the following field.
+But still, we must be careful on the exposing data in the family type (the compiled module type) for the following field.
 
@@
 Let's look at \cref{fig:plugin-example2}, the example of boolean and negation.
-When we define inductive type "b", the following "Field neg_handler.ff" can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface'' module type "b_3".
+When we define inductive type "b", the following "Field neg_handler.ff" can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface'' the compiled module type "b_3".
 We cannot export the eliminator of "b" inside "b_3" because doing so "neg_handler.ff" (and "ff_24") will not be able to be inherited to the context where "b" is extended with a third constructor, since the existence of recursor alone will assert that there are only two constructors in "b", not ``extensible'' at all!
@@
 However, as we pointed out earlier, the recursor of an inductive type cannot be exposed into the context for the sake of extensibility.
-Thus we need to alter the formulation of the recursor i.e., we need to have a better characterization of ``extensible inductive type'' exported in the module type as interface.
+Thus we need to alter the formulation of the recursor i.e., we need to have a better characterization of ``extensible inductive type'' exported in the compiled module type as a part of the interface for the following fields to use.
 
@@
 Because "reflect" is actually a left inverse of an injection "bool → self_B.b".
-A bit more formally, and restrict our focus to non-indexed inductve types,
+A bit more formally, and restrict our focus to non-indexed inductive types,
 
@@
 We provide "Closing Fact" as a command to carry out this global reasoning---we can attach "Closing Fact" with either (proof) terms or Coq's proof script, then only when compilation to module happens, the term/proof script will put into type-check to see if a "Closing Fact" statement can be proved.
-Note that, "Closing Fact" cannot immeidately check the correctness of the term/proof script because this verification only happens when we finished defining a family, and thus a bit harder to handle.
+Note that, "Closing Fact" cannot immediately check the correctness of the term/proof script because this verification only happens when we finished defining a family, and thus a bit harder to handle.
 
@@
 \textbf{Hack: ``Meta-data''.} Due to our implementation of the family, the family data is invisible for Coq internal and thus cause some inconvenience.
-For example, currently it is not possible to define a vanilla inductive type dependent on our ``extensible'' inductive type due to the invisiblity, unless we compile the whole family.
+For example, currently it is not possible to define a vanilla inductive type dependent on our ``extensible'' inductive type due to the invisibility, unless we compile the whole family.
 Similarly, customized tactic expression requested by \ref{langdesign-req5} cannot use lemmas or theorems in the fields of the current defining family as well.
@@
 To remedy this, we facilitate a "MetaData" command.
-This command will allow to bundle any original Coq primitive (in the current family context) into the module type, and thus visible to the following defining field.
+This command will allow to bundle any original Coq primitive (in the current family context) into the compiled module type, and thus visible to the following defining field.
 Doing so, we can define customized tactic as in \ref{langdesign-req5} that uses lemmas proved in the family.

### 2022-06-23 lang-design.tex
Save
Language Design Section, second attempt.

@@
+\subsection{Design Space}
+
 \YZ{looks more like a 'design space' rather than 'language design' section}\EDJreply{Good point. Do I just change the section name?s}
@@
 Finally, \textlabel{Req (5)}{langdesign-req5}: \textbf{we desire tactic
-programming} to relieve us from the hassle of direct manipulation of
-proof terms.
+programming} to relieve us from the hassle of direct manipulation of proof terms.
+
+\subsection{Language Design}
+Inspired by our above requirements and Family Polymorphism, we propose the following extended syntax for Coq Vernacular Commands. 
+% basically write down the specification of surface Coq commandline
+%   and then explain  how the language design makes the programmer’s life easier
+\begin{align*}
+  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
+  \quad |\quad "EndFamily" \\
+  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
+  & |\quad "Extend"~"Family"~\langle ident\rangle \\ 
+  & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
+  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
+  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle
+\end{align*}
+
+We use "Family" and "EndFamily" to scope.
+Each field in one family is declared and defined by "Field".
+We can define nested family.
+Family will also be used to encode recursive handlers.
+Each case handling of one inductive type will be aggregated into one family, then "FRecursor" will be used to construct a recursor using these handlers.
+For example, \mintinline{Coq}{FRecursor recur about T motive f} "using" "handlers" will generate a new field "recur" that is doing induction (using elimination principle of "T") using motive "f", and the handlers are located in the family "handlers".
+
+We can inherit a top level family "parent" using \mintinline{Coq}{Family ...
+extends parent} as a new top level family.
+As expected, "Override" will override/a field from the parent.
+"Inherit" can inherit nested family, inductive type and fields.
+The fields inside a family are order-sensitive because we have dependent type here.
+So "Inherit Until" is used if the user want to inherit a range of fields.
+If the user wants to extend an inductive type or a nested family from the parent, then "Extend FInductive" or "Extend Family" is used.
+Thus nested family can only be inherit or extended together with its outer family (and thus top level family), thus our support of family polymorphism is incomplete.
+
+This incomplete support can already be helpful for extending functions.
+Take \cref{fig:plugin-example2} as example, if we want to extend boolean inductive type and its negation, we only need to specify the new constructor for boolean and how negation deal with the new constructor.

### 2022-06-23 coqimpl.tex
Save Language Design.
Reply all the comments.

@@
+In this section, we describe how we can implement the proposed Vernacular commands.
+
 After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
@@
   := eq_refl.
+Final Field a' : nat 
+  := 1. 
+Field b'
+  : self_A.a' = 1
+  := eq_refl.
 EndFamily.
@@
 module.\YZ{'For example, ...'}\EDJreply{I don't think explaining a Coq feature inside our paper is a good idea?
-But I have added this in the following line, please check and retain it if you find this ok.
+But I have added this in the following line, please check.
+Keep it and resolve the comment if you find this ok.
 }
@@
 I wanted to make the example small but I guess I can add one `Final' statement into our example.
-}
+Don't resolve it until I add it into Figure 2.}
 

### 2022-06-23 syn-translate.tex
Adding syntactical translation.

The implementation described in \cref{sec:coqimpl} can be formalized in our metatheory. The described compilation strategy can be considered as a map from our syntactical (QIIT-)model to a subpart of our syntactical (QIIT-)model \textit{without linkages}. Basically, we can compile away the judgement $\goodSig{\Gamma}{v}{n}$, the type $\cL$ and $\cC$, and related terms, by transforming them into the corresponding sigma type.

Ultimately, we construct a model for our QIIT-syntax using only subparts of our QIIT-syntax -- without these linkages. This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers. This also justify the intuition -- a linkage is actually a module \textit{with universal quantification wrapping its fields to achieve late binding}.

It is also possible to compile away $\goodWSig{\Gamma}{w}{n}$ but the recursor will always require a sigma type as handler and thus the elimination rule for our W-type is still non-conventional.

After syntactical translation, we directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the subpart}. However, due to our non-conventional formulation of the inductive type, it is better we don't impose such assumption and directly give a consistency and canonicity proof.

### 2022-06-23 syn-translate.tex
Type-setting syntactical translation.

@@
 The described compilation strategy can be considered as a map from our syntactical (QIIT-)model to a subpart of our syntactical (QIIT-)model \textit{without linkages}.
+This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
+
 Basically, we can compile away the judgement $\goodSig{\Gamma}{v}{n}$, the type $\cL$ and $\cC$, and related terms, by transforming them into the corresponding sigma type.
 
+\newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
+\newcommand{\Sigr}[2]{{ "Sig"_r~{#1}~{#2} }}
+
+We use $\denotesT{}$ as the translation.
+This translation is invariant under the subpart.
+Since we are in an intrinsic setting, everything here is well-typed (well-formed)including this translation.
+Thus $\denotesS{T}$ means the same as $\denotesS{\goodType{\Gamma}{T}{}}$ (i.e.
+every type is tracking its context).
+In other words, every derivation of any judgement will track its ``context''.
+This will be clearer in the (pseudo-)Agda formalization.
+
+\begin{align*}
+  \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} & \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
+  \denotesT{\goodSig{\Gamma}{\sigma}{n}},&\denotesT{\Sigr{\Gamma}{n}}, \text{ and } \denotesT{\cC} \text{ are mutually recursively defined together, } \\
+  & \text{and they define inductively on the signature length} \\  
+  \denotesT{\goodSig{\Gamma}{\sigma}{n}} &= \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
+  \denotesT{\Sigr{\Gamma}{n+1}} &= \Sigma~(\Sigr{\Gamma}{n})~(\Sigma)
+\end{align*}
+
 Ultimately, we construct a model for our QIIT-syntax using only subparts of our QIIT-syntax -- without these linkages.
-This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
 This also justify the intuition -- a linkage is actually a module \textit{with universal quantification wrapping its fields to achieve late binding}.
@@
 
-After syntactical translation, we directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the subpart}.
+After syntactical translation, we should directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the subpart}.
 However, due to our non-conventional formulation of the inductive type, it is better we don't impose such assumption and directly give a consistency and canonicity proof.

### 2022-06-24 syn-translate.tex
Save syn-translate

@@
 \newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
-\newcommand{\Sigr}[2]{{ "Sig"_r~{#1}~{#2} }}
+\newcommand{\Sigr}[2]{{ "Sig"^r~{#1}~{#2} }}
 
@@
 Since we are in an intrinsic setting, everything here is well-typed (well-formed)including this translation.
-Thus $\denotesS{T}$ means the same as $\denotesS{\goodType{\Gamma}{T}{}}$ (i.e.
+Thus $\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$ (i.e.
 every type is tracking its context).
@@
 \begin{align*}
-  \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} & \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
-  \denotesT{\goodSig{\Gamma}{\sigma}{n}},&\denotesT{\Sigr{\Gamma}{n}}, \text{ and } \denotesT{\cC} \text{ are mutually recursively defined together, } \\
+  \text{We will define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
+  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursively defined together, } \\
   & \text{and they define inductively on the signature length} \\  
-  \denotesT{\goodSig{\Gamma}{\sigma}{n}} &= \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
-  \denotesT{\Sigr{\Gamma}{n+1}} &= \Sigma~(\Sigr{\Gamma}{n})~(\Sigma)
+  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, s.t. } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
+  {\Sigr{\Gamma}{(n+1)}} &= 
+    \Sigma~(\Sigr{\Gamma}{n})
+          ~(\Sigma~\cU
+                  ~(\Sigma~(\Pi(\cC (\pi_2[\pi_1]))(El~\pi_2[\pi_1]))~\cU)) \\
+  \Sigr{\Gamma}{0} &= \top \\
+  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &\text{ defines upon } \denotesT{\sigma} \\
+  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
+    \Sigma~(\denotesT{\cC ("pjl"~\sigma)})~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
+      &\text{when } \goodSig{\Gamma}{\sigma}{n+1} \\
+  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \top \quad
+      \text{when } \goodSig{\Gamma}{\sigma}{0} \\ 
+  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} & \text{ is defined upon } \denotesT{\sigma} \\
+  & \text{ and inductively on the signature length} \\ 
+  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &=
+  \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
+  &\text{when } \goodSig{\Gamma}{\sigma}{n+1} \\
+  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &= \top \quad \text{when } \goodSig{\Gamma}{\sigma}{0} \\
+  \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
+  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\cL(\nu^+~\sigma~T)}} &= (m, "lam"~t) \\ 
+  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} & \text{ is defined upon } \denotesT{m}\\
+  & \text{ and inductively on the signature length} \\ 
+  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} &= 
+  (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
+  & \text{when } \goodSig{\Gamma}{\sigma}{n+1}, \text{ and } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
+  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{when } \goodSig{\Gamma}{\sigma}{0}
 \end{align*}

### 2022-06-24 syntax.tex
Finishing Language Design Section. Try to fix graph embedding and format.
Save syn-translate

@@
 
-\begin{figure}[H]
+\begin{figure}[!htb]
+  \begin{align*}
+    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
+    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
+  \end{align*}
+\caption{Judgements and their Agda Encoding}
+\end{figure}
+
+\begin{figure}[!htb]
   \begin{minipage}[b]{0.3\linewidth}
@@
   \end{minipage}
-\begin{align*}
-  &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
-  &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
-\end{align*}
-\caption{Well-formedness of Four New Judgements, and Agda-encoding-Judgements correspondence}
+
+\caption{Well-formedness of Four New Judgements}
 \end{figure}
@@
 \goodType{\Gamma}{\cC (\nu^+ \ \sigma \ \{f\} \ T) \equiv 
-    \Sigma (\cC \sigma) T[sf \ f]}{}
-\\\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t[sf \ f][(id, \cCt \ o)]) }{}
+    \Sigma (\cC \sigma) T["sf" \ f]}{}
+\\\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}
 }
@@
 
+\begin{figure}[H]
+
   \judgebox{[\goodSeal{\Gamma}{s}{\sigma}{A}] 
@@
   $$
-\begin{figure}[H]
 
@@
 doing exhaustiveness checking when using recursor), and thus the "Seal-Id" is used.
-For notation clearance, we will sometime use $sf'$ to emphasize that we consider that seal $s$ as a term $(sf' \ s)$.
+For notation clearance, we will sometime use $"sf'"$ to emphasize that we consider that seal $s$ as a term $("sf'" \ s)$.
 

### 2022-06-24 syn-translate.tex
Syntactical Translation added. Please check.

@@
   & \text{and they define inductively on the signature length} \\  
-  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, s.t. } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
+  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
   {\Sigr{\Gamma}{(n+1)}} &= 
@@
     \Sigma~(\denotesT{\cC ("pjl"~\sigma)})~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
-      &\text{when } \goodSig{\Gamma}{\sigma}{n+1} \\
+      &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
   \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \top \quad
-      \text{when } \goodSig{\Gamma}{\sigma}{0} \\ 
+      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
   \denotesT{\goodType{\Gamma}{\cL\sigma}{}} & \text{ is defined upon } \denotesT{\sigma} \\
@@
   \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
-  &\text{when } \goodSig{\Gamma}{\sigma}{n+1} \\
-  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &= \top \quad \text{when } \goodSig{\Gamma}{\sigma}{0} \\
+  &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
+  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &= \top \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\
   \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
@@
   (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
-  & \text{when } \goodSig{\Gamma}{\sigma}{n+1}, \text{ and } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
-  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{when } \goodSig{\Gamma}{\sigma}{0}
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
+  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
+\end{align*}
+\begin{align*}
+  &\denotesT{\goodSig{\Gamma}{p_1\nu~\sigma}{n}} = "pjl"~\sigma 
+  &&\denotesT{\goodType{\Gamma}{p_1\nu'~\sigma}{}} = El~("pjl"~("pjr"~\sigma)) \\
+  &\denotesT{\goodSeal{\Gamma}{p_f\nu~\sigma}{p_1\nu~\sigma}{p_1\nu'~\sigma}} = "app"~("pjl"~("pjr"^2~\sigma)) 
+  &&\denotesT{\goodSig{\Gamma}{p_2\nu~\sigma}{n}} = El~("app"~("pjr"^3~\sigma)) \\ 
+  &\denotesT{\goodTerm{\Gamma}{p_1\mu~o}{\cL(p_1\nu\sigma)}} = "pjl"~o 
+  &&\denotesT{\goodTerm{\Gamma, p_1\nu'~\sigma}{p_2\mu'~o}{p_2\nu~\sigma}} = "app"~("pjr"~o)
+\end{align*}
+\begin{align*}
+  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[\pi_1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[\pi_1]} \\
+  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}} &= \mu^+~(h[(\pi_1,\pi_1\mu~\pi_2)])\\ &\quad \quad \{f_2[\pi_1]\}~(p_2\mu'~\pi_2)[(\pi_1, \uparrow^s[\pi_1^{\uparrow}])] \\
+  \denotesT{\goodInh{\Gamma}{"inhov"~h~t}{(\nu^+ \  \sigma_1 \  T_1)}{(\nu^+ \  \sigma_2\  T_2)}} &= \mu^+ (h[(\pi_1, p_1\mu~\pi_2)])~\{f_2[\pi_1]\}~t[\pi_1^{\uparrow}]
 \end{align*}
 
-Ultimately, we construct a model for our QIIT-syntax using only subparts of our QIIT-syntax -- without these linkages.
+The main idea of this syntactical translation is that, we make sure $\goodSig{\Gamma}{\nu^+ \_ }{n+1}$ and $\goodSig{\Gamma}{\mu^+ \_}{n+1}$ wrapped into a dependent pair.
+These two definitions decide the syntactical translation for $\goodSig{\Gamma}{\_}{n}$ and $\cL$.
+The syntactical translation for $\cC$ and $\cCt$ are decided by their $\beta$-rules.
+
+We use the local morphism (the same one for encoding the Sealing judgements) to translate inheritance judgement.
+This is mainly due to the existence of \ruleref{Inh} rule.
+Here we only typeset the overriding and inheritance rule, but all of the translation for inh judgment are induced by the $\beta$-rule of \ruleref{Inh}.
+
+We omit all the equational rules ($\beta,\eta$ and substitution) here.
+However, for coherence, when we mutual recursively define $\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and $\denotesT{\cC}$ above, we actually have to prove the substitution laws $(\cC\sigma)[\tau]\equiv \cC(\sigma[\tau])$ and $\Sigr{\Delta}{n}[\tau] \equiv \Sigr{\Gamma}{n}$ together.
+The details (of all the equational rules) can be found in the appendices written in (pseudo-)Agda form.
+
+Ultimately, we have constructed a model for our QIIT-syntax using only subparts of our QIIT-syntax -- without these linkages.
 This also justify the intuition -- a linkage is actually a module \textit{with universal quantification wrapping its fields to achieve late binding}.
@@
 
-After syntactical translation, we should directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the subpart}.
+After syntactical translation, we can directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the subpart}.
 However, due to our non-conventional formulation of the inductive type, it is better we don't impose such assumption and directly give a consistency and canonicity proof.

### 2022-06-24 lang-design.tex
Save Language Design.
Adding "accessing family components outside that family" into lang-design section Slight change on syntactic translation section.

@@
 
-First and foremost, \textlabel{Req (0)}{langdesign-req0}\textbf{we desire incremental(modular) type checking}.
+First and foremost, we don't want to throw away some basic good properties from Coq.
+\textlabel{Req (0A)}{langdesign-req0}~\textbf{we want to retain the incremental(modular) type checking}.
 Notice that, the modularity here in Coq is a bit different from other language, because Coq is \textit{interactive} theorem proving so we actually need \textit{statement-wise} incremental type checking, not only for saving the compilation time, but also for the immediate feedback and type-checking for the Coq developer.
+We don't want our family to ruin this conventional routine of interactive development.
+\textlabel{Req (0B)}{langdesign-req0b}~\textbf{we also want to keep the computational property of Coq when use family}.
+Coq, as a constructive logic, has been famous for its ability to do calculations and extract programs.
+We don't want our Family facility to ruin this: the developers should be able to project field as first-class value just like Coq's Module.
 
@@
 Take \cref{fig:plugin-example2} as example, if we want to extend boolean inductive type and its negation, we only need to specify the new constructor for boolean and how negation deal with the new constructor.
+This support provides the user a way to waive the necessity for duplicating code -- otherwise the user will need to copy and paste the whole original negation definition and add a new pattern matching cases.

### 2022-06-24 coqimpl.tex
Adding "accessing family components outside that family" into lang-design section Slight change on syntactic translation section.
took a pass on lang-design and the early-going of coqimpl

@@
 Please comment back so that I know what corresponding modification is needed on the paper. 
-}
+}\YZreply{My original question was more directed towards the fact
+that currently type checking is deferred to the Coq base. Generally,
+when one uses a statically typed language as a compilation target, they
+also want to have type checking at the source level, (1) because there
+could be type errors at the source level that could not otherwise be
+caught at the target level and (2) because it allows better error messages.
+So it needs to be addressed why deferring type checking to Coq is OK.}
 
 \textbf{Family compiled into module.} This is the main idea of compiling
-family components into Coq's primitives: we compile family into Coq
-module, family type into module type, and the context of a given term
+family components into Coq's primitives: we compile families into Coq
+modules, family types into module types, and the context of a given term
 into parameters of the module. Unlike module, we need to faithfully
-reflect the \textit{late binding} nature of family: for each defining
+reflect the \textit{late binding} nature of families: for each defining
 field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
@@
 I think altering example is fine.
-If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}
+If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}
   \begin{minipage}[t]{0.4\linewidth}
@@
 translate this statement into the "Module a_4", "Module Type a_5, a_6"
-three components.\YZ{Typo? Are they a_4, a_5, a_6 in Figure 2?}\EDJreply{Thanks, fixed.}
+three components.
 {The "EmptySig" is just empty module type.} Then,
 these modules and module types will be generated and type-checked by Coq
-immediately. The computed "Module a_10" is then part of the context of
-"Field b" (as the module type of "self__A") for "b" to refer to. The
-internal representation of family "A" is simply the list of these
-``compiled'' modules and module types. We can simply achieve incremental
+immediately. The compiled "Module a_10"\YZ{should this be a_4?} is then part of the context of
+"Field b" (as the module type of "self_A") for "b" to refer to. The
+internal representation of family "A" is the list of these
+compiled modules and module types. We can achieve incremental
 checking and overriding as in \ref{langdesign-req0}---for example, if we
@@
   \begin{minipage}[t]{0.3\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
+\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
 Family B.

### 2022-06-24 lang-design.tex
took a pass on lang-design and the early-going of coqimpl

--- 2022-06-23 lang-design.tex
@@
 
-\YZ{looks more like a 'design space' rather than 'language design' section}\EDJreply{Good point. Do I just change the section name?s}
 When adapting Family Polymorphism into dependent types, we choose to
@@
 need it to have good compatibility with inductive types, because we
-don't want to lose too much reasoning power of Coq (and its tactic
-programming).
+don't want to lose too much reasoning power of Coq\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}
+(and its tactic programming).
 
@@
   \begin{minipage}[t]{0.47\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Family STLC.
+\begin{minted}[fontsize=\footnotesize,escapeinside=||]{Coq}
+Family STLC.|\YZ{Does minted support adding extra keywords such as 'Family' and 'extends'?}|
 Inductive ty : Set :=
@@
 
-First and foremost, \textlabel{Req (0)}{langdesign-req0}\textbf{we desire incremental(modular) type checking}.
-Notice that, the modularity here in Coq is a bit different from other language, because Coq is \textit{interactive} theorem proving so we actually need \textit{statement-wise} incremental type checking, not only for saving the compilation time, but also for the immediate feedback and type-checking for the Coq developer.
+First and foremost, \textlabel{Req (0)}{langdesign-req0}: \textbf{we
+desire incremental (modular) type checking}. Notice that, the modularity
+here in Coq is a bit different from other languages, because Coq
+supports \textit{interactive} theorem proving, so we actually need
+\textit{statement-wise} incremental type checking, not only for avoiding
+re\-compilation, but also for the immediate feedback and
+type-checking for the Coq developer.
 
@@
 has_type Γ u A ->
-has_type Γ (subst i x tt) T. But this final goal requires to know the computation of subst.}
+has_type Γ (subst i x tt) T. But this final goal requires to know the
+computation of subst.}\YZreply{Seems like what's missing here is to
+convince the reader why exposing the definition of subst could be a
+problem for family polymorphism.}
 
@@
 \subsection{Language Design}
-Inspired by our above requirements and Family Polymorphism, we propose the following extended syntax for Coq Vernacular Commands. 
+Inspired by our above requirements and Family Polymorphism, we propose
+the following extension to Coq's Vernacular Commands.
 % basically write down the specification of surface Coq commandline
@@
 We use "Family" and "EndFamily" to scope.
-Each field in one family is declared and defined by "Field".
-We can define nested family.
-Family will also be used to encode recursive handlers.
-Each case handling of one inductive type will be aggregated into one family, then "FRecursor" will be used to construct a recursor using these handlers.
-For example, \mintinline{Coq}{FRecursor recur about T motive f} "using" "handlers" will generate a new field "recur" that is doing induction (using elimination principle of "T") using motive "f", and the handlers are located in the family "handlers".
+Each field in a family is declared and defined as a "Field".\YZ{This sentence sounds a bit recursive}
+We can define nested families.
+Family will also be used to encode recursive handlers. Each case
+handling of one inductive type will be aggregated into one family, then
+"FRecursor" will be used to construct a recursor using these handlers.
+For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
+"handlers" will generate a new field "recur" that is doing induction
+(using elimination principle of "T") using motive "f", and the handlers
+are located in the family "handlers".
 
-We can inherit a top level family "parent" using \mintinline{Coq}{Family ...
-extends parent} as a new top level family.
-As expected, "Override" will override/a field from the parent.
-"Inherit" can inherit nested family, inductive type and fields.
-The fields inside a family are order-sensitive because we have dependent type here.
-So "Inherit Until" is used if the user want to inherit a range of fields.
-If the user wants to extend an inductive type or a nested family from the parent, then "Extend FInductive" or "Extend Family" is used.
-Thus nested family can only be inherit or extended together with its outer family (and thus top level family), thus our support of family polymorphism is incomplete.
+We can inherit a top-level family "parent" using \mintinline{Coq}{Family
+... extends parent} as a new top-level family. As expected, "Override"
+will override a field from the parent. "Inherit" can be used to inherit
+nested families, inductive types, and fields. The fields inside a family
+are order-sensitive because we have dependent types here. So "Inherit
+Until" is used if the user wants to inherit a range of fields. If the
+user wants to extend an inductive type or a nested family from the
+parent, then "Extend FInductive" or "Extend Family" is used. Thus nested
+family can only be inherited or extended together with its outer family
+(and thus top level family), thus our support of family polymorphism is
+incomplete.
 
-This incomplete support can already be helpful for extending functions.
-Take \cref{fig:plugin-example2} as example, if we want to extend boolean inductive type and its negation, we only need to specify the new constructor for boolean and how negation deal with the new constructor.
+This incomplete support can already be useful for extending functions.
+Take \cref{fig:plugin-example2} as example, if we want to extend boolean
+inductive type and its negation, we only need to specify the new
+constructor for boolean and how negation deals with the new constructor.
+As a result, the programmer does not have to duplicate code---otherwise
+the user will need to copy and paste the whole original negation
+definition and add a new pattern matching case.

### 2022-06-24 lang-design.tex

@@
 
-First and foremost, \textlabel{Req (0)}{langdesign-req0}: \textbf{we
-desire incremental (modular) type checking}. Notice that, the modularity
-here in Coq is a bit different from other languages, because Coq
-supports \textit{interactive} theorem proving, so we actually need
+First and foremost, we don't want to throw away some basic good
+properties of Coq. \textlabel{Req~(0A)}{langdesign-req0}~\textbf{we
+want to retain incremental (modular) type checking}. Notice that, the
+modularity here in Coq is a bit different from other languages, because
+Coq supports \textit{interactive} theorem proving, so we actually need
 \textit{statement-wise} incremental type checking, not only for avoiding
-re\-compilation, but also for the immediate feedback and
-type-checking for the Coq developer.
+re\-compilation, but also to enable immediate feedback and incremental
+type-checking for the Coq developer. We don't want our family to ruin
+this conventional routine of interactive development.
+\textlabel{Req~(0B)}{langdesign-req0b}~\textbf{we also want to keep the
+computational property of Coq when using families}. Coq, as a constructive
+logic, has been famous for its ability to do calculations and extract
+programs. We don't want our Family facility to ruin this: the developers
+should be able to project field as first-class value just like Coq's
+Module.
 

### 2022-06-24 coq-example.tex
Adding more info for "accessing family field"

@@
 
-\subsection{Type Safety for STLC}
+\subsection{Type Safety for STLC}\EDJ{We need a fuller view of the two examples. Especially show "accessing family components outside that family".}
 We start with the very basic type safety proof of STLC, largely adapted from Software Foundation~\cite{pierce2014software}, which is also one of the primal motivating example of this project.

### 2022-06-24 lang-design.tex
Adding more info for "accessing family field"

--- 2022-06-24 lang-design.tex
@@
 Coq, as a constructive logic, has been famous for its ability to do calculations and extract programs.
-We don't want our Family facility to ruin this: the developers should be able to project field as first-class value just like Coq's Module.
+We don't want our Family facility to ruin this: the developers should be able to project field from families as first-class value just like how we can treat fields of Coq's Module as first-class values.
 

### 2022-06-24 syntactic-example.tex
Adding more info for "accessing family field"

@@
 
+The resulting constructed families are $\goodTerm{\cdot}{}{}$
+
 Here we explain how  our detailed construction in Figure \cref{fig:example-construction} is constructed.
@@
 After that, $\uparrow^s_2$ again witnesses the compatibility of the abstract interface, and the two are both used for constructing $"inh"_6$, which is responsible for inheriting field "k".
+
+Finally we make some comments on how a family is encoded in our metatheory.
+A family object will have a signature that expose the type of fields and the details of inductive type .
+This exposure can make sure inheritance can override and inherit field correctly.
+Sealing is only hiding information \textbf{inside one family}, i.e.
+hiding information of the former fields for the later fields, and thus the later fields is defined on the abstract interface of the former fields.
+The signature will expose the sealing itself and the inductive type.
+(See $"sig"_5$ has all the sealing information and inductive definition).
+A former sealing might not be related to any later sealing, for example, $"sig"_3$ is constructed using $"sl"_1$ but $"sig"_4$ is using trivial sealing $id_s$.

### 2022-06-24 lang-design.tex
Rephrase Wording.

--- 2022-06-24 lang-design.tex
@@
 \textlabel{Req~(0B)}{langdesign-req0b}~\textbf{we also want to keep the
-computational property of Coq when using families}. Coq, as a constructive
-logic, has been famous for its ability to do calculations and extract
-programs. We don't want our Family facility to ruin this: the developers
-should be able to project field as first-class value just like Coq's
-Module.
+computational ability of Coq when using families}. Coq, based on constructive
+logic, can consider proof as programs. 
+We don't want our Family facility to ruin this when incorporated with 
+the remaining parts of Coq: 
+the developers should be able to project fields of families 
+as first-class value just like how they can project fields of Coq's Module.
 

### 2022-06-24 coqimpl.tex
Reply part of the comments.

@@
 caught at the target level and (2) because it allows better error messages.
-So it needs to be addressed why deferring type checking to Coq is OK.}
+So it needs to be addressed why deferring type checking to Coq is OK.}\EDJreply{Good point.
+I think (1) is largely mediated by the fact the semantic of family is very close to that of the module.
+And the supporting feature is simple enough that there won't appear some problems like type error at source-level but not target level(2) is happening for sure but some basic error message is almost "transparent" i.e.
+the error-messages from Coq can be directly interpreted for the user.
+But my justification is solely empirical and as you know the empirical study for this paper is not really strong enough to support anything.}
 
@@
 Field b 
-  : self_A.a = self_A.a
+  : a = a
   := eq_refl.
@@
 Field b'
-  : self_A.a' = 1
+  : a' = 1
   := eq_refl.
@@
 I think altering example is fine.
-If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}
+If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}\EDJreply{Omitting self_ is easy and not much engineering effort is required.
+But omitting self_ will lead to `A.a' instead of `a' actually.
+So if we really want `a' as you last time mentioned, then a resolution is required and more engineering effort is required.
+But I don't see any technical difficulties.
+I have removed all the self, please check.}
   \begin{minipage}[t]{0.4\linewidth}
@@
 these modules and module types will be generated and type-checked by Coq
-immediately. The compiled "Module a_10"\YZ{should this be a_4?} is then part of the context of
+immediately. The compiled "Module a_4"\YZ{should this be a_4?}\EDJreply{Yes. Thanks!} is then part of the context of
 "Field b" (as the module type of "self_A") for "b" to refer to. The
@@
 This is satisfied by the \mintinline{Coq}{a : nat} that is included earlier.
-The resulting module is just like mundane Coq's module and thus
-enjoy all the computational mechanism from Coq for free.
+The resulting compiled module is just like mundane Coq's module and thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.
 
@@
 therefore we need non-overridables. This causal flow should be made
-more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph.}
+more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph. Please check and resolve if find satisfied.}
 

### 2022-06-24 lang-design.tex
Reply part of the comments.

@@
 Interface, will be removed. In this case, it will look like module with
-late-binding and inheritance as code and proof reuse for module. We also
+late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
 need it to have good compatibility with inductive types, because we
-don't want to lose too much reasoning power of Coq\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}
+don't want to diverge too much from the routine Coq programming experience\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness.
+I rephrase it now, please check.}
 (and its tactic programming).
@@
 
-We use "Family" and "EndFamily" to scope.
-Each field in a family is declared and defined as a "Field".\YZ{This sentence sounds a bit recursive}
+We use "Family" and "EndFamily" to scope, and "Field" is used to declare new field members in family.\YZ{This sentence sounds a bit recursive}\EDJreply{What about now?}
 We can define nested families.

### 2022-06-24 syntax.tex
Syntactical Translation added. Please check.
Reply part of the comments.

@@
 function type and dependent pair type.
+The subsitution law for these two require \textit{de bruijn lifting}: for $\goodTerm{\Gamma}{\sigma}{\Delta}$, we define its de bruijn lifting $\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow}{\Delta, A}$ given $\goodType{\Delta}{A}{}$.
 
@@
 
-\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{Working on it, don't resolve}
+\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{it is done in the example section, please check.}
 

### 2022-06-25 lang-design.tex
resolve comments and add new comments

@@
 need it to have good compatibility with inductive types, because we
-don't want to diverge too much from the routine Coq programming experience\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness.
+don't want to retain the idiomatic Coq programming experience
+%\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness.
 I rephrase it now, please check.}
@@
 
-We use "Family" and "EndFamily" to scope, and "Field" is used to declare new field members in family.\YZ{This sentence sounds a bit recursive}\EDJreply{What about now?}
+We use "Family" and "EndFamily" to scope, and "Field" is used to declare new members in a family.
 We can define nested families.

### 2022-06-25 lang-design.tex
Save comments from discussion.

--- 2022-06-24 lang-design.tex
@@
-\subsection{Design Space}
+\subsection{Design Space}\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).
+I think most readers will benefit from a recap of this encoding.\\
+Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
+}
 

### 2022-06-25 coqimpl.tex

@@
   := 1.
-Field b 
-  : a = a
+Field b : a = a
   := eq_refl.
@@
   := 1. 
-Field b'
-  : a' = 1
+Field b' : a' = 1
   := eq_refl.
@@
 \end{minted}
-  \end{minipage}\YZ{Can the self's be omitted?}\EDJreply{Yes.
+  \end{minipage}%\YZ{Can the self's be omitted?}\EDJreply{Yes.
 But in our current plugin implementation it is not omitted.
@@
 these modules and module types will be generated and type-checked by Coq
-immediately. The compiled "Module a_4"\YZ{should this be a_4?}\EDJreply{Yes. Thanks!} is then part of the context of
+immediately. The compiled "Module a_4" is then part of the context of
 "Field b" (as the module type of "self_A") for "b" to refer to. The
@@
 
-Now that we know a family data structure is actually a list of compiled
-modules, we may want to aggregate them into one complete compiled
-module. We achieve that by repeatedly using Coq's "Include" command as
-the example illustrated. Note that, in our example, we include two
-functors without arguments---this is due to a perculiarity of Coq's
+Now that we know a family declaration is compiled to a list of Coq
+modules and module types, we may want to aggregate them into one complete compiled
+module.\YZ{The reader wonders why this aggregation is needed. Seems like the place to talk about how the programmer can use A outside Family A.}
+We achieve that by repeatedly using Coq's "Include" command as
+the example illustrates. Note that, in our example, we include two
+functors without arguments---this is due to a peculiarity of Coq's
 "Include": non-instantiated parameters will be automatically
 instantiated by the appropriate fields in the surrounding defining
-module.\YZ{'For example, ...'}\EDJreply{I don't think explaining a Coq feature inside our paper is a good idea?
-But I have added this in the following line, please check.
-Keep it and resolve the comment if you find this ok.
-}
-For example, "b_7" requires a module of type "a_6".
-Inside module "A", when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat} inside the current surrounding "A".
-Then when we \mintinline{Coq}{Include b_7} without specifying the module parameters, Coq will try to find fields in the surrounding module, (i.e.
-module "A") to instantiate \mintinline{Coq}{self_A: a_6}.
-This is satisfied by the \mintinline{Coq}{a : nat} that is included earlier.
+module.
+For example, "b_7" requires a module of type "a_6". Inside module "A",
+when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat}
+inside the current surrounding "A". Then when we
+\mintinline{Coq}{Include b_7} without specifying the module parameter,
+Coq will try to find fields in the surrounding module (i.e., module
+"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
+the \mintinline{Coq}{a : nat} that is included earlier.
 The resulting compiled module is just like mundane Coq's module and thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.
@@
 Besides standalone family, we also need to implement family inheritance.
-To simplify the implementation, we can consider all families as inheritance and the standalone families are extending empty family.
-Thus we only need to deal with inheritance during interactive proving.
+To simplify the implementation, we can consider all families as
+inheritance and the standalone families are extending the empty family.
+Thus we need to deal with only inheritance during interactive proving.
 
-An inheritance data will be annotated with the type for the input family
-and output family, and generally there are three kinds of
-inheritance---extend, override and inherit. Extending corresponds to
-adding new fields, overriding corresponds to modifying existent fields,
-and inheriting corresponds to directly taking fields from the parent.
-Thus, it is possible to consider inheritance as solely a data rather than
-a functor, and this actually provides us the possibility to
+An inheritance data\YZ{what is 'inheritance data'?} will be annotated with the type for the input family
+and output family. There are three kinds of inheritance---extend,
+override, and inherit. Extending corresponds to adding new fields,
+overriding corresponds to modifying existent fields, and inheriting
+corresponds to directly taking fields from the parent.
+Thus, it is possible to consider inheritance as solely a data rather
+than a functor, and this actually provides us the possibility to
 \text{mix-in} inheritance data.
@@
 can allow fields to be type-dependent on former fields thus fulfilling
-\ref{langdesign-req1}. However, there are some cases we need to expose the concrete definition of a given fields (especially when we are proving properties about it).
+\ref{langdesign-req1}. However, there are cases where we need to expose
+the concrete definition of a given field (especially when we are proving
+properties about it).
 
@@
 future, once "add" is overridden into a subtraction function then
-"comm" will have problem to be inherited.
+"comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
 A more concrete example is in \cref{fig:plugin-example1}, consider the
-case we have "Field b : self_A.a = 1 := .." instead. Then our
+case we have "Field b : a = 1 := .." instead. Then our
 compilation and Coq's type checking will both fail: in the context of
@@
 more (and this is necessary so that "a" can be overridden in the
-future).
+future).\YZ{Not clear if you want to be able to override a or want to be able to have 'Field b : a = 1'.}
 
@@
 need to expose the concrete definition "a = 1" to the context of "b".
-\YZ{I had a hard time working through these two paras. I guess they are
-saying that there is a need to expose definitional equalities, and
-therefore we need non-overridables. This causal flow should be made
-more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph. Please check and resolve if find satisfied.}
+%\YZ{I had a hard time working through these two paras. I guess they are
+%saying that there is a need to expose definitional equalities, and
+%therefore we need non-overridables. This causal flow should be made
+%more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph. Please check and resolve if find satisfied.}
 
-In our Coq plugin, we can achieve such exposure of definition by
+In our Coq plugin, we can achieve such exposure of definitional equality by
 exposing the whole definition directly into the compiled module type. Doing so we also prohibit overriding because

### 2022-06-25 coqimpl.tex
more comments

@@
 
-\textbf{Reasoning Requires Non-Overridable.} This compilation strategy
+\textbf{Reasoning Requires Non-Overridable.}\YZ{At this point in this section, overriding has not been discussed yet, but it is already discussing non-overridables.
+Should probably discuss the extensibility first.}
+This compilation strategy
 can allow fields to be type-dependent on former fields thus fulfilling
@@
 \textbf{Special Fields: Overridable and Sealed Family as Interface.} 
-Non-overridable field is still not expressive enough to achieve the full extent of the dependent type programming.
-There are cases reasoning and overriding are both required, especially when we need to use Coq to do computation.
-For example, we may want a field to be an arbitrary monotonic function.
-Again, we cannot split into two fields \mintinline{Coq}{(f : nat → nat; p : monotonic f)} because to provide "p" we need to have concrete definition of "f" in the family type, which makes "f" non-overridable at all.
-One option is to use sigma type \mintinline{Coq}{{f | monotonic f}} but that is cumbersome when we have a bundle of stuff with multiple refinements.
+Non-overridable fields are still not expressive enough to achieve the full
+extent of dependent type programming. There are cases overriding and
+equational reasoning are both needed, especially when we use Coq to do
+computation. For example, we may want a field to be an overridable 
+monotonic function. We cannot split into two fields
+\mintinline{Coq}{(f : nat → nat; p : monotonic f)}\YZ{Is this semicolon syntax valid Coq syntax?} because to provide
+"p" we need to have concrete definition of "f" in the family type, which
+makes "f" non-overridable at all. One option is to use sigma type
+\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
+a bundle of stuff with multiple refinements.
 
-To resolve this, we provide family sealing mechanism---basically we allow the users to specify the family/compiled module type, and thus we can have a \textit{locally transparent}, \textit{non-overridable} field located in a \textit{overridable} family.
-Using the above example, we simply make \mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an \textit{overridable} family, but "f" to be non-overridable and transparent on its definition, and thus we can prove "p".
-Once we want to swap the implementation, we just override the whole ``sealed'' family.
+To resolve this, we provide family sealing mechanism---we allow the
+users to specify the family/compiled module type\YZ{I don't understand "specify the family/compiled module type"}, and thus we can have a
+\textit{locally transparent}, \textit{non-overridable} field located in
+an \textit{overridable} family. Using the above example, we simply make
+\mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an
+\textit{overridable} family, but "f" to be non-overridable and
+transparent on its definition, and thus we can prove "p". Once we want
+to swap the implementation of "f", we just override the whole ``sealed''
+family.
 
-This ``sealed'' family actually simulate interface to a certain level---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.
+``Sealed'' families actually simulate interfaces in certain sense\YZ{I was actually thinking sealed families simulate sigma types, no?}---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.\YZ{I don't understand what it means to "assign each field with an axiom".}
 % Basically, {a : int -> int; property : a is monotonic}, in this case

### 2022-06-25 coqimpl.tex
Reply comments.

--- 2022-06-25 coqimpl.tex
@@
 
-An inheritance data\YZ{what is 'inheritance data'?} will be annotated with the type for the input family
+An inheritance data\YZ{what is 'inheritance data'?}\EDJreply{It is just inheritance judgement, but before meta-theory is introduced, I can only say some "first-order data that represent the information of inheritance".
+I emphasize first order to distinguish from functor in ML community.}
+will be annotated with the type for the input family
 and output family. There are three kinds of inheritance---extend,

### 2022-06-25 coqimpl.tex
Reply more comments.

@@
 monotonic function. We cannot split into two fields
-\mintinline{Coq}{(f : nat → nat; p : monotonic f)}\YZ{Is this semicolon syntax valid Coq syntax?} because to provide
+\mintinline{Coq}{(f : nat → nat; p : monotonic f)}\YZ{Is this semicolon syntax valid Coq syntax?}\EDJreply{No, this part is just pseudo-code for two consecutive fields.
+I have no good idea on what to use other than semicolon, apparently I cannot use comma.} because to provide
 "p" we need to have concrete definition of "f" in the family type, which
@@
 To resolve this, we provide family sealing mechanism---we allow the
-users to specify the family/compiled module type\YZ{I don't understand "specify the family/compiled module type"}, and thus we can have a
+users to specify the compiled module type of a given family\YZ{I don't understand "specify the family/compiled module type"}\EDJreply{Sorry.
+I rephase it into "compiled module type of a given family".}, and thus we can have a
 \textit{locally transparent}, \textit{non-overridable} field located in
@@
 
-``Sealed'' families actually simulate interfaces in certain sense\YZ{I was actually thinking sealed families simulate sigma types, no?}---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.\YZ{I don't understand what it means to "assign each field with an axiom".}
+``Sealed'' families can actually simulate interfaces in certain sense\YZ{I was actually thinking sealed families simulate sigma types, no?}\EDJreply{You are absolutely right.
+I should add a "can" here.
+I can use sealed family to simulate interface is by creating a sigma type, but making the impleemntation of sigma type trivial (using place holder/using axiom as I specified below).
+Then once I want to instantiate with real implementation, I override the whole family -- as long as it has the same type as the sealed family, everything still works.}---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.\YZ{I don't understand what it means to "assign each field with an axiom".}\EDJreply{It really means ``place holder'' in general programming language (Python's NonImplement()), defer to future.
+Because Coq's Axiom can be used to instantiate anything.}
 % Basically, {a : int -> int; property : a is monotonic}, in this case

### 2022-06-25 coqimpl.tex
edits and replies to comments

--- 2022-06-25 coqimpl.tex
@@
 An inheritance data\YZ{what is 'inheritance data'?}\EDJreply{It is just inheritance judgement, but before meta-theory is introduced, I can only say some "first-order data that represent the information of inheritance".
-I emphasize first order to distinguish from functor in ML community.}
+I emphasize first order to distinguish from functor in ML community.}\YZreply{I think this is the place where you can use an example to demonstrate these three kinds of inheritance.}
 will be annotated with the type for the input family
@@
 Non-overridable fields are still not expressive enough to achieve the full
-extent of dependent type programming. There are cases overriding and
+extent of dependent type programming. There are cases where overriding and
 equational reasoning are both needed, especially when we use Coq to do
@@
 monotonic function. We cannot split into two fields
-\mintinline{Coq}{(f : nat → nat; p : monotonic f)}\YZ{Is this semicolon syntax valid Coq syntax?} because to provide
+\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
+because to provide
 "p" we need to have concrete definition of "f" in the family type, which
@@
 To resolve this, we provide family sealing mechanism---we allow the
-users to specify the family/compiled module type\YZ{I don't understand "specify the family/compiled module type"}, and thus we can have a
+users to specify the compiled module type of a given family,\YZ{I don't understand "specify the family/compiled module type"}\EDJreply{Sorry.
+I rephase it into "compiled module type of a given family".}\YZreply{Still cannot see what is going on and how it is related to the sentence that follows.
+An example is due.}
+and thus we can have a
 \textit{locally transparent}, \textit{non-overridable} field located in
-an \textit{overridable} family. Using the above example, we simply make
-\mintinline{Coq}{(f : nat → nat, p : monotonic f)} together inside an
-\textit{overridable} family, but "f" to be non-overridable and
-transparent on its definition, and thus we can prove "p". Once we want
-to swap the implementation of "f", we just override the whole ``sealed''
-family.
+an \textit{overridable} family. Using the above example, we simply seal
+\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
+together inside an overridable family that does not allow "f"
+to be overridden in any of its derived families.
+Because "f" is non-overridable inside the family, its definition is
+transparent and can be used to prove "p".
+However, because the enclosing family of "f" is overridable, to swap the
+implementation of "f", we just override the whole ``sealed'' family.
 
-``Sealed'' families actually simulate interfaces in certain sense\YZ{I was actually thinking sealed families simulate sigma types, no?}---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.\YZ{I don't understand what it means to "assign each field with an axiom".}
+``Sealed'' families can actually simulate interfaces in certain sense\YZ{I was actually thinking sealed families simulate sigma types, no?}\EDJreply{You are absolutely right.
+I should add a "can" here.
+I can use sealed family to simulate interface is by creating a sigma type, but making the implemntation of sigma type trivial (using place holder/using axiom as I specified below).
+Then once I want to instantiate with real implementation, I override the whole family -- as long as it has the same type as the sealed family, everything still works.}\YZreply{Can you explain this with an example?
+I recall the AbsInt example uses it?}---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.\YZ{I don't understand what it means to "assign each field with an axiom".}\EDJreply{It really means ``place holder'' in general programming language (Python's NonImplement()), defer to future.
+Because Coq's Axiom can be used to instantiate anything.}
 % Basically, {a : int -> int; property : a is monotonic}, in this case

### 2022-06-25 coqimpl.tex
Adding example for CoqImpl.

@@
   := 1. 
-Field b' : a' = 1
-  := eq_refl.
 EndFamily.
@@
 three components.
-{The "EmptySig" is just empty module type.} Then,
+{The "EmptySig" is just an empty module type.} Then,
 these modules and module types will be generated and type-checked by Coq
@@
 
-Now that we know a family declaration is compiled to a list of Coq
-modules and module types, we may want to aggregate them into one complete compiled
-module.\YZ{The reader wonders why this aggregation is needed. Seems like the place to talk about how the programmer can use A outside Family A.}
-We achieve that by repeatedly using Coq's "Include" command as
-the example illustrates. Note that, in our example, we include two
+Once the user closes a family with "EndFamily", they can access its fields by using module syntax.
+For example, they can access "A.a" to get "1" after "Family A" is defined.
+This is achieved by a complete compilation from a family to a module, via aggregating the list of the modules in the internal representation: the plugin will repeatedly using Coq's "Include" command as the "Module A" in \cref{fig:plugin-example1} illustrates.
+Now, the user can use "Module A" anywhere as they are using "Family A".
+\YZ{The reader wonders why this aggregation is needed. 
+Seems like the place to talk about how the programmer can use A outside Family A.}
+
+Note that, in our example, we include two
 functors without arguments---this is due to a peculiarity of Coq's
@@
 the \mintinline{Coq}{a : nat} that is included earlier.
-The resulting compiled module is just like mundane Coq's module and thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.
+The resulting compiled module is just like mundane Coq's module and 
+thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.
+
+\begin{figure}[!htb]
+  \begin{minipage}[t]{0.45\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family I.
+Field a : nat := 1.
+Field b : a = a := eq_refl.
+Fail Field c : a = 1 := eq_refl. (*Fail 1*)
+
+Final Field a' : nat := 1.
+Field c : a = 1 := eq_refl.
+Field add : nat -> nat -> nat := (+).
+Fail Field comm_add : ...
+
+Family MonotonicF.
+  Final Field f : nat -> nat := non_impl.
+  Field p : monotonic f := non_impl.
+SealFamily.
+
+EndFamily.
+\end{minted}
+  \end{minipage}
+\begin{minipage}[t]{0.45\linewidth}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Family I2 extends I.
+Override Field a : nat := 1. (* Override *)
+Inherit b.                   (* Inherit  *)
+
+Field d : nat := a.          (* Extend   *)
+Fail Override Field a' : nat := 2. (*Fail 2*)
+
+Override Field add := (-).
+
+Inherits Until MonotonicF.
+Override Family MonotonicF.
+  Final Field f : nat := (+ 1).
+  Field p : monotonic f := ...
+SealFamily. 
+EndFamily.
+\end{minted}
+  \end{minipage}
+  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
+\end{figure}
 
@@
 inheritance and the standalone families are extending the empty family.
-Thus we need to deal with only inheritance during interactive proving.
+Thus we only need to deal with inheritance during interactive theorem proving.
 
-An inheritance data\YZ{what is 'inheritance data'?}\EDJreply{It is just inheritance judgement, but before meta-theory is introduced, I can only say some "first-order data that represent the information of inheritance".
-I emphasize first order to distinguish from functor in ML community.}\YZreply{I think this is the place where you can use an example to demonstrate these three kinds of inheritance.}
-will be annotated with the type for the input family
-and output family. There are three kinds of inheritance---extend,
-override, and inherit. Extending corresponds to adding new fields,
-overriding corresponds to modifying existent fields, and inheriting
-corresponds to directly taking fields from the parent.
-Thus, it is possible to consider inheritance as solely a data rather
-than a functor, and this actually provides us the possibility to
-\text{mix-in} inheritance data.
+There are only three kinds of inheritance we are dealing with---extend, override, and inherit.
+We use first-order data to encode inheritance data inside the implementation of the plugin.
+The inheritance data will include the information of (1) the ``type'' of the parent family, (2) a list of operation indicating how each field of the parent will be dealt with (either inherited or overrided), (3) the operations indicating newly extended fields, and (4) the ``type'' of the children family.\YZ{what is 'inheritance data'?}\EDJreply{It is just inheritance judgement, but before meta-theory is introduced, I can only say some "first-order data that represent the information of inheritance".
+I emphasize first order to distinguish from functor in ML community.}\YZreply{I think this is the place where you can use an example to demonstrate these three kinds of inheritance.} The surface syntax for the three kinds of inheritance are demonstrated in \cref*{fig:plugin-example3}.
+The implementation of overriding, inheritance, and extension is achieved by simply manipulating the internal representation of a family correspondingly---we can swap the module in the list for overriding, retain the module in the list for inheritance and add new module into the list for extension.
+The compilation from a family (internal representation) to a module still acts the same.
+
+Since it is possible to consider inheritance as solely a first-order data rather
+than a higher-order functor, it is actually plausible to
+\text{mix-in} two inheritance data -- by carefully ``mix'' the internal list.
 
@@
 This compilation strategy
-can allow fields to be type-dependent on former fields thus fulfilling
-\ref{langdesign-req1}. However, there are cases where we need to expose
+has already allowed fields to be type-dependent on former fields thus fulfilling
+\ref{langdesign-req1}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
 the concrete definition of a given field (especially when we are proving
@@
 
-For example, we may
-define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
-and intend to prove its commutativity in a following field $"comm"$.
-But to prove so, we need to know more than just a type interface of
-"add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
-future, once "add" is overridden into a subtraction function then
-"comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
-A more concrete example is in \cref{fig:plugin-example1}, consider the
-case we have "Field b : a = 1 := .." instead. Then our
-compilation and Coq's type checking will both fail: in the context of
-"b" (module type "a_6") we only know "a" is of type "nat" but nothing
-more (and this is necessary so that "a" can be overridden in the
-future).\YZ{Not clear if you want to be able to override a or want to be able to have 'Field b : a = 1'.}
+% For example, we may
+% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
+% and intend to prove its commutativity in a following field $"comm"$.
+% But to prove so, we need to know more than just a type interface of
+% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
+% future, once "add" is overridden into a subtraction function then
+% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
+% A more concrete example is in \cref{fig:plugin-example1}, consider the
+% case we have "Field b : a = 1 := .." instead. Then our
+% compilation and Coq's type checking will both fail: in the context of
+% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
+% more (and this is necessary so that "a" can be overridden in the
+% future).
+Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the example, we can see that the first attempt of proving "a = 1" fails.
+This should fail, actually---because we want to retain the possibility to override "a" in the future, but the mere existence of the field "c" will stop that, since "c" witnesses "a" fix to "1".
+From the other perspective, the proof of "c" will require a concrete definition of "a" instead of just the exposure of its type.
+\textbf{We can see a conflict between \textit{overridable} and \textit{(definition-relevant) reasoning}}.
+We enable the exposure of the concrete definition by providing "Final" keyword upon "a'".
 
-To fix that, we need more information on "a" than its type---we
-need to expose the concrete definition "a = 1" to the context of "b".
+This example might sounds trivial and non-practical.
+Let's look at the example field "add".
+Now that we only expose the type of "add", we cannot prove its commutativity in "comm_add" (because we don't have access to its concrete definition) but we can override "add" with subtraction function in "Family I2".
+This makes sense because we shouldn't have (inherited) commutativity property for subtraction anyway.
+This is contrary to how "a'" is declared "Final", the second "c" can prove that it's equal to 1 but "a'" is not overridable in the "Family I2".
+\YZ{Not clear if you want to be able to override a or want to be able to have 'Field b : a = 1'.}\EDJreply{Rephrased. Please Check.}
+
+% To fix that, we need more information on "a" than its type---we
+% need to expose the concrete definition "a = 1" to the context of "b".
 %\YZ{I had a hard time working through these two paras. I guess they are
@@
 
-\textbf{Special Fields: Overridable and Sealed Family as Interface.} 
+\textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.} 
 Non-overridable fields are still not expressive enough to achieve the full
 extent of dependent type programming. There are cases where overriding and
-equational reasoning are both needed, especially when we use Coq to do
-computation. For example, we may want a field to be an overridable 
-monotonic function. We cannot split into two fields
+(definition-relevant) reasoning are both needed. For example, we may want a field to be an overridable monotonic function. We cannot split into two fields
 \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
@@
 
-To resolve this, we provide family sealing mechanism---we allow the
+To resolve this, we provide family sealing mechanism to simulate sigma type---we allow the
 users to specify the compiled module type of a given family,\YZ{I don't understand "specify the family/compiled module type"}\EDJreply{Sorry.
 I rephase it into "compiled module type of a given family".}\YZreply{Still cannot see what is going on and how it is related to the sentence that follows.
-An example is due.}
+An example is due.}\EDJreply{Done.
+Please check.}
 and thus we can have a
-\textit{locally transparent}, \textit{non-overridable} field located in
-an \textit{overridable} family. Using the above example, we simply seal
+ \textit{non-overridable} (thus locally transparent) field located in
+an \textit{overridable} family. Check the "MonotonicF" in \cref{fig:plugin-example3}, we simply seal
 \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
@@
 However, because the enclosing family of "f" is overridable, to swap the
-implementation of "f", we just override the whole ``sealed'' family.
+implementation of "f", we just override the whole ``sealed'' family, like we do in "I2.MonotonicF".
 
-``Sealed'' families can actually simulate interfaces in certain sense\YZ{I was actually thinking sealed families simulate sigma types, no?}\EDJreply{You are absolutely right.
+In our example, the seal family "I.MonotonicF" is actually simulating interfaces as well\YZ{I was actually thinking sealed families simulate sigma types, no?}\EDJreply{You are absolutely right.
 I should add a "can" here.
@@
 Then once I want to instantiate with real implementation, I override the whole family -- as long as it has the same type as the sealed family, everything still works.}\YZreply{Can you explain this with an example?
-I recall the AbsInt example uses it?}---we can assign each field of the sealed family with \mintinline{Coq}{Axiom non_implement : forall {T : Type}, T.}, and later instantiate them with concrete implementation by overriding this family.\YZ{I don't understand what it means to "assign each field with an axiom".}\EDJreply{It really means ``place holder'' in general programming language (Python's NonImplement()), defer to future.
+I recall the AbsInt example uses it?}\EDJreply{Done.
+Please check.}.
+We have \\ ``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}'', and we have assigned it to both "f" and "p", and later instantiate in "I2" with concrete implementation by overriding this family "I.MonotonicF".
+\YZ{I don't understand what it means to "assign each field with an axiom".}\EDJreply{It really means ``place holder'' in general programming language (Python's NonImplement()), defer to future.
 Because Coq's Axiom can be used to instantiate anything.}

### 2022-06-25 lang-design.tex
Adding example for CoqImpl.

@@
   & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
-  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle
+  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad | \quad ...
 \end{align*}
 
-We use "Family" and "EndFamily" to scope, and "Field" is used to declare new members in a family.
+These are some core extension we proposed to support \cref{fig:STLC-example}. We use "Family" and "EndFamily" to scope, and "Field" is used to declare new members in a family.
 We can define nested families.

### 2022-06-25 coqimpl.tex
Reply Comments.

@@
 The inheritance data will include the information of (1) the ``type'' of the parent family, (2) a list of operation indicating how each field of the parent will be dealt with (either inherited or overrided), (3) the operations indicating newly extended fields, and (4) the ``type'' of the children family.\YZ{what is 'inheritance data'?}\EDJreply{It is just inheritance judgement, but before meta-theory is introduced, I can only say some "first-order data that represent the information of inheritance".
-I emphasize first order to distinguish from functor in ML community.}\YZreply{I think this is the place where you can use an example to demonstrate these three kinds of inheritance.} The surface syntax for the three kinds of inheritance are demonstrated in \cref*{fig:plugin-example3}.
+I emphasize first order to distinguish from functor in ML community.}\YZreply{I think this is the place where you can use an example to demonstrate these three kinds of inheritance.} The surface syntax for the three kinds of inheritance are demonstrated in the exemplar \cref*{fig:plugin-example3}.
 The implementation of overriding, inheritance, and extension is achieved by simply manipulating the internal representation of a family correspondingly---we can swap the module in the list for overriding, retain the module in the list for inheritance and add new module into the list for extension.
@@
 \textbf{Reasoning Requires Non-Overridable.}\YZ{At this point in this section, overriding has not been discussed yet, but it is already discussing non-overridables.
-Should probably discuss the extensibility first.}
+Should probably discuss the extensibility first.}\EDJreply{Added.
+Please Check.}
 This compilation strategy
@@
 implementation of "f", we just override the whole ``sealed'' family, like we do in "I2.MonotonicF".
+And the compiled module type of "I2.MonotonicF" will be fixed throughout the extension---it is always these two fields "f, p".
 

### 2022-06-25 lang-design.tex
Reply Comments.

@@
 convince the reader why exposing the definition of subst could be a
-problem for family polymorphism.}
+problem for family polymorphism.}\EDJreply{Actually I don't see why I need to emphasize this "as a problem".
+This is just a requirement like any other requirements earlier.
+I don't know how to show its difficulty at this early stage.}
 

### 2022-06-25 syntax.tex
Reply Comments.

@@
 Do you want me to replace the terminology using exhaustiveness checking?
-}\YZreply{Yes, I think exhaustiveness is standard parlance.}
+}\YZreply{Yes, I think exhaustiveness is standard parlance.}\EDJreply{Rephrased.
+}
 Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden.
@@
 
-\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{it is done in the example section, please check.}
+\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{it is done in the metatheory example section, please check.}
 

### 2022-06-25 syntactic-example.tex
Save comments from discussion.
Fix meta-theory example.

@@
   "sig"_3 &:= \goodSig{\cdot}{\nu^+ "sig"_2 \{"sl"_1\} \ \cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)}{3} \\ 
+  "obj"_3 &: \goodTerm{\cdot}{\mu^+~"obj"_2~"handler"_1}{\cL"sig"_3}\\
   "predT" := (\Pi & \ El ("W"("pjr" ("pjl"^2 \pi_2))) \\  
   & (El ("W" ("pjr" ("pjl"^2  \pi_2)))[\pi_1])) \\
-  "pred"_1 & := \goodTerm{\cdot, \cC "sig"_3}{\lambda ("Wrec" \_)}{predT} \\ 
+  "pred"_1 & := \goodTerm{\cdot, \cC "sig"_3}{\lambda ("Wrec" \_)}{"predT"} \\ 
   "sig"_4 &:= \nu^+ \ "sig"_3 \ \{id_s\} \ "predT" \\ 
+  "obj"_4 &: \goodTerm{\cdot}{\mu^+~"obj"_3~"pred"_1}{\cL"sig"_4}\\
   "sl"_4 &:= \goodSeal{\cdot}{\_}{"sig"_4}{"sig"_4'} \\
-  "pN"_2 &:= \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
-  "sig"_5 &:= \nu^+ \ "sig"_4 \ \{"sl"_4\} \  "pN"_2 \\ 
+  "pN"_2 &:= \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{}
 \end{align*}
@@
 \begin{align*}
+  "sig"_5 &:= \nu^+ \ "sig"_4 \ \{"sl"_4\} \  "pN"_2 \\ 
+  "obj"_5 &: \goodTerm{\cdot}{\mu^+~"obj"_4~\_}{\cL"sig"_4}\\
+  \\ 
+  \text{Here we}&\text{ start to construct Family } "A"_2 \\ 
   "Nsig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  "Osig"}{1} \\ 
@@
 
-The resulting constructed families are $\goodTerm{\cdot}{}{}$
-
 Here we explain how  our detailed construction in Figure \cref{fig:example-construction} is constructed.
@@
 
+The resulting constructed families are $\goodTerm{\cdot}{"obj"_5}{\cL"sig"_5}$ as "Family A" and \\ $\goodTerm{\cdot}{("inh"~"inh"_6~"obj"_5)}{\cL((\nu^+ \ "sig"_{41} \{"sl"_{41}\} \ "pN"_2[(\pi_1, \uparrow^s_2)]))}$ as "Family" $"A"_2$.
+
 Finally we make some comments on how a family is encoded in our metatheory.

### 2022-06-25 coqimpl.tex
Minor Fix. Please Check coqimpl.
Fix meta-theory example.
edits & new comments

--- 2022-06-25 coqimpl.tex
@@
 functors without arguments---this is due to a peculiarity of Coq's
-"Include": non-instantiated parameters will be automatically
+"Include": uninstantiated parameters will be automatically
 instantiated by the appropriate fields in the surrounding defining
@@
   \begin{minipage}[t]{0.45\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
+\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
 Family I.
@@
 Final Field a' : nat := 1.
-Field c : a = 1 := eq_refl.
+Field c : a@\YZ{should this be a'?}@ = 1 := eq_refl.
 Field add : nat -> nat -> nat := (+).
@@
 \begin{minipage}[t]{0.45\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
+\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
 Family I2 extends I.
@@
 
-Override Field add := (-).
+Override Field add := (-).@\YZ{'add' is not a good choice of name---it doesn't make sense to override add with subtraction.}@
 
@@
 Override Family MonotonicF.
-  Final Field f : nat := (+ 1).
+  Final Field f : nat@\YZ{nat -> nat ?}@ := (+ 1).
   Field p : monotonic f := ...
@@
 inheritance and the standalone families are extending the empty family.
-Thus we only need to deal with inheritance during interactive theorem proving.
+Thus, we need to deal with only inheritance during interactive theorem proving.
 
-There are only three kinds of inheritance we are dealing with---extend, override, and inherit.
-We use first-order data to encode inheritance data inside the implementation of the plugin.
-The inheritance data will include the information of (1) the ``type'' of the parent family, (2) a list of operation indicating how each field of the parent will be dealt with (either inherited or overrided), (3) the operations indicating newly extended fields, and (4) the ``type'' of the children family.\YZ{what is 'inheritance data'?}\EDJreply{It is just inheritance judgement, but before meta-theory is introduced, I can only say some "first-order data that represent the information of inheritance".
-I emphasize first order to distinguish from functor in ML community.}\YZreply{I think this is the place where you can use an example to demonstrate these three kinds of inheritance.} The surface syntax for the three kinds of inheritance are demonstrated in \cref*{fig:plugin-example3}.
-The implementation of overriding, inheritance, and extension is achieved by simply manipulating the internal representation of a family correspondingly---we can swap the module in the list for overriding, retain the module in the list for inheritance and add new module into the list for extension.
-The compilation from a family (internal representation) to a module still acts the same.
+There are three kinds of inheritance---extend, override, and inherit. We
+use first-order data to encode inheritance data inside the
+implementation of the plugin. The inheritance data will include the
+information of (1) the ``type'' of the parent family, (2) a list of
+operation indicating how each field of the parent will be dealt with
+(either inherited or overriden), (3) the operations indicating newly
+extended fields, and (4) the ``type'' of the children family.
+The surface syntax for the three kinds of inheritance
+are demonstrated in the exemplar \cref*{fig:plugin-example3}. The
+implementation of overriding, inheritance, and extension is achieved by
+simply manipulating the internal representation of a family
+correspondingly---we can swap the module in the list for overriding,
+retain the module in the list for inheritance, and add new module into
+the list for extension. The compilation from a family (internal
+representation) to a module still acts the same.
 
 Since it is possible to consider inheritance as solely a first-order data rather
-than a higher-order functor, it is actually plausible to
-\text{mix-in} two inheritance data -- by carefully ``mix'' the internal list.
+than a higher-order functor, it is possible to
+\text{mix-in} two inheritance data---by carefully ``mix'' the internal list.\YZ{How about making mixin composition one of your Req's? Also does any example in the paper illustrate how to do this?}
 
-\textbf{Reasoning Requires Non-Overridable.}\YZ{At this point in this section, overriding has not been discussed yet, but it is already discussing non-overridables.
-Should probably discuss the extensibility first.}
+\textbf{Reasoning Requires Non-Overridable.}
 This compilation strategy
@@
 % future).
-Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the example, we can see that the first attempt of proving "a = 1" fails.
-This should fail, actually---because we want to retain the possibility to override "a" in the future, but the mere existence of the field "c" will stop that, since "c" witnesses "a" fix to "1".
-From the other perspective, the proof of "c" will require a concrete definition of "a" instead of just the exposure of its type.
-\textbf{We can see a conflict between \textit{overridable} and \textit{(definition-relevant) reasoning}}.
-We enable the exposure of the concrete definition by providing "Final" keyword upon "a'".
+Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the
+example. We can see that the first attempt of proving "a = 1" fails.
+This should fail, actually---because we want to retain the possibility
+to override "a" in the future, but the mere existence of the field "c"
+will stop that, since "c" witnesses "a" fix to "1". From the other
+perspective, the proof of "c" will require a concrete definition of "a"
+instead of just the exposure of its type. \textbf{We can see a conflict
+between \textit{overridable} and \textit{(definition-relevant)
+reasoning}}.  We enable the exposure of the concrete definition by
+providing "Final" keyword upon "a'".
 
-This example might sounds trivial and non-practical.
-Let's look at the example field "add".
-Now that we only expose the type of "add", we cannot prove its commutativity in "comm_add" (because we don't have access to its concrete definition) but we can override "add" with subtraction function in "Family I2".
-This makes sense because we shouldn't have (inherited) commutativity property for subtraction anyway.
-This is contrary to how "a'" is declared "Final", the second "c" can prove that it's equal to 1 but "a'" is not overridable in the "Family I2".
-\YZ{Not clear if you want to be able to override a or want to be able to have 'Field b : a = 1'.}\EDJreply{Rephrased. Please Check.}
+This example might sound trivial and contrived. Let's look at the
+example field "add". Now that we only expose the type of "add", we
+cannot prove its commutativity in "comm_add" (because we don't have
+access to its concrete definition) but we can override "add" with
+subtraction function in "Family I2". This makes sense because we
+shouldn't have (inherited) commutativity property for subtraction
+anyway. This is contrary to how "a'" is declared "Final", the second "c"
+can prove that it's equal to 1 but "a'" is not overridable in the
+"Family I2".
 
@@
 To resolve this, we provide family sealing mechanism to simulate sigma type---we allow the
-users to specify the compiled module type of a given family,\YZ{I don't understand "specify the family/compiled module type"}\EDJreply{Sorry.
-I rephase it into "compiled module type of a given family".}\YZreply{Still cannot see what is going on and how it is related to the sentence that follows.
-An example is due.}\EDJreply{Done.
-Please check.}
-and thus we can have a
- \textit{non-overridable} (thus locally transparent) field located in
-an \textit{overridable} family. Check the "MonotonicF" in \cref{fig:plugin-example3}, we simply seal
+users to specify the compiled module type of a given family, and thus we
+can have a \textit{non-overridable} (thus locally transparent) field
+located in an \textit{overridable} family.
+See, for example, the "MonotonicF" in \cref{fig:plugin-example3}.
+We simply seal
 \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
@@
 However, because the enclosing family of "f" is overridable, to swap the
-implementation of "f", we just override the whole ``sealed'' family, like we do in "I2.MonotonicF".
+implementation of "f", we just override the whole ``sealed'' family,
+like we do in "I2.MonotonicF". And the compiled module type of
+"MonotonicF" will be fixed throughout the extension in different
+families---it is always these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}
 
-In our example, the seal family "I.MonotonicF" is actually simulating interfaces as well\YZ{I was actually thinking sealed families simulate sigma types, no?}\EDJreply{You are absolutely right.
-I should add a "can" here.
-I can use sealed family to simulate interface is by creating a sigma type, but making the implemntation of sigma type trivial (using place holder/using axiom as I specified below).
-Then once I want to instantiate with real implementation, I override the whole family -- as long as it has the same type as the sealed family, everything still works.}\YZreply{Can you explain this with an example?
-I recall the AbsInt example uses it?}\EDJreply{Done.
-Please check.}.
-We have \\ ``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}'', and we have assigned it to both "f" and "p", and later instantiate in "I2" with concrete implementation by overriding this family "I.MonotonicF".
-\YZ{I don't understand what it means to "assign each field with an axiom".}\EDJreply{It really means ``place holder'' in general programming language (Python's NonImplement()), defer to future.
-Because Coq's Axiom can be used to instantiate anything.}
+In our example, the seal family "I.MonotonicF" is actually simulating
+interfaces as well.  We have
+\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
+and we have assigned it to both "f" and "p", and later instantiate in
+"I2" with concrete implementation by overriding this family
+"I.MonotonicF".
 % Basically, {a : int -> int; property : a is monotonic}, in this case

### 2022-06-25 coqimpl.tex

@@
 Final Field a' : nat := 1.
-Field c : a@\YZ{should this be a'?}@ = 1 := eq_refl.
-Field add : nat -> nat -> nat := (+).
+Field c : a'@\YZ{should this be a'?}\EDJreply{Yes Thanks.}@ = 1 := eq_refl.
+Field bop : nat -> nat -> nat := (+).
 Fail Field comm_add : ...
@@
  SealFamily. 
-
 EndFamily.
@@
 
-Override Field add := (-).@\YZ{'add' is not a good choice of name---it doesn't make sense to override add with subtraction.}@
+Override Field bop := (-).@\YZ{'add' is not a good choice of name---it doesn't make sense to override add with subtraction.}\EDJreply{Changed to b(inary)op now}@
 
@@
   Final Field f : nat@\YZ{nat -> nat ?}@ := (+ 1).
+
   Field p : monotonic f := ...

### 2022-06-25 coqimpl.tex
Reply the comments

@@
 \YZ{The reader wonders why this aggregation is needed. 
-Seems like the place to talk about how the programmer can use A outside Family A.}
+Seems like the place to talk about how the programmer can use A outside Family A.}\EDJreply{Please check.}
 
@@
 Override Family MonotonicF.
-  Final Field f : nat@\YZ{nat -> nat ?}@ := (+ 1).
-
+  Final Field f : nat -> nat@\YZ{nat -> nat ?}\EDJreply{Yes. Thanks}@ := (+ 1).
   Field p : monotonic f := ...
- SealFamily. 
+ EndFamily. 
 EndFamily.
@@
 than a higher-order functor, it is possible to
-\text{mix-in} two inheritance data---by carefully ``mix'' the internal list.\YZ{How about making mixin composition one of your Req's? Also does any example in the paper illustrate how to do this?}
+\text{mix-in} two inheritance data---by carefully ``mix'' the internal list.However, a good definition of mixin is still under investigation.\YZ{How about making mixin composition one of your Req's?
+Also does any example in the paper illustrate how to do this?}\EDJreply{mixin is not really formalized.
+There is no example illustrate it.
+The only example is the STLC example at the very end and I just kind of ``mention'' it.
+I think here I just want to emphasize, inheritance as first order data is still possible to support composition.}
 
@@
 this special treatment.\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
-Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on taking about meta-theory.
+Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory.
 Now everything is about judgemental so I don't have to emphasize judgemental.
@@
 "MonotonicF" will be fixed throughout the extension in different
-families---it is always these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}
+families---it is always these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }
 

### 2022-06-25 coqimpl.tex
Took a pass over paras on recursors

--- 2022-06-25 coqimpl.tex
@@
 \textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
-We don't actually support extensible inductive type from \ref{langdesign-req2} in Coq (because that require non-trivial research effort on theory side---i.e.
-{What is the semantic of extensible inductive type?}---and engineering), but we simulate extensible inductive type using overriding---when extending an inductive type $A$ with new constructors $c$, our plugin will generate a new inductive definition $A'$ with this new $c$ on the fly and feed to Coq.
-But still, we must be careful on the exposing data in the family type (the compiled module type) for the following field.
+We don't actually support extensible inductive types from
+\ref{langdesign-req2} in Coq (because that require non-trivial research
+effort on theory side---i.e. {What is the semantics of extensible
+inductive types?}---and engineering), but we simulate extensible
+inductive type using overriding---when extending an inductive type $A$
+with new constructors $c$, our plugin will generate a new inductive
+definition $A'$ with this new $c$ on the fly and feed it to Coq.  But
+still, we must be careful on the exposing data in the family type (the
+compiled module type) for the following field.
 
@@
   \begin{minipage}[t]{0.65\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
+\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
   (* Abstraction for Inductive Type *)
@@
 Module ff_24 (self_B: b_6) 
-             (self__neg_handler: tt_23 self_B).
+             (self__neg_handler@\YZ{'self_neg_handler' instead?}@: tt_23 self_B).
 Definition ff : self_B.b := self_B.tt. End ff_24.
@@
 
-Let's look at \cref{fig:plugin-example2}, the example of boolean and negation.
-When we define inductive type "b", the following "Field neg_handler.ff" can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface'' the compiled module type "b_3".
-We cannot export the eliminator of "b" inside "b_3" because doing so "neg_handler.ff" (and "ff_24") will not be able to be inherited to the context where "b" is extended with a third constructor, since the existence of recursor alone will assert that there are only two constructors in "b", not ``extensible'' at all!
+Let's look at \cref{fig:plugin-example2}, the example of boolean and
+negation. When we define inductive type "b", the following "Field
+neg_handler.ff" can only know "b" is a type (of "Set") and there are at
+least two constructors for it, just like what we declare in ``the
+abstracted interface'' the compiled module type "b_3". We cannot export\YZ{export or put?}
+the eliminator of "b" inside "b_3" because doing so "neg_handler.ff"
+(and "ff_24") will not be able to be inherited to the context where "b"
+is extended with a third constructor, since the existence of recursor
+alone will assert that there are only two constructors in "b", not
+``extensible'' at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}
 
-But we do have to provide a way to construct an (extensible) recursor (as in \ref{langdesign-req3}), and execute type checking (and exhaustiveness checking).
-Thus we need to extract a elimination principle for future usage.
-Here we show one exemplar extraction of the recursor in \mintinline{Coq}{Set} as module "b_rec_12".
+But we do have to provide a way to construct an (extensible) recursor
+(as in \ref{langdesign-req3}), and perform type checking (and
+exhaustiveness checking). Thus, we need to extract an elimination
+principle for future usage. Here we show one exemplar extraction of the
+recursor in \mintinline{Coq}{Set} as module "b_rec_12".
 
-To construct a recursor as in \ref{langdesign-req3}, we make for each constructor a handler and aggregate them inside a family.
-The whole family "neg_handler" can be considered as a large matching case.
-Doing so, family inheritance can function as matching-case reuse.
-Then we use "FRecursor" to specify the motive and inductive type.
-Here "FRecursor" will feed Coq
-a module "v_33_34" for doing type-checking---what it does is basically assert the existence of the recursor and apply to see if handlers can pass the type checking by Coq.
+To construct a recursor as in \ref{langdesign-req3}, we create for each
+constructor a handler and aggregate them inside a family. The whole
+family "neg_handler" can be considered as a large matching case. Doing
+so, family inheritance can function as matching-case reuse. Then we use
+"FRecursor" to specify the motive and inductive type. Here "FRecursor"
+will feed Coq
+a module "v_33_34" for doing type-checking---what it does is basically
+assert the existence of the recursor and apply to see if handlers can
+pass the type checking by Coq.
 
-Since the recursor is constructed knowing a concrete inductive type, thus not inheritable, for all the children, we have to reconstruct and override with a new recursor (acting like an extra exhaustiveness checking).
-This is affordable since each recursor handler doesn't need to be rechecked---look at "B2.neg_handler", there is only one new handler for "uu" and the handlers from the parent "B.neg_handler.tt,B.neg_handler.ff" are already type-checked and inherited(thanks to the fact that they only rely on ``the abstracted interface'' module "b_3").
-Of course, our plugin will ``inherits'' the recursor "neg" but what the plugin does is doing exhaustiveness checking, again.
+Since the recursor is constructed knowing a concrete inductive type,
+thus not inheritable, for all the children, we have to reconstruct and
+override with a new recursor (acting like an extra exhaustiveness
+checking\YZ{Why is it 'extra'? Is there already exhaustiveness checking anywhere?}). This is affordable since each recursor handler doesn't need
+to be rechecked---look at "B2.neg_handler", there is only one new
+handler for "uu" and the handlers from the parent
+"B.neg_handler.tt,B.neg_handler.ff" have already been type-checked and
+inherited (thanks to the fact that they only rely on ``the abstracted
+interface'' module "b_3"). Of course, our plugin will ``inherit'' the
+recursor "neg" but what the plugin does is doing exhaustiveness
+checking, again.\YZ{It'd be good to show *concretely* in Figure 4
+(1) code that uses the recursor and how the using code is compiled, (2)
+how B2 is compiled and that compiling B2 does not require recompiling
+fields inherited from B.}
 
@@
 % explain ftheorem wrapping this complicated recursor
-We also provide a command "FTheorem" when the users want to use tactic programming instead.
-This tool can avoid most boiler-plate code (e.g.
-specify the handler family) when writing a recursor like above, invoke proof interaction mode, allow tactic usage, and is also open to extension like "FRecursor".
+We also provide a command "FTheorem" when the users want to use tactic
+programming instead. This tool can avoid most boilerplate code (e.g.,
+specify the handler family) when writing a recursor like above, invoke
+proof interaction mode, allow tactic usage, and is also open to
+extension like "FRecursor".\YZ{example?}
 
-To summarize, the key insight here is that, to make each recursor handler inheritable, we need to seal an abstraction around the inductive type (e.g.
-module "b_3"); but to construct a recursor (and carry out the exhaustiveness checking), we need to break this abstraction and see the concrete definition of the inductive type (e.g.
-module "b_rec_12").
-Our meta-theory need to handler these two seemingly contradicting ideas simultaneously.
+To summarize, the key insight here is that, to make each recursor
+handler inheritable, we need to seal an abstraction around the inductive
+type (e.g. module "b_3"); but to construct a recursor (and carry out the
+exhaustiveness checking), we need to break this abstraction and see the
+concrete definition of the inductive type (e.g. module "b_rec_12"). Our
+meta-theory need to handle these two seemingly contradicting ideas
+simultaneously.
 
 \textbf{Propositional Partial Recursor.}
-Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is not expressive enough to really fulfill \ref{langdesign-req2}.
-For example, we cannot prove the constructors are injective, i.e.
-\mintinline{Coq}{self_B.tt ≠ self_B.ff}.
-In the original setting, to prove injectiveness of constructor of (vanilla) inductive type, we only require a recursor and its computational rules.
-However, as we pointed out earlier, the recursor of an inductive type cannot be exposed into the context for the sake of extensibility.
-Thus we need to alter the formulation of the recursor i.e., we need to have a better characterization of ``extensible inductive type'' exported in the compiled module type as a part of the interface for the following fields to use.
+Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
+not expressive enough to really fulfill \ref{langdesign-req2}. For
+example, we cannot prove the constructors are injective, i.e.,
+\mintinline{Coq}{self_B.tt ≠ self_B.ff}. In the original setting, to
+prove injectivity of constructor of (vanilla) inductive type, we only
+require a recursor and its computational rules. However, as we pointed
+out earlier, the recursor of an inductive type cannot be exposed into
+the context for the sake of extensibility. Thus we need to alter the
+formulation of the recursor i.e., we need to have a better
+characterization of ``extensible inductive type'' exported in the
+compiled module type as a part of the interface for the following fields
+to use.
 

### 2022-06-25 coqimpl.tex

@@
 Override Family MonotonicF.
-  Final Field f : nat@\YZ{nat -> nat ?}@ := (+ 1).
+  Final Field f : nat -> nat@\YZ{nat -> nat ?}\EDJreply{Yes. Thanks}@ := (+ 1).
   Field p : monotonic f := ...
-SealFamily. 
+ EndFamily. 
 EndFamily.

### 2022-06-25 coqimpl.tex
resolve comments

@@
 Final Field a' : nat := 1.
-Field c : a'@\YZ{should this be a'?}\EDJreply{Yes Thanks.}@ = 1 := eq_refl.
+Field c : a' = 1 := eq_refl.
 Field bop : nat -> nat -> nat := (+).
@@
 
-Override Field bop := (-).@\YZ{'add' is not a good choice of name---it doesn't make sense to override add with subtraction.}\EDJreply{Changed to b(inary)op now}@
+Override Field bop := (-).
 
@@
 Override Family MonotonicF.
-  Final Field f : nat -> nat@\YZ{nat -> nat ?}\EDJreply{Yes. Thanks}@ := (+ 1).
+  Final Field f : nat -> nat := (+ 1).
   Field p : monotonic f := ...
@@
 users to decorate field as `Final' so that the plugin will proceed with
-this special treatment.\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
+this special treatment.
+%\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
 Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory.

### 2022-06-25 coqimpl.tex
Replying comments

--- 2022-06-25 coqimpl.tex
@@
 Module ff_24 (self_B: b_6) 
-             (self__neg_handler@\YZ{'self_neg_handler' instead?}@: tt_23 self_B).
+             (self_neg_handler@\YZ{'self_neg_handler' instead?}\EDJreply{Yes thx. Though self_neg_handler is not used at all}@: tt_23 self_B).
 Definition ff : self_B.b := self_B.tt. End ff_24.
@@
 least two constructors for it, just like what we declare in ``the
-abstracted interface'' the compiled module type "b_3". We cannot export\YZ{export or put?}
+abstracted interface'' the compiled module type "b_3".
+We cannot export\YZ{export or put?}\EDJreply{Ok.
+I just figure out how my wording is confusing.
+put is absolutely correct but by put I want to mean ``export''.
+I think I use ``export'' in several places...}
 the eliminator of "b" inside "b_3" because doing so "neg_handler.ff"
@@
 thus not inheritable, for all the children, we have to reconstruct and
-override with a new recursor (acting like an extra exhaustiveness
-checking\YZ{Why is it 'extra'? Is there already exhaustiveness checking anywhere?}). This is affordable since each recursor handler doesn't need
+override with a new recursor (acting like another exhaustiveness
+checking in children family\YZ{Why is it 'extra'?
+Is there already exhaustiveness checking anywhere?}\EDJreply{In the parent family.
+Maybe extra is a bad wording }).
+This is affordable since each recursor handler doesn't need
 to be rechecked---look at "B2.neg_handler", there is only one new

### 2022-06-26 standardmodel.tex
Finishing Language Design Section. Try to fix graph embedding and format.
Save syn-translate
Fix syn-translation and others.

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
+We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
 we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.
@@
   \denotesS{\goodSig{\Gamma}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gamma}{\cC \tau}{}} \denotesS{T[sf \ f]} \\
+  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gamma}{\cC \tau}{}} \denotesS{T["sf" \ f]} \\
   & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
@@
   \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
-  &= \denotesS{\goodTerm{\Gamma}{\cCt (p_1\mu \ o)}{...}}, \denotesS{(p_2\mu \ o)[sf \ f][(id, \cCt (p_1\mu \ o))]} \\
+  &= \denotesS{\goodTerm{\Gamma}{\cCt (p_1\mu \ o)}{...}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \cCt (p_1\mu \ o))]} \\
   & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
@@
 
-We encourage the reader to read the (fake) Agda-style proof in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation.
+We again encourage the reader to read the (pseudo-)Agda-style proof mentioned in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation.
 Here, we need to clarify some ambiguity in our conventional formulation.
@@
 
-\begin{figure}
+\begin{figure}[!htb]
   \centering

### 2022-06-26 coqimpl.tex
took a pass on partial-recursor paras

--- 2022-06-25 coqimpl.tex
@@
 \mintinline{Coq}{self_B.tt ≠ self_B.ff}. In the original setting, to
-prove injectivity of constructor of (vanilla) inductive type, we only
+prove injectivity of constructors of a (vanilla) inductive type, we only
 require a recursor and its computational rules. However, as we pointed
 out earlier, the recursor of an inductive type cannot be exposed into
-the context for the sake of extensibility. Thus we need to alter the
-formulation of the recursor i.e., we need to have a better
-characterization of ``extensible inductive type'' exported in the
+the context for the sake of extensibility. Hence, we need to alter the
+formulation of the recursor: we need to have a better
+characterization of ``extensible inductive types'' exported in the
 compiled module type as a part of the interface for the following fields
@@
 
-The simplest idea is to allow partiality---i.e.
-returning \mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
+The simplest idea is to allow partiality---i.e., returning
+\mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
 We call this \textit{propositional partial recursor}.
 For example, the propositional partial recursor for \mintinline{Coq}{B.b} is 
-\begin{minted}{Coq}
-b_prec : forall P : b -> Set, P tt -> P ff 
+\begin{minted}[escapeinside=@@]{Coq}
+b_prec : forall P : b@\YZ{b or self_B.b?}@ -> Set, P tt -> P ff
     -> forall x : b, option (P x).
 \end{minted}
-with appropriate computational axioms. The partiality doesn't break extensibility (i.e. all the future extension of "b" can support this "b_prec").
+with appropriate computational axioms. The partiality doesn't break
+extensibility (i.e. all future extensions of "b" can support this
+"b_prec").
 
-What's more, partial recursor is theoretically enough for the "injection" and "discriminate" tactics.
-The rough idea to achieve this is to use "b_prec" to reflect our special inductive type into vanilla inductive type, and use the orignal "injection" and "discrimination" respectively.
-For example, we want to prove \mintinline{Coq}{self_B.tt ≠ self_B.ff}, we simply
+What's more, partial recursor is theoretically enough for the
+"injection" and "discriminate" tactics. The rough idea to achieve this
+is to use "b_prec" to reflect our special inductive type into vanilla
+inductive type, and use the original "injection" and "discriminate"
+tactics. For example, we want to prove \mintinline{Coq}{self_B.tt ≠
+self_B.ff}, we simply 
 \begin{minted}{Coq}
@@
 
-In fact, the above "reflect" function implies that propositional partial recursor is good enough to some extent, at least for non-indexed inductive type, because we can ``embed'' vanilla inductive type into types supporting partial recursor.
-Because "reflect" is actually a left inverse of an injection "bool → self_B.b".
-A bit more formally, and restrict our focus to non-indexed inductive types,
+In fact, the above "reflect" function implies that propositional partial
+recursor is good enough to some extent, at least for non-indexed
+inductive type, because we can ``embed'' vanilla inductive types into
+types supporting a partial recursor. Because "reflect" is actually a left
+inverse of an injection "bool → self_B.b". A bit more formally, and
+restricting our focus to non-indexed inductive types,
 
@@
   \begin{itemize}
-    \item Define $"C"$ as the inductive type using this list, i.e. 
+    \item Define $"C"$ as the inductive type using this list, i.e.,
     \begin{minted}{Coq}
@@
     \end{minted}
-    \item Assume an arbitrary type \mintinline{Coq}{D : Set} that has ...
+    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
     \begin{itemize}
-      \item following $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
+      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
       \item a partial recursor
-      \begin{minted}{Coq}
-prec : forall (R : Set), ..,  
+      \begin{minted}[escapeinside=@@]{Coq}
+prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}@
   (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> R) -> R)), .., D -> R
       \end{minted}
-      \item all the following $n$ (propositional) computational axioms: for all $i$, 
+      \item and $n$ (propositional) computational axioms: for all $i$, 
       \begin{minted}{Coq}
@@
   \end{itemize}
-  We will have an \textbf{embedding} from "C" to "D", more concretely, we can have \mintinline{Coq}{inj : C -> D} and
-
-  \mintinline{Coq}{linv : D -> option C} defined using eliminator and "prec", (both act similar to identity) such that
+  Then there exists an \textbf{embedding} from "C" to "D".
+  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
+  \mintinline{Coq}{linv : D -> option C} defined using eliminator and "prec"
+  (both acting like identity) such that
   \begin{minted}{Coq}
@@
 
-We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can actually be considered as a constructor, because they can be reflected to real constructors of "C" using "linv".
-Thus, \cref{thm:prec-complete} implies that (1) every future extension of the inductive type can support this partial recursion (trivially); (2) every type support this partial recursor with its computational axiom at least support these constructors because of the embedding.
-In other words, \textit{a type (at least) supports these constructors if and only if this type supports the corresponding partial recursor (with the computational axiom)}.
-Thus we can argue that the partial recursor gives a sound and complete characterization of all the extension of a given (non-indexed) inductive type.
+We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
+actually be considered as a constructor, because they can be reflected
+to real constructors of "C" using "linv". Thus, \cref{thm:prec-complete}
+implies that (1) every future extension of the inductive type can
+support this partial recursion (trivially); (2) every type supporting
+this partial recursor with its computational axioms at least supports
+these constructors because of the embedding. In other words, \textit{a
+type (at least) supports these constructors if and only if this type
+supports the corresponding partial recursor (with the computational
+axioms)}.  Thus, we can argue that the partial recursor gives a sound and
+complete characterization of all the extension of a given (non-indexed)
+inductive type.\YZ{What about indexed inductive types?}
 

### 2022-06-26 body.tex
Adding related-work section.

--- 2022-06-20 body.tex
@@
 
+\section{Related Work}\label{sec:related-work}
+\input{related-work.tex}
+
 \section{Conclusion}
@@
 \section{Metatheory in (pseudo-)Agda}
-We make three agda source code files (publicly and anonymously) openly avaliable that include the corresponding QIIT syntax definition and models.
+We make four agda source code files (publicly and anonymously) openly avaliable, and they include the corresponding QIIT syntax definition and models.
 
-For syntax, please check \href{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}{here}.
+For syntax, please check \\ \href{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}.
+This is done by mainly following the formulation \citet{altkap2016}.
+This is also very close to \citet{coquand2018canonicity} and \citet{sterling2019algebraic} but we formulate our syntax in a type-theoretic framework.
 
-For consistency model, please check \href{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}{here}.
+For syntactical translation, please check \\ \href{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}.
+This is the first QIIT-model we give.
+A QIIT-model can be consider as an arbitrary mapping from the QIIT data(our syntax in this case), also respecting the quotient(judgemental equality in this case).
+Here we use (a subpart of) QIIT data itself as the destination of the mapping.
 
-For canonicity model, please check \href{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}{here}.
+For consistency model, please check \\ \href{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}.
+This is following the standard model from \citet{altkap2016,kaposi2017type,kaposi2019gluing}.
 
-We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins.
+For canonicity model, please check \\ \href{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}.
+This is mainly following \citet{coquand2018canonicity,sterling2019algebraic}, but we carried out our argument in a type-theortic framework.
+We call it canonicity model to emphasize that it is a model used to prove canonicity, it is also called reducibility argument, as an instance of proof-relevant logical relation.
+This argument can be considered as a verbose/more concrete version of \citet{kaposi2019gluing}.
+
+We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.

### 2022-06-26 metatheory.tex
Adding syntactical translation.

@@
 our meta-logic a type theoretical framework (like Agda), 
-together with \textit{Quotient Inductive Inductive Type} (QIIT). Each
-judgement will actually be represented by a QIIT type, which are
+together with \textit{Quotient Inductive Inductive Type} (QIIT).
+Our work can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}.
+We base our formulation of the syntax in this type theory instead of a more conventional and informal representation, becase (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out.
+
+In our syntax, each judgement will actually be represented by a QIIT type, which are
 mutually recursively defined together. Quotient is used to represent
@@
 
+\subsection{Implementation Formalized as Syntactical Translation}
+
+\input{syn-translate.tex}
+
 \subsection{Standard Model for Consistency}

### 2022-06-26 syn-translate.tex
Adding "accessing family components outside that family" into lang-design section Slight change on syntactic translation section.

@@
 The implementation described in \cref{sec:coqimpl} can be formalized in our metatheory.
-The described compilation strategy can be considered as a map from our syntactical (QIIT-)model to a subpart of our syntactical (QIIT-)model \textit{without linkages}.
-This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
+The described compilation strategy can be considered as a \textit{syntactical translation}, a function mapping from our syntactical (QIIT-)model to a subpart of our syntactical (QIIT-)model \textit{without linkages}.\footnote{Here, syntactical QIIT-model can be roughly understood as just AST, with equational rules acting like quotient.
+So this mapping is mapping AST to AST, respecting quotient/judgemental equality.} This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
 
@@
 The details (of all the equational rules) can be found in the appendices written in (pseudo-)Agda form.
+We encourage the reader to read that (pseudo-)Agda-style proof because it is more friendly to type theorists, easy to do type checking and validate the proof and it is clearer on the mathematical object each translation uses.
 
@@
 However, due to our non-conventional formulation of the inductive type, it is better we don't impose such assumption and directly give a consistency and canonicity proof.
+
+% the link for syntactical translation
+% https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing

### 2022-06-26 coqimpl.tex
took a pass on ClosingFact and MetaData paras

@@
   More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using eliminator and "prec"
+  \mintinline{Coq}{linv : D -> option C} defined using eliminator\YZ{eliminator of what?} and "prec"
   (both acting like identity) such that
@@
 % 2. Computational rule for the total recursor
-However, to verify and prove the recursor and partial recursor and their computational rules (as in \ref{langdesign-req4}), our plugin has to compile the extensible family into a non-extensible Coq Module and prove the recursors by instantiating the content of Coq module with concrete inductive type.
+However, to verify and prove the recursor and partial recursor and their
+computational rules (as in \ref{langdesign-req4}), our plugin has to
+compile the extensible family into a non-extensible Coq Module and define
+the recursors by instantiating the content of Coq module with concrete
+inductive types.
 
-This style of ``Global Reasoning'' is that we can only prove the corresponding theorem when we know the ``full pictures'', especially the inductive type.
-We provide "Closing Fact" as a command to carry out this global reasoning---we can attach "Closing Fact" with either (proof) terms or Coq's proof script, then only when compilation to module happens, the term/proof script will put into type-check to see if a "Closing Fact" statement can be proved.
-Note that, "Closing Fact" cannot immediately check the correctness of the term/proof script because this verification only happens when we finished defining a family, and thus a bit harder to handle.
+This style of ``Global Reasoning'' says that we can prove the
+corresponding theorem only when we know the ``full picture'', in particular,
+all the constructors of an inductive type. We provide "Closing Fact" as
+a command to carry out this global reasoning\YZ{Are recursors and partial recursors automatically generated and proved by the plugin, or must they be asserted and proved by the programmer?}---we can attach to "Closing
+Fact" either (proof) terms or Coq's proof script, then only when
+compilation to module happens, is the term/proof script
+type-checked to see if a "Closing Fact" statement can be proved. Note
+that, "Closing Fact" cannot immediately check the correctness of the
+term/proof script because this verification only happens when we
+finished defining a family, and thus a bit harder to handle.
 
-This "Closing Fact" can ease us from the necessity to let plugin generate proof for recursors and partial recursors and their computational rules (as in \ref{langdesign-req4}), and thus making plugin development easier.
-For example, the computation rule for a concrete recursor like "B.neg" and partial recursors in \cref{fig:plugin-example2} can be directly asserted by using "Closing Fact" and verified later during compilation.
-This ``Global Reasoning'' will be again useful in the later examples.
+This "Closing Fact" can ease us from the necessity to let plugin
+generate proof for recursors and partial recursors and their
+computational rules (as in \ref{langdesign-req4}), and thus making
+plugin development easier.\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy. Question is if it makes the programmer's life easier.}
+For example, the computation rule for a
+concrete recursor like "B.neg" and partial recursors in
+\cref{fig:plugin-example2} can be directly asserted by using "Closing
+Fact" and their proofs verified later during compilation.\YZ{What is the alternative if not using "Closing Fact"?}
+This ``Global Reasoning'' will be again useful in later examples.\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}
 
-\textbf{Hack: ``Meta-data''.} Due to our implementation of the family, the family data is invisible for Coq internal and thus cause some inconvenience.
-For example, currently it is not possible to define a vanilla inductive type dependent on our ``extensible'' inductive type due to the invisibility, unless we compile the whole family.
-Similarly, customized tactic expression requested by \ref{langdesign-req5} cannot use lemmas or theorems in the fields of the current defining family as well.
+\textbf{Hack: ``Meta-data''.} Due to our implementation of the family,
+the family data is invisible to Coq internal and thus causes some
+inconvenience.  For example, currently it is not possible to define a
+vanilla inductive type dependent on our ``extensible'' inductive type
+due to the invisibility, unless we compile the whole family. Similarly,
+customized tactic expressions as required by \ref{langdesign-req5} cannot
+use lemmas or theorems in the fields of the current defining family as
+well.
 
-To remedy this, we facilitate a "MetaData" command.
-This command will allow to bundle any original Coq primitive (in the current family context) into the compiled module type, and thus visible to the following defining field.
-Doing so, we can define customized tactic as in \ref{langdesign-req5} that uses lemmas proved in the family.
+To remedy this, we support a "MetaData" command. This command
+bundles any original Coq primitive (in the current family
+context) into the compiled module type, and makes them visible to the
+following defining field. Doing so, we can define customized tactic as
+in \ref{langdesign-req5} that uses lemmas proved in the family.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}

### 2022-06-26 coqimpl.tex
resolve comments

@@
 Module ff_24 (self_B: b_6) 
-             (self_neg_handler@\YZ{'self_neg_handler' instead?}\EDJreply{Yes thx. Though self_neg_handler is not used at all}@: tt_23 self_B).
+             (self_neg_handler: tt_23 self_B).
 Definition ff : self_B.b := self_B.tt. End ff_24.
@@
 thus not inheritable, for all the children, we have to reconstruct and
-override with a new recursor (acting like another exhaustiveness
-checking in children family\YZ{Why is it 'extra'?
-Is there already exhaustiveness checking anywhere?}\EDJreply{In the parent family.
-Maybe extra is a bad wording }).
-This is affordable since each recursor handler doesn't need
+override with a new recursor (therefore serving as exhaustiveness
+checking for child families). This is affordable since each recursor handler doesn't need
 to be rechecked---look at "B2.neg_handler", there is only one new

### 2022-06-26 coqimpl.tex
Rewrite global reasoning section.

@@
 \begin{minted}[escapeinside=@@]{Coq}
-b_prec : forall P : b@\YZ{b or self_B.b?}@ -> Set, P tt -> P ff
+b_prec : forall P : b@\YZ{b or self_B.b?}\EDJreply{it is mainly used as self_B.b, but to make option(P x) at the end doesn't require to be self_B.b so it is kind of orthognal}@ -> Set, P tt -> P ff
     -> forall x : b, option (P x).
@@
   More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using eliminator\YZ{eliminator of what?} and "prec"
+  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of "C"\YZ{eliminator of what?}\EDJreply{"D" is not inductive type so it doesn't have eliminator} and "prec" of "D"
   (both acting like identity) such that
@@
 % 2. Computational rule for the total recursor
-However, to verify and prove the recursor and partial recursor and their
-computational rules (as in \ref{langdesign-req4}), our plugin has to
-compile the extensible family into a non-extensible Coq Module and define
-the recursors by instantiating the content of Coq module with concrete
-inductive types.
+Unfortunately, there are cases the users just want global reasoning instead of an extensible proof.
+For example, in STLC, we should expect`` \mintinline{Coq}{~ value (tm_app x y)}'' holds in all possible extension of STLC (i.e.
+application is never a terminal value).
+We know the correct way of proving this proposition \mintinline{Coq}{Field app_not_value : value (tm_app x y) -> False.} is to use "FRecursor" to inductively reason "value (tm_app x y)".
+However, this means that everytime we extend "value" with new terminal value, we will need to extend corresponding inductive cases for "app_not_value".
+We can imagine the newly extended proof to be boring case analysis that the newly extended terminal value is not "tm_app".
+What's worse, similar scenario can happen on other statement like \mintinline{Coq}{~ value (tm_if cond x y)}.
+Everything here is extended correctly but we end up having boilerplate code, \textbf{semantically}.
 
-This style of ``Global Reasoning'' says that we can prove the
-corresponding theorem only when we know the ``full picture'', in particular,
-all the constructors of an inductive type. We provide "Closing Fact" as
-a command to carry out this global reasoning\YZ{Are recursors and partial recursors automatically generated and proved by the plugin, or must they be asserted and proved by the programmer?}---we can attach to "Closing
-Fact" either (proof) terms or Coq's proof script, then only when
-compilation to module happens, is the term/proof script
-type-checked to see if a "Closing Fact" statement can be proved. Note
-that, "Closing Fact" cannot immediately check the correctness of the
-term/proof script because this verification only happens when we
+Our proposal is to allow global reasoning, which will look like a constraint inside the family.
+We provide "Closing Fact" as a command to carry out this global reasoning, attached with a proof script and proof term.
+This "Closing Fact" will look like an assertion during the construction of the surrounding family, and this assertion will only be verified when the whole family is closed and compiled into a module, with the help of the attached proof script and proof terms.
+
+For example,
+
+The downside is that "Closing Fact" cannot immediately check the correctness of the
+attached proof term/proof script because this verification only happens when we
 finished defining a family, and thus a bit harder to handle.
 
-This "Closing Fact" can ease us from the necessity to let plugin
-generate proof for recursors and partial recursors and their
-computational rules (as in \ref{langdesign-req4}), and thus making
-plugin development easier.\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy. Question is if it makes the programmer's life easier.}
-For example, the computation rule for a
-concrete recursor like "B.neg" and partial recursors in
-\cref{fig:plugin-example2} can be directly asserted by using "Closing
-Fact" and their proofs verified later during compilation.\YZ{What is the alternative if not using "Closing Fact"?}
-This ``Global Reasoning'' will be again useful in later examples.\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}
+% However, to verify and prove the recursor and partial recursor and their
+% computational rules (as in \ref{langdesign-req4}), our plugin has to
+% compile the extensible family into a non-extensible Coq Module and define
+% the recursors by instantiating the content of Coq module with concrete
+% inductive types. We provide "Closing Fact" as a command to carry out this global reasoning,
+
+% This style of ``Global Reasoning'' says that we can prove the
+% corresponding theorem only when we know the ``full picture'', in particular,
+% all the constructors of an inductive type. 
+% We provide "Closing Fact" as
+% a command to carry out this global reasoning
+\YZ{Are recursors and partial recursors automatically generated and proved by the plugin, or must they be asserted and proved by the programmer?}\EDJreply{(1) Closing Fact is not only used to generate recursor and partial recursor.
+I think the intro of this section is just giving people wrong impression.
+So I rewrite the whole sectio now.
+(2)  Recursors and Partial recursors are possible to be automatically generated but my plugin didn't implement it.
+Current stage is that user only need to assert them.
+Nobody has to prove it (plugin will prove it).
+Here when I say user has to provide proof script, it means for arbitrary proposition if the user want to carry out global reasoning.}
+% ---we can attach to "Closing
+% Fact" either (proof) terms or Coq's proof script, then only when
+% compilation to module happens, is the term/proof script
+% type-checked to see if a "Closing Fact" statement can be proved. Note
+% that, "Closing Fact" cannot immediately check the correctness of the
+% term/proof script because this verification only happens when we
+% finished defining a family, and thus a bit harder to handle.
+
+% This "Closing Fact" also can ease us from the necessity to let plugin
+% generate proof for recursors and partial recursors and their
+% computational rules (as in \ref{langdesign-req4}), and thus making
+% plugin development easier.
+\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy.
+Question is if it makes the programmer's life easier.}\EDJreply{You are right.
+It does sound like a bad excuse.
+I rewrite the whole section now.
+Closing Fact is initially used when the user wants global reasoning instead of extensible proof by induction, for example, the inversion lemma.
+\\ It just happens that they can be used by me so that I can postpone generation of these rule plugin (which is hard for me to do).
+This paragraph is just me being honest that my plugin is not mature enough to generate everything.}
+% For example, the computation rule for a
+% concrete recursor like "B.neg" and partial recursors in
+% \cref{fig:plugin-example2} can be directly asserted by using "Closing
+% Fact" and their proofs verified later during compilation.
+\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{To be a legit plugin, my plugin should generate them...}
+% This ``Global Reasoning'' will be again useful in later examples.\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}
 

### 2022-06-26 coq-example.tex

@@
 \subsection{Type Safety for STLC}\EDJ{We need a fuller view of the two examples. Especially show "accessing family components outside that family".}
-We start with the very basic type safety proof of STLC, largely adapted from Software Foundation~\cite{pierce2014software}, which is also one of the primal motivating example of this project.
+We start with the very basic type safety proof of STLC, largely adapted
+from Software Foundation~\cite{pierce2014software}, which is also one of
+the primal motivating example of this project.
 
-Our emphasis here is that, we singled out some programming language features from STLC and prove the type safety for each feature separately, following the style of the examples in MTC~\cite{delaware2013,forsta2020}.
-And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature to mix the semantic and properties of two programming language feature---product and boolean (in \cref{fig:STLC-example})---with vanilla STLC.
-This way, we can say precisely, \textit{a programming language feature itself is a piece of data/inheritance judgement/inherited family}.
-Compared to MTC, our example uses small-step operational semantic by exploiting the extensible inductive type, and most of the proof is directly adapted from the one in Software Foundation, resulting a more accessible proof.
+Our emphasis here is that, we singled out some language features from
+STLC and prove the type safety for each feature separately, following
+the style of the examples in MTC~\cite{delaware2013,forsta2020}.
+And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature
+to mix the semantics and properties of two language features---product
+and boolean (in \cref{fig:STLC-example})---with vanilla STLC. This way,
+we can say precisely, \textit{a programming language feature itself is a
+piece of data/inheritance judgement/inherited family}. Compared with MTC,
+our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?} by exploiting
+extensible inductive types, and most of the proof is directly adapted
+from the one in Software Foundation, resulting in a more accessible proof.\YZ{Can you say more about how your plugin enables more code reuse and modularity?}
 % In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach.
@@
 % We have problem on inversion lemma. Check if it is the same problem as MTC
-We use "Closing Fact" to state and prove the inversion lemma instead of using the extensible proving mechanism "FTheorem".
-The reasons are that (1) it introduce much less boilerplate code because the proof for these inversion lemma should be just simple case analysis and we should rely on Coq to auto-generate them; (2) it shouldn't bother us in the future because any extension on the syntax should still satisfy these inversion lemmas; (3) most importantly, we believe this inversion ``lemma'' should be part of the definition of the syntax instead of considering the syntax as a mere concrete inductive type.
-We should postulate this inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify that our inductive definition did satisfy the constraint, which is exactly what we expect from "Closing Fact".
+We use "Closing Fact" to state and prove the inversion lemma instead of
+using the extensible proving mechanism "FTheorem". The reasons are that
+(1) it introduces much less boilerplate code because the proof for these
+inversion lemmas should be just simple case analysis and we should rely
+on Coq to auto-generate them;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?} (2) it shouldn't bother us in the future
+because any extension on the syntax should still satisfy these inversion
+lemmas; (3) most importantly, we believe this inversion ``lemma'' should
+be part of the definition of the syntax instead of considering the
+syntax as a mere concrete inductive type. We should postulate this
+inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify
+that our inductive definition does satisfy the constraint, which is
+exactly what we expect from "Closing Fact".
 
@@
 
-The biggest difference in the proof script comes from the fact that we are handling ``extensible'' inductive type instead of real inductive type, and thus inversion tactic is not working and we have to manually lift these tactic into inversion lemma.
-This cost a mild code bloat.
+The biggest difference in the proof script comes from the fact that we
+are handling ``extensible'' inductive types instead of real inductive
+types, and thus the inversion tactic is not working and we have to manually
+create inversion lemmas. This leads to a mild code bloat.
 
-Another difference comes from the fact that we need to use "Closing Fact" to manually verify the computational rules for each recursor and use tactic to ``run'' the recursor by repeatedly "rewrite" using those verified computational rules.
-This part is possible to be automatically handled by the plugin, but still it will result propositional computational rules.
+Another difference comes from the fact that we need to use "Closing
+Fact" to manually verify the computational rules for each recursor and
+use tactics to ``run'' the recursor by repeatedly "rewrite" using those
+verified computational rules. This part is possible to be automatically
+handled by the plugin, but still it will require propositional
+computational rules.
 
 \subsection{Abstract Interpreters for \texttt{Imp}}
-The second example is adapted and modified from \citet{zm2017}
---- contrary to our first example, we use big-step interpreter and fuel to indicate the operational semantic on an imperative language with side-effect, and we specify the abstract interpretation and prove its soundness, with some of postulation on both computation and property.
-Then we extend the language feature and we instantiate the postulation on computation for both concrete interpreters and abstract interpreters.
-Thanks to the compilation, we can directly run the resulting abstract interpreters.
+The second example is adapted and modified from \citet{zm2017}.
+Contrary to our first example, we use a big-step interpreter with fuel
+as the operational semantics of an imperative language with
+side effects, and we specify the abstract interpretation and prove its
+soundness, with some postulation on both computation and property.
+Then we extend the base language with new features, and we instantiate
+the postulation on computation for both concrete interpreters and
+abstract interpreters.
+Thanks to the compilation to Coq, we can directly run the resulting
+abstract interpreters.
 
-We construct four families, in one inheritance chain.
+We construct four families, in one inheritance chain.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}
 

### 2022-06-26 coqimpl.tex
Rewriting global reasoning.
Add example for global reasoning and MetaData. Reply comments. Add "FieldClaim" info

@@
 proof interaction mode, allow tactic usage, and is also open to
-extension like "FRecursor".\YZ{example?}
+extension like "FRecursor".\YZ{example?}\EDJreply{FTheorem is really just a short hand for "FRecursor" + handler family.
+Nothing more conceptual here.
+If the reader is interested in the concrete syntax, I think I need to make the two examples avaliable in the appendix.}
 
@@
 application is never a terminal value).
-We know the correct way of proving this proposition \mintinline{Coq}{Field app_not_value : value (tm_app x y) -> False.} is to use "FRecursor" to inductively reason "value (tm_app x y)".
+We know the correct way of proving this proposition \mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use "FRecursor" to inductively reason "value (tm_app x y)".
 However, this means that everytime we extend "value" with new terminal value, we will need to extend corresponding inductive cases for "app_not_value".
@@
 Everything here is extended correctly but we end up having boilerplate code, \textbf{semantically}.
+\begin{figure}[!htb]
+\begin{minted}{Coq}
+Family STLC.
+  FInductive tm : Set := ... 
+  FInductive value : tm -> Prop := ... 
+  Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
+      by { intros x y H; inversion H; eauto }.
+  Fail Ltac inv := 
+    match goal with 
+    | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
+    (* Fail: value, tm_app, value_not_app unfound *)
+    end. 
+  MetaData tactic1.
+    Ltac inv := (* ... the same definition as above ... *)
+  EndMetaData.
+EndFamily.
+\end{minted}  
+\caption{Example for Global Reasoning and MetaData}\label{fig:plugin-example-global-reasoning-meta-data}
+\end{figure}
 
@@
 We provide "Closing Fact" as a command to carry out this global reasoning, attached with a proof script and proof term.
-This "Closing Fact" will look like an assertion during the construction of the surrounding family, and this assertion will only be verified when the whole family is closed and compiled into a module, with the help of the attached proof script and proof terms.
+This "Closing Fact" will look like an assertion during the construction of the surrounding family.
+This assertion will only be verified when the whole family is closed and compiled into a module, with the help of the attached proof script and proof terms.
 
-For example,
+Here we show the example to prove "value_not_app" in \cref{fig:plugin-example-global-reasoning-meta-data}, the script inside curly bracket after "by" is the attached proof script that will be executed when the whole family is being compiled.
+Since we are dealing with a concrete inductive type during compilation, we can use inversion tactic on the vanilla inductive type.
+This "Closing Fact", as an assertion, will have no problem being inherited but the proof script will be executed everytime during each compilation of the children families.
 
@@
 finished defining a family, and thus a bit harder to handle.
+
+Ultimately, "Closing Fact" is asserting a constraint upon all the possible extension 
+of the surrounding family and claiming a proof script can solve the constraints. 
+We expect it to be used in a light-weighted form for maintainability.
 
@@
 I think the intro of this section is just giving people wrong impression.
-So I rewrite the whole sectio now.
+So I rewrite the whole section now.
 (2)  Recursors and Partial recursors are possible to be automatically generated but my plugin didn't implement it.
@@
 Nobody has to prove it (plugin will prove it).
-Here when I say user has to provide proof script, it means for arbitrary proposition if the user want to carry out global reasoning.}
+Here when I say the user has to provide proof script, it means for arbitrary proposition if the user want to carry out global reasoning.}
 % ---we can attach to "Closing
@@
 % Fact" and their proofs verified later during compilation.
-\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{To be a legit plugin, my plugin should generate them...}
+\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know.
+There must be some other ways in the literature.
+}
 % This ``Global Reasoning'' will be again useful in later examples.\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}
@@
 the family data is invisible to Coq internal and thus causes some
-inconvenience.  For example, currently it is not possible to define a
-vanilla inductive type dependent on our ``extensible'' inductive type
-due to the invisibility, unless we compile the whole family. Similarly,
-customized tactic expressions as required by \ref{langdesign-req5} cannot
-use lemmas or theorems in the fields of the current defining family as
-well.
+inconvenience.
+For example, in \cref{fig:plugin-example-global-reasoning-meta-data} the first attempt of defining tactic inv fails.
+Because the earlier defining "tm", "value" and "value_not_app" are currently just data structure in the plugin and not visible for Coq internal (only visible after the family is compiled into module).
 
-To remedy this, we support a "MetaData" command. This command
-bundles any original Coq primitive (in the current family
-context) into the compiled module type, and makes them visible to the
-following defining field. Doing so, we can define customized tactic as
-in \ref{langdesign-req5} that uses lemmas proved in the family.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}
+Thus to refer to them, we have to scope the defining tactic with "MetaData" and our plugin will setup the appropriate environment.
+After a "MetaData" block is closed, the plugin will aggregate the content into the current defining family so that they can also be used when defining the following fields of the family.
+This can fulfill \ref{langdesign-req5}.
+
+To fulfill \ref{langdesign-req5}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
+
+% For example,
+
+%   For example, currently it is not possible to define a
+% vanilla inductive type dependent on our ``extensible'' inductive type
+% due to the invisibility, unless we compile the whole family. Similarly,
+% customized tactic expressions as required by \ref{langdesign-req5} cannot
+% use lemmas or theorems in the fields of the current defining family as
+% well.
+
+% To remedy this, we support a "MetaData" command. This command
+% bundles any original Coq primitive (in the current family
+% context) into the compiled module type, and makes them visible to the
+% following defining field. Doing so, we can define customized tactic as
+% in \ref{langdesign-req5} that uses lemmas proved in the family.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}

### 2022-06-26 lang-design.tex
Add example for global reasoning and MetaData. Reply comments. Add "FieldClaim" info

@@
   & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
+  & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
   & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad | \quad ...
@@
 
-These are some core extension we proposed to support \cref{fig:STLC-example}. We use "Family" and "EndFamily" to scope, and "Field" is used to declare new members in a family.
+These are some core extension we proposed to support \cref{fig:STLC-example}.
+We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
+"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.}
 We can define nested families.

### 2022-06-26 coq-example.tex
Reply some of the comments. Fixed Closing Fact part.

@@
 piece of data/inheritance judgement/inherited family}. Compared with MTC,
-our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?} by exploiting
+our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?}\EDJreply{I thought they use big step with fuel.
+Though Coq a la carte use small-step in one example} by exploiting
 extensible inductive types, and most of the proof is directly adapted
@@
 inversion lemmas should be just simple case analysis and we should rely
-on Coq to auto-generate them;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?} (2) it shouldn't bother us in the future
+on Coq to generate most of the boilerplate code;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?}\EDJreply{I have make the above Closing Fact section more about this detail.
+Please check.} (2) it shouldn't bother us in the future
 because any extension on the syntax should still satisfy these inversion
@@
 
-We construct four families, in one inheritance chain.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}
+We construct four families, in one inheritance chain.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful.
+They are here to show family inheritance is useful, and can extend stuff to all sorts of things.
+I also add some details below on how each family can be extended and what each family means.}
 
-The first family Lang define the big step operator (using fuel) of a while language, with expression. But it postulates the computational information of State used by Lang. It takes about 200+ LOCs.
+The first family "Lang" define the big step operator (using fuel) of a while language, with expression.
+But it postulates the computational information of "State" used by Lang.
+It takes about 200+ LOCs.
+We can swap the implementation of "State" to have different memory/state management by extending family "Lang".
 
-The second family LangwAbs define the abstract interpreter for it, together with a bunch of postulates, to prove the correctness of the abstract interpreter. It takes about 300+ LOCs.
+The second family "LangwAbs" define the abstract interpreter for it, together with a bunch of postulates, to prove the correctness of the abstract interpreter.
+It takes about 300+ LOCs.
+This example shows how to use family inheritance to extend a concrete interpreter with abstract interpreters.
+This example also shows in this family polymoprhism framework we can reason about "Lang" with computation details being abstracted.
 
-The third Family LangMore extend LangwAbs with nat constant and addition, and if-then-else control flow, of course retaining all the soundness theorem. It takes about 200+ LOCs.
+The third Family "LangMore" extend LangwAbs with nat constant and addition, and if-then-else control flow, of course retaining all the soundness theorem.
+It takes about 200+ LOCs.
+This is one example of using family inheritance to support new language feature, and compatible with the existent reasoning.
 
-The fourth family LangRun instantiate some of the postulates of the LangMore and recover computational information of Lang.
+The fourth family "LangRun" instantiate some of the postulates of the LangMore and recover computational information of Lang.
 The concrete interpreter is doing as expected; the abstract interpreter is expected to act as a type-analyzer.
 It takes about 200+ LOCs.
+This is one example of instantiate the detail computation.
+In this example, we use a simple Coq function to simulate state-mapping information.
 
-And at the end, we compile LangRun into module and use Coq to run it on some simple queries.
+And at the end, we compile LangRun into module and use Coq to run it on some simple queries. This "LangRun" also illustrates that we still have computability in the presence of family.

### 2022-06-26 coqimpl.tex
Reply some of the comments. Fixed Closing Fact part.

@@
 There must be some other ways in the literature.
-}
-% This ``Global Reasoning'' will be again useful in later examples.\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}
+}\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}\EDJreply{Please check.}
+% This ``Global Reasoning'' will be again useful in later examples.
 
@@
 After a "MetaData" block is closed, the plugin will aggregate the content into the current defining family so that they can also be used when defining the following fields of the family.
-This can fulfill \ref{langdesign-req5}.
+This can fulfill \ref{langdesign-req5}.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}
 
-To fulfill \ref{langdesign-req5}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
+% To fulfill \ref{langdesign-req5}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
 
@@
 % following defining field. Doing so, we can define customized tactic as
-% in \ref{langdesign-req5} that uses lemmas proved in the family.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}
+% in \ref{langdesign-req5} that uses lemmas proved in the family.

### 2022-06-26 coq-example.tex
comments in examples section

@@
 They are here to show family inheritance is useful, and can extend stuff to all sorts of things.
-I also add some details below on how each family can be extended and what each family means.}
+I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family?
+What is lost by putting the two interpreters in the same family?}
 
-The first family "Lang" define the big step operator (using fuel) of a while language, with expression.
-But it postulates the computational information of "State" used by Lang.
-It takes about 200+ LOCs.
-We can swap the implementation of "State" to have different memory/state management by extending family "Lang".
+The first family "Lang" define the big step operator (using fuel) of a
+while language, with expression. But it postulates the computational
+information of "State" used by Lang. It takes about 200+ LOCs. We
+can\YZ{In general, in a result section like this, you want to show what
+you *have done* rather than what you *can do*.}
+swap the implementation of "State" to have different memory/state
+management by extending family "Lang".\YZ{What is 'different
+memory/state management'?}
 
-The second family "LangwAbs" define the abstract interpreter for it, together with a bunch of postulates, to prove the correctness of the abstract interpreter.
-It takes about 300+ LOCs.
-This example shows how to use family inheritance to extend a concrete interpreter with abstract interpreters.
-This example also shows in this family polymoprhism framework we can reason about "Lang" with computation details being abstracted.
+The second family "LangwAbs" define the abstract interpreter for it,
+together with a bunch of postulates, to prove the correctness of the
+abstract interpreter. It takes about 300+ LOCs. This example shows how
+to use family inheritance to extend a concrete interpreter with abstract
+interpreters. This example also shows in this family polymoprhism
+framework we can reason about "Lang" with computation details being
+abstracted.
 
-The third Family "LangMore" extend LangwAbs with nat constant and addition, and if-then-else control flow, of course retaining all the soundness theorem.
-It takes about 200+ LOCs.
-This is one example of using family inheritance to support new language feature, and compatible with the existent reasoning.
+The third Family "LangMore" extend LangwAbs with nat constant and
+addition, and if-then-else control flow, of course retaining all the
+soundness theorem. It takes about 200+ LOCs. This is one example of
+using family inheritance to support new language feature, and compatible
+with the existent reasoning.
 
-The fourth family "LangRun" instantiate some of the postulates of the LangMore and recover computational information of Lang.
-The concrete interpreter is doing as expected; the abstract interpreter is expected to act as a type-analyzer.
-It takes about 200+ LOCs.
-This is one example of instantiate the detail computation.
-In this example, we use a simple Coq function to simulate state-mapping information.
+The fourth family "LangRun" instantiate some of the postulates of the
+LangMore and recover computational information of Lang. The concrete
+interpreter is doing as expected; the abstract interpreter is expected
+to act as a type-analyzer. It takes about 200+ LOCs. This is one example
+of instantiate the detail computation. In this example, we use a simple
+Coq function to simulate state-mapping information.
 
-And at the end, we compile LangRun into module and use Coq to run it on some simple queries. This "LangRun" also illustrates that we still have computability in the presence of family.
+And at the end, we compile LangRun into module and use Coq to run it on
+some simple queries. This "LangRun" also illustrates that we still have
+computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}

### 2022-06-26 coq-example.tex
Adding things for coq-example

--- 2022-06-26 coq-example.tex
@@
 Though Coq a la carte use small-step in one example} by exploiting
-extensible inductive types, and most of the proof is directly adapted
-from the one in Software Foundation, resulting in a more accessible proof.\YZ{Can you say more about how your plugin enables more code reuse and modularity?}
+extensible inductive types.
+
+Most of the proof is directly adapted
+from the one in Software Foundation, resulting in a similarly accessible proof, while the proof for each extended feature are scattered in the extended families resulting better modularity.
+One kind of examples are about the computation, and here we mainly focus on substitution functions.
+Substitutions dealing with product and projection are defined in the corresponding children family.
+This is mainly done by inheriting and extending the handler family from the parents.
+The other kind of examples are about proof-irrelevant propositions and we directly use "(Extend) FTheorem" to handle them.
+In these cases we don't even to additionally handle a auxillary handler family like we did for substitution---we simply use "Extend FTheorem"
+
+For example, the proof on progress on product projection is carried out in the children family that extend "tm" with products.
+
+\YZ{Can you say more about how your plugin enables more code reuse and modularity?}
 % In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach.

### 2022-06-26 coq-example.tex
Reply the comments.

@@
 I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family?
-What is lost by putting the two interpreters in the same family?}
+What is lost by putting the two interpreters in the same family?}\EDJreply{Nothing is lost.
+This is just simulating people starting with bare (big-step) operational semantic.
+Because they may not working on abstract interpretation from the first place.
+Those people may extend "Lang" with other stuff like another type-safety proof or something.}
 
@@
 while language, with expression. But it postulates the computational
-information of "State" used by Lang. It takes about 200+ LOCs. We
-can\YZ{In general, in a result section like this, you want to show what
-you *have done* rather than what you *can do*.}
-swap the implementation of "State" to have different memory/state
-management by extending family "Lang".\YZ{What is 'different
-memory/state management'?}
+information of "State" (environment mapping) used by Lang. It takes about 200+ LOCs. 
+% We
+% can
+\YZ{In general, in a result section like this, you want to show what
+you *have done* rather than what you *can do*.}\EDJreply{Ok.
+deleted.
+I initially want to emphasize ``I **can** use arbitrary implementation for this environment mapping interface'' but I don't know how to say that without using ``can''.}
+% swap the implementation of "State" to have different memory/state
+% management by extending family "Lang".
+\YZ{What is 'different
+memory/state management'?}\EDJreply{Actually I should say environment mapping/dictionary.
+I instantiate it with function(closure) as environment mapping in "LangRun".
+It is possible to use other mapping like Coq's primitive dictionary.}
 
@@
 to act as a type-analyzer. It takes about 200+ LOCs. This is one example
-of instantiate the detail computation. In this example, we use a simple
-Coq function to simulate state-mapping information.
+of instantiate the detail computation.
+Thanks to the posulation we have done in "Lang", now we instantiate that postulation with function type "Ident -> option exp" as the "State" (environment mapping) for simplicity.
 

### 2022-06-26 coq-example.tex
Fix and Reply the comments.

@@
 
-Most of the proof is directly adapted
+Most of the proof and computation is directly adapted
 from the one in Software Foundation, resulting in a similarly accessible proof, while the proof for each extended feature are scattered in the extended families resulting better modularity.
-One kind of examples are about the computation, and here we mainly focus on substitution functions.
+One kind of examples is about the computation, and here we mainly focus on substitution functions.
 Substitutions dealing with product and projection are defined in the corresponding children family.
@@
 The other kind of examples are about proof-irrelevant propositions and we directly use "(Extend) FTheorem" to handle them.
-In these cases we don't even to additionally handle a auxillary handler family like we did for substitution---we simply use "Extend FTheorem"
-
+In these cases we don't even need to additionally handle a auxillary handler family like we did for substitution---we simply use "Extend FTheorem" and our plugin will prepare the goal to prove.
+At the end, it will look like scattered proof scripts in a modular manner.
 For example, the proof on progress on product projection is carried out in the children family that extend "tm" with products.
-
-\YZ{Can you say more about how your plugin enables more code reuse and modularity?}
+\YZ{Can you say more about how your plugin enables more code reuse and modularity?}\EDJreply{please check}
 % In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach.
@@
 The first family "Lang" define the big step operator (using fuel) of a
-while language, with expression. But it postulates the computational
-information of "State" (environment mapping) used by Lang. It takes about 200+ LOCs. 
+while language, with expression. However, it postulates the definition of "State" (environment mapping) used by Lang. It takes about 200+ LOCs. 
 % We

### 2022-06-26 coqimpl.tex
Fix and Reply the comments.

@@
 
-Since it is possible to consider inheritance as solely a first-order data rather
-than a higher-order functor, it is possible to
-\text{mix-in} two inheritance data---by carefully ``mix'' the internal list.However, a good definition of mixin is still under investigation.\YZ{How about making mixin composition one of your Req's?
+Though inheritance is solely a first-order data rather
+than a higher-order functor, it is still possible to
+\text{mix-in} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance.
+However, a good definition of mixin is still under investigation.\YZ{How about making mixin composition one of your Req's?
 Also does any example in the paper illustrate how to do this?}\EDJreply{mixin is not really formalized.
@@
 \begin{figure}[!htb]
-  \begin{minipage}[t]{0.3\linewidth}
+  \begin{minipage}[t]{0.32\linewidth}
 \begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
@@
     by _rec.
+  Field example := 
+    self_B.neg self_B.tt. 
 EndFamily.
@@
 I think I use ``export'' in several places...}
-the eliminator of "b" inside "b_3" because doing so "neg_handler.ff"
-(and "ff_24") will not be able to be inherited to the context where "b"
+the eliminator of "b" inside "b_3" because doing so all of the following fields (e.g. "neg_handler.ff" and the corresponding "ff_24") will not be able to be inherited to the context where "b"
 is extended with a third constructor, since the existence of recursor
 alone will assert that there are only two constructors in "b", not
-``extensible'' at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}
+\textbf{extensible} at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}\EDJreply{You are right.
+It is not only about "ff_24".I rephrased it.
+Please check.}
 
@@
 constructor a handler and aggregate them inside a family. The whole
-family "neg_handler" can be considered as a large matching case. Doing
-so, family inheritance can function as matching-case reuse. Then we use
+family "neg_handler" can be considered as a large matching case. 
+Then we use
 "FRecursor" to specify the motive and inductive type. Here "FRecursor"
@@
 
+Once recursor is created it will look like any other function fields (and compiled in the same manner), here we show "self_B.neg" to apply to "self_B.tt" for "B.example".
+While in the implementation of our plugin, the concrete recursor for this "neg" will only be created during the compilation of the surrounding family into module---it will look exactly "term_for_type_checking" but uses concrete recursor.
+For example, at the appropriate place, our plugin will insert a definition into the compiled module:
+\begin{minted}{Coq}
+  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff.
+\end{minted}
+where "b_rec" and "b" are the corresponding recursor and the (vanilla) inductive type in that compiled module.
+
 Since the recursor is constructed knowing a concrete inductive type,
@@
 override with a new recursor (therefore serving as exhaustiveness
-checking for child families). This is affordable since each recursor handler doesn't need
+checking for children families). 
+This is affordable since each recursor handler doesn't need
 to be rechecked---look at "B2.neg_handler", there is only one new
@@
 "B.neg_handler.tt,B.neg_handler.ff" have already been type-checked and
-inherited (thanks to the fact that they only rely on ``the abstracted
-interface'' module "b_3"). Of course, our plugin will ``inherit'' the
-recursor "neg" but what the plugin does is doing exhaustiveness
-checking, again.\YZ{It'd be good to show *concretely* in Figure 4
+inherited (since they only rely on ``the abstracted
+interface'' module "b_3"). 
+Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
+the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg".
+
+Of course, our plugin will ``inherit'' the
+recursor "neg" in the surface 
+but what the plugin does is doing a second exhaustiveness
+checking in children family.\YZ{It'd be good to show *concretely* in Figure 4
 (1) code that uses the recursor and how the using code is compiled, (2)
 how B2 is compiled and that compiling B2 does not require recompiling
-fields inherited from B.}
+fields inherited from B.}\EDJreply{I added both. Please check.}
 
@@
 After a "MetaData" block is closed, the plugin will aggregate the content into the current defining family so that they can also be used when defining the following fields of the family.
-This can fulfill \ref{langdesign-req5}.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}
+This can fulfill \ref{langdesign-req5}.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}\EDJreply{Added.
+Please check}
 

### 2022-06-26 coqimpl.tex
resolve comments and add new comments

--- 2022-06-26 coqimpl.tex
@@
 We call this \textit{propositional partial recursor}.
-For example, the propositional partial recursor for \mintinline{Coq}{B.b} is 
+For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}
 \begin{minted}[escapeinside=@@]{Coq}
-b_prec : forall P : b@\YZ{b or self_B.b?}\EDJreply{it is mainly used as self_B.b, but to make option(P x) at the end doesn't require to be self_B.b so it is kind of orthognal}@ -> Set, P tt -> P ff
+b_prec : forall P : b -> Set, P tt -> P ff
     -> forall x : b, option (P x).
@@
   More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of "C"\YZ{eliminator of what?}\EDJreply{"D" is not inductive type so it doesn't have eliminator} and "prec" of "D"
-  (both acting like identity) such that
-  \begin{minted}{Coq}
-    forall c : C, linv (inj c) = some c
+  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of
+  "C" and "prec" of "D" (both acting like identity) such that
+  \begin{minted}[escapeinside=@@]{Coq}
+    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above? If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}@
   \end{minted}
@@
 % 2. Computational rule for the total recursor
-Unfortunately, there are cases the users just want global reasoning instead of an extensible proof.
-For example, in STLC, we should expect`` \mintinline{Coq}{~ value (tm_app x y)}'' holds in all possible extension of STLC (i.e.
-application is never a terminal value).
-We know the correct way of proving this proposition \mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use "FRecursor" to inductively reason "value (tm_app x y)".
-However, this means that everytime we extend "value" with new terminal value, we will need to extend corresponding inductive cases for "app_not_value".
-We can imagine the newly extended proof to be boring case analysis that the newly extended terminal value is not "tm_app".
-What's worse, similar scenario can happen on other statement like \mintinline{Coq}{~ value (tm_if cond x y)}.
-Everything here is extended correctly but we end up having boilerplate code, \textbf{semantically}.
+Unfortunately, there are cases the users just want global reasoning
+instead of an extensible proof. For example, in STLC, we expect
+``\mintinline{Coq}{~ value (tm_app x y)}'' to hold in all future
+extensions of STLC---applications will never be considered as values
+by an extension.
+We know the correct way to prove the proposition
+\mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use
+"FRecursor" to inductively reason "value (tm_app x y)". However, this
+means that every time we extend "value" with new value forms, we will
+need to extend corresponding inductive cases for "app_not_value". We can
+imagine the newly extended proof to be boring case analysis that the
+newly added value forms are not "tm_app". What's worse, similar
+scenario can happen on other statement like \mintinline{Coq}{~ value
+(tm_if cond x y)}. Everything here is extended correctly but we end up
+having boilerplate code, \textbf{semantically}.
 \begin{figure}[!htb]
@@
 
-Our proposal is to allow global reasoning, which will look like a constraint inside the family.
-We provide "Closing Fact" as a command to carry out this global reasoning, attached with a proof script and proof term.
-This "Closing Fact" will look like an assertion during the construction of the surrounding family.
-This assertion will only be verified when the whole family is closed and compiled into a module, with the help of the attached proof script and proof terms.
+Our proposal is to allow global reasoning, which will look like a
+constraint inside the family. We provide "Closing Fact" as a command to
+carry out this global reasoning, attached with a proof script and proof
+term. This "Closing Fact" will look like an assertion during the
+construction of the enclosing family. This assertion will only be
+verified when the whole family is closed and compiled into a module,
+with the help of the attached proof script and proof terms.
 
-Here we show the example to prove "value_not_app" in \cref{fig:plugin-example-global-reasoning-meta-data}, the script inside curly bracket after "by" is the attached proof script that will be executed when the whole family is being compiled.
-Since we are dealing with a concrete inductive type during compilation, we can use inversion tactic on the vanilla inductive type.
-This "Closing Fact", as an assertion, will have no problem being inherited but the proof script will be executed everytime during each compilation of the children families.
+Here we show the example to prove "value_not_app" in
+\cref{fig:plugin-example-global-reasoning-meta-data}, the script inside
+curly brackets after "by" is the attached proof script that will be
+executed upon "EndFamily", which triggers the compilation of the family
+being defined.\YZ{Please check if my wording is correct.}
+Since we are dealing with a concrete inductive type during compilation,
+we can use Coq's "inversion" tactic to complete this proof.
+This "Closing Fact", as an assertion, will have no problem being
+inherited, but the proof script will be executed every time a derived
+family is compiled.
 
@@
 attached proof term/proof script because this verification only happens when we
-finished defining a family, and thus a bit harder to handle.
+finished defining a family, and thus a bit harder to handle.\YZ{'harder' in what way? I suppose you mean the "Fail Ltac" in Figure 5?}
 
-Ultimately, "Closing Fact" is asserting a constraint upon all the possible extension 
-of the surrounding family and claiming a proof script can solve the constraints. 
-We expect it to be used in a light-weighted form for maintainability.
+Ultimately, "Closing Fact" is asserting a constraint upon all the
+possible extension of the surrounding family and claiming a proof script
+can solve the constraints.\YZ{This is a good one-liner that summarizes the intended usage scenarios of "Closing Fact". Should probably bring it out earlier.}
+We expect it to be used only sporadically for maintainability.
 
@@
 % a command to carry out this global reasoning
-\YZ{Are recursors and partial recursors automatically generated and proved by the plugin, or must they be asserted and proved by the programmer?}\EDJreply{(1) Closing Fact is not only used to generate recursor and partial recursor.
+
+%\YZ{Are recursors and partial recursors automatically generated and proved by the plugin, or must they be asserted and proved by the programmer?}\EDJreply{(1) Closing Fact is not only used to generate recursor and partial recursor.
 I think the intro of this section is just giving people wrong impression.
@@
 Here when I say the user has to provide proof script, it means for arbitrary proposition if the user want to carry out global reasoning.}
+
 % ---we can attach to "Closing
@@
 % plugin development easier.
-\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy.
+%\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy.
 Question is if it makes the programmer's life easier.}\EDJreply{You are right.
@@
 % Fact" and their proofs verified later during compilation.
-\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know.
+%\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know.
 There must be some other ways in the literature.
@@
 the family data is invisible to Coq internal and thus causes some
-inconvenience.
-For example, in \cref{fig:plugin-example-global-reasoning-meta-data} the first attempt of defining tactic inv fails.
-Because the earlier defining "tm", "value" and "value_not_app" are currently just data structure in the plugin and not visible for Coq internal (only visible after the family is compiled into module).
+inconvenience. For example, in
+\cref{fig:plugin-example-global-reasoning-meta-data} the first attempt
+of defining tactic "inv" fails. Because the earlier defining "tm",
+"value", and "value_not_app" are currently just data structure in the
+plugin and not visible for Coq internal (only visible after the family
+is compiled into module).
 
-Thus to refer to them, we have to scope the defining tactic with "MetaData" and our plugin will setup the appropriate environment.
-After a "MetaData" block is closed, the plugin will aggregate the content into the current defining family so that they can also be used when defining the following fields of the family.
-This can fulfill \ref{langdesign-req5}.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}
+Thus to refer to them, we have to scope the defining tactic with
+"MetaData" and our plugin will set up the appropriate environment. After
+a "MetaData" block is closed, the plugin will aggregate the content into
+the family being defined so that they can also be used when defining
+the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}
+This can fulfill \ref{langdesign-req5}.
 

### 2022-06-26 coqimpl.tex
Fixing partial recursor

@@
 prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}@
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> R) -> R)), .., D -> R
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> R) -> R)), .., 
+      D -> option R
       \end{minted}
@@
 the family data is invisible to Coq internal and thus causes some
+
 inconvenience. For example, in
@@
 the family being defined so that they can also be used when defining
-the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}
+the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}\EDJreply{Not sure.
+This auto-insertion will practically look like hijack the original Coq command.
+I am not sure this kind of hijacking can happen.}
 This can fulfill \ref{langdesign-req5}.

### 2022-06-26 coqimpl.tex
comments on partial recursors

--- 2022-06-26 coqimpl.tex
@@
 example, we cannot prove the constructors are injective, i.e.,
-\mintinline{Coq}{self_B.tt ≠ self_B.ff}. In the original setting, to
+\mintinline{Coq}{self_B.tt ≠ self_B.ff}.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only?
+My confusion after reading these paras is that I still don't know why partial recursors are needed.}
+In the original setting, to
 prove injectivity of constructors of a (vanilla) inductive type, we only
@@
 \mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
-We call this \textit{propositional partial recursor}.
+We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors? I suppose it has to do with the computational axioms?}
 For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}
@@
 b_prec : forall P : b -> Set, P tt -> P ff
-    -> forall x : b, option (P x).
+    -> forall x : b, option (P x).@\YZ{Not clear what it takes to define/prove partial recursors. The reader needs to be told if they require breaking abstraction like recursors do.}@
 \end{minted}
 with appropriate computational axioms. The partiality doesn't break
-extensibility (i.e. all future extensions of "b" can support this
+extensibility (i.e., all future extensions of "b" can support this
 "b_prec").

### 2022-06-26 coqimpl.tex
Reply some comments.

--- 2022-06-26 coqimpl.tex
@@
 We call this \textit{propositional partial recursor}.
-For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}
+For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}\EDJreply{Added below please check.}
 \begin{minted}[escapeinside=@@]{Coq}
@@
 \end{minted}
-with appropriate computational axioms. The partiality doesn't break
+with appropriate computational axioms.
+This partial recursor differs from original recursor only by adding an "option" at the end thus possible to be automatically generated (the case is similar for those computational axiom).
+The partiality doesn't break
 extensibility (i.e. all future extensions of "b" can support this
@@
   \begin{minted}[escapeinside=@@]{Coq}
-    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above? If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}@
+    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
+If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
+I mentioned left inverse 'linv' is the generalization}@
   \end{minted}
@@
 actually be considered as a constructor, because they can be reflected
-to real constructors of "C" using "linv". Thus, \cref{thm:prec-complete}
+to real constructors of "C" using the left inverse "linv" (directly generalization of "reflect : self_B.b → option bool" of the above). Thus, \cref{thm:prec-complete}
 implies that (1) every future extension of the inductive type can
@@
 executed upon "EndFamily", which triggers the compilation of the family
-being defined.\YZ{Please check if my wording is correct.}
+being defined.
+% \YZ{Please check if my wording is correct.}\EDJreply{Yes. Resolved}
 Since we are dealing with a concrete inductive type during compilation,
@@
 attached proof term/proof script because this verification only happens when we
-finished defining a family, and thus a bit harder to handle.\YZ{'harder' in what way? I suppose you mean the "Fail Ltac" in Figure 5?}
+finished defining a family, and thus a bit harder to handle.\YZ{'harder' in what way?
+I suppose you mean the "Fail Ltac" in Figure 5?}\EDJreply{The proof script cannot be run immediately to reply the user whether the proof script is even correct.
+This is bad engineering to some extent.
+In those bad cases, because error only happens at the the very end ``EndFamily'' , the user will need to go back to the place of Closing Fact to adjust the proof script, or to adjust other parts.
+I am not sure if early check is possible to be done (engineering-wise) I think it is possible (imagine I just do an early compilation, i.e.
+only compile part of the family to the point closing fact is used) but it is a big engineering difficulty for me.}
 

### 2022-06-26 coqimpl.tex

@@
 \mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
-We call this \textit{propositional partial recursor}.
+
+We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors?
+I suppose it has to do with the computational axioms?}\EDJreply{Yes.
+I want to emphasize the equality used by computational axiom is not judgemental equality.
+So Coq cannot automatically compute/reduce when a partial recursor applied to a term.
+This reduction can only be semi-automatically done by using [rewrite tactic and the computational axiom].}
 For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}\EDJreply{Added below please check.}
+
 \begin{minted}[escapeinside=@@]{Coq}
 b_prec : forall P : b -> Set, P tt -> P ff
-    -> forall x : b, option (P x).
+    -> forall x : b, option (P x).@\YZ{Not clear what it takes to define/prove partial recursors. The reader needs to be told if they require breaking abstraction like recursors do.}@
 \end{minted}

### 2022-06-27 coqimpl.tex
comments

@@
 This example might sound trivial and contrived. Let's look at the
-example field "add". Now that we only expose the type of "add", we
+example field "add".\YZ{This 'add' field is no longer there.} Now that we only expose the type of "add", we
 cannot prove its commutativity in "comm_add" (because we don't have
@@
 effort on theory side---i.e. {What is the semantics of extensible
-inductive types?}---and engineering), but we simulate extensible
+inductive types?}---and engineering),\YZ{I do consider that your plugin supports extensible inductive types. It's just that they are implemented as a plugin, rather than in the Coq core.}
+but we simulate extensible
 inductive type using overriding---when extending an inductive type $A$
@@
 
-To construct a recursor as in \ref{langdesign-req3}, we create for each
-constructor a handler and aggregate them inside a family. The whole
+To construct a recursor as in \ref{langdesign-req3}, we\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?} create for each
+constructor a handler and aggregate them into a family. The whole
 family "neg_handler" can be considered as a large matching case. 
@@
 
-Once recursor is created it will look like any other function fields (and compiled in the same manner), here we show "self_B.neg" to apply to "self_B.tt" for "B.example".
-While in the implementation of our plugin, the concrete recursor for this "neg" will only be created during the compilation of the surrounding family into module---it will look exactly "term_for_type_checking" but uses concrete recursor.
-For example, at the appropriate place, our plugin will insert a definition into the compiled module:
+Once recursor is created it will look like any other function fields
+(and compiled in the same manner). Here we show "self_B.neg" to apply to
+"self_B.tt" for "B.example".
+While in the implementation of our plugin, the concrete recursor for
+this "neg" will only be created during the compilation of the enclosing
+family into module---it will look exactly like "term_for_type_checking"
+but uses concrete recursor. For example, at the appropriate place, our
+plugin will insert a definition into the compiled module: 
 \begin{minted}{Coq}
@@
 override with a new recursor (therefore serving as exhaustiveness
-checking for children families). 
+checking for children families).\YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}
 This is affordable since each recursor handler doesn't need
@@
 but what the plugin does is doing a second exhaustiveness
-checking in children family.\YZ{It'd be good to show *concretely* in Figure 4
-(1) code that uses the recursor and how the using code is compiled, (2)
-how B2 is compiled and that compiling B2 does not require recompiling
-fields inherited from B.}\EDJreply{I added both. Please check.}
+checking in children family.
 

### 2022-06-27 lang-design.tex
comments

@@
 \cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
-this feature.\YZ{rephrase Req to Challenges}\EDJreply{But it is not really challenge because I didn't show their difficulty. It is more like requirements.}
+this feature.
 
@@
 information requires exposing \textit{computational rules} from the
-recursors.\YZ{What goes wrong if subst is not invariant or you do not
-know whether subst is invariant?}\EDJreply{Then it is unprovable because substitution lemma need to know this. You can just instantiate the whole lemma -- we want  
-has_type (Γ; x ↦ A) tt T ->
-has_type Γ u A ->
-has_type Γ (subst i x tt) T. But this final goal requires to know the
-computation of subst.}\YZreply{Seems like what's missing here is to
+recursors.\YZ{Seems like what's missing here is to
 convince the reader why exposing the definition of subst could be a
@@
 This is just a requirement like any other requirements earlier.
-I don't know how to show its difficulty at this early stage.}
+I don't know how to show its difficulty at this early stage.}\YZreply{There is apparently a tension between the need to expose definitional equality and the ability to override.
+Looking at the bigger picture, I think this section reads better if phrased as discussing challenges posed by adding family polymorphism to a proof assistant.
+I have some initial ideas that I need to discuss with you.}
 

### 2022-06-27 coqimpl.tex
Reply some of the comments.

@@
  Field bop : nat -> nat -> nat := (+).
- Fail Field comm_add : ...
+ Fail Field comm_bop : ...
 
@@
 
-\textbf{Reasoning Requires Non-Overridable.}
+\textbf{Reasoning Needs Non-Overridable.}
 This compilation strategy
@@
 This example might sound trivial and contrived. Let's look at the
-example field "add".\YZ{This 'add' field is no longer there.} Now that we only expose the type of "add", we
-cannot prove its commutativity in "comm_add" (because we don't have
-access to its concrete definition) but we can override "add" with
+example field binary operation "bop" initially set to be addition.\YZ{This 'add' field is no longer there.}\EDJreply{Thanks. Fixed} Now that we only expose the type of "bop", we
+cannot prove its commutativity in "comm_bop" (because we don't have
+access to its concrete definition) but we can override "bop" with
 subtraction function in "Family I2". This makes sense because we
-shouldn't have (inherited) commutativity property for subtraction
+shouldn't (have the ability to) inherit commutativity property for subtraction
 anyway. This is contrary to how "a'" is declared "Final", the second "c"
@@
 \textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
-We don't actually support extensible inductive types from
-\ref{langdesign-req2} in Coq (because that require non-trivial research
-effort on theory side---i.e. {What is the semantics of extensible
-inductive types?}---and engineering),\YZ{I do consider that your plugin supports extensible inductive types. It's just that they are implemented as a plugin, rather than in the Coq core.}
-but we simulate extensible
+We support ``extensible'' inductive types from
+\ref{langdesign-req2} in Coq to some extent\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort},\YZ{I do consider that your plugin supports extensible inductive types.
+It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it.
+It is more like I don't think this is really the final form of extensible inductive type that everyone is happy about}
+by simulating extensible
 inductive type using overriding---when extending an inductive type $A$
@@
 
-To construct a recursor as in \ref{langdesign-req3}, we\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?} create for each
+To construct a recursor as in \ref{langdesign-req3}, we\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
+Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''} create for each
 constructor a handler and aggregate them into a family. The whole
@@
 example, we cannot prove the constructors are injective, i.e.,
-\mintinline{Coq}{self_B.tt ≠ self_B.ff}.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only?
-My confusion after reading these paras is that I still don't know why partial recursors are needed.}
+\mintinline{Coq}{self_B.tt ≠ self_B.ff} because from the perspective of Coq they are just two terms declared in the module type.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only?
+My confusion after reading these paras is that I still don't know why partial recursors are needed.}\EDJreply{The issue is exactly, at that point of programming, for extensibility, there is no access to the concrete recursor.
+Concrete recursors also cannot be postulated as part of the interface because it breaks extensibility.}
 In the original setting, to

### 2022-06-27 coqimpl.tex
comments

--- 2022-06-27 coqimpl.tex
@@
 handler inheritable, we need to seal an abstraction around the inductive
-type (e.g. module "b_3"); but to construct a recursor (and carry out the
+type (e.g., module "b_3"); but to construct a recursor (and carry out the
 exhaustiveness checking), we need to break this abstraction and see the
-concrete definition of the inductive type (e.g. module "b_rec_12"). Our
+concrete definition of the inductive type (e.g., module "b_rec_12"). Our
 meta-theory need to handle these two seemingly contradicting ideas

### 2022-06-27 syntax.tex
comments

--- 2022-06-24 syntax.tex
@@
 After these,
-we will have four new kinds of judgement to model our linkage.
+we will have four new kinds of judgements to model our linkages.
 
@@
 Please check}
-Latex-wise, we spell out these \textit{well-formedness rule} of type
-judgements and their Agda counterpart to make things more accessible. Of
+Latex-wise, we spell out these \textit{well-formedness rules} of type
+judgements and their Agda counterparts to make things more accessible. Of
 course, a well-formed judgement is not necessarily derivable, just like
@@
 
-Due to our intrinsic setting, \textbf{we can and will omit a lot of presumptions of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly---because it is not possible to have "not-well-formed" context in our setting.
+Due to our intrinsic setting, \textbf{we can and will omit a lot of
+presumptions of the typing rules}. For example, we don't spell out that
+the rule "Type Universe" requires a well-typed context, even though it
+does implicitly, because it is not possible to have ill-formed
+contexts in our intrinsic setting.
 
-Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.
+Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the
+most special one because it is actually just a shorthand of term
+judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.\YZ{You haven't introduced the notation A[pi_1] yet at this point. And what is A[pi_1]?}
 
@@
     &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
-    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
+    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
+Besides, later you also use gamma to range over subsitutions and s to range over W-types, which add to the confusion.}
   \end{align*}
-\caption{Judgements and their Agda Encoding}
+\caption{Judgement forms, as Agda types and as presented in this section.}
 \end{figure}
@@
       {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
-      $$
+      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments?
+Besides, are these rules used anywhere in the following?
+Seems a bit unusual to have well-formedness rules for judgments per se.}
   \end{minipage}
@@
 
-\caption{Well-formedness of Four New Judgements}
+\caption{Well-formedness rules of four new judgements}
 \end{figure}
@@
 function type and dependent pair type.
+The substitution law for these two
+require \textit{de bruijn lifting}: for
+$\goodTerm{\Gamma}{\sigma}{\Delta}$, we define its de Bruijn lifting
+$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow}{\Delta, A}$ given
+$\goodType{\Delta}{A}{}$.\YZ{What is A[sigma] and what is de Bruijn lifting?
+Seems like if the reader is not familiar with the substitution calculus, then they will get lost starting here.
+Broadly speaking, a gentle intro to the intrinsic setting would help the reader a great deal.}
 
@@
 {\goodType{\Gamma}{T}{j}}
-{\goodTerm{\Gamma}{"c" \ T}{\cU_j}
+{\goodTerm{\Gamma}{"c" \ T}{\cU_j}\YZ{what is c? Why is this rule called Univ-1?Didn't you write that you omit universe levels?}
 }\quad
@@
 {\goodTerm{\Gamma}{T}{\cU_j}}
-{\goodType{\Gamma}{El \ T}{j}}
+{\goodType{\Gamma}{El \ T}{j}\YZ{'El T' isn't explained only until it makes a second appearance in Ind-Term.}}
 \quad
@@
 
-We have two kinds of signatures, one for inductive type, the other for linkages.
+We have two kinds of signatures, one for inductive types, the other for linkages.
 
@@
 
-Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type.
-To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
-Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig-Proj} to extract and corresponding $A_i, B_i$.
+Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A
+\vdash B(x)$ decides a W-type. To simulate extensible inductive types, we
+have to equip our inductive types with multiple constructors, and for
+simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ :
+(x : Aᵢ) → (Bᵢ x → W) → W}.
+Thus, we use a list of pairs ($A_i, B_i$) as the signature of one
+inductive type, and given an inductive type signature, we use
+\ruleref{WSig-Proj} to extract the corresponding $A_i, B_i$.
 
-Of course we need to put an eye on the substitution law for each piece of the syntax.
-Here $\bW \sigma$ is a large singleton type~\cite{stone2000}, and it only has one ``element'', the inductive type itself, inside.
-We formulate in this cumbersome way because we want to have inductive type to be a field element and exposing the constructors at the type level for later \textit{pattern-matching coverage checking} (i.e.
-if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type.
+Of course we need to put an eye on the substitution law for each piece
+of the syntax. Here $\bW \sigma$ is a large singleton
+type~\cite{stone2000}, and it only has one ``element'', the inductive
+type itself, inside. We formulate in this cumbersome way because we want
+to have inductive type to be a field element and exposing the
+constructors at the type level for later \textit{pattern-matching
+exhaustiveness checking} (i.e. if an inductive type as field element has type
+$\cU$, then no information about the inductive type is exposed).
+\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type.
 They are basically the same thing but (W T) cannot be at the type position in a judgement, i.e.
@@
 BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general.
-I just didn't write out the two equations c (El T) = T, El (c T) = T}.
-We follow the style of \textit{type universe à la Tarski} and distinguish types from their \textit{codes or names}, and thus we use $El\ T$ as the type given type name $T : \cU$---in these cases, $T$ is not allowed to locate at type position (i.e.
-at the right-most position of a term judgement) but $El \ T$ is.
+I just didn't write out the two equations c (El T) = T, El (c T) = T}\YZreply{Seems harmless to list these equations and move the reference to 'type universe a la Tarski' earlier.}
+We follow the style of \textit{type universe à la Tarski} and
+distinguish types from their \textit{codes or names}, and thus we use
+$El\ T$ as the type given type name $T : \cU$---in these cases, $T$ is
+not allowed to locate at type position (i.e. at the right-most position
+of a term judgement) but $El \ T$ is.
 
-The formulation of our recursor is closely related to the linkage, so we will postpone their description.
+The formulation of our recursor is closely related to that of linkages,
+so we will postpone their description.
 
@@
 {\goodSig{\Gamma}{\sigma}{n} 
- \quad \goodSeal{\Gamma}{f}{\sigma}{A}
+ \quad \goodSeal{\Gamma}{f}{\sigma}{A}\YZ{How should I understand 'f' and 'A'?}
  \quad \goodType{\Gamma, A}{T}{i}}
@@
 
-$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields.
-We omit the naming/label information for fields in the meta-theory.
-We can use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).
+$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature
+for linkages in the context $\Gamma$ with $n$ fields. We omit the
+naming/label information for fields in the meta\-theory. We can use $\cL$
+to retrieve the corresponding types of linkage of a given signature, and
+$\cC$ to get the compiled linkage type (an existential type).\YZ{Why is Csigma needed? How is it different from Lsigma?}
 
-The signature (for the linkage) and the linkage rules are quite similar the signature for inductive type.
-They are both constructed from the empty stuff, and respect substitution lemma for sure.
-We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony~\cite{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules).
-We omit them and the substitution laws for either field addition and projection.
-For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.\footnote{This omission can be supported by Agda using Implicit Arguments}
+The signature (for the linkage) and the linkage rules are quite similar to
+the signature for inductive types.  They are both constructed from the
+empty stuff, and respect substitution lemma for sure. We have
+\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component
+due to harmony~\cite{pfenning2009lecture}, so we have local soundness
+and local completeness ($\beta,\eta$ rules). We omit them and the
+substitution laws for either field addition and projection. For
+$\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.\footnote{This
+omission can be supported by Agda using Implicit Arguments}
 
@@
 }
-{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
+{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}\YZ{What is R[pi1 circ pi1]?}}
 \quad 
@@
 
-In our meta-theory, we only formulate non-dependent eliminator, i.e. the
-\textit{recursor}. Each recursor is constructed by a linkage containing
+In our meta-theory, we only formulate non-dependent eliminator, i.e., the
+\textit{recursor}.\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}
+Each recursor is constructed by a linkage containing
 functions/handlers/pattern-matching cases dealing with each
 constructor. Thus code reuse for recursor is delegated to linkage
-inheritance----to reuse one particular pattern-matching case,
+inheritance---to reuse one particular pattern-matching case,
 we just inherit the function/handler. This greatly simplifies the
@@
 
-Intuitively \ruleref{Hdler} is indicating the type of recursor handler for one specific constructor, and \ruleref{Hdlers} is the linkage type of the bundle of all recursor handlers of one given inductive type.
-We omit the definition of "RecSig" but \ruleref{Hdler-Proj} indicates its (apparent) internal definition.
-"prjR" is also the handler selection from the handler linkage, and is helpful for the formulation of $\beta$-rule of "Wrec" (omitted here).
-Finally \ruleref{Rec-Constr} is the used to construct a concrete recursor given a specific linkage of handlers.
-All these rules satisfy substitution laws and omitted here.
+Intuitively \ruleref{Hdler} is indicating the type of recursor handler
+for one specific constructor, and \ruleref{Hdlers} is the linkage type
+of the bundle of all recursor handlers of one given inductive type. We
+omit the definition of "RecSig" but \ruleref{Hdler-Proj} indicates its
+(apparent) internal definition. "prjR" is also the handler selection
+from the handler linkage, and is helpful for the formulation of
+$\beta$-rule of "Wrec" (omitted here). Finally \ruleref{Rec-Constr} is
+the used to construct a concrete recursor given a specific linkage of
+handlers. All these rules satisfy substitution laws and omitted here.
+\YZ{Seems like you omit many things here, which causes hardship for the reader (at least for me!). I think it's better to list all the rules and decide later what to throw to the appendix.}
 
@@
 
-\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a short hand of the other)}
+\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a shorthand of the other)}
 
-\begin{minted}[]{agda}
+\begin{minted}[escapeinside=@@]{agda}
 Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
-Seal Γ A τ = Tm (Γ, 𝓛σ) A[p₁]
+Seal Γ A τ = Tm (Γ, 𝓛σ) A[p₁]@\YZ{tau?}@
 
@@
 As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
-a short hand of term judgement $\goodTerm{\Gamma , \cC
-\sigma}{\_}{A[\pi_1]}$, a special form of \textit{local morphism}~\cite{abbott2003category}. The substitution law should be understood as a
-lemma, and it is easily derivable using $\gamma^\uparrow$.
+a shorthand of term judgement $\goodTerm{\Gamma , \cC
+\sigma}{\_}{A[\pi_1]}$,\YZ{Not quite. Previously you used cL I think.}
+a special form of \textit{local morphism}~\cite{abbott2003category}.
+The substitution law should be understood as a lemma, and it is easily
+derivable using $\gamma^\uparrow$.
 
-The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance---due to the presence of inductive type in the context, not everything is inheritable.
-For example, $ X_1 : \bW \sigma_1 \vdash t : T$ in the parent family will generally not be able to be inherited to the children once we override the inductive type.
-In other words, there is no way ``transplanting'' $t$ into the context $X_2 : \bW \sigma_2$.
-This makes all the fields after this inductive type non-inheritable---which shouldn't be the case because a lot of fields might not be directly relying on this concrete $\bW \sigma_1$.
+The motivation for sealing judgement is acting like syntactic sugar
+helpful for inheritance---due to the presence of inductive type in the
+context, not everything is inheritable. For example, $ X_1 : \bW
+\sigma_1 \vdash t : T$ in the parent family will generally not be able
+to be inherited to the children once we override the inductive type. In
+other words, there is no way to ``transplant'' $t$ into the context $X_2
+: \bW \sigma_2$.\YZ{Is sigma_2 the overriding of sigma_1? The reader has to guess.}
+This makes all the fields after this inductive type
+non-inheritable---which shouldn't be the case because a lot of fields
+might not be directly relying on this concrete $\bW \sigma_1$.
 
-Sealing is used to exploit this fact---with sealing, we can add well-typed terms $X_1 : \cU \vdash t : T$, into the parent family, in the context of $ X_1 : \bW \sigma_1$.
-Later when we want to inherit $t$, because $t$ only consider $X_1$ as an arbitrary type, inheritance is plausible.
+Sealing\YZ{It occurs to me that you use the word "sealing" to mean different (though tangentially related) things.
+Previously you used it as in "sealed families".} is used to exploit this fact---with sealing, given $X_1 : \bW
+\sigma_1$, we can add well-typed terms $X_1 : \cU \vdash t : T$, into
+the parent family. Later when we want to inherit $t$,
+because $t$ only consider $X_1$ as an arbitrary type, inheritance is
+possible.
 
-Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and \ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}.
-These fields together with inductive types are the one sealing responsible for hiding.
-After compilation of linkage type into existential type, we can easily hide the \textit{concrete} definition of the inductive type while exposing all the fields using "Wsup" and "Wrec" like an abstract interface.
-Apparently, future fields relies on this abstract interface will have no problem on inheritance.
-\YZ{Does coverage checking mean exhaustiveness checking?}\EDJreply{Yes.
-The one make sure every case is handled in pattern-matching.
-Do you want me to replace the terminology using exhaustiveness checking?
-}\YZreply{Yes, I think exhaustiveness is standard parlance.}
+Of course, any fields using "Wsup" and "Wrec" directly relies on the
+$\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and
+\ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup"
+can be understood as concrete constructors and "Wrec" are
+\textit{exhaustiveness checking}. These fields together with inductive
+types are the ones sealing is responsible for hiding. After compilation
+of linkage types into existential types, we can easily hide the
+\textit{concrete} definition of the inductive type while exposing all
+the fields using "Wsup" and "Wrec" like an abstract interface.
+Apparently, future fields relying on this abstract interface will have no
+problem on inheritance.
 Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden.
@@
 
-\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{Working on it, don't resolve}
-
-And these are also the cases we don't need to seal anything (i.e.
-doing exhaustiveness checking when using recursor), and thus the "Seal-Id" is used.
-For notation clearance, we will sometime use $"sf'"$ to emphasize that we consider that seal $s$ as a term $("sf'" \ s)$.
+There are also cases we don't need to seal anything (i.e., doing
+exhaustiveness checking when using recursor), and thus the "Seal-Id" is
+used. For notation clarity, we will sometimes use $"sf'"$ to emphasize
+that we consider that seal $s$ as a term $("sf'" \ s)$.
 

### 2022-06-27 syntactic-example.tex
comments

@@
-We use an example to better demonstrate different components of the syntax of our formulation. Our example will be the meta-theory encoding of the following pseudo-code.
+We use an example to better demonstrate different components of the
+syntax of our formulation. Our example will be the meta-theory encoding
+of the following pseudocode.
 
-We provide one example in its surface syntax as pseudocode, and their meta-theory encoding.
-This example is computing the predecessor of a mundane natural number, but we split the inductive definition into two parts and making "O" and "S" standing alone.
-"pred" is defined as $n \mapsto ("pred" n, n)$, and thus a predecessor.
+We provide one example in its surface syntax as pseudocode, and their
+meta-theory encoding. This example is computing the predecessor of a
+mundane natural number, but we split the inductive definition into two
+parts and making "O" and "S" standing alone.
+"pred" is defined as $n \mapsto ("pred" n, n)$\YZ{looks like an ill-founded definition}, and thus a predecessor.
 
@@
 \begin{minipage}[t]{0.5\linewidth}
-\begin{minted}{Coq}
+\begin{minted}[escapeinside=@@]{Coq}
 Family A.
@@
  EndFamily. (* pN₁, handler₁, sig₃ *)
- pred 
+ pred@\YZ{Why is there no FRecursor?}@
   : self[A].N → self[A].N × self[A].N
@@
   \begin{minipage}[t]{0.45\linewidth}
-\begin{minted}{Coq}
+\begin{minted}[escapeinside=@@]{Coq}
 Family A₂ extends A.

### 2022-06-28 coqimpl.tex
reorganize extensible inductive type section

@@
 We support ``extensible'' inductive types from
-\ref{langdesign-req2} in Coq to some extent\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort},\YZ{I do consider that your plugin supports extensible inductive types.
+\ref{langdesign-req2} in Coq to some extent.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types.
 It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it.
 It is more like I don't think this is really the final form of extensible inductive type that everyone is happy about}
-by simulating extensible
-inductive type using overriding---when extending an inductive type $A$
-with new constructors $c$, our plugin will generate a new inductive
-definition $A'$ with this new $c$ on the fly and feed it to Coq.  But
-still, we must be careful on the exposing data in the family type (the
-compiled module type) for the following field.
 
-% Insert one pseudo-code example for explanation of the mechanism
-% Use the natural number example
 \begin{figure}[!htb]
@@
 
-Let's look at \cref{fig:plugin-example2}, the example of boolean and
-negation. When we define inductive type "b", the following "Field
-neg_handler.ff" can only know "b" is a type (of "Set") and there are at
-least two constructors for it, just like what we declare in ``the
-abstracted interface'' the compiled module type "b_3".
+We start with the programming interface.
+We use "FInductive" to define a extensible inductive type and in the children family we use "Extend Finductive ..." to extend the inductive type with new constructors.
+In \cref{fig:plugin-example2}, we make the example of boolean and three-valued boolean.
+
+To construct a recursor as in \ref{langdesign-req3}, we need to create for each
+constructor a handler (function) and aggregate them into a family, and thus that family (e.g.
+"neg_handler" in \cref{fig:plugin-example2}) can be considered as a large matching-case statement.
+We use "FRecursor" to declare a recursor (e.g.
+"neg") by specifying the motive, the aiming inductive type and the handler family.
+Once recursor is created it will look like any other function fields (e.g.
+"neg").
+Here in \cref{fig:plugin-example2} we apply "neg" to "tt" as an exemplar field.
+The inheritance of recursor is mainly delegated by the inheritance of the handler family---if the user extend with new constructor (e.g.
+"uu"), then the inherited family (e.g.
+"neg_handler") will need to extend correspondingly (e.g.
+"neg_handler.uu").
+If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".
+
+The plugin actually realizes the "Extend FInductive" via overridding---when extending an inductive type $A$
+with new constructors $c$, our plugin will generate a new (syntactic) inductive definition $A'$ with this new $c$ on the fly and feed it to Coq.
+But we must be careful on the exposing data in the family type (the compiled module type) for the following field.
+In other words, we must be careful on what interface (context) the following fields are defined based on.
+
+Let's look at the exemplar \cref{fig:plugin-example2}, the example of boolean and negation.
+When we define the extensible inductive type "b", all the following fields (e.g.
+"neg_handler.ff") can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface''---the compiled module type "b_3".
 We cannot export\YZ{export or put?}\EDJreply{Ok.
@@
 I think I use ``export'' in several places...}
-the eliminator of "b" inside "b_3" because doing so all of the following fields (e.g. "neg_handler.ff" and the corresponding "ff_24") will not be able to be inherited to the context where "b"
-is extended with a third constructor, since the existence of recursor
-alone will assert that there are only two constructors in "b", not
+the eliminator of the inductive type "b" inside "b_3" because doing so all of the following fields (e.g.
+"neg_handler.ff" and the corresponding "ff_24") will not be able to be inherited to the context where "b"
+is extended with a third constructor, \textbf{since the existence of recursor
+alone will assert that there are only two constructors} in "b", not
 \textbf{extensible} at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}\EDJreply{You are right.
@@
 
-But we do have to provide a way to construct an (extensible) recursor
-(as in \ref{langdesign-req3}), and perform type checking (and
-exhaustiveness checking). Thus, we need to extract an elimination
-principle for future usage. Here we show one exemplar extraction of the
-recursor in \mintinline{Coq}{Set} as module "b_rec_12".
-
-To construct a recursor as in \ref{langdesign-req3}, we\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
-Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''} create for each
-constructor a handler and aggregate them into a family. The whole
-family "neg_handler" can be considered as a large matching case. 
-Then we use
-"FRecursor" to specify the motive and inductive type. Here "FRecursor"
-will feed Coq
-a module "v_33_34" for doing type-checking---what it does is basically
+However, we later still need the concrete eliminator "b" for doing type-checking (as in \ref{langdesign-req3}), thus the information doesn't totally disappear but store elsewhere.
+Here we show one exemplar extraction of the recursor to \mintinline{Coq}{Set} as module "b_rec_12".
+This concrete recursor will be used by our "FRecursor".
+For example, in \cref{fig:plugin-example2}, "FRecursor" command will feed Coq a module "v_33_34" for doing type-checking with specified handlers (e.g.
+"neg_handler") during compilation---what it does is basically
 assert the existence of the recursor and apply to see if handlers can
 pass the type checking by Coq.
-
-Once recursor is created it will look like any other function fields
-(and compiled in the same manner). Here we show "self_B.neg" to apply to
-"self_B.tt" for "B.example".
-While in the implementation of our plugin, the concrete recursor for
-this "neg" will only be created during the compilation of the enclosing
-family into module---it will look exactly like "term_for_type_checking"
-but uses concrete recursor. For example, at the appropriate place, our
-plugin will insert a definition into the compiled module: 
+Note that, at this point of the example, "FRecursor" will only do type check.
+The compilation of "FRecursor" into a component of the compiled module only happens when closing the whole family, by inserting the following into the compiled module:
 \begin{minted}{Coq}
@@
 where "b_rec" and "b" are the corresponding recursor and the (vanilla) inductive type in that compiled module.
+Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields.
 
-Since the recursor is constructed knowing a concrete inductive type,
-thus not inheritable, for all the children, we have to reconstruct and
-override with a new recursor (therefore serving as exhaustiveness
-checking for children families).\YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}
+Importantly, the recursor field (e.g.
+"neg") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) module, a recursor field is not compiled like that.
+For all the children family, we have to reconstruct and override with a ``new'' recursor using the (maybe inherited) handler family.
 This is affordable since each recursor handler doesn't need
 to be rechecked---look at "B2.neg_handler", there is only one new
-handler for "uu" and the handlers from the parent
+handler for "uu" and the handlers (mundane fields) from the parent
 "B.neg_handler.tt,B.neg_handler.ff" have already been type-checked and
@@
 Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
-the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg".
+the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg".  Of course, in the implementation, our plugin will ``inherit'' the
+recursor "neg" in the surface \YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}
+but what the plugin actually does is doing a second exhaustiveness
+checking in children family like mentioned above, with the ``same'' motive and ``same'' handler family.\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
+Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}
 
-Of course, our plugin will ``inherit'' the
-recursor "neg" in the surface 
-but what the plugin does is doing a second exhaustiveness
-checking in children family.
+% Insert one pseudo-code example for explanation of the mechanism
+% Use the natural number example
 
@@
 % explain ftheorem wrapping this complicated recursor
-We also provide a command "FTheorem" when the users want to use tactic
-programming instead. This tool can avoid most boilerplate code (e.g.,
-specify the handler family) when writing a recursor like above, invoke
-proof interaction mode, allow tactic usage, and is also open to
-extension like "FRecursor".\YZ{example?}\EDJreply{FTheorem is really just a short hand for "FRecursor" + handler family.
+Finally, we also provide a command "FTheorem" in programming interface when the users want to use tactic
+programming instead. This command (1) can avoid most boilerplate code (e.g., specify the handler family) when writing a recursor like above, (2) invoke
+proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor".\YZ{example?}\EDJreply{FTheorem is really just a short hand for "FRecursor" + handler family.
 Nothing more conceptual here.

### 2022-06-28 coqimpl.tex
Fix reorganization.

@@
 Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
-the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg".  Of course, in the implementation, our plugin will ``inherit'' the
-recursor "neg" in the surface \YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}
+the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg".
+For example, for the new "neg", when compiling family "B2" into a module, at the appropriate place our plugin will insert
+\begin{minted}{Coq}
+  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff neg_handler.uu.
+\end{minted}
+where "b_rec" and "b" are the corresponding recursor and the (vanilla but extended and modified) inductive type in that compiled module.
+Note that, even though "b_rec" and "b" has the same name as before, they are in the newly-compiled module so they refer to totally different things.
+
+Of course, in the implementation, our plugin will ``inherit'' the
+recursor "neg" in the surface \YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}\EDJreply{Added above. Please Check.}
 but what the plugin actually does is doing a second exhaustiveness
 checking in children family like mentioned above, with the ``same'' motive and ``same'' handler family.\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
-Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}
+Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}\EDJreply{Please check.}
 

### 2022-06-28 syntax.tex
comment

@@
 {\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-  \quad \goodTerm{\Gamma, \cL \sigma_2'}{t}{T}}
+  \quad \goodTerm{\Gamma, \cL \sigma_2'}{t}{T}\YZ{What is sigma2'?}}
 {\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}
@@
 
-Finally we introduce the inheritance judgement.
-Inheritance are judgement so naturally second class citizen data (i.e.
-not something function can return and cannot be assigned to variables), and it looks very much alike a function that transform a given linkage.
+Finally we introduce the inheritance judgement. Inheritance are
+judgement so naturally second class citizen data (i.e. not something
+function can return and cannot be assigned to variables), and it looks
+very much like a function that transform a given linkage.
 
-The syntax of the inheritance judgement is very close to how Family is defined in the surface syntax---a Family with no parent is an inheritance with empty input; an extended Family is exactly an inheritance---because the parent of an extended Family can be overridden thus inheritance should only define upon the ``interface'' instead of a concrete family.
-However, since our current formulation doesn't support further-binding (i.e.
-we cannot do inheritance inside a family), our formulation still have some distance to the real family polymorphism.
+The syntax of the inheritance judgement is very close to how Family is
+defined in the surface syntax---a Family with no parent is an
+inheritance with empty input; an extended Family is exactly an
+inheritance---because the parent of an extended Family can be overridden
+thus inheritance should only define upon the ``interface'' instead of a
+concrete family. However, since our current formulation doesn't support
+further-binding (i.e. we cannot do inheritance inside a family), our
+formulation still has some distance to the real family polymorphism.
 
-"inhid" and "inhext" are simple as expected---the former corresponds to empty inheritance, and the latter gives the programmer the ability to add new field.
+"inhid" and "inhext" are simple as expected---the former corresponds to
+empty inheritance, and the latter gives the programmer the ability to
+add new fields.
 
-"inhinh" is special because it requires user to provide a proof of ``upcasting'' from the context of the children to that of the parent.
-In other words, if I want to inherit a specific field from the parent, I have to convince the type checker that my current context can ``accommodate'' the inherited field.
+"inhinh" is special because it requires user to provide a proof of
+``upcasting'' from the context of the child to that of the parent.
+In other words, if I want to inherit a specific field from the parent, I
+have to convince the type checker that my current context can
+``accommodate'' the inherited field.
 

### 2022-06-28 coqimpl.tex
Fixing Propositional Partial Recursor

@@
 \begin{minted}[escapeinside=@@]{Coq}
-b_prec : forall P : b -> Set, P tt -> P ff
-    -> forall x : b, option (P x).@\YZ{Not clear what it takes to define/prove partial recursors. The reader needs to be told if they require breaking abstraction like recursors do.}@
+b_prec : forall R, option R -> option R -> b -> option R.@\YZ{Not clear what it takes to define/prove partial recursors.
+The reader needs to be told if they require breaking abstraction like recursors do.}\EDJreply{What do you mean?
+recursor doesn't break abstraction but break extensibility.
+Below mentioned partial recursor doesn't break extensibility}@
 \end{minted}
 with appropriate computational axioms.
-This partial recursor differs from original recursor only by adding an "option" at the end thus possible to be automatically generated (the case is similar for those computational axiom).
+This partial recursor differs from original recursor only by decorating an "option" with the return type thus possible to be automatically generated (and the case is similar for those computational axiom).
 The partiality doesn't break
 extensibility (i.e. all future extensions of "b" can support this
-"b_prec").
+"b_prec") because "b_prec" can just return "None" when bump into extended constructor in the future.
 
@@
 Definition reflect : self_B.b → option bool 
-                   := b_prec (fun _ => bool) true false.
+                   := b_prec (fun _ => bool) (Some true) (Some false).
 Definition tt_neq_ff : self_B.tt = self_B.ff → False.
 (* Because self_B.tt = self_B.ff → reflect self_B.tt = reflect self_B.ff 
-                → true = false    (by computational axiom)
-                → False           (by discriminate of bool) *)
+      → Some true = Some false    (by computational axiom)
+      → False           (by discriminate of bool and injection of Some) *)
 \end{minted}
@@
 recursor is good enough to some extent, at least for non-indexed
-inductive type, because we can ``embed'' vanilla inductive types into
+inductive type: because we can ``embed'' vanilla inductive types into
 types supporting a partial recursor. Because "reflect" is actually a left
 inverse of an injection "bool → self_B.b". A bit more formally, and
-restricting our focus to non-indexed inductive types,
+restricting our focus to non-indexed inductive types:
 
@@
       \begin{minted}[escapeinside=@@]{Coq}
-prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}@
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> R) -> R)), .., 
+prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{Yes. Discussed below. Basically constructing left inverse for simple inductive type, this weaker form is enough}@
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), .., 
       D -> option R
@@
       \end{minted}
+    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
     \end{itemize}
@@
 inductive type.\YZ{What about indexed inductive types?}
+
+We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
+\begin{minted}{Coq}
+  pdelim : ∀ (P : D -> Set), .., 
+  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), P (w b)) -> (P (dᵢ x w)))),  
+  .., ∀ (d : D) -> option (P d)
+\end{minted}
+To generalize to indexed inductive type, we base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}. Please refer to
 

### 2022-06-29 coqimpl.tex
Reply comments first so that yizhou can continue to read

@@
       \begin{minted}[escapeinside=@@]{Coq}
-prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{Yes. Discussed below. Basically constructing left inverse for simple inductive type, this weaker form is enough}@
+prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint. Let me fix other parts first}@
   (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), .., 
@@
   pdelim : ∀ (P : D -> Set), .., 
-  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), P (w b)) -> (P (dᵢ x w)))),  
+  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) -> option (P (dᵢ x w)))),  
   .., ∀ (d : D) -> option (P d)

### 2022-06-29 metatheory.tex
comments
Reply comments first so that yizhou can continue to read

@@
-We start with the formulation of the syntax, which is based on the formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT) given by~\citet{coquand2018canonicity}.
-Thus we will omit a great deal of details about MLTT here but focus on the new facility we introduced.
+% Motivation of using QIIT
+We start with the formulation of the syntax, which is based on the
+formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT)
+given by~\citet{coquand2018canonicity}. Thus we will omit a great deal
+of details about MLTT here\YZ{What details are omitted?} but focus on the new facility we introduced.
 
@@
 
-In our syntax, each judgement will actually be represented by a QIIT type, which are
-mutually recursively defined together. Quotient is used to represent
-judgemental(definitional) equality, thanks to which we can have a
-concise representation. This way, for all four kinds of judgement we are
-using, there are always equality among these data. Doing so, we can omit
-some obvious equality judgement. Most of the reasoning can be formulated
-using the \textit{algebra} of QIIT (mapping out function from this
-quotient data-type).
+% In our syntax, each judgement will actually be represented by a QIIT type, which are
+% mutually recursively defined together. Quotient is used to represent
+% judgemental(definitional) equality, thanks to which we can have a
+% concise representation. This way, for all four kinds of judgement we are
+% using, there are always equality among these data. Doing so, we can omit
+% some obvious equality judgement. Most of the reasoning can be formulated
+% using the \textit{algebra} of QIIT (mapping out function from this
+% quotient data-type).
 
-We also highlight some special feature in this style of formulation of dependent type theory: 1. this modern formulation of type theory does't have operational semantic but only equality 
-%We will recover canonicity afterwards, which implies the computational ability of our theory. 
-;
-2.
+% We also highlight some special feature in this style of formulation of dependent type theory: 1. this modern formulation of type theory does't have operational semantic but only equality 
+% %We will recover canonicity afterwards, which implies the computational ability of our theory. 
+% ;
+% 2.
 instead of using meta-level substitution, we use explicit substitution and de Bruijn indices in our intrinsic formulation.
 This is also called substitution calculus in the literature and is favoured due to its clear categorical semantic;
-3. we are dealing with quotient data, so we are actually manipulating equivalence classes of terms.
+% 3. we are dealing with quotient data, so we are actually manipulating equivalence classes of terms.
 
-We will write out the syntax of the whole type theory for completeness.
+% We will write out the syntax of the whole type theory for completeness.
 

### 2022-06-29 syntactic-example.tex
Reply comments first so that yizhou can continue to read

@@
 parts and making "O" and "S" standing alone.
-"pred" is defined as $n \mapsto ("pred" n, n)$\YZ{looks like an ill-founded definition}, and thus a predecessor.
+"pred" will map $S~n \mapsto (n, S~n)$ and $O \mapsto (O, O)$\YZ{looks like an ill-founded definition}\EDJreply{Fixed}, and thus a predecessor.
 
@@
  EndFamily. (* pN₁, handler₁, sig₃ *)
- pred@\YZ{Why is there no FRecursor?}@
-  : self[A].N → self[A].N × self[A].N
-  := ... 
+ FRecursor pred about self[A].N 
+  motive (fun _ => self[A].N × self[A].N)
+  using handler.@\YZ{Why is there no FRecursor?}\EDJreply{I wanted to save space and emphasize the type. Now fixed}@
+ (* pred
+  : self[A].N → self[A].N × self[A].N *)
   (* predT, pred₁, sig₄, sig₄', sl₄ *) 
@@
  EndFamily. (* inhᵢₙ, sig₃₁, inh₄*)
- pred 
-  : self[A₂].N → self[A₂].N×self[A₂].N 
-  := ... 
+ Inherit pred.
+
   (* sig₄₁, inh₅, sl₄₁ *)

### 2022-06-29 syntax.tex
Reply comments first so that yizhou can continue to read

@@
-Firstly, there are four standard judgement forms in MLTT,
+We formalize our language as a deep embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms.
+(Otherwise how can they be encoded in Agda?) There are four standard judgement forms in MLTT,
 well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
 $\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
-well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. Our presentation
+well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. These four judgements (QIIT types) are
+mutually recursively defined together. Our presentation
 follows the style of~\citet{kaposi2019gluing}, but for readability we
 omit universe levels attached to judgements.
-After these,
-we will have four new kinds of judgements to model our linkages.
-
-Because we formalize our language as a deep embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms (otherwise how can they be encoded in Agda?)
 %\YZ{What's the distinction of judgments vs.
@@
 presumptions of the typing rules}. For example, we don't spell out that
-the rule "Type Universe" requires a well-typed context, even though it
+the rule \ruleref{Type Universe} requires a well-typed context, even though it
 does implicitly, because it is not possible to have ill-formed
@@
 
-Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the
+Quotient is used to represent
+judgemental(definitional) equality (for each QIIT type), thanks to which we can have a
+concise representation. This way, for all four kinds of judgement we are
+using, there are always equality among these data. Doing so, we can omit
+some obvious equality judgement. Most of the reasoning can be formulated
+using the \textit{algebra} of QIIT (mapping out function from this
+quotient data-type).
+
+Now we highlight some other special feature contrary to more conventional setting: 1.
+this modern formulation of type theory does't have operational semantic but only equality\footnote{Later we will prove canonicity which can imply a certain level of computational ability of our theory}
+%We will recover canonicity afterwards, which implies the computational ability of our theory. 
+;
+2.
+instead of using meta-level substitution, we use explicit substitution and de Bruijn indices in our intrinsic formulation.
+This is also called substitution calculus in the literature and is favoured due to its closedness categorical semantic;
+3. we are dealing with quotient data, so we are actually manipulating equivalence classes of terms.
+
+We quickly review the substitution calculus,
+
+Based upon the basic MLTT,
+we will have four new kinds of judgements to model our linkages. Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the
 most special one because it is actually just a shorthand of term

### 2022-06-29 syntax.tex
Adding intro for substitution calculus

@@
-We formalize our language as a deep embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms.
+We formalize our language as a deep embedding in a type theory with QIIT types.
+Every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms.
 (Otherwise how can they be encoded in Agda?) There are four standard judgement forms in MLTT,
@@
 
-We quickly review the substitution calculus,
+We quickly review the substitution calculus.
+Substitution is very (if not the most) important because function is first-class citizen, and can be applied (almost) everywhere and return (almost) everything, thus we require the resulting substitution to commutate with most (if not all) type/term constructs.
+
+A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
+A substitution $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly understood in the way that $\gamma$ will output a \textbf{series} of (dependently-typed related) terms of types $\Delta$, parametrized by $\Gamma$ (i.e.
+we can consider there are holes in $\gamma$ of type $\Gamma$).
+Thus (1) \ruleref{Sub Extend} looks like concatenating one series of terms with a new term (the weird looking type of $\goodTerm{\Gamma}{t}{A[\sigma]}$ is because we need $A$ to appear in the context of $\Delta$ so that we can use \ruleref{Context Extension} to create new context); (2) we have \ruleref{Proj Subst-1} to extract a subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why substitution really looks like a (deeply-nested) pair.
+
+Since the $\Gamma$ before the $\vdash$ is always indicating the types of ``holes'', this explains the \ruleref{Term Subst} and \ruleref{Type Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution $\goodSub{\Gamma}{\gamma}{\Delta}$, we will have a new term $\goodTerm{\Gamma}{t[\gamma]}{A[\gamma]}$ after the substitution---this substitution can be understood as ``filling all the holes''.
+Once all the holes are filled with a given substitution, the new holes all come from the subsitution itself.
+Thus consecutive subsitution can be composed as \ruleref{Sub Comp}, because all the holes are always from the last substitutions.
+
+Note that, for simplicity, we will (a lot of time) writing $\pi_1$ as a \textbf{shorthand} of $\goodSub{\Gamma,A}{\pi_1~"id"}{\Gamma}$, and similarly another shorthand for $\goodTerm{\Gamma, A}{\pi_2 := \pi_2~"id"}{A[\pi_1]}$.
 
@@
 
-Now we show selected typing rules: we will have dependent function types
+We will have dependent function types
 and dependent pair types. Dependent pair type is used to model Coq
@@
 The substitution law for these two
-require \textit{de bruijn lifting}: for
-$\goodTerm{\Gamma}{\sigma}{\Delta}$, we define its de Bruijn lifting
-$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow}{\Delta, A}$ given
-$\goodType{\Delta}{A}{}$.\YZ{What is A[sigma] and what is de Bruijn lifting?
+require \textit{de bruijn lifting}: for a given substitution
+$\goodTerm{\Gamma}{\sigma}{\Delta}$ and a type
+$\goodType{\Delta}{A}{}$, we define its de Bruijn lifting
+$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ .
+Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.\YZ{What is A[sigma] and what is de Bruijn lifting?
 Seems like if the reader is not familiar with the substitution calculus, then they will get lost starting here.
-Broadly speaking, a gentle intro to the intrinsic setting would help the reader a great deal.}
+Broadly speaking, a gentle intro to the intrinsic setting would help the reader a great deal.}\EDJreply{A quick intuitive explanation of substitution-calculus is given above.
+Please check.
+}
 
@@
 $$
-\Rule[name=Empty Subst]
+\Rule[name=Ept Subst]
 {}{\goodSub{\Gamma}{\epsilon}{\cdot}}
 \quad
-\Rule[name=Id Subst]
+\Rule[]
 {}{\goodSub{\Gamma}{"id"}{\Gamma}}
 \quad
-\Rule[name=Subst Extension]
+\Rule[name=Sub Comp]{
+  \goodSub{\Delta}{\delta}{\Theta}
+  \quad \goodSub{\Gamma}{\gamma}{\Delta} 
+}{\goodSub{\Gamma}{\delta \circ \gamma}{\Theta}}
+\Rule[name=Sub Extend]
 {\goodSub{\Gamma}{\sigma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}

### 2022-06-29 syntax.tex
Replying part of the comments.

@@
 Due to our intrinsic setting, \textbf{we can and will omit a lot of
-presumptions of the typing rules}. For example, we don't spell out that
+presumptions of the typing rules}. This can work because we are using QIIT and everything is intrinsically well-typed. For example, we don't spell out that
 the rule \ruleref{Type Universe} requires a well-typed context, even though it
-does implicitly, because it is not possible to have ill-formed
-contexts in our intrinsic setting.
+does implicitly, \textbf{because it is not possible to have ill-formed (not-well-typed)
+contexts in our intrinsic setting}.
+This is a bit unusual for conventional extrinsic setting where they will put well-formedness of context everywhere in a lot of term judgements.
+We consider this as one advantage of using intrinsic setting---this well-formed (well-typed) context is usually attached to some other presumptions thus inferrable and can be considered as ``boilerplate code''.
+Removed from these boilerplate code, our intrinsic setting provides a more concise representation.
 
@@
 most special one because it is actually just a shorthand of term
-judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.\YZ{You haven't introduced the notation A[pi_1] yet at this point. And what is A[pi_1]?}
+judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.\YZ{You haven't introduced the notation A[pi_1] yet at this point.
+And what is A[pi_1]?}\EDJreply{added for the intro.
+It is a shorthand for pi_1(id)}
 
@@
 Besides, are these rules used anywhere in the following?
-Seems a bit unusual to have well-formedness rules for judgments per se.}
+Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes.
+Right column show all the well-formedness rule for all the judgement.
+That is unusual because it is a direct translation from QIIT.
+}
   \end{minipage}
@@
 
-\caption{Well-formedness rules of four new judgements}
+\caption{Well-formedness rules of the judgements}
 \end{figure}
@@
 {}
-{\goodType{\Gamma}{\cU_j}{j + 1}}
+{\goodType{\Gamma}{\cU}{j + 1}}
 \quad 
@@
 {\goodSub{\Gamma}{\gamma}{\Delta}}
-{\goodType{\Gamma}{\cU_j[\gamma] \equiv \cU_j }{j + 1} \quad
+{\goodType{\Gamma}{\cU[\gamma] \equiv \cU }{j + 1} \quad
   \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
@@
 {\goodType{\Gamma}{T}{j}}
-{\goodTerm{\Gamma}{"c" \ T}{\cU_j}\YZ{what is c? Why is this rule called Univ-1?Didn't you write that you omit universe levels?}
+{\goodTerm{\Gamma}{"c" \ T}{\cU}\YZ{what is c?
+Why is this rule called Univ-1?Didn't you write that you omit universe levels?}\EDJreply{I just kind of afraid they will think there is one universe.
+Let me remove all the indices}
 }\quad
 \Rule[name=Univ-2]
-{\goodTerm{\Gamma}{T}{\cU_j}}
+{\goodTerm{\Gamma}{T}{\cU}}
 {\goodType{\Gamma}{El \ T}{j}\YZ{'El T' isn't explained only until it makes a second appearance in Ind-Term.}}

### 2022-06-29 syntax.tex
All the comments should be replied now

@@
 
+Because in dependent type, function now can return type, of which the type is ``type of type'', \textit{the universe} $\cU$.
+We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
+distinguish types from their \textit{codes or names}, and thus we use
+$El\ T$ as the type given \textit{type name} $T : \cU$. In this case, (1) $\cU$ is the type including the \textit{codes or names} of types , and (2) $T$ is
+not allowed to locate at type position (i.e.
+at the right-most position of a term judgement) but $El \ T$ is.
+"c" is the ``inverse'' of $El$ that getting \underline{"c"}ode of a type, with the following two equations $"c" (El~T) \equiv T$ and $El ("c"~T) \equiv T$.
+Recall that we are actually using predicative universe and thus we actually have different level of universes $\cU_j$.
+Here level omitted for conciseness.
+
   \label{fig:rules:well-typed-ctx}
@@
 $$
-\Rule[name=Univ-1]
+\Rule[]
 {\goodType{\Gamma}{T}{j}}
 {\goodTerm{\Gamma}{"c" \ T}{\cU}\YZ{what is c?
-Why is this rule called Univ-1?Didn't you write that you omit universe levels?}\EDJreply{I just kind of afraid they will think there is one universe.
-Let me remove all the indices}
+Why is this rule called Univ-1?Didn't you write that you omit universe levels?}\EDJreply{I think c stands for code.
+I am just following convention.
+Now I remove their name because that is a bad name.
+About the level, I just kind of afraid they will think there is one universe.
+Now I've removed all the indices}
 }\quad
-\Rule[name=Univ-2]
+\Rule[]
 {\goodTerm{\Gamma}{T}{\cU}}
-{\goodType{\Gamma}{El \ T}{j}\YZ{'El T' isn't explained only until it makes a second appearance in Ind-Term.}}
+{\goodType{\Gamma}{El \ T}{j}\YZ{'El T' isn't explained only until it makes a second appearance in Ind-Term.}\EDJreply{I have made El and universe explanation earlier, please check.}}
 \quad
@@
 BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general.
-I just didn't write out the two equations c (El T) = T, El (c T) = T}\YZreply{Seems harmless to list these equations and move the reference to 'type universe a la Tarski' earlier.}
-We follow the style of \textit{type universe à la Tarski} and
-distinguish types from their \textit{codes or names}, and thus we use
-$El\ T$ as the type given type name $T : \cU$---in these cases, $T$ is
-not allowed to locate at type position (i.e. at the right-most position
-of a term judgement) but $El \ T$ is.
+I just didn't write out the two equations c (El T) = T, El (c T) = T}\YZreply{Seems harmless to list these equations and move the reference to 'type universe a la Tarski' earlier.}\EDJreply{Done.
+Please check.}
 
@@
 {\goodSig{\Gamma}{\sigma}{n} 
- \quad \goodSeal{\Gamma}{f}{\sigma}{A}\YZ{How should I understand 'f' and 'A'?}
+ \quad \goodSeal{\Gamma}{f}{\sigma}{A}\YZ{How should I understand 'f' and 'A'?}\EDJreply{`f' is the seal, 'A' is the result of sealing.
+I think I wrote down the motivation of sealing...
+Is it still too hard?
+Do you think I need to think of a more friendly way?}
  \quad \goodType{\Gamma, A}{T}{i}}
@@
 to retrieve the corresponding types of linkage of a given signature, and
-$\cC$ to get the compiled linkage type (an existential type).\YZ{Why is Csigma needed? How is it different from Lsigma?}
+$\cC$ to get the compiled linkage type (an existential type).\YZ{Why is Csigma needed?
+How is it different from Lsigma?}\EDJreply{It corresponds to the compilation of the family ``type'', which is part of the implementation detail.
+Basically we know every field of a family is dependent on the former fields.
+The implementation detail here is that, every field of a family is dependent on the compiled module of the former family (and thus includes all the former fields)} $\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, accoding to \ruleref{Compile}, is really a deeply nested dependent pair.
+This is reflecting the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}.
+Thus we need $\cC\sigma$ in our context.
 
@@
 }
-{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}\YZ{What is R[pi1 circ pi1]?}}
+{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}\YZ{What is R[pi1 circ pi1]?}\EDJreply{It is the composition of the substitution.
+Thanks for pointing out now I included circ into the typesetting} }
 \quad 
@@
 
-In our meta-theory, we only formulate non-dependent eliminator, i.e., the
-\textit{recursor}.\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}
+In our meta-theory, we only formulate non-dependent eliminator for simplicity, i.e., the
+\textit{recursor}.\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}\EDJreply{I don't think there is technical difficulties to do that.
+The formulation of course will be much more horrifying if use dependent eliminator here.
+}
 Each recursor is constructed by a linkage containing
@@
 handlers. All these rules satisfy substitution laws and omitted here.
-\YZ{Seems like you omit many things here, which causes hardship for the reader (at least for me!). I think it's better to list all the rules and decide later what to throw to the appendix.}
+\YZ{Seems like you omit many things here, which causes hardship for the reader (at least for me!).
+I think it's better to list all the rules and decide later what to throw to the appendix.}\EDJreply{Do you mean you want to see substitution laws and beta and eta laws?
+They are just commutating as expected I can write down them if you really want.
+And what else do you think is helpful}
 
@@
 Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
-Seal Γ A τ = Tm (Γ, 𝓛σ) A[p₁]@\YZ{tau?}@
+Seal Γ A τ = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi. I fixed it now}@
 
@@
 a shorthand of term judgement $\goodTerm{\Gamma , \cC
-\sigma}{\_}{A[\pi_1]}$,\YZ{Not quite. Previously you used cL I think.}
+\sigma}{\_}{A[\pi_1]}$,\YZ{Not quite. Previously you used cL I think.}\EDJreply{Sorry. It is fixed. It should be $\cC$ in the context.}
 a special form of \textit{local morphism}~\cite{abbott2003category}.
@@
 \sigma_1 \vdash t : T$ in the parent family will generally not be able
-to be inherited to the children once we override the inductive type. In
+to be inherited to the children once we override the inductive type $\bW
+\sigma_1$ with $\bW \sigma_2$. In
 other words, there is no way to ``transplant'' $t$ into the context $X_2
-: \bW \sigma_2$.\YZ{Is sigma_2 the overriding of sigma_1? The reader has to guess.}
+: \bW \sigma_2$.\YZ{Is sigma_2 the overriding of sigma_1? The reader has to guess.}\EDJreply{Thanks. Fixed}
 This makes all the fields after this inductive type
@@
 Sealing\YZ{It occurs to me that you use the word "sealing" to mean different (though tangentially related) things.
-Previously you used it as in "sealed families".} is used to exploit this fact---with sealing, given $X_1 : \bW
+Previously you used it as in "sealed families".}\EDJreply{They are actually the same thing.
+This sealing in metatheory is much stronger than that of the implemented plugin.
+Downside of the strong sealing is that if the user really use this version of sealing, the user will have to annotate what the sealed family type is.
+Now in the plugin, the sealed family type is almost fixed (directly inferred using a fixed rules).
+So basically this complete version of sealing (in metatheory) can deduce that of seal (in plugin implementation) } is used to exploit this fact---with sealing, given $X_1 : \bW
 \sigma_1$, we can add well-typed terms $X_1 : \cU \vdash t : T$, into
@@
 \quad \goodType{\Gamma, A_2}{T_2}{}
-  \quad \goodTerm{\Gamma, \cL \sigma_2'}{t}{T_2}}
+  \quad \goodTerm{\Gamma, A_2}{t}{T_2}}
 {\goodInh{\Gamma}{"inhov" \ h \ t}{(\nu^+ \  \sigma_1 \  T_1)}{(\nu^+ \  \sigma_2\  T_2)}}
@@
 {\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-  \quad \goodTerm{\Gamma, \cL \sigma_2'}{t}{T}\YZ{What is sigma2'?}}
+  \quad \goodTerm{\Gamma, A_2}{t}{T}\YZ{What is sigma2'?}\EDJreply{Sorry this should be A2, i.e.
+the result of sealing.
+Earlier version sigma2' is the result of sealing.
+It was type-set before the last meta-theory change}}
 {\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}

### 2022-06-29 syntax.tex
Fix typo.

@@
 most special one because it is actually just a shorthand of term
-judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.\YZ{You haven't introduced the notation A[pi_1] yet at this point.
+judgement $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$.\YZ{You haven't introduced the notation A[pi_1] yet at this point.
 And what is A[pi_1]?}\EDJreply{added for the intro.
@@
 possible.
+This sealing is the general version of the sealed family implemented in the plugin.
+Our plugin will deduce a sealed ``interface'' for a sealed family using a fixed set of rules but here sealing in metatheory can be arbitrary local morphism.
 

### 2022-06-29 lang-design.tex
Reply comments first so that yizhou can continue to read
Save. Rewriting

@@
+\subsection{Challenges}
+
+We want our metatheory mechanization to allow adding new function/proving new lemmas (Expression Problem), and adding new constructors for inductive type at the same time.
+Coq support the former but not the later---because when an existent inductive type is extended with new constructor, we have to modify the pattern matching in the existent functions.
+This motivates us to add inheritance/Family Polymorphism into Coq.
+
+However, to add inheritance into a dependent type system like Coq is not simple and there are some challenges we need to resolve.
+
 \subsection{Design Space}\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).

### 2022-06-29 syntax.tex

@@
 judgements and their Agda counterparts to make things more accessible. Of
-course, a well-formed judgement is not necessarily derivable, just like
-a well-typed QIIT type is not necessarily inhabited.
+course, a well-formed judgement is not necessarily derivable, just as
+a well-formed QIIT type is not necessarily inhabited.
 
 Due to our intrinsic setting, \textbf{we can and will omit a lot of
-presumptions of the typing rules}. This can work because we are using QIIT and everything is intrinsically well-typed. For example, we don't spell out that
-the rule \ruleref{Type Universe} requires a well-typed context, even though it
-does implicitly, \textbf{because it is not possible to have ill-formed (not-well-typed)
-contexts in our intrinsic setting}.
-This is a bit unusual for conventional extrinsic setting where they will put well-formedness of context everywhere in a lot of term judgements.
-We consider this as one advantage of using intrinsic setting---this well-formed (well-typed) context is usually attached to some other presumptions thus inferrable and can be considered as ``boilerplate code''.
-Removed from these boilerplate code, our intrinsic setting provides a more concise representation.
+presumptions of the typing rules}. This can work because we are using
+QIIT and everything is intrinsically well-formed. For example, we don't
+spell out that the rule \ruleref{Type Universe} requires a well-formed
+context, even though it does implicitly, \textbf{because it is not
+possible to have ill-formed contexts in our intrinsic setting}. This is
+a bit unusual for conventional extrinsic setting where they will put
+well-formedness of context everywhere in a lot of term judgements. We
+consider this as one advantage of using intrinsic setting---this
+well-formed context is usually attached to some other presumptions thus
+inferrable and can be considered as ``boilerplate code''. Free of
+such boilerplate code, our intrinsic setting provides a more concise
+representation.
 
-Quotient is used to represent
-judgemental(definitional) equality (for each QIIT type), thanks to which we can have a
-concise representation. This way, for all four kinds of judgement we are
-using, there are always equality among these data. Doing so, we can omit
-some obvious equality judgement. Most of the reasoning can be formulated
+Quotient is used to represent definitional equality (aka judgemental
+equality) between QIIT types concisely.
+This way, for all four kinds of judgements we are
+using, there is always equality among these data. Doing so, we can omit
+some obvious equality judgements. Most of the reasoning can be formulated
 using the \textit{algebra} of QIIT (mapping out function from this
@@
 
-Now we highlight some other special feature contrary to more conventional setting: 1.
-this modern formulation of type theory does't have operational semantic but only equality\footnote{Later we will prove canonicity which can imply a certain level of computational ability of our theory}
+Now we highlight some other special features that differ from a more
+conventional setting: (1) This modern formulation of type theory does not
+have an operational semantics but only equality.\footnote{Later we will prove
+canonicity which can imply a certain level of computational ability of
+our theory}
 %We will recover canonicity afterwards, which implies the computational ability of our theory. 
-;
-2.
-instead of using meta-level substitution, we use explicit substitution and de Bruijn indices in our intrinsic formulation.
-This is also called substitution calculus in the literature and is favoured due to its closedness categorical semantic;
-3. we are dealing with quotient data, so we are actually manipulating equivalence classes of terms.
+(2) Instead of using meta-level substitution, we use explicit
+substitution and de Bruijn indices in our intrinsic formulation. This is
+also called substitution calculus in the literature and is favoured due
+to its closedness categorical semantic.
+(3) We are dealing with quotient data, so we are actually manipulating
+equivalence classes of terms.
 
 We quickly review the substitution calculus.
-Substitution is very (if not the most) important because function is first-class citizen, and can be applied (almost) everywhere and return (almost) everything, thus we require the resulting substitution to commutate with most (if not all) type/term constructs.
+Substitution is very (if not the most) important because function is
+first-class citizens, and can be applied (almost) everywhere and can return
+(almost) everything, thus we require the resulting substitution to
+commute with most (if not all) type/term constructs.
 
@@
     &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
-Besides, later you also use gamma to range over subsitutions and s to range over W-types, which add to the confusion.}
+Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}
   \end{align*}

### 2022-06-29 coqimpl.tex
First draft of challenges (replacing design space section)

@@
 compiled modules and module types. We can achieve incremental
-checking and overriding as in \ref{langdesign-req0}---for example, if we
+checking and overriding as in \ref{chg:software-engineering}---for example, if we
 want to override "Field a", we simply replace that compiled module with
@@
 The resulting compiled module is just like mundane Coq's module and 
-thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.
+thus maintains all the computational mechanism from Coq, resolving \ref{chg:software-engineering}.
 
@@
 has already allowed fields to be type-dependent on former fields thus fulfilling
-\ref{langdesign-req1}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
+\ref{chg:definition-relevant-reasoning}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
 the concrete definition of a given field (especially when we are proving
@@
 We support ``extensible'' inductive types from
-\ref{langdesign-req2} in Coq to some extent.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types.
+\ref{chg:extensible-inductive-type} in Coq to some extent.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types.
 It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it.
@@
 
-To construct a recursor as in \ref{langdesign-req3}, we need to create for each
+To construct a recursor as in \ref{chg:extensible-inductive-type}, we need to create for each
 constructor a handler (function) and aggregate them into a family, and thus that family (e.g.
@@
 
-However, we later still need the concrete eliminator "b" for doing type-checking (as in \ref{langdesign-req3}), thus the information doesn't totally disappear but store elsewhere.
+However, we later still need the concrete eliminator "b" for doing type-checking (as in \ref{chg:extensible-inductive-type}), thus the information doesn't totally disappear but store elsewhere.
 Here we show one exemplar extraction of the recursor to \mintinline{Coq}{Set} as module "b_rec_12".
@@
 Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
-not expressive enough to really fulfill \ref{langdesign-req2}. For
+not expressive enough to really fulfill \ref{chg:extensible-inductive-type}. For
 example, we cannot prove the constructors are injective, i.e.,
@@
 % However, to verify and prove the recursor and partial recursor and their
-% computational rules (as in \ref{langdesign-req4}), our plugin has to
+% computational rules (as in \ref{chg:definition-relevant-reasoning}), our plugin has to
 % compile the extensible family into a non-extensible Coq Module and define
@@
 % generate proof for recursors and partial recursors and their
-% computational rules (as in \ref{langdesign-req4}), and thus making
+% computational rules (as in \ref{chg:definition-relevant-reasoning}), and thus making
 % plugin development easier.
@@
 I am not sure this kind of hijacking can happen.}
-This can fulfill \ref{langdesign-req5}.
+This can fulfill \ref{chg:software-engineering}.
 
-% To fulfill \ref{langdesign-req5}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
+% To fulfill \ref{chg:software-engineering}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
 
@@
 % due to the invisibility, unless we compile the whole family. Similarly,
-% customized tactic expressions as required by \ref{langdesign-req5} cannot
+% customized tactic expressions as required by \ref{chg:software-engineering} cannot
 % use lemmas or theorems in the fields of the current defining family as
@@
 % following defining field. Doing so, we can define customized tactic as
-% in \ref{langdesign-req5} that uses lemmas proved in the family.
+% in \ref{chg:software-engineering} that uses lemmas proved in the family.

### 2022-06-29 lang-design.tex
First draft of challenges (replacing design space section)

--- 2022-06-27 lang-design.tex
@@
-\subsection{Design Space}\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).
-I think most readers will benefit from a recap of this encoding.\\
-Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
-}
+\subsection{Challenges}
 
-When adapting Family Polymorphism into dependent types, we choose to
-focus only on the essence of family (and inheritance)
-structure in \citet{zm2017}, and thus a lot of unrelated features, like
-Interface, will be removed. In this case, it will look like module with
-late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
-need it to have good compatibility with inductive types, because we
-don't want to retain the idiomatic Coq programming experience
-%\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness.
-I rephrase it now, please check.}
-(and its tactic programming).
-
-Let's start with a basic example---STLC and its extension in
-\cref{fig:STLC-example}---to consider what kind of features are required
-and how much of them can be supported by family polymorphism.
-% add the example code here
+We want our metatheory mechanization to allow adding new function/proving new lemmas, and adding new constructors for inductive type at the same time  (Expression Problem~\cite{wadler-ep}).
+Coq support the former but not the later---because when an existent inductive type is extended with new constructor, we have to modify the pattern matching in the existent functions.
+This motivates us to add inheritance/Family Polymorphism into Coq.
 
@@
   | lam : id -> term -> term ...
+  | tunit : term 
 Fixpoint subst 
@@
   | tif : term -> term -> term -> term
+
 Extend Fixpoint subst (* Need to handle new term *)
@@
 
-First and foremost, we don't want to throw away some basic good
-properties of Coq. \textlabel{Req~(0A)}{langdesign-req0}~\textbf{we
-want to retain incremental (modular) type checking}. Notice that, the
-modularity here in Coq is a bit different from other languages, because
-Coq supports \textit{interactive} theorem proving, so we actually need
-\textit{statement-wise} incremental type checking, not only for avoiding
-re\-compilation, but also to enable immediate feedback and incremental
-type-checking for the Coq developer. We don't want our family to ruin
-this conventional routine of interactive development.
-\textlabel{Req~(0B)}{langdesign-req0b}~\textbf{we also want to keep the
-computational ability of Coq when using families}. Coq, based on constructive
-logic, can consider proof as programs. 
-We don't want our Family facility to ruin this when incorporated with 
-the remaining parts of Coq: 
-the developers should be able to project fields of families 
-as first-class value just like how they can project fields of Coq's Module.
+However, to add inheritance into a dependent type system like Coq is not simple and there are various challenges and tension we need to resolve.
+
+\textlabel{Challenge (1)}{chg:extensible-inductive-type}~\textbf{Extensible Inductive Type vs.
+Exhaustiveness of Inductive Reasoning}.
+Unlike most general purposed programming languages, Coq, as a dependent type theory, is a total language and thus each pattern matching cases must fully ``cover'' an inductive type.
+This process is \textit{dynamic} for inheritance---when inhertiance extend the inductive type in the children family but the user forget to extend the corresponding inductive reasoning, our compiler need to signal an error correspondingly, or better---guide the user to fill the incomplete inductive reasoning.
+For example, in \cref{fig:STLC-example} when "term" is extended with "tt" in the children family "STLC_bool", our compiler need to make sure "subst" and "subst_lemma" are also extended correctly otherwise signaling an error.
+Additionally, our compiler must support direct inheritance (and overriding) of the old case handlers, to waive the boilerplate code.
+
+\textlabel{Challenge (2)}{chg:definition-relevant-reasoning}~\textbf{Definition-awared Reasoning vs.
+Overridability}.
+In family polymorphism for general purposed programming language, overriding is allowed to happen anywhere.
+But the problem becomes much more subtle in the case of dependent type---the later fields can have a type dependent on the former fields!
+Consider a small example where we have consecutive fields \mintinline{Coq}{{..
+a : nat := 1; a-def : Eq a 1 := eq_refl; ..}}, should we support override "a"?
+If we do support it, how can our compiler detect that "a-def" is not inheritable and signal an error when such thing happens?
+
+A similar example is about "subst" and "subst_lemma" in \cref{fig:STLC-example}.
+For "subst_lemma" to hold throughout two families, we shouldnot allow overriding on the definition of "subst".
+For example, we should still stick with the fact that "subst tunit x y = tunit", as this fact is used when proving "subst_lemma" and we want this lemma to be inherited to the children family.
+
+These two examples both show the incompatibility between exposing the concrete definition of a field for reasoning (this second example even involves the definition of pattern matching) and overridability.
+We need to be careful about this tension when designing the language.
+
+\textlabel{Challenge (3)}{chg:consistency}~\textbf{Consistency vs.
+Self Parametrization}.
+The metatheoretical formulation in~\citet{zm2017} for family polymorphism is very natural as they consider extension of the family as the concatenation of a new linkage.
+However, this formulation is not directly adaptable for dependent type theory because it can easily encode non-termination (mutual calling of two function member) sabotaging consistency by the presence of "self".
+This conflict between self parametrization and consistency will be another challenge---we need change the original formulation of family and inheritance for dependent type theory.
+
+\textlabel{Challenge (4)}{chg:software-engineering}~\textbf{Engineering and Ecosystem}.
+Not only for the metatheory, we also want our implementation of family to have good compatibility with the rest of Coq.
+For example, we cannot compromise \textbf{tactic programming} as it is already a widely accepted part of Coq programming.
+Additionally, the interactive development style requires \textbf{incremental type checking}---our compiler has to be able to type check each Coq Vernacular command.
+What's more, we want our family ``term'' to incorporate Coq \textbf{without breaking its other functionality}.
+For example, the user may want to project the family member anywhere inside an arbitrary Gallina term, that is later being computed and normalized.
+However, families themselves can be second class citizens just like \citet{zm2017}.
+
+\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).
+I think most readers will benefit from a recap of this encoding.\\
+Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
+}
+
+% \subsection{Design Space}
+
+% When adapting Family Polymorphism into dependent types, we choose to
+% focus only on the essence of family (and inheritance)
+% structure in \citet{zm2017}, and thus a lot of unrelated features, like
+% Interface, will be removed. In this case, it will look like module with
+% late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
+% need it to have good compatibility with inductive types, because we
+% don't want to retain the idiomatic Coq programming experience
+% %\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness.
+I rephrase it now, please check.}
+% (and its tactic programming).
+
+% Let's start with a basic example---STLC and its extension in
+% \cref{fig:STLC-example}---to consider what kind of features are required
+% and how much of them can be supported by family polymorphism.
+% add the example code here
+
+% First and foremost, we don't want to throw away some basic good
+% properties of Coq. \textlabel{Req~(0A)}{chg:software-engineering}~\textbf{we
+% want to retain incremental (modular) type checking}. Notice that, the
+% modularity here in Coq is a bit different from other languages, because
+% Coq supports \textit{interactive} theorem proving, so we actually need
+% \textit{statement-wise} incremental type checking, not only for avoiding
+% re\-compilation, but also to enable immediate feedback and incremental
+% type-checking for the Coq developer. We don't want our family to ruin
+% this conventional routine of interactive development.
+% \textlabel{Req~(0B)}{chg:software-engineeringb}~\textbf{we also want to keep the
+% computational ability of Coq when using families}. Coq, based on constructive
+% logic, can consider proof as programs. 
+% We don't want our Family facility to ruin this when incorporated with 
+% the remaining parts of Coq: 
+% the developers should be able to project fields of families 
+% as first-class value just like how they can project fields of Coq's Module.
 
 % We want to reason about fields in a family.
-\textlabel{Req (1)}{langdesign-req1}:\textbf{ we want to be able to
-reason about fields in a family}, just like in a module, where a field
-can reason about its former fields. More generally, we want to allow a
-later field to be type-dependent on the former fields. In the example of
-\cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
-this feature.
+% \textlabel{Req (1)}{chg:definition-relevant-reasoning}:\textbf{ we want to be able to
+% reason about fields in a family}, just like in a module, where a field
+% can reason about its former fields. More generally, we want to allow a
+% later field to be type-dependent on the former fields. In the example of
+% \cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
+% this feature.
 
 % we want extensible inductive type
-\textlabel{Req (2)}{langdesign-req2}:\textbf{ we need extensible
-inductive types}, to extend "term" and "ty" in \cref{fig:STLC-example}.
-Extensible inductive types are not supported by \citet{zm2017}, so
-we need further consideration about it on both implementation and
-metatheory.
+% \textlabel{Req (2)}{chg:extensible-inductive-type}:\textbf{ we need extensible
+% inductive types}, 
+% to extend "term" and "ty" in \cref{fig:STLC-example}.
+% Extensible inductive types are not supported by \citet{zm2017}, so
+% we need further consideration about it on both implementation and
+% metatheory.
 
-\textlabel{Req (3)}{langdesign-req3}: \textbf{we also want extensible
-pattern matching and induction reasoning}, to extend "subst" in
-\cref{fig:STLC-example}.
-There are actually two kinds of ``pattern matching'', one for data and
-the other for induction reasoning, i.e., one uses the eliminator to
-\mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the
-eliminator to \mintinline{Coq}{Prop}. Luckily, in this setting,
-extensible pattern matching can be easily expressed with family
-extension---we just aggregate all case handlers of pattern matching
-into one family (as a bunch of functions), and then one family can
-encode one pattern matching, and family inheritance can express adding
-case handlers. Then we just need to introduce a primitive that will
-``wrap'' that family into a recursive function. Induction reasoning can
-be handled in an almost identical way.
+% \textlabel{Req (3)}{chg:extensible-inductive-type}: \textbf{we also want extensible
+% pattern matching and induction reasoning}, to extend "subst" in
+% \cref{fig:STLC-example}.
+% There are actually two kinds of ``pattern matching'', one for data and
+% the other for induction reasoning, i.e., one uses the eliminator to
+% \mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the
+% eliminator to \mintinline{Coq}{Prop}. Luckily, in this setting,
+% extensible pattern matching can be easily expressed with family
+% extension---we just aggregate all case handlers of pattern matching
+% into one family (as a bunch of functions), and then one family can
+% encode one pattern matching, and family inheritance can express adding
+% case handlers. Then we just need to introduce a primitive that will
+% ``wrap'' that family into a recursive function. Induction reasoning can
+% be handled in an almost identical way.
 
 % we want to be able to reason pattern matching as well
-However, there is still difference between data recursion and induction
-reasoning, because the former one is \textit{proof-relevant}. This
-difference leads to another issue: \textlabel{Req (4)}{langdesign-req4}:
-\textbf{we need to reason about (the computation about) extensible
-pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
-When proving "subst_lemma", we have to know "subst" is invariant on
-"tt", i.e., \mintinline{Coq}{(subst i x tt) = tt}. This kind of
-information requires exposing \textit{computational rules} from the
-recursors.\YZ{Seems like what's missing here is to
+% However, there is still difference between data recursion and induction
+% reasoning, because the former one is \textit{proof-relevant}. This
+% difference leads to another issue: \textlabel{Req (4)}{chg:definition-relevant-reasoning}:
+% \textbf{we need to reason about (the computation about) extensible
+% pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
+% When proving "subst_lemma", we have to know "subst" is invariant on
+% "tt", i.e., \mintinline{Coq}{(subst i x tt) = tt}. This kind of
+% information requires exposing \textit{computational rules} from the
+% recursors.
+\YZ{Seems like what's missing here is to
 convince the reader why exposing the definition of subst could be a
@@
 % we want tactic programming and certain level of automation
-Finally, \textlabel{Req (5)}{langdesign-req5}: \textbf{we desire tactic
-programming} to relieve us from the hassle of direct manipulation of proof terms.
+% Finally, \textlabel{Req (5)}{chg:software-engineering}: \textbf{we desire tactic
+% programming} to relieve us from the hassle of direct manipulation of proof terms.
 

### 2022-06-30 coqimpl.tex
Fixing coqimpl according to challenge section Move lang-design into coqimpl.

@@
+\subsection{Language Design}
+Inspired by our above challenges and Family Polymorphism, we propose
+the following extension to Coq's Vernacular Commands.
+% basically write down the specification of surface Coq commandline
+%   and then explain  how the language design makes the programmer’s life easier
+\begin{align*}
+  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
+  \quad |\quad "EndFamily" \\
+  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
+  & |\quad "Extend"~"Family"~\langle ident\rangle \\ 
+  & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
+  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
+  & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
+  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad | \quad ...
+\end{align*}
+
+These are some core extensions we proposed to support \cref{fig:STLC-example}.
+We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
+"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.}
+We can define nested families.
+
+To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness.
+The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family.
+Then
+"FRecursor" will be used to construct a recursor using the handler family.
+For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
+"handlers" will generate a new field "recur" that is doing induction
+(using elimination principle of "T") using motive "f", and the handlers
+are located in the family "handlers". This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family.
+
+We can inherit a top-level family "parent" using \mintinline{Coq}{Family
+... extends parent} as a new top-level family. As expected, "Override"
+will override a field from the parent. "Inherit" can be used to inherit
+nested families, inductive types, and fields. The fields inside a family
+are order-sensitive because we have dependent types here. So "Inherit
+Until" is used if the user wants to inherit a range of fields. If the
+user wants to extend an inductive type or a nested family from the
+parent, then "Extend FInductive" or "Extend Family" is used. Thus nested
+family can only be inherited or extended together with its outer family
+(and thus top level family), thus our support of family polymorphism is
+incomplete.
+
+% This incomplete support can already be useful for extending functions.
+% Take \cref{fig:plugin-example2} as example, if we want to extend boolean
+% inductive type and its negation, we only need to specify the new
+% constructor for boolean and how negation deals with the new constructor.
+% As a result, the programmer does not have to duplicate code---otherwise
+% the user will need to copy and paste the whole original negation
+% definition and add a new pattern matching case.
+
+\subsection{Plugin Implementation}
+
 In this section, we describe how we can implement the proposed Vernacular commands.
@@
 
-\textbf{Family compiled into module.} This is the main idea of compiling
-family components into Coq's primitives: we compile families into Coq
-modules, family types into module types, and the context of a given term
-into parameters of the module. Unlike module, we need to faithfully
+\textbf{Family compiled into module.} We support family by compiling
+family components into Coq modules, family ``types'' into module types, and the context of a given term
+into parameters of the module. We need to faithfully
 reflect the \textit{late binding} nature of families: for each defining
@@
 compiled modules and module types. We can achieve incremental
-checking and overriding as in \ref{chg:software-engineering}---for example, if we
+checking as in \ref{chg:software-engineering}; overriding is also directly supported---for example, if we
 want to override "Field a", we simply replace that compiled module with
@@
 The resulting compiled module is just like mundane Coq's module and 
-thus maintains all the computational mechanism from Coq, resolving \ref{chg:software-engineering}.
+thus maintains all the computational mechanism from Coq, fulfilling requirements in \ref{chg:software-engineering}.
 
@@
 This compilation strategy
-has already allowed fields to be type-dependent on former fields thus fulfilling
-\ref{chg:definition-relevant-reasoning}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
-the concrete definition of a given field (especially when we are proving
-properties about it).
+has already allowed fields to be type-dependent on former fields thus we can see concretely how
+\ref{chg:definition-relevant-reasoning} appears in our current formulation and what we needs to solve this challenge.
+
+% However, the current implementation is not satisfactory enough. There are cases where we need to expose
+% the concrete definition of a given field (especially when we are proving
+% properties about it).
 

### 2022-06-30 lang-design.tex
Fixing coqimpl according to challenge section Move lang-design into coqimpl.

@@
 % programming} to relieve us from the hassle of direct manipulation of proof terms.
-
-\subsection{Language Design}
-Inspired by our above requirements and Family Polymorphism, we propose
-the following extension to Coq's Vernacular Commands.
-% basically write down the specification of surface Coq commandline
-%   and then explain  how the language design makes the programmer’s life easier
-\begin{align*}
-  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
-  \quad |\quad "EndFamily" \\
-  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
-  & |\quad "Extend"~"Family"~\langle ident\rangle \\ 
-  & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
-  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
-  & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
-  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad | \quad ...
-\end{align*}
-
-These are some core extension we proposed to support \cref{fig:STLC-example}.
-We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
-"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.}
-We can define nested families.
-Family will also be used to encode recursive handlers. Each case
-handling of one inductive type will be aggregated into one family, then
-"FRecursor" will be used to construct a recursor using these handlers.
-For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
-"handlers" will generate a new field "recur" that is doing induction
-(using elimination principle of "T") using motive "f", and the handlers
-are located in the family "handlers".
-
-We can inherit a top-level family "parent" using \mintinline{Coq}{Family
-... extends parent} as a new top-level family. As expected, "Override"
-will override a field from the parent. "Inherit" can be used to inherit
-nested families, inductive types, and fields. The fields inside a family
-are order-sensitive because we have dependent types here. So "Inherit
-Until" is used if the user wants to inherit a range of fields. If the
-user wants to extend an inductive type or a nested family from the
-parent, then "Extend FInductive" or "Extend Family" is used. Thus nested
-family can only be inherited or extended together with its outer family
-(and thus top level family), thus our support of family polymorphism is
-incomplete.
-
-This incomplete support can already be useful for extending functions.
-Take \cref{fig:plugin-example2} as example, if we want to extend boolean
-inductive type and its negation, we only need to specify the new
-constructor for boolean and how negation deals with the new constructor.
-As a result, the programmer does not have to duplicate code---otherwise
-the user will need to copy and paste the whole original negation
-definition and add a new pattern matching case.

### 2022-06-30 coqimpl.tex
Adapt Sealed Family Section

@@
  Family MonotonicF.
-  Final Field f : nat -> nat := non_impl.
+  Field f : nat -> nat := non_impl.
   Field p : monotonic f := non_impl.
@@
  Override Family MonotonicF.
-  Final Field f : nat -> nat := (+ 1).
+  Field f : nat -> nat := (+ 1).
   Field p : monotonic f := ...
@@
 example. We can see that the first attempt of proving "a = 1" fails.
-This should fail, actually---because we want to retain the possibility
+\textit{This should fail}---because we want to retain the possibility
 to override "a" in the future, but the mere existence of the field "c"
-will stop that, since "c" witnesses "a" fix to "1". From the other
-perspective, the proof of "c" will require a concrete definition of "a"
-instead of just the exposure of its type. \textbf{We can see a conflict
-between \textit{overridable} and \textit{(definition-relevant)
-reasoning}}.  We enable the exposure of the concrete definition by
-providing "Final" keyword upon "a'".
+will stop that, since "c" witnesses "a" fix to "1". In our implementation, \textit{this will fail}---the proof of "c" will require a concrete definition of "a",
+but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in the context where we only know \mintinline{Coq}{a : nat} thus failed.
+
+To enable the exposure of the concrete definition,
+our plugin provide "Final" keyword. Once we decorate it upon "a'", the second attempt \mintinline{Coq}{c : a' = 1 := eq_refl} will succeed. Contrary to "a", our "a'" is not overridable in "Family I2".
 
@@
 
-In our Coq plugin, we can achieve such exposure of definitional equality by
-exposing the whole definition directly into the compiled module type. Doing so we also prohibit overriding because
+We implement "Final" by
+exposing the whole definition directly into the compiled module type. Doing so we will prohibit overriding because
 subtraction doesn't have the same definition as $add$. We require the
-users to decorate field as `Final' so that the plugin will proceed with
-this special treatment.
+users to manual decorate field as `Final' so that the plugin will proceed with this special treatment.
 %\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
@@
 \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
-because to provide
-"p" we need to have concrete definition of "f" in the family type, which
+because to prove
+the latter we need to have concrete definition of "f" in the family type, which
 makes "f" non-overridable at all. One option is to use sigma type
@@
 
-To resolve this, we provide family sealing mechanism to simulate sigma type---we allow the
-users to specify the compiled module type of a given family, and thus we
-can have a \textit{non-overridable} (thus locally transparent) field
-located in an \textit{overridable} family.
-See, for example, the "MonotonicF" in \cref{fig:plugin-example3}.
-We simply seal
+To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
+In a sealed family, every field will expose its definition inside the family but a family as a whole will be sealed into a relaxed type.
+Thus we can override one sealed family as a whole and thus change the internal implementation.
+
+For example, the "MonotonicF" in \cref{fig:plugin-example3}:
+we simply seal
 \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
-together inside an overridable family that does not allow "f"
-to be overridden in any of its derived families.
-Because "f" is non-overridable inside the family, its definition is
+together inside an sealed family.
+Because the definition of "f" is exposed inside the family, 
+its definition is
 transparent and can be used to prove "p".
@@
 implementation of "f", we just override the whole ``sealed'' family,
-like we do in "I2.MonotonicF". And the compiled module type of
+like we do in "I2.MonotonicF".
+
+And the compiled module type of
 "MonotonicF" will be fixed throughout the extension in different
-families---it is always these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }
+families.
+Overriding field will always have these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
+}
 
@@
 %   this refinement/property
+The implementation of sealed family is not so different from that of the normal family, but we need to be careful about the compiled module type.
 

### 2022-06-30 syntax.tex
resolve comments

@@
 
-A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
-A substitution $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly understood in the way that $\gamma$ will output a \textbf{series} of (dependently-typed related) terms of types $\Delta$, parametrized by $\Gamma$ (i.e.
-we can consider there are holes in $\gamma$ of type $\Gamma$).
-Thus (1) \ruleref{Sub Extend} looks like concatenating one series of terms with a new term (the weird looking type of $\goodTerm{\Gamma}{t}{A[\sigma]}$ is because we need $A$ to appear in the context of $\Delta$ so that we can use \ruleref{Context Extension} to create new context); (2) we have \ruleref{Proj Subst-1} to extract a subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why substitution really looks like a (deeply-nested) pair.
+A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
+type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
+A substitution $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
+understood in the way that $\gamma$ will output a \textbf{series} of
+(dependently-typed related) terms of types $\Delta$, parametrized by
+$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
+Thus (1) \ruleref{Sub Extend} looks like concatenating one
+series of terms with a new term (the weird looking type of
+$\goodTerm{\Gamma}{t}{A[\sigma]}$ is because we need $A$ to be typed in
+the context $\Delta$ so that we can use \ruleref{Context Extension} to
+create new context); (2) we have \ruleref{Proj Subst-1} to extract a
+subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
+substitution really looks like a (deeply-nested) pair.
 
-Since the $\Gamma$ before the $\vdash$ is always indicating the types of ``holes'', this explains the \ruleref{Term Subst} and \ruleref{Type Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution $\goodSub{\Gamma}{\gamma}{\Delta}$, we will have a new term $\goodTerm{\Gamma}{t[\gamma]}{A[\gamma]}$ after the substitution---this substitution can be understood as ``filling all the holes''.
-Once all the holes are filled with a given substitution, the new holes all come from the subsitution itself.
-Thus consecutive subsitution can be composed as \ruleref{Sub Comp}, because all the holes are always from the last substitutions.
+Since the $\Gamma$ before the $\vdash$ is always indicating the types of
+``holes'', this explains the \ruleref{Term Subst} and \ruleref{Type
+Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution
+$\goodSub{\Gamma}{\gamma}{\Delta}$, we will have a new term
+$\goodTerm{\Gamma}{t[\gamma]}{A[\gamma]}$ after the substitution---this
+substitution can be understood as ``filling all the holes''. Once all
+the holes are filled with a given substitution, the new holes all come
+from the substitution itself.
+Consecutive substitutions can be composed using \ruleref{Sub Comp}, because
+all the holes are always from the last substitutions.
 
-Note that, for simplicity, we will (a lot of time) writing $\pi_1$ as a \textbf{shorthand} of $\goodSub{\Gamma,A}{\pi_1~"id"}{\Gamma}$, and similarly another shorthand for $\goodTerm{\Gamma, A}{\pi_2 := \pi_2~"id"}{A[\pi_1]}$.
+Note that, for simplicity, we will often write $\pi_1$ as a
+shorthand for ${\pi_1~"id"}$, typed as
+$\goodSub{\Gamma,A}{\pi_1~"id"}{\Gamma}$, and $\pi_2$ as a shorthand
+for $\pi_2~"id"$, typed as $\goodTerm{\Gamma, A}{\pi_2~"id"}{A[\pi_1]}$.
 
-Based upon the basic MLTT,
-we will have four new kinds of judgements to model our linkages. Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the
-most special one because it is actually just a shorthand of term
-judgement $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$.\YZ{You haven't introduced the notation A[pi_1] yet at this point.
-And what is A[pi_1]?}\EDJreply{added for the intro.
-It is a shorthand for pi_1(id)}
+The core language builds on MLTT and adds four new judgement forms.
+Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
+shorthand for term judgement
+$\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$.
 
@@
 The substitution law for these two
-require \textit{de bruijn lifting}: for a given substitution
+require \textit{de Bruijn lifting}: for a given substitution
 $\goodTerm{\Gamma}{\sigma}{\Delta}$ and a type
 $\goodType{\Delta}{A}{}$, we define its de Bruijn lifting
-$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ .
-Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.\YZ{What is A[sigma] and what is de Bruijn lifting?
-Seems like if the reader is not familiar with the substitution calculus, then they will get lost starting here.
-Broadly speaking, a gentle intro to the intrinsic setting would help the reader a great deal.}\EDJreply{A quick intuitive explanation of substitution-calculus is given above.
-Please check.
-}
+$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ . Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.
 
@@
 distinguish types from their \textit{codes or names}, and thus we use
-$El\ T$ as the type given \textit{type name} $T : \cU$. In this case, (1) $\cU$ is the type including the \textit{codes or names} of types , and (2) $T$ is
-not allowed to locate at type position (i.e.
-at the right-most position of a term judgement) but $El \ T$ is.
-"c" is the ``inverse'' of $El$ that getting \underline{"c"}ode of a type, with the following two equations $"c" (El~T) \equiv T$ and $El ("c"~T) \equiv T$.
-Recall that we are actually using predicative universe and thus we actually have different level of universes $\cU_j$.
-Here level omitted for conciseness.
+$El\ T$ as the type given \textit{type name} $T : \cU$.
+In this case, (1) $\cU$ is the type including the \textit{codes or
+names} of types, and (2) $T$ is not allowed to locate at type position
+(i.e., at the rightmost position of a term judgement) but $El \ T$ is.
+"c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of a
+type, with the following two equations $"c" (El~T) \equiv T$ and $El
+("c"~T) \equiv T$.
+Recall that we are actually using predicative universe, so we actually
+have different levels of universes $\cU_j$.
+We omit universe levels for conciseness.
 
@@
 {\goodType{\Gamma}{T}{j}}
-{\goodTerm{\Gamma}{"c" \ T}{\cU}\YZ{what is c?
-Why is this rule called Univ-1?Didn't you write that you omit universe levels?}\EDJreply{I think c stands for code.
-I am just following convention.
-Now I remove their name because that is a bad name.
-About the level, I just kind of afraid they will think there is one universe.
-Now I've removed all the indices}
+{\goodTerm{\Gamma}{"c" \ T}{\cU}
 }\quad
@@
 {\goodTerm{\Gamma}{T}{\cU}}
-{\goodType{\Gamma}{El \ T}{j}\YZ{'El T' isn't explained only until it makes a second appearance in Ind-Term.}\EDJreply{I have made El and universe explanation earlier, please check.}}
+{\goodType{\Gamma}{El \ T}{j}}
 \quad
@@
 
-We have shown case some exemplar rules for all four standard type
-judgments, and now we focus on the newly introduced facility.
+We have shown some exemplar rules for all four standard types of
+judgments. Now we focus on the newly introduced facility.
 
@@
 
-\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a shorthand of the other)}
+\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a shorthand for the other)}
 
@@
 As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
-a shorthand of term judgement $\goodTerm{\Gamma , \cC
+a shorthand for term judgement $\goodTerm{\Gamma , \cC
 \sigma}{\_}{A[\pi_1]}$,\YZ{Not quite. Previously you used cL I think.}\EDJreply{Sorry. It is fixed. It should be $\cC$ in the context.}

### 2022-06-30 coqimpl.tex

@@
 \textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
-We support ``extensible'' inductive types from
-\ref{chg:extensible-inductive-type} in Coq to some extent.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types.
+We support extensible inductive types in Coq mentioned in
+\ref{chg:extensible-inductive-type} but we will not be aiming at the research of its formal semantic.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types.
 It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it.
@@
 
-To construct a recursor as in \ref{chg:extensible-inductive-type}, we need to create for each
+Recall, to construct a recursor, we need to create for each
 constructor a handler (function) and aggregate them into a family, and thus that family (e.g.
@@
 
-To summarize, the key insight here is that, to make each recursor
+To summarize, (1) to make each recursor
 handler inheritable, we need to seal an abstraction around the inductive
@@
 simultaneously.
-
-\textbf{Propositional Partial Recursor.}
-Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
-not expressive enough to really fulfill \ref{chg:extensible-inductive-type}. For
-example, we cannot prove the constructors are injective, i.e.,
-\mintinline{Coq}{self_B.tt ≠ self_B.ff} because from the perspective of Coq they are just two terms declared in the module type.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only?
-My confusion after reading these paras is that I still don't know why partial recursors are needed.}\EDJreply{The issue is exactly, at that point of programming, for extensibility, there is no access to the concrete recursor.
-Concrete recursors also cannot be postulated as part of the interface because it breaks extensibility.}
-In the original setting, to
-prove injectivity of constructors of a (vanilla) inductive type, we only
-require a recursor and its computational rules. However, as we pointed
-out earlier, the recursor of an inductive type cannot be exposed into
-the context for the sake of extensibility. Hence, we need to alter the
-formulation of the recursor: we need to have a better
-characterization of ``extensible inductive types'' exported in the
-compiled module type as a part of the interface for the following fields
-to use.
-
-The simplest idea is to allow partiality---i.e., returning
-\mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
-
-We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors?
-I suppose it has to do with the computational axioms?}\EDJreply{Yes.
-I want to emphasize the equality used by computational axiom is not judgemental equality.
-So Coq cannot automatically compute/reduce when a partial recursor applied to a term.
-This reduction can only be semi-automatically done by using [rewrite tactic and the computational axiom].}
-For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}\EDJreply{Added below please check.}
-
-\begin{minted}[escapeinside=@@]{Coq}
-b_prec : forall R, option R -> option R -> b -> option R.@\YZ{Not clear what it takes to define/prove partial recursors.
-The reader needs to be told if they require breaking abstraction like recursors do.}\EDJreply{What do you mean?
-recursor doesn't break abstraction but break extensibility.
-Below mentioned partial recursor doesn't break extensibility}@
-\end{minted}
-with appropriate computational axioms.
-This partial recursor differs from original recursor only by decorating an "option" with the return type thus possible to be automatically generated (and the case is similar for those computational axiom).
-The partiality doesn't break
-extensibility (i.e. all future extensions of "b" can support this
-"b_prec") because "b_prec" can just return "None" when bump into extended constructor in the future.
-
-What's more, partial recursor is theoretically enough for the
-"injection" and "discriminate" tactics. The rough idea to achieve this
-is to use "b_prec" to reflect our special inductive type into vanilla
-inductive type, and use the original "injection" and "discriminate"
-tactics. For example, we want to prove \mintinline{Coq}{self_B.tt ≠
-self_B.ff}, we simply 
-\begin{minted}{Coq}
-Definition reflect : self_B.b → option bool 
-                   := b_prec (fun _ => bool) (Some true) (Some false).
-Definition tt_neq_ff : self_B.tt = self_B.ff → False.
-(* Because self_B.tt = self_B.ff → reflect self_B.tt = reflect self_B.ff 
-      → Some true = Some false    (by computational axiom)
-      → False           (by discriminate of bool and injection of Some) *)
-\end{minted}
-
-In fact, the above "reflect" function implies that propositional partial
-recursor is good enough to some extent, at least for non-indexed
-inductive type: because we can ``embed'' vanilla inductive types into
-types supporting a partial recursor. Because "reflect" is actually a left
-inverse of an injection "bool → self_B.b". A bit more formally, and
-restricting our focus to non-indexed inductive types:
-
-\begin{theorem}\label{thm:prec-complete} Given a list of $n$ pairs of types $\{ x : "A"_i \vdash "B"_i(x) \}_{i}$
-  \begin{itemize}
-    \item Define $"C"$ as the inductive type using this list, i.e.,
-    \begin{minted}{Coq}
-Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
-    \end{minted}
-    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
-    \begin{itemize}
-      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
-      \item a partial recursor
-      \begin{minted}[escapeinside=@@]{Coq}
-prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint. Let me fix other parts first}@
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), .., 
-      D -> option R
-      \end{minted}
-      \item and $n$ (propositional) computational axioms: for all $i$, 
-      \begin{minted}{Coq}
-prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
-    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
-      \end{minted}
-    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
-    \end{itemize}
-  \end{itemize}
-  Then there exists an \textbf{embedding} from "C" to "D".
-  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of
-  "C" and "prec" of "D" (both acting like identity) such that
-  \begin{minted}[escapeinside=@@]{Coq}
-    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
-If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
-I mentioned left inverse 'linv' is the generalization}@
-  \end{minted}
-\end{theorem}
-
-We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
-actually be considered as a constructor, because they can be reflected
-to real constructors of "C" using the left inverse "linv" (directly generalization of "reflect : self_B.b → option bool" of the above). Thus, \cref{thm:prec-complete}
-implies that (1) every future extension of the inductive type can
-support this partial recursion (trivially); (2) every type supporting
-this partial recursor with its computational axioms at least supports
-these constructors because of the embedding. In other words, \textit{a
-type (at least) supports these constructors if and only if this type
-supports the corresponding partial recursor (with the computational
-axioms)}.  Thus, we can argue that the partial recursor gives a sound and
-complete characterization of all the extension of a given (non-indexed)
-inductive type.\YZ{What about indexed inductive types?}
-
-We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
-\begin{minted}{Coq}
-  pdelim : ∀ (P : D -> Set), .., 
-  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) -> option (P (dᵢ x w)))),  
-  .., ∀ (d : D) -> option (P d)
-\end{minted}
-To generalize to indexed inductive type, we base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}. Please refer to
+(2) \ref{chg:extensible-inductive-type} is resolved by this \textit{decoupling} of the syntax of implementation of recursor handlers and exhaustiveness checking.
+The former is handled by family inheritance and thus avoid the boilerplate code; and the non-inheritance of the compiled recursor ensures the exhaustiveness checking happens for every recursor.
 
@@
 % 2. Computational rule for the total recursor
-Unfortunately, there are cases the users just want global reasoning
-instead of an extensible proof. For example, in STLC, we expect
+Unfortunately, there are cases an extensible proof brings more boilerplate code. For example, in STLC, we expect
 ``\mintinline{Coq}{~ value (tm_app x y)}'' to hold in all future
 extensions of STLC---applications will never be considered as values
-by an extension.
+by any extension.
 We know the correct way to prove the proposition
@@
 
-Our proposal is to allow global reasoning, which will look like a
-constraint inside the family. We provide "Closing Fact" as a command to
-carry out this global reasoning, attached with a proof script and proof
-term. This "Closing Fact" will look like an assertion during the
-construction of the enclosing family. This assertion will only be
-verified when the whole family is closed and compiled into a module,
-with the help of the attached proof script and proof terms.
+Our proposal is to allow global reasoning by providing "Closing Fact" command. Ultimately, "Closing Fact" is asserting a constraint upon all the
+possible extension of the surrounding family and claiming that a proof script
+can solve the constraints.\YZ{This is a good one-liner that summarizes the intended usage scenarios of "Closing Fact".
+Should probably bring it out earlier.}\EDJreply{I put it here as you suggested but I think this one-liner at here will cause more confusion because it is too abstract.}
+The assertion will only be verified when the whole family is closed and compiled into a module.
+
+% We provide "Closing Fact" as a command to
+% carry out this global reasoning, attached with a proof script and proof
+% term. This "Closing Fact" will look like an assertion during the
+% construction of the enclosing family. This assertion will only be
+% verified when the whole family is closed and compiled into a module,
+% with the help of the attached proof script and proof terms.
 
@@
 % \YZ{Please check if my wording is correct.}\EDJreply{Yes. Resolved}
-Since we are dealing with a concrete inductive type during compilation,
+Since we are dealing with a concrete (vanilla) inductive type during compilation,
 we can use Coq's "inversion" tactic to complete this proof.
@@
 only compile part of the family to the point closing fact is used) but it is a big engineering difficulty for me.}
-
-Ultimately, "Closing Fact" is asserting a constraint upon all the
-possible extension of the surrounding family and claiming a proof script
-can solve the constraints.\YZ{This is a good one-liner that summarizes the intended usage scenarios of "Closing Fact". Should probably bring it out earlier.}
-We expect it to be used only sporadically for maintainability.
+We expect it to be used only sporadically for maintainability, and as an alternative to using "FRecursor".
+This is called global reasoning because this is a post-hoc reasoning when the family as a \textit{whole} is compiled into a module.
 
@@
 the family data is invisible to Coq internal and thus causes some
-
 inconvenience. For example, in
@@
 % in \ref{chg:software-engineering} that uses lemmas proved in the family.
+
+% Following needs a rewrite
+
+\textbf{Propositional Partial Recursor.} Notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''--they will possibly disrupt arbitrary extension of inductive type.
+For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)} in the "Family STLC" either using "FRecursor" or "Closing", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
+
+This raise an interesting question: what propositions are conservative extension?
+For example, apparently we know injectiveness of constructor, (i.e.
+\mintinline{Coq}{STLC.tt ≠ STLC.ff}) must be conservative because it holds for all extension of the inductive type "STLC.tm".
+But is there a ``best'' proposition (i.e.
+the proposition that everything the proposition derives are exactly conservative extension)?
+
+Since injectiveness of the constructors are derived by recursor in vanilla inductive type, an educated guess would be related to recursor.
+But we know the original recursor won't work because the mere existence of the original recursor will break extensibility.
+Thus we need to modify it.
+The simplest idea is to allow partiality---i.e., returning
+\mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
+
+We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors?
+I suppose it has to do with the computational axioms?}\EDJreply{Yes.
+I want to emphasize the equality used by computational axiom is not judgemental equality.
+So Coq cannot automatically compute/reduce when a partial recursor applied to a term.
+This reduction can only be semi-automatically done by using [rewrite tactic and the computational axiom].}
+For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}\EDJreply{Added below please check.}
+
+\begin{minted}[escapeinside=@@]{Coq}
+b_prec : forall R, option R -> option R -> b -> option R.@\YZ{Not clear what it takes to define/prove partial recursors.
+The reader needs to be told if they require breaking abstraction like recursors do.}\EDJreply{What do you mean?
+recursor doesn't break abstraction but break extensibility.
+Below mentioned partial recursor doesn't break extensibility}@
+\end{minted}
+with appropriate computational axioms.
+This partial recursor differs from original recursor only by decorating an "option" with the return type thus possible to be automatically generated (and the case is similar for those computational axiom).
+The partiality doesn't break
+extensibility (i.e. all future extensions of "b" can support this
+"b_prec") because "b_prec" can just return "None" when bump into extended constructor in the future.
+
+The first thing to check is that, if partial recursor is theoretically enough for injectiveness of the constructor. The rough idea to achieve this
+is to use "b_prec" to reflect our special inductive type into vanilla
+inductive type, and use the original "injection" and "discriminate"
+tactics. For example, we want to prove \mintinline{Coq}{self_B.tt ≠
+self_B.ff}, we simply 
+\begin{minted}{Coq}
+Definition reflect : self_B.b → option bool 
+                   := b_prec (fun _ => bool) (Some true) (Some false).
+Definition tt_neq_ff : self_B.tt = self_B.ff → False.
+(* Because self_B.tt = self_B.ff → reflect self_B.tt = reflect self_B.ff 
+      → Some true = Some false    (by computational axiom)
+      → False           (by discriminate of bool and injection of Some) *)
+\end{minted}
+
+Even though we cannot prove propositional partial recursor is the best,
+the above "reflect" function implies that propositional partial
+recursor is a good \textit{extensional characterization} of (non-indexed) extensible inductive type (with certain constraints): because we can ``embed'' vanilla inductive types into
+types supporting a partial recursor: "reflect" is actually a left
+inverse of an injection "bool → self_B.b". A bit more formally, and
+restricting our focus to non-indexed inductive types:
+
+\begin{theorem}\label{thm:prec-complete} Given a list of $n$ pairs of types $\{ x : "A"_i \vdash "B"_i(x) \}_{i}$ s.t.
+
+  \begin{itemize}
+    \item \textlabel{(Detectable Partiality)}{prop:detectable-partiality} for each pair $x : A_i \vdash B_i(x)$,
+    \begin{minted}{Coq}
+      Axiom detectable:
+      forall {T} {a} (f: Bᵢ a -> option T),
+        {forall x, f x <> None} + {exists x, f x = None}.
+    \end{minted} 
+    \item Define $"C"$ as the inductive type using this list, i.e.,
+    \begin{minted}{Coq}
+Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
+    \end{minted}
+    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
+    \begin{itemize}
+      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
+      \item a partial recursor
+      \begin{minted}[escapeinside=@@]{Coq}
+prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint. Let me fix other parts first}@
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), .., 
+      D -> option R
+      \end{minted}
+      \item and $n$ (propositional) computational axioms: for all $i$, 
+      \begin{minted}{Coq}
+prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
+    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
+      \end{minted}
+    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
+    \end{itemize}
+  \end{itemize}
+  Then there exists an \textbf{embedding} from "C" to "D".
+  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
+  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of
+  "C" and "prec" of "D" (both acting like identity) such that
+  \begin{minted}[escapeinside=@@]{Coq}
+    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
+If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
+I mentioned left inverse 'linv' is the generalization}@
+  \end{minted}
+\end{theorem}
+The reason we have this weird looking \ref{prop:detectable-partiality} is because the construction of "linv"---at one point, using "prec" to construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option C) → option C}, generally not provable, only if we have \mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
+This \ref{prop:detectable-partiality} is exactly transforming former to the later: recall that, according to the semantic of inductive type, \mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using "linv" on the substructure of "D".
+This function points out each of the translation result from "D" to "C" and some of them are "None" because of failing.
+For us, we only care about if \textbf{any} of the substructure of "D" is failed to translate---if any one substructure fails, then the whole transltion fails and thus we should have "None" as return.
+
+\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that means all of the examples shown in this paper.
+\ref{prop:detectable-partiality} is not supported when using function (infinitary branching) in the constructors, for example, the case of using HOAS to represent functions in "tm".
+In those cases, we still have the fact that partial recursor is a conservative extension, but we cannot show that partial recursor is the extensional characterization any more。
+
+We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
+actually be considered as a constructor, because they can be reflected
+to real constructors of "C" using the left inverse "linv" (directly generalization of "reflect : self_B.b → option bool" of the above). Thus, \cref{thm:prec-complete}
+implies that (1) every future extension of the inductive type can
+support this partial recursion (trivially); (2) every type supporting
+this partial recursor with its computational axioms at least supports
+these constructors because of the embedding. In other words, \textit{a
+type (at least) supports these constructors if and only if this type
+supports the corresponding partial recursor (with the computational
+axioms)}.  Thus, we can argue that the partial recursor gives a sound and
+complete extensional characterization of all the extension of a given (non-indexed)
+inductive type.\YZ{What about indexed inductive types?}
+
+We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
+\begin{minted}{Coq}
+  pdelim : ∀ (P : D -> Set), .., 
+  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) -> option (P (dᵢ x w)))),  
+  .., ∀ (d : D) -> option (P d)
+\end{minted}
+To generalize to indexed inductive type, we base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}. Please refer to

### 2022-06-30 syntactic-example.tex
reply to comments

--- 2022-06-27 syntactic-example.tex
@@
 parts and making "O" and "S" standing alone.
-"pred" is defined as $n \mapsto ("pred" n, n)$\YZ{looks like an ill-founded definition}, and thus a predecessor.
+"pred" will map $S~n \mapsto (n, S~n)$ and $O \mapsto (O, O)$, and thus a predecessor.
 
@@
  EndFamily. (* pN₁, handler₁, sig₃ *)
- pred@\YZ{Why is there no FRecursor?}@
-  : self[A].N → self[A].N × self[A].N
-  := ... 
+ FRecursor pred about self[A].N 
+  motive (fun _ => self[A].N × self[A].N)
+  using handler.
+ (* pred
+  : self[A].N → self[A].N × self[A].N *)
   (* predT, pred₁, sig₄, sig₄', sl₄ *) 

### 2022-06-30 syntax.tex
reply to comments

@@
 $\cU$, then no information about the inductive type is exposed).
-\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type.
-They are basically the same thing but (W T) cannot be at the type position in a judgement, i.e.
-k |- q : (W T) is not allowed, we only have k |- q : El (W T).
-This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}\YZreply{I think explanation is warranted; readers unfamiliar with that line of work may find it confusing.
-BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general.
-I just didn't write out the two equations c (El T) = T, El (c T) = T}\YZreply{Seems harmless to list these equations and move the reference to 'type universe a la Tarski' earlier.}\EDJreply{Done.
-Please check.}
 
@@
 Is it still too hard?
-Do you think I need to think of a more friendly way?}
+Do you think I need to think of a more friendly way?}\YZreply{Yes, still too rough on the reader.
+I sense that this part is central to the metatheory, so it is worth explaining the details well.}
  \quad \goodType{\Gamma, A}{T}{i}}
@@
 {\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
-\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}
+\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}\YZ{What is C_t? I don't think you talk about it anywhere.}
 \\
@@
 to retrieve the corresponding types of linkage of a given signature, and
-$\cC$ to get the compiled linkage type (an existential type).\YZ{Why is Csigma needed?
+$\cC$ to get the compiled linkage type (an existential type).
+$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, accoding to \ruleref{Compile}, is really a deeply nested dependent pair.
+This is reflecting the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}.
+Thus we need $\cC\sigma$ in our context.
+\YZ{Why is Csigma needed?
 How is it different from Lsigma?}\EDJreply{It corresponds to the compilation of the family ``type'', which is part of the implementation detail.
 Basically we know every field of a family is dependent on the former fields.
-The implementation detail here is that, every field of a family is dependent on the compiled module of the former family (and thus includes all the former fields)} $\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, accoding to \ruleref{Compile}, is really a deeply nested dependent pair.
-This is reflecting the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}.
-Thus we need $\cC\sigma$ in our context.
+The implementation detail here is that, every field of a family is dependent on the compiled module of the former family (and thus includes all the former fields)}\YZreply{I don't think you should explain the metatheory from the angle of how things are implemented in your plugin.
+Rather, when describing your Coq plugin, you can say it matches the metatheory.
+PS: you use two terminologies: 'compile' and 'seal', but I guess they mean very similar things. So avoid saying 'compile' here.
+PPS: Sealing means hiding, so what is being hided? Why do you need to hide it? And how does the 'Compile' and 'Seal' rules perform hiding?}
 
@@
 the signature for inductive types.  They are both constructed from the
-empty stuff, and respect substitution lemma for sure. We have
+empty stuff, and respect substitution lemmas. We have
 \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component
@@
 Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
-Seal Γ A τ = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi. I fixed it now}@
+Seal Γ A τ = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi. I fixed it now}\YZreply{Why is tau not used in the definition?}@
 
@@
 a shorthand for term judgement $\goodTerm{\Gamma , \cC
-\sigma}{\_}{A[\pi_1]}$,\YZ{Not quite. Previously you used cL I think.}\EDJreply{Sorry. It is fixed. It should be $\cC$ in the context.}
+\sigma}{\_}{A[\pi_1]}$,
 a special form of \textit{local morphism}~\cite{abbott2003category}.
@@
 other words, there is no way to ``transplant'' $t$ into the context $X_2
-: \bW \sigma_2$.\YZ{Is sigma_2 the overriding of sigma_1? The reader has to guess.}\EDJreply{Thanks. Fixed}
+: \bW \sigma_2$.
 This makes all the fields after this inductive type
@@
 {\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-  \quad \goodTerm{\Gamma, A_2}{t}{T}\YZ{What is sigma2'?}\EDJreply{Sorry this should be A2, i.e.
-the result of sealing.
-Earlier version sigma2' is the result of sealing.
-It was type-set before the last meta-theory change}}
+  \quad \goodTerm{\Gamma, A_2}{t}{T}}
 {\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}

### 2022-06-30 syntax.tex
Reply comments. Compile is very different from Seal.

@@
 {\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
-\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}\YZ{What is C_t? I don't think you talk about it anywhere.}
+\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}\YZ{What is C_t?
+I don't think you talk about it anywhere.}\EDJreply{I added the explanation together with more details on C.
+Ct is really just intro rule for C}
 \\
@@
 
+Before we dive in the details of our metatheory, we have to mention that, the challenge of incorporating family polymorphism with inductive type is quite different in metatheory than in surface syntax because of de Bruijn indices.
+(Using surface synax), given a linkage $\{a~:~\bW \sigma; b_1 : a \to T;..\}$ with a bunch of $b_i$ \textit{using} $a$.
+The tension we need to solve here is between exhaustiveness checking and inheritance---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to let $b_i$ term construct in the context of $\bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that.
+Our metatheory formulation of linkage is mainly trying to solve this tension.
+
 $\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature
@@
 $\cC$ to get the compiled linkage type (an existential type).
-$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, accoding to \ruleref{Compile}, is really a deeply nested dependent pair.
-This is reflecting the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}.
+$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, accoding to \ruleref{Compile}, is a deeply nested dependent pair.
+Then $\cCt \sigma$ provide the concrete module of type $\cC\sigma$ given a linkage of type $\cL\sigma$.
+
+The reason we need this additional $\cC\sigma$ here is because, the rule of linkages hinders its ability to do abstraction.
+Concretely speaking, (using surface syntax) we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to T\})}{h}{\cL(\{a~:~\cU; b~:~a \to T\})}$\footnote{If we had one, consider the case $T$ is bottom type and $\sigma$ is empty record and thus $a$ is initially bottom type.
+Now we further override $h.a$ using unit type we can derive inconsistency}.
+This term is important for us to abstract the concrete inductive type so that the later field can depend on the abstract interface.
+On the other hand, we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T) }{h}{(\Sigma a : \cU, a \to T)}$ because of the rule for existential (also existential type just don't allow overriding).
+Thus to allow abstraction happens, we need  $\cC\sigma$ to compile a linkage to module and we also need \ruleref{Lnkg Compile} to specify the compilation from linkage to module.
+
+This is also reflected in the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}.
 Thus we need $\cC\sigma$ in our context.
@@
 PS: you use two terminologies: 'compile' and 'seal', but I guess they mean very similar things. So avoid saying 'compile' here.
-PPS: Sealing means hiding, so what is being hided? Why do you need to hide it? And how does the 'Compile' and 'Seal' rules perform hiding?}
+PPS: Sealing means hiding, so what is being hided?
+Why do you need to hide it?
+And how does the 'Compile' and 'Seal' rules perform hiding?}\EDJreply{  Compile is very different from seal.
+Compile is making a linkage into a module, depriving the ability to override---if we understand a linkage as a module with a lot of universal quantifier inside.
+Compile cannot hide things, but only translate a linkage into a module.
+Seal is really something that does hiding---because existential type can hide stuff.
+Seal is an term between existential types.
+\\
+The ultimate reason we have this two notion is that, the introduction rule for linkage fails to provide abstraction mechanism so we have to translate them into module. \\
+Generally speaking Compile : Linkage -> Module, Seal : Module -> Module, we need module here because linkage cannot do abstraction of inductive type. A concrete detail is at the above. \\ 
+I now describe the motivation of compile in the earlier paragraph, right before the "reflecting" section.
+Note that, this compile is inspired by plugin implementation.
+This is the reason I say the metatheory is really close and inspired by the implementation.
+Similarly as I say ``there is no extensible inductive type but only overriding of new inductive type'' because in metatheory it is using overriding to ``extend'' (same as plugin implementation), seal means the same thing, and ``compile'' in metatheory happens in plugin implementation
+}
 
@@
 {\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}\YZ{What is R[pi1 circ pi1]?}\EDJreply{It is the composition of the substitution.
-Thanks for pointing out now I included circ into the typesetting} }
+Thanks for pointing out now I have included circ into the typesetting} }
 \quad 
@@
 Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
-Seal Γ A τ = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi. I fixed it now}\YZreply{Why is tau not used in the definition?}@
+Seal Γ σ A = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi.
+I fixed it now}\YZreply{Why is tau not used in the definition?}\EDJreply{Thanks.
+I wanted to remove tau but I accidentally removed sigma instead}@
 
@@
 
-The motivation for sealing judgement is acting like syntactic sugar
+The motivation for sealing judgement is still about the important term we mentioned earlier: it is acting like syntactic sugar
 helpful for inheritance---due to the presence of inductive type in the

### 2022-06-30 syntax.tex
Fix the introduction of sealing Fix meta-theory section.

@@
 
-Before we dive in the details of our metatheory, we have to mention that, the challenge of incorporating family polymorphism with inductive type is quite different in metatheory than in surface syntax because of de Bruijn indices.
-(Using surface synax), given a linkage $\{a~:~\bW \sigma; b_1 : a \to T;..\}$ with a bunch of $b_i$ \textit{using} $a$.
-The tension we need to solve here is between exhaustiveness checking and inheritance---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to let $b_i$ term construct in the context of $\bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that.
+$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature
+for linkages in the context $\Gamma$ with $n$ fields. We omit the
+naming/label information for fields in the meta\-theory. The signature (for the linkage) and the linkage rules are quite similar to
+the signature for inductive types.  They are both constructed from the
+empty stuff, and respect substitution lemmas. We have
+\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component
+due to harmony~\cite{pfenning2009lecture}, so we have local soundness
+and local completeness ($\beta,\eta$ rules). We omit them and the
+substitution laws for either field addition and projection. For
+$\mu^+,\nu^+$ rules, we will omit the sealing $\{f\}$ if possible.\footnote{This
+omission can be supported by Agda using Implicit Arguments}
+
+A side note: before we dive in more details of our metatheory for linkage (especially the meaning of sealing and $\cC$), we have to mention that, the challenge of incorporating family polymorphism with inductive type mentioned in \ref{chg:extensible-inductive-type} mutates in metatheory because of de Bruijn indices.
+(Using surface synax), consider the example of a linkage $\{a~:~\bW \sigma; b_1 : a \to T;..\}$ with a bunch of $b_i$ \textit{using} $a$.
+The tension we need to solve here is between \textit{exhaustiveness checking} and \textit{inheritance (code reuse)}---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to construct $b_i$ term in the context of $\bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that.
 Our metatheory formulation of linkage is mainly trying to solve this tension.
 
-$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature
-for linkages in the context $\Gamma$ with $n$ fields. We omit the
-naming/label information for fields in the meta\-theory. We can use $\cL$
-to retrieve the corresponding types of linkage of a given signature, and
-$\cC$ to get the compiled linkage type (an existential type).
-$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, accoding to \ruleref{Compile}, is a deeply nested dependent pair.
+We use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).
+$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, according to \ruleref{Compile}, is a deeply nested dependent pair.
 Then $\cCt \sigma$ provide the concrete module of type $\cC\sigma$ given a linkage of type $\cL\sigma$.
 
-The reason we need this additional $\cC\sigma$ here is because, the rule of linkages hinders its ability to do abstraction.
+The reason we need this additional type construct of $\cC\sigma$ here is to support \textit{code reuse}, as the rule of linkages hinders its ability to do abstraction.
 Concretely speaking, (using surface syntax) we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to T\})}{h}{\cL(\{a~:~\cU; b~:~a \to T\})}$\footnote{If we had one, consider the case $T$ is bottom type and $\sigma$ is empty record and thus $a$ is initially bottom type.
 Now we further override $h.a$ using unit type we can derive inconsistency}.
-This term is important for us to abstract the concrete inductive type so that the later field can depend on the abstract interface.
+This term is important for us to abstract the concrete inductive type so that the later field can depend on the abstract interface and achieve \textit{code reuse} reuse.
 On the other hand, we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T) }{h}{(\Sigma a : \cU, a \to T)}$ because of the rule for existential (also existential type just don't allow overriding).
-Thus to allow abstraction happens, we need  $\cC\sigma$ to compile a linkage to module and we also need \ruleref{Lnkg Compile} to specify the compilation from linkage to module.
+Thus to allow abstraction happens, we need  $\cC\sigma$ to compile a linkage to module as a part of the context and we also need \ruleref{Lnkg Compile} to specify the compilation from linkage to module.
 
@@
 
-The signature (for the linkage) and the linkage rules are quite similar to
-the signature for inductive types.  They are both constructed from the
-empty stuff, and respect substitution lemmas. We have
-\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component
-due to harmony~\cite{pfenning2009lecture}, so we have local soundness
-and local completeness ($\beta,\eta$ rules). We omit them and the
-substitution laws for either field addition and projection. For
-$\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.\footnote{This
-omission can be supported by Agda using Implicit Arguments}
+However, $\cC$ can only make sure we have $\cL(\{a~:~\bW \sigma; b~:~a \to T\}) \to (\Sigma a:\bW \sigma, a \to T)$, we still need $(\Sigma a:\bW \sigma, a \to T) \to (\Sigma a:\cU \sigma, a \to T)$ to complete this shifting of the context.
+This is sealing's job.
 
-Since our recursor handlers are closely related to linkages, we can finally talk about the rules of recursor now.
+\begin{figure}[H]
+
+  \judgebox{[\goodSeal{\Gamma}{s}{\sigma}{A}] 
+  = [\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}] }
+  $$
+  \Rule[name=Seal Subst]
+  {\goodSeal{\Delta}{s}{\sigma}{A}
+    \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+  {\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{A[\gamma]}}
+  =
+  \frac
+  {\goodTerm{\Delta, \cC\sigma}{s}{A[\pi_1]}
+    \quad  \goodSub{\Gamma}{\gamma}{\Delta}  }
+  {\goodTerm{\Gamma,\cC\sigma[\gamma]}{s[\gamma^\uparrow]}{A[\pi_1][\gamma^\uparrow]\equiv A[\gamma][\pi_1]}}
+  $$
+  $$
+  \Rule[name=Seal-Id]
+  {}
+  {\goodSeal{\Gamma}{id_s}{\sigma}{\cC \sigma}}
+  = \goodTerm{\Gamma, \cC\sigma}{\pi_2}{\cC\sigma[\pi_1]}
+  $$
+
+\centering
+\captionsetup{justification=centering}
+
+\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a shorthand for the other)}
+
+\begin{minted}[escapeinside=@@]{agda}
+Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
+Seal Γ σ A = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi.
+I fixed it now}\YZreply{Why is tau not used in the definition?}\EDJreply{Thanks.
+I wanted to remove tau but I accidentally removed sigma instead}@
+
+-[-]S : (f : Seal Γ σ A) → (τ : Tms Θ Γ) → Seal Θ σ[τ] A[τ]
+f[τ]S = t[τ ↑ A]
+
+idₛ : Seal Γ σ 𝓒σ
+idₛ = π₁
+\end{minted}
+
+\end{figure}
+
+As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
+a shorthand for term judgement $\goodTerm{\Gamma , \cC
+\sigma}{\_}{A[\pi_1]}$,
+a special form of \textit{local morphism}~\cite{abbott2003category}.
+The substitution law should be understood as a lemma, and it is easily
+derivable using $\gamma^\uparrow$.
+
+Sealing judgement is still about the important term $h$ we mentioned earlier.
+Thanks to $\cC$, a newly added term $t_{new}$ is now in the context of $\cC\{a : \bW \sigma; ..\}$.
+But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is extended, we would like $t_{new}$ to be in the context  of $\cC\{a : \cU; ..\}$ instead.
+The sealing is responsible for this because we have a sealing $\goodTerm{\Gamma, \cC\{a : \bW \sigma; ..\}}{s}{\cC\{a : \cU; ..\}}$ = $\goodTerm{\Gamma, (\Sigma a : \bW \sigma,..)}{s}{(\Sigma a : \cU,..)}$.
+Then we can construct $t_{new}$ in the context of  $(\Sigma a : \cU,..)$, ready for inheritance.
+Apparently $t_{new}$ in this case cannot be doing \textit{exhaustiveness checking}, but it can freely use any \textit{exhaustiveness-checked} recursor.
+\YZ{It occurs to me that you use the word "sealing" to mean different (though tangentially related) things.
+Previously you used it as in "sealed families".}\EDJreply{They are actually the same thing.
+This sealing in metatheory is much stronger than that of the implemented plugin.
+Downside of the strong sealing is that if the user really use this version of sealing, the user will have to annotate what the sealed family type is.
+Now in the plugin, the sealed family type is almost fixed (directly inferred using a fixed rules).
+So basically this complete version of sealing (in metatheory) can deduce that of seal (in plugin implementation) }
+
+% For example, $ X_1 : \bW
+% \sigma_1 \vdash t : T$ in the parent family will generally not be able
+% to be inherited to the children once we override the inductive type $\bW
+% \sigma_1$ with $\bW \sigma_2$. 
+% In
+% other words, there is no way to ``transplant'' $t$ into the context $X_2
+% : \bW \sigma_2$.
+% This makes all the fields after this inductive type
+% non-inheritable---which shouldn't be the case because a lot of fields
+% might not be directly relying on this concrete $\bW \sigma_1$.
+
+% Sealing is used to exploit this fact---with sealing, given $X_1 : \bW
+% \sigma_1$, we can add well-typed terms $X_1 : \cU \vdash t : T$, into
+% the parent family. Later when we want to inherit $t$,
+% because $t$ only consider $X_1$ as an arbitrary type, inheritance is
+% possible.
+
+This sealing is the general version of the sealed family implemented in the plugin.
+Our plugin will deduce a sealed ``interface'' for a sealed family using a fixed set of rules but here sealing in metatheory can be arbitrary local morphism.
+With the definition of sealing judgement in mind, we can see
+\ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the
+implementation. In the implementation, we will have "self__" all around
+when defining new fields in our plugin, and this "self__" is exactly the
+$A$ in the context of the two fields---especially when we don't do any
+sealing, this $A = \cC \sigma$, which is exactly how the implementation
+exposes "self__" as a module parameter.
+
+In summary, with the help of both sealing and $\cC$, we can shift the context from $\cL(\{a~:~\bW \sigma; ..\})$ to $\Sigma a : \cU,..$ and thus achieve code reuse.
+However, this is just allow possible \textit{code reuse}, we still need to solve the mutated tension in \ref{chg:extensible-inductive-type}.
+To make things concrete, we need to introduce the rules for recursors.
+
+% Since our recursor handlers are closely related to linkages, we can finally talk about the rules of recursor now.
 
@@
 
-Now, one missing piece for linkage is the sealing judgement, and both of the $\mu^+, \nu^+$ rules relies on that---then what is this sealing judgement for?
-
-\begin{figure}[H]
-
-  \judgebox{[\goodSeal{\Gamma}{s}{\sigma}{A}] 
-  = [\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}] }
-  $$
-  \Rule[name=Seal Subst]
-  {\goodSeal{\Delta}{s}{\sigma}{A}
-    \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-  {\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{A[\gamma]}}
-  =
-  \frac
-  {\goodTerm{\Delta, \cC\sigma}{s}{A[\pi_1]}
-    \quad  \goodSub{\Gamma}{\gamma}{\Delta}  }
-  {\goodTerm{\Gamma,\cC\sigma[\gamma]}{s[\gamma^\uparrow]}{A[\pi_1][\gamma^\uparrow]\equiv A[\gamma][\pi_1]}}
-  $$
-  $$
-  \Rule[name=Seal-Id]
-  {}
-  {\goodSeal{\Gamma}{id_s}{\sigma}{\cC \sigma}}
-  = \goodTerm{\Gamma, \cC\sigma}{\pi_2}{\cC\sigma[\pi_1]}
-  $$
-
-\centering
-\captionsetup{justification=centering}
-
-\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a shorthand for the other)}
-
-\begin{minted}[escapeinside=@@]{agda}
-Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
-Seal Γ σ A = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi.
-I fixed it now}\YZreply{Why is tau not used in the definition?}\EDJreply{Thanks.
-I wanted to remove tau but I accidentally removed sigma instead}@
-
--[-]S : (f : Seal Γ σ A) → (τ : Tms Θ Γ) → Seal Θ σ[τ] A[τ]
-f[τ]S = t[τ ↑ A]
-
-idₛ : Seal Γ σ 𝓒σ
-idₛ = π₁
-\end{minted}
-
-\end{figure}
-
-As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
-a shorthand for term judgement $\goodTerm{\Gamma , \cC
-\sigma}{\_}{A[\pi_1]}$,
-a special form of \textit{local morphism}~\cite{abbott2003category}.
-The substitution law should be understood as a lemma, and it is easily
-derivable using $\gamma^\uparrow$.
-
-The motivation for sealing judgement is still about the important term we mentioned earlier: it is acting like syntactic sugar
-helpful for inheritance---due to the presence of inductive type in the
-context, not everything is inheritable. For example, $ X_1 : \bW
-\sigma_1 \vdash t : T$ in the parent family will generally not be able
-to be inherited to the children once we override the inductive type $\bW
-\sigma_1$ with $\bW \sigma_2$. In
-other words, there is no way to ``transplant'' $t$ into the context $X_2
-: \bW \sigma_2$.
-This makes all the fields after this inductive type
-non-inheritable---which shouldn't be the case because a lot of fields
-might not be directly relying on this concrete $\bW \sigma_1$.
-
-Sealing\YZ{It occurs to me that you use the word "sealing" to mean different (though tangentially related) things.
-Previously you used it as in "sealed families".}\EDJreply{They are actually the same thing.
-This sealing in metatheory is much stronger than that of the implemented plugin.
-Downside of the strong sealing is that if the user really use this version of sealing, the user will have to annotate what the sealed family type is.
-Now in the plugin, the sealed family type is almost fixed (directly inferred using a fixed rules).
-So basically this complete version of sealing (in metatheory) can deduce that of seal (in plugin implementation) } is used to exploit this fact---with sealing, given $X_1 : \bW
-\sigma_1$, we can add well-typed terms $X_1 : \cU \vdash t : T$, into
-the parent family. Later when we want to inherit $t$,
-because $t$ only consider $X_1$ as an arbitrary type, inheritance is
-possible.
-This sealing is the general version of the sealed family implemented in the plugin.
-Our plugin will deduce a sealed ``interface'' for a sealed family using a fixed set of rules but here sealing in metatheory can be arbitrary local morphism.
-
-Of course, any fields using "Wsup" and "Wrec" directly relies on the
+Again, we still use \textit{decoupling} of the syntax of the case handlers and exhaustiveness checking to resolve \ref{chg:extensible-inductive-type}.
+The linkage is the part of defining case handlers and thus inheritable like other family.
+While any fields using "Wsup" and "Wrec" directly relies on the
 $\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and
@@
 
-There are also cases we don't need to seal anything (i.e., doing
+There are also cases we don't need to seal anything (i.e., for doing
 exhaustiveness checking when using recursor), and thus the "Seal-Id" is
@@
 that we consider that seal $s$ as a term $("sf'" \ s)$.
-
-With the definition of sealing judgement in mind, we can see
-\ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the
-implementation. In the implementation, we will have "self__" all around
-when defining new fields in our plugin, and this "self__" is exactly the
-$A$ in the context of the two fields---especially when we don't do any
-sealing, this $A = \cC \sigma$, which is exactly how the implementation
-exposes "self__" as a module parameter.
 
@@
 
-Finally we introduce the inheritance judgement. Inheritance are
+Inheritance are
 judgement so naturally second class citizen data (i.e. not something
@@
 Without ``subclassing'', there is no way of confusing inheritance and subtyping.
+Note that, "inhov" is also responsible for ``extending'' inductive type---the extended inductive type directly override the old ones when using metatheory to program, just like how plugin implementation works.
 

### 2022-06-30 coqimpl.tex
Reply comments. Compile is very different from Seal.
Code Highlighting for Family Use lstlisting insteadm but I cannot control the line spacing.

@@
 \begin{figure}[!htb]
-  \begin{minipage}[t]{0.25\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
+  \begin{minipage}[t]{0.30\linewidth}
+\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
 Family A.
-Field a : nat 
-  := 1.
-Field b : a = a
-  := eq_refl.
-Final Field a' : nat 
-  := 1. 
+Field a : nat := 1.
+Field b : a = a := eq_refl.
+Final Field a' : nat := 1. 
 EndFamily.
-\end{minted}
+\end{lstlisting}
   \end{minipage}%\YZ{Can the self's be omitted?}\EDJreply{Yes.
@@
 I have removed all the self, please check.}
-  \begin{minipage}[t]{0.4\linewidth}
+  \begin{minipage}[t]{0.35\linewidth}
 \begin{minted}[fontsize=\footnotesize]{Coq}
@@
   \begin{minipage}[t]{0.45\linewidth}
-\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
+\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
 Family I.
@@
 EndFamily.
-\end{minted}
+\end{lstlisting}
   \end{minipage}
 \begin{minipage}[t]{0.45\linewidth}
-\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
 Family I2 extends I.
@@
 EndFamily.
-\end{minted}
+\end{lstlisting}
   \end{minipage}
@@
   \begin{minipage}[t]{0.32\linewidth}
-\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
 Family B.
@@
 EndFamily.
-
 Family B2 extends B.
@@
 \begin{figure}[!htb]
-\begin{minted}{Coq}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
 Family STLC.
@@
 EndFamily.
-\end{minted}  
+\end{lstlisting}  
 \caption{Example for Global Reasoning and MetaData}\label{fig:plugin-example-global-reasoning-meta-data}
@@
 
-\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that means all of the examples shown in this paper.
-\ref{prop:detectable-partiality} is not supported when using function (infinitary branching) in the constructors, for example, the case of using HOAS to represent functions in "tm".
+\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that includes all of the examples shown in this paper.
+\ref{prop:detectable-partiality} is not supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
 In those cases, we still have the fact that partial recursor is a conservative extension, but we cannot show that partial recursor is the extensional characterization any more。
@@
   pdelim : ∀ (P : D -> Set), .., 
-  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) -> option (P (dᵢ x w)))),  
-  .., ∀ (d : D) -> option (P d)
+  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
+    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
 \end{minted}

### 2022-06-30 lang-design.tex
Code Highlighting for Family Use lstlisting insteadm but I cannot control the line spacing.

@@
   \begin{minipage}[t]{0.47\linewidth}
-\begin{minted}[fontsize=\footnotesize,escapeinside=||]{Coq}
-Family STLC.|\YZ{Does minted support adding extra keywords such as 'Family' and 'extends'?}|
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+Family STLC.
 Inductive ty : Set :=
@@
   | var : id -> term 
-  | lam : id -> term -> term ...
-  | tunit : term 
+  | lam : id -> term -> term 
+  | tunit : term ...
 Fixpoint subst 
@@
 Proposition subst_lemma :
-  has_type (Γ; x ↦ A) t T ->
-  has_type Γ u A ->
-  has_type Γ (subst x u t) T.
+  has_type (G; x ↦ A) t T ->
+  has_type G u A ->
+  has_type G (subst x u t) T.
 Inductive step : term -> term -> Prop 
@@
 EndFamily.
-\end{minted}
+\end{lstlisting}
   \end{minipage}
   \begin{minipage}[t]{0.47\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
 Family STLC_bool extends STLC.

### 2022-06-30 syntax.tex
resolve & add comments

@@
 {\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
-\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}\YZ{What is C_t?
-I don't think you talk about it anywhere.}\EDJreply{I added the explanation together with more details on C.
-Ct is really just intro rule for C}
+\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}
 \\
@@
 
-The reason we need this additional type construct of $\cC\sigma$ here is to support \textit{code reuse}, as the rule of linkages hinders its ability to do abstraction.
-Concretely speaking, (using surface syntax) we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to T\})}{h}{\cL(\{a~:~\cU; b~:~a \to T\})}$\footnote{If we had one, consider the case $T$ is bottom type and $\sigma$ is empty record and thus $a$ is initially bottom type.
-Now we further override $h.a$ using unit type we can derive inconsistency}.
-This term is important for us to abstract the concrete inductive type so that the later field can depend on the abstract interface and achieve \textit{code reuse} reuse.
-On the other hand, we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T) }{h}{(\Sigma a : \cU, a \to T)}$ because of the rule for existential (also existential type just don't allow overriding).
-Thus to allow abstraction happens, we need  $\cC\sigma$ to compile a linkage to module as a part of the context and we also need \ruleref{Lnkg Compile} to specify the compilation from linkage to module.
+The reason we need this additional type construct of $\cC\sigma$ here is
+to support \textit{code reuse}, as the rule of linkages hinders its
+ability to do abstraction. Concretely speaking, (using surface syntax)
+we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to
+T\})}{h}{\cL(\{a~:~\cU; b~:~a \to T\})}$\footnote{If we had one,
+consider the case $T$ is bottom type and $\sigma$ is empty record and
+thus $a$ is initially bottom type. Now we further override $h.a$ using
+unit type we can derive inconsistency\YZ{You are talking about
+overriding here, but you haven't told the reader what overriding means
+in your metatheory.}}. This term is important for us to abstract the
+concrete inductive type so that the later field can depend on the
+abstract interface and achieve \textit{code reuse}. On the other hand,
+we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T)
+}{h}{(\Sigma a : \cU, a \to T)}$ because of the rule for existential
+(also existential type just don't allow overriding\YZ{Why is it desired that this type in the context does not allow overriding (since it already says 'a : U')?}).
+Thus to allow abstraction happens, we need $\cC\sigma$ to compile a
+linkage to module as a part of the context and we also need
+\ruleref{Lnkg Compile} to specify the compilation from linkage to
+module.
 
@@
 Thus we need $\cC\sigma$ in our context.
-\YZ{Why is Csigma needed?
+%\YZ{Why is Csigma needed?
 How is it different from Lsigma?}\EDJreply{It corresponds to the compilation of the family ``type'', which is part of the implementation detail.
@@
 Rather, when describing your Coq plugin, you can say it matches the metatheory.
-PS: you use two terminologies: 'compile' and 'seal', but I guess they mean very similar things. So avoid saying 'compile' here.
-PPS: Sealing means hiding, so what is being hided?
+%PS: you use two terminologies: 'compile' and 'seal', but I guess they mean very similar things. So avoid saying 'compile' here.
+%PPS: Sealing means hiding, so what is being hided?
 Why do you need to hide it?
@@
 \\
-The ultimate reason we have this two notion is that, the introduction rule for linkage fails to provide abstraction mechanism so we have to translate them into module. \\
-Generally speaking Compile : Linkage -> Module, Seal : Module -> Module, we need module here because linkage cannot do abstraction of inductive type. A concrete detail is at the above. \\ 
-I now describe the motivation of compile in the earlier paragraph, right before the "reflecting" section.
+%The ultimate reason we have this two notion is that, the introduction rule for linkage fails to provide abstraction mechanism so we have to translate them into module. \\
+%Generally speaking Compile : Linkage -> Module, Seal : Module -> Module, we need module here because linkage cannot do abstraction of inductive type. A concrete detail is at the above. \\ 
+%I now describe the motivation of compile in the earlier paragraph, right before the "reflecting" section.
 Note that, this compile is inspired by plugin implementation.
@@
 Similarly as I say ``there is no extensible inductive type but only overriding of new inductive type'' because in metatheory it is using overriding to ``extend'' (same as plugin implementation), seal means the same thing, and ``compile'' in metatheory happens in plugin implementation
-}
+%}
 
-However, $\cC$ can only make sure we have $\cL(\{a~:~\bW \sigma; b~:~a \to T\}) \to (\Sigma a:\bW \sigma, a \to T)$, we still need $(\Sigma a:\bW \sigma, a \to T) \to (\Sigma a:\cU \sigma, a \to T)$ to complete this shifting of the context.
-This is sealing's job.
+However, $\cC$ can only make sure we have $\cL(\{a~:~\bW \sigma; b~:~a
+\to T\}) \to (\Sigma a:\bW \sigma, a \to T)$, we still need $(\Sigma
+a:\bW \sigma, a \to T) \to (\Sigma a:\cU \sigma, a \to T)$ to complete
+this shifting of the context. This is sealing's job.
 
@@
 Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
-Seal Γ σ A = Tm (Γ, 𝓒σ) A[π₁]@\YZ{tau?}\EDJreply{it should be pi.
-I fixed it now}\YZreply{Why is tau not used in the definition?}\EDJreply{Thanks.
-I wanted to remove tau but I accidentally removed sigma instead}@
+Seal Γ σ A = Tm (Γ, 𝓒σ) A[π₁]
 
@@
 
-Sealing judgement is still about the important term $h$ we mentioned earlier.
-Thanks to $\cC$, a newly added term $t_{new}$ is now in the context of $\cC\{a : \bW \sigma; ..\}$.
-But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is extended, we would like $t_{new}$ to be in the context  of $\cC\{a : \cU; ..\}$ instead.
-The sealing is responsible for this because we have a sealing $\goodTerm{\Gamma, \cC\{a : \bW \sigma; ..\}}{s}{\cC\{a : \cU; ..\}}$ = $\goodTerm{\Gamma, (\Sigma a : \bW \sigma,..)}{s}{(\Sigma a : \cU,..)}$.
-Then we can construct $t_{new}$ in the context of  $(\Sigma a : \cU,..)$, ready for inheritance.
-Apparently $t_{new}$ in this case cannot be doing \textit{exhaustiveness checking}, but it can freely use any \textit{exhaustiveness-checked} recursor.
-\YZ{It occurs to me that you use the word "sealing" to mean different (though tangentially related) things.
-Previously you used it as in "sealed families".}\EDJreply{They are actually the same thing.
-This sealing in metatheory is much stronger than that of the implemented plugin.
-Downside of the strong sealing is that if the user really use this version of sealing, the user will have to annotate what the sealed family type is.
-Now in the plugin, the sealed family type is almost fixed (directly inferred using a fixed rules).
-So basically this complete version of sealing (in metatheory) can deduce that of seal (in plugin implementation) }
+Sealing judgement is still about the important term $h$ we mentioned
+earlier. Thanks to $\cC$, a newly added term $t_{new}$\YZ{By 'a newly
+added term', do you mean a term that follows field 'a'?} is now in the
+context containing $\cC\{a : \bW \sigma; ..\}$.
+But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is
+extended, we would like $t_{new}$ to be in the context containing
+$\cC\{a : \cU; ..\}$ instead. The sealing is responsible for this
+because we have a sealing $\goodTerm{\Gamma, \cC\{a : \bW \sigma;
+..\}}{s}{\cC\{a : \cU; ..\}}$ = $\goodTerm{\Gamma, (\Sigma a : \bW
+\sigma,..)}{s}{(\Sigma a : \cU,..)}$. Then we can construct
+$t_{new}$ in the context containing $(\Sigma a : \cU,..)$, ready for
+inheritance.
+Apparently $t_{new}$ in this case cannot be doing \textit{exhaustiveness
+checking}, but it can freely use any \textit{exhaustiveness-checked}
+recursor.\YZ{It occurs to me that you use the word "sealing" to mean
+different (though tangentially related) things. Previously you used it
+as in "sealed families".}\EDJreply{They are actually the same thing.
+This sealing in metatheory is much stronger than that of the implemented
+plugin. Downside of the strong sealing is that if the user really use
+this version of sealing, the user will have to annotate what the sealed
+family type is. Now in the plugin, the sealed family type is almost
+fixed (directly inferred using a fixed rules). So basically this
+complete version of sealing (in metatheory) can deduce that of seal (in
+plugin implementation) }
 

### 2022-06-30 syntax.tex
escape

@@
 
-\centering
-\captionsetup{justification=centering}
-
-\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a shorthand for the other)}
+\medskip
 
@@
 Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
-Seal Γ σ A = Tm (Γ, 𝓒σ) A[π₁]
+Seal Γ σ A = Tm (Γ, @$\mathcal{C}$@σ) A[π₁]
 
@@
 
-idₛ : Seal Γ σ 𝓒σ
+idₛ : Seal Γ σ @$\mathcal{C}$@σ
 idₛ = π₁
 \end{minted}
+
+\caption{Seal Judgement, and its Agda Representation (or What do we mean when we say one judgement is a shorthand for the other)}
 

### 2022-06-30 syntax.tex
Save, fixing notation.

--- 2022-06-30 syntax.tex
@@
   \begin{align*}
-    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
-    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
-Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}
+    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\gamma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\gamma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
+    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{T}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
+Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
+I will fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
   \end{align*}

### 2022-06-30 syntax.tex
Reply comments and update.

@@
 consider the case $T$ is bottom type and $\sigma$ is empty record and
-thus $a$ is initially bottom type. Now we further override $h.a$ using
-unit type we can derive inconsistency\YZ{You are talking about
+thus $a$ is initially bottom type.
+Now the direct projection using \ruleref{Sig Proj} $\goodTerm{\Gamma, a : \cU}{h.b}{a \to \bot}$ will lead to problem because every field in linkage has this implicit universal quantification, and for this case we can substitute $a$ with arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top \to \bot}$ is a contradiction\YZ{You are talking about
 overriding here, but you haven't told the reader what overriding means
-in your metatheory.}}. This term is important for us to abstract the
+in your metatheory.}\EDJreply{Ok. I use the projection now instead.}}. This term is important for us to abstract the
 concrete inductive type so that the later field can depend on the
@@
 Sealing judgement is still about the important term $h$ we mentioned
-earlier. Thanks to $\cC$, a newly added term $t_{new}$\YZ{By 'a newly
-added term', do you mean a term that follows field 'a'?} is now in the
+earlier. Thanks to $\cC$, a newly added field $t_{new}$ after $a$ \YZ{By 'a newly
+added term', do you mean a term that follows field 'a'?}\EDJreply{Thanks. Fixed} is now in the
 context containing $\cC\{a : \bW \sigma; ..\}$.

### 2022-06-30 syntax.tex
resolve & add comments

--- 2022-06-30 syntax.tex
@@
   \begin{align*}
-    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\gamma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\gamma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
-    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{T}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
-Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
-I will fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
+    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
+    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
+Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}
   \end{align*}
@@
 checking}, but it can freely use any \textit{exhaustiveness-checked}
-recursor.\YZ{It occurs to me that you use the word "sealing" to mean
-different (though tangentially related) things. Previously you used it
-as in "sealed families".}\EDJreply{They are actually the same thing.
-This sealing in metatheory is much stronger than that of the implemented
-plugin. Downside of the strong sealing is that if the user really use
-this version of sealing, the user will have to annotate what the sealed
-family type is. Now in the plugin, the sealed family type is almost
-fixed (directly inferred using a fixed rules). So basically this
-complete version of sealing (in metatheory) can deduce that of seal (in
-plugin implementation) }
+recursor.
+%\YZ{It occurs to me that you use the word "sealing" to mean
+%different (though tangentially related) things. Previously you used it
+%as in "sealed families".}\EDJreply{They are actually the same thing.
+%This sealing in metatheory is much stronger than that of the implemented
+%plugin. Downside of the strong sealing is that if the user really use
+%this version of sealing, the user will have to annotate what the sealed
+%family type is. Now in the plugin, the sealed family type is almost
+%fixed (directly inferred using a fixed rules). So basically this
+%complete version of sealing (in metatheory) can deduce that of seal (in
+%plugin implementation)}
 
@@
 }
-{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}\YZ{What is R[pi1 circ pi1]?}\EDJreply{It is the composition of the substitution.
-Thanks for pointing out now I have included circ into the typesetting} }
+{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
 \quad 
@@
 }
-{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}}
+{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}\YZ{Why does this Sig not have a superscript?}}
 $$
@@
 $\beta$-rule of "Wrec" (omitted here). Finally \ruleref{Rec-Constr} is
-the used to construct a concrete recursor given a specific linkage of
+used to construct a concrete recursor given a specific linkage of
 handlers. All these rules satisfy substitution laws and omitted here.
-\YZ{Seems like you omit many things here, which causes hardship for the reader (at least for me!).
-I think it's better to list all the rules and decide later what to throw to the appendix.}\EDJreply{Do you mean you want to see substitution laws and beta and eta laws?
-They are just commutating as expected I can write down them if you really want.
-And what else do you think is helpful}
 
-Again, we still use \textit{decoupling} of the syntax of the case handlers and exhaustiveness checking to resolve \ref{chg:extensible-inductive-type}.
-The linkage is the part of defining case handlers and thus inheritable like other family.
-While any fields using "Wsup" and "Wrec" directly relies on the
-$\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and
+Again, we still use \textit{decoupling} of the syntax of the case
+handlers and exhaustiveness checking to resolve
+\ref{chg:extensible-inductive-type}. We use linkages to define
+case handlers and thus they are inheritable like other families.
+While any field using "Wsup" and "Wrec" directly relies on the
+$\bW\sigma$ (can be spotted from \ruleref{Ind-Term} and
 \ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup"
@@
 There are also cases we don't need to seal anything (i.e., for doing
-exhaustiveness checking when using recursor), and thus the "Seal-Id" is
+exhaustiveness checking when using recursor), and thus \ruleref{Seal-Id} is
 used. For notation clarity, we will sometimes use $"sf'"$ to emphasize
-that we consider that seal $s$ as a term $("sf'" \ s)$.
+that we consider that seal $s$ as a term $("sf'" \ s)$.\YZ{What is sf and sf'? sf first appears in Compile rules, so explain sf there.}
 

### 2022-06-30 body.tex

@@
 \maketitle
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{ltblue}{rgb}{0,0.4,0.4}
+\definecolor{dkviolet}{rgb}{0.54,0.17,0.89}
+\definecolor{dkblue}{rgb}{0.0,0.53,0.74}
+\definecolor{dkred}{rgb}{0.76,0.13,0.28}
+
+% lstlisting coq style (inspired from a file of Assia Mahboubi)
+% https://tex.stackexchange.com/questions/434523/coq-syntax-highlighting
+\lstdefinelanguage{Coq}{ 
+    % Anything betweeen $ becomes LaTeX math mode
+    mathescape=true,
+    % Comments may or not include Latex commands
+    texcl=false, 
+    % Vernacular commands
+    morekeywords=[1]{Section, Module, End, Require, Import, Export,
+        Variable, Variables, Parameter, Parameters, Axiom, Hypothesis,
+        Hypotheses, Notation, Local, Tactic, Reserved, Scope, Open, Close,
+        Bind, Delimit, Definition, Let, Ltac, Fixpoint, CoFixpoint, Add,
+        Morphism, Relation, Implicit, Arguments, Unset, Contextual,
+        Strict, Prenex, Implicits, Inductive, CoInductive, Record,
+        Structure, Canonical, Coercion, Context, Class, Global, Instance,
+        Program, Infix, Theorem, Lemma, Corollary, Proposition, Fact,
+        Remark, Example, Proof, Goal, Save, Qed, Defined, Hint, Resolve,
+        Rewrite, View, Search, Show, Print, Printing, All, Eval, Check,
+        Projections, inside, outside, Def,
+        % For Family Poly
+        Family, Extend, Field, EndFamily, SealFamily, FInductive,
+        FRecursor, MetaData, EndMetaData, Closing, Final, Inherit
+        },
+    % Gallina
+    morekeywords=[2]{forall, exists, exists2, fun, fix, cofix, struct,
+        match, with, end, as, in, return, let, if, is, then, else, for, of,
+        nosimpl, when},
+    % Sorts
+    morekeywords=[3]{Type, Prop, Set, true, false, option},
+    % Various tactics, some are std Coq subsumed by ssr, for the manual purpose
+    morekeywords=[4]{pose, set, move, case, elim, apply, clear, hnf,
+        intro, intros, generalize, rename, pattern, after, destruct,
+        induction, using, refine, inversion, injection, rewrite, congr,
+        unlock, compute, ring, field, fourier, replace, fold, unfold,
+        change, cutrewrite, simpl, have, suff, wlog, suffices, without,
+        loss, nat_norm, assert, cut, trivial, revert, bool_congr, nat_congr,
+        symmetry, transitivity, auto, split, left, right, autorewrite},
+    % Terminators
+    morekeywords=[5]{by, done, exact, reflexivity, tauto, romega, omega,
+        assumption, solve, contradiction, discriminate},
+    % Control
+    morekeywords=[6]{do, last, first, try, idtac, repeat},
+    % Comments delimiters, we do turn this off for the manual
+    morecomment=[s]{(*}{*)},
+    % Spaces are not displayed as a special character
+    showstringspaces=false,
+    % String delimiters
+    morestring=[b]",
+    morestring=[d],
+    % Size of tabulations
+    tabsize=3,
+    % Enables ASCII chars 128 to 255
+    extendedchars=false,
+    % Case sensitivity
+    sensitive=true,
+    % Automatic breaking of long lines
+    breaklines=false,
+    % Default style fors listings
+    basicstyle=\linespread{0.5}\ttfamily\footnotesize,
+    % Position of captions is bottom
+    captionpos=b,
+    % flexible columns
+    columns=[l]flexible,
+    % Style for (listings') identifiers
+    identifierstyle={\ttfamily\color{black}},
+    % Style for declaration keywords
+    keywordstyle=[1]{\ttfamily\color{dkviolet}},
+    % Style for gallina keywords
+    keywordstyle=[2]{\ttfamily\color{dkgreen}},
+    % Style for sorts keywords
+    keywordstyle=[3]{\ttfamily\color{ltblue}},
+    % Style for tactics keywords
+    keywordstyle=[4]{\ttfamily\color{dkblue}},
+    % Style for terminators keywords
+    keywordstyle=[5]{\ttfamily\color{dkred}},
+    %Style for iterators
+    %keywordstyle=[6]{\ttfamily\color{dkpink}},
+    % Style for strings
+    stringstyle=\ttfamily,
+    % Style for comments
+    commentstyle={\ttfamily\color{dkgreen}},
+    %moredelim=**[is][\ttfamily\color{red}]{/&}{&/},
+    literate=
+    {\\forall}{{\color{dkgreen}{$\forall\;$}}}1
+    {\\exists}{{$\exists\;$}}1
+    {<-}{{$\leftarrow\;$}}1
+    {=>}{{$\Rightarrow\;$}}1
+    {==}{{\code{==}\;}}1
+    {==>}{{\code{==>}\;}}1
+    %    {:>}{{\code{:>}\;}}1
+    {->}{{$\rightarrow\;$}}1
+    {<->}{{$\leftrightarrow\;$}}1
+    {<==}{{$\leq\;$}}1
+    {\#}{{$^\star$}}1 
+    {\\o}{{$\circ\;$}}1 
+    {\@}{{$\cdot$}}1 
+    {\/\\}{{$\wedge\;$}}1
+    {\\\/}{{$\vee\;$}}1
+    {++}{{\code{++}}}1
+    {~}{{\ }}1
+    {\@\@}{{$@$}}1
+    {\\mapsto}{{$\mapsto\;$}}1
+    {\\hline}{{\rule{\linewidth}{0.5pt}}}1
+    %
+}[keywords,comments,strings]
 
@@
 
-% Macro for all use
-
+% Macro for all use, the macro for metatheory syntax
 \newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}

### 2022-06-30 syntax.tex
comments

@@
 {\goodSig{\Gamma}{\sigma}{n} 
- \quad \goodSeal{\Gamma}{f}{\sigma}{A}\YZ{How should I understand 'f' and 'A'?}\EDJreply{`f' is the seal, 'A' is the result of sealing.
-I think I wrote down the motivation of sealing...
-Is it still too hard?
-Do you think I need to think of a more friendly way?}\YZreply{Yes, still too rough on the reader.
-I sense that this part is central to the metatheory, so it is worth explaining the details well.}
+ \quad \goodSeal{\Gamma}{f}{\sigma}{A}
  \quad \goodType{\Gamma, A}{T}{i}}
@@
 consider the case $T$ is bottom type and $\sigma$ is empty record and
-thus $a$ is initially bottom type.
-Now the direct projection using \ruleref{Sig Proj} $\goodTerm{\Gamma, a : \cU}{h.b}{a \to \bot}$ will lead to problem because every field in linkage has this implicit universal quantification, and for this case we can substitute $a$ with arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top \to \bot}$ is a contradiction\YZ{You are talking about
-overriding here, but you haven't told the reader what overriding means
-in your metatheory.}\EDJreply{Ok. I use the projection now instead.}}. This term is important for us to abstract the
+thus $a$ is initially bottom type. Now the direct projection using
+\ruleref{Sig Proj} $\goodTerm{\Gamma, a : \cU}{h.b}{a \to \bot}$ will
+lead to problem because every field in linkage has this implicit
+universal quantification, and for this case we can substitute $a$ with
+arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top
+\to \bot}$ is a contradiction}.
+This term is important for us to abstract the
 concrete inductive type so that the later field can depend on the
-abstract interface and achieve \textit{code reuse}. On the other hand,
+abstract interface and achieve \textit{code reuse}.
+\YZ{It reads as if you want to have 'h : L\{a : U, b : a -> T\}'. But it is problematic because it leads to inconsistency, as you describe in a footnote.}
+On the other hand,
 we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T)
 }{h}{(\Sigma a : \cU, a \to T)}$ because of the rule for existential
-(also existential type just don't allow overriding\YZ{Why is it desired that this type in the context does not allow overriding (since it already says 'a : U')?}).
+(also existential type just don't allow overriding\YZ{This sentence still speaks of overriding. Explain in terms of projection instead?}).
 Thus to allow abstraction happens, we need $\cC\sigma$ to compile a
@@
 Sealing judgement is still about the important term $h$ we mentioned
-earlier. Thanks to $\cC$, a newly added field $t_{new}$ after $a$ \YZ{By 'a newly
-added term', do you mean a term that follows field 'a'?}\EDJreply{Thanks. Fixed} is now in the
-context containing $\cC\{a : \bW \sigma; ..\}$.
+earlier. Thanks to $\cC$, a field $t_{new}$ that depends on $a$ is now
+in the context containing $\cC\{a : \bW \sigma; ..\}$.
 But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is

### 2022-06-30 syntax.tex
Small fix.

--- 2022-06-30 syntax.tex
@@
 (Using surface synax), consider the example of a linkage $\{a~:~\bW \sigma; b_1 : a \to T;..\}$ with a bunch of $b_i$ \textit{using} $a$.
-The tension we need to solve here is between \textit{exhaustiveness checking} and \textit{inheritance (code reuse)}---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to construct $b_i$ term in the context of $\bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that.
+The tension we need to solve here is between \textit{exhaustiveness checking} and \textit{inheritance (code reuse)}---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to construct $b_i$ term in the context of $a : \bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that and thus we hope $b_i$ constructed in the context of $a : \cU$ (and thus later this $a$ can be compatible with inductive type of any signature).
 Our metatheory formulation of linkage is mainly trying to solve this tension.
@@
 concrete inductive type so that the later field can depend on the
-abstract interface and achieve \textit{code reuse}. On the other hand,
+``abstract interface'' (i.e. $\cL(\{a:\cU~;~b:a \to T\})$) and achieve \textit{code reuse}. \\ On the other hand,
 we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T)
@@
 when defining new fields in our plugin, and this "self__" is exactly the
-$A$ in the context of the two fields---especially when we don't do any
-sealing, this $A = \cC \sigma$, which is exactly how the implementation
-exposes "self__" as a module parameter.
+$A$ in the context of the two fields---in our implementation, there is only sealing of inductive type happening (where each inductive type we only know \mintinline{Coq}{T : Set}).
 

### 2022-07-01 syntax.tex

@@
 consider the case $T$ is bottom type and $\sigma$ is empty record and
-thus $a$ is initially bottom type.
-Now the direct projection using \ruleref{Sig Proj} $\goodTerm{\Gamma, a : \cU}{h.b}{a \to \bot}$ will lead to problem because every field in linkage has this implicit universal quantification, and for this case we can substitute $a$ with arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top \to \bot}$ is a contradiction\YZ{You are talking about
-overriding here, but you haven't told the reader what overriding means
-in your metatheory.}\EDJreply{Ok. I use the projection now instead.}}. This term is important for us to abstract the
+thus $a$ is initially bottom type. Now the direct projection using
+\ruleref{Sig Proj} $\goodTerm{\Gamma, a : \cU}{h.b}{a \to \bot}$ will
+lead to problem because every field in linkage has this implicit
+universal quantification, and for this case we can substitute $a$ with
+arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top
+\to \bot}$ is a contradiction}.
+However, this underivable term is important for us to abstract the
 concrete inductive type so that the later field can depend on the
-``abstract interface'' (i.e. $\cL(\{a:\cU~;~b:a \to T\})$) and achieve \textit{code reuse}. \\ On the other hand,
+``abstract interface'' (i.e. something like $\cL(\{a:\cU~;~b:a \to T\})$) and achieve \textit{code reuse}. 
+\YZ{It reads as if you want to have 'h : L\{a : U, b : a -> T\}'.
+But it is problematic because it leads to inconsistency, as you describe in a footnote.}\EDJreply{Good point.
+I rephrased it so the introduction of sigma type is smoother.
+Please check.}
+\\ The good news, we don't have to stick with $\cL(\sigma)$ to express the abstract interface. We can also use sigma type, and actually 
 we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T)
-}{h}{(\Sigma a : \cU, a \to T)}$ because of the rule for existential
-(also existential type just don't allow overriding\YZ{Why is it desired that this type in the context does not allow overriding (since it already says 'a : U')?}).
+}{h}{(\Sigma a : \cU, a \to T)}$ because of the rules for sigma type.\footnote{The intro rule for sigma type decides that there is no such implicit universal quantification happening.}
+\YZ{This sentence still speaks of overriding. Explain in terms of projection instead?}\EDJreply{Fixed. Please check.}
 Thus to allow abstraction happens, we need $\cC\sigma$ to compile a
@@
 }
-{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}\YZ{Why does this Sig not have a superscript?}}
+{\goodSig{\Gamma}{"RecSig"~\sigma~R}{n}\YZ{Why does this Sig not have a superscript?}}
 $$

### 2022-07-01 coqimpl.tex
Fixing notation: sigma is for linkage signature gamma is for substitution tau is for inductive signature

@@
-\subsection{Language Design}
+\subsection{Language Design}\YZ{I think in §3.1, you should use a figure to show how the actual code the programmer will write to solve the programming challenge outlined in Figure 1.}
 Inspired by our above challenges and Family Polymorphism, we propose

### 2022-07-01 standardmodel.tex
Fixing notation: sigma is for linkage signature gamma is for substitution tau is for inductive signature

@@
   \denotesS{\goodWSig{\Gamma}{\_}{n}} &\text{ is a Set of list of pairs of type interpretation of length } n \\
-  & \text{ and we define } \goodWSigS{\Gamma}{w}{n} :\iff \denotesS{\goodWSig{\Gamma}{w}{n}} \in \denotesS{\goodWSig{\Gamma}{\_}{n}}\\
+  & \text{ and we define } \goodWSigS{\Gamma}{\tau}{n} :\iff \denotesS{\goodWSig{\Gamma}{\tau}{n}} \in \denotesS{\goodWSig{\Gamma}{\_}{n}}\\
   \denotesS{\goodWSig{\Gamma}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 
-  \denotesS{\goodWSig{\Gamma}{(w^+ \ A \ B \ w)}{n+1}}
-  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gamma}{w}{n}} \\ 
-  \denotesS{\goodType{\Gamma}{\pi^j_1~\sigma}{}} &= "fst"~({\denotesS{\sigma}}_{j})\\
-  \denotesS{\goodType{\Gamma, \pi^j_1~\sigma}{\pi^j_2~\sigma}{}} &= "snd"~({\denotesS{\sigma}}_j)\\
+  \denotesS{\goodWSig{\Gamma}{(w^+ \ A \ B \ \tau)}{n+1}}
+  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gamma}{\tau}{n}} \\ 
+  \denotesS{\goodType{\Gamma}{\pi^j_1~\tau}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
+  \denotesS{\goodType{\Gamma, \pi^j_1~\tau}{\pi^j_2~\tau}{}} &= "snd"~({\denotesS{\tau}}_j)\\
 \end{align*}
@@
   \begin{minted}[mathescape, escapeinside=||]{Agda}
-    data MW {i}{j} (σ : |$\denotesS{\texttt{WSig i }\cdot\texttt{ S}}$|) : Set (lsuc i) where
-      MWsup : |$n$| < S → (a : |$\pi^n_1$| σ) → ((|$\pi^n_2$| σ) a → MW σ) 
-              → MW σ
+    data MW {i}{j} (τ : |$\denotesS{\texttt{WSig i }\cdot\texttt{ S}}$|) : Set (lsuc i) where
+      MWsup : |$n$| < S → (a : |$\pi^n_1$|τ) → ((|$\pi^n_2$|τ) a → MW τ) 
+              → MW τ
       \end{minted}

### 2022-07-01 syntax.tex
Fixing notation: sigma is for linkage signature gamma is for substitution tau is for inductive signature

@@
 follows the style of~\citet{kaposi2019gluing}, but for readability we
-omit universe levels attached to judgements.
+omit the universe levels. We still working on predicative MLTT.
 %\YZ{What's the distinction of judgments vs.
@@
 shorthand for term judgement
-$\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$.
+$\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
 
@@
     &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\gamma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\gamma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
-    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{T}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
+    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
 Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
@@
 
-\judgebox{\goodWSig{\Gamma}{\sigma}{n}}
+\judgebox{\goodWSig{\Gamma}{\tau}{n}}
 $$
@@
 \Rule[name=WSig-Add]
-{\goodWSig{\Gamma}{s}{n}
+{\goodWSig{\Gamma}{\tau}{n}
   \quad \goodType{\Gamma}{A}{i}
   \quad \goodType{\Gamma, A}{B}{i}}
-{\goodWSig{\Gamma}{w^+ \  s \  A \  B}{n+1}}
+{\goodWSig{\Gamma}{w^+ \  \tau \  A \  B}{n+1}}
 \quad
 \Rule[name=Ind-Univ]
-{\goodWSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{\bW \sigma}{i}}
+{\goodWSig{\Gamma}{\tau}{n}}
+{\goodType{\Gamma}{\bW \tau}{i}}
 $$
@@
 \Rule[name=WSig-Proj]
-{\goodWSig{\Gamma}{s}{n} \quad j < n}
-{\goodType{\Gamma}{\pi^j_1 s}{i} \quad \goodType{\Gamma, \pi^j_1 s}{\pi^j_2  s}{i}}
+{\goodWSig{\Gamma}{\tau}{n} \quad j < n}
+{\goodType{\Gamma}{\pi^j_1 \tau}{i} \quad \goodType{\Gamma, \pi^j_1 \tau}{\pi^j_2  \tau}{i}}
 \quad
 \Rule[name=Ind-Sig]
-{\goodWSig{\Delta}{\sigma}{n}
+{\goodWSig{\Delta}{\tau}{n}
   \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodWSig{\Gamma}{\sigma[\gamma]}{n}
-  \quad \goodType{\Gamma}{W (\sigma[\gamma]) \equiv (W \sigma)[\gamma]}{i}}
+{\goodWSig{\Gamma}{\tau[\gamma]}{n}
+  \quad \goodType{\Gamma}{W (\tau[\gamma]) \equiv (W \tau)[\gamma]}{i}}
 $$
@@
 \Rule[name=Ind-Type]
-{\goodTerm{\Gamma}{T}{\bW \sigma}}
+{\goodTerm{\Gamma}{T}{\bW \tau}}
 {\goodTerm{\Gamma}{"W" \ T}{\cU}}
@@
 \Rule[name=Ind-Term]
-{\goodTerm{\Gamma}{T}{\bW \sigma}
-  \quad \goodTerm{\Gamma}{a}{\pi^j_1\sigma}
-  \quad \goodTerm{\Gamma, \pi^j_2\sigma[(id, a)]}{b}{El~("W"~T)}}
-{\goodTerm{\Gamma}{"Wsup" \ T \ a \ b}{El~("W"~T)} }
+{\goodTerm{\Gamma}{T}{\bW \tau}
+  \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
+  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)}}
+{\goodTerm{\Gamma}{"Wsup"~T~a~b}{El~("W"~T)} }
 $$
@@
 -[-]S : (f : Seal Γ σ A) → (τ : Tms Θ Γ) → Seal Θ σ[τ] A[τ]
-f[τ]S = t[τ ↑ A]
+f[τ]S = t[@$\tau^\uparrow$@]
 

### 2022-07-01 syntax.tex
Explaining sf

@@
 Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
-I will fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
+I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
   \end{align*}
@@
 
+For notation clarity, we will sometimes use $"sf'"$ to emphasize
+that we consider that seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term $\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]} $ and "sf" notation (used in \ruleref{Compile}) is defined as $\goodSub{\Gamma, \cC \sigma}{sf~s := (\pi_1, sf'~s)}{\Gamma, A}$.\YZ{What is sf and sf'?
+sf first appears in Compile rules, so explain sf there.}\EDJreply{Fixed.
+They both are explained at here now}
+
 Sealing judgement is still about the important term $h$ we mentioned
@@
 }
-{\goodSig{\Gamma}{"RecSig"~\sigma~R}{n}\YZ{Why does this Sig not have a superscript?}}
+{\goodSig{\Gamma}{"RecSig"~\sigma~R}{n}\YZ{Why does this Sig not have a superscript?}\EDJreply{Fixed now.}}
 $$
@@
 exhaustiveness checking when using recursor), and thus \ruleref{Seal-Id} is
-used. For notation clarity, we will sometimes use $"sf'"$ to emphasize
-that we consider that seal $s$ as a term $("sf'" \ s)$.\YZ{What is sf and sf'? sf first appears in Compile rules, so explain sf there.}
+used.
 

### 2022-07-01 coqimpl.tex
Explaining sf
Adding more refering back to consistency issue.

@@
 We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
-"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.}
+"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
 We can define nested families.
@@
 % explaining inheritance
-Besides standalone family, we also need to implement family inheritance.
+\textbf{Implementing Inheritance.}
+To resolve \ref{chg:consistency}, we make sure when inheritance happens, not only the content will be inherited, but the context of the corresponding field is also ``inherited''.
+Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty and has no "g" in "self".
+
 To simplify the implementation, we can consider all families as
@@
 There are three kinds of inheritance---extend, override, and inherit. We
-use first-order data to encode inheritance data inside the
+use a list to encode inheritance data inside the
 implementation of the plugin. The inheritance data will include the
@@
 \text{mix-in} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance.
-However, a good definition of mixin is still under investigation.\YZ{How about making mixin composition one of your Req's?
+However, a good definition of mixin is still under investigation.
+We will show one practical example of \textit{mixin} later.\YZ{How about making mixin composition one of your Req's?
 Also does any example in the paper illustrate how to do this?}\EDJreply{mixin is not really formalized.
@@
 % explain ftheorem wrapping this complicated recursor
-Finally, we also provide a command "FTheorem" in programming interface when the users want to use tactic
-programming instead. This command (1) can avoid most boilerplate code (e.g., specify the handler family) when writing a recursor like above, (2) invoke
+This design of decoupling of the case handlers and exhaustiveness checking handles the semantic well, however, it brings a lot of boilerplate code---we need to create an extra (handler) family, manually specifying the type of each case handlers, and then using "FRecursor" to ``tie the knot'' and construct a recursor field from the family.
+We thus provide a command "FTheorem" to remove these boilerplate in programming interface when using tactic programming instead. 
+This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
 proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor".\YZ{example?}\EDJreply{FTheorem is really just a short hand for "FRecursor" + handler family.
 Nothing more conceptual here.
-If the reader is interested in the concrete syntax, I think I need to make the two examples avaliable in the appendix.}
+If the reader is interested in the concrete syntax, I think I need to make the two examples avaliable in the appendix.} We expect this "FTheorem" to be used in theorem proving but not general programming.
 
@@
 being defined.
-% \YZ{Please check if my wording is correct.}\EDJreply{Yes. Resolved}
 Since we are dealing with a concrete (vanilla) inductive type during compilation,

### 2022-07-01 lang-design.tex
Save, fixing notation.
Adding more refering back to consistency issue.

@@
 The metatheoretical formulation in~\citet{zm2017} for family polymorphism is very natural as they consider extension of the family as the concatenation of a new linkage.
-However, this formulation is not directly adaptable for dependent type theory because it can easily encode non-termination (mutual calling of two function member) sabotaging consistency by the presence of "self".
+However, this formulation is not directly adaptable for dependent type theory because it can easily encode non-termination sabotaging consistency by the presence of "self".
+Informally, in the orignal formulation, we can override the "f" in family $\{"f" := id; "g" := "call self.f"\}$ with "call self.g", resulting the mutual calling of "f" and "g" thus a deadloop.
 This conflict between self parametrization and consistency will be another challenge---we need change the original formulation of family and inheritance for dependent type theory.
@@
 Looking at the bigger picture, I think this section reads better if phrased as discussing challenges posed by adding family polymorphism to a proof assistant.
-I have some initial ideas that I need to discuss with you.}
+I have some initial ideas that I need to discuss with you.}\EDJreply{Design Space section is rewritten into Challenge now.}
 

### 2022-07-01 syntax.tex
Adding more refering back to consistency issue.

@@
 
+Now, together with sealing, we can read \ruleref{Sig Add} and \ruleref{Lnkg Add} out clearly: given a constructed linkage $o : \cL\sigma$ of signature $\sigma$, the next added field $t : T$ will have to be in the context of the abtraction $A$ of the compilation $\cC\sigma$ (witnessed by the seal $f$).
+The ``concatenation'' of the $o$ and $t$ leads to a new linkage, of which the signature is the ``concatenation'' of the $\sigma$ and $T$ (we also record the sealing $f$ for simplicity).
+The construction of linkage in \ruleref{Lnkg Add} requires $t$ to \textbf{only} refer to the (abstraction $A$ of) former part of the defining family $o : \cL\sigma$, this is to avoid inconsistency mentioned in \ref{chg:consistency} in metatheory.
+
 This sealing is the general version of the sealed family implemented in the plugin.
@@
 very much like a function that transform a given linkage.
+Like plugin implementation and \ruleref{Lnkg Add}, we resolve \ref{chg:consistency} by making sure corresponding fields are in corresponding context during inheritance, and an overriding term can only refer to fields before it.
 
@@
 concrete family. However, since our current formulation doesn't support
-further-binding (i.e. we cannot do inheritance inside a family), our
+nested inheritance (i.e. we cannot do inheritance inside a family but only top level), our
 formulation still has some distance to the real family polymorphism.

### 2022-07-01 coq-example.tex
Add diagram for abs-int-example.

@@
 The second example is adapted and modified from \citet{zm2017}.
-Contrary to our first example, we use a big-step interpreter with fuel
+Contrary to our first example, here, we use a big-step interpreter with fuel
 as the operational semantics of an imperative language with
@@
 
-We construct four families, in one inheritance chain.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful.
+We construct four families, in one inheritance chain, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful.
 They are here to show family inheritance is useful, and can extend stuff to all sorts of things.
@@
 The first family "Lang" define the big step operator (using fuel) of a
-while language, with expression. However, it postulates the definition of "State" (environment mapping) used by Lang. It takes about 200+ LOCs. 
+while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200+ LOCs. 
 % We
@@
 
-The second family "LangwAbs" define the abstract interpreter for it,
-together with a bunch of postulates, to prove the correctness of the
-abstract interpreter. It takes about 300+ LOCs. This example shows how
+The second family "LangwAbs" define the abstract interpreter for "Lang" in "LangwAbs.AbsInterpreter",
+together with a bunch of postulates as a big sealed family "absStateData". Then we prove the correctness of the
+abstract interpretation function "analyze". It takes about 300+ LOCs. This example shows how
 to use family inheritance to extend a concrete interpreter with abstract
-interpreters. This example also shows in this family polymoprhism
-framework we can reason about "Lang" with computation details being
-abstracted.
+interpreters. 
+% This example also shows in this family polymoprhism framework we can reason about "Lang" with computation details being abstracted.
 
-The third Family "LangMore" extend LangwAbs with nat constant and
+The third Family "LangMore" extend "LangwAbs" with nat constant and
 addition, and if-then-else control flow, of course retaining all the
@@
 using family inheritance to support new language feature, and compatible
-with the existent reasoning.
+with the existent reasoning in "LangwAbs.AbsInterpreter".
 
 The fourth family "LangRun" instantiate some of the postulates of the
-LangMore and recover computational information of Lang. The concrete
-interpreter is doing as expected; the abstract interpreter is expected
+"LangMore" and recover computational information of "analyze". The "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
 to act as a type-analyzer. It takes about 200+ LOCs. This is one example
-of instantiate the detail computation.
-Thanks to the posulation we have done in "Lang", now we instantiate that postulation with function type "Ident -> option exp" as the "State" (environment mapping) for simplicity.
+of instantiate the detail computation of abstract interpretation.
 
@@
 some simple queries. This "LangRun" also illustrates that we still have
-computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}
+computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}\EDJreply{Done.
+Please check}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
+  \caption{Inheritance Diagram for Abstract Interpretation Example.
+Black arrows are inheritance.
+Purple arrow are extending.
+Green arrows are overriding.
+Grey box are sealed families.}\label{fig:abstract-interpretation-example}
+\end{figure}

### 2022-07-01 coqimpl.tex
Add diagram for abs-int-example.

@@
 are located in the family "handlers". This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family.
+
+\begin{figure}[!htb]
+  \begin{minipage}[t]{0.47\linewidth}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+Family STLC.
+FInductive ty : Set :=
+  | unit : ty
+  | arrow : ty -> ty -> ty.
+FInductive term : Set := 
+  | var : id -> term 
+  | lam : id -> term -> term ...
+ Family subst_internal.
+ EndFamily.
+
+  : id -> term -> term := ...
+Inductive has_type 
+  : context -> term -> type := ...
+Proposition subst_lemma :
+  has_type (G; x ↦ A) t T ->
+  has_type G u A ->
+  has_type G (subst x u t) T.
+Inductive step : term -> term -> Prop 
+  := ...
+(* ... and more, end with type safety *)
+EndFamily.
+\end{lstlisting}
+  \end{minipage}
+  \begin{minipage}[t]{0.47\linewidth}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+Family STLC_bool extends STLC.
+Extend FInductive ty : Set :=
+  | bool : ty.
+
+Extend FInductive term : Set := 
+  | tt : term | ff : term 
+  | tif : term -> term -> term -> term
+Extend Fixpoint subst (* Need to handle new term *)
+  : id -> term -> term := ...
+Extend Inductive has_type (* .. and new ty too *)
+  : context -> term -> type := ...
+Extend Proposition subst_lemma :
+  ... (* Need to prove extra cases *)
+
+Extend Inductive step : term -> term -> Prop 
+  := ... (* Need to expand this binary relation *)
+(* ... and more extension *)
+EndFamily.
+\end{lstlisting}
+  \end{minipage}
+  \caption{Example STLC and its extension, in designed syntax}\label{fig:STLC-example2}
+\end{figure}
 

### 2022-07-01 syntax.tex
Add diagram for abs-int-example.

@@
 very much like a function that transform a given linkage.
-Like plugin implementation and \ruleref{Lnkg Add}, we resolve \ref{chg:consistency} by making sure corresponding fields are in corresponding context during inheritance, and an overriding term can only refer to fields before it.
+Similar to plugin implementation and \ruleref{Lnkg Add}, we resolve \ref{chg:consistency} by making sure corresponding fields are in corresponding context during inheritance, and thus an overriding term can only refer to fields before the overriden field.
 

### 2022-07-02 coqimpl.tex
Update STLC example in 3.1

@@
   & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
+  & | \quad "FTheorem"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle \\
   & |\quad "Extend"~"Family"~\langle ident\rangle \\ 
@@
 \end{align*}
-
-These are some core extensions we proposed to support \cref{fig:STLC-example}.
-We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
-"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
-We can define nested families.
-
-To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness.
-The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family.
-Then
-"FRecursor" will be used to construct a recursor using the handler family.
-For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
-"handlers" will generate a new field "recur" that is doing induction
-(using elimination principle of "T") using motive "f", and the handlers
-are located in the family "handlers". This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family.
 
@@
  Family subst_internal.
+  Final Field var := ..
+  Final Field lam := ..
+   ...
  EndFamily.
-
-  : id -> term -> term := ...
-Inductive has_type 
+ FRecursor subst
+  about tm motive ..
+  using subst_internal.
+ ...
+ FInductive has_type 
   : context -> term -> type := ...
-Proposition subst_lemma :
-  has_type (G; x ↦ A) t T ->
-  has_type G u A ->
-  has_type G (subst x u t) T.
-Inductive step : term -> term -> Prop 
+ FTheorem subst_lemma 
+  about has_type 
+  motive ...
+  ..
+ Qed.
+ FInductive step : term -> term -> Prop 
   := ...
@@
   | tif : term -> term -> term -> term
-Extend Fixpoint subst (* Need to handle new term *)
-  : id -> term -> term := ...
-Extend Inductive has_type (* .. and new ty too *)
+ Extend Family subst_internal.
+  Final Field tt := ..
+  Final Field ff := ..
+  Final Field tif := ..
+ EndFamily.
+ Inherit subst.
+(* Plugin will redo exhaustiveness check, 
+  in case subst_internal is not 
+  extended correctly *)
+ Extend FInductive has_type (* .. and new ty too *)
   : context -> term -> type := ...
-Extend Proposition subst_lemma :
+ Extend FTheorem subst_lemma.
   ... (* Need to prove extra cases *)
+ Qed.
 
-Extend Inductive step : term -> term -> Prop 
+ Extend FInductive step : term -> term -> Prop 
   := ... (* Need to expand this binary relation *)
@@
 \end{figure}
+
+These are some core extensions we proposed to support \cref{fig:STLC-example}.
+We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
+"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
+We can define nested families.
+
+To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness.
+The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family.
+Then
+"FRecursor" will be used to construct a recursor using the handler family.
+For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
+"handlers" will generate a new field "recur" that is doing induction
+(using elimination principle of "T") using motive "f", and the handlers
+are located in the family "handlers". This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family. "FTheorem" is used for theorem proving
 

### 2022-07-02 coqimpl.tex
Going through before metatheory for small-fixes.

@@
-\subsection{Language Design}\YZ{I think in §3.1, you should use a figure to show how the actual code the programmer will write to solve the programming challenge outlined in Figure 1.}
-Inspired by our above challenges and Family Polymorphism, we propose
-the following extension to Coq's Vernacular Commands.
+\subsection{Language Design}\YZ{I think in §3.1, you should use a figure to show how the actual code the programmer will write to solve the programming challenge outlined in Figure 1.}\EDJreply{Done.}
+Inspired by the above challenges and Family Polymorphism, we propose
+the following extension to Coq's Vernacular Commands, and the corresponding STLC example.
 % basically write down the specification of surface Coq commandline
@@
   Final Field tif := ..
- EndFamily.
+ EndFamily. (* Automatic inheriting others 
+              from STLC.subst_internal *)
  Inherit subst.
-(* Plugin will redo exhaustiveness check, 
-  in case subst_internal is not 
-  extended correctly *)
+(* Plugin will redo exhaustiveness check, in case 
+ subst_internal is not extended correctly *)
  Extend FInductive has_type (* .. and new ty too *)
@@
 
-These are some core extensions we proposed to support \cref{fig:STLC-example}.
+These are the core extensions we proposed to support \cref{fig:STLC-example}.
 We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
 "FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
-We can define nested families.
+Nested families are also allowed.
 
@@
 Then
-"FRecursor" will be used to construct a recursor using the handler family.
-For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
-"handlers" will generate a new field "recur" that is doing induction
-(using elimination principle of "T") using motive "f", and the handlers
-are located in the family "handlers". This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family. "FTheorem" is used for theorem proving
+"FRecursor" will be used to construct a recursor by refering to the handler family.
+\mintinline{Coq}{FRecursor recur about T motive f } "using"
+"handlers" will generate a new field "recur" that is really doing computation
+(using elimination principle of the inductive type "T") with motive "f". The handlers
+are located in the family "handlers", similar to a large pattern-matching statement.
+This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family.
+For example, in \cref{fig:STLC-example2},  "FTheorem" has a similar functionality but is expected to be used for theorem proving---the handler and the family will be automatically constructed during proving.
 
@@
 are order-sensitive because we have dependent types here. So "Inherit
-Until" is used if the user wants to inherit a range of fields. If the
-user wants to extend an inductive type or a nested family from the
-parent, then "Extend FInductive" or "Extend Family" is used. Thus nested
-family can only be inherited or extended together with its outer family
-(and thus top level family), thus our support of family polymorphism is
+Until" is used if the user wants to inherit a range of fields.  "Extend Family" is used only inside a inheriting family. Thus nested
+family can only be inherited or extended together with its belonging family
+(i.e. ultimately top level family), thus our support of family polymorphism is
 incomplete.
@@
 
-After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
+After considering the pros and cons, we decide to implement a Coq Plugin where we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the orignal surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
 
@@
 
-\textbf{Family compiled into module.} We support family by compiling
+\textbf{Family compiled into module.} The main idea of our plugin is to support family by compiling
 family components into Coq modules, family ``types'' into module types, and the context of a given term
@@
 choose module and module parameters to achieve this wrapping---each
-family type in the context of the judgement will be one module parameter
+family type in the context will be one module parameter
 with the compiled module type, and thus each field of the family will
@@
 % Concrete details of compilation of a family term data structure
-Taking \cref{fig:plugin-example1} as an example: when the user
-interactively inputs the line "Field a : nat := ..", our plugin will
+Taking \cref{fig:plugin-example1} as an example: immediately after the user inputs the line "Field a : nat := ..", our plugin will
 translate this statement into the "Module a_4", "Module Type a_5, a_6"
@@
 To resolve \ref{chg:consistency}, we make sure when inheritance happens, not only the content will be inherited, but the context of the corresponding field is also ``inherited''.
-Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty and has no "g" in "self".
+Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty thus "self.g" doesn't type-check.
 
@@
 
-There are three kinds of inheritance---extend, override, and inherit. We
-use a list to encode inheritance data inside the
-implementation of the plugin. The inheritance data will include the
+There are three kinds of inheritance---extend, override, and inherit. Simiarly, we
+use a list to encode inheritance data in the
+implementation. The inheritance data will include the
 information of (1) the ``type'' of the parent family, (2) a list of
@@
 Though inheritance is solely a first-order data rather
-than a higher-order functor, it is still possible to
+than a ``mapping'', it is still possible to
 \text{mix-in} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance.
@@
 % future).
-Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the
+Take the field "c" of "Family I" in \cref{fig:plugin-example3} as the
 example. We can see that the first attempt of proving "a = 1" fails.
@@
 to override "a" in the future, but the mere existence of the field "c"
-will stop that, since "c" witnesses "a" fix to "1". In our implementation, \textit{this will fail}---the proof of "c" will require a concrete definition of "a",
-but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in the context where we only know \mintinline{Coq}{a : nat} thus failed.
+will stop that, since "c" witnesses "a" fix to "1". \textit{This will fail} in our implementation---the proof of "c" will require a concrete definition of "a",
+but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in (a module with) the context where we only know \mintinline{Coq}{a : nat} thus failed.
 
@@
 \textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.} 
-Non-overridable fields are still not expressive enough to achieve the full
-extent of dependent type programming. There are cases where overriding and
-(definition-relevant) reasoning are both needed. For example, we may want a field to be an overridable monotonic function. We cannot split into two fields
+Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming. There are cases where overriding and
+(definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
 \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
@@
 To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
-In a sealed family, every field will expose its definition inside the family but a family as a whole will be sealed into a relaxed type.
-Thus we can override one sealed family as a whole and thus change the internal implementation.
+In a sealed family, every field will expose its definition inside the family but a family as a whole will be sealed into a ``relaxed'' type.
+Thus we can override one sealed family as a whole and change the internal implementation.
 
@@
 families.
-Overriding field will always have these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
+Any overriding family will always have these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
 }
@@
 \mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
-and we have assigned it to both "f" and "p", and later instantiate in
+and we assigned it to both "f" and "p", and later instantiate in
 "I2" with concrete implementation by overriding this family
@@
 We start with the programming interface.
-We use "FInductive" to define a extensible inductive type and in the children family we use "Extend Finductive ..." to extend the inductive type with new constructors.
+We use "FInductive" to define an extensible inductive type and in the children family we use "Extend Finductive ..." to extend the type with new constructors.
 In \cref{fig:plugin-example2}, we make the example of boolean and three-valued boolean.
@@
 Recall, to construct a recursor, we need to create for each
-constructor a handler (function) and aggregate them into a family, and thus that family (e.g.
-"neg_handler" in \cref{fig:plugin-example2}) can be considered as a large matching-case statement.
+constructor a handler (function) and aggregate them into a family (e.g.
+"neg_handler" in \cref{fig:plugin-example2}).
 We use "FRecursor" to declare a recursor (e.g.
@@
 "neg_handler.uu").
-If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".
+If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".\EDJ{This paragraph is duplicating but I am not sure how to make it concise.}
 
@@
 I think I use ``export'' in several places...}
-the eliminator of the inductive type "b" inside "b_3" because doing so all of the following fields (e.g.
+the eliminator of the inductive type "b" to "b_3" because doing so all of the following fields (e.g.
 "neg_handler.ff" and the corresponding "ff_24") will not be able to be inherited to the context where "b"
-is extended with a third constructor, \textbf{since the existence of recursor
-alone will assert that there are only two constructors} in "b", not
+is extended with a third constructor, \textbf{since the mere existence of the recursor will assert that there are only two constructors} in "b", not
 \textbf{extensible} at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}\EDJreply{You are right.
@@
 
-However, we later still need the concrete eliminator "b" for doing type-checking (as in \ref{chg:extensible-inductive-type}), thus the information doesn't totally disappear but store elsewhere.
-Here we show one exemplar extraction of the recursor to \mintinline{Coq}{Set} as module "b_rec_12".
-This concrete recursor will be used by our "FRecursor".
+However, we later still need the concrete eliminator "b" for doing exhaustiveness checking (as in \ref{chg:extensible-inductive-type}), thus we will store the corresponding info somewhere.
+Here we show one exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module "b_rec_12".
+This concrete eliminator will be used by our "FRecursor".
 For example, in \cref{fig:plugin-example2}, "FRecursor" command will feed Coq a module "v_33_34" for doing type-checking with specified handlers (e.g.
@@
 Importantly, the recursor field (e.g.
-"neg") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) module, a recursor field is not compiled like that.
-For all the children family, we have to reconstruct and override with a ``new'' recursor using the (maybe inherited) handler family.
-This is affordable since each recursor handler doesn't need
+"neg") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) modules, a recursor field is not compiled like that.
+For all the children family, we have to reconstruct and override the "FRecursor" field with a ``new'' recursor using (maybe the same) handler family.
+This is affordable since each aggregated recursor handler doesn't need
 to be rechecked---look at "B2.neg_handler", there is only one new
@@
 but what the plugin actually does is doing a second exhaustiveness
-checking in children family like mentioned above, with the ``same'' motive and ``same'' handler family.\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
+checking in children family like mentioned above, with the ``same'' motive and the ``same'' handler family.\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
 Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}\EDJreply{Please check.}
@@
 % explain ftheorem wrapping this complicated recursor
-This design of decoupling of the case handlers and exhaustiveness checking handles the semantic well, however, it brings a lot of boilerplate code---we need to create an extra (handler) family, manually specifying the type of each case handlers, and then using "FRecursor" to ``tie the knot'' and construct a recursor field from the family.
-We thus provide a command "FTheorem" to remove these boilerplate in programming interface when using tactic programming instead. 
+This design of decoupling of the case handlers and exhaustiveness checking handles the semantic well, however, it brings a lot of boilerplate code---we need to create an extra (handler) family, manually specifying the type of each case handlers, and then using "FRecursor" to ``tie the knot'' and construct a recursor field from the handlers.
+We thus provide a command "FTheorem" to avoid these boilerplate in programming interface when using tactic programming. 
 This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
@@
 meta-theory need to handle these two seemingly contradicting ideas
-simultaneously.
-(2) \ref{chg:extensible-inductive-type} is resolved by this \textit{decoupling} of the syntax of implementation of recursor handlers and exhaustiveness checking.
-The former is handled by family inheritance and thus avoid the boilerplate code; and the non-inheritance of the compiled recursor ensures the exhaustiveness checking happens for every recursor.
+simultaneously when solving \ref{chg:extensible-inductive-type} .
+(2) \ref{chg:extensible-inductive-type} is also resolved by this \textit{decoupling} of the syntax of implementation of recursor handlers and exhaustiveness checking.
+The former is handled by family inheritance and thus avoiding the boilerplate code; and the non-inheritance of the compiled recursor ensures the exhaustiveness checking happens for every recursor.
 
@@
 newly added value forms are not "tm_app". What's worse, similar
-scenario can happen on other statement like \mintinline{Coq}{~ value
+scenario can happen on other statements like \mintinline{Coq}{~ value
 (tm_if cond x y)}. Everything here is extended correctly but we end up
@@
 I am not sure this kind of hijacking can happen.}
-This can fulfill \ref{chg:software-engineering}.
+This can fulfill \ref{chg:software-engineering} and define new customized tactic.
 
@@
 
-\textbf{Propositional Partial Recursor.} Notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''--they will possibly disrupt arbitrary extension of inductive type.
+\textbf{Propositional Partial Recursor.} Notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''---they will possibly disrupt arbitrary extension of inductive type.
 For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)} in the "Family STLC" either using "FRecursor" or "Closing", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
@@
 \end{minted}
-To generalize to indexed inductive type, we base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}. Please refer to
+% To generalize to indexed inductive type, we base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}. Please refer to

### 2022-07-02 lang-design.tex
Going through before metatheory for small-fixes.

@@
 
-We want our metatheory mechanization to allow adding new function/proving new lemmas, and adding new constructors for inductive type at the same time  (Expression Problem~\cite{wadler-ep}).
-Coq support the former but not the later---because when an existent inductive type is extended with new constructor, we have to modify the pattern matching in the existent functions.
-This motivates us to add inheritance/Family Polymorphism into Coq.
+We want the metatheory mechanization to allow adding new function/proving new lemmas, and adding new constructors for inductive type at the same time  (Expression Problem~\cite{wadler-ep}).
+Coq support the former but not the later---when an existent inductive type is extended with new constructor, we have to modify the pattern matching in the existent functions.
+This motivates us to add inheritance/Family Polymorphism into Coq as a solution to the Expression Problem.
 
@@
 
+Refering to the syntax of \citet{zm2017}, we will come up the above example.
 However, to add inheritance into a dependent type system like Coq is not simple and there are various challenges and tension we need to resolve.
@@
 Unlike most general purposed programming languages, Coq, as a dependent type theory, is a total language and thus each pattern matching cases must fully ``cover'' an inductive type.
-This process is \textit{dynamic} for inheritance---when inhertiance extend the inductive type in the children family but the user forget to extend the corresponding inductive reasoning, our compiler need to signal an error correspondingly, or better---guide the user to fill the incomplete inductive reasoning.
+This process is ``\textit{interactive}'' for inheritance---when inhertiance extend the inductive type in the children family but the user forgets to extend the corresponding inductive reasoning, our compiler need to signal an error correspondingly, or better---guide the user to fill the incomplete inductive reasoning.
 For example, in \cref{fig:STLC-example} when "term" is extended with "tt" in the children family "STLC_bool", our compiler need to make sure "subst" and "subst_lemma" are also extended correctly otherwise signaling an error.
-Additionally, our compiler must support direct inheritance (and overriding) of the old case handlers, to waive the boilerplate code.
+Additionally, our compiler must support direct inheritance (and overriding) of the old case handlers for "var", "lam" and etc, to avoid the boilerplate code.
 
@@
 In family polymorphism for general purposed programming language, overriding is allowed to happen anywhere.
-But the problem becomes much more subtle in the case of dependent type---the later fields can have a type dependent on the former fields!
+But the problem becomes more subtle in the case of dependent type since the later fields can have a type dependent on the former fields.
 Consider a small example where we have consecutive fields \mintinline{Coq}{{..
@@
 
-These two examples both show the incompatibility between exposing the concrete definition of a field for reasoning (this second example even involves the definition of pattern matching) and overridability.
+These two examples both show the incompatibility between exposing the concrete definition of a field for reasoning (this second example even involves the definition of fixpoint and pattern matching) and overridability.
 We need to be careful about this tension when designing the language.
@@
 Self Parametrization}.
-The metatheoretical formulation in~\citet{zm2017} for family polymorphism is very natural as they consider extension of the family as the concatenation of a new linkage.
+The metatheoretical formulation in~\citet{zm2017} for family polymorphism is very natural as they consider an extension of the family as the concatenation of a new linkage.
 However, this formulation is not directly adaptable for dependent type theory because it can easily encode non-termination sabotaging consistency by the presence of "self".
 Informally, in the orignal formulation, we can override the "f" in family $\{"f" := id; "g" := "call self.f"\}$ with "call self.g", resulting the mutual calling of "f" and "g" thus a deadloop.
-This conflict between self parametrization and consistency will be another challenge---we need change the original formulation of family and inheritance for dependent type theory.
+This conflict between self parametrization and consistency requests changing the original formulation of family and inheritance for dependent type theory.
 
@@
 Not only for the metatheory, we also want our implementation of family to have good compatibility with the rest of Coq.
-For example, we cannot compromise \textbf{tactic programming} as it is already a widely accepted part of Coq programming.
-Additionally, the interactive development style requires \textbf{incremental type checking}---our compiler has to be able to type check each Coq Vernacular command.
+For example, we shouldn't compromise \textbf{tactic programming} as it is part of the idiomatic Coq programming.
+Additionally, the interactive development style requires \textbf{incremental type checking}---our design has to be able to type check each Coq Vernacular command once inputted.
 What's more, we want our family ``term'' to incorporate Coq \textbf{without breaking its other functionality}.
-For example, the user may want to project the family member anywhere inside an arbitrary Gallina term, that is later being computed and normalized.
-However, families themselves can be second class citizens just like \citet{zm2017}.
-
-\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).
+For example, the user may want to project the family member anywhere inside an arbitrary Gallina term, that may later be computed and normalized.
+However, families themselves can be second class citizens just like \citet{zm2017}.\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).
 I think most readers will benefit from a recap of this encoding.\\

### 2022-07-03 canonicitymodel.tex
Small fixes throughout metatheory section

@@
 Thus once we can ``propagate'' $(\cdot)^C$ into every syntactic piece, and construct this model, we are done.
-In our (fake) Agda formulation, we use $(\cdot)_2$ to replace the notation of $(\cdot)^C$ because we consider most parts of this model are actually a dependent pair.
+In our (pseudo) Agda formulation, we use $(\cdot)_2$ to replace the notation of $(\cdot)^C$ because we consider most parts of this model are actually a dependent pair.
 
 Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
-The latter is more conventional and the former can be understood as type theoretic encoding of the latter\footnote{We usually use predicate to encode "belong to" relation in type theory}.
-From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out~\cite{skorstengaard2019introduction}---using the terminology of the conventional formulation, $\Gamma_p$ will be the set of substitution into closed reducible terms, and we will prove $t[\gamma] \in T_p$, a set of closed reducible term of type $T$ when $\gamma \in \Gamma_p$.
+The latter is more conventional and the former can be understood as type theoretic encoding of the latter\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory}.
+From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out~\cite{skorstengaard2019introduction}---using the terminology of the conventional formulation, $\Gamma_p$ will witness the set of substitution into closed reducible terms of types $\Gamma$.
+Then we will prove $t[\gamma] \in T_p$---a set of closed reducible term of type $T$---when $\gamma \in \Gamma_p$.
 

### 2022-07-03 coqimpl.tex
Small fixes throughout metatheory section

@@
 \mintinline{Coq}{STLC.tt ≠ STLC.ff}) must be conservative because it holds for all extension of the inductive type "STLC.tm".
-But is there a ``best'' proposition (i.e.
+But is there a \textbf{``best'' proposition} (i.e.
 the proposition that everything the proposition derives are exactly conservative extension)?
@@
 
-Even though we cannot prove propositional partial recursor is the best,
+Even though we cannot prove propositional partial recursor is \textit{the best},
 the above "reflect" function implies that propositional partial
-recursor is a good \textit{extensional characterization} of (non-indexed) extensible inductive type (with certain constraints): because we can ``embed'' vanilla inductive types into
+recursor is a good \textit{extensional characterization} of (non-indexed) extensible inductive type (with certain constraints)---because we can ``embed'' vanilla inductive types into
 types supporting a partial recursor: "reflect" is actually a left

### 2022-07-03 metatheory.tex
Small fixes throughout metatheory section

@@
 Our work can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}.
-We base our formulation of the syntax in this type theory instead of a more conventional and informal representation, becase (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out.
+We base our formulation of the syntax in this type theory instead of a more conventional and informal representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out.
 

### 2022-07-03 standardmodel.tex
Small fixes throughout metatheory section

@@
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1.
-we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2.
-we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma.
-We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e.
+We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.}
+
+The main idea of the proof is that: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma.
+We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo) Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e.
 the judgementally-equal syntax needs to be mapped to the same data for well-definedness).
 
-Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gamma}{t}{T}}$ (i.e. every term is tracking its context and type)
+The main intuition of the constructed consistency model is that: (1) since we have used sigma type to translate the linkage, using the model for sigma type for linkage will work; (2) directly using the inductive facility of (pseudo)-Agda to justify the intro and elim rule of our W-type.
+
+Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gamma}{t}{T}}$ (i.e.
+every term is tracking its context and type).\footnote{And our (pseudo) Agda formulation is well-typed as well.
+Latex-formulation should be well-typed but we discarded all the type information in the paper.}
 
@@
 
-Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
+Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
 
@@
 
-Since all the compatibility lemmas hold (by simply changing $\vdash$ into $\models$ in the syntactic rules then we would have all the compatibility lemmas\footnote{Apparently, in our (fake) Agda formulation, we don't need this extra step.}), we have constructed a model.
+Since all the compatibility lemmas hold (by simply changing $\vdash$ into $\models$ in the syntactic rules then we would have all the compatibility lemmas\footnote{Apparently, in our (pseudo) Agda formulation, we don't need this extra step.}), we have constructed a model.
 Finally, notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
@@
 
-The full model formulated in (fake) Agda syntax can be found in the appendix.
+The full model formulated in (pseudo) Agda syntax can be found in the appendix.

### 2022-07-03 syn-translate.tex
Small fixes throughout metatheory section

@@
 The implementation described in \cref{sec:coqimpl} can be formalized in our metatheory.
-The described compilation strategy can be considered as a \textit{syntactical translation}, a function mapping from our syntactical (QIIT-)model to a subpart of our syntactical (QIIT-)model \textit{without linkages}.\footnote{Here, syntactical QIIT-model can be roughly understood as just AST, with equational rules acting like quotient.
-So this mapping is mapping AST to AST, respecting quotient/judgemental equality.} This subpart is the standard MLTT with $\goodWSig{}{}{}$, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
+The described compilation strategy can be considered as a \textit{syntactical translation}, a function mapping from our syntactical (QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only difference lies in the "Wrec", in our metatheory, it is constructed using linkage; in ``subpart'', it will be using sigma types (translation of the linkage).} of our syntactical (QIIT-)model \textit{without linkages}.\footnote{Here, syntactical QIIT-model can be roughly understood as just AST, with equational rules acting like quotient.
+So this mapping is mapping AST to AST, respecting quotient/judgemental equality.} This ``subpart'' is the standard MLTT with inductive type signature, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
 
@@
 
-Ultimately, we have constructed a model for our QIIT-syntax using only subparts of our QIIT-syntax -- without these linkages.
-This also justify the intuition -- a linkage is actually a module \textit{with universal quantification wrapping its fields to achieve late binding}.
+Ultimately, we have constructed a model for our QIIT-syntax using only a ``subpart'' of our QIIT-syntax -- without these linkages.
+It is also possible to compile our $\goodWSig{\Gamma}{w}{n}$ and W-type into a more conventional formulation, but the recursor will always require a sigma type as handler and thus the elimination rule for our W-type is still non-conventional.
 
-It is also possible to compile away $\goodWSig{\Gamma}{w}{n}$ but the recursor will always require a sigma type as handler and thus the elimination rule for our W-type is still non-conventional.
+This translation justifies the intuition -- a linkage is actually a module \textbf{with universal quantification wrapping its fields to achieve late binding}.
 
-After syntactical translation, we can directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the subpart}.
+After syntactical translation, we can directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the ``subpart''}.
 However, due to our non-conventional formulation of the inductive type, it is better we don't impose such assumption and directly give a consistency and canonicity proof.

### 2022-07-03 syntax.tex
Small fixes throughout metatheory section

@@
 mutually recursively defined together. Our presentation
-follows the style of~\citet{kaposi2019gluing}, but for readability we
-omit the universe levels. We still working on predicative MLTT.
+follows the style of~\citet{kaposi2019gluing} and thus a predicative MLTT, but for readability we
+omit the universe levels.
 %\YZ{What's the distinction of judgments vs.
@@
 Here I want to say "even underivable judgement has to be well-typed".
-That's the main reason I spell out the fake Agda code there.
+That's the main reason I spell out the (pseudo) Agda code there.
 I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}\EDJreply{I rephrase that so there won't be confusion for the future readers.
@@
 course, a well-formed judgement is not necessarily derivable, just as
-a well-formed QIIT type is not necessarily inhabited.
+a well-formed QIIT type is not necessarily inhabited. These well-formedness rules are direct translations of the \textit{well-typedness} of QIIT.
 
-Due to our intrinsic setting, \textbf{we can and will omit a lot of
-presumptions of the typing rules}. This can work because we are using
-QIIT and everything is intrinsically well-formed. For example, we don't
+Due to our intrinsic setting, \textbf{we can and will omit a lot of premises of the typing rules}. This can work because we are using
+QIIT and everything is intrinsically well-formed(well-typed). For example, we don't
 spell out that the rule \ruleref{Type Universe} requires a well-formed
 context, even though it does implicitly, \textbf{because it is not
-possible to have ill-formed contexts in our intrinsic setting}. This is
+possible to have ill-formed(ill-typed) contexts in our intrinsic setting}. This is
 a bit unusual for conventional extrinsic setting where they will put
@@
 consider this as one advantage of using intrinsic setting---this
-well-formed context is usually attached to some other presumptions thus
+``well-formed context'' condition is usually attached to some other premises thus
 inferrable and can be considered as ``boilerplate code''. Free of
@@
 also called substitution calculus in the literature and is favoured due
-to its closedness categorical semantic.
+to its closedness to categorical semantic.
 (3) We are dealing with quotient data, so we are actually manipulating
-equivalence classes of terms.
+equivalence classes of terms instead of one representative (term) from the equivalence class.
 
@@
 first-class citizens, and can be applied (almost) everywhere and can return
-(almost) everything, thus we require the resulting substitution to
-commute with most (if not all) type/term constructs.
+(almost) everything, thus we require \textbf{the resulting substitution to
+commute} with most (if not all) type/term constructs.
 
@@
 series of terms with a new term (the weird looking type of
-$\goodTerm{\Gamma}{t}{A[\sigma]}$ is because we need $A$ to be typed in
+$\goodTerm{\Gamma}{t}{A[\sigma]}$ in the premise is due to our need of type $A$ in
 the context $\Delta$ so that we can use \ruleref{Context Extension} to
@@
 What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda.
-For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer fake agda formulation) }\YZreply{I was thinking about 'Delta |- A'.
+For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer (pseudo) agda formulation) }\YZreply{I was thinking about 'Delta |- A'.
 I suppose it is also implicitly enforced via intrinsically typed syntax.
@@
 We have shown some exemplar rules for all four standard types of
-judgments. Now we focus on the newly introduced facility.
-
-We have two kinds of signatures, one for inductive types, the other for linkages.
+judgments. Now we focus on the newly introduced facility. We have two kinds of signatures, one for inductive types, the other for linkages.
 
@@
 (x : Aᵢ) → (Bᵢ x → W) → W}.
-Thus, we use a list of pairs ($A_i, B_i$) as the signature of one
-inductive type, and given an inductive type signature, we use
+Thus, we use a `list' of pairs ($\goodWSig{\Gamma}{w^+~A_i~B_i~\tau}{}$) as the signature of one
+inductive type. Given an inductive type signature, we can use
 \ruleref{WSig-Proj} to extract the corresponding $A_i, B_i$.
@@
 of the syntax. Here $\bW \sigma$ is a large singleton
-type~\cite{stone2000}, and it only has one ``element'', the inductive
+type~\cite{stone2000}, and it only has one ``element''---the inductive
 type itself, inside. We formulate in this cumbersome way because we want
 to have inductive type to be a field element and exposing the
-constructors at the type level for later \textit{pattern-matching
-exhaustiveness checking} (i.e. if an inductive type as field element has type
+constructors at the type level for later \textit{exhaustiveness checking} (Since otherwise, if an inductive type as field element has type
 $\cU$, then no information about the inductive type is exposed).
@@
 empty stuff, and respect substitution lemmas. We have
-\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component
-due to harmony~\cite{pfenning2009lecture}, so we have local soundness
-and local completeness ($\beta,\eta$ rules). We omit them and the
+\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component.
+We do have harmony~\cite{pfenning2009lecture} (i.e. local soundness
+and local completeness, or $\beta,\eta$ rules). We omit them and the
 substitution laws for either field addition and projection. For
-$\mu^+,\nu^+$ rules, we will omit the sealing $\{f\}$ if possible.\footnote{This
-omission can be supported by Agda using Implicit Arguments}
+$\mu^+,\nu^+$ rules, we will omit the sealing $\{f\}$ if possible.
+% \footnote{This omission can be supported
+%  by Agda using Implicit Arguments}
 
-A side note: before we dive in more details of our metatheory for linkage (especially the meaning of sealing and $\cC$), we have to mention that, the challenge of incorporating family polymorphism with inductive type mentioned in \ref{chg:extensible-inductive-type} mutates in metatheory because of de Bruijn indices.
-(Using surface synax), consider the example of a linkage $\{a~:~\bW \sigma; b_1 : a \to T;..\}$ with a bunch of $b_i$ \textit{using} $a$.
+\textit{A side note}: before we dive in more details of our metatheory for linkage (especially the meaning of sealing and $\cC$), we have to mention that, the challenge of incorporating family polymorphism with inductive type mentioned in \ref{chg:extensible-inductive-type} mutates in metatheory because of de Bruijn indices.
+(Using named variables), consider the example of a linkage $\{a~:~\bW \sigma; b_1 : a \to T;b_2 ..\}$ with a bunch of $b_i$ \textit{using} $a$.
 The tension we need to solve here is between \textit{exhaustiveness checking} and \textit{inheritance (code reuse)}---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to construct $b_i$ term in the context of $a : \bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that and thus we hope $b_i$ constructed in the context of $a : \cU$ (and thus later this $a$ can be compatible with inductive type of any signature).
@@
 $\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, according to \ruleref{Compile}, is a deeply nested dependent pair.
-Then $\cCt \sigma$ provide the concrete module of type $\cC\sigma$ given a linkage of type $\cL\sigma$.
+$\cCt \sigma$ provide the concrete module of type $\cC\sigma$ given a linkage of type $\cL\sigma$.
 
@@
 to support \textit{code reuse}, as the rule of linkages hinders its
-ability to do abstraction. Concretely speaking, (using surface syntax)
+ability to do abstraction. Concretely speaking, (using named variables)
 we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to
@@
 This is also reflected in the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}.
-Thus we need $\cC\sigma$ in our context.
 %\YZ{Why is Csigma needed?
@@
 For notation clarity, we will sometimes use $"sf'"$ to emphasize
-that we consider that seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term $\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]} $ and "sf" notation (used in \ruleref{Compile}) is defined as $\goodSub{\Gamma, \cC \sigma}{sf~s := (\pi_1, sf'~s)}{\Gamma, A}$.\YZ{What is sf and sf'?
+that we consider that seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term $\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]} $ and "sf" notation (used in \ruleref{Compile}) is defined as \\ $\goodSub{\Gamma, \cC \sigma}{(sf~s) := (\pi_1, sf'~s)}{\Gamma, A}$.\YZ{What is sf and sf'?
 sf first appears in Compile rules, so explain sf there.}\EDJreply{Fixed.
@@
 $t_{new}$ in the context containing $(\Sigma a : \cU,..)$, ready for
-inheritance.
-Apparently $t_{new}$ in this case cannot be doing \textit{exhaustiveness
-checking}, but it can freely use any \textit{exhaustiveness-checked}
-recursor.
+inheritance. Thus sealing itself can be considered as proving 
+abstractions on sigma type.
+
 %\YZ{It occurs to me that you use the word "sealing" to mean
@@
 The ``concatenation'' of the $o$ and $t$ leads to a new linkage, of which the signature is the ``concatenation'' of the $\sigma$ and $T$ (we also record the sealing $f$ for simplicity).
+The key here is our signature, will \textit{record} this abstraction witness $f$, thus any projected member will still be in the abstracted context and thus possibly \textit{inheritable}.
 The construction of linkage in \ruleref{Lnkg Add} requires $t$ to \textbf{only} refer to the (abstraction $A$ of) former part of the defining family $o : \cL\sigma$, this is to avoid inconsistency mentioned in \ref{chg:consistency} in metatheory.
@@
 In summary, with the help of both sealing and $\cC$, we can shift the context from $\cL(\{a~:~\bW \sigma; ..\})$ to $\Sigma a : \cU,..$ and thus achieve code reuse.
-However, this is just allow possible \textit{code reuse}, we still need to solve the mutated tension in \ref{chg:extensible-inductive-type}.
+However, this is just allow \textit{possible} {code reuse}, we still need to solve the mutated tension in \ref{chg:extensible-inductive-type}.
 To make things concrete, we need to introduce the rules for recursors.
@@
 }
-Each recursor is constructed by a linkage containing
-functions/handlers/pattern-matching cases dealing with each
-constructor. Thus code reuse for recursor is delegated to linkage
-inheritance---to reuse one particular pattern-matching case,
-we just inherit the function/handler. This greatly simplifies the
-metatheory development.
+Like our plugin implementation, each recursor is constructed by a linkage containing handlers dealing with each
+constructor and code reuse for recursor is delegated to linkage
+inheritance. This greatly simplifies the metatheory development.
 
@@
 $\beta$-rule of "Wrec" (omitted here). Finally \ruleref{Rec-Constr} is
-used to construct a concrete recursor given a specific linkage of
-handlers. All these rules satisfy substitution laws and omitted here.
+used to construct a concrete recursor given a specific linkage of aggregated
+handlers. All these rules satisfy substitution laws and omitted here. Thus, we are still using \textit{decoupling} of the syntax of the case
+handlers and exhaustiveness checking to resolve
+\ref{chg:extensible-inductive-type}.
 
-Again, we still use \textit{decoupling} of the syntax of the case
-handlers and exhaustiveness checking to resolve
-\ref{chg:extensible-inductive-type}. We use linkages to define
-case handlers and thus they are inheritable like other families.
-While any field using "Wsup" and "Wrec" directly relies on the
-$\bW\sigma$ (can be spotted from \ruleref{Ind-Term} and
-\ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup"
+Then any field using "Wsup" and "Wrec", directly relying on the
+$\bW\sigma$ in the context (can be spotted from \ruleref{Ind-Term} and
+\ruleref{Rec-Constr}), and responsible for exhaustiveness checking will be non-inheritable---in these cases, "Wsup"
 can be understood as concrete constructors and "Wrec" are
@@
 problem on inheritance. 
+For example, $t_{new}$ above cannot be doing \textit{exhaustiveness
+checking}, but it can freely use any \textit{exhaustiveness-checked}
+recursor, and also inheritable.
 Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden.
 For "Wsup" we only need to override each constructor and for "Wrec" it is just doing exhaustiveness checking inside each children family.
-The recursor handlers however, can be safely inherited.
+The recursor handlers however, like plugin implementation, can be safely inherited.
 \YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
@@
 Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, and in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e.
-we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}$, $\goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption of judgements when needed.
+we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}$, $\goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the premises of judgements when needed.
 
@@
 $$
+\Rule[name=Inh]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+\quad \goodTerm{\Gamma}{l}{\cL \sigma_1}
+}
+{\goodTerm{\Gamma}{("inh" \ h \ l)}{\cL \sigma_2}} 
+\quad
 \Rule[name=Inh+Inh]
@@
 \quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}
-\quad 
+\\\\
 \goodInh{\Gamma, A_2}{i}{\tau_1[(\pi_1, \uparrow^s)]}{\tau_2}}
@@
 $$
-\Rule[name=Inh]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodTerm{\Gamma}{l}{\cL \sigma_1}
-}
-{\goodTerm{\Gamma}{("inh" \ h \ l)}{\cL \sigma_2}} 
-\quad 
 \Rule[name=Inh-Inh-beta]
@@
   \quad \goodTerm{\Gamma, A_1}{t}{T}
+  \quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}
 }
-{\goodTerm{\Gamma}{"inh" ("inhinh" \ h) (\mu^+ \ m \ t) \equiv \mu^+ \ ("inh" \ h \ m) \ t[(\pi_1, \uparrow^s)]}{}} 
+{\goodTerm{\Gamma}{"inh" ("inhinh" \ h) (\mu^+ \ m \  \{f_1\} \ t) \newline  \equiv \mu^+ \ ("inh" \ h \ m) \  \{f_2\} \ t[(\pi_1, \uparrow^s)]}{...}} 
 $$
@@
 defined in the surface syntax---a Family with no parent is an
-inheritance with empty input; an extended Family is exactly an
-inheritance---because the parent of an extended Family can be overridden
-thus inheritance should only define upon the ``interface'' instead of a
-concrete family. However, since our current formulation doesn't support
-nested inheritance (i.e. we cannot do inheritance inside a family but only top level), our
-formulation still has some distance to the real family polymorphism.
+inheritance with empty input; an extended Family can be considered as an
+inheritance.
+% ---because the parent of an extended Family can be overridden
+% thus inheritance should only define upon the ``interface'' instead of a
+% concrete family. However, since our current formulation doesn't support
+% nested inheritance (i.e. we cannot do inheritance inside a family but only top level), our
+% formulation still has some distance to the real family polymorphism.
 
@@
 Override "inhov" is even more special because in mundane OO programming language, we usually require that the overriding term has the same type/interface as the overriden one.
-Here we don't since the very reason we want the same type/interface is that we hope other inherited fields can use this overridden/late-binding field without breaking abstraction, and this reason is already managed by the assumption of "inhinh", the upcasting proof $\uparrow^s$.
+Here we don't since the very reason we want the same type/interface is that we hope other inherited fields can use this overridden/late-binding field without breaking abstraction, and this reason is already managed by the premises of "inhinh", the upcasting proof $\uparrow^s$.
 Thus "inhov" is quite like "inhext", where overriding is just throwing away one parent field.
@@
 
-Finally we need to introduce nested inheritance "inhnest" for dealing with nested families.
-The rule is mostly direct and we need "upcasting" again when dealing with nested family in different but inheriting context.
+Next, we introduce nested inheritance "inhnest" for dealing with nested families.
+The rule is mostly direct and we need ``upcasting'' again when dealing with nested family in different but inheriting context.
 We also show how to use an inheritance judgement to derive children family and how the beta rule is defined.
@@
 
+Finally, we show \ruleref{Inh-Inh-beta} to again emphasize the importance of sealing and its interaction with inheritance.
+$\goodTerm{\Gamma, A_1}{t}{T}$ is again well-typed in the new abstract context $\goodCtx{\Gamma, A_2}{}$, with the help of upcasting $\uparrow^s$ which is showing compatibility between abstraction.
+To make sure we have the full information of $t$ including the context $\Gamma, A_1$, we need to have a good sealing $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}$.
+
 \subsection{Exemplar Usage of Meta Theory}

### 2022-07-03 coq-example.tex
Small fixes through out.

@@
 to mix the semantics and properties of two language features---product
-and boolean (in \cref{fig:STLC-example})---with vanilla STLC. This way,
-we can say precisely, \textit{a programming language feature itself is a
+and boolean (in \cref{fig:STLC-example})---with vanilla STLC.
+
+This \textit{mixin} is implemented roughly by making inheritance judgement/data supporting \textit{weakening} and composition.
+Then, using notation from metatheory, given $\goodInh{\cdot}{i_1}{\sigma_1}{\sigma_2}$ and $\goodInh{\cdot}{i_2}{\sigma_1}{\sigma_3}$, we can weaken $i_2$ to $\goodInh{\cdot}{i_2'}{\sigma_2}{\sigma_3'}$ and the composition of $i_1$ and $i_2'$ is the resulting mixin---this implementation breaks commutativity and will need to solve (design) some compatibility solution: for example, when $i_1$ and $i_2$ are both extending with field of the same name, do we signal error or use $i_2$'s extension as considering it as an `override'?
+However, our current example of STLC only uses a subset of the mixin that is with a clear semantic (i.e.
+every sensible design of mixin should agree on).
+
+Using this example,
+we hope one day we can say precisely, \textit{a programming language feature itself is a
 piece of data/inheritance judgement/inherited family}. Compared with MTC,
@@
 
-Most of the proof and computation is directly adapted
+Most of the proof and computation are directly adapted
 from the one in Software Foundation, resulting in a similarly accessible proof, while the proof for each extended feature are scattered in the extended families resulting better modularity.
-One kind of examples is about the computation, and here we mainly focus on substitution functions.
+One kind of examples is about the computation, for example, substitution functions.
 Substitutions dealing with product and projection are defined in the corresponding children family.
@@
 The other kind of examples are about proof-irrelevant propositions and we directly use "(Extend) FTheorem" to handle them.
-In these cases we don't even need to additionally handle a auxillary handler family like we did for substitution---we simply use "Extend FTheorem" and our plugin will prepare the goal to prove.
+In these cases we don't need to additionally define a auxillary handler family like we did for substitution---we simply use "Extend FTheorem" and our plugin will prepare the goal to prove.
 At the end, it will look like scattered proof scripts in a modular manner.
@@
 that our inductive definition does satisfy the constraint, which is
-exactly what we expect from "Closing Fact".
+exactly what we expect from "Closing Fact"---if directly working on inductive definition doesn't bring us benefit, then decompose the property (the constraint) with the concrete definition might be a good idea.
 
@@
 verified computational rules. This part is possible to be automatically
-handled by the plugin, but still it will require propositional
+generated by the plugin, but still it will require propositional
 computational rules.
@@
 Then we extend the base language with new features, and we instantiate
-the postulation on computation for both concrete interpreters and
+the postulation on computation for
 abstract interpreters.
-Thanks to the compilation to Coq, we can directly run the resulting
+Thanks to the compilation to Coq module, we can directly run the resulting
 abstract interpreters.
+
+\begin{figure}[!htb]
+  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
+  \caption{Inheritance Diagram for Abstract Interpretation Example.
+Black arrows are inheritance.
+Purple arrow are extending.
+Green arrows are overriding.
+Grey box are sealed families.}\label{fig:abstract-interpretation-example}
+\end{figure}
 
@@
 Please check}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
-  \caption{Inheritance Diagram for Abstract Interpretation Example.
-Black arrows are inheritance.
-Purple arrow are extending.
-Green arrows are overriding.
-Grey box are sealed families.}\label{fig:abstract-interpretation-example}
-\end{figure}

### 2022-07-03 related-work.tex
Small fixes through out.

@@
 
-\textbf{Modular Mechanized Metatheory: Syntactic Approach} We group the work that achieves modular mechanized metatheory by extending surface syntax of proof assistants as \textit{syntactic approach}.
+\textbf{Modular Mechanized Metatheory: Linguistic Approach} We group the work that achieves modular mechanized metatheory by extending surface syntax of proof assistants to provide better abstraction as \textit{linguistic approach}.
 
-The closest work to ours might be \citet{boite2004proo}--they try to support \textit{extended} inductive type and related proof reuse in Coq.
-This biggest difference lies in our (incomplete) adaption of Family Polymorphism -- apart from supporting proof reuse and extension by inheritance, we also support implementation swapping using overriding, and we also argue that Family Polymorphism can provide a better and cleaner way of organizing mechanized metatheory development.
+The closest work to ours might be \citet{boite2004proof}--they try to support \textit{extended} inductive type and related proof reuse in Coq.
+This biggest difference lies in our (incomplete) adaption of Family Polymorphism -- apart from supporting proof reuse and extension by inheritance, we also support implementation swapping using overriding, and we also argue that Family Polymorphism can provide a better and cleaner way of organizing mechanized metatheory development a well as solving related expression problem.
 The other difference lies in the handling of the extension of the inductive type.
@@
 In their work, without using Family to organized the code, the extended inductive type can be considered as a different inductive type than the original one, possibly leading to more expression problem~\cite{wadler-ep}.
+The situation is similar for \citet{levin2003tinkertype}---they aim for modular manipulation of metatheory but didn't consider expression problem.
+What's more, our implementation is based on Coq plugin and is thus potentially more friendly to Coq community.
 % Another similar work is proof weaving \citep{mulhern2006proof} but I don't see a evaluation section in that paper
 
-The other work
-
-\textbf{Modular Mechanized Metatheory: Semantic Approach} We group the work using certain algebraic encoding (e.g.
-Church Encoding, datatype-generic programming) without modify the surface syntax of the proof assistants as \textit{semantic approach}, that includes \citet{delaware2013,forsta2020,liwei2022,schwaab2013modular, keuchel2013generic} and so on.
-Their advantages are various: this technique is more general and thus can be applied across different (dependent-typed) proof assistants and this technique can be easily distributed like a library and thus doesn't need to alter the core of the proof assistants.
+\textbf{Modular Mechanized Metatheory: Semantic Approach} We group the work using the semantic as encoding (e.g.
+Church Encoding, datatype-generic programming) to achieve certain design patterns without modify the surface syntax of the proof assistants as \textit{semantic approach}, that includes \citet{delaware2013,forsta2020,liwei2022,schwaab2013modular, keuchel2013generic} and so on.
+Their advantages are various: their techniques are more general and thus can be applied across different (dependent-typed) proof assistants and their technique can be easily distributed like a library and thus doesn't need to alter the core of the proof assistants.
 The downsides are mainly about the threshold: they are not friendly for the junior Coq development and these indirect encoding of datatype might lead to worse readability and accessibility compared to the direct style like ours.

### 2022-07-03 coq-example.tex
Add code for Section 5.1

@@
 
+\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{lstlisting}[language=Coq,frame = single, escapeinside={@}{@}]
+Family STLC_prod.
+ Extend FInductive ty : Set :=
+  | ty_prod : ty -> ty -> ty.
+ Extend FInductive tm : Set :=
+  | tm_prod : tm -> tm -> tm  
+  | tm_pi1 : tm -> tm  
+  | tm_pi2 : tm -> tm.
+ Extend Family subst_internal.
+  Final Field tm_prod := ..
+  Final Field tm_pi1 := ..
+  Final Field tm_pi2 := ..
+ EndFamily.
+
+ Inherits subst.
+ ...
+ Extend FInductive has_type 
+  : context -> term -> type := ...
+ Extend FTheorem subst_lemma.
+  ... 
+ Qed.
+(* ... and more, end with type safety *)
+(* Next, Invoking Mixin Command *)
+EndFamily.
+Family STLC_bool_prod 
+    extends STLC 
+    using STLC_bool using STLC_prod.
+\end{lstlisting}
+\caption{Example STLC extended with product, and Mixin with \cref{fig:STLC-example2}}
+\end{wrapfigure}
+
 Our emphasis here is that, we singled out some language features from
@@
 the style of the examples in MTC~\cite{delaware2013,forsta2020}.
-And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature
+And we use a \textit{not-yet-formally-defined} \textbf{mixin} feature
 to mix the semantics and properties of two language features---product
@@
 
-This \textit{mixin} is implemented roughly by making inheritance judgement/data supporting \textit{weakening} and composition.
-Then, using notation from metatheory, given $\goodInh{\cdot}{i_1}{\sigma_1}{\sigma_2}$ and $\goodInh{\cdot}{i_2}{\sigma_1}{\sigma_3}$, we can weaken $i_2$ to $\goodInh{\cdot}{i_2'}{\sigma_2}{\sigma_3'}$ and the composition of $i_1$ and $i_2'$ is the resulting mixin---this implementation breaks commutativity and will need to solve (design) some compatibility solution: for example, when $i_1$ and $i_2$ are both extending with field of the same name, do we signal error or use $i_2$'s extension as considering it as an `override'?
-However, our current example of STLC only uses a subset of the mixin that is with a clear semantic (i.e.
-every sensible design of mixin should agree on).
+\textbf{Mixin} is implemented by making inheritance judgement/data supporting \textit{weakening} and composition.
+Then, using notation from metatheory, given $\goodInh{\cdot}{i_1}{\sigma_1}{\sigma_2}$ and $\goodInh{\cdot}{i_2}{\sigma_1}{\sigma_3}$, we can weaken $i_2$ to $\goodInh{\cdot}{i_2'}{\sigma_2}{\sigma_3'}$ and the composition of $i_1$ and $i_2'$ is the resulting mixin---this implementation breaks commutativity and will need to solve design problems on compatibility (mostly related to the \textit{diamond problem}): for example, when $i_1$ and $i_2$ are both extending with a field of the same name, do we signal error or use $i_2$'s extension and consider it as an `override'?
+However, our current example of STLC only uses a subset of the mixin with a clear semantic (i.e.
+every sensible design of mixin should agree on)---because our STLC examples only use extension of fields with different names and no overriding is used.
+
+Note that, every top level family can be considered as an inheritance because (1) our implementation is simplified by considering every family as an inheritance, and (2) we want to align with the complete family polymorphism advocated by \citet{zm2017} where the user can override the parent of the inheriting family (inside a family due to late-binding), and in these case the nested inheriting family is more a inheritance data than a ``standalone'' linkage.
+Thus we will store for every top level family in our plugin the corresponding inheritance data together with a ``reference'' to the parent family.
+Then our mixin command can single out the inheritance data from "STLC_bool" and "STLC_prod" and ``mixin'' the inheritance to reach "STLC_bool_prod".
 
@@
 This is mainly done by inheriting and extending the handler family from the parents.
-The other kind of examples are about proof-irrelevant propositions and we directly use "(Extend) FTheorem" to handle them.
-In these cases we don't need to additionally define a auxillary handler family like we did for substitution---we simply use "Extend FTheorem" and our plugin will prepare the goal to prove.
+The other kind of examples are about propositions and we directly use "(Extend) FTheorem" to handle them.
+In these cases we don't need to additionally define a auxillary handler family like we did for substitution---we simply use "Extend FTheorem" and our plugin will prepare the goals to prove.
 At the end, it will look like scattered proof scripts in a modular manner.
@@
 that our inductive definition does satisfy the constraint, which is
-exactly what we expect from "Closing Fact"---if directly working on inductive definition doesn't bring us benefit, then decompose the property (the constraint) with the concrete definition might be a good idea.
+exactly what we expect from "Closing Fact".
+If directly working on inductive definition doesn't bring us benefit during proof engineering, then decompose the property (the constraint) with the concrete definition might be a good idea.
 

### 2022-07-03 coq-example.tex
Add diagram for Sec 5.1

@@
 \subsection{Type Safety for STLC}\EDJ{We need a fuller view of the two examples. Especially show "accessing family components outside that family".}
-We start with the very basic type safety proof of STLC, largely adapted
-from Software Foundation~\cite{pierce2014software}, which is also one of
-the primal motivating example of this project.
 
@@
 \end{lstlisting}
+\includegraphics[width=0.5\textwidth]{coqexmaple/STLC-inheritances.pdf}
 \caption{Example STLC extended with product, and Mixin with \cref{fig:STLC-example2}}
 \end{wrapfigure}
+
+We start with the very basic type safety proof of STLC, largely adapted
+from Software Foundation~\cite{pierce2014software}, which is also one of
+the primal motivating example of this project.
 

### 2022-07-04 lang-design.tex
comment

@@
 
-We want the metatheory mechanization to allow adding new function/proving new lemmas, and adding new constructors for inductive type at the same time  (Expression Problem~\cite{wadler-ep}).
-Coq support the former but not the later---when an existent inductive type is extended with new constructor, we have to modify the pattern matching in the existent functions.
-This motivates us to add inheritance/Family Polymorphism into Coq as a solution to the Expression Problem.
+We want the metatheory mechanization to allow adding new
+function/proving new lemmas, and adding new constructors for inductive
+type at the same time  (Expression Problem~\cite{wadler-ep}). Coq
+support the former but not the later---when an existent inductive type
+is extended with new constructor, we have to modify the pattern matching
+in the existent functions. This motivates us to add inheritance/Family
+Polymorphism into Coq as a solution to the Expression Problem.
 
@@
 
-Refering to the syntax of \citet{zm2017}, we will come up the above example.
-However, to add inheritance into a dependent type system like Coq is not simple and there are various challenges and tension we need to resolve.
+Refering to the syntax of \citet{zm2017}, we will come up the above
+example. However, to add inheritance into a dependent type system like
+Coq is not simple and there are various challenges and tension we need
+to resolve.
 
-\textlabel{Challenge (1)}{chg:extensible-inductive-type}~\textbf{Extensible Inductive Type vs.
-Exhaustiveness of Inductive Reasoning}.
-Unlike most general purposed programming languages, Coq, as a dependent type theory, is a total language and thus each pattern matching cases must fully ``cover'' an inductive type.
-This process is ``\textit{interactive}'' for inheritance---when inhertiance extend the inductive type in the children family but the user forgets to extend the corresponding inductive reasoning, our compiler need to signal an error correspondingly, or better---guide the user to fill the incomplete inductive reasoning.
-For example, in \cref{fig:STLC-example} when "term" is extended with "tt" in the children family "STLC_bool", our compiler need to make sure "subst" and "subst_lemma" are also extended correctly otherwise signaling an error.
-Additionally, our compiler must support direct inheritance (and overriding) of the old case handlers for "var", "lam" and etc, to avoid the boilerplate code.
+\textlabel{Challenge (1)}{chg:extensible-inductive-type}~\textbf{Extensible Inductive Type vs. Exhaustiveness of Inductive Reasoning}.
+Unlike most general purposed programming languages, Coq, as a dependent
+type theory, is a total language and thus each pattern matching cases
+must fully ``cover'' an inductive type. This process is
+``\textit{interactive}'' for inheritance---when inhertiance extend the
+inductive type in the children family but the user forgets to extend the
+corresponding inductive reasoning, our compiler need to signal an error
+correspondingly, or better---guide the user to fill the incomplete
+inductive reasoning. For example, in \cref{fig:STLC-example} when "term"
+is extended with "tt" in the children family "STLC_bool", our compiler
+need to make sure "subst" and "subst_lemma" are also extended correctly
+otherwise signaling an error.
+Additionally, our compiler must support direct inheritance (and
+overriding) of the old case handlers for "var", "lam" and etc, to avoid
+the boilerplate code.
 
-\textlabel{Challenge (2)}{chg:definition-relevant-reasoning}~\textbf{Definition-awared Reasoning vs.
-Overridability}.
-In family polymorphism for general purposed programming language, overriding is allowed to happen anywhere.
-But the problem becomes more subtle in the case of dependent type since the later fields can have a type dependent on the former fields.
-Consider a small example where we have consecutive fields \mintinline{Coq}{{..
-a : nat := 1; a-def : Eq a 1 := eq_refl; ..}}, should we support override "a"?
-If we do support it, how can our compiler detect that "a-def" is not inheritable and signal an error when such thing happens?
+\textlabel{Challenge (2)}{chg:definition-relevant-reasoning}~\textbf{Definition-awared Reasoning vs. Overridability}.
+In family polymorphism for general purposed programming language,
+overriding is allowed to happen anywhere. But the problem becomes more
+subtle in the case of dependent type since the later fields can have a
+type dependent on the former fields.  Consider a small example where we
+have consecutive fields \mintinline{Coq}{{.. a : nat := 1; a-def : Eq a
+1 := eq_refl; ..}}, should we support override "a"? If we do support it,
+how can our compiler detect that "a-def" is not inheritable and signal
+an error when such thing happens?
 
-A similar example is about "subst" and "subst_lemma" in \cref{fig:STLC-example}.
-For "subst_lemma" to hold throughout two families, we shouldnot allow overriding on the definition of "subst".
-For example, we should still stick with the fact that "subst tunit x y = tunit", as this fact is used when proving "subst_lemma" and we want this lemma to be inherited to the children family.
+A similar example is about "subst" and "subst_lemma" in
+\cref{fig:STLC-example}. For "subst_lemma" to hold throughout two
+families, we shouldnot allow overriding on the definition of "subst".
+For example, we should still stick with the fact that "subst tunit x y =
+tunit", as this fact is used when proving "subst_lemma" and we want this
+lemma to be inherited to the children family.
 
-These two examples both show the incompatibility between exposing the concrete definition of a field for reasoning (this second example even involves the definition of fixpoint and pattern matching) and overridability.
-We need to be careful about this tension when designing the language.
+These two examples both show the incompatibility between exposing the
+concrete definition of a field for reasoning (this second example even
+involves the definition of fixpoint and pattern matching) and
+overridability. We need to be careful about this tension when designing
+the language.\YZ{
+  The two examples seem to suggest there should not be any overriding allowed at all.
+  Is there an example where you actually want the ability to override?
+}
 
-\textlabel{Challenge (3)}{chg:consistency}~\textbf{Consistency vs.
-Self Parametrization}.
-The metatheoretical formulation in~\citet{zm2017} for family polymorphism is very natural as they consider an extension of the family as the concatenation of a new linkage.
-However, this formulation is not directly adaptable for dependent type theory because it can easily encode non-termination sabotaging consistency by the presence of "self".
-Informally, in the orignal formulation, we can override the "f" in family $\{"f" := id; "g" := "call self.f"\}$ with "call self.g", resulting the mutual calling of "f" and "g" thus a deadloop.
-This conflict between self parametrization and consistency requests changing the original formulation of family and inheritance for dependent type theory.
+\textlabel{Challenge (3)}{chg:consistency}~\textbf{Consistency vs. Self Parametrization}.
+The meta\-theoretical formulation in~\citet{zm2017} for family
+polymorphism is very natural as they consider an extension of the family
+as the concatenation of a new linkage. However, this formulation is not
+directly adaptable for dependent type theory because it can easily
+encode non-termination sabotaging consistency by the presence of "self".
+Informally, in the orignal formulation, we can override the "f" in
+family $\{"f" := id; "g" := "call self.f"\}$ with "call self.g",
+resulting the mutual calling of "f" and "g" thus a deadloop. This
+conflict between self parametrization and consistency requests changing
+the original formulation of family and inheritance for dependent type
+theory.
 
 \textlabel{Challenge (4)}{chg:software-engineering}~\textbf{Engineering and Ecosystem}.
-Not only for the metatheory, we also want our implementation of family to have good compatibility with the rest of Coq.
-For example, we shouldn't compromise \textbf{tactic programming} as it is part of the idiomatic Coq programming.
-Additionally, the interactive development style requires \textbf{incremental type checking}---our design has to be able to type check each Coq Vernacular command once inputted.
-What's more, we want our family ``term'' to incorporate Coq \textbf{without breaking its other functionality}.
-For example, the user may want to project the family member anywhere inside an arbitrary Gallina term, that may later be computed and normalized.
-However, families themselves can be second class citizens just like \citet{zm2017}.\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).
-I think most readers will benefit from a recap of this encoding.\\
+Not only for the metatheory, we also want our implementation of family
+to have good compatibility with the rest of Coq. For example, we
+shouldn't compromise \textbf{tactic programming} as it is part of the
+idiomatic Coq programming experience. Additionally, the interactive development
+style requires \textbf{incremental type checking}---our design has to be
+able to type check each Coq Vernacular command once inputted. What's
+more, we want our family ``term'' to be incorporated into Coq \textbf{without
+breaking its other functionality}. For example, the user may want to
+project the family member anywhere inside an arbitrary Gallina term,
+that may later be computed and normalized. However, families themselves
+can be second class citizens just like \citet{zm2017}.\YZ{I was thinking
+if the paper should start by talking about the functional encoding of
+inheritance (namely exposing a self parameter to allow late binding, and
+later taking the fixpoint of the extensible module when it needs to be
+accessed outside the family).  I think most readers will benefit from a
+recap of this encoding.\\
 Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).

### 2022-07-04 coqimpl.tex
Add code for Section 5.1
resolve comments

@@
-\subsection{Language Design}\YZ{I think in §3.1, you should use a figure to show how the actual code the programmer will write to solve the programming challenge outlined in Figure 1.}\EDJreply{Done.}
+\subsection{Language Design}
+
 Inspired by the above challenges and Family Polymorphism, we propose
@@
 than a ``mapping'', it is still possible to
-\text{mix-in} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance.
+\text{mixin} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance.
 However, a good definition of mixin is still under investigation.

### 2022-07-04 lang-design.tex
resolve comments

@@
 % recursors.
-\YZ{Seems like what's missing here is to
-convince the reader why exposing the definition of subst could be a
-problem for family polymorphism.}\EDJreply{Actually I don't see why I need to emphasize this "as a problem".
-This is just a requirement like any other requirements earlier.
-I don't know how to show its difficulty at this early stage.}\YZreply{There is apparently a tension between the need to expose definitional equality and the ability to override.
-Looking at the bigger picture, I think this section reads better if phrased as discussing challenges posed by adding family polymorphism to a proof assistant.
-I have some initial ideas that I need to discuss with you.}\EDJreply{Design Space section is rewritten into Challenge now.}
 

### 2022-07-04 lang-design.tex
Reply comments.

@@
 
-These two examples both show the incompatibility between exposing the
+Note we cannot simply just abandon overriding because we do want it sometimes---say we want a monotonic function \mintinline{Coq}{{..
+f := ..
+; p : monotonic f; ..}}.
+The proof of "p" likely requires the concrete definition of "f" but in the future we might want to override "f" and "p" together to swap the definition.
+
+These three examples both show the incompatibility between exposing the
 concrete definition of a field for reasoning (this second example even

### 2022-07-04 coqimpl.tex
Fail to move some parts from sec 3.2 to sec 3.1.

@@
   V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
-  \quad |\quad "EndFamily" \\
+  \quad |\quad "EndFamily" \quad | \quad "SealFamily" \\
   & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
   & | \quad "FTheorem"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle \\
-  & |\quad "Extend"~"Family"~\langle ident\rangle \\ 
+  & |\quad ["Extend" | "Override"]~"Family"~\langle ident\rangle \\ 
   & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
@@
   & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
-  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad | \quad ...
+  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad \\
+  & | \quad "Closing Fact"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\ 
+  & | \quad "Closing Fact"~\langle ident\rangle~":"~\langle term \rangle~"by"~"\{"\langle \text{one tactic expression}\rangle"\}" \\ 
+  & | \quad "MetaData"~\langle ident\rangle \quad | \quad "EndMetaData" \\ 
 \end{align*}
@@
   motive ...
-  ..
+ Proof.  ...
  Qed.
@@
   : context -> term -> type := ...
- Extend FTheorem subst_lemma.
+ Extend FTheorem subst_lemma. Proof.
   ... (* Need to prove extra cases *)
@@
 
-These are the core extensions we proposed to support \cref{fig:STLC-example}.
-We use "Family" and "EndFamily" to scope, and "Field" is used to define new members in a family.
+These are the core extensions we proposed to support \cref{fig:STLC-example}, and we adapt them into \cref{fig:STLC-example2}.
+We use "Family" and "EndFamily" to scope.
+"Field" is used to define new members in a family.
 "FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
@@
 
-We can inherit a top-level family "parent" using \mintinline{Coq}{Family
-... extends parent} as a new top-level family. As expected, "Override"
+We can also use "Family" and "SealFamily" to scope a (nested) family.
+The resulting family only support direct inheritance or "Override" (i.e.
+we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:definitional-relevant-reasoning}.
+What's more, "Override" family can only apply to sealed family as well (not allowed on normal family).
+
+We can inherit a top-level family "parent" using \mintinline{Coq}{Family ... extends parent} as a new top-level family. As expected, "Override"
 will override a field from the parent. "Inherit" can be used to inherit
-nested families, inductive types, and fields. The fields inside a family
+nested families, inductive types, and fields.
+"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve\ref{chg:definitional-relevant-reasoning}.
+The fields inside a family
 are order-sensitive because we have dependent types here. So "Inherit
@@
 
+We will explain the usage of "Closing Fact" and "MetaData" later---they are here to resolve \ref{chg:software-engineering} to some extent.
 % This incomplete support can already be useful for extending functions.
@@
 has already allowed fields to be type-dependent on former fields thus we can see concretely how
-\ref{chg:definition-relevant-reasoning} appears in our current formulation and what we needs to solve this challenge.
+\ref{chg:definition-relevant-reasoning} appears in our current formulation and how "Final" can help to solve this challenge.
 
@@
 To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
-In a sealed family, every field will expose its definition inside the family but a family as a whole will be sealed into a ``relaxed'' type.
-Thus we can override one sealed family as a whole and change the internal implementation.
+In a sealed family, every field will expose its definition inside the sealed family but a family as a whole will be sealed into a ``relaxed family type''.
+Thus we can override one sealed family as a whole and change the internal implementation as long as they can satisfy the ``relaxed family type''.
 
@@
 \begin{minted}{Coq}
-  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff.
+Definition neg := b_rec (λ _ ⇒ b) neg_handler.tt neg_handler.ff.
 \end{minted}
@@
 \begin{minted}{Coq}
-  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff neg_handler.uu.
+Definition neg := b_rec (λ _ ⇒ b) neg_handler.tt neg_handler.ff neg_handler.uu.
 \end{minted}

### 2022-07-04 coqimpl.tex
Fix typo.

@@
 The resulting family only support direct inheritance or "Override" (i.e.
-we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:definitional-relevant-reasoning}.
+we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:definition-relevant-reasoning}.
 What's more, "Override" family can only apply to sealed family as well (not allowed on normal family).
@@
 nested families, inductive types, and fields.
-"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve\ref{chg:definitional-relevant-reasoning}.
+"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:definition-relevant-reasoning}.
 The fields inside a family

### 2022-07-04 coq-example.tex
Fix statistics

@@
 
-Our formulation for bare STLC is around 400+ LOCs; the two families implementing product and boolean features both take around 300+ LOCs. 
+Our formulation for bare STLC is around 400+ LOCs; the two families implementing product and boolean features both take around 250+ LOCs. 
 % need comparison with the original implementation
@@
 types, and thus the inversion tactic is not working and we have to manually
-create inversion lemmas. This leads to a mild code bloat.
+create inversion lemmas.
 
@@
 generated by the plugin, but still it will require propositional 
-computational rules.
+computational rules.  What's more, we don't yet support the \mintinline{Coq}{Hint}. These three differences on experience lead to a mild code bloat.
 

### 2022-07-05 introduction.tex
baby steps

@@
 % contribution
+
+It is hard to write modular, extensible code.
+It is even harder to write modular, extensible proofs.
+While there exist many linguistic solutions to code reuse and
+extensibility in object-oriented and functional languages,
+linguistic solutions are lacking in proof assistants.
+
+Meanwhile, language designers and researchers are increasingly
+using proof assistants to mechanize meta\-theories, building on each
+other's work.
+Therefore, it has become ever more important that proof assistants
+make it easy to write reusable, extensible code and proofs.

### 2022-07-05 coqimpl.tex
Update propositional partial recursor Update intro with structure and contribution.

@@
 complete extensional characterization of all the extension of a given (non-indexed)
-inductive type.\YZ{What about indexed inductive types?}
+inductive type.\YZ{What about indexed inductive types?}\EDJreply{Done.
+I also have a formalization of this proof now, for indexed type and dependent eliminator.
+Though the formalization is about single constructor.}
 
@@
 \end{minted}
-% To generalize to indexed inductive type, we base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}. Please refer to
+To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable.
+Please refer to the supplementary material and appendix for a formalization of the left inverse in the context of indexed inductive type.

### 2022-07-05 introduction.tex
Update propositional partial recursor Update intro with structure and contribution.

--- 2022-05-18 introduction.tex
@@
 % contribution
+
+\large{\textbf{Structure of the paper}} We will quickly introduce Family Polymorphism and the challenges to adapt it into dependent type setting in \cref{sec:background+challenge}.
+After that, we will talk about the language design of family polymorphism in dependent type setting and the implementation of our Coq plugin in \cref{sec:coqimpl}.
+Then, we consider the meta-theory of incorporating family polymorphism into predicative MLTT, and deriving consistency and canonicity results in \cref{sec:metatheory}.
+\ref{sec:related-work} discusses related works and \ref{sec:conclusion} concludes.
+
+\large{\textbf{Contributions}} include the following
+\begin{itemize}
+  \item To our knowledge, this is the first attempt incorporating family polymorphism with theorem proving and higher order logic
+  \item We contribute a prototypical Coq plugin for the interested users and developers 
+  \item We provide some first-step meta-theoretical results, including consistency and canonicity
+\end{itemize}

### 2022-07-05 introduction.tex

@@
 % contribution
+It is hard to write modular, extensible code.
+It is even harder to write modular, extensible proofs.
+While there exist many linguistic solutions to code reuse and
+extensibility in object-oriented and functional languages,
+linguistic solutions are lacking in proof assistants.
+
+Meanwhile, language designers and researchers are increasingly
+using proof assistants to mechanize meta\-theories, building on each
+other's work.
+Therefore, it has become ever more important that proof assistants
+make it easy to write reusable, extensible code and proofs.
 

### 2022-07-12 introduction.tex
use \paragraph

@@
 other's work.
-Therefore, it has become ever more important that proof assistants
+As a result, it has become ever more important that proof assistants
 make it easy to write reusable, extensible code and proofs.
 
-\large{\textbf{Structure of the paper}} We will quickly introduce Family Polymorphism and the challenges to adapt it into dependent type setting in \cref{sec:background+challenge}.
+\paragraph{Structure of the paper} 
+We will quickly introduce Family Polymorphism and the challenges to adapt it into dependent type setting in \cref{sec:background+challenge}.
 After that, we will talk about the language design of family polymorphism in dependent type setting and the implementation of our Coq plugin in \cref{sec:coqimpl}.
@@
 
-\large{\textbf{Contributions}} include the following
+\paragraph{Contributions} include the following
 \begin{itemize}

### 2022-07-12 introduction.tex
intro tweak

@@
 % contribution
-It is hard to write modular, extensible code.
-It is even harder to write modular, extensible proofs.
-While there exist many linguistic solutions to code reuse and
-extensibility in object-oriented and functional languages,
-linguistic solutions are lacking in proof assistants.
 
-Meanwhile, language designers and researchers are increasingly
+Language designers and researchers are increasingly
 using proof assistants to mechanize meta\-theories, building on each
@@
 make it easy to write reusable, extensible code and proofs.
+
+However, linguistic solutions to type-safe code reuse and extensibility
+are lacking in proof assistants.
+This situation is in contrast to modern languages,
+such as Scala \cite{scala-oopsla05}, that have a good supply of
+linguistic features which solve the Expression problem \cite{wadler-ep}.
 

### 2022-07-13 coq-example.tex
Adding comments.

@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{wrapfigure}{r}{0.5\textwidth}\YZ{It is more convincing if you have four or five family mixin together, positive product, negative product, sum type, boolean, and natural number, general recursion and one more}
 \begin{lstlisting}[language=Coq,frame = single, escapeinside={@}{@}]
@@
 
-\subsection{Abstract Interpreters for \texttt{Imp}}
+\subsection{Abstract Interpreters for \texttt{Imp}}\YZ{Try to have two runnable interpreter -- choose another simple one like constant propagation. and remove LangMore}
 The second example is adapted and modified from \citet{zm2017}.

### 2022-07-13 coqimpl.tex
Adding comments.

@@
 Once the user closes a family with "EndFamily", they can access its fields by using module syntax.
-For example, they can access "A.a" to get "1" after "Family A" is defined.
-This is achieved by a complete compilation from a family to a module, via aggregating the list of the modules in the internal representation: the plugin will repeatedly using Coq's "Include" command as the "Module A" in \cref{fig:plugin-example1} illustrates.
+For example, they can access "A.a" to get "1" after "Family A" is defined.\YZ{1.
+move some of sec 3.2 ~> 3.1; 2.
+hint vaguely on how to do type checking in sec 3.1 (user interface)..
+metatheory part should also mention this part in sec 3.1} This is achieved by a complete compilation from a family to a module, via aggregating the list of the modules in the internal representation: the plugin will repeatedly using Coq's "Include" command as the "Module A" in \cref{fig:plugin-example1} illustrates.
 Now, the user can use "Module A" anywhere as they are using "Family A".
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]\YZ{Change this one into STLC example, to see the layout}
   \begin{minipage}[t]{0.32\linewidth}
@@
 
-\textbf{Hack: Global Reasoning.}
+\textbf{Hack: Global Reasoning.}\YZ{This is more like an escape hatch. Escape from family poly into vanilla Coq. Maybe these two hacks should be grouped into one section}
 % explain the places using ``Closing Fact''
@@
 having boilerplate code, \textbf{semantically}.
-\begin{figure}[!htb]
+\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}
 \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]

### 2022-07-13 coqimpl.tex
Move a lot from sec3.2 to sec3.1

@@
 
-These are the core extensions we proposed to support \cref{fig:STLC-example}, and we adapt them into \cref{fig:STLC-example2}.
+\textbf{Family and Defining Fields.} These are the core extensions we proposed to support \cref{fig:STLC-example}, and we adapt them into \cref{fig:STLC-example2}.
 We use "Family" and "EndFamily" to scope.
@@
 Nested families are also allowed.
+
+The user can use "self__A.<path>" to access the former defined fields inside a surrounding defining "Family A".\footnote{Note that, in most of our examples in the paper, these "self__" prefix is too long and verbose so we omitted them.} Note that, "self__A" is only visible in the Vernacular commands from our plugin.\footnote{(In other words, using "Definition" from Coq will not invoke the plugin to prepare "self__A" to be accessible---most original Vernacular command from vanilla Coq will not do such preparation.)}  Once the user closes a \textbf{top} family with "EndFamily", they can access the fields of the closed family by using module syntax.
+For example, they can access "A.a" to get "1" after "Family A" is defined.
+This "A" is truly accessible anywhere by vanilla Coq without any plugin preparation, just like a mundane module.
+
+Most programming experiences are preserved from Coq---each term inside the field definition is a well-typed Gallina term in vanilla Coq.
+Note that, a Family is not a first class value and thus we need to project out the Gallina terms from them.
+
+% Inheritance here
+\textbf{Inheritance on Fields.} "Family A extends B" has to appear at the top level (not inside any Family) and will start to construct the children "Family A" given the parent "Family B".
+Every field inside the inheriting family will be one of the three kinds of ``inheritance''---extending, overriding, and inheriting.
+The syntax of extending a new field is to directly use "Field", while overriding requires a prefix of "Override" before "Field f : .." and "Inherit f" will inherits field "f" from the parent.
+
+To resolve \ref{chg:consistency}, we make sure when inheriting or overriding on "f" happens, the context of the field "f" is also ``inherited''.
+Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty thus "self.g" doesn't type-check.
+Thus, before we inherit "g", the dependee "f" needs to be inherited.
+
+\begin{figure}[!htb]
+  \begin{minipage}[t]{0.45\linewidth}
+\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
+Family I.
+ Field a : nat := 1.
+ Field b : a = a := eq_refl.
+ Fail Field c : a = 1 := eq_refl. (*Fail 1*)
+
+ Final Field a' : nat := 1.
+ Field c : a' = 1 := eq_refl.
+ Field bop : nat -> nat -> nat := (+).
+ Fail Field comm_bop : ...
+
+ Family MonotonicF.
+  Field f : nat -> nat := non_impl.
+  Field p : monotonic f := non_impl.
+ SealFamily. 
+EndFamily.
+\end{lstlisting}
+  \end{minipage}
+\begin{minipage}[t]{0.45\linewidth}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+Family I2 extends I.
+ Override Field a : nat := 1. (* Override *)
+ Inherit b.                   (* Inherit  *)
+
+ Field d : nat := a.          (* Extend   *)
+ Fail Override Field a' : nat := 2. (*Fail 2*)
+
+ Override Field bop := (-).
+
+ Inherits Until MonotonicF.
+ Override Family MonotonicF.
+  Field f : nat -> nat := (+ 1).
+  Field p : monotonic f := ...
+ EndFamily. 
+EndFamily.
+\end{lstlisting}
+  \end{minipage}
+  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
+\end{figure}
+
+% Talk about Final
+\textbf{Final as non-overridable.} Every field is by default overridable, reminding us \ref{chg:definition-relevant-reasoning}.
+To resolve it, we provide the keyword "Final" to allow the user to specify \textit{non-overridability} and give the user the opportunity to choose from overriding and exposing the definition.
+
+Take the field "c" of "Family I" in \cref{fig:plugin-example3} as the
+example. We can see that the first attempt of proving "a = 1" fails.
+This should fail---because we want to retain the possibility
+to override "a" and inheriting "c" in the future, but they are conflicting, since "c" witnesses "a" fix to "1".
+On the other hand, once we decorate it upon "a'", the second attempt \mintinline{Coq}{c : a' = 1 := eq_refl} will succeed.
+Contrary to "a", our "a'" is not overridable in "Family I2", mediating the \ref{chg:definition-relevant-reasoning}.
+
+This example might sound trivial and contrived. Let's look at the
+example field binary operation "bop" initially set to be addition.\YZ{This 'add' field is no longer there.}\EDJreply{Thanks. Fixed} Now that we only expose the type of "bop", we
+cannot prove its commutativity in "comm_bop" (because we don't have
+access to its concrete definition) but we can override "bop" with
+subtraction function in "Family I2". This makes sense because we
+shouldn't (have the ability to) inherit commutativity property for subtraction
+anyway. This is contrary to how "a'" is declared "Final", the second "c"
+can prove that it's equal to 1 but "a'" is not overridable in the
+"Family I2".
+
+In our example, the seal family "I.MonotonicF" is actually simulating
+interfaces as well.  We have
+\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
+and we assigned it to both "f" and "p", and later instantiate in
+"I2" with concrete implementation by overriding this family
+"I.MonotonicF".
+
+\textbf{Overridable/Sealed Family as better Sigma Type.} Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming.
+There are cases where overriding and
+(definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
+\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
+because to prove
+the latter we need to have concrete definition of "f" in the family type, which
+makes "f" non-overridable at all. One option is to use sigma type
+\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
+a bundle of stuff with multiple refinements.
+
+To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
+In a sealed family, every field will expose its definition inside the sealed family but a family as a whole will be ``sealed'' with a ``relaxed family type''---it is an aggregation of the types instead of the concrete definitions of all the field.
+Thus we can override one sealed family as a whole and change the internal implementation as long as they can satisfy the ``relaxed family type''.
+
+For example, the "MonotonicF" in \cref{fig:plugin-example3}:
+we simply seal
+\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
+together inside an sealed family.
+Because the definition of "f" is exposed inside the family, 
+its definition is
+transparent and can be used to prove "p".
+However, because the enclosing family of "f" is overridable, to swap the
+implementation of "f", we just override the whole ``sealed'' family,
+like we do in "I2.MonotonicF".
+
+% if the user wants a given field to be non-overridable. We will later elaborate its detail application.
+
+\textbf{Inheritance on Nested Families.} We also provide extending, inheriting and overriding for nested family (member).
+Extension is done by simply declaring a nested family; the same syntax for inheriting a field is used in inheriting a family.
+However, overriding is special---overriding can only apply to \textit{sealed} nested family in the parent family.
+A family is sealed when using "SealFamily" as the scoping instead of "EndFamily".
+We will later elaborate the motivation and details of sealed family.
+
+% Currently, every sealed family can only be direct inherited or overridden as a whole. It cannot be extended.
+
+\textbf{Extensible Inductive Type.} We provide the keyword "FInductive" to enable extensible inductive type.
+"Extend FInductive" is used to support extension of the parent inductive type inside a children family.
+The syntax of these two is close to the mundane "Inductive" keyword.
 
@@
 This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family.
-For example, in \cref{fig:STLC-example2},  "FTheorem" has a similar functionality but is expected to be used for theorem proving---the handler and the family will be automatically constructed during proving.
+Once recursor is created it will look like any other function fields(e.g.
+"subst" in \cref{fig:STLC-example2}).
 
-We can also use "Family" and "SealFamily" to scope a (nested) family.
+"FTheorem" has a similar functionality but is expected to be used for theorem proving---the handler and the family will be automatically constructed during proving.
+
+% We can also use "Family" and "SealFamily" to scope a (nested) family.
 The resulting family only support direct inheritance or "Override" (i.e.
@@
 
-We can inherit a top-level family "parent" using \mintinline{Coq}{Family ... extends parent} as a new top-level family. As expected, "Override"
-will override a field from the parent. "Inherit" can be used to inherit
-nested families, inductive types, and fields.
+% We can inherit a top-level family "parent" using \mintinline{Coq}{Family ... extends parent} as a new top-level family. As expected, "Override"
+% will override a field from the parent. "Inherit" can be used to inherit
+% nested families, inductive types, and fields.
 "Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:definition-relevant-reasoning}.
 The fields inside a family
-are order-sensitive because we have dependent types here. So "Inherit
-Until" is used if the user wants to inherit a range of fields.  "Extend Family" is used only inside a inheriting family. Thus nested
-family can only be inherited or extended together with its belonging family
-(i.e. ultimately top level family), thus our support of family polymorphism is
-incomplete.
+% are order-sensitive because we have dependent types here. So "Inherit
+% Until" is used if the user wants to inherit a range of fields.  "Extend Family" is used only inside a inheriting family. Thus nested
+% family can only be inherited or extended together with its belonging family
+% (i.e. ultimately top level family), thus our support of family polymorphism is
+% incomplete.
 
@@
 
-Once the user closes a family with "EndFamily", they can access its fields by using module syntax.
-For example, they can access "A.a" to get "1" after "Family A" is defined.\YZ{1.
+\YZ{1.
 move some of sec 3.2 ~> 3.1; 2.
@@
 
-\begin{figure}[!htb]
-  \begin{minipage}[t]{0.45\linewidth}
-\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
-Family I.
- Field a : nat := 1.
- Field b : a = a := eq_refl.
- Fail Field c : a = 1 := eq_refl. (*Fail 1*)
-
- Final Field a' : nat := 1.
- Field c : a' = 1 := eq_refl.
- Field bop : nat -> nat -> nat := (+).
- Fail Field comm_bop : ...
-
- Family MonotonicF.
-  Field f : nat -> nat := non_impl.
-  Field p : monotonic f := non_impl.
- SealFamily. 
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-\begin{minipage}[t]{0.45\linewidth}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family I2 extends I.
- Override Field a : nat := 1. (* Override *)
- Inherit b.                   (* Inherit  *)
-
- Field d : nat := a.          (* Extend   *)
- Fail Override Field a' : nat := 2. (*Fail 2*)
-
- Override Field bop := (-).
-
- Inherits Until MonotonicF.
- Override Family MonotonicF.
-  Field f : nat -> nat := (+ 1).
-  Field p : monotonic f := ...
- EndFamily. 
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
-\end{figure}
-
 % explaining inheritance
 \textbf{Implementing Inheritance.}
-To resolve \ref{chg:consistency}, we make sure when inheritance happens, not only the content will be inherited, but the context of the corresponding field is also ``inherited''.
-Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty thus "self.g" doesn't type-check.
-
 To simplify the implementation, we can consider all families as
@@
 
-There are three kinds of inheritance---extend, override, and inherit. Simiarly, we
-use a list to encode inheritance data in the
+Simiarly, we
+use a list to encode the three kinds of inheritance data in the
 implementation. The inheritance data will include the
@@
 
-\textbf{Reasoning Needs Non-Overridable.}
-This compilation strategy
-has already allowed fields to be type-dependent on former fields thus we can see concretely how
-\ref{chg:definition-relevant-reasoning} appears in our current formulation and how "Final" can help to solve this challenge.
-
-% However, the current implementation is not satisfactory enough. There are cases where we need to expose
-% the concrete definition of a given field (especially when we are proving
-% properties about it).
-
-% For example, we may
-% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
-% and intend to prove its commutativity in a following field $"comm"$.
-% But to prove so, we need to know more than just a type interface of
-% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
-% future, once "add" is overridden into a subtraction function then
-% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
-% A more concrete example is in \cref{fig:plugin-example1}, consider the
-% case we have "Field b : a = 1 := .." instead. Then our
-% compilation and Coq's type checking will both fail: in the context of
-% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
-% more (and this is necessary so that "a" can be overridden in the
-% future).
-Take the field "c" of "Family I" in \cref{fig:plugin-example3} as the
-example. We can see that the first attempt of proving "a = 1" fails.
-\textit{This should fail}---because we want to retain the possibility
-to override "a" in the future, but the mere existence of the field "c"
-will stop that, since "c" witnesses "a" fix to "1". \textit{This will fail} in our implementation---the proof of "c" will require a concrete definition of "a",
+Our implementation also make sure the first attempt of defining "c" failed successfully in \cref{fig:plugin-example3}---the proof of "c" will require a concrete definition of "a",
 but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in (a module with) the context where we only know \mintinline{Coq}{a : nat} thus failed.
 
-To enable the exposure of the concrete definition,
-our plugin provide "Final" keyword. Once we decorate it upon "a'", the second attempt \mintinline{Coq}{c : a' = 1 := eq_refl} will succeed. Contrary to "a", our "a'" is not overridable in "Family I2".
-
-This example might sound trivial and contrived. Let's look at the
-example field binary operation "bop" initially set to be addition.\YZ{This 'add' field is no longer there.}\EDJreply{Thanks. Fixed} Now that we only expose the type of "bop", we
-cannot prove its commutativity in "comm_bop" (because we don't have
-access to its concrete definition) but we can override "bop" with
-subtraction function in "Family I2". This makes sense because we
-shouldn't (have the ability to) inherit commutativity property for subtraction
-anyway. This is contrary to how "a'" is declared "Final", the second "c"
-can prove that it's equal to 1 but "a'" is not overridable in the
-"Family I2".
-
-% To fix that, we need more information on "a" than its type---we
-% need to expose the concrete definition "a = 1" to the context of "b".
-%\YZ{I had a hard time working through these two paras. I guess they are
-%saying that there is a need to expose definitional equalities, and
-%therefore we need non-overridables. This causal flow should be made
-%more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph. Please check and resolve if find satisfied.}
-
+\textbf{Implemening "Final" and "Sealed Family".}
 We implement "Final" by
-exposing the whole definition directly into the compiled module type. Doing so we will prohibit overriding because
-subtraction doesn't have the same definition as $add$. We require the
+exposing the whole definition directly into the compiled module type.
+Doing so will prohibit overriding.
+Say, the example "bop" in \cref{fig:plugin-example3}, will be prohibited to be overridden by $(-)$
+subtraction doesn't have the same definition as $(+)$. We require the
 users to manual decorate field as `Final' so that the plugin will proceed with this special treatment.
@@
 
-\textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.} 
-Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming. There are cases where overriding and
-(definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
-\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
-because to prove
-the latter we need to have concrete definition of "f" in the family type, which
-makes "f" non-overridable at all. One option is to use sigma type
-\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
-a bundle of stuff with multiple refinements.
+% \textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.}
 
-To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
-In a sealed family, every field will expose its definition inside the sealed family but a family as a whole will be sealed into a ``relaxed family type''.
-Thus we can override one sealed family as a whole and change the internal implementation as long as they can satisfy the ``relaxed family type''.
-
-For example, the "MonotonicF" in \cref{fig:plugin-example3}:
-we simply seal
-\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
-together inside an sealed family.
-Because the definition of "f" is exposed inside the family, 
-its definition is
-transparent and can be used to prove "p".
-However, because the enclosing family of "f" is overridable, to swap the
-implementation of "f", we just override the whole ``sealed'' family,
-like we do in "I2.MonotonicF".
-
-And the compiled module type of
-"MonotonicF" will be fixed throughout the extension in different
-families.
-Any overriding family will always have these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
+The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type.
+The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
+For example, the compiled module type of
+"MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}".
+Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
 }
 
-In our example, the seal family "I.MonotonicF" is actually simulating
-interfaces as well.  We have
-\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
-and we assigned it to both "f" and "p", and later instantiate in
-"I2" with concrete implementation by overriding this family
-"I.MonotonicF".
 % Basically, {a : int -> int; property : a is monotonic}, in this case
@@
 %   this refinement/property
-The implementation of sealed family is not so different from that of the normal family, but we need to be careful about the compiled module type.
 
@@
 
-We start with the programming interface.
+% We start with the programming interface.
 We use "FInductive" to define an extensible inductive type and in the children family we use "Extend Finductive ..." to extend the type with new constructors.

### 2022-07-13 coqimpl.tex
Move sec3.1 ~> sec3, sec3.2~>sec 4.

@@
-\subsection{Language Design}
+% \subsection{Language Design}
 
-Inspired by the above challenges and Family Polymorphism, we propose
-the following extension to Coq's Vernacular Commands, and the corresponding STLC example.
-% basically write down the specification of surface Coq commandline
-%   and then explain  how the language design makes the programmer’s life easier
-\begin{align*}
-  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
-  \quad |\quad "EndFamily" \quad | \quad "SealFamily" \\
-  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
-  & | \quad "FTheorem"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle \\
-  & |\quad ["Extend" | "Override"]~"Family"~\langle ident\rangle \\ 
-  & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
-  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
-  & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
-  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad \\
-  & | \quad "Closing Fact"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\ 
-  & | \quad "Closing Fact"~\langle ident\rangle~":"~\langle term \rangle~"by"~"\{"\langle \text{one tactic expression}\rangle"\}" \\ 
-  & | \quad "MetaData"~\langle ident\rangle \quad | \quad "EndMetaData" \\ 
-\end{align*}
-
-\begin{figure}[!htb]
-  \begin{minipage}[t]{0.47\linewidth}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family STLC.
- FInductive ty : Set :=
-  | unit : ty
-  | arrow : ty -> ty -> ty.
- FInductive term : Set := 
-  | var : id -> term 
-  | lam : id -> term -> term ...
- Family subst_internal.
-  Final Field var := ..
-  Final Field lam := ..
-   ...
- EndFamily.
- FRecursor subst
-  about tm motive ..
-  using subst_internal.
- ...
- FInductive has_type 
-  : context -> term -> type := ...
- FTheorem subst_lemma 
-  about has_type 
-  motive ...
- Proof.  ...
- Qed.
- FInductive step : term -> term -> Prop 
-  := ...
-(* ... and more, end with type safety *)
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-  \begin{minipage}[t]{0.47\linewidth}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family STLC_bool extends STLC.
- Extend FInductive ty : Set :=
-  | bool : ty.
-
- Extend FInductive term : Set := 
-  | tt : term | ff : term 
-  | tif : term -> term -> term -> term
- Extend Family subst_internal.
-  Final Field tt := ..
-  Final Field ff := ..
-  Final Field tif := ..
- EndFamily. (* Automatic inheriting others 
-              from STLC.subst_internal *)
- Inherit subst.
-(* Plugin will redo exhaustiveness check, in case 
- subst_internal is not extended correctly *)
- Extend FInductive has_type (* .. and new ty too *)
-  : context -> term -> type := ...
- Extend FTheorem subst_lemma. Proof.
-  ... (* Need to prove extra cases *)
- Qed.
-
- Extend FInductive step : term -> term -> Prop 
-  := ... (* Need to expand this binary relation *)
-(* ... and more extension *)
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-  \caption{Example STLC and its extension, in designed syntax}\label{fig:STLC-example2}
-\end{figure}
-
-\textbf{Family and Defining Fields.} These are the core extensions we proposed to support \cref{fig:STLC-example}, and we adapt them into \cref{fig:STLC-example2}.
-We use "Family" and "EndFamily" to scope.
-"Field" is used to define new members in a family.
-"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
-Nested families are also allowed.
-
-The user can use "self__A.<path>" to access the former defined fields inside a surrounding defining "Family A".\footnote{Note that, in most of our examples in the paper, these "self__" prefix is too long and verbose so we omitted them.} Note that, "self__A" is only visible in the Vernacular commands from our plugin.\footnote{(In other words, using "Definition" from Coq will not invoke the plugin to prepare "self__A" to be accessible---most original Vernacular command from vanilla Coq will not do such preparation.)}  Once the user closes a \textbf{top} family with "EndFamily", they can access the fields of the closed family by using module syntax.
-For example, they can access "A.a" to get "1" after "Family A" is defined.
-This "A" is truly accessible anywhere by vanilla Coq without any plugin preparation, just like a mundane module.
-
-Most programming experiences are preserved from Coq---each term inside the field definition is a well-typed Gallina term in vanilla Coq.
-Note that, a Family is not a first class value and thus we need to project out the Gallina terms from them.
-
-% Inheritance here
-\textbf{Inheritance on Fields.} "Family A extends B" has to appear at the top level (not inside any Family) and will start to construct the children "Family A" given the parent "Family B".
-Every field inside the inheriting family will be one of the three kinds of ``inheritance''---extending, overriding, and inheriting.
-The syntax of extending a new field is to directly use "Field", while overriding requires a prefix of "Override" before "Field f : .." and "Inherit f" will inherits field "f" from the parent.
-
-To resolve \ref{chg:consistency}, we make sure when inheriting or overriding on "f" happens, the context of the field "f" is also ``inherited''.
-Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty thus "self.g" doesn't type-check.
-Thus, before we inherit "g", the dependee "f" needs to be inherited.
-
-\begin{figure}[!htb]
-  \begin{minipage}[t]{0.45\linewidth}
-\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
-Family I.
- Field a : nat := 1.
- Field b : a = a := eq_refl.
- Fail Field c : a = 1 := eq_refl. (*Fail 1*)
-
- Final Field a' : nat := 1.
- Field c : a' = 1 := eq_refl.
- Field bop : nat -> nat -> nat := (+).
- Fail Field comm_bop : ...
-
- Family MonotonicF.
-  Field f : nat -> nat := non_impl.
-  Field p : monotonic f := non_impl.
- SealFamily. 
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-\begin{minipage}[t]{0.45\linewidth}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family I2 extends I.
- Override Field a : nat := 1. (* Override *)
- Inherit b.                   (* Inherit  *)
-
- Field d : nat := a.          (* Extend   *)
- Fail Override Field a' : nat := 2. (*Fail 2*)
-
- Override Field bop := (-).
-
- Inherits Until MonotonicF.
- Override Family MonotonicF.
-  Field f : nat -> nat := (+ 1).
-  Field p : monotonic f := ...
- EndFamily. 
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
-\end{figure}
-
-% Talk about Final
-\textbf{Final as non-overridable.} Every field is by default overridable, reminding us \ref{chg:definition-relevant-reasoning}.
-To resolve it, we provide the keyword "Final" to allow the user to specify \textit{non-overridability} and give the user the opportunity to choose from overriding and exposing the definition.
-
-Take the field "c" of "Family I" in \cref{fig:plugin-example3} as the
-example. We can see that the first attempt of proving "a = 1" fails.
-This should fail---because we want to retain the possibility
-to override "a" and inheriting "c" in the future, but they are conflicting, since "c" witnesses "a" fix to "1".
-On the other hand, once we decorate it upon "a'", the second attempt \mintinline{Coq}{c : a' = 1 := eq_refl} will succeed.
-Contrary to "a", our "a'" is not overridable in "Family I2", mediating the \ref{chg:definition-relevant-reasoning}.
-
-This example might sound trivial and contrived. Let's look at the
-example field binary operation "bop" initially set to be addition.\YZ{This 'add' field is no longer there.}\EDJreply{Thanks. Fixed} Now that we only expose the type of "bop", we
-cannot prove its commutativity in "comm_bop" (because we don't have
-access to its concrete definition) but we can override "bop" with
-subtraction function in "Family I2". This makes sense because we
-shouldn't (have the ability to) inherit commutativity property for subtraction
-anyway. This is contrary to how "a'" is declared "Final", the second "c"
-can prove that it's equal to 1 but "a'" is not overridable in the
-"Family I2".
-
-In our example, the seal family "I.MonotonicF" is actually simulating
-interfaces as well.  We have
-\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
-and we assigned it to both "f" and "p", and later instantiate in
-"I2" with concrete implementation by overriding this family
-"I.MonotonicF".
-
-\textbf{Overridable/Sealed Family as better Sigma Type.} Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming.
-There are cases where overriding and
-(definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
-\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
-because to prove
-the latter we need to have concrete definition of "f" in the family type, which
-makes "f" non-overridable at all. One option is to use sigma type
-\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
-a bundle of stuff with multiple refinements.
-
-To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
-In a sealed family, every field will expose its definition inside the sealed family but a family as a whole will be ``sealed'' with a ``relaxed family type''---it is an aggregation of the types instead of the concrete definitions of all the field.
-Thus we can override one sealed family as a whole and change the internal implementation as long as they can satisfy the ``relaxed family type''.
-
-For example, the "MonotonicF" in \cref{fig:plugin-example3}:
-we simply seal
-\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
-together inside an sealed family.
-Because the definition of "f" is exposed inside the family, 
-its definition is
-transparent and can be used to prove "p".
-However, because the enclosing family of "f" is overridable, to swap the
-implementation of "f", we just override the whole ``sealed'' family,
-like we do in "I2.MonotonicF".
-
-% if the user wants a given field to be non-overridable. We will later elaborate its detail application.
-
-\textbf{Inheritance on Nested Families.} We also provide extending, inheriting and overriding for nested family (member).
-Extension is done by simply declaring a nested family; the same syntax for inheriting a field is used in inheriting a family.
-However, overriding is special---overriding can only apply to \textit{sealed} nested family in the parent family.
-A family is sealed when using "SealFamily" as the scoping instead of "EndFamily".
-We will later elaborate the motivation and details of sealed family.
-
-% Currently, every sealed family can only be direct inherited or overridden as a whole. It cannot be extended.
-
-\textbf{Extensible Inductive Type.} We provide the keyword "FInductive" to enable extensible inductive type.
-"Extend FInductive" is used to support extension of the parent inductive type inside a children family.
-The syntax of these two is close to the mundane "Inductive" keyword.
-
-To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness.
-The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family.
-Then
-"FRecursor" will be used to construct a recursor by refering to the handler family.
-\mintinline{Coq}{FRecursor recur about T motive f } "using"
-"handlers" will generate a new field "recur" that is really doing computation
-(using elimination principle of the inductive type "T") with motive "f". The handlers
-are located in the family "handlers", similar to a large pattern-matching statement.
-This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family.
-Once recursor is created it will look like any other function fields(e.g.
-"subst" in \cref{fig:STLC-example2}).
-
-"FTheorem" has a similar functionality but is expected to be used for theorem proving---the handler and the family will be automatically constructed during proving.
-
-% We can also use "Family" and "SealFamily" to scope a (nested) family.
-The resulting family only support direct inheritance or "Override" (i.e.
-we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:definition-relevant-reasoning}.
-What's more, "Override" family can only apply to sealed family as well (not allowed on normal family).
-
-% We can inherit a top-level family "parent" using \mintinline{Coq}{Family ... extends parent} as a new top-level family. As expected, "Override"
-% will override a field from the parent. "Inherit" can be used to inherit
-% nested families, inductive types, and fields.
-"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:definition-relevant-reasoning}.
-The fields inside a family
-% are order-sensitive because we have dependent types here. So "Inherit
-% Until" is used if the user wants to inherit a range of fields.  "Extend Family" is used only inside a inheriting family. Thus nested
-% family can only be inherited or extended together with its belonging family
-% (i.e. ultimately top level family), thus our support of family polymorphism is
-% incomplete.
-
-We will explain the usage of "Closing Fact" and "MetaData" later---they are here to resolve \ref{chg:software-engineering} to some extent.
-% This incomplete support can already be useful for extending functions.
-% Take \cref{fig:plugin-example2} as example, if we want to extend boolean
-% inductive type and its negation, we only need to specify the new
-% constructor for boolean and how negation deals with the new constructor.
-% As a result, the programmer does not have to duplicate code---otherwise
-% the user will need to copy and paste the whole original negation
-% definition and add a new pattern matching case.
-
-\subsection{Plugin Implementation}
+% \subsection{Plugin Implementation}
 

### 2022-07-13 lang-design.tex
Adding comments.
Move sec3.1 ~> sec3, sec3.2~>sec 4.

--- 2022-07-04 lang-design.tex
@@
-\subsection{Challenges}
-
-We want the metatheory mechanization to allow adding new
-function/proving new lemmas, and adding new constructors for inductive
-type at the same time  (Expression Problem~\cite{wadler-ep}). Coq
-support the former but not the later---when an existent inductive type
-is extended with new constructor, we have to modify the pattern matching
-in the existent functions. This motivates us to add inheritance/Family
-Polymorphism into Coq as a solution to the Expression Problem.
+Inspired by the above challenges and Family Polymorphism, we propose
+the following extension to Coq's Vernacular Commands, and the corresponding STLC example.
+% basically write down the specification of surface Coq commandline
+%   and then explain  how the language design makes the programmer’s life easier
+\begin{align*}
+  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
+  \quad |\quad "EndFamily" \quad | \quad "SealFamily" \\
+  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
+  & | \quad "FTheorem"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle \\
+  & |\quad ["Extend" | "Override"]~"Family"~\langle ident\rangle \\ 
+  & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
+  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
+  & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
+  & | \quad "Inherit"~\langle ident\rangle \quad | \quad "Inherit Until"~\langle ident\rangle \quad \\
+  & | \quad "Closing Fact"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\ 
+  & | \quad "Closing Fact"~\langle ident\rangle~":"~\langle term \rangle~"by"~"\{"\langle \text{one tactic expression}\rangle"\}" \\ 
+  & | \quad "MetaData"~\langle ident\rangle \quad | \quad "EndMetaData" \\ 
+\end{align*}
 
@@
 Family STLC.
-Inductive ty : Set :=
+ FInductive ty : Set :=
   | unit : ty
   | arrow : ty -> ty -> ty.
-Inductive term : Set := 
+ FInductive term : Set := 
   | var : id -> term 
-  | lam : id -> term -> term 
-  | tunit : term ...
-Fixpoint subst 
-  : id -> term -> term := ...
-Inductive has_type 
+  | lam : id -> term -> term ...
+ Family subst_internal.
+  Final Field var := ..
+  Final Field lam := ..
+   ...
+ EndFamily.
+ FRecursor subst
+  about tm motive ..
+  using subst_internal.
+ ...
+ FInductive has_type 
   : context -> term -> type := ...
-Proposition subst_lemma :
-  has_type (G; x ↦ A) t T ->
-  has_type G u A ->
-  has_type G (subst x u t) T.
-Inductive step : term -> term -> Prop 
+ FTheorem subst_lemma 
+  about has_type 
+  motive ...
+ Proof.  ...
+ Qed.
+ FInductive step : term -> term -> Prop 
   := ...
@@
 Family STLC_bool extends STLC.
-Extend Inductive ty : Set :=
+ Extend FInductive ty : Set :=
   | bool : ty.
 
-Extend Inductive term : Set := 
+ Extend FInductive term : Set := 
   | tt : term | ff : term 
   | tif : term -> term -> term -> term
+ Extend Family subst_internal.
+  Final Field tt := ..
+  Final Field ff := ..
+  Final Field tif := ..
+ EndFamily. (* Automatic inheriting others 
+              from STLC.subst_internal *)
+ Inherit subst.
+(* Plugin will redo exhaustiveness check, in case 
+ subst_internal is not extended correctly *)
+ Extend FInductive has_type (* .. and new ty too *)
+  : context -> term -> type := ...
+ Extend FTheorem subst_lemma. Proof.
+  ... (* Need to prove extra cases *)
+ Qed.
 
-Extend Fixpoint subst (* Need to handle new term *)
-  : id -> term -> term := ...
-Extend Inductive has_type (* .. and new ty too *)
-  : context -> term -> type := ...
-Extend Proposition subst_lemma :
-  ... (* Need to prove extra cases *)
-
-Extend Inductive step : term -> term -> Prop 
+ Extend FInductive step : term -> term -> Prop 
   := ... (* Need to expand this binary relation *)
@@
   \end{minipage}
-  \caption{Example STLC and its extension}\label{fig:STLC-example}
+  \caption{Example STLC and its extension, in designed syntax}\label{fig:STLC-example2}
 \end{figure}
 
-Refering to the syntax of \citet{zm2017}, we will come up the above
-example. However, to add inheritance into a dependent type system like
-Coq is not simple and there are various challenges and tension we need
-to resolve.
+\textbf{Family and Defining Fields.} These are the core extensions we proposed to support \cref{fig:STLC-example}, and we adapt them into \cref{fig:STLC-example2}.
+We use "Family" and "EndFamily" to scope.
+"Field" is used to define new members in a family.
+"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
+Nested families are also allowed.
 
-\textlabel{Challenge (1)}{chg:extensible-inductive-type}~\textbf{Extensible Inductive Type vs. Exhaustiveness of Inductive Reasoning}.
-Unlike most general purposed programming languages, Coq, as a dependent
-type theory, is a total language and thus each pattern matching cases
-must fully ``cover'' an inductive type. This process is
-``\textit{interactive}'' for inheritance---when inhertiance extend the
-inductive type in the children family but the user forgets to extend the
-corresponding inductive reasoning, our compiler need to signal an error
-correspondingly, or better---guide the user to fill the incomplete
-inductive reasoning. For example, in \cref{fig:STLC-example} when "term"
-is extended with "tt" in the children family "STLC_bool", our compiler
-need to make sure "subst" and "subst_lemma" are also extended correctly
-otherwise signaling an error.
-Additionally, our compiler must support direct inheritance (and
-overriding) of the old case handlers for "var", "lam" and etc, to avoid
-the boilerplate code.
+The user can use "self__A.<path>" to access the former defined fields inside a surrounding defining "Family A".\footnote{Note that, in most of our examples in the paper, these "self__" prefix is too long and verbose so we omitted them.} Note that, "self__A" is only visible in the Vernacular commands from our plugin.\footnote{(In other words, using "Definition" from Coq will not invoke the plugin to prepare "self__A" to be accessible---most original Vernacular command from vanilla Coq will not do such preparation.)}  Once the user closes a \textbf{top} family with "EndFamily", they can access the fields of the closed family by using module syntax.
+For example, they can access "A.a" to get "1" after "Family A" is defined.
+This "A" is truly accessible anywhere by vanilla Coq without any plugin preparation, just like a mundane module.
 
-\textlabel{Challenge (2)}{chg:definition-relevant-reasoning}~\textbf{Definition-awared Reasoning vs. Overridability}.
-In family polymorphism for general purposed programming language,
-overriding is allowed to happen anywhere. But the problem becomes more
-subtle in the case of dependent type since the later fields can have a
-type dependent on the former fields.  Consider a small example where we
-have consecutive fields \mintinline{Coq}{{.. a : nat := 1; a-def : Eq a
-1 := eq_refl; ..}}, should we support override "a"? If we do support it,
-how can our compiler detect that "a-def" is not inheritable and signal
-an error when such thing happens?
+Most programming experiences are preserved from Coq---each term inside the field definition is a well-typed Gallina term in vanilla Coq.
+Note that, a Family is not a first class value and thus we need to project out the Gallina terms from them.
 
-A similar example is about "subst" and "subst_lemma" in
-\cref{fig:STLC-example}. For "subst_lemma" to hold throughout two
-families, we shouldnot allow overriding on the definition of "subst".
-For example, we should still stick with the fact that "subst tunit x y =
-tunit", as this fact is used when proving "subst_lemma" and we want this
-lemma to be inherited to the children family.
+% Inheritance here
+\textbf{Inheritance on Fields.} "Family A extends B" has to appear at the top level (not inside any Family) and will start to construct the children "Family A" given the parent "Family B".
+Every field inside the inheriting family will be one of the three kinds of ``inheritance''---extending, overriding, and inheriting.
+The syntax of extending a new field is to directly use "Field", while overriding requires a prefix of "Override" before "Field f : .." and "Inherit f" will inherits field "f" from the parent.
 
-These two examples both show the incompatibility between exposing the
-concrete definition of a field for reasoning (this second example even
-involves the definition of fixpoint and pattern matching) and
-overridability. We need to be careful about this tension when designing
-the language.\YZ{
-  The two examples seem to suggest there should not be any overriding allowed at all.
-  Is there an example where you actually want the ability to override?
-}
+To resolve \ref{chg:consistency}, we make sure when inheriting or overriding on "f" happens, the context of the field "f" is also ``inherited''.
+Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty thus "self.g" doesn't type-check.
+Thus, before we inherit "g", the dependee "f" needs to be inherited.
 
-\textlabel{Challenge (3)}{chg:consistency}~\textbf{Consistency vs. Self Parametrization}.
-The meta\-theoretical formulation in~\citet{zm2017} for family
-polymorphism is very natural as they consider an extension of the family
-as the concatenation of a new linkage. However, this formulation is not
-directly adaptable for dependent type theory because it can easily
-encode non-termination sabotaging consistency by the presence of "self".
-Informally, in the orignal formulation, we can override the "f" in
-family $\{"f" := id; "g" := "call self.f"\}$ with "call self.g",
-resulting the mutual calling of "f" and "g" thus a deadloop. This
-conflict between self parametrization and consistency requests changing
-the original formulation of family and inheritance for dependent type
-theory.
+\begin{figure}[!htb]
+  \begin{minipage}[t]{0.45\linewidth}
+\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
+Family I.
+ Field a : nat := 1.
+ Field b : a = a := eq_refl.
+ Fail Field c : a = 1 := eq_refl. (*Fail 1*)
 
-\textlabel{Challenge (4)}{chg:software-engineering}~\textbf{Engineering and Ecosystem}.
-Not only for the metatheory, we also want our implementation of family
-to have good compatibility with the rest of Coq. For example, we
-shouldn't compromise \textbf{tactic programming} as it is part of the
-idiomatic Coq programming experience. Additionally, the interactive development
-style requires \textbf{incremental type checking}---our design has to be
-able to type check each Coq Vernacular command once inputted. What's
-more, we want our family ``term'' to be incorporated into Coq \textbf{without
-breaking its other functionality}. For example, the user may want to
-project the family member anywhere inside an arbitrary Gallina term,
-that may later be computed and normalized. However, families themselves
-can be second class citizens just like \citet{zm2017}.\YZ{I was thinking
-if the paper should start by talking about the functional encoding of
-inheritance (namely exposing a self parameter to allow late binding, and
-later taking the fixpoint of the extensible module when it needs to be
-accessed outside the family).  I think most readers will benefit from a
-recap of this encoding.\\
-Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
-}
+ Final Field a' : nat := 1.
+ Field c : a' = 1 := eq_refl.
+ Field bop : nat -> nat -> nat := (+).
+ Fail Field comm_bop : ...
 
-% \subsection{Design Space}
+ Family MonotonicF.
+  Field f : nat -> nat := non_impl.
+  Field p : monotonic f := non_impl.
+ SealFamily. 
+EndFamily.
+\end{lstlisting}
+  \end{minipage}
+\begin{minipage}[t]{0.45\linewidth}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+Family I2 extends I.
+ Override Field a : nat := 1. (* Override *)
+ Inherit b.                   (* Inherit  *)
 
-% When adapting Family Polymorphism into dependent types, we choose to
-% focus only on the essence of family (and inheritance)
-% structure in \citet{zm2017}, and thus a lot of unrelated features, like
-% Interface, will be removed. In this case, it will look like module with
-% late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
-% need it to have good compatibility with inductive types, because we
-% don't want to retain the idiomatic Coq programming experience
-% %\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness.
-I rephrase it now, please check.}
-% (and its tactic programming).
+ Field d : nat := a.          (* Extend   *)
+ Fail Override Field a' : nat := 2. (*Fail 2*)
 
-% Let's start with a basic example---STLC and its extension in
-% \cref{fig:STLC-example}---to consider what kind of features are required
-% and how much of them can be supported by family polymorphism.
-% add the example code here
+ Override Field bop := (-).
 
-% First and foremost, we don't want to throw away some basic good
-% properties of Coq. \textlabel{Req~(0A)}{chg:software-engineering}~\textbf{we
-% want to retain incremental (modular) type checking}. Notice that, the
-% modularity here in Coq is a bit different from other languages, because
-% Coq supports \textit{interactive} theorem proving, so we actually need
-% \textit{statement-wise} incremental type checking, not only for avoiding
-% re\-compilation, but also to enable immediate feedback and incremental
-% type-checking for the Coq developer. We don't want our family to ruin
-% this conventional routine of interactive development.
-% \textlabel{Req~(0B)}{chg:software-engineeringb}~\textbf{we also want to keep the
-% computational ability of Coq when using families}. Coq, based on constructive
-% logic, can consider proof as programs. 
-% We don't want our Family facility to ruin this when incorporated with 
-% the remaining parts of Coq: 
-% the developers should be able to project fields of families 
-% as first-class value just like how they can project fields of Coq's Module.
+ Inherits Until MonotonicF.
+ Override Family MonotonicF.
+  Field f : nat -> nat := (+ 1).
+  Field p : monotonic f := ...
+ EndFamily. 
+EndFamily.
+\end{lstlisting}
+  \end{minipage}
+  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
+\end{figure}
 
-% We want to reason about fields in a family.
-% \textlabel{Req (1)}{chg:definition-relevant-reasoning}:\textbf{ we want to be able to
-% reason about fields in a family}, just like in a module, where a field
-% can reason about its former fields. More generally, we want to allow a
-% later field to be type-dependent on the former fields. In the example of
-% \cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
-% this feature.
+% Talk about Final
+\textbf{Final as non-overridable.} Every field is by default overridable, reminding us \ref{chg:definition-relevant-reasoning}.
+To resolve it, we provide the keyword "Final" to allow the user to specify \textit{non-overridability} and give the user the opportunity to choose from overriding and exposing the definition.
 
-% we want extensible inductive type
-% \textlabel{Req (2)}{chg:extensible-inductive-type}:\textbf{ we need extensible
-% inductive types}, 
-% to extend "term" and "ty" in \cref{fig:STLC-example}.
-% Extensible inductive types are not supported by \citet{zm2017}, so
-% we need further consideration about it on both implementation and
-% metatheory.
+Take the field "c" of "Family I" in \cref{fig:plugin-example3} as the
+example. We can see that the first attempt of proving "a = 1" fails.
+This should fail---because we want to retain the possibility
+to override "a" and inheriting "c" in the future, but they are conflicting, since "c" witnesses "a" fix to "1".
+On the other hand, once we decorate it upon "a'", the second attempt \mintinline{Coq}{c : a' = 1 := eq_refl} will succeed.
+Contrary to "a", our "a'" is not overridable in "Family I2", mediating the \ref{chg:definition-relevant-reasoning}.
 
-% \textlabel{Req (3)}{chg:extensible-inductive-type}: \textbf{we also want extensible
-% pattern matching and induction reasoning}, to extend "subst" in
-% \cref{fig:STLC-example}.
-% There are actually two kinds of ``pattern matching'', one for data and
-% the other for induction reasoning, i.e., one uses the eliminator to
-% \mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the
-% eliminator to \mintinline{Coq}{Prop}. Luckily, in this setting,
-% extensible pattern matching can be easily expressed with family
-% extension---we just aggregate all case handlers of pattern matching
-% into one family (as a bunch of functions), and then one family can
-% encode one pattern matching, and family inheritance can express adding
-% case handlers. Then we just need to introduce a primitive that will
-% ``wrap'' that family into a recursive function. Induction reasoning can
-% be handled in an almost identical way.
+This example might sound trivial and contrived. Let's look at the
+example field binary operation "bop" initially set to be addition.\YZ{This 'add' field is no longer there.}\EDJreply{Thanks. Fixed} Now that we only expose the type of "bop", we
+cannot prove its commutativity in "comm_bop" (because we don't have
+access to its concrete definition) but we can override "bop" with
+subtraction function in "Family I2". This makes sense because we
+shouldn't (have the ability to) inherit commutativity property for subtraction
+anyway. This is contrary to how "a'" is declared "Final", the second "c"
+can prove that it's equal to 1 but "a'" is not overridable in the
+"Family I2".
 
-% we want to be able to reason pattern matching as well
-% However, there is still difference between data recursion and induction
-% reasoning, because the former one is \textit{proof-relevant}. This
-% difference leads to another issue: \textlabel{Req (4)}{chg:definition-relevant-reasoning}:
-% \textbf{we need to reason about (the computation about) extensible
-% pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
-% When proving "subst_lemma", we have to know "subst" is invariant on
-% "tt", i.e., \mintinline{Coq}{(subst i x tt) = tt}. This kind of
-% information requires exposing \textit{computational rules} from the
-% recursors.
+In our example, the seal family "I.MonotonicF" is actually simulating
+interfaces as well.  We have
+\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
+and we assigned it to both "f" and "p", and later instantiate in
+"I2" with concrete implementation by overriding this family
+"I.MonotonicF".
 
-% we want tactic programming and certain level of automation
-% Finally, \textlabel{Req (5)}{chg:software-engineering}: \textbf{we desire tactic
-% programming} to relieve us from the hassle of direct manipulation of proof terms.
+\textbf{Overridable/Sealed Family as better Sigma Type.} Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming.
+There are cases where overriding and
+(definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
+\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
+because to prove
+the latter we need to have concrete definition of "f" in the family type, which
+makes "f" non-overridable at all. One option is to use sigma type
+\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
+a bundle of stuff with multiple refinements.
+
+To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
+In a sealed family, every field will expose its definition inside the sealed family but a family as a whole will be ``sealed'' with a ``relaxed family type''---it is an aggregation of the types instead of the concrete definitions of all the field.
+Thus we can override one sealed family as a whole and change the internal implementation as long as they can satisfy the ``relaxed family type''.
+
+For example, the "MonotonicF" in \cref{fig:plugin-example3}:
+we simply seal
+\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
+together inside an sealed family.
+Because the definition of "f" is exposed inside the family, 
+its definition is
+transparent and can be used to prove "p".
+However, because the enclosing family of "f" is overridable, to swap the
+implementation of "f", we just override the whole ``sealed'' family,
+like we do in "I2.MonotonicF".
+
+% if the user wants a given field to be non-overridable. We will later elaborate its detail application.
+
+\textbf{Inheritance on Nested Families.} We also provide extending, inheriting and overriding for nested family (member).
+Extension is done by simply declaring a nested family; the same syntax for inheriting a field is used in inheriting a family.
+However, overriding is special---overriding can only apply to \textit{sealed} nested family in the parent family.
+A family is sealed when using "SealFamily" as the scoping instead of "EndFamily".
+We will later elaborate the motivation and details of sealed family.
+
+% Currently, every sealed family can only be direct inherited or overridden as a whole. It cannot be extended.
+
+\textbf{Extensible Inductive Type.} We provide the keyword "FInductive" to enable extensible inductive type.
+"Extend FInductive" is used to support extension of the parent inductive type inside a children family.
+The syntax of these two is close to the mundane "Inductive" keyword.
+
+To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness.
+The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family.
+Then
+"FRecursor" will be used to construct a recursor by refering to the handler family.
+\mintinline{Coq}{FRecursor recur about T motive f } "using"
+"handlers" will generate a new field "recur" that is really doing computation
+(using elimination principle of the inductive type "T") with motive "f". The handlers
+are located in the family "handlers", similar to a large pattern-matching statement.
+This way, the reuse and the extension of the case handlers will be delegated by the inheritance of the family.
+Once recursor is created it will look like any other function fields(e.g.
+"subst" in \cref{fig:STLC-example2}).
+
+"FTheorem" has a similar functionality but is expected to be used for theorem proving---the handler and the family will be automatically constructed during proving.
+
+% We can also use "Family" and "SealFamily" to scope a (nested) family.
+The resulting family only support direct inheritance or "Override" (i.e.
+we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:definition-relevant-reasoning}.
+What's more, "Override" family can only apply to sealed family as well (not allowed on normal family).
+
+% We can inherit a top-level family "parent" using \mintinline{Coq}{Family ... extends parent} as a new top-level family. As expected, "Override"
+% will override a field from the parent. "Inherit" can be used to inherit
+% nested families, inductive types, and fields.
+"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:definition-relevant-reasoning}.
+The fields inside a family
+% are order-sensitive because we have dependent types here. So "Inherit
+% Until" is used if the user wants to inherit a range of fields.  "Extend Family" is used only inside a inheriting family. Thus nested
+% family can only be inherited or extended together with its belonging family
+% (i.e. ultimately top level family), thus our support of family polymorphism is
+% incomplete.
+
+We will explain the usage of "Closing Fact" and "MetaData" later---they are here to resolve \ref{chg:software-engineering} to some extent.
+% This incomplete support can already be useful for extending functions.
+% Take \cref{fig:plugin-example2} as example, if we want to extend boolean
+% inductive type and its negation, we only need to specify the new
+% constructor for boolean and how negation deals with the new constructor.
+% As a result, the programmer does not have to duplicate code---otherwise
+% the user will need to copy and paste the whole original negation
+% definition and add a new pattern matching case.

### 2022-07-13 challenge.tex
Alter the simple negation example into the complicated STLC example In Coqimpl, implementation of extensible inductive type. Please check

\subsection{Challenges}

We want the metatheory mechanization to allow adding new
function/proving new lemmas, and adding new constructors for inductive
type at the same time  (Expression Problem~\cite{wadler-ep}). Coq
support the former but not the later---when an existent inductive type
is extended with new constructor, we have to modify the pattern matching
in the existent functions. This motivates us to add inheritance/Family
Polymorphism into Coq as a solution to the Expression Problem.

\begin{figure}[!htb]
  \begin{minipage}[t]{0.47\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC.
Inductive ty : Set :=
  | unit : ty
  | arrow : ty -> ty -> ty.
Inductive term : Set := 
  | tm_var : id -> term 
  | tm_abs : id -> term -> term 
  | tunit : term ...
Fixpoint subst 
  : id -> term -> term := ...
Inductive has_type 
  : context -> term -> type := ...
Proposition subst_lemma :
  has_type (G; x ↦ A) t T ->
  has_type G u A ->
  has_type G (subst x u t) T.
Inductive step : term -> term -> Prop 
  := ...
(* ... and more, end with type safety *)
EndFamily.
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{0.47\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC_bool extends STLC.
Extend Inductive ty : Set :=
  | bool : ty.

Extend Inductive term : Set := 
  | tt : term | ff : term 
  | tif : term -> term -> term -> term

Extend Fixpoint subst (* Need to handle new term *)
  : id -> term -> term := ...
Extend Inductive has_type (* .. and new ty too *)
  : context -> term -> type := ...
Extend Proposition subst_lemma :
  ... (* Need to prove extra cases *)

Extend Inductive step : term -> term -> Prop 
  := ... (* Need to expand this binary relation *)
(* ... and more extension *)
EndFamily.
\end{lstlisting}
  \end{minipage}
  \caption{Example STLC and its extension}\label{fig:STLC-example}
\end{figure}

Refering to the syntax of \citet{zm2017}, we will come up the above
example. However, to add inheritance into a dependent type system like
Coq is not simple and there are various challenges and tension we need
to resolve.

\textlabel{Challenge (1)}{chg:extensible-inductive-type}~\textbf{Extensible Inductive Type vs. Exhaustiveness of Inductive Reasoning}.
Unlike most general purposed programming languages, Coq, as a dependent
type theory, is a total language and thus each pattern matching cases
must fully ``cover'' an inductive type. This process is
``\textit{interactive}'' for inheritance---when inhertiance extend the
inductive type in the children family but the user forgets to extend the
corresponding inductive reasoning, our compiler need to signal an error
correspondingly, or better---guide the user to fill the incomplete
inductive reasoning. For example, in \cref{fig:STLC-example} when "term"
is extended with "tt" in the children family "STLC_bool", our compiler
need to make sure "subst" and "subst_lemma" are also extended correctly
otherwise signaling an error.
Additionally, our compiler must support direct inheritance (and
overriding) of the old case handlers for "tm_var", "tm_abs" and etc, to avoid
the boilerplate code.

\textlabel{Challenge (2)}{chg:definition-relevant-reasoning}~\textbf{Definition-awared Reasoning vs. Overridability}.
In family polymorphism for general purposed programming language,
overriding is allowed to happen anywhere. But the problem becomes more
subtle in the case of dependent type since the later fields can have a
type dependent on the former fields.  Consider a small example where we
have consecutive fields \mintinline{Coq}{{.. a : nat := 1; a-def : Eq a
1 := eq_refl; ..}}, should we support override "a"? If we do support it,
how can our compiler detect that "a-def" is not inheritable and signal
an error when such thing happens?

A similar example is about "subst" and "subst_lemma" in
\cref{fig:STLC-example}. For "subst_lemma" to hold throughout two
families, we shouldnot allow overriding on the definition of "subst".
For example, we should still stick with the fact that "subst tunit x y =
tunit", as this fact is used when proving "subst_lemma" and we want this
lemma to be inherited to the children family.

Note we cannot simply just abandon overriding because we do want it sometimes---say we want a monotonic function \mintinline{Coq}{{.. f := .. ; p : monotonic f; ..}}. The proof of "p" likely requires the concrete definition of "f" but in the future we might want to override "f" and "p" together to swap the definition.

These three examples both show the incompatibility between exposing the
concrete definition of a field for reasoning (this second example even
involves the definition of fixpoint and pattern matching) and
overridability. We need to be careful about this tension when designing
the language.\YZ{
  The two examples seem to suggest there should not be any overriding allowed at all.
  Is there an example where you actually want the ability to override?
}\EDJreply{Added a new example. Please check.}

\textlabel{Challenge (3)}{chg:consistency}~\textbf{Consistency vs. Self Parametrization}.
The meta\-theoretical formulation in~\citet{zm2017} for family
polymorphism is very natural as they consider an extension of the family
as the concatenation of a new linkage. However, this formulation is not
directly adaptable for dependent type theory because it can easily
encode non-termination sabotaging consistency by the presence of "self".
Informally, in the orignal formulation, we can override the "f" in
family $\{"f" := id; "g" := "call self.f"\}$ with "call self.g",
resulting the mutual calling of "f" and "g" thus a deadloop. This
conflict between self parametrization and consistency requests changing
the original formulation of family and inheritance for dependent type
theory.

\textlabel{Challenge (4)}{chg:software-engineering}~\textbf{Engineering and Ecosystem}.
Not only for the metatheory, we also want our implementation of family
to have good compatibility with the rest of Coq. For example, we
shouldn't compromise \textbf{tactic programming} as it is part of the
idiomatic Coq programming experience. Additionally, the interactive development
style requires \textbf{incremental type checking}---our design has to be
able to type check each Coq Vernacular command once inputted. What's
more, we want our family ``term'' to be incorporated into Coq \textbf{without
breaking its other functionality}. For example, the user may want to
project the family member anywhere inside an arbitrary Gallina term,
that may later be computed and normalized. However, families themselves
can be second class citizens just like \citet{zm2017}.\YZ{I was thinking
if the paper should start by talking about the functional encoding of
inheritance (namely exposing a self parameter to allow late binding, and
later taking the fixpoint of the extensible module when it needs to be
accessed outside the family).  I think most readers will benefit from a
recap of this encoding.\\
Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
}\EDJreply{Feel free to add these. I don't have many knowledge about this part.}

% \subsection{Design Space}

% When adapting Family Polymorphism into dependent types, we choose to
% focus only on the essence of family (and inheritance)
% structure in \citet{zm2017}, and thus a lot of unrelated features, like
% Interface, will be removed. In this case, it will look like module with
% late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
% need it to have good compatibility with inductive types, because we
% don't want to retain the idiomatic Coq programming experience
% %\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness. I rephrase it now, please check.}
% (and its tactic programming).

% Let's start with a basic example---STLC and its extension in
% \cref{fig:STLC-example}---to consider what kind of features are required
% and how much of them can be supported by family polymorphism.
% add the example code here

% First and foremost, we don't want to throw away some basic good
% properties of Coq. \textlabel{Req~(0A)}{chg:software-engineering}~\textbf{we
% want to retain incremental (modular) type checking}. Notice that, the
% modularity here in Coq is a bit different from other languages, because
% Coq supports \textit{interactive} theorem proving, so we actually need
% \textit{statement-wise} incremental type checking, not only for avoiding
% re\-compilation, but also to enable immediate feedback and incremental
% type-checking for the Coq developer. We don't want our family to ruin
% this conventional routine of interactive development.
% \textlabel{Req~(0B)}{chg:software-engineeringb}~\textbf{we also want to keep the
% computational ability of Coq when using families}. Coq, based on constructive
% logic, can consider proof as programs. 
% We don't want our Family facility to ruin this when incorporated with 
% the remaining parts of Coq: 
% the developers should be able to project fields of families 
% as first-class value just like how they can project fields of Coq's Module.

% We want to reason about fields in a family.
% \textlabel{Req (1)}{chg:definition-relevant-reasoning}:\textbf{ we want to be able to
% reason about fields in a family}, just like in a module, where a field
% can reason about its former fields. More generally, we want to allow a
% later field to be type-dependent on the former fields. In the example of
% \cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
% this feature.

% we want extensible inductive type
% \textlabel{Req (2)}{chg:extensible-inductive-type}:\textbf{ we need extensible
% inductive types}, 
% to extend "term" and "ty" in \cref{fig:STLC-example}.
% Extensible inductive types are not supported by \citet{zm2017}, so
% we need further consideration about it on both implementation and
% metatheory.

% \textlabel{Req (3)}{chg:extensible-inductive-type}: \textbf{we also want extensible
% pattern matching and induction reasoning}, to extend "subst" in
% \cref{fig:STLC-example}.
% There are actually two kinds of ``pattern matching'', one for data and
% the other for induction reasoning, i.e., one uses the eliminator to
% \mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the
% eliminator to \mintinline{Coq}{Prop}. Luckily, in this setting,
% extensible pattern matching can be easily expressed with family
% extension---we just aggregate all case handlers of pattern matching
% into one family (as a bunch of functions), and then one family can
% encode one pattern matching, and family inheritance can express adding
% case handlers. Then we just need to introduce a primitive that will
% ``wrap'' that family into a recursive function. Induction reasoning can
% be handled in an almost identical way.

% we want to be able to reason pattern matching as well
% However, there is still difference between data recursion and induction
% reasoning, because the former one is \textit{proof-relevant}. This
% difference leads to another issue: \textlabel{Req (4)}{chg:definition-relevant-reasoning}:
% \textbf{we need to reason about (the computation about) extensible
% pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
% When proving "subst_lemma", we have to know "subst" is invariant on
% "tt", i.e., \mintinline{Coq}{(subst i x tt) = tt}. This kind of
% information requires exposing \textit{computational rules} from the
% recursors.

% we want tactic programming and certain level of automation
% Finally, \textlabel{Req (5)}{chg:software-engineering}: \textbf{we desire tactic
% programming} to relieve us from the hassle of direct manipulation of proof terms.

### 2022-07-13 coqimpl.tex
Alter the simple negation example into the complicated STLC example In Coqimpl, implementation of extensible inductive type. Please check

@@
 
-\begin{figure}[!htb]\YZ{Change this one into STLC example, to see the layout}
+% \begin{figure}[!htb]\YZ{Change this one into STLC example, to see the layout}
+%   \begin{minipage}[t]{0.32\linewidth}
+% \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+% Family B.
+%   FInductive b : Set 
+%     := tt : b | ff : b.
+%   Family neg_handler. 
+%     Field tt : self_B.b 
+%       := self_B.ff.
+%     Field ff : self_B.b 
+%       := self_B.tt.
+%   EndFamily.
+%   FRecursor neg 
+%     about self_B.b 
+%     motive (fun _ => self_B.b)
+%     using self_B.neg_handler
+%     by _rec.
+%   Field example := 
+%     self_B.neg self_B.tt. 
+% EndFamily.
+% Family B2 extends B.
+%   Extend FInductive b : Set 
+%     := uu : b.
+%   Extend Family neg_handler.
+%     Field uu : self_B2.b 
+%       := self_B2.uu.
+%   EndFamily. 
+% EndFamily.
+% \end{lstlisting}
+%   \end{minipage}
+%   \begin{minipage}[t]{0.65\linewidth}
+% \begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
+%   (* Abstraction for Inductive Type *)
+% Module Type b_3 (self_B: EmptySig_4).
+% Parameter (b : Set). Parameter (tt ff : b). End b_3.      
+% Module Type b_6 := b_3 EmptyMod.
+%   (* Recursor to Set for b *)
+% Module b_rec_12 (self_B: b_6).
+% Definition __recursor_type_b_rec :=
+%   forall P : self_B.b -> Set,
+%   P self_B.tt -> P self_B.ff 
+%   -> forall __i : self_B.b, P __i. End b_rec_12. 
+%   (* Field B.neg_handler.ff *)
+% Module ff_24 (self_B: b_6) 
+%              (self_neg_handler: tt_23 self_B).
+% Definition ff : self_B.b := self_B.tt. End ff_24.
+%   (* Intermediate Module solely for type checking *)
+% Module v_33_34 (self_B: neg_handler_26).
+% Include b_rec_12 self_B.
+% Parameter (recursor_for_type_checking : __recursor_type_b_rec).
+% Definition term_for_type_checking :=
+%   recursor_for_type_checking (fun _ : self_B.b => self_B.b)
+%     self_B.neg_handler.tt
+%     self_B.neg_handler.ff.  
+% End v_33_34.
+% \end{minted}
+%   \end{minipage}
+% \caption{Example Code for Inductive Type}\label{fig:plugin-example2}
+% \end{figure}
+
+\begin{figure}[!htb]
   \begin{minipage}[t]{0.32\linewidth}
 \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family B.
-  FInductive b : Set 
-    := tt : b | ff : b.
-  Family neg_handler. 
-    Field tt : self_B.b 
-      := self_B.ff.
-    Field ff : self_B.b 
-      := self_B.tt.
+Family STLC.
+(* ... *)
+  Family subst_internal. 
+  Final Field tm_var : ...
+:= fun s x t => 
+    if (eqb x s) 
+    then t else (tm_var s).
+  Final Field tm_abs : ...
+:= fun s (b : tm)   
+   (recb : id -> tm -> tm) 
+   x t => 
+    if (eqb x s) 
+    then (tm_abs s body)
+    else (tm_abs s (recb x t)).
+(* ... *)
   EndFamily.
-  FRecursor neg 
-    about self_B.b 
-    motive (fun _ => self_B.b)
-    using self_B.neg_handler
+  FRecursor subst 
+    about tm 
+    motive (fun _ => id -> tm -> tm)
+    using subst_internal
     by _rec.
-  Field example := 
-    self_B.neg self_B.tt. 
+(*
+Field test := (subst (tm_var 0) 0 (tm_var 0)).
+*)
+(* ... *)
 EndFamily.
-Family B2 extends B.
-  Extend FInductive b : Set 
-    := uu : b.
-  Extend Family neg_handler.
-    Field uu : self_B2.b 
-      := self_B2.uu.
+Family STLC_bool extends STLC.
+(* ... *)
+Extend FInductive tm : Set :=
+  | tm_true : tm ...
+
+Extend Family subst_internal.
+Final Field tm_true : ...
+  := fun x t => tm_true.
+(* ... *)
 EndFamily. 
+Inherits subst.
+(* ... *)
 EndFamily.
@@
 \begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
-  (* Abstraction for Inductive Type *)
-Module Type b_3 (self_B: EmptySig_4).
-Parameter (b : Set). Parameter (tt ff : b). End b_3.      
-Module Type b_6 := b_3 EmptyMod.
-  (* Recursor to Set for b *)
-Module b_rec_12 (self_B: b_6).
-Definition __recursor_type_b_rec :=
-  forall P : self_B.b -> Set,
-  P self_B.tt -> P self_B.ff 
-  -> forall __i : self_B.b, P __i. End b_rec_12. 
-  (* Field B.neg_handler.ff *)
-Module ff_24 (self_B: b_6) 
-             (self_neg_handler: tt_23 self_B).
-Definition ff : self_B.b := self_B.tt. End ff_24.
+
+Module Type STLC_409.
+  Parameter (tm : Set).
+  Parameter (tm_var : id -> tm).
+  Parameter (tm_abs : id -> tm -> tm).
+  (* ... and all the fields defined before
+  and the abstraction for inductive type ...*)
+End STLC_409.
+(* Collect eliminator to Set for tm *)
+Module tm_rec_425 (STLC: STLC_409).
+  Definition __recursor_type_tm_rec :=
+    forall P : STLC.tm -> Set,
+    (forall n : id, P (STLC.tm_var n)) ->
+    (forall (n : id) (i : STLC.tm),
+    P i -> P (STLC.tm_abs n i)) ->
+    ... -> forall i : STLC.tm, P i.
+End tm_rec_425.
+(* Compiled Field STLC.subst_internal.tm_var *)
+Module tm_var_411 (STLC: STLC_409)
+  (subst_internal: EmptySig).
+  Definition tm_var :
+    forall (s : id) (x : id) (t : STLC.tm), STLC.tm :=
+    fun s x t => if eqb x s then t else STLC.var s.
+End tm_var_411. (* ... and more *)
+(* Aggregate into one subst_internal*)
+Module subst_internal_410 (STLC: STLC_409). 
+  Include tm_var_411 STLC. (* And other cases ... *)
+End subst_internal_410.
+Module v_33_34 (STLC: STLC_409).
 (* Intermediate Module solely for type checking *)
-Module v_33_34 (self_B: neg_handler_26).
-Include b_rec_12 self_B.
-Parameter (recursor_for_type_checking : __recursor_type_b_rec).
+  Include tm_rec_425 STLC.
+  Parameter (recursor_for_type_checking : __recursor_type_tm_rec).
   Definition term_for_type_checking :=
-  recursor_for_type_checking (fun _ : self_B.b => self_B.b)
-    self_B.neg_handler.tt
-    self_B.neg_handler.ff.  
+    recursor_for_type_checking (fun _  => id -> STLC.tm -> STLC.tm)
+      subst_internal_410.tm_var
+      subst_internal_410.tm_abs ... .
 End v_33_34.
@@
   \end{minipage}
-\caption{Example Code for Inductive Type}\label{fig:plugin-example2}
-\end{figure}
+\caption{Exempler STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
+\end{figure}\YZ{Change this one into STLC example, to see the layout}\EDJreply{Changed. I think it only make things worse.}
 
@@
 
-Recall, to construct a recursor, we need to create for each
-constructor a handler (function) and aggregate them into a family (e.g.
-"neg_handler" in \cref{fig:plugin-example2}).
+% Recall, to construct a recursor, we need to create for each
+% constructor a handler (function) and aggregate them into a family (e.g.
+"subst_internal" in \cref{fig:plugin-example2}).
 We use "FRecursor" to declare a recursor (e.g.
-"neg") by specifying the motive, the aiming inductive type and the handler family.
+"subst") by specifying the motive, the aiming inductive type and the handler family.
 Once recursor is created it will look like any other function fields (e.g.
-"neg").
-Here in \cref{fig:plugin-example2} we apply "neg" to "tt" as an exemplar field.
+"subst").
+Here in \cref{fig:plugin-example2} we can apply the function "subst" to "tm_var 0" as a new field.
 The inheritance of recursor is mainly delegated by the inheritance of the handler family---if the user extend with new constructor (e.g.
-"uu"), then the inherited family (e.g.
-"neg_handler") will need to extend correspondingly (e.g.
-"neg_handler.uu").
-If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".\EDJ{This paragraph is duplicating but I am not sure how to make it concise.}
+"tm_true"), then the inherited family (e.g.
+"subst_handler") will need to extend correspondingly (e.g.
+"subst_internal.tm_true").
+If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "subst".\EDJ{This paragraph is duplicating but I am not sure how to make it concise.}
 
-The plugin actually realizes the "Extend FInductive" via overridding---when extending an inductive type $A$
-with new constructors $c$, our plugin will generate a new (syntactic) inductive definition $A'$ with this new $c$ on the fly and feed it to Coq.
-But we must be careful on the exposing data in the family type (the compiled module type) for the following field.
-In other words, we must be careful on what interface (context) the following fields are defined based on.
-
-Let's look at the exemplar \cref{fig:plugin-example2}, the example of boolean and negation.
-When we define the extensible inductive type "b", all the following fields (e.g.
-"neg_handler.ff") can only know "b" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface''---the compiled module type "b_3".
+Let's look at the exemplar \cref{fig:plugin-example2}, the example of "subst" on STLC.
+When we define the extensible inductive type "tm", all the following fields (e.g.
+"subst_internal.tm_var") can only know "tm" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface''---the compiled module type "STLC_409".
 We cannot export\YZ{export or put?}\EDJreply{Ok.
@@
 I think I use ``export'' in several places...}
-the eliminator of the inductive type "b" to "b_3" because doing so all of the following fields (e.g.
-"neg_handler.ff" and the corresponding "ff_24") will not be able to be inherited to the context where "b"
-is extended with a third constructor, \textbf{since the mere existence of the recursor will assert that there are only two constructors} in "b", not
+the eliminator of the inductive type "tm" to "STLC_409" because doing so all of the following fields (e.g.
+"subst_internal.tm_var" and the corresponding "tm_var_411") will not be able to be inherited to the context where "tm"
+is extended with a third constructor, \textbf{since the mere existence of the recursor will assert that there are only two constructors} in "STLC_bool", not
 \textbf{extensible} at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}\EDJreply{You are right.
@@
 
-However, we later still need the concrete eliminator "b" for doing exhaustiveness checking (as in \ref{chg:extensible-inductive-type}), thus we will store the corresponding info somewhere.
-Here we show one exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module "b_rec_12".
+However, we later still need the concrete eliminator on "tm" for doing exhaustiveness checking (as in \ref{chg:extensible-inductive-type}), thus we will store the corresponding info somewhere.
+Here we show one exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module "tm_rec_425".
 This concrete eliminator will be used by our "FRecursor".
@@
 \begin{minted}{Coq}
-Definition neg := b_rec (λ _ ⇒ b) neg_handler.tt neg_handler.ff.
+Definition subst := tm_rec (λ _ ⇒ id -> tm -> tm) subst_internal.tm_var ...
 \end{minted}
-where "b_rec" and "b" are the corresponding recursor and the (vanilla) inductive type in that compiled module.
-Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields.
+where "tm_rec" and "tm" are the corresponding recursor and the (vanilla) inductive type in that compiled module.
+Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields in the programming interface.
+
+The plugin actually realizes the "Extend FInductive" via overridding---when extending an inductive type $A$
+with new constructors $c$, our plugin will generate a new (syntactic) inductive definition $A'$ with this new $c$ on the fly and feed it to Coq.
+But we must be careful on the exposing data in the family type (the compiled module type) for the following field.
+In other words, we must be careful on what interface (context) the following fields are defined based on.
 
 Importantly, the recursor field (e.g.
-"neg") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) modules, a recursor field is not compiled like that.
+"subst") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) modules, a recursor field is not compiled like that.
 For all the children family, we have to reconstruct and override the "FRecursor" field with a ``new'' recursor using (maybe the same) handler family.
 This is affordable since each aggregated recursor handler doesn't need
-to be rechecked---look at "B2.neg_handler", there is only one new
-handler for "uu" and the handlers (mundane fields) from the parent
-"B.neg_handler.tt,B.neg_handler.ff" have already been type-checked and
+to be rechecked---look at "STLC_bool.neg_handler", there are only  new
+handlers for "tm_true" and the handlers (mundane fields) from the parent
+"B.subst_internal.tm_var" have already been type-checked and
 inherited (since they only rely on ``the abstracted
-interface'' module "b_3"). 
+interface'' module "STLC_409"). 
 Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
-the module "ff_24"---doesn't require another type-check and is re-used during the type-checking and compilation of "B2.neg".
-For example, for the new "neg", when compiling family "B2" into a module, at the appropriate place our plugin will insert
+the module "tm_var_411"---doesn't require another type-check and is re-used during the type-checking and compilation of "STLC_bool.subst".
+For example, for the new "subst", when compiling family "STLC_bool" into a module, at the appropriate place our plugin will insert
 \begin{minted}{Coq}
-Definition neg := b_rec (λ _ ⇒ b) neg_handler.tt neg_handler.ff neg_handler.uu.
+Definition subst := tm_rec (λ _ ⇒ id -> tm -> tm) subst_internal.tm_var ...
 \end{minted}
-where "b_rec" and "b" are the corresponding recursor and the (vanilla but extended and modified) inductive type in that compiled module.
-Note that, even though "b_rec" and "b" has the same name as before, they are in the newly-compiled module so they refer to totally different things.
+where "tm_rec" and "tm" are the corresponding recursor and the (vanilla but extended and modified) inductive type in that compiled module.
+Note that, even though "tm_rec" and "tm" has the same name as before, they are in the newly-compiled module so they refer to totally different things.
 

### 2022-07-13 coqimpl.tex
Working on meta-theory rewriting now.

@@
 having boilerplate code, \textbf{semantically}.
-\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}
+\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}\EDJreply{Done}
 \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
@@
   EndMetaData.
+  (* ... STLC example continue ... *)
 EndFamily.
 \end{lstlisting}  
-\caption{Example for Global Reasoning and MetaData}\label{fig:plugin-example-global-reasoning-meta-data}
+\caption{Example for Global Reasoning and MetaData, using in STLC example}\label{fig:plugin-example-global-reasoning-meta-data}
 \end{figure}

### 2022-07-14 coqimpl.tex
Fix typo.

@@
 := fun s (b : tm)   
-   (recb : id -> tm -> tm) 
+   (recb : id → tm → tm) 
    x t => 
@@
     about tm 
-    motive (fun _ => id -> tm -> tm)
+    motive (fun _ => id → tm → tm)
     using subst_internal
@@
   Parameter (tm : Set).
-  Parameter (tm_var : id -> tm).
-  Parameter (tm_abs : id -> tm -> tm).
+  Parameter (tm_var : id → tm).
+  Parameter (tm_abs : id → tm → tm).
   (* ... and all the fields defined before
@@
     forall (s : id) (x : id) (t : STLC.tm), STLC.tm :=
-    fun s x t => if eqb x s then t else STLC.var s.
+    fun s x t ⇒ if eqb x s then t else STLC.var s.
 End tm_var_411. (* ... and more *)
@@
   Definition term_for_type_checking :=
-    recursor_for_type_checking (fun _  => id -> STLC.tm -> STLC.tm)
+    recursor_for_type_checking (fun _  ⇒ id → STLC.tm → STLC.tm)
       subst_internal_410.tm_var
@@
 \begin{minted}{Coq}
-Definition subst := tm_rec (λ _ ⇒ id -> tm -> tm) subst_internal.tm_var ...
+Definition subst := tm_rec (λ _ ⇒ id → tm → tm) subst_internal.tm_var ...
 \end{minted}
@@
 \begin{minted}[escapeinside=@@]{Coq}
+(* Inside Family B *)
+FInductive b : Set := tt : b | ff : b.
 b_prec : forall R, option R -> option R -> b -> option R.@\YZ{Not clear what it takes to define/prove partial recursors.
@@
 inductive type, and use the original "injection" and "discriminate"
-tactics. For example, we want to prove \mintinline{Coq}{self_B.tt ≠
-self_B.ff}, we simply 
+tactics. For example, we want to prove \mintinline{Coq}{B.tt ≠
+B.ff}, we simply 
 \begin{minted}{Coq}
-Definition reflect : self_B.b → option bool 
+Definition reflect : B.b → option bool 
                    := b_prec (fun _ => bool) (Some true) (Some false).
-Definition tt_neq_ff : self_B.tt = self_B.ff → False.
-(* Because self_B.tt = self_B.ff → reflect self_B.tt = reflect self_B.ff 
+Definition tt_neq_ff : B.tt = B.ff → False.
+(* Because B.tt = B.ff → reflect B.tt = reflect B.ff 
       → Some true = Some false    (by computational axiom)

### 2022-07-16 metatheory.tex
Adding comments.
Save, rewriting meta-theory

--- 2022-06-29 metatheory.tex
@@
 % Motivation of using QIIT
-We start with the formulation of the syntax, which is based on the
+We start with the formulation of the syntax in the declarative style, which is based on the
 formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT)
-given by~\citet{coquand2018canonicity}. Thus we will omit a great deal
-of details about MLTT here\YZ{What details are omitted?} but focus on the new facility we introduced.
+given by~\citet{coquand2018canonicity}. 
+% Thus we will omit a great deal of details about MLTT here
+\YZ{What details are omitted?} 
+We will mainly focus on the new facility that we introduced
 
 Though phrased in Latex, we are following~\citet{altkap2016} to use as
-our meta-logic a type theoretical framework (like Agda), 
+our meta-logic an extensional type theoretical framework (with Agda-like syntax), 
 together with \textit{Quotient Inductive Inductive Type} (QIIT).
 Our work can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}.
-We base our formulation of the syntax in this type theory instead of a more conventional and informal representation, becase (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out.
+We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\citet{gratzer-sterling-birkedal:2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the judgemental rules in QIIT.
 
@@
 
-\subsection{Syntax}
+\subsection{Syntax}\YZ{Rewrite syntax part into name-variable style instead of currently debruijn indices, and do that for all the models we introduced}
 

### 2022-07-16 syntax.tex
Save, rewriting meta-theory

@@
+\subsubsection{Warm Up}
+\textit{Quotient Inductive Inductive Type}~(QIIT) is an extension of \textit{inductive-inductive type} (IIT)~\cite{nordvall2010inductive}.
+IIT can be roughly considered as several inductive type mutually defined, in a more expressible way than mutual-inductive type.
+QIIT is allowing to have \textit{quotient}, in other words, equality results as part of the constructors.
+For example, in \mintinline{Agda}{data N2: Set where O : N2 | S : N2 → N2 | B : S (S x) ≡ x }, "B" will be a valid (equality) constructor, and "N2" is a set of two elements in our extensional setting.
+
+We highlight some special features of declarative style in QIIT that differ from a
+conventional setting: 
+(1) This declarative formulation of type theory does not
+have an operational semantics but only judgemental equality---$\beta$-reduction is now $\beta$-equality.\footnote{Later we will prove
+canonicity which can imply a certain level of computational ability of
+our theory} 
+(2) We directly use the equality from QIIT to represent judgemental equality in our syntax, relieve from us the burden to explicit mentioning equivalence closure; however, this also mean that, \textit{every piece of the syntax data}, including all kinds of judgements (QIIT type) and their derivations will have equality information upon them
+%We will recover canonicity afterwards, which implies the computational ability of our theory. 
+(3) We are dealing with quotient data, so we are actually manipulating
+\textit{equivalence classes of terms} instead of one representative (term) from the equivalence class.
+(4) Instead of using meta-level substitution, we use explicit
+substitution and de Bruijn indices in our intrinsic formulation. This is
+also called substitution calculus in the literature and is favoured due
+to its closedness to categorical semantic. We will go through these four details one-by-one.
+
+\textbf{Judgements}
 We formalize our language as a deep embedding in a type theory with QIIT types.
@@
 Latex-wise, we spell out these \textit{well-formedness rules} of type
-judgements and their Agda counterparts to make things more accessible. Of
+judgements and their (pseudo-)Agda counterparts to make things more accessible. Of
 course, a well-formed judgement is not necessarily derivable, just as
@@
 quotient data-type).
-
-Now we highlight some other special features that differ from a more
-conventional setting: (1) This modern formulation of type theory does not
-have an operational semantics but only equality.\footnote{Later we will prove
-canonicity which can imply a certain level of computational ability of
-our theory}
-%We will recover canonicity afterwards, which implies the computational ability of our theory. 
-(2) Instead of using meta-level substitution, we use explicit
-substitution and de Bruijn indices in our intrinsic formulation. This is
-also called substitution calculus in the literature and is favoured due
-to its closedness to categorical semantic.
-(3) We are dealing with quotient data, so we are actually manipulating
-equivalence classes of terms instead of one representative (term) from the equivalence class.
 

### 2022-07-16 metatheory.tex
Save.

@@
 Our work can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}.
-We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\citet{gratzer-sterling-birkedal:2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the judgemental rules in QIIT.
+We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\cite{gratzer-sterling-birkedal-2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the judgemental rules in QIIT.
 
@@
 
-\subsection{Syntax}\YZ{Rewrite syntax part into name-variable style instead of currently debruijn indices, and do that for all the models we introduced}
+\subsection{Syntax in Declarative Style}
+% \YZ{Rewrite syntax part into name-variable style instead of currently debruijn indices, and do that for all the models we introduced}
 

### 2022-07-16 syntax.tex
Save.

@@
-\subsubsection{Warm Up}
+\subsubsection{\textbf{Warm Up}}
 \textit{Quotient Inductive Inductive Type}~(QIIT) is an extension of \textit{inductive-inductive type} (IIT)~\cite{nordvall2010inductive}.
@@
 
-We highlight some special features of declarative style in QIIT that differ from a
+We highlight some specialities of declarative style in QIIT that differ from a
 conventional setting: 
@@
 %We will recover canonicity afterwards, which implies the computational ability of our theory. 
-(3) We are dealing with quotient data, so we are actually manipulating
-\textit{equivalence classes of terms} instead of one representative (term) from the equivalence class.
-(4) Instead of using meta-level substitution, we use explicit
+% (3) We are dealing with quotient data, so we are actually manipulating
+% \textit{equivalence classes of terms} instead of one representative (term) from the equivalence class.
+(3) Instead of using meta-level substitution, we use explicit
 substitution and de Bruijn indices in our intrinsic formulation. This is
 also called substitution calculus in the literature and is favoured due
-to its closedness to categorical semantic. We will go through these four details one-by-one.
+to its closedness to categorical semantic. We will go through these details one-by-one.
 
-\textbf{Judgements}
+\underline{\textit{\textbf{Judgements and Universe.}}}
+% Judgements in QIIT
 We formalize our language as a deep embedding in a type theory with QIIT types.
 Every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms.
-(Otherwise how can they be encoded in Agda?) There are four standard judgement forms in MLTT,
+(Otherwise how can they be encoded in Agda?) There are four standard judgement forms from (vanilla) MLTT,
 well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
@@
 course, a well-formed judgement is not necessarily derivable, just as
-a well-formed QIIT type is not necessarily inhabited. These well-formedness rules are direct translations of the \textit{well-typedness} of QIIT.
+a well-formed QIIT type is not necessarily inhabited. 
+These well-formedness rules are direct translations of the \textit{well-typedness} of QIIT.
 
-Due to our intrinsic setting, \textbf{we can and will omit a lot of premises of the typing rules}. This can work because we are using
-QIIT and everything is intrinsically well-formed(well-typed). For example, we don't
-spell out that the rule \ruleref{Type Universe} requires a well-formed
-context, even though it does implicitly, \textbf{because it is not
+% omitting presupposition
+Due to our intrinsic setting, \textbf{we can and will omit a lot of premises of the typing rules}~(the presuppositions). 
+For example, we don't explicitly
+spell out that the rule \ruleref{Type Universe} (that is stating a universe is a proper type) requires a well-formed
+context, even though it does implicitly inferrable, \textbf{because it is not
 possible to have ill-formed(ill-typed) contexts in our intrinsic setting}. This is
-a bit unusual for conventional extrinsic setting where they will put
-well-formedness of context everywhere in a lot of term judgements. We
+a bit unusual for conventional extrinsic setting where they will have
+well-formedness of context everywhere in most of the term judgements. We
 consider this as one advantage of using intrinsic setting---this
@@
 inferrable and can be considered as ``boilerplate code''. Free of
-such boilerplate code, our intrinsic setting provides a more concise
+such boilerplate code, the intrinsic setting provides a more concise
 representation.
 
+% How judgemental equality is used
 Quotient is used to represent definitional equality (aka judgemental
@@
 This way, for all four kinds of judgements we are
-using, there is always equality among these data. Doing so, we can omit
-some obvious equality judgements. Most of the reasoning can be formulated
-using the \textit{algebra} of QIIT (mapping out function from this
-quotient data-type).
+using, there is always equality among these data but only those equalities between terms and types are emphasized due to their non-triviality.
 
+% Example : how judgemental equality is used for the universe
+We review the concept of ``Universe'' and at the same time showing how equality is representing semantic.
+In dependent type setting, function now can return type, of which the type is ``type of type'', i.e., \textit{the universe} $\cU$.
+We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
+distinguish types from their \textit{codes or names} as showin in \ruleref{Codes/Get Type}: given a \textit{code} (actually a term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
+$\goodType{\Gamma}{El~T}{}$, and vice-versa (given a
+
+In other words, (1) a universe $\cU$ is a type including the \textit{codes or names} of types.
+(2) a code (term) $\goodTerm{\Gamma}{T}{\cU}$ is not allowed to locate at type position
+(i.e., at the rightmost position of a term judgement) but $El \ T$ is (i.e. we distinguish the code and the type). (3)
+"c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of a
+type, with the following \textit{two (omitted) equations representing their semantic} $"c" (El~T) \equiv T$ and $El
+("c"~T) \equiv T$---notice how the two sides of each equality is term of the same type thus well-typed. More concretely,
+
+Note that we are actually using predicative universe, so we actually
+have different levels of universes $\cU_j$. However, we omit universe levels in our latex presentation for conciseness and accessibility.
+% Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).
+
+\underline{\textit{\textbf{Substitution Calculus.}}}
 We quickly review the substitution calculus.
-Substitution is very (if not the most) important because function is
-first-class citizens, and can be applied (almost) everywhere and can return
+Substitution is very (if not the most) important because (dependent) functions are
+first-class citizens, and lambdas can wrap (almost) every parts and can return
 (almost) everything, thus we require \textbf{the resulting substitution to
@@
 
+\subsubsection{\textbf{Core Syntax of the Extended MLTT}}
 The core language builds on MLTT and adds four new judgement forms.
@@
 
-Because in dependent type, function now can return type, of which the type is ``type of type'', \textit{the universe} $\cU$.
-We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
-distinguish types from their \textit{codes or names}, and thus we use
-$El\ T$ as the type given \textit{type name} $T : \cU$.
-In this case, (1) $\cU$ is the type including the \textit{codes or
-names} of types, and (2) $T$ is not allowed to locate at type position
-(i.e., at the rightmost position of a term judgement) but $El \ T$ is.
-"c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of a
-type, with the following two equations $"c" (El~T) \equiv T$ and $El
-("c"~T) \equiv T$.
-Recall that we are actually using predicative universe, so we actually
-have different levels of universes $\cU_j$.
-We omit universe levels for conciseness.
-
   \label{fig:rules:well-typed-ctx}
@@
 $$
-\Rule[]
+\Rule[name=Code/Get Type]
 {\goodType{\Gamma}{T}{j}}

### 2022-07-16 metatheory.tex
Save

@@
 We start with the formulation of the syntax in the declarative style, which is based on the
-formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT)
+formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT)~\cite{martin1982constructive}
 given by~\citet{coquand2018canonicity}. 
@@
 together with \textit{Quotient Inductive Inductive Type} (QIIT).
-Our work can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}.
-We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a proof because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real theorem prover so that the reader can have an idea about how mechanized formalization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\cite{gratzer-sterling-birkedal-2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the judgemental rules in QIIT.
+
+\textit{Quotient Inductive Inductive Type}~(QIIT) is an extension of \textit{inductive-inductive type} (IIT)~\cite{nordvall2010inductive}.
+IIT can be roughly considered as several inductive type mutually defined, in a more expressible way than mutual-inductive type.
+QIIT is allowing to have \textit{quotient}, in other words, equality results as part of the constructors.
+For example, in \\ \mintinline{Agda}{data N2: Set where O : N2 | S : N2 → N2 | B : S (S x) ≡ x }, the "B" here is a valid (equality) constructor, and "N2" is a set of two elements in our extensional setting.
+
+Our formulation can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}.
+We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a pen-and-paper proof now because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real mechanized one so that the reader can have an idea about how mechanization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\cite{gratzer-sterling-birkedal-2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the judgemental rules in QIIT.
 

### 2022-07-16 syntax.tex
Save

@@
-\subsubsection{\textbf{Warm Up}}
-\textit{Quotient Inductive Inductive Type}~(QIIT) is an extension of \textit{inductive-inductive type} (IIT)~\cite{nordvall2010inductive}.
-IIT can be roughly considered as several inductive type mutually defined, in a more expressible way than mutual-inductive type.
-QIIT is allowing to have \textit{quotient}, in other words, equality results as part of the constructors.
-For example, in \mintinline{Agda}{data N2: Set where O : N2 | S : N2 → N2 | B : S (S x) ≡ x }, "B" will be a valid (equality) constructor, and "N2" is a set of two elements in our extensional setting.
+\subsubsection{\textbf{Warm Up: Review Declarative MLTT}}
 
@@
 (1) This declarative formulation of type theory does not
-have an operational semantics but only judgemental equality---$\beta$-reduction is now $\beta$-equality.\footnote{Later we will prove
+have an operational semantics but only judgemental equality (e.g. $\beta$-reduction is now $\beta$-equality).\footnote{Later we will prove
 canonicity which can imply a certain level of computational ability of
@@
 substitution and de Bruijn indices in our intrinsic formulation. This is
-also called substitution calculus in the literature and is favoured due
+also called \textit{substitution calculus} in the literature and is favoured due
 to its closedness to categorical semantic. We will go through these details one-by-one.
 
+\newcommand\mathboxtext[1]{
+  \fcolorbox{black}{faint-gray}{\ensuremath{#1}}
+}
+
+\begin{figure}[!htb]
+  \begin{align*}
+    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\gamma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\gamma}{\Delta}}  && t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
+  \end{align*}
+\caption{Judgement forms, as Agda types and as presented in this section.}\label{fig:vanillaMLTTjudgeQIIT}
+\end{figure}
+\begin{figure}[!htb]
+  \begin{minipage}[b]{0.3\linewidth}
+      $$
+      \Rule[name=Tm]
+      {\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{T}{j}}
+      {\goodTerm{\Gamma}{\_}{T}}
+      $$
+      $$
+      \Rule[name=Tms]
+      {\goodCtx{\Gamma}{i} \quad \goodCtx{\Delta}{i}}
+      {\goodSub{\Gamma}{\_}{\Delta} \text{ is well-formed}}
+      $$
+  \end{minipage}
+  \begin{minipage}[b]{0.6\linewidth}
+    \begin{minted}[]{agda}
+      data Con   : Set 
+      data Ty    : Con → Set   
+      data Tms   : Con → Con → Set 
+      data Tm    : (Γ : Con) → Ty Γ  → Set 
+    \end{minted}
+  \end{minipage}
+
+\caption{Well-formedness rules of the judgements}\label{fig:vanillaMLTTjudgewellformed}
+\end{figure}
 \underline{\textit{\textbf{Judgements and Universe.}}}
 % Judgements in QIIT
-We formalize our language as a deep embedding in a type theory with QIIT types.
+We formalize our language as a deep embedding in a type theory using QIIT types.
 Every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms.
@@
 $\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
-well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. These four judgements (QIIT types) are
-mutually recursively defined together. Our presentation
+well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. The QIITs corresponding of these four judgements are
+mutually recursively defined together, and declared in \cref{fig:vanillaMLTTjudgeQIIT}. Our presentation
 follows the style of~\citet{kaposi2019gluing} and thus a predicative MLTT, but for readability we
@@
 Please check}
-Latex-wise, we spell out these \textit{well-formedness rules} of type
-judgements and their (pseudo-)Agda counterparts to make things more accessible. Of
+Latex-wise, in \cref{fig:vanillaMLTTjudgewellformed}, we spell out these \textit{well-formedness rules} of type
+judgements and their (pseudo-)Agda counterparts to make things more accessible.
+In (pseudo-)Agda, well-formedness rules are just well-typedness of (indexed) QIIT family---(the judgement)"Tm~$\Gamma$~$T$" is well-typed (well-formed) only when $\Gamma$ and $T$ are both well-typed (well-formed).
+We omitted a bunch for latex version.
+Of
 course, a well-formed judgement is not necessarily derivable, just as
@@
 We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
-distinguish types from their \textit{codes or names} as showin in \ruleref{Codes/Get Type}: given a \textit{code} (actually a term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
-$\goodType{\Gamma}{El~T}{}$, and vice-versa (given a
+distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
+$\goodType{\Gamma}{El~T}{}$, and vice-versa.
 
@@
 type, with the following \textit{two (omitted) equations representing their semantic} $"c" (El~T) \equiv T$ and $El
-("c"~T) \equiv T$---notice how the two sides of each equality is term of the same type thus well-typed. More concretely,
+("c"~T) \equiv T$---notice how the two sides of each equality is term of the same type thus a well-typed equality. More concretely,
 
@@
 \underline{\textit{\textbf{Substitution Calculus.}}}
-We quickly review the substitution calculus.
+Now we review the substitution calculus.
 Substitution is very (if not the most) important because (dependent) functions are
 first-class citizens, and lambdas can wrap (almost) every parts and can return
-(almost) everything, thus we require \textbf{the resulting substitution to
-commute} with most (if not all) type/term constructs.
+(almost) everything---once we apply the functions,  \textbf{the substitution will happen and needs to
+commute} with most all type/term constructs in the function body.
 
@@
 \subsubsection{\textbf{Core Syntax of the Extended MLTT}}
-The core language builds on MLTT and adds four new judgement forms.
-Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
-shorthand for term judgement
-$\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
-
-\newcommand\mathboxtext[1]{
-  \fcolorbox{black}{faint-gray}{\ensuremath{#1}}
-}
-
 \begin{figure}[!htb]
   \begin{align*}
-    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\gamma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\gamma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}}
-    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
+&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
 Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
@@
   \end{align*}
-\caption{Judgement forms, as Agda types and as presented in this section.}
+\caption{Judgement forms, as Agda types, continued}
 \end{figure}
-
 \begin{figure}[!htb]
   \begin{minipage}[b]{0.3\linewidth}
-      $$
-      \Rule[name=Tm]
-      {\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{T}{j}}
-      {\goodTerm{\Gamma}{\_}{T}}
-      $$
-      $$
-      \Rule[name=Tms]
-      {\goodCtx{\Gamma}{i} \quad \goodCtx{\Delta}{i}}
-      {\goodSub{\Gamma}{\_}{\Delta} \text{ is well-formed}}
-      $$
       $$
@@
     \begin{minted}[]{agda}
-      data Con   : Set 
-      data Ty    : Con → Set   
-      data Tms   : Con → Con → Set 
-      data Tm    : (Γ : Con) → Ty Γ  → Set 
       data Sig  : Con → ℕ → Set
@@
 
-\caption{Well-formedness rules of the judgements}
+\caption{Well-formedness rules of the extended judgements}
 \end{figure}
+
+The core language builds on MLTT and adds four new judgement forms.
+Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
+shorthand for term judgement
+$\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
 

### 2022-07-16 syntax.tex
Warmup section rewritten Different topics for Warmup is stratified for accessibility make first four judgements into warmup section

@@
 also called \textit{substitution calculus} in the literature and is favoured due
-to its closedness to categorical semantic. We will go through these details one-by-one.
+to its closedness to categorical semantic and easir to handle in mechanization. We will go through these details one-by-one.
 
@@
 For example, we don't explicitly
-spell out that the rule \ruleref{Type Universe} (that is stating a universe is a proper type) requires a well-formed
+spell out that the rule \ruleref{Boolean} (that is stating a boolean is a proper type) requires a well-formed
 context, even though it does implicitly inferrable, \textbf{because it is not
@@
 % Example : how judgemental equality is used for the universe
-We review the concept of ``Universe'' and at the same time showing how equality is representing semantic.
+We review the concept of ``universe'' and at the same time showing how equality is representing semantic.
 In dependent type setting, function now can return type, of which the type is ``type of type'', i.e., \textit{the universe} $\cU$.
@@
 "c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of a
-type, with the following \textit{two (omitted) equations representing their semantic} $"c" (El~T) \equiv T$ and $El
-("c"~T) \equiv T$---notice how the two sides of each equality is term of the same type thus a well-typed equality. More concretely,
+type. More concretely, using QIIT, the corresponding constructors will look like, (notice that the type on both sides of the equalities are always the same and thus a well-typed equality) 
+\begin{align*}
+  &\mintinline{Agda}{c : Ty Γ → Tm Γ U} &&\mintinline{Agda}{El : (a : Tm Γ U) → Ty } \\
+  &\mintinline{Agda}{Elc : El (c T) ≡ T} && \mintinline{Agda}{cEl : c (El T) ≡ T}
+\end{align*}
+These two equations indicate how \textit{dynamic} should react when $El~"c"~T$ appears. Later we will see other similar $\beta$-equations but we won't display their (pseudo-)Agda formulation.
 
@@
 % Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).
-
-\underline{\textit{\textbf{Substitution Calculus.}}}
-Now we review the substitution calculus.
-Substitution is very (if not the most) important because (dependent) functions are
-first-class citizens, and lambdas can wrap (almost) every parts and can return
-(almost) everything---once we apply the functions,  \textbf{the substitution will happen and needs to
-commute} with most all type/term constructs in the function body.
-
-A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
-type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
-A substitution $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
-understood in the way that $\gamma$ will output a \textbf{series} of
-(dependently-typed related) terms of types $\Delta$, parametrized by
-$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
-Thus (1) \ruleref{Sub Extend} looks like concatenating one
-series of terms with a new term (the weird looking type of
-$\goodTerm{\Gamma}{t}{A[\sigma]}$ in the premise is due to our need of type $A$ in
-the context $\Delta$ so that we can use \ruleref{Context Extension} to
-create new context); (2) we have \ruleref{Proj Subst-1} to extract a
-subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
-substitution really looks like a (deeply-nested) pair.
-
-Since the $\Gamma$ before the $\vdash$ is always indicating the types of
-``holes'', this explains the \ruleref{Term Subst} and \ruleref{Type
-Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution
-$\goodSub{\Gamma}{\gamma}{\Delta}$, we will have a new term
-$\goodTerm{\Gamma}{t[\gamma]}{A[\gamma]}$ after the substitution---this
-substitution can be understood as ``filling all the holes''. Once all
-the holes are filled with a given substitution, the new holes all come
-from the substitution itself.
-Consecutive substitutions can be composed using \ruleref{Sub Comp}, because
-all the holes are always from the last substitutions.
-
-Note that, for simplicity, we will often write $\pi_1$ as a
-shorthand for ${\pi_1~"id"}$, typed as
-$\goodSub{\Gamma,A}{\pi_1~"id"}{\Gamma}$, and $\pi_2$ as a shorthand
-for $\pi_2~"id"$, typed as $\goodTerm{\Gamma, A}{\pi_2~"id"}{A[\pi_1]}$.
-
-\subsubsection{\textbf{Core Syntax of the Extended MLTT}}
-\begin{figure}[!htb]
-  \begin{align*}
-&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
-Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
-I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
-  \end{align*}
-\caption{Judgement forms, as Agda types, continued}
-\end{figure}
-\begin{figure}[!htb]
-  \begin{minipage}[b]{0.3\linewidth}
-      $$
-      \Rule[name=Sig]
-      {\goodCtx{\Gamma}{i} \quad n \in \nat}
-      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
-      $$
-      $$
-      \Rule[name=Seal]
-      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
-      \quad \goodType{\Gamma}{A}{} }
-      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
-      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments?
-Besides, are these rules used anywhere in the following?
-Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes.
-Right column show all the well-formedness rule for all the judgement.
-That is unusual because it is a direct translation from QIIT.
-}
-  \end{minipage}
-  \begin{minipage}[b]{0.6\linewidth}
-    \begin{minted}[]{agda}
-      data Sig  : Con → ℕ → Set
-      data WSig : Con → ℕ → Set 
-      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
-      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
-    \end{minted}
-  \end{minipage}
-
-\caption{Well-formedness rules of the extended judgements}
-\end{figure}
-
-The core language builds on MLTT and adds four new judgement forms.
-Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
-shorthand for term judgement
-$\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
-
-% \begin{minted}[texcomments]{Coq}
-%   Γ : Con  (* $\goodCtx{\Gamma}{}$ *) 
-% \end{minted}
-% T : Ty Γ (* $\goodType{\Gamma}{T}{}$ *)       σ : Tms Γ Δ {- $\goodSub{\Gamma}{\sigma}{\Delta}$ -}        t : Tm Γ T {- $\goodTerm{\Gamma}{t}{T}$ -}
-
-We will have dependent function types
-and dependent pair types. Dependent pair type is used to model Coq
-module, which is useful when we need to model the ``compilation'' from
-linkages to modules. We also have $\beta,\eta$-rule for both dependent
-function type and dependent pair type.
-The substitution law for these two
-require \textit{de Bruijn lifting}: for a given substitution
-$\goodTerm{\Gamma}{\sigma}{\Delta}$ and a type
-$\goodType{\Delta}{A}{}$, we define its de Bruijn lifting
-$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ . Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.
 
@@
 \quad 
-\Rule
-{}
-{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
-\quad 
 \Rule[name=Term Subst]
@@
 \Rule[]
+{\goodTerm{\Gamma}{t}{\Pi A B}}
+{\goodTerm{\Gamma, A}{"app"~t}{B}}
+\quad 
+\Rule[]
 {\goodTerm{\Gamma}{u}{A} 
@@
 {\goodTerm{\Gamma}{"pjl" \ t}{A}
-\quad \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
+\\\\ \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
 }
@@
 $$
-\Rule[name=Base Type/Term Subst]
+\Rule[name=Bool]
+{}
+{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
+\quad
+\Rule[]
 {\goodSub{\Gamma}{\gamma}{\Delta}}
@@
 
-We have shown some exemplar rules for all four standard types of
-judgments. Now we focus on the newly introduced facility. We have two kinds of signatures, one for inductive types, the other for linkages.
+\underline{\textit{\textbf{Substitution Calculus.}}}
+We first review the substitution calculus.
+Substitution is very (if not the most) important because (dependent) functions are
+first-class citizens, and lambdas can wrap all possible terms and once we apply the functions,  \textbf{the substitution will happen on the wrapped body and the substitution needs to
+commute} with all type/term constructs in the function body.
+However, our substitution \textbf{is part of the syntax} instead of a meta-operation in conventional setting.
+We review the key to substitution calculus, mainly surrounding the judgements $\goodSub{\Gamma}{\gamma}{\Delta},\goodTerm{\Gamma}{t}{T},\goodType{\Gamma}{T}{}$.
+
+A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
+type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
+A \textbf{simultaneous substitution} $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
+understood in the way that $\gamma$ will output a \textbf{series} of
+(dependently-typed related) terms of types $\Delta$, parametrized by
+$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
+Thus (1) \ruleref{Sub Extend} looks like concatenating one
+series of terms with a new term (the weird looking type of
+$\goodTerm{\Gamma}{t}{A[\sigma]}$ in the premise is due to our need of type $A$ in
+the context $\Delta$ so that we can use \ruleref{Context Extension} to
+create new context); (2) we have \ruleref{Proj Subst-1} to extract a
+subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
+substitution really looks like a (deeply-nested) pair.
+
+Since the $\Gamma$ before the $\vdash$ is always indicating the types of
+``holes'', this explains the \ruleref{Term Subst} and \ruleref{Type
+Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution
+$\goodSub{\Gamma}{\gamma}{\Delta}$, we will have a new term
+$\goodTerm{\Gamma}{t[\gamma]}{A[\gamma]}$ after the substitution---this
+substitution can be understood as ``filling all the holes''. Once all
+the holes are filled with a given substitution, the new holes all come
+from the substitution itself.
+Consecutive substitutions can be composed using \ruleref{Sub Comp}, because
+all the holes are always from the last substitutions.
+
+Note that, for simplicity, we will often write $\pi_1$ as a
+shorthand for ${\pi_1~"id"}$, typed as
+$\goodSub{\Gamma,A}{\pi_1~"id"}{\Gamma}$, and $\pi_2$ as a shorthand
+for $\pi_2~"id"$, typed as $\goodTerm{\Gamma, A}{\pi_2~"id"}{A[\pi_1]}$.
+
+\underline{\textit{\textbf{Original MLTT Constructs.}}}
+We only include  the dependent function types, dependent pair types, boolean type (as a base type later for canonicity), unit type, empty type and the universe type from the original MLTT.
+Dependent pair type is used to model Coq module, which is useful when we need to later model the ``compilation'' from
+linkages to modules. We also have $\beta,\eta$-rule for both dependent
+function type and dependent pair type.
+
+For these two types using binder, de Bruijn indices emphasizes its difference from the conventional setting---we directly use $\lambda t$ as a function term where $t$ is the body, no binder indication is required---to refer to variables in the context, we use $\pi_1,\pi_2$ from substitution calculus instead of using their names in the conventional setting.
+For example, the identity function $\lambda x.
+x$ is actually $\lambda(\pi_2 id)$.
+$"app"~t$ is even more different than that of the original setting---we don't require an application argument now.
+The reason is, (1) we asked for argument in conventional setting so we can use meta-level substitution as part of the dynamic, but here we use explicit substitution.
+If the user want to substitute some arguments, they can just use $[\gamma]$ instead.
+(2) Formulated in this way, "app" and $\lambda$ is symmetrical inverse to each other (just like how $El$ and "c" are to each other).
+This makes the $\beta\eta$-rule elegantly straightforward.
+Besides $\lambda$, $\Pi$ and $\Sigma$ are also using binders and are formulated in a similar style.
+
+Apparently, all of the type constructs have to commute with substitution somehow. The substitution law for dependent function and dependent pair
+requires \textit{de Bruijn lifting}: for a given substitution
+$\goodTerm{\Gamma}{\sigma}{\Delta}$ and a type
+$\goodType{\Delta}{A}{}$, we define its de Bruijn lifting
+$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ . Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.
+
+\subsubsection{\textbf{Core Syntax of the Extended MLTT}}
+\begin{figure}[!htb]
+  \begin{align*}
+&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
+Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
+I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
+  \end{align*}
+\caption{Extended Judgement forms, as Agda types}
+\end{figure}
+\begin{figure}[!htb]
+  \begin{minipage}[b]{0.3\linewidth}
+      $$
+      \Rule[name=Sig]
+      {\goodCtx{\Gamma}{i} \quad n \in \nat}
+      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
+      $$
+      $$
+      \Rule[name=Seal]
+      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
+      \quad \goodType{\Gamma}{A}{} }
+      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
+      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments?
+Besides, are these rules used anywhere in the following?
+Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes.
+Right column show all the well-formedness rule for all the judgement.
+That is unusual because it is a direct translation from QIIT.
+}
+  \end{minipage}
+  \begin{minipage}[b]{0.6\linewidth}
+    \begin{minted}[]{agda}
+      data Sig  : Con → ℕ → Set
+      data WSig : Con → ℕ → Set 
+      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
+      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
+    \end{minted}
+  \end{minipage}
+
+\caption{Well-formedness rules of the extended judgements}
+\end{figure}
+
+The core language builds on MLTT and adds four new judgement forms.
+% Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
+% shorthand for term judgement
+% $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
+% We have shown some exemplar rules for all four standard types of
+% judgments. Now we focus on the newly introduced facility.
+
+We have two kinds of signatures, one for inductive types, the other for linkages. We start with inductive type signature.
+
+% \begin{minted}[texcomments]{Coq}
+%   Γ : Con  (* $\goodCtx{\Gamma}{}$ *) 
+% \end{minted}
+% T : Ty Γ (* $\goodType{\Gamma}{T}{}$ *)       σ : Tms Γ Δ {- $\goodSub{\Gamma}{\sigma}{\Delta}$ -}        t : Tm Γ T {- $\goodTerm{\Gamma}{t}{T}$ -}
 

### 2022-07-17 syntax.tex
Add overview of syntax pieces.

@@
 % Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).
+
+\begin{figure}[!htb]
+  \begin{alignat*}{3}
+    \text{Context} \quad & \Gamma, \Delta &&= \cdot~|~ \Gamma, A \\ 
+    \text{Types}   \quad & A, B, T        &&= \cU~|~\cB~|~\bot~|~\top~|~\Pi A B~|~\Sigma A B~|~El~t~|~T[\gamma]\\
+    \text{Terms}   \quad & t, t_1, t_2    &&= "c"~T~|~t[\gamma]~|~\lambda t~|~"app"~t~|~(t_1,t_2)~|~"pjl"~t~|~"pjr"~t~|~"tt"~|~"ff"~|~\pi_2 \gamma  \\
+    \text{Substitution}   \quad & \gamma, \gamma_1, \gamma_2    &&= \epsilon~|~"id"~|~\gamma_1 \circ \gamma_2~|~(\gamma, t)~|~\pi_1 \gamma \\
+  \end{alignat*}
+\caption[Caption for LOF]{Overview of Instrinsically-typed Syntax for vanilla MLTT\protect\footnotemark}
+\end{figure}
+\footnotetext{This is not the definition of our syntax, just a grand overview of the intrinsically-typed syntax, to expose the reader with an idea on, for example, whether a syntax piece is a term or type, the general impression of the expressiveness of the calculus}
 
@@
 \Rule[name=Sub Extend]
-{\goodSub{\Gamma}{\sigma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
-{\goodSub{\Gamma}{(\sigma, t)}{(\Delta, A)}}
+{\goodSub{\Gamma}{\gamma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
+{\goodSub{\Gamma}{(\gamma, t)}{(\Delta, A)}}
 $$
@@
 \Rule[name=Proj Subst-1]
-{\goodSub{\Gamma}{\sigma}{(\Delta, A)}}
-{\goodSub{\Gamma}{\pi_1 \sigma}{\Delta}}
+{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
+{\goodSub{\Gamma}{\pi_1 \gamma}{\Delta}}
 \quad
 \Rule[name=Proj Subst-2]
-{\goodSub{\Gamma}{\sigma}{(\Delta, A)}}
-{\goodSub{\Gamma}{\pi_2 \sigma}{A[\pi_1 \sigma]}}
+{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
+{\goodSub{\Gamma}{\pi_2 \gamma}{A[\pi_1 \gamma]}}
 \quad
@@
 {}
-{\goodSub{\Gamma}{(\pi_1 \sigma, \pi_2 \sigma) \equiv \sigma}{\Delta}}
+{\goodSub{\Gamma}{(\pi_1 \gamma, \pi_2 \gamma) \equiv \gamma}{\Delta}}
 $$
@@
 For example, the identity function $\lambda x.
-x$ is actually $\lambda(\pi_2 id)$.
+x$ is actually $\lambda(\pi_2 "id")$.
 $"app"~t$ is even more different than that of the original setting---we don't require an application argument now.
@@
 \subsubsection{\textbf{Core Syntax of the Extended MLTT}}
+
+\begin{figure}[!htb]
+  \begin{alignat*}{3}
+    \text{Inductive Type Signature}   \quad & \tau       &&= w\cdot~|~w^+~\tau~A~B~|\tau[\gamma] \\
+    \text{Linkage Signature}          \quad & \sigma       &&= \nu\cdot~|~\nu^+~\sigma~\{f\}~T~|~\sigma[\gamma]~|~p_1\nu~\sigma~| \\
+    \text{Seal}          \quad & f       &&= t~|~p_f\nu'~\sigma\\    
+    \text{Types}   \quad & A, B, T        &&= \bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\\
+    \text{Terms}   \quad & t, t_1, t_2    &&= "W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t  \\
+  \end{alignat*}
+\caption{Overview of Instrinsically-typed Syntax for extended MLTT}
+\end{figure}
 \begin{figure}[!htb]
@@
 % judgments. Now we focus on the newly introduced facility. 
-
 We have two kinds of signatures, one for inductive types, the other for linkages. We start with inductive type signature.

### 2022-07-17 introduction.tex
more in intro

--- 2022-07-12 introduction.tex
@@
 % contribution
-It is hard to write modular, extensible code.
-It is even harder to write modular, extensible proofs.
-While there exist many linguistic solutions to code reuse and
-extensibility in object-oriented and functional languages,
-linguistic solutions are lacking in proof assistants.
 
-Meanwhile, language designers and researchers are increasingly
-using proof assistants to mechanize meta\-theories, building on each
-other's work.
-As a result, it has become ever more important that proof assistants
+There is a growing trend among semantics researchers to
+use proof assistants to mechanize meta\-theories.
+Consequently, it has become ever more important that proof assistants
 make it easy to write reusable, extensible code and proofs.
+
+However, the programmer runs into an old problem,
+the Expression Problem (EP) \cite{wadler-ep},
+that has deep roots in language design.
+%
+The EP is a programming challenge
+epitomizing the difficulty of writing type-safe, extensible code.
+To define an expression language that can be reused for future extension,
+the programmer faces a fundamental tension \cite{reynolds1975} between
+adding new constructors of a data type (e.g., new abstract syntax) and
+adding new functions over the data type (e.g., new compiler passes).
+
+The EP is well studied in the conventional setting of functional and
+object-oriented (OO) languages.
+Modern languages, such as Scala \cite{scala-oopsla05}, now have a good
+supply of linguistic features that offer expressive power to solve the
+EP.
+
+In contrast, proof assistants offer few linguistic solutions to
+programmers who want to mechanize their metatheories.
+Yet the challenge of writing extensible, type-safe code (and proofs!) is
+as real.
+As an example, consider \cref{fig:STLC-example}, the simply typed
+lambda calculus (STLC) mechanized in Coq.
+A programming challenge is to define distinct extensions of this STLC
+formalization to support distinct features,
+and then selectively compose these extensions to form new STLC variants.
+
+Because in Coq, code and proofs are closed to extension, the programmer
+cannot easily reuse them.
+The common practice is still to copy the existing mechanization and
+modify it for each extension.
+But having to maintain multiple copies is highly non-modular and
+inimical to good software engineering.
+%
+The programmer could also turn to design patterns~\cite{delaware2011,delaware2013}.
+But they require heavy lifting from the programmer to make code
+extensible, often leading to complex, non-idiomatic programming styles.
+
+A question naturally arises: can linguistic solutions to EP, as studied
+in the conventional setting, be adapted to metatheory mechanization in
+a proof assistant?
+
+At the core of many linguistic solutions to EP is \emph{inheritance}.
+Inheritance is sometimes interpreted narrowly as a subclass'
+inheriting methods and fields from a superclass.
+But its language-theoretic essence is more general:
+inheritance is a linguistic approach to incrementally
+modifying \emph{recursive, record-like structures}~\cite{cook1990inheritance}.
+
+Language mechanisms including
+mixins~\cite{mixin-1990},
+virtual classes~\cite{virtualclasses-1989,vc-calculus-2006},
+virtual types~\cite{thorup97} and associated types~\cite{ckj05},
+extensible cases~\cite{bac2006},
+and so on, are all based on the idea of inheritance.
+
+In particular, when a language mechanism applies inheritance to
+recursive, record-like definitions containing related types and terms,
+it is said to support \emph{family polymorphism}~\cite{ernst2001family}:
+code is polymorphic to a family it is nested within.
+
+\paragraph{Main ideas.}
+
+We contribute a language design that integrates family polymorphism into
+a proof assistant.
+Because code and proofs are polymorphic to a family it is nested
+within, they can be inherited and reused by a derived family.
+Extensible metatheory mechanization is thus made possible.
+
+Integrating family polymorphism into a dependent type theory poses
+significant technical challenges, however.
+\TODO
 

### 2022-07-17 abstract.tex
tweaks

With the growing practice of mechanizing language metatheories,
it has become ever more important that proof assistants
make it easy to write reusable, extensible code and proofs.
We ...

### 2022-07-17 introduction.tex
tweaks

@@
 
-There is a growing trend among semantics researchers to
-use proof assistants to mechanize meta\-theories.
-Consequently, it has become ever more important that proof assistants
-make it easy to write reusable, extensible code and proofs.
-
+There is a growing trend among programming languages researchers
+to use proof assistants to mechanize meta\-theories.
+%
 However, the programmer runs into an old problem,
 the Expression Problem (EP) \cite{wadler-ep},
-that has deep roots in language design.
-%
-The EP is a programming challenge
-epitomizing the difficulty of writing type-safe, extensible code.
+in the new setting of proof engineering.
+
+The EP is a programming challenge that
+epitomizes the difficulty of writing type-safe, extensible code.
 To define an expression language that can be reused for future extension,
@@
 
-In contrast, proof assistants offer few linguistic solutions to
-programmers who want to mechanize their metatheories.
+In contrast, proof assistants offer few linguistic solutions that
+address the EP.
 Yet the challenge of writing extensible, type-safe code (and proofs!) is
-as real.
+as real, especially for metatheory mechanization.
 As an example, consider \cref{fig:STLC-example}, the simply typed
 lambda calculus (STLC) mechanized in Coq.
-A programming challenge is to define distinct extensions of this STLC
-formalization to support distinct features,
+A programming challenge there is to define distinct extensions of this
+STLC formalization to support distinct features,
 and then selectively compose these extensions to form new STLC variants.
@@
 
-In particular, when a language mechanism applies inheritance to
+In particular, when a language mechanism enables inheritance over
 recursive, record-like definitions containing related types and terms,
@@
 within, they can be inherited and reused by a derived family.
-Extensible metatheory mechanization is thus made possible.
+%
+Family polymorphism enables extensible metatheory mechanization.
+As \cref{fig:?} shows,
+different extensions to STLC all inherit from the base STLC family:
+they reuse proofs of metatheories including the progress and
+preservation theorems, only adding new cases to the induction proofs as
+needed by the extension.
 
-Integrating family polymorphism into a dependent type theory poses
-significant technical challenges, however.
+Integrating family polymorphism into a dependent type theory with
+inductive types poses significant technical challenges, however.
 \TODO

### 2022-07-17 introduction.tex
intro

--- 2022-07-12 introduction.tex
@@
 
-Language designers and researchers are increasingly
-using proof assistants to mechanize meta\-theories, building on each
-other's work.
-As a result, it has become ever more important that proof assistants
-make it easy to write reusable, extensible code and proofs.
+There is a growing trend among programming languages researchers
+to use proof assistants to mechanize meta\-theories.
+%
+However, the programmer runs into an old problem
+in the new setting of proof engineering:
+the Expression Problem (EP) \cite{wadler-ep}.
 
-However, linguistic solutions to type-safe code reuse and extensibility
-are lacking in proof assistants.
-This situation is in contrast to modern languages,
-such as Scala \cite{scala-oopsla05}, that have a good supply of
-linguistic features which solve the Expression problem \cite{wadler-ep}.
+The EP is a programming challenge that
+epitomizes the difficulty of writing type-safe, extensible code.
+To define an expression language that can be reused for future extensions,
+the programmer faces a fundamental tension \cite{reynolds1975} between
+adding new constructors of a data type (e.g., new abstract syntax) and
+adding new functions over the data type (e.g., new compiler passes).
 
-\paragraph{Structure of the paper} 
-We will quickly introduce Family Polymorphism and the challenges to adapt it into dependent type setting in \cref{sec:background+challenge}.
-After that, we will talk about the language design of family polymorphism in dependent type setting and the implementation of our Coq plugin in \cref{sec:coqimpl}.
-Then, we consider the meta-theory of incorporating family polymorphism into predicative MLTT, and deriving consistency and canonicity results in \cref{sec:metatheory}.
-\ref{sec:related-work} discusses related works and \ref{sec:conclusion} concludes.
+The EP is well studied in the conventional setting of functional
+programming and object-oriented (OO) programming.
+Modern languages, such as Scala \cite{scala-oopsla05}, have a good
+supply of linguistic features that offer expressive power to solve the
+EP.
 
-\paragraph{Contributions} include the following
-\begin{itemize}
-  \item To our knowledge, this is the first attempt incorporating family polymorphism with theorem proving and higher order logic
-  \item We contribute a prototypical Coq plugin for the interested users and developers 
-  \item We provide some first-step meta-theoretical results, including consistency and canonicity
+In contrast, proof assistants offer few linguistic solutions that
+address the EP.
+Yet the challenge of writing extensible, type-safe code (and proofs!) is
+as real, especially for metatheory mechanization.
+As an example, consider \cref{fig:STLC-example}, the simply typed
+lambda calculus (STLC) mechanized in Coq.
+A programming challenge there is to define distinct extensions of this
+STLC formalization to support distinct features,
+and then selectively compose these extensions to form new STLC variants.
+
+In Coq, inductive types, as well as functions and theorems over
+inductive types, are closed to extension.
+So to reuse mechanized metatheories,
+the common practice is still to copy and modify them for each extension.
+But having to maintain multiple copies is highly non-modular and
+antithetical to good software engineering.
+%
+The programmer could also turn to design patterns~\cite{delaware2011,delaware2013}.
+But they require heavy lifting from the programmer to make code
+extensible, often leading to verbose, non-idiomatic programming styles.
+
+A question thus arises: can linguistic solutions to the EP, as studied
+in the conventional setting, be adapted to metatheory mechanization in
+a proof assistant?
+
+At the core of many linguistic solutions to the EP is \emph{inheritance}.
+Inheritance is sometimes interpreted narrowly as a subclass'
+inheriting methods and fields from a superclass.
+But its language-theoretic essence is more general:
+inheritance is a linguistic approach to incrementally
+modifying \emph{recursive, record-like definitions}~\cite{cook1990inheritance}.
+
+Language mechanisms including
+mixins~\cite{mixin-1990},
+virtual classes~\cite{virtualclasses-1989,vc-calculus-2006},
+virtual types~\cite{thorup97} and associated types~\cite{ckj05},
+extensible cases~\cite{bac2006},
+and so on, are all based on the essential idea of inheritance.
+%
+In particular, when a language mechanism enables inheritance over
+recursive definitions containing related types and terms,
+it is said to support \emph{family polymorphism}~\cite{ernst2001family}:
+types and terms are polymorphic to a family they are nested within.
+
+\paragraph{Main ideas.}
+
+We contribute a language design that integrates family polymorphism into
+a proof assistant.
+Because code and proofs are polymorphic to a family they are nested
+within, they can be inherited and reused by a derived family.
+%
+Family polymorphism enables extensible metatheory mechanization.
+As \cref{fig:?} shows,
+different extensions to STLC can all inherit from the base STLC family:
+they reuse mechanized metatheories from syntax to the type safety
+theorem, only adding new constructors to inductive types
+and new cases to the induction proofs as needed by an extension.
+
+Integrating family polymorphism into a dependent type theory with
+inductive types poses significant technical challenges, however.
+As we analyze, inductive types, definitional equalities, and logical
+consistency, are all inimical to the kind of family polymorphism
+found in existing language designs.
+Thus, our core contributions include design recipes for dealing with
+these issues and strong metatheoretical guarantees on consistency and
+canonicity.
+
+\paragraph{Contributions} To our knowledge, we offer the first
+language design that enables extensible metatheory mechanization
+in a higher-order, dependent type theory with inductive types.
+
+\begin{itemize}[leftmargin=3.5ex]
+
+\item We show how to integrate family polymorphism into a proof
+assistant (\cref{sec:lang-design}).
+The language design reconciles the expressiveness enabled by
+family polymorphism with the exactness of a proof assistant,
+while retaining an idiomatic programming style.
+
+\item We contribute a prototypical implementation of our language
+mechanism as a Coq plugin (\cref{sec:coqimpl}). The plugin works by
+compiling surface-language terms into Gallina terms parameterized by
+extensibility hooks.
+
+\item We capture the new language mechanism formally by extending
+Martin–Löf type theory with family polymorphism and extensible inductive
+types (\cref{sec:metatheory}).
+We derive strong metatheoretical results including on consistency and
+canonicity.
+We also formalize the translation implemented by the Coq plugin.
+
+\item We present case studies of using our Coq plugin to mechanize
+language metatheories.
+They show how our linguistic approach naturally solves the EP and
+enables a high degree of reusability and extensibility
+for proof engineering.
+
 \end{itemize}
+
+%\paragraph{Structure of the paper} We will quickly introduce Family
+%Polymorphism and the challenges to adapt it into dependent type setting
+%in \cref{sec:background+challenge}. After that, we will talk about the
+%language design of family polymorphism in dependent type setting and the
+%implementation of our Coq plugin in \cref{sec:coqimpl}. Then, we
+%consider the meta-theory of incorporating family polymorphism into
+%predicative MLTT, and deriving consistency and canonicity results in
+%\cref{sec:metatheory}. \ref{sec:related-work} discusses related works
+%and \ref{sec:conclusion} concludes.

### 2022-07-19 introduction.tex
pass on lang-design sec

@@
 inductive types poses significant technical challenges, however.
-As we analyze, inductive types, definitional equalities, and logical
-consistency, are all inimical to the kind of family polymorphism
-found in existing language designs.
-Thus, our core contributions include design recipes for dealing with
-these issues and strong metatheoretical guarantees on consistency and
-canonicity.
+As we analyze, pillars of such a dependent type theory—including
+inductive types, definitional equalities, and logical consistency—are
+all inimical to the kind of extensibility and family polymorphism
+found in existing OO language designs.
+Thus, our contributions include novel design recipes for dealing with
+these challenges and strong meta\-theoretical guarantees on the
+underlying logic.
 
 \paragraph{Contributions} To our knowledge, we offer the first
-language design that enables extensible metatheory mechanization
+language-based solution that enables extensible metatheory mechanization
 in a higher-order, dependent type theory with inductive types.
@@
 types (\cref{sec:metatheory}).
-We derive strong metatheoretical results including on consistency and
+We derive strong meta\-theoretical results including on consistency and
 canonicity.
@@
 language metatheories.
-They show how our linguistic approach naturally solves the EP and
-enables a high degree of reusability and extensibility
+They show how our language design naturally solves the EP and
+enables a high degree of reuse and extensibility
 for proof engineering.

### 2022-07-19 lang-design.tex
Alter the simple negation example into the complicated STLC example In Coqimpl, implementation of extensible inductive type. Please check
pass on lang-design sec

@@
  FInductive term : Set := 
-  | var : id -> term 
-  | lam : id -> term -> term ...
+  | tm_var : id -> term 
+  | tm_abs : id -> term -> term ...
  Family subst_internal.
-  Final Field var := ..
-  Final Field lam := ..
+  Final Field tm_var := ..
+  Final Field tm_abs := ..
    ...
@@
 % Inheritance here
-\textbf{Inheritance on Fields.} "Family A extends B" has to appear at the top level (not inside any Family) and will start to construct the children "Family A" given the parent "Family B".
-Every field inside the inheriting family will be one of the three kinds of ``inheritance''---extending, overriding, and inheriting.
-The syntax of extending a new field is to directly use "Field", while overriding requires a prefix of "Override" before "Field f : .." and "Inherit f" will inherits field "f" from the parent.
+\textbf{Inheritance on Fields.} "Family A extends B" has to appear at
+the top level (not inside any Family) and will start to construct the
+children "Family B" given the parent "Family A". Every field inside the
+derived family will be one of the three kinds of
+``inheritance''---extending, overriding, and inheriting.
+The syntax of extending a new field is to directly use "Field", while
+overriding requires a prefix of "Override" before "Field f : .." and
+"Inherit f" will inherit field "f" from the parent.
 
-To resolve \ref{chg:consistency}, we make sure when inheriting or overriding on "f" happens, the context of the field "f" is also ``inherited''.
-Thus the mentioned informal example in \ref{chg:consistency} will not apply here because the overriding "f" is also in the context where "self" is empty thus "self.g" doesn't type-check.
-Thus, before we inherit "g", the dependee "f" needs to be inherited.
+To resolve \ref{chg:consistency}, we make sure when inheriting or
+overriding on "f" happens, the context of the field "f" is also
+``inherited''. Thus the mentioned informal example in
+\ref{chg:consistency} will not apply here, because the overriding "f" is
+in a context where "self" is empty thus "self.g" doesn't
+type-check. Thus, before we inherit "g", the dependee "f" needs to be
+inherited.
 
@@
 % Talk about Final
-\textbf{Final as non-overridable.} Every field is by default overridable, reminding us \ref{chg:definition-relevant-reasoning}.
-To resolve it, we provide the keyword "Final" to allow the user to specify \textit{non-overridability} and give the user the opportunity to choose from overriding and exposing the definition.
+\textbf{Final as non-overridable.} Every field is by default
+overridable, reminding us of \ref{chg:definition-relevant-reasoning}.  To
+resolve it, we provide the keyword "Final" to allow the user to specify
+\textit{non-overridability} and give the user the opportunity to choose
+from overriding and exposing the definition.
 
-Take the field "c" of "Family I" in \cref{fig:plugin-example3} as the
+Take the field "c" of "Family I" in \cref{fig:plugin-example3} as an 
 example. We can see that the first attempt of proving "a = 1" fails.
 This should fail---because we want to retain the possibility
-to override "a" and inheriting "c" in the future, but they are conflicting, since "c" witnesses "a" fix to "1".
-On the other hand, once we decorate it upon "a'", the second attempt \mintinline{Coq}{c : a' = 1 := eq_refl} will succeed.
-Contrary to "a", our "a'" is not overridable in "Family I2", mediating the \ref{chg:definition-relevant-reasoning}.
+to override "a" and inheriting "c" in the future, but they are
+conflicting, since "c" witnesses that "a" is fixed to "1". On the other
+hand, once we decorate \lstinline{a'} as final, the second attempt
+\lstinline{c : a' = 1 := eq_refl} will succeed. Contrary to "a",
+our \lstinline{a'} is not overridable in "Family I2", mediating the
+\ref{chg:definition-relevant-reasoning}.
 
 This example might sound trivial and contrived. Let's look at the
-example field binary operation "bop" initially set to be addition.\YZ{This 'add' field is no longer there.}\EDJreply{Thanks. Fixed} Now that we only expose the type of "bop", we
+example field binary operation "bop" initially set to be addition.
+Now that we only expose the type of "bop", we
 cannot prove its commutativity in "comm_bop" (because we don't have
@@
 subtraction function in "Family I2". This makes sense because we
-shouldn't (have the ability to) inherit commutativity property for subtraction
-anyway. This is contrary to how "a'" is declared "Final", the second "c"
-can prove that it's equal to 1 but "a'" is not overridable in the
-"Family I2".
+shouldn't (have the ability to) inherit commutativity property for
+subtraction anyway. This is contrary to how \lstinline{a'} is declared
+"Final": the second "c" can prove that it's equal to 1 but
+\lstinline{a'} is not overridable in the "Family I2".
 
@@
 
-\textbf{Inheritance on Nested Families.} We also provide extending, inheriting and overriding for nested family (member).
-Extension is done by simply declaring a nested family; the same syntax for inheriting a field is used in inheriting a family.
-However, overriding is special---overriding can only apply to \textit{sealed} nested family in the parent family.
-A family is sealed when using "SealFamily" as the scoping instead of "EndFamily".
-We will later elaborate the motivation and details of sealed family.
+\textbf{Inheritance on Nested Families.} We also provide extending,
+inheriting and overriding for nested family (member). Extension is done
+by simply declaring a nested family; the same syntax for inheriting a
+field is used in inheriting a family. However, overriding is
+special---overriding can only apply to \textit{sealed} nested family in
+the parent family. A family is sealed when using "SealFamily" as the
+scoping instead of "EndFamily". We will later elaborate the motivation
+and details of sealed family.
 
@@
 
-\textbf{Extensible Inductive Type.} We provide the keyword "FInductive" to enable extensible inductive type.
-"Extend FInductive" is used to support extension of the parent inductive type inside a children family.
+\textbf{Extensible Inductive Type.} We provide the keyword "FInductive"
+to enable extensible inductive type. "Extend FInductive" is used to
+support extension of the parent inductive type inside a children family.
 The syntax of these two is close to the mundane "Inductive" keyword.
 
-To resolve \ref{chg:extensible-inductive-type}, our first step is to \textit{decouple} the definition of recursor handlers and the checking of exhaustiveness.
-The users will define recursive handlers for one inductive type by aggregating the (function) handlers into one family.
+To resolve \ref{chg:extensible-inductive-type}, our first step is to
+\textit{decouple} the definition of recursor handlers and the checking
+of exhaustiveness. The users will define recursive handlers for one
+inductive type by aggregating the (function) handlers into one family.
 Then
-"FRecursor" will be used to construct a recursor by refering to the handler family.
+"FRecursor" will be used to construct a recursor by referring to the handler family.
 \mintinline{Coq}{FRecursor recur about T motive f } "using"
@@
 
-"FTheorem" has a similar functionality but is expected to be used for theorem proving---the handler and the family will be automatically constructed during proving.
+"FTheorem" has a similar functionality but is expected to be used for
+theorem proving---the handler and the family will be automatically
+constructed during proving.\YZ{Cool. Can the same be done to FRecursor?}
 

### 2022-07-19 metatheory.tex
comment

@@
 formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT)~\cite{martin1982constructive}
-given by~\citet{coquand2018canonicity}. 
+given by~\citet{coquand2018canonicity}.\YZ{
+    What's the relation to the CIC underlying Coq, given that the implementation is a Coq plugin?
+}
 % Thus we will omit a great deal of details about MLTT here
-\YZ{What details are omitted?} 
-We will mainly focus on the new facility that we introduced
+We will mainly focus on the new facility that we introduced.
 

### 2022-07-22 coqimpl.tex
Add limitation subsection

@@
 Please refer to the supplementary material and appendix for a formalization of the left inverse in the context of indexed inductive type.
+
+\subsection{Limitation}
+We list some of the current limitation of our implementation. We believe some of these limitations can be remedied by more engineering efforts.
+\begin{itemize}
+  \item We don't support the full power of inductive facility in Coq, that includes mutual inductive type, notation for inductive type and parameters in the inductive type (we only support indexed inductive type right now)
+  \item We don't support mutual fixpoint definition
+  \item We don't support incorporating "Hint" Vernacular commands, and thus the users cannot specify the database hints
+  \item Unicode (i.e. "Require Import Coq.Unicode.Utf8.") is not compatible with our plugin 
+  \item We don't consider the compatibility of universe polymorphism mechanism with our plugin
+\end{itemize}

### 2022-07-25 syntax.tex
Adding references for W-type

@@
 
-Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A
+We base our formalization of the inductive type on W-type~\cite{martin1982constructive}. 
+W-type, despite of the simplicity of its formulation, is able to construct a wide range of inductive type in extensional setting~\cite{martin1984intuitionistic,dybjer1997representing,abbott2004representing} and intensional setting (with the help of identity type~\cite{hugunin2021not}).
+Recall in W-type, a pair of type $x : A
 \vdash B(x)$ decides a W-type. To simulate extensible inductive types, we

### 2022-07-25 coqimpl.tex
edits

--- 2022-07-14 coqimpl.tex
@@
 
-\YZ{1.
-move some of sec 3.2 ~> 3.1; 2.
-hint vaguely on how to do type checking in sec 3.1 (user interface)..
-metatheory part should also mention this part in sec 3.1} This is achieved by a complete compilation from a family to a module, via aggregating the list of the modules in the internal representation: the plugin will repeatedly using Coq's "Include" command as the "Module A" in \cref{fig:plugin-example1} illustrates.
-Now, the user can use "Module A" anywhere as they are using "Family A".
-\YZ{The reader wonders why this aggregation is needed. 
-Seems like the place to talk about how the programmer can use A outside Family A.}\EDJreply{Please check.}
+This is achieved by a complete compilation from a family to a module,
+via aggregating the list of the modules in the internal representation:
+the plugin will repeatedly using Coq's "Include" command as the "Module
+A" as \cref{fig:plugin-example1} illustrates.
+References to family "A" outside "A" will be compiled to references to
+this Coq module "A".
 
@@
 The resulting compiled module is just like mundane Coq's module and 
-thus maintains all the computational mechanism from Coq, fulfilling requirements in \ref{chg:software-engineering}.
+thus maintains all the computational mechanism from Coq, fulfilling
+requirements in \ref{chg:software-engineering}.\YZ{
+  Do I understand correctly that an alternative to individually
+  compiling family members to modules is to compile a family as a whole,
+  but this alternative would not satisfy the "incremental type checking"
+  or "instant feedback" requirement?
+}\YZreply{
+  Looks like individually compiling family members addresses not only
+  this requirement, but also the "self vs. consistency" challenge.
+}
 
@@
 The surface syntax for the three kinds of inheritance
-are demonstrated in the exemplar \cref*{fig:plugin-example3}. The
+are demonstrated in the exemplar \cref{fig:plugin-example3}. The
 implementation of overriding, inheritance, and extension is achieved by
@@
 Though inheritance is solely a first-order data rather
-than a ``mapping'', it is still possible to
-\text{mixin} two inheritance data---by carefully ``mix'' the internal list, we can compose the inheritance.
-However, a good definition of mixin is still under investigation.
-We will show one practical example of \textit{mixin} later.\YZ{How about making mixin composition one of your Req's?
-Also does any example in the paper illustrate how to do this?}\EDJreply{mixin is not really formalized.
-There is no example illustrate it.
-The only example is the STLC example at the very end and I just kind of ``mention'' it.
-I think here I just want to emphasize, inheritance as first order data is still possible to support composition.}
+than a ``mapping'', it is still possible to \text{mixin} two inheritance
+data---by carefully ``mix'' the internal list, we can compose the
+inheritance. However, a good definition of mixin is still under
+investigation. We will show one practical example of \textit{mixin}
+later.\YZ{How about making mixin composition one of your Req's? Also
+does any example in the paper illustrate how to do this?}\EDJreply{mixin
+is not really formalized. There is no example illustrate it. The only
+example is the STLC example at the very end and I just kind of
+``mention'' it. I think here I just want to emphasize, inheritance as
+first order data is still possible to support composition.}
 
-Our implementation also make sure the first attempt of defining "c" failed successfully in \cref{fig:plugin-example3}---the proof of "c" will require a concrete definition of "a",
-but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in (a module with) the context where we only know \mintinline{Coq}{a : nat} thus failed.
+Our implementation also makes sure the first attempt of defining "c"
+failed successfully in \cref{fig:plugin-example3}---proving "c"
+would require a concrete definition of "a", and our plugin rejects
+\mintinline{Coq}{c : a = 1 := eq_refl} in a context where we
+only know \mintinline{Coq}{a : nat}.
 
-\textbf{Implemening "Final" and "Sealed Family".}
+\textbf{Implementing "Final" and "Sealed Family".}
 We implement "Final" by
 exposing the whole definition directly into the compiled module type.
-Doing so will prohibit overriding.
-Say, the example "bop" in \cref{fig:plugin-example3}, will be prohibited to be overridden by $(-)$
-subtraction doesn't have the same definition as $(+)$. We require the
-users to manual decorate field as `Final' so that the plugin will proceed with this special treatment.
+Doing so will prohibit overriding. Say, "bop", declared "Final" in
+\cref{fig:plugin-example3}, is not overridable.
+We require the users to manual decorate field as `Final' so that the
+plugin will proceed with this special treatment.
 %\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
@@
 We support extensible inductive types in Coq mentioned in
-\ref{chg:extensible-inductive-type} but we will not be aiming at the research of its formal semantic.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types.
+\ref{chg:extensible-inductive-type}, but we will not be aiming at the
+research of its formal semantic.\footnote{Because to give a satisfactory
+formal definition and semantic for extensible inductive type is not
+something this paper aims at, and obviously requires more theoretical
+research effort}
+%\YZ{I do consider that your plugin supports extensible inductive types.
 It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it.
@@
   \end{minipage}
-\caption{Exempler STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
-\end{figure}\YZ{Change this one into STLC example, to see the layout}\EDJreply{Changed. I think it only make things worse.}
+\caption{Exemplar STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
+\end{figure}
 
@@
 
-Let's look at the exemplar \cref{fig:plugin-example2}, the example of "subst" on STLC.
-When we define the extensible inductive type "tm", all the following fields (e.g.
-"subst_internal.tm_var") can only know "tm" is a type (of "Set") and there are at least two constructors for it, just like what we declare in ``the abstracted interface''---the compiled module type "STLC_409".
-We cannot export\YZ{export or put?}\EDJreply{Ok.
-I just figure out how my wording is confusing.
-put is absolutely correct but by put I want to mean ``export''.
-I think I use ``export'' in several places...}
-the eliminator of the inductive type "tm" to "STLC_409" because doing so all of the following fields (e.g.
-"subst_internal.tm_var" and the corresponding "tm_var_411") will not be able to be inherited to the context where "tm"
-is extended with a third constructor, \textbf{since the mere existence of the recursor will assert that there are only two constructors} in "STLC_bool", not
-\textbf{extensible} at all!\YZ{I can see how the existence of the recursor alone prevents extensibility, but I don't see why it has to do with ff_24.}\EDJreply{You are right.
-It is not only about "ff_24".I rephrased it.
-Please check.}
+Let's look at the exemplar \cref{fig:plugin-example2}, the example of
+"subst" on STLC. When we define the extensible inductive type "tm", all
+the following fields (e.g., "subst_internal.tm_var") can only know "tm"
+is a type (of "Set") and there are at least two constructors for it,
+just like what we declare in ``the abstracted interface''---the compiled
+module type "STLC_409". We cannot export
+%\YZ{export or put?}\EDJreply{Ok.
+%I just figure out how my wording is confusing. put is absolutely correct
+%but by put I want to mean ``export''. I think I use ``export'' in
+%several places...}
+the eliminator of the inductive type "tm" to "STLC_409".
+Otherwise all the subsequent fields (e.g. "subst_internal.tm_var" and
+the corresponding "tm_var_411") would not be able to be inherited to the
+context where "tm" is extended with a third constructor, \textbf{since
+the mere existence of the recursor will assert that there are only two
+constructors} in "STLC_bool", not \textbf{extensible} at all!
 
-However, we later still need the concrete eliminator on "tm" for doing exhaustiveness checking (as in \ref{chg:extensible-inductive-type}), thus we will store the corresponding info somewhere.
-Here we show one exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module "tm_rec_425".
-This concrete eliminator will be used by our "FRecursor".
-For example, in \cref{fig:plugin-example2}, "FRecursor" command will feed Coq a module "v_33_34" for doing type-checking with specified handlers (e.g.
-"neg_handler") during compilation---what it does is basically
+However, we later still need the concrete eliminator on "tm" for doing
+exhaustiveness checking (as in \ref{chg:extensible-inductive-type}),
+thus we will store the corresponding info somewhere.  Here we show one
+exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module
+"tm_rec_425". This concrete eliminator will be used by our "FRecursor".
+For example, in \cref{fig:plugin-example2}, "FRecursor" command will
+feed Coq a module "v_33_34" for doing type-checking with specified
+handlers (e.g. "neg_handler") during compilation---what it does is
+basically
 assert the existence of the recursor and apply to see if handlers can
-pass the type checking by Coq.
-Note that, at this point of the example, "FRecursor" will only do type check.
-The compilation of "FRecursor" into a component of the compiled module only happens when closing the whole family, by inserting the following into the compiled module:
+pass the type checking by Coq.\YZ{
+  Sounds like we can still claim the plugin is doing type checking.
+  It's just that the checking is done by generating some Coq code and
+  then having Coq check it.
+}
+Note that, at this point of the example,
+"FRecursor" will only do type check. The compilation of "FRecursor" into
+a component of the compiled module only happens when closing the whole
+family, by inserting the following into the compiled module: 
 \begin{minted}{Coq}
@@
 \end{minted}
-where "tm_rec" and "tm" are the corresponding recursor and the (vanilla) inductive type in that compiled module.
-Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields in the programming interface.
+where "tm_rec" and "tm" are the corresponding recursor and
+the (vanilla) inductive type in that compiled module. Thus a "FRecursor"
+field is unlike other fields, even though it is exposed as a mundane
+function for the following fields in the programming interface.\YZ{
+  Why is this special treatment needed for compiling FRecursor?
+}
 
-The plugin actually realizes the "Extend FInductive" via overridding---when extending an inductive type $A$
-with new constructors $c$, our plugin will generate a new (syntactic) inductive definition $A'$ with this new $c$ on the fly and feed it to Coq.
-But we must be careful on the exposing data in the family type (the compiled module type) for the following field.
-In other words, we must be careful on what interface (context) the following fields are defined based on.
+The plugin actually realizes the "Extend FInductive" via overriding---%
+when extending an inductive type $A$ with new constructor $c$, our
+plugin will generate a new (syntactic) inductive definition $A'$ with
+this new $c$ on the fly and feed it to Coq. But we must be careful on
+the exposing data in the family type (the compiled module type) for the
+following field. In other words, we must be careful on what interface
+(context) the following fields are defined based on.
 
-Importantly, the recursor field (e.g.
-"subst") is not actually inheritable, because it is not compiled like any other fields---recall how other fields are compiled into reusable pieces of (parametrized) modules, a recursor field is not compiled like that.
-For all the children family, we have to reconstruct and override the "FRecursor" field with a ``new'' recursor using (maybe the same) handler family.
-This is affordable since each aggregated recursor handler doesn't need
-to be rechecked---look at "STLC_bool.neg_handler", there are only  new
-handlers for "tm_true" and the handlers (mundane fields) from the parent
-"B.subst_internal.tm_var" have already been type-checked and
-inherited (since they only rely on ``the abstracted
-interface'' module "STLC_409"). 
+Importantly, the recursor field (e.g. "subst") is not actually
+inheritable, because it is not compiled like any other fields.
+Recall how other fields are compiled into reusable pieces of
+(parametrized) modules; a recursor field is not compiled like that.
+For all the children family, we have to reconstruct and override the
+"FRecursor" field with a ``new'' recursor using (maybe the same) handler
+family.  This is affordable since each aggregated recursor handler
+doesn't need to be rechecked---look at "STLC_bool.neg_handler", there
+are only  new handlers for "tm_true" and the handlers (mundane fields)
+from the parent "B.subst_internal.tm_var" have already been type-checked
+and inherited (since they only rely on ``the abstracted interface''
+module "STLC_409"). 
 Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
-the module "tm_var_411"---doesn't require another type-check and is re-used during the type-checking and compilation of "STLC_bool.subst".
-For example, for the new "subst", when compiling family "STLC_bool" into a module, at the appropriate place our plugin will insert
+the module "tm_var_411"---doesn't require another type-check and is
+re-used during the type-checking and compilation of "STLC_bool.subst".
+For example, for the new "subst", when compiling family "STLC_bool" into
+a module, at the appropriate place our plugin will insert
 \begin{minted}{Coq}
@@
 \end{minted}
-where "tm_rec" and "tm" are the corresponding recursor and the (vanilla but extended and modified) inductive type in that compiled module.
-Note that, even though "tm_rec" and "tm" has the same name as before, they are in the newly-compiled module so they refer to totally different things.
+where "tm_rec" and "tm" are the corresponding recursor and the (vanilla
+but extended and modified) inductive type in that compiled module. Note
+that, even though "tm_rec" and "tm" has the same name as before, they
+are in the newly-compiled module so they refer to totally different
+things.
 
 Of course, in the implementation, our plugin will ``inherit'' the
-recursor "neg" in the surface \YZ{Show this recursor, as you do above for the recursor "Definition neg := ...".}\EDJreply{Added above. Please Check.}
+recursor "neg" in the surface,
 but what the plugin actually does is doing a second exhaustiveness
-checking in children family like mentioned above, with the ``same'' motive and the ``same'' handler family.\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
-Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}\EDJreply{Please check.}
+checking in children family like mentioned above, with the ``same'' motive and the ``same'' handler family.
+%\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
+Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}
 

### 2022-07-25 introduction.tex
edits

@@
 Inheritance is sometimes interpreted narrowly as a subclass'
-inheriting methods and fields from a superclass.
-But its language-theoretic essence is more general:
-inheritance is a linguistic approach to incrementally
-modifying \emph{recursive, record-like definitions}~\cite{cook1990inheritance}.
+inheriting methods and instance variables from its superclass.
+But the language-theoretic essence of inheritance is more general:
+it is a linguistic approach to incrementally modifying
+\emph{recursive, record-like definitions}~\cite{cook1990inheritance}.
 
@@
 
-Integrating family polymorphism into a dependent type theory with
-inductive types poses significant technical challenges, however.
-As we analyze, pillars of such a dependent type theory—including
-inductive types, definitional equalities, and logical consistency—are
+Integrating family polymorphism into a dependent type theory for
+logical reasoning, however, poses significant technical challenges.
+As we analyze, pillars of dependent type theories—including
+inductive types, definitional equality, and logical consistency—are
 all inimical to the kind of extensibility and family polymorphism

### 2022-07-25 syntax.tex
edits

--- 2022-07-17 syntax.tex
@@
 
-In other words, (1) a universe $\cU$ is a type including the \textit{codes or names} of types.
-(2) a code (term) $\goodTerm{\Gamma}{T}{\cU}$ is not allowed to locate at type position
-(i.e., at the rightmost position of a term judgement) but $El \ T$ is (i.e. we distinguish the code and the type). (3)
-"c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of a
-type. More concretely, using QIIT, the corresponding constructors will look like, (notice that the type on both sides of the equalities are always the same and thus a well-typed equality) 
+In other words, (1) a universe $\cU$ is a type inhabited by the \textit{codes or names} of types.
+(2) a code (term) $\goodTerm{\Gamma}{T}{\cU}$ cannot be used at type position
+(i.e., at the rightmost position of a term judgement) but $El \ T$ can (i.e., we distinguish the code and the type).
+(3) "c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of
+a type. More concretely, using QIIT, the corresponding constructors will
+look like, (notice that the type on both sides of the equalities are
+always the same and thus a well-typed equality) 
 \begin{align*}

### 2022-07-26 syntax.tex
bugfix

@@
 Note that we are actually using predicative universe, so we actually
-have different levels of universes $\cU_j$. However, we omit universe levels in our latex presentation for conciseness and accessibility.
+have different levels of universes $\cU_j$. However, we omit universe levels in our latex presentation for conciseness and readability.
 % Most of the reasoning can be formulated using the \textit{algebra} of QIIT (mapping out function from this quotient data-type).
@@
 \Rule[name=Sub Extend]
-{\goodSub{\Gamma}{\gamma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
+{\goodSub{\Gamma}{\gamma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\gamma]}}
 {\goodSub{\Gamma}{(\gamma, t)}{(\Delta, A)}}
@@
 Thus (1) \ruleref{Sub Extend} looks like concatenating one
-series of terms with a new term (the weird looking type of
-$\goodTerm{\Gamma}{t}{A[\sigma]}$ in the premise is due to our need of type $A$ in
+series of terms with a new term (the weird-looking type of
+$\goodTerm{\Gamma}{t}{A[\gamma]}$ in the premise is due to our need of type $A$ in
 the context $\Delta$ so that we can use \ruleref{Context Extension} to

### 2022-07-26 syntax.tex
Save. Add Identity Type and Based Path Induction.
edits and comments

@@
 \end{align*}
-These two equations indicate how \textit{dynamic} should react when $El~"c"~T$ appears. Later we will see other similar $\beta$-equations but we won't display their (pseudo-)Agda formulation.
+These two equations indicate how the \textit{dynamics} should react when
+$El~"c"~T$ appears. Later we will see other similar $\beta$-equations
+but we won't display their (pseudo-)Agda formulation.
 
@@
 Since the $\Gamma$ before the $\vdash$ is always indicating the types of
-``holes'', this explains the \ruleref{Term Subst} and \ruleref{Type
+``holes'', this explains rules \ruleref{Term Subst} and \ruleref{Type
 Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution
@@
 \underline{\textit{\textbf{Original MLTT Constructs.}}}
-We only include  the dependent function types, dependent pair types, boolean type (as a base type later for canonicity), unit type, empty type and the universe type from the original MLTT.
-Dependent pair type is used to model Coq module, which is useful when we need to later model the ``compilation'' from
-linkages to modules. We also have $\beta,\eta$-rule for both dependent
-function type and dependent pair type.
+We only include  the dependent function types, dependent pair types,
+the boolean type (as a base type later for canonicity), the unit type,
+the empty type, and the universe type from the original MLTT. Dependent
+pair types are used to model Coq module, which is useful when we need to
+later model the ``compilation'' from families to modules. We also have
+$\beta$- and $\eta$-rules for both dependent function types and
+dependent pair types.
 
-For these two types using binder, de Bruijn indices emphasizes its difference from the conventional setting---we directly use $\lambda t$ as a function term where $t$ is the body, no binder indication is required---to refer to variables in the context, we use $\pi_1,\pi_2$ from substitution calculus instead of using their names in the conventional setting.
-For example, the identity function $\lambda x.
-x$ is actually $\lambda(\pi_2 "id")$.
-$"app"~t$ is even more different than that of the original setting---we don't require an application argument now.
-The reason is, (1) we asked for argument in conventional setting so we can use meta-level substitution as part of the dynamic, but here we use explicit substitution.
-If the user want to substitute some arguments, they can just use $[\gamma]$ instead.
-(2) Formulated in this way, "app" and $\lambda$ is symmetrical inverse to each other (just like how $El$ and "c" are to each other).
-This makes the $\beta\eta$-rule elegantly straightforward.
-Besides $\lambda$, $\Pi$ and $\Sigma$ are also using binders and are formulated in a similar style.
+For these two types using binder, de Bruijn indices emphasizes its
+difference from the conventional setting---we directly use $\lambda t$
+as a function term where $t$ is the body, no binder indication is
+required---to refer to variables in the context, we use $\pi_1,\pi_2$
+from substitution calculus instead of using their names in the
+conventional setting.  For example, the identity function $\lambda x. x$
+is actually $\lambda(\pi_2 "id")$. $"app"~t$ is even more different than
+that in the original setting---we don't require an application argument
+now. The reason is, (1) we asked for argument in conventional setting so
+we can use meta-level substitution as part of the dynamics, but here we
+use explicit substitution. If the user wants to substitute some
+arguments, they can just use $[\gamma]$ instead. (2) Formulated in this
+way, "app" and $\lambda$ are symmetrical inverse to each other (just like
+how $El$ and "c" are to each other). This makes the $\beta\eta$-rule
+elegantly straightforward. Besides $\lambda$, $\Pi$ and $\Sigma$ are
+also using binders and are formulated in a similar style.
 
@@
 requires \textit{de Bruijn lifting}: for a given substitution
-$\goodTerm{\Gamma}{\sigma}{\Delta}$ and a type
+$\goodTerm{\Gamma}{\gamma}{\Delta}$ and a type
 $\goodType{\Delta}{A}{}$, we define its de Bruijn lifting
-$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ . Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.
+$\goodTerm{\Gamma,A[\gamma]}{\gamma^\uparrow := (\gamma \circ \pi_1, \pi_2)}{\Delta, A}$ . Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.
 
@@
   \begin{alignat*}{3}
-    \text{Inductive Type Signature}   \quad & \tau       &&= w\cdot~|~w^+~\tau~A~B~|\tau[\gamma] \\
-    \text{Linkage Signature}          \quad & \sigma       &&= \nu\cdot~|~\nu^+~\sigma~\{f\}~T~|~\sigma[\gamma]~|~p_1\nu~\sigma~| \\
+    \text{Inductive Type Signature}   \quad & \tau       &&= w\cdot~|~w^+~\tau~A~B~|~\tau[\gamma] \\
+    \text{Linkage Signature}          \quad & \sigma       &&= \nu\cdot~|~\nu^+~\sigma~\{f\}~T~|~\sigma[\gamma]~|~p_1\nu~\sigma \\
     \text{Seal}          \quad & f       &&= t~|~p_f\nu'~\sigma\\    
-    \text{Types}   \quad & A, B, T        &&= \bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\\
-    \text{Terms}   \quad & t, t_1, t_2    &&= "W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t  \\
+    \text{Types}   \quad & A, B, T        &&= \bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\YZ{No universes? And what is p1 nu′?}\\
+    \text{Terms}   \quad & t, t_1, t_2    &&= "W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t
   \end{alignat*}
@@
 
-We base our formalization of the inductive type on W-type~\cite{martin1982constructive}. 
-W-type, despite of the simplicity of its formulation, is able to construct a wide range of inductive type in extensional setting~\cite{martin1984intuitionistic,dybjer1997representing,abbott2004representing} and intensional setting (with the help of identity type~\cite{hugunin2021not}).
-Recall in W-type, a pair of type $x : A
-\vdash B(x)$ decides a W-type. To simulate extensible inductive types, we
+We base our formalization of inductive types on W-type~\cite{martin1982constructive}.
+W-types, despite their simplicity, are able to construct a wide range of
+inductive types in extensional setting~\cite{martin1984intuitionistic,dybjer1997representing,abbott2004representing}
+and intensional setting (with the help of identity types~\cite{hugunin2021not}).
+Recall with W-type, a pair of type $x : A \vdash B(x)$ decides a W-type.
+To simulate extensible inductive types, we
 have to equip our inductive types with multiple constructors, and for
-simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ :
-(x : Aᵢ) → (Bᵢ x → W) → W}.
+simplicity, each constructor has a fixed form
+\mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}.
 Thus, we use a `list' of pairs ($\goodWSig{\Gamma}{w^+~A_i~B_i~\tau}{}$) as the signature of one
-inductive type. Given an inductive type signature, we can use
-\ruleref{WSig-Proj} to extract the corresponding $A_i, B_i$.
+inductive type. Given the signature $\tau$ of an inductive type, \ruleref{WSig-Proj}
+extracts the corresponding $A_i, B_i$ via $\pi_1^i~\tau$ and $\pi_2^i~\tau$.
 
-Of course we need to put an eye on the substitution law for each piece
-of the syntax. Here $\bW \sigma$ is a large singleton
+Of course, we need to put an eye on the substitution law for each piece
+of the syntax. Here, $\bW \tau$ is a large singleton
 type~\cite{stone2000}, and it only has one ``element''---the inductive
 type itself, inside. We formulate in this cumbersome way because we want
-to have inductive type to be a field element and exposing the
-constructors at the type level for later \textit{exhaustiveness checking} (Since otherwise, if an inductive type as field element has type
-$\cU$, then no information about the inductive type is exposed).
+to have inductive types as family members and expose the
+constructors at the type level for later \textit{exhaustiveness checking}
+(otherwise inductive types as family members would have type~$\cU$, and
+no information about the inductive types could be exposed).
 
@@
 {\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
-\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}
 \\
@@
     \Sigma (\cC \sigma) T["sf" \ f]}{}
-\\\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}
+\\\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}\YZ{cL or cC?}
+\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}\YZ{nothing after colon?}
 }
@@
 
-\textit{A side note}: before we dive in more details of our metatheory for linkage (especially the meaning of sealing and $\cC$), we have to mention that, the challenge of incorporating family polymorphism with inductive type mentioned in \ref{chg:extensible-inductive-type} mutates in metatheory because of de Bruijn indices.
-(Using named variables), consider the example of a linkage $\{a~:~\bW \sigma; b_1 : a \to T;b_2 ..\}$ with a bunch of $b_i$ \textit{using} $a$.
-The tension we need to solve here is between \textit{exhaustiveness checking} and \textit{inheritance (code reuse)}---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to construct $b_i$ term in the context of $a : \bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that and thus we hope $b_i$ constructed in the context of $a : \cU$ (and thus later this $a$ can be compatible with inductive type of any signature).
-Our metatheory formulation of linkage is mainly trying to solve this tension.
+\textit{A side note}: before we dive in more details of our metatheory
+for linkage (especially the meaning of sealing and $\cC$), we have to
+mention that, the challenge of incorporating family polymorphism with
+inductive type mentioned in \ref{chg:extensible-inductive-type} mutates
+in metatheory because of de Bruijn indices. (Using named variables),
+consider the example of a linkage $\{a~:~\bW \sigma;\ b_1 : a \to T;\
+b_2 ..\}$ with a bunch of $b_i$ \textit{using} $a$. The tension we need
+to solve here is between \textit{exhaustiveness checking} and
+\textit{inheritance (code reuse)}---on one hand, if we want to make sure
+$b_i$ is pattern matching correctly, then we have to construct $b_i$
+term in the context of $a : \bW\sigma$; on the hand, if we want to let
+$b_i$ to be inherited into a derived family with extended inductive type
+$a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the
+context of $b_i$ is stopping that and thus we hope $b_i$ constructed in
+the context of $a : \cU$ (and thus later this $a$ can be compatible with
+inductive type of any signature). Our metatheory formulation of linkage
+is mainly trying to solve this tension.
 
-We use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).
-$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, according to \ruleref{Compile}, is a deeply nested dependent pair.
-$\cCt \sigma$ provide the concrete module of type $\cC\sigma$ given a linkage of type $\cL\sigma$.
+We use $\cL$ to retrieve the type of some linkage of a given signature,
+and $\cC$ to get the (existential) type of the compiled linkage.
+$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$,
+according to \ruleref{Compile}, is a deeply nested dependent pair.
+$\cCt \sigma$ provides the concrete module of type $\cC\sigma$ given a
+linkage of type $\cL\sigma$.
 
@@
 ability to do abstraction. Concretely speaking, (using named variables)
-we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to
-T\})}{h}{\cL(\{a~:~\cU; b~:~a \to T\})}$\footnote{If we had one,
-consider the case $T$ is bottom type and $\sigma$ is empty record and
+we cannot have $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma;~b~:~a \to
+T\})}{h}{\cL(\{a~:~\cU;~b~:~a \to T\})}$, which would mean that any
+subsequent field such as $b$ would be universally quantified under
+$a\,:\,\cU$ and thus leads to inconsistency
+\footnote{%
+Consider the case $T$ is bottom type and $\sigma$ is empty record and
 thus $a$ is initially bottom type. Now the direct projection using
@@
 ``abstract interface'' (i.e. something like $\cL(\{a:\cU~;~b:a \to T\})$) and achieve \textit{code reuse}.
-\YZ{It reads as if you want to have 'h : L\{a : U, b : a -> T\}'.
-But it is problematic because it leads to inconsistency, as you describe in a footnote.}\EDJreply{Good point.
-I rephrased it so the introduction of sigma type is smoother.
-Please check.}
-\\ The good news, we don't have to stick with $\cL(\sigma)$ to express the abstract interface. We can also use sigma type, and actually 
+
+The good news is that we don't have to use $\cL(\sigma)$ to
+express the abstract interface. We can also use $\Sigma$-types, and actually
 we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T)
-}{h}{(\Sigma a : \cU, a \to T)}$ because of the rules for sigma type.\footnote{The intro rule for sigma type decides that there is no such implicit universal quantification happening.}
-\YZ{This sentence still speaks of overriding. Explain in terms of projection instead?}\EDJreply{Fixed. Please check.}
-Thus to allow abstraction happens, we need $\cC\sigma$ to compile a
-linkage to module as a part of the context and we also need
+}{h}{(\Sigma a : \cU, a \to T)}$ because of the rules for
+$\Sigma$-types.\footnote{The intro rule for $\Sigma$-types decides that
+there is no such implicit universal quantification happening.}
+So to enable abstraction, we need $\cC\sigma$ to compile a
+linkage to module as part of the context, and we also need
 \ruleref{Lnkg Compile} to specify the compilation from linkage to
@@
 
-This is also reflected in the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}.
+This is also reflected in the implementation of our plugin: we know
+every field is defined in the context of the former fields, and our
+implementation makes every field depend on \textbf{the
+compiled module of the already-defined-part of the family}. 
 %\YZ{Why is Csigma needed?
@@
 \to T\}) \to (\Sigma a:\bW \sigma, a \to T)$, we still need $(\Sigma
-a:\bW \sigma, a \to T) \to (\Sigma a:\cU \sigma, a \to T)$ to complete
+a:\bW \sigma, a \to T) \to (\Sigma a:\cU, a \to T)$ to complete
 this shifting of the context. This is sealing's job.

### 2022-07-26 coqimpl.tex

@@
   this requirement, but also the "self vs. consistency" challenge.
-}
+}\EDJreply{Yes. Indivially compiling family member is to achieve both requirements}
 
@@
   then having Coq check it.
-}
+}\EDJreply{Yes! If my phrasing doesn't make things clear (i.e. our plugin is doing typechecking),
+feel free to change the phrasing for me}
 Note that, at this point of the example,
@@
   Why is this special treatment needed for compiling FRecursor?
-}
+}\EDJreply{because the non-special treatment cannot be applied to recursor --
+  the non-special treatment will require the recursor to be *exported* at the interface}
 

### 2022-07-26 syntax.tex
Reply the comments.

@@
 
+We also have identity type as part of our core syntax from the original MLTT.
+but we omit its presentations here---we follow~\citet{kaposi2019gluing} and use based path induction~\citet{program2013homotopy} as the elimination rule.
+
 \subsubsection{\textbf{Core Syntax of the Extended MLTT}}
@@
     \text{Seal}          \quad & f       &&= t~|~p_f\nu'~\sigma\\    
-    \text{Types}   \quad & A, B, T        &&= \bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\YZ{No universes? And what is p1 nu′?}\\
-    \text{Terms}   \quad & t, t_1, t_2    &&= "W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t
+    \text{Types}   \quad & A, B, T        &&= ...~|~\bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\YZ{No universes?
+And what is p1 nu′?}\EDJreply{Universe is described in Figure 9.``Overview of Instrinsically-typed Syntax for vanilla MLTT``.
+p_1 nu' is getting the resulting abstracted type after sealing.
+You can see that from Sig Proj rule.}\\
+    \text{Terms}   \quad & t, t_1, t_2    &&= ...~|~"W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t
   \end{alignat*}
@@
     \Sigma (\cC \sigma) T["sf" \ f]}{}
-\\\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}\YZ{cL or cC?}
-\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}\YZ{nothing after colon?}
+\\\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cC \nu\cdot}\YZ{cL or cC?}\EDJreply{cC. You are right.}
+\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}\YZ{nothing after colon?}\EDJreply{There should be something but the space is not enough.}
 }

### 2022-07-27 syntax.tex
Adding comments.
Add Beta equational rules for linkage.
edits & comments

--- 2022-07-26 syntax.tex
@@
 ability to do abstraction. Concretely speaking, (using named variables)
-we cannot have $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma;~b~:~a \to
-T\})}{h}{\cL(\{a~:~\cU;~b~:~a \to T\})}$, which would mean that any
+we cannot have $\goodTerm{\Gamma} %, \cL(\{a~:~\bW \sigma;~b~:~a \to T\})}
+{h}{\cL(\{a~:~\cU;~b~:~a \to T\})}$. It would otherwise mean that any
 subsequent field such as $b$ would be universally quantified under
-$a\,:\,\cU$ and thus leads to inconsistency
-\footnote{%
+$a\,:\,\cU$, leading to inconsistency.\footnote{%
 Consider the case $T$ is bottom type and $\sigma$ is empty record and
@@
 arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top
-\to \bot}$ is a contradiction}.
+\to \bot}$ is a contradiction}
 However, this underivable term is important for us to abstract the
-concrete inductive type so that the later field can depend on the
-``abstract interface'' (i.e. something like $\cL(\{a:\cU~;~b:a \to T\})$) and achieve \textit{code reuse}.
+concrete inductive type so that a later field can depend on the
+``abstract interface'' (i.e., something like $\cL(\{a:\cU~;~b:a \to T\})$)
+and achieve \textit{code reuse}.
 
@@
 
-For notation clarity, we will sometimes use $"sf'"$ to emphasize
-that we consider that seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term $\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]} $ and "sf" notation (used in \ruleref{Compile}) is defined as \\ $\goodSub{\Gamma, \cC \sigma}{(sf~s) := (\pi_1, sf'~s)}{\Gamma, A}$.\YZ{What is sf and sf'?
-sf first appears in Compile rules, so explain sf there.}\EDJreply{Fixed.
-They both are explained at here now}
+For notation clarity, we will sometimes use $sf'$ to emphasize that we
+consider that the seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term
+$\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]}$.\YZ{sf' seems unnecessary, since sf' s ad s are the same?}
+Notation "sf"
+(used in \ruleref{Compile}) is defined as the substitution $\goodSub{\Gamma,
+\cC \sigma}{(sf~s) := (\pi_1, sf'~s)}{\Gamma, A}$.
 
@@
 inheritance. Thus sealing itself can be considered as proving 
-abstractions on sigma type.
+abstractions on $\Sigma$-types.
 
@@
 
-Now, together with sealing, we can read \ruleref{Sig Add} and \ruleref{Lnkg Add} out clearly: given a constructed linkage $o : \cL\sigma$ of signature $\sigma$, the next added field $t : T$ will have to be in the context of the abtraction $A$ of the compilation $\cC\sigma$ (witnessed by the seal $f$).
-The ``concatenation'' of the $o$ and $t$ leads to a new linkage, of which the signature is the ``concatenation'' of the $\sigma$ and $T$ (we also record the sealing $f$ for simplicity).
-The key here is our signature, will \textit{record} this abstraction witness $f$, thus any projected member will still be in the abstracted context and thus possibly \textit{inheritable}.
-The construction of linkage in \ruleref{Lnkg Add} requires $t$ to \textbf{only} refer to the (abstraction $A$ of) former part of the defining family $o : \cL\sigma$, this is to avoid inconsistency mentioned in \ref{chg:consistency} in metatheory.
+Now, together with sealing, we can read \ruleref{Sig Add} and
+\ruleref{Lnkg Add} out clearly: given a constructed linkage $o : \cL\sigma$
+of signature $\sigma$, the next added field $t : T$ will have
+to be in the context of the abstraction $A$\YZ{Why do Lnkg Add and Sig Add leave A unspecified? I thought this A corresponds to 'self'.}
+of the compilation $\cC\sigma$ (witnessed by the seal $f$).
+The ``concatenation'' of the
+$o$ and $t$ leads to a new linkage, of which the signature is the
+``concatenation'' of the $\sigma$ and $T$ (we also record the sealing
+$f$ for simplicity). The key here is our signature, will \textit{record}
+this abstraction witness $f$, thus any projected member will still be in
+the abstracted context and thus possibly \textit{inheritable}.  The
+construction of linkage in \ruleref{Lnkg Add} requires $t$ to
+\textbf{only} refer to the (abstraction $A$ of) former part of the
+defining family $o : \cL\sigma$, this is to avoid inconsistency
+mentioned in \ref{chg:consistency} in metatheory.
 
-This sealing is the general version of the sealed family implemented in the plugin.
-Our plugin will deduce a sealed ``interface'' for a sealed family using a fixed set of rules but here sealing in metatheory can be arbitrary local morphism.
+This sealing is the general version of the sealed family implemented in
+the plugin. Our plugin will deduce a sealed ``interface''\YZ{Does this 'interface' correspond to A?}
+for a sealed family using a fixed set of rules, but here sealing in
+metatheory can be arbitrary local morphism.\YZ{why?}
 With the definition of sealing judgement in mind, we can see
@@
 when defining new fields in our plugin, and this "self__" is exactly the
-$A$ in the context of the two fields---in our implementation, there is only sealing of inductive type happening (where each inductive type we only know \mintinline{Coq}{T : Set}).
+$A$ in the context of the two fields---in our implementation, sealing
+happens only to inductive types (where each inductive type we only know
+\mintinline{Coq}{T : Set}).
 

### 2022-07-27 syntax.tex

@@
 \\ \goodSeal{\Gamma}{p_f\nu \  \sigma}{p_1\nu \  \sigma}{p_1\nu' \sigma}
-\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}
+\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}\YZ{Add beta or eta rules here to explain syntax term clearer}
+}
+$$
+
+$$ 
+\Rule[name=Sig Proj Beta]
+{\goodSig{\Gamma}{\sigma}{n} 
+\quad \goodSeal{\Gamma}{f}{\sigma}{A}
+\quad \goodType{\Gamma, A}{T}{i}}
+{\goodSig{\Gamma}{p_1\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv \sigma}{n}
+\\ \goodType{\Gamma}{p_1\nu' \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv A}{}
+\\ \goodSeal{\Gamma}{p_f\nu \  (\nu^+ \ \sigma \ \{f\} \ T) \equiv f}{\sigma}{A}
+\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv T}{}\YZ{Add beta or eta rules here to explain syntax term clearer}\EDJreply{Done}
 }
@@
 $$
-
+$$ 
+\Rule[name=Lnkg Proj Beta]
+{\goodTerm{\Gamma}{o}{\cL \sigma} 
+\quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
+ \quad \goodTerm{\Gamma, A}{t}{T}}
+{\goodTerm{\Gamma}{p_1\mu \ (\mu^+ \ o \ \{f\} \ t) \equiv o}{\cL \sigma}
+\\ \goodTerm{\Gamma, A}{p_2\mu \  (\mu^+ \ o \ \{f\} \ t) \equiv t}{T}
+}
+$$
 $$

### 2022-07-27 syntax.tex
Reply the comments. Do changes.

@@
 {\goodSig{\Gamma}{\sigma}{n} 
+ \quad \goodType{\Gamma}{A}{}
  \quad \goodSeal{\Gamma}{f}{\sigma}{A}
@@
 \\ \goodSeal{\Gamma}{p_f\nu \  \sigma}{p_1\nu \  \sigma}{p_1\nu' \sigma}
-\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}\YZ{Add beta or eta rules here to explain syntax term clearer}
+\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}\YZ{Add beta or eta rules here to explain syntax term clearer}\EDJreply{Done}
 }
@@
 {\goodSig{\Gamma}{\sigma}{n} 
+\quad \goodType{\Gamma}{A}{}
 \quad \goodSeal{\Gamma}{f}{\sigma}{A}
@@
 \\ \goodSeal{\Gamma}{p_f\nu \  (\nu^+ \ \sigma \ \{f\} \ T) \equiv f}{\sigma}{A}
-\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv T}{}\YZ{Add beta or eta rules here to explain syntax term clearer}\EDJreply{Done}
+\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv T}{}
 }
@@
 { \goodTerm{\Gamma}{o}{\cL \sigma} 
+\quad \goodType{\Gamma}{A}{}
 \quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
@@
 {\goodTerm{\Gamma}{o}{\cL \sigma} 
+\quad \goodType{\Gamma}{A}{}
 \quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
@@
 consider that the seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term
-$\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]}$.\YZ{sf' seems unnecessary, since sf' s ad s are the same?}
+$\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]}$.\YZ{sf' seems unnecessary, since sf' s ad s are the same?}\EDJreply{Yes.
+Technically it is pretty unnecessary.
+It is just a bit weird to have $\goodTerm{\Gamma, \cC \sigma}{s}{A[\pi_1]}$ and $\goodSeal{\Gamma}{s}{\sigma}{A}$ at the same time because it is one derivation but turns out to be two judgements, for intrinsically typed.
+I can also remove ``sf'' if you don't find it weird.
+I initially having ``sf'' to just *emphasize* ``sf s'' should be considered as a term  and ``s'' a sealing}
 Notation "sf"
@@
 of signature $\sigma$, the next added field $t : T$ will have
-to be in the context of the abstraction $A$\YZ{Why do Lnkg Add and Sig Add leave A unspecified? I thought this A corresponds to 'self'.}
+to be in the context of the abstraction $A$\YZ{Why do Lnkg Add and Sig Add leave A unspecified?
+I thought this A corresponds to 'self'.}\EDJreply{It is not unspecified.
+I now change it so that it is more clear.
+It is parametric as well.
+Just note that when specifying the abstraction `A', the seal `f' has to be specified at the sametime.
+(Thus I omitted it initially because it is implicitly specified w.r.t.
+the speicifcation of `f') }
 of the compilation $\cC\sigma$ (witnessed by the seal $f$).
@@
 This sealing is the general version of the sealed family implemented in
-the plugin. Our plugin will deduce a sealed ``interface''\YZ{Does this 'interface' correspond to A?}
+the plugin. Our plugin will deduce a sealed ``interface'' (which is the "A" in \ruleref{Sig Add}) \YZ{Does this 'interface' correspond to A?}\EDJreply{Yes.}
 for a sealed family using a fixed set of rules, but here sealing in
-metatheory can be arbitrary local morphism.\YZ{why?}
+metatheory can be arbitrary local morphism  $\goodTerm{\Gamma , \cC
+\sigma}{\_}{A[\pi_1]}$.\YZ{why?}\EDJreply{local morphism is exactly term of this form}
 With the definition of sealing judgement in mind, we can see

### 2022-07-27 syntax.tex
resolve comments

--- 2022-07-27 syntax.tex
@@
 }
-{\goodSig{\Gamma}{"RecSig"~\sigma~R}{n}\YZ{Why does this Sig not have a superscript?}\EDJreply{Fixed now.}}
+{\goodSig{\Gamma}{"RecSig"~\sigma~R}{n}}
 $$
@@
 
-In our meta-theory, we only formulate non-dependent eliminator for simplicity, i.e., the
-\textit{recursor}.\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}\EDJreply{I don't think there is technical difficulties to do that.
+In our meta-theory, for simplicity we only formulate non-dependent eliminators, i.e.,
+\textit{recursors}.
+%\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}\EDJreply{I don't think there is technical difficulties to do that.
 The formulation of course will be much more horrifying if use dependent eliminator here.

### 2022-07-27 syntax.tex
resolve comments

@@
     \text{Seal}          \quad & f       &&= t~|~p_f\nu'~\sigma\\    
-    \text{Types}   \quad & A, B, T        &&= ...~|~\bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\YZ{No universes?
-And what is p1 nu′?}\EDJreply{Universe is described in Figure 9.``Overview of Instrinsically-typed Syntax for vanilla MLTT``.
-p_1 nu' is getting the resulting abstracted type after sealing.
-You can see that from Sig Proj rule.}\\
+    \text{Types}   \quad & A, B, T        &&= ...~|~\bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\\
     \text{Terms}   \quad & t, t_1, t_2    &&= ...~|~"W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t
@@
   \begin{align*}
-&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over.
-Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point.
-I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
+&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
   \end{align*}
@@
 \\ \goodSeal{\Gamma}{p_f\nu \  \sigma}{p_1\nu \  \sigma}{p_1\nu' \sigma}
-\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}\YZ{Add beta or eta rules here to explain syntax term clearer}\EDJreply{Done}
+\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}
 }
@@
     \Sigma (\cC \sigma) T["sf" \ f]}{}
-\\\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cC \nu\cdot}\YZ{cL or cC?}\EDJreply{cC. You are right.}
+\\\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cC \nu\cdot}
 \\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}\YZ{nothing after colon?}\EDJreply{There should be something but the space is not enough.}

### 2022-07-27 syntax.tex
comments

@@
 \\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv T}{}
-}
+}\YZ{p_f nu' or p_f nu? In the syntax figure, you only have p_f nu, I think.}
 $$
@@
 I can also remove ``sf'' if you don't find it weird.
-I initially having ``sf'' to just *emphasize* ``sf s'' should be considered as a term  and ``s'' a sealing}
+I initially having ``sf'' to just *emphasize* ``sf s'' should be considered as a term  and ``s'' a sealing}\YZreply{
+  I'm still confused about the difference among s, sf s and sf' s.
+  I thought s and sf' s were technically the same; they are both terms of type A[pi1] in a context that contains cC sigma.
+  I thought sf s was different because it is a substitution.
+}
 Notation "sf"

### 2022-07-27 syntax.tex
Reply the comments.

@@
 \\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv T}{}
-}\YZ{p_f nu' or p_f nu? In the syntax figure, you only have p_f nu, I think.}
+}\YZ{p_f nu' or p_f nu? In the syntax figure, you only have p_f nu, I think.}\EDJreply{p_f nu, I don't have p_f nu' here. the prime is just for specifying we are taken out the abstraction result.}
 $$
@@
   I thought sf s was different because it is a substitution.
+}\EDJreply{You are absolutely right. s and sf' s are techincally the same. sf s is a substitution.
+  About s and sf' s, I can say that, I don't allow something like $\goodSeal{\Gamma}{sf'~s}{\sigma}{A}$. The reason I don't allow it is only notational consideration. 
+  I mean technically we can consider them the same. I can just remove all the appearances of sf' if you want and find it less confusing.
 }

### 2022-07-27 syntax.tex
resolve comments

@@
     \text{Linkage Signature}          \quad & \sigma       &&= \nu\cdot~|~\nu^+~\sigma~\{f\}~T~|~\sigma[\gamma]~|~p_1\nu~\sigma \\
-    \text{Seal}          \quad & f       &&= t~|~p_f\nu'~\sigma\\    
+    \text{Seal}          \quad & f       &&= t~|~p_f\nu~\sigma\\    
     \text{Types}   \quad & A, B, T        &&= ...~|~\bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\\
@@
 \\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv T}{}
-}\YZ{p_f nu' or p_f nu? In the syntax figure, you only have p_f nu, I think.}\EDJreply{p_f nu, I don't have p_f nu' here. the prime is just for specifying we are taken out the abstraction result.}
+}
 $$
@@
   I mean technically we can consider them the same. I can just remove all the appearances of sf' if you want and find it less confusing.
-}
+}\YZreply{Yes, sf' is confusing to me.}
 Notation "sf"

### 2022-07-27 syntax.tex
comments

@@
 (Thus I omitted it initially because it is implicitly specified w.r.t.
-the speicifcation of `f') }
+the speicifcation of `f') }\YZreply{
+  Changed what? I think Lnkg Add and Sig Add were not changed.
+  I still don't see how sealing converts \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}
+  and how the latter is now used as the abstracted A.
+}
 of the compilation $\cC\sigma$ (witnessed by the seal $f$).
@@
 $f$ for simplicity). The key here is our signature, will \textit{record}
-this abstraction witness $f$, thus any projected member will still be in
+this abstraction witness $f$,\YZ{Why is it necessary to record the witness f? I don't see how p_f nu is needed anywhere.}
+thus any projected member will still be in
 the abstracted context and thus possibly \textit{inheritable}.  The
@@
 This sealing is the general version of the sealed family implemented in
-the plugin. Our plugin will deduce a sealed ``interface'' (which is the "A" in \ruleref{Sig Add}) \YZ{Does this 'interface' correspond to A?}\EDJreply{Yes.}
+the plugin. Our plugin will deduce a sealed ``interface'' (which is the "A" in \ruleref{Sig Add})
 for a sealed family using a fixed set of rules, but here sealing in
 metatheory can be arbitrary local morphism  $\goodTerm{\Gamma , \cC
-\sigma}{\_}{A[\pi_1]}$.\YZ{why?}\EDJreply{local morphism is exactly term of this form}
+\sigma}{\_}{A[\pi_1]}$.
 With the definition of sealing judgement in mind, we can see

### 2022-07-27 syntax.tex
comments

@@
 }
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
-$$
-
-In our meta-theory, for simplicity we only formulate non-dependent eliminators, i.e.,
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}\YZ{
+(1) Use \tau for inductive types and \sigma for linkage signatures.
+(2) Try using the same metavariables (j and n) as in the rules on W-types.
+(3) In Rec-Constr, no checking of \Gamma \vdash T : \bW \tau?
+(4) Why are RecSig1, RecSig, Wrec, prjR not listed in the syntax figure?
+}
+$$
+
+In our metatheory, for simplicity we only formulate non-dependent eliminators, i.e.,
 \textit{recursors}.
@@
 }
-Like our plugin implementation, each recursor is constructed by a linkage containing handlers dealing with each
-constructor and code reuse for recursor is delegated to linkage
-inheritance. This greatly simplifies the metatheory development.
+Like our plugin implementation, each recursor is constructed by a
+linkage containing handlers dealing with each constructor, and code
+reuse for recursor is delegated to linkage inheritance.
+This greatly simplifies the metatheory development.
 
@@
 used to construct a concrete recursor given a specific linkage of aggregated
-handlers. All these rules satisfy substitution laws and omitted here. Thus, we are still using \textit{decoupling} of the syntax of the case
+handlers. All these rules satisfy substitution laws, which are omitted here.
+Thus, we are still using \textit{decoupling} of the syntax of the case
 handlers and exhaustiveness checking to resolve

### 2022-07-27 syntax.tex
Reply the comments.

@@
   and how the latter is now used as the abstracted A.
-}
+}\EDJreply{I added the $\goodType{\Gamma}{A}{}$  as you suggested. 
+\cC \{a : \bW sigma ; ..\} is definitionally equal to a sigma type as Compile rule suggested. 
+If you expand the definition, you will see two sigma type. 
+You are asking where is the function \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}, which by definition is requesting (Sigma (a : \bW sigma ; ..)) to (Sigma (a : \cU ; ..)). 
+This is something doable in vanilla MLTT, with the help of transforming \bW sigma to \cU, by rule `Ind-Type'. }
 of the compilation $\cC\sigma$ (witnessed by the seal $f$).
@@
 $f$ for simplicity). The key here is our signature, will \textit{record}
-this abstraction witness $f$,\YZ{Why is it necessary to record the witness f? I don't see how p_f nu is needed anywhere.}
+this abstraction witness $f$,\YZ{Why is it necessary to record the witness f?
+I don't see how p_f nu is needed anywhere.}\EDJreply{1.
+this is for the sake of eta/beta rule (local soundness + local completeness); 2.It is actually used in the rule Compile (the last one), if Ct is compile a variable, then the f will be replace by p_f}
 thus any projected member will still be in
@@
 { \goodWSig{\Gamma}{\sigma}{N}
-\quad \goodType{\Gamma}{T}{\bW \sigma}
+\quad \goodTerm{\Gamma}{T}{\bW \sigma}
 \quad \goodType{\Gamma}{R}{}
@@
 (4) Why are RecSig1, RecSig, Wrec, prjR not listed in the syntax figure?
-}
+}\EDJreply{Thx. Working on it. I seem to forgot to add inheritance as well.}
 $$

### 2022-07-27 syntax.tex
Save.

@@
   and how the latter is now used as the abstracted A.
-}\EDJreply{I added the $\goodType{\Gamma}{A}{}$  as you suggested. 
-\cC \{a : \bW sigma ; ..\} is definitionally equal to a sigma type as Compile rule suggested. 
+}\EDJreply{I added the $\goodType{\Gamma}{A}{}$  as you suggested. \\
+\cC \{a : \bW sigma ; ..\} is definitionally equal to a sigma type as Compile rule suggested. \\
 If you expand the definition, you will see two sigma type. 
-You are asking where is the function \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}, which by definition is requesting (Sigma (a : \bW sigma ; ..)) to (Sigma (a : \cU ; ..)). 
+You are asking where is the function \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}, which by definition is requesting (Sigma (a : \bW sigma ; ..)) to (Sigma (a : \cU ; ..)). \\
 This is something doable in vanilla MLTT, with the help of transforming \bW sigma to \cU, by rule `Ind-Type'. }
@@
 I don't see how p_f nu is needed anywhere.}\EDJreply{1.
-this is for the sake of eta/beta rule (local soundness + local completeness); 2.It is actually used in the rule Compile (the last one), if Ct is compile a variable, then the f will be replace by p_f}
+this is for the sake of eta/beta rule (local soundness + local completeness); 2.It is actually used in the rule Compile (the last one), if Ct is compile a variable, say ``Ct v'', then this ``Ct v = (Ct p_1 v), (t[sf p_f v][...])'', the f will actually be replaced by p_f; 3.
+p_f can be used by the user.
+It doesn't have to interact with other typing rules.
+}
 thus any projected member will still be in

### 2022-07-28 syntax.tex
Reply the comments
reply to comment

@@
 You are asking where is the function \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}, which by definition is requesting (Sigma (a : \bW sigma ; ..)) to (Sigma (a : \cU ; ..)). \\
-This is something doable in vanilla MLTT, with the help of transforming \bW sigma to \cU, by rule `Ind-Type'. }
+This is something doable in vanilla MLTT, with the help of transforming \bW sigma to \cU, by rule `Ind-Type'. And this function's type should be directly the sealing type.}\YZreply{
+  My confusion is this: it seems that in Lnkg Add and Sig Add, A could well be the top type and f a value of the top type. In this sense, A seems unspecified.
+}
 of the compilation $\cC\sigma$ (witnessed by the seal $f$).
@@
 $f$ for simplicity). The key here is our signature, will \textit{record}
-this abstraction witness $f$,\YZ{Why is it necessary to record the witness f?
+this abstraction witness $f$,
+%\YZ{Why is it necessary to record the witness f?
 I don't see how p_f nu is needed anywhere.}\EDJreply{1.

### 2022-07-28 syntax.tex
maybe?
Reply the comment. The maybe is correct

@@
   My confusion is this: it seems that in Lnkg Add and Sig Add, A could well be the top type and f a value of the top type. In this sense, A seems unspecified.
-}
+  Or maybe in the metatheory, it is the programmer's responsibility to find this A?
+}\EDJreply{You are absolutely right.
+In metatheory it is programmer's responsibility to specify the A they want.
+If they set A to top, they can do that but it will be hard for them to achieve meaningful program.
+\\
+A bit off-topic : this is not the only place that such specification is postponed to the user (in metatheory) another place is in inheritance when specifying upcasting function.}
 of the compilation $\cC\sigma$ (witnessed by the seal $f$).

### 2022-07-28 syntax.tex
additional repling comments
comment

@@
 to be in the context of the abstraction $A$\YZ{Why do Lnkg Add and Sig Add leave A unspecified?
-I thought this A corresponds to 'self'.}\EDJreply{It is not unspecified.
-I now change it so that it is more clear.
-It is parametric as well.
-Just note that when specifying the abstraction `A', the seal `f' has to be specified at the sametime.
-(Thus I omitted it initially because it is implicitly specified w.r.t.
-the speicifcation of `f') }\YZreply{
-  Changed what? I think Lnkg Add and Sig Add were not changed.
-  I still don't see how sealing converts \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}
-  and how the latter is now used as the abstracted A.
-}\EDJreply{I added the $\goodType{\Gamma}{A}{}$  as you suggested. \\
-\cC \{a : \bW sigma ; ..\} is definitionally equal to a sigma type as Compile rule suggested. \\
-If you expand the definition, you will see two sigma type. 
-You are asking where is the function \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}, which by definition is requesting (Sigma (a : \bW sigma ; ..)) to (Sigma (a : \cU ; ..)). \\
+I thought this A corresponds to 'self'.}\EDJreply{You are asking where is the function \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}, which by definition is requesting (Sigma (a : \bW sigma ; ..)) to (Sigma (a : \cU ; ..)).
+\\
 This is something doable in vanilla MLTT, with the help of transforming \bW sigma to \cU, by rule `Ind-Type'. And this function's type should be directly the sealing type.}\YZreply{
@@
 \\
+Also the `f' is specified together with `A' by the user as well. \\
 A bit off-topic : this is not the only place that such specification is postponed to the user (in metatheory) another place is in inheritance when specifying upcasting function.}
@@
 \textit{exhaustiveness checking}. These fields together with inductive
-types are the ones sealing is responsible for hiding. After compilation
+types are the ones sealing is responsible for hiding.\YZ{Is sealing really hiding these fields (that use Wsup and Wrec), or is it that the concrete definition of an inductive type is not hidden for these fields so that they can assume knowledge about the inductive type's signature?}
+After compilation
 of linkage types into existential types, we can easily hide the
 \textit{concrete} definition of the inductive type while exposing all
-the fields using "Wsup" and "Wrec" like an abstract interface.
+the fields that use "Wsup" and "Wrec" as an abstract interface.
 Apparently, future fields relying on this abstract interface will have no

### 2022-07-28 syntax.tex
Reply the comments.

@@
 \textit{exhaustiveness checking}. These fields together with inductive
-types are the ones sealing is responsible for hiding.\YZ{Is sealing really hiding these fields (that use Wsup and Wrec), or is it that the concrete definition of an inductive type is not hidden for these fields so that they can assume knowledge about the inductive type's signature?}
+types are the ones sealing is responsible for hiding.\YZ{Is sealing really hiding these fields (that use Wsup and Wrec), or is it that the concrete definition of an inductive type is not hidden for these fields so that they can assume knowledge about the inductive type's signature?}\EDJreply{More rigourously speaking, sealing is only hiding the $\tau$ appearing in $\bW \tau$, it never tries to hide (or make absence of) any field.
+But for the people that are only looking at the interface/type, the people cannot if a field is a constructor (using Wsup) or a normal function.
+\\
+The concrete definition of an inductive type (i.e.
+$\bW \tau$) is definitely hidden and those inductive type will thus have type $\cU$.
+After the $\tau$ is hidden, the programmer is assuming knowledge of the inductive type.
+}
 After compilation

### 2022-07-28 syntax.tex
i.e. → e.g.

--- 2022-07-28 syntax.tex
@@
 
-There are also cases we don't need to seal anything (i.e., for doing
+There are also cases we don't need to seal anything (e.g., for doing
 exhaustiveness checking when using recursor), and thus \ruleref{Seal-Id} is

### 2022-07-28 syntax.tex
reply

@@
 After the $\tau$ is hidden, the programmer is assuming knowledge of the inductive type.
-}
+}\YZreply{The sentence feels confusing because "these fields" are not sealed; rather, as you explain below, there's no need to seal anything when, say, doing exhaustiveness checking using Wrec.}
 After compilation

### 2022-07-28 syntax.tex
Reply the comments

--- 2022-07-28 syntax.tex
@@
 After the $\tau$ is hidden, the programmer is assuming knowledge of the inductive type.
+}\YZreply{The sentence feels confusing because "these fields" are not sealed; rather, as you explain below, there's no need to seal anything when, say, doing exhaustiveness checking using Wrec.}\EDJreply{Oh I see.
+The sentence here I say "fields and the inductive type" are sealed together because the sealing function will be between big sigma (Sig T:$\bW \sigma$, f : El (W T) -> El (W T)) and (Sig T:$\cU$, f : El T -> El T).
+That is the reason I say ``sealed together'' here.
+The (constructor) fields are not literally hidden, but their type are changed *together* with the former type T.
 }

### 2022-07-28 syntax.tex
Additional Comments.

--- 2022-07-28 syntax.tex
@@
 After the $\tau$ is hidden, the programmer is assuming knowledge of the inductive type.
+}\YZreply{The sentence feels confusing because "these fields" are not sealed; rather, as you explain below, there's no need to seal anything when, say, doing exhaustiveness checking using Wrec.}\EDJreply{Oh I see.
+The sentence here I say "fields and the inductive type" are sealed together because the sealing function will be between big sigma (Sig T:$\bW \sigma$, c1 : El (W T) -> El (W T)) and (Sig T:$\cU$, c1 : El T -> El T).
+That is the reason I say ``sealed together'' here.
+The (constructor) fields (e.g.
+c1) are not literally hidden, but their type are changed *together* with the former type T, because of type-dependency.
 }

### 2022-07-28 syntax.tex
resolve comment

--- 2022-07-28 syntax.tex
@@
 \textit{exhaustiveness checking}. These fields together with inductive
-types are the ones sealing is responsible for hiding.\YZ{Is sealing really hiding these fields (that use Wsup and Wrec), or is it that the concrete definition of an inductive type is not hidden for these fields so that they can assume knowledge about the inductive type's signature?}\EDJreply{More rigourously speaking, sealing is only hiding the $\tau$ appearing in $\bW \tau$, it never tries to hide (or make absence of) any field.
-But for the people that are only looking at the interface/type, the people cannot if a field is a constructor (using Wsup) or a normal function.
-\\
-The concrete definition of an inductive type (i.e.
-$\bW \tau$) is definitely hidden and those inductive type will thus have type $\cU$.
-After the $\tau$ is hidden, the programmer is assuming knowledge of the inductive type.
-}\YZreply{The sentence feels confusing because "these fields" are not sealed; rather, as you explain below, there's no need to seal anything when, say, doing exhaustiveness checking using Wrec.}
+types are the ones sealing is responsible for hiding.
+%\YZ{Is sealing really hiding these fields (that use Wsup and Wrec), or is it that the concrete definition of an inductive type is not hidden for these fields so that they can assume knowledge about the inductive type's signature?}\EDJreply{More rigourously speaking, sealing is only hiding the $\tau$ appearing in $\bW \tau$, it never tries to hide (or make absence of) any field.
+%}\YZreply{The sentence feels confusing because "these fields" are not sealed; rather, as you explain below, there's no need to seal anything when, say, doing exhaustiveness checking using Wrec.}\EDJreply{Oh I see.
+The sentence here I say "fields and the inductive type" are sealed together because the sealing function will be between big sigma (Sig T:$\bW \sigma$, c1 : El (W T) -> El (W T)) and (Sig T:$\cU$, c1 : El T -> El T).
+That is the reason I say ``sealed together'' here.
+The (constructor) fields (e.g.
+c1) are not literally hidden, but their type are changed *together* with the former type T, because of type-dependency.
+}
 After compilation

### 2022-07-28 syntax.tex
minor

@@
 
-Next we talk about inheritance judgement.
-Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, and in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e.
-we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}$, $\goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the premises of judgements when needed.
+Next we talk about inheritance judgements. Note that, there will be a lot
+of $\sigma_1, A_1$ and $\sigma_2, A_2$, and in these cases, $A_1$ is the
+result of the sealing of $\sigma_1$, i.e. we assume we have a term
+$\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}$,
+$\goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the premises of judgements
+when needed.
 
@@
 
-Inheritance are
-judgement so naturally second class citizen data (i.e. not something
-function can return and cannot be assigned to variables), and it looks
-very much like a function that transform a given linkage.
-Similar to plugin implementation and \ruleref{Lnkg Add}, we resolve \ref{chg:consistency} by making sure corresponding fields are in corresponding context during inheritance, and thus an overriding term can only refer to fields before the overriden field.
+Inheritance are judgements so naturally second class citizen data (i.e.,
+not something function can return and cannot be assigned to variables),
+and it looks very much like a function that transforms a given linkage.
+Similar to plugin implementation and \ruleref{Lnkg Add}, we resolve
+\ref{chg:consistency} by making sure that a derived family preserves the
+ordering of fields in its base family:
+in an inheritance judgment, an overriding term is checked in a context
+that contains those fields preceding it in the base family.
 
@@
 
-"inhinh" is special because it requires user to provide a proof of
+"inhinh" is special because it requires the user to provide a proof of
 ``upcasting'' from the context of the child to that of the parent.
@@
 
-Override "inhov" is even more special because in mundane OO programming language, we usually require that the overriding term has the same type/interface as the overriden one.
-Here we don't since the very reason we want the same type/interface is that we hope other inherited fields can use this overridden/late-binding field without breaking abstraction, and this reason is already managed by the premises of "inhinh", the upcasting proof $\uparrow^s$.
-Thus "inhov" is quite like "inhext", where overriding is just throwing away one parent field.
-This difference also show one of the distinct feature of our vanilla family inheritance compared to the mainstream OO inheritance---family inheritance don't need to introduce the concept of ``subclassing'' between families and thus no bounded polymorphism yet\footnote{Of course once trait and interface comes in, we can have bounded polymorphism back}, and code reuse is mainly achieved by late binding.
-Without ``subclassing'', there is no way of confusing inheritance and subtyping.
-Note that, "inhov" is also responsible for ``extending'' inductive type---the extended inductive type directly override the old ones when using metatheory to program, just like how plugin implementation works.
+Override "inhov" is even more special because in mundane OO programming
+language, we usually require that the overriding term has the same
+type/interface as the overriden one.  Here we don't since the very
+reason we want the same type/interface is that we hope other inherited
+fields can use this overridden/late-binding field without breaking
+abstraction, and this reason is already managed by the premises of
+"inhinh", the upcasting proof $\uparrow^s$. Thus "inhov" is quite like
+"inhext", where overriding is just throwing away one parent field. This
+difference also shows one of the distinct features of our vanilla family
+inheritance compared to the mainstream OO inheritance---family
+inheritance need not introduce the concept of ``subclassing''
+between families and thus no bounded polymorphism yet\footnote{Of course
+once trait and interface comes in, we can have bounded polymorphism
+back}, and code reuse is mainly achieved by late binding. Without
+``subclassing'', there is no way of confusing inheritance and subtyping.
+Note that, "inhov" is also responsible for ``extending'' inductive
+types---an extended inductive type directly overrides the old one when
+using metatheory to program, just like how plugin implementation works.
 

### 2022-07-28 syntactic-example.tex
wrap text

@@
 
-Here we explain how  our detailed construction in Figure \cref{fig:example-construction} is constructed.
-We first elaborate that how our inductive type "N" is written in the specific format (the style of W-type).
-We use 1, 0 here to indicate unit type and bottom type.
+Here we explain how  our detailed construction in Figure
+\cref{fig:example-construction} is constructed. We first elaborate that
+how our inductive type "N" is written in the specific format (the style
+of W-type). We use 1, 0 here to indicate unit type and bottom type.
 
-Next, we alias each part of the derivation with a name. Every derivation here is well-typed term. We also annotate in the original pseudocode the name of the related derivation.
+Next, we alias each part of the derivation with a name. Every derivation
+here is well-typed term. We also annotate in the original pseudocode the
+name of the related derivation.
 
-We start with the construction of "Osig" which is the signature of a standalone "O" constructor.
-With this we will have $"sig"_1, "obj"_1, "sig"_2, "obj"_2$ as the signature and object of the linkage with that inductive type, and the signature and the linkage once we export the concrete constructor $"O"_1$.
+We start with the construction of "Osig" which is the signature of a
+standalone "O" constructor. With this we will have $"sig"_1, "obj"_1,
+"sig"_2, "obj"_2$ as the signature and object of the linkage with that
+inductive type, and the signature and the linkage once we export the
+concrete constructor $"O"_1$.
 
-Now sealing comes in to seal these concrete component into an abstract interface $"sig"_2'$ for decoupling the following fields.
-To make sure the sealing exists, we take a look at the resulting $\cC "sig"_2$, and doing abstraction on it to get $"sl"_1$.
+Now sealing comes in to seal these concrete component into an abstract
+interface $"sig"_2'$ for decoupling the following fields. To make sure
+the sealing exists, we take a look at the resulting $\cC "sig"_2$, and
+doing abstraction on it to get $"sl"_1$.
 
-Now we construct the recursor, with the help of the handler module $"handler"_1$, with which we can construct the real recursor $"pred"_1$.
-Both detailed constructions are omitted.
-Then we again abstract the concrete recursor $"pred"_1$ away using $"sl"_4$ so that the field $"k"$ can applys it without resorting to the concrete recursor.
-Finally we have $"sig"_5$ as final signature.
+Now we construct the recursor, with the help of the handler module
+$"handler"_1$, with which we can construct the real recursor $"pred"_1$.
+Both detailed constructions are omitted. Then we again abstract the
+concrete recursor $"pred"_1$ away using $"sl"_4$ so that the field $"k"$
+can applys it without resorting to the concrete recursor. Finally we
+have $"sig"_5$ as final signature.
 
-Now we construct the inheritance judgement for $"A"_2$.
-Still, we start with signature for inductive type, "Nsig".
-But to ``extend'' the inductive type, we actually \textit{override} the inductive type with the enriched inductive type in $"inh"_1$.
-Because inductive type is overridden, inheritance on the constructors are not pausible and thus we need to override the old $"O"_1$ constructor using $"O"_2$ and extend with new constructor for "S", resulting $"inh"_3$.
-These newly overridden constructors are sealed by $"sl"_{22}$.
-However, look closely, our sealed abstract interface in the children still has more fields than the sealed abstract interface in the parent.
-We still need $\uparrow^s$ to indicate the ``compatibility'' between two abstract interfaces so that the inheritance from the parents can work.
-It can be understood as ``upcasting'', but this ``upcasting'' is used to upcast the context of the children.
-We immeidately see its usage in $"pN"_3$ and $"inh"_{in0}$.
-Upon this, we construct the complete inner-inheritance $"inh"_{in}$ that is responsible for extending "handler".
-Then we continue the construction for $"inh"_4$ that ``nest'' the inner-inheritance into the whole $"inh"_3$, and we re-do exhaustiveness checking when constructing $"inh"_5$ using $"inhov"$ and "Wrec".
-Since it is another recursor requiring concrete inductive definition, we do another abstraction on it using $"sl"_41$.
-After that, $\uparrow^s_2$ again witnesses the compatibility of the abstract interface, and the two are both used for constructing $"inh"_6$, which is responsible for inheriting field "k".
+Now we construct the inheritance judgement for $"A"_2$. Still, we start
+with signature for inductive type, "Nsig". But to ``extend'' the
+inductive type, we actually \textit{override} the inductive type with
+the enriched inductive type in $"inh"_1$. Because inductive type is
+overridden, inheritance on the constructors are not pausible and thus we
+need to override the old $"O"_1$ constructor using $"O"_2$ and extend
+with new constructor for "S", resulting $"inh"_3$. These newly
+overridden constructors are sealed by $"sl"_{22}$. However, look
+closely, our sealed abstract interface in the children still has more
+fields than the sealed abstract interface in the parent. We still need
+$\uparrow^s$ to indicate the ``compatibility'' between two abstract
+interfaces so that the inheritance from the parents can work. It can be
+understood as ``upcasting'', but this ``upcasting'' is used to upcast
+the context of the children. We immeidately see its usage in $"pN"_3$
+and $"inh"_{in0}$. Upon this, we construct the complete
+inner-inheritance $"inh"_{in}$ that is responsible for extending
+"handler". Then we continue the construction for $"inh"_4$ that ``nest''
+the inner-inheritance into the whole $"inh"_3$, and we re-do
+exhaustiveness checking when constructing $"inh"_5$ using $"inhov"$ and
+"Wrec". Since it is another recursor requiring concrete inductive
+definition, we do another abstraction on it using $"sl"_41$. After that,
+$\uparrow^s_2$ again witnesses the compatibility of the abstract
+interface, and the two are both used for constructing $"inh"_6$, which
+is responsible for inheriting field "k".
 
-The resulting constructed families are $\goodTerm{\cdot}{"obj"_5}{\cL"sig"_5}$ as "Family A" and \\ $\goodTerm{\cdot}{("inh"~"inh"_6~"obj"_5)}{\cL((\nu^+ \ "sig"_{41} \{"sl"_{41}\} \ "pN"_2[(\pi_1, \uparrow^s_2)]))}$ as "Family" $"A"_2$.
+The resulting constructed families are
+$\goodTerm{\cdot}{"obj"_5}{\cL"sig"_5}$ as "Family A" and \\
+$\goodTerm{\cdot}{("inh"~"inh"_6~"obj"_5)}{\cL((\nu^+ \ "sig"_{41}
+\{"sl"_{41}\} \ "pN"_2[(\pi_1, \uparrow^s_2)]))}$ as "Family" $"A"_2$.
 
-Finally we make some comments on how a family is encoded in our metatheory.
-A family object will have a signature that expose the type of fields and the details of inductive type .
-This exposure can make sure inheritance can override and inherit field correctly.
-Sealing is only hiding information \textbf{inside one family}, i.e.
-hiding information of the former fields for the later fields, and thus the later fields is defined on the abstract interface of the former fields.
+Finally we make some comments on how a family is encoded in our
+metatheory. A family object will have a signature that expose the type
+of fields and the details of inductive type . This exposure can make
+sure inheritance can override and inherit field correctly. Sealing is
+only hiding information \textbf{inside one family}, i.e. hiding
+information of the former fields for the later fields, and thus the
+later fields is defined on the abstract interface of the former fields.
 The signature will expose the sealing itself and the inductive type.
-(See $"sig"_5$ has all the sealing information and inductive definition).
-A former sealing might not be related to any later sealing, for example, $"sig"_3$ is constructed using $"sl"_1$ but $"sig"_4$ is using trivial sealing $id_s$.
+(See $"sig"_5$ has all the sealing information and inductive
+definition). A former sealing might not be related to any later sealing,
+for example, $"sig"_3$ is constructed using $"sl"_1$ but $"sig"_4$ is
+using trivial sealing $id_s$.

### 2022-07-28 syntactic-example.tex
bugfix

@@
 \begin{align*}
-  "Osig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  w\cdot}{1} \\
+  "Osig" &:= \goodWSig{\cdot}{w^+\ w\cdot\ \top\ \bot}{1} \\
   "sig"_1 &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Osig"[\pi_1])}{1}  \\

### 2022-07-28 syntactic-example.tex
\coloneqq

@@
   \begin{minipage}[t]{0.4\linewidth}
+  \small
 \begin{align*}
-  "Osig" &:= \goodWSig{\cdot}{w^+\ w\cdot\ \top\ \bot}{1} \\
-  "sig"_1 &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Osig"[\pi_1])}{1}  \\
-  "obj"_1 &:= \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ *}{\cL"sig"_1}\\
+  "Osig" &\coloneqq \goodWSig{\cdot}{w^+\ w\cdot\ \top\ \bot}{1} \\
+  "sig"_1 &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Osig"[\pi_1])}{1}  \\
+  "obj"_1 &\coloneqq \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ *}{\cL"sig"_1}\\
   \cC"sig"_1 &= \goodType{\cdot}{\Sigma \ \top \ (\bW \ "Osig"[\pi_1]) }{} \\
-  "sig"_2 &:=  \goodSig{\cdot}{\nu^+ \ "sig"_1 \ {id_s} \ El ("W" ("pjr" \ \pi_2))}{2} \\ 
-  "O"_1 &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  "sig"_2 &\coloneqq  \goodSig{\cdot}{\nu^+ \ "sig"_1 \ {id_s} \ El ("W" ("pjr" \ \pi_2))}{2} \\ 
+  "O"_1 &\coloneqq {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
   & \text{ and thus }  \goodTerm{\cdot, \cC"sig"_1}{"O"_1}{El ("W" ("pjr" \pi_2))} \\
-  "obj"_2 &:= \goodTerm{\cdot}{\mu^+ \ "obj"_1 \ "O"_1}{\cL"sig"_2} \\
-  "sig"_2' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU)El ("pjr" \pi_2)}{} \\ 
-  \cC"sig"_2 &= \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El \ W("pjr" \pi_2) }{} \\ 
-  "sl"_1 &:= ((*, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
+  "obj"_2 &\coloneqq \goodTerm{\cdot}{\mu^+ \ "obj"_1 \ "O"_1}{\cL"sig"_2} \\
+  "sig"_2' &\coloneqq \goodType{\cdot}{\Sigma (\Sigma \top \cU)El ("pjr" \pi_2)}{} \\ 
+  \cC"sig"_2 &\coloneqq \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El \ W("pjr" \pi_2) }{} \\ 
+  "sl"_1 &\coloneqq ((*, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
   & \text{ and thus }  \goodSeal{\cdot}{"sl"_1}{"sig"_2}{"sig"_2'} \\
-  "pN"_1 &:= \goodType{\cdot, "sig"_2'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
-  "handler"_1 & := \goodTerm{\cdot, "sig"_2'}{\_}{\cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)} \\
-  "sig"_3 &:= \goodSig{\cdot}{\nu^+ "sig"_2 \{"sl"_1\} \ \cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)}{3} \\ 
-  "obj"_3 &: \goodTerm{\cdot}{\mu^+~"obj"_2~"handler"_1}{\cL"sig"_3}\\
-  "predT" := (\Pi & \ El ("W"("pjr" ("pjl"^2 \pi_2))) \\  
-  & (El ("W" ("pjr" ("pjl"^2  \pi_2)))[\pi_1])) \\
-  "pred"_1 & := \goodTerm{\cdot, \cC "sig"_3}{\lambda ("Wrec" \_)}{"predT"} \\ 
-  "sig"_4 &:= \nu^+ \ "sig"_3 \ \{id_s\} \ "predT" \\ 
-  "obj"_4 &: \goodTerm{\cdot}{\mu^+~"obj"_3~"pred"_1}{\cL"sig"_4}\\
-  "sl"_4 &:= \goodSeal{\cdot}{\_}{"sig"_4}{"sig"_4'} \\
-  "pN"_2 &:= \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{}
+  "pN"_1 &\coloneqq \goodType{\cdot, "sig"_2'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "handler"_1 &\coloneqq \goodTerm{\cdot, "sig"_2'}{\_}{\cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)} \\
+  "sig"_3 &\coloneqq \goodSig{\cdot}{\nu^+ "sig"_2 \{"sl"_1\} \ \cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)}{3} \\ 
+  "obj"_3 &\coloneqq \goodTerm{\cdot}{\mu^+~"obj"_2~"handler"_1}{\cL"sig"_3}\\
+  "predT" &\coloneqq \Pi \ El ("W"("pjr" ("pjl"^2 \pi_2)))\ El ("W" ("pjr" ("pjl"^2  \pi_2)))[\pi_1] \\
+  "pred"_1 &\coloneqq \goodTerm{\cdot, \cC "sig"_3}{\lambda ("Wrec" \_)}{"predT"} \\ 
+  "sig"_4 &\coloneqq \nu^+ \ "sig"_3 \ \{id_s\} \ "predT" \\ 
+  "obj"_4 &\coloneqq \goodTerm{\cdot}{\mu^+~"obj"_3~"pred"_1}{\cL"sig"_4}\\
+  "sl"_4 &\coloneqq \goodSeal{\cdot}{\_}{"sig"_4}{"sig"_4'} \\
+  "pN"_2 &\coloneqq \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{}
 \end{align*}
@@
 \begin{align*}
-  "sig"_5 &:= \nu^+ \ "sig"_4 \ \{"sl"_4\} \  "pN"_2 \\ 
+  "sig"_5 &\coloneqq \nu^+ \ "sig"_4 \ \{"sl"_4\} \  "pN"_2 \\ 
   "obj"_5 &: \goodTerm{\cdot}{\mu^+~"obj"_4~\_}{\cL"sig"_4}\\
@@
   \text{Here we}&\text{ start to construct Family } "A"_2 \\ 
-  "Nsig" &:= \goodWSig{\cdot}{w+ \ \top \ \bot \  "Osig"}{1} \\ 
-  "sig"_{11} &:= \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
-  "inh"_1 &:= \goodInh{\cdot}{"inhov" \  "inhid" \ *}{"sig"_1}{"sig"_{11}}\\
-  "O"_2 &:= {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
+  "Nsig" &\coloneqq \goodWSig{\cdot}{w+ \ \top \ \bot \  "Osig"}{1} \\ 
+  "sig"_{11} &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
+  "inh"_1 &\coloneqq \goodInh{\cdot}{"inhov" \  "inhid" \ *}{"sig"_1}{"sig"_{11}}\\
+  "O"_2 &\coloneqq {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
   & \text{ but }  \goodTerm{\cdot, \cC"sig"_{11}}{"O"_2}{El ("W" ("pjr" \pi_2))} \\
-  "sig"_{21} &:= \nu^+ \ "sig"_{11} \ \{id_s\} (El (W (pjr p₂))) \\ 
-  "inh"_2 &:= \goodInh{\cdot}{"inhov" \ "inh"_1 \ "O"_2}{"sig"_2}{"sig"_{21}}\\
+  "sig"_{21} &\coloneqq \nu^+ \ "sig"_{11} \ \{id_s\} (El (W (pjr p₂))) \\ 
+  "inh"_2 &\coloneqq \goodInh{\cdot}{"inhov" \ "inh"_1 \ "O"_2}{"sig"_2}{"sig"_{21}}\\
   "S"_T := (\Pi & \ El ("W"("pjr"  ("pjl" \pi_2))) \\  
   & (El ("W" ("pjr" ("pjl"  \pi_2)))[\pi_1])) \\
-  "sig"_{22} &:= \nu^+ \ "sig"_{21} \ \{id_s\} \ "S"_T \\ 
-  "sig"_{21}' &:= \goodType{\cdot}{\Sigma (\Sigma \top \cU) (El ("pjr" \pi_2))}{} \\
-  "sig"_{22}' &:= \Sigma "sig"_{21}' (\Pi \ El ("pjr"  ("pjl" \pi_2)) \  ...) \\
-  "sl"_{22} &:= \goodSeal{\cdot}{\_}{"sig"_{22}}{"sig"_{22}'} \\ 
-  "inh"_3 &:= \goodInh{\cdot}{\_}{"sig"_2}{"sig"_{22}} \\ 
-  \uparrow^s &:= \goodTerm{\cdot, "sig"_{22}'}{\_}{"sig"_2'[\pi_1]} \\ 
-  "pN"_3 &:= "pN"_1[(\pi_1, \uparrow^s)] \\ 
+  "sig"_{22} &\coloneqq \nu^+ \ "sig"_{21} \ \{id_s\} \ "S"_T \\ 
+  "sig"_{21}' &\coloneqq \goodType{\cdot}{\Sigma (\Sigma \top \cU) (El ("pjr" \pi_2))}{} \\
+  "sig"_{22}' &\coloneqq \Sigma "sig"_{21}' (\Pi \ El ("pjr"  ("pjl" \pi_2)) \  ...) \\
+  "sl"_{22} &\coloneqq \goodSeal{\cdot}{\_}{"sig"_{22}}{"sig"_{22}'} \\ 
+  "inh"_3 &\coloneqq \goodInh{\cdot}{\_}{"sig"_2}{"sig"_{22}} \\ 
+  \uparrow^s &\coloneqq \goodTerm{\cdot, "sig"_{22}'}{\_}{"sig"_2'[\pi_1]} \\ 
+  "pN"_3 &\coloneqq "pN"_1[(\pi_1, \uparrow^s)] \\ 
 \end{align*}
@@
     \begin{align*}
-      "inh"_{in0} &:= \goodInh{\cdot, "sig"_{22}'}{"inhid"}{(\nu^+ \ \nu\cdot \ ("pN"_1[\pi_1]))[(\pi_1, \uparrow^s)]}{(\nu^+ \cdot ("pN"_3[\pi_1]))} \\
-      "inh"_{in} &:= \goodInh{\cdot , "sig"_{22}'}{\_}{(\nu^+ \ \nu\cdot \ "pN"_1[\pi_1])[(\pi_1, \uparrow^s)]}{(\nu^+ \ \cdot \ ("pN"_3[\pi_1]) \ \{id_s\} (\Pi "pN"_3 ("pN"_3[\pi_1])) )[\pi_1]}\\
-      "sig"_{31} &:=  \nu^+ \ "sig"_{22} \ {"sl"_{22}}\  \cL(\nu^+ \cdot ("pN"_3[\pi_1])  (\Pi "pN"_3 ("pN"_3[\pi_1]))[\pi_1]) \\ 
-      "inh"_4 &:= \goodInh{\cdot}{"inhnest" \ "inh"_3 \uparrow^s "inh"_{in}}{"sig"_3}{"sig"_{31}} \\
-      "sig"_{41} &:= {\nu^+ \ "sig"_{31} \ (\Pi (El ("W" ("pjr" ("pjl"^4 \pi_2)))) (El ("W" ("pjr" ("pjl"^4 \pi_2))))[\pi_1])} \\ 
-      "inh"_5 &:= \goodInh{\cdot}{inhov \ \_}{"sig"_4}{"sig"_{41}} \\ 
-      "sl"_{41} &:= \goodSeal{\cdot}{\_}{"sig"_{41}}{"sig"_{41}'} \\ 
-      \uparrow^s_2 &:= \goodTerm{\cdot, "sig"_{41}}{\_}{"sig"_4'[\pi_1]} \\ 
-      "inh"_6 &:= \goodInh{\cdot}{"inhinh" \ "inh"_5 \ \uparrow^s_2}{"sig"_5}{(\nu^+ \ "sig"_{41} \{"sl"_{41}\} \ "pN"_2[(\pi_1, \uparrow^s_2)])}
+      "inh"_{in0} &\coloneqq \goodInh{\cdot, "sig"_{22}'}{"inhid"}{(\nu^+ \ \nu\cdot \ ("pN"_1[\pi_1]))[(\pi_1, \uparrow^s)]}{(\nu^+ \cdot ("pN"_3[\pi_1]))} \\
+      "inh"_{in} &\coloneqq \goodInh{\cdot , "sig"_{22}'}{\_}{(\nu^+ \ \nu\cdot \ "pN"_1[\pi_1])[(\pi_1, \uparrow^s)]}{(\nu^+ \ \cdot \ ("pN"_3[\pi_1]) \ \{id_s\} (\Pi "pN"_3 ("pN"_3[\pi_1])) )[\pi_1]}\\
+      "sig"_{31} &\coloneqq  \nu^+ \ "sig"_{22} \ {"sl"_{22}}\  \cL(\nu^+ \cdot ("pN"_3[\pi_1])  (\Pi "pN"_3 ("pN"_3[\pi_1]))[\pi_1]) \\ 
+      "inh"_4 &\coloneqq \goodInh{\cdot}{"inhnest" \ "inh"_3 \uparrow^s "inh"_{in}}{"sig"_3}{"sig"_{31}} \\
+      "sig"_{41} &\coloneqq {\nu^+ \ "sig"_{31} \ (\Pi (El ("W" ("pjr" ("pjl"^4 \pi_2)))) (El ("W" ("pjr" ("pjl"^4 \pi_2))))[\pi_1])} \\ 
+      "inh"_5 &\coloneqq \goodInh{\cdot}{inhov \ \_}{"sig"_4}{"sig"_{41}} \\ 
+      "sl"_{41} &\coloneqq \goodSeal{\cdot}{\_}{"sig"_{41}}{"sig"_{41}'} \\ 
+      \uparrow^s_2 &\coloneqq \goodTerm{\cdot, "sig"_{41}}{\_}{"sig"_4'[\pi_1]} \\ 
+      "inh"_6 &\coloneqq \goodInh{\cdot}{"inhinh" \ "inh"_5 \ \uparrow^s_2}{"sig"_5}{(\nu^+ \ "sig"_{41} \{"sl"_{41}\} \ "pN"_2[(\pi_1, \uparrow^s_2)])}
     \end{align*}  

### 2022-07-28 syntactic-example.tex
bugfixes

@@
   "Osig" &\coloneqq \goodWSig{\cdot}{w^+\ w\cdot\ \top\ \bot}{1} \\
-  "sig"_1 &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Osig"[\pi_1])}{1}  \\
+  "sig"_1 &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot \ {id_s} \ (\bW \ "Osig"[\pi_1])}{1}  \\
   "obj"_1 &\coloneqq \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ *}{\cL"sig"_1}\\
-  \cC"sig"_1 &= \goodType{\cdot}{\Sigma \ \top \ (\bW \ "Osig"[\pi_1]) }{} \\
+  \cC"sig"_1 &\coloneqq \goodType{\cdot}{\Sigma \ \top \ (\bW \ "Osig"[\pi_1]) }{} \\
   "sig"_2 &\coloneqq  \goodSig{\cdot}{\nu^+ \ "sig"_1 \ {id_s} \ El ("W" ("pjr" \ \pi_2))}{2} \\ 
@@
 \begin{minipage}[t]{0.4\linewidth}
+  \small
 \begin{align*}
@@
   "inh"_2 &\coloneqq \goodInh{\cdot}{"inhov" \ "inh"_1 \ "O"_2}{"sig"_2}{"sig"_{21}}\\
-  "S"_T := (\Pi & \ El ("W"("pjr"  ("pjl" \pi_2))) \\  
-  & (El ("W" ("pjr" ("pjl"  \pi_2)))[\pi_1])) \\
+  "S"_T &\coloneqq \Pi \ El ("W"("pjr"  ("pjl" \pi_2))) \ El ("W" ("pjr" ("pjl"  \pi_2)))[\pi_1] \\
   "sig"_{22} &\coloneqq \nu^+ \ "sig"_{21} \ \{id_s\} \ "S"_T \\ 
@@
   \begin{minipage}{0.8\linewidth}
+    \small
     \centering

### 2022-07-29 syntactic-example.tex
comment

@@
   "sig"_2' &\coloneqq \goodType{\cdot}{\Sigma (\Sigma \top \cU)El ("pjr" \pi_2)}{} \\ 
-  \cC"sig"_2 &\coloneqq \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El \ W("pjr" \pi_2) }{} \\ 
+  \cC"sig"_2 &\coloneqq \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El ("W"("pjr" \pi_2)) }{} \\ 
   "sl"_1 &\coloneqq ((*, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
@@
 
-Here we explain how  our detailed construction in Figure
-\cref{fig:example-construction} is constructed. We first elaborate that
+Here we explain how  our detailed construction in
+\cref{fig:example-construction}. We first elaborate that
 how our inductive type "N" is written in the specific format (the style
@@
 concrete recursor $"pred"_1$ away using $"sl"_4$ so that the field $"k"$
-can applys it without resorting to the concrete recursor. Finally we
-have $"sig"_5$ as final signature.
+can apply it without resorting to the concrete recursor. Finally we
+have $"sig"_5$ as final signature.\YZ{
+  Maybe it is a good place to point out what El(pjr (pjl pi_2)) in pN_2 stands for
+  and what El(W(pjr(pjl^2 pi_2))) in predT stands for.
+}
 
@@
 understood as ``upcasting'', but this ``upcasting'' is used to upcast
-the context of the children. We immeidately see its usage in $"pN"_3$
+the context of the children. We immediately see its usage in $"pN"_3$
 and $"inh"_{in0}$. Upon this, we construct the complete

### 2022-07-29 syntactic-example.tex
bugfix

@@
   \text{Here we}&\text{ start to construct Family } "A"_2 \\ 
-  "Nsig" &\coloneqq \goodWSig{\cdot}{w+ \ \top \ \bot \  "Osig"}{1} \\ 
+  "Nsig" &\coloneqq \goodWSig{\cdot}{w^+\ "Osig"\ \top \ \top}{1} \\ 
   "sig"_{11} &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\

### 2022-07-29 syntactic-example.tex
minor

@@
   \text{Here we}&\text{ start to construct Family } "A"_2 \\ 
-  "Nsig" &\coloneqq \goodWSig{\cdot}{w^+\ "Osig"\ \top \ \top}{1} \\ 
+  "Nsig" &\coloneqq \goodWSig{\cdot}{w^+\ "Osig"\ \top \ \top}{2} \\ 
   "sig"_{11} &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
@@
 
-Finally we make some comments on how a family is encoded in our
-metatheory. A family object will have a signature that expose the type
-of fields and the details of inductive type . This exposure can make
-sure inheritance can override and inherit field correctly. Sealing is
-only hiding information \textbf{inside one family}, i.e. hiding
+Finally we comment on how a family is encoded in our metatheory.
+A family object will have a signature that exposes the type
+of fields and the details of inductive types. This exposure can make
+sure inheritance can override and inherit fields correctly. Sealing is
+only hiding information \textbf{inside one family}, i.e., hiding
 information of the former fields for the later fields, and thus the
-later fields is defined on the abstract interface of the former fields.
+later fields can be defined relying on only the abstract interface of
+the former fields.
 The signature will expose the sealing itself and the inductive type.

### 2022-07-29 body.tex
Update STLC example in 3.1
Update propositional partial recursor Update intro with structure and contribution.
intro tweak
Move sec3.1 ~> sec3, sec3.2~>sec 4.
Save, rewriting meta-theory
syntactical → syntactic

@@
         Family, Extend, Field, EndFamily, SealFamily, FInductive,
-        FRecursor, MetaData, EndMetaData, Closing, Final, Inherit
+        FRecursor, MetaData, EndMetaData, Closing, Final, Inherit, FTheorem
         },
@@
 
-\section{Language Design}
+\section{Background and Challenges}
 % what feature we want
@@
 % Before/After Comparison of examples (with/without Fampoly)
-\label{sec:lang-design}
-\input{lang-design.tex}
+\label{sec:background+challenge}
+\input{challenge.tex}
 
@@
 \newcommand{\bW}{{\mathbb{W}}}
+
+\section{Language Design}
+% we need to make this section before Metatheory
+% because a lot of our formulation in the metatheory
+%   is inspired by the implementation
+\label{sec:lang-design}
+\input{lang-design.tex}
 
@@
 
+It is hard to write modular, extensible code and proofs.
+We have presented ...
+
 \setlength{\bibsep}{.8ex}
@@
 
-For syntactical translation, please check \\ \href{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}.
+For syntactic translation, please check \\ \href{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}.
 This is the first QIIT-model we give.

### 2022-07-29 syn-translate.tex
syntactical → syntactic

@@
 The implementation described in \cref{sec:coqimpl} can be formalized in our metatheory.
-The described compilation strategy can be considered as a \textit{syntactical translation}, a function mapping from our syntactical (QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only difference lies in the "Wrec", in our metatheory, it is constructed using linkage; in ``subpart'', it will be using sigma types (translation of the linkage).} of our syntactical (QIIT-)model \textit{without linkages}.\footnote{Here, syntactical QIIT-model can be roughly understood as just AST, with equational rules acting like quotient.
+The described compilation strategy can be considered as a \textit{syntactic translation}, a function mapping from our syntactic (QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only difference lies in the "Wrec", in our metatheory, it is constructed using linkage; in ``subpart'', it will be using sigma types (translation of the linkage).} of our syntactic (QIIT-)model \textit{without linkages}.\footnote{Here, syntactic QIIT-model can be roughly understood as just AST, with equational rules acting like quotient.
 So this mapping is mapping AST to AST, respecting quotient/judgemental equality.} This ``subpart'' is the standard MLTT with inductive type signature, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
@@
 
-The main idea of this syntactical translation is that, we make sure $\goodSig{\Gamma}{\nu^+ \_ }{n+1}$ and $\goodSig{\Gamma}{\mu^+ \_}{n+1}$ wrapped into a dependent pair.
-These two definitions decide the syntactical translation for $\goodSig{\Gamma}{\_}{n}$ and $\cL$.
-The syntactical translation for $\cC$ and $\cCt$ are decided by their $\beta$-rules.
+The main idea of this syntactic translation is that, we make sure $\goodSig{\Gamma}{\nu^+ \_ }{n+1}$ and $\goodSig{\Gamma}{\mu^+ \_}{n+1}$ wrapped into a dependent pair.
+These two definitions decide the syntactic translation for $\goodSig{\Gamma}{\_}{n}$ and $\cL$.
+The syntactic translation for $\cC$ and $\cCt$ are decided by their $\beta$-rules.
 
@@
 
-After syntactical translation, we can directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the ``subpart''}.
+After syntactic translation, we can directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the ``subpart''}.
 However, due to our non-conventional formulation of the inductive type, it is better we don't impose such assumption and directly give a consistency and canonicity proof.
 
-% the link for syntactical translation
+% the link for syntactic translation
 % https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing

### 2022-07-29 syn-translate.tex
wrap lines

@@
-The implementation described in \cref{sec:coqimpl} can be formalized in our metatheory.
-The described compilation strategy can be considered as a \textit{syntactic translation}, a function mapping from our syntactic (QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only difference lies in the "Wrec", in our metatheory, it is constructed using linkage; in ``subpart'', it will be using sigma types (translation of the linkage).} of our syntactic (QIIT-)model \textit{without linkages}.\footnote{Here, syntactic QIIT-model can be roughly understood as just AST, with equational rules acting like quotient.
-So this mapping is mapping AST to AST, respecting quotient/judgemental equality.} This ``subpart'' is the standard MLTT with inductive type signature, a (non-conventional) W-type and a recursor that using sigma type to aggregate the handlers.
+The implementation described in \cref{sec:coqimpl} can be formalized in
+our metatheory. The described compilation strategy can be considered as
+a \textit{syntactic translation}, a function mapping from our syntactic
+(QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only
+difference lies in the "Wrec", in our metatheory, it is constructed
+using linkage; in ``subpart'', it will be using sigma types (translation
+of the linkage).} of our syntactic (QIIT-)model \textit{without
+linkages}.\footnote{Here, syntactic QIIT-model can be roughly understood
+as just AST, with equational rules acting like quotient. So this mapping
+is mapping AST to AST, respecting quotient/judgemental equality.} This
+``subpart'' is the standard MLTT with inductive type signature, a
+(non-conventional) W-type and a recursor that using sigma type to
+aggregate the handlers.
 
-Basically, we can compile away the judgement $\goodSig{\Gamma}{v}{n}$, the type $\cL$ and $\cC$, and related terms, by transforming them into the corresponding sigma type.
+Basically, we can compile away the judgement $\goodSig{\Gamma}{v}{n}$,
+the type $\cL$ and $\cC$, and related terms, by transforming them into
+the corresponding sigma type.
 
@@
 
-We use $\denotesT{}$ as the translation.
-This translation is invariant under the subpart.
-Since we are in an intrinsic setting, everything here is well-typed (well-formed)including this translation.
-Thus $\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$ (i.e.
-every type is tracking its context).
-In other words, every derivation of any judgement will track its ``context''.
-This will be clearer in the (pseudo-)Agda formalization.
+We use $\denotesT{}$ as the translation. This translation is invariant
+under the subpart. Since we are in an intrinsic setting, everything here
+is well-typed (well-formed)including this translation. Thus
+$\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$
+(i.e. every type is tracking its context). In other words, every
+derivation of any judgement will track its ``context''. This will be
+clearer in the (pseudo-)Agda formalization.
 
@@
 
-The main idea of this syntactic translation is that, we make sure $\goodSig{\Gamma}{\nu^+ \_ }{n+1}$ and $\goodSig{\Gamma}{\mu^+ \_}{n+1}$ wrapped into a dependent pair.
-These two definitions decide the syntactic translation for $\goodSig{\Gamma}{\_}{n}$ and $\cL$.
-The syntactic translation for $\cC$ and $\cCt$ are decided by their $\beta$-rules.
+The main idea of this syntactic translation is that, we make sure
+$\goodSig{\Gamma}{\nu^+ \_ }{n+1}$ and $\goodSig{\Gamma}{\mu^+ \_}{n+1}$
+wrapped into a dependent pair. These two definitions decide the
+syntactic translation for $\goodSig{\Gamma}{\_}{n}$ and $\cL$. The
+syntactic translation for $\cC$ and $\cCt$ are decided by their
+$\beta$-rules.
 
-We use the local morphism (the same one for encoding the Sealing judgements) to translate inheritance judgement.
-This is mainly due to the existence of \ruleref{Inh} rule.
-Here we only typeset the overriding and inheritance rule, but all of the translation for inh judgment are induced by the $\beta$-rule of \ruleref{Inh}.
+We use the local morphism (the same one for encoding the Sealing
+judgements) to translate inheritance judgement. This is mainly due to
+the existence of \ruleref{Inh} rule. Here we only typeset the overriding
+and inheritance rule, but all of the translation for inh judgment are
+induced by the $\beta$-rule of \ruleref{Inh}.
 
 We omit all the equational rules ($\beta,\eta$ and substitution) here.
-However, for coherence, when we mutual recursively define $\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and $\denotesT{\cC}$ above, we actually have to prove the substitution laws $(\cC\sigma)[\tau]\equiv \cC(\sigma[\tau])$ and $\Sigr{\Delta}{n}[\tau] \equiv \Sigr{\Gamma}{n}$ together.
-The details (of all the equational rules) can be found in the appendices written in (pseudo-)Agda form.
-We encourage the reader to read that (pseudo-)Agda-style proof because it is more friendly to type theorists, easy to do type checking and validate the proof and it is clearer on the mathematical object each translation uses.
+However, for coherence, when we mutual recursively define
+$\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and
+$\denotesT{\cC}$ above, we actually have to prove the substitution laws
+$(\cC\sigma)[\tau]\equiv \cC(\sigma[\tau])$ and $\Sigr{\Delta}{n}[\tau]
+\equiv \Sigr{\Gamma}{n}$ together. The details (of all the equational
+rules) can be found in the appendices written in (pseudo-)Agda form. We
+encourage the reader to read that (pseudo-)Agda-style proof because it
+is more friendly to type theorists, easy to do type checking and
+validate the proof and it is clearer on the mathematical object each
+translation uses.
 
-Ultimately, we have constructed a model for our QIIT-syntax using only a ``subpart'' of our QIIT-syntax -- without these linkages.
-It is also possible to compile our $\goodWSig{\Gamma}{w}{n}$ and W-type into a more conventional formulation, but the recursor will always require a sigma type as handler and thus the elimination rule for our W-type is still non-conventional.
+Ultimately, we have constructed a model for our QIIT-syntax using only a
+``subpart'' of our QIIT-syntax -- without these linkages. It is also
+possible to compile our $\goodWSig{\Gamma}{w}{n}$ and W-type into a more
+conventional formulation, but the recursor will always require a sigma
+type as handler and thus the elimination rule for our W-type is still
+non-conventional.
 
-This translation justifies the intuition -- a linkage is actually a module \textbf{with universal quantification wrapping its fields to achieve late binding}.
+This translation justifies the intuition -- a linkage is actually a
+module \textbf{with universal quantification wrapping its fields to
+achieve late binding}.
 
-After syntactic translation, we can directly have consistency and canonicity \textit{if we assume the consistency and canonicity of the ``subpart''}.
-However, due to our non-conventional formulation of the inductive type, it is better we don't impose such assumption and directly give a consistency and canonicity proof.
+After syntactic translation, we can directly have consistency and
+canonicity \textit{if we assume the consistency and canonicity of the
+``subpart''}. However, due to our non-conventional formulation of the
+inductive type, it is better we don't impose such assumption and
+directly give a consistency and canonicity proof.
 

### 2022-07-29 syn-translate.tex
edits

@@
 (QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only
-difference lies in the "Wrec", in our metatheory, it is constructed
-using linkage; in ``subpart'', it will be using sigma types (translation
+difference lies in "Wrec". In our metatheory, it is constructed
+using linkage; in ``subpart'', it will be using $\Sigma$-types (translation
 of the linkage).} of our syntactic (QIIT-)model \textit{without
@@
 is mapping AST to AST, respecting quotient/judgemental equality.} This
-``subpart'' is the standard MLTT with inductive type signature, a
-(non-conventional) W-type and a recursor that using sigma type to
-aggregate the handlers.
+``subpart'' is the standard MLTT with inductive type signatures, an
+(unconventional) formulation of W-types,
+and recursors that use $\Sigma$-types to aggregate handlers.
 
@@
 the type $\cL$ and $\cC$, and related terms, by transforming them into
-the corresponding sigma type.
+corresponding $\Sigma$-types.
 
@@
 
-We use $\denotesT{}$ as the translation. This translation is invariant
+We use $\denotesT{}$ as the translation. This translation is invariant\YZ{What does "invariant" mean?}
 under the subpart. Since we are in an intrinsic setting, everything here
@@
   \text{We will define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
-  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursively defined together, } \\
-  & \text{and they define inductively on the signature length} \\  
+  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursive, } \\
+  & \text{defined by induction on the signature length} \\  
   \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
@@
   \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
-    \Sigma~(\denotesT{\cC ("pjl"~\sigma)})~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
+    \Sigma~\denotesT{\cC ("pjl"~\sigma)}~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
       &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\

### 2022-07-29 syntactic-example.tex
syntactical → syntactic
Reply the comments.

@@
 can apply it without resorting to the concrete recursor. Finally we
-have $"sig"_5$ as final signature.\YZ{
+have $"sig"_5$ as final signature.
+Recall from \ruleref{Sigma Elim}, we have "pjl" and "pjr" as two projection from the sigma type (dependent pair).
+We will also use $"pjl"^n$ as a consecutive composition of "pjl", (e.g.
+$"pjl"^2~t$ is exactly $("pjl" ("pjl"~t))$ ).
+\YZ{
   Maybe it is a good place to point out what El(pjr (pjl pi_2)) in pN_2 stands for
   and what El(W(pjr(pjl^2 pi_2))) in predT stands for.
-}
+}\EDJreply{Done.}
 

### 2022-07-30 coqimpl.tex
comments

@@
 (*
-Field test := (subst (tm_var 0) 0 (tm_var 0)).
+Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}@
 *)
@@
       subst_internal_410.tm_abs ... .
-End v_33_34.
+End v_33_34.@\YZ{It'd be useful to (1) show how the parts of the STLC family are assembled and recursor inserted, and (2) show the code compiled for STLC\_bool and how code from STLC is shared with STLC\_bool.}@
 \end{minted}

### 2022-07-30 syntactic-example.tex
reply

@@
 how our inductive type "N" is written in the specific format (the style
-of W-type). We use 1, 0 here to indicate unit type and bottom type.
+of W-type).
+We use 1, 0 here to indicate unit type and bottom type.
+Recall from \ruleref{Sigma Elim}, we have "pjl" and "pjr" as two projection from the sigma type (dependent pair).
+We will also use $"pjl"^n$ as a consecutive composition of "pjl", (e.g.
+$"pjl"^2~t$ is exactly $("pjl" ("pjl"~t))$ ).
 
@@
 $"handler"_1$, with which we can construct the real recursor $"pred"_1$.
-Both detailed constructions are omitted. Then we again abstract the
-concrete recursor $"pred"_1$ away using $"sl"_4$ so that the field $"k"$
-can apply it without resorting to the concrete recursor. Finally we
+Both detailed constructions are omitted.
+$"obj"_3$ and $"sig"_3$ is the resulting linkage and linkage signature aggregating the nested family $"handler"_1$.
+$"pred"_1$ is then the constructed recursor over $"sig"_3$ , without any abstraction, for the sake of exhaustiveness checking.
+We show the example $"sig"_4$ as the result of adding this new recursor.
+
+Then we again abstract the concrete recursor $"pred"_1$ away using $"sl"_4$ so that the field $"k"$ (of type $"pN"_2$)
+can apply the recursor without resorting to any concrete recursor, thus achieving late-binding. Finally we
 have $"sig"_5$ as final signature.
-Recall from \ruleref{Sigma Elim}, we have "pjl" and "pjr" as two projection from the sigma type (dependent pair).
-We will also use $"pjl"^n$ as a consecutive composition of "pjl", (e.g.
-$"pjl"^2~t$ is exactly $("pjl" ("pjl"~t))$ ).
+
+Note that, since in our example there is only one type, the complicated debruijn indices $El("pjr" ("pjl" \pi_2)) $ in $"pN"_2$ and  $El("W"("pjr"("pjl"^2 \pi_2)))$ in "predT" are both referring to the type "N".
+The distinguishing difference is we have "W" (\ruleref{Ind-Type}) for the latter as "predT" is constructed in the context knowing the concrete definition of the inductive type, thus generally not inheritable.
+On the contrary,  $"pN"_2$ has as its context only $"N" : \cU$ and manifests its late-binding and its overridability for its former fields.
 \YZ{
@@
   and what El(W(pjr(pjl^2 pi_2))) in predT stands for.
-}\EDJreply{Done.}
+}\EDJreply{Done.}\YZreply{
+  I was actually alluding to how the type of pN_2 manifests late binding
+  while the type of predT refers to the concrete signature of the
+  inductive type.
+}\YZreply{
+  It’s good to be consistent in notations.  You later use * for products definition of pN_1, but you also use * as the term of type ($\bW$ $\sigma$) 
+}\EDJreply{Sry. Now I attempt another try. I also change the above paragraph a bit to explain the details on sig₃, sig₄, and etc.
+I also copy the comments in our slack to here to remind me about the changes.
+Also let me paste the full example here for future reference: \\
+% following is the body of pred₁
+pred_ = Wrec ...  
+  : Tm (⋅ ▹ 𝓒sig₃ ▹ (El (W (pjr (pjl (pjl π₂)))))) (El (W (pjr (pjl (pjl π₂)))))[π₁] \\
+% we do lambda abstraction now
+pred₁ = λ pred_ :   Tm (⋅ ▹ 𝓒sig₃) (∏ El (W (pjr (pjl (pjl π₂)))) (El (W (pjr (pjl (pjl p₂))))[π₁])) \\
+% we can see that, the type of pred₁ is predT 
+% now we construct sig4, this should not have any abstraction inside
+sig₄ = ∨+ sig₃ id_s (∏ El (W (pjr (pjl (pjl π₂)))) (El (W (pjr (pjl (pjl π₂))))[π₁])) \\ 
+% we omit the construction of obj₄
+obj₄ : Tm ⋅ 𝓛sig₄ \\
+% since pred is define in context 𝓒sig₃, we need to abstract 𝓒sig₄ into sig₄'
+% again for later field,
+sig₄' = 
+  Σ (Σ ((Σ (Σ ⊤ 𝕌) (El (pjr π₂))) 𝓛(∨+ ⋅ (pN[π₁]))[sf sl₁])) 
+    (∏ El (pjr (pjl (pjl π₂))) (El (pjr (pjl (pjl π₂)))[π₁]))
+% and we have, omitting the detail construction 
+sl₄ : Seal ⋅ sig₄ sig₄' \\
+% pN₂ now is constructed in the context of the sealed context sig₄'
+pN₂ = El (pjr (pjl π₂)) * El (pjr (pjl π₂)) : Ty (⋅ ▹ sig₄') 
+% the product type \\ 
+sig₅ = ∨+ sig₄ {sl₄} pN₂ \\ % the full example is in my gist last time I sent to you (Attempt 9), if you like it I can send the link again but note that the omission is still there and a lot of naming convention is different.
+It seems that as you last night complained, I forget to add sig₄' causing problems.  Do I also need to make a note that those primed variables are always the result of abstraction?
+}
 

### 2022-08-06 syntactic-example.tex
Fix notation.

@@
   "sig"_1 &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot \ {id_s} \ (\bW \ "Osig"[\pi_1])}{1}  \\
-  "obj"_1 &\coloneqq \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ *}{\cL"sig"_1}\\
+  "obj"_1 &\coloneqq \goodTerm{\cdot}{\mu^+ \ \mu\cdot \ {id_s} \ \star}{\cL"sig"_1}\\
   \cC"sig"_1 &\coloneqq \goodType{\cdot}{\Sigma \ \top \ (\bW \ "Osig"[\pi_1]) }{} \\
@@
   \cC"sig"_2 &\coloneqq \goodType{\cdot}{\Sigma (\Sigma \top (\bW \ "Osig"[\pi_1]))El ("W"("pjr" \pi_2)) }{} \\ 
-  "sl"_1 &\coloneqq ((*, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
+  "sl"_1 &\coloneqq ((\star, "W" ("pjr" ("pjl" \pi_2)) ), "pjr" \pi_2) \\ 
   & \text{ and thus }  \goodSeal{\cdot}{"sl"_1}{"sig"_2}{"sig"_2'} \\
-  "pN"_1 &\coloneqq \goodType{\cdot, "sig"_2'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{} \\
+  "pN"_1 &\coloneqq \goodType{\cdot, "sig"_2'}{El ("pjr" ("pjl" \pi_2)) \times El ("pjr" ("pjl" \pi_2))}{} \\
   "handler"_1 &\coloneqq \goodTerm{\cdot, "sig"_2'}{\_}{\cL(\nu^+ \ \nu\cdot \ \{id_s\} "pN"_1)} \\
@@
   "sl"_4 &\coloneqq \goodSeal{\cdot}{\_}{"sig"_4}{"sig"_4'} \\
-  "pN"_2 &\coloneqq \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) * El ("pjr" ("pjl" \pi_2))}{}
+  "pN"_2 &\coloneqq \goodType{\cdot, "sig"_4'}{El ("pjr" ("pjl" \pi_2)) \times El ("pjr" ("pjl" \pi_2))}{}
 \end{align*}
@@
   "sig"_{11} &\coloneqq \goodSig{\cdot}{\nu^+ \ \nu\cdot  \ (\bW \ "Nsig"[\pi_1])}{1} \\
-  "inh"_1 &\coloneqq \goodInh{\cdot}{"inhov" \  "inhid" \ *}{"sig"_1}{"sig"_{11}}\\
+  "inh"_1 &\coloneqq \goodInh{\cdot}{"inhov" \  "inhid" \ \star}{"sig"_1}{"sig"_{11}}\\
   "O"_2 &\coloneqq {"Wsup" \ ("pjr" \pi_2) \ () \ ("elim-"\bot \ \pi_2)}  \\ 
@@
   It’s good to be consistent in notations.  You later use * for products definition of pN_1, but you also use * as the term of type ($\bW$ $\sigma$) 
-}\EDJreply{Sry. Now I attempt another try. I also change the above paragraph a bit to explain the details on sig₃, sig₄, and etc.
+}\EDJreply{Notation replaced. First * replaced with $\star$, second * replaced with $\times$}
+\EDJreply{Sry. Now I attempt another try. I also change the above paragraph a bit to explain the details on sig₃, sig₄, and etc.
 I also copy the comments in our slack to here to remind me about the changes.

### 2022-08-06 syntax.tex
Reply the comments.
Fix notation.
Remove Notation sf'

@@
 \quad 
-\Rule[]
+\Rule[name=Sigma Elim]
 {\goodTerm{\Gamma}{t}{\Sigma A B}}
@@
 {\goodWSig{\Gamma}{\tau}{n}}
-{\goodType{\Gamma}{\bW \tau}{i}}
+{\goodType{\Gamma}{\bW \tau}{i}
+}
 $$
@@
 {\goodTerm{\Gamma}{T}{\bW \tau}}
-{\goodTerm{\Gamma}{"W" \ T}{\cU}}
+{\goodTerm{\Gamma}{"W" \ T}{\cU}
+}
+\quad 
+\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}
 \quad
@@
 
-For notation clarity, we will sometimes use $sf'$ to emphasize that we
-consider that the seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term
-$\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]}$.\YZ{sf' seems unnecessary, since sf' s ad s are the same?}\EDJreply{Yes.
+% For notation clarity, we will sometimes use $sf'$ to emphasize that we
+% consider that the seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term
+% $\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]}$.
+
+\YZ{sf' seems unnecessary, since sf' s ad s are the same?}\EDJreply{Yes.
 Technically it is pretty unnecessary.
@@
   I mean technically we can consider them the same. I can just remove all the appearances of sf' if you want and find it less confusing.
-}\YZreply{Yes, sf' is confusing to me.}
-Notation "sf"
+}\YZreply{Yes, sf' is confusing to me.}\EDJreply{sf' notation removed. Please check.}
+Note that, given $\goodSeal{\Gamma}{s}{\sigma}{A} = \goodTerm{\Gamma, \cC\sigma}{s}{A}$, notation "sf"
 (used in \ruleref{Compile}) is defined as the substitution $\goodSub{\Gamma,
-\cC \sigma}{(sf~s) := (\pi_1, sf'~s)}{\Gamma, A}$.
+\cC \sigma}{(sf~s) \coloneqq (\pi_1, s)}{\Gamma, A}$.
 

### 2022-08-06 coqimpl.tex
Reply some comments.

@@
   this requirement, but also the "self vs. consistency" challenge.
-}\EDJreply{Yes. Indivially compiling family member is to achieve both requirements}
+}\EDJreply{Yes. Indivially compiling family member can achieve both requirements}
 
@@
 (*
-Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}@
+Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}\EDJreply{Because I want to show I can do it but I don't really want to d -- if I really do it the compilation result shown later will need to have this information.}@
 *)

### 2022-08-07 metatheory.tex
syntactical → syntactic
Add the part why we use predicative MLTT.

--- 2022-07-16 metatheory.tex
@@
 formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT)~\cite{martin1982constructive}
-given by~\citet{coquand2018canonicity}. 
+given by~\citet{coquand2018canonicity}. We will mainly focus on the new facility that we introduced.
+
+The reason we choose this version of predicative MLTT is mainly because it is simpler on the formulation and metatheoretical reasoning, easy to extend and is enough as a background for setting up the formalization of the key feature for our language, like overriding and inheritance about the families.
+Compared to a more complete Coq's formal foundation like \textit{Predicative Calculus of Cumulative Inductive Constructions}~\cite{timany2018cumulative}, our formulation lacks cumulativity (though our formulation still has multi-level universe) and impredicative universe (the \mintinline{Coq}{Prop : Type}).
+\YZ{
+    What's the relation to the CIC underlying Coq, given that the implementation is a Coq plugin?
+}\EDJreply{Add the comparison between this version of predicative MLTT and PCUIC. Please Check.}
 % Thus we will omit a great deal of details about MLTT here
-\YZ{What details are omitted?} 
-We will mainly focus on the new facility that we introduced
 
@@
 
-\subsection{Implementation Formalized as Syntactical Translation}
+\subsection{Implementation Formalized as Syntactic Translation}
 

### 2022-08-07 lang-design.tex
Reply the comments.
Rewriting overridaibility chain.

@@
 \begin{figure}[!htb]
-  \begin{minipage}[t]{0.45\linewidth}
+  \begin{minipage}[t]{0.47\linewidth}
 \begin{lstlisting}[language=Coq, escapeinside={@}{@}]
@@
  Field a : nat := 1.
- Field b : a = a := eq_refl.
- Fail Field c : a = 1 := eq_refl. (*Fail 1*)
-
- Final Field a' : nat := 1.
- Field c : a' = 1 := eq_refl.
- Field bop : nat -> nat -> nat := (+).
- Fail Field comm_bop : ...
-
- Family MonotonicF.
-  Field f : nat -> nat := non_impl.
+ Field b : a = 1 := eq_refl.
+ Overridable Field a' : nat := 1.
+ Fail Field c : a' = 1 := eq_refl. (* Fail 1*)
+ Overridable Field f : nat -> nat := non_impl.
+ Overridable 
   Field p : monotonic f := non_impl.
- SealFamily. 
 EndFamily.
@@
 Family I2 extends I.
- Override Field a : nat := 1. (* Override *)
- Inherit b.                   (* Inherit  *)
 
- Field d : nat := a.          (* Extend   *)
- Fail Override Field a' : nat := 2. (*Fail 2*)
+ Inherit Field b.
+ Override Field a' : nat := 2.  (* Override *)
 
- Override Field bop := (-).
-
- Inherits Until MonotonicF.
- Override Family MonotonicF.
-  Field f : nat -> nat := (+ 1).
+ Override Field f : nat -> nat := (+ 1).
+ Override pins {f} 
   Field p : monotonic f := ...
@@
 
-% Talk about Final
-\textbf{Final as non-overridable.} Every field is by default
-overridable, reminding us of \ref{chg:definition-relevant-reasoning}.  To
-resolve it, we provide the keyword "Final" to allow the user to specify
-\textit{non-overridability} and give the user the opportunity to choose
-from overriding and exposing the definition.
+% % Talk about Final
+% \textbf{Final as non-overridable.} Every field is by default
+% overridable, reminding us of \ref{chg:definition-relevant-reasoning}.  To
+% resolve it, we provide the keyword "Final" to allow the user to specify
+% \textit{non-overridability} and give the user the opportunity to choose
+% from overriding and exposing the definition.
 
-Take the field "c" of "Family I" in \cref{fig:plugin-example3} as an 
-example. We can see that the first attempt of proving "a = 1" fails.
-This should fail---because we want to retain the possibility
-to override "a" and inheriting "c" in the future, but they are
-conflicting, since "c" witnesses that "a" is fixed to "1". On the other
-hand, once we decorate \lstinline{a'} as final, the second attempt
-\lstinline{c : a' = 1 := eq_refl} will succeed. Contrary to "a",
-our \lstinline{a'} is not overridable in "Family I2", mediating the
-\ref{chg:definition-relevant-reasoning}.
+% Take the field "c" of "Family I" in \cref{fig:plugin-example3} as an 
+% example. We can see that the first attempt of proving "a = 1" fails.
+% This should fail---because we want to retain the possibility
+% to override "a" and inheriting "c" in the future, but they are
+% conflicting, since "c" witnesses that "a" is fixed to "1". On the other
+% hand, once we decorate \lstinline{a'} as final, the second attempt
+% \lstinline{c : a' = 1 := eq_refl} will succeed. Contrary to "a",
+% our \lstinline{a'} is not overridable in "Family I2", mediating the
+% \ref{chg:definition-relevant-reasoning}.
 
-This example might sound trivial and contrived. Let's look at the
-example field binary operation "bop" initially set to be addition.
-Now that we only expose the type of "bop", we
-cannot prove its commutativity in "comm_bop" (because we don't have
-access to its concrete definition) but we can override "bop" with
-subtraction function in "Family I2". This makes sense because we
-shouldn't (have the ability to) inherit commutativity property for
-subtraction anyway. This is contrary to how \lstinline{a'} is declared
-"Final": the second "c" can prove that it's equal to 1 but
-\lstinline{a'} is not overridable in the "Family I2".
+% This example might sound trivial and contrived. Let's look at the
+% example field binary operation "bop" initially set to be addition.
+% Now that we only expose the type of "bop", we
+% cannot prove its commutativity in "comm_bop" (because we don't have
+% access to its concrete definition) but we can override "bop" with
+% subtraction function in "Family I2". This makes sense because we
+% shouldn't (have the ability to) inherit commutativity property for
+% subtraction anyway. This is contrary to how \lstinline{a'} is declared
+% "Final": the second "c" can prove that it's equal to 1 but
+% \lstinline{a'} is not overridable in the "Family I2".
 
-In our example, the seal family "I.MonotonicF" is actually simulating
-interfaces as well.  We have
-\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
-and we assigned it to both "f" and "p", and later instantiate in
-"I2" with concrete implementation by overriding this family
-"I.MonotonicF".
+% In our example, the seal family "I.MonotonicF" is actually simulating
+% interfaces as well.  We have
+% \mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
+% and we assigned it to both "f" and "p", and later instantiate in
+% "I2" with concrete implementation by overriding this family
+% "I.MonotonicF".
 
-\textbf{Overridable/Sealed Family as better Sigma Type.} Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming.
+% \textbf{Overridable/Sealed Family as better Sigma Type.} Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming.
 There are cases where overriding and
-(definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
-\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
-because to prove
-the latter we need to have concrete definition of "f" in the family type, which
-makes "f" non-overridable at all. One option is to use sigma type
-\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
-a bundle of stuff with multiple refinements.
+% (definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
+% \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
+% because to prove
+% the latter we need to have concrete definition of "f" in the family type, which
+% makes "f" non-overridable at all. One option is to use sigma type
+% \mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
+% a bundle of stuff with multiple refinements.
 
-To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
+% To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
 In a sealed family, every field will expose its definition inside the sealed family but a family as a whole will be ``sealed'' with a ``relaxed family type''---it is an aggregation of the types instead of the concrete definitions of all the field.
@@
 
-For example, the "MonotonicF" in \cref{fig:plugin-example3}:
-we simply seal
-\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
-together inside an sealed family.
-Because the definition of "f" is exposed inside the family, 
-its definition is
-transparent and can be used to prove "p".
-However, because the enclosing family of "f" is overridable, to swap the
-implementation of "f", we just override the whole ``sealed'' family,
-like we do in "I2.MonotonicF".
+% For example, the "MonotonicF" in \cref{fig:plugin-example3}:
+% we simply seal
+% \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
+% together inside an sealed family.
+% Because the definition of "f" is exposed inside the family, 
+% its definition is
+% transparent and can be used to prove "p".
+% However, because the enclosing family of "f" is overridable, to swap the
+% implementation of "f", we just override the whole ``sealed'' family,
+% like we do in "I2.MonotonicF".
+
+% Non-overridable fields are still not expressive enough to achieve the full power of dependent type programming. There are cases where overriding and
+% (definition-relevant) reasoning are both needed. For example, we may want a monotonic function as a overridable field member. We cannot split into two fields
+% \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
+% because to prove
+% the latter we need to have concrete definition of "f" in the family type, which
+% makes "f" non-overridable at all. One option is to use sigma type
+% \mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
+% a bundle of stuff with multiple refinements.
+
+% To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type.
+In a sealed family, every field will expose its definition inside the sealed family but a family as a whole will be ``sealed'' with a ``relaxed family type''---it is an aggregation of the types instead of the concrete definitions of all the field.
+Thus we can override one sealed family as a whole and change the internal implementation as long as they can satisfy the ``relaxed family type''.
+
+% For example, the "MonotonicF" in \cref{fig:plugin-example3}:
+% we simply seal
+% \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
+% together inside an sealed family.
+% Because the definition of "f" is exposed inside the family, 
+% its definition is
+% transparent and can be used to prove "p".
+% However, because the enclosing family of "f" is overridable, to swap the
+% implementation of "f", we just override the whole ``sealed'' family,
+% like we do in "I2.MonotonicF".
+
+\textbf{Overridable Fields with dependency.}  Each field is by default non-overridable.
+By the discussion of \ref{chg:definition-relevant-reasoning}, we can simply reason and access the definition about the former fields in a later field, just like Field "a" and "b" in \cref{fig:plugin-example3} and we lose the ability to override.
+
+However, overridability is useful especially in Coq as programs in Coq are runnable.
+Consider the sceanrio when we use arbitrary monotonic function as field members \mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}, and we make sure that all the following reasoning are only about the monotonicity.
+Then it should be allowed to swap the implementation of that monotonic function via overriding.
+
+To resolve, we introduce overridable fields.
+Each overridable field will only expose its type instead of the concrete definition to the following default (non-overridable) fields.
+For example, the first failure in \cref{fig:plugin-example3} per the discussion of \ref{chg:definition-relevant-reasoning}: since overridability will break the inheritability of "c" and thus we don't allow such field to exist in the first place.
+Our plugin will only expose the fact that "a'" is an arbitrary natural number but hiding its concrete definition, and thus "eq_refl" will fail to type-check.
+
+In this stage, this feature is yet too weak since we cannot do definitional-relevant reasoning about overridable fields, like field "c" in the first failure.
+Thus we introduce \textit{dependency} between overridable fields---in the syntax of \mintinline{Coq}{Overridable pins {x,y,...} Field f := fexp}.
+In this case, field "f" will depend on the (overridable) field "x,y,...", and semantically, this means the concrete definition of "x,y,..." is \textit{exposed} to "fexp" during the type checking of "fexp".
+Every overridable field can only depend on former \textit{overridable} and \textit{definition-transparent} (non-"Qed"-Coq-term) fields.
+
+To override an overridable field, we use the syntax \\ \mintinline{Coq}{Override [pins {x',y',...}] Field f := fexp'} in the definition of the children family.
+We can change the dependency \textbf{arbitrarily} in the children family if we want, just by specifying the new "pins" component and there is no restriction on the "x',y',..." we choose.
+The only restriction is that the overriden "f" has to have the same type as the parent field "f".
+
+However, if we want to inherit any overridable field in the children family, we need to make sure every dependent (overridable) field is \textit{inherited instead of overridden} in the children family.
+That means, every overriding will lead to a chain of overriding on the depender fields.
+This coincides with our intuition about the example---when the dependee monotonic function is overridden, the proof about its monotonicity needs to be changed accordingly.
+Look at the field "f" and "p" in \cref{fig:plugin-example3} in both families.
+Both fields are initially filled with place-holder in the parent family, but then in children family "I2", "f" is overridden by "(+ 1)" and then to prove its monotonicity, we need to expose the concrete definition of "f" and thus we need to \textit{newly} pins "f" when providing the overriding "p".
 
@@
 theorem proving---the handler and the family will be automatically
-constructed during proving.\YZ{Cool. Can the same be done to FRecursor?}
+constructed during proving.\YZ{Cool. Can the same be done to FRecursor?}\EDJreply{I think last time I have answered it. FRecursor for fixpoint command}
 

### 2022-08-07 coqimpl.tex
Save. Changing family compilation part.
Rewrite CoqImpl to coincide with the latest implementation.

@@
 family components into Coq modules, family ``types'' into module types, and the context of a given term
-into parameters of the module. We need to faithfully
-reflect the \textit{late binding} nature of families: for each defining
-field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
+into parameters of the module.
+
+Since by default every field is inheritable, for each defining field (informally) $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
 ``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
-\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
-choose module and module parameters to achieve this wrapping---each
+\cL \sigma.
+T$, so that field $t$ can be inherited to all the future family with strengthen context $\sigma^+$ with the help of weakening from $\sigma^+$ to $\sigma$.
+However, instead of using Coq's universal quantifier, we choose module and module parameters to achieve this wrapping---each
 family type in the context will be one module parameter
@@
 field interactively with Coq.
+Note that, when compiling the context during the  compilation of a default (non-overridable)  field, all the default (non-overridable) former fields will expose their definition in the context while overridable fields will only expose their types.
 % Insert one pseudo-code example for explanation of the mechanism
@@
 Definition a : nat := 1. End a_4.
-Module Type a_5 (self_A: EmptySig).
-Parameter (a : nat). End a_5.
-Module Type a_6 := a_5 EmptyMod.
+Module Type a_6 (self_A: EmptySig).
+Include a_4. End a_5.
+
 \end{minted}
@@
 Definition b 
-  : self_A.a = self_A.a 
+  : self_A.a = 1
   := eq_refl. End b_7.
@@
 Taking \cref{fig:plugin-example1} as an example: immediately after the user inputs the line "Field a : nat := ..", our plugin will
-translate this statement into the "Module a_4", "Module Type a_5, a_6"
-three components.
+translate this statement into the "Module a_4", "Module Type a_6"
+two components.
 {The "EmptySig" is just an empty module type.} Then,
 these modules and module types will be generated and type-checked by Coq
-immediately. The compiled "Module a_4" is then part of the context of
-"Field b" (as the module type of "self_A") for "b" to refer to. The
+immediately. The compiled "Module a_6" is then part of the context of
+"Field b" (as the module type of "self_A") for "b" to refer to, so that "b" knows "a"'s concrete definition. The
 internal representation of family "A" is the list of these
 compiled modules and module types. We can achieve incremental
-checking as in \ref{chg:software-engineering}; overriding is also directly supported---for example, if we
+checking as in \ref{chg:software-engineering}. This makes all kinds of inheritance plausible---for example, if we
 want to override "Field a", we simply replace that compiled module with
-the new module inside the list. This compiled module type serves as an
-abstraction that enables overriding.
+the new module inside the list. 
+% This compiled module type serves as an abstraction that enables overriding.
 
-This is achieved by a complete compilation from a family to a module,
+After closing the top family, the user can access family "A" like it is a module. This is achieved by a complete compilation from a family to a module,
 via aggregating the list of the modules in the internal representation:
@@
 For example, "b_7" requires a module of type "a_6". Inside module "A",
-when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat}
+when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat := 1}
 inside the current surrounding "A". Then when we
@@
 }\EDJreply{Yes. Indivially compiling family member can achieve both requirements}
+
+\textbf{Implementing "Overridable" and Overriding Chain.}
+We implement "Overridable" by simply annotating the corresponding field type with a mark and its dependency.
+Then when compute the context during the compilation of the term for that field, we make sure only all the dependee and the non-overridable field will expose its definition in the context.
+Most parts of the implementation is quite similar to the default fields, except that when inheriting an overridable field, we need to make sure all its dependee is inherited earlier.
+
+When aggregating the fields into a module, we will make sure all the overridable fields to expose its definition for simiplicity in the compilation.
 
@@
 
-\textbf{Implementing "Final" and "Sealed Family".}
-We implement "Final" by
-exposing the whole definition directly into the compiled module type.
-Doing so will prohibit overriding. Say, "bop", declared "Final" in
-\cref{fig:plugin-example3}, is not overridable.
-We require the users to manual decorate field as `Final' so that the
-plugin will proceed with this special treatment.
-%\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
+% \textbf{Implementing "Final" and "Sealed Family".}
+% We implement "Final" by
+% exposing the whole definition directly into the compiled module type.
+% Doing so will prohibit overriding. Say, "bop", declared "Final" in
+% \cref{fig:plugin-example3}, is not overridable.
+% We require the users to manual decorate field as `Final' so that the
+% plugin will proceed with this special treatment.
+% %\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
 Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory.
@@
 
-% \textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.}
+% % \textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.}
 
-The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type.
+% The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type.
 The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
-For example, the compiled module type of
-"MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}".
+% For example, the compiled module type of
+% "MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}".
+Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
+}
+
+% The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type.
+The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
+% For example, the compiled module type of
+% "MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}".
 Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.

### 2022-08-07 coqimpl.tex
Fix Example in Overridable section.

@@
 Field a : nat := 1.
-Field b : a = a := eq_refl.
-Final Field a' : nat := 1. 
+Field b : a = 1 := eq_refl.
+Overridable Field a' : nat 
+                     := 1. 
+Overridable pins {a'}
+  Field b' : a' = 1 
+          := eq_refl.
 EndFamily.
@@
 Definition a : nat := 1. End a_4.
-Module Type a_6 (self_A: EmptySig).
+Module Type a_6.
 Include a_4. End a_5.
-
+Module b_7 (self_A: a_6).
+Definition b 
+  : self_A.a = 1
+  := eq_refl. End b_7.
+Module Type b_9.
+Include a_6. Include b_7. End b_9.
+Module a'_10 (self_A: b_9).
+Definition a' : nat := 1. End a_4.
 \end{minted}
@@
 \begin{minted}[fontsize=\footnotesize]{Coq}
-Module b_7 (self_A: a_6).
-Definition b 
-  : self_A.a = 1
-  := eq_refl. End b_7.
+Module Type a'_12.
+Include b_9. Include a'_10. End a'_12.
+Module b'_16 (self_A: a'_12).
+Definition b' : self_A.a' = 1 := eq_refl. End b'_16.
+Module Type b'_19.
+Include b_9. Parameter a': nat.
+Parameter b': a' = 1. End b'_19.
+Module test_20 (self_A: b'_19).
+Definition test : nat := 1. End test_20.
 
 Module A. (* Final Aggregation *)
-Include a_4. Include b_7. End A.
+Include a_4. Include b_7. Include a'_10.
+Include b'_16. Include test_20. End A.
 \end{minted}
@@
 Most parts of the implementation is quite similar to the default fields, except that when inheriting an overridable field, we need to make sure all its dependee is inherited earlier.
+For example, in \cref{fig:plugin-example1}, the module "b'_16" will compile "b'" in the context of "a'_12" where the concrete definition of "a'" is exposed since "b'" depends on "a'"; while the module "test_20" is compiled in the context "b'_19" only knowing the type of "a'" and "b'".
 
-When aggregating the fields into a module, we will make sure all the overridable fields to expose its definition for simiplicity in the compilation.
+When aggregating the overridable fields into a module, since everything has concrete definition now, there won't be any problem---we just use the same way that aggregates the default fields.
 

### 2022-08-07 coq-example.tex
Change the final example illustration parts.

@@
     using STLC_bool using STLC_prod.
+Family STLC_bool_prod_sum 
+    extends STLC 
+    using STLC_sum using STLC_bool_prod.
+Family STLC_bool_prod_sum_fix 
+    extends STLC 
+    using STLC_bool_prod_sum using STLC_fix.
+(* ... and more, 
+   there are 11 mixins families ... *)
 \end{lstlisting}
@@
 And we use a \textit{not-yet-formally-defined} \textbf{mixin} feature
-to mix the semantics and properties of two language features---product
-and boolean (in \cref{fig:STLC-example})---with vanilla STLC.
+to mix the semantics and properties of language features---product
+, boolean (in \cref{fig:STLC-example}), sum type and iso-recursive type---with vanilla STLC.
 
@@
 
-Our formulation for bare STLC is around 400+ LOCs; the two families implementing product and boolean features both take around 250+ LOCs. 
+Our formulation for bare STLC is around 400+ LOCs; each families implementing product and boolean features both take around 300+ LOCs. 
 % need comparison with the original implementation
@@
 soundness, with some postulation on both computation and property.
-Then we extend the base language with new features, and we instantiate
-the postulation on computation for
-abstract interpreters.
+Then we instantiate our abstract interpretation with two concrete computation---one for type inference and the other for constant propagation.
+Our analyzer will return the analysis result about the end of the program, e.g.
+our constant propagator will return what constant value each variable is at the end of the program.
 Thanks to the compilation to Coq module, we can directly run the resulting
@@
 
+% Then we extend the base language with new features, and we instantiate
+% the postulation on computation for
+% abstract interpreters.
+
 \begin{figure}[!htb]
-  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
+  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp3.pdf}
   \caption{Inheritance Diagram for Abstract Interpretation Example.
@@
 Purple arrow are extending.
-Green arrows are overriding.
-Grey box are sealed families.}\label{fig:abstract-interpretation-example}
+Green arrows mean there are overriding.}\label{fig:abstract-interpretation-example}
 \end{figure}
 
-We construct four families, in one inheritance chain, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful.
+We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful.
 They are here to show family inheritance is useful, and can extend stuff to all sorts of things.
@@
 Those people may extend "Lang" with other stuff like another type-safety proof or something.}
-
 The first family "Lang" define the big step operator (using fuel) of a
-while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200+ LOCs. 
+while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200 LOCs. 
 % We
@@
 
-The third Family "LangMore" extend "LangwAbs" with nat constant and
-addition, and if-then-else control flow, of course retaining all the
-soundness theorem. It takes about 200+ LOCs. This is one example of
-using family inheritance to support new language feature, and compatible
-with the existent reasoning in "LangwAbs.AbsInterpreter".
+% The third Family "LangMore" extend "LangwAbs" with nat constant and
+% addition, and if-then-else control flow, of course retaining all the
+% soundness theorem. It takes about 200+ LOCs. This is one example of
+% using family inheritance to support new language feature, and compatible
+% with the existent reasoning in "LangwAbs.AbsInterpreter".
 
-The fourth family "LangRun" instantiate some of the postulates of the
-"LangMore" and recover computational information of "analyze". The "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
-to act as a type-analyzer. It takes about 200+ LOCs. This is one example
-of instantiate the detail computation of abstract interpretation.
+The third family and the fourth family "LangTyping" and "LangCP" instantiate some of the postulates of the
+"LangwAbs" and recover computational information of "analyze". The corresponding fields in "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
+to act as a type-analyzer and constant-propagator respectively. It takes about 200+ LOCs. 
+% These are two examples of instantiate the detail computation of abstract interpretation.
 
-And at the end, we compile LangRun into module and use Coq to run it on
-some simple queries. This "LangRun" also illustrates that we still have
+And at the end, the compiled module for "LangTyping" and "LangCP" is put into tests over simple queries. These "LangTyping" and "LangCP" also illustrates that we still have
 computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}\EDJreply{Done.

### 2022-08-08 syntax.tex
Add the part why we use predicative MLTT.
Add a paragraph on how seal corresponds to override

@@
     \text{Inductive Type Signature}   \quad & \tau       &&= w\cdot~|~w^+~\tau~A~B~|~\tau[\gamma] \\
-    \text{Linkage Signature}          \quad & \sigma       &&= \nu\cdot~|~\nu^+~\sigma~\{f\}~T~|~\sigma[\gamma]~|~p_1\nu~\sigma \\
+    \text{Linkage Signature}          \quad & \sigma       &&= \nu\cdot~|~\nu^+~\sigma~\{f\}~T~|~\sigma[\gamma]~|~p_1\nu~\sigma~|~"RecSig"~\tau~T \\
     \text{Seal}          \quad & f       &&= t~|~p_f\nu~\sigma\\    
-    \text{Types}   \quad & A, B, T        &&= ...~|~\bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\\
-    \text{Terms}   \quad & t, t_1, t_2    &&= ...~|~"W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t
+    \text{Types}   \quad & A, B, T        &&= ...~|~\bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma~|~"RecSig1"~A~B~T~|\\
+    \text{Terms}   \quad & t, t_1, t_2    &&= ...~|~"W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t~|\\ & &&"prjR"~j~o~|~"Wrec"~t_1~t_2~|~"inh"~h~t\\
+    \text{Inheritance Structure} \quad & h, h' &&= "inhid"~|~"inhov"~h~t~|~"inhext"~h~t~|~"inhinh"~h~|~"inh"~h~|~"inhnest"~h~h'
   \end{alignat*}
@@
 
-This sealing is the general version of the sealed family implemented in
-the plugin. Our plugin will deduce a sealed ``interface'' (which is the "A" in \ruleref{Sig Add})
-for a sealed family using a fixed set of rules, but here sealing in
-metatheory can be arbitrary local morphism  $\goodTerm{\Gamma , \cC
-\sigma}{\_}{A[\pi_1]}$.
-With the definition of sealing judgement in mind, we can see
-\ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the
-implementation. In the implementation, we will have "self__" all around
-when defining new fields in our plugin, and this "self__" is exactly the
-$A$ in the context of the two fields---in our implementation, sealing
-happens only to inductive types (where each inductive type we only know
-\mintinline{Coq}{T : Set}).
+% This sealing is the general version of the sealed family implemented in
+% the plugin. Our plugin will deduce a sealed ``interface'' (which is the "A" in \ruleref{Sig Add})
+% for a sealed family using a fixed set of rules, but here sealing in
+% metatheory can be arbitrary local morphism  $\goodTerm{\Gamma , \cC
+% \sigma}{\_}{A[\pi_1]}$.
+% With the definition of sealing judgement in mind, we can see
+% \ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the
+% implementation. In the implementation, we will have "self__" all around
+% when defining new fields in our plugin, and this "self__" is exactly the
+% $A$ in the context of the two fields---in our implementation, sealing
+% happens only to inductive types (where each inductive type we only know
+% \mintinline{Coq}{T : Set}).
 
@@
 To make things concrete, we need to introduce the rules for recursors.
+
+Finally, we make a note here on how sealing can easily express the overidability in our implementation.
+To expose the definition in our metatheory, we model that by adding equality reflection into our metatheory (and thus change our setting into extensional type theory).
+Then we can consider exposing the concrete definition as the existence of an identity type in the context.
+The key insight of the sealing in our metatheory, is that it gives the user the ability to manually specify which former fields expose its definition and which  former fields only expose their types---for those only expose their types, the corresponding identity type will be relaxed into a unit type.
+Now look back at the overridability mechanism in our implementation, it is a much restricted version: each overridable field can only see the concrete definition of the former non-overridable(default) fields and depending overridable fields; for each non-overridable(default) field, only the concrete definition of former non-overrdiable(default) fields are exposed.
 
@@
 \Rule[name=Hdlers]
-{\goodWSig{\Gamma}{\sigma}{n}
+{\goodWSig{\Gamma}{\tau}{n}
 \quad \goodType{\Gamma}{R}{}
 }
-{\goodSig{\Gamma}{"RecSig"~\sigma~R}{n}}
+{\goodSig{\Gamma}{"RecSig"~\tau~R}{n}}
 $$
@@
 \Rule[name=Hdler-Proj]
-{ \goodWSig{\Gamma}{\sigma}{N}
-\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\sigma~R)}
-\quad n < N
-}
-{\goodTerm{\Gamma}{"prjR"~n~o}{("RecSig1"~(\pi_1^n \sigma)~(\pi_2^n \sigma)~R)[\pi_1]}}
+{ \goodWSig{\Gamma}{\tau}{N}
+\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\tau~R)}
+\quad j < N
+}
+{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)[\pi_1]}}
 $$
@@
 \Rule[name=Rec-Constr]
-{ \goodWSig{\Gamma}{\sigma}{N}
-\quad \goodTerm{\Gamma}{T}{\bW \sigma}
+{ \goodWSig{\Gamma}{\tau}{N}
+\quad \goodTerm{\Gamma}{T}{\bW \tau}
 \quad \goodType{\Gamma}{R}{}
-\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\sigma~R)}
+\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
 \quad \goodTerm{\Gamma}{w}{El~("W"~T)}
@@
 (4) Why are RecSig1, RecSig, Wrec, prjR not listed in the syntax figure?
-}\EDJreply{Thx. Working on it. I seem to forgot to add inheritance as well.}
+}\EDJreply{Thx. Working on it. I seem to forgot to add inheritance as well.}\EDJreply{Done. Please Check.}
 $$

### 2022-08-23 coqimpl.tex
Rewriting Coqimpl 1. adding overridable pins inductive 2. reintroducing closing fact 3. mixin 4. partial recursor

@@
 
-% Mention still needing total recursor, for exhaustiveness checking
+% Attempt to rewrite the section for Overridable pins inductive/Closing Fact , and also introduce the concept of mixin
 
-\textbf{Hack: Global Reasoning.}\YZ{This is more like an escape hatch. Escape from family poly into vanilla Coq. Maybe these two hacks should be grouped into one section}
-% explain the places using ``Closing Fact''
-% 1. No need to export partial recursor
-% 2. Computational rule for the total recursor
-Unfortunately, there are cases an extensible proof brings more boilerplate code. For example, in STLC, we expect
-``\mintinline{Coq}{~ value (tm_app x y)}'' to hold in all future
-extensions of STLC---applications will never be considered as values
-by any extension.
-We know the correct way to prove the proposition
-\mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use
-"FRecursor" to inductively reason "value (tm_app x y)". However, this
-means that every time we extend "value" with new value forms, we will
-need to extend corresponding inductive cases for "app_not_value". We can
-imagine the newly extended proof to be boring case analysis that the
-newly added value forms are not "tm_app". What's worse, similar
-scenario can happen on other statements like \mintinline{Coq}{~ value
-(tm_if cond x y)}. Everything here is extended correctly but we end up
-having boilerplate code, \textbf{semantically}.
-\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}\EDJreply{Done}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family STLC.
-  FInductive tm : Set := ... 
-  FInductive value : tm -> Prop := ... 
-  Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
-      by { intros x y H; inversion H; eauto }.
-  Fail Ltac inv := 
-    match goal with 
-    | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
-    (* Fail: value, tm_app, value_not_app unfound *)
-    end. 
-  MetaData tactic1.
-    Ltac inv := (* ... the same definition as above ... *)
-  EndMetaData.
-  (* ... STLC example continue ... *)
-EndFamily.
-\end{lstlisting}  
-\caption{Example for Global Reasoning and MetaData, using in STLC example}\label{fig:plugin-example-global-reasoning-meta-data}
-\end{figure}
+However, "FRecursor" provides barely enough functionality for manipulating (extensible) inductive type, and it is far from satisfactory to be a legitimate programming interface---for example, the necessary tactic "inversion", "discriminate", "injection" is not applicable on (extensible) inductive type.
 
-Our proposal is to allow global reasoning by providing "Closing Fact" command. Ultimately, "Closing Fact" is asserting a constraint upon all the
-possible extension of the surrounding family and claiming that a proof script
-can solve the constraints.\YZ{This is a good one-liner that summarizes the intended usage scenarios of "Closing Fact".
+To resolve this problem, we notice that, the power of "FRecursor" comes from its distinction from the mundane fields and its ability to stab through the abstraction of "FInductive".
+However, the latter is quite similar on how "Overridable ..
+pins {..}" can see through the abstraction of former pinned overrdiable fields.
+This inspires us to generalize pins to not only apply to overrdiable fields, but also "FInductive" and "FRecursor" fields.
+
+\textbf{Overridable pins Inductive Type and FRecursor}.
+
+Also, when pins "FInductive" and "FRecursor", we can have the primitive coq-providing recursor and judgemental equality for computation ability for "FRecursor".
+
+Upon this design, it is natural to make "Overridable" fields to provide ``automatic overriding'' as it is likely the inherited tactic is able to provide a proof even in the context with the extended inductive type.
+
+But a question arises: do we consider an ``automatic overridden'' field \textit{inheriting} the parent field or \textit{overriding} the parent field? 
+% TODO: Mention still needing total recursor, for exhaustiveness checking
+
+\textbf{Overridable Field and Closing Fact}
+It is natural to consider the ``automatic overridden'' resulting inheritance, as we are using the same tactic (or the same untyped Gallina term to provide the proof).
+But for type theorists, these are definitely overriding because these are definitely different typed Gallina term!
+
+Instead of making a decision here, we introduce "Closing Fact":  "Closing Fact" will make a field tied with the untyped proof script and only allows inheritance.
+Of course, during inheritance, the untyped proof script will always be re-run and introduce (potentially) \textbf{different} typed terms\footnote{Thus called "Closing Fact" as its semantic depends on the closed  surrounding family in a global sense}.
+Thus, we encourage the user to use it in a proof-irrelevant way, where the difference between typed terms can be ignored.
+Then, the ``automatic overridden'' will always lead to overriding.
+
+This is similar to "FRecursor"---we only allow inheritance for "FRecursor" as its semantic can be considered as a tie between a handler family and an extensible inductive type, where during each inheritance, the tie will be rechecked.
+
+Just like "Overridable ..
+pins {..}", "Closing Fact" provides the ability to stab through all the abstraction of the former fields, and thus has the similar functionality  as "Overridable ..
+pins {..}".
+
+\textbf{Mixin}. The reason we introduce two different semantics for two similar functionality is due to the fact that, we want to introduce mixin in our facility.
+
+% TODO: mixin implementation sketch
+
+However, we don't allow any overriding happening in either mix-edin inheritance, because it will break the commutativity and complicate the semantic without much benefit.
+Overriding is not allowed in this case, but "FRecursor" and "Closing Fact" is still allowed since they only allow to be inherited.
+
+% \textbf{Hack: Global Reasoning.}\YZ{This is more like an escape hatch. Escape from family poly into vanilla Coq. Maybe these two hacks should be grouped into one section}
+% % explain the places using ``Closing Fact''
+% % 1. No need to export partial recursor
+% % 2. Computational rule for the total recursor
+% Unfortunately, there are cases an extensible proof brings more boilerplate code. For example, in STLC, we expect
+% ``\mintinline{Coq}{~ value (tm_app x y)}'' to hold in all future
+% extensions of STLC---applications will never be considered as values
+% by any extension.
+% We know the correct way to prove the proposition
+% \mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use
+% "FRecursor" to inductively reason "value (tm_app x y)". However, this
+% means that every time we extend "value" with new value forms, we will
+% need to extend corresponding inductive cases for "app_not_value". We can
+% imagine the newly extended proof to be boring case analysis that the
+% newly added value forms are not "tm_app". What's worse, similar
+% scenario can happen on other statements like \mintinline{Coq}{~ value
+% (tm_if cond x y)}. Everything here is extended correctly but we end up
+% having boilerplate code, \textbf{semantically}.
+% \begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}\EDJreply{Done}
+% \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+% Family STLC.
+%   FInductive tm : Set := ... 
+%   FInductive value : tm -> Prop := ... 
+%   Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
+%       by { intros x y H; inversion H; eauto }.
+%   Fail Ltac inv := 
+%     match goal with 
+%     | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
+%     (* Fail: value, tm_app, value_not_app unfound *)
+%     end. 
+%   MetaData tactic1.
+%     Ltac inv := (* ... the same definition as above ... *)
+%   EndMetaData.
+%   (* ... STLC example continue ... *)
+% EndFamily.
+% \end{lstlisting}  
+% \caption{Example for Global Reasoning and MetaData, using in STLC example}\label{fig:plugin-example-global-reasoning-meta-data}
+% \end{figure}
+
+% Our proposal is to allow global reasoning by providing "Closing Fact" command. Ultimately, "Closing Fact" is asserting a constraint upon all the
+% possible extension of the surrounding family and claiming that a proof script
+% can solve the constraints.\YZ{This is a good one-liner that summarizes the intended usage scenarios of "Closing Fact".
 Should probably bring it out earlier.}\EDJreply{I put it here as you suggested but I think this one-liner at here will cause more confusion because it is too abstract.}
-The assertion will only be verified when the whole family is closed and compiled into a module.
+% The assertion will only be verified when the whole family is closed and compiled into a module.
 
-% We provide "Closing Fact" as a command to
-% carry out this global reasoning, attached with a proof script and proof
-% term. This "Closing Fact" will look like an assertion during the
-% construction of the enclosing family. This assertion will only be
-% verified when the whole family is closed and compiled into a module,
-% with the help of the attached proof script and proof terms.
+% % We provide "Closing Fact" as a command to
+% % carry out this global reasoning, attached with a proof script and proof
+% % term. This "Closing Fact" will look like an assertion during the
+% % construction of the enclosing family. This assertion will only be
+% % verified when the whole family is closed and compiled into a module,
+% % with the help of the attached proof script and proof terms.
 
-Here we show the example to prove "value_not_app" in
-\cref{fig:plugin-example-global-reasoning-meta-data}, the script inside
-curly brackets after "by" is the attached proof script that will be
-executed upon "EndFamily", which triggers the compilation of the family
-being defined.
-Since we are dealing with a concrete (vanilla) inductive type during compilation,
-we can use Coq's "inversion" tactic to complete this proof.
-This "Closing Fact", as an assertion, will have no problem being
-inherited, but the proof script will be executed every time a derived
-family is compiled.
+% Here we show the example to prove "value_not_app" in
+% \cref{fig:plugin-example-global-reasoning-meta-data}, the script inside
+% curly brackets after "by" is the attached proof script that will be
+% executed upon "EndFamily", which triggers the compilation of the family
+% being defined.
+% Since we are dealing with a concrete (vanilla) inductive type during compilation,
+% we can use Coq's "inversion" tactic to complete this proof.
+% This "Closing Fact", as an assertion, will have no problem being
+% inherited, but the proof script will be executed every time a derived
+% family is compiled.
 
-The downside is that "Closing Fact" cannot immediately check the correctness of the
-attached proof term/proof script because this verification only happens when we
-finished defining a family, and thus a bit harder to handle.\YZ{'harder' in what way?
+% The downside is that "Closing Fact" cannot immediately check the correctness of the
+% attached proof term/proof script because this verification only happens when we
+% finished defining a family, and thus a bit harder to handle.\YZ{'harder' in what way?
 I suppose you mean the "Fail Ltac" in Figure 5?}\EDJreply{The proof script cannot be run immediately to reply the user whether the proof script is even correct.
@@
 only compile part of the family to the point closing fact is used) but it is a big engineering difficulty for me.}
-We expect it to be used only sporadically for maintainability, and as an alternative to using "FRecursor".
+% We expect it to be used only sporadically for maintainability, and as an alternative to using "FRecursor".
 This is called global reasoning because this is a post-hoc reasoning when the family as a \textit{whole} is compiled into a module.
@@
 
-\textbf{Propositional Partial Recursor.} Notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''---they will possibly disrupt arbitrary extension of inductive type.
+\textbf{Propositional Partial Recursor.} This section is introducing the concept of \textit{partial recursor}.
+The motivations are various: (a) we want to have some semantic characterization of extensible inductive type.
+(b) We want to show that, some tactics reasoning like "inversion" is still valid even in extensible inductive type.
+(c) More generally speaking, notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''---they will possibly disrupt arbitrary extension of inductive type.
 For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)} in the "Family STLC" either using "FRecursor" or "Closing", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
+We want to see what leads to a conservative extension---and the partial recursor is helpful for this question.
 
-This raise an interesting question: what propositions are conservative extension?
+% This raise an interesting question: what propositions are conservative extension? 
 For example, apparently we know injectiveness of constructor, (i.e.

### 2022-08-29 coqimpl.tex
Save.

@@
 
-However, "FRecursor" provides barely enough functionality for manipulating (extensible) inductive type, and it is far from satisfactory to be a legitimate programming interface---for example, the necessary tactic "inversion", "discriminate", "injection" is not applicable on (extensible) inductive type.
+However, "FRecursor" provides barely enough functionality for manipulating our(extensible) inductive type, and it is far from satisfactory to be a legitimate programming interface.
+(1) We do not yet have any computational information using this "FRecursor" until its compilation into module, i.e.
+we cannot do computation with it during the construction of the family using the current compilation strategy.
+(2) The necessary tactic like "discriminate" and "injection" is not applicable on (extensible) inductive type.
 
-To resolve this problem, we notice that, the power of "FRecursor" comes from its distinction from the mundane fields and its ability to stab through the abstraction of "FInductive".
+\textit{Generating Scheme for Propositional Computation Axiom for FRecursor}.
+We support "FScheme" that our plugin will automatically generate(postulate) the computational ``axiom'' for the specified  into the family.
+During compilation, these ``axiom'' can be easily proved/instantiated with concrete/vanilla inductive type.
+
+However, this method, though remedies problem (1) to some extent, is only introducing computational information \textbf{propositionally} but not judgementally.
+Thus, for example, during theorem proving, we might need the help of rewriting tactic on these computational ``axiom'' to justify the computation.
+And in the full dependent type theory, propositional equality is sometimes not enough and bringing a lot of coding challenges\footnote{For example, for "u : F a"， "v : F b", "u = v" is not even well-typed when "a = b" only propositionally hold.
+Transport is required during reasoning.}.
+
+% To resolve (2), we can simply  Our following three sections will try to resolve these problems. 
+% we notice that, the power of "FRecursor" comes from its distinction from the mundane fields and its ability to stab through the abstraction of "FInductive".
 However, the latter is quite similar on how "Overridable ..
@@
 
-\textbf{Overridable pins Inductive Type and FRecursor}.
+\textit{Overridable pins Inductive Type and FRecursor}.
+We notice that, the power of "FRecursor" comes from its distinction from the mundane fields and its ability to stab through the abstraction of "FInductive".
+However, the latter is quite similar on how "Overridable ..
+pins {..}" can see through the abstraction of former pinned overrdiable fields.
+This inspires us to generalize pins---we support pinning inductive type and "FRecursor".
+Once "FInductive" and "FRecursor" fields are pinned, we can stab through its abstraction and have the primitive coq-providing recursor and \textbf{judgemental equality} for the computational information for "FRecursor".
 
-Also, when pins "FInductive" and "FRecursor", we can have the primitive coq-providing recursor and judgemental equality for computation ability for "FRecursor".
+The compilation mechanism is simply make the underlying, compiled definition of inductive type and recursor included inside the parameters of the module, just like how the the concrete definition of overridable fields are included inside the parameters when pinned.
+This way, the computational information from vanilla Coq can be transimitted into the context.
 
-Upon this design, it is natural to make "Overridable" fields to provide ``automatic overriding'' as it is likely the inherited tactic is able to provide a proof even in the context with the extended inductive type.
+% However, only overridable fields can pin the (extensible) inductive type and "FRecursor".
 
-But a question arises: do we consider an ``automatic overridden'' field \textit{inheriting} the parent field or \textit{overriding} the parent field? 
+% Upon this design, it is natural to make "Overridable" fields to provide ``automatic overriding'' as it is likely the inherited tactic is able to provide a proof even in the context with the extended inductive type.
+
+% But a question arises: do we consider an ``automatic overridden'' field \textit{inheriting} the parent field or \textit{overriding} the parent field? 
 % TODO: Mention still needing total recursor, for exhaustiveness checking
+
+% To fulfill \ref{chg:software-engineering}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
+
+% For example,
+
+%   For example, currently it is not possible to define a
+% vanilla inductive type dependent on our ``extensible'' inductive type
+% due to the invisibility, unless we compile the whole family. Similarly,
+% customized tactic expressions as required by \ref{chg:software-engineering} cannot
+% use lemmas or theorems in the fields of the current defining family as
+% well.
+
+% To remedy this, we support a "MetaData" command. This command
+% bundles any original Coq primitive (in the current family
+% context) into the compiled module type, and makes them visible to the
+% following defining field. Doing so, we can define customized tactic as
+% in \ref{chg:software-engineering} that uses lemmas proved in the family.
+
+% Following needs a rewrite
+To achieve "injection" in problem (2), we can easily generate/postulate $n$ propositions for $n$ constructors, just like how we did for computational information.
+But for "discriminate" tactic, this way will generate $O(n^2)$ propositions.
+
+What's more, to manipulate inductive type, it is not necessarily enough to only support "injection" and "discriminate" tactic (and their corresponding axioms).
+Recall in vanilla inductive type, we believe that eliminator is the strong-enough ``toolkit'' because the eliminator (with its computational information) is the \textit{universal property} of the inductive type---it extensionally describes the inductive type up to (certain) isomorphism.
+This universal property is much more powerful than "injection" and "discriminate" and maybe the most powerful tool we can have to reason about inductive type.
+
+For example, given a new inductive type, to prove "discriminate" between two constructors, i.e., "c$_i$ ..
+$\neq$ c$_j$ .." for two different constructors, we simply use the eliminator to map each constructor to different natural number, assuming we can "discriminate" different natural number.\footnote{Rigorously, to avoid any suspicious circular reasoning, we can use large elimination to map them to $\top$ and $\bot$, and because $\exists a : ?, \top$ holds for the former not the later, thus $\top \neq \bot$.} "injection" can be supported in a similar way---we can simply let eliminator map each inductive type to the arguments of the constructors.
+
+Thus it is natural for us to think if (an altered form of) eliminator can be postulated into the context, since apparently we cannot directly postulate the original eliminator here---because the sole existence of the eliminator prevents any possible future extension of the inductive type.
+
+Thus we alter eliminators into \text{partial recursor}, such that (1) it will not hinder the future extension, and (2) it is still powerful enough to derive injectiveness like tactic "inject" and "discriminate".
+What's more, (3) we can actually prove this \text{partial recursor} is actually good enough because it can act as an \textit{extensional characterization} for extensible inductive type.
+We will call it \textit{propositional} because we only prpositionally postulate its computational informations in Coq.s
+
+\textbf{Propositional Partial Recursor.} Partial recursor is simply adding partially into the signature of the eliminator, by wrapping each motive with "option". 
+% This section is introducing the concept of \textit{partial recursor}.
+The motivations are various: (a) we want to have some semantic characterization of extensible inductive type.
+(b) We want to show that, some tactics reasoning like "inversion" is still valid even in extensible inductive type.
+(c) More generally speaking, notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''---they will possibly disrupt arbitrary extension of inductive type.
+For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)} in the "Family STLC" either using "FRecursor" or "Closing", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
+We want to see what leads to a conservative extension---and the partial recursor is helpful for this question.
+
+% This raise an interesting question: what propositions are conservative extension? 
+% For example, apparently we know injectiveness of constructor, (i.e.
+\mintinline{Coq}{STLC.tt ≠ STLC.ff}) must be conservative because it holds for all extension of the inductive type "STLC.tm".
+But is there a \textbf{``best'' proposition} (i.e.
+the proposition that everything the proposition derives are exactly conservative extension)?
+
+% Since injectiveness of the constructors are derived by recursor in vanilla inductive type, an educated guess would be related to recursor.
+But we know the original recursor won't work because the mere existence of the original recursor will break extensibility.
+Thus we need to modify it.
+% The simplest idea is to allow partiality---i.e., returning
+% \mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
+
+% We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors?
+I suppose it has to do with the computational axioms?}\EDJreply{Yes.
+I want to emphasize the equality used by computational axiom is not judgemental equality.
+So Coq cannot automatically compute/reduce when a partial recursor applied to a term.
+This reduction can only be semi-automatically done by using [rewrite tactic and the computational axiom].}
+For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}\EDJreply{Added below please check.}
+
+\begin{minted}[escapeinside=@@]{Coq}
+(* Inside Family B *)
+FInductive b : Set := tt : b | ff : b.
+b_prec : forall R, option R -> option R -> b -> option R.@\YZ{Not clear what it takes to define/prove partial recursors.
+The reader needs to be told if they require breaking abstraction like recursors do.}\EDJreply{What do you mean?
+recursor doesn't break abstraction but break extensibility.
+Below mentioned partial recursor doesn't break extensibility}@
+\end{minted}
+with appropriate computational axioms.
+This partial recursor differs from original recursor only by decorating an "option" with the return type thus possible to be automatically generated (and the case is similar for those computational axiom).
+Thanks to the partiality, partial recursor doesn't break
+extensibility (i.e. all future extensions of "b" can support this
+"b_prec") because for example, "b_prec" can just return "None" when bump into extended constructor in the future.
+
+The second thing to check is that, if partial recursor is theoretically enough for injectiveness of the constructor. The rough idea to achieve this
+is to use "b_prec" to reflect our special inductive type into vanilla
+inductive type, and use the original "injection" and "discriminate"
+tactics. For example, we want to prove \mintinline{Coq}{B.tt ≠
+B.ff}, we simply 
+\begin{minted}{Coq}
+Definition reflect : B.b → option bool 
+                   := b_prec (fun _ => bool) (Some true) (Some false).
+Definition tt_neq_ff : B.tt = B.ff → False.
+(* Because B.tt = B.ff → reflect B.tt = reflect B.ff 
+      → Some true = Some false    (by computational axiom)
+      → False           (by discriminate of bool and injection of Some) *)
+\end{minted}
+
+Thirdly, even though we cannot prove partial recursor is \textit{the best} reasoning tool for extensible inductive type,
+the above "reflect" function implies that propositional partial
+recursor is a good \textit{extensional characterization} of (non-indexed) extensible inductive type (under certain constraints)---because we can ``embed'' vanilla inductive types into
+types supporting a partial recursor: "reflect" is actually a left
+inverse of an injection "bool → self_B.b". A bit more formally, and
+restricting our focus to non-indexed inductive types:
+
+\begin{theorem}\label{thm:prec-complete} Given a list of $n$ pairs of types $\{ x : "A"_i \vdash "B"_i(x) \}_{i}$ s.t.
+
+  \begin{itemize}
+    \item \textlabel{(Detectable Partiality)}{prop:detectable-partiality} for each pair $x : A_i \vdash B_i(x)$,
+    \begin{minted}{Coq}
+      Axiom detectable:
+      forall {T} {a} (f: Bᵢ a -> option T),
+        {forall x, f x <> None} + {exists x, f x = None}.
+    \end{minted} 
+    \item Define $"C"$ as the inductive type using this list, i.e.,
+    \begin{minted}{Coq}
+Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
+    \end{minted}
+    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
+    \begin{itemize}
+      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
+      \item a partial recursor
+      \begin{minted}[escapeinside=@@]{Coq}
+prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint. Let me fix other parts first}@
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), .., 
+      D -> option R
+      \end{minted}
+      \item and $n$ (propositional) computational axioms: for all $i$, 
+      \begin{minted}{Coq}
+prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
+    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
+      \end{minted}
+    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
+    \end{itemize}
+  \end{itemize}
+  Then there exists an \textbf{embedding} from "C" to "D".
+  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
+  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of
+  "C" and "prec" of "D" (both acting like identity) such that
+  \begin{minted}[escapeinside=@@]{Coq}
+    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
+If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
+I mentioned left inverse 'linv' is the generalization}@
+  \end{minted}
+\end{theorem}
+The reason we have this weird looking \ref{prop:detectable-partiality} is because the construction of "linv"---at one point, using "prec" to construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option C) → option C}, generally not provable, only if we have \mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
+This \ref{prop:detectable-partiality} is exactly transforming former to the later: recall that, according to the semantic of inductive type, \mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using "linv" on the substructure of "D".
+This function points out each of the translation result from "D" to "C" and some of them are "None" because of failing.
+For us, we only care about if \textbf{any} of the substructure of "D" is failed to translate---if any one substructure fails, then the whole transltion fails and thus we should have "None" as return.
+
+\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that includes all of the examples shown in this paper.
+\ref{prop:detectable-partiality} is not supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
+In those cases, we still have the fact that partial recursor is a conservative extension, but we cannot show that partial recursor is the extensional characterization any more。
+
+We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
+actually be considered as a constructor, because they can be reflected
+to real constructors of "C" using the left inverse "linv" (directly generalization of "reflect : self_B.b → option bool" of the above). Thus, \cref{thm:prec-complete}
+implies that (1) every future extension of the inductive type can
+support this partial recursion (trivially); (2) every type supporting
+this partial recursor with its computational axioms at least supports
+these constructors because of the embedding. In other words, \textit{a
+type (at least) supports these constructors if and only if this type
+supports the corresponding partial recursor (with the computational
+axioms)}.  Thus, we can argue that the partial recursor gives a sound and
+complete extensional characterization of all the extension of a given (non-indexed)
+inductive type.\YZ{What about indexed inductive types?}\EDJreply{Done.
+I also have a formalization of this proof now, for indexed type and dependent eliminator.
+Though the formalization is about single constructor.}
+
+We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
+\begin{minted}{Coq}
+  pdelim : ∀ (P : D -> Set), .., 
+  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
+    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
+\end{minted}
+To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable.
+Please refer to the supplementary material and appendix for a formalization of the left inverse in the context of indexed inductive type.
+
+Though partial recursor is provable by simply using "FRecursor", for convenience, we support using "FScheme" to automatic generate it.
+At the meantime, we implement a tactic "prec_discriminate" that when fed with the appropriate partial recursor, it can discriminate an absurd equation between two different constructors of extensible inductive type.
+
+Finally, we should point out that, the mere postulation of partial recursor is \textit{conservative}.
+In other word, generally speaking, using "FRecursor" to prove theorems is not necessarily a ``\textit{conservative extension}''---the proved theorems might hinder future extensions of the inductive type.
+For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)}\footnote{Achieved by three steps: first proving partial recursor using "FRecursor", then use partial recursor to prove "tm_var i $\neq$ tm_lam v body" and finally using "FRecursor" prove "var_not_value" by induction on "value"} in the "Family STLC" either using "FRecursor", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
+But partial recursor will not cause this problem.
 
@@
 
-% To fulfill \ref{chg:software-engineering}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
-
-% For example,
-
-%   For example, currently it is not possible to define a
-% vanilla inductive type dependent on our ``extensible'' inductive type
-% due to the invisibility, unless we compile the whole family. Similarly,
-% customized tactic expressions as required by \ref{chg:software-engineering} cannot
-% use lemmas or theorems in the fields of the current defining family as
-% well.
-
-% To remedy this, we support a "MetaData" command. This command
-% bundles any original Coq primitive (in the current family
-% context) into the compiled module type, and makes them visible to the
-% following defining field. Doing so, we can define customized tactic as
-% in \ref{chg:software-engineering} that uses lemmas proved in the family.
-
-% Following needs a rewrite
-
-\textbf{Propositional Partial Recursor.} This section is introducing the concept of \textit{partial recursor}.
-The motivations are various: (a) we want to have some semantic characterization of extensible inductive type.
-(b) We want to show that, some tactics reasoning like "inversion" is still valid even in extensible inductive type.
-(c) More generally speaking, notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''---they will possibly disrupt arbitrary extension of inductive type.
-For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)} in the "Family STLC" either using "FRecursor" or "Closing", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
-We want to see what leads to a conservative extension---and the partial recursor is helpful for this question.
-
-% This raise an interesting question: what propositions are conservative extension? 
-For example, apparently we know injectiveness of constructor, (i.e.
-\mintinline{Coq}{STLC.tt ≠ STLC.ff}) must be conservative because it holds for all extension of the inductive type "STLC.tm".
-But is there a \textbf{``best'' proposition} (i.e.
-the proposition that everything the proposition derives are exactly conservative extension)?
-
-Since injectiveness of the constructors are derived by recursor in vanilla inductive type, an educated guess would be related to recursor.
-But we know the original recursor won't work because the mere existence of the original recursor will break extensibility.
-Thus we need to modify it.
-The simplest idea is to allow partiality---i.e., returning
-\mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
-
-We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors?
-I suppose it has to do with the computational axioms?}\EDJreply{Yes.
-I want to emphasize the equality used by computational axiom is not judgemental equality.
-So Coq cannot automatically compute/reduce when a partial recursor applied to a term.
-This reduction can only be semi-automatically done by using [rewrite tactic and the computational axiom].}
-For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}\EDJreply{Added below please check.}
-
-\begin{minted}[escapeinside=@@]{Coq}
-(* Inside Family B *)
-FInductive b : Set := tt : b | ff : b.
-b_prec : forall R, option R -> option R -> b -> option R.@\YZ{Not clear what it takes to define/prove partial recursors.
-The reader needs to be told if they require breaking abstraction like recursors do.}\EDJreply{What do you mean?
-recursor doesn't break abstraction but break extensibility.
-Below mentioned partial recursor doesn't break extensibility}@
-\end{minted}
-with appropriate computational axioms.
-This partial recursor differs from original recursor only by decorating an "option" with the return type thus possible to be automatically generated (and the case is similar for those computational axiom).
-The partiality doesn't break
-extensibility (i.e. all future extensions of "b" can support this
-"b_prec") because "b_prec" can just return "None" when bump into extended constructor in the future.
-
-The first thing to check is that, if partial recursor is theoretically enough for injectiveness of the constructor. The rough idea to achieve this
-is to use "b_prec" to reflect our special inductive type into vanilla
-inductive type, and use the original "injection" and "discriminate"
-tactics. For example, we want to prove \mintinline{Coq}{B.tt ≠
-B.ff}, we simply 
-\begin{minted}{Coq}
-Definition reflect : B.b → option bool 
-                   := b_prec (fun _ => bool) (Some true) (Some false).
-Definition tt_neq_ff : B.tt = B.ff → False.
-(* Because B.tt = B.ff → reflect B.tt = reflect B.ff 
-      → Some true = Some false    (by computational axiom)
-      → False           (by discriminate of bool and injection of Some) *)
-\end{minted}
-
-Even though we cannot prove propositional partial recursor is \textit{the best},
-the above "reflect" function implies that propositional partial
-recursor is a good \textit{extensional characterization} of (non-indexed) extensible inductive type (with certain constraints)---because we can ``embed'' vanilla inductive types into
-types supporting a partial recursor: "reflect" is actually a left
-inverse of an injection "bool → self_B.b". A bit more formally, and
-restricting our focus to non-indexed inductive types:
-
-\begin{theorem}\label{thm:prec-complete} Given a list of $n$ pairs of types $\{ x : "A"_i \vdash "B"_i(x) \}_{i}$ s.t.
-
-  \begin{itemize}
-    \item \textlabel{(Detectable Partiality)}{prop:detectable-partiality} for each pair $x : A_i \vdash B_i(x)$,
-    \begin{minted}{Coq}
-      Axiom detectable:
-      forall {T} {a} (f: Bᵢ a -> option T),
-        {forall x, f x <> None} + {exists x, f x = None}.
-    \end{minted} 
-    \item Define $"C"$ as the inductive type using this list, i.e.,
-    \begin{minted}{Coq}
-Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
-    \end{minted}
-    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
-    \begin{itemize}
-      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
-      \item a partial recursor
-      \begin{minted}[escapeinside=@@]{Coq}
-prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint. Let me fix other parts first}@
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), .., 
-      D -> option R
-      \end{minted}
-      \item and $n$ (propositional) computational axioms: for all $i$, 
-      \begin{minted}{Coq}
-prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
-    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
-      \end{minted}
-    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
-    \end{itemize}
-  \end{itemize}
-  Then there exists an \textbf{embedding} from "C" to "D".
-  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of
-  "C" and "prec" of "D" (both acting like identity) such that
-  \begin{minted}[escapeinside=@@]{Coq}
-    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
-If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
-I mentioned left inverse 'linv' is the generalization}@
-  \end{minted}
-\end{theorem}
-The reason we have this weird looking \ref{prop:detectable-partiality} is because the construction of "linv"---at one point, using "prec" to construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option C) → option C}, generally not provable, only if we have \mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
-This \ref{prop:detectable-partiality} is exactly transforming former to the later: recall that, according to the semantic of inductive type, \mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using "linv" on the substructure of "D".
-This function points out each of the translation result from "D" to "C" and some of them are "None" because of failing.
-For us, we only care about if \textbf{any} of the substructure of "D" is failed to translate---if any one substructure fails, then the whole transltion fails and thus we should have "None" as return.
-
-\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that includes all of the examples shown in this paper.
-\ref{prop:detectable-partiality} is not supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
-In those cases, we still have the fact that partial recursor is a conservative extension, but we cannot show that partial recursor is the extensional characterization any more。
-
-We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
-actually be considered as a constructor, because they can be reflected
-to real constructors of "C" using the left inverse "linv" (directly generalization of "reflect : self_B.b → option bool" of the above). Thus, \cref{thm:prec-complete}
-implies that (1) every future extension of the inductive type can
-support this partial recursion (trivially); (2) every type supporting
-this partial recursor with its computational axioms at least supports
-these constructors because of the embedding. In other words, \textit{a
-type (at least) supports these constructors if and only if this type
-supports the corresponding partial recursor (with the computational
-axioms)}.  Thus, we can argue that the partial recursor gives a sound and
-complete extensional characterization of all the extension of a given (non-indexed)
-inductive type.\YZ{What about indexed inductive types?}\EDJreply{Done.
-I also have a formalization of this proof now, for indexed type and dependent eliminator.
-Though the formalization is about single constructor.}
-
-We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
-\begin{minted}{Coq}
-  pdelim : ∀ (P : D -> Set), .., 
-  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
-    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
-\end{minted}
-To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable.
-Please refer to the supplementary material and appendix for a formalization of the left inverse in the context of indexed inductive type.
-
-\subsection{Limitation}
+\subsection{Engineering Limitation}
 We list some of the current limitation of our implementation. We believe some of these limitations can be remedied by more engineering efforts.

### 2022-08-29 coqimpl.tex
Reorganizing Coqimpl.tex

@@
 
-\textbf{Overridable Field and Closing Fact}
-It is natural to consider the ``automatic overridden'' resulting inheritance, as we are using the same tactic (or the same untyped Gallina term to provide the proof).
+\textbf{Escape Hatches: "Meta-data" and "Closing Fact".} There are cases the developers want to refer to global reasoning or just escape from the programming paradigm of family polymorphism.
+Thus we provide the following mechanism.
+
+\textit{"MetaData".} Due to our implementation of the family,
+the family data is invisible to Coq internal and thus causes some
+inconvenience. For example, in
+\cref{fig:plugin-example-global-reasoning-meta-data} the first attempt
+of defining tactic "inv" fails. Because the earlier defining "tm",
+"value", and "value_not_app" are currently just data structure in the
+plugin and not visible for Coq internal (only visible after the family
+is compiled into module).
+
+Thus to refer to them, we have to scope the defining tactic with
+"MetaData" and our plugin will set up the appropriate environment. After
+a "MetaData" block is closed, the plugin will aggregate the content into
+the family being defined so that they can also be used when defining
+the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}\EDJreply{Not sure.
+This auto-insertion will practically look like hijack the original Coq command.
+I am not sure this kind of hijacking can happen.}
+This can fulfill \ref{chg:software-engineering} and define new customized tactic.
+
+\textit{"Closing Fact".} We have described how to use partial recursor to "discriminate" different constructors of an extensible inductive type.
+Practically, "discriminate" is used when proving inversion lemma, like forementioned "var_not_value".
+
+% \textbf{Overridable Field and Closing Fact}
+% It is natural to consider the ``automatic overridden'' resulting inheritance, as we are using the same tactic (or the same untyped Gallina term to provide the proof).
 But for type theorists, these are definitely overriding because these are definitely different typed Gallina term!
 
-Instead of making a decision here, we introduce "Closing Fact":  "Closing Fact" will make a field tied with the untyped proof script and only allows inheritance.
+% Instead of making a decision here, we introduce "Closing Fact":  "Closing Fact" will make a field tied with the untyped proof script and only allows inheritance.
 Of course, during inheritance, the untyped proof script will always be re-run and introduce (potentially) \textbf{different} typed terms\footnote{Thus called "Closing Fact" as its semantic depends on the closed  surrounding family in a global sense}.
@@
 
-This is similar to "FRecursor"---we only allow inheritance for "FRecursor" as its semantic can be considered as a tie between a handler family and an extensible inductive type, where during each inheritance, the tie will be rechecked.
+% This is similar to "FRecursor"---we only allow inheritance for "FRecursor" as its semantic can be considered as a tie between a handler family and an extensible inductive type, where during each inheritance, the tie will be rechecked.
 
-Just like "Overridable ..
+% Just like "Overridable ..
 pins {..}", "Closing Fact" provides the ability to stab through all the abstraction of the former fields, and thus has the similar functionality  as "Overridable ..
@@
 
-\textbf{Hack: ``Meta-data''.} Due to our implementation of the family,
-the family data is invisible to Coq internal and thus causes some
-inconvenience. For example, in
-\cref{fig:plugin-example-global-reasoning-meta-data} the first attempt
-of defining tactic "inv" fails. Because the earlier defining "tm",
-"value", and "value_not_app" are currently just data structure in the
-plugin and not visible for Coq internal (only visible after the family
-is compiled into module).
-
-Thus to refer to them, we have to scope the defining tactic with
-"MetaData" and our plugin will set up the appropriate environment. After
-a "MetaData" block is closed, the plugin will aggregate the content into
-the family being defined so that they can also be used when defining
-the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}\EDJreply{Not sure.
-This auto-insertion will practically look like hijack the original Coq command.
-I am not sure this kind of hijacking can happen.}
-This can fulfill \ref{chg:software-engineering} and define new customized tactic.
-
 \subsection{Engineering Limitation}

### 2022-08-29 coq-example.tex
Add new mixin diagram.
Rewriting Coqimpl to include 1. Adding partial recursor 2. adding its generation for computation info 3. reorganize the whole implementation part

@@
+\newpage\EDJ{Please delete this newpage command later}
 We provide two examples here for illustration and evaluation of our Coq plugin.
@@
 \end{lstlisting}
-\includegraphics[width=0.5\textwidth]{coqexmaple/STLC-inheritances.pdf}
+\includegraphics[width=0.5\textwidth]{coqexmaple/Mixin-Venn-Diagram.pdf}
 \caption{Example STLC extended with product, and Mixin with \cref{fig:STLC-example2}}
@@
 , boolean (in \cref{fig:STLC-example}), sum type and iso-recursive type---with vanilla STLC.
-
-\textbf{Mixin} is implemented by making inheritance judgement/data supporting \textit{weakening} and composition.
-Then, using notation from metatheory, given $\goodInh{\cdot}{i_1}{\sigma_1}{\sigma_2}$ and $\goodInh{\cdot}{i_2}{\sigma_1}{\sigma_3}$, we can weaken $i_2$ to $\goodInh{\cdot}{i_2'}{\sigma_2}{\sigma_3'}$ and the composition of $i_1$ and $i_2'$ is the resulting mixin---this implementation breaks commutativity and will need to solve design problems on compatibility (mostly related to the \textit{diamond problem}): for example, when $i_1$ and $i_2$ are both extending with a field of the same name, do we signal error or use $i_2$'s extension and consider it as an `override'?
-However, our current example of STLC only uses a subset of the mixin with a clear semantic (i.e.
-every sensible design of mixin should agree on)---because our STLC examples only use extension of fields with different names and no overriding is used.
-
-Note that, every top level family can be considered as an inheritance because (1) our implementation is simplified by considering every family as an inheritance, and (2) we want to align with the complete family polymorphism advocated by \citet{zm2017} where the user can override the parent of the inheriting family (inside a family due to late-binding), and in these case the nested inheriting family is more a inheritance data than a ``standalone'' linkage.
-Thus we will store for every top level family in our plugin the corresponding inheritance data together with a ``reference'' to the parent family.
-Then our mixin command can single out the inheritance data from "STLC_bool" and "STLC_prod" and ``mixin'' the inheritance to reach "STLC_bool_prod".
 

### 2022-08-29 coqimpl.tex
Rewriting Coqimpl to include 1. Adding partial recursor 2. adding its generation for computation info 3. reorganize the whole implementation part

@@
 
+However, it seems verbose to prove this simple inversion lemma using "FRecursor", let alone the fact this "FRecursor" will need to be extended every time the inductive type (e.g.
+"value") is extended.
+Thus we provide a command "Closing Fact".
+"Closing Fact" will prove the given lemma in the context where all the definitions \textbf{are exposed}.
+Thus the above "var_not_value" can be simply achieved by
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Closing Fact var_not_value : 
+  forall i, ~ self__STLC.value (self__STLC.tm_var i) 
+    by { intros i H; inversion H; eauto }.
+\end{minted}
+We can use "inversion" tactic here thanks to the exposure of the vanilla inductive type in the context.
+
+We simply support inheritance for "Closing Fact".
+However, we fix the semantic of the "Closing Fact" by this one-line \textbf{untyped} proof script---and thus we restrict "Closing Fact" to be used in a proof-irrelevant way.
+
+\textit{"Overridable ..
+pins {..}" vs.
+"Closing Fact".} A keen reader will notice a great deal of similarity between "Overridable ..
+pins {..}" vs.
+"Closing Fact" as they both expose the concrete definition of the former fields (including extensible inductive type).
+In fact, using "Overridable" can prove this "var_not_value" in an almost identical way.
+The only difference seems to be the ``auto-inheritance'' of "Closing Fact" with the help of the proof script where even if the extensible inductive type is extended---but this difference seems superficial as we can implement something similar for "Overridable".
+
+The main difference is that, the semantic of a "Closing Fact" field is the \textbf{untyped} proof script but the semantic of a "Overridable" field is always a \textbf{typed} proof term.
+Thus when a pinned extensible inductive type/former field is extended/overridden, we can still \textbf{inherit} the "Closing Fact" but we can only \textbf{override} the overridable field.
+
+This difference might seem small and irrelevant to the user from the first sight but this becomes quite relevant when we introduce \textit{mixin} of two inheritances.
 % \textbf{Overridable Field and Closing Fact}
@@
 
-\textbf{Mixin}. The reason we introduce two different semantics for two similar functionality is due to the fact that, we want to introduce mixin in our facility.
+% \textbf{Mixin}.
+The reason we introduce two different semantics for two similar functionality for "Overridable ...
+pins {...}" and "Closing Fact" is due to the fact that, we want to introduce mixin in our facility.
 
-% TODO: mixin implementation sketch
+% \textit{Mixin} is implemented by making inheritance judgement/data supporting \textit{weakening} and composition.  
+\textbf{Mixin} is implemented by making inheritance judgement/data supporting \textit{weakening} and composition.More concretely,  for an inheritance from parent family type $\sigma_1$ to children family type ${\sigma_2}$ and another inheritance from ${\sigma_1}$ to ${\sigma_3}$, we can weaken(expand the input of) the second inheritance to ${\sigma_2}$ then the output children family type becomes ${\sigma_3'}$.
+Then the ``composition'' of the very first inheritance and the altered second inheritance is the resulting mixin.
 
-However, we don't allow any overriding happening in either mix-edin inheritance, because it will break the commutativity and complicate the semantic without much benefit.
+However, this implementation breaks commutativity: for example, when the two inheritances are overriding a given field in different ways, the order of composition becomes important.
+Thus to avoid this ambiguity, we simply \textit{don't allow overriding fields} inside any inheritance when we do mixin.
+In this case, a proof irrelevant "Closing Fact" is a useful alternative for overriding.
+
+Note that, every top level family can be considered as an inheritance because (1) our implementation is simplified by considering every family as an inheritance, and (2) we want to align with the complete family polymorphism advocated by \citet{zm2017} where the user can override the parent of the inheriting family (inside a family due to late-binding), and in these case the nested inheriting family is more a inheritance data than a ``standalone'' linkage.
+Thus we will store for every top level family in our plugin the corresponding inheritance data together with a ``reference'' to the parent family.
+
+% Mixin example:
+
+% Then our mixin command can single out the inheritance data from "STLC_bool" and "STLC_prod" and ``mixin'' the inheritance to reach "STLC_bool_prod".
+
+% However, we don't allow any overriding happening in either mix-edin inheritance, because it will break the commutativity and complicate the semantic without much benefit.
 Overriding is not allowed in this case, but "FRecursor" and "Closing Fact" is still allowed since they only allow to be inherited.
@@
 
-\subsection{Engineering Limitation}
+\subsection{Implementation Limitation}
 We list some of the current limitation of our implementation. We believe some of these limitations can be remedied by more engineering efforts.
 \begin{itemize}
-  \item We don't support the full power of inductive facility in Coq, that includes mutual inductive type, notation for inductive type and parameters in the inductive type (we only support indexed inductive type right now)
+  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, notation for inductive type and parameters in the inductive type (we only support indexed inductive type right now)
   \item We don't support mutual fixpoint definition
@@
   \item We don't consider the compatibility of universe polymorphism mechanism with our plugin
+  \item We only support automatic generation of partial recursor and the computation information for non-indexed inductive type 
 \end{itemize}

### 2022-09-02 lang-design.tex
Save.

@@
 \begin{align*}
-  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ident\rangle] 
+  V := \quad &"Family"~\langle ident\rangle~["extends"~\langle ref\rangle] 
   \quad |\quad "EndFamily" \quad | \quad "SealFamily" \\
-  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle~"using"~\langle ident\rangle \\
+  & |\quad "FRecursor"~\langle ident\rangle~"about"~\langle ref\rangle~"motive"~\langle term\rangle~"using"~\langle ref\rangle \\
+  & |\quad "FScheme CompEq about"~\langle ref \rangle~|\quad "FScheme " \langle ident\rangle " PRecT about"~\langle ref \rangle\\
   & | \quad "FTheorem"~\langle ident\rangle~"about"~\langle ident\rangle~"motive"~\langle term\rangle \\
-  & |\quad ["Extend" | "Override"]~"Family"~\langle ident\rangle \\ 
+  & |\quad ["Extend"]~"Family"~\langle ident\rangle \\ 
   & |\quad ["Extend"]~"FInductive"~\langle ident\rangle \\ 
-  & |\quad ["Final" | "Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
+  & |\quad ["Overridable " ["pins " \{ \langle ident\rangle ... \}]]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
+  & |\quad ["Override"]~"Field"~\langle ident\rangle~":"~\langle term \rangle~":="~\langle term \rangle \\
   & | \quad "FieldClaim"~\langle ident\rangle~":"~\langle term \rangle \\
@@
   & | \quad "Closing Fact"~\langle ident\rangle~":"~\langle term \rangle~"by"~"\{"\langle \text{one tactic expression}\rangle"\}" \\ 
+  & | \quad "Family"~\langle ident\rangle~"extends"~\langle ref\rangle~"using"~\langle ref\rangle~"using"~\langle ref\rangle \\ 
   & | \quad "MetaData"~\langle ident\rangle \quad | \quad "EndMetaData" \\ 
@@
 
+Unlike how Coq's "Inductive" will automatically generate the eliminator for the inductive type, we require the user to explicit use "FScheme ..
+PRecT" to generate a \textit{partial recursor} for the "FInductive".
+We will later explain what a partial recursor is, at this point the reader can just consider it as a mundane eliminator.
+
 To resolve \ref{chg:extensible-inductive-type}, our first step is to
@@
 
+% Explain FScheme
+However, "FRecursor" doesn't specify any reduction/computation information until a family is defined completely and closed (e.g.
+"subst" in \cref{fig:STLC-example2} cannot reduce when application during the definition of the surrounding family).
+To remedy this, we introduce "FScheme CompEq" that will generate proposition of equations indicating reductions.
+For example, for "subst", it will generate the equation
+\mintinline{Coq}{subst tm_unit = subst_internal.tm_unit}.
+
+Similarly, the generated partial recursor using "FScheme .. PRecT" doesn't have computation information either, and require "FScheme CompEq" to generate them.
+
+% Explain Mixin
+
+\textbf{Mixin.} For two inheritance with the same parent family, we support mixin the two using "Family ...
+extends ...
+using ...
+using ...".
+However, the candidates for mixin cannot have any overriding term (i.e.
+only allow inheritance and extension).
+
 % We can also use "Family" and "SealFamily" to scope a (nested) family.
@@
 
-We will explain the usage of "Closing Fact" and "MetaData" later---they are here to resolve \ref{chg:software-engineering} to some extent.
+We will explain the usage of "Closing Fact" and "MetaData" later---they are \textit{escape-hatches} to resolve \ref{chg:software-engineering} to some extent.
 % This incomplete support can already be useful for extending functions.

### 2022-09-02 coqimpl.tex
Save.
FScheme command description added. Wait for reviewing.

@@
 Then when compute the context during the compilation of the term for that field, we make sure only all the dependee and the non-overridable field will expose its definition in the context.
+
 Most parts of the implementation is quite similar to the default fields, except that when inheriting an overridable field, we need to make sure all its dependee is inherited earlier.
@@
 
-\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
+\textbf{``Extensible'' Inductive Type and Defining Recursor/Eliminator.}
 We support extensible inductive types in Coq mentioned in
@@
 
-\textit{Generating Scheme for Propositional Computation Axiom for FRecursor}.
-We support "FScheme" that our plugin will automatically generate(postulate) the computational ``axiom'' for the specified  into the family.
-During compilation, these ``axiom'' can be easily proved/instantiated with concrete/vanilla inductive type.
+% \textit{Generating Scheme for Propositional Computation Axiom for FRecursor}.
 
-However, this method, though remedies problem (1) to some extent, is only introducing computational information \textbf{propositionally} but not judgementally.
+Thus we support "FScheme" to remedy problem (1): our plugin can automatically generate(postulate) the computational ``axiom'' for the specified  into the family and prove them during compilation into modules.
+However, it is only introducing computational information \textbf{propositionally} but not judgementally, and we might need the help of rewriting tactic on these computational ``axiom'' to justify the computation.
+
+% However, this method, though remedies problem (1) to some extent, is only introducing computational information \textbf{propositionally} but not judgementally.
 Thus, for example, during theorem proving, we might need the help of rewriting tactic on these computational ``axiom'' to justify the computation.
@@
 We notice that, the power of "FRecursor" comes from its distinction from the mundane fields and its ability to stab through the abstraction of "FInductive".
-However, the latter is quite similar on how "Overridable ..
+Interestingly, the latter is quite similar on how "Overridable ..
 pins {..}" can see through the abstraction of former pinned overrdiable fields.
-This inspires us to generalize pins---we support pinning inductive type and "FRecursor".
-Once "FInductive" and "FRecursor" fields are pinned, we can stab through its abstraction and have the primitive coq-providing recursor and \textbf{judgemental equality} for the computational information for "FRecursor".
+This inspires us to generalize "pins"---we support pinning inductive type and "FRecursor".
+Once "FInductive" and "FRecursor" fields are pinned, we are directly exposed with the primitive coq-providing recursor and \textbf{judgemental equality} of the computational information for "FRecursor".
 
-The compilation mechanism is simply make the underlying, compiled definition of inductive type and recursor included inside the parameters of the module, just like how the the concrete definition of overridable fields are included inside the parameters when pinned.
+The compilation mechanism is simply make the underlying, compiled definition of inductive type and recursor included as part of the parameters of the module, just like how the the concrete definition of overridable fields are included inside the parameters when pinned.
 This way, the computational information from vanilla Coq can be transimitted into the context.
@@
 % Following needs a rewrite
-To achieve "injection" in problem (2), we can easily generate/postulate $n$ propositions for $n$ constructors, just like how we did for computational information.
-But for "discriminate" tactic, this way will generate $O(n^2)$ propositions.
+To achieve "injection" in problem (2), we can easily generate/postulate $n$ ``injection'' propositions for $n$ constructors, just like how we did for computational information.
+But for "discriminate" tactic, this way will generate $O(n^2)$ propositions in a verbose way.
 
@@
 For example, given a new inductive type, to prove "discriminate" between two constructors, i.e., "c$_i$ ..
-$\neq$ c$_j$ .." for two different constructors, we simply use the eliminator to map each constructor to different natural number, assuming we can "discriminate" different natural number.\footnote{Rigorously, to avoid any suspicious circular reasoning, we can use large elimination to map them to $\top$ and $\bot$, and because $\exists a : ?, \top$ holds for the former not the later, thus $\top \neq \bot$.} "injection" can be supported in a similar way---we can simply let eliminator map each inductive type to the arguments of the constructors.
+$\neq$ c$_j$ .." for two different constructors, we simply use the eliminator to map each constructor to different natural number, assuming we can "discriminate" different natural number.\footnote{Rigorously, to avoid any suspicious circular reasoning, we can use large elimination to map them to $\top$ and $\bot$, and because $\exists a : \_, \top$ holds for the former not the later, thus $\top \neq \bot$.} "injection" can be supported in a similar way---we can simply let eliminator map each inductive type to the arguments of the constructors.
 
-Thus it is natural for us to think if (an altered form of) eliminator can be postulated into the context, since apparently we cannot directly postulate the original eliminator here---because the sole existence of the eliminator prevents any possible future extension of the inductive type.
+Thus it is natural to think if (an altered form of) eliminator can be postulated into our context, since apparently we cannot directly postulate the original eliminator here---because the sole existence of the eliminator prevents any possible future extension of the inductive type.
 
 Thus we alter eliminators into \text{partial recursor}, such that (1) it will not hinder the future extension, and (2) it is still powerful enough to derive injectiveness like tactic "inject" and "discriminate".
-What's more, (3) we can actually prove this \text{partial recursor} is actually good enough because it can act as an \textit{extensional characterization} for extensible inductive type.
-We will call it \textit{propositional} because we only prpositionally postulate its computational informations in Coq.s
+What's more, (3) we can actually prove this \text{partial recursor} is actually good enough because it can act as an \textit{extensional characterization} for extensible inductive type \textit{to some extent}.
+We will call it \textit{propositional} because we only prpositionally postulate its computational informations for Coq.
 
-\textbf{Propositional Partial Recursor.} Partial recursor is simply adding partially into the signature of the eliminator, by wrapping each motive with "option". 
+\textbf{Propositional Partial Recursor.} Partial recursor is simply adding partiality into the signature of the eliminator, by wrapping each motive with "option". 
 % This section is introducing the concept of \textit{partial recursor}.
@@
 \ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that includes all of the examples shown in this paper.
-\ref{prop:detectable-partiality} is not supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
-In those cases, we still have the fact that partial recursor is a conservative extension, but we cannot show that partial recursor is the extensional characterization any more。
+
+\ref{prop:detectable-partiality} is not generally supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
+However, in those cases, we still have partial recursor provable and useful---we just cannot see the partial recursor as a strong toolkit anymore.
 
@@
 For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)}\footnote{Achieved by three steps: first proving partial recursor using "FRecursor", then use partial recursor to prove "tm_var i $\neq$ tm_lam v body" and finally using "FRecursor" prove "var_not_value" by induction on "value"} in the "Family STLC" either using "FRecursor", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
-But partial recursor will not cause this problem.
+But the postulation of partial recursor will not cause this problem.
 
@@
 I am not sure this kind of hijacking can happen.}
-This can fulfill \ref{chg:software-engineering} and define new customized tactic.
+
+This can not only fulfill \ref{chg:software-engineering} and define new customized tactic but also integrate any vanilla Coq's feature into family.
+For example, if the user don't want to use "FInductive" but only the non-extensible "Inductive" type, the user can simply declare the inductive type inside "MetaData" to be part of the family.
 
@@
 % having boilerplate code, \textbf{semantically}.
-% \begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}\EDJreply{Done}
-% \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-% Family STLC.
-%   FInductive tm : Set := ... 
-%   FInductive value : tm -> Prop := ... 
-%   Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
-%       by { intros x y H; inversion H; eauto }.
-%   Fail Ltac inv := 
-%     match goal with 
-%     | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
-%     (* Fail: value, tm_app, value_not_app unfound *)
-%     end. 
-%   MetaData tactic1.
-%     Ltac inv := (* ... the same definition as above ... *)
-%   EndMetaData.
-%   (* ... STLC example continue ... *)
-% EndFamily.
-% \end{lstlisting}  
-% \caption{Example for Global Reasoning and MetaData, using in STLC example}\label{fig:plugin-example-global-reasoning-meta-data}
-% \end{figure}
+\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}\EDJreply{Done}
+\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+Family STLC.
+  FInductive tm : Set := ... 
+  FInductive value : tm -> Prop := ... 
+  Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
+      by { intros x y H; inversion H; eauto }.
+  Fail Ltac inv := 
+    match goal with 
+    | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
+    (* Fail: value, tm_app, value_not_app unfound *)
+    end. 
+  MetaData tactic1.
+    Ltac inv := (* ... the same definition as above ... *)
+  EndMetaData.
+  (* ... STLC example continue ... *)
+EndFamily.
+\end{lstlisting}  
+\caption{Example for Global Reasoning and MetaData, using in STLC example}\label{fig:plugin-example-global-reasoning-meta-data}
+\end{figure}
 

### 2022-09-02 lang-design.tex
FScheme command description added. Wait for reviewing.

@@
 
-% Explain Mixin
-
-\textbf{Mixin.} For two inheritance with the same parent family, we support mixin the two using "Family ...
-extends ...
-using ...
-using ...".
-However, the candidates for mixin cannot have any overriding term (i.e.
-only allow inheritance and extension).
-
 % We can also use "Family" and "SealFamily" to scope a (nested) family.
@@
 % definition and add a new pattern matching case.
+
+% Explain Mixin
+
+Finally, \textbf{Mixin.} For two children family with the same parent family, we support mixin the two using "Family ...
+extends ...
+using ...
+using ...".
+(In this case, it is better to consider the children family itself as \textit{an inheritance object}.) However, the candidates for mixin cannot have any overriding term (i.e.
+only allow inheritance and extension).

### 2022-09-04 introduction.tex
reply
intro

@@
 the programmer faces a fundamental tension \cite{reynolds1975} between
-adding new constructors of a data type (e.g., new abstract syntax) and
+adding new constructors to a data type (e.g., new abstract syntax) and
 adding new functions over the data type (e.g., new compiler passes).
@@
 address the EP.
-Yet the challenge of writing extensible, type-safe code (and proofs!) is
+Yet, the challenge of writing extensible, type-safe code is
 as real, especially for metatheory mechanization.
-As an example, consider \cref{fig:STLC-example}, the simply typed
-lambda calculus (STLC) mechanized in Coq.
-A programming challenge there is to define distinct extensions of this
-STLC formalization to support distinct features,
-and then selectively compose these extensions to form new STLC variants.
+The programmer faces a tension between adding new constructors to an inductive data type
+(e.g., new abstract syntax) and adding new functions and theorems over
+the data type (e.g., new meta\-theoretical results).
 
-In Coq, inductive types, as well as functions and theorems over
-inductive types, are closed to extension.
+In the Coq proof assistant~\cite{coq}, inductive types, as well as functions and
+theorems over inductive types, are closed to extension.
 So to reuse mechanized metatheories,
-the common practice is still to copy and modify them for each extension.
+the common practice is still to copy code and proofs and then modify them in each extension.
 But having to maintain multiple copies is highly non-modular and
@@
 The programmer could also turn to design patterns~\cite{delaware2011,delaware2013}.
-But they require heavy lifting from the programmer to make code
+But they tend to require heavy lifting from the programmer to make code
 extensible, often leading to verbose, non-idiomatic programming styles.
 
-A question thus arises: can linguistic solutions to the EP, as studied
-in the conventional setting, be adapted to metatheory mechanization in
-a proof assistant?
+We answer the following question:
+can linguistic features designed to solve the EP (in
+conventional OO and functional languages)
+be adapted to metatheory mechanization in a proof assistant?
 
@@
 But the language-theoretic essence of inheritance is more general:
-it is a linguistic approach to incrementally modifying
-\emph{recursive, record-like definitions}~\cite{cook1990inheritance}.
+it is a linguistic approach to \emph{incrementally modifying
+recursive definitions}~\cite{cook1990inheritance}.
 
@@
 extensible cases~\cite{bac2006},
-and so on, are all based on the essential idea of inheritance.
+and so on, are all based on this essential idea of inheritance.
 %
@@
 
-\paragraph{Main ideas.}
+\begingroup
+
+\paragraph{Contributions}
 
@@
 Because code and proofs are polymorphic to a family they are nested
-within, they can be inherited and reused by a derived family.
-%
-Family polymorphism enables extensible metatheory mechanization.
-As \cref{fig:?} shows,
-different extensions to STLC can all inherit from the base STLC family:
-they reuse mechanized metatheories from syntax to the type safety
+within, they can be
+
+\begin{wrapfigure}[6]{r}{.215\textwidth}
+\includegraphics[scale=.48]{graphics/stlc-intro.pdf}
+\end{wrapfigure}
+
+\noindent
+inherited and reused by a derived family—%
+family polymorphism enables extensible metatheory mechanization.
+
+%As an example, consider \cref{fig:STLC-example}, the simply typed
+%lambda calculus (STLC) mechanized in Coq.
+%A programming challenge there is to define distinct extensions of this
+%STLC formalization to support distinct features,
+%and then selectively compose these extensions to form new STLC variants.
+As an example, the diagram to the right depicts an extensible
+mechanization of the simply typed λ-calculus (STLC), using family
+polymorphism.
+Different extensions to STLC can all inherit from the base STLC family:
+they reuse mechanized metatheories, from syntax to the type safety
 theorem, only adding new constructors to inductive types
-and new cases to the induction proofs as needed by an extension.
+and new cases to the induction proofs and recursive functions as needed
+by an extension.
 
@@
 
-\paragraph{Contributions} To our knowledge, we offer the first
-language-based solution that enables extensible metatheory mechanization
-in a higher-order, dependent type theory with inductive types.
+\endgroup
+
+Specifically, we make the following contributions.
 
@@
 
-\item We show how to integrate family polymorphism into a proof
-assistant (\cref{sec:lang-design}).
+\item We present the first language design that enables extensible
+metatheory mechanization in a higher-order, dependent type theory with
+inductive types (\cref{sec:lang-design}).
 The language design reconciles the expressiveness enabled by
@@
 
-\item We capture the new language mechanism formally by extending
+\item We capture the essence of the new language mechanism formally by extending
 Martin–Löf type theory with family polymorphism and extensible inductive
 types (\cref{sec:metatheory}).
-We derive strong meta\-theoretical results including on consistency and
+We derive strong meta\-theoretical results including consistency and
 canonicity.

### 2022-09-14 abstract.tex
abstract

@@
 With the growing practice of mechanizing language metatheories,
-it has become ever more important that proof assistants
+it has become ever more pressing that interactive theorem provers 
 make it easy to write reusable, extensible code and proofs.
-We ...
+%
+This paper presents a language design useful for extensible metatheory
+mechanization in a proof assistant.
+The new design achieves reuse and extensibility via a form of family
+polymorphism, a seemingly object-oriented idea, that allows code and
+proofs to be polymorphic to their enclosing families.
+Our development addresses significant technical challenges that arise
+from the underlying language being simultaneously functional,
+dependently typed, a logic, and an interactive tool.
+%
+Our results include
+\begin{enumerate*}
+\item a prototypical implementation of the language design as a Coq plugin,
+\item a formalization of the language design as a constructive type theory with inductive types,
+\item proofs of strong guarantees afforded by the type theory,
+and 
+\item case studies showing how the new expressiveness naturally addresses real
+programming challenges in metatheory mechanization.
+\end{enumerate*}

### 2022-09-15 challenge.tex
intro
Challenge#1

@@
-\subsection{Challenges}
+Supporting inheritance and family polymorphism in a proof assistant poses challenges
+far beyond those encountered in an object-oriented setting~\cite{ncm2004,vc-calculus-2006,zm2017},
+because the underlying programming language is simultaneously
+functional, dependently typed, a logic, and an interactive tool.
+
+\noindentparagraph{Extensible inductive types vs.\ exhaustive induction reasoning.}
+
+Inductive types, generalizing algebraic data types found in
+functional languages, are a central feature of any proof assistant being
+used for mechanizing language metatheories.
+They offer a means to define abstract syntax, typing rules, reduction
+rules, and so on.
+Unfortunately, inductive types are closed to extension.
+
+A family-polymorphism mechanism could potentially support extensible
+inductive types, by allowing a \emph{derived family} to add new constructors
+to inductive types inherited from a \emph{base family}.
+Such capability would be useful for extending mechanized languages.
+For example, \cref{fig:?} shows the part of an STLC extension that could
+potentially be mechanized using extensible inductive types: a derived family
+inherits constructors of inductive types from the base family (viz.\ the $\cdots$'s)
+and adds new constructors to model iso-recursive types.
+
+\begin{figure}
+\fontsize{8.5}{9}\selectfont
+\def\MathparLineskip{\lineskip=0.50ex}
+
+\fcolorbox{black}{black!10}{
+\begin{minipage}{0.38\textwidth}
+
+\noindent%
+Types \quad
+$ \tau \ \Coloneqq\ \mathbb{1} \bnf  \tau_1 \to \tau_2 $
+\medskip
+
+Terms \quad
+$ e \ \Coloneqq\ x \bnf () \bnf \lambda x.\,e \bnf e_1~e_2 $
+\medskip
+
+Typing rules
+\begin{mathpar}
+\Rule{ \Gamma(x)=\tau }{ \Gamma \vdash x : \tau }
+
+\Rule{ \Gamma, x:\tau_1 \vdash e : \tau_2 }{ \Gamma \vdash \lambda x.\,e : \tau_1 \to \tau_2 }
+
+\Rule{}{ \Gamma \vdash () : \mathbb{1} }
+\quad
+\Rule{
+  \Gamma \vdash e_1 : \tau_1 \to \tau_2
+  \quad
+  \Gamma \vdash e_2 : \tau_1
+}{
+  \Gamma \vdash e_1~e_2 : \tau_2
+}
+
+\end{mathpar}
+
+Values
+\begin{mathpar}
+\Rule{}{\textit{Val}(())}
+
+\Rule{}{\textit{Val}(\lambda x.\,e)}
+\end{mathpar}
+
+Reduction rules
+\begin{mathpar}
+\Rule{\textit{Val}(e_2)}{ (\lambda x.\,e_1)~e_2 \longrightarrow \subst{e_2}{e_1}{x} }
+
+\Rule{ e_1 \longrightarrow e_1' }{ e_1~e_2 \longrightarrow e_1'~e_2 }
+
+\Rule{ \textit{Val}(e_1) \quad e_2 \longrightarrow e_2' }{ e_1~e_2 \longrightarrow e_1~e_2' }
+
+\end{mathpar}
+
+\end{minipage}
+}%
+%\tikz \draw [arrows = {-Latex[width'=0pt .5, length=10pt, fill=white]}] (0,0) -- (1,0);
+\tikz \draw [
+  line width = .5pt,
+  arrows = {-Stealth[inset=0pt, length=5pt, fill=white, angle'=60, scale=2]}
+] (1,0) -- (0,0);%
+%https://tikz.dev/tikz-arrows
+\fcolorbox{black}{black!10}{
+\begin{minipage}{0.49\textwidth}
+
+\noindent%
+Types \quad $ \tau \ \Coloneqq\ \cdots \bnf \alpha \bnf \mu\alpha.\,\tau$
+\medskip
+
+Terms \quad $ e \ \Coloneqq\ \cdots \bnf \mathrm{fold}\;e \bnf \mathrm{unfold}\;e $
+\medskip
+
+Typing rules
+\begin{mathpar}
+\cdots
+\quad
+\Rule{
+  \Gamma \vdash e : \subst{\tau}{\mu\alpha.\,\tau}{\alpha}
+}{
+  \Gamma \vdash \mathrm{fold}\;e : \mu\alpha.\,\tau
+}
+\quad
+\Rule{
+  \Gamma \vdash e : \mu\alpha.\,\tau
+}{
+  \Gamma \vdash \mathrm{unfold}\;e : \subst{\tau}{\mu\alpha.\,\tau}{\alpha}
+}
+
+\end{mathpar}
+
+Values
+\begin{mathpar}
+\cdots
+
+\Rule{ \textit{Val}(e) }{ \textit{Val}(\mathrm{fold}\;e) }
+
+\end{mathpar}
+
+Reduction rules
+\begin{mathpar}
+\cdots
+
+\Rule{ e \longrightarrow e' }{ \mathrm{unfold}\;e \longrightarrow \mathrm{unfold}\;e' }
+
+\Rule{ e \longrightarrow e' }{ \mathrm{fold}\;e \longrightarrow \mathrm{fold}\;e' }
+
+\Rule{ \textit{Val}(e) }{ \mathrm{unfold}\;(\mathrm{fold}\;e) \longrightarrow e }
+
+\end{mathpar}
+
+\end{minipage}
+}
+
+\caption{%
+  Statics and dynamics of STLC (left) and its extension with iso-recursive types (right).
+}
+\label{fig:?}
+\end{figure}
+
+However, there is a tension between extensibility of inductive types
+and exhaustivity of induction reasoning.
+The substitution function and the substitution lemma are examples that
+require induction reasoning.
+In the derived family, to ensure that induction remains exhaustive,
+how could the programmer account for the new cases arising from the new
+constructors, without redefining or forcing the rechecking of those cases
+already handled by the base family?
+
+The tension results from the duality between variants and records.
+Existing family-polymorphism mechanisms are designed for record-like
+language constructs like classes and objects; inductive types, by
+contrast, are variant-like.
+
+\noindentparagraph{Late binding vs.\ dependent typing.}
+
+\noindentparagraph{Essence of inheritance vs.\ consistency of a logic.}
+
+\noindentparagraph{Family vs.\ interactivity.}
+
+\newpage
 

### 2022-09-15 coq-example.tex
Challenge#1

@@
 And at the end, the compiled module for "LangTyping" and "LangCP" is put into tests over simple queries. These "LangTyping" and "LangCP" also illustrates that we still have
-computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}\EDJreply{Done.
-Please check}
+computability in the presence of family.\YZ{Mention the alternative approach of using parameterized modules.}

### 2022-09-15 introduction.tex
Challenge#1

@@
 %
-In particular, when a language mechanism enables inheritance over
+In particular, when a language mechanism allows inheritance over
 recursive definitions containing related types and terms,
@@
 types and terms are polymorphic to a family they are nested within.
-
-\begingroup
 
@@
 within, they can be
+inherited and reused by a derived family.
+Hence, family polymorphism allows for extensible metatheory mechanization.
 
-\begin{wrapfigure}[6]{r}{.215\textwidth}
+\begingroup
+
+\begin{wrapfigure}[6]{r}{.200\textwidth}
 \includegraphics[scale=.48]{graphics/stlc-intro.pdf}
@@
 
-\noindent
-inherited and reused by a derived family—%
-family polymorphism enables extensible metatheory mechanization.
-
+%\noindent
 %As an example, consider \cref{fig:STLC-example}, the simply typed
@@
 polymorphism.
-Different extensions to STLC can all inherit from the base STLC family:
-they reuse mechanized metatheories, from syntax to the type safety
-theorem, only adding new constructors to inductive types
-and new cases to the induction proofs and recursive functions as needed
-by an extension.
+An extension of STLC with products and another with iso-recursive types
+can both inherit from the base STLC family:
+they reuse mechanized metatheories,
+from abstract syntax all the way to the type-safety theorem,
+only adding new constructors to inductive types
+and adding new cases to recursive functions and induction proofs
+\emph{as needed} by an extension.
 
@@
 \item We present case studies of using our Coq plugin to mechanize
-language metatheories.
+language metatheories (\cref{sec:coqexample}).
 They show how our language design naturally solves the EP and

### 2022-09-15 challenge.tex
reword

@@
 The tension results from the duality between variants and records.
-Existing family-polymorphism mechanisms are designed for record-like
-language constructs like classes and objects; inductive types, by
-contrast, are variant-like.
+For record-like language constructs (e.g., classes and objects), it is
+safe for an extension to add more fields.
+By contrast, variant-like language constructs (e.g., inductive types)
+do not automatically enjoy safe, modular addition of constructors.
 

### 2022-09-15 abstract.tex
isorec -> fix

@@
 Our development addresses significant technical challenges that arise
-from the underlying language being simultaneously functional,
-dependently typed, a logic, and an interactive tool.
+from the underlying language of a proof assistant being simultaneously
+functional, dependently typed, a logic, and an interactive tool.
 %

### 2022-09-15 challenge.tex
isorec -> fix

--- 2022-07-13 challenge.tex
@@
-\subsection{Challenges}
+Supporting inheritance and family polymorphism in a proof assistant poses challenges
+far beyond those encountered in an object-oriented setting~\cite{ncm2004,vc-calculus-2006,zm2017},
+because the underlying programming language is simultaneously
+functional, dependently typed, a logic, and an interactive tool.
+
+\noindentparagraph{Extensible inductive types vs.\ exhaustive induction reasoning.}
+
+Inductive types, generalizing algebraic data types found in
+functional languages, are a central feature of any proof assistant being
+used for mechanizing language metatheories.
+They offer a means to define abstract syntax, typing rules, reduction
+rules, and so on.
+Unfortunately, inductive types are closed to extension.
+
+A family-polymorphism mechanism could potentially support extensible
+inductive types, by allowing a \emph{derived family} to add new constructors
+to inductive types inherited from a \emph{base family}.
+Such capability would be useful for extending mechanized languages.
+For example, \cref{fig:?} shows the part of an STLC extension that could
+potentially be mechanized using such extensible inductive types: a derived family
+inherits constructors of inductive types from the base family (viz.\ the $\cdots$'s)
+and adds new constructors to model the new language feature.
+
+\input{stlc-nonmechanized.tex}
+
+However, there is a tension between extensibility of inductive types
+and exhaustivity of induction reasoning.
+In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
+the substitution function, require induction reasoning.
+In the derived family, to ensure that induction remains exhaustive,
+how could the programmer account for the new cases arising from the new
+constructors, without redefining or forcing the rechecking of those cases
+already handled by the base family?
+
+The tension results from the duality between variants and records.
+For record-like language constructs (e.g., classes and objects), it is
+safe for an extension to add more fields.
+By contrast, variant-like language constructs (e.g., inductive types)
+do not automatically enjoy safe, modular addition of constructors.
+
+\noindentparagraph{Late binding vs.\ dependent typing.}
+
+\noindentparagraph{Essence of inheritance vs.\ consistency of a logic.}
+
+\noindentparagraph{Family vs.\ interactivity.}
+
+\newpage
 

### 2022-09-26 challenge.tex
Challenge#2

@@
-Supporting inheritance and family polymorphism in a proof assistant poses challenges
-far beyond those encountered in an object-oriented setting~\cite{ncm2004,vc-calculus-2006,zm2017},
+Integrating family polymorphism into an interactive theorem prover poses challenges
+far beyond those found in an object-oriented setting~\cite{ncm2004,vc-calculus-2006,zm2017},
 because the underlying programming language is simultaneously
@@
 Such capability would be useful for extending mechanized languages.
-For example, \cref{fig:?} shows the part of an STLC extension that could
+For example, \cref{fig:stlc-nonmechanized} shows the part of an STLC extension that could
 potentially be mechanized using such extensible inductive types: a derived family
-inherits constructors of inductive types from the base family (viz.\ the $\cdots$'s)
+inherits constructors of inductive types from the base family (viz.\ the ellipsis under ``Typing rule'')
 and adds new constructors to model the new language feature.
@@
 the substitution function, require induction reasoning.
-In the derived family, to ensure that induction remains exhaustive,
-how could the programmer account for the new cases arising from the new
-constructors, without redefining or forcing the rechecking of those cases
-already handled by the base family?
+A language design must enforce that induction remains exhaustive in the
+derived family despite the new constructors, without requiring
+redefinition or rechecking of those cases already handled by the base
+family.
+%
+The tension reflects a duality between variants and records.
+For record-like language constructs (e.g., classes and objects), it is
+safe for an extension to add more fields: projection of existent fields
+still works.
+But variant-like constructs (e.g., inductive types) do not automatically
+enjoy safe, modular addition of constructors:
+pattern matching could potentially become non-exhaustive.
 
-The tension results from the duality between variants and records.
-For record-like language constructs (e.g., classes and objects), it is
-safe for an extension to add more fields.
-By contrast, variant-like language constructs (e.g., inductive types)
-do not automatically enjoy safe, modular addition of constructors.
+\noindentparagraph{Overriding vs.\ dependent typing.}
 
-\noindentparagraph{Late binding vs.\ dependent typing.}
+Common to many family-polymorphism mechanisms is the ability for a
+derived family to override fields defined in the base family.
+As we show later, this expressiveness is also useful for metatheory
+mechanization.
+However, it is in conflict with dependent typing, wherein type checking
+often involves unfolding definitions.
+In \cref{fig:stlc-nonmechanized}, the proof of the substitution lemma is
+checked against the definition of the substitution function.
+If a derived family were allowed to arbitrarily override how
+substitution is defined, then the proof, when inherited, would be
+ill-formed.
+A language design should only allow those overridings that are safe.
 
-\noindentparagraph{Essence of inheritance vs.\ consistency of a logic.}
+\noindentparagraph{Self reference vs.\ logical consistency.}
 

### 2022-09-26 introduction.tex
Challenge#2

@@
 
-In the Coq proof assistant~\cite{coq}, inductive types, as well as functions and
+In the Coq proof assistant~\cite{coq}, for instance, inductive types, as well as functions and
 theorems over inductive types, are closed to extension.
@@
 %
-The programmer could also turn to design patterns~\cite{delaware2011,delaware2013}.
+The programmer could also turn to design patterns~\cite{delaware2011,delaware2013} that
+use software product lines or Church encodings.
 But they tend to require heavy lifting from the programmer to make code
-extensible, often leading to verbose, non-idiomatic programming styles.
+extensible, often leading to non-idiomatic programming styles.
 
-We answer the following question:
-can linguistic features designed to solve the EP (in
-conventional OO and functional languages)
-be adapted to metatheory mechanization in a proof assistant?
+We thus set out to answer the following question:
+can extensible metatheory mechanization be made easier by
+having a proof assistant support new \emph{linguistic features} that address the EP?
+%can linguistic features designed to solve the EP for designed to solve the EP (in
+%conventional OO and functional languages)
+%be adapted to metatheory mechanization in a proof assistant?
 
@@
 virtual classes~\cite{virtualclasses-1989,vc-calculus-2006},
-virtual types~\cite{thorup97} and associated types~\cite{ckj05},
-extensible cases~\cite{bac2006},
-and so on, are all based on this essential idea of inheritance.
+virtual types~\cite{thorup97}, % and associated types~\cite{ckj05},
+extensible cases~\cite{bac2006}, etc.\ 
+are all based on this essential idea of inheritance.
 %
@@
 polymorphism.
-An extension of STLC with products and another with iso-recursive types
+An extension of STLC with products and another with fixpoints
 can both inherit from the base STLC family:

### 2022-09-27 challenge.tex
Challenge#3

@@
 redefinition or rechecking of those cases already handled by the base
-family.
-%
+family.\footnote{%
 The tension reflects a duality between variants and records.
 For record-like language constructs (e.g., classes and objects), it is
-safe for an extension to add more fields: projection of existent fields
-still works.
+safe for an extension to add more fields: existing fields can still be
+projected.
 But variant-like constructs (e.g., inductive types) do not automatically
 enjoy safe, modular addition of constructors:
-pattern matching could potentially become non-exhaustive.
+pattern matching could potentially become non-exhaustive.}
 
@@
 often involves unfolding definitions.
-In \cref{fig:stlc-nonmechanized}, the proof of the substitution lemma is
-checked against the definition of the substitution function.
-If a derived family were allowed to arbitrarily override how
-substitution is defined, then the proof, when inherited, would be
-ill-formed.
+For example, in an STLC mechanization, the substitution lemma is checked
+against the definition of the substitution function.
+If a derived family were allowed to arbitrarily override the
+substitution function, then the proof of the substitution lemma, when
+inherited, would be ill-formed (even in the absence of any inductive-type extensions).
 A language design should only allow those overridings that are safe.
@@
 \noindentparagraph{Self reference vs.\ logical consistency.}
+
+Inheritance and family polymorphism are mechanisms for
+incrementally modifying self-referential definitions~\cite{cook1990inheritance}:
+the implicit self parameterization is how they achieve late binding and thus extensibility.
+However, self reference could easily lead to divergence.
+Divergence is not a concern for the design of ordinary OO or
+functional languages, but it means logical inconsistency---and
+hence unsoundness---for a language aimed at logical reasoning.
+A language design must ensure that the added extensibility does not
+go rogue.
 
@@
 overridability. We need to be careful about this tension when designing
-the language.\YZ{
-  The two examples seem to suggest there should not be any overriding allowed at all.
-  Is there an example where you actually want the ability to override?
-}\EDJreply{Added a new example. Please check.}
+the language.
 
@@
 that may later be computed and normalized. However, families themselves
-can be second class citizens just like \citet{zm2017}.\YZ{I was thinking
-if the paper should start by talking about the functional encoding of
-inheritance (namely exposing a self parameter to allow late binding, and
-later taking the fixpoint of the extensible module when it needs to be
-accessed outside the family).  I think most readers will benefit from a
-recap of this encoding.\\
-Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
-}\EDJreply{Feel free to add these. I don't have many knowledge about this part.}
+can be second class citizens just like \citet{zm2017}.
+%\YZ{I was thinking
+%if the paper should start by talking about the functional encoding of
+%inheritance (namely exposing a self parameter to allow late binding, and
+%later taking the fixpoint of the extensible module when it needs to be
+%accessed outside the family).  I think most readers will benefit from a
+%recap of this encoding.\\
+%Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
+%}\EDJreply{Feel free to add these. I don't have many knowledge about this part.}
 

### 2022-09-27 challenge.tex
Challenge#4

@@
 
-\noindentparagraph{Family vs.\ interactivity.}
+\noindentparagraph{User experience.}
+
+Interactive theorem proving and tactic programming are
+part of the idiomatic programming experience in a proof assistant.
+A language design integrating family polymorphism should be compatible
+with these programming styles.
+For instance, it should be possible in our tool to
+incrementally navigate through vernacular commands and construct proofs
+using tactics with instant feedback on the proof state,
+even in the middle of a family definition.
 

### 2022-09-27 challenge.tex
tooling

@@
 
-\noindentparagraph{User experience.}
+\noindentparagraph{Tooling and user experience.}
 
@@
 with these programming styles.
-For instance, it should be possible in our tool to
-incrementally navigate through vernacular commands and construct proofs
+In particular, it should be possible in our tool to
+incrementally navigate through vernacular commands and, moreover, construct proofs
 using tactics with instant feedback on the proof state,

### 2022-09-29 challenge.tex
small edits

@@
-Integrating family polymorphism into an interactive theorem prover poses challenges
+Integrating family polymorphism into a proof assistant presents challenges
 far beyond those found in an object-oriented setting~\cite{ncm2004,vc-calculus-2006,zm2017},
@@
 
-\noindentparagraph{Extensible inductive types vs.\ exhaustive induction reasoning.}
+\noindentparagraph{%
+\headinglabel{C1}{chg:extensible-v-exhaustive}.\,
+Extensible inductive types vs.\ exhaustive induction reasoning.%
+}
 
 Inductive types, generalizing algebraic data types found in
-functional languages, are a central feature of any proof assistant being
-used for mechanizing language metatheories.
-They offer a means to define abstract syntax, typing rules, reduction
-rules, and so on.
-Unfortunately, inductive types are closed to extension.
+functional languages, are a central feature of any proof assistant
+in use for mechanizing language metatheories.
+They offer a means to define abstract syntax and inference rules.
+But unfortunately, inductive types are closed to extension by design.
 
-A family-polymorphism mechanism could potentially support extensible
+A family-polymorphism design could potentially support extensible
 inductive types, by allowing a \emph{derived family} to add new constructors
 to inductive types inherited from a \emph{base family}.
-Such capability would be useful for extending mechanized languages.
+Such a feature would be useful for extending mechanized languages.
 For example, \cref{fig:stlc-nonmechanized} shows the part of an STLC extension that could
 potentially be mechanized using such extensible inductive types: a derived family
-inherits constructors of inductive types from the base family (viz.\ the ellipsis under ``Typing rule'')
+inherits constructors of inductive types from the base family (e.g., the ellipsis under ``Typing rules'')
 and adds new constructors to model the new language feature.
@@
 However, there is a tension between extensibility of inductive types
-and exhaustivity of induction reasoning.
-In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
-the substitution function, require induction reasoning.
-A language design must enforce that induction remains exhaustive in the
-derived family despite the new constructors, without requiring
-redefinition or rechecking of those cases already handled by the base
-family.\footnote{%
+and exhaustivity of induction reasoning.\footnote{%
 The tension reflects a duality between variants and records.
@@
 enjoy safe, modular addition of constructors:
-pattern matching could potentially become non-exhaustive.}
+existing pattern matches could potentially become non-exhaustive.}
+In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
+the substitution function, require induction reasoning.
+A language design must enforce that induction remains exhaustive in the
+face of the new constructors in the derived family.
+For modularity, it should do so without requiring
+redefinition or rechecking of those cases already handled by the base
+family.
 
-\noindentparagraph{Overriding vs.\ dependent typing.}
+\noindentparagraph{%
+\headinglabel{C2}{chg:override-v-dependent}.\,
+Overriding vs.\ dependent typing.%
+}
 
@@
 
-\noindentparagraph{Self reference vs.\ logical consistency.}
+\noindentparagraph{%
+\headinglabel{C3}{chg:self-v-consistency}.\,
+Self reference vs.\ logical consistency.%
+}
 
@@
 incrementally modifying self-referential definitions~\cite{cook1990inheritance}:
-the implicit self parameterization is how they achieve late binding and thus extensibility.
+the implicit self parameterization is how they achieve late binding and
+thus offer extensibility.
 However, self reference could easily lead to divergence.
 Divergence is not a concern for the design of ordinary OO or
-functional languages, but it means logical inconsistency---and
+functional languages, but it would mean logical inconsistency---and
 hence unsoundness---for a language aimed at logical reasoning.
-A language design must ensure that the added extensibility does not
-go rogue.
+A family-polymorphism design must tame self reference to guarantee consistency.
 
-\noindentparagraph{Tooling and user experience.}
+\noindentparagraph{%
+\headinglabel{C4}{chg:tooling}.\,
+User experience and tooling.%
+}
 
@@
 even in the middle of a family definition.
+In addition, it should remain possible to perform computation on Gallina terms.
+
+\ifShowOldWriting
 
@@
 % programming} to relieve us from the hassle of direct manipulation of proof terms.
+
+\newpage
+
+\fi

### 2022-09-29 coqimpl.tex
small edits

@@
 the (vanilla) inductive type in that compiled module. Thus a "FRecursor"
-field is unlike other fields, even though it is exposed as a mundane
-function for the following fields in the programming interface.\YZ{
+field is unlike other fields, even though it is exposed as a plain
+function for the succeeding fields in the programming interface.\YZ{
   Why is this special treatment needed for compiling FRecursor?
@@
 the exposing data in the family type (the compiled module type) for the
-following field. In other words, we must be careful on what interface
-(context) the following fields are defined based on.
+succeeding field. In other words, we must be careful on the interface
+(context) based on which the succeeding fields are defined.
 
@@
 
-The compilation mechanism is simply make the underlying, compiled definition of inductive type and recursor included as part of the parameters of the module, just like how the the concrete definition of overridable fields are included inside the parameters when pinned.
+The compilation mechanism is simply make the underlying, compiled definition of inductive type and recursor included as part of the parameters of the module, just like how the concrete definition of overridable fields are included inside the parameters when pinned.
 This way, the computational information from vanilla Coq can be transimitted into the context.

### 2022-09-29 lang-design.tex
small edits

@@
+We demonstrate key ingredients of our design as a plugin to the Coq proof assistant,
+though we believe our design could be adapted to other proof assistants such as Lean.
+
+\ref{chg:extensible-v-exhaustive}
+
+\input{stlc-mechanized.tex}
+
+\newpage
+
 Inspired by the above challenges and Family Polymorphism, we propose
@@
 \end{align*}
-
-\begin{figure}[!htb]
-  \begin{minipage}[t]{0.47\linewidth}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family STLC.
- FInductive ty : Set :=
-  | unit : ty
-  | arrow : ty -> ty -> ty.
- FInductive term : Set := 
-  | tm_var : id -> term 
-  | tm_abs : id -> term -> term ...
- Family subst_internal.
-  Final Field tm_var := ..
-  Final Field tm_abs := ..
-   ...
- EndFamily.
- FRecursor subst
-  about tm motive ..
-  using subst_internal.
- ...
- FInductive has_type 
-  : context -> term -> type := ...
- FTheorem subst_lemma 
-  about has_type 
-  motive ...
- Proof.  ...
- Qed.
- FInductive step : term -> term -> Prop 
-  := ...
-(* ... and more, end with type safety *)
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-  \begin{minipage}[t]{0.47\linewidth}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family STLC_bool extends STLC.
- Extend FInductive ty : Set :=
-  | bool : ty.
-
- Extend FInductive term : Set := 
-  | tt : term | ff : term 
-  | tif : term -> term -> term -> term
- Extend Family subst_internal.
-  Final Field tt := ..
-  Final Field ff := ..
-  Final Field tif := ..
- EndFamily. (* Automatic inheriting others 
-              from STLC.subst_internal *)
- Inherit subst.
-(* Plugin will redo exhaustiveness check, in case 
- subst_internal is not extended correctly *)
- Extend FInductive has_type (* .. and new ty too *)
-  : context -> term -> type := ...
- Extend FTheorem subst_lemma. Proof.
-  ... (* Need to prove extra cases *)
- Qed.
-
- Extend FInductive step : term -> term -> Prop 
-  := ... (* Need to expand this binary relation *)
-(* ... and more extension *)
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-  \caption{Example STLC and its extension, in designed syntax}\label{fig:STLC-example2}
-\end{figure}
 
@@
 theorem proving---the handler and the family will be automatically
-constructed during proving.\YZ{Cool. Can the same be done to FRecursor?}\EDJreply{I think last time I have answered it. FRecursor for fixpoint command}
+constructed during proving.
 

### 2022-09-29 stlc-mechanized.tex
small edits

\begin{figure}
  \begin{minipage}[t]{0.47\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC.
 FInductive ty : Set :=
  | unit : ty
  | arrow : ty -> ty -> ty.
 FInductive term : Set := 
  | tm_var : id -> term 
  | tm_abs : id -> term -> term ...
 Family subst_internal.
  Final Field tm_var := ..
  Final Field tm_abs := ..
   ...
 EndFamily.
 FRecursor subst
  about tm motive ..
  using subst_internal.
 ...
 FInductive has_type 
  : context -> term -> type := ...
 FTheorem subst_lemma 
  about has_type 
  motive ...
 Proof.  ...
 Qed.
 FInductive step : term -> term -> Prop 
  := ...
(* ... and more, end with type safety *)
EndFamily.
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{0.47\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC_bool extends STLC.
 Extend FInductive ty : Set :=
  | bool : ty.

 Extend FInductive term : Set := 
  | tt : term | ff : term 
  | tif : term -> term -> term -> term
 Extend Family subst_internal.
  Final Field tt := ..
  Final Field ff := ..
  Final Field tif := ..
 EndFamily. (* Automatic inheriting others 
              from STLC.subst_internal *)
 Inherit subst.
(* Plugin will redo exhaustiveness check, in case 
 subst_internal is not extended correctly *)
 Extend FInductive has_type (* .. and new ty too *)
  : context -> term -> type := ...
 Extend FTheorem subst_lemma. Proof.
  ... (* Need to prove extra cases *)
 Qed.

 Extend FInductive step : term -> term -> Prop 
  := ... (* Need to expand this binary relation *)
(* ... and more extension *)
EndFamily.
\end{lstlisting}
  \end{minipage}
  \caption{Example STLC and its extension, in designed syntax}\label{fig:STLC-example2}
\end{figure}

### 2022-09-29 coqimpl.tex
a question

@@
 where "tm_rec" and "tm" are the corresponding recursor and the (vanilla
-but extended and modified) inductive type in that compiled module. Note
+but extended and modified) inductive type in that compiled module.\YZ{Where is this "tm" inductive type in figure 6?} Note
 that, even though "tm_rec" and "tm" has the same name as before, they

### 2022-09-29 coqimpl.tex
took a pass on coqimpl

@@
 field is unlike other fields, even though it is exposed as a plain
-function for the succeeding fields in the programming interface.\YZ{
-  Why is this special treatment needed for compiling FRecursor?
-}\EDJreply{because the non-special treatment cannot be applied to recursor --
-  the non-special treatment will require the recursor to be *exported* at the interface}
+function for the succeeding fields in the programming interface.
 
@@
 % explain ftheorem wrapping this complicated recursor
-This design of decoupling of the case handlers and exhaustiveness checking handles the semantic well, however, it brings a lot of boilerplate code---we need to create an extra (handler) family, manually specifying the type of each case handlers, and then using "FRecursor" to ``tie the knot'' and construct a recursor field from the handlers.
+This design of decoupling of the case handlers and exhaustiveness
+checking handles the semantic well, however, it brings a lot of
+boilerplate code---we need to create an extra (handler) family, manually
+specifying the type of each case handlers, and then using "FRecursor" to
+``tie the knot'' and construct a recursor field from the handlers.  
 We thus provide a command "FTheorem" to avoid these boilerplate in programming interface when using tactic programming. 
 This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
-proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor".\YZ{example?}\EDJreply{FTheorem is really just a short hand for "FRecursor" + handler family.
-Nothing more conceptual here.
-If the reader is interested in the concrete syntax, I think I need to make the two examples avaliable in the appendix.} We expect this "FTheorem" to be used in theorem proving but not general programming.
+proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor". We expect this "FTheorem" to be used in theorem proving but not general programming.
 
@@
 meta-theory need to handle these two seemingly contradicting ideas
-simultaneously when solving \ref{chg:extensible-inductive-type} .
-(2) \ref{chg:extensible-inductive-type} is also resolved by this \textit{decoupling} of the syntax of implementation of recursor handlers and exhaustiveness checking.
-The former is handled by family inheritance and thus avoiding the boilerplate code; and the non-inheritance of the compiled recursor ensures the exhaustiveness checking happens for every recursor.
+simultaneously when solving \ref{chg:extensible-inductive-type}.
+(2) \ref{chg:extensible-inductive-type} is also resolved by this
+\textit{decoupling} of the syntax of implementation of recursor handlers
+and exhaustiveness checking. The former is handled by family inheritance
+and thus avoiding the boilerplate code; and the non-inheritance of the
+compiled recursor ensures the exhaustiveness checking happens for every
+recursor.
 
@@
 
-However, "FRecursor" provides barely enough functionality for manipulating our(extensible) inductive type, and it is far from satisfactory to be a legitimate programming interface.
-(1) We do not yet have any computational information using this "FRecursor" until its compilation into module, i.e.
-we cannot do computation with it during the construction of the family using the current compilation strategy.
-(2) The necessary tactic like "discriminate" and "injection" is not applicable on (extensible) inductive type.
+\paragraph{Tactics and propositional recursors.}
+
+However, "FRecursor" provides barely enough functionality for
+manipulating our(extensible) inductive type, and it is far from
+satisfactory to be a legitimate programming interface. (1) We do not yet
+have any computational information using this "FRecursor" until its
+compilation into module, i.e. we cannot do computation with it during
+the construction of the family using the current compilation strategy.
+(2) The necessary tactic like "discriminate" and "injection" is not
+applicable on (extensible) inductive type.
 
@@
 
-\textit{Overridable pins Inductive Type and FRecursor}.
-We notice that, the power of "FRecursor" comes from its distinction from the mundane fields and its ability to stab through the abstraction of "FInductive".
-Interestingly, the latter is quite similar on how "Overridable ..
-pins {..}" can see through the abstraction of former pinned overrdiable fields.
-This inspires us to generalize "pins"---we support pinning inductive type and "FRecursor".
-Once "FInductive" and "FRecursor" fields are pinned, we are directly exposed with the primitive coq-providing recursor and \textbf{judgemental equality} of the computational information for "FRecursor".
-
-The compilation mechanism is simply make the underlying, compiled definition of inductive type and recursor included as part of the parameters of the module, just like how the concrete definition of overridable fields are included inside the parameters when pinned.
-This way, the computational information from vanilla Coq can be transimitted into the context.
+\textit{Overridable pins Inductive Type and FRecursor}. We notice that,
+the power of "FRecursor" comes from its distinction from the mundane
+fields and its ability to stab through the abstraction of "FInductive".
+Interestingly, the latter is quite similar to how "Overridable .. pins
+{..}" can see through the abstraction of former pinned overrdiable
+fields. This inspires us to generalize "pins"---we support pinning
+inductive type and "FRecursor". Once "FInductive" and "FRecursor" fields
+are pinned, we are directly exposed with the primitive coq-providing
+recursor and \textbf{judgemental equality} of the computational
+information for "FRecursor".
+
+The compilation mechanism is to simply make the underlying, compiled
+definition of inductive type and recursor included as part of the
+parameters of the module, just like how the concrete definition of
+overridable fields are included inside the parameters when pinned. This
+way, the computational information from vanilla Coq can be transmitted
+into the context.
 
@@
 % Following needs a rewrite
-To achieve "injection" in problem (2), we can easily generate/postulate $n$ ``injection'' propositions for $n$ constructors, just like how we did for computational information.
-But for "discriminate" tactic, this way will generate $O(n^2)$ propositions in a verbose way.
-
-What's more, to manipulate inductive type, it is not necessarily enough to only support "injection" and "discriminate" tactic (and their corresponding axioms).
-Recall in vanilla inductive type, we believe that eliminator is the strong-enough ``toolkit'' because the eliminator (with its computational information) is the \textit{universal property} of the inductive type---it extensionally describes the inductive type up to (certain) isomorphism.
-This universal property is much more powerful than "injection" and "discriminate" and maybe the most powerful tool we can have to reason about inductive type.
+To achieve "injection" in problem (2), we can easily generate/postulate
+$n$ ``injection'' propositions for $n$ constructors, just like how we
+did for computational information. But for "discriminate" tactic, this
+way will generate $O(n^2)$ propositions in a verbose way.
+
+What's more, to manipulate inductive type, it is not necessarily enough
+to only support "injection" and "discriminate" tactic (and their
+corresponding axioms). Recall in vanilla inductive type, we believe that
+eliminator is the strong-enough ``toolkit'' because the eliminator (with
+its computational information) is the \textit{universal property} of the
+inductive type---it extensionally describes the inductive type up to
+(certain) isomorphism. This universal property is much more powerful
+than "injection" and "discriminate" and maybe the most powerful tool we
+can have to reason about inductive type.
 
 For example, given a new inductive type, to prove "discriminate" between two constructors, i.e., "c$_i$ ..
-$\neq$ c$_j$ .." for two different constructors, we simply use the eliminator to map each constructor to different natural number, assuming we can "discriminate" different natural number.\footnote{Rigorously, to avoid any suspicious circular reasoning, we can use large elimination to map them to $\top$ and $\bot$, and because $\exists a : \_, \top$ holds for the former not the later, thus $\top \neq \bot$.} "injection" can be supported in a similar way---we can simply let eliminator map each inductive type to the arguments of the constructors.
-
-Thus it is natural to think if (an altered form of) eliminator can be postulated into our context, since apparently we cannot directly postulate the original eliminator here---because the sole existence of the eliminator prevents any possible future extension of the inductive type.
-
-Thus we alter eliminators into \text{partial recursor}, such that (1) it will not hinder the future extension, and (2) it is still powerful enough to derive injectiveness like tactic "inject" and "discriminate".
-What's more, (3) we can actually prove this \text{partial recursor} is actually good enough because it can act as an \textit{extensional characterization} for extensible inductive type \textit{to some extent}.
-We will call it \textit{propositional} because we only prpositionally postulate its computational informations for Coq.
+$\neq$ c$_j$ .." for two different constructors, we simply use the eliminator to map each constructor to a different natural number, assuming we can "discriminate" different natural numbers.\footnote{Rigorously, to avoid any suspicious circular reasoning, we can use large elimination to map them to $\top$ and $\bot$, and because $\exists a : \_, \top$ holds for the former not the latter, thus $\top \neq \bot$.} "injection" can be supported in a similar way---we can simply let eliminator map each inductive type to the arguments of the constructors.
+
+Thus it is natural to think if (an altered form of) eliminator can be
+postulated into our context, since apparently we cannot directly
+postulate the original eliminator here---because the sole existence of
+the eliminator prevents any possible future extension of the inductive
+type.
+
+Thus we alter eliminators into \text{partial recursor}, such that (1) it
+will not hinder the future extension, and (2) it is still powerful
+enough to derive injectivity ("inject") and disjointness ("discriminate") of constructors.
+What's more, (3) we can prove this \text{partial recursor} is
+actually good enough because it can act as an \textit{extensional
+characterization} for extensible inductive types \textit{to some extent}.
+We will call it \textit{propositional} because we only propositionally
+postulate its computational information for Coq.
 
@@
 % This raise an interesting question: what propositions are conservative extension? 
-% For example, apparently we know injectiveness of constructor, (i.e.
+% For example, apparently we know injectivity of constructor, (i.e.
 \mintinline{Coq}{STLC.tt ≠ STLC.ff}) must be conservative because it holds for all extension of the inductive type "STLC.tm".
@@
 
-% Since injectiveness of the constructors are derived by recursor in vanilla inductive type, an educated guess would be related to recursor.
+% Since injectivity of the constructors are derived by recursor in vanilla inductive type, an educated guess would be related to recursor.
 But we know the original recursor won't work because the mere existence of the original recursor will break extensibility.
@@
 This reduction can only be semi-automatically done by using [rewrite tactic and the computational axiom].}
-For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}\EDJreply{Added below please check.}
+For example, the propositional partial recursor for \mintinline{Coq}{B.b} is
 
@@
 FInductive b : Set := tt : b | ff : b.
-b_prec : forall R, option R -> option R -> b -> option R.@\YZ{Not clear what it takes to define/prove partial recursors.
-The reader needs to be told if they require breaking abstraction like recursors do.}\EDJreply{What do you mean?
-recursor doesn't break abstraction but break extensibility.
-Below mentioned partial recursor doesn't break extensibility}@
-\end{minted}
-with appropriate computational axioms.
-This partial recursor differs from original recursor only by decorating an "option" with the return type thus possible to be automatically generated (and the case is similar for those computational axiom).
-Thanks to the partiality, partial recursor doesn't break
-extensibility (i.e. all future extensions of "b" can support this
+b_prec : forall R, option R -> option R -> b -> option R.
+\end{minted}
+with appropriate computational axioms.  This partial
+recursor differs from original recursor only by decorating the return
+type with an "option" and can thus be automatically generated (and
+the case is similar for those computational axiom). Thanks to the
+partiality, partial recursor doesn't break
+extensibility (i.e., all future extensions of "b" can support this
 "b_prec") because for example, "b_prec" can just return "None" when bump into extended constructor in the future.
 
-The second thing to check is that, if partial recursor is theoretically enough for injectiveness of the constructor. The rough idea to achieve this
+The second thing to check is that, if partial recursor is theoretically enough for injectivity of the constructor. The rough idea to achieve this
 is to use "b_prec" to reflect our special inductive type into vanilla
@@
 
-Thirdly, even though we cannot prove partial recursor is \textit{the best} reasoning tool for extensible inductive type,
+Thirdly, even though we cannot prove partial recursors are \textit{the
+best} reasoning tool for extensible inductive types,
 the above "reflect" function implies that propositional partial
-recursor is a good \textit{extensional characterization} of (non-indexed) extensible inductive type (under certain constraints)---because we can ``embed'' vanilla inductive types into
+recursors are a good \textit{extensional characterization} of
+(non-indexed) extensible inductive type (under certain
+constraints)---because we can ``embed'' vanilla inductive types into
 types supporting a partial recursor: "reflect" is actually a left
@@
       \begin{minted}[escapeinside=@@]{Coq}
-prec : forall (R : Set), ..,  @\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint. Let me fix other parts first}@
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), .., 
+prec : forall (R : Set), ..,
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,@\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint.
+Let me fix other parts first}\YZreply{Is this constraint Detectable Partiality?}@
       D -> option R
@@
   More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of
-  "C" and "prec" of "D" (both acting like identity) such that
+  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of\/
+  "C" and "prec" of\/ "D" (both acting like identity) such that
   \begin{minted}[escapeinside=@@]{Coq}
@@
 \end{theorem}
-The reason we have this weird looking \ref{prop:detectable-partiality} is because the construction of "linv"---at one point, using "prec" to construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option C) → option C}, generally not provable, only if we have \mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
-This \ref{prop:detectable-partiality} is exactly transforming former to the later: recall that, according to the semantic of inductive type, \mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using "linv" on the substructure of "D".
-This function points out each of the translation result from "D" to "C" and some of them are "None" because of failing.
-For us, we only care about if \textbf{any} of the substructure of "D" is failed to translate---if any one substructure fails, then the whole transltion fails and thus we should have "None" as return.
+The reason we have this weird looking \ref{prop:detectable-partiality}
+is that the construction of "linv"---at one point, using "prec" to
+construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
+C) → option C}, generally not provable, unless we have
+\mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.\YZ{Don't you really mean
+"unless we have the Detectable Partiality assumption instead"?}
+This \ref{prop:detectable-partiality} is exactly transforming former to
+the latter:
+recall that, according to the semantic of inductive type,
+\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
+"linv" on the substructure of "D". This function points out each of the
+translation result from "D" to "C" and some of them are "None" because
+of failing. For us, we only care about if \textbf{any} of the
+substructure of "D" fails to translate---if any substructure fails,
+then the whole translation fails and thus we should have "None" as
+return.
 

### 2022-10-01 body.tex
Challenge#1
Challenge#4
pWIP: redo §3 stlc figure

--- 2022-06-26 body.tex
@@
 
+% lstlisting coq style (inspired from a file of Assia Mahboubi)
+% https://tex.stackexchange.com/questions/434523/coq-syntax-highlighting
 \section{Introduction}
@@
 
-\section{Language Design}
+\section{Design Requirements and Challenges}
 % what feature we want
@@
 % Before/After Comparison of examples (with/without Fampoly)
-\label{sec:lang-design}
-\input{lang-design.tex}
+\label{sec:background+challenge}
+\input{challenge.tex}
 
-% Macro for all use
-
+% Macro for all use, the macro for metatheory syntax
 \newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}
@@
 
+\section{Language Design}
+% we need to make this section before Metatheory
+% because a lot of our formulation in the metatheory
+%   is inspired by the implementation
+\label{sec:lang-design}
+\input{lang-design.tex}
+
 \section{Prototypical Coq Implementation}
@@
 
-\section{Conclusion}
+\section{Limitations and Conclusion}
 \label{sec:conclusion}
+
+It is hard to write modular, extensible code and proofs.
+We have presented the first language design that
+
+... make exciting directions for future work.
 

### 2022-10-01 challenge.tex
took a pass on coqimpl
pWIP: redo §3 stlc figure

@@
 functional languages, but it would mean logical inconsistency---and
-hence unsoundness---for a language aimed at logical reasoning.
+hence unsoundness---for a language aimed at logical reasoning!
 A family-polymorphism design must tame self reference to guarantee consistency.
@@
 incrementally navigate through vernacular commands and, moreover, construct proofs
-using tactics with instant feedback on the proof state,
+with common tactics while getting instant feedback on the proof state,
 even in the middle of a family definition.
-In addition, it should remain possible to perform computation on Gallina terms.
+In addition, it should remain possible to perform computation in the extended Gallina language.
 

### 2022-10-01 lang-design.tex
pWIP: redo §3 stlc figure

@@
-We demonstrate key ingredients of our design as a plugin to the Coq proof assistant,
+We demonstrate the key ingredients of our design as a plugin to the Coq proof assistant,
 though we believe our design could be adapted to other proof assistants such as Lean.
@@
 "Field" is used to define new members in a family.
-"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:consistency}.
+"FieldClaim" is similar as "Field" but will invoke proof-mode in Coq and thus the user can use tactic programming to define new members.\footnote{We expect "FieldClaim" to be used in a \textit{proof-irrelevant} situation and thus whether overridable or not doesn't matter.} Both "Field" and "FieldClaim" are only allowed to refer to the former fields, to avoid possible inconsistency mentioned in \ref{chg:self-v-consistency}.
 Nested families are also allowed.
@@
 
-To resolve \ref{chg:consistency}, we make sure when inheriting or
+To resolve \ref{chg:self-v-consistency}, we make sure when inheriting or
 overriding on "f" happens, the context of the field "f" is also
 ``inherited''. Thus the mentioned informal example in
-\ref{chg:consistency} will not apply here, because the overriding "f" is
+\ref{chg:self-v-consistency} will not apply here, because the overriding "f" is
 in a context where "self" is empty thus "self.g" doesn't
@@
 % \textbf{Final as non-overridable.} Every field is by default
-% overridable, reminding us of \ref{chg:definition-relevant-reasoning}.  To
+% overridable, reminding us of \ref{chg:override-v-dependent}.  To
 % resolve it, we provide the keyword "Final" to allow the user to specify
@@
 % our \lstinline{a'} is not overridable in "Family I2", mediating the
-% \ref{chg:definition-relevant-reasoning}.
+% \ref{chg:override-v-dependent}.
 
@@
 \textbf{Overridable Fields with dependency.}  Each field is by default non-overridable.
-By the discussion of \ref{chg:definition-relevant-reasoning}, we can simply reason and access the definition about the former fields in a later field, just like Field "a" and "b" in \cref{fig:plugin-example3} and we lose the ability to override.
+By the discussion of \ref{chg:override-v-dependent}, we can simply reason and access the definition about the former fields in a later field, just like Field "a" and "b" in \cref{fig:plugin-example3} and we lose the ability to override.
 
@@
 Each overridable field will only expose its type instead of the concrete definition to the following default (non-overridable) fields.
-For example, the first failure in \cref{fig:plugin-example3} per the discussion of \ref{chg:definition-relevant-reasoning}: since overridability will break the inheritability of "c" and thus we don't allow such field to exist in the first place.
+For example, the first failure in \cref{fig:plugin-example3} per the discussion of \ref{chg:override-v-dependent}: since overridability will break the inheritability of "c" and thus we don't allow such field to exist in the first place.
 Our plugin will only expose the fact that "a'" is an arbitrary natural number but hiding its concrete definition, and thus "eq_refl" will fail to type-check.
@@
 
-To resolve \ref{chg:extensible-inductive-type}, our first step is to
+To resolve \ref{chg:extensible-v-exhaustive}, our first step is to
 \textit{decouple} the definition of recursor handlers and the checking
@@
 The resulting family only support direct inheritance or "Override" (i.e.
-we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:definition-relevant-reasoning}.
+we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:override-v-dependent}.
 What's more, "Override" family can only apply to sealed family as well (not allowed on normal family).
@@
 % nested families, inductive types, and fields.
-"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:definition-relevant-reasoning}.
+"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:override-v-dependent}.
 The fields inside a family
@@
 
-We will explain the usage of "Closing Fact" and "MetaData" later---they are \textit{escape-hatches} to resolve \ref{chg:software-engineering} to some extent.
+We will explain the usage of "Closing Fact" and "MetaData" later---they are \textit{escape-hatches} to resolve \ref{chg:tooling} to some extent.
 % This incomplete support can already be useful for extending functions.

### 2022-10-01 stlc-mechanized.tex
pWIP: redo §3 stlc figure

@@
 \begin{figure}
+\begin{minipage}{\textwidth}
+\begin{multicols}{2}
+
+\begin{lstlisting}
+Family STLC.
+FInductive ty : Set :=[@\hfill\codecomment{Types}@]
+| ty_unit  : ty
+| ty_arrow : ty -> ty -> ty.
+
+FInductive tm : Set :=[@\hfill\codecomment{Terms}@]
+| tm_var  : id -> tm
+| tm_unit : tm
+| tm_abs  : id -> tm -> tm
+| tm_app  : tm -> tm -> tm.
+
+Field env : Type := id -> option ty.
+Field empty : env := \lambda _, None.
+
+FInductive ofty : env -> tm -> ty -> Prop :=
+| ot_var  : [@\dadada@][@\hfill\codecomment{Typing rules}@]
+| ot_unit : [@\dadada@]
+| ot_abs  : [@\dadada@]
+| ot_app  : [@\dadada@].
+
+FInductive value : tm -> Prop :=[@\hfill\codecomment{Value forms}@]
+| v_unit : value tm_unit
+| v_abs  : \forall x e, value (tm_abs x e).
+
+FRecursion subst on tm[@\hfill\codecomment{Subst function}@] 
+  motive \lambda(_ : tm), id -> tm -> tm.
+Case tm_var := \lambda y x e,
+  if (eqb x y) then e else (tm_var y)
+Case tm_unit := \lambda x e, tm_unit
+Case tm_abs := \lambda y e' IHe' x e, [@\dadada@]
+Case tm_app := \lambda e1 IHe1 e2 IHe2 x e,
+  tm_app (IHe1 x e) (IHe2 x e).
+End subst.
+
+FInductive step : tm -> tm -> Prop :=
+| st_app1 : [@\dadada@][@\hfill\codecomment{Reduction rules}@]
+| st_app2 : [@\dadada@]
+| st_beta : \forall x e v, value v ->
+  step (tm_app (tm_abs x e) v) (subst e x v).
+
+FInduction subst_lemma on ofty[@\hfill\codecomment{Subst lemma}@]
+  motive \lambda G' e T (_ : ofty G' e T),
+  \forall G x e' T', G' = extend G x T' ->
+  ofty empty e' T' -> ofty G (subst e x e') T.
+Case ot_var.  [@\dadada@] Qed.
+Case ot_unit. [@\dadada@] Qed.
+Case ot_abs.  [@\dadada@] Qed.
+Case ot_app.  [@\dadada@] Qed.
+End subst_lemma.
+
+End STLC.
+
+Family STLC_fix extends STLC.
+
+End STLC_fix.
+
+\end{lstlisting}
+
+\end{multicols}
+\end{minipage}
+
+\vspace{2em}
+
   \begin{minipage}[t]{0.47\linewidth}

### 2022-10-02 stlc-mechanized.tex
stlc figure

--- 2022-09-29 stlc-mechanized.tex
@@
+\begin{figure}
+\begin{minipage}{\textwidth}
+\begin{multicols}{2}
+
+\begin{lstlisting}[numbers=left]
+Family STLC.
+
+FInductive ty : Set :=[@\hfill\codecomment{Types}@]
+| ty_unit  : ty
+| ty_arrow : ty -> ty -> ty.
+
+FInductive tm : Set :=[@\hfill\codecomment{Terms}@]
+| tm_var  : id -> tm
+| tm_unit : tm
+| tm_abs  : id -> tm -> tm
+| tm_app  : tm -> tm -> tm.
+
+Field env : Type := id -> option ty.
+Field empty : env := \lambda _, None.
+
+FInductive hasty : env -> tm -> ty -> Prop :=
+| ht_var  : [@\dadada@]  | ht_unit : [@\dadada@][@\hfill\codecomment{Typing rules}@]
+| ht_abs  : [@\dadada@]  | ht_app  : [@\dadada@].
+
+FInductive value : tm -> Prop :=[@\hfill\codecomment{Value forms}@]
+| v_unit : value tm_unit
+| v_abs  : \forall x e, value (tm_abs x e).
+
+FRecursion subst on tm[@\hfill\codecomment{Substitution function}@] 
+motive \lambda(_ : tm), id -> tm -> tm.
+Case tm_var := \lambda y x e,
+  if (eqb x y) then e else (tm_var y).
+Case tm_unit := \lambda x e, tm_unit.
+Case tm_abs := \lambda y e' IHe' x e, [@\dadada@].
+Case tm_app := \lambda e1 IHe1 e2 IHe2 x e,
+  tm_app (IHe1 x e) (IHe2 x e).
+End subst.
+
+FInductive step : tm -> tm -> Prop :=
+| st_app1 : [@\dadada@][@\hfill\codecomment{Reduction rules}@]
+| st_app2 : [@\dadada@]
+| st_beta : \forall x e v, value v ->
+  step (tm_app (tm_abs x e) v) (subst e x v).
+
+Field steps := clos_refl_trans step.
+
+FInduction weakening on hasty[@\hfill\codecomment{Weaken.\ lemma}@]
+motive \lambda G e T (_ : hasty G e T),
+  \forall G', includedin G G' -> hasty G' e T.
+Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app.  [@\dadada@] Qed.
+End weakening.
+
+FInduction substlemma on hasty[@\hfill\codecomment{Subst.\ lemma}@]
+motive \lambda G' e T (_ : hasty G' e T),
+  \forall G x e' T', G' = extend G x T' ->
+  hasty empty e' T' -> hasty G (subst e x e') T.
+Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app.  [@\dadada@] Qed.
+End subst_lemma.
+
+FInduction preserve on hasty[@\hfill\codecomment{Preserv.\ theorem}@]
+motive \lambda G e T (_ : hasty G e T),
+  G = empty -> \forall e', step e e' -> hasty empty e' T.
+Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app.  [@\dadada@] Qed.
+End preserve.
+
+FInduction progress on hasty[@\hfill\codecomment{Progress theorem}@]
+motive \lambda G e T (_ : hasty G e T),
+  G = empty -> value e \/ \exists e', step e e'.
+Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app.  [@\dadada@] Qed.
+End progress.
+
+Field typesafe :[@\hfill\codecomment{Type-safety theorem}@]
+  \forall e e' T, steps e e' -> hasty empty e T ->
+  value e' \/ \exists e'', step e' e''.
+Proof. [@\dadada@] Qed.
+
+End STLC.
+\end{lstlisting}
+\medskip
+
+\hrule
+\hrule
+\medskip
+
+\begin{lstlisting}[numbers=left,firstnumber=77]
+Family STLC_fix extends STLC.
+
+FInductive tm : Set +=
+| tm_fix : id -> tm -> tm.
+
+FInductive hasty : env -> tm -> ty -> Prop +=
+| ht_fix : \forall G x e T, hasty (extend G x T) e T ->
+  hasty G (tm_fix x e) T.
+
+FRecursion subst on tm motive \lambda _, id -> tm -> tm.
+Case tm_fix := [@\dadada@].
+End subst.
+
+FInductive step : tm -> tm -> Prop +=
+| st_fix : \forall x e,
+  step (tm_fix x e) (subst e x (tm_fix x e)).
+
+FInduction weakening on hasty motive [@\dadada@].
+Case ht_fix. [@\dadada@] Qed.
+End weakening.
+
+FInduction substlemma on hasty motive [@\dadada@].
+Case ht_fix. [@\dadada@] Qed.
+End subst_lemma.
+
+FInduction preserve on hasty motive [@\dadada@].
+Case ht_fix. [@\dadada@] Qed.
+End preserve.
+
+FInduction progress on hasty motive [@\dadada@].
+Case ht_fix. [@\dadada@] Qed.
+End progress.
+
+End STLC_fix.
+\end{lstlisting}
+
+\end{multicols}
+\end{minipage}
+\caption{Using our plugin to mechanize STLC and the fixpoints extension
+as depicted in \cref{fig:stlc-nonmechanized}.}
+\label{fig:stlc-mechanized}
+\end{figure}
+
+\ifShowOldWriting
+
 \begin{figure}
@@
 \end{figure}
+
+\fi

### 2022-10-02 challenge.tex
start lang-design

@@
 \noindentparagraph{%
-\headinglabel{C2}{chg:override-v-dependent}.\,
+\headinglabel{C2}{chg:self-v-consistency}.\,
+Self reference vs.\ logical consistency.%
+}
+
+Inheritance and family polymorphism are mechanisms for
+incrementally modifying self-referential definitions~\cite{cook1990inheritance}:
+the implicit self parameterization is how they achieve late binding and
+thus offer extensibility.
+However, self reference could easily lead to divergence.
+Divergence is not a concern for the design of ordinary OO or
+functional languages, but it would mean logical inconsistency---and
+hence unsoundness---for a language aimed at logical reasoning!
+A family-polymorphism design must tame self reference to guarantee consistency.
+
+\noindentparagraph{%
+\headinglabel{C3}{chg:override-v-dependent}.\,
 Overriding vs.\ dependent typing.%
@@
 A language design should only allow those overridings that are safe.
-
-\noindentparagraph{%
-\headinglabel{C3}{chg:self-v-consistency}.\,
-Self reference vs.\ logical consistency.%
-}
-
-Inheritance and family polymorphism are mechanisms for
-incrementally modifying self-referential definitions~\cite{cook1990inheritance}:
-the implicit self parameterization is how they achieve late binding and
-thus offer extensibility.
-However, self reference could easily lead to divergence.
-Divergence is not a concern for the design of ordinary OO or
-functional languages, but it would mean logical inconsistency---and
-hence unsoundness---for a language aimed at logical reasoning!
-A family-polymorphism design must tame self reference to guarantee consistency.
 

### 2022-10-02 lang-design.tex
start lang-design

@@
-We demonstrate the key ingredients of our design as a plugin to the Coq proof assistant,
+We use a plugin to the Coq proof assistant to demonstrate the key
+ingredients of our design,
 though we believe our design could be adapted to other proof assistants such as Lean.
 
+\input{stlc-mechanized.tex}
+
+\cref{fig:stlc-mechanized} shows how STLC and its extension with
+fixpoints can be mechanized using our plugin, in a style envisioned in
+\cref{fig:stlc-nonmechanized}.
+The base family "STLC" hosts the STLC metatheories, from abstract syntax
+to the type-safety theorem.
+Family "STLC_fix", derived from "STLC", makes adjustments as needed by a fixpoints extension:
+it adds new constructors to the inductive types ("FInductive") and
+adds new cases to the recursive functions ("FRecursion") and induction proofs ("FInduction").
+Existing constructors and cases, as well as those definitions and
+theorems that need no adjustments ("ty", "env", "empty", "steps", and
+"typesafe"), are automatically inherited and reused.
+
+\noindentparagraph{Inductive types are extensible.}
+
+In family "SLTC_fix", inductive type "tm" is formed by the four
+constructors inherited from "STLC" and a fifth constructor "tm_fix".
+
+Crucially, the types of the constructors are polymorphic to their enclosing family.
+For example, although "tm_app" is defined in the base family,
+it can be used to construct applications of fixpoints in the derived family:
+%such as \lstinline[mathescape]{tm_app (tm_fix "f" (tm_abs "x" $\dadada$)) tm_unit},
+its type \lstinline{tm -> tm -> tm} says "tm_app" can be applied to any
+Gallina terms of type "tm",
+which in family "SLTC_fix" include those constructed from "tm_fix".
+
+\noindentparagraph{Induction remains exhaustive.}
+
 \ref{chg:extensible-v-exhaustive}
-
-\input{stlc-mechanized.tex}
 

### 2022-10-02 stlc-mechanized.tex
start lang-design

@@
 \begin{multicols}{2}
+
+\lstset{
+  basicstyle=\fontsize{8}{8.5}\ttfamily,
+}
 
@@
 | tm_abs  : id -> tm -> tm
-| tm_app  : tm -> tm -> tm.
+| tm_app  : tm -> tm -> tm.[@\label{line:tm-app}@]
 
@@
 \end{minipage}
-\caption{Using our plugin to mechanize STLC and the fixpoints extension
-as depicted in \cref{fig:stlc-nonmechanized}.}
+\caption{Using our plugin to mechanize STLC and the fixpoints extension,
+as envisioned in \cref{fig:stlc-nonmechanized}.}
 \label{fig:stlc-mechanized}

### 2022-10-03 introduction.tex
tooling
exhaustivity

@@
 But the language-theoretic essence of inheritance is more general:
-it is a linguistic approach to \emph{incrementally modifying
-recursive definitions}~\cite{cook1990inheritance}.
+it is a linguistic approach to incrementally modifying
+record-like definitions by allowing \emph{late binding}~\cite{cook1990inheritance}.
 
@@
 %
-In particular, when a language mechanism allows inheritance over
-recursive definitions containing related types and terms,
+In particular, when a language mechanism allows late binding of the
+meaning of nested types and terms,
 it is said to support \emph{family polymorphism}~\cite{ernst2001family}:
@@
 underlying logic.
+%
+Specifically, we make the following contributions.
 
 \endgroup
-
-Specifically, we make the following contributions.
 

### 2022-10-03 lang-design.tex
exhaustivity

@@
 Family "STLC_fix", derived from "STLC", makes adjustments as needed by a fixpoints extension:
-it adds new constructors to the inductive types ("FInductive") and
-adds new cases to the recursive functions ("FRecursion") and induction proofs ("FInduction").
+it adds new constructors to the inductive types (\lsti{FInductive}) and
+adds new cases to the recursive functions (\lsti{FRecursion}) and
+induction proofs (\lsti{FInduction}).
 Existing constructors and cases, as well as those definitions and
@@
 
-\noindentparagraph{Inductive types are extensible.}
+\noindentparagraph{Extensibility of inductive types.}
 
-In family "SLTC_fix", inductive type "tm" is formed by the four
-constructors inherited from "STLC" and a fifth constructor "tm_fix".
+In family "STLC_fix", inductive type "tm" \emph{further binds} the "tm" type
+in family "STLC". It is generated by five constructors: four inherited from
+"STLC" and a fifth called "tm_fix".
 
-Crucially, the types of the constructors are polymorphic to their enclosing family.
+Crucially, the types of the constructors are polymorphic to their enclosing family---%
+their meaning is \emph{late bound}.
 For example, although "tm_app" is defined in the base family,
@@
 %such as \lstinline[mathescape]{tm_app (tm_fix "f" (tm_abs "x" $\dadada$)) tm_unit},
-its type \lstinline{tm -> tm -> tm} says "tm_app" can be applied to any
+its type \lsti{tm -> tm -> tm} says "tm_app" can be applied to any
 Gallina terms of type "tm",
-which in family "SLTC_fix" include those constructed from "tm_fix".
+which in family "STLC_fix" include those constructed from "tm_fix".
 
-\noindentparagraph{Induction remains exhaustive.}
+\noindentparagraph{Exhaustivity of induction.}
 
-\ref{chg:extensible-v-exhaustive}
+Ordinarily, an inductive type is not extensible: it is exhaustively
+generated by its constructors and has no more inhabitants beyond those
+they construct.
+This idea is captured by the eliminator (aka recursor) associated with
+an inductive type.
+For example, if "tm" were defined as an ordinary inductive type, then
+its eliminator would have the following type:
+
+\begingroup
+\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
+\begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
+\forall (P : tm -> Type), [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] -> [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] -> [@\hlcase{tm-recursor-case-abs}@](\forall x t, P t -> P (tm_abs x t))[@\tikzmarkend{tm-recursor-case-abs}@] ->
+  [@\hlcase{tm-recursor-case-app}@](\forall t1 t2, P t1 -> P t2 -> P (tm_app t1 t2))[@\tikzmarkend{tm-recursor-case-app}@] -> \forall t, P t
+\end{lstlisting}
+\endgroup
+
+\noindent
+The eliminator would allow for function definitions by recursion
+and proofs by induction, over "tm", that exhaustively
+handle the four cases corresponding to each constructor.
+The (dependent) return type~"P" is called the \emph{motive} of the recursion.
+%Exhaustivity of induction would follow, trivially.
+
+With inductive types made extensible, exhaustivity of induction becomes
+in question, however.
+To reconcile the tension without requiring redefinition or rechecking of
+case handlers (challenge \ref{chg:extensible-v-exhaustive}),
+our design introduces the \lsti{FRecursion} and \lsti{FInduction} commands.
+The key idea is to allow case handlers to be added retroactively, should
+inductive types be extended, and to allow recursion and induction (which are defined in terms of case handlers) to be late bound.
+
+As an example, consider the substitution function "subst", defined using
+\lsti!FRecursion!.
+The \lsti{on} clause specifies that recursion is over "tm".
+The \lsti{motive} clause suggests that the recursive function being
+defined has type \lsti{tm -> id -> tm -> tm}.
+%
+The "subst" function in "STLC_fix" further binds that in "STLC":
+the four cases from "STLC" are automatically inherited and reused,
+with "STLC_fix" retroactively adding a fifth case to form the new "subst" function.
+%
+For exhaustivity, it is a static error if the programmer fails to
+further bind "subst" and define this fifth case.
+The type system does this check by examining if the inductive type "tm",
+over which "subst" is recursively defined, is further bound in the same family.
+
+Like every other field of the family, the meaning of "subst" is late bound.
+For example, the type of "st_beta" (modeling the $\beta$-reduction rule)
+in family "STLC" refers to "subst".
+When "st_beta" is inherited into family "STLC_fix", it has a type that now
+refers to the "subst" function in the same family, where "subst" is defined
+by handling all the five cases known to that family.
+
+The \lstinline{FInduction} command is similar to \lstinline{FRecursion} but
+allows cases to be defined in proof mode.
+For example, "weakenlem" is proven by induction on the typing relation
+"hasty".
+Its \lsti{motive} clause suggests that the lemma reads as
+\lsti{\forall G t T, hasty G t T -> \forall G', includedin G G' -> hasty G' t T}.
+Upon entering case "ht_abs", for instance, Coq enters proof mode with the goal
+
+\begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
+  \forall ...
+\end{lstlisting}
+
+\noindent
+The programmer can discharge the goal using tactic programming.
+The idiomatic programming experience with Coq is largely preserved.
 

### 2022-10-03 stlc-mechanized.tex
exhaustivity

@@
 FInductive tm : Set :=[@\hfill\codecomment{Terms}@]
+| tm_unit : tm
 | tm_var  : id -> tm
-| tm_unit : tm
 | tm_abs  : id -> tm -> tm
@@
 FInductive hasty : env -> tm -> ty -> Prop :=
-| ht_var  : [@\dadada@]  | ht_unit : [@\dadada@][@\hfill\codecomment{Typing rules}@]
+| ht_unit : [@\dadada@]  | ht_var : [@\dadada@][@\hfill\codecomment{Typing rules}@]
 | ht_abs  : [@\dadada@]  | ht_app : [@\dadada@].
@@
 | v_unit : value tm_unit
-| v_abs  : \forall x e, value (tm_abs x e).
+| v_abs  : \forall x t, value (tm_abs x t).
 
@@
 motive \lambda(_ : tm), id -> tm -> tm.
-Case tm_var := \lambda y x e,
-  if (eqb x y) then e else (tm_var y).
-Case tm_unit := \lambda x e, tm_unit.
-Case tm_abs := \lambda y e' IHe' x e, [@\dadada@].
-Case tm_app := \lambda e1 IHe1 e2 IHe2 x e,
-  tm_app (IHe1 x e) (IHe2 x e).
+Case tm_unit := \lambda x t, tm_unit.
+Case tm_var := \lambda y x t,
+  if (eqb x y) then t else (tm_var y).
+Case tm_abs := \lambda y t' IHt' x t,
+  tm_abs y (if (eqb x y) then t' else IHt' x t).
+Case tm_app := \lambda t1 IHt1 t2 IHt2 x t,
+  tm_app (IHt1 x t) (IHt2 x t).
 End subst.
@@
 | st_app2 : [@\dadada@]
-| st_beta : \forall x e v, value v ->
-  step (tm_app (tm_abs x e) v) (subst e x v).
+| st_beta : \forall x t v, value v ->
+  step (tm_app (tm_abs x t) v) (subst t x v).
 
@@
 
-FInduction weakening on hasty[@\hfill\codecomment{Weaken.\ lemma}@]
-motive \lambda G e T (_ : hasty G e T),
-  \forall G', includedin G G' -> hasty G' e T.
-Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+FInduction weakenlem on hasty[@\hfill\codecomment{Weaken.\ lemma}@]
+motive \lambda G t T (_ : hasty G t T),
+  \forall G', includedin G G' -> hasty G' t T.
+Case ht_unit. [@\dadada@] Qed.[@\hfill@]Case ht_var. [@\dadada@] Qed.
 Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app. [@\dadada@] Qed.
-End weakening.
+End weakenlem.
 
-FInduction substlemma on hasty[@\hfill\codecomment{Subst.\ lemma}@]
-motive \lambda G' e T (_ : hasty G' e T),
-  \forall G x e' T', G' = extend G x T' ->
-  hasty empty e' T' -> hasty G (subst e x e') T.
-Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+FInduction substlem on hasty[@\hfill\codecomment{Subst.\ lemma}@]
+motive \lambda G' t T (_ : hasty G' t T),
+  \forall G x t' T', G' = extend G x T' ->
+  hasty empty t' T' -> hasty G (subst t x t') T.
+Case ht_unit. [@\dadada@] Qed.[@\hfill@]Case ht_var. [@\dadada@] Qed.
 Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app. [@\dadada@] Qed.
-End subst_lemma.
+End substlem.
 
 FInduction preserve on hasty[@\hfill\codecomment{Preserv.\ theorem}@]
-motive \lambda G e T (_ : hasty G e T),
-  G = empty -> \forall e', step e e' -> hasty empty e' T.
-Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+motive \lambda G t T (_ : hasty G t T),
+  G = empty -> \forall t', step t t' -> hasty empty t' T.
+Case ht_unit. [@\dadada@] Qed.[@\hfill@]Case ht_var. [@\dadada@] Qed.
 Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app. [@\dadada@] Qed.
@@
 FInduction progress on hasty[@\hfill\codecomment{Progress theorem}@]
-motive \lambda G e T (_ : hasty G e T),
-  G = empty -> value e \/ \exists e', step e e'.
-Case ht_var.  [@\dadada@] Qed.[@\hfill@]Case ht_unit. [@\dadada@] Qed.
+motive \lambda G t T (_ : hasty G t T),
+  G = empty -> value t \/ \exists t', step t t'.
+Case ht_unit. [@\dadada@] Qed.[@\hfill@]Case ht_var. [@\dadada@] Qed.
 Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app. [@\dadada@] Qed.
@@
 Field typesafe :[@\hfill\codecomment{Type-safety theorem}@]
-  \forall e e' T, steps e e' -> hasty empty e T ->
-  value e' \/ \exists e'', step e' e''.
+  \forall t t' T, steps t t' -> hasty empty t T ->
+  value t' \/ \exists t'', step t' t''.
 Proof. [@\dadada@] Qed.
@@
 FInductive hasty : env -> tm -> ty -> Prop +=
-| ht_fix : \forall G x e T, hasty (extend G x T) e T ->
-  hasty G (tm_fix x e) T.
+| ht_fix : \forall G x t T, hasty (extend G x T) t T ->
+  hasty G (tm_fix x t) T.
 
 FRecursion subst on tm motive \lambda _, id -> tm -> tm.
-Case tm_fix := [@\dadada@].
+Case tm_fix := \lambda y t' IHt' x t, [@\dadada@].
 End subst.
@@
 FInductive step : tm -> tm -> Prop +=
-| st_fix : \forall x e,
-  step (tm_fix x e) (subst e x (tm_fix x e)).
+| st_fix : \forall x t,
+  step (tm_fix x t) (subst t x (tm_fix x t)).
 
-FInduction weakening on hasty motive [@\dadada@].
+FInduction weakenlem on hasty motive [@\dadada@].
 Case ht_fix. [@\dadada@] Qed.
-End weakening.
+End weakenlem.
 
-FInduction substlemma on hasty motive [@\dadada@].
+FInduction substlem on hasty motive [@\dadada@].
 Case ht_fix. [@\dadada@] Qed.
-End subst_lemma.
+End substlem.
 

### 2022-10-05 challenge.tex
rework C2

--- 2022-10-01 challenge.tex
@@
 Such a feature would be useful for extending mechanized languages.
-For example, \cref{fig:stlc-nonmechanized} shows the part of an STLC extension that could
-potentially be mechanized using such extensible inductive types: a derived family
-inherits constructors of inductive types from the base family (e.g., the ellipsis under ``Typing rules'')
-and adds new constructors to model the new language feature.
+As an example, \cref{fig:stlc-nonmechanized} shows an STLC extension,
+the mechanization of which would be made easier by extensible inductive types.
+Code would be organized into two families, with the
+derived family inheriting constructors from the base family (e.g., the
+ellipsis under ``Typing rules'') and adding new constructors to model
+the syntax and semantics of a fixpoint construct.
 
@@
 In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
-the substitution function, require induction reasoning.
+the substitution function, require induction (i.e., elimination of inductive types).
 A language design must enforce that induction remains exhaustive in the
 face of the new constructors in the derived family.
-For modularity, it should do so without requiring
+For modularity, the type system should do so without requiring
 redefinition or rechecking of those cases already handled by the base
@@
 \noindentparagraph{%
-\headinglabel{C2}{chg:override-v-dependent}.\,
-Overriding vs.\ dependent typing.%
+\headinglabel{C2}{chg:latebind-v-defeq}.\,
+Late binding vs.\ definitional equality.%
 }
 
-Common to many family-polymorphism mechanisms is the ability for a
-derived family to override fields defined in the base family.
-As we show later, this expressiveness is also useful for metatheory
-mechanization.
-However, it is in conflict with dependent typing, wherein type checking
-often involves unfolding definitions.
-For example, in an STLC mechanization, the substitution lemma is checked
-against the definition of the substitution function.
-If a derived family were allowed to arbitrarily override the
-substitution function, then the proof of the substitution lemma, when
-inherited, would be ill-formed (even in the absence of any inductive-type extensions).
-A language design should only allow those overridings that are safe.
+Family polymorphism enables modular reuse via late binding:
+code of a base family can be reused by a derived family,
+because what fields referenced by that code unfold to is polymorphic to
+the enclosing family.
+
+This flexibility, however, prevents \emph{definitional equality}
+that one takes for granted when programming in a proof assistant.
+In \cref{fig:stlc-nonmechanized}, a proof assistant supporting family
+polymorphism cannot equate references to the substitution function and
+its definition (i.e., a pattern match against four cases),
+as a derived family may modify the definition of the function by adding
+new cases.
+Without the ability to unfold the substitution function, how can the
+programmer even prove the substitution lemma?
+%
+The problem is compounded by the occasional need to override fields in derived families,
+which also precludes the unfolding of definitions.
+
+%Common to family-polymorphism mechanisms in the OO realm is the ability
+%for a derived family to override fields defined in the base family.
+%As we show later, this expressiveness is also useful for metatheory
+%mechanization.
+%However, it is in conflict with dependent typing, wherein type checking
+%often involves unfolding definitions.
+%For example, in an STLC mechanization, the substitution lemma is checked
+%against the definition of the substitution function.
+%If a derived family were allowed to arbitrarily override the
+%substitution function, then the proof of the substitution lemma, when
+%inherited, would be ill-formed (even in the absence of any inductive-type extensions).
+%A language design should only allow those overridings that are safe.
 
@@
 functional languages, but it would mean logical inconsistency---and
-hence unsoundness---for a language aimed at logical reasoning!
+hence unsoundness!---for a language aimed at logical reasoning.
 A family-polymorphism design must tame self reference to guarantee consistency.
@@
 
-Interactive theorem proving and tactic programming are
-part of the idiomatic programming experience in a proof assistant.
+Interactive theorem proving and tactic programming are central to
+the idiomatic programming experience in a proof assistant.
 A language design integrating family polymorphism should be compatible

### 2022-10-05 lang-design.tex
rework C2

@@
 % \textbf{Final as non-overridable.} Every field is by default
-% overridable, reminding us of \ref{chg:override-v-dependent}.  To
+% overridable, reminding us of \ref{chg:latebind-v-defeq}.  To
 % resolve it, we provide the keyword "Final" to allow the user to specify
@@
 % our \lstinline{a'} is not overridable in "Family I2", mediating the
-% \ref{chg:override-v-dependent}.
+% \ref{chg:latebind-v-defeq}.
 
@@
 \textbf{Overridable Fields with dependency.}  Each field is by default non-overridable.
-By the discussion of \ref{chg:override-v-dependent}, we can simply reason and access the definition about the former fields in a later field, just like Field "a" and "b" in \cref{fig:plugin-example3} and we lose the ability to override.
+By the discussion of \ref{chg:latebind-v-defeq}, we can simply reason and access the definition about the former fields in a later field, just like Field "a" and "b" in \cref{fig:plugin-example3} and we lose the ability to override.
 
@@
 Each overridable field will only expose its type instead of the concrete definition to the following default (non-overridable) fields.
-For example, the first failure in \cref{fig:plugin-example3} per the discussion of \ref{chg:override-v-dependent}: since overridability will break the inheritability of "c" and thus we don't allow such field to exist in the first place.
+For example, the first failure in \cref{fig:plugin-example3} per the discussion of \ref{chg:latebind-v-defeq}: since overridability will break the inheritability of "c" and thus we don't allow such field to exist in the first place.
 Our plugin will only expose the fact that "a'" is an arbitrary natural number but hiding its concrete definition, and thus "eq_refl" will fail to type-check.
@@
 The resulting family only support direct inheritance or "Override" (i.e.
-we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:override-v-dependent}.
+we can only override a sealed family as a whole together)---we will later explain this design decision and its usage, especially on resolving \ref{chg:latebind-v-defeq}.
 What's more, "Override" family can only apply to sealed family as well (not allowed on normal family).
@@
 % nested families, inductive types, and fields.
-"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:override-v-dependent}.
+"Final" is used to claim that the decorated field is non-overridable, and later we will show how it is used to resolve \ref{chg:latebind-v-defeq}.
 The fields inside a family

### 2022-10-06 lang-design.tex
stuff

@@
 it can be used to construct applications of fixpoints in the derived family:
-%such as \lstinline[mathescape]{tm_app (tm_fix "f" (tm_abs "x" $\dadada$)) tm_unit},
+%such as \lsti[mathescape]{tm_app (tm_fix "f" (tm_abs "x" $\dadada$)) tm_unit},
 its type \lsti{tm -> tm -> tm} says "tm_app" can be applied to any
@@
 \newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
-\begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
 \forall (P : tm -> Type), [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] -> [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] -> [@\hlcase{tm-recursor-case-abs}@](\forall x t, P t -> P (tm_abs x t))[@\tikzmarkend{tm-recursor-case-abs}@] ->
@@
 %
-The "subst" function in "STLC_fix" further binds that in "STLC":
+The "subst" function in "STLC" is further bound by the "subst" in "STLC_fix":
 the four cases from "STLC" are automatically inherited and reused,
@@
 
-Like every other field of the family, the meaning of "subst" is late bound.
+Like all references to the fields of a family, the meaning of "subst" is late bound.
 For example, the type of "st_beta" (modeling the $\beta$-reduction rule)
@@
 When "st_beta" is inherited into family "STLC_fix", it has a type that now
-refers to the "subst" function in the same family, where "subst" is defined
+refers to the "subst" function in "STLC_fix", where "subst" is defined
 by handling all the five cases known to that family.
 
-The \lstinline{FInduction} command is similar to \lstinline{FRecursion} but
+The \lsti{FInduction} command is similar to \lsti{FRecursion} but
 allows cases to be defined in proof mode.
@@
 \noindent
-The programmer can discharge the goal using tactic programming.
-The idiomatic programming experience with Coq is largely preserved.
+The programmer can use tactic programming to discharge the goal.
+%The idiomatic programming experience with Coq is largely preserved.
+
+\noindentparagraph{Equality on late bound definitions.}
+
+Consider proving the "ht_unit" case of the substitution lemma, "substlem".
+The goal is seemingly trivial: the programmer is asked to prove
+
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+\forall G' G x t' T', G' = extend G x T' -> hasty empty t' T' -> hasty G (subst tm_unit x t') ty_unit.
+\end{lstlisting}
+
+\noindent
+If "subst" were an ordinary Coq function, then the programmer could
+discharge the goal with
+\lsti{intros; simpl subst; apply ht_unit.}
+The Ltac term \lsti{simpl subst} unfolds \lsti{subst} using its definition
+and simplifies \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
+
+However, with \lsti{subst} being late bound, \lsti{subst} can no longer
+be unfolded: the definition of \lsti{subst}, as a recursive function,
+varies across families, yet the proof of the \lsti{ht_unit} case should be
+reusable by a derived family regardless of how that family may add
+new cases to the definition of \lsti{subst}.
+How can the programmer change the goal to \lsti{hasty G tm_unit ty_unit}, then?
+
+A key observation is that although late binding
+prevents \emph{definitional equality} on \lsti{subst}, it does not
+affect \emph{propositional equality}.
+That is, \lsti{subst tm_unit x t' = tm_unit}, as a proposition
+(\lsti{Prop}), can still be proved.
+After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
+from a base family to a derived family.
+%
+Hence, our plugin provides a tactic \lsti{fsimpl}
+with which the programmer can rewrite \lsti{subst tm_unit x t'} to
+\lsti{tm_unit}.
 

### 2022-10-06 lang-design.tex
stuff

@@
 However, with \lsti{subst} being late bound, \lsti{subst} can no longer
-be unfolded: the definition of \lsti{subst}, as a recursive function,
-varies across families, yet the proof of the \lsti{ht_unit} case should be
-reusable by a derived family regardless of how that family may add
+be unfolded (challenge~\ref{chg:latebind-v-defeq}):
+the definition of \lsti{subst}, as a recursive function,
+varies across families, yet a derived family should be able to reuse the
+proof of the \lsti{ht_unit} case regardless of how it may add
 new cases to the definition of \lsti{subst}.
-How can the programmer change the goal to \lsti{hasty G tm_unit ty_unit}, then?
+Without the ability to unfold,
+how can the programmer make progress in this proof, then?
 
@@
 %
-Hence, our plugin provides a tactic \lsti{fsimpl}
-with which the programmer can rewrite \lsti{subst tm_unit x t'} to
-\lsti{tm_unit}.
+Based on this insight, our plugin automatically generates propositional
+equalities for recursive functions defined using \lsti{FRecursion}.
+It also provides a tactic \lsti{fsimpl} that enables, for instance,
+simplifying \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
+Under the hood, \lsti{fsimpl} works by rewriting using the generated
+propositional equalities.
+
+Note that only within those families that contain "subst" is definitional
+equality on \lsti{subst} unavailable.
+Within those families, the meaning of \lsti{subst} is late bound; it is
+polymorphic to the enclosing family.
+In contrast, outside those families, \lsti{subst} can be referenced only
+by specifying a family that contains it,
+such as \lsti{STLC.subst} or \lsti{STLC_fix.subst}.
+So Coq, by unfolding \lsti{STLC_fix.subst}, is able to consider
+\lsti{STLC_fix.subst tm_unit x t'} and \lsti{tm_unit}
+as definitionally equal.
 

### 2022-10-06 stlc-mechanized.tex
stuff

@@
 FInductive hasty : env -> tm -> ty -> Prop :=
-| ht_unit : [@\dadada@]  | ht_var : [@\dadada@][@\hfill\codecomment{Typing rules}@]
-| ht_abs  : [@\dadada@]  | ht_app : [@\dadada@].
+| ht_var  : [@\dadada@][@\hfill\codecomment{Typing rules}@]
+| ht_unit : \forall G, hasty G tm_unit ty_unit
+| ht_abs  : [@\dadada@]
+| ht_app  : [@\dadada@].
 
@@
 \end{lstlisting}
+\medskip
+
+\hrule
+\hrule
+\medskip
+
+\begin{lstlisting}[numbers=left,firstnumber=111]
+Check STLC_fix.typesafe.
+\end{lstlisting}
 

### 2022-10-07 lang-design.tex
stuff
late binding

--- 2022-10-06 lang-design.tex
@@
 "typesafe"), are automatically inherited and reused.
+For example, executing the last line, \lsti{Check STLC_fix.typesafe}, will
+display the type-safety theorem of the fixpoints extension.
 
-\noindentparagraph{Extensibility of inductive types.}
+\noindentparagraph{Extending inductive types.}
 
@@
 
-\noindentparagraph{Exhaustivity of induction.}
+\noindentparagraph{Ensuring exhaustivity of induction.}
 
@@
 With inductive types made extensible, exhaustivity of induction becomes
-in question, however.
+in jeopardy, however.
 To reconcile the tension without requiring redefinition or rechecking of
@@
 
-Like all references to the fields of a family, the meaning of "subst" is late bound.
-For example, the type of "st_beta" (modeling the $\beta$-reduction rule)
-in family "STLC" refers to "subst".
-When "st_beta" is inherited into family "STLC_fix", it has a type that now
-refers to the "subst" function in "STLC_fix", where "subst" is defined
-by handling all the five cases known to that family.
-
 The \lsti{FInduction} command is similar to \lsti{FRecursion} but
 allows cases to be defined in proof mode.
-For example, "weakenlem" is proven by induction on the typing relation
-"hasty".
+Consider "weakenlem" as an example.
+It is proven by induction on the typing relation "hasty".
 Its \lsti{motive} clause suggests that the lemma reads as
-\lsti{\forall G t T, hasty G t T -> \forall G', includedin G G' -> hasty G' t T}.
+
+\begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
+\forall G t T, hasty G t T -> \forall G', includedin G G' -> hasty G' t T.
+\end{lstlisting}
+
+\noindent
 Upon entering case "ht_abs", for instance, Coq enters proof mode with the goal
@@
 %The idiomatic programming experience with Coq is largely preserved.
+Because the lemma is by induction on "hasty" and because "STLC_fix" adds
+a constructor to "hasty", the programmer is required to handle this extra case
+in the proof of "substlem" in family "STLC_fix".
 
-\noindentparagraph{Equality on late bound definitions.}
+\noindentparagraph{Late binding of nested names.}
+
+OO inheritance allows the late binding of method names.
+Family polymorphism generalizes the power of OO inheritance by allowing the
+late binding of all names nested within families, including those referring to types.
+We have seen that late binding of "tm" allows the "tm" constructors in "STLC"
+to be reused in "STLC_fix" to construct terms containing "tm_fix".
+
+As another example, consider the type of "st_beta" in family "STLC".
+It refers to "subst", whose meaning is late bound.
+When "st_beta" is inherited into family "STLC_fix", "st_beta" has a type that now
+refers to the "subst" function in "STLC_fix", where "subst" is defined
+by handling all the five cases known to that family.
+Thus, late binding of "subst" allows the derived family to reuse "st_beta"
+to $\beta$-reduce abstractions whose body is a fixpoint.
+
+Importantly, a name is late bound only within a family that defines or extends it.
+Outside such families, the name can be accessed only by explicitly
+specifying a family that contains it.
+For example, the last line of \cref{fig:stlc-mechanized} accesses "typesafe"
+with "STLC_fix.typesafe".
+Executing the command prints
+
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+STLC_fix.typesafe: \forall t t' T, STLC_fix.steps t t' -> STLC_fix.hasty STLC_fix.empty t T ->
+                   STLC_fix.value t' \/ \exists t'', STLC_fix.step t' t''.
+\end{lstlisting}
+
+\noindent
+In this type, all references to nested names are qualified by "STLC_fix", as desired.
+
+\noindentparagraph{Equality on late bound names.}
 
@@
 However, with \lsti{subst} being late bound, \lsti{subst} can no longer
-be unfolded: the definition of \lsti{subst}, as a recursive function,
-varies across families, yet the proof of the \lsti{ht_unit} case should be
-reusable by a derived family regardless of how that family may add
-new cases to the definition of \lsti{subst}.
-How can the programmer change the goal to \lsti{hasty G tm_unit ty_unit}, then?
+be unfolded (challenge~\ref{chg:latebind-v-defeq}):
+the definition of \lsti{subst}, as a recursive function,
+varies across families, yet a derived family should be able to reuse the
+proof of the \lsti{ht_unit} case
+even when it has to modify the definition of \lsti{subst}.
+Without the ability to unfold \lsti{subst},
+how can the programmer make progress in this proof, then?
 
@@
 %
-Hence, our plugin provides a tactic \lsti{fsimpl}
-with which the programmer can rewrite \lsti{subst tm_unit x t'} to
-\lsti{tm_unit}.
+Based on this insight, our plugin automatically generates and proves
+propositional equalities for recursive functions defined with \lsti{FRecursion}.
+The plugin also provides a tactic \lsti{fsimpl} that enables, for instance,
+simplifying \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
+Under the hood, \lsti{fsimpl} works by rewriting using the generated
+propositional equalities.
+
+Note that definitional equality on \lsti{subst} is available
+outside those families that contain "subst".
+Within those families, the meaning of \lsti{subst} is late bound.
+In contrast, outside those families, \lsti{subst} can be referenced only
+by specifying a family that contains it,
+such as \lsti{STLC.subst} or \lsti{STLC_fix.subst}.
+So as far as Coq's type checker is concerned, \lsti{tm_unit} and
+\lsti{STLC_fix.subst tm_unit x t'} are the same thing---Coq equates them
+definitionally by unfolding \lsti{STLC_fix.subst} and performing
+$\beta$-reduction.
+
+\noindentparagraph{Overriding.}
+
+\noindentparagraph{Ensuring consistency.}
+
+Example of a circular proof
+
+\noindentparagraph{Composing families.}
+
+\noindentparagraph{Tactics for eliminating inductive types.}
+
+\noindentparagraph{Limitations.}
 

### 2022-10-07 lang-design.tex
overriding

@@
 \noindentparagraph{Overriding.}
+
+In an OO language, a subclass can override methods of a superclass.
+Similar expressiveness is useful for metatheory mechanization, too.
+%
+For example, in a derived family, the programmer may want to override
+a lemma that was proved by induction in the base family,
+by a new proof that uses a different proof strategy.
+
+Coq distinguishes \emph{opaque} definitions from \emph{transparent} ones.
+Our plugin supports the overriding of opaque definitions, which
+include most proofs.
+It is safe to override opaque definitions, because Coq will never attempt to unfold them.
+For transparent definitions, the common case is that the programmer does
+not want to override them.
+In \cref{fig:stlc-mechanized}, "env", "empty", "subst", and "steps" are
+transparent---that is, they are not defined with \lsti{Qed}---and thus
+the plugin treats them as non-overridable by default.
+As a result, the definitional equalities on "env", "empty", and "steps",
+as well as the propositional equality on "subst", are available to the
+type checker for type-checking the families.
+
+Our plugin does allow the overriding of transparent definitions that are
+explicitly marked as \lsti{Overridable} by the programmer.
+Overriding is made safe by requiring that when an overridable field is
+overridden, code whose type checking requires unfolding that definition
+should be overridden too.
+We give an example in a supplemental appendix.
 

### 2022-10-09 challenge.tex
sound logical reasoning

@@
 
-Inheritance and family polymorphism are mechanisms for
-incrementally modifying self-referential definitions~\cite{cook1990inheritance}:
-the implicit self parameterization is how they achieve late binding and
-thus offer extensibility.
+The language-theoretic essence of late binding is self reference;
+inheritance and family polymorphism are mechanisms for
+incrementally modifying self-referential definitions~\cite{cook1990inheritance}.
+%the implicit self parameterization is how they achieve late binding and
+%thus offer extensibility.
 However, self reference could easily lead to divergence.

### 2022-10-09 lang-design.tex
sound logical reasoning

--- 2022-10-07 lang-design.tex
@@
 
-Note that definitional equality on \lsti{subst} is available
+Note that the definitional equality on \lsti{subst} is available
 outside those families that contain "subst".
@@
 
-\noindentparagraph{Ensuring consistency.}
+In an OO language, a subclass can override methods of a superclass.
+Similar expressiveness is useful for metatheory mechanization, too.
+%
+For example, in a derived family, the programmer may want to override
+the proof of a lemma with a different proof strategy than it was proved
+in the base family.
+%
+Overriding is potentially in conflict with having equalities on late bound names.
 
-Example of a circular proof
+Coq distinguishes \emph{opaque} definitions from \emph{transparent} ones.
+Our plugin supports the overriding of opaque definitions, which
+include most proofs.
+It is safe to override opaque definitions, because Coq will never attempt to unfold them.
+For transparent definitions, the common case is that the programmer does
+not want to override them.
+In \cref{fig:stlc-mechanized}, "env", "empty", "subst", and "steps" are
+transparent---that is, they are not defined with \lsti{Qed}.
+Our plugin treats transparent definitions as non-overridable by default.
+Thus, the definitional equalities on "env", "empty", and "steps",
+as well as the propositional equality on "subst", are available to the
+type checker for type-checking the families.
+
+Our plugin does allow the overriding of transparent definitions that are
+explicitly marked as \lsti{Overridable} by the programmer.
+Overriding is made safe by requiring that when an overridable field is
+overridden, code whose type checking requires unfolding that field
+should be overridden too.
+We give an example in a supplemental appendix.
+
+\noindentparagraph{Sound logical reasoning.}
+
+In \cref{fig:stlc-mechanized}, just because the "typesafe" theorem in
+"STLC" is inherited and reused by "STLC_fix", it does not follow that
+in "STLC_fix" the programmer can use "typesafe" to prove "progress".
+If the programmer did, then they would be committing the logical fallacy
+of circular reasoning: the proof of "progress" would depend on "typesafe",
+yet the proof of "typesafe" depends on "progress".
+%
+Such circularity would easily lead to logical inconsistency:
+
+\begin{centered}
+\begin{minipage}{36ex}
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+Family A.
+FLemma f : False. Proof. Admitted.
+FLemma g : False. Proof. apply f. Qed.
+End A.
+\end{lstlisting}
+\end{minipage}
+\hspace{3.5em}
+\begin{minipage}{36ex}
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+Family B extends A.
+FLemma f : False. Proof. apply g. Qed.
+End A.
+[@~@]
+\end{lstlisting}
+\end{minipage}
+\end{centered}
+
+\noindent
+In the program above, family~"B" extends~"A", overriding lemma~"f" by
+proving it using lemma~"g".
+Lemma~"g" is in turn inherited from family "A", where it is proven using
+a late bound reference to lemma~"f".
+Circularity between "f" and "g" allows proving \lsti{False}!
+
+To ensure the soundness of logical reasoning, the type system requires
+that in a derived family, the context in which a field is defined be
+preserved from the base family.
+In "STLC", "progress" is in the context of "typesafe".
+Per the requirement, this relationship must be preserved into
+"STLC_fix", which prevents the proof of "progress" from depending on
+"typesafe" in "STLC_fix".
+
+Note that the requirement still allows a derived family to introduce
+new dependencies into the context of a field.
+For example, consider an extension of STLC with iso-recursive types
+(left column of \cref{fig:stlc-isorec-prod}).
+Family "STLC_isorec" introduces "tysubst" into the context of "hasty".
+\input{stlc-isorec-prod.tex}
 

### 2022-10-09 stlc-isorec-prod.tex
sound logical reasoning

\begin{figure}
\lstset{
    basicstyle=\fontsize{8}{8.5}\ttfamily,
% numbers=left,
}

\begin{minipage}{\textwidth}
\begin{multicols}{2}

\definecolor{codecomment-color}{HTML}{FF007F}

\begin{lstlisting}
Family STLC_isorec extends STLC.[@\tanc{STLCisorec-start}@]
[@\hfill\codecomment{STLC extended with iso-recursive types}@]
FInductive ty : Set +=
| ty_var : id -> ty | ty_rec : id -> ty -> ty.

FInductive tm : Set +=
| tm_fold : tm -> tm | tm_unfold : tm -> tm.

FRecursion tysubst on ty[@\hfill\codecomment{Type-level substitution}@]
motive \lambda(_ : ty), id -> ty -> ty.
Case ty_unit  := [@\dadada@].   Case ty_arrow := [@\dadada@].
Case ty_var   := [@\dadada@].   Case ty_rec   := [@\dadada@].
End tysubst.

FInductive hasty : env -> tm -> ty -> Prop +=
| ht_fold : \forall G t \alpha T,
    hasty G t (tysubst T \alpha (ty_rec \alpha T)) ->
    hasty G (tm_fold t) (ty_rec \alpha T)
| ht_unfold : [@\dadada@].

[@\dadada@]

End STLC_isorec.[@\tanc{STLCisorec-end}@]
\end{lstlisting}

\makeline[0pt]{STLCisorec-start}{STLCisorec-end}[codecomment-color!50]

\columnbreak

\definecolor{codecomment-color}{HTML}{0DA3FF}

\begin{lstlisting}
Family STLC_prod extends STLC.[@\tanc{STLCprod-start}@]
[@\hfill\codecomment{STLC extended with products}@]
FInductive ty : Set +=
| ty_prod : ty -> ty -> ty.

FInductive tm : Set +=
| tm_pair : tm -> tm -> tm
| tm_fst : tm -> tm | tm_snd : tm -> tm.

[@\dadada@]

End STLC_prod.[@\tanc{STLCprod-end}@]
\end{lstlisting}

\makeline[.5\textwidth+9pt]{STLCprod-start}{STLCprod-end}[codecomment-color!50]

\vspace{-5pt}

\definecolor{codecomment-color}{HTML}{E8A20C}

\begin{lstlisting}
Family STLC_isorec_prod extends STLC[@\tanc{STLCisorecprod-start}@]
using STLC_prod, STLC_isorec.
[@\hfill\ \ \codecomment{STLC extended with iso-recursive types and products}@]
FRecursion tysubst on ty
motive \lambda(_ : ty), id -> ty -> ty.
Case ty_prod := [@\dadada@].[@\hfill\codecomment{Substitution on product types}@]
End tysubst.

End STLC_isorec_prod.[@\tanc{STLCisorecprod-end}@]
\end{lstlisting}

\makeline[.5\textwidth+9pt]{STLCisorecprod-start}{STLCisorecprod-end}[codecomment-color!50]

\end{multicols}
\end{minipage}

\caption{Composing extensions of STLC.}
\label{fig:stlc-isorec-prod}
\end{figure}

### 2022-10-09 stlc-mechanized.tex
sound logical reasoning

@@
   basicstyle=\fontsize{8}{8.5}\ttfamily,
+% numbers=left,
 }
 
-\begin{lstlisting}[numbers=left]
-Family STLC.
+\colorlet{codecomment-color}{dkgreen}
+
+\begin{lstlisting}
+Family STLC.[@\hfill\codecomment{The base STLC}@][@\tanc{STLC-start}@]
 
@@
 
-Field env : Type := id -> option ty.
-Field empty : env := \lambda _, None.
+FDefinition env : Type := id -> option ty.
+FDefinition empty : env := \lambda _, None.
 
@@
 
-Field steps := clos_refl_trans step.
+FDefinition steps := clos_refl_trans step.
 
@@
 Case ht_abs.  [@\dadada@] Qed.[@\hfill@]Case ht_app. [@\dadada@] Qed.
-End substlem.
-
-FInduction preserve on hasty[@\hfill\codecomment{Preserv.\ theorem}@]
+End substlem.[@\tanc{STLC-fstcol-end}\columnbreak@]
+FInduction preserve on hasty[@\hfill\codecomment{Preserv.\ theorem}@][@\tanc{STLC-sndcol-start}@]
 motive \lambda G t T (_ : hasty G t T),
@@
 
-Field typesafe :[@\hfill\codecomment{Type-safety theorem}@]
+FTheorem typesafe :[@\hfill\codecomment{Type-safety theorem}@]
   \forall t t' T, steps t t' -> hasty empty t T ->
@@
 
-End STLC.
+End STLC.[@\tanc{STLC-end}@]
 \end{lstlisting}
-\medskip
 
-\hrule
-\hrule
-\medskip
+\makeline[0pt]{STLC-start}{STLC-fstcol-end}[codecomment-color!50]
+\makeline[.5\textwidth+9pt]{STLC-sndcol-start}{STLC-end}[codecomment-color!50]
 
-\begin{lstlisting}[numbers=left,firstnumber=77]
-Family STLC_fix extends STLC.
+\vspace{-8pt}
 
+\colorlet{codecomment-color}{orange}
+
+\begin{lstlisting}[firstnumber=77]
+Family STLC_fix extends STLC.[@\tanc{STLCfix-start}@]
+[@\hfill\codecomment{STLC extended with fixpoints}@]
 FInductive tm : Set +=
@@
 
-End STLC_fix.
+End STLC_fix.[@\tanc{STLCfix-end}@]
 \end{lstlisting}
-\medskip
 
-\hrule
-\hrule
-\medskip
+\makeline[.5\textwidth+9pt]{STLCfix-start}{STLCfix-end}[codecomment-color!50]
 
-\begin{lstlisting}[numbers=left,firstnumber=111]
-Check STLC_fix.typesafe.
+\vspace{-8pt}
+
+\colorlet{codecomment-color}{black}
+
+\begin{lstlisting}[firstnumber=111]
+Check STLC_fix.typesafe.[@\tanc{Check-STLCfix-typesafe}@]
 \end{lstlisting}
+
+\makeline[.5\textwidth+9pt]{Check-STLCfix-typesafe}{Check-STLCfix-typesafe}[codecomment-color!35]
 
@@
 \end{minipage}
+
 \caption{Example STLC and its extension, in designed syntax}\label{fig:STLC-example2}

### 2022-10-09 coqimpl.tex
Elaborate Compilation Process. Example for prec_discriminate.

@@
 %   (* Intermediate Module solely for type checking *)
-% Module v_33_34 (self_B: neg_handler_26).
+% Module do_tc (self_B: neg_handler_26).
 % Include b_rec_12 self_B.
@@
 %     self_B.neg_handler.ff.  
-% End v_33_34.
+% End do_tc.
 % \end{minted}
@@
 Family STLC.
-(* ... *)
+  FInductive tm : Set := ... 
+  FScheme tm_prec PRecT about self__STLC.tm.
+  FScheme Comp about self__STLC.tm_prec.
+  (* Generate Partial Recursor 
+  and the computational axiom *)
   Family subst_internal. 
@@
     by _rec.
-(*
-Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}\EDJreply{Because I want to show I can do it but I don't really want to d -- if I really do it the compilation result shown later will need to have this information.}@
-*)
+(* Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}\EDJreply{Because I want to show I can do it but I don't really want to d -- if I really do it the compilation result shown later will need to have this information.}@ *)
 (* ... *)
@@
   | tm_true : tm ...
-
 Extend Family subst_internal.
@@
 \begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
-
-Module Type STLC_409.
+Module Type STLC_Interface.
   Parameter (tm : Set).
@@
   and the abstraction for inductive type ...*)
-End STLC_409.
+End STLC_Interface.
 (* Collect eliminator to Set for tm *)
-Module tm_rec_425 (STLC: STLC_409).
+Module tm_rec_type (STLC: STLC_Interface).
   Definition __recursor_type_tm_rec :=
@@
     ... -> forall i : STLC.tm, P i.
-End tm_rec_425.
+End tm_rec_type.
 (* Compiled Field STLC.subst_internal.tm_var *)
-Module tm_var_411 (STLC: STLC_409)
-  (subst_internal: EmptySig).
+Module tm_var_handle (self__STLC: STLC_Interface)
+  (self__subst_internal: EmptySig_84).
 Definition tm_var :
-    forall (s : id) (x : id) (t : STLC.tm), STLC.tm :=
-    fun s x t ⇒ if eqb x s then t else STLC.var s.
-End tm_var_411. (* ... and more *)
-(* Aggregate into one subst_internal*)
-Module subst_internal_410 (STLC: STLC_409). 
-  Include tm_var_411 STLC. (* And other cases ... *)
+  forall (s : ident) (x : ident) (t : self__STLC.tm), self__STLC.tm :=
+  fun s x t => if eqb x s then t else self__STLC.tm_var s.
+End tm_var_handle. (* ... and more *)
+(* Assembling STLC.subst_internal when compilation *)
+Module subst_internal_410 (STLC: STLC_Interface). 
+  Module subst_internal.
+  Include tm_var_handle STLC. Include tm_abs_handle STLC.
+  Include tm_app_handle STLC. Include tm_unit_handle STLC.
+  End subst_internal.
 End subst_internal_410.
-Module v_33_34 (STLC: STLC_409).
+Module do_tc (STLC: STLC_Interface).
 (* Intermediate Module solely for type checking *)
-  Include tm_rec_425 STLC.
+  Include tm_rec_type STLC.
   Parameter (recursor_for_type_checking : __recursor_type_tm_rec).
@@
       subst_internal_410.tm_abs ... .
-End v_33_34.@\YZ{It'd be useful to (1) show how the parts of the STLC family are assembled and recursor inserted, and (2) show the code compiled for STLC\_bool and how code from STLC is shared with STLC\_bool.}@
+End do_tc.
+(* Ctx that expose concrete inductive type *)
+Module Type ctx_subst_113.
+Inductive __internal_tm : Set := ... 
+Include subst_internal_410. (* ... and more inclusion *)
+End ctx_subst_113.
+(* Compilation of recursor, not reusable
+    when inductive type extended due to the ctx *)
+Module subst_128 (self__STLC: ctx_subst_113).
+Include __motive_of_subst_107 self__STLC.
+Definition subst :=
+  self__STLC.__internal_tm_rec __motiveTsubst
+	self__STLC.subst_internal.tm_var self__STLC.subst_internal.tm_abs
+    self__STLC.subst_internal.tm_app self__STLC.subst_internal.tm_unit.
+End subst_128.
+(* Assembling the whole STLC *)
+Module STLC.
+Include tm_5. Include tm_prec_72. Include tm_prectm_81.
+Include subst_internal_410. Include subst_128.
+End v_129.
+(* Assembling STLC_bool.subst_internal via reusing 
+    most of the former compilation results *)
+Module subst_internal_248 (STLC_bool: STLC_Interface).
+Include tm_true_197 STLC_bool.
+Include tm_false_200 STLC_bool.
+Include tm_if_203 STLC_bool.
+(* The Reused Compilation result *)
+Include tm_var_handle STLC_bool.
+Include tm_abs_handle STLC_bool.
+Include tm_app_handle STLC_bool.
+Include tm_unit_handle STLC_bool.
+End v_248.
+@\YZ{It'd be useful to (1) show how the parts of the STLC family are assembled and recursor inserted, and (2) show the code compiled for STLC\_bool and how code from STLC is shared with STLC\_bool.}@
 \end{minted}
@@
 just like what we declare in ``the abstracted interface''---the compiled
-module type "STLC_409". We cannot export
+module type "STLC_Interface". We cannot export
 %\YZ{export or put?}\EDJreply{Ok.
@@
 %several places...}
-the eliminator of the inductive type "tm" to "STLC_409".
+the eliminator of the inductive type "tm" to "STLC_Interface".
 Otherwise all the subsequent fields (e.g. "subst_internal.tm_var" and
@@
 exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module
-"tm_rec_425". This concrete eliminator will be used by our "FRecursor".
+"tm_rec_type". This concrete eliminator will be used by our "FRecursor".
 For example, in \cref{fig:plugin-example2}, "FRecursor" command will
-feed Coq a module "v_33_34" for doing type-checking with specified
+feed Coq a module "do_tc" for doing type-checking with specified
 handlers (e.g. "neg_handler") during compilation---what it does is
@@
 and inherited (since they only rely on ``the abstracted interface''
-module "STLC_409"). 
+module "STLC_Interface"). 
 Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
@@
 Though partial recursor is provable by simply using "FRecursor", for convenience, we support using "FScheme" to automatic generate it.
-At the meantime, we implement a tactic "prec_discriminate" that when fed with the appropriate partial recursor, it can discriminate an absurd equation between two different constructors of extensible inductive type.
+At the meantime, we implement a tactic "prec_discriminate" that when fed with the appropriate partial recursor (with computational axiom generated), it can discriminate an absurd equation between two different constructors of extensible inductive type.
+\begin{minted}{Coq}
+  FInductive tm : Set := tunit : tm | tvar : ident -> tm | ..
+  FScheme tm_prec PRecT about self__STLC.tm.
+  FScheme Rec2D about self__STLC.tm_prec.
+  (* ... *)
+  Goal ∀ i, tvar i = tunit -> False.
+  intros i H. prec_discriminate self__STLC.tm_prec H. Qed.
+\end{minted}
 

### 2022-10-09 coqimpl.tex
Enhacning Compilation Code Example.

@@
 
-\begin{figure}[!htb]
-  \begin{minipage}[t]{0.32\linewidth}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family STLC.
-  FInductive tm : Set := ... 
-  FScheme tm_prec PRecT about self__STLC.tm.
-  FScheme Comp about self__STLC.tm_prec.
-  (* Generate Partial Recursor 
-  and the computational axiom *)
-  Family subst_internal. 
-  Final Field tm_var : ...
-:= fun s x t => 
-    if (eqb x s) 
-    then t else (tm_var s).
-  Final Field tm_abs : ...
-:= fun s (b : tm)   
-   (recb : id → tm → tm) 
-   x t => 
-    if (eqb x s) 
-    then (tm_abs s body)
-    else (tm_abs s (recb x t)).
-(* ... *)
-  EndFamily.
-  FRecursor subst 
-    about tm 
-    motive (fun _ => id → tm → tm)
-    using subst_internal
-    by _rec.
-(* Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}\EDJreply{Because I want to show I can do it but I don't really want to d -- if I really do it the compilation result shown later will need to have this information.}@ *)
-(* ... *)
-EndFamily.
-Family STLC_bool extends STLC.
-(* ... *)
-Extend FInductive tm : Set :=
-  | tm_true : tm ...
-Extend Family subst_internal.
-Final Field tm_true : ...
-  := fun x t => tm_true.
-(* ... *)
-EndFamily. 
-Inherits subst.
-(* ... *)
-EndFamily.
-\end{lstlisting}
-  \end{minipage}
-  \begin{minipage}[t]{0.65\linewidth}
-\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
-Module Type STLC_Interface.
-  Parameter (tm : Set).
-  Parameter (tm_var : id → tm).
-  Parameter (tm_abs : id → tm → tm).
-  (* ... and all the fields defined before
-  and the abstraction for inductive type ...*)
-End STLC_Interface.
-(* Collect eliminator to Set for tm *)
-Module tm_rec_type (STLC: STLC_Interface).
-  Definition __recursor_type_tm_rec :=
-    forall P : STLC.tm -> Set,
-    (forall n : id, P (STLC.tm_var n)) ->
-    (forall (n : id) (i : STLC.tm),
-    P i -> P (STLC.tm_abs n i)) ->
-    ... -> forall i : STLC.tm, P i.
-End tm_rec_type.
-(* Compiled Field STLC.subst_internal.tm_var *)
-Module tm_var_handle (self__STLC: STLC_Interface)
-  (self__subst_internal: EmptySig_84).
-Definition tm_var :
-  forall (s : ident) (x : ident) (t : self__STLC.tm), self__STLC.tm :=
-  fun s x t => if eqb x s then t else self__STLC.tm_var s.
-End tm_var_handle. (* ... and more *)
-(* Assembling STLC.subst_internal when compilation *)
-Module subst_internal_410 (STLC: STLC_Interface). 
-  Module subst_internal.
-  Include tm_var_handle STLC. Include tm_abs_handle STLC.
-  Include tm_app_handle STLC. Include tm_unit_handle STLC.
-  End subst_internal.
-End subst_internal_410.
-Module do_tc (STLC: STLC_Interface).
-(* Intermediate Module solely for type checking *)
-  Include tm_rec_type STLC.
-  Parameter (recursor_for_type_checking : __recursor_type_tm_rec).
-  Definition term_for_type_checking :=
-    recursor_for_type_checking (fun _  ⇒ id → STLC.tm → STLC.tm)
-      subst_internal_410.tm_var
-      subst_internal_410.tm_abs ... .
-End do_tc.
-(* Ctx that expose concrete inductive type *)
-Module Type ctx_subst_113.
-Inductive __internal_tm : Set := ... 
-Include subst_internal_410. (* ... and more inclusion *)
-End ctx_subst_113.
-(* Compilation of recursor, not reusable
-    when inductive type extended due to the ctx *)
-Module subst_128 (self__STLC: ctx_subst_113).
-Include __motive_of_subst_107 self__STLC.
-Definition subst :=
-  self__STLC.__internal_tm_rec __motiveTsubst
-	self__STLC.subst_internal.tm_var self__STLC.subst_internal.tm_abs
-    self__STLC.subst_internal.tm_app self__STLC.subst_internal.tm_unit.
-End subst_128.
-(* Assembling the whole STLC *)
-Module STLC.
-Include tm_5. Include tm_prec_72. Include tm_prectm_81.
-Include subst_internal_410. Include subst_128.
-End v_129.
-(* Assembling STLC_bool.subst_internal via reusing 
-    most of the former compilation results *)
-Module subst_internal_248 (STLC_bool: STLC_Interface).
-Include tm_true_197 STLC_bool.
-Include tm_false_200 STLC_bool.
-Include tm_if_203 STLC_bool.
-(* The Reused Compilation result *)
-Include tm_var_handle STLC_bool.
-Include tm_abs_handle STLC_bool.
-Include tm_app_handle STLC_bool.
-Include tm_unit_handle STLC_bool.
-End v_248.
-@\YZ{It'd be useful to (1) show how the parts of the STLC family are assembled and recursor inserted, and (2) show the code compiled for STLC\_bool and how code from STLC is shared with STLC\_bool.}@
-\end{minted}
-  \end{minipage}
-\caption{Exemplar STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
-\end{figure}
+% \begin{figure}[!htb]
+%   \begin{minipage}[t]{0.32\linewidth}
+% \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
+% Family STLC.
+%   FInductive tm : Set := tm_var : ... 
+%   FScheme tm_prec PRecT about self__STLC.tm.
+%   FScheme Comp about self__STLC.tm_prec.
+%   (* Generate Partial Recursor 
+%   and the computational axiom *)
+%   Family subst_internal. 
+%   Final Field tm_var : ...
+% := fun s x t => 
+%     if (eqb x s) 
+%     then t else (tm_var s).
+% (* ... other three cases *)
+%   EndFamily.
+%   FRecursor subst 
+%     about tm 
+%     motive (fun _ => id → tm → tm)
+%     using subst_internal
+%     by _rec.
+% (* Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}\EDJreply{Because I want to show I can do it but I don't really want to d -- if I really do it the compilation result shown later will need to have this information.}@ *)
+% (* ... *)
+% EndFamily.
+% Family STLC_bool extends STLC.
+% (* ... *)
+% Extend FInductive tm : Set :=
+%   | tm_true : tm ...
+% Extend Family subst_internal.
+% Final Field tm_true : ...
+%   := fun x t => tm_true.
+% (* ... *)
+% EndFamily. 
+% Inherits subst.
+% (* ... *)
+% EndFamily.
+% \end{lstlisting}
+%   \end{minipage}
+%   \begin{minipage}[t]{0.65\linewidth}
+% \begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
+% Module Type STLC_Interface.
+%   Parameter (tm : Set).
+%   Parameter (tm_var : id → tm).
+%   Parameter (tm_abs : id → tm → tm).
+%   (* ... and all the fields defined before
+%   and the abstraction for inductive type ...*)
+% End STLC_Interface.
+% (* Collect eliminator to Set for tm *)
+% Module tm_rec_type (STLC: STLC_Interface).
+%   Definition __recursor_type_tm_rec :=
+%     forall P : STLC.tm -> Set,
+%     (forall n : id, P (STLC.tm_var n)) ->
+%     (forall (n : id) (i : STLC.tm),
+%     P i -> P (STLC.tm_abs n i)) ->
+%     ... -> forall i : STLC.tm, P i.
+% End tm_rec_type.
+% (* Compiled Field STLC.subst_internal.tm_var *)
+% Module tm_var_handle (self__STLC: STLC_Interface)
+%   (self__subst_internal: EmptySig_84).
+% Definition tm_var :
+%   forall (s : ident) (x : ident) (t : self__STLC.tm), self__STLC.tm :=
+%   fun s x t => if eqb x s then t else self__STLC.tm_var s.
+% End tm_var_handle. (* ... and more *)
+% (* Assembling STLC.subst_internal when compilation *)
+% Module subst_internal_410 (STLC: STLC_Interface). 
+%   Module subst_internal.
+%   Include tm_var_handle STLC. Include tm_abs_handle STLC.
+%   Include tm_app_handle STLC. Include tm_unit_handle STLC.
+%   End subst_internal.
+% End subst_internal_410.
+% Module do_tc (STLC: STLC_Interface).
+% (* Intermediate Module solely for type checking *)
+%   Include tm_rec_type STLC.
+%   Parameter (recursor_for_type_checking : __recursor_type_tm_rec).
+%   Definition term_for_type_checking :=
+%     recursor_for_type_checking (fun _  ⇒ id → STLC.tm → STLC.tm)
+%       subst_internal_410.tm_var ... .
+% End do_tc.
+% (* Ctx that expose concrete inductive type *)
+% Module Type ctx_subst_113.
+% Inductive __internal_tm : Set := ... 
+% Include subst_internal_410. (* ... and more inclusion *)
+% End ctx_subst_113.
+% (* Compilation of recursor, not reusable
+%     when inductive type extended due to the ctx *)
+% Module subst_128 (self__STLC: ctx_subst_113).
+% Include __motive_of_subst_107 self__STLC.
+% Definition subst :=
+%   self__STLC.__internal_tm_rec __motiveTsubst
+% 	self__STLC.subst_internal.tm_var ... .
+% End subst_128.
+% (* Assembling the whole STLC *)
+% Module STLC.
+% Include tm_5. Include tm_prec_72. Include tm_prectm_81.
+% Include subst_internal_410. Include subst_128.
+% End v_129.
+% (* Assembling STLC_bool.subst_internal via reusing 
+%     most of the former compilation results *)
+% Module subst_internal_248 (STLC_bool: STLC_Interface).
+% Include tm_true_197 STLC_bool. Include ...
+% (* The Reused Compilation result *)
+% Include tm_var_handle STLC_bool. Include ...
+% End v_248.
+% @\YZ{It'd be useful to (1) show how the parts of the STLC family are assembled and recursor inserted, and (2) show the code compiled for STLC\_bool and how code from STLC is shared with STLC\_bool.}@
+% \end{minted}
+%   \end{minipage}
+% \caption{Exemplar STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
+% \end{figure}
+\include{ext-ind-type-compilation-example.tex}
 

### 2022-10-09 ext-ind-type-compilation-example.tex
Enhacning Compilation Code Example.

\begin{figure}
\lstset{
    basicstyle=\fontsize{8}{8.5}\ttfamily,
% numbers=left,
}

\begin{minipage}{\textwidth}
\begin{multicols}{2}

\definecolor{codecomment-color}{HTML}{FF007F}

\begin{lstlisting}
Family STLC.[@\tanc{Ext-Ind-Cmp-start}@]
 FInductive tm : Set := tm_var : [@\dadada@] 
 FScheme tm_prec PRecT about self__STLC.tm.
 FScheme Comp about self__STLC.tm_prec.
 (* Generate Partial Recursor 
    and the computational axiom *)
 Family subst_internal. 
  Final Field tm_var : [@\dadada@]
  := fun s x t => 
    if (eqb x s) 
    then t else (tm_var s).
 (* [@\dadada@] other three cases *)
 EndFamily.
 FRecursor subst 
  about tm 
  motive (fun _ => id → tm → tm)
  using subst_internal
  by _rec.
(* Field test := (subst (tm_var 0) 0 (tm_var 0)). *)
(*  *)
EndFamily.
Family STLC_bool extends STLC.
(* [@\dadada@] *)
 Extend FInductive tm : Set :=
  | tm_true : tm [@\dadada@]
 Extend Family subst_internal.
  Final Field tm_true : [@\dadada@]
   := fun x t => tm_true.
  (* [@\dadada@] *)
 EndFamily. 
 Inherits subst.
(* [@\dadada@] *)
EndFamily.[@\tanc{Ext-Ind-Cmp-end}@]
\end{lstlisting}

\makeline[0pt]{Ext-Ind-Cmp-start}{Ext-Ind-Cmp-end}[codecomment-color!50]

\columnbreak

\definecolor{codecomment-color}{HTML}{0DA3FF}

\begin{lstlisting}
Module Type STLC_Interface.[@\tanc{Cmp-Result-start}@]
  Parameter (tm : Set).
  Parameter (tm_var : id → tm).
  Parameter (tm_abs : id → tm → tm).
  (* [@\dadada@] and all the fields for inductive type [@\dadada@] *)
End STLC_Interface.
(* Collect eliminator to Set for tm *)
Module tm_rec_type (STLC: STLC_Interface).
  Definition __recursor_type_tm_rec :=
    forall P : STLC.tm -> Set,
    (forall n : id, P (STLC.tm_var n)) ->
    (forall (n : id) (i : STLC.tm),
    P i -> P (STLC.tm_abs n i)) ->
    [@\dadada@] -> forall i : STLC.tm, P i.
End tm_rec_type.
(* Compiled Field STLC.subst_internal.tm_var *)
Module tm_var_handle (self__STLC: STLC_Interface)
  (self__subst_internal: EmptySig_84).
Definition tm_var :
  forall (s : ident) (x : ident) (t : self__STLC.tm), self__STLC.tm :=
  fun s x t => if eqb x s then t else self__STLC.tm_var s.
End tm_var_handle. (* [@\dadada@] and more *)
(* Assembling STLC.subst_internal when compilation *)
Module subst_internal_410 (STLC: STLC_Interface). 
  Module subst_internal.
  Include tm_var_handle STLC. Include tm_abs_handle STLC.
  Include tm_app_handle STLC. Include tm_unit_handle STLC.
  End subst_internal.
End subst_internal_410.
Module do_tc (STLC: STLC_Interface).
(* Intermediate Module solely for type checking *)
  Include tm_rec_type STLC.
  Parameter (recursor_for_type_checking : __recursor_type_tm_rec).
  Definition term_for_type_checking :=
    recursor_for_type_checking (fun _  ⇒ id → STLC.tm → STLC.tm)
      subst_internal_410.tm_var [@\dadada@] .
End do_tc.
(* Ctx that expose concrete inductive type *)
Module Type ctx_subst_113.
Inductive __internal_tm : Set := [@\dadada@] 
Include subst_internal_410. (* [@\dadada@] and more inclusion *)
End ctx_subst_113.
(* Compilation of recursor, not reusable
    when inductive type extended due to the ctx *)
Module subst_128 (self__STLC: ctx_subst_113).
Include __motive_of_subst_107 self__STLC.
Definition subst :=
  self__STLC.__internal_tm_rec __motiveTsubst
	self__STLC.subst_internal.tm_var [@\dadada@] .
End subst_128.
(* Assembling the whole STLC *)
Module STLC.
Include tm_5. Include tm_prec_72. Include tm_prectm_81.
Include subst_internal_410. Include subst_128.
End v_129.
(* Assembling STLC_bool.subst_internal via reusing 
    most of the former compilation results *)
Module subst_internal_248 (STLC_bool: STLC_Interface).
Include tm_true_197 STLC_bool. Include [@\dadada@]
(* The Reused Compilation result *)
Include tm_var_handle STLC_bool. Include [@\dadada@]
End v_248. [@\tanc{Cmp-Result-end}@]

\end{lstlisting}

\makeline[.5\textwidth+9pt]{Cmp-Result-start}{Cmp-Result-end}[codecomment-color!50]

\vspace{-5pt}

\end{multicols}
\end{minipage}

\caption{Exemplar STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
\end{figure}

### 2022-10-09 ext-ind-type-compilation-example.tex
Fix example diagram

@@
  FInductive tm : Set := tm_var : [@\dadada@]
+ (* [@\dadada@] other constructors *) 
  FScheme tm_prec PRecT about self__STLC.tm.
@@
  (* Generate Partial Recursor 
-    and the computational axiom *)
+    and the computational axiom for tm *)
  Family subst_internal. 
@@
   := fun s x t => 
-    if (eqb x s) 
-    then t else (tm_var s).
- (* [@\dadada@] other three cases *)
+    if (eqb x s) then t else (tm_var s).
+  (* [@\dadada@] other cases *)
  EndFamily.
@@
 (* Field test := (subst (tm_var 0) 0 (tm_var 0)). *)
-(*  *)
+(* [@\dadada@] other things in STLC *)
 EndFamily.
@@
   | tm_true : tm [@\dadada@]
+  (* [@\dadada@] other new constructors *)
  Extend Family subst_internal.
@@
    := fun x t => tm_true.
-  (* [@\dadada@] *)
+  (* [@\dadada@] other extended cases *)
  EndFamily. 
  Inherits subst.
-(* [@\dadada@] *)
+(* [@\dadada@] other things in STLC_Bool *)
 EndFamily.[@\tanc{Ext-Ind-Cmp-end}@]
@@
 \makeline[0pt]{Ext-Ind-Cmp-start}{Ext-Ind-Cmp-end}[codecomment-color!50]
-
-\columnbreak
+\vspace{-3pt}
 
@@
 \begin{lstlisting}
-Module Type STLC_Interface.[@\tanc{Cmp-Result-start}@]
+(* Compilation Of the above *) [@\tanc{Cmp-Result0-start}@]
+Module Type STLC_Interface. 
   Parameter (tm : Set).
@@
     [@\dadada@] -> forall i : STLC.tm, P i.
-End tm_rec_type.
-(* Compiled Field STLC.subst_internal.tm_var *)
+End tm_rec_type. [@\tanc{Cmp-Result0-end}@]
+\end{lstlisting}
+
+\makeline[0pt]{Cmp-Result0-start}{Cmp-Result0-end}[codecomment-color!50]
+
+\columnbreak
+
+\definecolor{codecomment-color}{HTML}{0DA3FF}
+
+\begin{lstlisting}
+(* Compiled Field STLC.subst_internal.tm_var *) [@\tanc{Cmp-Result-start}@]
 Module tm_var_handle (self__STLC: STLC_Interface)
@@
   Definition tm_var :
-  forall (s : ident) (x : ident) (t : self__STLC.tm), self__STLC.tm :=
-  fun s x t => if eqb x s then t else self__STLC.tm_var s.
+    forall (s : ident) (x : ident) 
+    (t : self__STLC.tm), self__STLC.tm :=
+    fun s x t => if eqb x s
+       then t else self__STLC.tm_var s.
 End tm_var_handle. (* [@\dadada@] and more *)
@@
 End subst_internal_410.
+(* Intermediate Module solely for type checking *)
 Module do_tc (STLC: STLC_Interface).
-(* Intermediate Module solely for type checking *)
   Include tm_rec_type STLC.
-  Parameter (recursor_for_type_checking : __recursor_type_tm_rec).
+  Parameter (recursor_for_type_checking 
+    : __recursor_type_tm_rec).
   Definition term_for_type_checking :=
-    recursor_for_type_checking (fun _  ⇒ id → STLC.tm → STLC.tm)
+    recursor_for_type_checking 
+      (fun _  ⇒ id → STLC.tm → STLC.tm)
       subst_internal_410.tm_var [@\dadada@] .
@@
 End v_248. [@\tanc{Cmp-Result-end}@]
-
 \end{lstlisting}
@@
 
-\vspace{-5pt}
-
 \end{multicols}
@@
 
-\caption{Exemplar STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
+\caption{Exemplar Compilation about Inductive Type}\label{fig:plugin-example2}
 \end{figure}

### 2022-10-09 coqimpl.tex
Fixing section Extensible Inductive Type.

@@
 We support extensible inductive types in Coq mentioned in
-\ref{chg:extensible-inductive-type}, but we will not be aiming at the
-research of its formal semantic.\footnote{Because to give a satisfactory
-formal definition and semantic for extensible inductive type is not
-something this paper aims at, and obviously requires more theoretical
-research effort}
+\ref{chg:extensible-inductive-type}.
+% \footnote{Because to give a satisfactory
+% formal definition and semantic for extensible inductive type is not
+% something this paper aims at, and obviously requires more theoretical
+% research effort}
 %\YZ{I do consider that your plugin supports extensible inductive types.
@@
 
-Let's look at the exemplar \cref{fig:plugin-example2}, the example of
+We illustrate the usage and plugin implementation of inductive type and reasoning in the exemplar \cref{fig:plugin-example2}, the example of
 "subst" on STLC. When we define the extensible inductive type "tm", all
@@
 just like what we declare in ``the abstracted interface''---the compiled
-module type "STLC_Interface". We cannot export
+module type "STLC_Interface". However, we will not export
 %\YZ{export or put?}\EDJreply{Ok.
@@
 %several places...}
-the eliminator of the inductive type "tm" to "STLC_Interface".
-Otherwise all the subsequent fields (e.g. "subst_internal.tm_var" and
-the corresponding "tm_var_411") would not be able to be inherited to the
-context where "tm" is extended with a third constructor, \textbf{since
-the mere existence of the recursor will assert that there are only two
-constructors} in "STLC_bool", not \textbf{extensible} at all!
-
-However, we later still need the concrete eliminator on "tm" for doing
-exhaustiveness checking (as in \ref{chg:extensible-inductive-type}),
-thus we will store the corresponding info somewhere.  Here we show one
+the eliminator of the inductive type "tm" to "STLC_Interface", \textbf{since
+the mere existence of the eliminator will assert that there are only two
+constructors} in "tm", lead to an \textbf{non-extensible} "tm"!
+Thus the "STLC_Interface" will only include the ``constructor'' themselves in the form of parameter function.
+Doing so, all the subsequent fields (e.g.
+"subst_internal.tm_var" and the corresponding compiled "tm_var_handle") can be inherited to the
+context where "tm" is extended with new constructors without re-typechecking (re-compilation).
+
+To construct recursive function on a given inductive type, we use "FRecursion".
+"FRecursion" is a syntax sugar with (1) a normal family that includes all the case handling as field members, and (2) a step of doing type-checking and exhaustiveness checking (as in \ref{chg:extensible-inductive-type}).
+By considering all the case handling gathered in a family, the extension of the case handling can be simplified as the extension of the family.
+In our example, the resulting handler family is "subst_internal_410".
+While to achieve type-checking, we require the signature of eliminator about "tm" to be exposed.
+Here we show one
 exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module
-"tm_rec_type". This concrete eliminator will be used by our "FRecursor".
-For example, in \cref{fig:plugin-example2}, "FRecursor" command will
-feed Coq a module "do_tc" for doing type-checking with specified
-handlers (e.g. "neg_handler") during compilation---what it does is
+"tm_rec_type". This concrete eliminator will be used by our "FRecursion".
+For example, in \cref{fig:plugin-example2}, after closing an "FRecursion" block, our plugin will feed Coq the module "do_tc" for doing type-checking with specified
+handlers (e.g. "subst_internal_410") during compilation---what it does is
 basically
@@
 feel free to change the phrasing for me}
-Note that, at this point of the example,
-"FRecursor" will only do type check. The compilation of "FRecursor" into
-a component of the compiled module only happens when closing the whole
-family, by inserting the following into the compiled module: 
-\begin{minted}{Coq}
-Definition subst := tm_rec (λ _ ⇒ id → tm → tm) subst_internal.tm_var ...
-\end{minted}
-where "tm_rec" and "tm" are the corresponding recursor and
-the (vanilla) inductive type in that compiled module. Thus a "FRecursor"
-field is unlike other fields, even though it is exposed as a plain
-function for the succeeding fields in the programming interface.
-
-The plugin actually realizes the "Extend FInductive" via overriding---%
+
+% Note that, at this point of the example,
+% "FRecursion" will only do type check via "do_tc". 
+% The compilation of "FRecursor"  only happens when closing the whole
+% family, by including "subst_18". 
+% However, notice the context (parameter) of the "subst_18": it is "ctx_subst_113", where the concrete inductive type is exposed in the context, instead of "STLC_Interface".
+This concrete exposition makes "__internal_tm_rec" accessible, but also makes "subst_18" non-inheritable when "tm" is extended.
+Thus in every children family, modules like "subst_18" needs to be regenerated.
+% Thus a "FRecursion" field is unlike normal fields, 
+% even though it is exposed as a plain
+% function for the succeeding fields in the programming interface.
+
+Note that, the "FRecursion" field (e.g. "subst") is not actually
+inheritable, because it is not compiled like any other fields.
+Recall how other fields are compiled into reusable pieces of
+(parametrized) modules; a recursor field is not compiled like that.
+
+In our example "STLC", when the "FRecursion" block is closed, it will only do type check via "do_tc".  The compilation of "FRecursion"  
+only happens when closing the top-level
+family "STLC", by generating "subst_18" on-the-fly and assembly it into "module STLC". .
+However, notice the context (parameter) of the "subst_18": it is "ctx_subst_113", where the concrete inductive type is exposed in the context, instead of "STLC_Interface".
+This concrete exposition makes "__internal_tm_rec" accessible, but also makes "subst_18" non-inheritable when "tm" is extended.
+Thus, for all the children family, we have to regenerate module and replace the
+module with the ``new'' recursor using ``the same'' handler
+family (e.g. "Module subst_242", whose context "self__STLC_bool" is also different from that of the parent).
+
+This is affordable since each aggregated recursor handler
+doesn't need to be rechecked: look at surface syntax, the user only need to handle the new constructors like "tm_true" in "STLC_bool.subst"; during compilation,"Module subst_internal_248" will directly include the old handlers
+from the parent (e.g. "STLC.subst_internal.tm_var") that have already been type-checked (since they only rely on ``the abstracted interface''
+module "STLC_Interface" and thus well-typed). 
+Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
+the module "tm_var_handler"---doesn't require another type-check and is
+re-used during the type-checking and compilation of "STLC_bool.subst".
+
+Of course, in the implementation, these details are all hidden and when the user ``inherit'' the recursor "subst" in the surface syntax,
+our plugin actually will actually do a second exhaustiveness
+checking in children family mentioned above, with the ``same'' motive and the ``same'' handler family.
+%\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
+Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}
+
+The plugin realizes the "Extend FInductive" via overriding---%
 when extending an inductive type $A$ with new constructor $c$, our
@@
 
-Importantly, the recursor field (e.g. "subst") is not actually
-inheritable, because it is not compiled like any other fields.
-Recall how other fields are compiled into reusable pieces of
-(parametrized) modules; a recursor field is not compiled like that.
-For all the children family, we have to reconstruct and override the
-"FRecursor" field with a ``new'' recursor using (maybe the same) handler
-family.  This is affordable since each aggregated recursor handler
-doesn't need to be rechecked---look at "STLC_bool.neg_handler", there
-are only  new handlers for "tm_true" and the handlers (mundane fields)
-from the parent "B.subst_internal.tm_var" have already been type-checked
-and inherited (since they only rely on ``the abstracted interface''
-module "STLC_Interface"). 
-Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
-the module "tm_var_411"---doesn't require another type-check and is
-re-used during the type-checking and compilation of "STLC_bool.subst".
-For example, for the new "subst", when compiling family "STLC_bool" into
-a module, at the appropriate place our plugin will insert
-\begin{minted}{Coq}
-Definition subst := tm_rec (λ _ ⇒ id → tm → tm) subst_internal.tm_var ...
-\end{minted}
-where "tm_rec" and "tm" are the corresponding recursor and the (vanilla
-but extended and modified) inductive type in that compiled module.\YZ{Where is this "tm" inductive type in figure 6?} Note
-that, even though "tm_rec" and "tm" has the same name as before, they
-are in the newly-compiled module so they refer to totally different
-things.
-
-Of course, in the implementation, our plugin will ``inherit'' the
-recursor "neg" in the surface,
-but what the plugin actually does is doing a second exhaustiveness
-checking in children family like mentioned above, with the ``same'' motive and the ``same'' handler family.
-%\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
-Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}
-
 % Insert one pseudo-code example for explanation of the mechanism
@@
 % explain ftheorem wrapping this complicated recursor
-This design of decoupling of the case handlers and exhaustiveness
-checking handles the semantic well, however, it brings a lot of
-boilerplate code---we need to create an extra (handler) family, manually
-specifying the type of each case handlers, and then using "FRecursor" to
-``tie the knot'' and construct a recursor field from the handlers.  
-We thus provide a command "FTheorem" to avoid these boilerplate in programming interface when using tactic programming. 
-This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
-proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor". We expect this "FTheorem" to be used in theorem proving but not general programming.
+% This design of decoupling of the case handlers and exhaustiveness
+% checking handles the semantic well, however, it brings a lot of
+% boilerplate code---we need to create an extra (handler) family, manually
+% specifying the type of each case handlers, and then using "FRecursor" to
+% ``tie the knot'' and construct a recursor field from the handlers.
+
+For theorem proving, we provide a command "FInductive" that is doing similarly as "FRecursion" in programming interface but using tactic programming.
+Like "FRecursion", "FInduction" can handle newly added constructor in the children family as well.
+
+% This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
+% proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor". We expect this "FTheorem" to be used in theorem proving but not general programming.
 
@@
 handler inheritable, we need to seal an abstraction around the inductive
-type (e.g., module "b_3"); but to construct a recursor (and carry out the
+type (e.g., module "STLC_Interface"); but to construct a recursor (and carry out the
 exhaustiveness checking), we need to break this abstraction and see the
-concrete definition of the inductive type (e.g., module "b_rec_12"). Our
+concrete definition of the inductive type (e.g., module "tm_rec_type"). Our
 meta-theory need to handle these two seemingly contradicting ideas
@@
 \textit{decoupling} of the syntax of implementation of recursor handlers
-and exhaustiveness checking. The former is handled by family inheritance
+and exhaustiveness checking under the syntax sugar. 
+The former is handled by family inheritance
 and thus avoiding the boilerplate code; and the non-inheritance of the
@@
 satisfactory to be a legitimate programming interface. (1) We do not yet
-have any computational information using this "FRecursor" until its
-compilation into module, i.e. we cannot do computation with it during
+have any computational information using this "FRecursion" until its
+compilation into module, i.e. we cannot do computation with "subst" during
 the construction of the family using the current compilation strategy.
@@
 
-% \textit{Generating Scheme for Propositional Computation Axiom for FRecursor}.
-
+\textit{Generating Scheme for Propositional Computation Axiom for FRecursor}.
 Thus we support "FScheme" to remedy problem (1): our plugin can automatically generate(postulate) the computational ``axiom'' for the specified  into the family and prove them during compilation into modules.

### 2022-10-09 ext-ind-type-compilation-example.tex
Fixing section Extensible Inductive Type.

@@
     and the computational axiom for tm *)
- Family subst_internal. 
-  Final Field tm_var : [@\dadada@]
-  := fun s x t => 
+ FRecursion subst on tm 
+ motive (fun _ => id → tm → tm). 
+  Case tm_var := fun s x t => 
     if (eqb x s) then t else (tm_var s).
   (* [@\dadada@] other cases *)
- EndFamily.
- FRecursor subst 
-  about tm 
-  motive (fun _ => id → tm → tm)
-  using subst_internal
-  by _rec.
+ End subst.
 (* Field test := (subst (tm_var 0) 0 (tm_var 0)). *)
@@
 Family STLC_bool extends STLC.
-(* [@\dadada@] *)
  Extend FInductive tm : Set :=
@@
   (* [@\dadada@] other new constructors *)
- Extend Family subst_internal.
-  Final Field tm_true : [@\dadada@]
-   := fun x t => tm_true.
+ FRecursion subst.
+  Case tm_true := fun x t => tm_true.
   (* [@\dadada@] other extended cases *)
- EndFamily. 
+ End subst. 
  Inherits subst.
 (* [@\dadada@] other things in STLC_Bool *)
-EndFamily.[@\tanc{Ext-Ind-Cmp-end}@]
+EndFamily.
+[@\tanc{Ext-Ind-Cmp-end}@]
 \end{lstlisting}
@@
 \makeline[0pt]{Ext-Ind-Cmp-start}{Ext-Ind-Cmp-end}[codecomment-color!50]
-\vspace{-3pt}
+% \vspace{-3pt}
 
@@
     [@\dadada@] -> forall i : STLC.tm, P i.
-End tm_rec_type. [@\tanc{Cmp-Result0-end}@]
+End tm_rec_type.  
+Module tm_prec_72 (self__STLC: STLC_Interface).
+Definition tm_prec :
+  forall i : tm, forall P : tm -> Type,
+  (forall n, option (P (tm_var n))) ->
+  [@\dadada@], option (P __i) :=
+End tm_prec_72. 
+Module Type substtm_139 (self__STLC: [@\dadada@]).
+Parameter
+  (subst_on_tm_unit :
+	 eq (self__STLC.subst (self__STLC.tm_unit))
+       self__STLC.subst_internal.tm_unit).
+(* [@\dadada@] subst equality on other constructors *)
+End substtm_139. 
+[@\tanc{Cmp-Result0-end}@]
 \end{lstlisting}
@@
 Include tm_var_handle STLC_bool. Include [@\dadada@]
-End v_248. [@\tanc{Cmp-Result-end}@]
+End v_248. 
+(* Regenerated FRecursion result *)
+Module subst_242 (self__STLC_bool: [@\dadada@]).
+Include __motive_of_subst_107 self__STLC_bool.
+Definition subst :=
+  self__STLC_bool.__internal_tm_rec __motiveTsubst
+	self__STLC_bool.subst_internal.tm_var [@\dadada@]
+    self__STLC_bool.subst_internal.tm_true [@\dadada@] .
+End subst_242. [@\tanc{Cmp-Result-end}@]
 \end{lstlisting}

### 2022-10-09 coqimpl.tex
Adding information for Partial recursor.

@@
 Thus we support "FScheme" to remedy problem (1): our plugin can automatically generate(postulate) the computational ``axiom'' for the specified  into the family and prove them during compilation into modules.
+For example, "FScheme Comp about subst" in \cref{fig:plugin-example2} will compile into "Module Type substtm_139", filled with all the computational axiom for "subst".
+There module type when included in the context can be used to reduce "subst"; they will also be proved after compilation.
+
 However, it is only introducing computational information \textbf{propositionally} but not judgementally, and we might need the help of rewriting tactic on these computational ``axiom'' to justify the computation.
@@
 Though partial recursor is provable by simply using "FRecursor", for convenience, we support using "FScheme" to automatic generate it.
+For example, "FScheme tm_prec PRecT" in \cref{fig:plugin-example2} can lead to compilation of "Module tm_prec_72".
 At the meantime, we implement a tactic "prec_discriminate" that when fed with the appropriate partial recursor (with computational axiom generated), it can discriminate an absurd equation between two different constructors of extensible inductive type.

### 2022-10-09 lang-design.tex
mixin

@@
 Note that the requirement still allows a derived family to introduce
-new dependencies into the context of a field.
-For example, consider an extension of STLC with iso-recursive types
-(left column of \cref{fig:stlc-isorec-prod}).
-Family "STLC_isorec" introduces "tysubst" into the context of "hasty".
+new declarations into the context of a field.
+For example, the left column of \cref{fig:stlc-isorec-prod} shows an
+extension of STLC with iso-recursive types, where "tysubst" is
+introduced into the context of "hasty".
+
 \input{stlc-isorec-prod.tex}
@@
 \noindentparagraph{Composing families.}
+
+Families can be readily reused to construct larger extensions that
+\emph{mix in}~\cite{mixin-1990} the functionalities of the individual
+families.
+In \cref{fig:stlc-isorec-prod}, family "STLC_fix_isorec" is declared as
+an "STLC" extension that mixes in "STLC_fix" and "STLC_isorec".
+As a result, \lsti{STLC_fix_isorec.typesafe} readily gives the
+type-safety theorem of STLC equipped with fixpoints and iso-recursive types.
+
+Mixin composition is a form of multiple inheritance, which may lead to
+name conflicts in general.
+Our plugin requires the programmer to resolve conflicts by
+explicitly overriding conflicted fields.
+
+In the presence of extensible inductive types, mixin composition may
+also create an obligation to retrofit the mixins with new case handlers.
+In \cref{fig:stlc-isorec-prod}, family "STLC_prod_isorec" is composed
+of two mixins: "STLC_prod", which extends the inductive type "ty" with a
+new constructor "ty_prod", and "STLC_isorec", which introduces a
+function "tysubst" recursively defined on "ty".
+Hence, for exhaustivity, it is required that a composition of
+"STLC_prod" and "STLC_isorec" should additionally handle the "ty_prod"
+case in "tysubst".
 

### 2022-10-09 stlc-isorec-prod.tex
mixin

@@
 
-\definecolor{codecomment-color}{HTML}{FF007F}
+\definecolor{codecomment-color}{HTML}{DD033F}
 
@@
 
-[@\dadada@]
-
+[@\dadada@][@\hfill\codecomment{Other adjustments}@]
 End STLC_isorec.[@\tanc{STLCisorec-end}@]
@@
 \makeline[0pt]{STLCisorec-start}{STLCisorec-end}[codecomment-color!50]
+
+\vspace{-12pt}
+
+\definecolor{codecomment-color}{HTML}{008B8B}
+
+\begin{lstlisting}
+Family STLC_fix_isorec extends STLC[@\tanc{STLCisorecfix-start}@]
+using STLC_fix, STLC_isorec.
+[@\hfill\ \ \codecomment{STLC extended with fixpoints and iso-recursive types}@]
+End STLC_fix_isorec.[@\tanc{STLCisorecfix-end}@]
+\end{lstlisting}
+
+\makeline[0pt]{STLCisorecfix-start}{STLCisorecfix-end}[codecomment-color!50]
 
@@
 
-[@\dadada@]
-
+[@\dadada@][@\hfill\codecomment{Other adjustments}@]
 End STLC_prod.[@\tanc{STLCprod-end}@]
@@
 
-\vspace{-5pt}
+\vspace{-10pt}
 
@@
 \begin{lstlisting}
-Family STLC_isorec_prod extends STLC[@\tanc{STLCisorecprod-start}@]
+Family STLC_prod_isorec extends STLC[@\tanc{STLCisorecprod-start}@]
 using STLC_prod, STLC_isorec.
-[@\hfill\ \ \codecomment{STLC extended with iso-recursive types and products}@]
+[@\hfill\ \ \codecomment{STLC extended with products and iso-recursive types}@]
 FRecursion tysubst on ty
@@
 
-End STLC_isorec_prod.[@\tanc{STLCisorecprod-end}@]
+End STLC_prod_isorec.[@\tanc{STLCisorecprod-end}@]
 \end{lstlisting}
@@
 \makeline[.5\textwidth+9pt]{STLCisorecprod-start}{STLCisorecprod-end}[codecomment-color!50]
+
+\vspace{-10pt}
+
+\definecolor{codecomment-color}{HTML}{708090}
+
+\begin{lstlisting}
+Family STLC_fix_prod_isorec extends STLC[@\tanc{STLCisorecfixprod-start}@]
+using STLC_fix, STLC_prod_isorec.[@\hfill\codecommentstart{STLC extended}@]
+[@\hfill\codecommentend{with fixpoints, products, and iso-recursive types}@]
+End STLC_fix_isorec.[@\tanc{STLCisorecfixprod-end}@]
+\end{lstlisting}
+
+\makeline[.5\textwidth+9pt]{STLCisorecfixprod-start}{STLCisorecfixprod-end}[codecomment-color!50]
 

### 2022-10-11 syntax.tex
Should mention how computational axiom etc is modelled.
cite "explicit subst" paper

@@
 (3) Instead of using meta-level substitution, we use explicit
-substitution and de Bruijn indices in our intrinsic formulation. This is
+substitution~\cite{abadi1989subst} and de Bruijn indices in our intrinsic formulation. This is
 also called \textit{substitution calculus} in the literature and is favoured due
@@
 In our metatheory, for simplicity we only formulate non-dependent eliminators, i.e.,
-\textit{recursors}.
+\textit{recursors}.\YZ{Metatheory should mention how computational axioms about case handlers, partial recursors, and axioms over partial recursors are dealt with in the metatheory.}  
 %\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}\EDJreply{I don't think there is technical difficulties to do that.

### 2022-10-12 challenge.tex
partial recursor

@@
 code of a base family can be reused by a derived family,
-because what fields referenced by that code unfold to is polymorphic to
-the enclosing family.
+because fields referenced by that code have meanings polymorphic to the
+enclosing family.
 
@@
 %
-The problem is compounded by the occasional need to override fields in derived families,
-which also precludes the unfolding of definitions.
+The problem is compounded by the occasional need in derived families to override fields,
+which is potentially at odds with being able to use equalities over the fields.
 
@@
 even in the middle of a family definition.
-In addition, it should remain possible to perform computation in the extended Gallina language.
+In addition, it should remain possible to normalize terms that mention families.
 

### 2022-10-12 lang-design.tex
partial recursor

@@
 to the type-safety theorem.
-Family "STLC_fix", derived from "STLC", makes adjustments as needed by a fixpoints extension:
+Family "STLCFix", derived from "STLC", makes adjustments as needed by a fixpoints extension:
 it adds new constructors to the inductive types (\lsti{FInductive}) and
@@
 "typesafe"), are automatically inherited and reused.
-For example, executing the last line, \lsti{Check STLC_fix.typesafe}, will
-display the type-safety theorem of the fixpoints extension.
+In particular, executing the last command, \lsti{Check STLCFix.typesafe},
+displays the type-safety theorem of the fixpoints extension.
 
@@
 
-In family "STLC_fix", inductive type "tm" \emph{further binds} the "tm" type
+In family "STLCFix", inductive type "tm" \emph{further binds} the "tm" type
 in family "STLC". It is generated by five constructors: four inherited from
@@
 
-Crucially, the types of the constructors are polymorphic to their enclosing family---%
-their meaning is \emph{late bound}.
+Crucially, the types of the constructors are polymorphic to their enclosing family:
+the types mention "tm", and the meaning of "tm" is \emph{late bound}.
 For example, although "tm_app" is defined in the base family,
@@
 Gallina terms of type "tm",
-which in family "STLC_fix" include those constructed from "tm_fix".
+which in family "STLCFix" include those constructed from "tm_fix".
 
@@
 they construct.
-This idea is captured by the eliminator (aka recursor) associated with
-an inductive type.
+This idea is captured by the eliminator (aka \emph{recursor})
+associated with an inductive type.
 For example, if "tm" were defined as an ordinary inductive type, then
@@
 
-\begingroup
+\begin{centered}
+\begin{minipage}{.71\textwidth}
 \newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-\forall (P : tm -> Type), [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] -> [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] -> [@\hlcase{tm-recursor-case-abs}@](\forall x t, P t -> P (tm_abs x t))[@\tikzmarkend{tm-recursor-case-abs}@] ->
-  [@\hlcase{tm-recursor-case-app}@](\forall t1 t2, P t1 -> P t2 -> P (tm_app t1 t2))[@\tikzmarkend{tm-recursor-case-app}@] -> \forall t, P t
+tm_rec : \forall (P : tm -> Set), [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] -> [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] ->
+         [@\hlcase{tm-recursor-case-abs}@](\forall x t, P t -> P (tm_abs x t))[@\tikzmarkend{tm-recursor-case-abs}@] ->
+         [@\hlcase{tm-recursor-case-app}@](\forall t1, P t1 -> \forall t2, P t2 -> P (tm_app t1 t2))[@\tikzmarkend{tm-recursor-case-app}@] -> \forall t, P t
 \end{lstlisting}
-\endgroup
+\end{minipage}
+\end{centered}
 
 \noindent
-The eliminator would allow for function definitions by recursion
+The eliminator would enable function definitions by recursion
 and proofs by induction, over "tm", that exhaustively
@@
 in jeopardy, however.
+In particular, recursor "tm_rec" should no longer be allowed, because
+its type mentions "tm", whose meaning is late bound, yet
+"tm_rec" purports to claim \lsti{\forall t, P t} given only four case handlers.
+
 To reconcile the tension without requiring redefinition or rechecking of
-case handlers (challenge \ref{chg:extensible-v-exhaustive}),
+case handlers~(\ref{chg:extensible-v-exhaustive}),
 our design introduces the \lsti{FRecursion} and \lsti{FInduction} commands.
@@
 %
-The "subst" function in "STLC" is further bound by the "subst" in "STLC_fix":
+The "subst" function in "STLC" is further bound by the "subst" in "STLCFix":
 the four cases from "STLC" are automatically inherited and reused,
-with "STLC_fix" retroactively adding a fifth case to form the new "subst" function.
+with "STLCFix" retroactively adding a fifth case to form the new "subst" function.
 %
@@
 %The idiomatic programming experience with Coq is largely preserved.
-Because the lemma is by induction on "hasty" and because "STLC_fix" adds
+Because the lemma is by induction on "hasty" and because "STLCFix" adds
 a constructor to "hasty", the programmer is required to handle this extra case
-in the proof of "substlem" in family "STLC_fix".
+in the proof of "substlem" in family "STLCFix".
 
@@
 We have seen that late binding of "tm" allows the "tm" constructors in "STLC"
-to be reused in "STLC_fix" to construct terms containing "tm_fix".
+to be reused in "STLCFix" to construct terms containing "tm_fix".
 
@@
 It refers to "subst", whose meaning is late bound.
-When "st_beta" is inherited into family "STLC_fix", "st_beta" has a type that now
-refers to the "subst" function in "STLC_fix", where "subst" is defined
+When "st_beta" is inherited into family "STLCFix", "st_beta" has a type that now
+refers to the "subst" function in "STLCFix", where "subst" is defined
 by handling all the five cases known to that family.
@@
 For example, the last line of \cref{fig:stlc-mechanized} accesses "typesafe"
-with "STLC_fix.typesafe".
+with "STLCFix.typesafe".
 Executing the command prints
@@
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-STLC_fix.typesafe: \forall t t' T, STLC_fix.steps t t' -> STLC_fix.hasty STLC_fix.empty t T ->
-                   STLC_fix.value t' \/ \exists t'', STLC_fix.step t' t''.
+STLCFix.typesafe: \forall t t' T, STLCFix.steps t t' -> STLCFix.hasty STLCFix.empty t T ->
+                   STLCFix.value t' \/ \exists t'', STLCFix.step t' t''.
 \end{lstlisting}
@@
 \noindent
-In this type, all references to nested names are qualified by "STLC_fix", as desired.
+In this type, all references to nested names are qualified by "STLCFix", as desired.
 
@@
 However, with \lsti{subst} being late bound, \lsti{subst} can no longer
-be unfolded (challenge~\ref{chg:latebind-v-defeq}):
+be unfolded~(\ref{chg:latebind-v-defeq}):
 the definition of \lsti{subst}, as a recursive function,
@@
 That is, \lsti{subst tm_unit x t' = tm_unit}, as a proposition
-(\lsti{Prop}), can still be proved.
+(\lsti{Prop}), should still hold.
 After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
@@
 %
-Based on this insight, our plugin automatically generates and proves
-propositional equalities for recursive functions defined with \lsti{FRecursion}.
+Based on this insight, our plugin automatically generates
+propositional equalities for each recursive function defined with \lsti{FRecursion},
+making them available as axioms for use by the rest of the current family.
 The plugin also provides a tactic \lsti{fsimpl} that enables, for instance,
 simplifying \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
-Under the hood, \lsti{fsimpl} works by rewriting using the generated
+Under the hood, \lsti{fsimpl} works by rewriting using the postulated
 propositional equalities.
@@
 outside those families that contain "subst".
-Within those families, the meaning of \lsti{subst} is late bound.
-In contrast, outside those families, \lsti{subst} can be referenced only
+Within those families, the meaning of \lsti{subst} is late bound, so only
+propositional equality is available.
+In contrast, outside those families, \lsti{subst} is always referenced
 by specifying a family that contains it,
-such as \lsti{STLC.subst} or \lsti{STLC_fix.subst}.
-So as far as Coq's type checker is concerned, \lsti{tm_unit} and
-\lsti{STLC_fix.subst tm_unit x t'} are the same thing---Coq equates them
-definitionally by unfolding \lsti{STLC_fix.subst} and performing
-$\beta$-reduction.
+such as \lsti{STLC.subst} or \lsti{STLCFix.subst}.
+As far as the type checker is concerned, \lsti{tm_unit} and
+\lsti{STLCFix.subst tm_unit x t'} are the same thing---the type checker
+equates them definitionally by unfolding \lsti{STLCFix.subst} and
+performing normalization.
 
@@
 %
-Overriding is potentially in conflict with having equalities on late bound names.
+Overriding is potentially incompatible with having equalities on late bound names.
 
@@
 Overriding is made safe by requiring that when an overridable field is
-overridden, code whose type checking requires unfolding that field
+overridden, code whose type checking involves unfolding that field
 should be overridden too.
-We give an example in a supplemental appendix.
+We expect this feature to be used only occasionally.
+We give an example in \cref{sec:coqexample}.
 
@@
 In \cref{fig:stlc-mechanized}, just because the "typesafe" theorem in
-"STLC" is inherited and reused by "STLC_fix", it does not follow that
-in "STLC_fix" the programmer can use "typesafe" to prove "progress".
+"STLC" is inherited and reused by "STLCFix", it does not follow that
+in "STLCFix" the programmer can use "typesafe" to prove "progress".
 If the programmer did, then they would be committing the logical fallacy
@@
 \begin{centered}
+\definecolor{codecomment-color}{HTML}{556B2F}
 \begin{minipage}{36ex}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-Family A.
+Family A.[@\tanc{A-inconsistency-start}@]
 FLemma f : False. Proof. Admitted.
 FLemma g : False. Proof. apply f. Qed.
-End A.
+End A.[@\tanc{A-inconsistency-end}@]
 \end{lstlisting}
 \end{minipage}
+\makeline[22pt]{A-inconsistency-start}{A-inconsistency-end}[codecomment-color!50]
 \hspace{3.5em}
+\definecolor{codecomment-color}{HTML}{FF00FF}
 \begin{minipage}{36ex}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-Family B extends A.
+Family B extends A.[@\tanc{B-inconsistency-start}@]
 FLemma f : False. Proof. apply g. Qed.
-End A.
+End B.[@\tanc{B-inconsistency-end}@]
 [@~@]
@@
 \end{minipage}
+\makeline[.5\textwidth+23pt]{B-inconsistency-start}{B-inconsistency-end}[codecomment-color!50]
 \end{centered}
@@
 
-To ensure the soundness of logical reasoning, the type system requires
-that in a derived family, the context in which a field is defined be
-preserved from the base family.
+To ensure the soundness of logical reasoning (\ref{chg:self-v-consistency}),
+the type system requires that in a derived family, the context in which
+a field is defined be preserved from the base family.
 In "STLC", "progress" is in the context of "typesafe".
 Per the requirement, this relationship must be preserved into
-"STLC_fix", which prevents the proof of "progress" from depending on
-"typesafe" in "STLC_fix".
+"STLCFix", which prevents the proof of "progress" from depending on
+"typesafe" in "STLCFix".
 
@@
 families.
-In \cref{fig:stlc-isorec-prod}, family "STLC_fix_isorec" is declared as
-an "STLC" extension that mixes in "STLC_fix" and "STLC_isorec".
-As a result, \lsti{STLC_fix_isorec.typesafe} readily gives the
-type-safety theorem of STLC equipped with fixpoints and iso-recursive types.
+In \cref{fig:stlc-isorec-prod}, family "STLCFixIsorec" is declared as
+an "STLC" extension that mixes in "STLCFix" and "STLCIsorec".
+As a result, \lsti{STLCFixIsorec.typesafe} is readily the proof of the type-safety
+theorem of an STLC equipped with fixpoints and iso-recursive types.
 
@@
 also create an obligation to retrofit the mixins with new case handlers.
-In \cref{fig:stlc-isorec-prod}, family "STLC_prod_isorec" is composed
-of two mixins: "STLC_prod", which extends the inductive type "ty" with a
-new constructor "ty_prod", and "STLC_isorec", which introduces a
+In \cref{fig:stlc-isorec-prod}, family "STLCProdIsorec" is composed
+of two mixins: "STLCProd", which extends the inductive type "ty" with a
+new constructor "ty_prod", and "STLCIsorec", which introduces a
 function "tysubst" recursively defined on "ty".
 Hence, for exhaustivity, it is required that a composition of
-"STLC_prod" and "STLC_isorec" should additionally handle the "ty_prod"
+"STLCProd" and "STLCIsorec" should additionally handle the "ty_prod"
 case in "tysubst".
 
-\noindentparagraph{Tactics for eliminating inductive types.}
+\noindentparagraph{Tactics support for constructors.}
+
+Coq provides tactics for proving injectivity and disjointness of constructors
+(i.e., \lsti{injection} and \lsti{discriminate}).
+The proof terms generated by the tactics involve exhaustively
+matching on the constructors of an inductive type, so
+they do not work for extensible inductive types (\ref{chg:extensible-v-exhaustive}).
+%
+In principle, the programmer could use \lsti{FInduction} to prove
+injectivity and disjointness.
+But that would be tedious, and the programmer would have to revisit
+the induction proof every time the inductive type is extended.
+%
+Having to use \lsti{FInduction} is unsatisfying, also because
+injectivity and disjointness ought to hold
+regardless of whether inductive types are extensible.
+
+To provide a streamlined programming experience (\ref{chg:tooling}),
+our plugin offers two tactics, \lsti{finjection} and \lsti{fdiscriminate}.
+For example, in a proof state that contains a manifestly false assumption
+\lsti{H : tm_var "x" = tm_abs "y" t}, the programmer can use
+\lsti{fdiscriminate H} to obtain \lsti{False} and thus discharge
+the current goal, just as they would with \lsti{discriminate H} if "tm"
+were not extensible.
+
+\noindentparagraph{Partial recursors.}
+
+The tactics \lsti{finjection} and \lsti{fdiscriminate} are powered by
+\emph{partial recursors} that our plugin automatically generates for all
+inductive types defined with \lsti{FInductive}.
+
+As analyzed earlier, ordinary recursors, such as "tm_rec", are impossible
+within a family in which the name of the inductive type is late bound.
+However,
+a key observation is that extensible inductive types still admit a
+weakened form of recursors where the motive is an \lsti{option} type.
+For example, within family "STLC", the partial recursor for
+"tm" has the following type:
+
+\begin{centered}
+\begin{minipage}{.96\textwidth}
+\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+STLC_tm_prec : \forall (P : tm -> Set), [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] -> [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] ->
+               [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
+               [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
+               \forall t, option (P t)
+\end{lstlisting}
+\end{minipage}
+\end{centered}
+
+\noindent
+As expected, "STLC_tm_prec" is axiomatized along with
+four (propositional) equalities, one for each constructor of "tm".
+For instance, the equality for the "tm_abs" constructor is as follows:
+
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+\forall x t P H1 H2 H3 H4,
+STLC_tm_prec P H1 H2 H3 H4 (tm_abs x t) = H3 x t (STLC_tm_prec P H1 H2 H3 H4 t)
+\end{lstlisting}
+
+%\noindent
+Importantly, the partial recursor is compatible with the late binding of "tm" in its type.
+When "STLC_tm_prec" is inherited into family "STLCFix", all the previous four equalities
+still hold, and a trivial, fifth equality is made available:
+
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+\forall x t P H1 H2 H3 H4, STLC_tm_prec P H1 H2 H3 H4 (tm_fix x t) = None
+\end{lstlisting}
+
+Although partial recursors appear weaker than ordinary recursors,
+there is power in restraint:
+they offer a principled way to derive injectivity and disjointness of
+constructors, while allowing extensibility.
+Partial recursors allow for injective mappings from values of a late bound inductive
+type, like "tm", to values of an ordinary inductive type, like \lsti{\nat},
+the injectivity and disjointness of whose constructors are readily
+available.
+%
+For example, in a proof state with the assumption \lsti{H : tm_var "x" = tm_abs "y" t},
+running \lsti{fdiscriminate H} first uses \lsti{H} to obtain the equality
+
+\begin{centered}
+\begin{minipage}{\textwidth}
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+STLC_tm_prec (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_var "x") =
+STLC_tm_prec (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_abs "y" t)
+\end{lstlisting}
+\end{minipage}
+\end{centered}
+
+\noindent
+and then uses the axiomatized equalities about "STLC_tm_prec" to
+rewrite the equality above, obtaining \lsti{Some 1 = Some 4},
+from which \lsti{False} easily follows.
 

### 2022-10-12 stlc-isorec-prod.tex
partial recursor

--- 2022-10-09 stlc-isorec-prod.tex
@@
 
-\definecolor{codecomment-color}{HTML}{FF007F}
+\definecolor{codecomment-color}{HTML}{DD033F}
 
 \begin{lstlisting}
-Family STLC_isorec extends STLC.[@\tanc{STLCisorec-start}@]
+Family STLCIsorec extends STLC.[@\tanc{STLCIsorec-start}@]
 [@\hfill\codecomment{STLC extended with iso-recursive types}@]
@@
 
-[@\dadada@]
-
-End STLC_isorec.[@\tanc{STLCisorec-end}@]
+[@\dadada@][@\hfill\codecomment{Other adjustments}@]
+End STLCIsorec.[@\tanc{STLCIsorec-end}@]
 \end{lstlisting}
 
-\makeline[0pt]{STLCisorec-start}{STLCisorec-end}[codecomment-color!50]
+\makeline[0pt]{STLCIsorec-start}{STLCIsorec-end}[codecomment-color!50]
+
+\vspace{-12pt}
+
+\definecolor{codecomment-color}{HTML}{008B8B}
+
+\begin{lstlisting}
+Family STLCFixIsorec extends STLC[@\tanc{STLCIsorecfix-start}@]
+using STLCFix, STLCIsorec.
+[@\hfill\ \ \codecomment{STLC extended with fixpoints and iso-recursive types}@]
+End STLCFixIsorec.[@\tanc{STLCIsorecfix-end}@]
+\end{lstlisting}
+
+\makeline[0pt]{STLCIsorecfix-start}{STLCIsorecfix-end}[codecomment-color!50]
 
@@
 \begin{lstlisting}
-Family STLC_prod extends STLC.[@\tanc{STLCprod-start}@]
+Family STLCProd extends STLC.[@\tanc{STLCProd-start}@]
 [@\hfill\codecomment{STLC extended with products}@]
@@
 
-[@\dadada@]
-
-End STLC_prod.[@\tanc{STLCprod-end}@]
+[@\dadada@][@\hfill\codecomment{Other adjustments}@]
+End STLCProd.[@\tanc{STLCProd-end}@]
 \end{lstlisting}
 
-\makeline[.5\textwidth+9pt]{STLCprod-start}{STLCprod-end}[codecomment-color!50]
+\makeline[.5\textwidth+9pt]{STLCProd-start}{STLCProd-end}[codecomment-color!50]
 
-\vspace{-5pt}
+\vspace{-10pt}
 
@@
 \begin{lstlisting}
-Family STLC_isorec_prod extends STLC[@\tanc{STLCisorecprod-start}@]
-using STLC_prod, STLC_isorec.
-[@\hfill\ \ \codecomment{STLC extended with iso-recursive types and products}@]
+Family STLCProdIsorec extends STLC[@\tanc{STLCIsorecProd-start}@]
+using STLCProd, STLCIsorec.
+[@\hfill\ \ \codecomment{STLC extended with products and iso-recursive types}@]
 FRecursion tysubst on ty
@@
 
-End STLC_isorec_prod.[@\tanc{STLCisorecprod-end}@]
+End STLCProdIsorec.[@\tanc{STLCIsorecProd-end}@]
 \end{lstlisting}
 
-\makeline[.5\textwidth+9pt]{STLCisorecprod-start}{STLCisorecprod-end}[codecomment-color!50]
+\makeline[.5\textwidth+9pt]{STLCIsorecProd-start}{STLCIsorecProd-end}[codecomment-color!50]
+
+\vspace{-10pt}
+
+\definecolor{codecomment-color}{HTML}{708090}
+
+\begin{lstlisting}
+Family STLCFixProdIsorec extends STLC[@\tanc{STLCIsorecFixProd-start}@]
+using STLCFix, STLCProdIsorec.[@\hfill\codecommentstart{STLC extended}@]
+[@\hfill\codecommentend{with fixpoints, products, and iso-recursive types}@]
+End STLCFixProdIsorec.[@\tanc{STLCIsorecFixProd-end}@]
+\end{lstlisting}
+
+\makeline[.5\textwidth+9pt]{STLCIsorecFixProd-start}{STLCIsorecFixProd-end}[codecomment-color!50]
 
@@
 
+\vspace{-10pt}
+
 \caption{Composing extensions of STLC.}

### 2022-10-12 stlc-mechanized.tex
partial recursor

@@
 \begin{lstlisting}[firstnumber=77]
-Family STLC_fix extends STLC.[@\tanc{STLCfix-start}@]
+Family STLCFix extends STLC.[@\tanc{STLCFix-start}@]
 [@\hfill\codecomment{STLC extended with fixpoints}@]
@@
 
-End STLC_fix.[@\tanc{STLCfix-end}@]
+End STLCFix.[@\tanc{STLCFix-end}@]
 \end{lstlisting}
 
-\makeline[.5\textwidth+9pt]{STLCfix-start}{STLCfix-end}[codecomment-color!50]
+\makeline[.5\textwidth+9pt]{STLCFix-start}{STLCFix-end}[codecomment-color!50]
 
@@
 \begin{lstlisting}[firstnumber=111]
-Check STLC_fix.typesafe.[@\tanc{Check-STLCfix-typesafe}@]
+Check STLCFix.typesafe.[@\tanc{Check-STLCFix-typesafe}@]
 \end{lstlisting}
 
-\makeline[.5\textwidth+9pt]{Check-STLCfix-typesafe}{Check-STLCfix-typesafe}[codecomment-color!35]
+\makeline[.5\textwidth+9pt]{Check-STLCFix-typesafe}{Check-STLCFix-typesafe}[codecomment-color!35]
 

### 2022-10-12 ext-ind-type-compilation-example.tex
FIxing Diagram
Save. Fixing

@@
 
-\definecolor{codecomment-color}{HTML}{FF007F}
-
-\begin{lstlisting}
-Family STLC.[@\tanc{Ext-Ind-Cmp-start}@]
- FInductive tm : Set := tm_var : [@\dadada@]
- (* [@\dadada@] other constructors *) 
- FScheme tm_prec PRecT about self__STLC.tm.
- FScheme Comp about self__STLC.tm_prec.
- (* Generate Partial Recursor 
-    and the computational axiom for tm *)
- FRecursion subst on tm 
- motive (fun _ => id → tm → tm). 
-  Case tm_var := fun s x t => 
-    if (eqb x s) then t else (tm_var s).
-  (* [@\dadada@] other cases *)
- End subst.
-(* Field test := (subst (tm_var 0) 0 (tm_var 0)). *)
-(* [@\dadada@] other things in STLC *)
-EndFamily.
-Family STLC_bool extends STLC.
- Extend FInductive tm : Set :=
-  | tm_true : tm [@\dadada@]
-  (* [@\dadada@] other new constructors *)
- FRecursion subst.
-  Case tm_true := fun x t => tm_true.
-  (* [@\dadada@] other extended cases *)
- End subst. 
- Inherits subst.
-(* [@\dadada@] other things in STLC_Bool *)
-EndFamily.
-[@\tanc{Ext-Ind-Cmp-end}@]
-\end{lstlisting}
-
-\makeline[0pt]{Ext-Ind-Cmp-start}{Ext-Ind-Cmp-end}[codecomment-color!50]
-% \vspace{-3pt}
-
 \definecolor{codecomment-color}{HTML}{0DA3FF}
@@
 \begin{lstlisting}
-(* Compilation Of the above *) [@\tanc{Cmp-Result0-start}@]
-Module Type STLC_Interface. 
+(* **Type Checking of Inductive Type** *) [@\tanc{Cmp-Result0-start}@]
+Module Type STLC_Ctx1. 
+(* The Context including Inductive Type *)
   Parameter (tm : Set).
@@
   (* [@\dadada@] and all the fields for inductive type [@\dadada@] *)
-End STLC_Interface.
-(* Collect eliminator to Set for tm *)
-Module tm_rec_type (STLC: STLC_Interface).
+End STLC_Ctx1.
+Module Type tm_prec_64 (self: tm_6). (* Type for partial recursor *)
+Parameter
+  (tm_prec :
+  forall i : self.tm, forall P : self.tm -> Type,
+  (forall n, option (P (self.tm_var n))) ->
+  [@\dadada@], option (P i)).
+End tm_prec_64.
+Module Type STLC_Ctx2. (* The Context including partial recursor *) 
+  Include STLC_Ctx1.
+  Include tm_prec_64.
+End STLC_Ctx2.
+(* [@\dadada@] etc, for every context, (everytime a new field is added) we will construct a new context as module type *)
+(* Collect eliminator TYPE for later type-check *)
+Module tm_rec_type (self: STLC_Ctx1).
   Definition __recursor_type_tm_rec :=
-    forall P : STLC.tm -> Set,
-    (forall n : id, P (STLC.tm_var n)) ->
-    (forall (n : id) (i : STLC.tm),
-    P i -> P (STLC.tm_abs n i)) ->
-    [@\dadada@] -> forall i : STLC.tm, P i.
+    forall P : self.tm -> Set,
+    (forall n : id, P (self.tm_var n)) ->
+    (forall (n : id) (i : self.tm),
+    P i -> P (self.tm_abs n i)) ->
+    [@\dadada@] -> forall i : self.tm, P i.
 End tm_rec_type.  
-Module tm_prec_72 (self__STLC: STLC_Interface).
-Definition tm_prec :
-  forall i : tm, forall P : tm -> Type,
-  (forall n, option (P (tm_var n))) ->
-  [@\dadada@], option (P __i) :=
-End tm_prec_72. 
-Module Type substtm_139 (self__STLC: [@\dadada@]).
-Parameter
-  (subst_on_tm_unit :
-	 eq (self__STLC.subst (self__STLC.tm_unit))
-       self__STLC.subst_internal.tm_unit).
-(* [@\dadada@] subst equality on other constructors *)
-End substtm_139. 
+(* **Type checking of Recursion** *)
+(* Typecheck each handler, here STLC.subst.tm_var *)
+(* Also an example of fields in nested family, thus with two contexts *) 
+Module tm_var_handle (self__STLC: STLC_Ctx1)
+    (self__subst_internal: EmptySig_84).
+  Definition tm_var :
+    forall (s : ident) (x : ident) 
+    (t : self__STLC.tm), self__STLC.tm :=
+    fun s x t => if eqb x s
+       then t else self__STLC.tm_var s.
+End tm_var_handle. (* [@\dadada@] and more *)
+(* Assembling tycked handlers in STLC.subst during compilation *)
+Module subst_internal_410 (STLC: STLC_Ctx1). 
+ Module subst_internal.
+ Include tm_var_handle STLC. Include tm_abs_handle STLC.
+ Include tm_app_handle STLC. Include tm_unit_handle STLC.
+ End subst_internal.
+End subst_internal_410.
+(* Intermediate Module solely for exhausitivity checking *) 
+Module do_tc (self: STLC_Ctx[@\dadada@]).
+  Include tm_rec_type self.
+  Parameter (recursor_for_type_checking 
+    : __recursor_type_tm_rec).
+  Definition term_for_type_checking :=
+    recursor_for_type_checking 
+      (fun _  ⇒ id → self.tm → self.tm)
+      subst_internal_410.tm_var [@\dadada@] .
+End do_tc.
 [@\tanc{Cmp-Result0-end}@]
@@
 \begin{lstlisting}
-(* Compiled Field STLC.subst_internal.tm_var *) [@\tanc{Cmp-Result-start}@]
-Module tm_var_handle (self__STLC: STLC_Interface)
-    (self__subst_internal: EmptySig_84).
-  Definition tm_var :
-    forall (s : ident) (x : ident) 
-    (t : self__STLC.tm), self__STLC.tm :=
-    fun s x t => if eqb x s
-       then t else self__STLC.tm_var s.
-End tm_var_handle. (* [@\dadada@] and more *)
-(* Assembling STLC.subst_internal when compilation *)
-Module subst_internal_410 (STLC: STLC_Interface). 
- Module subst_internal.
- Include tm_var_handle STLC. Include tm_abs_handle STLC.
- Include tm_app_handle STLC. Include tm_unit_handle STLC.
- End subst_internal.
-End subst_internal_410.
-(* Intermediate Module solely for type checking *)
-Module do_tc (STLC: STLC_Interface).
-  Include tm_rec_type STLC.
-  Parameter (recursor_for_type_checking 
-    : __recursor_type_tm_rec).
-  Definition term_for_type_checking :=
-    recursor_for_type_checking 
-      (fun _  ⇒ id → STLC.tm → STLC.tm)
-      subst_internal_410.tm_var [@\dadada@] .
-End do_tc.
+[@\tanc{Cmp-Result-start}@]
+(* Exposing Computational Axiom into Ctx *)
+Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).
+Parameter
+  (subst_on_tm_unit :
+	 (self.subst (self.tm_unit)) = 
+      self.subst_internal.tm_unit).
+(* [@\dadada@] subst equality on other constructors *)
+End substtm_139.
+
+(* Compilation, only when closing a top-level family *)
 (* Ctx that expose concrete inductive type *)
@@
 End ctx_subst_113.
-(* Compilation of recursor, not reusable
-    when inductive type extended due to the ctx *)
-Module subst_128 (self__STLC: ctx_subst_113).
-Include __motive_of_subst_107 self__STLC.
+(* Compilation of recursor, 
+  require concrete inductive type in the ctx *)
+Module subst_128 (self: ctx_subst_113).
+Include __motive_of_subst_107 self.
 Definition subst :=
-  self__STLC.__internal_tm_rec __motiveTsubst
-	self__STLC.subst_internal.tm_var [@\dadada@] .
+  self.__internal_tm_rec __motiveTsubst
+	self.subst_internal.tm_var [@\dadada@] .
 End subst_128.
-(* Assembling the whole STLC *)
+(* Compilation of Partial Recursor *)
+Module tm_prec_72 (self: STLC_Ctx1).
+Definition tm_prec :
+  forall i : self.tm, forall P : self.tm -> Type,
+  (forall n, option (P (self.tm_var n))) ->
+  [@\dadada@], option (P i) :=
+End tm_prec_72. 
+(* Assembling/Compiling the whole STLC *)
 Module STLC.
@@
     most of the former compilation results *)
-Module subst_internal_248 (STLC_bool: STLC_Interface).
+Module subst_internal_248 (STLC_bool: STLC_Ctx1).
 Include tm_true_197 STLC_bool. Include [@\dadada@]

### 2022-10-12 ext-ind-type-compilation-example.tex
Tidy up the example, with one new ctx.

@@
 \begin{lstlisting}
-(* **Type Checking of Inductive Type** *) [@\tanc{Cmp-Result0-start}@]
+(* Type Checking of Inductive Type *) [@\tanc{Tyck-IndTy-start}@]
 Module Type STLC_Ctx1. 
@@
 End STLC_Ctx1.
-Module Type tm_prec_64 (self: tm_6). (* Type for partial recursor *)
-Parameter
-  (tm_prec :
-  forall i : self.tm, forall P : self.tm -> Type,
-  (forall n, option (P (self.tm_var n))) ->
-  [@\dadada@], option (P i)).
+Module Type tm_prec_64 (self: tm_6). 
+Parameter (* Type for partial recursor *)
+(tm_prec : ∀ t, ∀ P : self.tm -> Type,
+  (∀ n, option (P (self.tm_var n))) -> 
+  [@\dadada@], option (P t)).
 End tm_prec_64.
-Module Type STLC_Ctx2. (* The Context including partial recursor *) 
+Module Type STLC_Ctx2.  
   Include STLC_Ctx1.
-  Include tm_prec_64.
+  Include tm_prec_64. (* New Ctx include prec *)
 End STLC_Ctx2.
-(* [@\dadada@] etc, for every context, (everytime a new field is added) we will construct a new context as module type *)
+(* [@\dadada@] etc, everytime a new field is added, a new 
+module type is created as the ctx with the new field *)
 (* Collect eliminator TYPE for later type-check *)
 Module tm_rec_type (self: STLC_Ctx1).
-  Definition __recursor_type_tm_rec :=
-    forall P : self.tm -> Set,
-    (forall n : id, P (self.tm_var n)) ->
-    (forall (n : id) (i : self.tm),
-    P i -> P (self.tm_abs n i)) ->
-    [@\dadada@] -> forall i : self.tm, P i.
-End tm_rec_type.  
-(* **Type checking of Recursion** *)
+  Definition _TYPE_tm_rec :=
+    ∀ P : self.tm -> Set,
+    ( ∀ n, P (self.tm_var n)) ->
+    ( ∀ n i, P i -> P (self.tm_abs n i)) ->
+    [@\dadada@] -> ∀ i : self.tm, P i.
+End tm_rec_type.  [@\tanc{Tyck-IndTy-end}@]
+\end{lstlisting}
+
+\makeline[0pt]{Tyck-IndTy-start}{Tyck-IndTy-end}[codecomment-color!50]
+
+\definecolor{codecomment-color}{HTML}{0DA30F}
+
+\begin{lstlisting}
+(* Type checking of Recursion *) [@\tanc{Tyck-RecTy-start}@]
 (* Typecheck each handler, here STLC.subst.tm_var *)
-(* Also an example of fields in nested family, thus with two contexts *) 
-Module tm_var_handle (self__STLC: STLC_Ctx1)
-    (self__subst_internal: EmptySig_84).
-  Definition tm_var :
-    forall (s : ident) (x : ident) 
-    (t : self__STLC.tm), self__STLC.tm :=
-    fun s x t => if eqb x s
-       then t else self__STLC.tm_var s.
-End tm_var_handle. (* [@\dadada@] and more *)
-(* Assembling tycked handlers in STLC.subst during compilation *)
-Module subst_internal_410 (STLC: STLC_Ctx1). 
+(* Also an example of fields in nested family [@\dadada@]*) 
+Module tm_var_handle (STLC: STLC_Ctx1)
+    (self: EmptySig). (* [@\dadada@] thus with two contexts *) 
+  Definition tm_var s x t := 
+    if eqb x s then t else STLC.tm_var s.
+End tm_var_handle. (* [@\dadada@] other handlers are alike *)
+(* Assembling tycked handlers in STLC.subst *)
+Module subst_internal_410 (M: STLC_Ctx1). 
  Module subst_internal.
- Include tm_var_handle STLC. Include tm_abs_handle STLC.
- Include tm_app_handle STLC. Include tm_unit_handle STLC.
+ Include tm_var_handle M. Include tm_abs_handle M.
+ Include tm_app_handle M. Include tm_unit_handle M.
  End subst_internal.
@@
   Include tm_rec_type self. 
-  Parameter (recursor_for_type_checking 
-    : __recursor_type_tm_rec).
+  Include subst_internal_410 self.
+  Parameter (for_tyck : _TYPE_tm_rec).
   Definition term_for_type_checking :=
-    recursor_for_type_checking 
-      (fun _  ⇒ id → self.tm → self.tm)
-      subst_internal_410.tm_var [@\dadada@] .
-End do_tc.
-[@\tanc{Cmp-Result0-end}@]
+    for_tyck  ( λ _  ⇒ (id → self.tm → self.tm))
+      subst_internal.tm_var [@\dadada@] .
+End do_tc.[@\tanc{Tyck-RecTy-end}@]
 \end{lstlisting}
 
-\makeline[0pt]{Cmp-Result0-start}{Cmp-Result0-end}[codecomment-color!50]
+\makeline[0pt]{Tyck-RecTy-start}{Tyck-RecTy-end}[codecomment-color!50]
 
@@
 
-\definecolor{codecomment-color}{HTML}{0DA3FF}
-
 \begin{lstlisting}
-[@\tanc{Cmp-Result-start}@]
-(* Exposing Computational Axiom into Ctx *)
+(* Exposing Computational Axiom into Ctx *) [@\tanc{Tyck-Recty-cont-start}@]
 Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).
@@
 (* [@\dadada@] subst equality on other constructors *)
-End substtm_139.
+End substtm_139. [@\tanc{Tyck-Recty-cont-end}@]
+\end{lstlisting}
 
-(* Compilation, only when closing a top-level family *)
+\makeline[.5\textwidth+9pt]{Tyck-Recty-cont-start}{Tyck-Recty-cont-end}[codecomment-color!50]
+
+\definecolor{codecomment-color}{HTML}{5D030F}
+
+\begin{lstlisting}
+(* Compilation, only when closing a top-level family *) [@\tanc{Cmp-Result-start}@]
 (* Ctx that expose concrete inductive type *)
-Module Type ctx_subst_113.
+Module Type STLC_CCtx1.
 Inductive __internal_tm : Set := [@\dadada@] 
 Include subst_internal_410. (* [@\dadada@] and more inclusion *)
-End ctx_subst_113.
-(* Compilation of recursor, 
-  require concrete inductive type in the ctx *)
-Module subst_128 (self: ctx_subst_113).
+End STLC_CCtx.
+(* The compilation of recursor and[@\dadada@] *)
+Module subst_128 (self: STLC_CCtx1).
 Include __motive_of_subst_107 self.
@@
 End subst_128.
-(* Compilation of Partial Recursor *)
-Module tm_prec_72 (self: STLC_Ctx1).
+(* [@\dadada@] the compilation of partial recursor, 
+both require concrete inductive type in the ctx *)
+Module tm_prec_72 (self: STLC_CCtx1).
 Definition tm_prec :
-  forall i : self.tm, forall P : self.tm -> Type,
-  (forall n, option (P (self.tm_var n))) ->
-  [@\dadada@], option (P i) :=
+  ∀ t, ∀ P : self.tm -> Type,
+  (∀ n, option (P (self.tm_var n))) ->
+  [@\dadada@], option (P t) := [@\dadada@]
 End tm_prec_72. 
-(* Assembling/Compiling the whole STLC *)
-Module STLC.
-Include tm_5. Include tm_prec_72. Include tm_prectm_81.
+Module STLC. (* Assembling/Compiling the whole STLC *)
+Include tm_5. Include tm_prec_72. 
+Include tm_prectm_81. (* Justification for comp axiom *)
 Include subst_internal_410. Include subst_128.
@@
 End v_248. 
-(* Regenerated FRecursion result *)
-Module subst_242 (self__STLC_bool: [@\dadada@]).
-Include __motive_of_subst_107 self__STLC_bool.
+(* Regenerated FRecursion Compilation *)
+Module subst_242 (self: [@\dadada@]).
+Include __motive_of_subst_107 self.
 Definition subst :=
-  self__STLC_bool.__internal_tm_rec __motiveTsubst
-	self__STLC_bool.subst_internal.tm_var [@\dadada@]
-    self__STLC_bool.subst_internal.tm_true [@\dadada@] .
+  self.__internal_tm_rec __motiveTsubst
+	self.subst_internal.tm_var [@\dadada@]
+    self.subst_internal.tm_true [@\dadada@] .
 End subst_242. [@\tanc{Cmp-Result-end}@]
@@
 
-\caption{Exemplar Compilation about Inductive Type}\label{fig:plugin-example2}
+\caption{Exemplar Compilation about Inductive Type, for STLC example}\label{fig:plugin-example2}
 \end{figure}

### 2022-10-12 ext-ind-type-compilation-example.tex
Small Fixes.

@@
 (* Type Checking of Inductive Type *) [@\tanc{Tyck-IndTy-start}@]
+(* Tyck happens during interactively creating families*)
 Module Type STLC_Ctx1. 
@@
 \begin{lstlisting}
-(* Compilation, only when closing a top-level family *) [@\tanc{Cmp-Result-start}@]
+(* Compilation, happens when closing a top-level family *) [@\tanc{Cmp-Result-start}@]
 (* Ctx that expose concrete inductive type *)
-Module Type STLC_CCtx1.
+Module Type STLC_CCtx.
 Inductive __internal_tm : Set := [@\dadada@] 
@@
 End STLC_CCtx.
-(* The compilation of recursor and[@\dadada@] *)
-Module subst_128 (self: STLC_CCtx1).
+(* Both the compilation of recursor and[@\dadada@] *)
+Module subst_128 (self: STLC_CCtx).
 Include __motive_of_subst_107 self.
@@
 End subst_128.
-(* [@\dadada@] the compilation of partial recursor, 
-both require concrete inductive type in the ctx *)
-Module tm_prec_72 (self: STLC_CCtx1).
+(* [@\dadada@]the compilation of partial recursor 
+require concrete inductive type exposed in the ctx *)
+Module tm_prec_72 (self: STLC_CCtx[@\dadada@]).
 Definition tm_prec :
@@
 Include tm_5. Include tm_prec_72. 
-Include tm_prectm_81. (* Justification for comp axiom *)
+Include tm_prectm_81. (*Justify comp axiom, mostly eq_refl*)
 Include subst_internal_410. Include subst_128.
@@
     most of the former compilation results *)
-Module subst_internal_248 (STLC_bool: STLC_Ctx1).
-Include tm_true_197 STLC_bool. Include [@\dadada@]
+(* Notice how handlers don't require explicitly
+   exposing the inductive type *)
+Module subst_internal_248 (self: STLC_Ctx1).
+(* The New Handlers *)
+Include tm_true_197 self. Include [@\dadada@]
 (* The Reused Compilation result *)
-Include tm_var_handle STLC_bool. Include [@\dadada@]
+Include tm_var_handle self. Include [@\dadada@]
 End v_248. 

### 2022-10-12 syntax.tex
Mention how partial recursor is provable in plugin and thus trivially modelled by Metatheory

@@
 
+% With sealing on hand, our metatheory is directly corresponding to our Coq's implementation.
+
+Partial recursor and computational axiom is modelled like any other recursors and theorems.\footnote{This is partially because the partial recursor is provable in our Plugin.} Recall that the sealed context is corresponding to a module type.
+Partial recursor is constructed directly using \ruleref{Rec-Constr} like any other recursors in our metatheory.
+With trivial sealing, and with concrete inductive type and partial recursor exposed in the context, we will have judgemental equality for beta-equality of the partial recursor, to prove the computational axiom.
+The only difference is that, partial recursor doesn't support user extension in plugin implementation but our encoding supports that.
+
 Next we talk about inheritance judgements. Note that, there will be a lot

### 2022-10-12 syntax.tex
typo fix.

@@
 
-Partial recursor and computational axiom is modelled like any other recursors and theorems.\footnote{This is partially because the partial recursor is provable in our Plugin.} Recall that the sealed context is corresponding to a module type.
+Partial recursor and computational axiom is modelled like any other recursors and theorems.\footnote{This is partially because the partial recursor is provable in our plugin.} Recall that the sealed context is corresponding to a module type.
 Partial recursor is constructed directly using \ruleref{Rec-Constr} like any other recursors in our metatheory.

### 2022-10-12 lang-design.tex
stuff

@@
 For example, in a derived family, the programmer may want to override
-the proof of a lemma with a different proof strategy than it was proved
+the proof of a lemma, using a different proof strategy than it was proved
 in the base family.
 %
-Overriding is potentially incompatible with having equalities on late bound names.
+Overriding is potentially incompatible with having equalities on late bound names, however.
 
@@
 injectivity and disjointness.
-But that would be tedious, and the programmer would have to revisit
-the induction proof every time the inductive type is extended.
-%
-Having to use \lsti{FInduction} is unsatisfying, also because
-injectivity and disjointness ought to hold
-regardless of whether inductive types are extensible.
+But this workaround is unsatisfying:
+it is tedious, it forces the programmer to revisit
+the induction proofs every time an inductive type is extended,
+and above all, why should a property like
+\lsti{tm_var "x" = tm_abs "y" t -> False} have anything to do with "tm_fix"?
 
@@
 
-The tactics \lsti{finjection} and \lsti{fdiscriminate} are powered by
-\emph{partial recursors} that our plugin automatically generates for all
-inductive types defined with \lsti{FInductive}.
+We make the observation
+that injectivity and disjointness of existing constructors ought to hold
+regardless of future addition of constructors.
+%
+This insight motivates the design of \emph{partial recursors}, which
+power the \lsti{finjection} and \lsti{fdiscriminate} tactics.
+Partial recursors are automatically generated for all inductive types
+defined with \lsti{FInductive}.
 
@@
 a key observation is that extensible inductive types still admit a
-weakened form of recursors where the motive is an \lsti{option} type.
+weakened elimination principle where the motive is an \lsti{option} type.
 For example, within family "STLC", the partial recursor for
@@
 %\noindent
-Importantly, the partial recursor is compatible with the late binding of "tm" in its type.
+Importantly, unlike the standard "tm_rec" recursor, the partial recursor
+"STLC_tm_prec" is compatible with the late binding of "tm" in its type.
 When "STLC_tm_prec" is inherited into family "STLCFix", all the previous four equalities
@@
 Although partial recursors appear weaker than ordinary recursors,
-there is power in restraint:
-they offer a principled way to derive injectivity and disjointness of
-constructors, while allowing extensibility.
-Partial recursors allow for injective mappings from values of a late bound inductive
+there is power in restraint.
+Partial recursors offer a principled, uniform way to derive injectivity and disjointness of
+constructors, while allowing extensibility:
+they enable injective mappings from values of a late bound inductive
 type, like "tm", to values of an ordinary inductive type, like \lsti{\nat},
@@
 from which \lsti{False} easily follows.
+
+Note that the proof term generated by \lsti{fdiscriminate H} in family "STLC"
+is reusable by family "STLCFix", because the partial recursor and its
+equalities are compatible with the late binding of "tm".
+Also note that within family "STLCFix", a second partial recursor
+(called "STLCFix_tm_prec") on "tm" is available;
+it can be used to derive properties about "tm_fix",
+such as \lsti{tm_var "x" = tm_fix "y" t -> False}.
 

### 2022-10-13 metatheory-condensed.tex
Start Reworking Metatheory QIIT possibly removable altogether

To formalize our metatheory, we follow the formulation of~\citet{altkap2016}, which provides a declarative and intrinsic style of the predicative \textit{Martin-Lof Type Theory} (MLTT)~\cite{martin1982constructive},
using \textbf{Quotient Inductive Inductive Type}(QIIT)\footnote{Inductive Inductive Type(IIT) is a generalization of mutual inductive type where (indexed-)types \mintinline{Coq}{A : Type, B : A -> Type} can be defined mutually. QIIT further enhances IIT by allowing equality constructor (i.e. mathematical quotient).}, \textbf{explicit substitution} and \textbf{debruijn indices} as the main tool. 
Generally speaking, we \textbf{use each QIIT type to represent each kind of judgements, and thus a term will represent a derivation of a judgement}. The classical formulation of dependent type theory requires a lot of \textit{presupposition} and quotient afterwards, causing a lot of duplicacy. Both issues can be concisely solved by using QIIT expressing intrinsically typed syntax instead. What's more, working \textbf{in a type-theoretical setting} using QIIT makes checking of the pen-and-paper proof easier. However, we will follow conventional notation and thus the reader might not notice the usage of QIIT explicitly.\EDJ{We might be able to remove QIIT in the main text totally. Let's try to do it later.}

% Explicit substitution
Instead of meta-level substitution, we use explicit substitution, i.e., substitution itself is a expression in the program and will be done in the runtime. Thus substitution has its own typing judgement.
% Debruijn Indices
Contrary to named variables, debruijn indices use number to indicate the variable in the context. For example, \mintinline{Agda}{}

Now, we will mainly focus on the new typing rules that we introduced. We refer the those curious about the complete formulation to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}. Our metatheory is in an extensional setting.

% linkage, signature

% Inductive type and recursor and handlers

% Inhertance

### 2022-10-13 ext-ind-type-compilation-example.tex
Questions in compilation figure

@@
 End STLC_Ctx1.
-Module Type tm_prec_64 (self: tm_6). 
+Module Type tm_prec_64 (self: tm_6).[@\YZ{tm_6? Do you mean STLC_Ctx1 instead?}@]
 Parameter (* Type for partial recursor *)
@@
 (* Also an example of fields in nested family [@\dadada@]*) 
-Module tm_var_handle (STLC: STLC_Ctx1)
-    (self: EmptySig). (* [@\dadada@] thus with two contexts *) 
+Module tm_var_handle (STLC: STLC_Ctx1)[@\YZ{(1) Why doesn't the module name contain "subst"? (2) Should it be STLC_Ctx2 instead of STLC_Ctx1?}@]
+    (self: EmptySig). (* [@\dadada@] thus with two contexts *)[@\YZ{(3) Are nested families really needed?}@]
   Definition tm_var s x t := 
@@
 (* Exposing Computational Axiom into Ctx *) [@\tanc{Tyck-Recty-cont-start}@]
-Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).
+Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).[@\YZ{What is "..."?}@]
 Parameter
   (subst_on_tm_unit :
-	 (self.subst (self.tm_unit)) = 
+	 (self.subst (self.tm_unit)) =[@\YZ{Where is this "subst" added to the context?}@]
       self.subst_internal.tm_unit).

### 2022-10-13 ext-ind-type-compilation-example.tex
Reply the comment.

@@
 End STLC_Ctx1.
-Module Type tm_prec_64 (self: tm_6).[@\YZ{tm_6? Do you mean STLC_Ctx1 instead?}@]
+Module Type tm_prec_64 (self: STLC_Ctx1).[@\YZ{tm_6? Do you mean STLC_Ctx1 instead?}\EDJreply{Yes Thanks.}@]
 Parameter (* Type for partial recursor *)
@@
 (* Also an example of fields in nested family [@\dadada@]*) 
-Module tm_var_handle (STLC: STLC_Ctx1)[@\YZ{(1) Why doesn't the module name contain "subst"? (2) Should it be STLC_Ctx2 instead of STLC_Ctx1?}@]
-    (self: EmptySig). (* [@\dadada@] thus with two contexts *)[@\YZ{(3) Are nested families really needed?}@]
+Module tm_var_handle (STLC: STLC_Ctx1)[@\YZ{(1) Why doesn't the module name contain "subst"?
+(2) Should it be STLC_Ctx2 instead of STLC_Ctx1?}\EDJreply{I rename to ctx so that I think it is easier to understand.
+Feel Free to change to the way you want.
+I think name them all into "_Ctx" suffix is easier to emphasize the fact they are the result of compilation of the context.
+(2) The fact is, no, it also will need to include tm_prec's computational axiom and ty, which I also forgot to mention.
+Also ty has its prec and computational axiom.
+So at least 6.
+Do you want me to write doesn STLC_Ctx(1,2,3,4,5,6)?
+It is unnecessary complication and the space is not enough.
+I proposed to move this part of information into the figure 5 toy example in the slack, what do you think?
+}@]
+    (self: EmptySig).
+(* [@\dadada@] thus with two contexts *)[@\YZ{(3) Are nested families really needed?}\EDJreply{Not really but if consider handlers as one family, it is nested family.
+What's more, the audience needs to know how nested family is compiled at some point.}@]
   Definition tm_var s x t := 
@@
 (* Exposing Computational Axiom into Ctx *) [@\tanc{Tyck-Recty-cont-start}@]
-Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).[@\YZ{What is "..."?}@]
+Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).[@\YZ{What is "..."?}\EDJreply{My intension is STLC_Ctx_n for some n where at this point the audience should able to see what the context of that STLC_Ctx is.
+Apparently I failed.}@]
 Parameter
   (subst_on_tm_unit :
-	 (self.subst (self.tm_unit)) =[@\YZ{Where is this "subst" added to the context?}@]
+	 (self.subst (self.tm_unit)) =[@\YZ{Where is this "subst" added to the context?}\EDJreply{The reason is above. Currently the space is already not enough.}@]
       self.subst_internal.tm_unit).

### 2022-10-13 ext-ind-type-compilation-example.tex
Fix Module tm_var_handle.

@@
 (* Also an example of fields in nested family [@\dadada@]*) 
-Module tm_var_handle (STLC: STLC_Ctx1)[@\YZ{(1) Why doesn't the module name contain "subst"?
+Module tm_var_handle (STLC: STLC_Ctx1).[@\YZ{(1) Why doesn't the module name contain "subst"?
 (2) Should it be STLC_Ctx2 instead of STLC_Ctx1?}\EDJreply{I rename to ctx so that I think it is easier to understand.
@@
 }@]
-    (self: EmptySig).
-(* [@\dadada@] thus with two contexts *)[@\YZ{(3) Are nested families really needed?}\EDJreply{Not really but if consider handlers as one family, it is nested family.
+[@\YZ{(3) Are nested families really needed?}\EDJreply{Not really but if consider handlers as one family, it is nested family.
 What's more, the audience needs to know how nested family is compiled at some point.}@]

### 2022-10-13 metatheory-condensed.tex
Save.
Metatheory Condensing.

@@
 
+$$
+\Rule[name=Tm Subst]
+{\goodType{\Delta}{T}{j} 
+  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+{\goodType{\Gamma}{T[\gamma]}{j}},
+\Rule[]
+{\goodTerm{\Delta}{t}{T}
+  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
+{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
+\quad
+\Rule[name=Dbj Indices]
+{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
+{\goodSub{\Gamma}{\pi_1 \gamma}{\Delta}},
+\Rule[]
+{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
+{\goodSub{\Gamma}{\pi_2 \gamma}{A[\pi_1 \gamma]}}
+$$
+
 % Explicit substitution
-Instead of meta-level substitution, we use explicit substitution, i.e., substitution itself is a expression in the program and will be done in the runtime.
-Thus substitution has its own typing judgement.
+Instead of meta-level substitution, we use \textit{explicit substitution}, i.e., substitution itself is a expression in the program and will be done in the runtime.
+Thus substitution has its own judgement $\goodSub{\Gamma}{\gamma}{\Delta}$.
+We can apply substitution to type and terms in MLTT.
+We also have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
 % Debruijn Indices
-Contrary to named variables, debruijn indices use number to indicate the variable in the context. For example, \mintinline{Agda}{}
+Contrary to named variables, \textit{debruijn indices} refer to the variable in the context like operating a stack.
+For example, we can get the very first variable in the context $\Gamma, x : T \vdash x : T$ by $\goodTerm{\Gamma, T}{\pi_2 "id"}{T[\pi_1"id"]}$.
+Similarly, to get the second and above variables, we only need to apply a sequence of $\pi_1$s.
 
 Now, we will mainly focus on the new typing rules that we introduced.
-We refer the those curious about the complete formulation to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
+For those curious about the complete formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
 Our metatheory is in an extensional setting.
 
-% linkage, signature
+% linkage, signature and compilation and seal
+$$
+\Rule[name=Ept Sig]
+{}
+{\goodSig{\Gamma}{\nu\cdot}{0}}\quad
+\Rule[name=Lnkg Type/Compile]
+{\goodSig{\Gamma}{\sigma}{n}}
+{\goodType{\Gamma}{\cL \sigma}{i}
+\quad \goodType{\Gamma}{\cC \sigma}{}}
+\quad
+\Rule[name=Lnkg Compile]
+{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+}
+{
+  \goodTerm{\Gamma}{\cCt o}{\cC \sigma}
+}
+$$
+$$
+\Rule[name=Sig Add]
+{\goodSig{\Gamma}{\sigma}{n} 
+ \quad \goodType{\Gamma}{A}{}
+ \quad \goodSeal{\Gamma}{f}{\sigma}{A}
+ \quad \goodType{\Gamma, A}{T}{i}}
+{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
+\quad
+\Biggl( \boxed{\goodSeal{\Gamma}{s}{\sigma}{A}} 
+:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}} \Biggr)
+$$
+$$
+\Rule[name=Ept Lnkg]
+{}
+{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
+\quad
+\Rule[name=Lnkg Add]
+{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+\quad \goodType{\Gamma}{A}{}
+\quad  {\goodSeal{\Gamma}{f}{\sigma}{A}}
+ \quad \boxed{\goodTerm{\Gamma, A}{t}{T}}
+}
+{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
+$$
+We start with signature (with a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
+A signature is basically a list of types used to indicate the type of \textit{linkages}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
+Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement and will be explained later.
+
+Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited the linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
+The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
+For example, $\cC $
+
+The difference between compilation type and linkage types lies in their introduction rules.
+The compilation types are just syntactic sugar for sigma types.
+However, the intro rules for the linkage types gives it the power of late-binding (also the fundation of inheritance and overriding).
+More concretely, the boxed premise in \ruleref{Lnkg Add} is making $t$ parametric on $A$, contrary to how in sigma type, the $t$ will be about a concrete $a : A$.
+This parametricity makes $t$ inheritable and the earlier stuff about $A$ late-bound.
+
+The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$.
+For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators).
+Then any later fields $\goodTerm{\Gamma, A}{t}{T}$ relies only on the \textit{interface} $A$ can be inherited to other contexts (e.g.
+using extended inductive type), as long as the targeting contexts can be abstracted into the surrounding \textit{interface} for $t$ (i.e.
+$\Gamma, A$)  as well.
+
+We also show one example illustrating the importance of the compilation type.
 
@@
 
-% Inhertance
+$$
+\Rule[]
+{}
+{\goodWSig{\Gamma}{w\cdot}{0}}
+\quad
+\Rule[name=WSig-Add]
+{\goodWSig{\Gamma}{\tau}{n}
+  \quad \goodType{\Gamma}{A}{i}
+  \quad \goodType{\Gamma, A}{B}{i}}
+{\goodWSig{\Gamma}{w^+ \  \tau \  A \  B}{n+1}}
+\quad
+\Rule[name=Ind-Univ]
+{\goodWSig{\Gamma}{\tau}{n}}
+{\goodType{\Gamma}{\bW \tau}{i}
+}
+$$
+$$
+\Rule[name=Ind-Term]
+{\goodTerm{\Gamma}{T}{\bW \tau}
+  \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
+  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)}}
+{\goodTerm{\Gamma}{"Wsup"~T~a~b}{El~("W"~T)} }
+$$
+$$
+\Rule[name=Hdler]
+{\goodType{\Gamma}{A}{}
+\quad \goodType{\Gamma, A}{B}{}
+\quad \goodType{\Gamma}{R}{}
+}
+{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
+\quad 
+\Rule[name=Hdlers]
+{\goodWSig{\Gamma}{\tau}{n}
+\quad \goodType{\Gamma}{R}{}
+}
+{\goodSig{\Gamma}{"RecSig"~\tau~R}{n}}
+$$
+$$
+\Rule[name=Hdler-Proj]
+{ \goodWSig{\Gamma}{\tau}{N}
+\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\tau~R)}
+\quad j < N
+}
+{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)[\pi_1]}}
+$$
+$$
+\Rule[name=Rec-Constr]
+{ \goodWSig{\Gamma}{\tau}{N}
+\quad \goodTerm{\Gamma}{T}{\bW \tau}
+\quad \goodType{\Gamma}{R}{}
+\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
+}
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}\YZ{
+(1) Use \tau for inductive types and \sigma for linkage signatures.
+(2) Try using the same metavariables (j and n) as in the rules on W-types.
+(3) In Rec-Constr, no checking of \Gamma \vdash T : \bW \tau?
+(4) Why are RecSig1, RecSig, Wrec, prjR not listed in the syntax figure?
+}\EDJreply{Thx. Working on it. I seem to forgot to add inheritance as well.}\EDJreply{Done. Please Check.}
+$$
+
+Our formulation of (extensible) inductive type is close to Wtype (\ruleref{WSig-Add} requires a pair $A \vdash B$ as well), but with mulitple constructors.
+Thus we have the signature of inductive type collecting each constructors, with a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$.
+Unconventionally, a signature can induce a "universe" of types $\bW$: in other words, $T : \bW$ will be the inductive type instead of $\bW$ itself.
+"Wsup" is used to construct a specific term of an given inductive type.
+
+Similar to plugin implementation, we delegate handlers for recursion using linkages.
+We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers.
+Specifying the handler family using "Wrec" on target inductive term $w$ leads to elimination.
+
+% Inhertance judgement

### 2022-10-13 metatheory-condensed.tex
Managed to Shrink Linkage + Ind Type into 1 page. Please Check.

@@
 % Explicit substitution
-Instead of meta-level substitution, we use \textit{explicit substitution}, i.e., substitution itself is a expression in the program and will be done in the runtime.
+Instead of meta-level substitution, we use \textbf{explicit substitution}, i.e., substitution itself is a expression in the program and will be done in the runtime.
 Thus substitution has its own judgement $\goodSub{\Gamma}{\gamma}{\Delta}$.
@@
 % Debruijn Indices
-Contrary to named variables, \textit{debruijn indices} refer to the variable in the context like operating a stack.
+Contrary to named variables, \textbf{debruijn indices} refer to the variable in the context like operating a stack.
 For example, we can get the very first variable in the context $\Gamma, x : T \vdash x : T$ by $\goodTerm{\Gamma, T}{\pi_2 "id"}{T[\pi_1"id"]}$.
@@
 We also show one example illustrating the importance of the compilation type.
+
+Looking closely, the sealing judgement helps to model most machinaries about abstraction in plugin implementation.
 
@@
 % Inhertance judgement
+In the plugin implementation, "Family"s are second class objects like Coq's "Module".
+What's more, a "Family", with or without "extends" argument can be considered as a standalone piece of code trying to \textit{inherit} an existent (but maybe empty) family.
+Thus we model them using \textit{inheritance judgement} instead of considering them as first class terms.
+
+In this way, we will consider the program in our metatheory similar to those vanilla MLTT program, except there are meta variables denoting inheirtance judgement, of which the only usage is to apply to linkages.
+We don't yet consider non-trivial equality between inheritance judgements.
+
+\subsection{Syntactic Translation that removes Linkages}
+To further provide the intuition why our core calculus is sound, we provide a syntactic translation from our core calculus to the subpart where linkages are absent.
+This can justify the \textit{consistency} and \textit{canonicity} modulo the quirky (but apparent) formulation of our inductive type.
+To be complete, we also provide a consistency and canonicity proof for our complete core calculus in the appendix.
+
+Thanks to QIIT formulation, syntactic translation becomes a function between QIIT type, which by definition respects all the judgemental equalities.
+
+\subsection{Relation to Plugin Implementation}

### 2022-10-14 metatheory-condensed.tex
Adding Inheritance Information.

@@
 $$
-\Rule[name=Tm Subst]
+\Rule[name=Subst]
 {\goodType{\Delta}{T}{j} 
@@
 Thus substitution has its own judgement $\goodSub{\Gamma}{\gamma}{\Delta}$.
-We can apply substitution to type and terms in MLTT.
+We can apply substitution to types\footnote{The type judgement is the form $\goodType{\Gamma}{T}{}$} and terms\footnote{The term judgement is the form $\goodTerm{\Gamma}{t}{T}$} in MLTT via \ruleref{Subst}.
 We also have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
@@
 
-The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$.
+The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
 For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators).
@@
 Thus we have the signature of inductive type collecting each constructors, with a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$.
-Unconventionally, a signature can induce a "universe" of types $\bW$: in other words, $T : \bW$ will be the inductive type instead of $\bW$ itself.
-"Wsup" is used to construct a specific term of an given inductive type.
+Unconventionally, a signature can induce a "universe" of types $\bW\tau$---in other words, $T : \bW\tau$ will be the inductive type instead of $\bW\tau$ itself.
+"Wsup" is used to construct a specific term of a given inductive type.
 
-Similar to plugin implementation, we delegate handlers for recursion using linkages.
+Similar to plugin implementation, we delegate handlers for recursion to linkages.
 We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers.
@@
 
+$$
+\Rule[name=Inh]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+\quad \goodTerm{\Gamma}{l}{\cL \sigma_1}
+}
+{\goodTerm{\Gamma}{("inh" \ h \ l)}{\cL \sigma_2}},
+\Rule[name=Inh-Id]
+{}
+{\goodInh{\Gamma}{"inhid"}{\sigma}{\sigma}},
+\Rule[name=Inh-Ext]
+{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
+  \quad \goodTerm{\Gamma, A_2}{t}{T}}
+{\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}
+$$
+
 % Inhertance judgement
 In the plugin implementation, "Family"s are second class objects like Coq's "Module".
-What's more, a "Family", with or without "extends" argument can be considered as a standalone piece of code trying to \textit{inherit} an existent (but maybe empty) family.
-Thus we model them using \textit{inheritance judgement} instead of considering them as first class terms.
+What's more, a "Family" (with or without "extends" clause) can be considered as a standalone piece of code trying to \textit{inherit} an existent (but maybe empty) family.
+Thus we model them using \textit{inheritance judgement} $\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}$ instead of considering them as first class terms.
+The only way to use $h$ is by \ruleref{Inh} and transforming a linkage of signature $\sigma_1$ to another linkage of $\sigma_2$.
 
-In this way, we will consider the program in our metatheory similar to those vanilla MLTT program, except there are meta variables denoting inheirtance judgement, of which the only usage is to apply to linkages.
+Inheritance judgement can be further enriched with inheriting operation, overriding operation, extending operation and etc.
+These operations correspond to inherit, override and extend command inside "Family" during plugin implementation.
+
+In this way, we will consider the program in our metatheory similar to those vanilla MLTT program, except there are meta variables denoting inheirtance judgement (i.e.
+"Family"), of which the only usage is to apply \ruleref{Inh}.
 We don't yet consider non-trivial equality between inheritance judgements.

### 2022-10-14 metatheory-condensed.tex
Adding Syn Translation

@@
 
-Our formulation of (extensible) inductive type is close to Wtype (\ruleref{WSig-Add} requires a pair $A \vdash B$ as well), but with mulitple constructors.
+Our formulation of (extensible) inductive type is close to Wtype (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
 Thus we have the signature of inductive type collecting each constructors, with a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$.
@@
 
+\newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
+\newcommand{\Sigr}[2]{{ "Sig"^r~{#1}~{#2} }}
+
 Thanks to QIIT formulation, syntactic translation becomes a function between QIIT type, which by definition respects all the judgemental equalities.
+We will now sketch this translation $\denotesT{\_}$, which keeps most parts untouched except for the linkages related.
+
+\begin{align*}
+  \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
+  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursive, } \\
+  & \text{defined by induction on the signature length} \\  
+  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\
+  {\Sigr{\Gamma}{(n+1)}} &= 
+    \Sigma~(\Sigr{\Gamma}{n})
+          ~(\Sigma~\cU
+                  ~(\Sigma~(\Pi(\cC (\pi_2[\pi_1]))(El~\pi_2[\pi_1]))~\cU)) \\
+  \Sigr{\Gamma}{0} &= \top \\
+  \denotesT{\goodType{\Gamma}{\cC \sigma}{}}, \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &\text{ defines upon } \denotesT{\sigma} 
+  \text{ and inductively on the signature length} \\
+  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesT{\goodType{\Gamma}{\cL\sigma}{}} = \top \quad
+      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
+  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
+    \Sigma~\denotesT{\cC ("pjl"~\sigma)}~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
+      &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
+  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &=
+  \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
+  \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
+  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\cL(\nu^+~\sigma~T)}} &= (m, "lam"~t) \\ 
+  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} & \text{ is defined upon } \denotesT{m} \text{ and inductively on the signature length} \\ 
+  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} &= 
+  (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
+  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
+\end{align*}
+\begin{align*}
+  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[\pi_1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[\pi_1]} \\
+  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}} &= \mu^+~(h[(\pi_1,\pi_1\mu~\pi_2)])\\ &\quad \quad \{f_2[\pi_1]\}~(p_2\mu'~\pi_2)[(\pi_1, \uparrow^s[\pi_1^{\uparrow}])] \\
+\end{align*}
+
+We omit most of the inheritance translation.
 

### 2022-10-14 metatheory-condensed.tex
Add Relations to plugin.

@@
 Thus substitution has its own judgement $\goodSub{\Gamma}{\gamma}{\Delta}$.
-We can apply substitution to types\footnote{The type judgement is the form $\goodType{\Gamma}{T}{}$} and terms\footnote{The term judgement is the form $\goodTerm{\Gamma}{t}{T}$} in MLTT via \ruleref{Subst}.
+We can apply substitution to types and terms\footnote{The type judgement is the form $\goodType{\Gamma}{T}{}$.
+The term judgement is the form $\goodTerm{\Gamma}{t}{T}$} in MLTT via \ruleref{Subst}.
 We also have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
@@
 
-Looking closely, the sealing judgement helps to model most machinaries about abstraction in plugin implementation.
+% Looking closely, the sealing judgement helps to model most machinaries about abstraction in plugin implementation.
 
@@
 We will now sketch this translation $\denotesT{\_}$, which keeps most parts untouched except for the linkages related.
+We omit most of the inheritance translation.
 
@@
 
-We omit most of the inheritance translation.
+\subsection{Relation to Plugin Implementation}
+Our core calculus can model most of plugin implementation.
 
-\subsection{Relation to Plugin Implementation}
+% Extending Inductive Type
+We extend inductive type by overriding the inductive type with a enriched one.
+Thanks to sealing, most stuff from the parent family can be inherited to the children because enriched inductive type can export strictly stronger interface when abstracting.
+
+% Partial Recursor + Computational Axiom
+The partial recursors and the computational axioms are no different from recursors and normal fields.
+We can prove partial recursor using current recursing mechanism and proving computational axiom on the trivial abstraction.
+After that, we can use sealing to abstract partial recursor and computational axiom again for the later fields.
+
+% Extending Recursion Handler
+The extension of recursion handlers is delegated by the inheritance of families. Adding new handlers for a new constructors is exactly extending families with new fields of certain types.
+
+% Overriding + Pins
+Finally, we pointed out how overridable and pins is modeled if we admit equality reflection.
+Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
+hide its definition when type-check t, and remaining former fields have their definition exposed in t.
+This above elaborated "hiding" syntax directly corresponding to the semantic of sealing---when definition is exposed, we expose propositional equality into the context, which will be a judgemental equality with the help of equality reflection; otherwise we only expose the type and no extra information.

### 2022-10-14 metatheory-condensed.tex
Add Information for compilation type.

@@
 Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement and will be explained later.
+We will use $\{a : T_1, b : T_2, ..\}$ as a shorthand for signature with type $T_1, T_2, ..$ for readability.
 
@@
 The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
-For example, $\cC $
+For example, $\cC\{a : \cU, b : a \to \bot \} = \Sigma(a : \cU).\Sigma (b : a \to \bot).
+\top$.
 
@@
 
-We also show one example illustrating the importance of the compilation type.
+We now sketch out the importance of the compilation type.
+First we assume a singleton type (universe)\footnote{Something like this singleton type universe will appear later in inductive type formulation.} $A : S(T) \iff A \equiv T$, i.e.
+a type of one single type.
+We can see that we have a term $x : \cC\{A : S(\bot), B : A \to \bot \} \vdash h_1 : \cC\{A : \cU, B:A \to \bot \}$ which is equivalent to $x : \Sigma (A: S(\bot)).
+A \to \bot \vdash h_1 : \Sigma (A: \cU).
+A \to \bot$.
+This can be considered as using seal to do an abstraction on $A:S(\bot)$.
+
+But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
+
+However, this form of abstracting a type is very important when programming in our metatheory---as later we will show inductive type of different constructors locates in different type (universe), and to make later fields inheritable we need to work on an interface of abstracted type.
+Thus the compilation type is indispensible.
 
@@
 Finally, we pointed out how overridable and pins is modeled if we admit equality reflection.
-Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
+Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
 hide its definition when type-check t, and remaining former fields have their definition exposed in t.

### 2022-10-14 ext-ind-type-compilation-example.tex
compilation of family STLC

@@
 End STLC_Ctx1.
-Module Type tm_prec_64 (self: STLC_Ctx1).[@\YZ{tm_6? Do you mean STLC_Ctx1 instead?}\EDJreply{Yes Thanks.}@]
+Module Type tm_prec_64 (self: STLC_Ctx1).
 Parameter (* Type for partial recursor *)
@@
 }@]
-[@\YZ{(3) Are nested families really needed?}\EDJreply{Not really but if consider handlers as one family, it is nested family.
-What's more, the audience needs to know how nested family is compiled at some point.}@]
   Definition tm_var s x t := 
@@
 (* Exposing Computational Axiom into Ctx *) [@\tanc{Tyck-Recty-cont-start}@]
-Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).[@\YZ{What is "..."?}\EDJreply{My intension is STLC_Ctx_n for some n where at this point the audience should able to see what the context of that STLC_Ctx is.
-Apparently I failed.}@]
+Module Type substtm_139 (self: STLC_Ctx[@\dadada@]).
 Parameter
   (subst_on_tm_unit :
-	 (self.subst (self.tm_unit)) =[@\YZ{Where is this "subst" added to the context?}\EDJreply{The reason is above. Currently the space is already not enough.}@]
+	 (self.subst (self.tm_unit)) =
       self.subst_internal.tm_unit).
@@
 (* Compilation, happens when closing a top-level family *) [@\tanc{Cmp-Result-start}@]
-(* Ctx that expose concrete inductive type *)
+(* Ctx that exposes concrete inductive type *)
 Module Type STLC_CCtx.

### 2022-10-14 lang-design.tex
compilation of family STLC

@@
 affect \emph{propositional equality}.
-That is, \lsti{subst tm_unit x t' = tm_unit}, as a proposition
-(\lsti{Prop}), should still hold.
+That is, \lsti{\forall x t, subst tm_unit x t = tm_unit},
+as a proposition (\lsti{Prop}), should still hold.
 After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
@@
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-STLC_tm_prec : \forall (P : tm -> Set), [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] -> [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] ->
+tm_prec_STLC : \forall (P : tm -> Set), [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] -> [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] ->
                [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
@@
 \noindent
-As expected, "STLC_tm_prec" is axiomatized along with
+As expected, "tm_prec_STLC" is axiomatized along with
 four (propositional) equalities, one for each constructor of "tm".
-For instance, the equality for the "tm_abs" constructor is as follows:
+For instance, the equality for constructor "tm_abs" is as follows:
 
+\begin{centered}
+\begin{minipage}{.99\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-\forall x t P H1 H2 H3 H4,
-STLC_tm_prec P H1 H2 H3 H4 (tm_abs x t) = H3 x t (STLC_tm_prec P H1 H2 H3 H4 t)
+tm_abs_eq_STLC: \forall x t P H1 H2 H3 H4,
+                tm_prec_STLC P H1 H2 H3 H4 (tm_abs x t) = H3 x t (tm_prec_STLC P H1 H2 H3 H4 t)
 \end{lstlisting}
+\end{minipage}
+\end{centered}
 
@@
 Importantly, unlike the standard "tm_rec" recursor, the partial recursor
-"STLC_tm_prec" is compatible with the late binding of "tm" in its type.
-When "STLC_tm_prec" is inherited into family "STLCFix", all the previous four equalities
+"tm_prec_STLC" is compatible with the late binding of "tm" in its type.
+When "tm_prec_STLC" is inherited into family "STLCFix", all the previous four equalities
 still hold, and a trivial, fifth equality is made available:
 
+\begin{centered}
+\begin{minipage}{.88\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-\forall x t P H1 H2 H3 H4, STLC_tm_prec P H1 H2 H3 H4 (tm_fix x t) = None
+tm_fix_eq_STLC : \forall x t P H1 H2 H3 H4, tm_prec_STLC P H1 H2 H3 H4 (tm_fix x t) = None
 \end{lstlisting}
+\end{minipage}
+\end{centered}
 
-Although partial recursors appear weaker than ordinary recursors,
-there is power in restraint.
+Partial recursors appear weaker than ordinary recursors,
+but there is power in restraint.
 Partial recursors offer a principled, uniform way to derive injectivity and disjointness of
-constructors, while allowing extensibility:
-they enable injective mappings from values of a late bound inductive
-type, like "tm", to values of an ordinary inductive type, like \lsti{\nat},
+constructors, while allowing for extensibility:
+they enable injective mappings from a late bound inductive
+type, like "tm", to an ordinary inductive type, like \lsti{\nat},
 the injectivity and disjointness of whose constructors are readily
@@
 For example, in a proof state with the assumption \lsti{H : tm_var "x" = tm_abs "y" t},
-running \lsti{fdiscriminate H} first uses \lsti{H} to obtain the equality
+running \lsti{fdiscriminate H} first applies the injective mapping to
+both sides of~\lsti{H}, obtaining
 
 \begin{centered}
-\begin{minipage}{\textwidth}
+\begin{minipage}{.96\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-STLC_tm_prec (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_var "x") =
-STLC_tm_prec (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_abs "y" t)
+tm_prec_STLC (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_var "x") =
+tm_prec_STLC (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_abs "y" t)
 \end{lstlisting}
@@
 \noindent
-and then uses the axiomatized equalities about "STLC_tm_prec" to
-rewrite the equality above, obtaining \lsti{Some 1 = Some 4},
+and then rewrites the above equality using two of the axiomatized
+equalities about "tm_prec_STLC",
+%(namely "tm_var_eq_STLC" and "tm_abs_eq_STLC"),
+obtaining \lsti{Some 1 = Some 4},
 from which \lsti{False} easily follows.
 
-Note that the proof term generated by \lsti{fdiscriminate H} in family "STLC"
+Note that the proof term generated by \lsti{fdiscriminate H} in "STLC"
 is reusable by family "STLCFix", because the partial recursor and its
@@
 Also note that within family "STLCFix", a second partial recursor
-(called "STLCFix_tm_prec") on "tm" is available;
-it can be used to derive properties about "tm_fix",
-such as \lsti{tm_var "x" = tm_fix "y" t -> False}.
+(called "tm_prec_STLCFix") and its equalities are available,
+allowing properties about "tm_fix",
+such as \lsti{\neg(tm_var "x" = tm_fix "y" t)}, to be proved.
 
-\noindentparagraph{Limitations.}
+%\noindentparagraph{Limitations.}
+
+\ifShowOldWriting
 
@@
 only allow inheritance and extension).
+
+\fi

### 2022-10-14 stlc-compiled.tex
compilation of family STLC

\begin{figure}
\begin{minipage}{\textwidth}
\begin{multicols}{2}

\lstset{
  basicstyle=\fontsize{8}{8.5}\ttfamily,
% numbers=left,
}

\colorlet{codecomment-color}{dkgreen}

\newcommand{\innersep}{\vspace{5pt}}

\begin{lstlisting}
Module Type STLC\_ty\_Ctx.[@\tanc{c-STLC-ty-start}@]
End STLC\_ty\_Ctx.[@\innersep@]
Module Type STLC\_ty (self : STLC\_ty\_Ctx).
Parameter ty : Set.
Parameter ty_unit : ty.
Parameter ty_arrow : ty -> ty -> ty.
Parameter ty_prec_STLC : [@\dadada@].[@\hfill\codecomment{partial recursor}@]
Parameter ty_unit_eq_STLC : [@\dadada@].[@\hfill\codecomment{eq.\ on recursor}@]
Parameter ty_arrow_eq_STLC : [@\dadada@].[@\codecomment{eq.\ on recursor}@]
End STLC\_ty.[@\tanc{c-STLC-ty-end}@]

Module Type STLC\_tm\_Ctx.[@\tanc{c-STLC-tm-start}@]
Include STLC\_ty\_Ctx.  Include STLC\_ty.
End STLC\_tm\_Ctx.[@\innersep@]
Module Type STLC\_tm (self : STLC\_tm\_Ctx).
Parameter tm : Set.
[@\dadada@][@\hfill\codecommentstart{axiomatized \texttt{tm} constructors, partial recursor}@]
[@\hfill\codecommentend{\texttt{tm\_prec\_STLC}, and its four equalities}@]
End STLC\_tm.[@\tanc{c-STLC-tm-end}@]

Module Type STLC\_env\_Ctx.[@\tanc{c-STLC-env-start}@]
Include STLC\_tm\_Ctx.  Include STLC\_tm.
End STLC\_env\_Ctx.[@\innersep@]
Module STLC\_env (self : STLC\_env\_Ctx).
Definition env : Type := id -> option self.ty.
End STLC\_env.[@\tanc{c-STLC-env-end}@]

Module Type STLC\_empty\_Ctx.[@\tanc{c-STLC-empty-start}@]
Include STLC\_env\_Ctx.  Include STLC\_env.
End STLC\_empty\_Ctx.[@\innersep@]
Module STLC\_empty (self : STLC\_empty\_Ctx).
Definition empty : self.env := \lambda _, None.
End STLC\_empty.[@\tanc{c-STLC-empty-end}@]

[@\dadada@][@\hfill\codecomment{compilation of \texttt{hasty}, \texttt{value}}@]

Module Type STLC\_subst\_Ctx.[@\tanc{c-STLC-subst-start}@]
Include STLC\_value\_Ctx.  Include STLC\_value.
End STLC\_subst\_Ctx.[@\innersep@]
Module STLC\_subst\_Cases (self : STLC\_subst\_Ctx).
Definition STLC\_subst\_tm_unit :=
  \lambda (x : id) (t : self.tm), self.tm_unit.
Definition STLC\_subst\_tm_var := [@\dadada@].
Definition STLC\_subst\_tm_abs := [@\dadada@].
Definition STLC\_subst\_tm_app := [@\dadada@].
End STLC\_subst\_Cases.[@\innersep@]
Module Type STLC\_subst (self : STLC\_subst\_Ctx).
Parameter subst : self.tm[@\,@]->[@\,@]id[@\,@]->[@\,@]self.tm[@\,@]->[@\,@]self.tm.
Parameter subst_tm_unit_eq : \forall x t,
  self.subst (self.tm_unit) x t = self.tm_unit.
Parameter subst_tm_var_eq : [@\dadada@].
Parameter subst_tm_abs_eq : [@\dadada@].
Parameter subst_tm_app_eq : [@\dadada@].
End STLC\_subst.[@\tanc{c-STLC-subst-end}@]

[@\dadada@][@\hfill\codecomment{compilation of rest of family \texttt{STLC}}@][@\columnbreak@]

[@~@]
\end{lstlisting}

\makeline[0pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!50]
\makeline[0pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!50]
\makeline[0pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!50]
\makeline[0pt]{c-STLC-empty-start}{c-STLC-empty-end}[codecomment-color!50]
\makeline[0pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!50]

\end{multicols}
\end{minipage}
\caption{Compilation of the STLC mechanizations in \cref{fig:stlc-mechanized} to parameterized Coq modules.}
\label{fig:stlc-compiled}
\end{figure}

### 2022-10-16 lang-design.tex
stlc compilation figure

@@
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-tm_rec : \forall (P : tm -> Set), [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] -> [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] ->
+tm_rect : \forall (P : tm -> Type), [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] -> [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] ->
          [@\hlcase{tm-recursor-case-abs}@](\forall x t, P t -> P (tm_abs x t))[@\tikzmarkend{tm-recursor-case-abs}@] ->
@@
 in jeopardy, however.
-In particular, recursor "tm_rec" should no longer be allowed, because
+In particular, recursor "tm_rect" should no longer be allowed, because
 its type mentions "tm", whose meaning is late bound, yet
-"tm_rec" purports to claim \lsti{\forall t, P t} given only four case handlers.
+"tm_rect" purports to claim \lsti{\forall t, P t} given only four case handlers.
 
@@
 and above all, why should a property like
-\lsti{tm_var "x" = tm_abs "y" t -> False} have anything to do with "tm_fix"?
+\lsti{\neg(tm_var "x" = tm_abs "y" t)} have anything to do with "tm_fix"?
 
@@
 
-As analyzed earlier, ordinary recursors, such as "tm_rec", are impossible
+As analyzed earlier, ordinary recursors, such as "tm_rect", are impossible
 within a family in which the name of the inductive type is late bound.
@@
 \begin{centered}
-\begin{minipage}{.96\textwidth}
+\begin{minipage}{.97\textwidth}
 \newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-tm_prec_STLC : \forall (P : tm -> Set), [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] -> [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] ->
+tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
                 [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
@@
 \noindent
-As expected, "tm_prec_STLC" is axiomatized along with
+As expected, "tm_prect_STLC" is axiomatized along with
 four (propositional) equalities, one for each constructor of "tm".
@@
 \begin{centered}
-\begin{minipage}{.99\textwidth}
+\begin{minipage}{\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
 tm_abs_eq_STLC: \forall x t P H1 H2 H3 H4,
-                tm_prec_STLC P H1 H2 H3 H4 (tm_abs x t) = H3 x t (tm_prec_STLC P H1 H2 H3 H4 t)
+                tm_prect_STLC P H1 H2 H3 H4 (tm_abs x t) = H3 x t (tm_prect_STLC P H1 H2 H3 H4 t)
 \end{lstlisting}
@@
 %\noindent
-Importantly, unlike the standard "tm_rec" recursor, the partial recursor
-"tm_prec_STLC" is compatible with the late binding of "tm" in its type.
-When "tm_prec_STLC" is inherited into family "STLCFix", all the previous four equalities
+Importantly, unlike the standard "tm_rect" recursor, the partial recursor
+"tm_prect_STLC" is compatible with the late binding of "tm" in its type.
+When "tm_prect_STLC" is inherited into family "STLCFix", all the previous four equalities
 still hold, and a trivial, fifth equality is made available:
@@
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-tm_fix_eq_STLC : \forall x t P H1 H2 H3 H4, tm_prec_STLC P H1 H2 H3 H4 (tm_fix x t) = None
+tm_fix_eq_STLC : \forall x t P H1 H2 H3 H4, tm_prect_STLC P H1 H2 H3 H4 (tm_fix x t) = None
 \end{lstlisting}
@@
 \begin{centered}
-\begin{minipage}{.96\textwidth}
+\begin{minipage}{.98\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-tm_prec_STLC (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_var "x") =
-tm_prec_STLC (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_abs "y" t)
+tm_prect_STLC (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_var "x") =
+tm_prect_STLC (\lambda[@\,@]_, \nat) (\lambda[@\,@]_, Some 1) (Some 2) (\lambda[@\,@]_[@\,@]_[@\,@]_, Some 3) (\lambda[@\,@]_[@\,@]_[@\,@]_[@\,@]_, Some 4) (tm_abs "y" t)
 \end{lstlisting}
@@
 and then rewrites the above equality using two of the axiomatized
-equalities about "tm_prec_STLC",
+equalities about "tm_prect_STLC",
 %(namely "tm_var_eq_STLC" and "tm_abs_eq_STLC"),
@@
 Also note that within family "STLCFix", a second partial recursor
-(called "tm_prec_STLCFix") and its equalities are available,
+(called "tm_prect_STLCFix") and its equalities are available,
 allowing properties about "tm_fix",

### 2022-10-16 stlc-compiled.tex
stlc compilation figure

@@
 \lstset{
-  basicstyle=\fontsize{8}{8.5}\ttfamily,
+  basicstyle=\fontsize{8}{7.8}\ttfamily,
 % numbers=left,
@@
 
-\newcommand{\innersep}{\vspace{5pt}}
+\newcommand{\innersep}{\vspace{4pt}}
+\newcommand{\outersep}{\vspace{4pt}}
+\newcommand{\outdent}{\!\!\!}
 
 \begin{lstlisting}
-Module Type STLC\_ty\_Ctx.[@\tanc{c-STLC-ty-start}@]
-End STLC\_ty\_Ctx.[@\innersep@]
-Module Type STLC\_ty (self : STLC\_ty\_Ctx).
-Parameter ty : Set.
-Parameter ty_unit : ty.
-Parameter ty_arrow : ty -> ty -> ty.
-Parameter ty_prec_STLC : [@\dadada@].[@\hfill\codecomment{partial recursor}@]
-Parameter ty_unit_eq_STLC : [@\dadada@].[@\hfill\codecomment{eq.\ on recursor}@]
-Parameter ty_arrow_eq_STLC : [@\dadada@].[@\codecomment{eq.\ on recursor}@]
-End STLC\_ty.[@\tanc{c-STLC-ty-end}@]
-
-Module Type STLC\_tm\_Ctx.[@\tanc{c-STLC-tm-start}@]
+[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{tm} in family \texttt{STLC}}@][@\tanc{c-STLC-tm-start}@]
+[@\outdent@]Module Type STLC\_tm\_Ctx.
+[@\outdent@]End STLC\_tm\_Ctx.[@\innersep@]
+[@\outdent@]Module Type STLC\_tm (self : STLC\_tm\_Ctx).
+Axiom tm : Set.
+Axiom tm_unit : tm.[@\hfill@]Axiom tm_var : id -> tm.
+Axiom tm_abs : id -> tm -> tm.
+Axiom tm_app : tm -> tm -> tm.
+Axiom tm_prect_STLC : [@\dadada@].
+Axiom [@\textls[-30]{tm\_unit\_eq\_STLC}@][@\;:\;@][@\dadada@]. Axiom [@\textls[-30]{tm\_abs\_eq\_STLC}@][@\;:\;@][@\dadada@].
+Axiom [@\textls[-30]{tm\_abs\_eq\_STLC}@][@\;:\;@][@\dadada@].  Axiom [@\textls[-30]{tm\_app\_eq\_STLC}@][@\;:\;@][@\dadada@].
+[@\outdent@]End STLC\_tm.[@\tanc{c-STLC-tm-end}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in family \texttt{STLC}}@][@\tanc{c-STLC-subst-start}@]
+[@\outdent@]Module Type STLC\_subst\_Ctx.
+Include STLC\_tm\_Ctx.  Include STLC\_tm.
+[@\outdent@]End STLC\_subst\_Ctx.[@\innersep@]
+[@\outdent@]Module STLC\_subst\_Cases (self : STLC\_subst\_Ctx).
+Def subst\_tm_unit :=
+  \lambda (x : id) (t : self.tm), self.tm_unit.
+Def subst\_tm_var := [@\dadada@].
+Def subst\_tm_abs := [@\dadada@].  Def subst\_tm_app := [@\dadada@].
+[@\outdent@]End STLC\_subst\_Cases.[@\innersep@]
+[@\outdent@]Module Type STLC\_subst (self : STLC\_subst\_Ctx).
+Axiom subst : self.tm[@\,@]->[@\,@]id[@\,@]->[@\,@]self.tm[@\,@]->[@\,@]self.tm.
+Axiom subst_tm_unit_eq : \forall x t,
+  self.subst (self.tm_unit) x t = self.tm_unit.
+Axiom subst_tm_var_eq : [@\dadada@].
+Axiom subst_tm_abs_eq : [@\dadada@].
+Axiom subst_tm_app_eq : [@\dadada@].
+[@\outdent@]End STLC\_subst.[@\tanc{c-STLC-subst-end}@][@\outersep@]
+[@\outdent@][@\dadada@][@\hfill\codecomment{Code emitted upon definition of \texttt{ty} in family \texttt{STLC}}@][@\tanc{c-STLC-ty-start}@][@\tanc{c-STLC-ty-end}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{env} in family \texttt{STLC}}@][@\tanc{c-STLC-env-start}@]
+[@\outdent@]Module Type STLC\_env\_Ctx.
 Include STLC\_ty\_Ctx.  Include STLC\_ty.
-End STLC\_tm\_Ctx.[@\innersep@]
-Module Type STLC\_tm (self : STLC\_tm\_Ctx).
-Parameter tm : Set.
-[@\dadada@][@\hfill\codecommentstart{axiomatized \texttt{tm} constructors, partial recursor}@]
-[@\hfill\codecommentend{\texttt{tm\_prec\_STLC}, and its four equalities}@]
-End STLC\_tm.[@\tanc{c-STLC-tm-end}@]
-
-Module Type STLC\_env\_Ctx.[@\tanc{c-STLC-env-start}@]
-Include STLC\_tm\_Ctx.  Include STLC\_tm.
-End STLC\_env\_Ctx.[@\innersep@]
-Module STLC\_env (self : STLC\_env\_Ctx).
-Definition env : Type := id -> option self.ty.
-End STLC\_env.[@\tanc{c-STLC-env-end}@]
-
-Module Type STLC\_empty\_Ctx.[@\tanc{c-STLC-empty-start}@]
-Include STLC\_env\_Ctx.  Include STLC\_env.
-End STLC\_empty\_Ctx.[@\innersep@]
-Module STLC\_empty (self : STLC\_empty\_Ctx).
-Definition empty : self.env := \lambda _, None.
-End STLC\_empty.[@\tanc{c-STLC-empty-end}@]
-
-[@\dadada@][@\hfill\codecomment{compilation of \texttt{hasty}, \texttt{value}}@]
-
-Module Type STLC\_subst\_Ctx.[@\tanc{c-STLC-subst-start}@]
-Include STLC\_value\_Ctx.  Include STLC\_value.
-End STLC\_subst\_Ctx.[@\innersep@]
-Module STLC\_subst\_Cases (self : STLC\_subst\_Ctx).
-Definition STLC\_subst\_tm_unit :=
-  \lambda (x : id) (t : self.tm), self.tm_unit.
-Definition STLC\_subst\_tm_var := [@\dadada@].
-Definition STLC\_subst\_tm_abs := [@\dadada@].
-Definition STLC\_subst\_tm_app := [@\dadada@].
-End STLC\_subst\_Cases.[@\innersep@]
-Module Type STLC\_subst (self : STLC\_subst\_Ctx).
-Parameter subst : self.tm[@\,@]->[@\,@]id[@\,@]->[@\,@]self.tm[@\,@]->[@\,@]self.tm.
-Parameter subst_tm_unit_eq : \forall x t,
-  self.subst (self.tm_unit) x t = self.tm_unit.
-Parameter subst_tm_var_eq : [@\dadada@].
-Parameter subst_tm_abs_eq : [@\dadada@].
-Parameter subst_tm_app_eq : [@\dadada@].
-End STLC\_subst.[@\tanc{c-STLC-subst-end}@]
-
-[@\dadada@][@\hfill\codecomment{compilation of rest of family \texttt{STLC}}@][@\columnbreak@]
-
-[@~@]
+[@\outdent@]End STLC\_env\_Ctx.[@\innersep@]
+[@\outdent@]Module STLC\_env (self : STLC\_env\_Ctx).
+Def env : Type := id -> option self.ty.
+[@\outdent@]End STLC\_env.[@\tanc{c-STLC-env-end}@][@\outersep@]
+[@\outdent@][@\dadada@][@\hfill\codecomment{Code emitted for other fields defined in family \texttt{STLC}}@][@\tanc{c-STLC-rest}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted upon completion of family \texttt{STLC}}@][@\tanc{c-STLC-start}@]
+[@\outdent@]Module STLC.[@\innersep@]
+Inductive tm\_ : Set :=[@\tanc{c-STLC-d-tm-start}@]
+| tm_unit\_ : tm\_    | tm_var\_ : id -> tm\_
+| tm_abs\_  : id -> tm\_ -> tm\_
+| tm_app\_  : tm\_ -> tm\_ -> tm\_.
+Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
+Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
+Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
+[@\hfill\codecomment{Instantiate \texttt{tm} partial recursor \& equalities}@]
+Def tm_prect_STLC P H1 H2 H3 H4 : \forall t, P t :=
+  fix F t := match t with
+  | tm_unit\_ => H1  | tm_var\_ x => H2 x
+  | tm_abs\_ x t => H3 x t (F t)
+  | tm_app\_ t1 t2 => H4 t1 (F t1) t2 (F t2) end.
+Fact[@\;@][@\textls[-35]{tm\_unit\_eq\_STLC}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
+Include STLC\_subst\_Cases.[@\tanc{c-STLC-d-subst-start}@]
+[@\hfill\codecomment{Instantiate \texttt{subst} \& equalities}@]
+Def subst[@\,@]:[@\,@]tm[@\,@]->[@\,@]id[@\,@]->[@\,@]tm[@\,@]->[@\,@]tm :=[@\tanc{c-STLC-firstcol-end}\columnbreak@]
+  tm\__rect _ subst\_tm_unit subst\_tm_var[@\tanc{c-STLC-sndcol-start}@]
+             subst\_tm_abs  subst\_tm_app.
+Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\innersep@]
+[@\dadada@][@\hfill\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@][@\tanc{c-STLC-d-ty}@][@\innersep@]
+Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLC-d-env}@][@\innersep@]
+[@\dadada@][@\hfill\codecomment{Include/Instantiate other fields of family \texttt{STLC}}@][@\tanc{c-STLC-d-rest}@][@\innersep@]
+[@\outdent@]End STLC.[@\tanc{c-STLC-end}@]
 \end{lstlisting}
 
-\makeline[0pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!50]
-\makeline[0pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!50]
-\makeline[0pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!50]
-\makeline[0pt]{c-STLC-empty-start}{c-STLC-empty-end}[codecomment-color!50]
-\makeline[0pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!50]
+\makeline[0pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!50][1.25pt]
+\makeline[0pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!50][1.25pt]
+\makeline[0pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!50][1.25pt]
+\makeline[0pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!50][1.25pt]
+\makeline[0pt]{c-STLC-rest}{c-STLC-rest}[codecomment-color!50][1.25pt]
+\makeline[0pt]{c-STLC-start}{c-STLC-firstcol-end}[codecomment-color!50][1.25pt]
+\makeline[4pt]{c-STLC-d-tm-start}{c-STLC-d-tm-end}[codecomment-color!25][1.25pt]
+\makeline[4pt]{c-STLC-d-subst-start}{c-STLC-firstcol-end}[codecomment-color!20][1.25pt]
+\makeline[.5\textwidth+9pt]{c-STLC-sndcol-start}{c-STLC-end}[codecomment-color!50][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLC-sndcol-start}{c-STLC-d-subst-end}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLC-d-ty}{c-STLC-d-ty}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLC-d-env}{c-STLC-d-env}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLC-d-rest}{c-STLC-d-rest}[codecomment-color!25][1.25pt]
+
+\vspace{-15pt}
+
+\colorlet{codecomment-color}{orange}
+
+\begin{lstlisting}
+[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{tm} in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-tm-start}@]
+[@\outdent@]Module Type STLCFix\_tm\_Ctx.
+[@\outdent@]End STLCFix\_tm\_Ctx.[@\innersep@]
+[@\outdent@]Module Type STLCFix\_tm (self : STLCFix\_tm\_Ctx).
+Include STLC\_tm(self).
+Axiom tm_fix : id -> tm -> tm.
+Axiom tm_fix_eq_STLC : \forall[@\,@][@\dadada@],[@\,@]tm_prect_STLC [@\dadada@][@\;@]=[@\;@]None.
+Axiom tm_prect_STLCFix : [@\dadada@].
+Axiom tm_fix_eq_STLCFix : [@\dadada@].
+[@\outdent@]End STLCFix\_tm.[@\tanc{c-STLCFix-tm-end}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-subst-start}@]
+[@\outdent@]Module Type STLCFix\_subst\_Ctx.
+Include STLCFix\_tm\_Ctx. Include STLCFix\_tm.
+[@\outdent@]End STLCFix\_subst\_Ctx.[@\innersep@]
+[@\outdent@]Module STLCFix\_subst\_Cases (self[@\,@]:[@\,@]STLCFix\_subst\_Ctx).
+Include STLC\_subst\_Cases(self).
+Def subst\_tm_fix := [@\dadada@].
+[@\outdent@]End STLCFix\_subst\_Cases.[@\innersep@]
+[@\outdent@]Module Type STLCFix\_subst (self[@\,@]:[@\,@]STLCFix\_subst\_Ctx).
+Include STLC\_subst(self).
+Axiom subst_tm_fix_eq : [@\dadada@].
+[@\outdent@]End STLCFix\_subst.[@\tanc{c-STLCFix-subst-end}@][@\outersep@]
+[@\outdent@][@\dadada@] [@\codecomment{Code emitted for other fields defined in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-rest}@][@\outersep@]
+[@\outdent\codecomment{Code emitted upon completion of family \texttt{STLCFix}}@][@\tanc{c-STLCFix-start}@]
+[@\outdent@]Module STLCFix.[@\innersep@]
+Inductive tm\_ : Set :=[@\tanc{c-STLCFix-d-tm-start}@]
+| tm_unit\_ : tm\_    | tm_var\_ : id -> tm\_
+| tm_abs\_  : id -> tm\_ -> tm\_
+| tm_app\_  : tm\_ -> tm\_ -> tm\_
+| tm_fix\_  : id -> tm\_ -> tm\_.
+Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
+Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
+Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
+Def tm_fix  := tm_fix\_.
+[@\dadada@][@\hfill\codecomment{Instantiate \texttt{tm} partial recursors \& equalities}@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
+Include STLCFix\_subst\_Cases. [@\tanc{c-STLCFix-d-subst-start}@]
+[@\hfill\codecomment{Instantiate \texttt{subst} \& equalities}@]
+Def subst[@\,@]:[@\,@]tm[@\,@]->[@\,@]id[@\,@]->[@\,@]tm[@\,@]->[@\,@]tm :=
+  tm\__rect _ subst\_tm_unit subst\_tm_var
+             subst\_tm_abs  subst\_tm_app subst\_tm_fix.
+Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]
+[@\dadada@][@\hfill\codecomment{\textls[-10]{Instantiate \texttt{ty}, its constructors, partial recursor \& eq.}}@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
+Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLCFix-d-env}@][@\innersep@]
+[@\dadada@][@\hfill\codecomment{Include/Instantiate other fields of family \texttt{STLCFix}}@][@\tanc{c-STLCFix-d-rest}@][@\innersep@]
+[@\outdent@]End STLCFix.[@\tanc{c-STLCFix-end}@]
+\end{lstlisting}
+
+\makeline[.5\textwidth+9pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!50][1.25pt]
+\makeline[.5\textwidth+9pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!50][1.25pt]
+\makeline[.5\textwidth+9pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!50][1.25pt]
+\makeline[.5\textwidth+9pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!50][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-tm-start}{c-STLCFix-d-tm-end}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-subst-start}{c-STLCFix-d-subst-end}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-ty}{c-STLCFix-d-ty}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-env}{c-STLCFix-d-env}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-rest}{c-STLCFix-d-rest}[codecomment-color!25][1.25pt]
+
+\vspace{-15pt}
+
+\definecolor{codecomment-color}{HTML}{000080}
+
+\begin{lstlisting}
+[@\outdent@][@\codecomment{Code emitted upon command \texttt{Check} \dadada}@][@\tanc{c-Check-STLCFix-typesafe-start}@]
+[@\outdent@]Check STLCFix.typesafe.[@\tanc{c-Check-STLCFix-typesafe-end}@]
+\end{lstlisting}
+
+\makeline[.5\textwidth+9pt]{c-Check-STLCFix-typesafe-start}{c-Check-STLCFix-typesafe-end}[codecomment-color!35][1.25pt]
 
@@
 \end{minipage}
-\caption{Compilation of the STLC mechanizations in \cref{fig:stlc-mechanized} to parameterized Coq modules.}
+
+\vspace{-8pt}
+
+\caption{Compilation of the mechanization of the STLCs in \cref{fig:stlc-mechanized} to parameterized Coq modules.}
 \label{fig:stlc-compiled}

### 2022-10-16 stlc-isorec-prod.tex
stlc compilation figure

@@
 }
+
+\newcommand{\innersep}{\vspace{5pt}}
 
@@
 [@\hfill\codecomment{STLC extended with iso-recursive types}@]
+FInductive tm : Set +=
+| tm_fold : tm -> tm | tm_unfold : tm -> tm.[@\innersep@]
+FRecursion subst on tm motive [@\dadada@]. [@\dadada@] End subst.[@\innersep@]
 FInductive ty : Set +=
-| ty_var : id -> ty | ty_rec : id -> ty -> ty.
-
-FInductive tm : Set +=
-| tm_fold : tm -> tm | tm_unfold : tm -> tm.
-
+| ty_var : id -> ty | ty_rec : id -> ty -> ty.[@\innersep@]
 FRecursion tysubst on ty[@\hfill\codecomment{Type-level substitution}@]
@@
 Case ty_var   := [@\dadada@].   Case ty_rec   := [@\dadada@].
-End tysubst.
-
+End tysubst.[@\innersep@]
 FInductive hasty : env -> tm -> ty -> Prop +=
@@
     hasty G (tm_fold t) (ty_rec \alpha T)
-| ht_unfold : [@\dadada@].
-
-[@\dadada@][@\hfill\codecomment{Other adjustments}@]
+| ht_unfold : [@\dadada@].[@\innersep@]
+[@\dadada@][@\hfill\codecomment{Other adjustments}@][@\innersep@]
 End STLCIsorec.[@\tanc{STLCIsorec-end}@]
@@
 Family STLCFixIsorec extends STLC[@\tanc{STLCIsorecfix-start}@]
-using STLCFix, STLCIsorec.
-[@\hfill\ \ \codecomment{STLC extended with fixpoints and iso-recursive types}@]
+using STLCFix, STLCIsorec.[@\hfill\codecomment{STLC extended}@]
+[@\hfill\codecommentend{with fixpoints and iso-recursive types}@]
 End STLCFixIsorec.[@\tanc{STLCIsorecfix-end}@]
@@
 [@\hfill\codecomment{STLC extended with products}@]
-FInductive ty : Set +=
-| ty_prod : ty -> ty -> ty.
-
 FInductive tm : Set +=
 | tm_pair : tm -> tm -> tm
-| tm_fst : tm -> tm | tm_snd : tm -> tm.
-
-[@\dadada@][@\hfill\codecomment{Other adjustments}@]
+| tm_fst : tm -> tm | tm_snd : tm -> tm.[@\innersep@]
+FRecursion subst on tm motive [@\dadada@]. [@\dadada@] End subst.[@\innersep@]
+FInductive ty : Set +=
+| ty_prod : ty -> ty -> ty.[@\innersep@]
+[@\dadada@][@\hfill\codecomment{Other adjustments}@][@\innersep@]
 End STLCProd.[@\tanc{STLCProd-end}@]
@@
 
-\vspace{-10pt}
+\vspace{-12pt}
 
@@
 Family STLCProdIsorec extends STLC[@\tanc{STLCIsorecProd-start}@]
-using STLCProd, STLCIsorec.
-[@\hfill\ \ \codecomment{STLC extended with products and iso-recursive types}@]
+using STLCProd, STLCIsorec.[@\hfill\codecommentstart{STLC extended}@]
+[@\hfill\codecommentend{with products and iso-recursive types}@]
 FRecursion tysubst on ty
@@
 Case ty_prod := [@\dadada@].[@\hfill\codecomment{Substitution on product types}@]
-End tysubst.
-
+End tysubst.[@\innersep@]
 End STLCProdIsorec.[@\tanc{STLCIsorecProd-end}@]

### 2022-10-16 coqimpl.tex
compilation of family STLC
simplify

@@
-% \subsection{Language Design}
-
-% \subsection{Plugin Implementation}
+\input{stlc-compiled.tex}
+
+\begin{itemize}
+\item Parameterized by enclosing family "self": explicit extensibility hooks
+\item "self" includes only those fields in the context
+\item Extensible inductive type $\Rightarrow$ axiomatize
+\item Further-bindable \lsti{FRecursion} $\Rightarrow$ axiomatize
+\item Non-overridable definitions $\Rightarrow$ expose definitional equality
+\item Instantiate inductive types, partial recursors, and equalities $\Rightarrow$ OK to use recursor; equalities easily discharged by \lsti{reflexivity}
+\item Instantiate recursion/induction and equalities $\Rightarrow$ OK to use recursor; equalities easily discharged by \lsti{reflexivity}\\
+      Code reuse
+\end{itemize}
+
+\newpage
 

### 2022-10-16 stlc-compiled.tex
simplify

@@
 \begin{figure}
+\vspace{-5pt}
 \begin{minipage}{\textwidth}
@@
 \lstset{
-  basicstyle=\fontsize{8}{7.8}\ttfamily,
+  basicstyle=\fontsize{8}{8.2}\ttfamily,
 % numbers=left,
@@
 
-\newcommand{\innersep}{\vspace{4pt}}
-\newcommand{\outersep}{\vspace{4pt}}
+\newcommand{\innersep}{\vspace{6pt}}
+\newcommand{\outersep}{\vspace{5pt}}
 \newcommand{\outdent}{\!\!\!}
@@
 Inductive tm\_ : Set :=[@\tanc{c-STLC-d-tm-start}@]
-| tm_unit\_ : tm\_    | tm_var\_ : id -> tm\_
-| tm_abs\_  : id -> tm\_ -> tm\_
-| tm_app\_  : tm\_ -> tm\_ -> tm\_.
+| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
+| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_.
 Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
@@
 Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
-[@\hfill\codecomment{Instantiate \texttt{tm} partial recursor \& equalities}@]
-Def tm_prect_STLC P H1 H2 H3 H4 : \forall t, P t :=
-  fix F t := match t with
-  | tm_unit\_ => H1  | tm_var\_ x => H2 x
-  | tm_abs\_ x t => H3 x t (F t)
-  | tm_app\_ t1 t2 => H4 t1 (F t1) t2 (F t2) end.
-Fact[@\;@][@\textls[-35]{tm\_unit\_eq\_STLC}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
+[@\hfill\codecomment{Instantiate \texttt{tm} partial recursor \& its equalities}@]
+Def tm_prect_STLC P := tm\__rect (\lambda[@\,@]t,[@\,@]option[@\,@](P[@\,@]t)).
+Fact[@\;@][@\textls[-20]{tm\_unit\_eq\_STLC}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
 Include STLC\_subst\_Cases.[@\tanc{c-STLC-d-subst-start}@]
-[@\hfill\codecomment{Instantiate \texttt{subst} \& equalities}@]
-Def subst[@\,@]:[@\,@]tm[@\,@]->[@\,@]id[@\,@]->[@\,@]tm[@\,@]->[@\,@]tm :=[@\tanc{c-STLC-firstcol-end}\columnbreak@]
-  tm\__rect _ subst\_tm_unit subst\_tm_var[@\tanc{c-STLC-sndcol-start}@]
+Def subst[@\,@] :=[@\hfill\codecomment{Instantiate \texttt{subst} \& its equalities}@]
+  tm\__rect _ subst\_tm_unit subst\_tm_var
              subst\_tm_abs  subst\_tm_app.
-Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\innersep@]
-[@\dadada@][@\hfill\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@][@\tanc{c-STLC-d-ty}@][@\innersep@]
+Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}\columnbreak@][@\innersep@]
+[@\dadada@][@\hfill\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@][@\tanc{c-STLC-d-ty}@][@\tanc{c-STLC-sndcol-start}@][@\innersep@]
 Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLC-d-env}@][@\innersep@]
@@
 
-\makeline[0pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!50][1.25pt]
-\makeline[0pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!50][1.25pt]
-\makeline[0pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!50][1.25pt]
-\makeline[0pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!50][1.25pt]
-\makeline[0pt]{c-STLC-rest}{c-STLC-rest}[codecomment-color!50][1.25pt]
-\makeline[0pt]{c-STLC-start}{c-STLC-firstcol-end}[codecomment-color!50][1.25pt]
-\makeline[4pt]{c-STLC-d-tm-start}{c-STLC-d-tm-end}[codecomment-color!25][1.25pt]
-\makeline[4pt]{c-STLC-d-subst-start}{c-STLC-firstcol-end}[codecomment-color!20][1.25pt]
-\makeline[.5\textwidth+9pt]{c-STLC-sndcol-start}{c-STLC-end}[codecomment-color!50][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLC-sndcol-start}{c-STLC-d-subst-end}[codecomment-color!25][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLC-d-ty}{c-STLC-d-ty}[codecomment-color!25][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLC-d-env}{c-STLC-d-env}[codecomment-color!25][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLC-d-rest}{c-STLC-d-rest}[codecomment-color!25][1.25pt]
+\makeline[-1pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!55][1.8pt]
+\makeline[-1pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!55][1.8pt]
+\makeline[-1pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!55][1.8pt]
+\makeline[-1pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!55][1.8pt]
+\makeline[-1pt]{c-STLC-rest}{c-STLC-rest}[codecomment-color!55][1.8pt]
+\makeline[-1pt]{c-STLC-start}{c-STLC-firstcol-end}[codecomment-color!55][1.8pt]
+\makeline[4pt]{c-STLC-d-tm-start}{c-STLC-d-tm-end}[codecomment-color!35][1.0pt]
+\makeline[4pt]{c-STLC-d-subst-start}{c-STLC-firstcol-end}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+8pt]{c-STLC-sndcol-start}{c-STLC-end}[codecomment-color!55][1.8pt]
+\makeline[.5\textwidth+12pt]{c-STLC-d-ty}{c-STLC-d-ty}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+12pt]{c-STLC-d-env}{c-STLC-d-env}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+12pt]{c-STLC-d-rest}{c-STLC-d-rest}[codecomment-color!35][1.0pt]
 
-\vspace{-15pt}
+\vspace{-12pt}
 
@@
 Inductive tm\_ : Set :=[@\tanc{c-STLCFix-d-tm-start}@]
-| tm_unit\_ : tm\_    | tm_var\_ : id -> tm\_
-| tm_abs\_  : id -> tm\_ -> tm\_
-| tm_app\_  : tm\_ -> tm\_ -> tm\_
-| tm_fix\_  : id -> tm\_ -> tm\_.
+| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
+| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_
+| tm_fix\_[@\,@]:[@\,@]id->tm\_->tm\_.
 Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
@@
 Def tm_fix  := tm_fix\_.
-[@\dadada@][@\hfill\codecomment{Instantiate \texttt{tm} partial recursors \& equalities}@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
+[@\dadada@][@\hfill\codecomment{Instantiate \texttt{tm} partial recursors \& their equalities}@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
 Include STLCFix\_subst\_Cases. [@\tanc{c-STLCFix-d-subst-start}@]
-[@\hfill\codecomment{Instantiate \texttt{subst} \& equalities}@]
-Def subst[@\,@]:[@\,@]tm[@\,@]->[@\,@]id[@\,@]->[@\,@]tm[@\,@]->[@\,@]tm :=
+Def subst :=[@\hfill\codecomment{Instantiate \texttt{subst} \& its equalities}@]
   tm\__rect _ subst\_tm_unit subst\_tm_var
@@
 
-\makeline[.5\textwidth+9pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!50][1.25pt]
-\makeline[.5\textwidth+9pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!50][1.25pt]
-\makeline[.5\textwidth+9pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!50][1.25pt]
-\makeline[.5\textwidth+9pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!50][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-tm-start}{c-STLCFix-d-tm-end}[codecomment-color!25][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-subst-start}{c-STLCFix-d-subst-end}[codecomment-color!25][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-ty}{c-STLCFix-d-ty}[codecomment-color!25][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-env}{c-STLCFix-d-env}[codecomment-color!25][1.25pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-rest}{c-STLCFix-d-rest}[codecomment-color!25][1.25pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!55][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!55][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!55][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!55][1.8pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-tm-start}{c-STLCFix-d-tm-end}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-subst-start}{c-STLCFix-d-subst-end}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-ty}{c-STLCFix-d-ty}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-env}{c-STLCFix-d-env}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+13pt]{c-STLCFix-d-rest}{c-STLCFix-d-rest}[codecomment-color!35][1.0pt]
 
-\vspace{-15pt}
+\vspace{-12pt}
 
@@
 \begin{lstlisting}
-[@\outdent@][@\codecomment{Code emitted upon command \texttt{Check} \dadada}@][@\tanc{c-Check-STLCFix-typesafe-start}@]
+[@\outdent@][@\codecomment{Code emitted upon command \texttt{\textls[-35]{Check STLCFix.typesafe}}}@][@\tanc{c-Check-STLCFix-typesafe-start}@]
 [@\outdent@]Check STLCFix.typesafe.[@\tanc{c-Check-STLCFix-typesafe-end}@]
@@
 
-\makeline[.5\textwidth+9pt]{c-Check-STLCFix-typesafe-start}{c-Check-STLCFix-typesafe-end}[codecomment-color!35][1.25pt]
+\makeline[.5\textwidth+8pt]{c-Check-STLCFix-typesafe-start}{c-Check-STLCFix-typesafe-end}[codecomment-color!50][1.8pt]
 
@@
 
-\vspace{-8pt}
+%\vspace{-10pt}
 

### 2022-10-20 metatheory-condensed.tex
Save. Metatheory.

@@
 
-Now, we will mainly focus on the new typing rules that we introduced.
-For those curious about the complete formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
+Now, we will mainly focus on the newly introduced typing rules, and illustrate their connection to the plugin implementation.
+For those curious about the complete meta-theory formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
 Our metatheory is in an extensional setting.
@@
 $$
-We start with signature (with a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
-A signature is basically a list of types used to indicate the type of \textit{linkages}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
-Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement and will be explained later.
+We start with signature (in a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
+A signature is a list of types contained by a \textit{linkage}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
 We will use $\{a : T_1, b : T_2, ..\}$ as a shorthand for signature with type $T_1, T_2, ..$ for readability.
+
+Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
+Intuitively, sealing can abstract any given type by indicating a morphism to the interface.
+Sealing is a convenient device when programming using meta-theory, \textit{to indicate whether each field is overridable, non-overridable (thus definitionally exposed) or extensible}.
+For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
+
+However, there is subtle difference compared to the plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
+In other words, every field in meta-theory has a right to choose how they perceive and abstract the prior fields.
+While in the plugin, "Overridable" will decide how the current defining field is percived by the following fields.
 
@@
 
-The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
+% The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
 For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators).
@@
 
-We now sketch out the importance of the compilation type.
-First we assume a singleton type (universe)\footnote{Something like this singleton type universe will appear later in inductive type formulation.} $A : S(T) \iff A \equiv T$, i.e.
-a type of one single type.
+The importance of the compilation type lies in its difference from the linkage type.
+First we have a singleton type $"s A" : S(T) \iff "A" \equiv T$, i.e.
+a type of one single term.
 We can see that we have a term $x : \cC\{A : S(\bot), B : A \to \bot \} \vdash h_1 : \cC\{A : \cU, B:A \to \bot \}$ which is equivalent to $x : \Sigma (A: S(\bot)).
@@
 But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
+This example illustrates the programming difference between the sigma type and linkage type.
 
@@
 
-Our formulation of (extensible) inductive type is close to Wtype (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
-Thus we have the signature of inductive type collecting each constructors, with a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$.
+Our formulation of (extensible) inductive type is close to W-type (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
+Thus we have a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$ indicating signature of inductive type by its constructors.
 Unconventionally, a signature can induce a "universe" of types $\bW\tau$---in other words, $T : \bW\tau$ will be the inductive type instead of $\bW\tau$ itself.
@@
 
-Similar to plugin implementation, we delegate handlers for recursion to linkages.
-We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers.
-Specifying the handler family using "Wrec" on target inductive term $w$ leads to elimination.
+When programming via meta-theory, we usually seal the inductive type, only expose its constructors (as a mundane function member) and hide the eliminator by abstracting $T : \bW\tau$ into $"W"~T : \cU$.
+Then the later fields---those only rely on this interface with $\cU$---can be inherited to other contexts (e.g.
+using enriched inductive type), as long as the targeting contexts can be abstracted into this {interface} again.
+
+Similar to plugin implementation, we delegate handlers (and their inheritance structure) for recursion to linkages.
+We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers for all constructors.
+\ruleref{Rec-Constr} uses "RecSig" to achieve exhaustiveness checking on elimination.
+The partial recursors and the computational axioms are treated as normal fields as expected, since we know they are provable in the plugin implementation.
 
@@
 Inheritance judgement can be further enriched with inheriting operation, overriding operation, extending operation and etc.
-These operations correspond to inherit, override and extend command inside "Family" during plugin implementation.
+These operations correspond to inherit, override and extend command inside "Family" in our plugin implementation.
 
@@
 "Family"), of which the only usage is to apply \ruleref{Inh}.
-We don't yet consider non-trivial equality between inheritance judgements.
+We don't yet consider non-trivial equality between derivations of inheritance judgements.
 
@@
 % Extending Inductive Type
-We extend inductive type by overriding the inductive type with a enriched one.
-Thanks to sealing, most stuff from the parent family can be inherited to the children because enriched inductive type can export strictly stronger interface when abstracting.
+
+% Focus on FRecursion and FInduction
 
@@
 % Overriding + Pins
-Finally, we pointed out how overridable and pins is modeled if we admit equality reflection.
+Finally, we pointed out how overridable and pins is modeled..
 Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
 hide its definition when type-check t, and remaining former fields have their definition exposed in t.
-This above elaborated "hiding" syntax directly corresponding to the semantic of sealing---when definition is exposed, we expose propositional equality into the context, which will be a judgemental equality with the help of equality reflection; otherwise we only expose the type and no extra information.
+This above elaborated "hiding" syntax directly corresponding to the semantic of sealing---when definition is exposed, we expose a term by using singleton type, which will provide us a judgemental equality to the definition.

### 2022-10-20 syntax.tex
Save. Metatheory.

@@
 Finally, we make a note here on how sealing can easily express the overidability in our implementation.
-To expose the definition in our metatheory, we model that by adding equality reflection into our metatheory (and thus change our setting into extensional type theory).
-Then we can consider exposing the concrete definition as the existence of an identity type in the context.
+To expose the definition in our metatheory, we model that by using singleton type, where $"s A" : S(T) \iff "A" \equiv T$ .
 The key insight of the sealing in our metatheory, is that it gives the user the ability to manually specify which former fields expose its definition and which  former fields only expose their types---for those only expose their types, the corresponding identity type will be relaxed into a unit type.

### 2022-10-20 metatheory-condensed.tex
Save. Metatheory

@@
 However, there is subtle difference compared to the plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
-In other words, every field in meta-theory has a right to choose how they perceive and abstract the prior fields.
-While in the plugin, "Overridable" will decide how the current defining field is percived by the following fields.
+In other words, every field in meta-theory has the right to choose how they perceive and abstract the prior fields.
+While in the plugin, "Overridable" keyword will decide how the current defining field is percived by the following fields.
+
+We can now point out how "Overridable" and "pins" is modeled.
+Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
+hide its definition when type-check t, and remaining former fields have their definition exposed in t.
+This above elaborated "hiding" syntax directly corresponding to the $f$ in $(\mu^+ ~o~\{f\}~t)$, where each fields decides the abstraction of the prior.
 
@@
 
-Similar to plugin implementation, we delegate handlers (and their inheritance structure) for recursion to linkages.
+Similar to plugin implementation, we delegate handlers (and their reuse) for recursion to linkages (and their inheritance).
 We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers for all constructors.
@@
 \end{align*}
-
-\subsection{Relation to Plugin Implementation}
-Our core calculus can model most of plugin implementation.
-
-% Extending Inductive Type
-
-% Focus on FRecursion and FInduction
-
-% Partial Recursor + Computational Axiom
-The partial recursors and the computational axioms are no different from recursors and normal fields.
-We can prove partial recursor using current recursing mechanism and proving computational axiom on the trivial abstraction.
-After that, we can use sealing to abstract partial recursor and computational axiom again for the later fields.
-
-% Extending Recursion Handler
-The extension of recursion handlers is delegated by the inheritance of families. Adding new handlers for a new constructors is exactly extending families with new fields of certain types.
-
-% Overriding + Pins
-Finally, we pointed out how overridable and pins is modeled..
-Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
-hide its definition when type-check t, and remaining former fields have their definition exposed in t.
-This above elaborated "hiding" syntax directly corresponding to the semantic of sealing---when definition is exposed, we expose a term by using singleton type, which will provide us a judgemental equality to the definition.

### 2022-10-20 coq-example.tex
Save. Third Example.

@@
 \newpage\EDJ{Please delete this newpage command later}
-We provide two examples here for illustration and evaluation of our Coq plugin.
+We provide three examples here for illustration and evaluation of our Coq plugin.
 
@@
 computability in the presence of family.\YZ{Mention the alternative approach of using parameterized modules.}
+
+\subsection{An Extensible Decision Procedure for Parsing}
+The third example is a decision procedure for parsing.
+The basic idea is that we can use Cop's Inductive Facility to encode CFG rules.
+Then extensible inductive family corresponds to extensible CFG rules.
+For example, indexed type \mintinline{Coq}{is_prog : strs → Type} indicate a token list is a syntacitcally well-formed program.
+
+To prove \mintinline{Coq}{forall s, (is_prog s) + (is_prog s -> False)}, we use strong induction on the length and the non-trivial part is at the inductive step, i.e.
+we need to show \mintinline{Coq}{(forall e, len e < len s -> (is_prog e) + (is_prog e -> False)) -> (is_prog s) + (is_prog s -> False)}.
+
+Then we provide an overridable predicate \mintinline{Coq}{alg_is_prog : strs -> Type}, decidable provided the inductive hypothesis and sound \& complete w.r.t.
+to the specification \mintinline{Coq}{alg_is_prog e  <-> is_prog e}.
+We require extension to take care of the soundness \& completeness and the conditional decidablity.
+
+Among them, completeness \mintinline{Coq}{is_prog e -> alg_is_prog e} is the most challenging part.
+This is proved by recursion on the predicate "is_prog e", but most of the handler on "is_prog" will not be reused because \mintinline{Coq}{alg_is_prog e} is overriden during inheritance.
+Unlike completeness, other parts of the construction, including soundness and and conditional decidability can be reused easily.

### 2022-10-20 metatheory-condensed.tex
Save. Third Example.

@@
 Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Intuitively, sealing can abstract any given type by indicating a morphism to the interface.
-Sealing is a convenient device when programming using meta-theory, \textit{to indicate whether each field is overridable, non-overridable (thus definitionally exposed) or extensible}.
+Intuitively, when programming using meta-theory, sealing is a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.
 For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
+Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.
 
@@
 The only way to use $h$ is by \ruleref{Inh} and transforming a linkage of signature $\sigma_1$ to another linkage of $\sigma_2$.
-
-Inheritance judgement can be further enriched with inheriting operation, overriding operation, extending operation and etc.
-These operations correspond to inherit, override and extend command inside "Family" in our plugin implementation.
+Inheritance judgement can be inductively constructed via inheriting operations, overriding operations, extending operations and etc, corresponding to "Inherit", "Override" and "Extend" command  in our plugin implementation.
 

### 2022-10-20 coq-example.tex
The Parser Example.

@@
 
-To prove \mintinline{Coq}{forall s, (is_prog s) + (is_prog s -> False)}, we use strong induction on the length and the non-trivial part is at the inductive step, i.e.
-we need to show \mintinline{Coq}{(forall e, len e < len s -> (is_prog e) + (is_prog e -> False)) -> (is_prog s) + (is_prog s -> False)}.
+\begin{figure}[!htb]
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
+  }
+
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{2}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting}
+  FInductive is_prog : strs -> Type := 
+  | ip_seq : forall a b, is_prog a -> is_prog b 
+      -> is_prog (a ++ [";"] ++ b) 
+  | ip_assign : forall i e, is_atom i -> is_exp e 
+      -> is_prog ([i; "="] ++ e).  
+  \end{lstlisting}
+
+  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
+
+  \columnbreak
+  \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+  FInductive is_prog : strs -> Type += 
+  | ip_if : forall a b, is_exp a -> is_prog b 
+      -> is_prog (["if"] ++ a ++ ["then"] ++ b). 
+  \end{lstlisting}
+
+  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+  \caption{"is_prog" and its extension}
+  \end{figure}
+
+To prove \mintinline{Coq}{forall s, (is_prog s) + (is_prog s -> False)}, we use strong induction on the length and the non-trivial part is  the inductive step, i.e. we need to show
+\begin{minted}[fontsize=\footnotesize]{Coq}
+  (forall e, len e < len s -> (is_prog e) + (is_prog e -> False)) -> (is_prog s) + (is_prog s -> False)
+\end{minted}
 

### 2022-10-20 metatheory-condensed.tex
Questions in metatheory section

--- 2022-10-20 metatheory-condensed.tex
@@
 
-Now, we will mainly focus on the newly introduced typing rules, and illustrate their connection to the plugin implementation.
-For those curious about the complete meta-theory formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
-Our metatheory is in an extensional setting.
+Now, we will mainly focus on the new typing rules that we introduced.
+For those curious about the complete formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
+Our metatheory is in an extensional setting.\YZ{
+    Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
+    not definitionally equal but propositionally equal in family STLC?
+}
 
@@
 $$
-We start with signature (in a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
-A signature is a list of types contained by a \textit{linkage}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
+We start with signature (with a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
+A signature is basically a list of types used to indicate the type of \textit{linkages}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
+Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement and will be explained later.
 We will use $\{a : T_1, b : T_2, ..\}$ as a shorthand for signature with type $T_1, T_2, ..$ for readability.
 
-Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Intuitively, sealing can abstract any given type by indicating a morphism to the interface.
-Sealing is a convenient device when programming using meta-theory, \textit{to indicate whether each field is overridable, non-overridable (thus definitionally exposed) or extensible}.
-For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
-
-However, there is subtle difference compared to the plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
-In other words, every field in meta-theory has a right to choose how they perceive and abstract the prior fields.
-While in the plugin, "Overridable" will decide how the current defining field is percived by the following fields.
-
-Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited the linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
+Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited by linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
 The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
@@
 
-% The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
+The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
 For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators).
@@
 using extended inductive type), as long as the targeting contexts can be abstracted into the surrounding \textit{interface} for $t$ (i.e.
-$\Gamma, A$)  as well.
+$\Gamma, A$)  as well.\YZ{Is sealing used only for late-binding the signatures of inductive types?
+Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}
 
-The importance of the compilation type lies in its difference from the linkage type.
-First we have a singleton type $"s A" : S(T) \iff "A" \equiv T$, i.e.
-a type of one single term.
+We now sketch out the importance of the compilation type.
+First we assume a singleton type (universe)\footnote{Something like this singleton type universe will appear later in inductive type formulation.} $A : S(T) \iff A \equiv T$, i.e.
+a type of one single type.
 We can see that we have a term $x : \cC\{A : S(\bot), B : A \to \bot \} \vdash h_1 : \cC\{A : \cU, B:A \to \bot \}$ which is equivalent to $x : \Sigma (A: S(\bot)).
@@
 
-But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
-This example illustrates the programming difference between the sigma type and linkage type.
+But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited\YZ{How is this type inhabited?
+Through an identity seal?}
+and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
 
@@
 }
-{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
+{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv \Pi A\; \Pi (\Pi B (R[\pi_1\circ\pi_1]))\; R[\pi_1\circ\pi_1]}{}}
 \quad 
@@
 }
-{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)[\pi_1]}}
+{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)[\pi_1]}}\YZ{Why is this pi_1 substitution needed?}
 $$
@@
 \quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}
 }
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}\YZ{
-(1) Use \tau for inductive types and \sigma for linkage signatures.
-(2) Try using the same metavariables (j and n) as in the rules on W-types.
-(3) In Rec-Constr, no checking of \Gamma \vdash T : \bW \tau?
-(4) Why are RecSig1, RecSig, Wrec, prjR not listed in the syntax figure?
-}\EDJreply{Thx. Working on it. I seem to forgot to add inheritance as well.}\EDJreply{Done. Please Check.}
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
 $$
 
-Our formulation of (extensible) inductive type is close to W-type (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
-Thus we have a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$ indicating signature of inductive type by its constructors.
+Our formulation of (extensible) inductive type is close to Wtype (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
+Thus we have the signature of inductive type collecting each constructors, with a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$.
 Unconventionally, a signature can induce a "universe" of types $\bW\tau$---in other words, $T : \bW\tau$ will be the inductive type instead of $\bW\tau$ itself.
@@
 
-When programming via meta-theory, we usually seal the inductive type, only expose its constructors (as a mundane function member) and hide the eliminator by abstracting $T : \bW\tau$ into $"W"~T : \cU$.
-Then the later fields---those only rely on this interface with $\cU$---can be inherited to other contexts (e.g.
-using enriched inductive type), as long as the targeting contexts can be abstracted into this {interface} again.
-
-Similar to plugin implementation, we delegate handlers (and their inheritance structure) for recursion to linkages.
-We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers for all constructors.
-\ruleref{Rec-Constr} uses "RecSig" to achieve exhaustiveness checking on elimination.
-The partial recursors and the computational axioms are treated as normal fields as expected, since we know they are provable in the plugin implementation.
+Similar to plugin implementation, we delegate handlers for recursion to linkages.
+We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers.
+Specifying the handler family using "Wrec" on target inductive term $w$ leads to elimination.
 
@@
 Inheritance judgement can be further enriched with inheriting operation, overriding operation, extending operation and etc.
-These operations correspond to inherit, override and extend command inside "Family" in our plugin implementation.
+These operations correspond to inherit, override and extend command inside "Family" during plugin implementation.
 
@@
 "Family"), of which the only usage is to apply \ruleref{Inh}.
-We don't yet consider non-trivial equality between derivations of inheritance judgements.
+We don't yet consider non-trivial equality between inheritance judgements.
 
@@
 % Extending Inductive Type
-
-% Focus on FRecursion and FInduction
+We extend inductive type by overriding the inductive type with a enriched one.
+Thanks to sealing, most stuff from the parent family can be inherited to the children because enriched inductive type can export strictly stronger interface when abstracting.
 
@@
 % Overriding + Pins
-Finally, we pointed out how overridable and pins is modeled..
+Finally, we pointed out how overridable and pins is modeled if we admit equality reflection.
 Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
 hide its definition when type-check t, and remaining former fields have their definition exposed in t.
-This above elaborated "hiding" syntax directly corresponding to the semantic of sealing---when definition is exposed, we expose a term by using singleton type, which will provide us a judgemental equality to the definition.
+This above elaborated "hiding" syntax directly corresponding to the semantic of sealing---when definition is exposed, we expose propositional equality into the context, which will be a judgemental equality with the help of equality reflection; otherwise we only expose the type and no extra information.

### 2022-10-20 syntax.tex
Questions in metatheory section

--- 2022-10-12 syntax.tex
@@
 distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
-$\goodType{\Gamma}{El~T}{}$, and vice-versa.
+$\goodType{\Gamma}{El~T}{}$, and vice-versa.\YZ{Why is it wanted to distinguish types from their codes?}
 
@@
 \quad 
-\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}
+\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}\YZ{What is star? Is it just tau?}
 \quad

### 2022-10-20 metatheory-condensed.tex

--- 2022-10-20 metatheory-condensed.tex
@@
 For those curious about the complete meta-theory formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
-Our metatheory is in an extensional setting.
+Our metatheory is in an extensional setting.\YZ{
+    Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
+    not definitionally equal but propositionally equal in family STLC?
+}
 
@@
 For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
-Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.
+Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.\YZ{Is sealing used only for late-binding the signatures of inductive types?
+Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}
 
@@
 
-Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited the linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
+Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited by linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
 The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
@@
 
-But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
+But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited\YZ{How is this type inhabited?
+Through an identity seal?} and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
 This example illustrates the programming difference between the sigma type and linkage type.
@@
 \quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}
 }
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}\YZ{
-(1) Use \tau for inductive types and \sigma for linkage signatures.
-(2) Try using the same metavariables (j and n) as in the rules on W-types.
-(3) In Rec-Constr, no checking of \Gamma \vdash T : \bW \tau?
-(4) Why are RecSig1, RecSig, Wrec, prjR not listed in the syntax figure?
-}\EDJreply{Thx. Working on it. I seem to forgot to add inheritance as well.}\EDJreply{Done. Please Check.}
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
 $$

### 2022-10-21 metatheory-condensed.tex
More questions

@@
 }
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}\YZ{I suppose Rec-Constr is only for exhaustivity checking?
+If so, then would it be more intuitive to not include the w in the term?
+And I suppose the programmer will want to use sealing to hide its type, or otherwise it would fail to normalize for lack of a beta rule?}
 $$

### 2022-10-21 syntax.tex
More questions

@@
 \\\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cC \nu\cdot}
-\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}\YZ{nothing after colon?}\EDJreply{There should be something but the space is not enough.}
+\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}
 }
@@
 
-\YZ{sf' seems unnecessary, since sf' s ad s are the same?}\EDJreply{Yes.
+%\YZ{sf' seems unnecessary, since sf' s ad s are the same?}\EDJreply{Yes.
 Technically it is pretty unnecessary.
@@
 I initially having ``sf'' to just *emphasize* ``sf s'' should be considered as a term  and ``s'' a sealing}\YZreply{
-  I'm still confused about the difference among s, sf s and sf' s.
-  I thought s and sf' s were technically the same; they are both terms of type A[pi1] in a context that contains cC sigma.
-  I thought sf s was different because it is a substitution.
-}\EDJreply{You are absolutely right. s and sf' s are techincally the same. sf s is a substitution.
-  About s and sf' s, I can say that, I don't allow something like $\goodSeal{\Gamma}{sf'~s}{\sigma}{A}$. The reason I don't allow it is only notational consideration. 
-  I mean technically we can consider them the same. I can just remove all the appearances of sf' if you want and find it less confusing.
-}\YZreply{Yes, sf' is confusing to me.}\EDJreply{sf' notation removed. Please check.}
+%  I'm still confused about the difference among s, sf s and sf' s.
+%  I thought s and sf' s were technically the same; they are both terms of type A[pi1] in a context that contains cC sigma.
+%  I thought sf s was different because it is a substitution.
+%}\EDJreply{You are absolutely right. s and sf' s are techincally the same. sf s is a substitution.
+%  About s and sf' s, I can say that, I don't allow something like $\goodSeal{\Gamma}{sf'~s}{\sigma}{A}$. The reason I don't allow it is only notational consideration. 
+%  I mean technically we can consider them the same. I can just remove all the appearances of sf' if you want and find it less confusing.
+%}\YZreply{Yes, sf' is confusing to me.}\EDJreply{sf' notation removed. Please check.}
 Note that, given $\goodSeal{\Gamma}{s}{\sigma}{A} = \goodTerm{\Gamma, \cC\sigma}{s}{A}$, notation "sf"
@@
 of signature $\sigma$, the next added field $t : T$ will have
-to be in the context of the abstraction $A$\YZ{Why do Lnkg Add and Sig Add leave A unspecified?
+to be in the context of the abstraction $A$
+% \YZ{Why do Lnkg Add and Sig Add leave A unspecified?
 I thought this A corresponds to 'self'.}\EDJreply{You are asking where is the function \cC \{a : \bW sigma ; ..\} to \cC\{a : \cU ; ..\}, which by definition is requesting (Sigma (a : \bW sigma ; ..)) to (Sigma (a : \cU ; ..)).
 \\
-This is something doable in vanilla MLTT, with the help of transforming \bW sigma to \cU, by rule `Ind-Type'. And this function's type should be directly the sealing type.}\YZreply{
-  My confusion is this: it seems that in Lnkg Add and Sig Add, A could well be the top type and f a value of the top type. In this sense, A seems unspecified.
-  Or maybe in the metatheory, it is the programmer's responsibility to find this A?
-}\EDJreply{You are absolutely right.
+% This is something doable in vanilla MLTT, with the help of transforming \bW sigma to \cU, by rule `Ind-Type'. And this function's type should be directly the sealing type.}\YZreply{
+%   My confusion is this: it seems that in Lnkg Add and Sig Add, A could well be the top type and f a value of the top type. In this sense, A seems unspecified.
+%   Or maybe in the metatheory, it is the programmer's responsibility to find this A?
+% }\EDJreply{You are absolutely right.
 In metatheory it is programmer's responsibility to specify the A they want.
@@
 \\
-Also the `f' is specified together with `A' by the user as well. \\
-A bit off-topic : this is not the only place that such specification is postponed to the user (in metatheory) another place is in inheritance when specifying upcasting function.}
+% Also the `f' is specified together with `A' by the user as well. \\
+% A bit off-topic : this is not the only place that such specification is postponed to the user (in metatheory) another place is in inheritance when specifying upcasting function.}
 of the compilation $\cC\sigma$ (witnessed by the seal $f$).
@@
 }
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}\YZ{
-(1) Use \tau for inductive types and \sigma for linkage signatures.
-(2) Try using the same metavariables (j and n) as in the rules on W-types.
-(3) In Rec-Constr, no checking of \Gamma \vdash T : \bW \tau?
-(4) Why are RecSig1, RecSig, Wrec, prjR not listed in the syntax figure?
-}\EDJreply{Thx. Working on it. I seem to forgot to add inheritance as well.}\EDJreply{Done. Please Check.}
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
 $$
@@
 }
-{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}
+{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited?}
 $$

### 2022-10-21 metatheory-condensed.tex
Answer questions.

@@
 For those curious about the complete meta-theory formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
-Our metatheory is in an extensional setting.\YZ{
+Our metatheory is formulated using QIIT in an extensional setting.\YZ{
     Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
     not definitionally equal but propositionally equal in family STLC?
+}\EDJreply{No, TLDR: it is a different level. \\ 
+There are two type theories, one is Agda-like (= Latex = Conventional math), the other is the one formulated in QIIT.
+This sentence is saying that the Agda-like one is extensional MLTT (i.e.
+we have function extensionality and etc here in Latex writing directly).
+\\ But the one you are asking about is in the metatheory in QIIT: if we have an identity type $Id x y$ (propositionally equal) can we have a judgemetnal equality for the two terms in QIIT $x \equiv y$.
+\\
+Apparently, we don't setup our QIIT formulation in this way (I thought we need, but now I don't think we need it. Our QIIT formulation is purely in intensional MLTT now).\\ 
+But even if we do setup like that, I will not say a conflict is happening (because it is to our advantage if we have judgemental equal in metatheory).
+I will say it is a bit off as to how Coq is formulated (Coq is closer to an intensional MLTT, so no judgemental equal only propositional one).
 }
@@
 Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.\YZ{Is sealing used only for late-binding the signatures of inductive types?
-Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}
+Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}\EDJreply{All late-bound are achieved by ``selective-abstraction'' imposed by sealing.
+For example, recursion will be consequently abstracted because of the abstraction of the inductive type it recurse on.
+}
 
@@
 But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited\YZ{How is this type inhabited?
-Through an identity seal?} and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
+Through an identity seal?}\EDJreply{Yes.
+a = $\bot$, b is identity.
+$S(\bot)$ is a singleton type with only bottom type as element inside} and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
 This example illustrates the programming difference between the sigma type and linkage type.
@@
 }
-{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)[\pi_1]}}\YZ{Why is this pi_1 substitution needed?}
+{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}\YZ{Why is this pi_1 substitution needed?}\EDJreply{You are right, I think it is a typo.}
 $$
@@
 \quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}\EDJreply{Yes.}
 }

### 2022-10-21 metatheory-condensed.tex
Answer more.

@@
 If so, then would it be more intuitive to not include the w in the term?
-And I suppose the programmer will want to use sealing to hide its type, or otherwise it would fail to normalize for lack of a beta rule?}
+And I suppose the programmer will want to use sealing to hide its type, or otherwise it would fail to normalize for lack of a beta rule?}\EDJreply{ 1.
+Yes.
+But not include w will cause $"Wrec"~r$ be a function type.
+we are not supposed to add a new intro-rule for function type(or for any other type), this is bad meta-theory design.
+\\
+  2.
+It is actually the other way around -- if things are abstracted, then no beta rule can apply.
+More concretely, $T : \bW\tau$ is abstracted into $T' : \cU$, then this later one we cannot use $"Wrec"$ on it for sure.
+But once we have $\{T : \bW\tau, "rec" : \cL(RecSig~\tau)\}$ we are okay, and this, after wrap "rec" with an lambda abstraction, we will have $\cC\{T : \bW\tau, "rec" : \cL(RecSig~\tau~R), "recf" : T -> R\}$ as compilation type.
+Then with sealing abstraction, we will have $\Sigma(T : U, \Sigma ("rec" : ?, \Sigma("ref" : T -> R)))$
+}
 $$

### 2022-10-21 metatheory-condensed.tex
More reply+ Rephrase sealing.

@@
 Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Intuitively, when programming using meta-theory, sealing is a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.
-For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
-Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.\YZ{Is sealing used only for late-binding the signatures of inductive types?
+Sealing is used as a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.
+\YZ{Is sealing used only for late-binding the signatures of inductive types?
 Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}\EDJreply{All late-bound are achieved by ``selective-abstraction'' imposed by sealing.
 For example, recursion will be consequently abstracted because of the abstraction of the inductive type it recurse on.
-}
+}\EDJreply{Sorry, I see what you mean.
+Yes.
+the box part is important for late-binding.
+But the functionality of sealing is always there.
+They are both required to achieve overriding.
+A illustrative example is , if the type is $S(\bot)$, then even with the boxed premise, it is still not overridable because the type itself restricts all the possibilities.}
 
-However, there is subtle difference compared to the plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
-In other words, every field in meta-theory has the right to choose how they perceive and abstract the prior fields.
-While in the plugin, "Overridable" keyword will decide how the current defining field is percived by the following fields.
+% For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
+Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.
 
-We can now point out how "Overridable" and "pins" is modeled.
+% We can now point out how "Overridable" and "pins" is modeled.
 Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
 hide its definition when type-check t, and remaining former fields have their definition exposed in t.
-This above elaborated "hiding" syntax directly corresponding to the $f$ in $(\mu^+ ~o~\{f\}~t)$, where each fields decides the abstraction of the prior.
+% This above elaborated "hiding" syntax directly corresponding to the $f$ in $(\mu^+ ~o~\{f\}~t)$, where each fields decides the abstraction of the prior.
+
+Intuitively, sealing corresponds to the following surface syntax\\
+\mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...} where every former field "x","y","z",...
+hide its definition when type-check "t", and remaining former fields have their definition exposed for "t".
+Thus, "x","y","z", ...
+can be overridden but other fields have to stay the same if we want to inherit "t".
+In meta-theory, we control the sealing $f$ so that "x","y","z",...
+only has type exposed and remaining fields are definitionally exposed using singleton type.
+
+The reader should feel this subtle difference compared to our current plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
+In other words, every field in meta-theory has the right to choose how they perceive and abstract the prior fields.
+While in the plugin, "Overridable" keyword will decide how the current defining field is percived by all the following fields.
 

### 2022-10-21 metatheory-condensed.tex
respond to or close comments

@@
 For those curious about the complete meta-theory formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
-Our metatheory is formulated using QIIT in an extensional setting.\YZ{
-    Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
-    not definitionally equal but propositionally equal in family STLC?
-}\EDJreply{No, TLDR: it is a different level. \\ 
-There are two type theories, one is Agda-like (= Latex = Conventional math), the other is the one formulated in QIIT.
+Our metatheory is formulated using QIIT in an extensional setting.
+%\YZ{
+%    Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
+%    not definitionally equal but propositionally equal in family STLC?
+%}\EDJreply{No, TLDR: it is a different level. \\ 
+%There are two type theories, one is Agda-like (= Latex = Conventional math), the other is the one formulated in QIIT.
 This sentence is saying that the Agda-like one is extensional MLTT (i.e.
@@
 \\
-Apparently, we don't setup our QIIT formulation in this way (I thought we need, but now I don't think we need it. Our QIIT formulation is purely in intensional MLTT now).\\ 
-But even if we do setup like that, I will not say a conflict is happening (because it is to our advantage if we have judgemental equal in metatheory).
+%Apparently, we don't setup our QIIT formulation in this way (I thought we need, but now I don't think we need it. Our QIIT formulation is purely in intensional MLTT now).\\ 
+%But even if we do setup like that, I will not say a conflict is happening (because it is to our advantage if we have judgemental equal in metatheory).
 I will say it is a bit off as to how Coq is formulated (Coq is closer to an intensional MLTT, so no judgemental equal only propositional one).
-}
+%}
 
@@
 Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Sealing is used as a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.
-\YZ{Is sealing used only for late-binding the signatures of inductive types?
+Sealing is used as a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.\YZ{Is sealing used only for late-binding the signatures of inductive types?
 Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}\EDJreply{All late-bound are achieved by ``selective-abstraction'' imposed by sealing.
@@
 They are both required to achieve overriding.
-A illustrative example is , if the type is $S(\bot)$, then even with the boxed premise, it is still not overridable because the type itself restricts all the possibilities.}
+A illustrative example is, if the type is $S(\bot)$, then even with the boxed premise, it is still not overridable because the type itself restricts all the possibilities.}\YZreply{
+  By "selective abstraction", do you mean in the metatheory the programmer chooses what and how to abstract?
+}
 
@@
 
-But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited\YZ{How is this type inhabited?
-Through an identity seal?}\EDJreply{Yes.
-a = $\bot$, b is identity.
-$S(\bot)$ is a singleton type with only bottom type as element inside} and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
+But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited
+and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
 This example illustrates the programming difference between the sigma type and linkage type.
@@
 However, this form of abstracting a type is very important when programming in our metatheory---as later we will show inductive type of different constructors locates in different type (universe), and to make later fields inheritable we need to work on an interface of abstracted type.
-Thus the compilation type is indispensible.
+Thus the compilation type is indispensable.
 
@@
 }
-{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}\YZ{Why is this pi_1 substitution needed?}\EDJreply{You are right, I think it is a typo.}
+{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}
 $$
@@
 \quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}\EDJreply{Yes.}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}\EDJreply{Yes.}\YZreply{Should the redundant premises be removed?}
 }
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}\YZ{I suppose Rec-Constr is only for exhaustivity checking?
+{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
+\YZ{I suppose Rec-Constr is only for exhaustivity checking?
 If so, then would it be more intuitive to not include the w in the term?
@@
 Then with sealing abstraction, we will have $\Sigma(T : U, \Sigma ("rec" : ?, \Sigma("ref" : T -> R)))$
+}\YZreply{
+  So there IS a beta rule for recursor (but you don't show it), right?
 }

### 2022-10-21 metatheory-condensed.tex
Reply more.

@@
   By "selective abstraction", do you mean in the metatheory the programmer chooses what and how to abstract?
-}
+}\EDJreply{Yes.}
 
@@
 \Rule[name=Rec-Constr]
-{ \goodWSig{\Gamma}{\tau}{N}
+{ 
+  % \goodWSig{\Gamma}{\tau}{N}
 \quad \goodTerm{\Gamma}{T}{\bW \tau}
-\quad \goodType{\Gamma}{R}{}
+% \quad \goodType{\Gamma}{R}{}
 \quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}\EDJreply{Yes.}\YZreply{Should the redundant premises be removed?}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}\EDJreply{Yes.}\YZreply{Should the redundant premises be removed?}\EDJreply{Let's remove it.
+Save Space for another rule.
+What other rules do you want?}
 }
@@
   So there IS a beta rule for recursor (but you don't show it), right?
+}\EDJreply{
+  Yes. 
+  % https://github.com/DKXXXL/Redstone/blob/46c2420982e1fac50138924cdd7e152786cd38c8/Sketch/Extensible-Module-1/Syntax.agda#L589 \\ 
+  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$.
 }

### 2022-10-21 canonicitymodel.tex
Small fixes through out.
comments

@@
 We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}}, 
-then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.
+then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.\YZ{Can we say something about strong/weak normalization?}
 
@@
 The proof can be found in appendix.
+The intuition of the model is similar to that of the consistency model---use the model of sigma type for linkage type and using Agda's inductive facility to justify our W-type.
 

### 2022-10-21 metatheory-condensed.tex
More reply.
comments

@@
 Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Sealing is used as a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.\YZ{Is sealing used only for late-binding the signatures of inductive types?
+Sealing is used as a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.
+%\YZ{Is sealing used only for late-binding the signatures of inductive types?
 Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}\EDJreply{All late-bound are achieved by ``selective-abstraction'' imposed by sealing.
 For example, recursion will be consequently abstracted because of the abstraction of the inductive type it recurse on.
-}\EDJreply{Sorry, I see what you mean.
+%}\EDJreply{Sorry, I see what you mean.
 Yes.
@@
 A illustrative example is, if the type is $S(\bot)$, then even with the boxed premise, it is still not overridable because the type itself restricts all the possibilities.}\YZreply{
-  By "selective abstraction", do you mean in the metatheory the programmer chooses what and how to abstract?
-}\EDJreply{Yes.}
+%  By "selective abstraction", do you mean in the metatheory the programmer chooses what and how to abstract?
+%}\EDJreply{Yes.}
 
@@
 \quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}\YZ{Is the 1st premise implied by the 2nd, and the 3rd implied by the 4th?}\EDJreply{Yes.}\YZreply{Should the redundant premises be removed?}\EDJreply{Let's remove it.
-Save Space for another rule.
-What other rules do you want?}
+\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
 }
@@
   % https://github.com/DKXXXL/Redstone/blob/46c2420982e1fac50138924cdd7e152786cd38c8/Sketch/Extensible-Module-1/Syntax.agda#L589 \\ 
-  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$.
-}
+  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$. \\
+  We also have canonicity (i.e. a closed term of boolean reduces to true or false), so there is definitely a beta rule for inductive type.
+}\YZreply{I'd add typeset this beta rule}
 $$

### 2022-10-21 standardmodel.tex
comments

@@
 Finally, notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
-and that is a contradiction.
+and that is a contradiction.\YZ{Explicitly state the theorem.}
 

### 2022-10-21 canonicitymodel.tex
More reply.

@@
 We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}}, 
-then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.\YZ{Can we say something about strong/weak normalization?}
+then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.\YZ{Can we say something about strong/weak normalization?}\EDJreply{
+  No. \\
+  1.
+In the declarative system, we cannot talk about strong/weak normalization (because we don't have operational semantic, if you recall the definition of strong/weak normalization, it based on ).
+So any review says strong normalization to you please stop them.
+\\
+  2. In declarative system, we only say ``normalization'' directly, which means, we construct a function that \\
+      for arbitrary class of (judgemental-equivalent) terms, it will return one syntax representation (called normal-form) for each class \\
+    (apparently most people will make the syntax representation  without beta redex, however, interestingly, they may do eta expansion instead of eta reduction) \\ 
+  3. So translation won't work because it compiles away the normal form of linkage (the $\mu^+$). This part requires direct proof \\
+  4. Of course we can conjecture we have that normalization function \\
+  5.
+Also we do have close term normalization, which is canonicity.
+Or more operationally speaking, we have termination (a runnable interpreter) for close term (since QIIT is runnable, the canoncity model ), \\
+  but we don't have termination (a runnable interpreter) for open terms.  \\ 
+  So earlier you say ``normalize'' a recursion on an inductive type, that is no doable -- if the programmer wants to program on meta-theory, they will have to manually write out the a chain of judgemental equality (with a bunch of beta rules inside) by themselves.
+}
 

### 2022-10-21 metatheory-condensed.tex
More reply.

@@
   We also have canonicity (i.e. a closed term of boolean reduces to true or false), so there is definitely a beta rule for inductive type.
-}\YZreply{I'd add typeset this beta rule}
+}\YZreply{I'd add typeset this beta rule}\EDJreply{I just realize you use the words normalize. I will say currently this system cannot ``reduce'' any open term. But can only ``reduce'' close term. \\ 
+When the programmer program on this meta-theory, they don't need canonicity or normalization at all (they write out a chain of  judgemental equality by themselves).
+If they program on a proof assistant, then our proof assistant will need normalization to do type checking.
+}
 $$

### 2022-10-21 canonicitymodel.tex
Modify Reply.

@@
   1.
-In the declarative system, we cannot talk about strong/weak normalization (because we don't have operational semantic, if you recall the definition of strong/weak normalization, it based on ).
-So any review says strong normalization to you please stop them.
+In the declarative system, we cannot talk about strong/weak normalization (because we don't have operational semantic, if you recall the definition of strong/weak normalization, it based on reduction/path of steps/op semantic).
+So any review says strong/weak normalization to you, please stop them.
 \\
@@
       for arbitrary class of (judgemental-equivalent) terms, it will return one syntax representation (called normal-form) for each class \\
-    (apparently most people will make the syntax representation  without beta redex, however, interestingly, they may do eta expansion instead of eta reduction) \\ 
-  3. So translation won't work because it compiles away the normal form of linkage (the $\mu^+$). This part requires direct proof \\
+    (apparently most people will make the normal form  without beta redex, however, interestingly, they may do eta expansion instead of eta reduction because theoretically easier) \\ 
+  3.
+So translation won't work because it compiles away the normal form of linkage (the $\mu^+$).
+This part requires a direct proof which construct normal form and neutral form for our calculus, and a normalization model from which we extract a runnable interpreter for open terms \\
   4. Of course we can conjecture we have that normalization function \\
@@
 Also we do have close term normalization, which is canonicity.
-Or more operationally speaking, we have termination (a runnable interpreter) for close term (since QIIT is runnable, the canoncity model ), \\
+Or more operationally speaking, we have termination (a runnable interpreter) for close term (since QIIT is runnable, the canoncity model itself is the runnable interpreter), \\
   but we don't have termination (a runnable interpreter) for open terms.  \\ 

### 2022-10-21 syntactic-example.tex
comments

@@
 parts and making "O" and "S" standing alone.
-"pred" will map $S~n \mapsto (n, S~n)$ and $O \mapsto (O, O)$, and thus a predecessor.
+"pred" will map $S~n \mapsto (n, S~n)$ and $O \mapsto (O, O)$, and thus a predecessor.\YZ{
+  Using family polymorphism for adding  a successor constructor is unnatural.
+  I suggest simply renaming the constructors and not mentioning the correspondence to natural numbers.
+}
 
@@
     \end{minted}
-  \end{minipage}
+  \end{minipage}\YZ{This figure needs updating; it should use the same syntax as in §3.}\YZreply{Also, this figure should contain an example of accessing a field outside its family.}
   \caption{Demonstrating Pseudo-Code}

### 2022-10-21 coq-example.tex
comments

@@
 
-Then we provide an overridable predicate \mintinline{Coq}{alg_is_prog : strs -> Type}, decidable provided the inductive hypothesis and sound \& complete w.r.t.
+Then we provide an overridable predicate \mintinline{Coq}{alg_is_prog : strs -> Type},\YZ{Why does it need to be overridden?}
+decidable provided the inductive hypothesis and sound \& complete w.r.t.
 to the specification \mintinline{Coq}{alg_is_prog e  <-> is_prog e}.
@@
 
-Among them, completeness \mintinline{Coq}{is_prog e -> alg_is_prog e} is the most challenging part.
-This is proved by recursion on the predicate "is_prog e", but most of the handler on "is_prog" will not be reused because \mintinline{Coq}{alg_is_prog e} is overriden during inheritance.
-Unlike completeness, other parts of the construction, including soundness and and conditional decidability can be reused easily.
+Among them, completeness \mintinline{Coq}{is_prog e -> alg_is_prog e} is
+the most challenging part. This is proved by recursion on the predicate
+"is_prog e", but most of the handler on "is_prog" will not be reused
+because \mintinline{Coq}{alg_is_prog e} is overriden during
+inheritance.\YZ{
+    If alg_is_prog e is overridden, then how can ANY handler be reused?
+}
+Unlike completeness, other parts of the construction, including
+soundness and and conditional decidability can be reused easily.
+\YZ{But is the reusability achieved through a derived family's reusing handlers in the base family? Doesn't sound like it because the induction seems to be on strs.}

### 2022-10-21 coq-example.tex
Reply the comments.

--- 2022-10-20 coq-example.tex
@@
 And at the end, the compiled module for "LangTyping" and "LangCP" is put into tests over simple queries. These "LangTyping" and "LangCP" also illustrates that we still have
-computability in the presence of family.\YZ{Mention the alternative approach of using parameterized modules.}
+computability in the presence of family.\YZ{Mention the alternative approach of using parameterized modules.}\EDJreply{What.
+I don't have that version.
+My earlier version without family is not extensible, just a POC to see if the proof idea itself is correct.}
 
@@
 
-Then we provide an overridable predicate \mintinline{Coq}{alg_is_prog : strs -> Type}, decidable provided the inductive hypothesis and sound \& complete w.r.t.
+Then we provide an overridable predicate \mintinline{Coq}{alg_is_prog : strs -> Type},\YZ{Why does it need to be overridden?}\EDJreply{Basically alg_is_prog is a big disjunction list, corresponding to each of the constructor of }
+decidable provided the inductive hypothesis and sound \& complete w.r.t.
 to the specification \mintinline{Coq}{alg_is_prog e  <-> is_prog e}.
-We require extension to take care of the soundness \& completeness and the conditional decidablity.
+"alg_is_prog e" will be a big disjunction, of which each disjunct $P_i$ corresponds to the argument of one constructor of "is_prog" (each CFG rule).
+Thus once "is_prog" is extended, we need to disjunct "alg_is_prog" with more condition, thus overridable.
+We require extension to take care of the disjuncts, their soundness \& completeness and the conditional decidablity.
 
-Among them, completeness \mintinline{Coq}{is_prog e -> alg_is_prog e} is the most challenging part.
-This is proved by recursion on the predicate "is_prog e", but most of the handler on "is_prog" will not be reused because \mintinline{Coq}{alg_is_prog e} is overriden during inheritance.
-Unlike completeness, other parts of the construction, including soundness and and conditional decidability can be reused easily.
+Among them, completeness \mintinline{Coq}{is_prog e -> alg_is_prog e} is
+the most challenging part. This is proved by recursion on the predicate
+"is_prog e", but none of the handler on "is_prog" will be reused
+because \mintinline{Coq}{alg_is_prog e} is overriden during
+inheritance.\YZ{
+    If alg_is_prog e is overridden, then how can ANY handler be reused?
+}\EDJreply{You are right. I rephrase into ``none of the handlers''.}
+Unlike completeness, other parts of the construction, including
+soundness and and conditional decidability can be reused easily.
+\YZ{But is the reusability achieved through a derived family's reusing handlers in the base family?
+Doesn't sound like it because the induction seems to be on strs.}\EDJreply{to prove something like alg_is_prog e -> is_prog e is not doing induction on string.
+It is just unfolding definition.
+\\
+alg_is_prog is a big disjunction, alg_is_prog e = p1 e \/ p2 e\/ ..., and during overriding this disunction list is increasing. \\
+So we organize the program by proving ``p1 e -> is_prog e'', ``p2 e -> is_prog e''.
+Everytime this disjunction list is increasing during overriding, the old proof ``p1 e -> is_prog e'' is reused in the new .}

### 2022-10-21 coq-example.tex
Questions

@@
 because \mintinline{Coq}{alg_is_prog e} is overriden during
-inheritance.\YZ{
-    If alg_is_prog e is overridden, then how can ANY handler be reused?
-}\EDJreply{You are right. I rephrase into ``none of the handlers''.}
+inheritance.
 Unlike completeness, other parts of the construction, including

### 2022-10-21 metatheory-condensed.tex
Questions

--- 2022-10-21 metatheory-condensed.tex
@@
 }\YZreply{
-  So there IS a beta rule for recursor (but you don't show it), right?
-}\EDJreply{
-  Yes. 
-  % https://github.com/DKXXXL/Redstone/blob/46c2420982e1fac50138924cdd7e152786cd38c8/Sketch/Extensible-Module-1/Syntax.agda#L589 \\ 
-  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$. \\
-  We also have canonicity (i.e. a closed term of boolean reduces to true or false), so there is definitely a beta rule for inductive type.
-}\YZreply{I'd add typeset this beta rule}
+  What happens if the programmer forgets to seal "rec"? I suppose rec
+  could still be inherited into a derived family. But it would not be
+  useful if the inductive type is extended there, because rec is
+  applicable only to the inductive type defined in the base family.
+  Orthogonally, can you add the beta rule?
+}
+%\YZreply{
+%  So there IS a beta rule for recursor (but you don't show it), right?
+%}\EDJreply{
+%  Yes. 
+%  % https://github.com/DKXXXL/Redstone/blob/46c2420982e1fac50138924cdd7e152786cd38c8/Sketch/Extensible-Module-1/Syntax.agda#L589 \\ 
+%  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$. \\
+%  We also have canonicity (i.e. a closed term of boolean reduces to true or false), so there is definitely a beta rule for inductive type.
+%}\YZreply{I'd add typeset this beta rule}\EDJreply{I just realize you use the words normalize.
+I will say currently this system cannot ``reduce'' any open term.
+But can only ``reduce'' close term.
+\\
+%When the programmer program on this meta-theory, they don't need canonicity or normalization at all (they write out a chain of  judgemental equality by themselves).
+If they program on a proof assistant, then our proof assistant will need normalization to do type checking.
+%
 $$

### 2022-10-21 metatheory-condensed.tex
Reply the comment

@@
   Orthogonally, can you add the beta rule?
+}\EDJreply{
+  You are right.
+Since "rec" has type "\cL(RecSig~\tau~R)", after inheritance, it is basically the same type "rec : \cL(RecSig~\tau[k]~R[k])" for some "k" stands for the upcasting.
+but the most important $\tau[k]$ has an  unchanged shape (like the number of constructors) so it is still a recursor for inductive type of signature $\tau$.
+Once the parent inductive type $\bW\tau$ is extended $\bW\tau_2$ in the children family, with enriched signature $\tau_2$, then because of the signature inconsistency with $\tau[k]$, this old recursor is not useful for this new inductive type.
+\\
+  In other words, this recursor linkage "rec" can be used for any inductive type with signature $\tau$.\\
+  Yes I will add beta rule. I see the last comment.
 }

### 2022-10-22 syntactic-example.tex
Change Syntax for the Examplar usage.

@@
   I suggest simply renaming the constructors and not mentioning the correspondence to natural numbers.
+}\EDJreply{I don't have a good idea for the renaming. What about you?}
+
+\begin{figure}[!htb]\label{fig:example-pseudocode}
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
   }
 
-\begin{figure}[H]\label{fig:example-pseudocode}
-\begin{minipage}[t]{0.5\linewidth}
-\begin{minted}[escapeinside=@@]{Coq}
-Family A.
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{2}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting} 
+  Family A. [@\tanc{Coq-Example1-start}@]
   FInductive N := O : N. 
- (* Osig, sig₁, obj₁, sig₂,
-    O₁, obj₂, sig₂', sl₁ *)
- Family handler.
-   O :self[A].N × self[A].N 
-     := (self[A].O, self[A].O).
+  (* [@ \tiny{Osig, sig₁, obj₁, sig₂,
+     O₁, obj₂, sig₂', sl₁} @] *)
 
- EndFamily. (* pN₁, handler₁, sig₃ *)
- FRecursor pred about self[A].N 
-  motive (fun _ => self[A].N × self[A].N)
-  using handler.
- (* pred
-  : self[A].N → self[A].N × self[A].N *)
-  (* predT, pred₁, sig₄, sig₄', sl₄ *) 
- k := self[A].pred self[A].O 
-(* pN₂, sig₅ *)
-EndFamily.
-\end{minted}
-  \end{minipage}
-  \begin{minipage}[t]{0.45\linewidth}
-\begin{minted}[escapeinside=@@]{Coq}
-Family A₂ extends A.
- Extend FInductive N := S : N → N.
- (* Nsig, sig₁₁, inh₁, O₂, sig₂₁, inh₂, 
-    sig₂₂, sig₂₁', sig₂₂', sl₂₂, inh₃ *)
- Extend Family handler.
-   Inherit O. (* inhᵢₙ₀, pN₃ *)
-   S : self[A].N × self[A].N
-       → self[A].N × self[A].N 
-     := fun (x,y) => (y, self[A₂].S y).
- EndFamily. (* inhᵢₙ, sig₃₁, inh₄*)
+  FRecursion pred on N
+    motive (fun _ => N × N).
+  Case O := (O, O). 
+  End pred. (* [@ \tiny{pN₁, handler₁, sig₃} @] *)
+  (* pred : self[A].N → self[A].N × self[A].N *)
+   (* [@ \tiny{predT, pred₁, sig₄, sig₄', sl₄} @] *) 
+  k := pred O 
+  (* [@ \tiny{pN₂, sig₅} @] *)  
+  End A.[@\tanc{Coq-Example1-end}@]
+  \end{lstlisting}
+
+  \makeline[0pt]{Coq-Example1-start}{Coq-Example1-end}[codecomment-color!50]
+
+  \columnbreak
+
+  \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+  Family [@ \tiny{"A"$_2$} @] extends A. [@\tanc{Coq-Example2-start}@]
+  FInductive N += S : N → N.
+  (* [@ \tiny{Nsig, sig₁₁, inh₁, O₂, sig₂₁, inh₂, 
+      sig₂₂, sig₂₁', sig₂₂', sl₂₂, inh₃} @] *)
+  FRecursion pred.
+  Inherit O. (* [@ \tiny{inhᵢₙ₀, pN₃} @] *)
+  Case S := fun (x,y) => (y, S y).
+  EndFamily. (* [@ \tiny{inhᵢₙ, sig₃₁, inh₄} @] *)
   Inherit pred.
-
-  (* sig₄₁, inh₅, sl₄₁ *)
+    (* [@ \tiny{sig₄₁, inh₅, sl₄₁} @] *)
   Inherits k. 
   (* inh₆ *)
-EndFamily.
-    \end{minted}
-  \end{minipage}\YZ{This figure needs updating; it should use the same syntax as in §3.}\YZreply{Also, this figure should contain an example of accessing a field outside its family.}
-  \caption{Demonstrating Pseudo-Code}
+  End [@\tiny{"A"$_2$}@][@\tanc{Coq-Example2-end}@]
+  \end{lstlisting}
+
+  \makeline[.5\textwidth+9pt]{Coq-Example2-start}{Coq-Example2-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+
+  \caption{Exemplar Compilation about Inductive Type, for STLC example}\label{fig:plugin-example2}
   \end{figure}
 
+\YZreply{Also, this figure should contain an example of accessing a field outside its family.}
 \begin{figure}

### 2022-10-23 metatheory-condensed.tex
Add beta rule for recursion
Add beta rule.
metatheory sec intro text

@@
+We contribute \TT, a core type theory that incorporates the core elements
+of family polymorphism into
+\citeauthor{martin1982constructive} dependent type theory~\cite{martin1982constructive},
+while maintaining consistency and canonicity.
+
+Most notably, \TT extends MLTT with what we call \emph{linkages}
+(a namesake of the construct through which \citet{zm2017} model family
+polymorphism in an OO setting).
+Linkages model families, so they are like tuples, but with a twist: they
+support late binding.
+
+\TT is intended as a foundational model. So unlike our user-facing plugin,
+typing is structural rather than nominal, and
+it supports family polymorphism by providing MLTT-style judgments that
+can be used to encode core aspects of the language supported by our
+plugin, rather than building them in directly.
+%
+For example, \TT supports the extension of inductive types through
+judgments that enable the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
+Also unlike our plugin, \TT does not feature automations such as
+the generation of propositional equalities and partial recursors,
+but it is expressive enough to encode their definitions.
+
+\newpage
+
 To formalize our metatheory, we follow the formulation of~\citet{altkap2016}, which provides a declarative and intrinsic style of the predicative \textit{Martin-Lof Type Theory} (MLTT)~\cite{martin1982constructive},
@@
   \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
-  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)}}
-{\goodTerm{\Gamma}{"Wsup"~T~a~b}{El~("W"~T)} }
+  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)[\pi_1]}}
+{\goodTerm{\Gamma}{"Wsup"~j~T~a~b}{El~("W"~T)} }
 $$
@@
   applicable only to the inductive type defined in the base family.
-  Orthogonally, can you add the beta rule?
 }\EDJreply{
@@
 \\
-  In other words, this recursor linkage "rec" can be used for any inductive type with signature $\tau$.\\
-  Yes I will add beta rule. I see the last comment.
+  In other words, this recursor linkage "rec" can be used for any inductive type with signature $\tau$.
 }
@@
 %
+$$
+$$
+\Rule[name=Rec-Beta]
+{\goodTerm{\Gamma}{h}{\cL("RecSig"~\tau~R)} 
+\\ \goodTerm{\Gamma}{T}{\bW\tau}}
+{\goodTerm{\Gamma}{"Wrec"~h~("Wsup"~j~T~a~b) \equiv 
+("app"~(("app"~("prjR"~j~h))[("id", a)]))[("id", "lam" ("Wrec"~h[\pi_1]~b))]
+}{R}}\EDJ{"app" and "lam" are lambda application and abstraction}
 $$

### 2022-10-23 metatheory-condensed.tex
Change into p and var Add one sentence explain substitution.

@@
 \Rule[name=Dbj Indices]
-{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
-{\goodSub{\Gamma}{\pi_1 \gamma}{\Delta}},
+{}
+{\goodSub{\Gamma, A}{p^1 \gamma}{\Gamma}},
 \Rule[]
-{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
-{\goodSub{\Gamma}{\pi_2 \gamma}{A[\pi_1 \gamma]}}
+{}
+{\goodSub{\Gamma, A}{"var"_0}{A[p^1]}}
 $$
@@
 The term judgement is the form $\goodTerm{\Gamma}{t}{T}$} in MLTT via \ruleref{Subst}.
-We also have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
+Note how substitution will change the context---we suggest intuitvely understanding context as holes of the term/types and substitution as filling those holes, where the substitution itself has holes.
+After filling the holes, then the new context is about the holes of the substitution.
+We have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
 % Debruijn Indices
 Contrary to named variables, \textbf{debruijn indices} refer to the variable in the context like operating a stack.
-For example, we can get the very first variable in the context $\Gamma, x : T \vdash x : T$ by $\goodTerm{\Gamma, T}{\pi_2 "id"}{T[\pi_1"id"]}$.
-Similarly, to get the second and above variables, we only need to apply a sequence of $\pi_1$s.
+We will use "var"$_i$ indicating the $i$-th variable (from right to left, $i$ starts with 0) in the context.
+On the other hand, we can also get the rest part of the stack, or ``subpart of "id"'' using $\goodSub{\Gamma, A}{p^1 \gamma}{\Gamma}$.
+For example, we can get the very first variable in the context $\Gamma, x : T \vdash x : T$ by $\goodTerm{\Gamma, T}{var_0}{T[p^1]}$.
+Note here $T$ is a type in $\Gamma$, but now shifted into $\goodType{\Gamma, T}{T[p^1]}{}$ using $p^1$.
+Similarly, to get the second and prior variables, we use "var"$_i$ and $p^{i+1}$ correspondingly.
 
@@
 \Biggl( \boxed{\goodSeal{\Gamma}{s}{\sigma}{A}} 
-:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}} \Biggr)
+:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[p^1]}} \Biggr)
 $$
@@
   \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
-  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)[\pi_1]}}
+  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)[p^1]}}
 {\goodTerm{\Gamma}{"Wsup"~j~T~a~b}{El~("W"~T)} }
@@
 }
-{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv \Pi A\; \Pi (\Pi B (R[\pi_1\circ\pi_1]))\; R[\pi_1\circ\pi_1]}{}}
+{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv \Pi A\; \Pi (\Pi B (R[p^2]))\; R[p^2]}{}}
 \quad 
@@
 {\goodTerm{\Gamma}{"Wrec"~h~("Wsup"~j~T~a~b) \equiv 
-("app"~(("app"~("prjR"~j~h))[("id", a)]))[("id", "lam" ("Wrec"~h[\pi_1]~b))]
+("app"~(("app"~("prjR"~j~h))[("id", a)]))[("id", "lam" ("Wrec"~h[p^1]~b))]
 }{R}}\EDJ{"app" and "lam" are lambda application and abstraction}
@@
           ~(\Sigma~\cU
-                  ~(\Sigma~(\Pi(\cC (\pi_2[\pi_1]))(El~\pi_2[\pi_1]))~\cU)) \\
+                  ~(\Sigma~(\Pi(\cC ("var"_0[p^1]))(El~"var"_0[p^1]))~\cU)) \\
   \Sigr{\Gamma}{0} &= \top \\
@@
   \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
-    \Sigma~\denotesT{\cC ("pjl"~\sigma)}~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
+    \Sigma~\denotesT{\cC ("pjl"~\sigma)}~(El \ ("app"~("pjr"^3~\sigma))[(p^1, "app"~("pjl"~("pjr"^2~t)))]) \\
       &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
@@
 \begin{align*}
-  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[\pi_1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[\pi_1]} \\
-  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}} &= \mu^+~(h[(\pi_1,\pi_1\mu~\pi_2)])\\ &\quad \quad \{f_2[\pi_1]\}~(p_2\mu'~\pi_2)[(\pi_1, \uparrow^s[\pi_1^{\uparrow}])] \\
+  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[p^1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[p^1]} \\
+  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(p^1, \uparrow^s)])}} &= \mu^+~(h[(p^1,p_1\mu~"var"_0)])\\ &\quad \quad \{f_2[p^1]\}~(p_2\mu'~"var"_0)[(p^1, \uparrow^s[{p^1}^{\uparrow}])] \\
 \end{align*}

### 2022-10-23 metatheory-condensed.tex
stuff

--- 2022-10-23 metatheory-condensed.tex
@@
 polymorphism in an OO setting).
-Linkages model families, so they are like tuples, but with a twist: they
+Linkages model families, so they are like tuples but with a twist: they
 support late binding.
 
-\TT is intended as a foundational model. So unlike our user-facing plugin,
+\TT is intended as a foundational model. So unlike our programmer-facing plugin,
 typing is structural rather than nominal, and
-it supports family polymorphism by providing MLTT-style judgments that
+it supports family polymorphism by providing MLTT-style typing rules that
 can be used to encode core aspects of the language supported by our
@@
 For example, \TT supports the extension of inductive types through
-judgments that enable the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
+formation rules that enable the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
 Also unlike our plugin, \TT does not feature automations such as
@@
 but it is expressive enough to encode their definitions.
+
+%\input{selected-typing.tex}
+
+\noindentparagraph{Syntax.}
 

### 2022-10-23 syntax.tex
Answer questions.
question
unanswered questions

@@
     \text{Inheritance Structure} \quad & h, h' &&= "inhid"~|~"inhov"~h~t~|~"inhext"~h~t~|~"inhinh"~h~|~"inh"~h~|~"inhnest"~h~h'
-  \end{alignat*}
+  \end{alignat*}\YZ{ What is 'W T'? Do you mean 'c T'?  }\YZreply{Why is Seal not merged into Terms?}
 \caption{Overview of Instrinsically-typed Syntax for extended MLTT}
@@
 \quad 
-\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}\YZ{What is star? Is it just tau?}
+\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}\YZ{What is star? Is it just tau??}
 \quad
@@
 }
-{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)[\pi_1]}}
+{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}
 $$
@@
 }
-{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited?}
+{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited??}
 $$

### 2022-10-23 syntax.tex
Answer questions.

@@
     \text{Types}   \quad & A, B, T        &&= ...~|~\bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|~\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma~|~"RecSig1"~A~B~T~|\\
-    \text{Terms}   \quad & t, t_1, t_2    &&= ...~|~"W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t~|\\ & &&"prjR"~j~o~|~"Wrec"~t_1~t_2~|~"inh"~h~t\\
+    \text{Terms, including Seal}   \quad & t, t_1, t_2, f    &&= ...~|~"W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t~|\\ & &&"prjR"~j~o~|~"Wrec"~t_1~t_2~|~"inh"~h~t\\
     \text{Inheritance Structure} \quad & h, h' &&= "inhid"~|~"inhov"~h~t~|~"inhext"~h~t~|~"inhinh"~h~|~"inh"~h~|~"inhnest"~h~h'
-  \end{alignat*}\YZ{ What is 'W T'? Do you mean 'c T'?  }\YZreply{Why is Seal not merged into Terms?}
+  \end{alignat*}\YZ{ What is 'W T'?
+Do you mean 'c T'?
+}\YZreply{Why is Seal not merged into Terms?}\EDJreply{You can see \ruleref{Ind-Type}.
+For $T : \bW\tau$, $W~T : \cU$, for $\goodType{\Gamma}{T}{}$, $c~T : \cU$ }\EDJreply{Seal can easily mergedd into terms.
+But seal has a new judgement.
+Do you want to merge them?
+}
 \caption{Overview of Instrinsically-typed Syntax for extended MLTT}
@@
 \quad 
-\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}\YZ{What is star? Is it just tau??}
+\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}\YZ{What is star?
+Is it just tau??}\EDJreply{star is a convention indicating this $\bW\tau$ is a singleton type, in other words, there is an eta(uniqueness) rule for arbitrary $a : \bW$,$a = \star$.
+But here I just want to show the constructor instead of the uniqueness rule.
+TBH I don't know what other good name (other than $\star$) can be used in  constructor/intro rule }
 \quad
@@
 }
-{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited??}
+{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited??}\EDJreply{$t$ is here the term in the field, not field name(we don't have name in metatheory) or type (type is T, it can be included).
+So no.
+Because the point of 'inhinh' in inheritance judgement is that when writing down inheritance, the children family doesn't have to include the original term.
+}
 $$

### 2022-10-23 metatheory.tex
Reply all the comments. Please check.
FIx language: distinguish judgement and the derivation of the judgement.

@@
     What's the relation to the CIC underlying Coq, given that the implementation is a Coq plugin?
-}\EDJreply{Add the comparison between this version of predicative MLTT and PCUIC. Please Check.}
+}\EDJreply{Add the comparison between this version of predicative MLTT and PCUIC. Please Check. }
 % Thus we will omit a great deal of details about MLTT here
@@
 Our formulation can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}.
-We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a pen-and-paper proof now because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real mechanized one so that the reader can have an idea about how mechanization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\cite{gratzer-sterling-birkedal-2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the judgemental rules in QIIT.
+We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a pen-and-paper proof now because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real mechanized one so that the reader can have an idea about how mechanization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\cite{gratzer-sterling-birkedal-2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the typing rules in QIIT.
 

### 2022-10-23 standardmodel.tex
FIx language: distinguish judgement and the derivation of the judgement.

@@
 The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
-\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.
+\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement also interpreted.
 

### 2022-10-23 syn-translate.tex
Fix notation.
Reply some comments.
Adding Syn Translation
FIx language: distinguish judgement and the derivation of the judgement.

@@
 
-Basically, we can compile away the judgement $\goodSig{\Gamma}{v}{n}$,
+Basically, we can compile away (the derivation of) judgement $\goodSig{\Gamma}{v}{n}$,
 the type $\cL$ and $\cC$, and related terms, by transforming them into
@@
 
-\newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
-\newcommand{\Sigr}[2]{{ "Sig"^r~{#1}~{#2} }}
+% \newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
+% \newcommand{\Sigr}[2]{{ "Sig"^r~{#1}~{#2} }}
 
-We use $\denotesT{}$ as the translation. This translation is invariant\YZ{What does "invariant" mean?}
+We use $\denotesT{}$ as the translation.
+This translation behaves like identity\YZ{What does "invariant" mean?}\EDJreply{behaves like identity.
+Sorry I should have said "subpart is invariant under this translation" instead}
 under the subpart. Since we are in an intrinsic setting, everything here

### 2022-10-23 syntax.tex
FIx language: distinguish judgement and the derivation of the judgement.

@@
   \end{align*}
-\caption{Judgement forms, as Agda types and as presented in this section.}\label{fig:vanillaMLTTjudgeQIIT}
+\caption{Judgements, as Agda types and as presented in this section.}\label{fig:vanillaMLTTjudgeQIIT}
 \end{figure}
@@
 We formalize our language as a deep embedding in a type theory using QIIT types.
-Every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms.
+Every judgement is a (dependent) QIIT \textit{type} and all derivations of judgement  are well-typed QIIT terms.
 (Otherwise how can they be encoded in Agda?) There are four standard judgement forms from (vanilla) MLTT,
@@
 Please check}
-Latex-wise, in \cref{fig:vanillaMLTTjudgewellformed}, we spell out these \textit{well-formedness rules} of type
-judgements and their (pseudo-)Agda counterparts to make things more accessible.
+Latex-wise, in \cref{fig:vanillaMLTTjudgewellformed}, we spell out these \textit{well-formedness rules} of each
+judgement and their (pseudo-)Agda counterparts to make things more accessible.
 In (pseudo-)Agda, well-formedness rules are just well-typedness of (indexed) QIIT family---(the judgement)"Tm~$\Gamma$~$T$" is well-typed (well-formed) only when $\Gamma$ and $T$ are both well-typed (well-formed).
@@
 course, a well-formed judgement is not necessarily derivable, just as
-a well-formed QIIT type is not necessarily inhabited. 
+a well-typed QIIT type is not necessarily inhabited. 
 These well-formedness rules are direct translations of the \textit{well-typedness} of QIIT.
@@
 This way, for all four kinds of judgements we are
-using, there is always equality among these data but only those equalities between terms and types are emphasized due to their non-triviality.
+using, there is always equality among the derivations but only those equalities between terms and types are emphasized due to their non-triviality.
 
@@
   \end{align*}
-\caption{Extended Judgement forms, as Agda types}
+\caption{Extended Judgements, as Agda types}
 \end{figure}
@@
 Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes.
-Right column show all the well-formedness rule for all the judgement.
+Right column show all the well-formedness rule for all the judgements.
 That is unusual because it is a direct translation from QIIT.
@@
 
-The core language builds on MLTT and adds four new judgement forms.
+The core language builds on MLTT and adds four new kinds of judgements.
 % Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
@@
 
-\caption{Seal Judgement, and its Agda Representation (or What do we mean when we say one judgement is a shorthand for the other)}
+\caption{Sealing Judgement, and its Agda Representation (or What do we mean when we say one judgement is a shorthand for the other)}
 
@@
 $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}$,
-$\goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the premises of judgements
+$\goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the premises of rule
 when needed.
@@
 
-The syntax of the inheritance judgement is very close to how Family is
+The deduction rules of the inheritance judgement is very close to how Family is
 defined in the surface syntax---a Family with no parent is an
@@
 The rule is mostly direct and we need ``upcasting'' again when dealing with nested family in different but inheriting context.
-We also show how to use an inheritance judgement to derive children family and how the beta rule is defined.
+We also show how to use an inheritance judgement and their derivation to construct children family and how the beta rule is defined.
 We omit most of the other substitution laws here.

### 2022-10-24 metatheory-condensed.tex
selected syntax

@@
 We contribute \TT, a core type theory that incorporates the core elements
-of family polymorphism into
-\citeauthor{martin1982constructive} dependent type theory~\cite{martin1982constructive},
+of \underline{f}amily polymorphism into
+\underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory~\cite{martin1982constructive},
 while maintaining consistency and canonicity.
@@
 Most notably, \TT extends MLTT with what we call \emph{linkages}
-(a namesake of the construct through which \citet{zm2017} model family
-polymorphism in an OO setting).
+(a namesake of the theoretical device through which \citet{zm2017} model
+family polymorphism in an OO setting).
 Linkages model families, so they are like tuples but with a twist: they
@@
 it supports family polymorphism by providing MLTT-style typing rules that
-can be used to encode core aspects of the language supported by our
+can be used to \emph{encode} core aspects of the language supported by our
 plugin, rather than building them in directly.
@@
 For example, \TT supports the extension of inductive types through
-formation rules that enable the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
+formation rules that enable the encoding of the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
 Also unlike our plugin, \TT does not feature automations such as
@@
 
-%\input{selected-typing.tex}
+\input{selected-typing.tex}
 

### 2022-10-24 syntax.tex
question
Reply the comment.
selected syntax

@@
 }
+  \YZ{The grammar suggests linkages are first-class terms.
+But aren't linkages second-class citizens?}\EDJreply{Linkages are first-class citizens in metatheory.
+Linkages are second class citizens in Coq plugin.
+Inheritance judgement is second class in plugin and metatheory.
+}
 \caption{Overview of Instrinsically-typed Syntax for extended MLTT}
@@
 \quad 
-\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}\YZ{What is star?
-Is it just tau??}\EDJreply{star is a convention indicating this $\bW\tau$ is a singleton type, in other words, there is an eta(uniqueness) rule for arbitrary $a : \bW$,$a = \star$.
-But here I just want to show the constructor instead of the uniqueness rule.
-TBH I don't know what other good name (other than $\star$) can be used in  constructor/intro rule }
+\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}
 \quad
@@
 }
-{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited??}\EDJreply{$t$ is here the term in the field, not field name(we don't have name in metatheory) or type (type is T, it can be included).
+{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}%\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited??}\EDJreply{$t$ is here the term in the field, not field name(we don't have name in metatheory) or type (type is T, it can be included).
 So no.
@@
 {\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}
+\quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}\YZ{This rule has a mysterious name.}
 \\\\

### 2022-10-24 abstract.tex
Challenge#2
Some Comments.

@@
 %
-This paper presents a language design useful for extensible metatheory
+This paper presents a language design geared towards extensible metatheory
 mechanization in a proof assistant.
@@
 \item a formalization of the language design as a constructive type theory with inductive types,
-\item proofs of strong guarantees afforded by the type theory,
+\item proofs of strong guarantees\EDJ{We don't yet have normalization though} afforded by the type theory,
 and 

### 2022-10-24 introduction.tex
Some Comments.

@@
 
-\item We present the first language design that enables extensible
+\item We present the first\EDJ{Are you sure? This claim is so bold.} language design that enables extensible
 metatheory mechanization in a higher-order, dependent type theory with
@@
 types (\cref{sec:metatheory}).
-We derive strong meta\-theoretical results including consistency and
+We derive strong meta\-theoretical\EDJ{I will suggest use ``foundational'' instead.
+Sterling at some places say canonicity is the most basic meta-theoretical property.} results including consistency and
 canonicity.

### 2022-10-24 lang-design.tex
Adding comments in first few sections.

@@
 After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
-from a base family to a derived family.
+from a base family to a derived family.\EDJ{I think the reviewer might ask : what if the subst late-bound tm_unit case/why the subst cases cannot be late-bound?
+it not necessary return tm_unit in those case.
+\\ What the plugin generates is actually subst tm_unit x t = subst.tm_unit.
+In other words, it points back to the handler instead of the concrete constructor.
+That is one of the advantage of looking FRecursion as a family.}
 %
@@
 In contrast, outside those families, \lsti{subst} is always referenced
-by specifying a family that contains it,
+by specifying a family that contains it, \EDJ{I think the reviewer without background of Family Polymorphism will be confused at this paragraph.
+i.e.
+Why inside family it is late-bound but not outside?
+}
 such as \lsti{STLC.subst} or \lsti{STLCFix.subst}.

### 2022-10-25 syntax.tex

@@
 \label{fig:rules:well-typed-ctx}
-\judgebox{\goodCtx{\Gamma}{i}}
+\judgebox{\goodCtx{\Gm}{i}}
 $$ 
@@
 \Rule[name=Context Extension]
-{\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{A}{j}}
-{\goodCtx{\Gamma, A}{i \cup j}}  
-$$
-
-\judgebox{ \goodType{\Gamma}{T}{i} }
+{\goodCtx{\Gm}{i} \quad \goodType{\Gm}{A}{j}}
+{\goodCtx{\Gm, A}{i \cup j}}  
+$$
+
+\judgebox{ \goodType{\Gm}{T}{i} }
 $$
@@
 {}
-{\goodType{\Gamma}{\cU}{j + 1}}
+{\goodType{\Gm}{\cU}{j + 1}}
 \quad 
@@
 {}
-{\goodType{\Gamma}{\cB}{0}}
+{\goodType{\Gm}{\cB}{0}}
 \quad 
@@
 {}
-{\goodType{\Gamma}{\bot}{0}}
+{\goodType{\Gm}{\bot}{0}}
 \quad 
 \Rule[name=Function]
-{\goodType{\Gamma}{A}{j} 
-  \quad \goodType{\Gamma, A}{B}{k}}
-{\goodType{\Gamma}{\Pi A B}{j \cup k}}
+{\goodType{\Gm}{A}{j} 
+  \quad \goodType{\Gm, A}{B}{k}}
+{\goodType{\Gm}{\TyPi{A}{B}}{j \cup k}}
 $$
@@
 \Rule[name=Func/DPair Subst]
-{\goodSub{\Gamma}{\gamma}{\Delta}
+{\goodSub{\Gm}{\gm}{\Delta}
 \quad \goodType{\Delta}{A}{} 
@@
 {
-  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
-  \quad 
-  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
+  \goodType{\Gm}{ \sub{(\TyPi {A} {B})}{\gm} \equiv \TyPi {\sub{A}{\gm}} {\sub{B}{\gm^\uparrow}} }{j \cup k}
+  \quad 
+  \goodType{\Gm}{ \sub{(\Sigma A B)}{\gm} \equiv \Sigma \sub{A}{\gm} \sub{B}{\gm^\uparrow} }{j \cup k}
 }
@@
 {\goodType{\Delta}{T}{j} 
-  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodType{\Gamma}{T[\gamma]}{j}}
+  \quad {\goodSub{\Gm}{\gm}{\Delta}}}
+{\goodType{\Gm}{\sub{T}{\gm}}{j}}
 \quad
 \Rule[name=Base Type Subst]
-{\goodSub{\Gamma}{\gamma}{\Delta}}
-{\goodType{\Gamma}{\cU[\gamma] \equiv \cU }{j + 1} \quad
-  \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
-  \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
-}
-$$
-\judgebox{ \goodTerm{\Gamma}{t}{T} }
+{\goodSub{\Gm}{\gm}{\Delta}}
+{\goodType{\Gm}{\sub{\cU}{\gm} \equiv \cU }{j + 1} \quad
+  \goodType{\Gm}{\sub{\cB}{\gm} \equiv \cB}{0} \quad 
+  \goodType{\Gm}{ \sub{\bot}{\gm} \equiv \bot}{0}
+}
+$$
+\judgebox{ \goodTerm{\Gm}{t}{T} }
 $$
 \Rule[name=Code/Get Type]
-{\goodType{\Gamma}{T}{j}}
-{\goodTerm{\Gamma}{"c" \ T}{\cU}
+{\goodType{\Gm}{T}{j}}
+{\goodTerm{\Gm}{\codety{T}}{\cU}
 }\quad
 \Rule[]
-{\goodTerm{\Gamma}{T}{\cU}}
-{\goodType{\Gamma}{El \ T}{j}}
+{\goodTerm{\Gm}{T}{\cU}}
+{\goodType{\Gm}{\El{T}}{j}}
 \quad 
@@
 {\goodTerm{\Delta}{t}{T}
-  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
+  \quad {\goodSub{\Gm}{\gm}{\Delta}}}
+{\goodTerm{\Gm}{ \sub{t}{\gm}}{ \sub{T}{\gm} }}
 $$
@@
 \Rule[]
-{\goodTerm{\Gamma, A}{t}{B}}
-{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
+{\goodTerm{\Gm, A}{t}{B}}
+{\goodTerm{\Gm}{\lam{t}}{\TyPi {A} {B}}}
 \quad 
 \Rule[]
-{\goodTerm{\Gamma}{t}{\Pi A B}}
-{\goodTerm{\Gamma, A}{"app"~t}{B}}
+{\goodTerm{\Gm}{t}{\TyPi {A} {B}}}
+{\goodTerm{\Gm, A}{\app{t}}{B}}
 \quad 
 \Rule[]
-{\goodTerm{\Gamma}{u}{A} 
-\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
-{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
+{\goodTerm{\Gm}{u}{A} 
+\quad \goodTerm{\Gm}{v}{B[(id, u)]}}
+{\goodTerm{\Gm}{(u,v)}{\Sigma A B}}
 \quad 
 \Rule[name=Sigma Elim]
-{\goodTerm{\Gamma}{t}{\Sigma A B}}
-{\goodTerm{\Gamma}{"pjl" \ t}{A}
-\\\\ \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
+{\goodTerm{\Gm}{t}{\Sigma A B}}
+{\goodTerm{\Gm}{\fst{t}}{A}
+\\\\ \goodTerm{\Gm}{\snd{t}}{B[(id, \fst{t})]}
 }
@@
 {}
-{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
-\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
-\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
+{\goodTerm{\Gm}{ \sub{(\lam{t})}{\gm} \equiv \lam{(\sub{t}{\gm^\uparrow})}}{\TyPi {A} {B}}
+\quad \goodTerm{\Gm}{ \sub{(u,v)}{\gm} \equiv (\sub{u}{\gm},\sub{v}{\gm})}{\Sigma A B}
+\quad \goodType{\Gm}{\El{(\sub{T}{\gm})} \equiv \sub{(\El{T})}{\gm}}{}
 }
@@
 {}
-{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
+{\goodTerm{\Gm}{\true, \false}{\cB}}
 \quad
 \Rule[]
-{\goodSub{\Gamma}{\gamma}{\Delta}}
-{\goodType{\Gamma}{(El \ T)[\gamma] \equiv El \ (T[\gamma]) }{j} \quad
- \goodTerm{\Gamma}{"tt"[\gamma] \equiv "tt"}{\cB} \quad 
- \goodTerm{\Gamma}{"ff"[\gamma] \equiv "ff"}{\cB} 
-}
-$$
-\judgebox{\goodSub{\Gamma}{\sigma}{\Delta}}
+{\goodSub{\Gm}{\gm}{\Delta}}
+{\goodType{\Gm}{ \sub{(\El{T})}{\gm} \equiv \El{(\sub{T}{\gm})} }{j} \quad
+ \goodTerm{\Gm}{ \sub{\true}{\gm} \equiv \true}{\cB} \quad 
+ \goodTerm{\Gm}{ \sub{\false}{\gm} \equiv \false}{\cB} 
+}
+$$
+\judgebox{\goodSub{\Gm}{\gm}{\Delta}}
 $$
 \Rule[name=Ept Subst]
-{}{\goodSub{\Gamma}{\epsilon}{\cdot}}
+{}{\goodSub{\Gm}{\epsilon}{\cdot}}
 \quad
 \Rule[]
-{}{\goodSub{\Gamma}{"id"}{\Gamma}}
+{}{\goodSub{\Gm}{"id"}{\Gm}}
 \quad
@@
   \goodSub{\Delta}{\delta}{\Theta}
-  \quad \goodSub{\Gamma}{\gamma}{\Delta} 
-}{\goodSub{\Gamma}{\delta \circ \gamma}{\Theta}}
+  \quad \goodSub{\Gm}{\gm}{\Delta} 
+}{\goodSub{\Gm}{ \SubstComp{\delta}{\gm}}{\Theta}}
 \Rule[name=Sub Extend]
-{\goodSub{\Gamma}{\gamma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\gamma]}}
-{\goodSub{\Gamma}{(\gamma, t)}{(\Delta, A)}}
+{\goodSub{\Gm}{\gm}{\Delta} \quad \goodTerm{\Gm}{t}{ \sub{A}{\gm}}}
+{\goodSub{\Gm}{(\gm, t)}{(\Delta, A)}}
 $$
@@
 \Rule[name=Proj Subst-1]
-{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
-{\goodSub{\Gamma}{\pi_1 \gamma}{\Delta}}
+{\goodSub{\Gm}{\gm}{(\Delta, A)}}
+{\goodSub{\Gm}{\pi_1 \gm}{\Delta}}
 \quad
 \Rule[name=Proj Subst-2]
-{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
-{\goodSub{\Gamma}{\pi_2 \gamma}{A[\pi_1 \gamma]}}
+{\goodSub{\Gm}{\gm}{(\Delta, A)}}
+{\goodSub{\Gm}{\pi_2 \gm}{ \sub{A}{\pi_1 \gm}}}
 \quad
@@
 {}
-{\goodSub{\Gamma}{(\pi_1 \gamma, \pi_2 \gamma) \equiv \gamma}{\Delta}}
+{\goodSub{\Gm}{(\pi_1 \gm, \pi_2 \gm) \equiv \gm}{\Delta}}
 $$
@@
 
-\judgebox{\goodWSig{\Gamma}{\tau}{n}}
+\judgebox{\goodWSig{\Gm}{\wsig}{n}}
 $$
@@
 {}
-{\goodWSig{\Gamma}{w\cdot}{0}}
+{\goodWSig{\Gm}{\WSigEmp}{0}}
 \quad
 \Rule[name=WSig-Add]
-{\goodWSig{\Gamma}{\tau}{n}
-  \quad \goodType{\Gamma}{A}{i}
-  \quad \goodType{\Gamma, A}{B}{i}}
-{\goodWSig{\Gamma}{w^+ \  \tau \  A \  B}{n+1}}
+{\goodWSig{\Gm}{\wsig}{n}
+  \quad \goodType{\Gm}{A}{i}
+  \quad \goodType{\Gm, A}{B}{i}}
+{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}}
 \quad
 \Rule[name=Ind-Univ]
-{\goodWSig{\Gamma}{\tau}{n}}
-{\goodType{\Gamma}{\bW \tau}{i}
-}
-$$
-%\YZ{Is there an introduction rule for terms of type 'bW sigma'?}
-%\EDJreply{Yes there is. I omit it because it is a singleton type, please see my newly added explanation below.}
+{\goodWSig{\Gm}{\wsig}{n}}
+{\goodType{\Gm}{\TyWSingle {\wsig}}{i}
+}
+$$
 
@@
 \Rule[name=WSig-Proj]
-{\goodWSig{\Gamma}{\tau}{n} \quad j < n}
-{\goodType{\Gamma}{\pi^j_1 \tau}{i} \quad \goodType{\Gamma, \pi^j_1 \tau}{\pi^j_2  \tau}{i}}
+{\goodWSig{\Gm}{\wsig}{n} \quad j < n}
+{\goodType{\Gm}{\wsigproj{j}{1}{\wsig} }{i} \quad \goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{i}}
 \quad
 \Rule[name=Ind-Sig]
-{\goodWSig{\Delta}{\tau}{n}
-  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodWSig{\Gamma}{\tau[\gamma]}{n}
-  \quad \goodType{\Gamma}{W (\tau[\gamma]) \equiv (W \tau)[\gamma]}{i}}
+{\goodWSig{\Delta}{\wsig}{n}
+  \quad {\goodSub{\Gm}{\gm}{\Delta}}}
+{\goodWSig{\Gm}{ \sub{\wsig}{\gm}}{n}
+  \quad \goodType{\Gm}{W ( \sub{\wsig}{\gm}) \equiv \sub{(W \wsig)}{\gm}}{i}}
 $$
@@
 \Rule[name=Ind-Type]
-{\goodTerm{\Gamma}{T}{\bW \tau}}
-{\goodTerm{\Gamma}{"W" \ T}{\cU}
-}
-\quad 
-\Rule[]{}{\goodTerm{\Gamma}{\star}{\bW \tau}}
+{\goodWSig{\Gm}{\wsig}{n}}
+{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}
+}
+\quad 
+\Rule[]{}{\goodTerm{\Gm}{\wsingle{\wsig}}{\TyWSingle{\wsig}}}
 \quad
 \Rule[name=Ind-Term]
-{\goodTerm{\Gamma}{T}{\bW \tau}
-  \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
-  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)}}
-{\goodTerm{\Gamma}{"Wsup"~T~a~b}{El~("W"~T)} }
+{{\goodWSig{\Gm}{\wsig}{n}}
+  \quad \goodTerm{\Gm}{a}{\wsigproj{j}{1}{\wsig}}
+  \quad \goodTerm{\Gm, \sub{\wsigproj{j}{2}{\wsig}}{(id, a)}}{b}{\El{(\wcode{\wsig})}}}
+{\goodTerm{\Gm}{\wsup{T}{a}{b}}{\El{(\wcode{\wsig})}} }
 $$
@@
 
-\judgebox{\goodSig{\Gamma}{\sigma}{n} }
+\judgebox{\goodSig{\Gm}{\lsig}{n} }
 $$
 \Rule[name=Lnkg Type/Compile]
-{\goodSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{\cL \sigma}{i}
-\quad \goodType{\Gamma}{\cC \sigma}{}}
+{\goodSig{\Gm}{\lsig}{n}}
+{\goodType{\Gm}{\TyLkg{\lsig}}{i}
+\quad \goodType{\Gm}{\TyTkg{\lsig}}{}}
 \quad
 \Rule[name=Sig/Lnkg Subst]
-{\goodSig{\Delta}{\sigma}{n}
-  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodSig{\Gamma}{\sigma[\gamma]}{n}
-  \quad \goodType{\Gamma}{\cL (\sigma[\gamma]) \equiv (\cL \sigma)[\gamma]}{i}}
+{\goodSig{\Delta}{\lsig}{n}
+  \quad {\goodSub{\Gm}{\gm}{\Delta}}}
+{\goodSig{\Gm}{ \sub {\lsig}{\gm}}{n}
+  \quad \goodType{\Gm}{\TyLkg{ \sub{\lsig}{\gm}} \equiv \sub{(\TyLkg{\lsig})}{\gm}}{i}}
 $$
@@
 {}
-{\goodSig{\Gamma}{\nu\cdot}{0}}
+{\goodSig{\Gm}{\LSigEmp}{0}}
 \quad
 \Rule[name=Sig Add]
-{\goodSig{\Gamma}{\sigma}{n} 
- \quad \goodType{\Gamma}{A}{}
- \quad \goodSeal{\Gamma}{f}{\sigma}{A}
- \quad \goodType{\Gamma, A}{T}{i}}
-{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
+{\goodSig{\Gm}{\lsig}{n} 
+ \quad \goodType{\Gm}{A}{}
+ \quad \goodSeal{\Gm}{\seal}{\lsig}{A}
+ \quad \goodType{\Gm, A}{T}{i}}
+{\goodSig{\Gm}{(\LSigAdd {\lsig}{\seal}{T})}{n+1}}
 $$
@@
 \Rule[name=Sig Proj]
-{\goodSig{\Gamma}{\sigma}{n+1}}
-{\goodSig{\Gamma}{p_1\nu \ \sigma}{n}
-\\ \goodType{\Gamma}{p_1\nu' \ \sigma}{}
-\\ \goodSeal{\Gamma}{p_f\nu \  \sigma}{p_1\nu \  \sigma}{p_1\nu' \sigma}
-\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}
+{\goodSig{\Gm}{\lsig}{n+1}}
+{\goodSig{\Gm}{\lsigproj{1} {\lsig}}{n}
+\\ \goodType{\Gm}{\lsigprojT {\lsig}}{}
+\\ \goodSeal{\Gm}{\lsigproj{s}{\lsig}}{\lsigproj{1} {\lsig}}{\lsigprojT {\lsig}}
+\\ \goodType{\Gm, \lsigprojT {\lsig}}{\lsigproj{2}{\lsig}}{}
 }
@@
 \Rule[name=Sig Proj Beta]
-{\goodSig{\Gamma}{\sigma}{n} 
-\quad \goodType{\Gamma}{A}{}
-\quad \goodSeal{\Gamma}{f}{\sigma}{A}
-\quad \goodType{\Gamma, A}{T}{i}}
-{\goodSig{\Gamma}{p_1\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv \sigma}{n}
-\\ \goodType{\Gamma}{p_1\nu' \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv A}{}
-\\ \goodSeal{\Gamma}{p_f\nu \  (\nu^+ \ \sigma \ \{f\} \ T) \equiv f}{\sigma}{A}
-\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ (\nu^+ \ \sigma \ \{f\} \ T) \equiv T}{}
+{\goodSig{\Gm}{\lsig}{n} 
+\quad \goodType{\Gm}{A}{}
+\quad \goodSeal{\Gm}{\seal}{\lsig}{A}
+\quad \goodType{\Gm, A}{T}{i}}
+{\goodSig{\Gm}{\lsigproj{1} {(\LSigAdd  {\lsig}  {\seal}  {T})} \equiv \lsig}{n}
+\\ \goodType{\Gm}{\lsigprojT {(\LSigAdd  {\lsig} {\seal} {T})} \equiv A}{}
+\\ \goodSeal{\Gm}{\lsigproj{s}{(\LSigAdd{\lsig} {\seal} {T})} \equiv \seal}{\lsig}{A}
+\\ \goodType{\Gm, \lsigprojT {\lsig}}{\lsigproj{2}{(\LSigAdd {\lsig} {\seal}  {T})} \equiv T}{}
 }
@@
 {}
-{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
+{\goodTerm{\Gm}{\LkgEmp}{\TyLkg {\LSigEmp}}}
 \quad
 \Rule[name=Lnkg Add]
-{ \goodTerm{\Gamma}{o}{\cL \sigma} 
-\quad \goodType{\Gamma}{A}{}
-\quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
- \quad \goodTerm{\Gamma, A}{t}{T}
-}
-{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
+{ \goodTerm{\Gm}{o}{\cL \lsig} 
+\quad \goodType{\Gm}{A}{}
+\quad  \goodSeal{\Gm}{\seal}{\lsig}{A} 
+ \quad \goodTerm{\Gm, A}{t}{T}
+}
+{\goodTerm{\Gm}{(\LkgAdd {o} {t})}{\TyLkg{\LSigAdd {\lsig} {\seal} {T}}}}
 $$
@@
 \Rule[name=Lnkg Proj]
-{\goodTerm{\Gamma}{o}{\cL\sigma}}
-{\goodTerm{\Gamma}{p_1\mu \ o}{\cL (p_1\nu \ \sigma)}
-\\ \goodTerm{\Gamma, p_1\nu' \ \sigma}{p_2\mu \ o}{p_2\nu \ \sigma}
+{\goodTerm{\Gm}{o}{\cL\lsig}}
+{\goodTerm{\Gm}{\lkgproj{1} {o}}{\cL (\lsigproj{1} {\lsig})}
+\\ \goodTerm{\Gm, \lsigprojT {\lsig}}{\lkgproj{2}{o}}{\lsigproj{2}{\lsig}}
 }
@@
 \Rule[name=Lnkg Compile]
-{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+{ \goodTerm{\Gm}{o}{\cL \lsig} 
 }
 {
-  \goodTerm{\Gamma}{\cCt o}{\cC \sigma}
+  \goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\lsig}}
 }
@@
 \Rule[name=Lnkg Proj Beta]
-{\goodTerm{\Gamma}{o}{\cL \sigma} 
-\quad \goodType{\Gamma}{A}{}
-\quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
- \quad \goodTerm{\Gamma, A}{t}{T}}
-{\goodTerm{\Gamma}{p_1\mu \ (\mu^+ \ o \ \{f\} \ t) \equiv o}{\cL \sigma}
-\\ \goodTerm{\Gamma, A}{p_2\mu \  (\mu^+ \ o \ \{f\} \ t) \equiv t}{T}
+{\goodTerm{\Gm}{o}{\cL \lsig} 
+\quad \goodType{\Gm}{A}{}
+\quad  \goodSeal{\Gm}{f}{\lsig}{A} 
+ \quad \goodTerm{\Gm, A}{t}{T}}
+{\goodTerm{\Gm}{\lkgproj{1} {(\LkgAdd {o} {t})} \equiv o}{\cL \lsig}
+\\ \goodTerm{\Gm, A}{\lkgproj{2}{(\LkgAdd {o}{t})} \equiv t}{T}
 }
@@
 {}
-{\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
+{\goodType{\Gm}{\TyTkg{\LSigEmp} \equiv \top}{} 
 \\
-\goodType{\Gamma}{\cC (\nu^+ \ \sigma \ \{f\} \ T) \equiv 
-    \Sigma (\cC \sigma) T["sf" \ f]}{}
-\\\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cC \nu\cdot}
-\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}
-}
-$$
+\goodType{\Gm}{\TyTkg {\LSigAdd  {\lsig} {f} {T}} \equiv 
+    \Sigma (\TyTkg {\lsig}) T["sf" \ f]}{}
+\\\\ \goodTerm{\Gm}{ \Tkg {\LkgEmp} \equiv ()}{\TyTkg{\LSigEmp}}
+\\ \goodTerm{\Gm}{\Tkg {(\LkgAdd {o} {t})} \equiv ((\Tkg {o}), t["sf" \ f][(id, \Tkg {o})]) }{}
+}
+$$
+
 % $$
@@
 
-  \judgebox{[\goodSeal{\Gamma}{s}{\sigma}{A}] 
-  = [\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}] }
+  \judgebox{[\goodSeal{\Gm}{s}{\lsig}{A}] 
+  = [\goodTerm{\Gm, \TyTkg{\lsig}}{s}{A[\pi_1]}] }
   $$
   \Rule[name=Seal Subst]
-  {\goodSeal{\Delta}{s}{\sigma}{A}
-    \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-  {\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{A[\gamma]}}
+  {\goodSeal{\Delta}{s}{\lsig}{A}
+    \quad {\goodSub{\Gm}{\gm}{\Delta}}}
+  {\goodSeal{\Gm}{ \sub{s}{\gm}}{ \sub{\lsig}{\gm}}{ \sub{A}{\gm}}}
   =
   \frac
-  {\goodTerm{\Delta, \cC\sigma}{s}{A[\pi_1]}
-    \quad  \goodSub{\Gamma}{\gamma}{\Delta}  }
-  {\goodTerm{\Gamma,\cC\sigma[\gamma]}{s[\gamma^\uparrow]}{A[\pi_1][\gamma^\uparrow]\equiv A[\gamma][\pi_1]}}
+  {\goodTerm{\Delta, \TyTkg{\lsig}}{s}{ \sub{A}{\pi_1}}
+    \quad  \goodSub{\Gm}{\gm}{\Delta}  }
+  {\goodTerm{\Gm,\TyTkg{\sub{\lsig}{\gm}}}{ \sub{s}{\gm^\uparrow}}{\sub{\sub{A}{\pi_1}}{\gm^\uparrow}\equiv \sub{\sub{A}{\gm}}{\pi_1}}}
   $$
@@
   {}
-  {\goodSeal{\Gamma}{id_s}{\sigma}{\cC \sigma}}
-  = \goodTerm{\Gamma, \cC\sigma}{\pi_2}{\cC\sigma[\pi_1]}
+  {\goodSeal{\Gm}{id_s}{\lsig}{\TyTkg {\lsig}}}
+  = \goodTerm{\Gm, \TyTkg{\lsig}}{\pi_2}{\TyTkg{\sub{\lsig}{\pi_1}}}
   $$
@@
 \Rule[name=Hdler]
-{\goodType{\Gamma}{A}{}
-\quad \goodType{\Gamma, A}{B}{}
-\quad \goodType{\Gamma}{R}{}
-}
-{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
+{\goodType{\Gm}{A}{}
+\quad \goodType{\Gm, A}{B}{}
+\quad \goodType{\Gm}{R}{}
+}
+{\goodType{\Gm}{\CaseSig{A}{B}{R} \equiv (\TyPi {A} {(\TyPi {(\TyPi {B} {(\sub{R}{ \SubstComp {\pi_1}{\pi_1}})})} {\sub{R}{\SubstComp{\pi_1}{\pi_1}}})} )}{}}
 \quad 
 \Rule[name=Hdlers]
-{\goodWSig{\Gamma}{\tau}{n}
-\quad \goodType{\Gamma}{R}{}
-}
-{\goodSig{\Gamma}{"RecSig"~\tau~R}{n}}
+{\goodWSig{\Gm}{\wsig}{n}
+\quad \goodType{\Gm}{R}{}
+}
+{\goodSig{\Gm}{\RecSig{\wsig}{R}}{n}}
 $$
@@
 \Rule[name=Hdler-Proj]
-{ \goodWSig{\Gamma}{\tau}{N}
-\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\tau~R)}
+{ \goodWSig{\Gm}{\wsig}{N}
+\quad \goodTerm{\Gm}{o}{\cL(\RecSig{\wsig}{R})}
 \quad j < N
 }
-{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}
+{\goodTerm{\Gm}{\Recproj{j}{o}}{\sub{(\CaseSig{(\pi_1^j \wsig)}{(\pi_2^j \wsig)}{R})}{\pi_1}}}
 $$
@@
 \Rule[name=Rec-Constr]
-{ \goodWSig{\Gamma}{\tau}{N}
-\quad \goodTerm{\Gamma}{T}{\bW \tau}
-\quad \goodType{\Gamma}{R}{}
-\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
-}
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
+{ \goodWSig{\Gm}{\wsig}{N}
+\quad \goodType{\Gm}{R}{}
+\quad \goodTerm{\Gm}{r}{\cL(\RecSig{\wsig}{R})}
+\quad \goodTerm{\Gm}{w}{\El{(\wcode{\wsig})}}
+}
+{\goodTerm{\Gm}{\Wrec{w}{r}}{R}}
 $$
@@
 
-\judgebox{\goodInh{\Gamma}{h}{\sigma}{\tau}}
+\judgebox{\goodInh{\Gm}{h}{\sigma_1}{\sigma_2}}
 
@@
 {}
-{\goodInh{\Gamma}{"inhid"}{\sigma}{\sigma}}
+{\goodInh{\Gm}{\InhId}{\lsig}{\lsig}}
 \quad
@@
 {
-\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}  
-\quad \goodType{\Gamma, A_1}{T_1}{}
-\quad \goodType{\Gamma, A_2}{T_2}{}
-  \quad \goodTerm{\Gamma, A_2}{t}{T_2}}
-{\goodInh{\Gamma}{"inhov" \ h \ t}{(\nu^+ \  \sigma_1 \  T_1)}{(\nu^+ \  \sigma_2\  T_2)}}
+\goodInh{\Gm}{h}{\sigma_1}{\sigma_2}  
+\quad \goodType{\Gm, A_1}{T_1}{}
+\quad \goodType{\Gm, A_2}{T_2}{}
+  \quad \goodTerm{\Gm, A_2}{t}{T_2}}
+{\goodInh{\Gm}{\InhOv{h}{t}}{(\LSigAdd  {\sigma_1 } {\seal_1} {T_1})}{(\LSigAdd   {\sigma_2} {\seal_2} {T_2})}}
 $$
@@
 \Rule[name=Inh-Ext]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-  \quad \goodTerm{\Gamma, A_2}{t}{T}}
-{\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}
+{\goodInh{\Gm}{h}{\sigma_1}{\sigma_2}
+  \quad \goodTerm{\Gm, A_2}{t}{T}}
+{\goodInh{\Gm}{\InhExt {h} {t}}{\sigma_1}{(\LSigAdd {\sigma_2} {\seal_2} { T})}}
 \quad
 \Rule[name=Inh-Inh]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodType{\Gamma, A_1}{T}{}
-\quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}
-}
-{\goodInh{\Gamma}{"inhinh" \ h}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}}%\YZ{Should the term be 'inhinh h t' instead of 'inhinh h', where t is the field being inherited??}\EDJreply{$t$ is here the term in the field, not field name(we don't have name in metatheory) or type (type is T, it can be included).
-So no.
-Because the point of 'inhinh' in inheritance judgement is that when writing down inheritance, the children family doesn't have to include the original term.
-}
+{\goodInh{\Gm}{h}{\sigma_1}{\sigma_2}
+\quad \goodType{\Gm, A_1}{T}{}
+\quad \goodTerm{\Gm, A_2}{\uparrow^s}{\sub{A_1}{\pi_1}}
+}
+{\goodInh{\Gm}{\InhInh{h}}{(\LSigAdd {\sigma_1} {\seal_1}  {T})}{(\LSigAdd   {\sigma_2} {\seal_2} {T[(\pi_1, \uparrow^s)]})}}
 $$
@@
 \Rule[name=Inh]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodTerm{\Gamma}{l}{\cL \sigma_1}
-}
-{\goodTerm{\Gamma}{("inh" \ h \ l)}{\cL \sigma_2}} 
+{\goodInh{\Gm}{h}{\sigma_1}{\sigma_2}
+\quad \goodTerm{\Gm}{l}{\cL \sigma_1}
+}
+{\goodTerm{\Gm}{(\inh {h} {l})}{\cL \sigma_2}} 
 \quad
 \Rule[name=Inh+Inh]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}\YZ{This rule has a mysterious name.}
+{\goodInh{\Gm}{h}{\sigma_1}{\sigma_2}
+\quad \goodTerm{\Gm, A_2}{\uparrow^s}{\sub{A_1}{\pi_1}}
 \\\\
-\goodInh{\Gamma, A_2}{i}{\tau_1[(\pi_1, \uparrow^s)]}{\tau_2}}
-{\goodInh{\Gamma}{"inhnest" \ h \ i}{(\nu^+ \  \sigma_1\  \cL\tau_1)}{(\nu^+ \  \sigma_2\  \cL\tau_2)}}
+\goodInh{\Gm, A_2}{i}{\tau_1[(\pi_1, \uparrow^s)]}{\tau_2}}
+{\goodInh{\Gm}{\InhNest{h}{i}}{(\LSigAdd {\sigma_1} {\seal_1}  {\cL\tau_1})}{(\LSigAdd   {\sigma_2} {\seal_2} {\cL\tau_2)}}}
 \quad
@@
 \Rule[name=Inh-Inh-beta]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-  \quad \goodTerm{\Gamma}{m}{\cL \sigma_1}
-  \quad \goodTerm{\Gamma, A_1}{t}{T}
-  \quad \goodTerm{\Gamma, A_2}{\uparrow^s}{A_1[\pi_1]}
-}
-{\goodTerm{\Gamma}{"inh" ("inhinh" \ h) (\mu^+ \ m \  \{f_1\} \ t) \newline  \equiv \mu^+ \ ("inh" \ h \ m) \  \{f_2\} \ t[(\pi_1, \uparrow^s)]}{...}} 
-$$
-
+{\goodInh{\Gm}{h}{\sigma_1}{\sigma_2}
+  \quad \goodTerm{\Gm}{m}{\cL \sigma_1}
+  \quad \goodTerm{\Gm, A_1}{t}{T}
+  \quad \goodTerm{\Gm, A_2}{\uparrow^s}{\sub{A_1}{\pi_1}}
+}
+{\goodTerm{\Gm}{\inh {(\InhInh{h})} {(\LkgAdd{m}{t})} \newline  \equiv \LkgAdd{(\inh {h} {m})}  {t[(\pi_1, \uparrow^s)]}}{...}} 
+$$
 Inheritance are judgements so naturally second class citizen data (i.e.,

### 2022-10-26 syntax.tex
Renaming Typing Rules, Part 1 Use this git change to rename all the legacy name from the new rule name

@@
 In dependent type setting, function now can return type, of which the type is ``type of type'', i.e., \textit{the universe} $\cU$.
-We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
+We follow the style of \textit{name=ty/universe à la Tarski}~\cite{hofmann1997syntax} and
 distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
@@
 $$ 
-\Rule[name=Empty Context]{}{\goodCtx{\cdot}{0}} 
-\quad
-\Rule[name=Context Extension]
+\Rule[name=ctx/ept]{}{\goodCtx{\cdot}{0}} 
+\quad
+\Rule[name=ctx/ext]
 {\goodCtx{\Gm}{i} \quad \goodType{\Gm}{A}{j}}
@@
 $$
-\Rule[name=Type Universe]
+\Rule[name=name=ty/universe]
 {}
@@
 \quad
-\Rule[name=Sub Comp]{
+\Rule[name=subst/comp]{
   \goodSub{\Delta}{\delta}{\Theta}
@@
 }{\goodSub{\Gm}{ \SubstComp{\delta}{\gm}}{\Theta}}
-\Rule[name=Sub Extend]
+\Rule[name=subst/ext]
 {\goodSub{\Gm}{\gm}{\Delta} \quad \goodTerm{\Gm}{t}{ \sub{A}{\gm}}}
@@
 $\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
-Thus (1) \ruleref{Sub Extend} looks like concatenating one
+Thus (1) \ruleref{subst/ext} looks like concatenating one
 series of terms with a new term (the weird-looking type of
 $\goodTerm{\Gamma}{t}{A[\gamma]}$ in the premise is due to our need of type $A$ in
-the context $\Delta$ so that we can use \ruleref{Context Extension} to
+the context $\Delta$ so that we can use \ruleref{ctx/ext} to
 create new context); (2) we have \ruleref{Proj Subst-1} to extract a
@@
 from the substitution itself.
-Consecutive substitutions can be composed using \ruleref{Sub Comp}, because
+Consecutive substitutions can be composed using \ruleref{subst/comp}, because
 all the holes are always from the last substitutions.

### 2022-10-26 metatheory-condensed.tex
Fix Syntactic Translation.

@@
 {
-  \goodTerm{\Gamma}{\cCt o}{\cC \sigma}
+  \goodTerm{\Gamma}{\Tkg {o}}{\cC \sigma}
 }
@@
   \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
-  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursive, } \\
+  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ are mutually recursive, } \\
   & \text{defined by induction on the signature length} \\  
-  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\
+  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
+  &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
   {\Sigr{\Gamma}{(n+1)}} &= 
@@
           ~(\Sigma~\cU
-                  ~(\Sigma~(\Pi(\cC ("var"_0[p^1]))(El~"var"_0[p^1]))~\cU)) \\
+                  ~(\Sigma~(\Pi(\TyTkg {("var"_0[p^1])})(El~"var"_0[p^1]))~\cU)) \\
   \Sigr{\Gamma}{0} &= \top \\
-  \denotesT{\goodType{\Gamma}{\cC \sigma}{}}, \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &\text{ defines upon } \denotesT{\sigma} 
+  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}}, \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ defines upon } \denotesT{\sigma} 
   \text{ and inductively on the signature length} \\
-  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesT{\goodType{\Gamma}{\cL\sigma}{}} = \top \quad
+  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} = \top \quad
       \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
-  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
-    \Sigma~\denotesT{\cC ("pjl"~\sigma)}~(El \ ("app"~("pjr"^3~\sigma))[(p^1, "app"~("pjl"~("pjr"^2~t)))]) \\
+  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= 
+    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(El \ (\app{("pjr"^3~\denotesT{\sigma})})[(p^1, \app{(\fst{("pjr"^2~{\denotesT{\sigma}})})})]) \\
+  \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &=
+  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(El~(\fst{(\snd{\denotesT{\sigma}})}))(El~(\app{("pjr"^3~\denotesT{\sigma})} )) \\
   &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
-  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &=
-  \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
-  \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
-  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\cL(\nu^+~\sigma~T)}} &= (m, "lam"~t) \\ 
-  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} & \text{ is defined upon } \denotesT{m} \text{ and inductively on the signature length} \\ 
-  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} &= 
-  (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
-  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
+  \denotesT{\goodSig{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{(\codety{T})}) \\ 
+  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\TyLkg{(\LSigAdd{\sigma}{s}{T})}}} &= (\denotesT{m}, \lam{t}) \\ 
+  \denotesT{\goodTerm{\Gamma}{\Tkg{m}}{\TyTkg {\sigma}}} & \text{ is defined upon } \denotesT{m} \text{ and inductively on the signature length} \\ 
+  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= 
+  (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{(\snd{\denotesT{m}})}, f = p_f\nu~\sigma \\ 
+  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
 \end{align*}
 \begin{align*}
-  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[p^1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[p^1]} \\
-  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(p^1, \uparrow^s)])}} &= \mu^+~(h[(p^1,p_1\mu~"var"_0)])\\ &\quad \quad \{f_2[p^1]\}~(p_2\mu'~"var"_0)[(p^1, \uparrow^s[{p^1}^{\uparrow}])] \\
+  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \TyLkg{\sigma_1}}{\_}{\TyLkg{\sigma_2[p^1]}} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \TyLkg{\sigma_1}}{h}{\TyLkg{\sigma_2[p^1]}} \\
+  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\LSigAdd {\sigma_1} {s_1}{T})}{(\LSigAdd {\sigma_2} {s_2} {T[(p^1, \uparrow^s)]})}} &= \mu^+~(h[(p^1,\lkgproj{1}{"var"_0})])\\ &\quad \quad \{f_2[p^1]\}~(\lkgproj{2}{"var"_0})[(p^1, \uparrow^s[{p^1}^{\uparrow}])] \\
 \end{align*}

### 2022-10-26 syn-translate.tex
Fix Syn Translation.

@@
 \begin{align*}
-  \text{We will define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
-  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursive, } \\
+  \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
+  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ are mutually recursive, } \\
   & \text{defined by induction on the signature length} \\  
-  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
+  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
+  &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
   {\Sigr{\Gamma}{(n+1)}} &= 
@@
           ~(\Sigma~\cU
-                  ~(\Sigma~(\Pi(\cC (\pi_2[\pi_1]))(El~\pi_2[\pi_1]))~\cU)) \\
+                  ~(\Sigma~(\Pi(\TyTkg {("var"_0[p^1])})(El~"var"_0[p^1]))~\cU)) \\
   \Sigr{\Gamma}{0} &= \top \\
-  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &\text{ defines upon } \denotesT{\sigma} \\
-  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
-    \Sigma~\denotesT{\cC ("pjl"~\sigma)}~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
+  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}}, \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ defines upon } \denotesT{\sigma} 
+  \text{ and inductively on the signature length} \\
+  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} = \top \quad
+      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
+  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= 
+    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(El \ (\app{("pjr"^3~\denotesT{\sigma})})[(p^1, \app{(\fst{("pjr"^2~{\denotesT{\sigma}})})})]) \\
+  \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &=
+  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(El~(\fst{(\snd{\denotesT{\sigma}})}))(El~(\app{("pjr"^3~\denotesT{\sigma})} )) \\
   &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
-  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \top \quad
-      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
-  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} & \text{ is defined upon } \denotesT{\sigma} \\
-  & \text{ and inductively on the signature length} \\ 
-  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &=
-  \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
-  &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
-  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &= \top \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\
-  \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
-  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\cL(\nu^+~\sigma~T)}} &= (m, "lam"~t) \\ 
-  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} & \text{ is defined upon } \denotesT{m}\\
-  & \text{ and inductively on the signature length} \\ 
-  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} &= 
-  (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
-  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
+  \denotesT{\goodSig{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{(\codety{T})}) \\ 
+  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\TyLkg{(\LSigAdd{\sigma}{s}{T})}}} &= (\denotesT{m}, \lam{t}) \\ 
+  \denotesT{\goodTerm{\Gamma}{\Tkg{m}}{\TyTkg {\sigma}}} & \text{ is defined upon } \denotesT{m} \text{ and inductively on the signature length} \\ 
+  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= 
+  (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{(\snd{\denotesT{m}})}, f = p_f\nu~\sigma \\ 
+  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
 \end{align*}
 \begin{align*}
-  &\denotesT{\goodSig{\Gamma}{p_1\nu~\sigma}{n}} = "pjl"~\sigma 
-  &&\denotesT{\goodType{\Gamma}{p_1\nu'~\sigma}{}} = El~("pjl"~("pjr"~\sigma)) \\
-  &\denotesT{\goodSeal{\Gamma}{p_f\nu~\sigma}{p_1\nu~\sigma}{p_1\nu'~\sigma}} = "app"~("pjl"~("pjr"^2~\sigma)) 
-  &&\denotesT{\goodSig{\Gamma}{p_2\nu~\sigma}{n}} = El~("app"~("pjr"^3~\sigma)) \\ 
-  &\denotesT{\goodTerm{\Gamma}{p_1\mu~o}{\cL(p_1\nu\sigma)}} = "pjl"~o 
-  &&\denotesT{\goodTerm{\Gamma, p_1\nu'~\sigma}{p_2\mu'~o}{p_2\nu~\sigma}} = "app"~("pjr"~o)
-\end{align*}
-\begin{align*}
-  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[\pi_1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[\pi_1]} \\
-  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}} &= \mu^+~(h[(\pi_1,\pi_1\mu~\pi_2)])\\ &\quad \quad \{f_2[\pi_1]\}~(p_2\mu'~\pi_2)[(\pi_1, \uparrow^s[\pi_1^{\uparrow}])] \\
-  \denotesT{\goodInh{\Gamma}{"inhov"~h~t}{(\nu^+ \  \sigma_1 \  T_1)}{(\nu^+ \  \sigma_2\  T_2)}} &= \mu^+ (h[(\pi_1, p_1\mu~\pi_2)])~\{f_2[\pi_1]\}~t[\pi_1^{\uparrow}]
+  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \TyLkg{\sigma_1}}{\_}{\TyLkg{\sigma_2[p^1]}} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \TyLkg{\sigma_1}}{h}{\TyLkg{\sigma_2[p^1]}} \\
+  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\LSigAdd {\sigma_1} {s_1}{T})}{(\LSigAdd {\sigma_2} {s_2} {T[(p^1, \uparrow^s)]})}} &= \mu^+~(h[(p^1,\lkgproj{1}{"var"_0})])\\ &\quad \quad \{f_2[p^1]\}~(\lkgproj{2}{"var"_0})[(p^1, \uparrow^s[{p^1}^{\uparrow}])] \\
 \end{align*}

### 2022-10-26 coq-example.tex
Restructure the introduction of Parsing Example.

@@
 \subsection{An Extensible Decision Procedure for Parsing}
-The third example is a decision procedure for parsing.
-The basic idea is that we can use Cop's Inductive Facility to encode CFG rules.
-Then extensible inductive family corresponds to extensible CFG rules.
-For example, indexed type \mintinline{Coq}{is_prog : strs → Type} indicate a token list is a syntacitcally well-formed program.
+The third example is a decision procedure for parsing, based on the fact that, we can use Cop's Inductive Facility to encode CFG rules.
+Thus extensible inductive family corresponds to extensible CFG rules.
+For example, indexed type \mintinline{Coq}{is_prog : strs → Type} is a predicate asserting if a token list is a syntactically well-formed program.
 
@@
 
-To prove \mintinline{Coq}{forall s, (is_prog s) + (is_prog s -> False)}, we use strong induction on the length and the non-trivial part is  the inductive step, i.e. we need to show
+We prove decidability of "is_prog" via strong induction on the length. The non-trivial part is  the inductive step, i.e. we need to show "is_prog" is ``conditionally decidable'':
 \begin{minted}[fontsize=\footnotesize]{Coq}
@@
 
-Then we provide an overridable predicate \mintinline{Coq}{alg_is_prog : strs -> Type},\YZ{Why does it need to be overridden?}\EDJreply{Basically alg_is_prog is a big disjunction list, corresponding to each of the constructor of }
-decidable provided the inductive hypothesis and sound \& complete w.r.t.
+This proposition should hold due to the shape of the inductive definition---because we can translate the definition of "is_prog" into an equivalence to a big disjunction: 
+\begin{minted}[fontsize=\footnotesize]{Coq}
+(∃ a ∃ b , is_prog a /\ is_prog b /\ s = a ++ [";"] ++ b)
+\/ (∃ i ∃ e, is_atom i /\ is_exp e /\ s = i ++ [":="] ++ e) 
+... 
+<-> is_prog s
+\end{minted}
+Then decidability is immediate via this equivalence, since on LHS every sub token list has length strictly smaller than "s".
+
+However, when maintaining this equivalence proof through out different families---calling this big disjunction on LHS as \mintinline{Coq}{alg_is_prog : strs -> Type}---"alg_is_prog" is changing when "is_prog" extends.
+Thus "alg_is_prog" should be encoded as an overridable field.
+
+We will have challenges maintaining this equivalence proof.
+From left to right (soundness), we can have incremental construction and proof reuse throughout different families---we can easily prove "is_prog l" using each constructor.
+However, from right to left (completeness), the proof from the parent family cannot be reused in the children as "alg_is_prog" is changing and the proposition itself is changing.
+
+% Then we provide an overridable predicate \mintinline{Coq}{alg_is_prog : strs -> Type},\YZ{Why does it need to be overridden?}\EDJreply{Basically alg_is_prog is a big disjunction list, corresponding to each of the constructor of }
+% decidable provided the inductive hypothesis and sound \& complete w.r.t.
 to the specification \mintinline{Coq}{alg_is_prog e  <-> is_prog e}.
@@
 
-Among them, completeness \mintinline{Coq}{is_prog e -> alg_is_prog e} is
-the most challenging part. This is proved by recursion on the predicate
-"is_prog e", but none of the handler on "is_prog" will be reused
-because \mintinline{Coq}{alg_is_prog e} is overriden during
-inheritance.
-Unlike completeness, other parts of the construction, including
-soundness and and conditional decidability can be reused easily.
+% Among them, completeness \mintinline{Coq}{is_prog e -> alg_is_prog e} is
+% the most challenging part. This is proved by recursion on the predicate
+% "is_prog e", but none of the handler on "is_prog" will be reused
+% because \mintinline{Coq}{alg_is_prog e} is overriden during
+% inheritance.
+
+% Unlike completeness, other parts of the construction, including
+% soundness and and conditional decidability can be reused easily.
 \YZ{But is the reusability achieved through a derived family's reusing handlers in the base family?

### 2022-10-27 introduction.tex
address/reply to comments in intro

--- 2022-10-03 introduction.tex
@@
 
-\item We present the first language design that enables extensible
+\item We present, to our knowledge, the first\EDJ{Are you sure? This claim is so bold.}\YZreply{Is there any previous language designs in this space?} language design that enables extensible
 metatheory mechanization in a higher-order, dependent type theory with
@@
 The language design reconciles the expressiveness enabled by
-family polymorphism with the exactness of a proof assistant,
+family polymorphism with the rigor of a proof assistant,
 while retaining an idiomatic programming style.
@@
 types (\cref{sec:metatheory}).
-We derive strong meta\-theoretical results including consistency and
-canonicity.
+We derive foundational meta\-theoretical results including consistency
+and canonicity.
+%\EDJ{I will suggest use ``foundational'' instead. Sterling at some places say canonicity is the most basic meta-theoretical property.}
 We also formalize the translation implemented by the Coq plugin.

### 2022-10-27 metatheory-condensed.tex
explain substitution

@@
 %
-For example, \TT supports the extension of inductive types through
+For example, \TT supports extending inductive types through
 formation rules that enable the encoding of the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
@@
 
-\noindentparagraph{Syntax.}
+\noindentparagraph{MLTT.}
+
+\cref{fig:typing-selected} presents the syntax of \TT.
+The new forms (relative to MLTT) are highlighted.
+
+As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
+substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
+of the syntax rather than meta-operations.
+%
+Variables are represented by de Bruijn indices:
+\var{n} is the variable bound by the $n$-th closest enclosing binder.
+For example, $\lambda x.\,\lambda y.\,x$ is $\lam{(\lam{\var{1}})}$.
+%
+Formation rules for substitutions have the form
+$\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
+valid in the context~$\Dl$ yields terms valid in~$\Gm$.
+%
+The two main forms of substitutions are weakening and extension:
+\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
+\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
+then applies~$\gm$.
+For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
+of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.
+
+%Note that we are actually using predicative universe, so we actually have different levels of universes $\cU_i$. 
+%We omit universe levels in our latex presentation for conciseness and readability.
 

### 2022-10-27 metatheory-condensed.tex
Changing sealing part.

@@
 Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Sealing is used as a convenient device \textit{to indicate whether each field is overridable(extensible) or non-overridable (thus definitionally exposed)}.
+Sealing is corresponding \textit{to indicating whether each field is overridable(extensible) or non-overridable (thus definitionally exposed) in the Coq Plugin}.
+Concretely, we encode our STLC example to show the motivation of the sealing.
 %\YZ{Is sealing used only for late-binding the signatures of inductive types?
@@
 
-Intuitively, sealing corresponds to the following surface syntax\\
-\mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...} where every former field "x","y","z",...
+% Intuitively, sealing corresponds to the following surface syntax\\
+% \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...} where every former field "x","y","z",...
 hide its definition when type-check "t", and remaining former fields have their definition exposed for "t".
@@
 
-The reader should feel this subtle difference compared to our current plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
+% The reader should feel this subtle difference compared to our current plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
 In other words, every field in meta-theory has the right to choose how they perceive and abstract the prior fields.

### 2022-10-27 metatheory-condensed.tex
Comment.

@@
 %
-Formation rules for substitutions have the form
+Formation rules for substitutions have\EDJ{It seems to me G |- g : D is a judgement not a rule.
+Rules usually have two parts (premises + conclusion).
+Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here).
+I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this.
+In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) } the form
 $\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms

### 2022-10-27 introduction.tex
Reply comments.

@@
 
-\item We present, to our knowledge, the first\EDJ{Are you sure? This claim is so bold.}\YZreply{Is there any previous language designs in this space?} language design that enables extensible
+\item We present, to our knowledge, the first\EDJ{Are you sure?
+This claim is so bold.}\YZreply{Is there any previous language designs in this space?}\EDJreply{Olivier Boite.
+2004.
+Proof reuse with extended inductive types.
+In International Conference on Theorem Proving in
+Higher Order Logics. Springer, 50–65.  I think I also mentioned several paper in the related work.} language design that enables extensible
 metatheory mechanization in a higher-order, dependent type theory with

### 2022-10-27 metatheory-condensed.tex
Save about sealing.

@@
 We will use $\{a : T_1, b : T_2, ..\}$ as a shorthand for signature with type $T_1, T_2, ..$ for readability.
-
+\newpage 
 Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
@@
 % This above elaborated "hiding" syntax directly corresponding to the $f$ in $(\mu^+ ~o~\{f\}~t)$, where each fields decides the abstraction of the prior.
+
+\begin{figure}[!htb]
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
+  }
+
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{3}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting}
+  Family STLC. 
+   FInductive ty (* no hiding *)
+
+   FInductive tm (* hiding ty *)
+
+   FRecursion subst
+     Case ... (* hiding ty tm *)
+     ...
+   End subst (* hiding ty *)
+
+   Field t1 := (tvar 0) 
+   (* hiding ty tm subst *)
+
+   Field t2 := (tapp t1 t1)
+   (* hiding ty tm subst *)
+
+  ...
+  \end{lstlisting}
+
+  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
+
+  \columnbreak
+  % \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+
+    (* s₁  : {} | {} *)
+    - o₁  = μ+ {} s₁ ⋆ : 𝕃{ty : 𝕎τ₁}
+    (* s₂  : {ty : 𝕎τ₁} | [ty : 𝕌; bool : ty; ...] *) 
+    - o₂  = μ+ o₁ s₂ ⋆ : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂}
+    (* s₃₁  : {ty : 𝕎τ₁; tm : 𝕎τ₂} 
+            | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
+    - o₃₁  = μ+ o₂ s₃ (μ+ ..) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig)}
+    (* s₃ : {ty : 𝕎(τ₁); tm : 𝕎τ₂; subst' : 𝕃(RecSig)} 
+          | [ty : 𝕌; ... ; tm : 𝕎τ₂; subst' : 𝕃(RecSig)] *)
+    - o₃ = μ+ o₃₁ s₃ (λ (Wrec ..)) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
+
+    (* s₄ : {ty : 𝕎τ₁; ... subst : tm → nat → tm → tm } 
+          | [ty : 𝕌; ... ; tm : 𝕌; subst : tm → nat → tm → tm] *)
+    - o₄ = μ+ o₃ s₄ .. : 𝕃{ty : 𝕎τ₁; ... t₁ : tm}
+
+    (* s₅ : {ty : 𝕎τ₁; tm : 𝕎τ₂; subst : 𝕃(RecSig); t1 : tm} 
+          | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : tm] *)
+    - o₅ = μ+ o₄ s₅ : 𝕃{ ... }
+  \end{lstlisting}
+
+  \columnbreak
+
+  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+  \caption{Sealing and Linkage}\label{fig:sealing+linkage}
+  \end{figure}
+
+In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.
+For "tm", "ty" is late-bound and thus extending "ty" won't cause disruption on the inheritance of "tm".
+For "subst" itself, the definition of "tm" is not hidden to achieve exhaustiveness checking.
+However, for each case handler inside "subst", "tm" is late-bound and thus each handler is inheritable.
+
+Each hiding comment corresponds to the textual-aligned sealing derivation on RHS.
+We use surface syntax for eligibility: curly brackets indiate a signature (omitting sealing at type level for readability) and square bracket indicate a nested sigma type, corresponding to sealing judgement $\goodSeal{}{s}{\sigma}{A}$.
+
+For "tm", "ty" is hidden and thus the sealing "s₂" will make "ty" into an arbitrary type "𝕌" and expose its constructor "bool", "arrow", and etc.
+Then when constructing "o₂", the newly added type "tm" as "⋆ : 𝕎τ₂" is constructed in the context where "ty : 𝕌", ignoring its original eliminator.
+(Because eliminator is only possible when "ty : 𝕎τ₁").
+On the other hand, for "subst", only "ty" is hidden and thus the sealing "s₃" leave "tm : 𝕎τ₂" untouched and "Wrec" becomes possible when constructing "o₃".
+
+Basically, the hidden prior fields---inductive type are sealed into arbitrary type with ``constructors only'', where the recursion is constructed without sealing the inductive type.
+Generally speaking, each extensible inductive type are hidden for all the other fields except for the recursion on itself.
+Furthermore, normal fields (e.g.
+"t1, t2") can expose its definition to all in meta-theory by using singleton type.
+(e.g.
+saying "t1 : S(tvar 0)" instead of the current "t1 : tm"); and then overridable fields are hidden for normal fields (by sealing "t1 : S(tvar 0)" into "t1 : tm").
+Generally speaking, anything with explicit "hiding" decoration can be understood as sealing.
 

### 2022-10-27 metatheory-condensed.tex
Condensing compilation type. Fixing typing rules in the maintext.

@@
 $$
-\Rule[name=Ept Sig]
-{}
-{\goodSig{\Gamma}{\nu\cdot}{0}}\quad
-\Rule[name=Lnkg Type/Compile]
-{\goodSig{\Gamma}{\sigma}{n}}
-{\goodType{\Gamma}{\cL \sigma}{i}
-\quad \goodType{\Gamma}{\cC \sigma}{}}
+\Rule[name=lsig/ept]
+{""}
+{\goodSig{\Gm}{\LSigEmp}{0}}
+\quad
+\Rule[name=ty/lkg]
+{\goodSig{\Gm}{\lsig}{n}}
+{\goodType{\Gm}{\TyLkg{\lsig}}{i}
+\\ \goodType{\Gm}{\TyTkg{\lsig}}{}}
 \quad
@@
 
-Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited by linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
+% Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited by linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
 The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
@@
 
-The difference between compilation type and linkage types lies in their introduction rules.
+Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) sigma type instead of a linkage type.
+For example, $\TyLkg{\{a : S(\bot), b : a \to \bot\}}$ can only be sealed into $\TyTkg{\{a : \cU, b : a \to \bot\}}$ but never $\TyLkg{\{a : \cU, b : a \to \bot\}}$.
+Because the later cannot be inhabitted---by projecting the second field of the $ t : \TyLkg{\{a : \cU, b : a \to \bot\}}$, we can get a proof of bottom type.
+
+% The difference between compilation type and linkage types lies in their introduction rules.
 The compilation types are just syntactic sugar for sigma types.
@@
 
-The importance of the compilation type lies in its difference from the linkage type.
+% The importance of the compilation type lies in its difference from the linkage type.
 First we have a singleton type $"s A" : S(T) \iff "A" \equiv T$, i.e.
@@
 
-But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited
-and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
+% But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited
+% and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
 This example illustrates the programming difference between the sigma type and linkage type.
 
-However, this form of abstracting a type is very important when programming in our metatheory---as later we will show inductive type of different constructors locates in different type (universe), and to make later fields inheritable we need to work on an interface of abstracted type.
+% However, this form of abstracting a type is very important when programming in our metatheory---as later we will show inductive type of different constructors locates in different type (universe), and to make later fields inheritable we need to work on an interface of abstracted type.
 Thus the compilation type is indispensable.

### 2022-10-27 metatheory-condensed.tex
Comments on how linkage already capture code reuse

@@
 Our metatheory is formulated using QIIT in an extensional setting.
-%\YZ{
-%    Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
-%    not definitionally equal but propositionally equal in family STLC?
-%}\EDJreply{No, TLDR: it is a different level. \\ 
-%There are two type theories, one is Agda-like (= Latex = Conventional math), the other is the one formulated in QIIT.
-This sentence is saying that the Agda-like one is extensional MLTT (i.e.
-we have function extensionality and etc here in Latex writing directly).
-\\ But the one you are asking about is in the metatheory in QIIT: if we have an identity type $Id x y$ (propositionally equal) can we have a judgemetnal equality for the two terms in QIIT $x \equiv y$.
-\\
-%Apparently, we don't setup our QIIT formulation in this way (I thought we need, but now I don't think we need it. Our QIIT formulation is purely in intensional MLTT now).\\ 
-%But even if we do setup like that, I will not say a conflict is happening (because it is to our advantage if we have judgemental equal in metatheory).
-I will say it is a bit off as to how Coq is formulated (Coq is closer to an intensional MLTT, so no judgemental equal only propositional one).
-%}
-
-% linkage, signature and compilation and seal
-$$
-\Rule[name=lsig/ept]
-{""}
-{\goodSig{\Gm}{\LSigEmp}{0}}
-\quad
-\Rule[name=ty/lkg]
-{\goodSig{\Gm}{\lsig}{n}}
-{\goodType{\Gm}{\TyLkg{\lsig}}{i}
-\\ \goodType{\Gm}{\TyTkg{\lsig}}{}}
-\quad
-\Rule[name=Lnkg Compile]
-{ \goodTerm{\Gamma}{o}{\cL \sigma} 
-}
-{
-  \goodTerm{\Gamma}{\Tkg {o}}{\cC \sigma}
-}
-$$
-$$
-\Rule[name=Sig Add]
-{\goodSig{\Gamma}{\sigma}{n} 
- \quad \goodType{\Gamma}{A}{}
- \quad \goodSeal{\Gamma}{f}{\sigma}{A}
- \quad \goodType{\Gamma, A}{T}{i}}
-{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
-\quad
-\Biggl( \boxed{\goodSeal{\Gamma}{s}{\sigma}{A}} 
-:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[p^1]}} \Biggr)
-$$
-$$
-\Rule[name=Ept Lnkg]
-{}
-{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
-\quad
-\Rule[name=Lnkg Add]
-{ \goodTerm{\Gamma}{o}{\cL \sigma} 
-\quad \goodType{\Gamma}{A}{}
-\quad  {\goodSeal{\Gamma}{f}{\sigma}{A}}
- \quad \boxed{\goodTerm{\Gamma, A}{t}{T}}
-}
-{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
-$$
-We start with signature (in a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
-A signature is a list of types contained by a \textit{linkage}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
-We will use $\{a : T_1, b : T_2, ..\}$ as a shorthand for signature with type $T_1, T_2, ..$ for readability.
-\newpage 
-Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Sealing is corresponding \textit{to indicating whether each field is overridable(extensible) or non-overridable (thus definitionally exposed) in the Coq Plugin}.
-Concretely, we encode our STLC example to show the motivation of the sealing.
-%\YZ{Is sealing used only for late-binding the signatures of inductive types?
-Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}\EDJreply{All late-bound are achieved by ``selective-abstraction'' imposed by sealing.
-For example, recursion will be consequently abstracted because of the abstraction of the inductive type it recurse on.
-%}\EDJreply{Sorry, I see what you mean.
-Yes.
-the box part is important for late-binding.
-But the functionality of sealing is always there.
-They are both required to achieve overriding.
-A illustrative example is, if the type is $S(\bot)$, then even with the boxed premise, it is still not overridable because the type itself restricts all the possibilities.}\YZreply{
-%  By "selective abstraction", do you mean in the metatheory the programmer chooses what and how to abstract?
-%}\EDJreply{Yes.}
-
-% For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
-Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.
-
-% We can now point out how "Overridable" and "pins" is modeled.
-Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
-hide its definition when type-check t, and remaining former fields have their definition exposed in t.
-% This above elaborated "hiding" syntax directly corresponding to the $f$ in $(\mu^+ ~o~\{f\}~t)$, where each fields decides the abstraction of the prior.
-
-\begin{figure}[!htb]
-  \lstset{
-      basicstyle=\fontsize{8}{8.5}\ttfamily,
-  % numbers=left,
-  }
-
-  \begin{minipage}{\textwidth}
-  \begin{multicols}{3}
-
-  \definecolor{codecomment-color}{HTML}{0DA3FF}
-
-  \begin{lstlisting}
-  Family STLC. 
-   FInductive ty (* no hiding *)
-
-   FInductive tm (* hiding ty *)
-
-   FRecursion subst
-     Case ... (* hiding ty tm *)
-     ...
-   End subst (* hiding ty *)
-
-   Field t1 := (tvar 0) 
-   (* hiding ty tm subst *)
-
-   Field t2 := (tapp t1 t1)
-   (* hiding ty tm subst *)
-
-  ...
-  \end{lstlisting}
-
-  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
-
-  \columnbreak
-  % \definecolor{codecomment-color}{HTML}{5D030F}
-
-  \begin{lstlisting}
-
-    (* s₁  : {} | {} *)
-    - o₁  = μ+ {} s₁ ⋆ : 𝕃{ty : 𝕎τ₁}
-    (* s₂  : {ty : 𝕎τ₁} | [ty : 𝕌; bool : ty; ...] *) 
-    - o₂  = μ+ o₁ s₂ ⋆ : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂}
-    (* s₃₁  : {ty : 𝕎τ₁; tm : 𝕎τ₂} 
-            | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
-    - o₃₁  = μ+ o₂ s₃ (μ+ ..) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig)}
-    (* s₃ : {ty : 𝕎(τ₁); tm : 𝕎τ₂; subst' : 𝕃(RecSig)} 
-          | [ty : 𝕌; ... ; tm : 𝕎τ₂; subst' : 𝕃(RecSig)] *)
-    - o₃ = μ+ o₃₁ s₃ (λ (Wrec ..)) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
-
-    (* s₄ : {ty : 𝕎τ₁; ... subst : tm → nat → tm → tm } 
-          | [ty : 𝕌; ... ; tm : 𝕌; subst : tm → nat → tm → tm] *)
-    - o₄ = μ+ o₃ s₄ .. : 𝕃{ty : 𝕎τ₁; ... t₁ : tm}
-
-    (* s₅ : {ty : 𝕎τ₁; tm : 𝕎τ₂; subst : 𝕃(RecSig); t1 : tm} 
-          | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : tm] *)
-    - o₅ = μ+ o₄ s₅ : 𝕃{ ... }
-  \end{lstlisting}
-
-  \columnbreak
-
-  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
-
-  \end{multicols}
-  \end{minipage}
-  \caption{Sealing and Linkage}\label{fig:sealing+linkage}
-  \end{figure}
-
-In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.
-For "tm", "ty" is late-bound and thus extending "ty" won't cause disruption on the inheritance of "tm".
-For "subst" itself, the definition of "tm" is not hidden to achieve exhaustiveness checking.
-However, for each case handler inside "subst", "tm" is late-bound and thus each handler is inheritable.
-
-Each hiding comment corresponds to the textual-aligned sealing derivation on RHS.
-We use surface syntax for eligibility: curly brackets indiate a signature (omitting sealing at type level for readability) and square bracket indicate a nested sigma type, corresponding to sealing judgement $\goodSeal{}{s}{\sigma}{A}$.
-
-For "tm", "ty" is hidden and thus the sealing "s₂" will make "ty" into an arbitrary type "𝕌" and expose its constructor "bool", "arrow", and etc.
-Then when constructing "o₂", the newly added type "tm" as "⋆ : 𝕎τ₂" is constructed in the context where "ty : 𝕌", ignoring its original eliminator.
-(Because eliminator is only possible when "ty : 𝕎τ₁").
-On the other hand, for "subst", only "ty" is hidden and thus the sealing "s₃" leave "tm : 𝕎τ₂" untouched and "Wrec" becomes possible when constructing "o₃".
-
-Basically, the hidden prior fields---inductive type are sealed into arbitrary type with ``constructors only'', where the recursion is constructed without sealing the inductive type.
-Generally speaking, each extensible inductive type are hidden for all the other fields except for the recursion on itself.
-Furthermore, normal fields (e.g.
-"t1, t2") can expose its definition to all in meta-theory by using singleton type.
-(e.g.
-saying "t1 : S(tvar 0)" instead of the current "t1 : tm"); and then overridable fields are hidden for normal fields (by sealing "t1 : S(tvar 0)" into "t1 : tm").
-Generally speaking, anything with explicit "hiding" decoration can be understood as sealing.
-
-% Intuitively, sealing corresponds to the following surface syntax\\
-% \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...} where every former field "x","y","z",...
-hide its definition when type-check "t", and remaining former fields have their definition exposed for "t".
-Thus, "x","y","z", ...
-can be overridden but other fields have to stay the same if we want to inherit "t".
-In meta-theory, we control the sealing $f$ so that "x","y","z",...
-only has type exposed and remaining fields are definitionally exposed using singleton type.
-
-% The reader should feel this subtle difference compared to our current plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
-In other words, every field in meta-theory has the right to choose how they perceive and abstract the prior fields.
-While in the plugin, "Overridable" keyword will decide how the current defining field is percived by all the following fields.
-
-% Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited by linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
-The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
-For example, $\cC\{a : \cU, b : a \to \bot \} = \Sigma(a : \cU).\Sigma (b : a \to \bot).
-\top$.
-
-Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) sigma type instead of a linkage type.
-For example, $\TyLkg{\{a : S(\bot), b : a \to \bot\}}$ can only be sealed into $\TyTkg{\{a : \cU, b : a \to \bot\}}$ but never $\TyLkg{\{a : \cU, b : a \to \bot\}}$.
-Because the later cannot be inhabitted---by projecting the second field of the $ t : \TyLkg{\{a : \cU, b : a \to \bot\}}$, we can get a proof of bottom type.
-
-% The difference between compilation type and linkage types lies in their introduction rules.
-The compilation types are just syntactic sugar for sigma types.
-However, the intro rules for the linkage types gives it the power of late-binding (also the fundation of inheritance and overriding).
-More concretely, the boxed premise in \ruleref{Lnkg Add} is making $t$ parametric on $A$, contrary to how in sigma type, the $t$ will be about a concrete $a : A$.
-This parametricity makes $t$ inheritable and the earlier stuff about $A$ late-bound.
-
-% The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
-For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators).
-Then any later fields $\goodTerm{\Gamma, A}{t}{T}$ relies only on the \textit{interface} $A$ can be inherited to other contexts (e.g.
-using extended inductive type), as long as the targeting contexts can be abstracted into the surrounding \textit{interface} for $t$ (i.e.
-$\Gamma, A$)  as well.
-
-% The importance of the compilation type lies in its difference from the linkage type.
-First we have a singleton type $"s A" : S(T) \iff "A" \equiv T$, i.e.
-a type of one single term.
-We can see that we have a term $x : \cC\{A : S(\bot), B : A \to \bot \} \vdash h_1 : \cC\{A : \cU, B:A \to \bot \}$ which is equivalent to $x : \Sigma (A: S(\bot)).
-A \to \bot \vdash h_1 : \Sigma (A: \cU).
-A \to \bot$.
-This can be considered as using seal to do an abstraction on $A:S(\bot)$.
-
-% But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited
-% and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
-This example illustrates the programming difference between the sigma type and linkage type.
-
-% However, this form of abstracting a type is very important when programming in our metatheory---as later we will show inductive type of different constructors locates in different type (universe), and to make later fields inheritable we need to work on an interface of abstracted type.
-Thus the compilation type is indispensable.
-
-% Looking closely, the sealing judgement helps to model most machinaries about abstraction in plugin implementation.
-
-% Inductive type and recursor and handlers
 
@@
 
-Our formulation of (extensible) inductive type is close to W-type (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
+We start with the formulation of (extensible) inductive type, which is close to W-type (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
 Thus we have a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$ indicating signature of inductive type by its constructors.
@@
 
-When programming via meta-theory, we usually seal the inductive type, only expose its constructors (as a mundane function member) and hide the eliminator by abstracting $T : \bW\tau$ into $"W"~T : \cU$.
-Then the later fields---those only rely on this interface with $\cU$---can be inherited to other contexts (e.g.
+% When programming via meta-theory, we usually seal the inductive type, only expose its constructors (as a mundane function member) and hide the eliminator by abstracting $T : \bW\tau$ into $"W"~T : \cU$.
+Then the following fields---those only rely on this interface with $\cU$---can be inherited to other contexts (e.g.
 using enriched inductive type), as long as the targeting contexts can be abstracted into this {interface} again.
@@
 The partial recursors and the computational axioms are treated as normal fields as expected, since we know they are provable in the plugin implementation.
+
+To introduce the concrete programming on inductive type in metatheory, we need to first introduce the formulation of families.
+
+%\YZ{
+%    Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
+%    not definitionally equal but propositionally equal in family STLC?
+%}\EDJreply{No, TLDR: it is a different level. \\ 
+%There are two type theories, one is Agda-like (= Latex = Conventional math), the other is the one formulated in QIIT.
+This sentence is saying that the Agda-like one is extensional MLTT (i.e.
+we have function extensionality and etc here in Latex writing directly).
+\\ But the one you are asking about is in the metatheory in QIIT: if we have an identity type $Id x y$ (propositionally equal) can we have a judgemetnal equality for the two terms in QIIT $x \equiv y$.
+\\
+%Apparently, we don't setup our QIIT formulation in this way (I thought we need, but now I don't think we need it. Our QIIT formulation is purely in intensional MLTT now).\\ 
+%But even if we do setup like that, I will not say a conflict is happening (because it is to our advantage if we have judgemental equal in metatheory).
+I will say it is a bit off as to how Coq is formulated (Coq is closer to an intensional MLTT, so no judgemental equal only propositional one).
+%}
+
+% linkage, signature and compilation and seal
+$$
+\Rule[name=lsig/ept]
+{""}
+{\goodSig{\Gm}{\LSigEmp}{0}}
+\quad
+\Rule[name=ty/lkg]
+{\goodSig{\Gm}{\lsig}{n}}
+{\goodType{\Gm}{\TyLkg{\lsig}}{i}
+\\ \goodType{\Gm}{\TyTkg{\lsig}}{}}
+\quad
+\Rule[name=Lnkg Compile]
+{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+}
+{
+  \goodTerm{\Gamma}{\Tkg {o}}{\cC \sigma}
+}
+$$
+$$
+\Rule[name=Sig Add]
+{\goodSig{\Gamma}{\sigma}{n} 
+ \quad \goodType{\Gamma}{A}{}
+ \quad \goodSeal{\Gamma}{f}{\sigma}{A}
+ \quad \goodType{\Gamma, A}{T}{i}}
+{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
+\quad
+\Biggl( \boxed{\goodSeal{\Gamma}{s}{\sigma}{A}} 
+:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[p^1]}} \Biggr)
+$$
+$$
+\Rule[name=Ept Lnkg]
+{}
+{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
+\quad
+\Rule[name=Lnkg Add]
+{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+\quad \goodType{\Gamma}{A}{}
+\quad  {\goodSeal{\Gamma}{f}{\sigma}{A}}
+ \quad \boxed{\goodTerm{\Gamma, A}{t}{T}}
+}
+{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
+$$
+$$
+\Rule[name=lkg/proj]
+{\goodTerm{\Gm}{o}{\TyLkg{\lsig}}}
+{\goodTerm{\Gm}{\lkgproj{1} {o}}{\TyLkg {(\lsigproj{1} {\lsig})}}
+\\ \boxed{\goodTerm{\Gm, \lsigprojT {\lsig}}{\lkgproj{2}{o}}{\lsigproj{2}{\lsig}}}
+}
+$$
+
+We start with signature (in a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
+A signature is a list of types contained by a \textit{linkage}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
+We will use $\{a : T_1, b : T_2, ..\}$ as a shorthand for signature with type $T_1, T_2, ..$ for readability.
+
+\ruleref{lkg/proj} indicates how linkage captures the idea of \textit{code reuse}, as $\lkgproj{2}{-}$ will lead to the boxed term that is parametric on the surrounding linkage $\lsigprojT{\lsig}$ .
+To code reuse, we simply project out a field from the parent and add back to the children family.
+\newpage 
+Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
+Sealing is corresponding \textit{to indicating whether each field is overridable(extensible) or non-overridable (thus definitionally exposed) in the Coq Plugin}.
+Concretely, we encode our STLC example to show the motivation of the sealing.
+%\YZ{Is sealing used only for late-binding the signatures of inductive types?
+Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}\EDJreply{All late-bound are achieved by ``selective-abstraction'' imposed by sealing.
+For example, recursion will be consequently abstracted because of the abstraction of the inductive type it recurse on.
+%}\EDJreply{Sorry, I see what you mean.
+Yes.
+the box part is important for late-binding.
+But the functionality of sealing is always there.
+They are both required to achieve overriding.
+A illustrative example is, if the type is $S(\bot)$, then even with the boxed premise, it is still not overridable because the type itself restricts all the possibilities.}\YZreply{
+%  By "selective abstraction", do you mean in the metatheory the programmer chooses what and how to abstract?
+%}\EDJreply{Yes.}
+
+% For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
+Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.
+
+% We can now point out how "Overridable" and "pins" is modeled.
+Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
+hide its definition when type-check t, and remaining former fields have their definition exposed in t.
+% This above elaborated "hiding" syntax directly corresponding to the $f$ in $(\mu^+ ~o~\{f\}~t)$, where each fields decides the abstraction of the prior.
+
+\begin{figure}[!htb]
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
+  }
+
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{3}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting}
+  Family STLC. 
+   FInductive ty (* no hiding *)
+
+   FInductive tm (* hiding ty *)
+
+   FRecursion subst
+     Case ... (* hiding ty tm *)
+     ...
+   End subst (* hiding ty *)
+
+   Field t1 := (tvar 0) 
+   (* hiding ty tm subst *)
+
+   Field t2 := (tapp t1 t1)
+   (* hiding ty tm subst *)
+
+  ...
+  \end{lstlisting}
+
+  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
+
+  \columnbreak
+  % \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+
+    (* s₁  : {} | {} *)
+    - o₁  = μ+ {} s₁ ⋆ : 𝕃{ty : 𝕎τ₁}
+    (* s₂  : {ty : 𝕎τ₁} | [ty : 𝕌; bool : ty; ...] *) 
+    - o₂  = μ+ o₁ s₂ ⋆ : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂}
+    (* s₃₁  : {ty : 𝕎τ₁; tm : 𝕎τ₂} 
+            | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
+    - o₃₁  = μ+ o₂ s₃ (μ+ ..) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig)}
+    (* s₃ : {ty : 𝕎(τ₁); tm : 𝕎τ₂; subst' : 𝕃(RecSig)} 
+          | [ty : 𝕌; ... ; tm : 𝕎τ₂; subst' : 𝕃(RecSig)] *)
+    - o₃ = μ+ o₃₁ s₃ (λ (Wrec ..)) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
+
+    (* s₄ : {ty : 𝕎τ₁; ... subst : tm → nat → tm → tm } 
+          | [ty : 𝕌; ... ; tm : 𝕌; subst : tm → nat → tm → tm] *)
+    - o₄ = μ+ o₃ s₄ .. : 𝕃{ty : 𝕎τ₁; ... t₁ : tm}
+
+    (* s₅ : {ty : 𝕎τ₁; tm : 𝕎τ₂; subst : 𝕃(RecSig); t1 : tm} 
+          | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : tm] *)
+    - o₅ = μ+ o₄ s₅ : 𝕃{ ... }
+  \end{lstlisting}
+
+  \columnbreak
+
+  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+  \caption{Sealing and Linkage}\label{fig:sealing+linkage}
+  \end{figure}
+
+In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.
+For "tm", "ty" is late-bound and thus extending "ty" won't cause disruption on the inheritance of "tm".
+For "subst" itself, the definition of "tm"  cannot hidden for the sake of exhaustiveness checking.
+However, for each case handler inside "subst", "tm" is late-bound and thus each handler is inheritable.
+
+Each hiding comment corresponds to the textual-aligned sealing derivation on RHS.
+For eligibility, curly brackets indiate a signature (omitting sealing at type level for readability) and square bracket indicate a nested sigma type, corresponding to sealing judgement $\goodSeal{}{s}{\sigma}{A}$.
+
+For "tm", "ty" is hidden and thus the sealing "s₂" will make "ty" into an arbitrary type "𝕌" and expose its constructor "bool", "arrow", and etc.
+Then when constructing "o₂", the newly added field "tm" as "⋆ : 𝕎τ₂" is constructed in the context where "ty : 𝕌", ignoring its original eliminator.
+(Because eliminator is only possible when "ty : 𝕎τ₁").
+On the other hand, for "subst", only "ty" is hidden and thus the sealing "s₃" leave "tm : 𝕎τ₂" untouched and "Wrec" becomes possible when constructing "o₃".
+
+Basically, the hidden prior fields---inductive type are sealed into arbitrary type with ``constructors only'', where the recursion is constructed without sealing the inductive type.
+Generally speaking, each extensible inductive type are hidden for all the other fields except for the recursion on itself.
+Furthermore, normal fields (e.g.
+"t1, t2") can expose its definition to all other fields in meta-theory by using singleton type.
+(e.g.
+using "t1 : S(tvar 0)" instead of the current "t1 : tm"); and then overridable fields are hidden for normal fields (by sealing "t1 : S(tvar 0)" into "t1 : tm").
+Generally speaking, anything with explicit "hiding" decoration can be understood as sealing.
+
+% Intuitively, sealing corresponds to the following surface syntax\\
+% \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...} where every former field "x","y","z",...
+hide its definition when type-check "t", and remaining former fields have their definition exposed for "t".
+Thus, "x","y","z", ...
+can be overridden but other fields have to stay the same if we want to inherit "t".
+In meta-theory, we control the sealing $f$ so that "x","y","z",...
+only has type exposed and remaining fields are definitionally exposed using singleton type.
+
+% The reader should feel this subtle difference compared to our current plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
+In other words, every field in meta-theory has the right to choose how they perceive and abstract the prior fields.
+While in the plugin, "Overridable" keyword will decide how the current defining field is percived by all the following fields.
+
+% Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited by linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
+The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
+For example, $\cC\{a : \cU, b : a \to \bot \} = \Sigma(a : \cU).\Sigma (b : a \to \bot).
+\top$.
+
+Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) sigma type instead of a linkage type.
+For example, $\TyLkg{\{a : S(\bot), b : a \to \bot\}}$ can only be sealed into $\TyTkg{\{a : \cU, b : a \to \bot\}}$ but never $\TyLkg{\{a : \cU, b : a \to \bot\}}$.
+Because the latter cannot be inhabitted---by projecting the second field of the $ t : \TyLkg{\{a : \cU, b : a \to \bot\}}$, we can get a proof of bottom type.
+For simplicity, we make sealing a term in the context of compilation type instead of linkage type.
+
+% The difference between compilation type and linkage types lies in their introduction rules.
+The compilation types are just syntactic sugar for sigma types.
+However, the intro rules for the linkage types gives it the power of late-binding (also the fundation of inheritance and overriding).
+More concretely, the boxed premise in \ruleref{Lnkg Add} is making $t$ parametric on $A$, contrary to how in sigma type, the $t$ will be about a concrete $a : A$.
+This parametricity makes $t$ inheritable and the earlier stuff about $A$ late-bound.
+
+% The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
+For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators).
+Then any later fields $\goodTerm{\Gamma, A}{t}{T}$ relies only on the \textit{interface} $A$ can be inherited to other contexts (e.g.
+using extended inductive type), as long as the targeting contexts can be abstracted into the surrounding \textit{interface} for $t$ (i.e.
+$\Gamma, A$)  as well.
+
+% The importance of the compilation type lies in its difference from the linkage type.
+First we have a singleton type $"s A" : S(T) \iff "A" \equiv T$, i.e.
+a type of one single term.
+We can see that we have a term $x : \cC\{A : S(\bot), B : A \to \bot \} \vdash h_1 : \cC\{A : \cU, B:A \to \bot \}$ which is equivalent to $x : \Sigma (A: S(\bot)).
+A \to \bot \vdash h_1 : \Sigma (A: \cU).
+A \to \bot$.
+This can be considered as using seal to do an abstraction on $A:S(\bot)$.
+
+% But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited
+% and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
+This example illustrates the programming difference between the sigma type and linkage type.
+
+% However, this form of abstracting a type is very important when programming in our metatheory---as later we will show inductive type of different constructors locates in different type (universe), and to make later fields inheritable we need to work on an interface of abstracted type.
+Thus the compilation type is indispensable.
+
+% Looking closely, the sealing judgement helps to model most machinaries about abstraction in plugin implementation.
+
+% Inductive type and recursor and handlers
 
@@
 % Inhertance judgement
-In the plugin implementation, "Family"s are second class objects like Coq's "Module".
+Currently we can achieve inheritance by projecting out the field, and then add back the projection into the children family.
+That means the formulation of linkage already captures the concept of \textit{code reuse}.
+
+However, in the plugin implementation, we use "Family"s to organize the code reuse.
+They are second class objects like Coq's "Module".
 What's more, a "Family" (with or without "extends" clause) can be considered as a standalone piece of code trying to \textit{inherit} an existent (but maybe empty) family.

### 2022-10-27 metatheory-condensed.tex
Comments on Linkage achieves code reuse Linkage Transformer is only a library function.

@@
 
-\ruleref{lkg/proj} indicates how linkage captures the idea of \textit{code reuse}, as $\lkgproj{2}{-}$ will lead to the boxed term that is parametric on the surrounding linkage $\lsigprojT{\lsig}$ .
+\ruleref{lkg/proj} indicates how linkage captures the idea of \textit{code reuse}, as $\lkgproj{2}{-}$ will lead to the boxed term that is parametric on the abstracted interface of linkage $\lsigprojT{\lsig}$ .
 To code reuse, we simply project out a field from the parent and add back to the children family.
@@
 Thus we model them using \textit{inheritance judgement} $\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}$ instead of considering them as first class terms.
+Inheritance judgement is like ``library function'' upon linkage for achieving inheritance and overriding.
 The only way to use $h$ is by \ruleref{Inh} and transforming a linkage of signature $\sigma_1$ to another linkage of $\sigma_2$.

### 2022-10-27 introduction.tex
soften claim in intro

--- 2022-10-27 introduction.tex
@@
 
-\item We present, to our knowledge, the first\EDJ{Are you sure? This claim is so bold.}\YZreply{Is there any previous language designs in this space?} language design that enables extensible
+\item We present a
+%\EDJ{Are you sure?
+This claim is so bold.}\YZreply{Is there any previous language designs in this space?}\EDJreply{Olivier Boite.
+2004.
+Proof reuse with extended inductive types.
+In International Conference on Theorem Proving in
+%Higher Order Logics. Springer, 50–65.  I think I also mentioned several paper in the related work.}
+language design that enables extensible
 metatheory mechanization in a higher-order, dependent type theory with

### 2022-10-27 metatheory-condensed.tex
Fixing typing rule in the main text
address a comment

@@
 
-\cref{fig:typing-selected} presents the syntax of \TT.
+\cref{fig:typing-selected} presents selected syntax and typing rules of \TT.
 The new forms (relative to MLTT) are highlighted.
+We briefly review the base MLTT fragment first.
 
@@
 %
-Formation rules for substitutions have\EDJ{It seems to me G |- g : D is a judgement not a rule.
+Typing judgments for substitutions have the form 
+%\EDJ{It seems to me G |- g : D is a judgement not a rule.
 Rules usually have two parts (premises + conclusion).
@@
 I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this.
-In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) } the form
+In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
 $\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
@@
 $$
-\Rule[]
-{}
-{\goodWSig{\Gamma}{w\cdot}{0}}
+\Rule[name=wsig/ept]
+{""}
+{\goodWSig{\Gm}{\WSigEmp}{0}}
 \quad 
-\Rule[name=WSig-Add]
-{\goodWSig{\Gamma}{\tau}{n}
-  \quad \goodType{\Gamma}{A}{i}
-  \quad \goodType{\Gamma, A}{B}{i}}
-{\goodWSig{\Gamma}{w^+ \  \tau \  A \  B}{n+1}}
+\Rule[name=wsigeq/ept/eta]
+{\goodWSig{\Gm}{\wsig}{0}}
+{\goodWSig{\Gm}{\wsig \equiv \WSigEmp}{0}}
 \quad
-\Rule[name=Ind-Univ]
-{\goodWSig{\Gamma}{\tau}{n}}
-{\goodType{\Gamma}{\bW \tau}{i}
+\Rule[name=wsig/add]
+{\goodWSig{\Gm}{\wsig}{n}
+  \quad \goodType{\Gm}{A}{i}
+  \quad \goodType{\Gm, A}{B}{i}}
+{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}}
+\quad
+\Rule[name=wsig/universe]
+{\goodWSig{\Gm}{\wsig}{n}}
+{\goodType{\Gm}{\TyWSingle {\wsig}}{i}
+\\ \goodTerm{\Gm}{\wsingle{\wsig}}{\TyWSingle{\wsig}}
 }
@@
 $$
-\Rule[name=Ind-Term]
-{\goodTerm{\Gamma}{T}{\bW \tau}
-  \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
-  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)[p^1]}}
-{\goodTerm{\Gamma}{"Wsup"~j~T~a~b}{El~("W"~T)} }
+\Rule[name=ind/cstr]
+{{\goodWSig{\Gm}{\wsig}{n}}
+  \\ \goodTerm{\Gm}{a}{\wsigproj{j}{1}{\wsig}}
+  \\ \goodTerm{\Gm, \sub{\wsigproj{j}{2}{\wsig}}{(id, a)}}{b}{\El{(\wcode{\wsig})}}}
+{\goodTerm{\Gm}{\wsup{T}{a}{b}}{\El{(\wcode{\wsig})}} }
 $$
 $$
-\Rule[name=Hdler]
-{\goodType{\Gamma}{A}{}
-\quad \goodType{\Gamma, A}{B}{}
-\quad \goodType{\Gamma}{R}{}
+\Rule[name=ty/casesig]
+{\goodType{\Gm}{A}{}
+\quad \goodType{\Gm, A}{B}{}
+\quad \goodType{\Gm}{R}{}
 }
-{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv \Pi A\; \Pi (\Pi B (R[p^2]))\; R[p^2]}{}}
+{\goodType{\Gm}{\CaseSig{A}{B}{R} \equiv (\TyPi {A} {(\TyPi {(\TyPi {B} {(\sub{R}{ \SubstComp {\pi_1}{\pi_1}})})} {\sub{R}{\SubstComp{\pi_1}{\pi_1}}})} )}{}}
 \quad 
@@
 $$
-\Rule[name=Hdler-Proj]
-{ \goodWSig{\Gamma}{\tau}{N}
-\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\tau~R)}
-\quad j < N
+\Rule[name=reclkg/proj]
+{ \goodWSig{\Gm}{\wsig}{N}
+\\ \goodTerm{\Gm}{o}{\TyLkg{(\RecSig{\wsig}{R})}}
+\\ j < N
 }
-{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}
+{\goodTerm{\Gm}{\Recproj{j}{o}}{\sub{(\CaseSig{(\pi_1^j \wsig)}{(\pi_2^j \wsig)}{R})}{\pi_1}}}
 $$
 $$
-\Rule[name=Rec-Constr]
-{ 
-  % \goodWSig{\Gamma}{\tau}{N}
-\quad \goodTerm{\Gamma}{T}{\bW \tau}
-% \quad \goodType{\Gamma}{R}{}
-\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
-\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
+\Rule[name=tm/wrec]
+{ \goodWSig{\Gm}{\wsig}{N}
+\\ \goodType{\Gm}{R}{}
+\\ \goodTerm{\Gm}{r}{\TyLkg{(\RecSig{\wsig}{R})}}
+\\ \goodTerm{\Gm}{w}{\El{(\wcode{\wsig})}}
 }
-{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
+{\goodTerm{\Gm}{\Wrec{w}{r}}{R}}
 \YZ{I suppose Rec-Constr is only for exhaustivity checking?
@@
 $$
-\Rule[name=Rec-Beta]
-{\goodTerm{\Gamma}{h}{\cL("RecSig"~\tau~R)} 
-\\ \goodTerm{\Gamma}{T}{\bW\tau}}
-{\goodTerm{\Gamma}{"Wrec"~h~("Wsup"~j~T~a~b) \equiv 
-("app"~(("app"~("prjR"~j~h))[("id", a)]))[("id", "lam" ("Wrec"~h[p^1]~b))]
+\Rule[name=tmeq/rec/beta]
+{\goodTerm{\Gamma}{h}{\cL(\RecSig{\tau}{R})} 
+\\ \goodTerm{\Gamma}{T}{\TyWSingle{\tau}}}
+{\goodTerm{\Gamma}{\Wrec{h}{(\wsup{\wsig}{a}{b})} \equiv 
+(\app{((\app{(\Recproj{j}{h})} )[("id", a)])} )[("id", \lam{(\Wrec{h[\pi_1]}{b})})]
 }{R}}\EDJ{"app" and "lam" are lambda application and abstraction}
@@
 \quad
-\Rule[name=Lnkg Compile]
-{ \goodTerm{\Gamma}{o}{\cL \sigma} 
+\Rule[name=lnkg/pkg]
+{ \goodTerm{\Gm}{o}{\TyLkg {\lsig}} 
 }
 {
-  \goodTerm{\Gamma}{\Tkg {o}}{\cC \sigma}
+  \goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\lsig}}
 }
@@
 $$
-\Rule[name=Sig Add]
-{\goodSig{\Gamma}{\sigma}{n} 
- \quad \goodType{\Gamma}{A}{}
- \quad \goodSeal{\Gamma}{f}{\sigma}{A}
- \quad \goodType{\Gamma, A}{T}{i}}
-{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
+\Rule[name=lsig/add]
+{\goodSig{\Gm}{\lsig}{n} 
+ \quad \goodType{\Gm}{A}{}
+ \quad \goodSeal{\Gm}{\seal}{\lsig}{A}
+ \quad \goodType{\Gm, A}{T}{i}}
+{\goodSig{\Gm}{(\LSigAdd {\lsig}{\seal}{T})}{n+1}}
 \quad
@@
 $$
-\Rule[name=Ept Lnkg]
-{}
-{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
+\Rule[name=lnkg/ept]
+{""}
+{\goodTerm{\Gm}{\LkgEmp}{\TyLkg {\LSigEmp}}  
+}
 \quad
-\Rule[name=Lnkg Add]
-{ \goodTerm{\Gamma}{o}{\cL \sigma} 
-\quad \goodType{\Gamma}{A}{}
-\quad  {\goodSeal{\Gamma}{f}{\sigma}{A}}
- \quad \boxed{\goodTerm{\Gamma, A}{t}{T}}
+\Rule[name=lkg/add]
+{ \goodTerm{\Gm}{o}{\TyLkg {\lsig}} 
+\\ \goodType{\Gm}{A}{}
+\\  \goodSeal{\Gm}{\seal}{\lsig}{A} 
+\\  \boxed{\goodTerm{\Gm, A}{t}{T}}
 }
-{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
+{\goodTerm{\Gm}{(\LkgAdd {o} {t})}{\TyLkg{\LSigAdd {\lsig} {\seal} {T}}}
+}
 $$

### 2022-10-27 metatheory-condensed.tex
Tarski-style universes

@@
 
-\cref{fig:typing-selected} presents selected syntax and typing rules of \TT.
+\cref{fig:typing-selected} presents the syntax and selected typing rules of \TT.
 The new forms (relative to MLTT) are highlighted.
@@
 
-%Note that we are actually using predicative universe, so we actually have different levels of universes $\cU_i$. 
-%We omit universe levels in our latex presentation for conciseness and readability.
+We use Tarski-style universes~\cite{hofmann1997syntax}: a universe~$\cU$
+is inhabited by the \emph{codes} of types, with $\El{t}$ decoding~$t$ and
+$\codety{T}$ encoding type~$T$ (rules \ruleref{ty/el} and \ruleref{tm/c}).
+MLTT has an infinite hierarchy of universes;
+we omit universe levels for conciseness.
 

### 2022-10-27 metatheory-condensed.tex
fix unicode chars in listing

@@
 We contribute \TT, a core type theory that incorporates the core elements
-of \underline{f}amily polymorphism into
+of \underline{f}a\underline{m}i\underline{l}y polymorphism into
 \underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory~\cite{martin1982constructive},
@@
 In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
-$\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
-valid in the context~$\Dl$ yields terms valid in~$\Gm$. 
+$\goodSub{\Gm}{\gm}{\Dl}$.
+The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
+yields terms valid in~$\Gm$ (e.g., rule \ruleref{tm/sub}).
 %
@@
 
-    (* s₁  : {} | {} *)
-    - o₁  = μ+ {} s₁ ⋆ : 𝕃{ty : 𝕎τ₁}
-    (* s₂  : {ty : 𝕎τ₁} | [ty : 𝕌; bool : ty; ...] *) 
-    - o₂  = μ+ o₁ s₂ ⋆ : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂}
-    (* s₃₁  : {ty : 𝕎τ₁; tm : 𝕎τ₂} 
+    (* s[@₁@]  : {} | {} *)
+    - o[@₁@]  = μ+ {} s[@₁@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]}
+    (* s[@₂@]  : {ty : 𝕎τ[@₁@]} | [ty : 𝕌; bool : ty; ...] *) 
+    - o[@₂@]  = μ+ o[@₁@] s[@₂@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]}
+    (* s[@₃@][@₁@]  : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]} 
             | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
-    - o₃₁  = μ+ o₂ s₃ (μ+ ..) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig)}
-    (* s₃ : {ty : 𝕎(τ₁); tm : 𝕎τ₂; subst' : 𝕃(RecSig)} 
-          | [ty : 𝕌; ... ; tm : 𝕎τ₂; subst' : 𝕃(RecSig)] *)
-    - o₃ = μ+ o₃₁ s₃ (λ (Wrec ..)) : 𝕃{ty : 𝕎τ₁; tm : 𝕎τ₂; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
+    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] (μ+ ..) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)}
+    (* s[@₃@] : {ty : 𝕎(τ[@₁@]); tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)} 
+          | [ty : 𝕌; ... ; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)] *)
+    - o[@₃@] = μ+ o[@₃@][@₁@] s[@₃@] (λ (Wrec ..)) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
 
-    (* s₄ : {ty : 𝕎τ₁; ... subst : tm → nat → tm → tm } 
+    (* s[@₄@] : {ty : 𝕎τ[@₁@]; ... subst : tm → nat → tm → tm } 
           | [ty : 𝕌; ... ; tm : 𝕌; subst : tm → nat → tm → tm] *)
-    - o₄ = μ+ o₃ s₄ .. : 𝕃{ty : 𝕎τ₁; ... t₁ : tm}
+    - o[@₄@] = μ+ o[@₃@] s[@₄@] .. : 𝕃{ty : 𝕎τ[@₁@]; ... t[@₁@] : tm}
 
-    (* s₅ : {ty : 𝕎τ₁; tm : 𝕎τ₂; subst : 𝕃(RecSig); t1 : tm} 
+    (* s[@₅@] : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst : 𝕃(RecSig); t1 : tm} 
           | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : tm] *)
-    - o₅ = μ+ o₄ s₅ : 𝕃{ ... }
+    - o[@₅@] = μ+ o[@₄@] s[@₅@] : 𝕃{ ... }
   \end{lstlisting}
@@
 Each hiding comment corresponds to the textual-aligned sealing derivation on RHS.
-For eligibility, curly brackets indiate a signature (omitting sealing at type level for readability) and square bracket indicate a nested sigma type, corresponding to sealing judgement $\goodSeal{}{s}{\sigma}{A}$.
+For legibility, curly brackets indicate a signature (omitting sealing at type level for readability) and square bracket indicate a nested sigma type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.
 

### 2022-10-28 metatheory-condensed.tex
W-types

@@
 We use Tarski-style universes~\cite{hofmann1997syntax}: a universe~$\cU$
-is inhabited by the \emph{codes} of types, with $\El{t}$ decoding~$t$ and
-$\codety{T}$ encoding type~$T$ (rules \ruleref{ty/el} and \ruleref{tm/c}).
+is inhabited by the \emph{codes} of types,
+with $\El{t}$ decoding~$t$ (rule \ruleref{ty/el})
+and $\codety{T}$ encoding type~$T$ (rule \ruleref{tm/c}).
 MLTT has an infinite hierarchy of universes;
-we omit universe levels for conciseness.
+we omit universe levels in the presentation for conciseness.
+
+\noindentparagraph{W-types.}
+
+W-types~\cite{martin1984intuitionistic} are a standard way to model
+inductive types in MLTT.
+A constructor of some inductive type~$T$ is given by a pair of types
+$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
+given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
+$\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
+it constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
+As the rules suggest, our formulation additionally supports
+\emph{W-type signatures} (judgment form \goodWSig{\Gm}{\wsig}{n})
+composed of $n$ pairs of types, each of which represents a constructor.
 
@@
             | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
-    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] (μ+ ..) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)}
-    (* s[@₃@] : {ty : 𝕎(τ[@₁@]); tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)} 
+    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] ([@μ@]+ ..) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)}
+    (* s[@₃@] : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)} 
           | [ty : 𝕌; ... ; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)] *)
@@
 Each hiding comment corresponds to the textual-aligned sealing derivation on RHS.
-For legibility, curly brackets indicate a signature (omitting sealing at type level for readability) and square bracket indicate a nested sigma type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.
+For legibility, curly brackets indicate a signature (omitting sealing at type level for readability) and square bracket indicate a nested $\Sigma$-type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.
 
@@
 
-Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) sigma type instead of a linkage type.
+Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) $\Sigma$-type instead of a linkage type.
 For example, $\TyLkg{\{a : S(\bot), b : a \to \bot\}}$ can only be sealed into $\TyTkg{\{a : \cU, b : a \to \bot\}}$ but never $\TyLkg{\{a : \cU, b : a \to \bot\}}$.

### 2022-10-28 metatheory-condensed.tex
Wrec

@@
 
-\noindentparagraph{W-types.}
+\noindentparagraph{Introducing and eliminating inductive types.}
 
@@
 inductive types in MLTT.
-A constructor of some inductive type~$T$ is given by a pair of types
-$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
+A constructor of an inductive type~$T$ is given by a pair of types
+$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (\ruleref{wsig/add}):
 given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
 $\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
-it constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
+it constructs a term of type~$T$ (\ruleref{tm/wsup}).
 As the rules suggest, our formulation additionally supports
@@
 composed of $n$ pairs of types, each of which represents a constructor.
+
+W-types are eliminated by $\Wrec{\lkg}{t}$, where $t$ is of the W-type
+$\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of case
+handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
+The type of each case handler has the form $\CaseSig{A}{B}{T}$, where
+$T$ is the motive of the recursion (\ruleref{ty/casety}).
+This collection of case handlers then encodes the \lsti{FRecursion}
+construct supported by our plugin.
+We choose to give it a linkage type $\TyLkg{\RecSig{\wsig}{T}}$,
+but it could be typed as an ordinary $n$-tuple, which is subsumed by
+linkages to be introduced next.
+
+\noindentparagraph{Linkages.}
 

### 2022-10-29 syntax.tex

@@
 distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
-$\goodType{\Gamma}{El~T}{}$, and vice-versa.\YZ{Why is it wanted to distinguish types from their codes?}
+$\goodType{\Gamma}{El~T}{}$, and vice-versa.\YZ{Why is it wanted to distinguish types from their codes?}\EDJreply{Because in formalization, for a judgement $G |- t : T$, code can only be placed at $t$ and type can only be placed at $T$.
+In Russel style, they don't do such distinction.
+But for the sake of formalization, we will distinguish code and types.}
 

### 2022-10-29 metatheory-condensed.tex
Update Sealing Example.

@@
     (* s[@₄@] : {ty : 𝕎τ[@₁@]; ... subst : tm → nat → tm → tm } 
-          | [ty : 𝕌; ... ; tm : 𝕌; subst : tm → nat → tm → tm] *)
-    - o[@₄@] = μ+ o[@₃@] s[@₄@] .. : 𝕃{ty : 𝕎τ[@₁@]; ... t[@₁@] : tm}
+          | [.. ; tm : 𝕌; tvar : nat → tm; .. subst : tm → nat → tm → tm] *)
+    - o[@₄@] = μ+ o[@₃@] s[@₄@] [@("tvar"~0)@] : 𝕃{ty : 𝕎τ[@₁@]; ... t[@₁@] : [@\TyS{"tvar"~0}@]}
 
-    (* s[@₅@] : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst : 𝕃(RecSig); t1 : tm} 
-          | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : tm] *)
-    - o[@₅@] = μ+ o[@₄@] s[@₅@] : 𝕃{ ... }
+    (* s[@₅@] : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst : 𝕃(RecSig); t1 : [@\TyS{"tvar"~0}@]} 
+          | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : [@\TyS{"tvar"~0}@]] *)
+    - o[@₅@] = μ+ o[@₄@] s[@₅@] .. : 𝕃{ ... }
   \end{lstlisting}
@@
 Generally speaking, each extensible inductive type are hidden for all the other fields except for the recursion on itself.
-Furthermore, normal fields (e.g.
-"t1, t2") can expose its definition to all other fields in meta-theory by using singleton type.
-(e.g.
-using "t1 : S(tvar 0)" instead of the current "t1 : tm"); and then overridable fields are hidden for normal fields (by sealing "t1 : S(tvar 0)" into "t1 : tm").
-Generally speaking, anything with explicit "hiding" decoration can be understood as sealing.
+
+We choose to expose the definition of $t_1$ to all the following fields including $t_2$ by using singleton type $\TyS{-}$.
+If other fields, say $t_3$, want to hide it, just seal $t_1$ into type "tm"; and then overriding $t_1$ will not disrupt the inheritablity of $t_3$.
+But such overriding will definitely disrupt the inheritablity of $t_2$ as $s_5$ doesn't hide $t_1$.
+Generally speaking, any field that has a clear "hiding" semantic on the prior fields can be understood as sealing.
+All of our plugin commands have clear "hiding" semantics.
 

### 2022-10-29 metatheory-condensed.tex
stuff

--- 2022-10-28 metatheory-condensed.tex
@@
 \var{n} is the variable bound by the $n$-th closest enclosing binder.
-For example, $\lambda x.\,\lambda y.\,x$ is $\lam{(\lam{\var{1}})}$.
+For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
 %
@@
 inductive types in MLTT.
-A constructor of some inductive type~$T$ is given by a pair of types
+A constructor of an inductive type~$T$ is given by a pair of types
 $\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
@@
 As the rules suggest, our formulation additionally supports
-\emph{W-type signatures} (judgment form \goodWSig{\Gm}{\wsig}{n})
-composed of $n$ pairs of types, each of which represents a constructor.
+\emph{W-type signatures}.
+A well formed W-type signature \goodWSig{\Gm}{\wsig}{n} is
+composed of $n$ pairs of types, each of which represents a constructor
+(rules \ruleref{wsig/empty} and \ruleref{wsig/add}).
+
+W-types are eliminated with the form $\Wrec{\lkg}{t}$, where $t$ is of a
+W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
+case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
+Each case handler has a type of the form $\CaseSig{A}{B}{T}$, where
+$T$ is the motive of the recursion (\ruleref{ty/casety}).
+This collection of case handlers then encodes those defined or inherited
+by an \lsti{FRecursion} command in our plugin.
+We choose to type it with a linkage type $\TyLkg{\RecSig{\wsig}{T}}$
+to avoid introducing $n$-tuples, which linkages generalize.
+
+Rules \ruleref{tm/wsup} and \ruleref{tm/wrec} require knowing~$\wsig$:
+the W-type is exhaustively generated by its constructors, and
+its elimination must exhaustively handle all the constructors in
+its signature.
+%Notice that rule \ruleref{tm/wrec} requires knowing~$\wsig$:
+%typing a recursive function $\lam{\Wrec{\lkg}{\var{0}}}$ must check
+%that the case handlers~$\lkg$ exhaustively handle all the constructors
+%in the W-type signature~$\wsig$.
+
+In contrast, $\wsig$ should be hidden from the typing context of any
+term that does not invoke $\texttt{Wsup}$ or $\texttt{Wrec}$, so that
+the term can be reused---without being rechecked---for a different
+W-type signature~$\wsig'$ that extends~$\wsig$ with additional
+constructors.
+Moreover, the typing of the term should be made polymorphic to
+the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
+so that the term can be reused---without being rechecked---when those fields
+are overridden to support the extended signature~$\wsig'$.
 

### 2022-10-29 metatheory-condensed.tex
make Wrec arity 3

@@
 
-W-types are eliminated with the form $\Wrec{\lkg}{t}$, where $t$ is of a
+W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a
 W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
@@
 its signature.
-%Notice that rule \ruleref{tm/wrec} requires knowing~$\wsig$:
-%typing a recursive function $\lam{\Wrec{\lkg}{\var{0}}}$ must check
-%that the case handlers~$\lkg$ exhaustively handle all the constructors
-%in the W-type signature~$\wsig$.
 
 In contrast, $\wsig$ should be hidden from the typing context of any
-term that does not invoke $\texttt{Wsup}$ or $\texttt{Wrec}$, so that
+term that does not invoke $\Wsup{\wsig}{\cdot}{\cdot}$ or $\Wrec{\wsig}{\cdot}{\cdot}$, so that
 the term can be reused---without being rechecked---for a different
@@
   \\ \goodTerm{\Gm, \sub{\wsigproj{j}{2}{\wsig}}{(id, a)}}{b}{\El{(\wcode{\wsig})}}}
-{\goodTerm{\Gm}{\wsup{T}{a}{b}}{\El{(\wcode{\wsig})}} }
+{\goodTerm{\Gm}{\Wsup{T}{a}{b}}{\El{(\wcode{\wsig})}} }
 $$
@@
 }
-{\goodTerm{\Gm}{\Wrec{w}{r}}{R}}
+{\goodTerm{\Gm}{\Wrec{\wsig}{w}{r}}{R}}
 \YZ{I suppose Rec-Constr is only for exhaustivity checking?
@@
 \\ \goodTerm{\Gamma}{T}{\TyWSingle{\tau}}}
-{\goodTerm{\Gamma}{\Wrec{h}{(\wsup{\wsig}{a}{b})} \equiv 
-(\app{((\app{(\Recproj{j}{h})} )[("id", a)])} )[("id", \lam{(\Wrec{h[\pi_1]}{b})})]
+{\goodTerm{\Gamma}{\Wrec{\wsig}{h}{(\Wsup{\wsig}{a}{b})} \equiv 
+(\app{((\app{(\Recproj{j}{h})} )[("id", a)])} )[("id", \lam{(\Wrec{\wsig}{h[\pi_1]}{b})})]
 }{R}}\EDJ{"app" and "lam" are lambda application and abstraction}

### 2022-10-30 metatheory-condensed.tex
Removing bWtau

@@
     (* s[@₁@]  : {} | {} *)
-    - o[@₁@]  = μ+ {} s[@₁@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]}
-    (* s[@₂@]  : {ty : 𝕎τ[@₁@]} | [ty : 𝕌; bool : ty; ...] *) 
-    - o[@₂@]  = μ+ o[@₁@] s[@₂@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]}
-    (* s[@₃@][@₁@]  : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]}
+    - o[@₁@]  = μ+ {} s[@₁@] [@$W\tau_1$@] : 𝕃{ty : [@$\TyS{W\tau_1}$@]}
+    (* s[@₂@]  : {ty : [@$\TyS{W\tau_1}$@]} | [ty : 𝕌; bool : ty; ...] *) 
+    - o[@₂@]  = μ+ o[@₁@] s[@₂@] [@$W\tau_2$@] : 𝕃{ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]}
+    (* s[@₃@][@₁@]  : {ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]}
             | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
-    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] ([@μ@]+ ..) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)}
-    (* s[@₃@] : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)} 
-          | [ty : 𝕌; ... ; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)] *)
-    - o[@₃@] = μ+ o[@₃@][@₁@] s[@₃@] (λ (Wrec ..)) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
+    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] ([@μ@]+ ..) : 𝕃{ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig)}
+    (* s[@₃@] : {ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig)} 
+          | [ty : 𝕌; ... ; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig)] *)
+    - o[@₃@] = μ+ o[@₃@][@₁@] s[@₃@] (λ (Wrec ..)) : 𝕃{ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
 
-    (* s[@₄@] : {ty : 𝕎τ[@₁@]; ... subst : tm → nat → tm → tm } 
+    (* s[@₄@] : {ty : [@$\TyS{W\tau_1}$@]; ... subst : tm → nat → tm → tm } 
           | [.. ; tm : 𝕌; tvar : nat → tm; .. subst : tm → nat → tm → tm] *)
-    - o[@₄@] = μ+ o[@₃@] s[@₄@] [@("tvar"~0)@] : 𝕃{ty : 𝕎τ[@₁@]; ... t[@₁@] : [@\TyS{"tvar"~0}@]}
+    - o[@₄@] = μ+ o[@₃@] s[@₄@] [@("tvar"~0)@] : 𝕃{ty : [@$\TyS{W\tau_1}$@]; ... t[@₁@] : [@\TyS{"tvar"~0}@]}
 
-    (* s[@₅@] : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst : 𝕃(RecSig); t1 : [@\TyS{"tvar"~0}@]} 
+    (* s[@₅@] : {ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst : 𝕃(RecSig); t1 : [@\TyS{"tvar"~0}@]} 
           | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : [@\TyS{"tvar"~0}@]] *)
@@
 Each hiding comment corresponds to the textual-aligned sealing derivation on RHS.
-For legibility, curly brackets indicate a signature (omitting sealing at type level for readability) and square bracket indicate a nested $\Sigma$-type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.
+For readability, RHS is informal---curly brackets indicate a signature (omitting sealing at type level) and square bracket indicate a nested $\Sigma$-type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.
 

### 2022-10-30 metatheory-condensed.tex
add linkage rules in selected-typing.tex

--- 2022-10-29 metatheory-condensed.tex
@@
 %
-The two main forms of substitutions are weakening and extension:
+The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
 \sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
@@
 For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
-of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.
+that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
+$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.
 
@@
 constructors.
-Moreover, the typing of the term should be made polymorphic to
+Moreover, the typing of the term should be made parametric to
 the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
@@
 are overridden to support the extended signature~$\wsig'$.
+%
+Such abstraction required by family polymorphism is supported via linkages.
+
+\noindentparagraph{Family polymorphism via linkages.}
+
+Family polymorphism requires late binding.
+In \TT, families are expressed through linkages, and late binding of
+field references is achieved by requiring that typing be polymorphic to
+the definition of that field.
 

### 2022-10-31 metatheory-condensed.tex
stuff

@@
 
-W-types~\cite{martin1984intuitionistic} are a standard way to model
-inductive types in MLTT.
-A constructor of an inductive type~$T$ is given by a pair of types
+W-types~\cite{martin1984intuitionistic} are a succinct way to model
+inductive types in MLTT;
+together with the identity type $\TyId{\cdot}{\cdot}$, they can express a
+whole host of inductive types~\cite{hugunin2021not}.
+%
+A constructor is given by a pair of types
 $\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
@@
 $\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
-it constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
+$\Wsup{\wsig}{t_1}{t_2}$ constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
 As the rules suggest, our formulation additionally supports
@@
 composed of $n$ pairs of types, each of which represents a constructor
-(rules \ruleref{wsig/empty} and \ruleref{wsig/add}).
+(rules \ruleref{wsig/empty} and \ruleref{wsig/add}),
+with $\wsigproj{i}{1}{\wsig}$ and $\wsigproj{i}{2}{\wsig}$ giving the
+$i$-th pair in~$\wsig$, and $\wcode{\wsig}$ giving the code of the W-type (\ruleref{tm/w}).
 
@@
 the definition of that field.
+
+Well-formedness judgments of \emph{linkage signatures} have form
+$\goodSig{\Gm}{\lsig}{n}$.
+As rules \ruleref{lsig/empty} and \ruleref{lsig/add} show, a linkage
+signature is a list of $n$~types.
+%
+Well-formedness judgments of \emph{linkages} have form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
+where $\TyLkg{\lsig}$ is a type formed by~$\lsig$.
+As rules \ruleref{l/empty} and \ruleref{l/add} show, a linkage is a list
+of $n$~terms, each representing a field of the family modeled by the linkage.
+
+In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
+responsible for late binding.
+Here, $A$ abstracts the fields preceding the current field~$t$
+(i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
+Crucially, unlike the second field in a dependent pair (\ruleref{tm/pair}),
+the typing of~$t$ in \ruleref{l/add} is polymorphic to how the fields in
+$t$'s context are defined, so that a different linkage that overrides
+fields in $t$'s context can reuse~$t$---without retyping it---by
+projecting it (\ruleref{l/proj2}) and then including it.
+
+This type~$A$ abstracting the context in \ruleref{l/add} and
+\ruleref{lsig/add} is not necessarily the same as those recorded by~$\lsig$,
+because a field representing the code $\wcode{\wsig}$ of a W-type has to
+expose different types to different succeeding fields.
+Succeeding fields that invoke
+$\Wsup{\wsig}{\cdot}{\cdot}$ or $\Wrec{\wsig}{\cdot}{\cdot}$
+should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
+and \ruleref{tm/wrec} stipulate, whereas $\wsig$ should be hidden
+from all other fields so that they can be reused
+for a W-type signature~$\wsig'$ extending~$\wsig$.
+
+We illustrate this with \cref{fig:stlc-linkage-typing}.
+On the left is code excerpted from \cref{fig:stlc-mechanized}, and on the right
+is how the corresponding fields are modeled and typed in \TT.
+%
+$\lkg_i$ is the linkage that adds the $i$-th field~$t_i$ with the typing
+$\goodTerm{A_i}{t_i}{T_i}$.
+%
+Here, we feel free to use a notation with explicit names.
+We name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
+and feel free to write its type, which is a dependent tuple type, as a
+dependent record type with explicit labels.
+%
+\begin{itemize}[labelsep=*,leftmargin=1pc]
+
+\item 
+The first field $t_1$ is defined as $\wcode{\wsig_{\texttt{tm}}}$
+and given the singleton type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$, where $\wsig_{\texttt{tm}}$
+is the W-type signature for \lsti{tm}.
+This typing is recorded by $\lsig_1$ and is thus available in all $\lsig_i$'s.
+
+\item 
+The next four fields model the four constructors of~\lsti{tm}.
+Constructor \lsti{tm_unit} is modeled as $\Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}$
+and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
+the typing context containing the first field~\lsti{tm}.
+Because the W-type signature $\wsig_{\texttt{tm}}$ is exposed to this typing context,
+$\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
+as required by rule \ruleref{tm/wsup}.
+
+\item
+Similarly, $\wsig_{\texttt{tm}}$ is exposed to the typing context of
+$t_{10}$, which models the recursive function \lsti{subst} by invoking
+the recursor $\Wrec{\wsig_{\texttt{tm}}}{\cdot}{\cdot}$.
+%
+Other fields automatically generated by the plugin, such as partial
+recursors and various propositional equalities, can also be modeled in
+this way by invoking $\texttt{Wrec}$.
+
+\item
+By contrast, $\wsig_{\texttt{tm}}$ is hidden from the typing context of all other fields.
+Their typing depends on the knowledge that \lsti!tm! has type~$\cU$,
+rather than \TyS{\wcode{\wsig_{\texttt{tm}}}},
+so they can be reused in a context where \lsti!tm! is defined as $\wcode{\wsig'_{\texttt{tm}}}$,
+where $\wsig'_{\texttt{tm}}$ extends~$\wsig_{\texttt{tm}}$ with additional constructors.
+
+\end{itemize}
+
+The third premise
+$\goodTerm{\Gm,\TyTkg{\lsig}}{\seal}{\sub{A}{\SubstWeak{1}}}$
+in rules \ruleref{l/add} and \ruleref{lsig/add} is
+responsible for hiding W-type signatures in~$\lsig$.
+Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
+packaging the types of all fields preceding the current field~$t$,
+and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
+$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new term of type $A$
+that hides W-type signatures, if necessary.
+When no hiding is needed, $\seal$ is the identity $\var{0}$ and $A$ is
+$\TyTkg{\lsig}$.
+
+\bigskip
+
+A supplemental appendix gives a more complete example showing how families
+are modeled and type-checked as linkages.
+
+\begin{figure}
+\small
+\newcolumntype{L}{>{$}l<{$}}
+\newcolumntype{C}{>{$}c<{$}}
+\newcolumntype{R}{>{$}r<{$}}
+\begin{tabular}{@{}lR@{\ }C@{\ }LR@{\ \,}C@{\,\ }L@{}}
+\lsti!Family STLC.!
+& \lkg_0 & \coloneqq & \LkgEmp
+\\
+\lsti!FInductive tm :=!
+& \lkg_1 & \coloneqq & \LkgAdd{\lkg_0}{\dadada}
+& \mathit{self} : []
+& \vdash
+& \wcode{\wsig_{\texttt{tm}}} : \TyS{\wcode{\wsig_{\texttt{tm}}}}
+\\
+\lsti!| tm_unit : tm | !\dadada
+& \lkg_2 & \coloneqq & \LkgAdd{\lkg_1}{\dadada}
+& \mathit{self} : [\texttt{tm}:\TyS{\wcode{\wsig_{\texttt{tm}}}}]
+& \vdash
+& \Wsup{\wsig_{\texttt{tm}}}{\top}{\bot} : \El{\mathit{self.}\texttt{tm}}
+\\
+\lsti!FRecursion subst !\dadada
+\\
+\lsti!Case tm_unit := !\dadada
+& & &
+& \mathit{self} : [\texttt{tm}:\cU, \texttt{tm\_unit}: \El{\texttt{tm}}, \dadada]
+& \vdash
+& \dadada : \CaseSig{\top}{\bot}{\texttt{id}\to\El{\textit{self.}\texttt{tm}}\to\El{\textit{self.}\texttt{tm}}}
+\\
+\lsti!End subst.!
+& & &
+& \mathit{self} : [\texttt{tm}:\TyS{\wcode{\wsig_{\texttt{tm}}}}, \texttt{tm\_unit}: \El{\texttt{tm}}, \dadada]
+\\
+\lsti!!
+\end{tabular}
+\caption{Typing of each field}
+\label{fig:stlc-linkage-typing}
+\end{figure}
 

### 2022-10-31 stlc-mechanized.tex
stlc compilation figure
simplify
stuff

@@
 
-FInductive ty : Set :=[@\hfill\codecomment{Types}@]
-| ty_unit  : ty
-| ty_arrow : ty -> ty -> ty.
-
 FInductive tm : Set :=[@\hfill\codecomment{Terms}@]
@@
 | tm_app  : tm -> tm -> tm.[@\label{line:tm-app}@]
-
-FDefinition env : Type := id -> option ty.
-FDefinition empty : env := \lambda _, None.
-
-FInductive hasty : env -> tm -> ty -> Prop :=
-| ht_var  : [@\dadada@][@\hfill\codecomment{Typing rules}@]
-| ht_unit : \forall G, hasty G tm_unit ty_unit
-| ht_abs  : [@\dadada@]
-| ht_app  : [@\dadada@].
-
-FInductive value : tm -> Prop :=[@\hfill\codecomment{Value forms}@]
-| v_unit : value tm_unit
-| v_abs  : \forall x t, value (tm_abs x t).
 
@@
 End subst.
+
+FInductive ty : Set :=[@\hfill\codecomment{Types}@]
+| ty_unit  : ty
+| ty_arrow : ty -> ty -> ty.
+
+FDefinition env : Type := id -> option ty.
+FDefinition empty : env := \lambda _, None.
+
+FInductive hasty : env -> tm -> ty -> Prop :=
+| ht_unit : \forall G, hasty G tm_unit ty_unit
+| ht_var  : [@\dadada@][@\hfill\codecomment{Typing rules}@]
+| ht_abs  : [@\dadada@]
+| ht_app  : [@\dadada@].
+
+FInductive value : tm -> Prop :=[@\hfill\codecomment{Value forms}@]
+| v_unit : value tm_unit
+| v_abs  : \forall x t, value (tm_abs x t).
 
@@
 
+FRecursion subst on tm motive \lambda _, id -> tm -> tm.
+Case tm_fix := \lambda y t' IHt' x t, [@\dadada@].
+End subst.
+
 FInductive hasty : env -> tm -> ty -> Prop +=
@@
   hasty G (tm_fix x t) T.
-
-FRecursion subst on tm motive \lambda _, id -> tm -> tm.
-Case tm_fix := \lambda y t' IHt' x t, [@\dadada@].
-End subst.
 
@@
 
-\colorlet{codecomment-color}{black}
+\definecolor{codecomment-color}{HTML}{000080}
 
@@
 
-\makeline[.5\textwidth+9pt]{Check-STLCFix-typesafe}{Check-STLCFix-typesafe}[codecomment-color!35]
+\makeline[.5\textwidth+9pt]{Check-STLCFix-typesafe}{Check-STLCFix-typesafe}[codecomment-color!50]
 

### 2022-10-31 metatheory-condensed.tex
linkage transformers

@@
 $\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new term of type $A$
-that hides W-type signatures, if necessary.
-When no hiding is needed, $\seal$ is the identity $\var{0}$ and $A$ is
+that hides W-type signatures behind $\cU$, if necessary.
+When no hiding is needed, $\seal$ is $\var{0}$ (the identity) and $A$ is
 $\TyTkg{\lsig}$.
 
-\bigskip
+When a family is closed and a complete linkage~$\lkg$ is formed, fields
+of the family can be accessed by projecting them out of the tuple
+$\Tkg{\lkg}$, as the last line of \cref{fig:stlc-linkage-typing} suggests.
 
-A supplemental appendix gives a more complete example showing how families
-are modeled and type-checked as linkages.
+\noindentparagraph{Linkage transformers.}
+
+Inheritance and code reuse can already be expressed through projection
+of fields out of linkages and their inclusion into new linkages.
+To make common patterns of linkage manipulations more convenient,
+\TT builds \emph{linkage transformers}, whose well-formedness judgments
+have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$.
+Linkage transformers are second-class; the only way to use them is to
+apply them to linkages to form new linkages (\ruleref{tm/inh}).
+%
+Derived families can be modeled as linkage transformers.
+%
+Linkage transformers are inductively constructed from several introduction
+rules.
+\cref{fig:typing-selected} shows the beta rule of an exemplary transformer,
+\ruleref{lxeq/ov/beta}.
+It can read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
+applying the transformer $\InhOv{h}{t_2}$ overrides the linkage's last
+field~$t_1$ with~$t_2$.
+$\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$, for instance, can be used
+in \cref{fig:stlc-linkage-typing} to form a derived family that
+overrides $\wsig_{\texttt{tm}}$ with an extended
+signature~$\wsig'_{\texttt{tm}}$.
+
+\noindentparagraph{The complete formalization.}
+
+The definitive version is formalized in a logical framework with
+quotient inductive-inductive types~\cite{nordvall2010inductive,altkap2016},
+available as a supplemental file in Agda syntax.
+%
+We also typeset a more complete set of rules than
+\cref{fig:mltt-typing-selected,fig:fmltt-typing-selected} allow
+and present it in a supplemental appendix.
+%
+Another supplemental appendix gives a larger-scale example than
+\cref{fig:stlc-linkage-typing} showing how families are modeled and
+type-checked using linkages and linkage transformers.
 

### 2022-10-31 standardmodel.tex
Highlight model.

@@
+\newcommand{\goodCtxS}[2]{{ {#1} \ \models_S }}
+\newcommand{\goodTypeS}[3]{{ {#1} \models_S {#2} }}
+\newcommand{\goodTermS}[3]{{ {#1} \models_S {#2} : {#3} }}
+\newcommand{\goodSubS}[3]{{ {#1} \models_S {#2} : {#3} }}
+\newcommand{\goodSigS}[3]{{ {#1} \models_S {#2} \ \  \texttt{LSig}^{#3} }}
+\newcommand{\goodWSigS}[3]{{ {#1} \models_S {#2} \ \ \texttt{WSig}^{#3} }}
+\newcommand{\goodSealS}[4]{{ {#1} \models_S {#2} : {#3} \  |\  {#4} }}
+\newcommand{\goodInhS}[4]{{ {#1} \models_S {#2} : {#3} \twoheadrightarrow {#4}}}
+
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.}
+We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of QIIT.
+However, in this section, for readability, we only inductively interpret each syntax piece into naive set theory that is also respecting judgemental equalities.
+Please refer to our pseduo-Agda formulation for a more formal and complete version.
 
-The main idea of the proof is that: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma.
-We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo) Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e.
-the judgementally-equal syntax needs to be mapped to the same data for well-definedness).
+The big picture of the proof is that: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma.
+The main goal is to interpret $\denotesS{\bot} = \emptyset$ which implies it is impossible $\goodTermS{\cdot}{t}{\bot}$ and thus $\bot$ is syntactically non-derivable.
+We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$ and that $\denotesS{}$ respecting judgemental equality.\footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo) Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}
 
-The main intuition of the constructed consistency model is that: (1) since we have used sigma type to translate the linkage, using the model for sigma type for linkage will work; (2) directly using the inductive facility of (pseudo)-Agda to justify the intro and elim rule of our W-type.
+The main intuition of the consistency model is that: (1) since we have used sigma type to translate the linkage, using the model for sigma type for linkage will work; (2) directly using the inductive facility of (pseudo)-Agda (or naive set theory) to justify the intro and elim rule of our W-type.
 
@@
 
-\newcommand{\goodCtxS}[2]{{ {#1} \ \models_S }}
-\newcommand{\goodTypeS}[3]{{ {#1} \models_S {#2} }}
-\newcommand{\goodTermS}[3]{{ {#1} \models_S {#2} : {#3} }}
-\newcommand{\goodSubS}[3]{{ {#1} \models_S {#2} : {#3} }}
-\newcommand{\goodSigS}[3]{{ {#1} \models_S {#2} \ \  Sig^{#3} }}
-\newcommand{\goodWSigS}[3]{{ {#1} \models_S {#2} \ \ WSig^{#3} }}
-\newcommand{\goodSealS}[4]{{ {#1} \models_S {#2} : {#3} \  |\  {#4} }}
-\newcommand{\goodInhS}[4]{{ {#1} \models_S {#2} : {#3} \twoheadrightarrow {#4}}}
+\definecolor{new}{HTML}{F5EFE6}
 
 \begin{align*}
-  \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set and we define } \goodCtxS{\Gamma}{i} :\iff \denotesS{\goodCtx{\Gamma}{i}} \text{ is defined} \\
-  \denotesS{\goodType{\Gamma}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gamma}{i}} \\
-  & \text{ and we define } \goodTypeS{\Gamma}{T}{i} : \iff \denotesS{\goodType{\Gamma}{T}{i}} \text{ is defined } \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod1}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set and we define } \goodCtxS{\Gamma}{i} :\iff \denotesS{\goodCtx{\Gamma}{i}} \text{ is defined}\tikzmarkend{stmod1} 
+  \\
   \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
@@
   \\
-  \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gamma}{\gamma}{\Delta}} : \iff \denotesS{\gamma} \in \denotesS{\goodSub{\Gamma}{\_}{\Delta}} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gamma}{\gamma}{\Delta}} : \iff \denotesS{\gamma} \in \denotesS{\goodSub{\Gamma}{\_}{\Delta}} \tikzmarkend{stmod2}
+  \\
   \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gamma}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
-  \denotesS{\goodTerm{\Gamma}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gamma}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gamma}{\_}{T}} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gamma}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gamma}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gamma}{\_}{T}}
+  \tikzmarkend{stmod3}
+  \\
   \denotesS{\goodTerm{\Gamma}{\_}{T}} & = \prod_{\gamma \in \denotesS{\goodCtx{\Gamma}{i}}}\denotesS{T}(\gamma) \\
-  \denotesS{\goodType{\Gamma}{\Sigma A B}{}} &\text{ is a family of sets, defined dependent on } \denotesS{A}, \denotesS{B} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod4}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodType{\Gamma}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gamma}{i}}  
+  \\
+  & \text{ and we define } \goodTypeS{\Gamma}{T}{i} : \iff \denotesS{\goodType{\Gamma}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 
+  \\
   \denotesS{\goodType{\Gamma}{\Sigma A B}{}}(\gamma) &:= \sum_{a \in \denotesS{A}\gamma} \denotesS{B}(\gamma, a) \\
-  \denotesS{\goodSig{\Gamma}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gamma}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gamma}{\_}{n}} \\
-  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} & \text{ is a family of sets, dependent on and defined inductively on } \denotesS{\sigma} \\   
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod5}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodSig{\Gamma}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gamma}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gamma}{\_}{n}} \tikzmarkend{stmod5} 
+  \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod6}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
   \denotesS{\goodSig{\Gamma}{\_}{n+1}} &= \{
@@
   \denotesS{\goodType{\Gamma}{\cC \sigma}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} & \text{ is a family of sets, dependent on } \denotesS{\sigma}  \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod7}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
   \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\gamma) \land t \in \prod_{m' \in \denotesS{A}(\gamma)}\denotesS{T}(\gamma, m')  \}  \\
@@
   \denotesS{\goodTerm{\Gamma}{p_1\mu~t}{\cL(p_1\nu~\sigma)}}(\gamma) &=  "fst"~(t\gamma) \\ 
-  \denotesS{\goodTerm{\Gamma}{\cCt \{\sigma\} o}{\cC\sigma}} & \text{ is a dependent function, defined inductively on } \denotesS{\sigma} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodTerm{\Gamma}{\cCt \{\sigma\} o}{\cC\sigma}} & \text{ is a dependent function, defined inductively on } {\sigma} 
+  \tikzmarkend{stmod8}
+  \\
   \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
-  &= \denotesS{\goodTerm{\Gamma}{\cCt (p_1\mu \ o)}{...}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \cCt (p_1\mu \ o))]} \\
+  &= \denotesS{{\cCt (p_1\mu \ o)}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \cCt (p_1\mu \ o))]} \\
   & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
@@
   \denotesS{\goodType{\Gamma}{\cB}{i}} \ (\gamma) &= \{0, 1\} \quad \text{a specific two element set} \\
-  \denotesS{\goodWSig{\Gamma}{\_}{n}} &\text{ is a Set of list of pairs of type interpretation of length } n \\
-  & \text{ and we define } \goodWSigS{\Gamma}{\tau}{n} :\iff \denotesS{\goodWSig{\Gamma}{\tau}{n}} \in \denotesS{\goodWSig{\Gamma}{\_}{n}}\\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod9}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodWSig{\Gamma}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
+  & \text{ and we define } \goodWSigS{\Gamma}{\tau}{n} :\iff \denotesS{\goodWSig{\Gamma}{\tau}{n}} \in \denotesS{\goodWSig{\Gamma}{\_}{n}}
+  \tikzmarkend{stmod9}
+  \\
   \denotesS{\goodWSig{\Gamma}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 

### 2022-10-31 canonicitymodel.tex
Fixing Consistency and Canonicity.

@@
 \begin{theorem}[Canonicity]
-  For a closed boolean term $\goodTerm{\cdot}{t}{\mathbb{B}}$, we have $\goodTerm{\cdot}{t \equiv tt}{\mathbb{B}}$ or $\goodTerm{\cdot}{t \equiv ff}{\mathbb{B}}$ hold.
+  For a closed boolean term $\goodTerm{\cdot}{t}{\cB}$, we have $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$ hold.
 \end{theorem}
@@
 Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
-We can even argue that, if this theorem is proven in a constructive meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}}, 
+We can even argue that, if this theorem is proven in a computable meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}}, 
 then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.\YZ{Can we say something about strong/weak normalization?}\EDJreply{
@@
 The canonicity model $"Tm"^C~\Gamma^C~"T"^C$ will map each syntax piece $("t : Tm Γ T") \mapsto "(t, tₚ) :" "Tm"^C~\Gamma^C~"T"^C = \sum~("t : Tm Γ T")."Tm"_p~\Gamma^C~"T"^C~"t"$ to a dependent pair\footnote{This $"Tm"_p$ is usually called dependent model.}, of which the first part will be the same as the input syntax.
-Using this mapping on closed boolean term $"t : Tm ⋅ "\mathbb{B}$ will give us $"Tm"_p~\cdot^C~\mathbb{B}^C~"t"$, which will unfolds to our final goal $"t" \equiv "tt" + "t" \equiv "ff"$.
+Using this mapping on closed boolean term $"t : Tm ⋅ "\mathbb{B}$ will give us a proof of $"Tm"_p~\cdot^C~\mathbb{B}^C~"t"$, which will unfolds to our final goal $"t" \equiv "tt" + "t" \equiv "ff"$.
 Roughly, this can read as ``every closed boolean term is reducible''.
@@
 
-Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
-The latter is more conventional and the former can be understood as type theoretic encoding of the latter\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory}.
-From this perspective, this part of the canonicity model is quite similar to how conventional (proof-irrelevant) logical relation is carried out~\cite{skorstengaard2019introduction}---using the terminology of the conventional formulation, $\Gamma_p$ will witness the set of substitution into closed reducible terms of types $\Gamma$.
-Then we will prove $t[\gamma] \in T_p$---a set of closed reducible term of type $T$---when $\gamma \in \Gamma_p$.
+Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as ``a reducible set'': $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
+The reformulation can be understood as an unfolding of the \textit{type-theoretic logical relation proof}.\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory} If we use the terminology of the conventional formulation, this reducible set means if we apply its element $t$ with a substitution $\gamma$ from the set $\Gamma$ where we have $\Gamma_p$ witnesses the ``reducibility'' of $\Gamma$, then we will have $t[\gamma] \in T_p$ as a closed and \textbf{reducible} term.
+Now if $\Gamma$ is actually empty $\cdot$ and $T$ is $\cB$, we are done.
+
+% $\Gamma_p$ witnesses the set of substitution $ \gamma \in \Gamma$ makes applied term $t[\gamma]$ closed and \textbf{reducible} of types $T_p$.
+Thus, we proved $t[\gamma] \in T_p$---the set of closed reducible term of type $T$---when $\gamma \in \Gamma_p$.
 
 The proof can be found in appendix.
-The intuition of the model is similar to that of the consistency model---use the model of sigma type for linkage type and using Agda's inductive facility to justify our W-type.
+The intuition of the model is similar to that of the consistency model---use the canonicity model of sigma type (from the earlier work) for linkage type and using Agda's inductive facility to justify our W-type.
 

### 2022-10-31 lang-design.tex
reminder to myself
Some Comments.

@@
 Such circularity would easily lead to logical inconsistency:
+\YZ{Mention non-exhaustivity/overriding already poses challenges to ensuring consistency.}
 
@@
 Note that the requirement still allows a derived family to introduce
-new declarations into the context of a field.
+new declarations into the context of a field. \EDJ{Maybe a better wording is ``the relative order between fields are preserved during inheritance''?}
 For example, the left column of \cref{fig:stlc-isorec-prod} shows an
@@
 
-\noindentparagraph{Composing families.}
+\noindentparagraph{Composing families.} \EDJ{I feel this paragraph will cause more confusion---because it is inheritance judgement/linkage transformer being mixed in, not family itself.
+Conceptually not distinguishing family and linkage transfomer will make reader wonder what this magic black box here actually is}
 

### 2022-10-31 standardmodel.tex

@@
 
-We again encourage the reader to read the (pseudo-)Agda-style proof mentioned in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation.
-Here, we need to clarify some ambiguity in our conventional formulation.
-For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on $\denotesS{\Sigma}$'', the best way to understand is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ also defined inductively on the length of the signature.
+We encourage the reader to read the (pseudo-)Agda-style proof with more type annotation so it is clearer about the mathematical object doing the interpretation.
+Here, we need to clarify some ambiguity in our latex formulation.
+For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on ${\sigma}$'', the best way to understand it is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$, inductively on the length of the signature.
 The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
@@
 
-Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
+Note that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
 

### 2022-10-31 metatheory-condensed.tex
figure

@@
 inductive types in MLTT;
-together with the identity type $\TyId{\cdot}{\cdot}$, they can express a
+together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
 whole host of inductive types~\cite{hugunin2021not}.
@@
 In contrast, $\wsig$ should be hidden from the typing context of any
-term that does not invoke $\Wsup{\wsig}{\cdot}{\cdot}$ or $\Wrec{\wsig}{\cdot}{\cdot}$, so that
+term that does not invoke $\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
 the term can be reused---without being rechecked---for a different
@@
 Succeeding fields that invoke
-$\Wsup{\wsig}{\cdot}{\cdot}$ or $\Wrec{\wsig}{\cdot}{\cdot}$
+$\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
 should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
@@
 %
-Here, we feel free to use a notation with explicit names.
-We name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
+For readability, we feel free to use a notation with explicit names
+in this typing judgment:
+we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
 and feel free to write its type, which is a dependent tuple type, as a
-dependent record type with explicit labels.
+dependent record type with labels.
 %
@@
 $t_{10}$, which models the recursive function \lsti{subst} by invoking
-the recursor $\Wrec{\wsig_{\texttt{tm}}}{\cdot}{\cdot}$.
+the recursor $\Wrec{\wsig_{\texttt{tm}}}{\cDot}{\cDot}$.
 %
@@
 \ruleref{lxeq/ov/beta}.
-It can read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
+It can be read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
 applying the transformer $\InhOv{h}{t_2}$ overrides the linkage's last
@@
 
-\begin{figure}
-\small
-\newcolumntype{L}{>{$}l<{$}}
-\newcolumntype{C}{>{$}c<{$}}
-\newcolumntype{R}{>{$}r<{$}}
-\begin{tabular}{@{}lR@{\ }C@{\ }LR@{\ \,}C@{\,\ }L@{}}
-\lsti!Family STLC.!
-& \lkg_0 & \coloneqq & \LkgEmp
-\\
-\lsti!FInductive tm :=!
-& \lkg_1 & \coloneqq & \LkgAdd{\lkg_0}{\dadada}
-& \mathit{self} : []
-& \vdash
-& \wcode{\wsig_{\texttt{tm}}} : \TyS{\wcode{\wsig_{\texttt{tm}}}}
-\\
-\lsti!| tm_unit : tm | !\dadada
-& \lkg_2 & \coloneqq & \LkgAdd{\lkg_1}{\dadada}
-& \mathit{self} : [\texttt{tm}:\TyS{\wcode{\wsig_{\texttt{tm}}}}]
-& \vdash
-& \Wsup{\wsig_{\texttt{tm}}}{\top}{\bot} : \El{\mathit{self.}\texttt{tm}}
-\\
-\lsti!FRecursion subst !\dadada
-\\
-\lsti!Case tm_unit := !\dadada
-& & &
-& \mathit{self} : [\texttt{tm}:\cU, \texttt{tm\_unit}: \El{\texttt{tm}}, \dadada]
-& \vdash
-& \dadada : \CaseSig{\top}{\bot}{\texttt{id}\to\El{\textit{self.}\texttt{tm}}\to\El{\textit{self.}\texttt{tm}}}
-\\
-\lsti!End subst.!
-& & &
-& \mathit{self} : [\texttt{tm}:\TyS{\wcode{\wsig_{\texttt{tm}}}}, \texttt{tm\_unit}: \El{\texttt{tm}}, \dadada]
-\\
-\lsti!!
-\end{tabular}
-\caption{Typing of each field}
-\label{fig:stlc-linkage-typing}
-\end{figure}
+\input{stlc-linkage-typing.tex}
 

### 2022-10-31 stlc-linkage-typing.tex
figure

\begin{figure}
\small
\renewcommand*{\arraystretch}{1.15}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{C}{>{$}c<{$}}
\newcolumntype{R}{>{$}r<{$}}
\begin{tabular}{@{}lL@{\ }@{}}
\lsti!Family STLC.!
&
\lsig_0 \coloneqq \LSigEmp
\quad
\lkg_0 \coloneqq \LkgEmp
\quad
\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
\\[4pt]
\multirow{2}{*}{\lsti!FInductive tm :=!}
&
\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{T_1}
\quad
\lkg_1 \coloneqq \LkgAdd{\lkg_0}{t_1}
\quad
\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
\quad
\goodTerm{\TyTkg{\lsig_0}}{\seal_1}{\sub{A_1}{\SubstWeak{1}}}
\\
&
\goodTerm{\mathit{self} : \big[\big]}{\wcode{\wsig_{\texttt{tm}}}}{\TyS{\wcode{\wsig_{\texttt{tm}}}}}
\\[4pt]
\multirow{2}{*}{\lsti!| tm_unit : tm |!\ \dadada}
&
\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{T_2}
\quad
\lkg_2 \coloneqq \LkgAdd{\lkg_1}{t_2}
\quad
\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
\quad
\goodTerm{\TyTkg{\lsig_1}}{\seal_2}{\sub{A_2}{\SubstWeak{1}}}
\\
&
\goodTerm{
    \mathit{self} : \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
}{
    \Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}
}{
    \El{\mathit{self.}\texttt{tm}}
}
\\[4pt]
\lsti!FRecursion subst!\ \dadada
\\
\multirow{2}{*}{\lsti!Case tm_unit :=!\ \dadada}
&
\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{T_6}
\quad
\lkg_6 \coloneqq \LkgAdd{\lkg_5}{t_6}
\quad
\goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
\quad
\goodTerm{\TyTkg{\lsig_5}}{\seal_6}{\sub{A_6}{\SubstWeak{1}}}
\\
&
\goodTerm{
    \mathit{self} : \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
}{
%   \lambda\_.\,\lambda\_.\,\lambda x.\,\lambda t.\,\mathit{self}.\texttt{tm\_unit}
    \dadada
}{
    \CaseSig{\top}{\bot}{\El{\mathit{self}.\texttt{tm}}\to\mathit{id}\to\El{\mathit{self}.\texttt{tm}}}
}
\\[4pt]
\lsti!Case!\ \dadada
\\
\multirow{2}{*}{\lsti!End subst.!}
&
\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{T_{10}}
\quad
\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{t_{10}}
\quad
\goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
\quad
\goodTerm{\TyTkg{\lsig_9}}{\seal_{10}}{\sub{A_{10}}{\SubstWeak{1}}}
\\
&
\goodTerm{
    \mathit{self} : \left[{%
    \begin{minipage}{.35\textwidth}%
        \vspace{-3.5pt}
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
        \dadada
        \\
        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
        \dadada
    \end{array}
    \end{gather*}
    \end{minipage}
    }\right]
}{
    \lambda t.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{t}
}{
    \begin{minipage}{.20\textwidth}%
        \vspace{-3.5pt}\scriptsize
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \El{\mathit{self}.\texttt{tm}} \to \El{\mathit{self}.\texttt{tm}}
        \\
        \to \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
    \end{array}
    \end{gather*}
    \end{minipage}
}
\\
\dadada
\\[4pt]
\lsti!End STLC.!
&
\goodTerm{\cdot}{\Tkg{\lkg_{20}}}{\TyTkg{\lsig_{20}}}
\end{tabular}

\caption{Typing of each field}
\label{fig:stlc-linkage-typing}
\end{figure}

### 2022-11-01 stlc-linkage-typing.tex
coloring

@@
 \newcolumntype{R}{>{$}r<{$}}
-\begin{tabular}{@{}lL@{\ }@{}}
+%
+\definecolor{A-color}{HTML}{E7EFC5}
+\definecolor{T-color}{HTML}{FFF1D0}
+\definecolor{t-color}{HTML}{FAE0E4}
+\colorlet{Bkg-color}{black!8}
+%
+\begin{tabular}{@{}l@{\ \ \ }L@{}}
+\cellcolor{Bkg-color}
 \lsti!Family STLC.!
@@
 \goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
-\\[4pt]
-\multirow{2}{*}{\lsti!FInductive tm :=!}
+\\
+\\[-8pt]
+\cellcolor{Bkg-color}
 &
-\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{T_1}
+\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-a}(-0.02,-0.10)(-0.02,0.28)
+    T_1
+    \tikzmarkend{T1-a}
+}
 \quad
-\lkg_1 \coloneqq \LkgAdd{\lkg_0}{t_1}
+\lkg_1 \coloneqq \LkgAdd{\lkg_0}{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-a}(-0.02,-0.10)(-0.02,0.28)
+    t_1
+    \tikzmarkend{t1-a}
+}
 \quad
@@
 \quad
-\goodTerm{\TyTkg{\lsig_0}}{\seal_1}{\sub{A_1}{\SubstWeak{1}}}
+\goodTerm{\TyTkg{\lsig_0}}{\seal_1}{\sub{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-a}(-0.02,-0.10)(-0.05,0.28)
+    A_1
+    \tikzmarkend{A1-a}
+}{\SubstWeak{1}}}
 \\
+\cellcolor{Bkg-color}
+\multirow{-2}{*}{%
+\lsti!FInductive!\ \ 
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.01,-0.10)(-0.02,0.28)%
+    \lsti!tm!%
+    \tikzmarkend{t1-c}%
+\ \lsti!:=!}
 &
-\goodTerm{\mathit{self} : \big[\big]}{\wcode{\wsig_{\texttt{tm}}}}{\TyS{\wcode{\wsig_{\texttt{tm}}}}}
-\\[4pt]
-\multirow{2}{*}{\lsti!| tm_unit : tm |!\ \dadada}
+\goodTerm{
+    \mathit{self} : 
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-b}(0.00,-0.15)(0.00,0.30)%
+    \big[\big]
+    \tikzmarkend{A1-b}%
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-b}(-0.02,-0.10)(-0.02,0.28)
+    \wcode{\wsig_{\texttt{tm}}}
+    \tikzmarkend{t1-b}
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-b}(-0.02,-0.10)(-0.02,0.28)
+    \TyS{\wcode{\wsig_{\texttt{tm}}}}
+    \tikzmarkend{T1-b}
+}
+\\
+\\[-8pt]
+\cellcolor{Bkg-color}
 &
-\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{T_2}
+\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-a}(-0.02,-0.10)(-0.02,0.28)
+    T_2
+    \tikzmarkend{T2-a}
+}
 \quad
-\lkg_2 \coloneqq \LkgAdd{\lkg_1}{t_2}
+\lkg_2 \coloneqq \LkgAdd{\lkg_1}{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-a}(-0.02,-0.10)(-0.02,0.28)
+    t_2
+    \tikzmarkend{t2-a}
+}
 \quad
@@
 \quad
-\goodTerm{\TyTkg{\lsig_1}}{\seal_2}{\sub{A_2}{\SubstWeak{1}}}
+\goodTerm{\TyTkg{\lsig_1}}{\seal_2}{\sub{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-a}(-0.02,-0.10)(-0.05,0.28)
+    A_2
+    \tikzmarkend{A2-a}
+}{\SubstWeak{1}}}
 \\
+\cellcolor{Bkg-color}
+\multirow{-2}{*}{%
+    \lsti!|!\ %
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-c}(0.01,-0.10)(-0.02,0.28)%
+    \lsti!tm_unit!%
+    \tikzmarkend{t2-c}%
+}
 &
 \goodTerm{
-    \mathit{self} : \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
+    \mathit{self} :
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-b}(0.00,-0.15)(0.00,0.30)%
+    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
+    \tikzmarkend{A2-b}%
 }{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-b}(-0.02,-0.10)(-0.02,0.28)
     \Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}
+    \tikzmarkend{t2-b}
 }{
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-b}(-0.02,-0.10)(-0.02,0.28)
     \El{\mathit{self.}\texttt{tm}}
+    \tikzmarkend{T2-b}
 }
+\\
+\\[-8pt]
+\lsti!| !\dadada\ \ \lsti!| !\dadada\ \ \lsti!| !\dadada\lsti!.!
+&
+\dadada
 \\[4pt]
-\lsti!FRecursion subst!\ \dadada
+\lsti!FRecursion subst!\ \dadada\lsti!.!
+&
+\textit{\color{codecomment-color}%
+    $\wcode{\wsig_{\texttt{tm}}}$ is abstracted into $\cU$ in the typing
+    context of the case handler below:%
+}
 \\
-\multirow{2}{*}{\lsti!Case tm_unit :=!\ \dadada}
+\cellcolor{Bkg-color}
 &
-\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{T_6}
+\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-a}(-0.02,-0.10)(-0.02,0.28)
+    T_6
+    \tikzmarkend{T6-a}
+}
 \quad
-\lkg_6 \coloneqq \LkgAdd{\lkg_5}{t_6}
+\lkg_6 \coloneqq \LkgAdd{\lkg_5}{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-a}(-0.02,-0.10)(-0.02,0.28)
+    t_6
+    \tikzmarkend{t6-a}
+}
 \quad
@@
 \quad
-\goodTerm{\TyTkg{\lsig_5}}{\seal_6}{\sub{A_6}{\SubstWeak{1}}}
+\goodTerm{\TyTkg{\lsig_5}}{\seal_6}{\sub{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-a}(-0.02,-0.10)(-0.05,0.28)
+    A_6
+    \tikzmarkend{A6-a}
+}{\SubstWeak{1}}}
 \\
+\cellcolor{Bkg-color}
+\multirow{-2}{*}{%
+    \lsti!Case tm_unit :=!\ %
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-c}(0.01,-0.10)(-0.02,0.28)%
+    \dadada%
+    \tikzmarkend{t6-c}%
+    \lsti!.!%
+}
 &
 \goodTerm{
-    \mathit{self} : \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+    \mathit{self} :
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-b}(0.00,-0.15)(0.00,0.30)%
+    \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+    \tikzmarkend{A6-b}%
 }{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-b}(-0.02,-0.10)(-0.02,0.28)
 %   \lambda\_.\,\lambda\_.\,\lambda x.\,\lambda t.\,\mathit{self}.\texttt{tm\_unit}
     \dadada
+    \tikzmarkend{t6-b}
 }{
-    \CaseSig{\top}{\bot}{\El{\mathit{self}.\texttt{tm}}\to\mathit{id}\to\El{\mathit{self}.\texttt{tm}}}
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-b}(-0.02,-0.10)(-0.02,0.28)
+    \CaseSig{\top}{\bot}{
+        \fontsize{8.5}{10.2}\selectfont
+        \text{$\El{\mathit{self}.\texttt{tm}}\to\mathit{id}\to\El{\mathit{self}.\texttt{tm}}$}
     }
-\\[4pt]
-\lsti!Case!\ \dadada
+    \tikzmarkend{T6-b}
+}
 \\
-\multirow{2}{*}{\lsti!End subst.!}
+\\[-8pt]
+\lsti!Case!\ \dadada\ \ 
+\lsti!Case!\ \dadada\ \ 
+\lsti!Case!\ \dadada%
 &
-\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{T_{10}}
+\dadada
+\\
+\cellcolor{Bkg-color}
+&
+\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-a}(-0.02,-0.10)(-0.02,0.28)
+    T_{10}
+    \tikzmarkend{T10-a}
+}
 \quad
-\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{t_{10}}
+\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-a}(-0.02,-0.10)(-0.02,0.28)
+    t_{10}
+    \tikzmarkend{t10-a}
+}
 \quad
@@
 \quad
-\goodTerm{\TyTkg{\lsig_9}}{\seal_{10}}{\sub{A_{10}}{\SubstWeak{1}}}
+\goodTerm{\TyTkg{\lsig_9}}{\seal_{10}}{\sub{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-a}(-0.02,-0.10)(-0.05,0.28)%
+    A_{10}
+    \tikzmarkend{A10-a}
+}{\SubstWeak{1}}}
 \\
+\cellcolor{Bkg-color}
+\multirow{-2}{*}{%
+    \lsti!End!\ \ %
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-c}(0.01,-0.10)(-0.02,0.28)%
+    \lsti!subst!%
+    \tikzmarkend{t10-c}%
+    \lsti!.!%
+}
 &
 \goodTerm{
-    \mathit{self} : \left[{%
-    \begin{minipage}{.35\textwidth}%
-        \vspace{-3.5pt}
+    \mathit{self} :
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-b}(0.00,-0.32)(0.00,0.47)%
+    \!\left[{%
+    \begin{minipage}{.31\textwidth}%
+    \vspace{-7.5pt}%
+    \fontsize{8.0}{9.2}\selectfont%
     \begin{gather*}
@@
         \texttt{tm\_unit} : \El{\texttt{tm}},\ 
-        \dadada
-        \\
+        \dadada,
+        \smallskip\\
         \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
@@
     \end{minipage}
-    }\right]
+    }\right]\!
+    \tikzmarkend{A10-b}%
 }{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-b}(-0.02,-0.10)(-0.02,0.28)
     \lambda t.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{t}
+    \tikzmarkend{t10-b}
 }{
     \begin{minipage}{.20\textwidth}%
-        \vspace{-3.5pt}\scriptsize
+    \vspace{-7.5pt}%
+    \fontsize{8.0}{9.2}\selectfont%
     \begin{gather*}
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.26)(-0.02,0.40)
     \begin{array}{@{}l@{}}
@@
     \end{array}
+    \tikzmarkend{T10-b}
     \end{gather*}
@@
 \dadada
-\\[4pt]
-\lsti!End STLC.!
 &
-\goodTerm{\cdot}{\Tkg{\lkg_{20}}}{\TyTkg{\lsig_{20}}}
+\dadada
+\\
+\cellcolor{Bkg-color}
+\lsti!End !%
+\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-c}(-0.02,-0.10)(-0.02,0.28)%
+\lsti!STLC!%
+\tikzmarkend{t-STLC-c}%
+\lsti!.!
+&
+\goodTerm{\cdot}{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-b}(-0.02,-0.10)(-0.02,0.28)%
+    \Tkg{\lkg_{20}}
+    \tikzmarkend{t-STLC-b}
+}{\TyTkg{\lsig_{20}}}
 \end{tabular}
 
-\caption{Typing of each field}
+\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.}
 \label{fig:stlc-linkage-typing}

### 2022-11-01 metatheory-condensed.tex
singleton type

@@
 we omit universe levels in the presentation for conciseness.
+
+A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
+in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
 
@@
 %
-\begin{itemize}[labelsep=*,leftmargin=1pc]
+\begin{itemize}
+  [leftmargin=4ex]
+  %[labelsep=*,leftmargin=1pc]
 
@@
 where $\wsig'_{\texttt{tm}}$ extends~$\wsig_{\texttt{tm}}$ with additional constructors.
+%
+For example, in \cref{fig:stlc-linkage-typing}, the typing of the case handlers of
+\lsti{subst} is agnostic of the definition of \lsti!tm!---it sees only $\text{\lsti!tm!}:\cU$---so
+they can be reused by a linkage modeling \lsti{STLCFix}.
 

### 2022-11-01 standardmodel.tex
Save. Changing Notation.

@@
 
-Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gamma}{t}{T}}$ (i.e.
+Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gm}{t}{T}}$ (i.e.
 every term is tracking its context and type).\footnote{And our (pseudo) Agda formulation is well-typed as well.
@@
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod1}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodCtx{\Gamma}{i}} & \text{ is a Set and we define } \goodCtxS{\Gamma}{i} :\iff \denotesS{\goodCtx{\Gamma}{i}} \text{ is defined}\tikzmarkend{stmod1} 
+  \denotesS{\goodCtx{\Gm}{i}} & \text{ is a Set and we define } \goodCtxS{\Gm}{i} :\iff \denotesS{\goodCtx{\Gm}{i}} \text{ is defined}\tikzmarkend{stmod1} 
   \\
   \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodCtx{\Gamma, A}{i}} &= \{(\gamma, a) : \gamma \in \denotesS{\goodCtx{\Gamma}{}}, a \in \denotesS{\goodType{\Gamma}{A}{}}(\gamma)\}
+  \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \Gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\Gm)\}
   \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gamma}{\gamma}{\Delta}} : \iff \denotesS{\gamma} \in \denotesS{\goodSub{\Gamma}{\_}{\Delta}} \tikzmarkend{stmod2}
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gm}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gm}{\Gm}{\Delta}} : \iff \denotesS{\Gm} \in \denotesS{\goodSub{\Gm}{\_}{\Delta}} \tikzmarkend{stmod2}
   \\
-  \denotesS{\goodSub{\Gamma}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gamma}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gamma}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gamma}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gamma}{\_}{T}}
+  \denotesS{\goodSub{\Gm}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gm}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gm}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gm}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gm}{\_}{T}}
   \tikzmarkend{stmod3}
   \\
-  \denotesS{\goodTerm{\Gamma}{\_}{T}} & = \prod_{\gamma \in \denotesS{\goodCtx{\Gamma}{i}}}\denotesS{T}(\gamma) \\
+  \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\Gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\Gm) \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod4}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gamma}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gamma}{i}}  
+  \denotesS{\goodType{\Gm}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gm}{i}}  
   \\
-  & \text{ and we define } \goodTypeS{\Gamma}{T}{i} : \iff \denotesS{\goodType{\Gamma}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 
+  & \text{ and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesS{\goodType{\Gm}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 
   \\
-  \denotesS{\goodType{\Gamma}{\Sigma A B}{}}(\gamma) &:= \sum_{a \in \denotesS{A}\gamma} \denotesS{B}(\gamma, a) \\
+  \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\Gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod5}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodSig{\Gamma}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gamma}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gamma}{\_}{n}} \tikzmarkend{stmod5} 
+  \denotesS{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gm}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gm}{\_}{n}} \tikzmarkend{stmod5} 
   \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod6}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
-  \denotesS{\goodSig{\Gamma}{\_}{n+1}} &= \{
+  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
+  \denotesS{\goodSig{\Gm}{\_}{n+1}} &= \{
     (\denotesS{\sigma}, \denotesS{A}, \denotesS{f}, \denotesS{T}) :
-      {\goodSigS{\Gamma}{\sigma}{n}}
-      \land  {\goodTypeS{\Gamma}{A}{}} \\  
-      & \quad \quad \quad \land  {\goodSealS{\Gamma}{f}{\sigma}{A}}
-      \land  {\goodTypeS{\Gamma, A}{T}{i}} 
+      {\goodSigS{\Gm}{\sigma}{n}}
+      \land  {\goodTypeS{\Gm}{A}{}} \\  
+      & \quad \quad \quad \land  {\goodSealS{\Gm}{f}{\sigma}{A}}
+      \land  {\goodTypeS{\Gm, A}{T}{i}} 
   \} \\ 
-  \denotesS{\goodSig{\Gamma}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gamma}{\cC \tau}{}} \denotesS{T["sf" \ f]} \\
-  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gamma}{\cC \sigma}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
+  \denotesS{\goodSig{\Gm}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
+  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gm}{\TyTkg {\tau}}{}} \denotesS{T["sf" \ f]} \\
+  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod7}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\gamma) \land t \in \prod_{m' \in \denotesS{A}(\gamma)}\denotesS{T}(\gamma, m')  \}  \\
-  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gamma}{\cL \sigma}{i}} (\gamma)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
-  \denotesS{\goodSig{\Gamma}{\nu\cdot}{0}} &= \star \\
-  \denotesS{\goodSig{\Gamma}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, A, f, T) \\
-  \denotesS{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot} } (\gamma) &= \star \\
-  \denotesS{\goodTerm{\Gamma}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\gamma)&= (o(\gamma), (\lambda \ m' . t'(\gamma, m'))) \\
-  \denotesS{\goodTerm{\Gamma}{p_1\mu~t}{\cL(p_1\nu~\sigma)}}(\gamma) &=  "fst"~(t\gamma) \\ 
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\Gm) \land t \in \prod_{m' \in \denotesS{A}(\Gm)}\denotesS{T}(\Gm, m')  \}  \\
+  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+  \denotesS{\goodSig{\Gm}{\nu\cdot}{0}} &= \star \\
+  \denotesS{\goodSig{\Gm}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, A, f, T) \\
+  \denotesS{\goodTerm{\Gm}{\mu\cdot}{\cL \nu\cdot} } (\Gm) &= \star \\
+  \denotesS{\goodTerm{\Gm}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
+  \denotesS{\goodTerm{\Gm}{p_1\mu~t}{\cL(p_1\nu~\sigma)}}(\Gm) &=  "fst"~(t\Gm) \\ 
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodTerm{\Gamma}{\cCt \{\sigma\} o}{\cC\sigma}} & \text{ is a dependent function, defined inductively on } {\sigma} 
+  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}} & \text{ is a dependent function, defined inductively on } {\sigma} 
   \tikzmarkend{stmod8}
   \\
-  \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
-  &= \denotesS{{\cCt (p_1\mu \ o)}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \cCt (p_1\mu \ o))]} \\
-  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{n+1}} \\
-  \denotesS{\goodTerm{\Gamma}{\cCt o}{\cC\sigma}}
-  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gamma}{\_}{0}} \\
-  \denotesS{\goodType{\Gamma}{\bot}{i}} \ (\gamma) &= \emptyset \\
-  \denotesS{\goodType{\Gamma}{\cB}{i}} \ (\gamma) &= \{0, 1\} \quad \text{a specific two element set} \\
+  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
+  &= \denotesS{{\Tkg {p_1\mu \ o}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {p_1\mu \ o})]} \\
+  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
+  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+  \denotesS{\goodType{\Gm}{\bot}{i}} \ (\Gm) &= \emptyset \\
+  \denotesS{\goodType{\Gm}{\cB}{i}} \ (\Gm) &= \{0, 1\} \quad \text{a specific two element set} \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod9}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodWSig{\Gamma}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
-  & \text{ and we define } \goodWSigS{\Gamma}{\tau}{n} :\iff \denotesS{\goodWSig{\Gamma}{\tau}{n}} \in \denotesS{\goodWSig{\Gamma}{\_}{n}}
+  \denotesS{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
+  & \text{ and we define } \goodWSigS{\Gm}{\tau}{n} :\iff \denotesS{\goodWSig{\Gm}{\tau}{n}} \in \denotesS{\goodWSig{\Gm}{\_}{n}}
   \tikzmarkend{stmod9}
   \\
-  \denotesS{\goodWSig{\Gamma}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 
-  \denotesS{\goodWSig{\Gamma}{(w^+ \ A \ B \ \tau)}{n+1}}
-  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gamma}{\tau}{n}} \\ 
-  \denotesS{\goodType{\Gamma}{\pi^j_1~\tau}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
-  \denotesS{\goodType{\Gamma, \pi^j_1~\tau}{\pi^j_2~\tau}{}} &= "snd"~({\denotesS{\tau}}_j)\\
+  \denotesS{\goodWSig{\Gm}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 
+  \denotesS{\goodWSig{\Gm}{(w^+ \ A \ B \ \tau)}{n+1}}
+  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gm}{\tau}{n}} \\ 
+  \denotesS{\goodType{\Gm}{\pi^j_1~\tau}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
+  \denotesS{\goodType{\Gm, \pi^j_1~\tau}{\pi^j_2~\tau}{}} &= "snd"~({\denotesS{\tau}}_j)\\
 \end{align*}
@@
 Here, we need to clarify some ambiguity in our latex formulation.
-For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on ${\sigma}$'', the best way to understand it is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$, inductively on the length of the signature.
-The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
-\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement also interpreted.
+For example, when we wrote ``$\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$ defined inductively on ${\sigma}$'', the best way to understand it is to consider we are defining a function $\denotesS{\goodType{\Gm}{\TyTkg {\_}}{}}$ as the interpretation for $\goodType{\Gm}{\TyTkg {\_}}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$, inductively on the length of the signature.
+The case is similar for $\denotesS{\cL \_}$, $\denotesS{\Tkg {\_}}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
+\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\TyTkg{} : "Sig"~\Gm~n \to "Ty"~\Gm$ as a constructor and our model needs to interpret this constructor to a function $\TyTkg{}^S : "Sig"^S \Gm^S~n \to "Ty"^S~\Gm^S$ with each judgement also interpreted.
 
@@
 
-Note that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.
+Note that $\denotesS{\goodSig{\Gm}{s}{n}}$ and $\denotesS{\goodType{\Gm}{\TyTkg{\sigma}}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gm}{f}{\sigma}{A}$ relies on $\TyTkg{}$.
 
@@
 The verification of validity of this model is largely omitted here and we refer the reader to the appendix.
-To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$ by function extensionality.
-One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together.
-The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{(p_1\mu~o)}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
+To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gm}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gm}{o}{\cL\sigma}}$ by function extensionality.
+One exception is for $\TyTkg{}$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together.
+The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gm}{\sigma}{n+1}\quad\goodTermS{\Gm}{o}{\cL\sigma}}{\goodTermS{\Gm}{(p_1\mu~o)}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
 

### 2022-11-01 canonicitymodel.tex
Fix notation for canonicity model.

@@
   \begin{itemize}
-    \item if $\goodTerm{\cdot}{t}{El~("W"~*)}$ with $\goodTerm{\cdot}{*}{\bW~\sigma}$ and $\goodWSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv "Wsup"~j~a~b}{El~("W"~*)}$ for some $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot, B[(id, a)]}{b}{El~("W"~*)}$ and $j < n$
+    \item if $\goodTerm{\cdot}{t}{El~(\wcode{\tau})}$ with $\goodWSig{\cdot}{\tau}{n}$, then $\goodTerm{\cdot}{t \equiv "Wsup"~j~a~b}{El~(\wcode{\tau})}$ for some $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot, B[(id, a)]}{b}{El~(\wcode{\tau})}$ and $j < n$
     \item if $\goodTerm{\cdot}{t}{\mathbb{B}}$ then $\goodTerm{\cdot}{t \equiv "tt"}{\mathbb{B}}$ or $\goodTerm{\cdot}{t \equiv "ff"}{\mathbb{B}}$ 
-    \item if $\goodTerm{\cdot}{t}{\cL\sigma}$ with $\goodSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv \mu^+~o~t}{\cL\sigma}$
+    \item if $\goodTerm{\cdot}{t}{\cL\sigma}$ with $\goodSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv \LkgAdd{o}{t}}{\cL\sigma}$
 
-      $\quad$ for some $\goodTerm{\cdot}{o}{\cL(p_1\nu~\sigma)}$ and $\goodTerm{\cdot, p_1\nu'~\sigma}{t}{p_2\nu~\sigma}$
+      $\quad$ for some $\goodTerm{\cdot}{o}{\cL(\lkgproj{1}{\sigma})}$ and $\goodTerm{\cdot, \lsigproj{2}{\sigma}}{t}{\lkgproj{2}{\sigma}}$
     \item if $\goodTerm{\cdot}{t}{\Sigma A B}$ then $\goodTerm{\cdot}{t = (a, b)}{\Sigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[(id, a)]}$

### 2022-11-01 standardmodel.tex

@@
   \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \denotesS{\goodSig{\Gm}{\nu\cdot}{0}} &= \star \\
-  \denotesS{\goodSig{\Gm}{\nu^+ \ \sigma \ f \ T}{n+1}} &= (\sigma, A, f, T) \\
-  \denotesS{\goodTerm{\Gm}{\mu\cdot}{\cL \nu\cdot} } (\Gm) &= \star \\
-  \denotesS{\goodTerm{\Gm}{\mu^+ \ o \ t}{\cL(\nu^+ \ \sigma  \ T)}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
-  \denotesS{\goodTerm{\Gm}{p_1\mu~t}{\cL(p_1\nu~\sigma)}}(\Gm) &=  "fst"~(t\Gm) \\ 
+  \denotesS{\goodSig{\Gm}{\LSigEmp}{0}} &= \star \\
+  \denotesS{\goodSig{\Gm}{\LSigAdd{\sigma}{s}{T}}{n+1}} &= (\sigma, A, f, T) \\
+  \denotesS{\goodTerm{\Gm}{\LkgEmp}{\cL \LSigEmp} } (\Gm) &= \star \\
+  \denotesS{\goodTerm{\Gm}{\LkgAdd{o}{t}}{\cL(\LSigAdd {\sigma}{s}  {T})}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
+  \denotesS{\goodTerm{\Gm}{\lkgproj{1}{t}}{\cL(\lsigproj{1}{\sigma})}}(\Gm) &=  "fst"~(t\Gm) \\ 
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
@@
   \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
-  &= \denotesS{{\Tkg {p_1\mu \ o}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {p_1\mu \ o})]} \\
+  &= \denotesS{{\Tkg {\lkgproj{1}{o}}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {\lkgproj{1} {o}})]} \\
   & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
@@
   \\
-  \denotesS{\goodWSig{\Gm}{w\cdot}{0}} &= [] \quad \text{ an empty list} \\ 
-  \denotesS{\goodWSig{\Gm}{(w^+ \ A \ B \ \tau)}{n+1}}
+  \denotesS{\goodWSig{\Gm}{\WSigEmp}{0}} &= [] \quad \text{ an empty list} \\ 
+  \denotesS{\goodWSig{\Gm}{(\WSigAdd{\tau}{A}{B})}{n+1}}
   &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gm}{\tau}{n}} \\ 
-  \denotesS{\goodType{\Gm}{\pi^j_1~\tau}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
-  \denotesS{\goodType{\Gm, \pi^j_1~\tau}{\pi^j_2~\tau}{}} &= "snd"~({\denotesS{\tau}}_j)\\
+  \denotesS{\goodType{\Gm}{\wsigproj{j}{1}{\tau}}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
+  \denotesS{\goodType{\Gm, \wsigproj{j}{1}{\tau}}{\wsigproj{j}{2}{\tau}}{}} &= "snd"~({\denotesS{\tau}}_j)\\
 \end{align*}
@@
     data MW {i}{j} (τ : |$\denotesS{\texttt{WSig i }\cdot\texttt{ S}}$|) : Set (lsuc i) where
-      MWsup : |$n$| < S → (a : |$\pi^n_1$|τ) → ((|$\pi^n_2$|τ) a → MW τ) 
+      MWsup : |$n$| < S → (a : |$\wsigproj{n}{1}{\tau}$|) → ((|$\wsigproj{n}{1}{\tau}$|τ) a → MW τ) 
               → MW τ
@@
 The verification of validity of this model is largely omitted here and we refer the reader to the appendix.
-To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gm}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gm}{o}{\cL\sigma}}$ by function extensionality.
+To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gm}{\lkgproj{1}{(\LkgAdd{o}{t})}}{\cL\sigma}} = \denotesS{\goodTerm{\Gm}{o}{\cL\sigma}}$ by function extensionality.
 One exception is for $\TyTkg{}$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together.
-The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gm}{\sigma}{n+1}\quad\goodTermS{\Gm}{o}{\cL\sigma}}{\goodTermS{\Gm}{(p_1\mu~o)}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.
+The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gm}{\sigma}{n+1}\quad\goodTermS{\Gm}{o}{\cL\sigma}}{\goodTermS{\Gm}{(\lkgproj{1}{o})}{\cL(\lsigproj{1}{\sigma})}}$ is proved by simply unfolding the definition.
 

### 2022-11-01 casestudies.tex
stuff

We

\bigskip

\cref{sec:coqexample-stlc}
\cref{sec:coqexample-analysis}
\cref{sec:coqexample-parser}

### 2022-11-01 coq-example.tex
Update statistic infomation about Parser example.
stuff

@@
 \subsection{Type Safety for STLC}\EDJ{We need a fuller view of the two examples. Especially show "accessing family components outside that family".}
+\label{sec:coqexample-stlc}
 
@@
 \subsection{Abstract Interpreters for \texttt{Imp}}\YZ{Try to have two runnable interpreter -- choose another simple one like constant propagation. and remove LangMore}
+\label{sec:coqexample-analysis}
 The second example is adapted and modified from \citet{zm2017}.
@@
 \subsection{An Extensible Decision Procedure for Parsing}
+\label{sec:coqexample-parser}
 The third example is a decision procedure for parsing, based on the fact that, we can use Cop's Inductive Facility to encode CFG rules.
@@
 However, from right to left (completeness), the proof from the parent family cannot be reused in the children as "alg_is_prog" is changing and the proposition itself is changing.
+So we use "overridable" to organize the proof of completeness.
+
+It is around 750 lines of codes using Family facility and around 900 lines of code using vanilla Coq. Around 350 lines of them are public helper function used by both implementations.
 

### 2022-11-01 coqimpl.tex
stuff

@@
 \item Instantiate inductive types, partial recursors, and equalities $\Rightarrow$ OK to use recursor; equalities easily discharged by \lsti{reflexivity}
-\item Instantiate recursion/induction and equalities $\Rightarrow$ OK to use recursor; equalities easily discharged by \lsti{reflexivity}\\
-      Code reuse
+\item Instantiate definition of recursive functions (induction proofs) and equalities $\Rightarrow$ OK to use recursor; equalities easily discharged by \lsti{reflexivity}\\
+      Code shared by code compiled for derived families
+\item Include 
 \end{itemize}
+
+\ifShowOldWriting
 
@@
 
+\fi
+
 \subsection{Implementation Limitation}

### 2022-11-01 metatheory-condensed.tex
stuff

@@
 
-\input{selected-typing.tex}
+\input{mltt-selected.tex}
+\input{fmltt-selected.tex}
 
@@
 
-\cref{fig:typing-selected} presents the syntax and selected typing rules of \TT.
-The new forms (relative to MLTT) are highlighted.
+\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT, and
+\cref{fig:fmltt-selected} \TT.
+%The new forms (relative to MLTT) are highlighted.
 We briefly review the base MLTT fragment first.
@@
 then applies~$\gm$.
-For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
+For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
 that has type $\TySigma{A}{B}$, then $\snd{t}$ has type

### 2022-11-01 stlc-compiled.tex
stuff

@@
 [@\outdent@]End STLC\_subst.[@\tanc{c-STLC-subst-end}@][@\outersep@]
-[@\outdent@][@\dadada@][@\hfill\codecomment{Code emitted upon definition of \texttt{ty} in family \texttt{STLC}}@][@\tanc{c-STLC-ty-start}@][@\tanc{c-STLC-ty-end}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{ty} in family \texttt{STLC}}@] [@\dadada@][@\tanc{c-STLC-ty-start}@][@\tanc{c-STLC-ty-end}@][@\outersep@]
 [@\outdent@][@\codecomment{Code emitted upon definition of \texttt{env} in family \texttt{STLC}}@][@\tanc{c-STLC-env-start}@]
@@
 [@\outdent@]End STLC\_env.[@\tanc{c-STLC-env-end}@][@\outersep@]
-[@\outdent@][@\dadada@][@\hfill\codecomment{Code emitted for other fields defined in family \texttt{STLC}}@][@\tanc{c-STLC-rest}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted for other fields defined in family \texttt{STLC}}@] [@\dadada@][@\tanc{c-STLC-rest}@][@\outersep@]
 [@\outdent@][@\codecomment{Code emitted upon completion of family \texttt{STLC}}@][@\tanc{c-STLC-start}@]
@@
 
-\makeline[-1pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!55][1.8pt]
-\makeline[-1pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!55][1.8pt]
-\makeline[-1pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!55][1.8pt]
-\makeline[-1pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!55][1.8pt]
-\makeline[-1pt]{c-STLC-rest}{c-STLC-rest}[codecomment-color!55][1.8pt]
-\makeline[-1pt]{c-STLC-start}{c-STLC-firstcol-end}[codecomment-color!55][1.8pt]
+\makeline[-1pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!50][1.8pt]
+\makeline[-1pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!50][1.8pt]
+\makeline[-1pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!50][1.8pt]
+\makeline[-1pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!50][1.8pt]
+\makeline[-1pt]{c-STLC-rest}{c-STLC-rest}[codecomment-color!50][1.8pt]
+\makeline[-1pt]{c-STLC-start}{c-STLC-firstcol-end}[codecomment-color!50][1.8pt]
 \makeline[4pt]{c-STLC-d-tm-start}{c-STLC-d-tm-end}[codecomment-color!35][1.0pt]
 \makeline[4pt]{c-STLC-d-subst-start}{c-STLC-firstcol-end}[codecomment-color!35][1.0pt]
-\makeline[.5\textwidth+8pt]{c-STLC-sndcol-start}{c-STLC-end}[codecomment-color!55][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLC-sndcol-start}{c-STLC-end}[codecomment-color!50][1.8pt]
 \makeline[.5\textwidth+12pt]{c-STLC-d-ty}{c-STLC-d-ty}[codecomment-color!35][1.0pt]
@@
 [@\outdent@]End STLCFix\_subst.[@\tanc{c-STLCFix-subst-end}@][@\outersep@]
-[@\outdent@][@\dadada@] [@\codecomment{Code emitted for other fields defined in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-rest}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted for other fields defined in family \texttt{STLCFix}}@] [@\dadada@][@\tanc{c-STLCFix-rest}@][@\outersep@]
 [@\outdent\codecomment{Code emitted upon completion of family \texttt{STLCFix}}@][@\tanc{c-STLCFix-start}@]
@@
 
-\makeline[.5\textwidth+8pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!55][1.8pt]
-\makeline[.5\textwidth+8pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!55][1.8pt]
-\makeline[.5\textwidth+8pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!55][1.8pt]
-\makeline[.5\textwidth+8pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!55][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!50][1.8pt]
 \makeline[.5\textwidth+13pt]{c-STLCFix-d-tm-start}{c-STLCFix-d-tm-end}[codecomment-color!35][1.0pt]

### 2022-11-02 coq-example.tex
Fixing STLC example.

@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}\YZ{It is more convincing if you have four or five family mixin together, positive product, negative product, sum type, boolean, and natural number, general recursion and one more}
-\begin{lstlisting}[language=Coq,frame = single, escapeinside={@}{@}]
-Family STLC_prod.
- Extend FInductive ty : Set :=
-  | ty_prod : ty -> ty -> ty.
- Extend FInductive tm : Set :=
-  | tm_prod : tm -> tm -> tm  
-  | tm_pi1 : tm -> tm  
-  | tm_pi2 : tm -> tm.
- Extend Family subst_internal.
-  Final Field tm_prod := ..
-  Final Field tm_pi1 := ..
-  Final Field tm_pi2 := ..
- EndFamily.
+  \includegraphics[width=0.5\textwidth]{coqexmaple/Mixin-Venn-Diagram.pdf}
 
- Inherits subst.
- ...
- Extend FInductive has_type 
-  : context -> term -> type := ...
- Extend FTheorem subst_lemma.
-  ... 
- Qed.
-(* ... and more, end with type safety *)
-(* Next, Invoking Mixin Command *)
-EndFamily.
-Family STLC_bool_prod 
-    extends STLC 
-    using STLC_bool using STLC_prod.
-Family STLC_bool_prod_sum 
-    extends STLC 
-    using STLC_sum using STLC_bool_prod.
-Family STLC_bool_prod_sum_fix 
-    extends STLC 
-    using STLC_bool_prod_sum using STLC_fix.
-(* ... and more, 
-   there are 11 mixins families ... *)
-\end{lstlisting}
-\includegraphics[width=0.5\textwidth]{coqexmaple/Mixin-Venn-Diagram.pdf}
-\caption{Example STLC extended with product, and Mixin with \cref{fig:STLC-example2}}
-\end{wrapfigure}
-
-We start with the very basic type safety proof of STLC, largely adapted
-from Software Foundation~\cite{pierce2014software}, which is also one of
-the primal motivating example of this project.
-
-Our emphasis here is that, we singled out some language features from
-STLC and prove the type safety for each feature separately, following
+The first example is the example throughout our paper---type safety proof of STLC, largely adapted
+from Software Foundation~\cite{pierce2014software}. We singled out each language extensions from
+STLC and prove the type safety for them separately, following
 the style of the examples in MTC~\cite{delaware2013,forsta2020}.
-And we use a \textit{not-yet-formally-defined} \textbf{mixin} feature
-to mix the semantics and properties of language features---product
+And we use the \textbf{mixin} feature---that is capable of mixing two linkage transformer---to mix the semantics and properties of language features---product
 , boolean (in \cref{fig:STLC-example}), sum type and iso-recursive type---with vanilla STLC.
 
-Using this example,
-we hope one day we can say precisely, \textit{a programming language feature itself is a
-piece of data/inheritance judgement/inherited family}. Compared with MTC,
+% Using this example,
+% we hope one day we can say precisely, \textit{a programming language feature itself is a
+% piece of data/inheritance judgement/inherited family}. 
+Compared with MTC,
 our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?}\EDJreply{I thought they use big step with fuel.
@@
 extensible inductive types.
+Most of the proof and computation are directly adapted from the one in Software Foundation, resulting in a similarly accessible proof.
+However, we highlight some special expriences when using our plugin.
 
-Most of the proof and computation are directly adapted
-from the one in Software Foundation, resulting in a similarly accessible proof, while the proof for each extended feature are scattered in the extended families resulting better modularity.
-One kind of examples is about the computation, for example, substitution functions.
+\begin{minted}[fontsize=\footnotesize]{Coq}
+Lemma subst_on_true :
+  forall G, hasty G (subst tm_true x k) ty_bool.
+Overridable pins {step}
+  FClaim not_step_tm_true:
+    forall x', ~ step tm_true x' 
+  by { intros x' h; inversion h }.
+\end{minted}
+
+Our proof is \textbf{more modular} as each part handling the extended feature are scattered in the extended families.
+
+We use \textbf{propositional computational axiom} when computing "FRecursion".
+For example, when we prove or extend substitution lemma for bool, we need to prove a subgoal like "subst_on_true" requiring a computation based on the definition of "subst".
+In vanilla Coq, that can be easily done via "simpl" but in our case, we need to use the postulated computation axiom to simplify the computation of "subst" instead.
+
+We handle \textbf{inversion lemma} differently.
+Inversion lemmas are those small lemmas while involving global reasoning.
+One example is "not_step_tm_true" which trivially asserts "tm_true" is not further reducible.
+It requires global reasoning because once the user errorneously add a constructor for "step tm_true tm_true" then this lemma cannot hold.
+Thus one way is to prove it by "FInduction" on the "step".
+However, this brings a lot of boilerplate code as everytime "step" is extended with some unrelated constructor (e.g.
+"st_if : step (tm_if tm_true b c) b"), "not_step_tm_true" needs to be extended correspondingly---we need to check "st_if" doesn't "step tm_true" requiring using \textbf{partial recursor and the computational axiom} to prove "(tm_if tm_true b c) <> tm_true" syntactically speaking using tactic "fdiscriminate".
+
+Thus we provide a small-scale \textbf{escape hatch via "Overridable"}.
+Once "pins {step}", we will know the \textit{concrete} (vanilla Coq) inductive type of "step" when proving "not_step_tm_true" and thus we can use all Coq's induction facility including "inversion".
+Once the definition of "step" extends during inheritance, due to the semantic of "Overridable", we need to provide a new proof but our plugin can also try to use the old proof script automatically.
+This trick can apply to small inversion lemma where we believe our inductive definition is ``well-formed'' enough and thus an "inversion" tactic can easily handle it.
+Another way to look at it is the fact that, this kind of inversion lemma should be part of the definition in the first hand, like a constraint---and the proof that our inductive definition satisfy this constraint should be trivial and any future extension should respect this constraint.
+
+% One kind of examples is about the computation, for example, substitution functions.
 Substitutions dealing with product and projection are defined in the corresponding children family.
@@
 At the end, it will look like scattered proof scripts in a modular manner.
-For example, the proof on progress on product projection is carried out in the children family that extend "tm" with products.
+% For example, the proof on progress on product projection is carried out in the children family that extend "tm" with products.
 \YZ{Can you say more about how your plugin enables more code reuse and modularity?}\EDJreply{please check}
@@
 % We have problem on inversion lemma. Check if it is the same problem as MTC
-We use "Closing Fact" to state and prove the inversion lemma instead of
-using the extensible proving mechanism "FTheorem". The reasons are that
-(1) it introduces much less boilerplate code because the proof for these
-inversion lemmas should be just simple case analysis and we should rely
-on Coq to generate most of the boilerplate code;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?}\EDJreply{I have make the above Closing Fact section more about this detail.
+% We use "Closing Fact" to state and prove the inversion lemma instead of
+% using the extensible proving mechanism "FTheorem". The reasons are that
+% (1) it introduces much less boilerplate code because the proof for these
+% inversion lemmas should be just simple case analysis and we should rely
+% on Coq to generate most of the boilerplate code;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?}\EDJreply{I have make the above Closing Fact section more about this detail.
 Please check.} (2) it shouldn't bother us in the future
-because any extension on the syntax should still satisfy these inversion
-lemmas; (3) most importantly, we believe this inversion ``lemma'' should
-be part of the definition of the syntax instead of considering the
-syntax as a mere concrete inductive type. We should postulate this
-inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify
-that our inductive definition does satisfy the constraint, which is
-exactly what we expect from "Closing Fact".
+% because any extension on the syntax should still satisfy these inversion
+% lemmas; (3) most importantly, we believe this inversion ``lemma'' should
+% be part of the definition of the syntax instead of considering the
+% syntax as a mere concrete inductive type. We should postulate this
+% inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify
+% that our inductive definition does satisfy the constraint, which is
+% exactly what we expect from "Closing Fact".
 If directly working on inductive definition doesn't bring us benefit during proof engineering, then decompose the property (the constraint) with the concrete definition might be a good idea.
 
-Our formulation for bare STLC is around 400+ LOCs; each families implementing product and boolean features both take around 300+ LOCs. 
+Our formulation for the base STLC family is around 400+ LOCs; each families implementing single features both take around 300+ LOCs.
+
+For each sub family, the main code bloat comes from the inversion lemma (global reasoning) and the fact we don't yet support  \mintinline{Coq}{Hint}.
 % need comparison with the original implementation
 
-The biggest difference in the proof script comes from the fact that we
-are handling ``extensible'' inductive types instead of real inductive
-types, and thus the inversion tactic is not working and we have to manually
-create inversion lemmas.
-
-Another difference comes from the fact that we need to use "Closing
-Fact" to manually verify the computational rules for each recursor and
-use tactics to ``run'' the recursor by repeatedly "rewrite" using those
-verified computational rules. This part is possible to be automatically
-generated by the plugin, but still it will require propositional 
-computational rules.  What's more, we don't yet support the \mintinline{Coq}{Hint}. These three differences on experience lead to a mild code bloat.
+% The biggest difference in the proof script comes from the fact that we
+% are handling ``extensible'' inductive types instead of real inductive
+% types, and thus the inversion tactic is not working and we have to manually
+% create inversion lemmas. 
+% What's more, we don't yet support the \mintinline{Coq}{Hint}. These three differences on experience lead to a mild code bloat.
 

### 2022-11-02 metatheory-condensed.tex
l/proj2
tighten fmltt figures a bit

@@
 Each case handler has a type of the form $\CaseSig{A}{B}{T}$, where
-$T$ is the motive of the recursion (\ruleref{ty/casety}).
+$T$ is the motive of the recursion (\ruleref{tyeq/casety}).
 This collection of case handlers then encodes those defined or inherited
@@
 fields in $t$'s context can reuse~$t$---without retyping it---by
-projecting it (\ruleref{l/proj2}) and then including it.
+projecting it (\ruleref{l/proj/2}) and then including it.
 
@@
 
+\definecolor{A-color}{HTML}{E7EFC5}
+\definecolor{T-color}{HTML}{FFF1D0}
+\definecolor{t-color}{HTML}{FAE0E4}
+
 We illustrate this with \cref{fig:stlc-linkage-typing}.
@@
 $\lkg_i$ is the linkage that adds the $i$-th field~$t_i$ with the typing
-$\goodTerm{A_i}{t_i}{T_i}$.
+$\goodTerm{%
+\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i}(-0.02,-0.10)(-0.02,0.28)
+A_i%
+\tikzmarkend{A-i}%
+}{%
+\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i}(-0.02,-0.10)(-0.02,0.28)%
+t_i%
+\tikzmarkend{t-i}%
+}{%
+\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i}(-0.02,-0.10)(-0.02,0.28)%
+T_i%
+\tikzmarkend{T-i}%
+}$.
 %
@@
 we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
-and feel free to write its type, which is a dependent tuple type, as a
+and feel free to write its type~$A_i$, which is a dependent tuple type, as a
 dependent record type with labels.
@@
 in rules \ruleref{l/add} and \ruleref{lsig/add} is
-responsible for hiding W-type signatures in~$\lsig$.
+responsible for hiding W-type signatures.
 Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
-packaging the types of all fields preceding the current field~$t$,
+that packages the types of all fields preceding the current field~$t$,
 and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
-$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new term of type $A$
+$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new tuple of type~$A$
 that hides W-type signatures behind $\cU$, if necessary.
-When no hiding is needed, $\seal$ is $\var{0}$ (the identity) and $A$ is
+%
+It is straightforward to find the~$\seal$ that fits the bill, though this process
+is not automated in \TT.
+In particular, when no hiding is needed, $\seal$ is $\var{0}$ (the identity) and $A$ is
 $\TyTkg{\lsig}$.
+In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are identities,
+and $\seal_6$ turns
+\begingroup
+\fontsize{9}{9}\selectfont%
+$\big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$
+\endgroup
+into
+\begingroup
+\fontsize{9}{9}\selectfont%
+$\big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$%
+\endgroup
+.
 
-When a family is closed and a complete linkage~$\lkg$ is formed, fields
-of the family can be accessed by projecting them out of the tuple
-$\Tkg{\lkg}$, as the last line of \cref{fig:stlc-linkage-typing} suggests.
+When a family is completed (e.g., \lsti!End!~\lsti!STLC!),
+a linkage~$\lkg$ containing all the fields is available
+(e.g., $\lkg_{20}$ in \cref{fig:stlc-linkage-typing}).
+Fields of the family can then be accessed by projecting them out of the tuple
+$\Tkg{\lkg}$.
 
@@
 To make common patterns of linkage manipulations more convenient,
-\TT builds \emph{linkage transformers}, whose well-formedness judgments
-have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$.
-Linkage transformers are second-class; the only way to use them is to
-apply them to linkages to form new linkages (\ruleref{tm/inh}).
+\TT provides a ``library'' of \emph{linkage transformers}, whose well-formedness judgments
+have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$. The idea is that
+applying~$h$ to a linkage of type~$\TyLkg{\lsig_1}$ yields a linkage of
+type~$\TyLkg{\lsig_2}$ (\ruleref{tm/inh}).
+%Linkage transformers are second-class; the only way to use them is to
+%apply them to linkages to form new linkages.
 %
-Derived families can be modeled as linkage transformers.
-%
-Linkage transformers are inductively constructed from several introduction
-rules.
-\cref{fig:typing-selected} shows the beta rule of an exemplary transformer,
-\ruleref{lxeq/ov/beta}.
+Derived families can be modeled as linkage transformers inductively
+constructed from the introduction forms $\InhId$, $\InhExt{h}{t}$, $\InhOv{h}{t}$, etc.
+\cref{fig:fmltt-selected} shows the beta rule of an exemplary transformer,
+\ruleref{tmeq/ov/beta}.
 It can be read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
@@
 We also typeset a more complete set of rules than
-\cref{fig:mltt-typing-selected,fig:fmltt-typing-selected} allow
-and present it in a supplemental appendix.
+the space allows and present it in a supplemental appendix.
 %

### 2022-11-02 cc-condensed.tex
Start Canonicity/Consistency.

% Canonicity / Consistency Condensed
\noindentparagraph{Consistency and Canonicity.}
Two of the most fundamental properties for dependent type theory are \textit{consistency} and \textit{canonicity}:
\begin{theorem}[Consistency]
  $\goodTerm{\cdot}{t}{\bot}$ is impossible, i.e. $\bot$ is not inhabitted in the empty context
\end{theorem}
\begin{theorem}[Canonicity]
  if $\goodTerm{\cdot}{t}{\cB}$ then it is either $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$
\end{theorem}
Consistency means our dependent type theory suffices as a logical system as not every proposition is trivially provable. Canonicity means our dependent type theory sufficies as ``a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}. We can even argue that, if this canonicity theorem is proven in a computable meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}}, 
then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.

The

### 2022-11-02 cc-condensed.tex
Now two models take 2.5 pages.

@@
 \begin{theorem}[Consistency]
-  $\goodTerm{\cdot}{t}{\bot}$ is impossible, i.e. $\bot$ is not inhabitted in the empty context
+  $\goodTerm{\cdot}{t}{\bot}$ is not derivable, i.e. $\bot$ is not inhabitted in the empty context
 \end{theorem}
@@
 Consistency means our dependent type theory suffices as a logical system as not every proposition is trivially provable.
-Canonicity means our dependent type theory sufficies as ``a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
+Canonicity implies our dependent type theory sufficies as ``a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
 We can even argue that, if this canonicity theorem is proven in a computable meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}},
@@
 
-The
+We refer to the attached (pseudo-)Agda style code for the complete consistency and canonicity proof. But here we sketch the big picture.
+
+\noindentparagraph{Consistency } is proved by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, by formally resorting to the concept of \textit{algebra/model} of QIIT.
+But if we were using naive set theory: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma for each rule.
+The main goal is to interpret $\denotesS{\bot} = \emptyset$ which implies it is impossible $\goodTermS{\cdot}{t}{\bot}$ and thus $\bot$ is syntactically non-derivable.\footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo-)Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation closer to interpretation in the paper}
+
+We only need to take care of the new feature for our consistency model.
+But (1) since we have used sigma type to translate the linkage, using the model for sigma type for linkage will work; (2) directly using the inductive facility of (pseudo)-Agda (or naive set theory) to for W-type is enough due to the similarity of our W-type to the conventional W-type.
+
+Here we will only show some interpretation and also omit the universe level to sketch out the idea.
+
+\definecolor{new}{HTML}{F5EFE6}
+
+\begin{align*}
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod1}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodCtx{\Gm}{i}} & \text{ is a Set and we define } \goodCtxS{\Gm}{i} :\iff \denotesS{\goodCtx{\Gm}{i}} \text{ is defined}\tikzmarkend{stmod1} 
+  \\
+  \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
+  \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \Gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\Gm)\}
+  \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gm}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gm}{\Gm}{\Delta}} : \iff \denotesS{\Gm} \in \denotesS{\goodSub{\Gm}{\_}{\Delta}} \tikzmarkend{stmod2}
+  \\
+  \denotesS{\goodSub{\Gm}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gm}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gm}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gm}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gm}{\_}{T}}
+  \tikzmarkend{stmod3}
+  \\
+  \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\Gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\Gm) \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod4}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodType{\Gm}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gm}{i}}  
+  \\
+  & \text{ and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesS{\goodType{\Gm}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 
+  \\
+  \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\Gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod5}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gm}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gm}{\_}{n}} \tikzmarkend{stmod5} 
+  \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod6}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
+  \denotesS{\goodSig{\Gm}{\_}{n+1}} &= \{
+    (\denotesS{\sigma}, \denotesS{A}, \denotesS{f}, \denotesS{T}) :
+      {\goodSigS{\Gm}{\sigma}{n}}
+      \land  {\goodTypeS{\Gm}{A}{}} \\  
+      & \quad \quad \quad \land  {\goodSealS{\Gm}{f}{\sigma}{A}}
+      \land  {\goodTypeS{\Gm, A}{T}{i}} 
+  \} \\ 
+  \denotesS{\goodSig{\Gm}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
+  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gm}{\TyTkg {\tau}}{}} \denotesS{T["sf" \ f]} \\
+  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod7}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\Gm) \land t \in \prod_{m' \in \denotesS{A}(\Gm)}\denotesS{T}(\Gm, m')  \}  \\
+  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+  \denotesS{\goodSig{\Gm}{\LSigEmp}{0}} &= \star \\
+  \denotesS{\goodSig{\Gm}{\LSigAdd{\sigma}{s}{T}}{n+1}} &= (\sigma, A, f, T) \\
+  \denotesS{\goodTerm{\Gm}{\LkgEmp}{\cL \LSigEmp} } (\Gm) &= \star \\
+  \denotesS{\goodTerm{\Gm}{\LkgAdd{o}{t}}{\cL(\LSigAdd {\sigma}{s}  {T})}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
+  \denotesS{\goodTerm{\Gm}{\lkgproj{1}{t}}{\cL(\lsigproj{1}{\sigma})}}(\Gm) &=  "fst"~(t\Gm) \\ 
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}} & \text{ is a dependent function, defined inductively on } {\sigma} 
+  \tikzmarkend{stmod8}
+  \\
+  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
+  &= \denotesS{{\Tkg {\lkgproj{1}{o}}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {\lkgproj{1} {o}})]} \\
+  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
+  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+  \denotesS{\goodType{\Gm}{\bot}{i}} \ (\Gm) &= \emptyset \\
+  \denotesS{\goodType{\Gm}{\cB}{i}} \ (\Gm) &= \{0, 1\} \quad \text{a specific two element set} \\
+  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod9}(0.05,-0.15)(-0.10,0.30)
+  \denotesS{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
+  & \text{ and we define } \goodWSigS{\Gm}{\tau}{n} :\iff \denotesS{\goodWSig{\Gm}{\tau}{n}} \in \denotesS{\goodWSig{\Gm}{\_}{n}}
+  \tikzmarkend{stmod9}
+  \\
+  \denotesS{\goodWSig{\Gm}{\WSigEmp}{0}} &= [] \quad \text{ an empty list} \\ 
+  \denotesS{\goodWSig{\Gm}{(\WSigAdd{\tau}{A}{B})}{n+1}}
+  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gm}{\tau}{n}} \\ 
+  \denotesS{\goodType{\Gm}{\wsigproj{j}{1}{\tau}}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
+  \denotesS{\goodType{\Gm, \wsigproj{j}{1}{\tau}}{\wsigproj{j}{2}{\tau}}{}} &= "snd"~({\denotesS{\tau}}_j)\\
+\end{align*}
+
+Here, we need to clarify some ambiguity in our latex formulation.
+For example, when we wrote ``$\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$ defined inductively on ${\sigma}$'', the best way to understand it is to consider we are defining a function $\denotesS{\goodType{\Gm}{\TyTkg {\_}}{}}$ as the interpretation for $\goodType{\Gm}{\TyTkg {\_}}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$, inductively on the length of the signature.
+The case is similar for $\denotesS{\cL \_}$, $\denotesS{\Tkg {\_}}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
+\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\TyTkg{} : "Sig"~\Gm~n \to "Ty"~\Gm$ as a constructor and our model needs to interpret this constructor to a function $\TyTkg{}^S : "Sig"^S \Gm^S~n \to "Ty"^S~\Gm^S$ with each judgement also interpreted.
+
+Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
+What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-definedness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
+
+Note that $\denotesS{\goodSig{\Gm}{s}{n}}$ and $\denotesS{\goodType{\Gm}{\TyTkg{\sigma}}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gm}{f}{\sigma}{A}$ relies on $\TyTkg{}$.
+
+Finally, notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
+and that is a contradiction.
+
+\noindentparagraph{Canonicity} is proved by following the canonicity argument from~\citet{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model.
+Canonicity is also called \textit{reducibility}.
+
+Now we sketch out the big picture of the proof.
+The canonicity model $"Tm"^C~\Gamma^C~"T"^C$ will map each syntax piece $("t : Tm Γ T") \mapsto "(t, tₚ) :" "Tm"^C~\Gamma^C~"T"^C = \sum~("t : Tm Γ T")."Tm"_p~\Gamma^C~"T"^C~"t"$ to a dependent pair\footnote{This $"Tm"_p$ is usually called dependent model.}, of which the first part will be the same as the input syntax.
+Using this mapping on closed boolean term $"t : Tm ⋅ "\mathbb{B}$ will give us a proof of $"Tm"_p~\cdot^C~\mathbb{B}^C~"t"$, which will unfolds to our final goal $"t" \equiv "tt" + "t" \equiv "ff"$.
+Roughly, this can read as ``every closed boolean term is reducible''.
+Thus once we can ``propagate'' $(\cdot)^C$ (i.e.
+construct this model completely) for every syntactic piece, we are done.
+In our (pseudo) Agda formulation, we use $(\cdot)_2$ to replace the notation of $(\cdot)^C$ because we consider most parts of this model are actually a dependent pair.
+
+Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as ``a reducible set'': $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
+The reformulation can be understood as an unfolding of the \textit{type-theoretic logical relation proof}.\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory} If we use the terminology of the conventional formulation, this reducible set means if we apply its element $t$ with a substitution $\gamma$ from the set $\Gamma$ (where we have $\Gamma_p$ witnesses the ``reducibility'' of $\Gamma$), then we will have $t[\gamma] \in T_p$ as a closed and \textbf{reducible} term.
+Now if $\Gamma$ is actually empty $\cdot$ and $T$ is $\cB$, we are done.
+
+Again, using the same insight from earlier, we reuse the model for sigma type and W-type for linkages and our inductive type.

### 2022-11-02 metatheory-condensed.tex
Inser CC-condensed inplace.
Venn diagram

@@
 a linkage~$\lkg$ containing all the fields is available
-(e.g., $\lkg_{20}$ in \cref{fig:stlc-linkage-typing}).
+(e.g., $\lkg_{53}$ in \cref{fig:stlc-linkage-typing}).
 Fields of the family can then be accessed by projecting them out of the tuple
@@
 \input{stlc-linkage-typing.tex}
+
+\input{cc-condensed.tex}
 

### 2022-11-02 cc-condensed.tex

@@
 
-We refer to the attached (pseudo-)Agda style code for the complete consistency and canonicity proof. But here we sketch the big picture.
+We refer to the attached (pseudo-)Agda style code for \textbf{the complete consistency and canonicity proof}. But here we sketch the big picture.
 
-\noindentparagraph{Consistency } is proved by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, by formally resorting to the concept of \textit{algebra/model} of QIIT.
+\noindentparagraph{Consistency} is proved by extending the standard model from~\citet{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization.
 But if we were using naive set theory: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma for each rule.
@@
 
-Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
+Basically in Agda, we are simply constructing a function that mapping our QIIT data-type(stands for syntax) to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
 What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-definedness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
-
 Note that $\denotesS{\goodSig{\Gm}{s}{n}}$ and $\denotesS{\goodType{\Gm}{\TyTkg{\sigma}}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gm}{f}{\sigma}{A}$ relies on $\TyTkg{}$.
@@
 Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as ``a reducible set'': $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
-The reformulation can be understood as an unfolding of the \textit{type-theoretic logical relation proof}.\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory} If we use the terminology of the conventional formulation, this reducible set means if we apply its element $t$ with a substitution $\gamma$ from the set $\Gamma$ (where we have $\Gamma_p$ witnesses the ``reducibility'' of $\Gamma$), then we will have $t[\gamma] \in T_p$ as a closed and \textbf{reducible} term.
+This reformulation can be understood as an unfolding of the \textit{type-theoretic logical relation proof}.\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory} If we use the terminology of the conventional formulation, this reducible set means if we apply its element $t$ with a substitution $\gamma$ from the set $\Gamma$ (where we have $\Gamma_p$ witnesses the ``reducibility'' of $\Gamma$), then we will have $t[\gamma] \in T_p$ as a closed and \textbf{reducible} term.
 Now if $\Gamma$ is actually empty $\cdot$ and $T$ is $\cB$, we are done.

### 2022-11-02 standardmodel.tex

@@
-\newcommand{\goodCtxS}[2]{{ {#1} \ \models_S }}
-\newcommand{\goodTypeS}[3]{{ {#1} \models_S {#2} }}
-\newcommand{\goodTermS}[3]{{ {#1} \models_S {#2} : {#3} }}
-\newcommand{\goodSubS}[3]{{ {#1} \models_S {#2} : {#3} }}
-\newcommand{\goodSigS}[3]{{ {#1} \models_S {#2} \ \  \texttt{LSig}^{#3} }}
-\newcommand{\goodWSigS}[3]{{ {#1} \models_S {#2} \ \ \texttt{WSig}^{#3} }}
-\newcommand{\goodSealS}[4]{{ {#1} \models_S {#2} : {#3} \  |\  {#4} }}
-\newcommand{\goodInhS}[4]{{ {#1} \models_S {#2} : {#3} \twoheadrightarrow {#4}}}
-
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
@@
 
-\begin{align*}
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod1}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodCtx{\Gm}{i}} & \text{ is a Set and we define } \goodCtxS{\Gm}{i} :\iff \denotesS{\goodCtx{\Gm}{i}} \text{ is defined}\tikzmarkend{stmod1} 
-  \\
-  \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \Gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\Gm)\}
-  \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gm}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gm}{\Gm}{\Delta}} : \iff \denotesS{\Gm} \in \denotesS{\goodSub{\Gm}{\_}{\Delta}} \tikzmarkend{stmod2}
-  \\
-  \denotesS{\goodSub{\Gm}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gm}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gm}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gm}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gm}{\_}{T}}
-  \tikzmarkend{stmod3}
-  \\
-  \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\Gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\Gm) \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod4}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gm}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gm}{i}}  
-  \\
-  & \text{ and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesS{\goodType{\Gm}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 
-  \\
-  \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\Gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod5}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gm}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gm}{\_}{n}} \tikzmarkend{stmod5} 
-  \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod6}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
-  \denotesS{\goodSig{\Gm}{\_}{n+1}} &= \{
-    (\denotesS{\sigma}, \denotesS{A}, \denotesS{f}, \denotesS{T}) :
-      {\goodSigS{\Gm}{\sigma}{n}}
-      \land  {\goodTypeS{\Gm}{A}{}} \\  
-      & \quad \quad \quad \land  {\goodSealS{\Gm}{f}{\sigma}{A}}
-      \land  {\goodTypeS{\Gm, A}{T}{i}} 
-  \} \\ 
-  \denotesS{\goodSig{\Gm}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gm}{\TyTkg {\tau}}{}} \denotesS{T["sf" \ f]} \\
-  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod7}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\Gm) \land t \in \prod_{m' \in \denotesS{A}(\Gm)}\denotesS{T}(\Gm, m')  \}  \\
-  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \denotesS{\goodSig{\Gm}{\LSigEmp}{0}} &= \star \\
-  \denotesS{\goodSig{\Gm}{\LSigAdd{\sigma}{s}{T}}{n+1}} &= (\sigma, A, f, T) \\
-  \denotesS{\goodTerm{\Gm}{\LkgEmp}{\cL \LSigEmp} } (\Gm) &= \star \\
-  \denotesS{\goodTerm{\Gm}{\LkgAdd{o}{t}}{\cL(\LSigAdd {\sigma}{s}  {T})}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
-  \denotesS{\goodTerm{\Gm}{\lkgproj{1}{t}}{\cL(\lsigproj{1}{\sigma})}}(\Gm) &=  "fst"~(t\Gm) \\ 
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}} & \text{ is a dependent function, defined inductively on } {\sigma} 
-  \tikzmarkend{stmod8}
-  \\
-  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
-  &= \denotesS{{\Tkg {\lkgproj{1}{o}}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {\lkgproj{1} {o}})]} \\
-  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
-  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
-  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \denotesS{\goodType{\Gm}{\bot}{i}} \ (\Gm) &= \emptyset \\
-  \denotesS{\goodType{\Gm}{\cB}{i}} \ (\Gm) &= \{0, 1\} \quad \text{a specific two element set} \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod9}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
-  & \text{ and we define } \goodWSigS{\Gm}{\tau}{n} :\iff \denotesS{\goodWSig{\Gm}{\tau}{n}} \in \denotesS{\goodWSig{\Gm}{\_}{n}}
-  \tikzmarkend{stmod9}
-  \\
-  \denotesS{\goodWSig{\Gm}{\WSigEmp}{0}} &= [] \quad \text{ an empty list} \\ 
-  \denotesS{\goodWSig{\Gm}{(\WSigAdd{\tau}{A}{B})}{n+1}}
-  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gm}{\tau}{n}} \\ 
-  \denotesS{\goodType{\Gm}{\wsigproj{j}{1}{\tau}}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
-  \denotesS{\goodType{\Gm, \wsigproj{j}{1}{\tau}}{\wsigproj{j}{2}{\tau}}{}} &= "snd"~({\denotesS{\tau}}_j)\\
-\end{align*}
+% \begin{align*}
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod1}(0.05,-0.15)(-0.10,0.30)
+%   \denotesS{\goodCtx{\Gm}{i}} & \text{ is a Set and we define } \goodCtxS{\Gm}{i} :\iff \denotesS{\goodCtx{\Gm}{i}} \text{ is defined}\tikzmarkend{stmod1} 
+%   \\
+%   \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
+%   \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \Gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\Gm)\}
+%   \\
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gm}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gm}{\Gm}{\Delta}} : \iff \denotesS{\Gm} \in \denotesS{\goodSub{\Gm}{\_}{\Delta}} \tikzmarkend{stmod2}
+%   \\
+%   \denotesS{\goodSub{\Gm}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gm}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gm}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gm}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gm}{\_}{T}}
+%   \tikzmarkend{stmod3}
+%   \\
+%   \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\Gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\Gm) \\
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod4}(0.05,-0.15)(-0.10,0.30)
+%   \denotesS{\goodType{\Gm}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gm}{i}}  
+%   \\
+%   & \text{ and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesS{\goodType{\Gm}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 
+%   \\
+%   \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\Gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod5}(0.05,-0.15)(-0.10,0.30)
+%   \denotesS{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gm}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gm}{\_}{n}} \tikzmarkend{stmod5} 
+%   \\
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod6}(0.05,-0.15)(-0.10,0.30)
+%   \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
+%   \denotesS{\goodSig{\Gm}{\_}{n+1}} &= \{
+%     (\denotesS{\sigma}, \denotesS{A}, \denotesS{f}, \denotesS{T}) :
+%       {\goodSigS{\Gm}{\sigma}{n}}
+%       \land  {\goodTypeS{\Gm}{A}{}} \\  
+%       & \quad \quad \quad \land  {\goodSealS{\Gm}{f}{\sigma}{A}}
+%       \land  {\goodTypeS{\Gm, A}{T}{i}} 
+%   \} \\ 
+%   \denotesS{\goodSig{\Gm}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
+%   \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gm}{\TyTkg {\tau}}{}} \denotesS{T["sf" \ f]} \\
+%   & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+%   \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod7}(0.05,-0.15)(-0.10,0.30)
+%   \denotesS{\goodType{\Gm}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
+%   \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\Gm) \land t \in \prod_{m' \in \denotesS{A}(\Gm)}\denotesS{T}(\Gm, m')  \}  \\
+%   & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+%   \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+%   \denotesS{\goodSig{\Gm}{\LSigEmp}{0}} &= \star \\
+%   \denotesS{\goodSig{\Gm}{\LSigAdd{\sigma}{s}{T}}{n+1}} &= (\sigma, A, f, T) \\
+%   \denotesS{\goodTerm{\Gm}{\LkgEmp}{\cL \LSigEmp} } (\Gm) &= \star \\
+%   \denotesS{\goodTerm{\Gm}{\LkgAdd{o}{t}}{\cL(\LSigAdd {\sigma}{s}  {T})}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
+%   \denotesS{\goodTerm{\Gm}{\lkgproj{1}{t}}{\cL(\lsigproj{1}{\sigma})}}(\Gm) &=  "fst"~(t\Gm) \\ 
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
+%   \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}} & \text{ is a dependent function, defined inductively on } {\sigma} 
+%   \tikzmarkend{stmod8}
+%   \\
+%   \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
+%   &= \denotesS{{\Tkg {\lkgproj{1}{o}}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {\lkgproj{1} {o}})]} \\
+%   & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
+%   \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
+%   &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+%   \denotesS{\goodType{\Gm}{\bot}{i}} \ (\Gm) &= \emptyset \\
+%   \denotesS{\goodType{\Gm}{\cB}{i}} \ (\Gm) &= \{0, 1\} \quad \text{a specific two element set} \\
+%   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod9}(0.05,-0.15)(-0.10,0.30)
+%   \denotesS{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
+%   & \text{ and we define } \goodWSigS{\Gm}{\tau}{n} :\iff \denotesS{\goodWSig{\Gm}{\tau}{n}} \in \denotesS{\goodWSig{\Gm}{\_}{n}}
+%   \tikzmarkend{stmod9}
+%   \\
+%   \denotesS{\goodWSig{\Gm}{\WSigEmp}{0}} &= [] \quad \text{ an empty list} \\ 
+%   \denotesS{\goodWSig{\Gm}{(\WSigAdd{\tau}{A}{B})}{n+1}}
+%   &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gm}{\tau}{n}} \\ 
+%   \denotesS{\goodType{\Gm}{\wsigproj{j}{1}{\tau}}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
+%   \denotesS{\goodType{\Gm, \wsigproj{j}{1}{\tau}}{\wsigproj{j}{2}{\tau}}{}} &= "snd"~({\denotesS{\tau}}_j)\\
+% \end{align*}
 
@@
 
-Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
+Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps constructor $"Con"$ to $"Con"^S$, and constructor "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
 What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-definedness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.

### 2022-11-02 coq-example.tex
Fixing all-typing-rule
Fixing Abstract Interpretation example.

@@
 \begin{minted}[fontsize=\footnotesize]{Coq}
-Lemma subst_on_true :
-  forall G, hasty G (subst tm_true x k) ty_bool.
-Overridable pins {step}
-  FClaim not_step_tm_true:
-    forall x', ~ step tm_true x' 
+Lemma subst_on_true : forall G, hasty G (subst tm_true x k) ty_bool.
+Overridable pins {step} FClaim not_step_tm_true: forall x', ~ step tm_true x' 
     by { intros x' h; inversion h }.
@@
 
-Our formulation for the base STLC family is around 400+ LOCs; each families implementing single features both take around 300+ LOCs.
+Our formulation for the base STLC family is around 400+ LOCs; each families implementing a single feature takes around 100~300 LOCs.
 
@@
 The second example is adapted and modified from \citet{zm2017}.
-Contrary to our first example, here, we use a big-step interpreter with fuel
+Contrary to our first example, here we use a big-step interpreter with fuel
 as the operational semantics of an imperative language with
 side effects, and we specify the abstract interpretation and prove its
-soundness, with some postulation on both computation and property.
-Then we instantiate our abstract interpretation with two concrete computation---one for type inference and the other for constant propagation.
+soundness. 
+% Then we instantiate our abstract interpretation with two concrete computation---one for type inference and the other for constant propagation. 
 Our analyzer will return the analysis result about the end of the program, e.g.
 our constant propagator will return what constant value each variable is at the end of the program.
-Thanks to the compilation to Coq module, we can directly run the resulting
-abstract interpreters.
+Thanks to the compilation to Coq module, we can directly run the resulting abstract interpreters.
 
@@
 
-We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful.
-They are here to show family inheritance is useful, and can extend stuff to all sorts of things.
-I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family?
-What is lost by putting the two interpreters in the same family?}\EDJreply{Nothing is lost.
-This is just simulating people starting with bare (big-step) operational semantic.
-Because they may not working on abstract interpretation from the first place.
-Those people may extend "Lang" with other stuff like another type-safety proof or something.}
+\begin{minted}[fontsize=\footnotesize]{Coq}
+  Lemma analyze_faithful : forall prog st end_st, 
+    interp prog st = end_st -> abs_interp prog (lift st) = (lift end_st)
+\end{minted}
+
+We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.
 The first family "Lang" define the big step operator (using fuel) of a
 while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200 LOCs.
-% We
-% can
-\YZ{In general, in a result section like this, you want to show what
-you *have done* rather than what you *can do*.}\EDJreply{Ok.
-deleted.
-I initially want to emphasize ``I **can** use arbitrary implementation for this environment mapping interface'' but I don't know how to say that without using ``can''.}
-% swap the implementation of "State" to have different memory/state
-% management by extending family "Lang".
-\YZ{What is 'different
-memory/state management'?}\EDJreply{Actually I should say environment mapping/dictionary.
-I instantiate it with function(closure) as environment mapping in "LangRun".
-It is possible to use other mapping like Coq's primitive dictionary.}
 
-The second family "LangwAbs" define the abstract interpreter for "Lang" in "LangwAbs.AbsInterpreter",
-together with a bunch of postulates as a big sealed family "absStateData". Then we prove the correctness of the
-abstract interpretation function "analyze". It takes about 300+ LOCs. This example shows how
-to use family inheritance to extend a concrete interpreter with abstract
-interpreters. 
+The second family "LangwAbs" define the abstract interpreter for "Lang" in "LangwAbs.AbsInterpreter".
+The idea of abstract interpretation is to lift concrete interpretation, their value and their states into their abstract correspondence.
+Then we come up with a faithful (one example would be "analyze_faithful") and generic abstract intepreter.
+Then, the user can freely choose the instantiation and lifting of this generic framework and end up with their aiming program analyzer.
+For example, lifting 1 into abstract value "integer" can give us a type inferer; lifting 1 into an element of a very flat lattice can gives us a constant propagator.
+There are also special mathematical structure required in the abstract value (like lattices) for the abstract interpretation.
+
+Thus, to implement this generic abstract interpreter, we need to postulate the abstract value and their mathematical property.
+Then we need to prove the abstract interpreter is \textbf{sound and terminating}  with the help of these postulations.
+Then for concrete analyzer, we only need to instantiate these postulation.
+The construction and proof of soundness uses "FInduction" and "FRecursion".
+"analyze_soundness" is saying if
+It takes about 300+ LOCs. 
+% This example shows how
+% to use family inheritance to extend a concrete interpreter with abstract
+% interpreters. 
 % This example also shows in this family polymoprhism framework we can reason about "Lang" with computation details being abstracted.
@@
 
-The third family and the fourth family "LangTyping" and "LangCP" instantiate some of the postulates of the
-"LangwAbs" and recover computational information of "analyze". The corresponding fields in "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
+The third family and the fourth family "LangTyping" and "LangCP" instantiate the postulates of the
+"LangwAbs" and recover computational information of "analyze". The corresponding fields in "absStateData" is overridden by concrete implementations. The resulting abstract interpreter is expected
 to act as a type-analyzer and constant-propagator respectively. It takes about 200+ LOCs. 

### 2022-11-02 coq-example.tex
Recheck the number Removing line number from type info.

@@
 
-Our formulation for the base STLC family is around 400+ LOCs; each families implementing a single feature takes around 100~300 LOCs.
+Our formulation for the base STLC family is around 400 LOCs; each families implementing a single feature takes 100\~250 LOCs.
 
@@
 Then we need to prove the abstract interpreter is \textbf{sound and terminating}  with the help of these postulations.
+In our implementation, this postulation happens in "absStateData" where we indicate the "lift" of concrete interpretation and the mathematical properties these lifted should satisfy.
 Then for concrete analyzer, we only need to instantiate these postulation.
-The construction and proof of soundness uses "FInduction" and "FRecursion".
-"analyze_soundness" is saying if
+The construction and proof of soundness uses "FInduction" and "FRecursion", mainly on "exp" and "stmt".
 It takes about 300+ LOCs. 

### 2022-11-02 coq-example.tex
Update number for Analysis Example.

@@
 In our implementation, this postulation happens in "absStateData" where we indicate the "lift" of concrete interpretation and the mathematical properties these lifted should satisfy.
+For example, "absState" is the lift of concrete state, "join" is doing lattice join on two abstract value...
 Then for concrete analyzer, we only need to instantiate these postulation.
 The construction and proof of soundness uses "FInduction" and "FRecursion", mainly on "exp" and "stmt".
-It takes about 300+ LOCs. 
+It takes about 450 LOCs. 
 % This example shows how
@@
 The third family and the fourth family "LangTyping" and "LangCP" instantiate the postulates of the
-"LangwAbs" and recover computational information of "analyze". The corresponding fields in "absStateData" is overridden by concrete implementations. The resulting abstract interpreter is expected
-to act as a type-analyzer and constant-propagator respectively. It takes about 200+ LOCs. 
+"LangwAbs" and thus recover computational information of "analyze".
+"LangCP" also extend "exp" with integer expression.
+The corresponding fields in "absStateData" is overridden by concrete implementations.
+The resulting abstract interpreter is expected to act as a type-analyzer and constant-propagator respectively.
+They takes 200 and 300 LOCs respectively.
 % These are two examples of instantiate the detail computation of abstract interpretation.

### 2022-11-02 casestudies.tex
Venn diagram
istart on stlc case study

@@
-We
+To demonstrate the applicability of our design and implementation, we
+conduct three case studies.
+
+\noindentparagraph{Type safety of STLCs.}
+
+The first case study is the mechanization of the type safety theorem of
+STLC and those of its extensions,
+which has been occurring in the examples in this paper.
+The code base is ported from Software Foundations~\cite{sf-pl}.
+%
+The base STLC family consists of about 400 lines of code.
+Lines of code in each of the four derived families
+($\mathrm{Y}$, $\times$, $+$, and $\mu$ in the Venn diagram)
+vary from 100 to 300, largely depending on how many constructors they
+add to the inductive types.
+
+\input{stlc-venn.tex}
+
+Using individual families to organize the mechanization of individual
+language features leads to a modular design that also facilitates code reuse.
+%
+Individually developed language features can be easily composed (as
+mixins) to form new STLC variants.
+%
+Composing features can lead to \emph{feature interactions}~\cite{batory2011feature}:
+features working correctly in isolation may require coordination when composed.
+For example, composing \lsti{STLCIsorec} and \lsti{STLCProd}
+(\cref{fig:stlc-isorec-prod}) creates a need to extend \lsti{tysubst} to
+handle \lsti{ty_prod}, which the type-checker enforces.
+
+The linguistic nature of our approach allows us to retain a programming
+style similar to the original proofs while achieving extensibility.
 

### 2022-11-02 coq-example.tex
istart on stlc case study

--- 2022-11-02 coq-example.tex
@@
-\newpage\EDJ{Please delete this newpage command later}
+\newpage
 We provide three examples here for illustration and evaluation of our Coq plugin.
 
-\subsection{Type Safety for STLC}\EDJ{We need a fuller view of the two examples. Especially show "accessing family components outside that family".}
+\subsection{Type Safety for STLC}
 \label{sec:coqexample-stlc}
 
-  \includegraphics[width=0.5\textwidth]{coqexmaple/Mixin-Venn-Diagram.pdf}
+% \includegraphics[width=0.5\textwidth]{coqexmaple/Mixin-Venn-Diagram.pdf}
 
@@
 And we use the \textbf{mixin} feature---that is capable of mixing two linkage transformer---to mix the semantics and properties of language features---product
-, boolean (in \cref{fig:STLC-example}), sum type and iso-recursive type---with vanilla STLC.
+, boolean, sum type and iso-recursive type---with vanilla STLC.
 
@@
 Though Coq a la carte use small-step in one example} by exploiting
-extensible inductive types.
-Most of the proof and computation are directly adapted from the one in Software Foundation, resulting in a similarly accessible proof.
-However, we highlight some special expriences when using our plugin.
+extensible inductive types. Most of the proof and computation are
+directly adapted from the one in Software Foundation, resulting in a
+similarly accessible proof. However, we highlight some special
+experiences when using our plugin.
 
 \begin{minted}[fontsize=\footnotesize]{Coq}
-Lemma subst_on_true : forall G, hasty G (subst tm_true x k) ty_bool.
-Overridable pins {step} FClaim not_step_tm_true: forall x', ~ step tm_true x' 
+Lemma subst_on_true :
+  forall G, hasty G (subst tm_true x k) ty_bool.
+Overridable pins {step}
+  FClaim not_step_tm_true:
+    forall x', ~ step tm_true x' 
   by { intros x' h; inversion h }.
@@
 One example is "not_step_tm_true" which trivially asserts "tm_true" is not further reducible.
-It requires global reasoning because once the user errorneously add a constructor for "step tm_true tm_true" then this lemma cannot hold.
-Thus one way is to prove it by "FInduction" on the "step".
-However, this brings a lot of boilerplate code as everytime "step" is extended with some unrelated constructor (e.g.
+It requires global reasoning because once the user erroneously add a constructor for "step tm_true tm_true" then this lemma cannot hold.
+Thus, one way is to prove it by "FInduction" on the "step".
+However, this brings a lot of boilerplate code as every time "step" is extended with some unrelated constructor (e.g.
 "st_if : step (tm_if tm_true b c) b"), "not_step_tm_true" needs to be extended correspondingly---we need to check "st_if" doesn't "step tm_true" requiring using \textbf{partial recursor and the computational axiom} to prove "(tm_if tm_true b c) <> tm_true" syntactically speaking using tactic "fdiscriminate".
@@
 Once "pins {step}", we will know the \textit{concrete} (vanilla Coq) inductive type of "step" when proving "not_step_tm_true" and thus we can use all Coq's induction facility including "inversion".
-Once the definition of "step" extends during inheritance, due to the semantic of "Overridable", we need to provide a new proof but our plugin can also try to use the old proof script automatically.
+Once the definition of "step" extends during inheritance, due to the semantic of "Overridable", we need to provide a new proof, but our plugin can also try to use the old proof script automatically.
 This trick can apply to small inversion lemma where we believe our inductive definition is ``well-formed'' enough and thus an "inversion" tactic can easily handle it.
@@
 
-Our formulation for the base STLC family is around 400+ LOCs; each families implementing a single feature takes around 100~300 LOCs.
+Our formulation for the base STLC family is around 400+ LOCs; each family implementing a single feature takes around 100~300 LoCs.
 
@@
 The second example is adapted and modified from \citet{zm2017}.
-Contrary to our first example, here we use a big-step interpreter with fuel
+Contrary to our first example, here, we use a big-step interpreter with fuel
 as the operational semantics of an imperative language with
 side effects, and we specify the abstract interpretation and prove its
-soundness. 
-% Then we instantiate our abstract interpretation with two concrete computation---one for type inference and the other for constant propagation. 
+soundness, with some postulation on both computation and property.
+Then we instantiate our abstract interpretation with two concrete computation---one for type inference and the other for constant propagation.
 Our analyzer will return the analysis result about the end of the program, e.g.
 our constant propagator will return what constant value each variable is at the end of the program.
-Thanks to the compilation to Coq module, we can directly run the resulting abstract interpreters.
+Thanks to the compilation to Coq module, we can directly run the resulting
+abstract interpreters.
 
@@
 
-\begin{minted}[fontsize=\footnotesize]{Coq}
-  Lemma analyze_faithful : forall prog st end_st, 
-    interp prog st = end_st -> abs_interp prog (lift st) = (lift end_st)
-\end{minted}
-
-We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.
+We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful.
+They are here to show family inheritance is useful, and can extend stuff to all sorts of things.
+I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family?
+What is lost by putting the two interpreters in the same family?}\EDJreply{Nothing is lost.
+This is just simulating people starting with bare (big-step) operational semantic.
+Because they may not working on abstract interpretation from the first place.
+Those people may extend "Lang" with other stuff like another type-safety proof or something.}
 The first family "Lang" define the big step operator (using fuel) of a
 while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200 LOCs. 
+% We
+% can
+\YZ{In general, in a result section like this, you want to show what
+you *have done* rather than what you *can do*.}\EDJreply{Ok.
+deleted.
+I initially want to emphasize ``I **can** use arbitrary implementation for this environment mapping interface'' but I don't know how to say that without using ``can''.}
+% swap the implementation of "State" to have different memory/state
+% management by extending family "Lang".
+\YZ{What is 'different
+memory/state management'?}\EDJreply{Actually I should say environment mapping/dictionary.
+I instantiate it with function(closure) as environment mapping in "LangRun".
+It is possible to use other mapping like Coq's primitive dictionary.}
 
-The second family "LangwAbs" define the abstract interpreter for "Lang" in "LangwAbs.AbsInterpreter".
-The idea of abstract interpretation is to lift concrete interpretation, their value and their states into their abstract correspondence.
-Then we come up with a faithful (one example would be "analyze_faithful") and generic abstract intepreter.
-Then, the user can freely choose the instantiation and lifting of this generic framework and end up with their aiming program analyzer.
-For example, lifting 1 into abstract value "integer" can give us a type inferer; lifting 1 into an element of a very flat lattice can gives us a constant propagator.
-There are also special mathematical structure required in the abstract value (like lattices) for the abstract interpretation.
-
-Thus, to implement this generic abstract interpreter, we need to postulate the abstract value and their mathematical property.
-Then we need to prove the abstract interpreter is \textbf{sound and terminating}  with the help of these postulations.
-Then for concrete analyzer, we only need to instantiate these postulation.
-The construction and proof of soundness uses "FInduction" and "FRecursion".
-"analyze_soundness" is saying if
-It takes about 300+ LOCs. 
-% This example shows how
-% to use family inheritance to extend a concrete interpreter with abstract
-% interpreters. 
+The second family "LangwAbs" define the abstract interpreter for "Lang" in "LangwAbs.AbsInterpreter",
+together with a bunch of postulates as a big sealed family "absStateData". Then we prove the correctness of the
+abstract interpretation function "analyze". It takes about 300+ LOCs. This example shows how
+to use family inheritance to extend a concrete interpreter with abstract
+interpreters. 
 % This example also shows in this family polymoprhism framework we can reason about "Lang" with computation details being abstracted.
@@
 
-The third family and the fourth family "LangTyping" and "LangCP" instantiate the postulates of the
-"LangwAbs" and recover computational information of "analyze". The corresponding fields in "absStateData" is overridden by concrete implementations. The resulting abstract interpreter is expected
+The third family and the fourth family "LangTyping" and "LangCP" instantiate some of the postulates of the
+"LangwAbs" and recover computational information of "analyze". The corresponding fields in "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
 to act as a type-analyzer and constant-propagator respectively. It takes about 200+ LOCs. 

### 2022-11-03 casestudies.tex

@@
 ($\mathrm{Y}$, $\times$, $+$, and $\mu$ in the Venn diagram)
-vary from 100 to 300, largely depending on how many constructors they
+vary from 100 to 250, largely depending on how many constructors they
 add to the inductive types.
@@
 
+\begin{itemize}
+\item 
+proofs similar to those in SF
+\item
+differences from SF:
+    \begin{itemize}
+    \item 
+    axiomatized equalities on subst
+    \item
+    inversion lemmas
+    \item
+    \end{itemize}
+\end{itemize}
+
 \bigskip

### 2022-11-03 coq-example.tex

--- 2022-11-02 coq-example.tex
@@
 Though Coq a la carte use small-step in one example} by exploiting
-extensible inductive types.
-Most of the proof and computation are directly adapted from the one in Software Foundation, resulting in a similarly accessible proof.
-However, we highlight some special expriences when using our plugin.
+extensible inductive types. Most of the proof and computation are
+directly adapted from the one in Software Foundation, resulting in a
+similarly accessible proof. However, we highlight some special
+experiences when using our plugin.
 
@@
 
-Our formulation for the base STLC family is around 400 LOCs; each families implementing a single feature takes 100\~250 LOCs.
+Our formulation for the base STLC family is around 400 LOCs; each family implementing a single feature takes 100\~250 LOCs.
 

### 2022-11-03 casestudies.tex
stlc case study

@@
 add to the inductive types.
+%
+The linguistic nature of our approach allows us to retain a programming
+style similar to the original proofs in Software Foundations.
 
@@
 %
-Individually developed language features can be easily composed (as
-mixins) to form new STLC variants.
-%
+Individually developed features can be easily composed (as mixins) to
+form new STLC variants.
+
 Composing features can lead to \emph{feature interactions}~\cite{batory2011feature}:
@@
 
-The linguistic nature of our approach allows us to retain a programming
-style similar to the original proofs while achieving extensibility.
-
-\begin{itemize}
-\item 
-proofs similar to those in SF
-\item
-differences from SF:
-    \begin{itemize}
-    \item 
-    axiomatized equalities on subst
-    \item
-    inversion lemmas
-    \item
-    \end{itemize}
-\end{itemize}
+The elimination of inductive types defined via \lsti{FInductive} is
+mostly via the \lsti{FRecursion} and \lsti{FInduction} commands.
+An exception is seemingly trivial ``inversion lemmas''.
+For example, consider the lemma \lsti!\forall t, \neg step tm_true t!
+stating that \lsti!tm_true! is irreducible.
+If \lsti{step} were an ordinary inductive type, then it could be proved in Coq
+with \lsti{intros t H; inversion H}.
+But since \lsti{step} is extensible, it seems that the lemma has to be
+proved by \lsti{FInduction} on \lsti{step}, and that the programmer
+has to manually verify that the lemma still holds in derived families
+that add new constructors to \lsti{step}.
+%
+However, we observe that it is lighter-weight to use overriding: 
+the programmer can specify that the proof of the lemma should be overridden
+in any derived family that extends \lsti{step} with new constructors,
+and in return, they are permitted to treat \lsti{step} as an ordinary
+inductive type in the proof and thus discharge it with \lsti{inversion},
+with the plugin automatically trying the same proof script in a derived
+family to override the proof.
 

### 2022-11-03 coq-example.tex

@@
 
-\begin{minted}[fontsize=\footnotesize]{Coq}
-  Lemma analyze_faithful : forall prog st end_st, 
-    interp prog st = end_st -> abs_interp prog (lift st) = (lift end_st)
-\end{minted}
+We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.
 
-We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.
 The first family "Lang" define the big step operator (using fuel) of a
-while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200 LOCs.
+"Imp" language, with side-effect-free expression "exp" and statements "stmt" defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200 LOCs.
 
 The second family "LangwAbs" define the abstract interpreter for "Lang" in "LangwAbs.AbsInterpreter".
-The idea of abstract interpretation is to lift concrete interpretation, their value and their states into their abstract correspondence.
-Then we come up with a faithful (one example would be "analyze_faithful") and generic abstract intepreter.
+The idea of abstract interpretation is to lift concrete interpretation, their value, their computation and their states into their abstract correspondence.
+Then we come up with a sound and generic abstract intepreter.
 Then, the user can freely choose the instantiation and lifting of this generic framework and end up with their aiming program analyzer.
@@
 
-Thus, to implement this generic abstract interpreter, we need to postulate the abstract value and their mathematical property.
-Then we need to prove the abstract interpreter is \textbf{sound and terminating}  with the help of these postulations.
+\begin{minted}[fontsize=\footnotesize]{Coq}
+  Field RExp : exp -> absExp -> Prop. (* lifting of expression *)
+  Field RState : state -> absState -> Prop. (* lifting of state *)
+  Field exp_analyze : exp -> absState -> absExp.
+  Field exp_analyze_sound : forall st ast,
+    RState st ast -> RExp (exp_eval e mem) (exp_analyze e amem)
+  Field analyze_sound : forall fuel st ast,
+    RState st ast -> RState (eval fuel stm st) (analyze stm ast)
+\end{minted}
+
+Here, to implement our generic abstract interpreter, we need to postulate the abstract value and their mathematical property.
+We let the user specify the ``lifting'' predicate, e.g., "RState st ast" means abstract state "ast" is the lifting of the concrete state "st".
+The concrete expression evaluator "exp_eval" is abstracted into "exp_analyze" with appropriate soundness.
+However, for each "stmt" that transform between states, their transfer functions need to be handled more delicately: for assignment, we have "aupdate"; for "while" we have "Afixf" calculating fixpoint of a given transfer function, and satisfying monotonicity; for "ifte" we have "join" with their property.
+With these transfer functions postulated, we can prove the abstract interpreter is \textbf{sound and terminating}  with the help of these postulations.
 In our implementation, this postulation happens in "absStateData" where we indicate the "lift" of concrete interpretation and the mathematical properties these lifted should satisfy.

### 2022-11-03 metatheory-condensed.tex
Save.

@@
 
+\begin{figure}[!htb]
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
+  }
+
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{3}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting}
+  Family STLC_bool extends STLC.
+
+   FInductive ty += ty_bool
+   FInductive tm += tm_true ..
+
+   FRecursion subst
+     (* Inherits all others *)
+
+     Case tm_true ...
+
+   End subst
+
+   Inherit t1
+
+  ...
+  \end{lstlisting}
+
+  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
+
+  \columnbreak
+  % \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+
+  i₀ = inh-id : Inh ⋅ {} {} 
+  i₁ = inh-ov i₀ (W τ₁') : Inh ⋅ {ty : 𝕊(W τ₁)} {ty : 𝕊(W τ₁')}
+  s₂' : {ty : 𝕊(W τ₁')} | [ty : 𝕌] 
+  i₂ = inh-ov i₁ (W τ₂') 
+     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂') }
+  (* recall s₃₁ : {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} 
+                | [ty : 𝕌;  ...] *)
+  s₃₁' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂')} | [ty : 𝕌; ...; tm_true : tm; ..]
+  ↑s₃ : Tm (⋅, [ty : 𝕌; ...; tm_true : tm; ..]) [ty : 𝕌; ...] (* the sealing before extension *)
+  ii₃₀₀ = inh-id 
+        : Inh (⋅, [ty : 𝕌; ...; tm_true : tm; ..])
+              (RecSig τ₂  tm)[p₀, ↑s₃]
+              (RecSig τ₂  tm)[p₀, ↑s₃]
+  ii₃₀₁ = inh-extend (inh-extend ii₃₀₀ {p₁} .. ) {p₁} ..
+        : Inh (⋅, [ty : 𝕌; tm :𝕌; tvar : nat → tm; ...; tm_true : tm]) (RecSig τ₂   tm)[p₀, ⇡s] (RecSig τ₂'  tm)
+  i₃₁ = inh+inh i₂ ↑s₃ ii₃₀₁ : 
+      Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂); subst' : 𝕃(RecSig τ₂ tm)} {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm)}
+
+  (* s₃' : {ty : 𝕎(τ₁'); tm : 𝕎(τ₂'); subst' : 𝕃(RecSig τ₂' tm) }
+          | [ty : 𝕌; ... ; tm : 𝕎(τ₂'); subst' : 𝕃(RecSig τ₂' tm)] *)
+  i₃ = inh-ov i₃₁ ..
+    : Inh ⋅ {ty : 𝕎(τ₁); tm : 𝕎(τ₂); subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm}  {ty : 𝕎(τ₁'); tm : 𝕎(τ₂'); subst' : 𝕃(RecSig τ₂' tm); subst : subst : tm → nat → tm → tm}
+  (* recall s₄ : {ty : 𝕎(τ₁); ... subst : tm → nat → tm → tm } 
+      | [ty : 𝕌; ... ; tm : 𝕌; ... ; subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm] *)
+  s₄' : {ty : 𝕎(τ₁'); ... subst : tm → nat → tm → tm } 
+        | [ty : 𝕌; ... ; tm : 𝕌; ... ; subst' : 𝕃(RecSig τ₂' tm); subst : tm → nat → tm → tm]
+  ↑s₄ : Tm (⋅, [ty : 𝕌; ... ; tm : 𝕌; ... ; subst' : 𝕃(RecSig τ₂' tm); subst : tm → nat → tm → tm]) [ty : 𝕌; ... ; tm : 𝕌; ... ; subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm]
+  i₄ = inh-inh i₃ ↑s₄ :
+    Inh ⋅ {ty : 𝕎(τ₁); ... subst : tm → nat → tm → tm; 𝕊(tm_var 0)} 
+          {ty : 𝕎(τ₁'); ... subst : tm → nat → tm → tm; 𝕊(tm_var 0)} 
+  \end{lstlisting}
+
+  \columnbreak
+
+  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+  \caption{Inheritance Metatheory}\label{fig:inheritance}
+  \end{figure}
+
 In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.

### 2022-11-03 metatheory-condensed.tex
Formatting Inheritance Example.

--- 2022-11-02 metatheory-condensed.tex
@@
 
+\begin{figure}[!htb]
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
+  }
+
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{3}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting}
+  Family STLC_bool extends STLC.
+
+   FInductive ty += ty_bool
+   FInductive tm += tm_true ..
+
+   FRecursion subst
+     (* Inherits all others *)
+
+     Case tm_true ...
+
+   End subst
+
+   Inherit t1
+
+  ...
+  \end{lstlisting}
+
+  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
+
+  \columnbreak
+  % \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+
+  i₀ = inh-id : Inh ⋅ {} {} 
+  i₁ = inh-ov i₀ (W τ₁') : Inh ⋅ {ty : 𝕊(W τ₁)} {ty : 𝕊(W τ₁')}
+  s₂' : {ty : 𝕊(W τ₁')} | [ty : 𝕌] 
+  i₂ = inh-ov i₁ (W τ₂') 
+     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂') }
+  (* recall s₃₁ : {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} 
+                | [ty : 𝕌;  ..] *)
+  s₃₁' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂')} | [ty : 𝕌; ..; tm_true : tm; ..]
+  ↑s₃ : Tm (⋅, [ty : 𝕌; ..; tm_true : tm; ..]) [ty : 𝕌; ..] 
+  (* ↑s₃ map (the type of) s₃₁' to the sealing before extension  s₃₁ *)
+  ii₃₀₀ = inh-id 
+        : Inh (⋅, [ty : 𝕌; ..; tm_true : tm; ..])
+              (RecSig τ₂  tm)[p₀, ↑s₃]
+              (RecSig τ₂  tm)[p₀, ↑s₃]
+  ii₃₀₁ = inh-extend (inh-extend ii₃₀₀ {p₁} .. ) {p₁} ..
+        : Inh (⋅, [ty : 𝕌; tm :𝕌; ..; tm_true : tm; ..]) 
+              (RecSig τ₂   tm)[p₀, ⇡s] (RecSig τ₂'  tm)
+  i₃₁ = inh+inh i₂ ↑s₃ ii₃₀₁ : 
+      Inh ⋅ {ty : 𝕊(W τ₁); ..; subst' : 𝕃(RecSig τ₂ tm)} 
+            {ty : 𝕊(W τ₁'); ..; subst' : 𝕃(RecSig τ₂' tm)}
+  (* s₃' : {ty : W(τ₁'); tm : W(τ₂'); subst' : 𝕃(RecSig τ₂' tm) }
+          | [ty : 𝕌; .. ; tm : W(τ₂'); subst' : 𝕃(RecSig τ₂' tm)] *)
+  i₃ = inh-ov i₃₁ ..
+     : Inh ⋅ {ty : W(τ₁); tm : W(τ₂); subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm}  
+             {ty : W(τ₁'); tm : W(τ₂'); subst' : 𝕃(RecSig τ₂' tm); subst : subst : tm → nat → tm → tm}
+  (* recall s₄ : {ty : W(τ₁); .. subst : .. } 
+      | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..] *)
+  s₄' : {ty : W(τ₁'); .. subst : .. } 
+        | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
+  ↑s₄ : Tm (⋅, [ty : 𝕌; .. ; tm : 𝕌; .. ; subst' : 𝕃(RecSig τ₂ tm);subst : ..]) 
+          [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
+  i₄ = inh-inh i₃ ↑s₄ :
+    Inh ⋅ {ty : W(τ₁); .. subst : ..; t1 : 𝕊(tm_var 0)} 
+          {ty : W(τ₁'); .. subst : ..; t1 : 𝕊(tm_var 0)} 
+  \end{lstlisting}
+
+  \columnbreak
+
+  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+  \caption{Inheritance Metatheory}\label{fig:inheritance}
+  \end{figure}
+
 In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.

### 2022-11-03 metatheory-condensed.tex
Fix subscript.

--- 2022-11-02 metatheory-condensed.tex
@@
 
+\begin{figure}[!htb]
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
+  }
+
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{3}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting}
+  Family STLC_bool extends STLC.
+
+   FInductive ty += ty_bool
+   FInductive tm += tm_true ..
+
+   FRecursion subst
+     (* Inherits all others *)
+
+     Case tm_true ...
+
+   End subst
+
+   Inherit t1
+
+  ...
+  \end{lstlisting}
+
+  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
+
+  \columnbreak
+  % \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+
+  i[@₀@] = inh-id : Inh ⋅ {} {} 
+  i[@₁@] = inh-ov i[@₀@] (W [@τ₁@]') : Inh ⋅ {ty : 𝕊(W τ₁)} {ty : 𝕊(W τ₁')}
+  s[@₂@]' : {ty : 𝕊(W τ₁')} | [ty : 𝕌] 
+  i[@₂@] = inh-ov i₁ (W τ₂') 
+     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂') }
+  (* recall s[@₃₁@] : {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} 
+                | [ty : 𝕌;  ..] *)
+  s[@₃₁@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂')} | [ty : 𝕌; ..; tm_true : tm; ..]
+  ↑s[@₃@] : Tm (⋅, [ty : 𝕌; ..; tm_true : tm; ..]) [ty : 𝕌; ..] 
+  (* ↑s[@₃@] map (the type of) s[@₃₁@]' to the sealing before inheritance s[@₃₁@] *)
+  ii[@₃₀₀@] = inh-id 
+           ⫶ Inh ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..])
+              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
+              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
+  ii[@₃₀₁@] = inh-extend (inh-extend ii[@₃₀₀@] {p[@₁@]} .. ) {p[@₁@]} ..
+        : Inh ( ⋅, [ty : 𝕌; tm :𝕌; ..; tm_true : tm; ..]) 
+              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]] (RecSig τ₂'  tm)
+  i[@₃₁@] = inh+inh i[@₂@] ↑s[@₃@] ii[@₃₀₁@] : 
+      Inh ⋅ {ty : 𝕊(W τ₁); ..; subst' : 𝕃(RecSig τ₂ tm)} 
+            {ty : 𝕊(W τ₁'); ..; subst' : 𝕃(RecSig τ₂' tm)}
+  s[@₃@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm) }
+      | [ty : 𝕌; .. ; tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm)] 
+  i[@₃@] = inh-ov i[@₃₁@] ..
+     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂); subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm}  
+             {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm); subst : subst : tm → nat → tm → tm}
+  (* recall s₄ : {ty : 𝕊(W τ₁); .. subst : .. } 
+      | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..] *)
+  s[@₄@]' : {ty : 𝕊(W τ₁'); .. subst : .. } 
+        | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
+  ↑s[@₄@] : Tm (⋅, [ty : 𝕌; .. ; tm : 𝕌; .. ; subst' : 𝕃(RecSig τ₂ tm);subst : ..]) 
+          [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
+  i[@₄@] = inh-inh i[@₃@] ↑s[@₄@] 
+         : Inh ⋅ {ty : 𝕊(W τ₁); .. subst : ..; t1 : 𝕊(tm_var 0)} 
+                 {ty : 𝕊(W τ₁'); .. subst : ..; t1 : 𝕊(tm_var 0)} 
+  \end{lstlisting}
+
+  \columnbreak
+
+  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+  \caption{Inheritance Metatheory}\label{fig:inheritance}
+  \end{figure}
+
 In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.

### 2022-11-03 metatheory-condensed.tex
Fix typo.

@@
   s[@₃₁@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂')} | [ty : 𝕌; ..; tm_true : tm; ..]
-  ↑s[@₃@] : Tm (⋅, [ty : 𝕌; ..; tm_true : tm; ..]) [ty : 𝕌; ..] 
-  (* ↑s[@₃@] map (the type of) s[@₃₁@]' to the sealing before inheritance s[@₃₁@] *)
+  ↑s[@₃@] : Tm ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..]) [ty : 𝕌; ..] 
+  (* ↑s[@₃@] map (the type of) s[@₃₁@]' to that of the sealing s[@₃₁@] before inheritance *)
   ii[@₃₀₀@] = inh-id 
@@
         | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
-  ↑s[@₄@] : Tm (⋅, [ty : 𝕌; .. ; tm : 𝕌; .. ; subst' : 𝕃(RecSig τ₂ tm);subst : ..]) 
+  ↑s[@₄@] : Tm ( ⋅, [ty : 𝕌; .. ; tm : 𝕌; .. ; subst' : 𝕃(RecSig τ₂ tm);subst : ..]) 
           [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]

### 2022-11-03 casestudies.tex
AI case study

@@
+%We report on our experience of applying the proposed design and implementation
+%to the mechanization of language metatheories.
+
 To demonstrate the applicability of our design and implementation, we
@@
 %
-The base STLC family consists of about 400 lines of code.
+The base STLC family consists of about 400~LOC.
 Lines of code in each of the four derived families
@@
 
-\bigskip
+\noindentparagraph{Abstract interpreters for imperative languages.}
 
-\cref{sec:coqexample-stlc}
-\cref{sec:coqexample-analysis}
-\cref{sec:coqexample-parser}
+Our second case study is a mechanization of abstract interpreters
+for simple imperative languages.
+In addition to the metatheories, this case study produces abstract
+interpreters that are directly runnable in Coq.
+
+The code is organized into four families.
+A base family, \lsti{Imp}, defines the abstract syntax of a while-language
+with pure expressions and impure statements, via \lsti{FInductive}.
+It also defines a big-step interpreter parameterized by a fuel value,
+via \lsti{FRecursion}.
+It contains \textasciitilde 200~LOC.
+
+A second family, \lsti{ImpGAI}, extends \lsti{Imp} and contains
+\textasciitilde 450~LOC.
+It exports a generic framework for deriving provably sound
+abstract interpreters.
+The soundness theorem of this generic abstract interpreter, \lsti{analyze},
+is with respect to the big-step interpreter, \lsti{eval}.
+It states that the abstraction relation \lsti{RState} over a
+concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
+by the analysis:
+
+\begin{centered}
+\begin{minipage}{.86\textwidth}
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+\forall stmt fuel S absS, RState S absS -> RState (eval fuel stmt S) (analyze stmt absS)
+\end{lstlisting}
+\end{minipage}
+\end{centered}
+
+\noindent
+\lsti{analyze} is defined via \lsti{FRecursion}, and the soundness theorem
+is proved via \lsti{FInduction}.
+This family leaves fields representing the abstract domain, the
+abstraction relation, monotonicity of transfer functions, etc.\ 
+unspecified or unproven---a derived family can instantiate these
+``parameters'' by overriding the fields (and also possibly extend the abstract syntax),
+to create a sound, runnable abstract interpreter for a (possibly extended) while-language.
+
+The other two families then extend \lsti{ImpGAI}.
+Family \lsti{ImpTyping} (\textasciitilde200~LOC) is an abstract
+interpreter that does type inference.
+Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
+with natural number arithmetic via \lsti{FInductive},
+and instantiates the generic abstract interpreter to perform constant propagation.
+%
+Gallina code generated for these two families are runnable;
+they are put to test over simple queries.
+
+We note that our plugin cannot yet allow the expression of two abstract
+interpreters nested within the same family.
+Work on nested inheritance~\cite{ncm2004,zm2017} points to a direction
+to further increase the expressive power of our language design.
+
+%\cref{sec:coqexample-stlc}
+%\cref{sec:coqexample-analysis}
+%\cref{sec:coqexample-parser}

### 2022-11-03 casestudies.tex
cite Types as AI

@@
 The soundness theorem of this generic abstract interpreter, \lsti{analyze},
-is with respect to the big-step interpreter, \lsti{eval}.
+is with respect to the big-step interpreter, \lsti{eval}, inherited from \lsti{Imp}.
 It states that the abstraction relation \lsti{RState} over a
@@
 Family \lsti{ImpTyping} (\textasciitilde200~LOC) is an abstract
-interpreter that does type inference.
+interpreter that does type inference~\cite{cousot1997types}.
 Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
-with natural number arithmetic via \lsti{FInductive},
+with natural number arithmetic,
 and instantiates the generic abstract interpreter to perform constant propagation.

### 2022-11-04 metatheory-condensed.tex
Update information on mechanization.

@@
 
-The definitive version is formalized in a logical framework with
-quotient inductive-inductive types~\cite{nordvall2010inductive,altkap2016},
-available as a supplemental file in Agda syntax.
+The definitive version is a \textbf{pen-and-paper} formalization assuming the facility of quotient inductive-inductive types~\cite{nordvall2010inductive,altkap2016} and extensional equality (equality reflection), available as a supplemental file in a \textbf{(pseudo-)Agda syntax}.
+Our formalization is directly following Agda syntax and type-checking (equality-checking) is done manually, leaving traces in the comments.
+We are currently working on mechanizing it and using Agda's rewriting mechanism to replace our limited application of equality reflection.
 %

### 2022-11-04 casestudies.tex
simplified inh fig for AI case study
mention extraction

@@
 
+\begin{wrapfigure}[13]{r}{0.13\textwidth}
+\includegraphics[scale=.68]{graphics/ai-casestudy.pdf}
+\end{wrapfigure}
+
 Our second case study is a mechanization of abstract interpreters
@@
 The other two families then extend \lsti{ImpGAI}.
-Family \lsti{ImpTyping} (\textasciitilde200~LOC) is an abstract
+Family \lsti{ImpTI} (\textasciitilde200~LOC) is an abstract
 interpreter that does type inference~\cite{cousot1997types}.
@@
 and instantiates the generic abstract interpreter to perform constant propagation.
-%
-Gallina code generated for these two families are runnable;
-they are put to test over simple queries.
+
+Our implementation of family polymorphism retains the computation
+ability of the host proof assistant;
+Gallina code generated for \lsti{ImpTI} and \lsti{ImpCP} are directly
+runnable in Coq.
+The two abstract interpreters can be extracted into OCaml;
+we put them to test over simple queries.
 

### 2022-11-04 cc-condensed.tex
Fix typo. Remove a seal
consistency

@@
 \noindentparagraph{Consistency and Canonicity.}
-Two of the most fundamental properties for dependent type theory are \textit{consistency} and \textit{canonicity}:
+Two of the most fundamental properties for dependent type theory are \textit{consistency} and \textit{canonicity}.
 \begin{theorem}[Consistency]
-  $\goodTerm{\cdot}{t}{\bot}$ is not derivable, i.e. $\bot$ is not inhabitted in the empty context
+  The typing judgment $\goodTerm{\cdot}{t}{\bot}$ is not derivable for any term~$t$.
 \end{theorem}
 \begin{theorem}[Canonicity]
-  if $\goodTerm{\cdot}{t}{\cB}$ then it is either $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$
+  If $\goodTerm{\cdot}{t}{\cB}$, then either $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$.
 \end{theorem}
-Consistency means our dependent type theory suffices as a logical system as not every proposition is trivially provable.
-Canonicity implies our dependent type theory sufficies as ``a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
+
+Consistency says that the type $\bot$ is uninhabitable.
+Thus, it is safe to use the type theory for logical reasoning, as not
+every proposition is trivially provable.
+Consistency is a consequence of a meta\-circular interpretation of all
+the well-formedness rules, which, in particular, interprets $\bot$ in
+\TT as an uninhabited type in the meta\-language.
+The construction of this interpretation follows the standard model of
+\citet{altkap2016}, extending it to handle linkages.
+We refer the reader to a supplemental Agda file for the complete proof.
+
+Canonicity implies our dependent type theory suffices as ``a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
 We can even argue that, if this canonicity theorem is proven in a computable meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}},
+
+\newpage
+
+\ifShowOldWriting
+
 then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.
@@
 
+\fi
+
 \noindentparagraph{Canonicity} is proved by following the canonicity argument from~\citet{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model.

### 2022-11-04 metatheory-condensed.tex
Phrasing
consistency

@@
 Our formalization is directly following Agda syntax and type-checking (equality-checking) is done manually, leaving traces in the comments.
-We are currently working on mechanizing it and using Agda's rewriting mechanism to replace our limited application of equality reflection.
+We are currently working on mechanizing it and relying on Agda's rewriting mechanism to replace our limited usage of equality reflection.
 %
@@
 %
-Another supplemental appendix gives a larger-scale example than
+Another supplemental appendix gives a more complete example than
 \cref{fig:stlc-linkage-typing} showing how families are modeled and

### 2022-11-04 casestudies.tex
edit on QIIT para

@@
 ability of the host proof assistant;
-Gallina code generated for \lsti{ImpTI} and \lsti{ImpCP} are directly
-runnable in Coq.
-The two abstract interpreters can be extracted into OCaml;
-we put them to test over simple queries.
+the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
+directly runnable in Coq.
+We also extract them to OCaml and put them to test over simple queries.
 

### 2022-11-04 metatheory-condensed.tex
edit on QIIT para

--- 2022-11-03 metatheory-condensed.tex
@@
 
-The definitive version is formalized in a logical framework with
-quotient inductive-inductive types~\cite{nordvall2010inductive,altkap2016},
-available as a supplemental file in Agda syntax.
+The definitive version containing all the rules in \TT is stated in
+terms of a meta\-language supporting quotient inductive-inductive types
+(QIITs)~\cite{nordvall2010inductive,altkap2016}.
+QIITs support equality constructors, which facilitate the expression of
+conversion rules (e.g., \ruleref{tmeq/sub/id}).
+The formalization is available as a supplemental file in Agda syntax,
+though it is not checked by Agda because Agda does not support QIITs
+natively.
+An effort to mechanize the formalization is ongoing, using Agda's
+\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
+%a \textbf{pen-and-paper} formalization
+%assuming the facility of quotient inductive-inductive
+%types~\cite{nordvall2010inductive,altkap2016} and extensional equality
+%(equality reflection), available as a supplemental file in a
+%\textbf{(pseudo-)Agda syntax}. Our formalization is directly following
+%Agda syntax and type-checking (equality-checking) is done manually,
+%leaving traces in the comments. We are currently working on mechanizing
+%it and relying on Agda's rewriting mechanism to replace our limited
+%usage of equality reflection.
+
+We also typeset a more complete set of rules than the space allows and
+present it in a supplemental appendix.
 %
-We also typeset a more complete set of rules than
-the space allows and present it in a supplemental appendix.
-%
-Another supplemental appendix gives a larger-scale example than
+Another supplemental appendix gives a more complete example than
 \cref{fig:stlc-linkage-typing} showing how families are modeled and

### 2022-11-04 cc-condensed.tex
canonicity

@@
-% Canonicity / Consistency Condensed
-\noindentparagraph{Consistency and Canonicity.}
-Two of the most fundamental properties for dependent type theory are \textit{consistency} and \textit{canonicity}.
+\noindentparagraph{Consistency.}
+One of the most fundamental properties of a dependent type theory is consistency.
 \begin{theorem}[Consistency]
+\label{thm:consistency}
   The typing judgment $\goodTerm{\cdot}{t}{\bot}$ is not derivable for any term~$t$.
 \end{theorem}
-\begin{theorem}[Canonicity]
-  If $\goodTerm{\cdot}{t}{\cB}$, then either $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$.
-\end{theorem}
 
-Consistency says that the type $\bot$ is uninhabitable.
+\noindent
+Consistency says that the type $\bot$ is not inhabited.
 Thus, it is safe to use the type theory for logical reasoning, as not
 every proposition is trivially provable.
-Consistency is a consequence of a meta\-circular interpretation of all
+\cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
 the well-formedness rules, which, in particular, interprets $\bot$ in
@@
 \citet{altkap2016}, extending it to handle linkages.
-We refer the reader to a supplemental Agda file for the complete proof.
+We refer the reader to a supplemental file (in Agda syntax) for the complete proof.
 
-Canonicity implies our dependent type theory suffices as ``a computational foundation for mathematics''\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}.
-We can even argue that, if this canonicity theorem is proven in a computable meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}},
+\noindentparagraph{Canonicity.}
 
-\newpage
+A second property we prove is canonicity.
 
-\ifShowOldWriting
+\begin{theorem}[Canonicity]
+\label{thm:canonicity}
+  If $\goodTerm{\cdot}{t}{\cB}$, then either $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$.
+\end{theorem}
 
-then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.
+\noindent
+This canonicity theorem says that every closed term of the ground
+type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
+When proved in a constructive meta\-logic, the proof of canonicity
+amounts to a normalization function for closed terms of the ground type.
+So canonicity serves to justify the computational nature of the type theory.
 
-We refer to the attached (pseudo-)Agda style code for \textbf{the complete consistency and canonicity proof}. But here we sketch the big picture.
-
-\noindentparagraph{Consistency} is proved by extending the standard model from~\citet{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization.
-But if we were using naive set theory: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma for each rule.
-The main goal is to interpret $\denotesS{\bot} = \emptyset$ which implies it is impossible $\goodTermS{\cdot}{t}{\bot}$ and thus $\bot$ is syntactically non-derivable.\footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo-)Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation closer to interpretation in the paper}
-
-We only need to take care of the new feature for our consistency model.
-But (1) since we have used sigma type to translate the linkage, using the model for sigma type for linkage will work; (2) directly using the inductive facility of (pseudo)-Agda (or naive set theory) to for W-type is enough due to the similarity of our W-type to the conventional W-type.
-
-Here we will only show some interpretation and also omit the universe level to sketch out the idea.
-
-\definecolor{new}{HTML}{F5EFE6}
-
-\begin{align*}
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod1}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodCtx{\Gm}{i}} & \text{ is a Set and we define } \goodCtxS{\Gm}{i} :\iff \denotesS{\goodCtx{\Gm}{i}} \text{ is defined}\tikzmarkend{stmod1} 
-  \\
-  \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \Gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\Gm)\}
-  \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod2}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodSub{\Gm}{\_}{\Delta}} & \text{ is a Set and we define }  {\goodSubS{\Gm}{\Gm}{\Delta}} : \iff \denotesS{\Gm} \in \denotesS{\goodSub{\Gm}{\_}{\Delta}} \tikzmarkend{stmod2}
-  \\
-  \denotesS{\goodSub{\Gm}{\_}{\Delta}} & = \denotesS{\goodCtx{\Gm}{i}} \rightarrow \denotesS{\goodCtx{\Delta}{j}} \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod3}(0.05,-0.15)(-0.10,0.30) \denotesS{\goodTerm{\Gm}{\_}{T}} & \text{ is a Set (of dependent function) } \\& \text{and we define }  {\goodTermS{\Gm}{t}{T}} : \iff \denotesS{t} \in \denotesS{\goodTerm{\Gm}{\_}{T}}
-  \tikzmarkend{stmod3}
-  \\
-  \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\Gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\Gm) \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod4}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gm}{T}{i}} & \text{ is a family of sets indexed by } \denotesS{\goodCtx{\Gm}{i}}  
-  \\
-  & \text{ and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesS{\goodType{\Gm}{T}{i}} \text{ is defined } \tikzmarkend{stmod4} 
-  \\
-  \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\Gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod5}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigS{\Gm}{\sigma}{n}} :\iff \denotesS{\sigma} \in  \denotesS{\goodSig{\Gm}{\_}{n}} \tikzmarkend{stmod5} 
-  \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod6}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod6} \\   
-  \denotesS{\goodSig{\Gm}{\_}{n+1}} &= \{
-    (\denotesS{\sigma}, \denotesS{A}, \denotesS{f}, \denotesS{T}) :
-      {\goodSigS{\Gm}{\sigma}{n}}
-      \land  {\goodTypeS{\Gm}{A}{}} \\  
-      & \quad \quad \quad \land  {\goodSealS{\Gm}{f}{\sigma}{A}}
-      \land  {\goodTypeS{\Gm, A}{T}{i}} 
-  \} \\ 
-  \denotesS{\goodSig{\Gm}{\_}{0}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gm}{\TyTkg {\tau}}{}} \denotesS{T["sf" \ f]} \\
-  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \denotesS{f} , \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{i}}&= \denotesS{\top}\quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod7}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\Gm) \land t \in \prod_{m' \in \denotesS{A}(\Gm)}\denotesS{T}(\Gm, m')  \}  \\
-  & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \denotesS{\goodSig{\Gm}{\LSigEmp}{0}} &= \star \\
-  \denotesS{\goodSig{\Gm}{\LSigAdd{\sigma}{s}{T}}{n+1}} &= (\sigma, A, f, T) \\
-  \denotesS{\goodTerm{\Gm}{\LkgEmp}{\cL \LSigEmp} } (\Gm) &= \star \\
-  \denotesS{\goodTerm{\Gm}{\LkgAdd{o}{t}}{\cL(\LSigAdd {\sigma}{s}  {T})}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
-  \denotesS{\goodTerm{\Gm}{\lkgproj{1}{t}}{\cL(\lsigproj{1}{\sigma})}}(\Gm) &=  "fst"~(t\Gm) \\ 
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}} & \text{ is a dependent function, defined inductively on } {\sigma} 
-  \tikzmarkend{stmod8}
-  \\
-  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
-  &= \denotesS{{\Tkg {\lkgproj{1}{o}}}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \Tkg {\lkgproj{1} {o}})]} \\
-  & \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
-  \denotesS{\goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\sigma}}}
-  &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \denotesS{\goodType{\Gm}{\bot}{i}} \ (\Gm) &= \emptyset \\
-  \denotesS{\goodType{\Gm}{\cB}{i}} \ (\Gm) &= \{0, 1\} \quad \text{a specific two element set} \\
-  \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod9}(0.05,-0.15)(-0.10,0.30)
-  \denotesS{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
-  & \text{ and we define } \goodWSigS{\Gm}{\tau}{n} :\iff \denotesS{\goodWSig{\Gm}{\tau}{n}} \in \denotesS{\goodWSig{\Gm}{\_}{n}}
-  \tikzmarkend{stmod9}
-  \\
-  \denotesS{\goodWSig{\Gm}{\WSigEmp}{0}} &= [] \quad \text{ an empty list} \\ 
-  \denotesS{\goodWSig{\Gm}{(\WSigAdd{\tau}{A}{B})}{n+1}}
-  &= (\denotesS{A}, \denotesS{B})~"::"~\denotesS{\goodWSig{\Gm}{\tau}{n}} \\ 
-  \denotesS{\goodType{\Gm}{\wsigproj{j}{1}{\tau}}{}} &= "fst"~({\denotesS{\tau}}_{j})\\
-  \denotesS{\goodType{\Gm, \wsigproj{j}{1}{\tau}}{\wsigproj{j}{2}{\tau}}{}} &= "snd"~({\denotesS{\tau}}_j)\\
-\end{align*}
-
-Here, we need to clarify some ambiguity in our latex formulation.
-For example, when we wrote ``$\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$ defined inductively on ${\sigma}$'', the best way to understand it is to consider we are defining a function $\denotesS{\goodType{\Gm}{\TyTkg {\_}}{}}$ as the interpretation for $\goodType{\Gm}{\TyTkg {\_}}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gm}{\TyTkg {\sigma}}{}}$, inductively on the length of the signature.
-The case is similar for $\denotesS{\cL \_}$, $\denotesS{\Tkg {\_}}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
-\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\TyTkg{} : "Sig"~\Gm~n \to "Ty"~\Gm$ as a constructor and our model needs to interpret this constructor to a function $\TyTkg{}^S : "Sig"^S \Gm^S~n \to "Ty"^S~\Gm^S$ with each judgement also interpreted.
-
-Basically in Agda, we are simply constructing a function that mapping our QIIT data-type(stands for syntax) to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
-What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-definedness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
-Note that $\denotesS{\goodSig{\Gm}{s}{n}}$ and $\denotesS{\goodType{\Gm}{\TyTkg{\sigma}}{i}}$ is mutual recursively defined together---because \\ $\goodSealS{\Gm}{f}{\sigma}{A}$ relies on $\TyTkg{}$.
-
-Finally, notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
-and that is a contradiction.
-
-\fi
-
-\noindentparagraph{Canonicity} is proved by following the canonicity argument from~\citet{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model.
-Canonicity is also called \textit{reducibility}.
-
-Now we sketch out the big picture of the proof.
-The canonicity model $"Tm"^C~\Gamma^C~"T"^C$ will map each syntax piece $("t : Tm Γ T") \mapsto "(t, tₚ) :" "Tm"^C~\Gamma^C~"T"^C = \sum~("t : Tm Γ T")."Tm"_p~\Gamma^C~"T"^C~"t"$ to a dependent pair\footnote{This $"Tm"_p$ is usually called dependent model.}, of which the first part will be the same as the input syntax.
-Using this mapping on closed boolean term $"t : Tm ⋅ "\mathbb{B}$ will give us a proof of $"Tm"_p~\cdot^C~\mathbb{B}^C~"t"$, which will unfolds to our final goal $"t" \equiv "tt" + "t" \equiv "ff"$.
-Roughly, this can read as ``every closed boolean term is reducible''.
-Thus once we can ``propagate'' $(\cdot)^C$ (i.e.
-construct this model completely) for every syntactic piece, we are done.
-In our (pseudo) Agda formulation, we use $(\cdot)_2$ to replace the notation of $(\cdot)^C$ because we consider most parts of this model are actually a dependent pair.
-
-Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as ``a reducible set'': $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
-This reformulation can be understood as an unfolding of the \textit{type-theoretic logical relation proof}.\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory} If we use the terminology of the conventional formulation, this reducible set means if we apply its element $t$ with a substitution $\gamma$ from the set $\Gamma$ (where we have $\Gamma_p$ witnesses the ``reducibility'' of $\Gamma$), then we will have $t[\gamma] \in T_p$ as a closed and \textbf{reducible} term.
-Now if $\Gamma$ is actually empty $\cdot$ and $T$ is $\cB$, we are done.
-
-Again, using the same insight from earlier, we reuse the model for sigma type and W-type for linkages and our inductive type.
+We prove \cref{thm:canonicity} by constructing a logical-relations model
+for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
+In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
+as a logical predicate on closed terms: the predicate includes all
+closed, reducible terms of type~$T$.
+The proof is available as a supplemental file in Agda syntax.

### 2022-11-04 cc-condensed.tex
Add one sentence for metatheory.
edit

@@
 \citet{altkap2016}, extending it to handle linkages.
+Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products,
+as rule \ruleref{l/add} indicates.
+%
 We refer the reader to a supplemental file (in Agda syntax) for the complete proof.
@@
 In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
-as a logical predicate on closed terms: the predicate includes all
+into a logical predicate on closed terms: the predicate includes all
 closed, reducible terms of type~$T$.

### 2022-11-04 lang-design.tex
move a comment

@@
 We expect this feature to be used only occasionally.
-We give an example in \cref{sec:coqexample}.
+Our case studies contain uses of overriding (\cref{sec:casestudies}).
 
@@
 
-\noindentparagraph{Composing families.} \EDJ{I feel this paragraph will cause more confusion---because it is inheritance judgement/linkage transformer being mixed in, not family itself.
-Conceptually not distinguishing family and linkage transfomer will make reader wonder what this magic black box here actually is}
+\noindentparagraph{Composing families.}
 
@@
 As a result, \lsti{STLCFixIsorec.typesafe} is readily the proof of the type-safety
-theorem of an STLC equipped with fixpoints and iso-recursive types.
+theorem of an STLC equipped with fixpoints and iso-recursive types.\EDJ{I feel this paragraph will cause more confusion---because it is inheritance judgement/linkage transformer being mixed in, not family itself.
+Conceptually not distinguishing family and linkage transfomer will make reader wonder what this magic black box here actually is}
 

### 2022-11-05 lang-design.tex
trying out a name for the language design and impl

@@
-We use a plugin to the Coq proof assistant to demonstrate the key
-ingredients of our design,
-though we believe our design could be adapted to other proof assistants such as Lean.
+We present the key ingredients of our design as an extension to the Coq
+proof assistant,
+though we believe it could be adapted to other proof assistants such as Lean.
+We call our design and implementation \Lang (\underline{f}amily
+\underline{\smash{po}}lymorphism for a \underline{\smash{p}}roof assistant).
+In this section, we focus on the language design of \Lang.
+\cref{sec:coqimpl} describes its implementation as a Coq plugin.
 
@@
 \cref{fig:stlc-mechanized} shows how STLC and its extension with
-fixpoints can be mechanized using our plugin, in a style envisioned in
+fixpoints can be mechanized using \Lang, in a style envisioned in
 \cref{fig:stlc-nonmechanized}.
@@
 %
-Based on this insight, our plugin automatically generates
+Based on this insight, \Lang automatically generates
 propositional equalities for each recursive function defined with \lsti{FRecursion},
 making them available as axioms for use by the rest of the current family.
-The plugin also provides a tactic \lsti{fsimpl} that enables, for instance,
+\Lang also provides a tactic \lsti{fsimpl} that enables, for instance,
 simplifying \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
@@
 Coq distinguishes \emph{opaque} definitions from \emph{transparent} ones.
-Our plugin supports the overriding of opaque definitions, which
+\Lang supports the overriding of opaque definitions, which
 include most proofs.
@@
 transparent---that is, they are not defined with \lsti{Qed}.
-Our plugin treats transparent definitions as non-overridable by default.
+\Lang treats transparent definitions as non-overridable by default.
 Thus, the definitional equalities on "env", "empty", and "steps",
@@
 
-Our plugin does allow the overriding of transparent definitions that are
+\Lang does allow the overriding of transparent definitions that are
 explicitly marked as \lsti{Overridable} by the programmer.
@@
 name conflicts in general.
-Our plugin requires the programmer to resolve conflicts by
+\Lang requires the programmer to resolve conflicts by
 explicitly overriding conflicted fields.
@@
 To provide a streamlined programming experience (\ref{chg:tooling}),
-our plugin offers two tactics, \lsti{finjection} and \lsti{fdiscriminate}.
+\Lang offers two tactics, \lsti{finjection} and \lsti{fdiscriminate}.
 For example, in a proof state that contains a manifestly false assumption

### 2022-11-05 coqimpl.tex
coqimpl

--- 2022-10-09 coqimpl.tex
@@
-% \subsection{Language Design}
-
-% \subsection{Plugin Implementation}
+\input{stlc-compiled.tex}
+\input{stlcfix-compiled.tex}
+
+We implement our language design as a Coq plugin.
+Rather than modifying the Coq kernel, a prototypical implementation as a
+plugin allows a clearly defined trusted base and
+improved compatibility with different versions of Coq.
+
+The implementation works by translating programs in \Lang syntax into
+programs that can be checked and evaluated by Coq.
+The translation is compatible with interactive theorem proving, in that
+a family is translated piece by piece, allowing each field to be defined
+and checked separately.
+The translation is modular and efficient, in that
+code inherited from a base family need not be retranslated or rechecked.
+
+\noindentparagraph{Explicit self parameterization.}
+
+The spirit of the translation is to take ``family polymorphism''
+literally: every field is translated into a Coq definition that is
+polymorphic to (i.e., universally quantified over) a representation of
+its enclosing family.
+While this universal quantification has been implicit with the \Lang
+syntax, it has to be made explicit in the translated Coq code.
+
+\cref{fig:stlc-compiled,fig:stlcfix-compiled} illustrate the translation of the
+\lsti{STLC} and \lsti{STLCFix} families from \cref{fig:stlc-mechanized}.
+Fields of a family are translated into parameterized Coq modules
+(or parameterized module types).
+
+For instance, consider field \lsti{env} in family \lsti{STLC}.
+It is translated into a top-level module named \lsti{STLC\_env}.
+This module has a parameter called \lsti{self} representing the enclosing
+family: fields of the current family in the context of \lsti{env} can be
+referenced through this \lsti{self}.
+In particular, \lsti{env} is defined as \lsti{id -> option ty},
+where \lsti{ty} is a late-bound reference to the \lsti{ty} field of the
+enclosing family.
+Hence, this reference to \lsti{ty} is translated in \lsti{STLC\_env} to \lsti{self.ty},
+which is manifestly polymorphic to the enclosing family.
+This translation of the \lsti{env} field can be shared
+with a derived family even if it extends \lsti{ty} (e.g., \lsti{STLCProd})---%
+no recompilation is needed because \lsti{self.ty} is not tied to a
+particular definition of \lsti{ty}.
+
+The type of \lsti{STLC\_env}'s \lsti{self} parameter is \lsti{STLC\_env\_Ctx},
+a module type constructed from the translation of \lsti{env}'s preceding
+field, \lsti{STLC\_ty}, and its context \lsti{STLC\_ty\_Ctx}.
+In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is
+constructed from the translation of \lsti{ty}'s preceding field,
+\lsti{STLC\_subst}, and its context \lsti{STLC\_subst\_Ctx}.
+Thus, the \lsti{self} parameter can be used to reference those and only
+those fields in the current field's typing context, which echoes the
+discussion in \cref{sec:sound}.
+
+\noindentparagraph{Translating extensible inductive types.}
+
+An \lsti{FInductive} definition is translated to a parameterized module type.
+Consider the inductive type \lsti{tm}.
+In \lsti{STLC}, it is translated to a top-level module type
+\lsti{STLC\_tm} that declares a \lsti{tm} type,
+its constructors (\lsti{tm_unit} etc.),
+its partial recursor (\lsti{tm_prect_STLC}),
+and the propositional equalities on the partial recursor (\lsti{tm_unit_eq_STLC} etc.).
+
+Importantly,
+\lsti{STLC\_tm} merely declares the existence of these names and their types;
+it does not specify their definitions or proofs.
+Knowing the existence of these names and their types
+suffices to type-check and translate the rest of the family.
+Leaving the definitions and proofs open enables \lsti{STLC} and \lsti{STLCFix}
+to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and \lsti{End}~\lsti{STLCFix}.
+
+The command \lsti{FInductive tm : Set += tm_fix : }\dadada~in family
+\lsti{STLCFix} is again translated to a module type \lsti{STLCFix\_tm} (\cref{fig:stlcfix-compiled}).
+It includes all names declared by \lsti{STLC\_tm}, and additionally declares
+the new constructor \lsti{tm_fix} and related equalities.
+
+\noindentparagraph{Translating recursion and induction.}
+
+An \lsti{FRecursion} definition is translated in two parts:
+first a module containing the definitions of all the case handlers,
+and then a module type declaring the existence of the recursive function as well as
+its propositional equalities.
+
+Consider the translation of \lsti{subst} in family \lsti{STLC}.
+First, a module named \lsti{STLC\_subst\_Cases} is generated interactively:
+every time the programmer completes a \lsti{Case}, a case handler (e.g., \lsti{Def subst\_tm_unit})
+is generated and added to the module.
+
+Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
+%
+As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
+so its definition is not exposed to the succeeding fields.
+Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
+and the existence of its propositional equalities, leaving \lsti{subst} undefined
+and the equalities unproven.
+%
+The propositional equalities are stated in terms of the case handlers,
+whose definitions \emph{are} available through the \lsti{self} parameter.
+For example, the type of \lsti{subst_tm_unit_eq} can be simplified to
+\lsti{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
+The equalities will be available for use in the checking and translation of
+the succeeding fields through their \lsti{self} parameters.
+
+Importantly, code generated for the case handlers is shared with derived families.
+In \cref{fig:stlcfix-compiled}, \lsti{STLCFix\_subst\_Cases}
+reuses---without rechecking---all the case handlers in
+\lsti{STLC\_subst\_Cases} via the command \lsti{Include STLC\_subst\_Cases(self)}.
+
+The translation of \lsti{FInduction} is similar, except that there is no need to
+register propositional equalities, as \lsti{FInduction} proofs are considered opaque.
+
+\noindentparagraph{Translation of further-bindables vs.\ non-further-bindables.}
+
+In family \lsti{STLC},
+field \lsti{env} and the case handlers for \lsti{subst}
+are not further-bindable by derived families.
+In contrast, \lsti{tm}, \lsti{subst}, and the related equalities
+can be further bound.
+The distinction is reflected in the translations.
+Further-bindable fields are translated to module types that export only types of the fields.
+Non-further-bindable fields are translated to modules that export definitional equalities
+on the fields.
+
+%\footnotetext{%
+%Axioms in Coq module types are more like parameters than real axioms.
+%They become part of the signature described by the module type.
+%}
+
+\noindentparagraph{Eliminating \lsti{self} by aggregation.}
+
+Upon the conclusion of a family definition (e.g., \lsti{End}~\lsti{STLCFix}),
+\lsti{self} parameters in the translations of the fields must be instantiated,
+so that references like \lsti{STLCFix.typesafe} can be translated.
+This is done by assembling the \lsti{self}-parameterized pieces into a single module.
+%
+For example, module \lsti{STLC} in \cref{fig:stlc-compiled} is generated
+upon \lsti{End}~\lsti{STLCFix}.
+All the fields are added to this module in the same order as they appear in the family definition.
+
+For the further-bindables, \lsti{Axiom}s declared in the module types must be instantiated.
+%
+\begin{itemize}[align=left,itemsep=5pt,labelsep=*,leftmargin=*]
+
+\item
+For \lsti{tm},
+an inductive type \lsti{tm\_} is generated, with \lsti{tm\_}
+instantiating the axiomatized \lsti{tm} type, and with
+its constructors (\lsti{tm_unit\_} etc.) instantiating 
+the axiomatized constructors (\lsti{tm_unit} etc.).
+%
+The partial recursor \lsti{tm_prect_STLC} is instantiated
+%to \lsti{\lambda}\,\lsti{P, tm\__rect (\lambda}\,\lsti{t, option (P t))},
+with the help of \lsti{tm\_rect}, the recursor of the generated \lsti{tm\_} type.
+Proving the equalities on \lsti{tm_prect_STLC} is then trivial, by \lsti{reflexivity}.
+
+\item
+Similarly, \lsti{subst} is instantiated by applying the recursor \lsti{tm\_rect}
+to the (already included) case handlers (\lsti{subst\_tm_unit}, etc.).
+Propositional equalities on \lsti{subst} are then trivial, by \lsti{reflexivity}.
+
+\end{itemize}
+
+For the non-further-bindables, the translated modules are directly included
+(e.g., \lsti{Include STLC\_subst\_Cases} and \lsti{Include STLC\_env}).
+%
+The instantiations of \lsti{self}s are implicit thanks to a Coq nicety:
+when including a higher-order module, Coq automatically instantiates its
+parameter with the current interactive module environment.
+For instance, command \lsti{Include STLC\_subst\_Cases} is successfully executed,
+because Coq automatically instantiates \lsti{self} using the current
+module environment, which contains all the fields required by \lsti{STLC\_subst\_Cases\_Ctx}.
+
+Module \lsti{STLCFix} in \cref{fig:stlcfix-compiled} is generated upon
+\lsti{End}~\lsti{STLCFix}, in the same fashion described above for \lsti{STLC}.
+The translation makes code and proof sharing evident.
+For example, the four case handlers for \lsti{subst} compiled for \lsti{STLC}
+are reused by module \lsti{STLCFix}. So is the proof of \lsti{typesafe}.
+
+Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)
+can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a module),
+as the last line of \cref{fig:stlcfix-compiled} shows.
+
+\ifShowOldWriting
+
+\newpage
 

### 2022-11-05 stlc-compiled.tex
coqimpl

@@
 [@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in family \texttt{STLC}}@][@\tanc{c-STLC-subst-start}@]
-[@\outdent@]Module Type STLC\_subst\_Ctx.
+[@\outdent@]Module Type STLC\_subst\_Cases\_Ctx.
 Include STLC\_tm\_Ctx.  Include STLC\_tm.
-[@\outdent@]End STLC\_subst\_Ctx.[@\innersep@]
-[@\outdent@]Module STLC\_subst\_Cases (self : STLC\_subst\_Ctx).
+[@\outdent@]End STLC\_subst\_Cases\_Ctx.[@\innersep@]
+[@\outdent@]Module [@\textls[-30]{STLC}@]\_[@\textls[-30]{subst}@]\_[@\textls[-30]{Cases}@] (self[@\,@]:[@\,@][@\textls[-30]{STLC}@]\_[@\textls[-30]{subst}@]\_[@\textls[-30]{Cases}@]\_[@\textls[-30]{Ctx}@]).
 Def subst\_tm_unit :=
@@
 [@\outdent@]End STLC\_subst\_Cases.[@\innersep@]
+[@\outdent@]Module Type STLC\_subst\_Ctx.
+Include STLC\_subst\_Cases\_Ctx.
+Include STLC\_subst\_Cases.
+[@\outdent@]End STLC\_subst\_Ctx.[@\innersep@]
 [@\outdent@]Module Type STLC\_subst (self : STLC\_subst\_Ctx).
 Axiom subst : self.tm[@\,@]->[@\,@]id[@\,@]->[@\,@]self.tm[@\,@]->[@\,@]self.tm.
-Axiom subst_tm_unit_eq : \forall x t,
-  self.subst (self.tm_unit) x t = self.tm_unit.
+Axiom subst_tm_unit_eq :
+  \forall x t, self.subst (self.tm_unit) x t =
+  self.subst\_tm_unit x t.
 Axiom subst_tm_var_eq : [@\dadada@].
@@
 Axiom subst_tm_app_eq : [@\dadada@].
-[@\outdent@]End STLC\_subst.[@\tanc{c-STLC-subst-end}@][@\outersep@]
-[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{ty} in family \texttt{STLC}}@] [@\dadada@][@\tanc{c-STLC-ty-start}@][@\tanc{c-STLC-ty-end}@][@\outersep@]
+[@\outdent@]End STLC\_subst.[@\tanc{c-STLC-subst-end}@][@\columnbreak@]
+[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{ty} in family \texttt{STLC}}@] [@\tanc{c-STLC-ty-start}@]
+[@\dadada@][@\tanc{c-STLC-ty-end}@][@\outersep@]
 [@\outdent@][@\codecomment{Code emitted upon definition of \texttt{env} in family \texttt{STLC}}@][@\tanc{c-STLC-env-start}@]
@@
 [@\outdent@]End STLC\_env.[@\tanc{c-STLC-env-end}@][@\outersep@]
-[@\outdent@][@\codecomment{Code emitted for other fields defined in family \texttt{STLC}}@] [@\dadada@][@\tanc{c-STLC-rest}@][@\outersep@]
-[@\outdent@][@\codecomment{Code emitted upon completion of family \texttt{STLC}}@][@\tanc{c-STLC-start}@]
+[@\outdent@][@\codecomment{Code emitted for other fields defined in family \texttt{STLC}}@][@\tanc{c-STLC-rest-start}@]
+[@\dadada@][@\tanc{c-STLC-rest-end}@][@\outersep@]
+[@\outdent@][@\codecomment{Code emitted upon conclusion of family \texttt{STLC}}@][@\tanc{c-STLC-start}@]
 [@\outdent@]Module STLC.[@\innersep@]
@@
 | tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_.
-Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
+[@\codecomment{Instantiate \texttt{tm} \& its constructors}@]
+Def tm := tm\_.
 Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
 Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
-[@\hfill\codecomment{Instantiate \texttt{tm} partial recursor \& its equalities}@]
-Def tm_prect_STLC P := tm\__rect (\lambda[@\,@]t,[@\,@]option[@\,@](P[@\,@]t)).
+[@\codecomment{Instantiate \texttt{tm} partial recursor \& its equalities}@]
+Def tm_prect_STLC := \lambda P, tm\__rect (\lambda[@\,@]t, option (P t)).
 Fact[@\;@][@\textls[-20]{tm\_unit\_eq\_STLC}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
 Include STLC\_subst\_Cases.[@\tanc{c-STLC-d-subst-start}@]
-Def subst[@\,@] :=[@\hfill\codecomment{Instantiate \texttt{subst} \& its equalities}@]
-  tm\__rect _ subst\_tm_unit subst\_tm_var
+[@\codecomment{Instantiate \texttt{subst} \& its equalities}@]
+Def subst[@\,@] := tm\__rect _ subst\_tm_unit subst\_tm_var
                         subst\_tm_abs  subst\_tm_app.
-Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}\columnbreak@][@\innersep@]
-[@\dadada@][@\hfill\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@][@\tanc{c-STLC-d-ty}@][@\tanc{c-STLC-sndcol-start}@][@\innersep@]
+Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}@][@\innersep@]
+[@\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@] [@\dadada@][@\tanc{c-STLC-d-ty}@][@\tanc{c-STLC-sndcol-start}@][@\innersep@]
 Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLC-d-env}@][@\innersep@]
-[@\dadada@][@\hfill\codecomment{Include/Instantiate other fields of family \texttt{STLC}}@][@\tanc{c-STLC-d-rest}@][@\innersep@]
+[@\codecomment{Include/Instantiate other fields of family \texttt{STLC}}@] [@\dadada@][@\tanc{c-STLC-d-rest}@][@\innersep@]
 [@\outdent@]End STLC.[@\tanc{c-STLC-end}@]
@@
 
+\makeline[-1pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!50][1.8pt]
 \makeline[-1pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!50][1.8pt]
-\makeline[-1pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!50][1.8pt]
-\makeline[-1pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!50][1.8pt]
-\makeline[-1pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!50][1.8pt]
-\makeline[-1pt]{c-STLC-rest}{c-STLC-rest}[codecomment-color!50][1.8pt]
-\makeline[-1pt]{c-STLC-start}{c-STLC-firstcol-end}[codecomment-color!50][1.8pt]
-\makeline[4pt]{c-STLC-d-tm-start}{c-STLC-d-tm-end}[codecomment-color!35][1.0pt]
-\makeline[4pt]{c-STLC-d-subst-start}{c-STLC-firstcol-end}[codecomment-color!35][1.0pt]
-\makeline[.5\textwidth+8pt]{c-STLC-sndcol-start}{c-STLC-end}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLC-rest-start}{c-STLC-rest-end}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+8pt]{c-STLC-start}{c-STLC-end}[codecomment-color!50][1.8pt]
+\makeline[.5\textwidth+12pt]{c-STLC-d-tm-start}{c-STLC-d-tm-end}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+12pt]{c-STLC-d-subst-start}{c-STLC-d-subst-end}[codecomment-color!35][1.0pt]
 \makeline[.5\textwidth+12pt]{c-STLC-d-ty}{c-STLC-d-ty}[codecomment-color!35][1.0pt]
@@
 
-\vspace{-12pt}
-
-\colorlet{codecomment-color}{orange}
-
-\begin{lstlisting}
-[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{tm} in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-tm-start}@]
-[@\outdent@]Module Type STLCFix\_tm\_Ctx.
-[@\outdent@]End STLCFix\_tm\_Ctx.[@\innersep@]
-[@\outdent@]Module Type STLCFix\_tm (self : STLCFix\_tm\_Ctx).
-Include STLC\_tm(self).
-Axiom tm_fix : id -> tm -> tm.
-Axiom tm_fix_eq_STLC : \forall[@\,@][@\dadada@],[@\,@]tm_prect_STLC [@\dadada@][@\;@]=[@\;@]None.
-Axiom tm_prect_STLCFix : [@\dadada@].
-Axiom tm_fix_eq_STLCFix : [@\dadada@].
-[@\outdent@]End STLCFix\_tm.[@\tanc{c-STLCFix-tm-end}@][@\outersep@]
-[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-subst-start}@]
-[@\outdent@]Module Type STLCFix\_subst\_Ctx.
-Include STLCFix\_tm\_Ctx. Include STLCFix\_tm.
-[@\outdent@]End STLCFix\_subst\_Ctx.[@\innersep@]
-[@\outdent@]Module STLCFix\_subst\_Cases (self[@\,@]:[@\,@]STLCFix\_subst\_Ctx).
-Include STLC\_subst\_Cases(self).
-Def subst\_tm_fix := [@\dadada@].
-[@\outdent@]End STLCFix\_subst\_Cases.[@\innersep@]
-[@\outdent@]Module Type STLCFix\_subst (self[@\,@]:[@\,@]STLCFix\_subst\_Ctx).
-Include STLC\_subst(self).
-Axiom subst_tm_fix_eq : [@\dadada@].
-[@\outdent@]End STLCFix\_subst.[@\tanc{c-STLCFix-subst-end}@][@\outersep@]
-[@\outdent@][@\codecomment{Code emitted for other fields defined in family \texttt{STLCFix}}@] [@\dadada@][@\tanc{c-STLCFix-rest}@][@\outersep@]
-[@\outdent\codecomment{Code emitted upon completion of family \texttt{STLCFix}}@][@\tanc{c-STLCFix-start}@]
-[@\outdent@]Module STLCFix.[@\innersep@]
-Inductive tm\_ : Set :=[@\tanc{c-STLCFix-d-tm-start}@]
-| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
-| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_
-| tm_fix\_[@\,@]:[@\,@]id->tm\_->tm\_.
-Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
-Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
-Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
-Def tm_fix  := tm_fix\_.
-[@\dadada@][@\hfill\codecomment{Instantiate \texttt{tm} partial recursors \& their equalities}@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
-Include STLCFix\_subst\_Cases. [@\tanc{c-STLCFix-d-subst-start}@]
-Def subst :=[@\hfill\codecomment{Instantiate \texttt{subst} \& its equalities}@]
-  tm\__rect _ subst\_tm_unit subst\_tm_var
-             subst\_tm_abs  subst\_tm_app subst\_tm_fix.
-Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]
-[@\dadada@][@\hfill\codecomment{\textls[-10]{Instantiate \texttt{ty}, its constructors, partial recursor \& eq.}}@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
-Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLCFix-d-env}@][@\innersep@]
-[@\dadada@][@\hfill\codecomment{Include/Instantiate other fields of family \texttt{STLCFix}}@][@\tanc{c-STLCFix-d-rest}@][@\innersep@]
-[@\outdent@]End STLCFix.[@\tanc{c-STLCFix-end}@]
-\end{lstlisting}
-
-\makeline[.5\textwidth+8pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!50][1.8pt]
-\makeline[.5\textwidth+8pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!50][1.8pt]
-\makeline[.5\textwidth+8pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!50][1.8pt]
-\makeline[.5\textwidth+8pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!50][1.8pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-tm-start}{c-STLCFix-d-tm-end}[codecomment-color!35][1.0pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-subst-start}{c-STLCFix-d-subst-end}[codecomment-color!35][1.0pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-ty}{c-STLCFix-d-ty}[codecomment-color!35][1.0pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-env}{c-STLCFix-d-env}[codecomment-color!35][1.0pt]
-\makeline[.5\textwidth+13pt]{c-STLCFix-d-rest}{c-STLCFix-d-rest}[codecomment-color!35][1.0pt]
+\end{multicols}
+\end{minipage}
 
@@
 
-\definecolor{codecomment-color}{HTML}{000080}
-
-\begin{lstlisting}
-[@\outdent@][@\codecomment{Code emitted upon command \texttt{\textls[-35]{Check STLCFix.typesafe}}}@][@\tanc{c-Check-STLCFix-typesafe-start}@]
-[@\outdent@]Check STLCFix.typesafe.[@\tanc{c-Check-STLCFix-typesafe-end}@]
-\end{lstlisting}
-
-\makeline[.5\textwidth+8pt]{c-Check-STLCFix-typesafe-start}{c-Check-STLCFix-typesafe-end}[codecomment-color!50][1.8pt]
-
-\end{multicols}
-\end{minipage}
-
-%\vspace{-10pt}
-
-\caption{Compilation of the mechanization of the STLCs in \cref{fig:stlc-mechanized} to parameterized Coq modules.}
+\caption{Compilation of family \lsti!STLC! (\cref{fig:stlc-mechanized}).}
 \label{fig:stlc-compiled}

### 2022-11-05 stlcfix-compiled.tex
coqimpl

\begin{figure}
\vspace{-5pt}
\begin{minipage}{\textwidth}
\begin{multicols}{2}

\lstset{
  basicstyle=\fontsize{8}{8.2}\ttfamily,
% numbers=left,
}

\colorlet{codecomment-color}{dkgreen}

\newcommand{\innersep}{\vspace{6pt}}
\newcommand{\outersep}{\vspace{5pt}}
\newcommand{\outdent}{\!\!\!}

\colorlet{codecomment-color}{orange}

\begin{lstlisting}
[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{tm} in \texttt{STLCFix}}@][@\tanc{c-STLCFix-tm-start}@]
[@\outdent@]Module Type STLCFix\_tm\_Ctx.
[@\outdent@]End STLCFix\_tm\_Ctx.[@\innersep@]
[@\outdent@]Module Type STLCFix\_tm (self : STLCFix\_tm\_Ctx).
Include STLC\_tm(self).
Axiom tm_fix : id -> tm -> tm.
Axiom tm_fix_eq_STLC : \forall[@\,@][@\dadada@],[@\,@]tm_prect_STLC [@\dadada@][@\;@]=[@\;@]None.
Axiom tm_prect_STLCFix : [@\dadada@].
Axiom tm_fix_eq_STLCFix : [@\dadada@].
[@\outdent@]End STLCFix\_tm.[@\tanc{c-STLCFix-tm-end}@][@\outersep@]
[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in \texttt{STLCFix}}@][@\tanc{c-STLCFix-subst-start}@]
[@\outdent@]Module Type STLCFix\_subst\_Cases\_Ctx.
Include STLCFix\_tm\_Ctx. Include STLCFix\_tm.
[@\outdent@]End STLCFix\_subst\_Cases\_Ctx.[@\innersep@]
[@\outdent@]Module STLCFix\_subst\_Cases
[@\outdent@](self : STLCFix\_subst\_Cases\_Ctx).
Include STLC\_subst\_Cases(self).
Def subst\_tm_fix := [@\dadada@].
[@\outdent@]End STLCFix\_subst\_Cases.[@\innersep@]
[@\outdent@]Module Type STLCFix\_subst\_Ctx.
Include STLCFix\_subst\_Cases\_Ctx.
Include STLCFix\_subst\_Cases.
[@\outdent@]End STLCFix\_subst\_Ctx.[@\innersep@]
[@\outdent@]Module Type STLCFix\_subst (self[@\,@]:[@\,@][@\textls[-30]{STLCFix}@]\_[@\textls[-30]{subst}@]\_[@\textls[-30]{Ctx}@]).
Include STLC\_subst(self).
Axiom subst_tm_fix_eq : [@\dadada@].
[@\outdent@]End STLCFix\_subst.[@\tanc{c-STLCFix-subst-end}@][@\outersep@]
[@\outdent@][@\codecomment{Code emitted for other fields defined in \texttt{STLCFix}}@] [@\dadada@][@\tanc{c-STLCFix-rest}@][@\outersep@]
[@\outdent\codecomment{Code emitted upon conclusion of \texttt{STLCFix}}@][@\tanc{c-STLCFix-start}@]
[@\outdent@]Module STLCFix.[@\innersep@]
Inductive tm\_ : Set :=[@\tanc{c-STLCFix-d-tm-start}@]
| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_
| tm_fix\_[@\,@]:[@\,@]id->tm\_->tm\_.
[@\codecomment{Instantiate \texttt{tm} \& its constructors}@]
Def tm := tm\_.
Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
Def tm_fix  := tm_fix\_.
[@\codecomment{Instantiate \texttt{tm} partial recursors \& their equalities}@] [@\dadada@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
Include STLCFix\_subst\_Cases. [@\tanc{c-STLCFix-d-subst-start}@]
[@\codecomment{Instantiate \texttt{subst} \& its equalities}@]
Def subst := tm\__rect _ subst\_tm_unit subst\_tm_var
             subst\_tm_abs  subst\_tm_app subst\_tm_fix.
Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]
[@\codecomment{\textls[-10]{Instantiate \texttt{ty}, its constructors, partial recursor \& eq.}}@] [@\dadada@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLCFix-d-env}@][@\innersep@]
[@\codecomment{Include/Instantiate other fields of \texttt{STLCFix}}@] [@\dadada@][@\tanc{c-STLCFix-d-rest}@][@\innersep@]
[@\outdent@]End STLCFix.[@\tanc{c-STLCFix-end}@]
\end{lstlisting}

\makeline[-1pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!50][1.8pt]
\makeline[-1pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!50][1.8pt]
\makeline[.5\textwidth+8pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!50][1.8pt]
\makeline[.5\textwidth+8pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!50][1.8pt]
\makeline[.5\textwidth+12pt]{c-STLCFix-d-tm-start}{c-STLCFix-d-tm-end}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+12pt]{c-STLCFix-d-subst-start}{c-STLCFix-d-subst-end}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+12pt]{c-STLCFix-d-ty}{c-STLCFix-d-ty}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+12pt]{c-STLCFix-d-env}{c-STLCFix-d-env}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+12pt]{c-STLCFix-d-rest}{c-STLCFix-d-rest}[codecomment-color!35][1.0pt]

\vspace{-8pt}

\definecolor{codecomment-color}{HTML}{000080}

\begin{lstlisting}
[@\outdent@][@\codecomment{Code emitted upon command \texttt{\textls[-35]{Check STLCFix.typesafe}}}@][@\tanc{c-Check-STLCFix-typesafe-start}@]
[@\outdent@]Check STLCFix.typesafe.[@\tanc{c-Check-STLCFix-typesafe-end}@]
\end{lstlisting}

\makeline[.5\textwidth+8pt]{c-Check-STLCFix-typesafe-start}{c-Check-STLCFix-typesafe-end}[codecomment-color!50][1.8pt]

\end{multicols}
\end{minipage}

\vspace{-12pt}

\caption{Compilation of family \lsti!STLCFix! (\cref{fig:stlc-mechanized}).}
\label{fig:stlcfix-compiled}
\end{figure}

### 2022-11-06 casestudies.tex
a la carte

@@
 
-The elimination of inductive types defined via \lsti{FInductive} is
+Elimination of inductive types defined via \lsti{FInductive} is
 mostly via the \lsti{FRecursion} and \lsti{FInduction} commands.
-An exception is seemingly trivial ``inversion lemmas''.
+An exception is some seemingly trivial ``inversion lemmas''.
 For example, consider the lemma \lsti!\forall t, \neg step tm_true t!
@@
 %
-However, we observe that it is lighter-weight to use overriding: 
+However, we observe that it is lighter-weight to use overriding (\cref{sec:override}):
 the programmer can specify that the proof of the lemma should be overridden
-in any derived family that extends \lsti{step} with new constructors,
+in any derived family that further binds \lsti{step},
 and in return, they are permitted to treat \lsti{step} as an ordinary
-inductive type in the proof and thus discharge it with \lsti{inversion},
-with the plugin automatically trying the same proof script in a derived
+inductive type in the proof and thus discharge the proof with \lsti{inversion}.
+The plugin then automatically tries the same proof script in a derived
 family to override the proof.
+
+We note that the current plugin implementation does not support
+mutually inductive types or mutual induction, which will be
+useful for modeling languages much more complex than STLCs.
+We believe that generalizing our design to mutually inductive types does
+not pose theoretical challenges, however.
 

### 2022-11-06 coqimpl.tex
a la carte

@@
 The distinction is reflected in the translations.
-Further-bindable fields are translated to module types that export only types of the fields.
-Non-further-bindable fields are translated to modules that export definitional equalities
+The further-bindable fields are translated to module types that export only types of the fields.
+The non-further-bindable fields are translated to modules that export definitional equalities
 on the fields.
+Opaque definitions can be further bound, too (\cref{sec:override});
+they are translated to modules that export opaque definitions.
 
@@
 
-\fi
-
 \subsection{Implementation Limitation}
@@
 \end{itemize}
+
+\fi

### 2022-11-06 lang-design.tex
a la carte

@@
 displays the type-safety theorem of the fixpoints extension.
+
+\subsection{Extensible Inductive Types and Exhaustive Recursion/Induction}
 
@@
 
+\subsection{Late Binding and Equalities}
+\label{sec:latebind}
+
 \noindentparagraph{Late binding of nested names.}
@@
 
-\noindentparagraph{Overriding.}
+\subsection{Overriding}
+\label{sec:override}
+%\noindentparagraph{Overriding.}
 
@@
 %
-For example, in a derived family, the programmer may want to override
-the proof of a lemma, using a different proof strategy than it was proved
-in the base family.
+For example, in a derived family, rather than adding new cases to an
+induction proof, the programmer may prefer overriding the proof
+entirely, as we observe in our case studies (\cref{sec:casestudies}).
 %
@@
 include most proofs.
-It is safe to override opaque definitions, because Coq will never attempt to unfold them.
+It is safe to override opaque definitions, because the type checker will
+never attempt to unfold them.
 For transparent definitions, the common case is that the programmer does
@@
 
-\Lang does allow the overriding of transparent definitions that are
+\Lang does allow the overriding of transparent definitions
 explicitly marked as \lsti{Overridable} by the programmer.
@@
 should be overridden too.
-We expect this feature to be used only occasionally.
-Our case studies contain uses of overriding (\cref{sec:casestudies}).
+We expect this feature to be used occasionally.
 
-\noindentparagraph{Sound logical reasoning.}
+\subsection{Sound Logical Reasoning}
+\label{sec:sound}
+
+%\noindentparagraph{Sound logical reasoning.}
 
@@
 Such circularity would easily lead to logical inconsistency:
-\YZ{Mention non-exhaustivity/overriding already poses challenges to ensuring consistency.}
+%\YZ{Mention non-exhaustivity/overriding already poses challenges to ensuring consistency.}
 
@@
 Note that the requirement still allows a derived family to introduce
-new declarations into the context of a field. \EDJ{Maybe a better wording is ``the relative order between fields are preserved during inheritance''?}
+new declarations into the context of a field. %\EDJ{Maybe a better wording is ``the relative order between fields are preserved during inheritance''?}
 For example, the left column of \cref{fig:stlc-isorec-prod} shows an
@@
 introduced into the context of "hasty".
+%
+In the event that it cannot be inferred where to place of a newly introduced field,
+programmer annotation is required.
 
@@
 
-\noindentparagraph{Composing families.}
+\subsection{Composing Families as Mixins}
+
+%\noindentparagraph{Composing families.}
 
@@
 families.
+A family like \lsti{STLCFix} can be viewed as a family-to-family functor---%
+and hence a mixin, in the sense of \cite{flatt1998mixin}---that
+transforms any family providing the base STLC functionalities into a new
+family additionally supporting fixpoints.
+%\EDJ{I feel this paragraph will cause more confusion---because it is inheritance judgement/linkage transformer being mixed in, not family itself.
+Conceptually not distinguishing family and linkage transfomer will make reader wonder what this magic black box here actually is}
+
 In \cref{fig:stlc-isorec-prod}, family "STLCFixIsorec" is declared as
 an "STLC" extension that mixes in "STLCFix" and "STLCIsorec".
-As a result, \lsti{STLCFixIsorec.typesafe} is readily the proof of the type-safety
-theorem of an STLC equipped with fixpoints and iso-recursive types.\EDJ{I feel this paragraph will cause more confusion---because it is inheritance judgement/linkage transformer being mixed in, not family itself.
-Conceptually not distinguishing family and linkage transfomer will make reader wonder what this magic black box here actually is}
+The family is declared with minimal verbiage, yet
+\lsti{STLCFixIsorec.typesafe} is automatically a proof of the type-safety
+theorem of an STLC equipped with fixpoints and iso-recursive types.
 
@@
 case in "tysubst".
+
+%\YZ{
+%Family IDENT.
+%
+%Family IDENT extends IDENT_0.
+%
+%Family IDENT extends IDENT_0 using IDENT_1, IDENT_2, ..., IDENT_n.
+%
+%Semantically, family IDENT is the composition λ IDENT_0. IDENT_n(…(IDENT_2(IDENT_1(IDENT_0)))…).
+%
+%IDENT_0 is called the immediate superfamily of IDENT.
+%
+%If the using-clause is present, then for each i > 0, IDENT_0 must be the same as, or a derived family of, the immediate superfamily of IDENT_i.
+%
+%If there is any name conflict among nested fields, then either the programmer has to resolve it or the composition is rejected upon attempting to close the family.
+%}
+
+\subsection{Injectivity and Disjointness of Constructors via Partial Recursors}
 

### 2022-11-06 related-work.tex
a la carte

@@
-This section discusses related work.
+There exist many approaches to modular mechanization or proof reuse,
+with different focuses and trade-offs.
+
+\noindentparagraph{\`A la carte.}
+
+The design pattern \emph{data types à la carte} (DTC) introduced for Haskell by~\citet{dtc}
+is a source of inspiration for many approaches
+\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
+The original DTC encoding requires type-level general recursion that
+fails the strict positivity check imposed by proof assistants including Coq and Agda.
+\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
+it overcomes the problem by using Church encodings for
+data types and using Mendler-style folds for evaluation,
+though it requires \lsti{Set} impredicativity.
+Feature composition is automated through heavy use of type classes.
+The framework is implemented as a Coq library.
+\citet{schwaab2013modular} adapt DTC to Agda by considering a restricted
+class of functors that admit least fixed points.
+\citet{keuchel2013generic} use datatype-generic programming techniques
+for the underlying representation of type-level fixed points and
+avoid \lsti{Set} impredicativity.
+
+All of these approaches are largely extra\-linguistic, in that they work
+within the confine of the language offered by a proof assistant, which
+comes with trade-offs.
+On the one hand,
+they can be conveniently distributed as libraries, and the encoding can
+be more easily adapted for new purposes.
+For example, MTC~\cite{delaware2013} has been applied to implementing
+composable program adverbs~\cite{liwei2022} and been adapted to support
+unanticipated feature extensions that require type changes
+\cite{delaware2013b}.
+
+On the other hand, the extra\-linguistic nature of the approaches tend
+to lead to non-idiomatic code and offset their user-friendliness.
+In particular, because data types have to be encoded (rather than expressed
+through natively supported inductive types), the resulting code is
+less readable and the resulting programming style inaccessible to non-experts.
+In addition, extra programmer effort may be required, such as
+having to manually prove additional well-formedness conditions on the algebras.
+
+\citet{forsta2020} introduce \emph{Coq à la carte}.
+It still follows DTC, but rather than embracing DTC's use of generic
+fixed points, it considers specific instantiations instead.
+The resulting mechanism appears more streamlined than prior \emph{à la
+carte} approaches particularly for its extensive tool support for generating boilerplate code.
+%
+The approach does not seem to fully address the expression problem,
+as \citet{liwei2022} observe, in that extending a composed
+mechanization requires recompilation.
+
+Our approach addresses the expression problem by extending the
+linguistic facilities offered by a proof assistant.
+Families, in particular, offer an organizational advantage.  They allow
+grouping and coevolving related types, functions, and proofs without
+explicit parameterization;
+all further-bindable fields and all opaque fields are automatically
+extensibility hooks.
+Because family polymorphism does not require explicit parameterization
+or complex encodings,
+the resulting programming experience and code are accessible to the
+working Coq programmer.
+The more OO aspects of family polymorphism, such as the ability to
+override and families as mixins, also facilitate extensibility and reuse
+with minimal programmer effort.
+
+\bigskip
 
@@
 
+\ifShowOldWriting
+
 \textbf{Modular Mechanized Metatheory: Semantic Approach} We group the work using the semantic as encoding (e.g.
@@
 % Have no idea how to write this :(
+
+\fi

### 2022-11-06 stlc-isorec-prod.tex
a la carte

@@
 Family STLCFixIsorec extends STLC[@\tanc{STLCIsorecfix-start}@]
-using STLCFix, STLCIsorec.[@\hfill\codecomment{STLC extended}@]
+using STLCFix, STLCIsorec.[@\hfill\codecommentstart{STLC extended}@]
 [@\hfill\codecommentend{with fixpoints and iso-recursive types}@]

### 2022-11-07 casestudies.tex
justify escape hatch

@@
 mostly via the \lsti{FRecursion} and \lsti{FInduction} commands.
-An exception is some seemingly trivial ``inversion lemmas''.
+An exception is a handful of trivial ``inversion lemmas''.
 For example, consider the lemma \lsti!\forall t, \neg step tm_true t!
@@
 If \lsti{step} were an ordinary inductive type, then it could be proved in Coq
-with \lsti{intros t H; inversion H}.
+simply by \lsti{intros t H; inversion H}.
 But since \lsti{step} is extensible, it seems that the lemma has to be
 proved by \lsti{FInduction} on \lsti{step}, and that the programmer
-has to manually verify that the lemma still holds in derived families
+has to verify that the lemma still holds in derived families
 that add new constructors to \lsti{step}.
 %
-However, we observe that it is lighter-weight to use overriding (\cref{sec:override}):
+We observe that it is lighter-weight to use overriding (\cref{sec:override}) instead:
 the programmer can specify that the proof of the lemma should be overridden
@@
 The plugin then automatically tries the same proof script in a derived
-family to override the proof.
+family to override the proof. Although proof scripts rather
+than proof terms are reused, rechecking of proof terms seems justified
+by the triviality of the lemmas and the terseness of the proofs.
 

### 2022-11-07 casestudies.tex
reword

@@
 proved by \lsti{FInduction} on \lsti{step}, and that the programmer
-has to verify that the lemma still holds in derived families
-that add new constructors to \lsti{step}.
+has to verify that a derived family does not accidentally make \lsti{tm_true} reducible.
+%lemma still holds in derived families that add new constructors to \lsti{step}.
 %

### 2022-11-07 related-work.tex
Boite

@@
 
-\noindentparagraph{\`A la carte.}
+\noindentparagraph{Encodings based on product lines or DTC.}
+
+\citet{delaware2011} engineer product lines of theorems and proofs built
+from feature modules. Feature composition is manual, which seems to have
+motivated later approaches based on DTC.
 
@@
 explicit parameterization;
-all further-bindable fields and all opaque fields are automatically
-extensibility hooks.
+all further-bindable fields are automatically extensibility hooks.
 Because family polymorphism does not require explicit parameterization
@@
 with minimal programmer effort.
+
+\noindentparagraph{Proof reuse, proof weaving, and proof repair.}
+
+\citet{boite2004proof} addresses proof reuse specifically in response
+to inductive types extended with new constructors. Proof reuse is
+via a tactic that adapts the original proof to the extended inductive
+type while generating proof obligations, so rechecking of proof terms is
+entailed.
+The design requires distinct names for a base inductive type and its
+extensions (including distinct names for constructors), while our
+approach supports the late binding of names.
 

### 2022-11-07 related-work.tex
more on proof reuse and repair

@@
 
-\noindentparagraph{Proof reuse, proof weaving, and proof repair.}
+\noindentparagraph{Proof reuse and proof repair.}
 
@@
 The design requires distinct names for a base inductive type and its
-extensions (including distinct names for constructors), while our
-approach supports the late binding of names.
+extensions (including distinct names for constructors), while family
+polymorphism in our approach allows names to be late bound.
+
+\citet{mulhern2006proof} introduces a heuristic approach that allows proofs
+for multiple small languages to be combined to yield proofs for
+composite languages, as long as the proof structure follows the same
+pattern.
+%
+\citet{johlut2004} enable proof reuse in Isabelle by adapting theorems
+from one setting for reuse in another: proof terms are transformed
+by first explicitly stating all assumptions, and then abstracting over
+function symbols and type constants.
+
+Pumpkin Pi~\cite{ringer2021pumpkin} is a Coq plug-in that helps
+repair proofs broken by changes in type definitions.
+Its decompiler from proof terms to proof scripts prioritizes suggesting
+useful tactics over soundness.
+While Pumpkin Pi focuses on refactoring existing proofs in response to
+changed definitions,
+the focus of our solution is preempting refactoring altogether by
+writing code that has built-in hooks for future extensions.
 
 \bigskip
+
+\ifShowOldWriting
 
@@
 
-\ifShowOldWriting
-
 \textbf{Modular Mechanized Metatheory: Semantic Approach} We group the work using the semantic as encoding (e.g.

### 2022-11-07 related-work.tex
related work on EP

@@
 changed definitions,
-the focus of our solution is preempting refactoring altogether by
-writing code that has built-in hooks for future extensions.
+our solution can be viewed as an effort to preempt refactoring by
+allowing the programmer to write code that has built-in hooks for
+future extensions.
 
-\bigskip
+\noindentparagraph{Solutions to the expression problem} abound.
+Our approach is the first that applies family polymorphism~\cite{ernst2001family},
+an idea found in the context of OO programming,
+to the context of mechanized proofs.
+Families in \Lang are second-class, in that they are not considered
+terms by themselves.
+Families as modeled in \TT (namely linkages) are first-class and thus more flexible.
+The distinction corresponds to class-based vs.\ object-base family
+polymorphism as summarized before~\cite{cdnw07-tribe}.
+
+\citet{bac2006} address the expression problem in a functional language,
+introducing a language mechanism that allows pattern-matching cases to
+be added retroactively in accordance with extensions to variant types.
+Our \lsti{FRecursion} and \lsti{FInduction} commands achieve a similar
+functionality, with families bestowing additional organizational power.
 
@@
 
-\textbf{Solution to The Expression Problem}
-% Have no idea how to write this :(
-
 \fi

### 2022-11-07 related-work.tex
adjust claim about Coq a la carte

@@
 %
-The approach does not seem to fully address the expression problem,
-as \citet{liwei2022} observe, in that extending a composed
-mechanization requires recompilation.
+But even with the tool support, components of individually developed
+feature extensions (e.g., "substlem") have to be composed separately by
+invoking the tool.
+%
+%The approach does not seem to fully address the expression problem,
+%as \citet{liwei2022} observe, in that further extending a composed
+%mechanization causes (a small amount of) boilerplate code that glues definitions
+%to be regenerated.
 
@@
 The more OO aspects of family polymorphism, such as the ability to
-override and families as mixins, also facilitate extensibility and reuse
+use families as mixins and the ability to grow a series of mechanized
+languages in integral increments, also facilitate extensibility and
+reuse
 with minimal programmer effort.

### 2022-11-07 conclusion.tex
conclusion

It is hard to write modular, extensible code and proofs.
We have presented a solution that equips a proof assistant with
linguistic facilities for family polymorphism.
The language design ensures that the expressive power brought by family polymorphism
is in harmony with the strictness of a proof assistant, while incurring
low cognitive overhead and allowing an idiomatic programming experience.
We implement the design via a translation to Coq and demonstrate its applicability
using case studies.
A novel dependent type theory formalizes the core aspects of the language mechanism
and is shown to enjoy consistency and canonicity.
Generalizing the mechanism to support nested family polymorphism and
scaling it to sizable proof engineering efforts make exciting directions for
future work.

### 2022-11-07 metatheory-condensed.tex
coqimpl
related work on EP
Nada's comment

@@
 \underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory~\cite{martin1982constructive},
-while maintaining consistency and canonicity.
+while maintaining consistency and canonicity.\NA{maybe you can give an overview of the salient points of the design and what they're trying to achieve? What is the core aspect modeled by the theory?}
 
@@
 \begin{itemize}
-  [leftmargin=4ex]
-  %[labelsep=*,leftmargin=1pc]
+% [leftmargin=4ex,itemsep=.5ex]
+  [labelsep=*,leftmargin=1pc,itemsep=3pt]
 
@@
 %
-Other fields automatically generated by the plugin, such as partial
-recursors and various propositional equalities, can also be modeled in
-this way by invoking $\texttt{Wrec}$.
+Partial recursors automatically generated by the plugin
+can also be modeled in this way by invoking $\texttt{Wrec}$.
 
@@
 
+\ifShowOldWriting
 \newpage
@@
 
-\newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
-\newcommand{\Sigr}[2]{{ "Sig"^r~{#1}~{#2} }}
-
 Thanks to QIIT formulation, syntactic translation becomes a function between QIIT type, which by definition respects all the judgemental equalities.
@@
 \end{align*}
+
+\fi

### 2022-11-08 coqimpl.tex
related work on EP
excuse

@@
 The translation is modular and efficient, in that
-code inherited from a base family need not be retranslated or rechecked.
+code compiled for non-further-bindable fields of a base family
+is shared with derived families without having to be rechecked.
 
@@
 For example, module \lsti{STLC} in \cref{fig:stlc-compiled} is generated
-upon \lsti{End}~\lsti{STLCFix}.
+upon \lsti{End}~\lsti{STLC}.
 All the fields are added to this module in the same order as they appear in the family definition.
@@
 are reused by module \lsti{STLCFix}. So is the proof of \lsti{typesafe}.
+
+One may argue that because the first four constructors of \lsti{tm} are
+repeated in \lsti{STLCFix}, the translation does not satisfy the
+modular compilation requirement. Compared with alternatives that create
+wrapper inductive types, we consider it a small price to pay in return
+for the clarity of implementation especially given that compiled case handlers
+are entirely reusable without rechecking.
 

### 2022-11-08 coqimpl.tex
edits

@@
 
-For the further-bindables, \lsti{Axiom}s declared in the module types must be instantiated.
-%
-\begin{itemize}[align=left,itemsep=5pt,labelsep=*,leftmargin=*]
-
-\item
-For \lsti{tm},
-an inductive type \lsti{tm\_} is generated, with \lsti{tm\_}
-instantiating the axiomatized \lsti{tm} type, and with
-its constructors (\lsti{tm_unit\_} etc.) instantiating 
-the axiomatized constructors (\lsti{tm_unit} etc.).
-%
-The partial recursor \lsti{tm_prect_STLC} is instantiated
-%to \lsti{\lambda}\,\lsti{P, tm\__rect (\lambda}\,\lsti{t, option (P t))},
-with the help of \lsti{tm\_rect}, the recursor of the generated \lsti{tm\_} type.
-Proving the equalities on \lsti{tm_prect_STLC} is then trivial, by \lsti{reflexivity}.
-
-\item
-Similarly, \lsti{subst} is instantiated by applying the recursor \lsti{tm\_rect}
-to the (already included) case handlers (\lsti{subst\_tm_unit}, etc.).
-Propositional equalities on \lsti{subst} are then trivial, by \lsti{reflexivity}.
-
-\end{itemize}
-
 For the non-further-bindables, the translated modules are directly included
@@
 
+For the further-bindables, \lsti{Axiom}s declared in the module types must be instantiated.
+%
+\begin{itemize}[align=left,itemsep=0pt,labelsep=*,leftmargin=*]
+
+\item
+An inductive type \lsti{tm} is generated, instantiating the axiomatized
+\lsti{tm} type and its constructors.
+%
+The partial recursor \lsti{tm_prect_STLC} is defined
+%to \lsti{\lambda}\,\lsti{P, tm_rect (\lambda}\,\lsti{t, option (P t))},
+with the help of \lsti{tm_rect}, the recursor Coq generates for \lsti{tm}.
+Proving the equalities on \lsti{tm_prect_STLC} is then trivial, by \lsti{reflexivity}.
+
+\item
+Similarly, \lsti{subst} is instantiated by applying the recursor \lsti{tm_rect}
+to the (already included) case handlers (\lsti{subst\_tm_unit}, etc.).
+Propositional equalities on \lsti{subst} are then trivial, by \lsti{reflexivity}.
+
+\end{itemize}
+
 Module \lsti{STLCFix} in \cref{fig:stlcfix-compiled} is generated upon
@@
 The translation makes code and proof sharing evident.
-For example, the four case handlers for \lsti{subst} compiled for \lsti{STLC}
-are reused by module \lsti{STLCFix}. So is the proof of \lsti{typesafe}.
+For example, the four case handlers (\lsti{subst\_tm_var} etc.) compiled for \lsti{STLC}
+are reused to instantiate \lsti{subst}.
+Translations of the definition of \lsti{env} and the proof of \lsti{typesafe},
+are also reused to construct module \lsti{STLCFix}.
 

### 2022-11-08 stlc-compiled.tex
edits

@@
 Axiom subst_tm_var_eq : [@\dadada@].
-Axiom subst_tm_abs_eq : [@\dadada@].
-Axiom subst_tm_app_eq : [@\dadada@].
+Axiom [@\dadada@]
 [@\outdent@]End STLC\_subst.[@\tanc{c-STLC-subst-end}@][@\columnbreak@]
@@
 [@\outdent@]Module STLC.[@\innersep@]
-Inductive tm\_ : Set :=[@\tanc{c-STLC-d-tm-start}@]
-| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
-| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_.
-[@\codecomment{Instantiate \texttt{tm} \& its constructors}@]
-Def tm := tm\_.
-Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
-Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
+[@\codecomment{Instantiate \texttt{tm} \& its constructors}@][@\tanc{c-STLC-d-tm-start}@]
+Inductive tm : Set :=
+| tm_unit : tm | tm_var : id -> tm
+| tm_abs : id -> tm -> tm | tm_app : tm -> tm -> tm.
 [@\codecomment{Instantiate \texttt{tm} partial recursor \& its equalities}@]
-Def tm_prect_STLC := \lambda P, tm\__rect (\lambda[@\,@]t, option (P t)).
-Fact[@\;@][@\textls[-20]{tm\_unit\_eq\_STLC}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
+Def tm_prect_STLC := \lambda P, tm_rect (\lambda[@\,@]t, option (P t)).
+Fact [@\textls[-20]{tm\_unit\_eq\_STLC}@] : [@\dadada@]. [@\textls[-40]{{\color{dkblue}reflexivity}.}@] Qed.
+Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
 Include STLC\_subst\_Cases.[@\tanc{c-STLC-d-subst-start}@]
 [@\codecomment{Instantiate \texttt{subst} \& its equalities}@]
-Def subst[@\,@] := tm\__rect _ subst\_tm_unit subst\_tm_var
-                        subst\_tm_abs  subst\_tm_app.
-Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}@][@\innersep@]
+Def subst[@\,@] := tm_rect _ subst\_tm_unit
+  subst\_tm_var subst\_tm_abs subst\_tm_app.
+Fact [@\textls[-40]{subst\_tm\_unit\_eq}@] : [@\dadada@]. [@\textls[-40]{{\color{dkblue}reflexivity}.}@] Qed.
+Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}@][@\innersep@]
 [@\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@] [@\dadada@][@\tanc{c-STLC-d-ty}@][@\tanc{c-STLC-sndcol-start}@][@\innersep@]

### 2022-11-08 stlcfix-compiled.tex
edits

@@
 [@\outdent@](self : STLCFix\_subst\_Cases\_Ctx).
-Include STLC\_subst\_Cases(self).
-Def subst\_tm_fix := [@\dadada@].
+Include STLC\_subst\_Cases(self).[@\hfill\codecomment{reuse}@]
+Def subst\_tm_fix := [@\dadada@].[@\hfill\codecomment{translation of new case}@]
 [@\outdent@]End STLCFix\_subst\_Cases.[@\innersep@]
@@
 [@\outdent@]Module STLCFix.[@\innersep@]
-Inductive tm\_ : Set :=[@\tanc{c-STLCFix-d-tm-start}@]
-| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
-| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_
-| tm_fix\_[@\,@]:[@\,@]id->tm\_->tm\_.
-[@\codecomment{Instantiate \texttt{tm} \& its constructors}@]
-Def tm := tm\_.
-Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
-Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
-Def tm_fix  := tm_fix\_.
+[@\codecomment{Instantiate \texttt{tm} \& its constructors}@][@\tanc{c-STLCFix-d-tm-start}@]
+Inductive tm : Set :=
+| tm_unit : tm | tm_var : id -> tm
+| tm_abs : id -> tm -> tm | tm_app : tm -> tm -> tm
+| tm_fix : id -> tm -> tm.
 [@\codecomment{Instantiate \texttt{tm} partial recursors \& their equalities}@] [@\dadada@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
@@
 [@\codecomment{Instantiate \texttt{subst} \& its equalities}@]
-Def subst := tm\__rect _ subst\_tm_unit subst\_tm_var
-             subst\_tm_abs  subst\_tm_app subst\_tm_fix.
-Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]
-[@\codecomment{\textls[-10]{Instantiate \texttt{ty}, its constructors, partial recursor \& eq.}}@] [@\dadada@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
-Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLCFix-d-env}@][@\innersep@]
+Def subst := tm_rect _ subst\_tm_unit
+subst\_tm_var subst\_tm_abs subst\_tm_app[@\hfill\codecomment{reuse}@]
+subst\_tm_fix.
+Fact [@\textls[-40]{subst\_tm\_unit\_eq}@] : [@\dadada@]. [@\textls[-40]{{\color{dkblue}reflexivity}.}@] Qed.
+Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]
+[@\codecomment{\textls[-10]{Include\;\texttt{ty}, its constructors, partial recursor \& eq.}}@] [@\dadada@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
+Include STLC\_env.[@\hfill\codecomment{reuse}@][@\tanc{c-STLCFix-d-env}@][@\innersep@]
 [@\codecomment{Include/Instantiate other fields of \texttt{STLCFix}}@] [@\dadada@][@\tanc{c-STLCFix-d-rest}@][@\innersep@]
+Include STLC\_typesafe.[@\hfill\codecomment{reuse}@][@\tanc{c-STLCFix-d-typesafe}@][@\innersep@]
 [@\outdent@]End STLCFix.[@\tanc{c-STLCFix-end}@]
@@
 \makeline[.5\textwidth+12pt]{c-STLCFix-d-rest}{c-STLCFix-d-rest}[codecomment-color!35][1.0pt]
+\makeline[.5\textwidth+12pt]{c-STLCFix-d-typesafe}{c-STLCFix-d-typesafe}[codecomment-color!35][1.0pt]
 

### 2022-11-08 metatheory-condensed.tex
executive summary of metatheory

@@
-We contribute \TT, a core type theory that incorporates the core elements
-of \underline{f}a\underline{m}i\underline{l}y polymorphism into
-\underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory~\cite{martin1982constructive},
-while maintaining consistency and canonicity.\NA{maybe you can give an overview of the salient points of the design and what they're trying to achieve? What is the core aspect modeled by the theory?}
+We contribute \TT, a core type theory that extends
+%\underline{f}a\underline{m}i\underline{l}y polymorphism
+%\underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory (MLTT)~\cite{martin1982constructive}
+Martin-Löf dependent type theory (MLTT)~\cite{martin1982constructive}
+with facilities to express family polymorphism
+while maintaining consistency and canonicity.
+%\NA{maybe you can give an overview of the salient points of the design and what they're trying to achieve? What is the core aspect modeled by the theory?}
+%
+We acknowledge that the presentation herein is dense for an audience
+without prior exposure to MLTT, so we summarize the salient points
+first before delving into the technical details.
 
-Most notably, \TT extends MLTT with what we call \emph{linkages}
-(a namesake of the theoretical device through which \citet{zm2017} model
-family polymorphism in an OO setting).
-Linkages model families, so they are like tuples but with a twist: they
-support late binding.
+\noindentparagraph{Summary.}
 
@@
 typing is structural rather than nominal, and
-it supports family polymorphism by providing MLTT-style typing rules that
-can be used to \emph{encode} core aspects of the language supported by our
-plugin, rather than building them in directly.
+partial recursors and propositional equalities are not automated but have to
+be explicitly defined.
+\TT provides MLTT-style constructs that can be used to \emph{encode}
+family polymorphism rather than building it in directly.
 %
-For example, \TT supports extending inductive types through
-formation rules that enable the encoding of the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
-Also unlike our plugin, \TT does not feature automations such as
-the generation of propositional equalities and partial recursors,
-but it is expressive enough to encode their definitions.
+Most notably, \TT extends MLTT with what we call \emph{linkages}.\footnote{
+Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
+family polymorphism in an OO setting, but the technical details differ significantly.}
+Linkages model families, so they are like tuples but with a twist: they
+support late binding.
+Unlike dependent tuples where a later component is \emph{existentially}
+quantified over the earlier ones, a linkage component is \emph{universally}
+quantified over---and thus polymorphic to---components preceding it.
+%
+\TT also features \emph{linkage transformers} that can be used to inductively
+construct linkages from other linkages, by inheriting or overriding components of a linkage,
+or by extending linkages with new components.
+%
+Inductive types are modeled as W-types~\cite{martin1984intuitionistic} and their
+extension as overriding.
+%
+Consistency and canonicity are proved by giving semantic models to the syntax.
+
+%\TT supports extending inductive types through
+%formation rules that enable the encoding of the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
+
+%Also unlike our plugin, \TT does not feature automations such as
+%the generation of propositional equalities and partial recursors,
+%but it is expressive enough to encode their definitions.
 
@@
 We briefly review the base MLTT fragment first.
-
+%
 As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:

### 2022-11-08 casestudies.tex
conclusion
shortening
edits

@@
 
-To demonstrate the applicability of our design and implementation, we
-conduct three case studies.
+%To demonstrate the applicability of our design and implementation, we
+%conduct two case studies.
 
@@
 simply by \lsti{intros t H; inversion H}.
-But since \lsti{step} is extensible, it seems that the lemma has to be
-proved by \lsti{FInduction} on \lsti{step}, and that the programmer
-has to verify that a derived family does not accidentally make \lsti{tm_true} reducible.
+But \lsti{step} is extensible. So one way to prove the lemma is
+by \lsti{FInduction} on \lsti{step} and verifying that
+a derived family does not accidentally make \lsti{tm_true} reducible.
 %lemma still holds in derived families that add new constructors to \lsti{step}.
@@
 and in return, they are permitted to treat \lsti{step} as an ordinary
-inductive type in the proof and thus discharge the proof with \lsti{inversion}.
+inductive type in the proof and thus use \lsti{inversion} to prove it.
 The plugin then automatically tries the same proof script in a derived
 family to override the proof. Although proof scripts rather
-than proof terms are reused, rechecking of proof terms seems justified
-by the triviality of the lemmas and the terseness of the proofs.
+than proof terms are reused, the practice seems justified
+by the triviality of the lemmas and the terseness of the proof scripts.
 
-We note that the current plugin implementation does not support
-mutually inductive types or mutual induction, which will be
+The current plugin implementation does not yet support
+mutually inductive types or mutual induction, which are
 useful for modeling languages much more complex than STLCs.
-We believe that generalizing our design to mutually inductive types does
-not pose theoretical challenges, however.
+However,
+we believe generalizing our design to mutually inductive types does
+not pose theoretical challenges.
 
 \noindentparagraph{Abstract interpreters for imperative languages.}
-
-\begin{wrapfigure}[13]{r}{0.13\textwidth}
-\includegraphics[scale=.68]{graphics/ai-casestudy.pdf}
-\end{wrapfigure}
 
@@
 
+\begin{wrapfigure}[13]{r}{0.13\textwidth}
+\includegraphics[scale=.68]{graphics/ai-casestudy.pdf}
+\end{wrapfigure}
+
 \noindent

### 2022-11-08 metatheory-condensed.tex
edits

@@
 construct linkages from other linkages, by inheriting or overriding components of a linkage,
-or by extending linkages with new components.
+or by extending a linkage with new components.
 %

### 2022-11-08 introduction.tex
fix broken refs

@@
 Martin–Löf type theory with family polymorphism and extensible inductive
-types (\cref{sec:metatheory}).
+types (\cref{sec:metatheory2}).
 We derive foundational meta\-theoretical results including consistency
@@
 \item We present case studies of using our Coq plugin to mechanize
-language metatheories (\cref{sec:coqexample}).
+language metatheories (\cref{sec:casestudies}).
 They show how our language design naturally solves the EP and

### 2022-11-08 casestudies.tex
edits and tightening

@@
 family to override the proof. Although proof scripts rather
-than proof terms are reused, the practice seems justified
+than proof terms are reused, this practice seems justified
 by the triviality of the lemmas and the terseness of the proof scripts.

### 2022-11-08 coqimpl.tex
edits and tightening

@@
-\input{stlc-compiled.tex}
-\input{stlcfix-compiled.tex}
-
 We implement our language design as a Coq plugin.
@@
 
-For instance, consider field \lsti{env} in family \lsti{STLC}.
+For instance,
+consider field \lsti{env} in family \lsti{STLC}.
 It is translated into a top-level module named \lsti{STLC\_env}.
@@
 family: fields of the current family in the context of \lsti{env} can be
-referenced through this \lsti{self}.
-In particular, \lsti{env} is defined as \lsti{id -> option ty},
+referenced through \lsti{self}.
+In particular, \lsti{env} is defined as \lsti{id}\,->\,\lsti{option ty},
 where \lsti{ty} is a late-bound reference to the \lsti{ty} field of the
@@
 discussion in \cref{sec:sound}.
+
+\input{stlc-compiled.tex}
+\input{stlcfix-compiled.tex}
 
@@
 The translation makes code and proof sharing evident.
-For example, the four case handlers (\lsti{subst\_tm_var} etc.) compiled for \lsti{STLC}
-are reused to instantiate \lsti{subst}.
-Translations of the definition of \lsti{env} and the proof of \lsti{typesafe},
+In particular, case handlers compiled for \lsti{STLC}
+are reused to instantiate \lsti{subst}, \lsti{substlem}, and alike.
+\lsti{STLC\_env} and \lsti{STLC\_typesafe}
 are also reused to construct module \lsti{STLCFix}.
-
+%
 One may argue that because the first four constructors of \lsti{tm} are
 repeated in \lsti{STLCFix}, the translation does not satisfy the
-modular compilation requirement. Compared with alternatives that create
-wrapper inductive types, we consider it a small price to pay in return
-for the clarity of implementation especially given that compiled case handlers
-are entirely reusable without rechecking.
+modular compilation requirement.
+We could address this issue by using wrapper types,
+but we consider restating constructors a small price to pay in return
+for the clarity and concision of implementation.
+We emphasize that compiled case handlers
+are entirely reusable without rechecking, even with restated constructors.
 

### 2022-11-08 introduction.tex
edits and tightening

@@
 mechanism as a Coq plugin (\cref{sec:coqimpl}). The plugin works by
-compiling surface-language terms into Gallina terms parameterized by
-extensibility hooks.
+compiling surface-language definitions into Coq modules parameterized by
+explicit extensibility hooks.
 
 \item We capture the essence of the new language mechanism formally by extending
-Martin–Löf type theory with family polymorphism and extensible inductive
-types (\cref{sec:metatheory2}).
-We derive foundational meta\-theoretical results including consistency
+Martin–Löf type theory with facilities to express family polymorphism (\cref{sec:metatheory2}).
+We prove foundational meta\-theoretical results including consistency
 and canonicity.
 %\EDJ{I will suggest use ``foundational'' instead. Sterling at some places say canonicity is the most basic meta-theoretical property.}
-We also formalize the translation implemented by the Coq plugin.
+%We also formalize the translation implemented by the Coq plugin.
 

### 2022-11-08 metatheory-condensed.tex
edits and tightening

@@
 family polymorphism in an OO setting, but the technical details differ significantly.}
-Linkages model families, so they are like tuples but with a twist: they
+Linkages model families, so they are tuples composed of fields but with a twist: they
 support late binding.
@@
 
+\input{stlc-linkage-typing.tex}
+
 We illustrate this with \cref{fig:stlc-linkage-typing}.
@@
 In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are identities,
-and $\seal_6$ turns
+and $\seal_6$ is from
 \begingroup
@@
 \endgroup
-into
+to
 \begingroup
@@
 
-When a family is completed (e.g., \lsti!End!~\lsti!STLC!),
+When a family is concluded (e.g., \lsti!End!~\lsti!STLC!),
 a linkage~$\lkg$ containing all the fields is available
@@
 type-checked using linkages and linkage transformers.
-
-\input{stlc-linkage-typing.tex}
 
@@
 
-\begin{figure}[!htb]
-  \lstset{
-      basicstyle=\fontsize{8}{8.5}\ttfamily,
-  % numbers=left,
-  }
-
-  \begin{minipage}{\textwidth}
-  \begin{multicols}{3}
-
-  \definecolor{codecomment-color}{HTML}{0DA3FF}
-
-  \begin{lstlisting}
-  Family STLC_bool extends STLC.
-
-   FInductive ty += ty_bool
-   FInductive tm += tm_true ..
-
-   FRecursion subst
-     (* Inherits all others *)
-
-     Case tm_true ...
-
-   End subst
-
-   Inherit t1
-
-  ...
-  \end{lstlisting}
-
-  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
-
-  \columnbreak
-  % \definecolor{codecomment-color}{HTML}{5D030F}
-
-  \begin{lstlisting}
-
-  i[@₀@] = inh-id : Inh ⋅ {} {} 
-  i[@₁@] = inh-ov i[@₀@] (W [@τ₁@]') : Inh ⋅ {ty : 𝕊(W τ₁)} {ty : 𝕊(W τ₁')}
-  s[@₂@]' : {ty : 𝕊(W τ₁')} | [ty : 𝕌] 
-  i[@₂@] = inh-ov i₁ (W τ₂') 
-     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂') }
-  (* recall s[@₃₁@] : {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} 
-                | [ty : 𝕌;  ..] *)
-  s[@₃₁@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂')} | [ty : 𝕌; ..; tm_true : tm; ..]
-  ↑s[@₃@] : Tm ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..]) [ty : 𝕌; ..] 
-  (* ↑s[@₃@] map (the type of) s[@₃₁@]' to that of the sealing s[@₃₁@] before inheritance *)
-  ii[@₃₀₀@] = inh-id 
-           ⫶ Inh ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..])
-              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
-              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
-  ii[@₃₀₁@] = inh-extend (inh-extend ii[@₃₀₀@] {p[@₁@]} .. ) {p[@₁@]} ..
-        : Inh ( ⋅, [ty : 𝕌; tm :𝕌; ..; tm_true : tm; ..]) 
-              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]] (RecSig τ₂'  tm)
-  i[@₃₁@] = inh+inh i[@₂@] ↑s[@₃@] ii[@₃₀₁@] : 
-      Inh ⋅ {ty : 𝕊(W τ₁); ..; subst' : 𝕃(RecSig τ₂ tm)} 
-            {ty : 𝕊(W τ₁'); ..; subst' : 𝕃(RecSig τ₂' tm)}
-  s[@₃@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm) }
-      | [ty : 𝕌; .. ; tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm)] 
-  i[@₃@] = inh-ov i[@₃₁@] ..
-     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂); subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm}  
-             {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm); subst : subst : tm → nat → tm → tm}
-  (* recall s₄ : {ty : 𝕊(W τ₁); .. subst : .. } 
-      | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..] *)
-  s[@₄@]' : {ty : 𝕊(W τ₁'); .. subst : .. } 
-        | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
-  ↑s[@₄@] : Tm ( ⋅, [ty : 𝕌; .. ; tm : 𝕌; .. ; subst' : 𝕃(RecSig τ₂ tm);subst : ..]) 
-          [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
-  i[@₄@] = inh-inh i[@₃@] ↑s[@₄@] 
-         : Inh ⋅ {ty : 𝕊(W τ₁); .. subst : ..; t1 : 𝕊(tm_var 0)} 
-                 {ty : 𝕊(W τ₁'); .. subst : ..; t1 : 𝕊(tm_var 0)} 
-  \end{lstlisting}
-
-  \columnbreak
-
-  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
-
-  \end{multicols}
-  \end{minipage}
-  \caption{Inheritance Metatheory}\label{fig:inheritance}
-  \end{figure}
-
 In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.
@@
 
+\begin{figure}[!htb]
+  \lstset{
+      basicstyle=\fontsize{8}{8.5}\ttfamily,
+  % numbers=left,
+  }
+
+  \begin{minipage}{\textwidth}
+  \begin{multicols}{3}
+
+  \definecolor{codecomment-color}{HTML}{0DA3FF}
+
+  \begin{lstlisting}
+  Family STLC_bool extends STLC.
+
+   FInductive ty += ty_bool
+   FInductive tm += tm_true ..
+
+   FRecursion subst
+     (* Inherits all others *)
+
+     Case tm_true ...
+
+   End subst
+
+   Inherit t1
+
+  ...
+  \end{lstlisting}
+
+  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
+
+  \columnbreak
+  % \definecolor{codecomment-color}{HTML}{5D030F}
+
+  \begin{lstlisting}
+
+  i[@₀@] = inh-id : Inh ⋅ {} {} 
+  i[@₁@] = inh-ov i[@₀@] (W [@τ₁@]') : Inh ⋅ {ty : 𝕊(W τ₁)} {ty : 𝕊(W τ₁')}
+  s[@₂@]' : {ty : 𝕊(W τ₁')} | [ty : 𝕌] 
+  i[@₂@] = inh-ov i₁ (W τ₂') 
+     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂') }
+  (* recall s[@₃₁@] : {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} 
+                | [ty : 𝕌;  ..] *)
+  s[@₃₁@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂')} | [ty : 𝕌; ..; tm_true : tm; ..]
+  ↑s[@₃@] : Tm ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..]) [ty : 𝕌; ..] 
+  (* ↑s[@₃@] map (the type of) s[@₃₁@]' to that of the sealing s[@₃₁@] before inheritance *)
+  ii[@₃₀₀@] = inh-id 
+           ⫶ Inh ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..])
+              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
+              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
+  ii[@₃₀₁@] = inh-extend (inh-extend ii[@₃₀₀@] {p[@₁@]} .. ) {p[@₁@]} ..
+        : Inh ( ⋅, [ty : 𝕌; tm :𝕌; ..; tm_true : tm; ..]) 
+              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]] (RecSig τ₂'  tm)
+  i[@₃₁@] = inh+inh i[@₂@] ↑s[@₃@] ii[@₃₀₁@] : 
+      Inh ⋅ {ty : 𝕊(W τ₁); ..; subst' : 𝕃(RecSig τ₂ tm)} 
+            {ty : 𝕊(W τ₁'); ..; subst' : 𝕃(RecSig τ₂' tm)}
+  s[@₃@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm) }
+      | [ty : 𝕌; .. ; tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm)] 
+  i[@₃@] = inh-ov i[@₃₁@] ..
+     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂); subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm}  
+             {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm); subst : subst : tm → nat → tm → tm}
+  (* recall s₄ : {ty : 𝕊(W τ₁); .. subst : .. } 
+      | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..] *)
+  s[@₄@]' : {ty : 𝕊(W τ₁'); .. subst : .. } 
+        | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
+  ↑s[@₄@] : Tm ( ⋅, [ty : 𝕌; .. ; tm : 𝕌; .. ; subst' : 𝕃(RecSig τ₂ tm);subst : ..]) 
+          [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
+  i[@₄@] = inh-inh i[@₃@] ↑s[@₄@] 
+         : Inh ⋅ {ty : 𝕊(W τ₁); .. subst : ..; t1 : 𝕊(tm_var 0)} 
+                 {ty : 𝕊(W τ₁'); .. subst : ..; t1 : 𝕊(tm_var 0)} 
+  \end{lstlisting}
+
+  \columnbreak
+
+  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
+
+  \end{multicols}
+  \end{minipage}
+  \caption{Inheritance Metatheory}\label{fig:inheritance}
+  \end{figure}
+
 \fi

### 2022-11-08 metatheory-condensed.tex
sneak peek of linkage transformer use

@@
 %apply them to linkages to form new linkages.
-%
+
 Derived families can be modeled as linkage transformers inductively
-constructed from the introduction forms $\InhId$, $\InhExt{h}{t}$, $\InhOv{h}{t}$, etc.
+constructed from the introduction forms
+$\InhId$, $\InhExt{h}{t}$, $\InhOv{h}{t}$, $\InhInh{h}$, etc.
 \cref{fig:fmltt-selected} shows the beta rule of an exemplary transformer,
@@
 field~$t_1$ with~$t_2$.
-$\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$, for instance, can be used
-in \cref{fig:stlc-linkage-typing} to form a derived family that
+For instance, the construction below shows that 
+$\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$ is used as the first step
+in creating a derived family that
 overrides $\wsig_{\texttt{tm}}$ with an extended
 signature~$\wsig'_{\texttt{tm}}$.
+\vspace{.9ex}
+
+\noindent
+\input{stlcfix-linkage-typing.tex}
+
+\noindent
+A supplemental appendix gives a more complete example
+showing how families are modeled and
+type-checked using linkages and linkage transformers.
 
@@
 %usage of equality reflection. 
-
+%
 We also typeset a more complete set of rules than the space allows and
 present it in a supplemental appendix.
-%
-Another supplemental appendix gives a more complete example than
-\cref{fig:stlc-linkage-typing} showing how families are modeled and
-type-checked using linkages and linkage transformers.
 

### 2022-11-08 stlc-linkage-typing.tex
singleton type
tighten fmltt figures a bit
Venn diagram
stlc case study
sneak peek of linkage transformer use

@@
 \renewcommand*{\arraystretch}{1.15}
+\setlength\tabcolsep{0pt}
 \newcolumntype{L}{>{$}l<{$}}
@@
 %
-\definecolor{A-color}{HTML}{E7EFC5}
-\definecolor{T-color}{HTML}{FFF1D0}
-\definecolor{t-color}{HTML}{FAE0E4}
 \colorlet{Bkg-color}{black!8}
+\colorlet{Bkg-color-2}{white}
 %
@@
 \lsti!FInductive!\ \ 
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.01,-0.10)(-0.02,0.28)%
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.03,-0.10)(-0.02,0.28)%
     \lsti!tm!%
     \tikzmarkend{t1-c}%
-\ \lsti!:=!}
+\ \;\lsti!:=!}
 &
@@
     \tikzmarkend{t2-c}%
+    \ \ \lsti!: tm!
 }
@@
 \\
-\\[-8pt]
+\cellcolor{Bkg-color-2}
 \lsti!| !\dadada\ \ \lsti!| !\dadada\ \ \lsti!| !\dadada\lsti!.!
@@
 \\[4pt]
-\lsti!FRecursion subst!\ \dadada\lsti!.!
+\cellcolor{Bkg-color-2}
+\lsti!FRecursion subst!\ \;\dadada\lsti!.!
 &
@@
     $\wcode{\wsig_{\texttt{tm}}}$ is abstracted into $\cU$ in the typing
-    context of the case handler below:%
+    context (i.e., $A_6$) of the case handler below:%
 }
@@
 \multirow{-2}{*}{%
-    \lsti!Case tm_unit :=!\ %
+    \lsti!Case tm_unit :=!\ \;%
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-c}(0.01,-0.10)(-0.02,0.28)%
@@
 \\
-\\[-8pt]
+\cellcolor{Bkg-color-2}
 \lsti!Case!\ \dadada\ \ 
@@
 \\
+\cellcolor{Bkg-color-2}
 \dadada
@@
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-b}(-0.02,-0.10)(-0.02,0.28)%
-    \Tkg{\lkg_{20}}
+    \Tkg{\lkg_{53}}
     \tikzmarkend{t-STLC-b}
-}{\TyTkg{\lsig_{20}}}
+}{\TyTkg{\lsig_{53}}}
 \end{tabular}

### 2022-11-08 metatheory-condensed.tex
columnize

@@
 constructed from the introduction forms
-$\InhId$, $\InhExt{h}{t}$, $\InhOv{h}{t}$, $\InhInh{h}$, etc.
-\cref{fig:fmltt-selected} shows the beta rule of an exemplary transformer,
+$\small\InhId$, $\small\InhExt{h}{t}$, $\small\InhOv{h}{t}$, $\small\InhInh{h}$, etc.
+\cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
 \ruleref{tmeq/ov/beta}.
-It can be read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
-applying the transformer $\InhOv{h}{t_2}$ overrides the linkage's last
-field~$t_1$ with~$t_2$.
+It states that applying the transformer $\small\InhOv{h}{t_2}$ to
+a linkage of form $\LkgAdd{\lkg}{t_1}$
+overrides the linkage's last field~$t_1$ with~$t_2$.
 For instance, the construction below shows that 
-$\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$ is used as the first step
-in creating a derived family that
+$\small\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$
+is used as the first step in creating a linkage transformer modeling a derived family that
 overrides $\wsig_{\texttt{tm}}$ with an extended

### 2022-11-08 casestudies.tex
add one sentence about parser case study

@@
 directly runnable in Coq.
-We also extract them to OCaml and put them to test over simple queries.
+We also extract them to OCaml and put them to test over queries.
 
@@
 
+In addition to the two case studies above, we also use extensible
+inductive types for modeling extensible context-free grammars and derive
+decision procedures for them.
+
 %\cref{sec:coqexample-stlc}

### 2022-11-08 casestudies.tex
Update Information on AI example.

@@
 In addition to the metatheories, this case study produces abstract
-interpreters that are directly runnable in Coq.
+interpreters that are directly ready for program extraction in Coq.
 
@@
 with pure expressions and impure statements, via \lsti{FInductive}.
-It also defines a big-step interpreter parameterized by a fuel value,
+The semantic is formulated as a functional interpreter like a CEK machine\cite{felleisen1986control}, parameterized by a fuel value,
 via \lsti{FRecursion}.
@@
 A second family, \lsti{ImpGAI}, extends \lsti{Imp} and contains
-\textasciitilde 450~LOC.
-It exports a generic framework for deriving provably sound
-abstract interpreters.
+\textasciitilde 550~LOC.
+It exports a generic framework for deriving
+abstract interpreters with partial correctness.
 The soundness theorem of this generic abstract interpreter, \lsti{analyze},
-is with respect to the big-step interpreter, \lsti{eval}, inherited from \lsti{Imp}.
+is with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
 It states that the abstraction relation \lsti{RState} over a
@@
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-\forall stmt fuel S absS, RState S absS -> RState (eval fuel stmt S) (analyze stmt absS)
+\forall stmt fuel S absS, RState S absS -> RState (eval fuel stmt S) (analyze fuel stmt absS)
 \end{lstlisting}
@@
 abstraction relation, monotonicity of transfer functions, etc.\ 
-unspecified or unproven---a derived family can instantiate these
-``parameters'' by overriding the fields (and also possibly extend the abstract syntax),
+largely unspecified or unproven\EDJ{It is not very relaxed.
+Currently everything is using a dicionary as memory model and stick with that.
+The abstract value for each expression is the one *really* unspecified.}---a derived family can instantiate these
+``parameters'' by overriding appropriate fields (and also possibly extend the abstract syntax),
 to create a sound, runnable abstract interpreter for a (possibly extended) while-language.
@@
 the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
-directly runnable in Coq.
-We also extract them to OCaml and put them to test over queries.
+directly ready for program extraction to OCaml and put them to test over queries.
 

### 2022-11-09 related-work.tex
back to 20 pages
Some remarks on first class/second class

@@
-There exist many approaches to modular mechanization or proof reuse,
+Approaches to modular mechanization or proof reuse exist,
 with different focuses and trade-offs.
@@
 terms by themselves.
-Families as modeled in \TT (namely linkages) are first-class and thus more flexible.
+Families as modeled in \TT (namely linkages) are first-class and thus more flexible.\EDJ{A more unified way is to consider Coq's Family (our surface syntax) second class corresponds to Inheritance Judgement/Linkage Transformer (secondclass in type theory).
+Linkage transformer is the helper function to construct first class linkage easily.
+Even non-derived (surface syntax) family can be considered in this way because they are just inheritance on empty family.}
 The distinction corresponds to class-based vs.\ object-base family

### 2022-11-09 cc-condensed.tex
move a comment
More Comments.

@@
 type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
-When proved in a constructive meta\-logic, the proof of canonicity
-amounts to a normalization function for closed terms of the ground type.
+When canonicity is proved in a constructive meta\-logic, its proof
+amounts to a normalization function for closed terms of the ground type\EDJ{Suggesting use "boolean type".
+Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat).
+We also does specify what is ground type.
+}.
 So canonicity serves to justify the computational nature of the type theory.
@@
 into a logical predicate on closed terms: the predicate includes all
-closed, reducible terms of type~$T$.
+closed, ``reducible''\EDJ{I will add a it is my bad to introduce "reducible" without defining it first.
+But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)} terms of type~$T$.
 The proof is available as a supplemental file in Agda syntax.

### 2022-11-09 related-work.tex
mute discussion about 1st v 2nd class families

--- 2022-11-07 related-work.tex
@@
 \noindentparagraph{Solutions to the expression problem} abound.
+Almost all solutions involve some form of either explicit or implicit parameterization.
 Our approach is the first that applies family polymorphism~\cite{ernst2001family},
@@
 to the context of mechanized proofs.
-Families in \Lang are second-class, in that they are not considered
-terms by themselves.
-Families as modeled in \TT (namely linkages) are first-class and thus more flexible.
-The distinction corresponds to class-based vs.\ object-base family
-polymorphism as summarized before~\cite{cdnw07-tribe}.
+%Families in \Lang are second-class, in that they are not considered
+%terms by themselves.
+%Families as modeled in \TT (namely linkages) are first-class and thus more flexible.\EDJ{A more unified way is to consider Coq's Family (our surface syntax) second class corresponds to Inheritance Judgement/Linkage Transformer (secondclass in type theory).
+Linkage transformer is the helper function to construct first class linkage easily.
+Even non-derived (surface syntax) family can be considered in this way because they are just inheritance on empty family.}
+%The distinction corresponds to class-based vs.\ object-base family
+%polymorphism as summarized before~\cite{cdnw07-tribe}.
 
-\citet{bac2006} address the expression problem in a functional language,
+\citet{bac2006} address the expression problem in a functional setting,
 introducing a language mechanism that allows pattern-matching cases to

### 2022-11-09 casestudies.tex
edits on AI case study

@@
 The code is organized into four families.
-A base family, \lsti{Imp}, defines the abstract syntax of a while-language
-with pure expressions and impure statements, via \lsti{FInductive}.
-The semantic is formulated as a functional interpreter like a CEK machine\cite{felleisen1986control}, parameterized by a fuel value,
-via \lsti{FRecursion}.
+A base family, \lsti{Imp}, defines via \lsti{FInductive} the abstract
+syntax of a while-language with pure expressions and impure statements.
+The semantics is given by an interpreter defined as a CEK-style abstract
+machine~\cite{felleisen1986control} and parameterized by a fuel value.
+Family \lsti{Imp} defines the interpreter via \lsti{FRecursion}.
 It contains \textasciitilde 200~LOC.
 
-A second family, \lsti{ImpGAI}, extends \lsti{Imp} and contains
+A second family, \lsti{ImpGAI}, extends \lsti{Imp} and uses
 \textasciitilde 550~LOC.
 It exports a generic framework for deriving
-abstract interpreters with partial correctness.
-The soundness theorem of this generic abstract interpreter, \lsti{analyze},
-is with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
-It states that the abstraction relation \lsti{RState} over a
+abstract interpreters with partial-correctness guarantees.
+Soundness of the abstract interpreter, \lsti{analyze},
+is stated with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
+The theorem says that the abstraction relation \lsti{RState} over a
 concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
@@
 \begin{centered}
-\begin{minipage}{.86\textwidth}
+\begin{minipage}{.91\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
@@
 abstraction relation, monotonicity of transfer functions, etc.\ 
-largely unspecified or unproven\EDJ{It is not very relaxed.
-Currently everything is using a dicionary as memory model and stick with that.
-The abstract value for each expression is the one *really* unspecified.}---a derived family can instantiate these
-``parameters'' by overriding appropriate fields (and also possibly extend the abstract syntax),
-to create a sound, runnable abstract interpreter for a (possibly extended) while-language.
+largely unspecified or unproven---a derived family can further bind
+these ``parameters'' by overriding appropriate fields (and also possibly
+extend the abstract syntax), to create a sound, runnable abstract
+interpreter for a (possibly extended) while-language.
+%\EDJ{It is not very relaxed. Currently everything is using a dicionary as memory model and stick with that. The abstract value for each expression is the one *really* unspecified.}
 
@@
 with natural number arithmetic,
-and instantiates the generic abstract interpreter to perform constant propagation.
+and further binds the generic abstract interpreter to perform constant propagation.
 
@@
 the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
-directly ready for program extraction to OCaml and put them to test over queries.
+ready for program extraction to OCaml.
+We put them to test over queries.
 
-We note that our plugin cannot yet allow the expression of two abstract
+We note that our plugin cannot yet allow the expression of two derived abstract
 interpreters nested within the same family.

### 2022-11-09 coqimpl.tex
edits on AI case study

@@
 Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)
-can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a module),
+can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a Coq module),
 as the last line of \cref{fig:stlcfix-compiled} shows.

### 2022-11-09 body.tex
Save.
Collect Typing Rules in One page.
Start Reworking Metatheory QIIT possibly removable altogether
compilation of family STLC
selected syntax
stuff
Start Canonicity/Consistency.
edits and tightening
sneak peek of linkage transformer use
Update Appendix.

@@
 
-% Macro for all use, the macro for metatheory syntax
-\newcommand{\denotes}[1]{{\llbracket {#1} \rrbracket}}
-\newcommand{\denotesS}[1]{{{\llbracket {#1} \rrbracket}_S}}
-\newcommand{\goodCtx}[2]{{ {#1} \ \vdash }}
-\newcommand{\goodType}[3]{{ {#1} \vdash {#2} }}
-\newcommand{\goodTerm}[3]{{ {#1} \vdash {#2} : {#3} }}
-\newcommand{\goodSub}[3]{{ {#1} \vdash {#2} : {#3} }}
-\newcommand{\goodSig}[3]{{ {#1} \vdash {#2} \ \  Sig^{#3} }}
-\newcommand{\goodWSig}[3]{{ {#1} \vdash {#2} \ \ WSig^{#3} }}
-\newcommand{\goodSeal}[4]{{ {#1} \vdash {#2} : {#3} \  |\  {#4} }}
-\newcommand{\goodInh}[4]{{ {#1} \vdash {#2} : {#3} \twoheadrightarrow {#4}}}
-\newcommand{\nat}{\mathbf{N}}
-
-\newcommand{\cU}{{\mathcal{U}}}
-\newcommand{\cB}{{\mathbb{B}}}
-\newcommand{\cL}{{\mathcal{L}}}
-\newcommand{\cC}{{\mathcal{C}}}
-\newcommand{\cCt}{{\mathcal{C}_t}}
-\newcommand{\bW}{{\mathbb{W}}}
-
 \section{Language Design}
@@
 
-\section{Prototypical Coq Implementation}
+\section{Compiling Family Polymorphism to Parameterized Modules}
 % we need to make this section before Metatheory
@@
 
-\section{Metatheory}
+\section{\TT: A Core Dependent Type Theory}
+\label{sec:metatheory2}
+\input{metatheory-condensed.tex}
 
-\label{sec:metatheory}
-\input{metatheory.tex}
-
-\section{Example}\label{sec:coqexample}
-\input{coq-example.tex}
+\section{Case Studies}
+\label{sec:casestudies}
+\input{casestudies.tex}
 
@@
 
-\section{Limitations and Conclusion}
+\section{Conclusion}
 \label{sec:conclusion}
 
-It is hard to write modular, extensible code and proofs.
-We have presented the first language design that
-
-... make exciting directions for future work.
+\input{conclusion.tex}
 
@@
 For syntactic translation, please check \\ \href{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}.
+The syntactic translation here is to translate most of the novel feature introduced in this paper into a standard MLTT.
+The essence here is to translate linkage into a sigma type  with dependent function inside---and the dependent function is simulating the nature of ``late-binding'' of each field.
+
 This is the first QIIT-model we give.
-A QIIT-model can be consider as an arbitrary mapping from the QIIT data(our syntax in this case), also respecting the quotient(judgemental equality in this case).
-Here we use (a subpart of) QIIT data itself as the destination of the mapping.
+A QIIT-model can be roughly consider as an arbitrary mapping from the QIIT data(our syntax in this case), also respecting the quotient(judgemental equality in this case).
+Here, the syntactic translation model is mapping to (a subpart of) the original QIIT data (itself).
 
@@
 This is mainly following \citet{coquand2018canonicity,sterling2019algebraic}, but we carried out our argument in a type-theortic framework.
-We call it canonicity model to emphasize that it is a model used to prove canonicity, it is also called reducibility argument, as an instance of proof-relevant logical relation.
-This argument can be considered as a verbose/more concrete version of \citet{kaposi2019gluing}.
+In other references, it is also called reducibility argument, as an instance of proof-relevant logical relation.
+Our model can be considered as a verbose/more concrete version of \citet{kaposi2019gluing}.
+
+For both consistency and canonicity model, the key part is still about handling the novel feature.
+Thanks to the insight from syntactic translation, we can always inspired by consistency and canonicity model of sigma type and dependent function in standard MLTT.
 
 We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.
+
+\section{Typing Rules in one Figure}
+\include{all-typing-rule.tex}
+
+\section{Metatheory, in detail}
+
+\label{sec:metatheory}
+\input{metatheory.tex}
+
+\section{Example in Metatheory}\label{sec:coqexample}
+\input{coq-example.tex}

### 2022-11-09 cc-condensed.tex
reply

--- 2022-11-04 cc-condensed.tex
@@
 type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
-When proved in a constructive meta\-logic, the proof of canonicity
-amounts to a normalization function for closed terms of the ground type.
+When canonicity is proved in a constructive meta\-logic, its proof
+amounts to a normalization function for closed terms of the ground type.\EDJ{Suggesting use "boolean type".
+Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat).
+We also does specify what is ground type.
+}\YZreply{I think it is clear from the context and from the article 'the' that 'the ground type' refers to B.}
 So canonicity serves to justify the computational nature of the type theory.
@@
 into a logical predicate on closed terms: the predicate includes all
-closed, reducible terms of type~$T$.
+closed, ``reducible'' terms of type~$T$.\EDJ{I will add a it is my bad to introduce "reducible" without defining it first.
+But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}
 The proof is available as a supplemental file in Agda syntax.

### 2022-11-09 abstract.tex
address comment in abstract
took a pass on §1–3

@@
 proofs to be polymorphic to their enclosing families.
-Our development addresses significant technical challenges that arise
+Our development addresses technical challenges that arise
 from the underlying language of a proof assistant being simultaneously
@@
 \item a formalization of the language design as a constructive type theory with inductive types,
-\item proofs of strong guarantees\EDJ{We don't yet have normalization though} afforded by the type theory,
+\item consistency and canonicity of the type theory,
+%\EDJ{We don't yet have normalization though}
 and 

### 2022-11-09 challenge.tex
AI case study
took a pass on §1–3

@@
 \headinglabel{C4}{chg:tooling}.\,
-User experience and tooling.%
+User experience and system implementation.%
 }
@@
 Interactive theorem proving and tactic programming are central to
-the idiomatic programming experience in a proof assistant.
+a typical programming experience with a proof assistant.
 A language design integrating family polymorphism should be compatible
-with these programming styles.
-In particular, it should be possible in our tool to
+with these forms of programming.
+In particular, it should be possible in our system to
 incrementally navigate through vernacular commands and, moreover, construct proofs
@@
 even in the middle of a family definition.
-In addition, it should remain possible to normalize terms that mention families.
+Last but not least, in addition to producing theorems,
+it should be possible for terms defined in families to possess computational content.
 

### 2022-11-09 introduction.tex
add one sentence about parser case study
took a pass on §1–3

@@
 %
-The programmer could also turn to design patterns~\cite{delaware2011,delaware2013} that
-use software product lines or Church encodings.
+The programmer could also turn to design patterns and encodings~\cite{delaware2011,delaware2013,schwaab2013modular,keuchel2013generic,forsta2020}.
 But they tend to require heavy lifting from the programmer to make code
-extensible, often leading to non-idiomatic programming styles.
+extensible and often lead to non-idiomatic programming styles.
 
@@
 
-\paragraph{Contributions}
+\noindentparagraph{Contributions}
 
@@
 
-\begin{itemize}[leftmargin=3.5ex]
+\begin{itemize}%[leftmargin=3.5ex]
+    [labelsep=*,leftmargin=1pc,itemsep=3pt]
 
@@
 family polymorphism with the rigor of a proof assistant,
-while retaining an idiomatic programming style.
+while largely retaining an idiomatic programming style.
 
@@
 They show how our language design naturally solves the EP and
-enables a high degree of reuse and extensibility
-for proof engineering.
+enables a good amount of reuse and extensibility
+for mechanizing proofs.
 

### 2022-11-09 stlc-linkTran-typing.tex
Working on Example for Linkage Transformer.

\begin{table}[]
  \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   &
    $i$ &
    $A_i$ &
    $A_i'$ &
    $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ &
    $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i}{A_i}$ &
    $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$ \\ \hline
  Family STLC\_bool &
    0 &
     &
     &
    $\{\}$ &
    $\{\}$ &
    $\InhId$ \\ \hline
  FInductive tm += &
    1 &
    $[]$ &
    $[]$ &
    $\{"tm" : \TyS{\wcode{\wsig_1}}\}$ &
    $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$ &
    $\InhOv{}{}$ \\ \hline
  | .. | tm\_false &
    2 &
    $["tm" : \cU,$ .. &
    $["tm" : \cU, ..$ &
     &
     &
     \\ \hline
   &
     &
    $"tm_unit"]$ &
    $"tm_unit", "tm_true" ]$ &
     &
     &
     \\ \hline
  FRecursion subst &
     &
     &
     &
     &
     &
     \\ \hline
  (* Inherit ... *) &
    $\alpha$ &
     &
     &
    $\RecSig{\wsig_1}{"tm"}$ &
    $\RecSig{\wsig_1}{"tm"}$ &
    $\InhId$ \\ \hline
  Case tm\_true &
    $\beta$ &
    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
    $\sigma_{\alpha}$ &
    $\{.. "tm:true": \}$ &
    $\InhExt{}{}$ \\ \hline
  End subst &
    3 &
    $A_2$ &
    $A_2'$ &
    $\{.. subst' : \TyLkg{\lsig_\alpha}\}$ &
    $\{.. subst' : \TyLkg{\lsig_\beta'}\}$ &
    $\InhNest{}{}$ \\ \hline
   &
    4 &
    $\TyTkg{\lsig_3}$ &
    $\TyTkg{\lsig_3'}$ &
    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
    $\InhOv{}{}$ \\ \hline
  (* Inherit t1 *) &
    5 &
    $["tm":\cU, $ &
    $["tm":\cU, $ &
    $\{.. "t"_1 : \El{"tm"}\}$ &
    $\{.. "t"_1 : \El{"tm"}\}$ &
    $\InhInh{}$ \\ \hline
   &
     &
    $.. subst]$ &
    $.. subst]$ &
     &
     &
     \\ \hline
  \end{tabular}
  \end{table}

### 2022-11-09 stlc-linkTran-typing.tex
Working on LT example
Working on LT example Next step is to beautify it.
Working on LT example.

@@
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
 \begin{table}[]
-  \begin{tabular}{|l|l|l|l|l|l|l|}
+  \begin{tabular}{|l|l|l|l|}
   \hline
-   &
-    $i$ &
-    $A_i$ &
-    $A_i'$ &
-    $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ &
-    $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i}{A_i}$ &
-    $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$ \\ \hline
-  Family STLC\_bool &
-    0 &
-     &
-     &
-    $\{\}$ &
-    $\{\}$ &
-    $\InhId$ \\ \hline
-  FInductive tm += &
-    1 &
-    $[]$ &
-    $[]$ &
-    $\{"tm" : \TyS{\wcode{\wsig_1}}\}$ &
-    $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$ &
-    $\InhOv{}{}$ \\ \hline
-  | .. | tm\_false &
-    2 &
-    $["tm" : \cU,$ .. &
-    $["tm" : \cU, ..$ &
-     &
-     &
-     \\ \hline
-   &
-     &
-    $"tm_unit"]$ &
-    $"tm_unit", "tm_true" ]$ &
-     &
-     &
-     \\ \hline
-  FRecursion subst &
-     &
-     &
-     &
-     &
-     &
-     \\ \hline
-  (* Inherit ... *) &
-    $\alpha$ &
-     &
-     &
-    $\RecSig{\wsig_1}{"tm"}$ &
-    $\RecSig{\wsig_1}{"tm"}$ &
-    $\InhId$ \\ \hline
-  Case tm\_true &
-    $\beta$ &
-    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
-    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
-    $\sigma_{\alpha}$ &
-    $\{.. "tm:true": \}$ &
-    $\InhExt{}{}$ \\ \hline
-  End subst &
-    3 &
-    $A_2$ &
-    $A_2'$ &
-    $\{.. subst' : \TyLkg{\lsig_\alpha}\}$ &
-    $\{.. subst' : \TyLkg{\lsig_\beta'}\}$ &
-    $\InhNest{}{}$ \\ \hline
-   &
-    4 &
-    $\TyTkg{\lsig_3}$ &
-    $\TyTkg{\lsig_3'}$ &
-    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
-    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
-    $\InhOv{}{}$ \\ \hline
-  (* Inherit t1 *) &
-    5 &
-    $["tm":\cU, $ &
-    $["tm":\cU, $ &
-    $\{.. "t"_1 : \El{"tm"}\}$ &
-    $\{.. "t"_1 : \El{"tm"}\}$ &
-    $\InhInh{}$ \\ \hline
-   &
-     &
-    $.. subst]$ &
-    $.. subst]$ &
-     &
-     &
-     \\ \hline
+                          & $i$      & $A_i$             & $A_i'$             \\ \hline
+  ***Family STLC\_bool... & 0        &                   &                    \\ \hline
+  ***FInductive tm +=     & 1        & $[]$              & $[]$               \\ \hline
+                          &          & $["tm" : \cU,$ .. & $["tm" : \cU, ..$  \\
+  \multirow{-2}{*}{*** .. | tm\_true | ..} & \multirow{-2}{*}{2} & $"tm_unit"]$                              & $"tm_unit", "tm_true", .. ]$              \\ \hline
+  FRecursion subst***     &          &                   &                    \\ \hline
+  \rowcolor[HTML]{E0D7D7} 
+  (* Inherit ... *)***    & $\alpha$ &                   &                    \\ \hline
+  \rowcolor[HTML]{E0D7D7} 
+  Case tm\_true***                         & $\beta$             & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          \\ \hline
+  End subst***            & 3        & $A_2$             & $A_2'$             \\ \hline
+                          & 4        & $\TyTkg{\lsig_3}$ & $\TyTkg{\lsig_3'}$ \\ \hline
+                                           &                     & $["tm":\cU, .. subst : \El("tm") \to id $ & $["tm":\cU, .. subst : \El("tm") \to id $ \\
+  \multirow{-2}{*}{(* Inherit t1 *)***}    & \multirow{-2}{*}{5} & $ \to \El(tm) \to \El(tm) ]$              & $ \to \El(tm) \to \El(tm) ]$              \\ \hline
   \end{tabular}
   \end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+  \begin{tabular}{|l|l|l|l|}
+  \hline
+   & $i$ & $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ & $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i}{A_i}$ \\ \hline
+  ***Family STLC\_bool...                  & 0                   & $\{\}$                                   & $\{\}$                                   \\ \hline
+  ***FInductive tm +=                      & 1                   & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$       & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$      \\ \hline
+                                           &                     &                                          &                                          \\
+  \multirow{-2}{*}{*** .. | tm\_true | ..} & \multirow{-2}{*}{2} &                                          &                                          \\ \hline
+  FRecursion subst***                      &                     &                                          &                                          \\ \hline
+  \rowcolor[HTML]{E0D7D7} 
+  (* Inherit ... *)***                     & $\alpha$            & $\RecSig{\wsig_1}{"tm"}$                 & $\RecSig{\wsig_1}{"tm"}$                 \\ \hline
+  \rowcolor[HTML]{E0D7D7} 
+  Case tm\_true***                         & $\beta$             & $\sigma_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \hline
+  End subst***                             & 3                   & $\{.. subst' : \TyLkg{\lsig_\alpha}\}$   & $\{.. subst' : \TyLkg{\lsig_\beta'}\}$   \\ \hline
+                                           & 4                   & $\{.. subst : \El{"tm"} \to id \to ..\}$ & $\{.. subst : \El{"tm"} \to id \to ..\}$ \\ \hline
+                                           &                     & $\{.. "t"_1 : \El{"tm"}\}$               & $\{.. "t"_1 : \El{"tm"}\}$               \\
+  \multirow{-2}{*}{(* Inherit t1 *)***}    & \multirow{-2}{*}{5} &                                          &                                          \\ \hline
+  \end{tabular}
+  \end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\begin{tabular}{|l|l|l|}
+\hline
+                                         & $i$                 & $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$             \\ \hline
+***Family STLC\_bool...                  & 0                   & $\InhId$                                             \\ \hline
+***FInductive tm +=                      & 1                   & $\InhOv{i_1}{\wcode{\wsig_1'}}$                      \\ \hline
+                                         &                     &                                                      \\
+\multirow{-2}{*}{*** .. | tm\_true | ..} & \multirow{-2}{*}{2} &                                                      \\ \hline
+FRecursion subst***                      &                     &                                                      \\ \hline
+\rowcolor[HTML]{E0D7D7} 
+(* Inherit ... *)***                     & $\alpha$            & $\InhId$                                             \\ \hline
+\rowcolor[HTML]{E0D7D7} 
+Case tm\_true***                         & $\beta$             & $\InhExt{i_\alpha}{..}$                              \\ \hline
+End subst***                             & 3                   & $\InhNest{i_2}{i_\beta}$ with $\uparrow s_3$         \\ \hline
+                                         & 4                   & $\InhOv{i_3}{\lambda t. \Wrec{\wsig_1'}{t}{subst'}}$ \\ \hline
+                                         &                     & $\InhInh{i_4}$ with $\uparrow s_4$                   \\
+\multirow{-2}{*}{(* Inherit t1 *)***}    & \multirow{-2}{*}{5} &                                                      \\ \hline
+\end{tabular}
+\end{table}

### 2022-11-09 stlc-linkTran-typing.tex
Working on LT example, Almost done.

--- 2022-11-09 stlc-linkTran-typing.tex
@@
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
 \begin{table}[]
-  \begin{tabular}{|l|l|l|l|l|l|l|}
+\begin{tabular}{|l|l|l|l|}
 \hline
-   &
-    $i$ &
-    $A_i$ &
-    $A_i'$ &
-    $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ &
-    $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i}{A_i}$ &
-    $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$ \\ \hline
-  Family STLC\_bool &
-    0 &
-     &
-     &
-    $\{\}$ &
-    $\{\}$ &
-    $\InhId$ \\ \hline
-  FInductive tm += &
-    1 &
-    $[]$ &
-    $[]$ &
-    $\{"tm" : \TyS{\wcode{\wsig_1}}\}$ &
-    $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$ &
-    $\InhOv{}{}$ \\ \hline
-  | .. | tm\_false &
-    2 &
-    $["tm" : \cU,$ .. &
-    $["tm" : \cU, ..$ &
-     &
-     &
-     \\ \hline
-   &
-     &
-    $"tm_unit"]$ &
-    $"tm_unit", "tm_true" ]$ &
-     &
-     &
-     \\ \hline
-  FRecursion subst &
-     &
-     &
-     &
-     &
-     &
-     \\ \hline
-  (* Inherit ... *) &
-    $\alpha$ &
-     &
-     &
-    $\RecSig{\wsig_1}{"tm"}$ &
-    $\RecSig{\wsig_1}{"tm"}$ &
-    $\InhId$ \\ \hline
-  Case tm\_true &
-    $\beta$ &
-    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
-    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
-    $\sigma_{\alpha}$ &
-    $\{.. "tm:true": \}$ &
-    $\InhExt{}{}$ \\ \hline
-  End subst &
-    3 &
-    $A_2$ &
-    $A_2'$ &
-    $\{.. subst' : \TyLkg{\lsig_\alpha}\}$ &
-    $\{.. subst' : \TyLkg{\lsig_\beta'}\}$ &
-    $\InhNest{}{}$ \\ \hline
-   &
-    4 &
-    $\TyTkg{\lsig_3}$ &
-    $\TyTkg{\lsig_3'}$ &
-    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
-    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
-    $\InhOv{}{}$ \\ \hline
-  (* Inherit t1 *) &
-    5 &
-    $["tm":\cU, $ &
-    $["tm":\cU, $ &
-    $\{.. "t"_1 : \El{"tm"}\}$ &
-    $\{.. "t"_1 : \El{"tm"}\}$ &
-    $\InhInh{}$ \\ \hline
-   &
-     &
-    $.. subst]$ &
-    $.. subst]$ &
-     &
-     &
-     \\ \hline
+\rowcolor[HTML]{FFFFE6} 
+Surface Syntax          & $i$      & $A_i$             & $A_i'$                       \\ \hline
+***Family STLC\_bool... & 0        &                   &                              \\ \cline{2-4} 
+***FInductive tm +=     & 1        & $[]$              & $[]$                         \\ \cline{2-4} 
+*** .. | tm\_true | .. & 2       & $["tm" : \cU,$ ..                         & $["tm" : \cU, ..$                         \\
+                        &          & $"tm_unit"]$      & $"tm_unit", "tm_true", .. ]$ \\ \cline{2-4} 
+FRecursion subst***     &          &                   &                              \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+(* Inherit ... *)***    & $\alpha$ &                   &                              \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+Case tm\_true***       & $\beta$ & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          \\ \cline{2-4} 
+End subst***            & 3        & $A_2$             & $A_2'$                       \\ \cline{2-4} 
+                        & 4        & $\TyTkg{\lsig_3}$ & $\TyTkg{\lsig_3'}$           \\ \cline{2-4} 
+(* Inherit t1 *)***    & 5       & $["tm":\cU, .. subst : \El{"tm"} \to id $ & $["tm":\cU, .. subst : \El{"tm"} \to id $ \\
+                       &         & $ \to \El{"tm"} \to \El{"tm"} ]$          & $ \to \El{"tm"} \to \El{"tm"} ]$          \\ \hline
 \end{tabular}
 \end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|}
+\hline
+\rowcolor[HTML]{FFFFE6} 
+Surface Syntax & $i$ & $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ & $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i}{A_i}$ \\ \hline
+***Family STLC\_bool... & 0        & $\{\}$                                   & $\{\}$                                   \\ \cline{2-4} 
+***FInductive tm +=     & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$       & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$      \\ \cline{2-4} 
+*** .. | tm\_true | ..  & 2        &                                          &                                          \\
+                        &          &                                          &                                          \\ \cline{2-4} 
+FRecursion subst***     &          &                                          &                                          \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+(* Inherit ... *)***    & $\alpha$ & $\RecSig{\wsig_1}{"tm"}$                 & $\RecSig{\wsig_1}{"tm"}$                 \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+Case tm\_true***        & $\beta$  & $\sigma_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \cline{2-4} 
+End subst***            & 3        & $\{.. subst' : \TyLkg{\lsig_\alpha}\}$   & $\{.. subst' : \TyLkg{\lsig_\beta'}\}$   \\ \cline{2-4} 
+                        & 4        & $\{.. subst : \El{"tm"} \to id \to ..\}$ & $\{.. subst : \El{"tm"} \to id \to ..\}$ \\ \cline{2-4} 
+(* Inherit t1 *)***     & 5        & $\{.. "t"_1 : \El{"tm"}\}$               & $\{.. "t"_1 : \El{"tm"}\}$               \\
+                        &          &                                          &                                          \\ \hline
+\end{tabular}
+\end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+  \begin{tabular}{|l|l|l|}
+  \hline
+  \rowcolor[HTML]{FFFFE6} 
+  Surface Syntax          & $i$      & $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$             \\ \hline
+  ***Family STLC\_bool... & 0        & $\InhId$                                             \\ \cline{2-3} 
+  ***FInductive tm +=     & 1        & $\InhOv{i_1}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
+  *** .. | tm\_true | ..  & 2        &                                                      \\
+                          &          &                                                      \\ \cline{2-3} 
+  FRecursion subst***     &          &                                                      \\ \cline{2-3} 
+  \rowcolor[HTML]{E0D7D7} 
+  (* Inherit ... *)***    & $\alpha$ & $\InhId$                                             \\ \cline{2-3} 
+  \rowcolor[HTML]{E0D7D7} 
+  Case tm\_true***        & $\beta$  & $\InhExt{i_\alpha}{..}$                              \\ \cline{2-3} 
+  End subst***            & 3        & $\InhNest{i_2}{i_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
+                          & 4        & $\InhOv{i_3}{\lambda t. \Wrec{\wsig_1'}{t}{subst'}}$ \\ \cline{2-3} 
+  (* Inherit t1 *)***     & 5        & $\InhInh{i_4}$ with $\uparrow s_4$                   \\
+                          &          &                                                      \\ \hline
+  \end{tabular}
+  \end{table}

### 2022-11-09 stlc-linkTran-typing.tex
LT example figure is done.
Finishing Linkage Transformer Example.

--- 2022-11-09 stlc-linkTran-typing.tex
@@
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
 \begin{table}[]
-  \begin{tabular}{|l|l|l|l|l|l|l|}
+\begin{tabular}{|l|l|l|l|}
 \hline
-   &
-    $i$ &
-    $A_i$ &
-    $A_i'$ &
-    $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ &
-    $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i}{A_i}$ &
-    $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$ \\ \hline
-  Family STLC\_bool &
-    0 &
-     &
-     &
-    $\{\}$ &
-    $\{\}$ &
-    $\InhId$ \\ \hline
-  FInductive tm += &
-    1 &
-    $[]$ &
-    $[]$ &
-    $\{"tm" : \TyS{\wcode{\wsig_1}}\}$ &
-    $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$ &
-    $\InhOv{}{}$ \\ \hline
-  | .. | tm\_false &
-    2 &
-    $["tm" : \cU,$ .. &
-    $["tm" : \cU, ..$ &
-     &
-     &
-     \\ \hline
-   &
-     &
-    $"tm_unit"]$ &
-    $"tm_unit", "tm_true" ]$ &
-     &
-     &
-     \\ \hline
-  FRecursion subst &
-     &
-     &
-     &
-     &
-     &
-     \\ \hline
-  (* Inherit ... *) &
-    $\alpha$ &
-     &
-     &
-    $\RecSig{\wsig_1}{"tm"}$ &
-    $\RecSig{\wsig_1}{"tm"}$ &
-    $\InhId$ \\ \hline
-  Case tm\_true &
-    $\beta$ &
-    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
-    $\TyTkg{\RecSig{\wsig_1}{"tm"}}$ &
-    $\sigma_{\alpha}$ &
-    $\{.. "tm:true": \}$ &
-    $\InhExt{}{}$ \\ \hline
-  End subst &
-    3 &
-    $A_2$ &
-    $A_2'$ &
-    $\{.. subst' : \TyLkg{\lsig_\alpha}\}$ &
-    $\{.. subst' : \TyLkg{\lsig_\beta'}\}$ &
-    $\InhNest{}{}$ \\ \hline
-   &
-    4 &
-    $\TyTkg{\lsig_3}$ &
-    $\TyTkg{\lsig_3'}$ &
-    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
-    $\{.. subst : \El{"tm"} \to \mathbb{N} \to ..\}$ &
-    $\InhOv{}{}$ \\ \hline
-  (* Inherit t1 *) &
-    5 &
-    $["tm":\cU, $ &
-    $["tm":\cU, $ &
-    $\{.. "t"_1 : \El{"tm"}\}$ &
-    $\{.. "t"_1 : \El{"tm"}\}$ &
-    $\InhInh{}$ \\ \hline
-   &
-     &
-    $.. subst]$ &
-    $.. subst]$ &
-     &
-     &
-     \\ \hline
+\rowcolor[HTML]{FFFFE6} 
+Surface Syntax          & $i$      & $A_i$             & $A_i'$                       \\ \hline
+\lsti!Family STLC\_bool...! & 0        &                   &                              \\ \cline{2-4} 
+\lsti!FInductive tm +=!     & 1        & $[]$              & $[]$                         \\ \cline{2-4} 
+\lsti!| .. | tm\_true | ..! & 2       & $["tm" : \cU, "tm_var" : id \to \El{"tm"} .. $                          & $["tm" : \cU,  ..$                         \\
+                      &          & $"tm_unit" : \El{"tm"}]$      & $"tm_unit", "tm_true", .. ]$ \\ \cline{2-4} 
+\lsti!FRecursion subst!     &          &                   &                              \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+\lsti!(* Inherit ... *)!    & $\alpha$ &                   &                              \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+\lsti!Case tm\_true!       & $\beta$ & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          \\ \cline{2-4} 
+\lsti!End subst!            & 3        & $A_2$             & $A_2'$                       \\ \cline{2-4} 
+                      & 4        & $\TyTkg{\lsig_3}$ & $\TyTkg{\lsig_3'}$           \\ \cline{2-4} 
+\lsti!(* Inherit t1 *)!    & 5       & $["tm":\cU, .. subst : \El{"tm"} \to id $ & $["tm":\cU, .. subst : \El{"tm"} \to id $ \\
+                      &         & $ \to \El{"tm"} \to \El{"tm"} ]$          & $ \to \El{"tm"} \to \El{"tm"} ]$          \\ \hline
 \end{tabular}
 \end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|}
+\hline
+\rowcolor[HTML]{FFFFE6} 
+Surface Syntax & $i$ & $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ & $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i'}{A_i'}$ \\ \hline
+\lsti!Family STLC\_bool...! & 0        & $\{\}$                                   & $\{\}$                                   \\ \cline{2-4} 
+\lsti!FInductive tm +=!     & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$       & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$      \\ \cline{2-4} 
+\lsti!| .. | tm\_true | ..!  & 2        &                                          &                                          \\
+                      &          &                                          &                                          \\ \cline{2-4} 
+\lsti!FRecursion subst!     &          &                                          &                                          \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+\lsti!(* Inherit ... *)!    & $\alpha$ & $\RecSig{\wsig_1}{"tm"}$                 & $\RecSig{\wsig_1}{"tm"}$                 \\ \cline{2-4} 
+\rowcolor[HTML]{E0D7D7} 
+\lsti!Case tm\_true!        & $\beta$  & $\sigma_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \cline{2-4} 
+\lsti!End subst!            & 3        & $\{.. subst' : \TyLkg{\lsig_\alpha}\}$   & $\{.. subst' : \TyLkg{\lsig_\beta'}\}$   \\ \cline{2-4} 
+                      & 4        & $\{.. subst : \El{"tm"} \to id \to ..\}$ & $\{.. subst : \El{"tm"} \to id \to ..\}$ \\ \cline{2-4} 
+\lsti!(* Inherit t1 *)!     & 5        & $\{.. "t"_1 : \El{"tm"}\}$               & $\{.. "t"_1 : \El{"tm"}\}$               \\
+                      &          &                                          &                                          \\ \hline
+\end{tabular}
+\end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\begin{tabular}{|l|l|l|}
+\hline
+\rowcolor[HTML]{FFFFE6} 
+Surface Syntax          & $i$      & $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$             \\ \hline
+\lsti!Family STLC\_bool...! & 0        & $\InhId$                                             \\ \cline{2-3} 
+\lsti!FInductive tm +=!     & 1        & $\InhOv{i_1}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
+\lsti!| .. | tm\_true | ..!  & 2        &                                                      \\
+                        &          &                                                      \\ \cline{2-3} 
+\lsti!FRecursion subst!     &          &                                                      \\ \cline{2-3} 
+\rowcolor[HTML]{E0D7D7} 
+\lsti!(* Inherit ... *)!    & $\alpha$ & $\InhId$                                             \\ \cline{2-3} 
+\rowcolor[HTML]{E0D7D7} 
+\lsti!Case tm\_true!        & $\beta$  & $\InhExt{i_\alpha}{..}$                              \\ \cline{2-3} 
+\lsti!End subst!            & 3        & $\InhNest{i_2}{i_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
+                        & 4        & $\InhOv{i_3}{\lambda t. \Wrec{\wsig_1'}{t}{subst'}}$ \\ \cline{2-3} 
+\lsti!(* Inherit t1 *)!     & 5        & $\InhInh{i_4}$ with $\uparrow s_4$                   \\
+                        &          &                                                      \\ \hline
+\end{tabular}
+\end{table}
+
+We take "Family STLC_bool" extending "STLC" as an example.
+We list out the surface syntax and their translation in out \TT.
+In our figure, $\lsig_i$ stands for the parent signature (the signature before the transformation); and $\lsig_i'$ stands for the signature after the transformation.
+
+Before we sketch out the signature, we need to specify \textit{the sealing premise} $s$ in \ruleref{lsig/add}. However, we omit $s$ as the information of $A$ is enough to specify $s$.
+
+When sketching signature $\lsig$, we use curly bracket as a shorthand for signature and square bracket as a shorthand for (deeply) nested sigma type (usually used in the case of $\TyTkg{\cdot}$).
+Because every $\lsig_i$ is always constructed from $\lsig_{i-1}$ so in the figure we omit the previous fields in each signature.
+The only exceptions are the rows indexed by $\alpha$ and $\beta$.
+These two rows are describing a nested family for case handlers, and $\lsig_\beta$ is constructed by $\lsig_\alpha$.
+
+Our final target is to construct a term $\goodInh{\cdot}{?}{\lsig_5}{\lsig_5'}$ using the four constructors.
+There are two upcasting operations required $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_4'}{\uparrow s_4}{A_4}$ (omitting debruijn indices here) for checking the compatibility between $A_i$ and $A_i'$.

### 2022-11-09 stlc-linkTran-typing.tex
Finishing Linkage Transformer example.

@@
 
-Our final target is to construct a term $\goodInh{\cdot}{?}{\lsig_5}{\lsig_5'}$ using the four constructors.
-There are two upcasting operations required $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_4'}{\uparrow s_4}{A_4}$ (omitting debruijn indices here) for checking the compatibility between $A_i$ and $A_i'$.
+Our final target is to construct a term $\goodInh{\cdot}{?}{\lsig_5}{\lsig_5'}$ using the four linkage transformers' rules.
+There are two upcasting operations $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_4'}{\uparrow s_4}{A_4}$ (omitting debruijn indices here) for checking the compatibility between $A_i$ and $A_i'$.

### 2022-11-09 stlc-linkTran-typing.tex
Fix typo in the example.

@@
 \lsti!| .. | tm\_true | ..! & 2       & $["tm" : \cU, "tm_var" : id \to \El{"tm"} .. $                          & $["tm" : \cU,  ..$                         \\
-                      &          & $"tm_unit" : \El{"tm"}]$      & $"tm_unit", "tm_true", .. ]$ \\ \cline{2-4} 
+                      &          & $"tm_unit" : \El{"tm"}]$      & $"tm_unit", "tm_true" : \El{"tm"}, .. ]$ \\ \cline{2-4} 
 \lsti!FRecursion subst!     &          &                   &                              \\ \cline{2-4} 
@@
 \lsti!Family STLC\_bool...! & 0        & $\InhId$                                             \\ \cline{2-3} 
-\lsti!FInductive tm +=!     & 1        & $\InhOv{i_1}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
+\lsti!FInductive tm +=!     & 1        & $\InhOv{i_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
 \lsti!| .. | tm\_true | ..!  & 2        &                                                      \\
@@
 \lsti!Case tm\_true!        & $\beta$  & $\InhExt{i_\alpha}{..}$                              \\ \cline{2-3} 
-\lsti!End subst!            & 3        & $\InhNest{i_2}{i_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
+\lsti!End subst!            & 3        & $\InhNest{i_1}{i_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
                         & 4        & $\InhOv{i_3}{\lambda t. \Wrec{\wsig_1'}{t}{subst'}}$ \\ \cline{2-3} 
@@
 There are two upcasting operations $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_4'}{\uparrow s_4}{A_4}$ (omitting debruijn indices here) for checking the compatibility between $A_i$ and $A_i'$.
+
+Note that in our table, the Row 2 only has information of $A_2$ and $A_2'$ (i.e.
+there is no $\lsig_2$ and $\lsig_2'$).
+We especially use one row to indicate that $s_2$ and $s_2'$ is ``sealing/abstracting'' $\lsig_1' = \{"tm" : \TyS{\wcode{\wsig_1'}}\}$, into a nested dependent pair of "tm" type and its constructors $A_2' = ["tm" : \cU, "tm_var" : id \to \El{"tm"}, "tm_lam" : id \to \El{"tm"} \to \El{"tm"}, "tm_app" : \El{"tm"} \to \El{"tm"} \to \El{"tm"}]$.

### 2022-11-09 related-work.tex
blume et al.

@@
 
-\citet{bac2006} address the expression problem in a functional setting,
-introducing a language mechanism that allows pattern-matching cases to
-be added retroactively in accordance with extensions to variant types.
+\citet{bac2006} address the expression problem for a core subset of Standard~ML
+by combining explicitly coded open recursion with a mechanism that
+allows pattern-matching cases to be defined separately and combined
+later.
 Our \lsti{FRecursion} and \lsti{FInduction} commands achieve a similar
-functionality, with families bestowing additional organizational power.
+functionality, with families making open recursion implicit and
+bestowing organizational power.
 

### 2022-11-09 stlc-linkTran-typing.tex
Fix typo in LT example.
\_ -> _
move tabulars
Fix typo in Table.
Fix typo in the LT example.

@@
+In this appendix, we ...
+
 % Please add the following required packages to your document preamble:
@@
 % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{table}[]
+\begin{centered}
 \begin{tabular}{|l|l|l|l|}
@@
 \rowcolor[HTML]{FFFFE6} 
-Surface Syntax          & $i$      & $A_i$             & $A_i'$                       \\ \hline
-\lsti!Family STLC\_bool...! & 0        &                   &                              \\ \cline{2-4} 
+surface-syntax program     & $i$      & $A_i$             & $A_i'$                       \\ \hline
+\lsti!Family STLC_bool...! & 0        &                   &                              \\ \cline{2-4} 
 \lsti!FInductive tm +=!     & 1        & $[]$              & $[]$                         \\ \cline{2-4} 
-\lsti!| .. | tm\_true | ..! & 2       & $["tm" : \cU, "tm_var" : id \to \El{"tm"} .. $                          & $["tm" : \cU,  ..$                         \\
+\lsti!| .. | tm_true | ..! & 2       & $["tm" : \cU, "tm_var" : id \to \El{"tm"} .. $                          & $["tm" : \cU,  ..$                         \\
                       &          & $"tm_unit" : \El{"tm"}]$      & $"tm_unit", "tm_true" : \El{"tm"}, .. ]$ \\ \cline{2-4} 
@@
 \rowcolor[HTML]{E0D7D7} 
-\lsti!Case tm\_true!       & $\beta$ & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          \\ \cline{2-4} 
+\lsti!Case tm_true ...!       & $\beta$ & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          \\ \cline{2-4} 
 \lsti!End subst!            & 3        & $A_2$             & $A_2'$                       \\ \cline{2-4} 
                       & 4        & $\TyTkg{\lsig_3}$ & $\TyTkg{\lsig_3'}$           \\ \cline{2-4} 
-\lsti!(* Inherit t1 *)!    & 5       & $["tm":\cU, .. subst : \El{"tm"} \to id $ & $["tm":\cU, .. subst : \El{"tm"} \to id $ \\
+\lsti!(* Inherit t1 *)!    & 5       & $["tm":\cU, .. "subst" : \El{"tm"} \to id $ & $["tm":\cU, .. "subst" : \El{"tm"} \to id $ \\
                       &         & $ \to \El{"tm"} \to \El{"tm"} ]$          & $ \to \El{"tm"} \to \El{"tm"} ]$          \\ \hline
 \end{tabular}
-\end{table}
+\end{centered}
 
@@
 \rowcolor[HTML]{FFFFE6} 
-Surface Syntax & $i$ & $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{A_i}$ & $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i'}{A_i'}$ \\ \hline
-\lsti!Family STLC\_bool...! & 0        & $\{\}$                                   & $\{\}$                                   \\ \cline{2-4} 
+surface-syntax program & $i$ & $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{T_i}$ & $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i'}{T_i'}$ \\ \hline
+\lsti!Family STLC_bool...! & 0        & $\{\}$                                   & $\{\}$                                   \\ \cline{2-4} 
 \lsti!FInductive tm +=!     & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$       & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$      \\ \cline{2-4} 
-\lsti!| .. | tm\_true | ..!  & 2        &                                          &                                          \\
+\lsti!| .. | tm_true | ..!  & 2        &                                          &                                          \\
                       &          &                                          &                                          \\ \cline{2-4} 
@@
 \rowcolor[HTML]{E0D7D7} 
-\lsti!Case tm\_true!        & $\beta$  & $\sigma_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \cline{2-4} 
-\lsti!End subst!            & 3        & $\{.. subst' : \TyLkg{\lsig_\alpha}\}$   & $\{.. subst' : \TyLkg{\lsig_\beta'}\}$   \\ \cline{2-4} 
-                      & 4        & $\{.. subst : \El{"tm"} \to id \to ..\}$ & $\{.. subst : \El{"tm"} \to id \to ..\}$ \\ \cline{2-4} 
+\lsti!Case tm_true ...!        & $\beta$  & $\sigma_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \cline{2-4} 
+\lsti!End subst!            & 3        & $\{.. "subst'" : \TyLkg{\lsig_\alpha}\}$   & $\{.. "subst"' : \TyLkg{\lsig_\beta'}\}$   \\ \cline{2-4} 
+                      & 4        & $\{.. "subst" : \El{"tm"} \to id \to ..\}$ & $\{.. "subst" : \El{"tm"} \to id \to ..\}$ \\ \cline{2-4} 
 \lsti!(* Inherit t1 *)!     & 5        & $\{.. "t"_1 : \El{"tm"}\}$               & $\{.. "t"_1 : \El{"tm"}\}$               \\
@@
 \end{tabular}
-\end{table}
+\end{centered}
 
@@
 % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{table}[]
+\begin{centered}
 \begin{tabular}{|l|l|l|}
@@
 \rowcolor[HTML]{FFFFE6} 
-Surface Syntax          & $i$      & $\goodInh{..}{i_i}{\sigma_i}{\sigma_i'}$             \\ \hline
-\lsti!Family STLC\_bool...! & 0        & $\InhId$                                             \\ \cline{2-3} 
-\lsti!FInductive tm +=!     & 1        & $\InhOv{i_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
-\lsti!| .. | tm\_true | ..!  & 2        &                                                      \\
+surface-syntax program     & $i$      & $\goodInh{..}{h_i}{\sigma_i}{\sigma_i'}$             \\ \hline
+\lsti!Family STLC_bool...! & 0        & $\InhId$                                             \\ \cline{2-3} 
+\lsti!FInductive tm +=!     & 1        & $\InhOv{h_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
+\lsti!| .. | tm_true | ..!  & 2        &                                                      \\
                         &          &                                                      \\ \cline{2-3} 
@@
 \rowcolor[HTML]{E0D7D7} 
-\lsti!Case tm\_true!        & $\beta$  & $\InhExt{i_\alpha}{..}$                              \\ \cline{2-3} 
-\lsti!End subst!            & 3        & $\InhNest{i_1}{i_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
-                        & 4        & $\InhOv{i_3}{\lambda t. \Wrec{\wsig_1'}{t}{subst'}}$ \\ \cline{2-3} 
-\lsti!(* Inherit t1 *)!     & 5        & $\InhInh{i_4}$ with $\uparrow s_4$                   \\
+\lsti!Case tm_true ...!        & $\beta$  & $\InhExt{h_\alpha}{..}$                              \\ \cline{2-3} 
+\lsti!End subst!            & 3        & $\InhNest{h_1}{h_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
+                        & 4        & $\InhOv{h_3}{\lambda t. \Wrec{\wsig_1'}{t}{"subst'"}}$ \\ \cline{2-3} 
+\lsti!(* Inherit t1 *)!     & 5        & $\InhInh{h_4}$ with $\uparrow s_4$                   \\
                         &          &                                                      \\ \hline
 \end{tabular}
-\end{table}
+\end{centered}
 
-We take "Family STLC_bool" extending "STLC" as an example.
-We list out the surface syntax and their translation in out \TT.
+We take \lsti!Family STLC_bool!
+extending \lsti!STLC!
+as an example.
+We list out the surface syntax and their translation into \TT.
 In our figure, $\lsig_i$ stands for the parent signature (the signature before the transformation); and $\lsig_i'$ stands for the signature after the transformation.

### 2022-11-09 stlc-linkTran-typing.tex
Fix typo in LT example.

@@
 \rowcolor[HTML]{FFFFE6} 
-surface-syntax program & $i$ & $\sigma_i = \LSigAdd{\sigma_{i-1}}{s_i}{T_i}$ & $\sigma_i' = \LSigAdd{\sigma_{i-1}'}{s_i'}{T_i'}$ \\ \hline
+surface-syntax program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i'}{T_i'}$ \\ \hline
 \lsti!Family STLC_bool...! & 0        & $\{\}$                                   & $\{\}$                                   \\ \cline{2-4} 
 \lsti!FInductive tm +=!     & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$       & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$      \\ \cline{2-4} 
-\lsti!| .. | tm_true | ..!  & 2        &                                          &                                          \\
+\lsti!| .. | tm_true | ..!  & 2        & N/A                                         & N/A                                         \\
                       &          &                                          &                                          \\ \cline{2-4} 
@@
 \rowcolor[HTML]{E0D7D7} 
-\lsti!Case tm_true ...!        & $\beta$  & $\sigma_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \cline{2-4} 
+\lsti!Case tm_true ...!        & $\beta$  & $\lsig_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \cline{2-4} 
 \lsti!End subst!            & 3        & $\{.. "subst'" : \TyLkg{\lsig_\alpha}\}$   & $\{.. "subst"' : \TyLkg{\lsig_\beta'}\}$   \\ \cline{2-4} 
@@
 \rowcolor[HTML]{FFFFE6} 
-surface-syntax program     & $i$      & $\goodInh{..}{h_i}{\sigma_i}{\sigma_i'}$             \\ \hline
+surface-syntax program     & $i$      & $\goodInh{..}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
 \lsti!Family STLC_bool...! & 0        & $\InhId$                                             \\ \cline{2-3} 
 \lsti!FInductive tm +=!     & 1        & $\InhOv{h_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
-\lsti!| .. | tm_true | ..!  & 2        &                                                      \\
+\lsti!| .. | tm_true | ..!  & 2        & N/A                                                     \\
                         &          &                                                      \\ \cline{2-3} 
@@
 Because every $\lsig_i$ is always constructed from $\lsig_{i-1}$ so in the figure we omit the previous fields in each signature.
-The only exceptions are the rows indexed by $\alpha$ and $\beta$.
-These two rows are describing a nested family for case handlers, and $\lsig_\beta$ is constructed by $\lsig_\alpha$.
+One exception are the rows indexed by $\alpha$ and $\beta$.
+These two rows are describing a nested family for case handlers, and $\lsig_\beta$ is based on $\lsig_\alpha$.
+The other exception is $\lsig_3$.
+$\lsig_3$ is based on $\lsig_1$.
 
@@
 Note that in our table, the Row 2 only has information of $A_2$ and $A_2'$ (i.e.
-there is no $\lsig_2$ and $\lsig_2'$).
-We especially use one row to indicate that $s_2$ and $s_2'$ is ``sealing/abstracting'' $\lsig_1' = \{"tm" : \TyS{\wcode{\wsig_1'}}\}$, into a nested dependent pair of "tm" type and its constructors $A_2' = ["tm" : \cU, "tm_var" : id \to \El{"tm"}, "tm_lam" : id \to \El{"tm"} \to \El{"tm"}, "tm_app" : \El{"tm"} \to \El{"tm"} \to \El{"tm"}]$.
+there is no $\lsig_2$ or $\lsig_2'$).
+We especially use one row to indicate that $s_2$ and $s_2'$ is ``sealing/abstracting'' $\lsig_1' = \{"tm" : \TyS{\wcode{\wsig_1'}}\}$, into a nested dependent pair of "tm" type and its constructors $A_2' = ["tm" : \cU, "tm_var" : id \to \El{"tm"}, "tm_lam" : id \to \El{"tm"} \to \El{"tm"}, "tm_app" : \El{"tm"} \to \El{"tm"} \to \El{"tm"}, "tm_true" : \El{"tm"}, "tm_false" : \El{"tm"}]$.

### 2022-11-09 stlc-linkTran-typing.tex
Reformatting LT example.

@@
   \hline
-\rowcolor[HTML]{FFFFE6} 
-surface-syntax program     & $i$      & $A_i$             & $A_i'$                       \\ \hline
-\lsti!Family STLC_bool...! & 0        &                   &                              \\ \cline{2-4} 
-\lsti!FInductive tm +=!     & 1        & $[]$              & $[]$                         \\ \cline{2-4} 
-\lsti!| .. | tm_true | ..! & 2       & $["tm" : \cU, "tm_var" : id \to \El{"tm"} .. $                          & $["tm" : \cU,  ..$                         \\
-                      &          & $"tm_unit" : \El{"tm"}]$      & $"tm_unit", "tm_true" : \El{"tm"}, .. ]$ \\ \cline{2-4} 
-\lsti!FRecursion subst!     &          &                   &                              \\ \cline{2-4} 
-\rowcolor[HTML]{E0D7D7} 
-\lsti!(* Inherit ... *)!    & $\alpha$ &                   &                              \\ \cline{2-4} 
-\rowcolor[HTML]{E0D7D7} 
-\lsti!Case tm_true ...!       & $\beta$ & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          & $\TyTkg{\RecSig{\wsig_1}{"tm"}}$          \\ \cline{2-4} 
-\lsti!End subst!            & 3        & $A_2$             & $A_2'$                       \\ \cline{2-4} 
-                      & 4        & $\TyTkg{\lsig_3}$ & $\TyTkg{\lsig_3'}$           \\ \cline{2-4} 
-\lsti!(* Inherit t1 *)!    & 5       & $["tm":\cU, .. "subst" : \El{"tm"} \to id $ & $["tm":\cU, .. "subst" : \El{"tm"} \to id $ \\
-                      &         & $ \to \El{"tm"} \to \El{"tm"} ]$          & $ \to \El{"tm"} \to \El{"tm"} ]$          \\ \hline
+  \rowcolor[HTML]{FFFC9E} 
+  surface syntax of program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i}{T_i}$ \\ \hline
+  \lsti!Family STLC\_bool...!   & 0        & $\{\}$                                     & $\{\}$                                     \\ \cline{2-4} 
+  \lsti!FInductive tm += !      & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$         & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$        \\ \cline{2-4} 
+  \lsti! (*earlier constr*)!    & 2        & $\{..; "tm_var": \El{"tm"}, .. \}$         & $\{..;"tm_var": \El{"tm"},.. \}$           \\ \cline{2-4} 
+  \lsti! .. | tm\_true | .. !   & 3        & $\lsig_2$                                  & $\{..,"tm_true":\El{"tm"} \}$              \\ \cline{2-4} 
+  \rowcolor[HTML]{CDCDCD} 
+  \lsti!  FRecursion subst!       &          &                                            &                                            \\ \cline{2-4} 
+  \rowcolor[HTML]{CDCDCD} 
+  \lsti!  (* Inherit ... *)!      & $\alpha$ & $\RecSig{\wsig_1}{"tm"}$                   & $\RecSig{\wsig_1}{"tm"}$                   \\ \cline{2-4} 
+  \rowcolor[HTML]{CDCDCD} 
+  \lsti!  Case tm\_true!          & $\beta$  & $\lsig_{\alpha}$                          & $\{.. "tm:true": \}$                       \\ \cline{2-4} 
+  \rowcolor[HTML]{FFFFFF} 
+  \lsti!  End subst!              & 4        & $\{.. "subst'" : \TyLkg{\lsig_\alpha}\}$   & $\{.. "subst'" : \TyLkg{\lsig_\beta'}\}$   \\ \cline{2-4} 
+                            & 5        & $\{.. "subst" : \El{"tm"} \to id \to ..\}$ & $\{.. "subst" : \El{"tm"} \to id \to ..\}$ \\ \cline{2-4} 
+  \lsti!(* Do the same on ty*)! & 6        & ...                                        & ...                                        \\ \cline{2-4} 
+  \lsti!(* Inherit env *)!      & 7        & $\{.. "env" : id \to "option"~"ty"\}$      & $\{.. "env" : id \to "option"~"ty"\}$      \\ \hline
   \end{tabular}
@@
 \begin{centered}
-\begin{tabular}{|l|l|l|l|}
-\hline
-\rowcolor[HTML]{FFFFE6} 
-surface-syntax program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i'}{T_i'}$ \\ \hline
-\lsti!Family STLC_bool...! & 0        & $\{\}$                                   & $\{\}$                                   \\ \cline{2-4} 
-\lsti!FInductive tm +=!     & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$       & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$      \\ \cline{2-4} 
-\lsti!| .. | tm_true | ..!  & 2        & N/A                                         & N/A                                         \\
-                      &          &                                          &                                          \\ \cline{2-4} 
-\lsti!FRecursion subst!     &          &                                          &                                          \\ \cline{2-4} 
-\rowcolor[HTML]{E0D7D7} 
-\lsti!(* Inherit ... *)!    & $\alpha$ & $\RecSig{\wsig_1}{"tm"}$                 & $\RecSig{\wsig_1}{"tm"}$                 \\ \cline{2-4} 
-\rowcolor[HTML]{E0D7D7} 
-\lsti!Case tm_true ...!        & $\beta$  & $\lsig_{\alpha}$                        & $\{.. "tm:true": \}$                     \\ \cline{2-4} 
-\lsti!End subst!            & 3        & $\{.. "subst'" : \TyLkg{\lsig_\alpha}\}$   & $\{.. "subst"' : \TyLkg{\lsig_\beta'}\}$   \\ \cline{2-4} 
-                      & 4        & $\{.. "subst" : \El{"tm"} \to id \to ..\}$ & $\{.. "subst" : \El{"tm"} \to id \to ..\}$ \\ \cline{2-4} 
-\lsti!(* Inherit t1 *)!     & 5        & $\{.. "t"_1 : \El{"tm"}\}$               & $\{.. "t"_1 : \El{"tm"}\}$               \\
-                      &          &                                          &                                          \\ \hline
-\end{tabular}
-\end{centered}
 
-% Please add the following required packages to your document preamble:
-% \usepackage[table,xcdraw]{xcolor}
-% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{centered}
 \begin{tabular}{|l|l|l|}
 \hline
-\rowcolor[HTML]{FFFFE6} 
-surface-syntax program     & $i$      & $\goodInh{..}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
-\lsti!Family STLC_bool...! & 0        & $\InhId$                                             \\ \cline{2-3} 
-\lsti!FInductive tm +=!     & 1        & $\InhOv{h_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
-\lsti!| .. | tm_true | ..!  & 2        & N/A                                                     \\
-                        &          &                                                      \\ \cline{2-3} 
-\lsti!FRecursion subst!     &          &                                                      \\ \cline{2-3} 
-\rowcolor[HTML]{E0D7D7} 
-\lsti!(* Inherit ... *)!    & $\alpha$ & $\InhId$                                             \\ \cline{2-3} 
-\rowcolor[HTML]{E0D7D7} 
-\lsti!Case tm_true ...!        & $\beta$  & $\InhExt{h_\alpha}{..}$                              \\ \cline{2-3} 
-\lsti!End subst!            & 3        & $\InhNest{h_1}{h_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
-                        & 4        & $\InhOv{h_3}{\lambda t. \Wrec{\wsig_1'}{t}{"subst'"}}$ \\ \cline{2-3} 
-\lsti!(* Inherit t1 *)!     & 5        & $\InhInh{h_4}$ with $\uparrow s_4$                   \\
-                        &          &                                                      \\ \hline
+\rowcolor[HTML]{FFFC9E} 
+surface syntax of program & $i$      & $\goodInh{..}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
+\lsti!Family STLC\_bool!   & 0        & $\InhId$                                             \\ \cline{2-3} 
+\lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
+\lsti! (*earlier constr*) !   & 2        & $\InhOv{h_1}{\Wsup{\wsig_1'}{\top}{\bot}}$           \\ \cline{2-3} 
+\lsti! .. | tm\_true | ..!    & 3        & $\InhExt{h_2}{\Wsup{\wsig_1'}{\top}{\bot}}$          \\ \cline{2-3} 
+\rowcolor[HTML]{CDCDCD} 
+\lsti!  FRecursion subst!       &          &                                                      \\ \cline{2-3} 
+\rowcolor[HTML]{CDCDCD} 
+\lsti!  (* Inherit ... *)!      & $\alpha$ & $\InhId$                                             \\ \cline{2-3} 
+\rowcolor[HTML]{CDCDCD} 
+\lsti!  Case tm\_true!          & $\beta$  & $\InhExt{h_\alpha}{..}$                              \\ \cline{2-3} 
+\rowcolor[HTML]{FFFFFF} 
+\lsti!  End subst!              & 4        & $\InhNest{h_3}{h_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
+                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_1'}{t}{subst'}}$ \\ \cline{2-3} 
+\lsti!(* Do the same on ty*)! & 6        & ...                                                  \\ \cline{2-3} 
+\lsti!(* Inherit env *)!      & 7        & $\InhInh{h_6}$ with $\uparrow s_4$                   \\ \hline
 \end{tabular}
@@
 In our figure, $\lsig_i$ stands for the parent signature (the signature before the transformation); and $\lsig_i'$ stands for the signature after the transformation.
+We omit the information for $s_i$ and $A_i$
 
-Before we sketch out the signature, we need to specify \textit{the sealing premise} $s$ in \ruleref{lsig/add}. However, we omit $s$ as the information of $A$ is enough to specify $s$.
-
-When sketching signature $\lsig$, we use curly bracket as a shorthand for signature and square bracket as a shorthand for (deeply) nested sigma type (usually used in the case of $\TyTkg{\cdot}$).
+When sketching signature $\lsig$, we use curly bracket as a shorthand for signature.
 Because every $\lsig_i$ is always constructed from $\lsig_{i-1}$ so in the figure we omit the previous fields in each signature.
@@
 These two rows are describing a nested family for case handlers, and $\lsig_\beta$ is based on $\lsig_\alpha$.
-The other exception is $\lsig_3$.
-$\lsig_3$ is based on $\lsig_1$.
+The other exception is $\lsig_3$
 
-Our final target is to construct a term $\goodInh{\cdot}{?}{\lsig_5}{\lsig_5'}$ using the four linkage transformers' rules.
+Our final target is to construct a term $\goodInh{\cdot}{?}{\lsig_7}{\lsig_7'}$ using the four linkage transformers' rules.
 There are two upcasting operations $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_4'}{\uparrow s_4}{A_4}$ (omitting debruijn indices here) for checking the compatibility between $A_i$ and $A_i'$.

### 2022-11-09 stlc-linkTran-typing.tex
Fix typo in LT example

@@
   surface syntax of program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i}{T_i}$ \\ \hline
-  \lsti!Family STLC\_bool...!   & 0        & $\{\}$                                     & $\{\}$                                     \\ \cline{2-4} 
+  \lsti!Family STLC_bool...!   & 0        & $\{\}$                                     & $\{\}$                                     \\ \cline{2-4} 
   \lsti!FInductive tm += !      & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$         & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$        \\ \cline{2-4} 
-  \lsti! (*earlier constr*)!    & 2        & $\{..; "tm_var": \El{"tm"}, .. \}$         & $\{..;"tm_var": \El{"tm"},.. \}$           \\ \cline{2-4} 
-  \lsti! .. | tm\_true | .. !   & 3        & $\lsig_2$                                  & $\{..,"tm_true":\El{"tm"} \}$              \\ \cline{2-4} 
+  \lsti! (*existent constr*)!    & 2        & $\{..; "tm_var": \El{"tm"}, .. \}$         & $\{..;"tm_var": \El{"tm"},.. \}$           \\ \cline{2-4} 
+  \lsti! .. | tm_true | .. !   & 3        & $\lsig_2$                                  & $\{..,"tm_true":\El{"tm"} \}$              \\ \cline{2-4} 
   \rowcolor[HTML]{CDCDCD} 
@@
   \rowcolor[HTML]{CDCDCD} 
-  \lsti!  Case tm\_true!          & $\beta$  & $\lsig_{\alpha}$                          & $\{.. "tm:true": \}$                       \\ \cline{2-4} 
+  \lsti!  Case tm_true!          & $\beta$  & $\lsig_{\alpha}$                          & $\{.. "tm:true": \}$                       \\ \cline{2-4} 
   \rowcolor[HTML]{FFFFFF} 
@@
 surface syntax of program & $i$      & $\goodInh{..}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
-\lsti!Family STLC\_bool!   & 0        & $\InhId$                                             \\ \cline{2-3} 
+\lsti!Family STLC_bool!   & 0        & $\InhId$                                             \\ \cline{2-3} 
 \lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
-\lsti! (*earlier constr*) !   & 2        & $\InhOv{h_1}{\Wsup{\wsig_1'}{\top}{\bot}}$           \\ \cline{2-3} 
-\lsti! .. | tm\_true | ..!    & 3        & $\InhExt{h_2}{\Wsup{\wsig_1'}{\top}{\bot}}$          \\ \cline{2-3} 
+\lsti! (*existent constr*) !   & 2        & $\InhOv{h_1}{\Wsup{\wsig_1'}{\top}{\bot}}$           \\ \cline{2-3} 
+\lsti! .. | tm_true | ..!    & 3        & $\InhExt{h_2}{\Wsup{\wsig_1'}{\top}{\bot}}$          \\ \cline{2-3} 
 \rowcolor[HTML]{CDCDCD} 
@@
 \rowcolor[HTML]{CDCDCD} 
-\lsti!  Case tm\_true!          & $\beta$  & $\InhExt{h_\alpha}{..}$                              \\ \cline{2-3} 
+\lsti!  Case tm_true!          & $\beta$  & $\InhExt{h_\alpha}{..}$                              \\ \cline{2-3} 
 \rowcolor[HTML]{FFFFFF} 
 \lsti!  End subst!              & 4        & $\InhNest{h_3}{h_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
-                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_1'}{t}{subst'}}$ \\ \cline{2-3} 
+                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_1'}{t}{"subst'"}}$ \\ \cline{2-3} 
 \lsti!(* Do the same on ty*)! & 6        & ...                                                  \\ \cline{2-3} 
-\lsti!(* Inherit env *)!      & 7        & $\InhInh{h_6}$ with $\uparrow s_4$                   \\ \hline
+\lsti!(* Inherit env *)!      & 7        & $\InhInh{h_6}$ with $\uparrow s_6$                   \\ \hline
 \end{tabular}
@@
 These two rows are describing a nested family for case handlers, and $\lsig_\beta$ is based on $\lsig_\alpha$.
-The other exception is $\lsig_3$
+The other exception is $\lsig_3$, which is directly $\lsig_2$ since the parent doesn't have those new constructor.
+Note for $\lsig_2$ and $\lsig_2'$, there are actually multiple constructors inserted into the signature (e.g.
+"tm_lam" and "tm_app") which we omitted here.
 
 Our final target is to construct a term $\goodInh{\cdot}{?}{\lsig_7}{\lsig_7'}$ using the four linkage transformers' rules.
-There are two upcasting operations $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_4'}{\uparrow s_4}{A_4}$ (omitting debruijn indices here) for checking the compatibility between $A_i$ and $A_i'$.
-
-Note that in our table, the Row 2 only has information of $A_2$ and $A_2'$ (i.e.
-there is no $\lsig_2$ or $\lsig_2'$).
-We especially use one row to indicate that $s_2$ and $s_2'$ is ``sealing/abstracting'' $\lsig_1' = \{"tm" : \TyS{\wcode{\wsig_1'}}\}$, into a nested dependent pair of "tm" type and its constructors $A_2' = ["tm" : \cU, "tm_var" : id \to \El{"tm"}, "tm_lam" : id \to \El{"tm"} \to \El{"tm"}, "tm_app" : \El{"tm"} \to \El{"tm"} \to \El{"tm"}, "tm_true" : \El{"tm"}, "tm_false" : \El{"tm"}]$.
+There are two upcasting operations $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_6'}{\uparrow s_6}{A_6}$ (omitting debruijn indices here) for checking the compatibility between the omitted $A_i$ and $A_i'$.

### 2022-11-10 lang-design.tex
edits in §3.1

@@
 proof assistant,
-though we believe it could be adapted to other proof assistants such as Lean.
+though we believe the design could be adapted to other proof assistants such as Lean.
 We call our design and implementation \Lang (\underline{f}amily
@@
 In family "STLCFix", inductive type "tm" \emph{further binds} the "tm" type
-in family "STLC". It is generated by five constructors: four inherited from
+in family "STLC". It has five constructors: four inherited from
 "STLC" and a fifth called "tm_fix".
 
-Crucially, the types of the constructors are polymorphic to their enclosing family:
-the types mention "tm", and the meaning of "tm" is \emph{late bound}.
-For example, although "tm_app" is defined in the base family,
-it can be used to construct applications of fixpoints in the derived family:
-%such as \lsti[mathescape]{tm_app (tm_fix "f" (tm_abs "x" $\dadada$)) tm_unit},
-its type \lsti{tm -> tm -> tm} says "tm_app" can be applied to any
-Gallina terms of type "tm",
+Crucially, the meanings of \lsti{tm} and its constructors are
+\emph{late bound}, depending on the family in which they are referenced.
+Consider \lsti{tm_app}.
+It is defined in family \lsti{STLC} and is thus unaware of \lsti{tm_fix}.
+Yet in family \lsti{STLCFix}, we can use \lsti{tm_app} to construct
+applications of fixpoints, as in
+\lsti[mathescape]{tm_app (tm_fix "f" $t_1$) $t_2$}.
+This use is justified by \lsti{tm_app}'s type,
+\lsti[mathescape]{tm$\,$->$\,$tm$\,$->$\,$tm}.
+It allows \lsti{tm_app} to be applied to anything of type \lsti{tm},
 which in family "STLCFix" include those constructed from "tm_fix".
@@
 further bind "subst" and define this fifth case.
-The type system does this check by examining if the inductive type "tm",
+The type system does this check in family \lsti{STLCFix}
+by examining if the inductive type "tm",
 over which "subst" is recursively defined, is further bound in the same family.
@@
 
+\begin{centered}
+\begin{minipage}{.63\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
@@
 \end{lstlisting}
+\end{minipage}
+\end{centered}
 
@@
 Because the lemma is by induction on "hasty" and because "STLCFix" adds
-a constructor to "hasty", the programmer is required to handle this extra case
-in the proof of "substlem" in family "STLCFix".
+a constructor to "hasty", the programmer is required in family "STLCFix" to
+extend the proof of "substlem" to handle this extra case.
 

### 2022-11-10 lang-design.tex
took a pass on §3

@@
 It is proven by induction on the typing relation "hasty".
-Its \lsti{motive} clause suggests that the lemma reads as
+Its \lsti{motive} clause shows that the lemma reads as
+\lstinline[basicstyle=\fontsize{8}{9}\ttfamily]!\forall G t T, hasty G t T -> \forall G', includedin G G' -> hasty G' t T!.
+
+%\begin{centered}
+%\begin{minipage}{.63\textwidth}
+%\begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
+%\forall G t T, hasty G t T -> \forall G', includedin G G' -> hasty G' t T.
+%\end{lstlisting}
+%\end{minipage}
+%\end{centered}
+
+\noindent
+Upon entering case "ht_abs", for instance, Coq enters proof mode with the goal
 
 \begin{centered}
-\begin{minipage}{.63\textwidth}
+\begin{minipage}{.635\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
-\forall G t T, hasty G t T -> \forall G', includedin G G' -> hasty G' t T.
+\forall G', includedin G G' -> hasty G' (tm_abs x t) (ty_arrow T1 T2)
 \end{lstlisting}
@@
 \noindent
-Upon entering case "ht_abs", for instance, Coq enters proof mode with the goal
-
-\begin{lstlisting}[basicstyle=\fontsize{8}{9}\ttfamily]
-  \forall ...
-\end{lstlisting}
-
-\noindent
+and with the induction hypothesis
+\lstinline[basicstyle=\fontsize{8}{9}\ttfamily]!\forall G', includedin (extend G x T1) G' -> hasty G' t T2!.
+%
 The programmer can use tactic programming to discharge the goal.
 %The idiomatic programming experience with Coq is largely preserved.
-Because the lemma is by induction on "hasty" and because "STLCFix" adds
-a constructor to "hasty", the programmer is required in family "STLCFix" to
+Because the lemma is by \lsti!FInduction! on "hasty" and because "STLCFix" adds
+a constructor "ht_fix" to "hasty", the programmer is required in family "STLCFix" to
 extend the proof of "substlem" to handle this extra case.
@@
 Thus, late binding of "subst" allows the derived family to reuse "st_beta"
-to $\beta$-reduce abstractions whose body is a fixpoint.
+as the $\beta$-reduction rule for applications possibly constructed from "tm_fix".
 
-Importantly, a name is late bound only within a family that defines or extends it.
+Importantly, a name is late bound only within a family that defines or further binds it.
 Outside such families, the name can be accessed only by explicitly
@@
 For example, the last line of \cref{fig:stlc-mechanized} accesses "typesafe"
-with "STLCFix.typesafe".
-Executing the command prints
+with a qualifier "STLCFix".
+The command prints
 
+\begin{centered}
+\begin{minipage}{.880\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
@@
 
+\begin{centered}
+\begin{minipage}{\textwidth}
 \begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
@@
 
-However, with \lsti{subst} being late bound, \lsti{subst} can no longer
+But with \lsti{subst} being late bound, \lsti{subst} cannot and should not
 be unfolded~(\ref{chg:latebind-v-defeq}):
@@
 After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
-from a base family to a derived family.\EDJ{I think the reviewer might ask : what if the subst late-bound tm_unit case/why the subst cases cannot be late-bound?
-it not necessary return tm_unit in those case.
-\\ What the plugin generates is actually subst tm_unit x t = subst.tm_unit.
+from a base family to a derived family.
+%\EDJ{I think the reviewer might ask : what if the subst late-bound tm_unit case/why the subst cases cannot be late-bound? it not necessary return tm_unit in those case. 
+%\\ What the plugin generates is actually subst tm_unit x t = subst.tm_unit.
 In other words, it points back to the handler instead of the concrete constructor.
-That is one of the advantage of looking FRecursion as a family.}
+That is one of the advantage of looking FRecursion as a family.}\YZreply{Currently the paper is taking the position that case handlers cannot be further bound, to make the story simpler.}
+What varies is \lsti{subst} itself as a recursive function combining the case handlers.
 %
 Based on this insight, \Lang automatically generates
-propositional equalities for each recursive function defined with \lsti{FRecursion},
-making them available as axioms for use by the rest of the current family.
+a propositional equality for each case handler defined within \lsti{FRecursion},
+making the equalities and the recursive function available as axioms for use by the rest of the current family.
 \Lang also provides a tactic \lsti{fsimpl} that enables, for instance,
 simplifying \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
-Under the hood, \lsti{fsimpl} works by rewriting using the postulated
-propositional equalities.
+Under the hood, \lsti{fsimpl} works by rewriting applications of the
+axiomatized recursive function using the axiomatized propositional equalities.
 
 Note that the definitional equality on \lsti{subst} is available
-outside those families that contain "subst".
-Within those families, the meaning of \lsti{subst} is late bound, so only
+\emph{outside} those families that contain "subst".
+Within those families, the meaning of \lsti{subst} is late bound
+(i.e., polymorphic to the enclosing family), so only
 propositional equality is available.
 In contrast, outside those families, \lsti{subst} is always referenced
-by specifying a family that contains it, \EDJ{I think the reviewer without background of Family Polymorphism will be confused at this paragraph.
+by specifying a family that contains it, %\EDJ{I think the reviewer without background of Family Polymorphism will be confused at this paragraph.
 i.e.
@@
 }
-such as \lsti{STLC.subst} or \lsti{STLCFix.subst}.
+as in \lsti{STLC.subst} and \lsti{STLCFix.subst}.
 As far as the type checker is concerned, \lsti{tm_unit} and
@@
 %
-In the event that it cannot be inferred where to place of a newly introduced field,
-programmer annotation is required.
+In the event that \Lang cannot infer where the programmer intends to place a new field,
+annotation is required.
 
@@
 
-In \cref{fig:stlc-isorec-prod}, family "STLCFixIsorec" is declared as
+In \cref{fig:stlc-isorec-prod}, "STLCFixIsorec" is declared as
 an "STLC" extension that mixes in "STLCFix" and "STLCIsorec".
@@
 but there is power in restraint.
-Partial recursors offer a principled, uniform way to derive injectivity and disjointness of
-constructors, while allowing for extensibility:
+Partial recursors serve as extensional characterizations of inductive types allowing future extension.
+In particular,
+they offer a principled, uniform way to derive injectivity and disjointness of
+constructors:
 they enable injective mappings from a late bound inductive
@@
 For example, in a proof state with the assumption \lsti{H : tm_var "x" = tm_abs "y" t},
-running \lsti{fdiscriminate H} first applies the injective mapping to
+running \lsti{fdiscriminate H} first applies an injective mapping to
 both sides of~\lsti{H}, obtaining

### 2022-11-10 stlc-linkTran-typing.tex
Fix typo in LT example.

@@
   \rowcolor[HTML]{FFFC9E} 
-  surface syntax of program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i}{T_i}$ \\ \hline
+  surface syntax of program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i'}{T_i'}$ \\ \hline
   \lsti!Family STLC_bool...!   & 0        & $\{\}$                                     & $\{\}$                                     \\ \cline{2-4} 
-  \lsti!FInductive tm += !      & 1        & $\{"tm" : \TyS{\wcode{\wsig_1}}\}$         & $\{"tm" : \TyS{\wcode{\wsig_1'}}\}$        \\ \cline{2-4} 
+  \lsti!FInductive tm += !      & 1        & $\{"tm" : \TyS{\wcode{\wsig_{\texttt{tm}}}}\}$         & $\{"tm" : \TyS{\wcode{\wsig_{\texttt{tm}}'}}\}$        \\ \cline{2-4} 
   \lsti! (*existent constr*)!    & 2        & $\{..; "tm_var": \El{"tm"}, .. \}$         & $\{..;"tm_var": \El{"tm"},.. \}$           \\ \cline{2-4} 
@@
   \rowcolor[HTML]{CDCDCD} 
-  \lsti!  (* Inherit ... *)!      & $\alpha$ & $\RecSig{\wsig_1}{"tm"}$                   & $\RecSig{\wsig_1}{"tm"}$                   \\ \cline{2-4} 
+  \lsti!  (* Inherit ... *)!      & $\alpha$ & $\RecSig{\wsig_{\texttt{tm}}}{"tm"}$                   & $\RecSig{\wsig_{\texttt{tm}}}{"tm"}$                   \\ \cline{2-4} 
   \rowcolor[HTML]{CDCDCD} 
-  \lsti!  Case tm_true!          & $\beta$  & $\lsig_{\alpha}$                          & $\{.. "tm:true": \}$                       \\ \cline{2-4} 
+  \lsti!  Case tm_true!          & $\beta$  & $\lsig_{\alpha}$                          & $\{.. "tm_true": .. \}$                       \\ \cline{2-4} 
   \rowcolor[HTML]{FFFFFF} 
@@
 \lsti!Family STLC_bool!   & 0        & $\InhId$                                             \\ \cline{2-3} 
-\lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_1'}}$                      \\ \cline{2-3} 
-\lsti! (*existent constr*) !   & 2        & $\InhOv{h_1}{\Wsup{\wsig_1'}{\top}{\bot}}$           \\ \cline{2-3} 
-\lsti! .. | tm_true | ..!    & 3        & $\InhExt{h_2}{\Wsup{\wsig_1'}{\top}{\bot}}$          \\ \cline{2-3} 
+\lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_{\texttt{tm}}'}}$                      \\ \cline{2-3} 
+\lsti! (*existent constr*) !   & 2        & $\InhOv{h_1}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$           \\ \cline{2-3} 
+\lsti! .. | tm_true | ..!    & 3        & $\InhExt{h_2}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$          \\ \cline{2-3} 
 \rowcolor[HTML]{CDCDCD} 
@@
 \lsti!  End subst!              & 4        & $\InhNest{h_3}{h_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
-                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_1'}{t}{"subst'"}}$ \\ \cline{2-3} 
+                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_{\texttt{tm}}'}{t}{"subst'"}}$ \\ \cline{2-3} 
 \lsti!(* Do the same on ty*)! & 6        & ...                                                  \\ \cline{2-3} 
@@
 One exception are the rows indexed by $\alpha$ and $\beta$.
-These two rows are describing a nested family for case handlers, and $\lsig_\beta$ is based on $\lsig_\alpha$.
+These two rows are describing a nested family for case handlers, and $\lsig_\beta'$ is based on $\lsig_\alpha'$ and $\lsig_\beta = \lsig_\alpha$.
 The other exception is $\lsig_3$, which is directly $\lsig_2$ since the parent doesn't have those new constructor.

### 2022-11-10 lang-design.tex
spacing

@@
 \newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
-\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
 tm_rect : \forall (P : tm -> Type), [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] -> [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] ->
@@
 \begin{centered}
-\begin{minipage}{.97\textwidth}
+\begin{minipage}{\textwidth}
 \newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
-\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
 tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
   [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
-                [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
-                \forall t, option (P t)
+  [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
 \end{lstlisting}
@@
 \end{centered}
+
+%\begin{centered}
+%\begin{minipage}{.97\textwidth}
+%\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
+%\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
+%tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
+%                [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
+%                [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
+%                \forall t, option (P t)
+%\end{lstlisting}
+%\end{minipage}
+%\end{centered}
 

### 2022-11-10 lang-design.tex
undo 'extensional characterization' remark

@@
 
+%\begin{centered}
+%\begin{minipage}{\textwidth}
+%\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
+%\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
+%tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
+%  [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
+%  [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
+%\end{lstlisting}
+%\end{minipage}
+%\end{centered}
+
 \begin{centered}
-\begin{minipage}{\textwidth}
+\begin{minipage}{.97\textwidth}
 \newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
@@
                 [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
-  [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
+                [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
+                \forall t, option (P t)
 \end{lstlisting}
@@
 \end{centered}
-
-%\begin{centered}
-%\begin{minipage}{.97\textwidth}
-%\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
-%\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
-%tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
-%                [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
-%                [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
-%                \forall t, option (P t)
-%\end{lstlisting}
-%\end{minipage}
-%\end{centered}
 
@@
 but there is power in restraint.
-Partial recursors serve as extensional characterizations of inductive types allowing future extension.
+%Partial recursors serve as extensional characterizations of inductive types allowing future extension.
 In particular,
 they offer a principled, uniform way to derive injectivity and disjointness of
-constructors:
+constructors, while supporting future extension:
 they enable injective mappings from a late bound inductive

### 2022-11-10 stlc-linkTran-typing.tex
Fix typo in LT

@@
 \rowcolor[HTML]{FFFFFF} 
-\lsti!  End subst!              & 4        & $\InhNest{h_3}{h_\beta}$ with $\uparrow s_3$         \\ \cline{2-3} 
+\lsti!  End subst!              & 4        & $\InhNest{h_3}{h_\beta}$ with $\uparrow^s_3$         \\ \cline{2-3} 
                           & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_{\texttt{tm}}'}{t}{"subst'"}}$ \\ \cline{2-3} 
 \lsti!(* Do the same on ty*)! & 6        & ...                                                  \\ \cline{2-3} 
-\lsti!(* Inherit env *)!      & 7        & $\InhInh{h_6}$ with $\uparrow s_6$                   \\ \hline
+\lsti!(* Inherit env *)!      & 7        & $\InhInh{h_6}$ with $\uparrow^s_6$                   \\ \hline
 \end{tabular}
@@
 Our final target is to construct a term $\goodInh{\cdot}{?}{\lsig_7}{\lsig_7'}$ using the four linkage transformers' rules.
-There are two upcasting operations $\goodTerm{A_3'}{\uparrow s_3}{A_3}$ and $\goodTerm{A_6'}{\uparrow s_6}{A_6}$ (omitting debruijn indices here) for checking the compatibility between the omitted $A_i$ and $A_i'$.
+There are two upcasting operations $\goodTerm{A_3'}{\uparrow^s_3}{A_3}$ and $\goodTerm{A_6'}{\uparrow^s_6}{A_6}$ (omitting debruijn indices here) for checking the compatibility between the omitted $A_i$ and $A_i'$.

### 2022-11-10 metatheory-condensed.tex
add one sentence about parser case study
edits in FMLTT sec

@@
 typing is structural rather than nominal, and
-partial recursors and propositional equalities are not automated but have to
-be explicitly defined.
-\TT provides MLTT-style constructs that can be used to \emph{encode}
-family polymorphism rather than building it in directly.
+fields automatically axiomatized by the plugin require explicit definitions in \TT.
 %
-Most notably, \TT extends MLTT with what we call \emph{linkages}.\footnote{
-Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
-family polymorphism in an OO setting, but the technical details differ significantly.}
-Linkages model families, so they are tuples composed of fields but with a twist: they
-support late binding.
+\TT provides MLTT-style constructs that can be used to express families and
+family polymorphism. %rather than building it in directly.
+%
+Most notably, it extends MLTT with what we call \emph{linkages}.
+(Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
+family polymorphism in an OO setting, but the technical details differ significantly from the prior work.)
+
+\begin{itemize}
+[labelsep=*,leftmargin=1pc,itemsep=3pt]
+\item 
+Linkages model families, so they are like tuples composed of the fields.
+But there is a twist: they support late binding.
 Unlike dependent tuples where a later component is \emph{existentially}
@@
 quantified over---and thus polymorphic to---components preceding it.
-%
-\TT also features \emph{linkage transformers} that can be used to inductively
-construct linkages from other linkages, by inheriting or overriding components of a linkage,
-or by extending a linkage with new components.
-%
+
+\item
+\TT also features \emph{linkage transformers} that enable inductive
+construction of linkages (and thus families): a linkage can be constructed by inheriting or
+overriding components of an existing linkage, or by extending an existing
+linkage with new components.
+
+\item
 Inductive types are modeled as W-types~\cite{martin1984intuitionistic} and their
 extension as overriding.
-%
-Consistency and canonicity are proved by giving semantic models to the syntax.
+
+\item
+Consistency and canonicity of \TT are proved by giving semantic interpretations to the syntactic typing judgments.
+\end{itemize}
 
@@
 
-\noindentparagraph{MLTT.}
+\noindentparagraph{Brief review of MLTT.}
 
-\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT, and
-\cref{fig:fmltt-selected} \TT.
-%The new forms (relative to MLTT) are highlighted.
-We briefly review the base MLTT fragment first.
+\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT
+(and \cref{fig:fmltt-selected} \TT).
+%We briefly review the base MLTT fragment first.
 %
-As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
+As is conventional with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
 substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
@@
 that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
-$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.
+$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
+where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
 
@@
 
-Rules \ruleref{tm/wsup} and \ruleref{tm/wrec} require knowing~$\wsig$:
+Rules \ruleref{tm/wsup} and \ruleref{tm/wrec} require access to~$\wsig$:
 the W-type is exhaustively generated by its constructors, and
@@
 because a field representing the code $\wcode{\wsig}$ of a W-type has to
-expose different types to different succeeding fields.
-Succeeding fields that invoke
+expose different types to different fields that come after it.
+Later fields that invoke
 $\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
@@
 %
-Partial recursors automatically generated by the plugin
-can also be modeled in this way by invoking $\texttt{Wrec}$.
+Like \lsti{subst}, partial recursors in \Lang are axiomatized by the plugin
+(\cref{sec:latebind,sec:prec}), and they can similarly be defined in \TT
+using $\texttt{Wrec}$.
 

### 2022-11-10 coqimpl.tex
edits

@@
 programs that can be checked and evaluated by Coq.
-The translation is compatible with interactive theorem proving, in that
+The translation is compatible with interactive theorem proving (\ref{chg:tooling}), in that
 a family is translated piece by piece, allowing each field to be defined
@@
 The translation is modular and efficient, in that
-code compiled for non-further-bindable fields of a base family
-is shared with derived families without having to be rechecked.
+code compiled for fields of a base family
+can be shared with derived families without having to be rechecked.
 
@@
 
-For instance,
+As an example,
 consider field \lsti{env} in family \lsti{STLC}.
@@
 referenced through \lsti{self}.
-In particular, \lsti{env} is defined as \lsti{id}\,->\,\lsti{option ty},
+In particular, \lsti{env} is defined as \lsti{id}\,\lsti{->}\,\lsti{option ty},
 where \lsti{ty} is a late-bound reference to the \lsti{ty} field of the
 enclosing family.
-Hence, this reference to \lsti{ty} is translated in \lsti{STLC\_env} to \lsti{self.ty},
+Hence, this reference to \lsti{ty} is translated to \lsti{self.ty},
 which is manifestly polymorphic to the enclosing family.
@@
 with a derived family even if it extends \lsti{ty} (e.g., \lsti{STLCProd})---%
-no recompilation is needed because \lsti{self.ty} is not tied to a
-particular definition of \lsti{ty}.
+no recompilation is needed because \lsti{self.ty} is not tied to any
+concrete definition of \lsti{ty}.
 
 The type of \lsti{STLC\_env}'s \lsti{self} parameter is \lsti{STLC\_env\_Ctx},
-a module type constructed from the translation of \lsti{env}'s preceding
-field, \lsti{STLC\_ty}, and its context \lsti{STLC\_ty\_Ctx}.
+a module type constructed from \lsti{STLC\_ty}, 
+the translation of the field before \lsti{env},
+and its context \lsti{STLC\_ty\_Ctx}.
 In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is
-constructed from the translation of \lsti{ty}'s preceding field,
-\lsti{STLC\_subst}, and its context \lsti{STLC\_subst\_Ctx}.
+constructed from \lsti{STLC\_subst},
+the translation of the field before \lsti{ty},
+and its context \lsti{STLC\_subst\_Ctx}.
 Thus, the \lsti{self} parameter can be used to reference those and only
@@
 \lsti{STLC\_tm} that declares a \lsti{tm} type,
-its constructors (\lsti{tm_unit} etc.),
+its constructors (e.g., \lsti{tm_abs}),
 its partial recursor (\lsti{tm_prect_STLC}),
-and the propositional equalities on the partial recursor (\lsti{tm_unit_eq_STLC} etc.).
+and the propositional equalities on the partial recursor
+(e.g., \lsti{tm_abs_eq_STLC}; \cref{sec:prec}).
 
@@
 \lsti{STLCFix} is again translated to a module type \lsti{STLCFix\_tm} (\cref{fig:stlcfix-compiled}).
-It includes all names declared by \lsti{STLC\_tm}, and additionally declares
+It includes all the names declared by \lsti{STLC\_tm}
+via command \lsti!Include STLC\_tm(self)!,
+and additionally declares
 the new constructor \lsti{tm_fix} and related equalities.
@@
 As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
-so its definition is not exposed to the succeeding fields.
+so its definition is not exposed to the fields that come after it.
 Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
@@
 whose definitions \emph{are} available through the \lsti{self} parameter.
-For example, the type of \lsti{subst_tm_unit_eq} can be simplified to
-\lsti{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
-The equalities will be available for use in the checking and translation of
-the succeeding fields through their \lsti{self} parameters.
+So Coq can simplify, for example, \lsti{subst_tm_unit_eq}'s type to
+\lsti!\forall x t, self.subst self.tm_unit x t = self.tm_unit!.
+The equalities will be included and available for use in the
+translations of the subsequent fields through their \lsti{self}
+parameters.
 
 Importantly, code generated for the case handlers is shared with derived families.
-In \cref{fig:stlcfix-compiled}, \lsti{STLCFix\_subst\_Cases}
+In \cref{fig:stlcfix-compiled}, module \lsti{STLCFix\_subst\_Cases}
 reuses---without rechecking---all the case handlers in
-\lsti{STLC\_subst\_Cases} via the command \lsti{Include STLC\_subst\_Cases(self)}.
+\lsti{STLC\_subst\_Cases} via command \lsti{Include STLC\_subst\_Cases(self)}.
 
@@
 on the fields.
-Opaque definitions can be further bound, too (\cref{sec:override});
-they are translated to modules that export opaque definitions.
+Opaque fields can be further bound, too (\cref{sec:override});
+they are translated to modules that export opaque fields.
 
@@
 
-Upon the conclusion of a family definition (e.g., \lsti{End}~\lsti{STLCFix}),
-\lsti{self} parameters in the translations of the fields must be instantiated,
-so that references like \lsti{STLCFix.typesafe} can be translated.
-This is done by assembling the \lsti{self}-parameterized pieces into a single module.
-%
+Upon the conclusion of a family definition,
+a representation of the family is created.
 For example, module \lsti{STLC} in \cref{fig:stlc-compiled} is generated
 upon \lsti{End}~\lsti{STLC}.
-All the fields are added to this module in the same order as they appear in the family definition.
+%
+This module can be viewed as the ``fixed point'' of the
+\lsti{self}-parameterized translations.
+The ``fixed point'' is taken step by step, by adding the translation of
+each field to this module in the same order as they appear in the family
+definition.
 
 For the non-further-bindables, the translated modules are directly included
-(e.g., \lsti{Include STLC\_subst\_Cases} and \lsti{Include STLC\_env}).
+(e.g., \lsti{Include STLC\_env} and \lsti{Include STLC\_subst\_Cases} in \cref{fig:stlc-compiled}).
 %
-The instantiations of \lsti{self}s are implicit thanks to a Coq nicety:
+The instantiation of \lsti{self}s for these modules is implicit, thanks to a Coq nicety:
 when including a higher-order module, Coq automatically instantiates its
@@
 For instance, command \lsti{Include STLC\_subst\_Cases} is successfully executed,
-because Coq automatically instantiates \lsti{self} using the current
-module environment, which contains all the fields required by \lsti{STLC\_subst\_Cases\_Ctx}.
+because Coq automatically instantiates the \lsti{self} parameter using the current
+module environment, which by construction contains all the fields required by \lsti{STLC\_subst\_Cases\_Ctx}.
 
@@
 %
-\begin{itemize}[align=left,itemsep=0pt,labelsep=*,leftmargin=*]
+\begin{itemize}
+  [align=left,itemsep=2pt,labelsep=*,leftmargin=*]
 
 \item
-An inductive type \lsti{tm} is generated, instantiating the axiomatized
+In \cref{fig:stlc-compiled},
+an inductive type \lsti{tm} is generated, instantiating the axiomatized
 \lsti{tm} type and its constructors.
@@
 Similarly, \lsti{subst} is instantiated by applying the recursor \lsti{tm_rect}
-to the (already included) case handlers (\lsti{subst\_tm_unit}, etc.).
+to the (already included) case handlers. %(\lsti{subst\_tm_unit}, etc.).
 Propositional equalities on \lsti{subst} are then trivial, by \lsti{reflexivity}.
@@
 
-Module \lsti{STLCFix} in \cref{fig:stlcfix-compiled} is generated upon
-\lsti{End}~\lsti{STLCFix}, in the same fashion described above for \lsti{STLC}.
-The translation makes code and proof sharing evident.
+Module \lsti{STLCFix} in \cref{fig:stlcfix-compiled} is emitted upon
+\lsti{End}~\lsti{STLCFix}, in the same way as described above for \lsti{STLC}.
+The translation makes sharing evident.
 In particular, case handlers compiled for \lsti{STLC}
@@
 \lsti{STLC\_env} and \lsti{STLC\_typesafe}
-are also reused to construct module \lsti{STLCFix}.
+are also reused in the construction of module \lsti{STLCFix}.
 %
-One may argue that because the first four constructors of \lsti{tm} are
+One may argue that since the first four constructors of \lsti{tm} are
 repeated in \lsti{STLCFix}, the translation does not satisfy the
 modular compilation requirement.
-We could address this issue by using wrapper types,
+We could address this concern by using wrapper types,
 but we consider restating constructors a small price to pay in return
 for the clarity and concision of implementation.
-We emphasize that compiled case handlers
+We emphasize that compiled case handlers, such as \lsti!subst\_tm_abs!,
 are entirely reusable without rechecking, even with restated constructors.
-
+%
 Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)

### 2022-11-10 metatheory-condensed.tex
edits

@@
 W-types~\cite{martin1984intuitionistic} are a succinct way to model
-inductive types in MLTT;
-together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
+inductive types in MLTT.
+Together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
 whole host of inductive types~\cite{hugunin2021not}.
@@
 $T$ is the motive of the recursion (\ruleref{tyeq/casety}).
-This collection of case handlers then encodes those defined or inherited
+For simplicity, we model non-dependent motives.
+This collection of case handlers~$\lkg$ then encodes those defined and inherited
 by an \lsti{FRecursion} command in our plugin.
 We choose to type it with a linkage type $\TyLkg{\RecSig{\wsig}{T}}$
-to avoid introducing $n$-tuples, which linkages generalize.
+to avoid introducing non-dependent $n$-tuples, which linkages generalize.
 
@@
 %
-Such abstraction required by family polymorphism is supported via linkages.
+Such abstraction required by family polymorphism is supported via linkages,
+which we discuss next.
 
@@
 fields in $t$'s context can reuse~$t$---without retyping it---by
-projecting it (\ruleref{l/proj/2}) and then including it.
+projecting it and then including it.
 
@@
 the typing context containing the first field~\lsti{tm}.
-Because the W-type signature $\wsig_{\texttt{tm}}$ is exposed to this typing context,
-$\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
+The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
+this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
+So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
 as required by rule \ruleref{tm/wsup}.
@@
 \item
-Similarly, $\wsig_{\texttt{tm}}$ is exposed to the typing context of
+Similarly, $\wsig_{\texttt{tm}}$ is exposed in the typing context of
 $t_{10}$, which models the recursive function \lsti{subst} by invoking
@@
 \item
-By contrast, $\wsig_{\texttt{tm}}$ is hidden from the typing context of all other fields.
+By contrast, $\wsig_{\texttt{tm}}$ is hidden from the typing of all other fields.
 Their typing depends on the knowledge that \lsti!tm! has type~$\cU$,
@@
 For example, in \cref{fig:stlc-linkage-typing}, the typing of the case handlers of
-\lsti{subst} is agnostic of the definition of \lsti!tm!---it sees only $\text{\lsti!tm!}:\cU$---so
-they can be reused by a linkage modeling \lsti{STLCFix}.
+\lsti{subst} (e.g., $t_6$) is oblivious to the definition of \lsti!tm!---it sees only $\text{\lsti!tm!}:\cU$---so
+the case handlers can be reused by a linkage modeling \lsti{STLCFix}.
 
@@
 
-Inheritance and code reuse can already be expressed through projection
+Inheritance and code reuse can already be expressed through the projection
 of fields out of linkages and their inclusion into new linkages.
@@
 overrides $\wsig_{\texttt{tm}}$ with an extended
-signature~$\wsig'_{\texttt{tm}}$.
+signature~$\wsig'_{\texttt{tm}}$:
 \vspace{.9ex}

### 2022-11-10 metatheory-condensed.tex
clean up linkage tranformer appendix

@@
 case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
-Each case handler has a type of the form $\CaseSig{A}{B}{T}$, where
-$T$ is the motive of the recursion (\ruleref{tyeq/casety}).
-For simplicity, we model non-dependent motives.
-This collection of case handlers~$\lkg$ then encodes those defined and inherited
+Each case handler has a type of form $\CaseSig{A}{B}{T}$, where
+$T$ is the motive of the recursion (\ruleref{tyeq/casety});
+for simplicity, we model non-dependent motives for recursion.
+The collection of case handlers~$\lkg$ encodes those defined and inherited
 by an \lsti{FRecursion} command in our plugin.

### 2022-11-10 stlc-linkTran-typing.tex
Add information for LT exmaple.
clean up linkage tranformer appendix

@@
-In this appendix, we ...
+In this appendix, we sketch how to use the ``library'' of linkage
+transformers in \TT to inductively construct a linkage that models a
+derived family of \lsti{STLC}.
 
+We use \lsti{STLCBool} as an example. \lsti{STLCBool} extends
+\lsti{STLC} with boolean values (\lsti{tm_true} and \lsti{tm_false})
+and if-then-else expressions.
+The left column of the table below shows initial code excerpted from
+this family.
+Each cell in the last column defines a linkage transformer~$h_i$
+inductively constructed from the linkage transformers $h_0$, ..., $h_{i-1}$.
+The goal is to eventually construct a linkage transformer $h_n$ representing
+the entire family \lsti{STLCBool}.
+
+Of note is the two grayed rows. They are constructing a linkage~$h_\beta$
+containing the case handlers for \lsti{subst}.
+This linkage $h_\beta$ is then appended as a nested linkage to $h_3$.
+
+\medskip
+
+\if 0
 % Please add the following required packages to your document preamble:
@@
 \begin{centered}
+  \small
+\renewcommand*{\arraystretch}{1.35}
 \begin{tabular}{|l|l|l|l|}
@@
   \rowcolor[HTML]{FFFC9E} 
-  surface syntax of program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i'}{T_i'}$ \\ \hline
-  \lsti!Family STLC_bool...!   & 0        & $\{\}$                                     & $\{\}$                                     \\ \cline{2-4} 
+  surface-syntax program & $i$ & $\lsig_i = \LSigAdd{\lsig_{i-1}}{s_i}{T_i}$ & $\lsig_i' = \LSigAdd{\lsig_{i-1}'}{s_i'}{T_i'}$ \\ \hline
+  \lsti!Family STLCBool...!   & 0        & $\{\}$                                     & $\{\}$                                     \\ \cline{2-4} 
   \lsti!FInductive tm += !      & 1        & $\{"tm" : \TyS{\wcode{\wsig_{\texttt{tm}}}}\}$         & $\{"tm" : \TyS{\wcode{\wsig_{\texttt{tm}}'}}\}$        \\ \cline{2-4} 
-  \lsti! (*existent constr*)!    & 2        & $\{..; "tm_var": \El{"tm"}, .. \}$         & $\{..;"tm_var": \El{"tm"},.. \}$           \\ \cline{2-4} 
+  \codecomment{existent constr.}    & 2        & $\{..; "tm_var": \El{"tm"}, .. \}$         & $\{..;"tm_var": \El{"tm"},.. \}$           \\ \cline{2-4} 
   \lsti! .. | tm_true | .. !   & 3        & $\lsig_2$                                  & $\{..,"tm_true":\El{"tm"} \}$              \\ \cline{2-4} 
@@
 
+\fi
+
 % Please add the following required packages to your document preamble:
@@
 \begin{centered}
-
+\small
+\renewcommand*{\arraystretch}{1.25}
 \begin{tabular}{|l|l|l|}
@@
 \rowcolor[HTML]{FFFC9E} 
-surface syntax of program & $i$      & $\goodInh{..}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
-\lsti!Family STLC_bool!   & 0        & $\InhId$                                             \\ \cline{2-3} 
+surface-syntax program & $i$      & $\goodInh{\cdot}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
+\lsti!Family STLCBool!   & 0        & $\InhId$                                             \\ \cline{2-3} 
 \lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_{\texttt{tm}}'}}$                      \\ \cline{2-3} 
-\lsti! (*existent constr*) !   & 2        & $\InhOv{h_1}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$           \\ \cline{2-3} 
-\lsti! .. | tm_true | ..!    & 3        & $\InhExt{h_2}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$          \\ \cline{2-3} 
+\codecomment{existing constructors}   & 2        & $\InhOv{h_1}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$           \\ \cline{2-3} 
+\lsti!| tm_true |!\ \;\dadada    & 3        & $\InhExt{h_2}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$          \\ \cline{2-3} 
 \rowcolor[HTML]{CDCDCD} 
-\lsti!  FRecursion subst!       &          &                                                      \\ \cline{2-3} 
+\lsti!FRecursion subst!\ \;\dadada\ \;\lsti!+=! & $\alpha$ & $\InhId$ \\ \cline{2-3} 
 \rowcolor[HTML]{CDCDCD} 
-\lsti!  (* Inherit ... *)!      & $\alpha$ & $\InhId$                                             \\ \cline{2-3} 
-\rowcolor[HTML]{CDCDCD} 
-\lsti!  Case tm_true!          & $\beta$  & $\InhExt{h_\alpha}{..}$                              \\ \cline{2-3} 
+\lsti!  Case tm_true := !\dadada. & $\beta$  & $\InhExt{h_\alpha}{...}$ \\ \cline{2-3} 
 \rowcolor[HTML]{FFFFFF} 
-\lsti!  End subst!              & 4        & $\InhNest{h_3}{h_\beta}$ with $\uparrow^s_3$         \\ \cline{2-3} 
-                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_{\texttt{tm}}'}{t}{"subst'"}}$ \\ \cline{2-3} 
-\lsti!(* Do the same on ty*)! & 6        & ...                                                  \\ \cline{2-3} 
-\lsti!(* Inherit env *)!      & 7        & $\InhInh{h_6}$ with $\uparrow^s_6$                   \\ \hline
+\lsti!End subst.!              & 4        & $\InhNest{h_3}{h_\beta}$
+  %with $\uparrow^s_3$
+  \\ \cline{2-3} 
+                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_{\texttt{tm}}'}{t}{"substCases"}}$ \\ \cline{2-3} 
+\lsti!FInductive ty += !\dadada. & 6        & ...                                                  \\ \cline{2-3} 
+\codecomment{Inherit "env"}   & 7        & $\InhInh{h_6}$
+  %with $\uparrow^s_6$
+  \\ \cline{2-3} 
+\codecomment{Inherit "empty"}   & 8        & $\InhInh{h_7}$
+  \\ \cline{2-3} 
+\dadada   & \dadada & \dadada
+  \\ \hline
 \end{tabular}
@@
 
+\if 0
+
 We take \lsti!Family STLC_bool!
 extending \lsti!STLC!
-as an example.
+as an example, based on the example \cref{fig:stlc-linkage-typing} and \cref{fig:stlc-mechanized}.
 We list out the surface syntax and their translation into \TT.
@@
 There are two upcasting operations $\goodTerm{A_3'}{\uparrow^s_3}{A_3}$ and $\goodTerm{A_6'}{\uparrow^s_6}{A_6}$ (omitting debruijn indices here) for checking the compatibility between the omitted $A_i$ and $A_i'$.
+
+\fi

### 2022-11-10 body.tex
Working on Example for Linkage Transformer.
move tabulars
clean up linkage tranformer appendix
working on all-typing-rule.tex

@@
 \section{Metatheory in (pseudo-)Agda}
+\label{sec:agda-files}
 We make four agda source code files (publicly and anonymously) openly avaliable, and they include the corresponding QIIT syntax definition and models.
@@
 
-\section{Typing Rules in one Figure}
-\include{all-typing-rule.tex}
+\section{A More Complete Set of Typing Rules for \TT}
+\input{all-typing-rule.tex}
 
+\section{Using \TT's Linkage Transformers to Model a Derived Family}
+\input{stlc-linkTran-typing.tex}
+
+\if 0
 \section{Metatheory, in detail}
-
 \label{sec:metatheory}
@@
 \input{coq-example.tex}
+\fi

### 2022-11-10 metatheory-condensed.tex
working on all-typing-rule.tex

@@
 \noindent
-A supplemental appendix gives a more complete example
-showing how families are modeled and
-type-checked using linkages and linkage transformers.
+A supplemental appendix sketches how the other introduction forms of linkage transformers
+can be used to model the construction of a derived family as a linkage transformer.
 

### 2022-11-10 agda-files.tex
appendix

--- 2022-11-09 body.tex
@@
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-\maketitle
-
-% lstlisting coq style (inspired from a file of Assia Mahboubi)
-% https://tex.stackexchange.com/questions/434523/coq-syntax-highlighting
-\section{Introduction}
-\label{sec:intro}
-\input{introduction.tex}
-
-% Concept 
-% Contribution
-
-\section{Design Requirements and Challenges}
-% what feature we want
-% what problem/formalization we want to achieve
-% Before/After Comparison of examples (with/without Fampoly)
-\label{sec:background+challenge}
-\input{challenge.tex}
-
-\section{Language Design}
-% we need to make this section before Metatheory
-% because a lot of our formulation in the metatheory
-%   is inspired by the implementation
-\label{sec:lang-design}
-\input{lang-design.tex}
-
-\section{Compiling Family Polymorphism to Parameterized Modules}
-% we need to make this section before Metatheory
-% because a lot of our formulation in the metatheory
-%   is inspired by the implementation
-\label{sec:coqimpl}
-\input{coqimpl.tex}
-
-\section{\TT: A Core Dependent Type Theory}
-\label{sec:metatheory2}
-\input{metatheory-condensed.tex}
-
-\section{Case Studies}
-\label{sec:casestudies}
-\input{casestudies.tex}
-
-\section{Related Work}\label{sec:related-work}
-\input{related-work.tex}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-\input{conclusion.tex}
-
-\setlength{\bibsep}{.8ex}
-\bibliographystyle{tex-macros/ACM-Reference-Format.bst}
-\bibliography{refs.bib}
-
-\appendix
-
-\section{Metatheory in (pseudo-)Agda}
 We make four agda source code files (publicly and anonymously) openly avaliable, and they include the corresponding QIIT syntax definition and models.
@@
 We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.
-
-\section{Typing Rules in one Figure}
-\include{all-typing-rule.tex}
-
-\section{Metatheory, in detail}
-
-\label{sec:metatheory}
-\input{metatheory.tex}
-
-\section{Example in Metatheory}\label{sec:coqexample}
-\input{coq-example.tex}

### 2022-11-10 body.tex
appendix

@@
 
-\section{Metatheory in (pseudo-)Agda}
-\label{sec:agda-files}
-We make four agda source code files (publicly and anonymously) openly avaliable, and they include the corresponding QIIT syntax definition and models.
+\newpage
 
-For syntax, please check \\ \href{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}.
-This is done by mainly following the formulation \citet{altkap2016}.
-This is also very close to \citet{coquand2018canonicity} and \citet{sterling2019algebraic} but we formulate our syntax in a type-theoretic framework.
+\section{More Typing Rules for \TT}
+\input{all-typing-rule.tex}
 
-For syntactic translation, please check \\ \href{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}.
-The syntactic translation here is to translate most of the novel feature introduced in this paper into a standard MLTT.
-The essence here is to translate linkage into a sigma type  with dependent function inside---and the dependent function is simulating the nature of ``late-binding'' of each field.
-
-This is the first QIIT-model we give.
-A QIIT-model can be roughly consider as an arbitrary mapping from the QIIT data(our syntax in this case), also respecting the quotient(judgemental equality in this case).
-Here, the syntactic translation model is mapping to (a subpart of) the original QIIT data (itself).
-
-For consistency model, please check \\ \href{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}.
-This is following the standard model from \citet{altkap2016,kaposi2017type,kaposi2019gluing}.
-
-For canonicity model, please check \\ \href{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}.
-This is mainly following \citet{coquand2018canonicity,sterling2019algebraic}, but we carried out our argument in a type-theortic framework.
-In other references, it is also called reducibility argument, as an instance of proof-relevant logical relation.
-Our model can be considered as a verbose/more concrete version of \citet{kaposi2019gluing}.
-
-For both consistency and canonicity model, the key part is still about handling the novel feature.
-Thanks to the insight from syntactic translation, we can always inspired by consistency and canonicity model of sigma type and dependent function in standard MLTT.
-
-We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.
-
-\section{A More Complete Set of Typing Rules for \TT}
-\input{all-typing-rule.tex}
+\newpage
 
@@
 \input{stlc-linkTran-typing.tex}
+
+\newpage
+
+\section{The Complete Formalization and Proofs Written in Agda Syntax}
+\label{sec:agda-files}
+\input{agda-files.tex}
 
@@
 
-\section{Example in Metatheory}\label{sec:coqexample}
+\section{Example in Metatheory}
+\label{sec:coqexample}
 \input{coq-example.tex}
+
 \fi

### 2022-11-10 metatheory-condensed.tex
rewording

@@
 \item
-\TT also features \emph{linkage transformers} that enable inductive
-construction of linkages (and thus families): a linkage can be constructed by inheriting or
+\TT also features a set of \emph{linkage transformers} that enable inductive
+construction of families: a linkage can be constructed by inheriting or
 overriding components of an existing linkage, or by extending an existing

### 2022-11-10 stlc-linkTran-typing.tex
tweaks

@@
 Each cell in the last column defines a linkage transformer~$h_i$
-inductively constructed from the linkage transformers $h_0$, ..., $h_{i-1}$.
+inductively constructed from the linkage transformers $h_0$, ..., $h_{i-1}$
+using one of the introduction forms.
 The goal is to eventually construct a linkage transformer $h_n$ representing
@@
 
-Of note is the two grayed rows. They are constructing a linkage~$h_\beta$
+Of note is the two grayed rows. They are constructing a linkage transformer~$h_\beta$
 containing the case handlers for \lsti{subst}.
-This linkage $h_\beta$ is then appended as a nested linkage to $h_3$.
+This $h_\beta$ is then appended to $h_3$ as a nested linkage transformer.
 
@@
 \rowcolor[HTML]{CDCDCD} 
-\lsti!  Case tm_true := !\dadada. & $\beta$  & $\InhExt{h_\alpha}{...}$ \\ \cline{2-3} 
+\lsti!  Case tm_true := !\dadada. \lsti! Case !\dadada & $\beta$  & $\InhExt{h_\alpha}{...}$ \\ \cline{2-3} 
 \rowcolor[HTML]{FFFFFF} 

### 2022-11-10 lang-design.tex
edits in FMLTT sec
Some Comments.

@@
 \subsection{Composing Families as Mixins}
+\label{sec:mixin}
 
@@
 families.
-A family like \lsti{STLCFix} can be viewed as a family-to-family functor---%
+A family like \lsti{STLCFix} can be viewed as a family-to-family functor\EDJ{The reader might be confused why look at them as functor, as the parent seems unchanged.
+One insight is that the parent of nested family, in Zhang OOPSLA 2017, is late-bound.
+Thus it has to be a functor.
+}---%
 and hence a mixin, in the sense of \cite{flatt1998mixin}---that
@@
 Mixin composition is a form of multiple inheritance, which may lead to
-name conflicts in general.
+name conflicts in general.\EDJ{Only name conflicts on overridable fields can be resolved like this. Name/definition conflicts for normal fields cannot be resolved (cannot be mixed in). }
 \Lang requires the programmer to resolve conflicts by
@@
 \subsection{Injectivity and Disjointness of Constructors via Partial Recursors}
+\label{sec:prec}
 

### 2022-11-10 casestudies.tex
More Comments.

@@
 inductive types for modeling extensible context-free grammars and derive
-decision procedures for them.
+decision procedures for them.\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}
 

### 2022-11-10 cc-condensed.tex
More Comments.

@@
 \citet{altkap2016}, extending it to handle linkages.
-Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products,
+Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products\EDJ{Suggesting changing to ``can be interpreted into non-dependent pair with dependent function for late-binding''},
 as rule \ruleref{l/add} indicates.
@@
 type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
-When canonicity is proved in a constructive meta\-logic, its proof
+When the canonicity theorem is proved in a constructive meta\-logic, its proof
 amounts to a normalization function for closed terms of the ground type.\EDJ{Suggesting use "boolean type".
@@
 into a logical predicate on closed terms: the predicate includes all
-closed, ``reducible'' terms of type~$T$.\EDJ{I will add a it is my bad to introduce "reducible" without defining it first.
-But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}
+closed, ``reducible'' terms of type~$T$.\EDJ{I will add a quote around reducible here.
+it is my bad to introduce "reducible" without defining it first.
+But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}\EDJreply{Because in the operational semantic, using ``reducible'' is totally fine.
+But in declarative style, only ``canonicity'' this word is making sense.
+But this ``canonicity'' is not friendly to the audience not familar with declartive style.}
 The proof is available as a supplemental file in Agda syntax.

### 2022-11-10 coqimpl.tex
Some Comments.
More Comments.

@@
 Importantly,
-\lsti{STLC\_tm} merely declares the existence of these names and their types;
+\lsti{STLC\_tm} merely declares the existence of these names and their types;\EDJ{What does this sentence mean?}
 it does not specify their definitions or proofs.
@@
 %
-As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
+As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,\EDJ{If we don't allow further bound of case handler(mentioned in your last reply), how do the user further bound the subst?
+I mean in the surface syntax}
 so its definition is not exposed to the fields that come after it.

### 2022-11-10 metatheory-condensed.tex
More Comments.

@@
 \TT is intended as a foundational model. So unlike our programmer-facing plugin,
-typing is structural rather than nominal, and
+typing is structural rather than nominal\EDJ{I don't actually know what do you mean by structural here. I think surface syntax (Coq) is pretty structual as well.}, and
 fields automatically axiomatized by the plugin require explicit definitions in \TT.
@@
 fields in $t$'s context can reuse~$t$---without retyping it---by
-projecting it and then including it.
+projecting it and then including it\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.
 
 This type~$A$ abstracting the context in \ruleref{l/add} and
-\ruleref{lsig/add} is not necessarily the same as those recorded by~$\lsig$,
+\ruleref{lsig/add} is not necessarily the same as those recorded by~$\lsig$\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
 because a field representing the code $\wcode{\wsig}$ of a W-type has to
@@
 from all other fields so that they can be reused
-for a W-type signature~$\wsig'$ extending~$\wsig$.
+for a W-type signature~$\wsig'$ extending~$\wsig$\EDJ{Suggest changing to ``so that they can be reused
+in a different context where W-type signature~$\wsig'$ is extended to~$\wsig$''}.
 

### 2022-11-10 metatheory-condensed.tex
small edits

--- 2022-11-10 metatheory-condensed.tex
@@
 \item
-\TT also features a set of \emph{linkage transformers} that enable inductive
-construction of families: a linkage can be constructed by inheriting or
+\TT also features \emph{linkage transformers} modeling the inductive
+construction of families from other families.
+They allow a linkage to be created by inheriting or
 overriding components of an existing linkage, or by extending an existing

### 2022-11-10 related-work.tex

@@
 In addition, extra programmer effort may be required, such as
-having to manually prove additional well-formedness conditions on the algebras.
+having to manually prove additional well-formedness conditions on the algebras\EDJ{What is this `algebra'?
+I mean if it is a very frequent word to the audience  familiar with DTC then removing explanation is fine.}.
 
@@
 
-\noindentparagraph{Solutions to the expression problem} abound.
+\noindentparagraph{Solutions to the expression problem} abound\EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}.
 Almost all solutions involve some form of either explicit or implicit parameterization.

### 2022-11-10 coqimpl.tex
reply

@@
 %
-As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,\EDJ{If we don't allow further bound of case handler(mentioned in your last reply), how do the user further bound the subst?
-I mean in the surface syntax}
+As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,\EDJ{If we don't allow further binding case handler(mentioned in your last reply), how do the user further bind the subst?
+I mean in the surface syntax}\YZreply{The position is that individual case handlers cannot be further bound, but the recursive function itself can be further bound by acquiring new case handlers.}
 so its definition is not exposed to the fields that come after it.

### 2022-11-10 lang-design.tex
More Comments.

@@
 
-In \cref{fig:stlc-isorec-prod}, "STLCFixIsorec" is declared as
+In \cref{fig:stlc-isorec-prod}, "STLCFixIsorec" is declared as \EDJ{I think this paragraph is confusing for the audience about ``what is the relationship between this graph and the above graph?'' or ``why did you mention functor in the first place?''.
+I think saying that mixin is an operation on two functors might be helpful, but does introduce a great deal of complexity in the text.
+What's more, w know mixin of two (standalone) families is not really making sense because two families might have a bunch of duplication.
+Thus mixin of two functors is making sense.
+I think this point needs to be mentioned (still the complexity of text is a problem).
+}
 an "STLC" extension that mixes in "STLCFix" and "STLCIsorec".
@@
 allowing properties about "tm_fix",
-such as \lsti{\neg(tm_var "x" = tm_fix "y" t)}, to be proved.
+such as \lsti{\neg(tm_var "x" = tm_fix "y" t)}, to be proved.\EDJ{Actually this one is provable using the first partial recursor because LHS is "Some sth" RHS is "None".
+An approrpiate example is STLC_bool where "tm_true = tm_false" because using first partial recursor both will reduced to "None", only the second partial recursor can distinguish them.}
 

### 2022-11-10 related-work.tex
reply

@@
 
-\noindentparagraph{Solutions to the expression problem} abound\EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}.
-Almost all solutions involve some form of either explicit or implicit parameterization.
+\noindentparagraph{Solutions to the expression problem} \!abound.\EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}\YZreply{Yes}
+Almost all involve some form of either explicit or implicit parameterization
+as extensibility hooks.
 Our approach is the first that applies family polymorphism~\cite{ernst2001family},
-an idea found in the context of OO programming,
+an idea from the context of OO languages,
 to the context of mechanized proofs.

### 2022-11-10 coqimpl.tex
More Comments.

@@
 the translation of the field before \lsti{env},
-and its context \lsti{STLC\_ty\_Ctx}.
+and its context \lsti{STLC\_ty\_Ctx}\EDJ{Why I see in the figure there are only two includes, i.e. only stlcty and stlctyctx}.
 In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is
@@
 \lsti{STLC\_tm} that declares a \lsti{tm} type,
-its constructors (e.g., \lsti{tm_abs}),
+its constructors (e.g., \lsti{tm_abs}) \EDJ{I am not sure if it is better we emphasize in the following way :`` that declares a tm type, its constructors (e.g.
+tmabs) as a normal function,..
+'' i.e.
+emphasize it is a normal function in the module type},
 its partial recursor (\lsti{tm_prect_STLC}),

### 2022-11-10 coqimpl.tex
Resolve comments.

@@
 %
-As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,\EDJ{If we don't allow further binding case handler(mentioned in your last reply), how do the user further bind the subst?
+As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
+% \EDJ{If we don't allow further binding case handler(mentioned in your last reply), how do the user further bind the subst?
 I mean in the surface syntax}\YZreply{The position is that individual case handlers cannot be further bound, but the recursive function itself can be further bound by acquiring new case handlers.}

### 2022-11-10 coqimpl.tex
More Comments.

@@
 its partial recursor (\lsti{tm_prect_STLC}),
-and the propositional equalities on the partial recursor
+and the propositional equalities\EDJ{I propose to change this sentence to ``and the prop eq about computational behaviour on the partial recursor''.
+And thus in the other parts of the paper, we can use the word ``computational behaviour'' instead of prop eq.
+Because just saying prop eq seems a bit of lack of context/details.} on the partial recursor
 (e.g., \lsti{tm_abs_eq_STLC}; \cref{sec:prec}).
@@
 and then a module type declaring the existence of the recursive function as well as
-its propositional equalities.
+its propositional equalities.\EDJ{Simiarly, here I think using ``computational behaviour'' is more accurate than using the word ``prop eq''}
 

### 2022-11-10 metatheory-condensed.tex
More Comments.

@@
 is not automated in \TT.
-In particular, when no hiding is needed, $\seal$ is $\var{0}$ (the identity) and $A$ is
+In particular, when no hiding is needed, $\seal$ is $\var{0}$ (the identity)\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right?
+Or you meant to say it is identity ``sealing''?} and $A$ is
 $\TyTkg{\lsig}$.

### 2022-11-10 metatheory-condensed.tex
no longer mentioning structural typing

--- 2022-11-10 metatheory-condensed.tex
@@
 \TT is intended as a foundational model. So unlike our programmer-facing plugin,
-typing is structural rather than nominal\EDJ{I don't actually know what do you mean by structural here. I think surface syntax (Coq) is pretty structual as well.}, and
+%typing is structural rather than nominal\EDJ{I don't actually know what do you mean by structural here. I think surface syntax (Coq) is pretty structual as well.}, and
 fields automatically axiomatized by the plugin require explicit definitions in \TT.
@@
 \item 
-Linkages model families, so they are like tuples composed of the fields.
-But there is a twist: they support late binding.
+Linkages model families; they are like tuples composed of the fields,
+with field names represented by variable bindings.
+But there is a twist: linkages support late binding.
 Unlike dependent tuples where a later component is \emph{existentially}

### 2022-11-10 coqimpl.tex
address a comment

--- 2022-11-10 coqimpl.tex
@@
 the translation of the field before \lsti{env},
-and its context \lsti{STLC\_ty\_Ctx}\EDJ{Why I see in the figure there are only two includes, i.e. only stlcty and stlctyctx}.
+and its context \lsti{STLC\_ty\_Ctx}.
 In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is
@@
 \lsti{STLC\_tm} that declares a \lsti{tm} type,
-its constructors (e.g., \lsti{tm_abs}) \EDJ{I am not sure if it is better we emphasize in the following way :`` that declares a tm type, its constructors (e.g.
-tmabs) as a normal function,..
-'' i.e.
-emphasize it is a normal function in the module type},
+its constructors (e.g., \lsti{tm_abs}),
 its partial recursor (\lsti{tm_prect_STLC}),
@@
 Importantly,
-\lsti{STLC\_tm} merely declares the existence of these names and their types;\EDJ{What does this sentence mean?}
-it does not specify their definitions or proofs.
-Knowing the existence of these names and their types
-suffices to type-check and translate the rest of the family.
-Leaving the definitions and proofs open enables \lsti{STLC} and \lsti{STLCFix}
+\lsti{STLC\_tm} merely declares the existence of these names and their types;
+it does not specify their definitions.
+Having these names and their types available through the context parameters (\lsti{self})
+suffices for the translations of the subsequent fields to be type-checked by Coq.
+Leaving the definitions open enables \lsti{STLC} and \lsti{STLCFix}
 to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and \lsti{End}~\lsti{STLCFix}.

### 2022-11-10 related-work.tex

@@
 
-\noindentparagraph{Solutions to the expression problem} \!abound.\EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}\YZreply{Yes}
+\noindentparagraph{Solutions to the expression problem} \!abound.
+% \EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}\YZreply{Yes}
 Almost all involve some form of either explicit or implicit parameterization

### 2022-11-10 lang-design.tex
bugfix
questino about comment

@@
 %(namely "tm_var_eq_STLC" and "tm_abs_eq_STLC"),
-obtaining \lsti{Some 1 = Some 4},
+obtaining \lsti{Some 1 = Some 3},
 from which \lsti{False} easily follows.
@@
 such as \lsti{\neg(tm_var "x" = tm_fix "y" t)}, to be proved.\EDJ{Actually this one is provable using the first partial recursor because LHS is "Some sth" RHS is "None".
-An approrpiate example is STLC_bool where "tm_true = tm_false" because using first partial recursor both will reduced to "None", only the second partial recursor can distinguish them.}
+An approrpiate example is STLC_bool where "tm_true = tm_false" because using first partial recursor both will reduced to "None", only the second partial recursor can distinguish them.}\YZreply{To avoid introducing another family, can we use \lsti{tm_fix x t = tm_fix x' t' -> t = t'} as an example?}
 

### 2022-11-10 lang-design.tex
Reply the comment.

@@
 such as \lsti{\neg(tm_var "x" = tm_fix "y" t)}, to be proved.\EDJ{Actually this one is provable using the first partial recursor because LHS is "Some sth" RHS is "None".
-An approrpiate example is STLC_bool where "tm_true = tm_false" because using first partial recursor both will reduced to "None", only the second partial recursor can distinguish them.}\YZreply{To avoid introducing another family, can we use \lsti{tm_fix x t = tm_fix x' t' -> t = t'} as an example?}
+An approrpiate example is STLC_bool where "tm_true = tm_false" because using first partial recursor both will reduced to "None", only the second partial recursor can distinguish them.}\YZreply{To avoid introducing another family, can we use \lsti{tm_fix x t = tm_fix x' t' -> t = t'} as an example?}\EDJreply{That is fine I think.
+I think you already declared the existence of ``finjection'' earlier.}
 

### 2022-11-10 lang-design.tex
address a comment about functors

--- 2022-11-10 lang-design.tex
@@
 families.
-A family like \lsti{STLCFix} can be viewed as a family-to-family functor\EDJ{The reader might be confused why look at them as functor, as the parent seems unchanged.
+A family like \lsti{STLCFix} can be viewed as a family-to-family functor---%\EDJ{The reader might be confused why look at them as functor, as the parent seems unchanged.
 One insight is that the parent of nested family, in Zhang OOPSLA 2017, is late-bound.
 Thus it has to be a functor.
-}---%
+}
 and hence a mixin, in the sense of \cite{flatt1998mixin}---that
-transforms any family providing the base STLC functionalities into a new
+transforms any family providing the base STLC functionalities
+(i.e., \lsti{STLC} or a derived family there of) into a new
 family additionally supporting fixpoints.

### 2022-11-10 coqimpl.tex

@@
 and checked separately.
-The translation is modular and efficient, in that
+The translation is modular and efficient\EDJ{I am very hesitant to say efficient :( }, in that
 code compiled for fields of a base family

### 2022-11-10 lang-design.tex
resolve a comment

--- 2022-11-10 lang-design.tex
@@
 (called "tm_prect_STLCFix") and its equalities are available,
-allowing properties about "tm_fix",
-such as \lsti{\neg(tm_var "x" = tm_fix "y" t)}, to be proved.\EDJ{Actually this one is provable using the first partial recursor because LHS is "Some sth" RHS is "None".
-An approrpiate example is STLC_bool where "tm_true = tm_false" because using first partial recursor both will reduced to "None", only the second partial recursor can distinguish them.}\YZreply{To avoid introducing another family, can we use \lsti{tm_fix x t = tm_fix x' t' -> t = t'} as an example?}
+allowing \lsti{tm_fix} properties such as \lsti[basicstyle=\fontsize{8.5}{9}]{tm_fix x t1 = tm_fix x t2 -> t1 = t2} to be proved.
 

### 2022-11-10 lang-design.tex
use "computational behaviors" in lang-design.tex

@@
 That is, \lsti{\forall x t, subst tm_unit x t = tm_unit},
-as a proposition (\lsti{Prop}), should still hold.
+as a proposition (\lsti{Prop}) about the computational behavior of \lsti{subst},
+should still hold.
 After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
-from a base family to a derived family.
+from a base family to a derived family;
 %\EDJ{I think the reviewer might ask : what if the subst late-bound tm_unit case/why the subst cases cannot be late-bound? it not necessary return tm_unit in those case. 
@@
 That is one of the advantage of looking FRecursion as a family.}\YZreply{Currently the paper is taking the position that case handlers cannot be further bound, to make the story simpler.}
-What varies is \lsti{subst} itself as a recursive function combining the case handlers.
-%
+what can vary is \lsti{subst} itself as a recursive function combining the case handlers.
+
 Based on this insight, \Lang automatically generates
@@
 simplifying \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
-Under the hood, \lsti{fsimpl} works by rewriting applications of the
-axiomatized recursive function using the axiomatized propositional equalities.
+\lsti{fsimpl} works by rewriting applications of the
+axiomatized recursive function using the axiomatized equalities about its computational behaviors.
 
@@
 As expected, "tm_prect_STLC" is axiomatized along with
-four (propositional) equalities, one for each constructor of "tm".
+four equalities describing its computational behaviors, one for each constructor.
 For instance, the equality for constructor "tm_abs" is as follows:
@@
 \noindent
-and then rewrites the above equality using two of the axiomatized
-equalities about "tm_prect_STLC",
+and then rewrites the above equality using the axiomatized computational behaviors
+of "tm_prect_STLC",
 %(namely "tm_var_eq_STLC" and "tm_abs_eq_STLC"),
@@
 from which \lsti{False} easily follows.
-
+%
 Note that the proof term generated by \lsti{fdiscriminate H} in "STLC"
 is reusable by family "STLCFix", because the partial recursor and its
-equalities are compatible with the late binding of "tm".
-Also note that within family "STLCFix", a second partial recursor
-(called "tm_prect_STLCFix") and its equalities are available,
-allowing \lsti{tm_fix} properties such as \lsti[basicstyle=\fontsize{8.5}{9}]{tm_fix x t1 = tm_fix x t2 -> t1 = t2} to be proved.
+computational behaviors are compatible with the late binding of "tm".
+
+Within family "STLCFix", a second partial recursor
+(called "tm_prect_STLCFix") and its computational behaviors are axiomatized,
+so properties such as \lsti[basicstyle=\fontsize{8.25}{9},columns=fixed,basewidth=0.9ex]{tm_fix x t1 = tm_fix x t2 ->  t1 = t2} can be proved.
 

### 2022-11-10 coqimpl.tex
resolve comment

@@
 The type of \lsti{STLC\_env}'s \lsti{self} parameter is \lsti{STLC\_env\_Ctx},
-a module type constructed from \lsti{STLC\_ty}, 
-the translation of the field before \lsti{env},
-and its context \lsti{STLC\_ty\_Ctx}\EDJ{Why I see in the figure there are only two includes, i.e. only stlcty and stlctyctx}.
+a module type constructed from \lsti{STLC\_ty}
+(i.e., the translation of the field before \lsti{env})
+and its context \lsti{STLC\_ty\_Ctx}.
 In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is

### 2022-11-10 lang-design.tex
address a comment

@@
 
-Mixin composition is a form of multiple inheritance, which may lead to
-name conflicts in general.\EDJ{Only name conflicts on overridable fields can be resolved like this. Name/definition conflicts for normal fields cannot be resolved (cannot be mixed in). }
+Mixin composition is a form of multiple inheritance, which may cause
+name conflicts in general. %\EDJ{Only name conflicts on overridable fields can be resolved like this. Name/definition conflicts for normal fields cannot be resolved (cannot be mixed in). }
 \Lang requires the programmer to resolve conflicts by
-explicitly overriding conflicted fields.
+overriding conflicted overridable fields.
 
@@
 
+\begin{centered}
+\begin{minipage}{\textwidth}
+\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
+\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
+tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
+  [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
+  [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
+\end{lstlisting}
+\end{minipage}
+\end{centered}
+
 %\begin{centered}
-%\begin{minipage}{\textwidth}
+%\begin{minipage}{.97\textwidth}
 %\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
@@
 %                [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
-%  [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
+%                [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
+%                \forall t, option (P t)
 %\end{lstlisting}
@@
 %\end{centered}
-
-\begin{centered}
-\begin{minipage}{.97\textwidth}
-\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
-\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
-tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
-                [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
-                [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
-                \forall t, option (P t)
-\end{lstlisting}
-\end{minipage}
-\end{centered}
 

### 2022-11-10 lang-design.tex
space hacks

@@
 its eliminator would have the following type:
+
+%\begin{centered}
+%\begin{minipage}{.995\textwidth}
+%\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
+%\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
+%tm_rect : \forall (P : tm -> Type), [@\hlcase{tm-recursor-case-unit}@]P tm_unit[@\tikzmarkend{tm-recursor-case-unit}@] -> [@\hlcase{tm-recursor-case-var}@](\forall x, P (tm_var x))[@\tikzmarkend{tm-recursor-case-var}@] ->
+%  [@\hlcase{tm-recursor-case-abs}@](\forall x t, P t -> P (tm_abs x t))[@\tikzmarkend{tm-recursor-case-abs}@] -> [@\hlcase{tm-recursor-case-app}@](\forall t1, P t1 -> \forall t2, P t2 -> P (tm_app t1 t2))[@\tikzmarkend{tm-recursor-case-app}@] -> \forall t, P t
+%\end{lstlisting}
+%\end{minipage}
+%\end{centered}
 
@@
 In an OO language, a subclass can override methods of a superclass.
-Similar expressiveness is useful for metatheory mechanization, too.
+Similar expressivity is useful for mechanizing proofs, too.
 %
@@
 entirely, as we observe in our case studies (\cref{sec:casestudies}).
+
+Overriding is potentially incompatible with having equalities on late bound names, however.
 %
-Overriding is potentially incompatible with having equalities on late bound names, however.
-
 Coq distinguishes \emph{opaque} definitions from \emph{transparent} ones.
@@
 %
-Such circularity would easily lead to logical inconsistency:
+Such circularity would easily lead to logical inconsistency.
+Consider the following two families, where "B" extends "A":
 %\YZ{Mention non-exhaustivity/overriding already poses challenges to ensuring consistency.}
@@
 \noindent
-In the program above, family~"B" extends~"A", overriding lemma~"f" by
-proving it using lemma~"g".
+"B" overrides lemma~"f" by proving it using~"g".
 Lemma~"g" is in turn inherited from family "A", where it is proven using
@@
 Within family "STLCFix", a second partial recursor
-(called "tm_prect_STLCFix") and its computational behaviors are axiomatized,
-so properties such as \lsti[basicstyle=\fontsize{8.25}{9},columns=fixed,basewidth=0.9ex]{tm_fix x t1 = tm_fix x t2 ->  t1 = t2} can be proved.
+(called "tm_prect_STLCFix") and its computational behaviors are automatically axiomatized,
+which allows properties of \lsti[basewidth=0.9ex]{tm_fix}, such as
+\lsti[columns=fixed,basewidth=1.2ex]{tm_fix x t1 = tm_fix x t2 ->  t1 = t2}, to be proved.
 

### 2022-11-10 coqimpl.tex
address two comments

@@
 Consider the inductive type \lsti{tm}.
-In \lsti{STLC}, it is translated to a top-level module type
+In \cref{fig:stlc-compiled}, it is translated to a top-level module type
 \lsti{STLC\_tm} that declares a \lsti{tm} type,
-its constructors (e.g., \lsti{tm_abs}) \EDJ{I am not sure if it is better we emphasize in the following way :`` that declares a tm type, its constructors (e.g.
+four functions standing for the constructors,
+%\EDJ{I am not sure if it is better we emphasize in the following way :`` that declares a tm type, its constructors (e.g.
 tmabs) as a normal function,..
@@
 emphasize it is a normal function in the module type},
-its partial recursor (\lsti{tm_prect_STLC}),
-and the propositional equalities\EDJ{I propose to change this sentence to ``and the prop eq about computational behaviour on the partial recursor''.
+a partial recursor (\lsti{tm_prect_STLC}),
+and the computational behaviors of the partial recursor
+(e.g., \lsti{tm_abs_eq_STLC}). %\cref{sec:prec}.
+%\EDJ{I propose to change this sentence to ``and the prop eq about computational behaviour on the partial recursor''.
 And thus in the other parts of the paper, we can use the word ``computational behaviour'' instead of prop eq.
-Because just saying prop eq seems a bit of lack of context/details.} on the partial recursor
-(e.g., \lsti{tm_abs_eq_STLC}; \cref{sec:prec}).
+Because just saying prop eq seems a bit of lack of context/details.}
 
@@
 suffices for the translations of the subsequent fields to be type-checked by Coq.
-Leaving the definitions open enables \lsti{STLC} and \lsti{STLCFix}
-to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and \lsti{End}~\lsti{STLCFix}.
+Leaving the definitions unspecified enables \lsti{STLC} and \lsti{STLCFix}
+to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and upon \lsti{End}~\lsti{STLCFix}.
 

### 2022-11-10 metatheory-condensed.tex
reply a comment
Reply the reply.

@@
 In particular, when no hiding is needed, $\seal$ is $\var{0}$ (the identity)\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right?
-Or you meant to say it is identity ``sealing''?} and $A$ is
+Or you meant to say it is identity ``sealing''?}\YZreply{Yes, I meant identity sealing.
+I cannot say 'sealing' because the paper currently does not use the word sealing anywhere.}\EDJreply{What about just remove the parenthesis and ``the identity''} and $A$ is
 $\TyTkg{\lsig}$.

### 2022-11-10 coqimpl.tex
changes of wording: equality -> comp behavior

@@
 via command \lsti!Include STLC\_tm(self)!,
-and additionally declares
-the new constructor \lsti{tm_fix} and related equalities.
+and additionally declares \lsti{tm_fix}, a partial recursor, and related equalities.
 
@@
 and then a module type declaring the existence of the recursive function as well as
-its propositional equalities.\EDJ{Simiarly, here I think using ``computational behaviour'' is more accurate than using the word ``prop eq''}
+its computational behaviors.
 
@@
 Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
-and the existence of its propositional equalities, leaving \lsti{subst} undefined
+and the equalities about its computational behaviors, leaving \lsti{subst} undefined
 and the equalities unproven.
 %
-The propositional equalities are stated in terms of the case handlers,
+The equalities are stated in terms of the case handlers,
 whose definitions \emph{are} available through the \lsti{self} parameter.
-So Coq can simplify, for example, \lsti{subst_tm_unit_eq}'s type to
-\lsti!\forall x t, self.subst self.tm_unit x t = self.tm_unit!.
-The equalities will be included and available for use in the
-translations of the subsequent fields through their \lsti{self}
-parameters.
+So Coq can simplify, for example, the type of \lsti{subst_tm_unit_eq} to
+\lsti[basicstyle=\fontsize{8.5}{9},basewidth=.95ex]{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
+These equalities about the computational bahaviors of \lsti{subst}
+will be included and available for use in the translations of the
+subsequent fields through their \lsti{self} parameters.
 
@@
 The translation of \lsti{FInduction} is similar, except that there is no need to
-register propositional equalities, as \lsti{FInduction} proofs are considered opaque.
+register computational behaviors, as \lsti{FInduction} proofs are considered opaque.
 

### 2022-11-10 stlc-compiled.tex
edits
changes of wording: equality -> comp behavior

@@
 | tm_abs : id -> tm -> tm | tm_app : tm -> tm -> tm.
-[@\codecomment{Instantiate \texttt{tm} partial recursor \& its equalities}@]
+[@\codecomment{Instantiate \texttt{tm} partial recursor \& its computational behaviors}@]
 Def tm_prect_STLC := \lambda P, tm_rect (\lambda[@\,@]t, option (P t)).
@@
 Include STLC\_subst\_Cases.[@\tanc{c-STLC-d-subst-start}@]
-[@\codecomment{Instantiate \texttt{subst} \& its equalities}@]
+[@\codecomment{Instantiate \texttt{subst} \& its computational behaviors}@]
 Def subst[@\,@] := tm_rect _ subst\_tm_unit
@@
 Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}@][@\innersep@]
-[@\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@] [@\dadada@][@\tanc{c-STLC-d-ty}@][@\tanc{c-STLC-sndcol-start}@][@\innersep@]
+[@\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor, etc.}@] [@\dadada@][@\tanc{c-STLC-d-ty}@][@\tanc{c-STLC-sndcol-start}@][@\innersep@]
 Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLC-d-env}@][@\innersep@]
@@
 
-\vspace{-12pt}
+\vspace{-14pt}
 

### 2022-11-10 stlcfix-compiled.tex
edits
changes of wording: equality -> comp behavior

@@
 | tm_fix : id -> tm -> tm.
-[@\codecomment{Instantiate \texttt{tm} partial recursors \& their equalities}@] [@\dadada@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
+[@\codecomment{Instantiate \texttt{tm} partial recursors \& their comp.\ behaviors}@] [@\dadada@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
 Include STLCFix\_subst\_Cases. [@\tanc{c-STLCFix-d-subst-start}@]
-[@\codecomment{Instantiate \texttt{subst} \& its equalities}@]
+[@\codecomment{Instantiate \texttt{subst} \& its computational behaviors}@]
 Def subst := tm_rect _ subst\_tm_unit
@@
 Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]
-[@\codecomment{\textls[-10]{Include\;\texttt{ty}, its constructors, partial recursor \& eq.}}@] [@\dadada@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
+[@\codecomment{\textls[-10]{Include\;\texttt{ty}, its constructors, partial recursor, etc.}}@] [@\dadada@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
 Include STLC\_env.[@\hfill\codecomment{reuse}@][@\tanc{c-STLCFix-d-env}@][@\innersep@]
@@
 
-\vspace{-8pt}
+\vspace{-3pt}
 
@@
 
-\caption{Compilation of family \lsti!STLCFix! (\cref{fig:stlc-mechanized}).}
+\caption{Compilation of family \lsti!STLCFix! and the final \lsti!Check! command (\cref{fig:stlc-mechanized}).}
 \label{fig:stlcfix-compiled}

### 2022-11-10 coqimpl.tex
more changes of wording: equality -> comp behavior

@@
 so its definition is not exposed to the fields that come after it.
-Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
+Accordingly, the translation \lsti{STLC\_subst} merely declares the types of \lsti{subst}
 and the equalities about its computational behaviors, leaving \lsti{subst} undefined
@@
 on the fields.
-Opaque fields can be further bound, too (\cref{sec:override});
-they are translated to modules that export opaque fields.
+Opaque fields in \Lang can be further bound (\cref{sec:override});
+they are translated to Coq modules that export opaque fields.
 
@@
 with the help of \lsti{tm_rect}, the recursor Coq generates for \lsti{tm}.
-Proving the equalities on \lsti{tm_prect_STLC} is then trivial, by \lsti{reflexivity}.
+Computational behaviors of \lsti{tm_prect_STLC} are immediate, by \lsti{reflexivity}.
 
@@
 to the (already included) case handlers. %(\lsti{subst\_tm_unit}, etc.).
-Propositional equalities on \lsti{subst} are then trivial, by \lsti{reflexivity}.
+Computational behaviors of \lsti{subst} are then immediate, by \lsti{reflexivity}.
 

### 2022-11-10 agda-files.tex
Update information about agda files. Also I cannot see appendices anymore.

@@
-We make four agda source code files (publicly and anonymously) openly avaliable, and they include the corresponding QIIT syntax definition and models.
+We provide four agda source code files in the ``other supplemental material'', and they include the corresponding QIIT syntax definition and models.
 
-For syntax, please check \\ \href{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}{https://drive.google.com/file/d/1aoG67rmXzP_x1MvZCIN3do0sucyqjwkn/view?usp=sharing}.
+For syntax, please check "syntax.agda" in the ``other supplemental material''.
 This is done by mainly following the formulation \citet{altkap2016}.
@@
 
-For syntactic translation, please check \\ \href{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}{https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing}.
+For syntactic translation, please check "SyntaxTranslation.agda" in the ``other supplemental material'' .
 The syntactic translation here is to translate most of the novel feature introduced in this paper into a standard MLTT.
@@
 
-For consistency model, please check \\ \href{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}{https://drive.google.com/file/d/1pNhnn125P5byAHDaSIlpbxMvr1F-9FRo/view?usp=sharing}.
-This is following the standard model from \citet{altkap2016,kaposi2017type,kaposi2019gluing}.
+For consistency model, please check "semantic.agda" in the ``other supplemental material''. This is following the standard model from \citet{altkap2016,kaposi2017type,kaposi2019gluing}.
 
-For canonicity model, please check \\ \href{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}{https://drive.google.com/file/d/1R6C7QNfyu8fbl6LE2ruvpqZ_ZDSRVo0c/view?usp=sharing}.
+For canonicity model, please check "canonicity.agda" in the ``other supplemental material''.
 This is mainly following \citet{coquand2018canonicity,sterling2019algebraic}, but we carried out our argument in a type-theortic framework.

### 2022-11-10 metatheory-condensed.tex
address comments

@@
 quantified over the earlier ones, a linkage component is \emph{universally}
-quantified over---and thus polymorphic to---components preceding it.
+quantified over---and thus polymorphic to---the components preceding it.
 
@@
 (i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
-Crucially, unlike the second field in a dependent pair (\ruleref{tm/pair}),
-the typing of~$t$ in \ruleref{l/add} is polymorphic to how the fields in
-$t$'s context are defined, so that a different linkage that overrides
-fields in $t$'s context can reuse~$t$---without retyping it---by
-projecting it and then including it\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.
+Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
+the premise $\goodTerm{\Gm, A}{t}{T}$ makes clear that
+the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified over
+how the fields in $t$'s context are defined.
+So a different linkage~$\ell'$ that overrides fields in $t$'s context (and thus models a derived family)
+can reuse~$t$---without retyping it---by first
+projecting~$t$ from $\LkgAdd{\lkg}{t}$ and then appending it to~$\lkg'$.
+%\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.
 
 This type~$A$ abstracting the context in \ruleref{l/add} and
-\ruleref{lsig/add} is not necessarily the same as those recorded by~$\lsig$\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
+\ruleref{lsig/add} do not necessarily contain the same types of the prior fields as recorded by~$\TyLkg{\lsig}$,
+%\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
 because a field representing the code $\wcode{\wsig}$ of a W-type has to
@@
 should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
-and \ruleref{tm/wrec} stipulate, whereas $\wsig$ should be hidden
-from all other fields so that they can be reused
-for a W-type signature~$\wsig'$ extending~$\wsig$\EDJ{Suggest changing to ``so that they can be reused
-in a different context where W-type signature~$\wsig'$ is extended to~$\wsig$''}.
+and \ruleref{tm/wrec} stipulate.
+By contrast, $\wsig$ should be hidden
+from all other fields, so that they can be reused
+in a different context where the W-type signature~$\wsig$ is replaced by an extended one~$\wsig'$.
 

### 2022-11-11 cc-condensed.tex
resolve comments

--- 2022-11-09 cc-condensed.tex
@@
 every proposition is trivially provable.
+
 \cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
@@
 \citet{altkap2016}, extending it to handle linkages.
-Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products,
+A linkage~$\LkgAdd{\lkg}{t}$ can be interpreted into a non-dependent pair where the
+second component is a dependent function (modeling late binding),
+%\EDJ{Suggesting changing to ``can be interpreted into non-dependent pair with dependent function for late-binding''}
 as rule \ruleref{l/add} indicates.
 %
-We refer the reader to a supplemental file (in Agda syntax) for the complete proof.
+We refer the reader to a supplemental file in Agda syntax for the complete proof.
 
@@
 type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
-When canonicity is proved in a constructive meta\-logic, its proof
-amounts to a normalization function for closed terms of the ground type\EDJ{Suggesting use "boolean type".
+When the canonicity theorem is proved in a constructive meta\-logic, its proof
+amounts to a normalization function for closed terms of the ground type.
+%\EDJ{Suggesting use "boolean type".
 Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat).
 We also does specify what is ground type.
-}.
+}\YZreply{I think it is clear from the context and from the article 'the' that 'the ground type' refers to B.}
 So canonicity serves to justify the computational nature of the type theory.
@@
 into a logical predicate on closed terms: the predicate includes all
-closed, ``reducible''\EDJ{I will add a it is my bad to introduce "reducible" without defining it first.
-But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)} terms of type~$T$.
+closed, ``reducible'' terms of type~$T$.
+%\EDJ{I will add a quote around reducible here.
+it is my bad to introduce "reducible" without defining it first.
+But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}\EDJreply{Because in the operational semantic, using ``reducible'' is totally fine.
+But in declarative style, only ``canonicity'' this word is making sense.
+But this ``canonicity'' is not friendly to the audience not familar with declartive style.}
 The proof is available as a supplemental file in Agda syntax.

### 2022-11-11 metatheory-condensed.tex
resolve comments

@@
 is not automated in \TT.
-In particular, when no hiding is needed, $\seal$ is $\var{0}$ (the identity)\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right?
+In particular, when no hiding is needed, $\seal$ is $\var{0}$
+%\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right?
 Or you meant to say it is identity ``sealing''?}\YZreply{Yes, I meant identity sealing.
-I cannot say 'sealing' because the paper currently does not use the word sealing anywhere.}\EDJreply{What about just remove the parenthesis and ``the identity''} and $A$ is
-$\TyTkg{\lsig}$.
-In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are identities,
+I cannot say 'sealing' because the paper currently does not use the word sealing anywhere.}\EDJreply{What about just remove the parenthesis and ``the identity''}
+and $A$ is $\TyTkg{\lsig}$.
+In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are $\var{0}$,
 and $\seal_6$ is from

### 2022-11-11 related-work.tex
resolve comments

@@
 In addition, extra programmer effort may be required, such as
-having to manually prove additional well-formedness conditions on the algebras\EDJ{What is this `algebra'?
+having to manually prove additional well-formedness conditions.
+%on the algebras\EDJ{What is this `algebra'?
 I mean if it is a very frequent word to the audience  familiar with DTC then removing explanation is fine.}.
@@
 %
-But even with the tool support, components of individually developed
-feature extensions (e.g., "substlem") have to be composed separately by
-invoking the tool.
+But even with the tool support, components (e.g., "substlem") of
+individually developed feature extensions have to be composed separately
+by invoking the tool.
 %
@@
 our solution can be viewed as an effort to preempt refactoring by
-allowing the programmer to write code that has built-in hooks for
+enabling the programmer to write code that has built-in hooks for
 future extensions.

### 2022-11-11 conclusion.tex
fix broken refs
assorted changes

@@
 using case studies.
-A novel dependent type theory formalizes the core aspects of the language mechanism
+A novel dependent type theory formalizes the essence of the language mechanism
 and is shown to enjoy consistency and canonicity.
-Generalizing the mechanism to support nested family polymorphism and
-scaling it to sizable proof engineering efforts make exciting directions for
-future work.
+Generalizing the mechanism to support nested family polymorphism
+will make extensibility scale to sizable proof engineering efforts and
+thus make exciting directions for future work.

### 2022-11-11 stlc-linkTran-typing.tex
assorted changes

@@
 In this appendix, we sketch how to use the ``library'' of linkage
-transformers in \TT to inductively construct a linkage that models a
+transformers in \TT to inductively construct a linkage transformer that models a
 derived family of \lsti{STLC}.
@@
 
+Most steps are self-explanatory.
 Of note is the two grayed rows. They are constructing a linkage transformer~$h_\beta$
@@
 surface-syntax program & $i$      & $\goodInh{\cdot}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
-\lsti!Family STLCBool!   & 0        & $\InhId$                                             \\ \cline{2-3} 
+\lsti!Family STLCBool extends STLC.!   & 0        & $\InhId$                                             \\ \cline{2-3} 
 \lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_{\texttt{tm}}'}}$                      \\ \cline{2-3} 

### 2022-11-11 agda-files.tex
Update Instruction in Appendix.

@@
 
-We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.
+We suggest using editors with proper syntax highlighting to read these source files.
+e.g.
+VSCode with \textit{language-agda} plugin.\footnote{\url{https://marketplace.visualstudio.com/items?itemName=j-mueller.agda}}

### 2022-11-11 metatheory-condensed.tex
trend

@@
 %
-As is conventional with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
+As is the current trend with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
 substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part

### 2022-11-11 agda-files.tex
itemize agda files

--- 2022-11-10 agda-files.tex
@@
+In the ``Other supplemental material'' section of the submission, we
+include the complete formalization of \TT.
+It consists of four files in Agda syntax.
+%
+
+\begin{description}[
+    itemsep=.5ex,topsep=.6ex,
+    labelsep=1.5ex,
+    leftmargin=6.5ex,
+%   labelindent=3.5ex,
+    font=\ttfamily\bfseries,
+]
+
+\item[Syntax.agda]
+
+This file contains the syntactic typing rules of \TT.
+We use the style of intrinsically typed syntax \cite{chapman2009type},
+which allows only well typed \TT types and terms to be constructed.
+The use of intrinsically typed syntax puts our formalization on a more
+rigorous footing.
+
+\item[Semantic.agda]
+
+This file defines the standard model from which
+\cref{thm:consistency} (consistency) follows.
+
+\item[Canonicity.agda]
+
+This file defines the logical-relations model from which
+\cref{thm:canonicity} (canonicity) follows.
+
+\item[SynTranslation.agda]
+
+We also provide a third model that serves as a translation that ``compiles
+away'' the novel features---most notably, linkages---introduced by \TT.
+As is similar to the first two models, 
+a linkage is translated to a non-dependent pair where the
+second component is a dependent function (modeling late binding).
+
+\end{description}
+
+All the definitions contained in these files are stated in
+terms of a meta\-language supporting quotient inductive-inductive types
+(QIITs)~\cite{nordvall2010inductive,altkap2016}.
+QIITs support equality constructors, which facilitate the expression of
+conversion rules (e.g., \ruleref{tmeq/sub/id}).
+%
+We note that the current formalization is not checked by Agda because
+Agda does not support QIITs natively.
+An effort to mechanize the formalization is underway, using Agda's
+\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
+
+\ifShowOldWriting
+
+\bigskip
+
 We provide four agda source code files in the ``other supplemental material'', and they include the corresponding QIIT syntax definition and models.
@@
 This is the first QIIT-model we give.
-A QIIT-model can be roughly consider as an arbitrary mapping from the QIIT data(our syntax in this case), also respecting the quotient(judgemental equality in this case).
+A QIIT-model can be roughly consider as an arbitrary mapping from the QIIT data (our syntax in this case), also respecting the quotient (judgmental equality in this case).
 Here, the syntactic translation model is mapping to (a subpart of) the original QIIT data (itself).
@@
 We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.
+
+\fi

### 2022-11-11 agda-files.tex

@@
 
+We suggest reading these files in an editor with syntax highlighting for Agda
+(e.g., \mbox{VS~Code} with the \textsf{language-agda}\footnote{\url{https://marketplace.visualstudio.com/items?itemName=j-mueller.agda}}
+extension enabled).
+
 \ifShowOldWriting
@@
 This is the first QIIT-model we give.
-A QIIT-model can be roughly consider as an arbitrary mapping from the QIIT data (our syntax in this case), also respecting the quotient (judgmental equality in this case).
+A QIIT-model can be roughly consider as an arbitrary mapping from the QIIT data(our syntax in this case), also respecting the quotient(judgemental equality in this case).
 Here, the syntactic translation model is mapping to (a subpart of) the original QIIT data (itself).
@@
 
-We suggest using editors with proper syntax highlighting, e.g. VSCode with agda-code plugins, to read these source files.
+We suggest using editors with proper syntax highlighting to read these source files.
+e.g.
+VSCode with \textit{language-agda} plugin.\footnote{\url{https://marketplace.visualstudio.com/items?itemName=j-mueller.agda}}
 

### 2022-11-11 abstract.tex
small edits
ref to appendices

@@
 %
-This paper presents a language design geared towards extensible metatheory
+This paper presents a novel language design geared towards extensible metatheory
 mechanization in a proof assistant.
@@
 \item a prototypical implementation of the language design as a Coq plugin,
-\item a formalization of the language design as a constructive type theory with inductive types,
-\item consistency and canonicity of the type theory,
+\item a dependent type theory capturing the essence of the language mechanism
+      and its consistency and canonicity results,
 %\EDJ{We don't yet have normalization though}

### 2022-11-11 metatheory-condensed.tex
itemize agda files
ref to appendices

@@
 \noindent
-A supplemental appendix sketches how the other introduction forms of linkage transformers
+\ifreport \cref{sec:lt-in-action}
+\else A supplemental appendix
+\fi
+sketches how the other introduction forms of linkage transformers
 can be used to model the construction of a derived family as a linkage transformer.
@@
 natively.
-An effort to mechanize the formalization is ongoing, using Agda's
+An effort to mechanize the formalization is underway, using Agda's
 \lsti{REWRITE} pragma to replace the limited uses of equality reflection.
@@
 We also typeset a more complete set of rules than the space allows and
-present it in a supplemental appendix.
+present it in
+\ifreport \cref{sec:more-rules}.
+\else a supplemental appendix.
+\fi
 

### 2022-11-11 metatheory-condensed.tex
Justify trend by ref. Fix ref for Coquand.

@@
 %
-As is the current trend with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
+As is the current trend~\cite{gratzer-sterling-birkedal-2019,altkap2016,coquand2018canonicity} with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
 substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part

### 2022-11-11 metatheory-condensed.tex
edits in FMLTT sec

--- 2022-11-11 metatheory-condensed.tex
@@
 responsible for late binding.
-Here, $A$ abstracts the fields preceding the current field~$t$
-(i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
+Here, $A$ abstracts the context of the current field~$t$, controlling
+how the types of the fields prior to~$t$ are exposed to the typing of~$t$.
 Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
 the premise $\goodTerm{\Gm, A}{t}{T}$ makes clear that
-the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified over
-how the fields in $t$'s context are defined.
-So a different linkage~$\ell'$ that overrides fields in $t$'s context (and thus models a derived family)
+the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
+rather than \emph{existentially} quantified as is in \ruleref{tm/pair}---%
+over how the fields in $t$'s context are defined.
+Late binding enables reuse.
+A different linkage~$\ell'$ that overrides fields in $t$'s context (and thus models a derived family)
 can reuse~$t$---without retyping it---by first
@@
 
-This type~$A$ abstracting the context in \ruleref{l/add} and
-\ruleref{lsig/add} do not necessarily contain the same types of the prior fields as recorded by~$\TyLkg{\lsig}$,
+%recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$
+The type~$A$ in \ruleref{l/add} and \ruleref{lsig/add},
+abstracting the types of the prior fields, does not necessarily contain the same types as those recorded by~$\TyLkg{\lsig}$,
 %\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
-because a field representing the code $\wcode{\wsig}$ of a W-type has to
+because a field defining the code $\wcode{\wsig}$ of a W-type has to
 expose different types to different fields that come after it.
@@
 
-We illustrate this with \cref{fig:stlc-linkage-typing}.
+We use \cref{fig:stlc-linkage-typing} to illustrate.
 On the left is code excerpted from \cref{fig:stlc-mechanized}, and on the right

### 2022-11-11 stlc-linkage-typing.tex
columnize
edits in FMLTT sec

@@
 %
-\begin{tabular}{@{}l@{\ \ \ }L@{}}
+\begin{tabular}{@{}l@{\ \ \ }L@{\quad}L@{\quad}L@{\quad}L@{}}
 \cellcolor{Bkg-color}
@@
 \lsig_0 \coloneqq \LSigEmp
-\quad
+&
 \lkg_0 \coloneqq \LkgEmp
-\quad
-\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
+&
+\goodTerm{\cdot}{\lkg_0}{\TyLkg{\lsig_0}}
+&
 \\
@@
 }
-\quad
+&
 \lkg_1 \coloneqq \LkgAdd{\lkg_0}{
@@
 }
-\quad
-\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
-\quad
+&
+\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
+&
 \goodTerm{\TyTkg{\lsig_0}}{\seal_1}{\sub{
@@
 &
+\multicolumn{4}{@{}l@{}}{
 \goodTerm{
@@
 }
+}
 \\
@@
 }
-\quad
+&
 \lkg_2 \coloneqq \LkgAdd{\lkg_1}{
@@
 }
-\quad
+&
 \goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
-\quad
+&
 \goodTerm{\TyTkg{\lsig_1}}{\seal_2}{\sub{
@@
 &
+\multicolumn{4}{@{}l@{}}{
 \textit{\color{codecomment-color}%
@@
     context (i.e., $A_6$) of the case handler below:%
+}
 }
@@
 }
-\quad
+&
 \lkg_6 \coloneqq \LkgAdd{\lkg_5}{
@@
 }
-\quad
+&
 \goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
-\quad
+&
 \goodTerm{\TyTkg{\lsig_5}}{\seal_6}{\sub{
@@
 }
-\quad
+&
 \lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
@@
 }
-\quad
+&
 \goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
-\quad
+&
 \goodTerm{\TyTkg{\lsig_9}}{\seal_{10}}{\sub{
@@
 
-\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.}
+\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.
+Each row $\goodTerm{\mathit{self}\!:\!
+\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i-cap}(-0.02,-0.10)(-0.02,0.28)
+A_i%
+\tikzmarkend{A-i-cap}%
+}{%
+\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i-cap}(-0.02,-0.10)(-0.02,0.28)%
+t_i%
+\tikzmarkend{t-i-cap}%
+}{%
+\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i-cap}(-0.02,-0.10)(-0.02,0.28)%
+T_i%
+\tikzmarkend{T-i-cap}%
+}$ types a field.
+The type $A_i$ controls how the typing of field $t_i$ sees the types of the fields prior to $t_i$.
+}
 \label{fig:stlc-linkage-typing}

### 2022-11-11 metatheory-condensed.tex
back on 20 pages

--- 2022-11-11 metatheory-condensed.tex
@@
 %
-As is the current trend with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
+As is the current trend~\cite{coquand2018canonicity,altkap2016,gratzer-sterling-birkedal-2019}
+with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
 substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
@@
 %
-Variables are represented by de Bruijn indices:
+Variables are in the form of de~Bruijn indices:
 \var{n} is the variable bound by the $n$-th closest enclosing binder.
@@
 %
-Typing judgments for substitutions have the form 
+Substitutions are typed with the form 
 %\EDJ{It seems to me G |- g : D is a judgement not a rule.
@@
 The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
-yields terms valid in~$\Gm$ (e.g., rule \ruleref{tm/sub}).
+yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
 %

### 2022-11-11 metatheory-condensed.tex
de bruijn
tweak

@@
 %
-Variables are in the form of de~Bruijn indices:
+Variables are in the form of \citeauthor{debruijn1964} indices:
 \var{n} is the variable bound by the $n$-th closest enclosing binder.
@@
 %\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
-because a field defining the code $\wcode{\wsig}$ of a W-type has to
+because a field defined as the code $\wcode{\wsig}$ of a W-type has to
 expose different types to different fields that come after it.

### 2022-11-11 casestudies.tex
resolve comments
tweaks

--- 2022-11-09 casestudies.tex
@@
 However,
-we believe generalizing our design to mutually inductive types does
+we believe accommodating mutually inductive types does
 not pose theoretical challenges.
@@
 In addition to the metatheories, this case study produces abstract
-interpreters that are directly ready for program extraction in Coq.
+interpreters that are directly ready for program extraction.
 
@@
 
-\begin{wrapfigure}[13]{r}{0.13\textwidth}
+\begin{wrapfigure}[13]{r}{0.138\textwidth}
 \includegraphics[scale=.68]{graphics/ai-casestudy.pdf}
@@
 decision procedures for them.
+%\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}
 

### 2022-11-11 metatheory-condensed.tex
tweaks

@@
 By contrast, $\wsig_{\texttt{tm}}$ is hidden from the typing of all other fields.
-Their typing depends on the knowledge that \lsti!tm! has type~$\cU$,
+Their typing should depend on the knowledge that \lsti!tm! has type~$\cU$,
 rather than \TyS{\wcode{\wsig_{\texttt{tm}}}},
-so they can be reused in a context where \lsti!tm! is defined as $\wcode{\wsig'_{\texttt{tm}}}$,
+so that they can be reused in a context where
+\lsti!tm! is defined as $\wcode{\wsig'_{\texttt{tm}}}$,
 where $\wsig'_{\texttt{tm}}$ extends~$\wsig_{\texttt{tm}}$ with additional constructors.

### 2022-11-11 stlc-compiled.tex
minor

@@
 [@\outdent@]End STLC\_tm.[@\tanc{c-STLC-tm-end}@][@\outersep@]
-[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in family \texttt{STLC}}@][@\tanc{c-STLC-subst-start}@]
+[@\outdent@][@\codecomment{Code emitted for definition of \texttt{subst} in family \texttt{STLC}}@][@\tanc{c-STLC-subst-start}@]
 [@\outdent@]Module Type STLC\_subst\_Cases\_Ctx.
@@
 [@\outdent@]Module [@\textls[-30]{STLC}@]\_[@\textls[-30]{subst}@]\_[@\textls[-30]{Cases}@] (self[@\,@]:[@\,@][@\textls[-30]{STLC}@]\_[@\textls[-30]{subst}@]\_[@\textls[-30]{Cases}@]\_[@\textls[-30]{Ctx}@]).
-Def subst\_tm_unit :=
+Def subst\_tm_unit :=[@\codecomment{\textls[-20]{emmited upon definition of case}}@]
   \lambda (x : id) (t : self.tm), self.tm_unit.
@@
 Def tm_prect_STLC := \lambda P, tm_rect (\lambda[@\,@]t, option (P t)).
-Fact [@\textls[-20]{tm\_unit\_eq\_STLC}@] : [@\dadada@]. [@\textls[-40]{{\color{dkblue}reflexivity}.}@] Qed.
+Fact [@\textls[-0]{tm\_unit\_eq\_STLC}@] : [@\dadada@]. reflexivity. Qed.
 Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
@@
   subst\_tm_var subst\_tm_abs subst\_tm_app.
-Fact [@\textls[-40]{subst\_tm\_unit\_eq}@] : [@\dadada@]. [@\textls[-40]{{\color{dkblue}reflexivity}.}@] Qed.
+Fact [@\textls[-0]{subst\_tm\_unit\_eq}@] : [@\dadada@]. reflexivity. Qed.
 Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}@][@\innersep@]

### 2022-11-11 agda-files.tex
add refs
small edits

@@
 In the ``Other supplemental material'' section of the submission, we
-include the complete formalization of \TT.
+include the definitive, complete formalization of \TT.
 It consists of four files in Agda syntax.
@@
 \cref{thm:consistency} (consistency) follows.
+%
+The construction largely follows the standard model of
+\cite{altkap2016,kaposi2017type,kaposi2019gluing}.
 
@@
 \cref{thm:canonicity} (canonicity) follows.
+%
+The construction largely follows that in
+\cite{coquand2018canonicity,sterling2019algebraic}.
 

### 2022-11-11 agda-files.tex
Rephrasing

@@
 We note that the current formalization is not checked by Agda because
-Agda does not support QIITs natively.
+Agda does not support QIITs natively and Agda does not support equality reflection.
 An effort to mechanize the formalization is underway, using Agda's
-\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
+\lsti{REWRITE} pragma to replace the limited uses of equality reflection, 
+and using axioms to postulate QIIT type following \cite{altkap2016}.
 

### 2022-11-11 agda-files.tex
Reorder.

@@
 We note that the current formalization is not checked by Agda because
-Agda does not support QIITs natively and Agda does not support equality reflection.
+Agda does not support equality reflection and Agda does not support QIITs natively.
 An effort to mechanize the formalization is underway, using Agda's

### 2023-02-07 pldi2023/review-reply-densed.md
Drafting the response.

We thank all the reviewers for their critical and encouraging feedback.
Below we first address the main concerns. We might paraphrase

* (Paragraph 3/4 from Reviewer D)**The presentation of the calculus seems to be geared so that it closely mirrors an Agda formalization, in lieu of a more conventional presentation (e.g. substitutions are part of the syntax, and the calculus uses De Bruijn indices). Presumably due to space concerns, the formalization also lacks sufficient examples, which makes it hard to understand how things fit together: As one example, the description of the TM/WSUP rule does not mention the role of the signature $\tau$ in the rule, leaving the reader to figure out its purpose. It would be helpful to provide an concrete example of how this rule is used. I appreciate the effort to ground things via Figure 8's mapping from a few definitions of the STLC family to the core calculus. Unfortunately, this listing is effectively a partial definition, in that it contain several intermediate terms ($l_x$) whose definitions are not given. At the end of the day, I did not emerge from this section with a good intuition of how 'linkages' work and how they are used. In a similar vein, the proof of the key metatheoretic properties assumes that the reader is familiar with similar proofs, stating that the key technical device (an interpretation into a metalanguage) is largely similar to e.g. Altenkirch and Kaposi, without clearly explaining the extension. More worrisome, the full proof is *only* given in a notation heavy, Agda-style format, which is both unchecked by a machine and hard for humans to read.  While the formalization tries to rebut these concerns by acknowledging that the presentation will be dense to those 'without prior exposure to MLTT', but it would be more accurate to stipulate 'non-experts in MLTT'.**

<!-- I am not sure if I want to reply this or how I would like to respond this. 
      The only part I want to reply is 
        "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format, which is both unchecked by a machine and hard for humans to read."
      But my reply will certainly upgrade to a notation war that leads to nowhere.

    The real reply I want to give is: For example, For normalization of MLTT, given Abel's thesis (written in conventional style) and Kaposi's thesis (written in fake Agda style), I will believe the latter proof even in pen-and-paper form. This is due to the personal taste and numerous error I have personally made when I use latex/human language to lay out any mathematical object. So I simply believe fake-Agda is better for Human development (at least for myself).

    Apart from the above sentence, I think I can make no reply but really just do some modification suggestion on the paper
 -->

* (Paragraph 6) Reviewer D gives an example that will lead to awkwardness in our paper, because it require `coe` or equality transpose between propositionally equal types (due to the loss of judgemental(definitional) equality).

<!-- He got a point. To avoid transpose we need overridablity/pins which is not properly covered in the paper. -->

* (Paragraph 9) Reviewer D suggests the inclusion of `fdestruct` tactic

<!-- Generally speaking this is either not possible or trivial.
      For example, for overridable/pins it is trivial,
                  for non-overdiable stuff, it is not possible because the inductive type is extensible
    But I don't know If I want to tell the reviewer at all  -->

* **What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching? Is it rejected? Can it be defined, but only works for STLC.tm? In the latter case, how does the plugin prevent other extensible definitions from referring to it?** (Review D)

* **"In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?** (Review D)

```C
// For example
Family A {Field a; Field b; Field c;}

Family B extends A {
  // Inherit a. // this command line will make a difference once uncommented. Because the fields are order sensitive. In the case when a2's definition is dependent on a, omitting the inherit won't type check a2 
  // Thus this ``annotation`` require the programmer's effort. 
  Field a2.
}
```

* (Reviewer C) **The utility of the FMLTT formalization in this paper is unclear to me. Don't you already have consistency and canonicity via the translation to Coq demonstrated by FPOP? Or put another way, instead of defining FMLTT directly, couldn't you translate it to MLTT following the approach in FPOP?**

The technical detail here is that, FMLTT also has this unconventional Wtype formulation, which is closer to the inductive type in the surface syntax for Coq and Agda.

Our translation doesn't translate the unconventional Wtype formulation back to the conventional Wtype one. So the core calculus after translation is MLTT + this unconventional Wtype.

We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.

* **looking at figure 2 this extension involves adding cases to the term language and adjusting the proofs, but not adding e.g. new things to the typing judgment.  Imagine we extend STLC to System F, using two type contexts, one for term variables, the other for type variables. Can such an extension be defined?**

* (Reviewer B)  **Would/does your approach also work in the presence of other approaches to binding, e.g. de Bruijn or locally nameless?**

* (Reviewer B) **looking at figure 2 this extension involves adding cases to the term language and adjusting the proofs, but not adding e.g. new things to the typing judgment.  Imagine we extend STLC to System F, using two type contexts, one for term variables, the other for type variables. Can such an extension be defined?** (Reviewer B also asked a question on extending the feature of reference)

To extend typing judgement is actually adding new indices for an existent inductive type, which is not covered in our paper. This not only happens for polymorphism extension, but also for the case of extending with the heap -- since reduction relation will have a new information on heap.

This is a quite challenging problem we are facing right now as well. The main challenge here for us is how to achieve code/proof reuse after this novel extension.

* (Reviewer B) **the fake-Agda formalization of translation is at any rate not a direct proof of correctness of the plugin implementation.**

We consider the formalized translation as an oversimplified fundation of the plugin implementation, also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.

* (Reviewer B) **Where does the A in rules like TyEq/PK/Add come from?**

At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
In practice, the plugin implementation always choose a "default" one. We didn't formalize the "default" one as it is usually just making inductive type into an abstracted form and make sure other parts stay the "same" (I.e. family {Nat : W(O, S); pred : Nat -> Nat} is abstracted into {Nat : U; O : Nat; S : Nat -> Nat; pred : Nat -> Nat}).

* (Reviewer A/B) **In the conclusion of rule Tyeq/Casety, R should be T?** 
Yes. Sorry for the confusing typo.

* (Reviewer A) **In terms of ergonomics, I wonder how powerful and usable is the FInduction construct is. In Coq proofs, it's not rare to use induction tactics inside a larger proof. Will users need to outline those into their explicit own lemmas?**

Currently at plugin level, we only support layout the subcases in a lemma forms. 
  I think the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself. 
  A. The former idea needs either 
      (1) extend the coq's (proof assistant's) kernel (i.e. make the pattern matching facility extensible) or 
      (2) maniuplate the AST of the proof term
  B. the later idea

  B breaks the expectation of incremental compilation a bit. 
  A.(1) and A.(2) both require sophisticated engineering effort. A(1) will also risk sabtaging the kernel itself and expanding the trusted base.
  Modulo the interesting semantic on the pattern matching itself (i.e. if we always consider the pattern matching can be elaborated to the usage of eliminator), I believe the current prototype is exploring the same semantic. So we didn't choose to explore this direction but this is definitely more practical.

* (Reviewer A) **Why is W(t) a term and not a type? What are bold-W and bold-P needed for?**

bold-W is a typo. We forgot to delete it.

ℙ is a technical device of the system mentioned on Line 788. Its functionality is to transform a linkage(overridable/extensible family) into a module(sigma type). The reason we need it is because, proper abstraction `A` cannot happen on linkage (𝕃) but only on a module (ℙ). (i.e. contrary to Line 788, the `Γ, 𝕃(𝜎) ⊢ s : A[p¹]` is usually unprovable for nontrivial `A`)

<!-- Do I want to expand it? -->

### 2023-02-07 pldi2023/review-reply-densed.md
Typo fix.
Adding Response.

@@
 We thank all the reviewers for their critical and encouraging feedback.
-Below we first address the main concerns. We might paraphrase
+Below we first address the main concerns, which might be paraphrased for the space.
 
-* (Paragraph 3/4 from Reviewer D)**The presentation of the calculus seems to be geared so that it closely mirrors an Agda formalization, in lieu of a more conventional presentation (e.g.
-substitutions are part of the syntax, and the calculus uses De Bruijn indices).
-Presumably due to space concerns, the formalization also lacks sufficient examples, which makes it hard to understand how things fit together: As one example, the description of the TM/WSUP rule does not mention the role of the signature $\tau$ in the rule, leaving the reader to figure out its purpose.
-It would be helpful to provide an concrete example of how this rule is used.
-I appreciate the effort to ground things via Figure 8's mapping from a few definitions of the STLC family to the core calculus.
-Unfortunately, this listing is effectively a partial definition, in that it contain several intermediate terms ($l_x$) whose definitions are not given.
-At the end of the day, I did not emerge from this section with a good intuition of how 'linkages' work and how they are used.
-In a similar vein, the proof of the key metatheoretic properties assumes that the reader is familiar with similar proofs, stating that the key technical device (an interpretation into a metalanguage) is largely similar to e.g.
-Altenkirch and Kaposi, without clearly explaining the extension.
-More worrisome, the full proof is *only* given in a notation heavy, Agda-style format, which is both unchecked by a machine and hard for humans to read.
-While the formalization tries to rebut these concerns by acknowledging that the presentation will be dense to those 'without prior exposure to MLTT', but it would be more accurate to stipulate 'non-experts in MLTT'.**
+* (Reviewer A) "In terms of ergonomics, I wonder how powerful and usable" this is ...
+; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."
 
-<!-- I am not sure if I want to reply this or how I would like to respond this. 
-      The only part I want to reply is 
-        "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format, which is both unchecked by a machine and hard for humans to read."
-      But my reply will certainly upgrade to a notation war that leads to nowhere.
+<!-- I am not sure if the following respond is appropriate -->
 
-    The real reply I want to give is: For example, For normalization of MLTT, given Abel's thesis (written in conventional style) and Kaposi's thesis (written in fake Agda style), I will believe the latter proof even in pen-and-paper form.
-This is due to the personal taste and numerous error I have personally made when I use latex/human language to lay out any mathematical object.
-So I simply believe fake-Agda is better for Human development (at least for myself).
+As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level.
+We only have `fdiscriminate` and `finjection` for intermediate fact.
+We acknowledge this limitation of our current research and are trying to advance more in this future direction.
 
-    Apart from the above sentence, I think I can make no reply but really just do some modification suggestion on the paper
+We believe the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself.
+A. The former idea needs either 
+    (1) extend the Coq's kernel (i.e. make the Coq's pattern matching facility extensible) or 
+    (2) maniuplate the AST of the proof term
+B. the latter will manipulate the AST of the proof script
+
+B will break incremental compilation "a bit more" (i.e.
+we cannot have incremental compilation for each recursor).
+We believe the promising future lies in the direction of A(1), by enhance proof assistants with native extensible pattern matching instead of this current encoding.
+As Reviewer D mentioned, there is much work to be done.
+
+<!-- 
+// I am not sure if I want to include the following
+
+Reviewer D points out the current lack of "fdestruct". In fact, we believe "fdestruct" doesn't exist. 
+Both "finjection" and "fdiscriminate" works on arbitrary extensible inductive type but "fdestruct" can only work on extensible inductive type with fixed number of constructors -- that is a contradiction because all the extensions right now are about extending with new constructors.
+
+One way to have destruct tactic usable is to use overridable/pins.
+This is not completely covered in Section 3.3, but this is used when we want overridable term to be transparent and thus an inductive type and "destruct"-able
+
  -->
 
-* (Paragraph 6) Reviewer D gives an example that will lead to awkwardness in our paper, because it require `coe` or equality transpose between propositionally equal types (due to the loss of judgemental(definitional) equality).
+* "Seems related to type theories for modules, but doesn't discuss the connection"
 
-<!-- He got a point. To avoid transpose we need overridablity/pins which is not properly covered in the paper. -->
+Finally, I note that a family has similarities with an ML-style module,
+including the fact that some of its members are opaque while others are
+abstract. This is modelled through the use of singleton types, just as
+dependent type theories for modules do (e.g., Dreyer, Crary, Harper, POPL
+2003), which are also extensions of MLTT. Furthermore, there has been work on
+extending ML modules with mixin composition (Dreyer, Rossberg, ICFP 2008).
+ML-style modules are the basis of Coq's module system as well, so I would have
+expected some comparison with this line of work.
 
-* (Paragraph 9) Reviewer D suggests the inclusion of `fdestruct` tactic
+  > We need to say something about "A type system for higher-order modules"
 
-<!-- Generally speaking this is either not possible or trivial.
-      For example, for overridable/pins it is trivial,
-                  for non-overdiable stuff, it is not possible because the inductive type is extensible
-    But I don't know If I want to tell the reviewer at all  -->
+* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition
 
-* **What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching?
-Is it rejected?
-Can it be defined, but only works for STLC.tm?
-In the latter case, how does the plugin prevent other extensible definitions from referring to it?** (Review D)
+Thanks for this feedback! We will definitely reorganize the whole structure for Section 5 to achieve better acessibility.
 
-* **"In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?** (Review D)
+### Review B
 
-```C
-// For example
-Family A {Field a; Field b; Field c;}
+* (Reviewer B)"One thing I don't really get from the rules/discussion (including the
+  appendix) is the exact mechanism for choosing the A in rules like TyEq/PK/Add
+  and LSig/Add where A is supposed to abstract out the possibly-extensible types
+  in deciding what to make available to later definitions. Is A always going to
+  be uniquely determinable from s? (or if not, does the nondeterminism matter?)
+  Is it up to the user (in this case, the plugin) to get this right, or do the
+  rules somehow make sure that the right things are abstracted in a way I can't
+  see?"
 
-Family B extends A {
-  // Inherit a.
-// this command line will make a difference once uncommented.
-Because the fields are order sensitive.
-In the case when a2's definition is dependent on a, omitting the inherit won't type check a2
-  // Thus this ``annotation`` require the programmer's effort. 
-  Field a2.
-}
-```
+At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
+In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type opauqe (Line 791) and make sure other parts stay the "same" 
+(I.e.
+s maps {Nat : W(O, S); pred : Nat -> Nat} into {Nat : U; O : Nat; S : Nat -> Nat; pred : Nat -> Nat}).
+It is possible to design an interface where the programmer can choose which `A` they would love to abstract to, but we haven't found any use case for this functionality yet.
 
-* (Reviewer C) **The utility of the FMLTT formalization in this paper is unclear to me.
-Don't you already have consistency and canonicity via the translation to Coq demonstrated by FPOP?
-Or put another way, instead of defining FMLTT directly, couldn't you translate it to MLTT following the approach in FPOP?**
+* (Reviewer B/D) "I find it somewhat surprising that neither the FPOP plugin / implementation
+  itself, nor the code of the case studies, is provided."
+
+We will definitely make the FPOP plugin/implementation available.
+
+* (Review B) "I think the paper would also be improved by delineating the limitations of the
+  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"
+
+Thanks for the feedback!
+We will definitely explicitize the limitation of the current implementation and meta-theory in the final revision.
+Especially the problem concretized and emphasized by Reviewer D involving equality coercion (the pervasive propositional equality on extensible inductive family lead to unnecessary equality coercion e.g.
+when `denoteTy t` is not definitionally equal to `nat`).
+
+We believe this can lead to exciting future work on both theory and practice -- a normaliation algorithm for an extended calculus admitting more property on extensible inductive type; and a proper proof assistant acknowledging native extensible inductive type instead of our current encoding by the plugin.
+
+And yes, in the final revision we will explicitize the trusted codebase as our plugin only translate the code into the vanilla Coq term without expanding the trusted base.
+
+### Review C
+
+* "The utility of the FMLTT formalization in this paper is unclear to me. Don't
+  you already have consistency and canonicity via the translation to Coq
+  demonstrated by FPOP? Or put another way, instead of defining FMLTT directly,
+  couldn't you translate it to MLTT following the approach in FPOP?"
 
@@
 
-Our translation doesn't translate the unconventional Wtype formulation back to the conventional Wtype one. So the core calculus after translation is MLTT + this unconventional Wtype.
+Our translation doesn't translate this unconventional Wtype formulation back to the conventional one.
+So the core calculus after translation is MLTT + this unconventional Wtype.
+In that case, translation function as a guide to the plugin implementation, and pedantically is not enough to show the consistency/canonicitiy for FMLTT.
 
-We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.
+<!-- We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.
+-->
+### Review D
+* (Reviewer D) 
+* > This expert reviewer's main concerns are (1) the incredible terseness of the theory
+  presentation and (2) a few current limitations (framed as future research challenges).
+  I don't think there are misunderstandings, so the strategy is to acknowledge them
+  and promise to somehow address them in the final revision.
 
-* **looking at figure 2 this extension involves adding cases to the term language and adjusting the proofs, but not adding e.g.
-new things to the typing judgment.
-Imagine we extend STLC to System F, using two type contexts, one for term variables, the other for type variables.
-Can such an extension be defined?**
+<!-- I think 
+  terseness is already covered above
+  challenge:
+  1. the equality coerciion thing, covered above
+  2. the limitedness of the current tactics, covered in the first question
 
-* (Reviewer B)  **Would/does your approach also work in the presence of other approaches to binding, e.g. de Bruijn or locally nameless?**
+  // TBH I don't know what are other tactics people usually use can be added here... But it is future work. I think this question
 
-* (Reviewer B) **looking at figure 2 this extension involves adding cases to the term language and adjusting the proofs, but not adding e.g.
-new things to the typing judgment.
-Imagine we extend STLC to System F, using two type contexts, one for term variables, the other for type variables.
-Can such an extension be defined?** (Reviewer B also asked a question on extending the feature of reference)
+  // so I think there is no big questions from Reviewer D anymore.
 
-To extend typing judgement is actually adding new indices for an existent inductive type, which is not covered in our paper.
-This not only happens for polymorphism extension, but also for the case of extending with the heap -- since reduction relation will have a new information on heap.
+-->
 
-This is a quite challenging problem we are facing right now as well. The main challenge here for us is how to achieve code/proof reuse after this novel extension.
+* Reviewer D reviewer also has some minor questions. Optionally respond to them.
 
-* (Reviewer B) **the fake-Agda formalization of translation is at any rate not a direct proof of correctness of the plugin implementation.**
-
-We consider the formalized translation as an oversimplified fundation of the plugin implementation, also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.
-
-* (Reviewer B) **Where does the A in rules like TyEq/PK/Add come from?**
-
-At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
-In practice, the plugin implementation always choose a "default" one.
-We didn't formalize the "default" one as it is usually just making inductive type into an abstracted form and make sure other parts stay the "same" (I.e.
-family {Nat : W(O, S); pred : Nat -> Nat} is abstracted into {Nat : U; O : Nat; S : Nat -> Nat; pred : Nat -> Nat}).
-
-* (Reviewer A/B) **In the conclusion of rule Tyeq/Casety, R should be T?** 
-Yes. Sorry for the confusing typo.
-
-* (Reviewer A) **In terms of ergonomics, I wonder how powerful and usable is the FInduction construct is.
-In Coq proofs, it's not rare to use induction tactics inside a larger proof.
-Will users need to outline those into their explicit own lemmas?**
-
-Currently at plugin level, we only support layout the subcases in a lemma forms. 
-  I think the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself.
-  A. The former idea needs either 
-      (1) extend the coq's (proof assistant's) kernel (i.e. make the pattern matching facility extensible) or 
-      (2) maniuplate the AST of the proof term
-  B. the later idea
-
-  B breaks the expectation of incremental compilation a bit. 
-  A.(1) and A.(2) both require sophisticated engineering effort. A(1) will also risk sabtaging the kernel itself and expanding the trusted base.
-  Modulo the interesting semantic on the pattern matching itself (i.e.
-if we always consider the pattern matching can be elaborated to the usage of eliminator), I believe the current prototype is exploring the same semantic.
-So we didn't choose to explore this direction but this is definitely more practical.
-
-* (Reviewer A) **Why is W(t) a term and not a type? What are bold-W and bold-P needed for?**
-
-bold-W is a typo. We forgot to delete it.
-
-ℙ is a technical device of the system mentioned on Line 788.
-Its functionality is to transform a linkage(overridable/extensible family) into a module(sigma type).
-The reason we need it is because, proper abstraction `A` cannot happen on linkage (𝕃) but only on a module (ℙ).
-(i.e.
-contrary to Line 788, the `Γ, 𝕃(𝜎) ⊢ s : A[p¹]` is usually unprovable for nontrivial `A`)
-
-<!-- Do I want to expand it? -->
+* Reviewer A asks five questions in the "comments to authors" section;
+  optionally answer them towards the end of the response letter.

### 2023-02-07 pldi2023/review-reply-densed.md
Fix typo.

@@
 
-We believe this can lead to exciting future work on both theory and practice -- a normaliation algorithm for an extended calculus admitting more property on extensible inductive type; and a proper proof assistant acknowledging native extensible inductive type instead of our current encoding by the plugin.
+We believe this can lead to exciting future work on both theory and practice -- a normaliation algorithm for an extended calculus admitting more convertibility on elimination of the extensible inductive type; and a proper proof assistant acknowledging native extensible inductive type (instead of our current encoding by the plugin), and thus support the reduction of open terms properly.
 

### 2023-02-07 pldi2023/review-reply-densed.md
Save comments.

--- 2023-02-07 pldi2023/review-reply-densed.md
@@
 
+<!-- 
+we hypothesize that nested induction can be lifted to the top level and then it will generate extra proof obligation balabla.
+
+requires significant engineering effort
+
+ -->
+
 As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level.
@@
 
+<!-- 
+Doesn't seem very relevant for our paper rebuttal
+
 We believe the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself.
@@
     (1) extend the Coq's kernel (i.e. make the Coq's pattern matching facility extensible) or 
-    (2) maniuplate the AST of the proof term
+    (2) maniuplate the AST of the proof term -> this will need to be shown at proof-script level
 B. the latter will manipulate the AST of the proof script
@@
 As Reviewer D mentioned, there is much work to be done.
+-->
+
+<!-- Basically the generated proof term, we look at each pattern matching point
+      and trying to give further proof obligation
+    // concretize this idea
+ -->
 
 <!--
+
+// the reviewer is actually asking nested pattern matching stuff here
 // I am not sure if I want to include the following
@@
 Reviewer D points out the current lack of "fdestruct". In fact, we believe "fdestruct" doesn't exist.
+
 Both "finjection" and "fdiscriminate" works on arbitrary extensible inductive type but "fdestruct" can only work on extensible inductive type with fixed number of constructors -- that is a contradiction because all the extensions right now are about extending with new constructors.
@@
 
-Thanks for this feedback! We will definitely reorganize the whole structure for Section 5 to achieve better acessibility.
+<!-- I love you chatgpt -->
+
+Thank you for your feedback. We will promptly reorganize the structure of Section 5 to improve accessibility
+<!-- sympethsize, dense ,hard to read, not giving too much
+
+    we will expand  it in the main text to clarify the question the reviewers have, and also use appendices to explain the  formalization  in greater details 
+         -->
 
@@
 At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
-In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type opauqe (Line 791) and make sure other parts stay the "same" 
+In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type opauqe (Line 791) and make sure other parts stay the "same".
+
+<!-- Use example on Pg 16. Concrete run it -->
 (I.e.
@@
 
-We will definitely make the FPOP plugin/implementation available.
+We will certainly make the FPOP plugin/implementation available.
 
@@
 
-Thanks for the feedback!
-We will definitely explicitize the limitation of the current implementation and meta-theory in the final revision.
-Especially the problem concretized and emphasized by Reviewer D involving equality coercion (the pervasive propositional equality on extensible inductive family lead to unnecessary equality coercion e.g.
-when `denoteTy t` is not definitionally equal to `nat`).
+Thank you for your feedback. We will clearly state the limitations of the current implementation and meta-theory in the final revision. In particular,
 
-We believe this can lead to exciting future work on both theory and practice -- a normaliation algorithm for an extended calculus admitting more property on extensible inductive type; and a proper proof assistant acknowledging native extensible inductive type instead of our current encoding by the plugin.
+1.
+we will emphasize the issue raised by Reviewer D regarding equality coercion and the pervasive propositional equality on extensible inductive families.
+This can result in unnecessary equality coercion when `denoteTy t` is not definitionally equal to `nat`.
 
-And yes, in the final revision we will explicitize the trusted codebase as our plugin only translate the code into the vanilla Coq term without expanding the trusted base.
+We believe that this opens up new opportunities for future work in both theory and practice,
+(a). such as a normalization algorithm for an extended calculus that allows for more convertibility upon elimination of the extensible inductive type; 
+(b).
+and a proper proof assistant that supports native extensible inductive types instead of the current encoding via our plugin, thus allowing for proper reduction of open terms.
+This will change the kernel of proof assistant for sure.
+
+In the final revision, we will also clarify that our plugin only translates code into vanilla Coq terms without expanding the trusted codebase.
 
@@
 
-The technical detail here is that, FMLTT also has this unconventional Wtype formulation, which is closer to the inductive type in the surface syntax for Coq and Agda.
+The technical detail here is that, FMLTT also has this unconventional Wtype formulation 
+<!-- explain how unconventional it is
+      what conventional about unconventional Wtype
+ -->
+, which is closer to the inductive type in the surface syntax for Coq and Agda.
 
@@
 In that case, translation function as a guide to the plugin implementation, and pedantically is not enough to show the consistency/canonicitiy for FMLTT.
+
+<!-- make it clear that we have MLTT + unconventional Wtype as target calculus -->
+
+<!-- acknowledge the reviewer's  idea of translating unconventional Wtype into conventional Wtype -->
 
@@
   optionally answer them towards the end of the response letter.
+
+* Reviewer B ask some questions as well
+
+* (Reviewer B) contradictory feature like polymorphism + heap causing unsoundness
+<!-- we can expect mixin will generate proof obligation that is not pluasible be proved
+
+one fixpoint extension, the other termination, definitely will fail on the reducibility argument.
+ -->

### 2023-02-07 pldi2023/review-reply-densed.md
Save.

@@
 
-<!-- I am not sure if the following respond is appropriate -->
+As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level.
+We only have `fdiscriminate` and `finjection` for deducing intermediate fact.
+The key problem is the lack of certain nested pattern matching.
 
-<!-- 
-we hypothesize that nested induction can be lifted to the top level and then it will generate extra proof obligation balabla.
-
-requires significant engineering effort
-
+<!-- The idea is 
+    fill in the new hole once pattern matching needs to extend the clauses
+    lifting nested induction to top level is not necessary
  -->
 
-As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level.
-We only have `fdiscriminate` and `finjection` for intermediate fact.
-We acknowledge this limitation of our current research and are trying to advance more in this future direction.
-
-<!-- 
-Doesn't seem very relevant for our paper rebuttal
-
-We believe the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself.
-A. The former idea needs either 
-    (1) extend the Coq's kernel (i.e. make the Coq's pattern matching facility extensible) or 
-    (2) maniuplate the AST of the proof term -> this will need to be shown at proof-script level
-B. the latter will manipulate the AST of the proof script
-
-B will break incremental compilation "a bit more" (i.e.
-we cannot have incremental compilation for each recursor).
-We believe the promising future lies in the direction of A(1), by enhance proof assistants with native extensible pattern matching instead of this current encoding.
-As Reviewer D mentioned, there is much work to be done.
--->
-
-<!-- Basically the generated proof term, we look at each pattern matching point
-      and trying to give further proof obligation
-    // concretize this idea
- -->
-
-<!--
-
-// the reviewer is actually asking nested pattern matching stuff here
-// I am not sure if I want to include the following
-
-Reviewer D points out the current lack of "fdestruct". In fact, we believe "fdestruct" doesn't exist.
-
-Both "finjection" and "fdiscriminate" works on arbitrary extensible inductive type but "fdestruct" can only work on extensible inductive type with fixed number of constructors -- that is a contradiction because all the extensions right now are about extending with new constructors.
-
-One way to have destruct tactic usable is to use overridable/pins.
-This is not completely covered in Section 3.3, but this is used when we want overridable term to be transparent and thus an inductive type and "destruct"-able
-
- -->
+We acknowledge this limitation of our current research and are trying to advance more in this direction.
+We hypothesize that code involving nested induction/pattern matching an be reused by fill in the hole after extension upon inductive type.
+The plugin will generate extra proof obligation to fill this hole.
+This certainly requires significant amount of engineering effort.
 
@@
 
-<!-- I love you chatgpt -->
+Thank you for your feedback. 
+We sympathize with the general impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the general understanding of the whole paper.
 
-Thank you for your feedback. We will promptly reorganize the structure of Section 5 to improve accessibility
-<!-- sympethsize, dense ,hard to read, not giving too much
-
-    we will expand  it in the main text to clarify the question the reviewers have, and also use appendices to explain the  formalization  in greater details 
-         -->
+We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization in greater details.
 
@@
 -->
-### Review D
-* (Reviewer D) 
-* > This expert reviewer's main concerns are (1) the incredible terseness of the theory
-  presentation and (2) a few current limitations (framed as future research challenges).
-  I don't think there are misunderstandings, so the strategy is to acknowledge them
-  and promise to somehow address them in the final revision.
-
-<!-- I think 
-  terseness is already covered above
-  challenge:
-  1. the equality coerciion thing, covered above
-  2. the limitedness of the current tactics, covered in the first question
-
-  // TBH I don't know what are other tactics people usually use can be added here... But it is future work. I think this question
-
-  // so I think there is no big questions from Reviewer D anymore.
-
--->
-
-* Reviewer D reviewer also has some minor questions. Optionally respond to them.
+### Other questions
 
@@
 
-* (Reviewer B) contradictory feature like polymorphism + heap causing unsoundness
-<!-- we can expect mixin will generate proof obligation that is not pluasible be proved
+* (Reviewer B) How will our plugin react when trying to mixin the contradictory features?
+For example, since STLC with polymorphism and  STLC + references both enjoying principal typing under HM type inference, but their composition doesn't.
 
-one fixpoint extension, the other termination, definitely will fail on the reducibility argument.
- -->
+We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.
+
+Another exmaple would be extend *STLC and its termination proof* with general recursion feature. Our plugin will generate unprovable proof obligation for the reducibility argument.
+
+ * Reviewer D reviewer also has some minor questions. Optionally respond to them.

### 2023-02-07 pldi2023/review-reply-densed.md
Save.

@@
 Thank you for your feedback. 
-We sympathize with the general impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the general understanding of the whole paper.
+We sympathize with this impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the readers.
 
@@
 
-Thank you for your feedback. We will clearly state the limitations of the current implementation and meta-theory in the final revision. In particular,
-
-1.
-we will emphasize the issue raised by Reviewer D regarding equality coercion and the pervasive propositional equality on extensible inductive families.
+Thank you for your feedback.
+We will clearly state the limitations of the current implementation and meta-theory in the final revision.
+In particular, we will emphasize the issue raised by Reviewer D regarding equality coercion and the pervasive propositional equality on extensible inductive families.
 This can result in unnecessary equality coercion when `denoteTy t` is not definitionally equal to `nat`.
@@
 * (Reviewer B) How will our plugin react when trying to mixin the contradictory features?
-For example, since STLC with polymorphism and  STLC + references both enjoying principal typing under HM type inference, but their composition doesn't.
+For example, since STLC with (DHM-)polymorphism and  STLC + references both enjoying type soundness, but their composition doesn't.
 
@@
 
-Another exmaple would be extend *STLC and its termination proof* with general recursion feature. Our plugin will generate unprovable proof obligation for the reducibility argument.
+Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature.
+Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.
 

### 2023-02-07 pldi2023/review-reply-densed.md
Save.

@@
 At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
-In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type opauqe (Line 791) and make sure other parts stay the "same".
+In practice, this A is decided by implementation.
+Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".
 
-<!-- Use example on Pg 16. Concrete run it -->
-(I.e.
-s maps {Nat : W(O, S); pred : Nat -> Nat} into {Nat : U; O : Nat; S : Nat -> Nat; pred : Nat -> Nat}).
-It is possible to design an interface where the programmer can choose which `A` they would love to abstract to, but we haven't found any use case for this functionality yet.
+Taking (Line 745) Figure.
+8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A` has `tm : 𝕌` instead.
+`s` will make sure other fields stay the same type.
+To show this explicitly at the plugin level, we look at Figure 4 (Line 494) `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding to `tm : 𝕌`).
+With this interface `STLC°tm` we cannot pattern match any term of type `tm : 𝕌` thus doing abstraction successfully.
+
+Generally speaking, all the (extensible) inductive type will be simply "wrapped" by a module type with its interface and only exposing the constructor (with no eliminators), just like how we generate `STLC°tm`.
 
@@
 * (Reviewer B) How will our plugin react when trying to mixin the contradictory features?
-For example, since STLC with (DHM-)polymorphism and  STLC + references both enjoying type soundness, but their composition doesn't.
+For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.
 

### 2023-02-07 pldi2023/review-reply-densed.md
Save.

@@
 
-The technical detail here is that, FMLTT also has this unconventional Wtype formulation 
-<!-- explain how unconventional it is
-      what conventional about unconventional Wtype
- -->
-, which is closer to the inductive type in the surface syntax for Coq and Agda.
+The technical detail here is that, FMLTT also has this unconventional Wtype formulation.
+While each classic Wtype is defined by a pair of type `A ⊢ B`, our unconventional Wtype is defined by a list of pair of types `Aᵢ ⊢ Bᵢ`, each pair corresponds to one constructor.
 
-Our translation doesn't translate this unconventional Wtype formulation back to the conventional one.
-So the core calculus after translation is MLTT + this unconventional Wtype.
-In that case, translation function as a guide to the plugin implementation, and pedantically is not enough to show the consistency/canonicitiy for FMLTT.
+One way to interpret it is the classical formluation encoding this list using a big sum type. So the conventional and unconventional Wtype can be translated to each other.
 
-<!-- make it clear that we have MLTT + unconventional Wtype as target calculus -->
+However, our translation doesn't translate this unconventional Wtype formulation back to the conventional one.
+So the target calculus after translation is **MLTT + this unconventional Wtype**.
+In that case, translation can only function as a guide to the plugin implementation, and **pedantically** is not enough to show the consistency/canonicitiy for FMLTT because **pedantically** target calculus is not proven to be consistency/canonicity.
 
-<!-- acknowledge the reviewer's  idea of translating unconventional Wtype into conventional Wtype -->
+We agree with Reviewer B's insight -- we will get consistency/canonicitiy when we translate unconventional Wtype into the conventional one.
+The reason we didn't choose to do so because we expect this translation a lot more verbose than the current proof because of the simplicity of Wtype itself compared to the rich functionality provided by (fake-)Agda's Inductive Facility.
+We only use the latter when constructing consistency/canonicity model.
 
-<!-- We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.
--->
 ### Other questions

### 2023-02-07 pldi2023/review-reply-densed.md
Save. Workin on optional and the module question

@@
 
-* Reviewer A asks five questions in the "comments to authors" section;
-  optionally answer them towards the end of the response letter.
+* (Reviewer A) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?
 
-* Reviewer B ask some questions as well
+In dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.
+
+𝕎(t) is a typo and should be deleted. Thanks for pointing out.
+
+ℙ is another technical detail of the system.
+Its functionality is to transform a linkage(overridable/extensible) into a module(sigma type).
+The reason it is irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but only on a module (ℙ).
+For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ : A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial `A₆`.
 
@@
 
- * Reviewer D reviewer also has some minor questions. Optionally respond to them.
+* (Reviewer D) **What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching?
+Is it rejected?
+In the latter case, how does the plugin prevent other extensible definitions from referring to it?**
+
+Coq will reject on the pattern matching on `tm`.
+For example, Figure 4, Line 494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the given inductive `tm`.
+The following field definition will be type-checked based on this interface, and for Coq it will only consider `tm` as an arbitrary type and will fail to pattern match it.
+
+<!-- I am not sure if I want to mention overridable/pins  -->

### 2023-02-07 pldi2023/review-reply-densed.md
Add comments for mentioning the proof.

@@
 
-* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition
+* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition.
+Reviewer D: "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"
 
@@
 
-We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization in greater details.
+We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater details aiming for readability.
 
@@
 
-* (Reviewer B)"One thing I don't really get from the rules/discussion (including the
+* *(Reviewer B)"One thing I don't really get from the rules/discussion (including the
   appendix) is the exact mechanism for choosing the A in rules like TyEq/PK/Add
@@
   rules somehow make sure that the right things are abstracted in a way I can't
-  see?"
+  see?"*
 
-At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
+  The typing rules do not commit to a particular choice of $A$, as the reviewer
+  points out.
   In practice, this A is decided by implementation.
@@
 
-Taking (Line 745) Figure.
-8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A` has `tm : 𝕌` instead.
-`s` will make sure other fields stay the same type.
-To show this explicitly at the plugin level, we look at Figure 4 (Line 494) `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding to `tm : 𝕌`).
-With this interface `STLC°tm` we cannot pattern match any term of type `tm : 𝕌` thus doing abstraction successfully.
+  Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A`
+  has `tm : 𝕌` instead. `s` will make sure other fields stay the same type. To
+  show this explicitly at the plugin level, we look at Figure 4 (Line 494)
+  `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding
+  to `tm : 𝕌`). With this interface `STLC°tm` we cannot pattern match any term
+  of type `tm : 𝕌` thus doing abstraction successfully.
 
-Generally speaking, all the (extensible) inductive type will be simply "wrapped" by a module type with its interface and only exposing the constructor (with no eliminators), just like how we generate `STLC°tm`.
+  Generally speaking, all the (extensible) inductive type will be simply
+  "wrapped" by a module type with its interface and only exposing the
+  constructor (with no eliminators), just like how we generate `STLC°tm`.
 
-* (Reviewer B/D) "I find it somewhat surprising that neither the FPOP plugin / implementation
-  itself, nor the code of the case studies, is provided."
+* *(Reviewer B/D) "I find it somewhat surprising that neither the FPOP plugin / implementation
+  itself, nor the code of the case studies, is provided."*
 
@@
 
-* (Review B) "I think the paper would also be improved by delineating the limitations of the
-  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"
+* *(Review B) "I think the paper would also be improved by delineating the limitations of the
+  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"*
 
-Thank you for your feedback.
-We will clearly state the limitations of the current implementation and meta-theory in the final revision.
-In particular, we will emphasize the issue raised by Reviewer D regarding equality coercion and the pervasive propositional equality on extensible inductive families.
-This can result in unnecessary equality coercion when `denoteTy t` is not definitionally equal to `nat`.
+  Thank you for your feedback. We will clearly state the limitations of the
+  current implementation and meta-theory in the final revision. In particular,
+  we will emphasize the issue raised by Reviewer D regarding equality coercion
+  and the pervasive propositional equality on extensible inductive families.
+  This can result in unnecessary equality coercion when `denoteTy t` is not
+  definitionally equal to `nat`.
 
-We believe that this opens up new opportunities for future work in both theory and practice,
-(a). such as a normalization algorithm for an extended calculus that allows for more convertibility upon elimination of the extensible inductive type; 
-(b).
-and a proper proof assistant that supports native extensible inductive types instead of the current encoding via our plugin, thus allowing for proper reduction of open terms.
-This will change the kernel of proof assistant for sure.
+  We believe that this opens up new opportunities for future work in both theory
+  and practice, (a). such as a normalization algorithm for an extended calculus
+  that allows for more convertibility upon elimination of the extensible
+  inductive type; (b). and a proper proof assistant that supports native
+  extensible inductive types instead of the current encoding via our plugin,
+  thus allowing for proper reduction of open terms. This will change the kernel
+  of proof assistant for sure.
 
-In the final revision, we will also clarify that our plugin only translates code into vanilla Coq terms without expanding the trusted codebase.
+  In the final revision, we will also clarify that our plugin only translates
+  code into vanilla Coq terms without expanding the trusted codebase.
 
@@
 
-* "The utility of the FMLTT formalization in this paper is unclear to me. Don't
+* *"The utility of the FMLTT formalization in this paper is unclear to me. Don't
   you already have consistency and canonicity via the translation to Coq
   demonstrated by FPOP? Or put another way, instead of defining FMLTT directly,
-  couldn't you translate it to MLTT following the approach in FPOP?"
+  couldn't you translate it to MLTT following the approach in FPOP?"*
 
-The technical detail here is that, FMLTT also has this unconventional Wtype formulation.
-While each classic Wtype is defined by a pair of type `A ⊢ B`, our unconventional Wtype is defined by a list of pair of types `Aᵢ ⊢ Bᵢ`, each pair corresponds to one constructor.
+  The technical detail here is that, FMLTT also has this unconventional Wtype
+  formulation. While each classic Wtype is defined by a pair of type `A ⊢ B`,
+  our unconventional Wtype is defined by a list of pair of types `Aᵢ ⊢ Bᵢ`, each
+  pair corresponds to one constructor.
 
-One way to interpret it is the classical formluation encoding this list using a big sum type. So the conventional and unconventional Wtype can be translated to each other.
+  One way to interpret it is the classical formluation encoding this list using
+  a big sum type. So the conventional and unconventional Wtype can be translated
+  to each other.
 
-However, our translation doesn't translate this unconventional Wtype formulation back to the conventional one.
-So the target calculus after translation is **MLTT + this unconventional Wtype**.
-In that case, translation can only function as a guide to the plugin implementation, and **pedantically** is not enough to show the consistency/canonicitiy for FMLTT because **pedantically** target calculus is not proven to be consistency/canonicity.
+  However, our translation doesn't translate this unconventional Wtype
+  formulation back to the conventional one. So the target calculus after
+  translation is **MLTT + this unconventional Wtype**. In that case, translation
+  can only function as a guide to the plugin implementation, and
+  **pedantically** is not enough to show the consistency/canonicitiy for FMLTT
+  because **pedantically** target calculus is not proven to be
+  consistency/canonicity.
 
-We agree with Reviewer B's insight -- we will get consistency/canonicitiy when we translate unconventional Wtype into the conventional one.
-The reason we didn't choose to do so because we expect this translation a lot more verbose than the current proof because of the simplicity of Wtype itself compared to the rich functionality provided by (fake-)Agda's Inductive Facility.
-We only use the latter when constructing consistency/canonicity model.
+  We agree with Reviewer B's insight -- we will get consistency/canonicitiy when
+  we translate unconventional Wtype into the conventional one. The reason we
+  didn't choose to do so because we expect this translation a lot more verbose
+  than the current proof because of the simplicity of Wtype itself compared to
+  the rich functionality provided by (fake-)Agda's Inductive Facility. We only
+  use the latter when constructing consistency/canonicity model.
 
@@
 
-* (Reviewer A) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?
+* *(Reviewer A) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?*
 
@@
 
-ℙ is another technical detail of the system.
-Its functionality is to transform a linkage(overridable/extensible) into a module(sigma type).
-The reason it is irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but only on a module (ℙ).
-For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ : A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial `A₆`.
+  ℙ is another technical detail of the system. Its functionality is to transform
+  a linkage(overridable/extensible) into a module(sigma type). The reason it is
+  irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but
+  only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
+  A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial
+  `A₆`.
 
-* (Reviewer B) How will our plugin react when trying to mixin the contradictory features?
-For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.
+* *(Reviewer B) How will our plugin react when trying to mixin the contradictory features?
+For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.*
 
-We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.
+  We can expect our plugin will generate unprovable proof obligation under
+  mixin, hindering qed'ed the proposition and thus closing the family.
 
-Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature.
-Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.
+  Another exmaple would be extend *a family of STLC and its termination proof*
+  with general recursion feature. Our plugin will generate unprovable proof
+  obligation inside the reducibility argument for the fixpoint feature.
 
@@
 
-Coq will reject on the pattern matching on `tm`.
-For example, Figure 4, Line 494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the given inductive `tm`.
-The following field definition will be type-checked based on this interface, and for Coq it will only consider `tm` as an arbitrary type and will fail to pattern match it.
+  Coq will reject on the pattern matching on `tm`. For example, Figure 4, Line
+  494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the
+  given inductive `tm`. The following field definition will be type-checked
+  based on this interface, and for Coq it will only consider `tm` as an
+  arbitrary type and will fail to pattern match it.
 

### 2023-02-08 pldi2023/review-reply-densed.md
indent

@@
 
-* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition.
-Reviewer D: "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"
+* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition
 
@@
 
-We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater details aiming for readability.
+We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization in greater details.
 

### 2023-02-08 coqimpl.tex
fix typos and awkward wordings

@@
 \lsti[basicstyle=\fontsize{8.5}{9},basewidth=.95ex]{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
-These equalities about the computational bahaviors of \lsti{subst}
+These equalities about the computational behaviors of \lsti{subst}
 will be included and available for use in the translations of the

### 2023-02-08 introduction.tex
fix typos and awkward wordings

@@
 
-\noindentparagraph{Contributions}
+\noindentparagraph{Contributions.}
 
@@
 Thus, our contributions include novel design recipes for dealing with
-these challenges and strong meta\-theoretical guarantees on the
+these challenges and foundational meta\-theoretical guarantees on the
 underlying logic.

### 2023-02-08 metatheory-condensed.tex
fix typos and awkward wordings

@@
 %
-For readability, we feel free to use a notation with explicit names
+For readability, we use a notation with explicit names
 in this typing judgment:
 we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
-and feel free to write its type~$A_i$, which is a dependent tuple type, as a
+and write its type~$A_i$, which is a dependent tuple type, as a
 dependent record type with labels.

### 2023-02-08 related-work.tex
add related work
fix typos and awkward wordings

@@
 \cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
+%\cite{vdr2022intrinsically}
 The original DTC encoding requires type-level general recursion that
@@
 In particular, because data types have to be encoded (rather than expressed
-through natively supported inductive types), the resulting code is
-less readable and the resulting programming style inaccessible to non-experts.
+through natively supported inductive types), the resulting code can be
+obtuse at first blush, making the programming style inaccessible to non-experts.
 In addition, extra programmer effort may be required, such as

### 2023-02-08 pldi2023/review-reply-densed.md
Still need to point out extending index is not possible currently.

--- 2023-02-07 pldi2023/review-reply-densed.md
@@
 
-* *(Reviewer A) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?*
+* (Reviewer A/B) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?
 
@@
 
-  We can expect our plugin will generate unprovable proof obligation under
-  mixin, hindering qed'ed the proposition and thus closing the family.
+<!-- I think we still need to clarify we doesn't support extending these -->
+We currently doesn't support extend STLC with polymorphism and reference, which requires extending existent inductive family with new indices.
 
-  Another exmaple would be extend *a family of STLC and its termination proof*
-  with general recursion feature. Our plugin will generate unprovable proof
-  obligation inside the reducibility argument for the fixpoint feature.
+Hypothetically speaking, We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.
 
-* (Reviewer D) **What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching?
+Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature.
+Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.
+
+* (Reviewer D) **"What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching?
 Is it rejected?
-In the latter case, how does the plugin prevent other extensible definitions from referring to it?**
+In the latter case, how does the plugin prevent other extensible definitions from referring to it?"**
 
@@
 
+* (Reviewer D) **""In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?"**
+
+For example
+```C
+Family A {Field a; Field b; Field c;}
+
+Family B extends A {
+  (* Inherit a. *) 
+  (* this vernacular command line will make a difference once uncommented. Before uncommented it, a2 will be the first field of the Family B.
+      The fields are order sensitive, especially when the definition of a2 dependent on a, placing a2 before a will cause program rejected. 
+      This ``annotation`` require the programmers effort. *) 
+  Field a2.
+}
+```
+
 <!-- I am not sure if I want to mention overridable/pins  -->
+
+* (Reviewer D) **"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation?
+Is it that modules are being generated in the background, while commands are being processed?**
+
+We should be more clear that we want to emphasize, whenever one vernacular command is emitted, our plugin will translate and type-check.
+Basically type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking happens after a whole family is closed
+
+Yes, modules/functors are generated in the background after each command is emitted.

### 2023-02-08 pldi2023/review-reply-densed.md
Save.

@@
   > We need to say something about "A type system for higher-order modules"
+
+Generally speaking, compared to ML-style module, Family focuses on overridability and code inheritance.
+The latter can be modeled by the functor in a verbose way.
+Module also has a clear distinction on the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to mundane classes.
+
+Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience.
+For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation).
+Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phase is acceptable.
+
+In fact, for plugin development, the family is compiled into module/functors.
+In meta-theory, we use sigma type as a conceptually simpler representation of modules.
+We also use singleton type to expose concrete type information in a family inspired by their work.
 

### 2023-02-08 pldi2023/review-reply-densed.md
Save.

@@
 
-* "Seems related to type theories for modules, but doesn't discuss the connection"
-
-Finally, I note that a family has similarities with an ML-style module,
-including the fact that some of its members are opaque while others are
-abstract. This is modelled through the use of singleton types, just as
-dependent type theories for modules do (e.g., Dreyer, Crary, Harper, POPL
-2003), which are also extensions of MLTT. Furthermore, there has been work on
-extending ML modules with mixin composition (Dreyer, Rossberg, ICFP 2008).
-ML-style modules are the basis of Coq's module system as well, so I would have
-expected some comparison with this line of work.
-
-  > We need to say something about "A type system for higher-order modules"
+* (Reviewer A) Some comparison between our work and [DCH 2003] [DR 2008]?
 
@@
 The latter can be modeled by the functor in a verbose way.
-Module also has a clear distinction on the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to mundane classes.
+Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.
 
@@
 We also use singleton type to expose concrete type information in a family inspired by their work.
+So our work is heavily influenced by ML-modules.
 

### 2023-02-08 pldi2023/review-reply-densed.md
Save. Formatting now.

@@
 We thank all the reviewers for their critical and encouraging feedback.
-Below we first address the main concerns, which might be paraphrased for the space.
+Below we address the main concerns, which might be paraphrased for the space.
 
-* (Reviewer A) "In terms of ergonomics, I wonder how powerful and usable" this is ...
-; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."
+* (Review A) __"In terms of ergonomics, I wonder how powerful and usable" this is ...
+; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."__
 
-As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level.
-We only have `fdiscriminate` and `finjection` for deducing intermediate fact.
-The key problem is the lack of certain nested pattern matching.
+  As Reviewer A/D highlights, the current method for inductively eliminating data only involves using "FInduction" at the top level.
+We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.
 
@@
   We acknowledge this limitation of our current research and are trying to advance more in this direction.
-We hypothesize that code involving nested induction/pattern matching an be reused by fill in the hole after extension upon inductive type.
+We hypothesize that code involving nested induction/pattern matching can be reused by fill in the hole after extension happens upon inductive type.
 The plugin will generate extra proof obligation to fill this hole.
@@
 
-* (Reviewer A) Some comparison between our work and [DCH 2003] [DR 2008]?
+* (Review A) __Some comparison between this paper and [DCH 2003] [DR 2008]?__
 
@@
 
-* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition.
-Reviewer D: "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"
+* (Review A/B/C/D) __Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition.
+"More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"__
 
@@
 
-We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater details aiming for readability.
+  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater details aiming for accessibility.
 
-### Review B
-
-* *(Reviewer B)"One thing I don't really get from the rules/discussion (including the
-  appendix) is the exact mechanism for choosing the A in rules like TyEq/PK/Add
-  and LSig/Add where A is supposed to abstract out the possibly-extensible types
-  in deciding what to make available to later definitions. Is A always going to
-  be uniquely determinable from s? (or if not, does the nondeterminism matter?)
-  Is it up to the user (in this case, the plugin) to get this right, or do the
-  rules somehow make sure that the right things are abstracted in a way I can't
-  see?"*
+* (Review B) __The `A` of _TyEq/PK/Add_
+  and _LSig/Add_ where `A` decides what to make available to later definitions. Is `A` non-determinstic? How does plugin decides `A`?__
 
@@
   Generally speaking, all the (extensible) inductive type will be simply
-  "wrapped" by a module type with its interface and only exposing the
+  "wrapped" by a module type only exposing the
   constructor (with no eliminators), just like how we generate `STLC°tm`.
 
-* *(Reviewer B/D) "I find it somewhat surprising that neither the FPOP plugin / implementation
-  itself, nor the code of the case studies, is provided."*
+* (Review B/D) __"I find it somewhat surprising that neither the FPOP plugin / implementation itself, nor the code of the case studies, is provided."__
 
@@
 
-* *(Review B) "I think the paper would also be improved by delineating the limitations of the
-  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"*
+* (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__
 
   Thank you for your feedback. We will clearly state the limitations of the
-  current implementation and meta-theory in the final revision. In particular,
-  we will emphasize the issue raised by Reviewer D regarding equality coercion
+  current implementation and meta-theory in the final revision.
+
+  In particular, we will emphasize the issue raised by Review D regarding equality coercion
   and the pervasive propositional equality on extensible inductive families.
@@
 
-### Review C
-
-* *"The utility of the FMLTT formalization in this paper is unclear to me. Don't
-  you already have consistency and canonicity via the translation to Coq
-  demonstrated by FPOP? Or put another way, instead of defining FMLTT directly,
-  couldn't you translate it to MLTT following the approach in FPOP?"*
+* (Review C) __The utility of the FMLTT formalization seems unclear since consistency and canonicity can already be deduced by the translation to MLTT following the approach in FPOP__
 
@@
   **pedantically** is not enough to show the consistency/canonicitiy for FMLTT
-  because **pedantically** target calculus is not proven to be
+  because, **pedantically**, target calculus is not proven to be
   consistency/canonicity.
@@
   we translate unconventional Wtype into the conventional one. The reason we
-  didn't choose to do so because we expect this translation a lot more verbose
+  didn't choose to do so is that we expect this translation a lot more verbose
   than the current proof because of the simplicity of Wtype itself compared to
@@
 
-### Other questions
+* (Review A/B) __In the conclusion of rule Tyeq/Casety, R should be T?__
 
-* (Reviewer A/B) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?
+Yes. Thank you for noticing this typo.
+
+* (Review A) __"Why is W(t) a term and not a type? What are bold-W and bold-P needed for?"__
 
@@
 
-* *(Reviewer B) How will our plugin react when trying to mixin the contradictory features?
+* *(Review B) How will our plugin react when trying to mixin the contradictory features?
 For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.*
@@
 
-* (Reviewer D) **"What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching?
+* (Review D) __"What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching?
 Is it rejected?
-In the latter case, how does the plugin prevent other extensible definitions from referring to it?"**
+In the latter case, how does the plugin prevent other extensible definitions from referring to it?"__
 
-  Coq will reject on the pattern matching on `tm`. For example, Figure 4, Line
-  494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the
-  given inductive `tm`. The following field definition will be type-checked
+  Coq will reject on the pattern matching on `tm`.
+For example, Figure 4, Line 494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the given inductive `tm`.
+The following field definition will be type-checked
   based on this interface, and for Coq it will only consider `tm` as an
@@
 
-* (Reviewer D) **""In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?"**
+* (Review D) __""In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?"__
 
@@
 
-* (Reviewer D) **"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation?
-Is it that modules are being generated in the background, while commands are being processed?**
+* (Review D) __"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation?
+Is it that modules are being generated in the background, while commands are being processed?__
 

### 2023-02-08 pldi2023/review-reply-densed.md
Please resolve the HTML comment. Only left with mixin module paper

@@
   current implementation and meta-theory in the final revision. 
-
+  <!-- do we want to mention overridability/pins? -->
   In particular, we will emphasize the issue raised by Review D regarding equality coercion
@@
 
-* *(Review B) How will our plugin react when trying to mixin the contradictory features?
-For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.*
+* (Review B) __How will our plugin react when trying to mixin the contradictory features?
+For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.__
 
@@
 
-<!-- I am not sure if I want to mention overridable/pins  -->
-
 * (Review D) __"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation?
@@
 
-  We should be more clear that we want to emphasize, whenever one vernacular command is emitted, our plugin will translate and type-check.
-Basically type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking happens after a whole family is closed
+  We should be clearer that, the point we want to emphasize is, whenever one vernacular command is emitted, our plugin will translate and type-check.
+Basically type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking only happens after a whole family is closed.
 

### 2023-02-08 pldi2023/review-reply-densed.md
Please resolve the HTML comments.
edits to two answers

@@
       fill in the new hole once pattern matching needs to extend the clauses
-      lifting nested induction to top level is not necessary
+
+      lifting nested induction to top level seems not necessary
   -->
@@
 
-* (Review B) __The `A` of _TyEq/PK/Add_
-  and _LSig/Add_ where `A` decides what to make available to later definitions. Is `A` non-determinstic? How does plugin decides `A`?__
+* (Review-B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
 
-  The typing rules do not commit to a particular choice of $A$, as the reviewer
-  points out.
+  The FMLTT typing rules do not commit to a particular choice of the context type $A$,
+  though there exists a simple algorithm for picking the right $A$ (lines 785–787): a
+  field is checked in a context of type $A$ that hides W-type signatures behind
+  $\mathbb{U}$, unless that field invokes a constructor or eliminator of some
+  W-type. Intuitively, the type system is sound without this algorithm because
+  it restricts the ways a field can be used if a wrong $A$ is chosen.
+
+<!--
    In practice, this A is decided by implementation.
@@
   constructor (with no eliminators), just like how we generate `STLC°tm`.
+-->
 
-* (Review B/D) __"I find it somewhat surprising that neither the FPOP plugin / implementation itself, nor the code of the case studies, is provided."__
+* (Reviews-B&D) __The plugin implementation is not included as a supplement.__
 
-  We will certainly make the FPOP plugin/implementation available.
+  We will ensure the release of our research artifacts for public access.
 

### 2023-02-08 pldi2023/review-reply-densed.md
Add comment on mixin up ML.
Finish all questions.

@@
 So our work is heavily influenced by ML-modules.
+
+  Our module and mixin has the similar semantic of that from [DCH 2003] [DR 2008].
+However, compared to their work, we mainly focus on the perspect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
+Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.
 

### 2023-02-08 pldi2023/review-reply-densed.md
Fix typo.

@@
   and practice, (a). such as a normalization algorithm for an extended calculus
-  that allows for more convertibility upon elimination of the extensible
+  that admits more conversion equation upon elimination of the extensible
   inductive type; (b). and a proper proof assistant that supports native

### 2023-02-08 pldi2023/review-reply-densed.md
Save.

@@
 
-  As Reviewer A/D highlights, the current method for inductively eliminating data only involves using "FInduction" at the top level.
+  As Reviewer A/D highlighted, the current method for inductively eliminating data only involves using "FInduction" at the top level.
 We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.
@@
   We acknowledge this limitation of our current research and are trying to advance more in this direction.
-We hypothesize that code involving nested induction/pattern matching can be reused by fill in the hole after extension happens upon inductive type.
-The plugin will generate extra proof obligation to fill this hole.
-This certainly requires significant amount of engineering effort.
+We hypothesize that we can reuse code involving nested induction/pattern matching, by filling in the new matching clauses after extension happens upon the inductive type.
+The plugin will generate an extra proof obligation to fill this hole.
+It certainly requires a significant amount of engineering effort.
 
@@
 
-  Generally speaking, compared to ML-style module, Family focuses on overridability and code inheritance.
-The latter can be modeled by the functor in a verbose way.
+  Generally speaking, compared to ML-style modules, families focus on overridability and code inheritance.
+The latter can be modelled by the functor in a verbose way.
 Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.
@@
 For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation).
-Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phase is acceptable.
+Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phases is acceptable.
 
-  In fact, for plugin development, the family is compiled into module/functors.
+  In fact, for plugin development, the family is compiled into modules/functors.
 In meta-theory, we use sigma type as a conceptually simpler representation of modules.
 We also use singleton type to expose concrete type information in a family inspired by their work.
-So our work is heavily influenced by ML-modules.
+So our work is heavily influenced by ML-module.
 
-  Our module and mixin has the similar semantic of that from [DCH 2003] [DR 2008].
-However, compared to their work, we mainly focus on the perspect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
+  Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008].
+However, compared to their work, we mainly focus on the prospect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
 Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.

### 2023-02-08 pldi2023/review-reply-densed.md
Fix typo

@@
   Thank you for your feedback. 
-  We sympathize with this impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the readers.
+  We sympathize with this impression in Section 5, and we agree it is dense, hard to read and not giving too much inspiration to the readers.
 
-  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater details aiming for accessibility.
+  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater detail aiming for accessibility.
 
@@
   and practice, (a). such as a normalization algorithm for an extended calculus
-  that admits more conversion equation upon elimination of the extensible
+  that admits more conversion equations upon elimination of the extensible
   inductive type; (b). and a proper proof assistant that supports native
@@
   thus allowing for proper reduction of open terms. This will change the kernel
-  of proof assistant for sure.
+  of the proof assistant for sure.
 

### 2023-02-08 pldi2023/review-reply-densed.md
Pass Grammarly.

@@
 
-  The technical detail here is that, FMLTT also has this unconventional Wtype
-  formulation. While each classic Wtype is defined by a pair of type `A ⊢ B`,
-  our unconventional Wtype is defined by a list of pair of types `Aᵢ ⊢ Bᵢ`, each
+  The technical detail here is that FMLTT also has this unconventional Wtype
+  formulation. While each classic Wtype is defined by a pair of types `A |- B`,
+  our unconventional Wtype is defined by a list of pairs of types `A_i ⊢ B_i`, and each
   pair corresponds to one constructor.
 
-  One way to interpret it is the classical formluation encoding this list using
-  a big sum type. So the conventional and unconventional Wtype can be translated
-  to each other.
+  The conventional and unconventional Wtype can be translated
+  to each other because a list is also an inductive type.
 
@@
   can only function as a guide to the plugin implementation, and
-  **pedantically** is not enough to show the consistency/canonicitiy for FMLTT
+  **pedantically** is not enough to show the consistency/canonicity for FMLTT
   because, **pedantically**, target calculus is not proven to be
@@
 
-  We agree with Reviewer B's insight -- we will get consistency/canonicitiy when
-  we translate unconventional Wtype into the conventional one. The reason we
+  We agree with Reviewer B's insight -- we will get consistency/canonicity when
+  we translate the unconventional Wtype into the conventional one. The reason we
   didn't choose to do so is that we expect this translation a lot more verbose
@@
   the rich functionality provided by (fake-)Agda's Inductive Facility. We only
-  use the latter when constructing consistency/canonicity model.
+  use the latter when constructing the consistency/canonicity model.
 
@@
 
-  In dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.
+  In a dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.
 
-  𝕎(t) is a typo and should be deleted. Thanks for pointing out.
+  𝕎(t) is a typo and should be deleted. Thanks for pointing that out.
 
@@
   a linkage(overridable/extensible) into a module(sigma type). The reason it is
-  irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but
+  irreplaceable is that proper abstraction cannot happen on linkage (𝕃) but
   only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
@@
   <!-- I think we still need to clarify we doesn't support extending these -->
-  We currently doesn't support extend STLC with polymorphism and reference, which requires extending existent inductive family with new indices.
+  We currently don't support extending STLC with polymorphism and reference, which requires extending the existent inductive family with new indices.
 
-  Hypothetically speaking, We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.
+  Hypothetically speaking, We can expect our plugin will generate unprovable proof obligation under mixin, hindering qeding the proposition and thus closing the family.
 
-  Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature.
-Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.
+  Another example would be extending *a family of STLC and its termination proof* with the general recursion feature.
+Our plugin will generate an unprovable proof obligation inside the reducibility argument for the fixpoint feature.
 
@@
 
-  Coq will reject on the pattern matching on `tm`.
+  Coq will reject the pattern matching on `tm`.
 For example, Figure 4, Line 494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the given inductive `tm`.
 The following field definition will be type-checked
-  based on this interface, and for Coq it will only consider `tm` as an
-  arbitrary type and will fail to pattern match it.
+  based on this interface. As for Coq, it will only consider `tm` as an
+  arbitrary type and will fail to pattern-match it.
 
@@
   (* Inherit a. *) 
-  (* this vernacular command line will make a difference once uncommented. Before uncommented it, a2 will be the first field of the Family B.
-      The fields are order sensitive, especially when the definition of a2 dependent on a, placing a2 before a will cause program rejected. 
-      This ``annotation`` require the programmers effort. *) 
+    (* this vernacular command line will make a difference once uncommented. 
+        Before uncommenting it, a2 will be the first field of Family B.
+        The fields are order sensitive, especially when the definition of a2 depends on a, placing a2 before a will fail the type-check. 
+      This ``annotation`` require the programmer`s effort. *) 
   Field a2.
@@
   We should be clearer that, the point we want to emphasize is, whenever one vernacular command is emitted, our plugin will translate and type-check.
-Basically type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking only happens after a whole family is closed.
+Basically, type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking only happens after a whole family is closed.
 

### 2023-02-08 pldi2023/review-reply-densed.md
Pass Grammarly. 1848 words right now, twice more.

@@
   We should be clearer that, the point we want to emphasize is, whenever one vernacular command is emitted, our plugin will translate and type-check.
-Basically, type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking only happens after a whole family is closed.
+Basically, type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where type-checking only happens after a whole family is closed.
 

### 2023-02-08 pldi2023/review-reply-densed.md
shortening

--- 2023-02-08 pldi2023/review-reply-densed.md
@@
 
-* (Review A) __"In terms of ergonomics, I wonder how powerful and usable" this is ...
-; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."__
+* (R-A, R-D) **Ergonomics is lacking because induction is allowed only at the top level.**
 
-  As Reviewer A/D highlighted, the current method for inductively eliminating data only involves using "FInduction" at the top level.
+  As Reviewer A/D highlights, the current method for inductively eliminating data only involves using "FInduction" at the top level.
 We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.
@@
       fill in the new hole once pattern matching needs to extend the clauses
-
       lifting nested induction to top level seems not necessary
@@
   We acknowledge this limitation of our current research and are trying to advance more in this direction.
-We hypothesize that we can reuse code involving nested induction/pattern matching, by filling in the new matching clauses after extension happens upon the inductive type.
-The plugin will generate an extra proof obligation to fill this hole.
-It certainly requires a significant amount of engineering effort.
+We hypothesize that code involving nested induction/pattern matching can be reused by fill in the hole after extension happens upon inductive type.
+The plugin will generate extra proof obligation to fill this hole.
+This certainly requires significant amount of engineering effort.
 
-* (Review A) __Some comparison between this paper and [DCH 2003] [DR 2008]?__
+* (A) __Connection to work on ML-style modules [DCH 2003, DR 2008]?__
 
-  Generally speaking, compared to ML-style modules, families focus on overridability and code inheritance.
-The latter can be modelled by the functor in a verbose way.
+  Thank you for the references.
+  Both ML-style modules and our families are modularity mechanisms.
+  ML-style modules are more about abstraction (caring about issues such as
+  signature ascription), while our families are more about extensibility.
+  Both use singletons to model and control the propagation of
+  definitions inside a module or family.
+  Work on mixin composition for ML-style modules focuses on making 
+  mixins play well with the peculiarities of ML-style modules,
+  while our work focuses on supporting mixins in the presence of extensible
+  inductive types.
+
+  <!--
+  Generally speaking, compared to ML-style module, Family focuses on overridability and code inheritance.
+The latter can be modeled by the functor in a verbose way.
 Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.
-
+  ...
   Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience.
 For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation).
-Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phases is acceptable.
-
-  In fact, for plugin development, the family is compiled into modules/functors.
+Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phase is acceptable.
+  ...
+  In fact, for plugin development, the family is compiled into module/functors.
 In meta-theory, we use sigma type as a conceptually simpler representation of modules.
 We also use singleton type to expose concrete type information in a family inspired by their work.
-So our work is heavily influenced by ML-module.
+So our work is heavily influenced by ML-modules.
+  ...
+  Our module and mixin has the similar semantic of that from [DCH 2003] [DR 2008].
+However, compared to their work, we mainly focus on the perspect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
+Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.
+  -->
 
-  Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008].
-However, compared to their work, we mainly focus on the prospect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
-Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.
+* (A, B, D) **The presentation of FMLTT is tough. The Agda-looking scripts are not readable.**
 
-* (Review A/B/C/D) __Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition.
-"More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"__
+  We sympathize with the reviewers. We will expand the section to clarify things
+  in response to the reviewers' questions. We will also use an appendix to
+  explain the formalization and proofs in greater detail.
 
-  Thank you for your feedback. 
-  We sympathize with this impression in Section 5, and we agree it is dense, hard to read and not giving too much inspiration to the readers.
-
-  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater detail aiming for accessibility.
-
-* (Review-B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
+* (B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
 
@@
 
-* (Reviews-B&D) __The plugin implementation is not included as a supplement.__
+* (B, D) __The plugin implementation is not included as a supplement.__
 
@@
   and practice, (a). such as a normalization algorithm for an extended calculus
-  that admits more conversion equations upon elimination of the extensible
+  that allows for more convertibility upon elimination of the extensible
   inductive type; (b). and a proper proof assistant that supports native
@@
   thus allowing for proper reduction of open terms. This will change the kernel
-  of the proof assistant for sure.
+  of proof assistant for sure.
 
@@
 
-* (Review C) __The utility of the FMLTT formalization seems unclear since consistency and canonicity can already be deduced by the translation to MLTT following the approach in FPOP__
+* (B, C) __Consistency is already implied for FMLTT by a translation to MLTT, and for FPOP by a translation to Coq__.
 
-  The technical detail here is that, FMLTT also has this unconventional Wtype
-  formulation. While each classic Wtype is defined by a pair of type `A ⊢ B`,
-  our unconventional Wtype is defined by a list of pair of types `Aᵢ ⊢ Bᵢ`, each
-  pair corresponds to one constructor.
+  We agree with the reviewers' insights.
 
-  One way to interpret it is the classical formluation encoding this list using
-  a big sum type. So the conventional and unconventional Wtype can be translated
-  to each other.
+  We note that the effort required to prove consistency directly for FMLTT is
+  almost identical to that required to prove a translation to MLTT is
+  type-preserving: both proofs are structured as type-preserving, metacircular
+  interpretations.
 
+  <!--
+  Our main excuse for doing the proofs is that we found it more
+  educational for ourselves to directly prove FMLTT consistent than doing the
+  translation.
+  ...
+  We decided to prove consistency and canonicity directly for FMLTT, because our
+  formulation of W-types, which are part of both FMLTT and our target MLTT, is
+  slightly unconventional.
+  We define a W-type as given by a list of pairs of types $A_i ⊢ B_i$; each pair
+  corresponds to one constructor.
+  The convention is to define a W-type as given by a single pair of types $A ⊢ B$,
+  which can be understood to encode multiple constructors using
+  using a big sum type. So the conventional and unconventional Wtype can be
+  translated to each other. 
+  ...
   However, our translation doesn't translate this unconventional Wtype
@@
   consistency/canonicity.
-
+  ...
   We agree with Reviewer B's insight -- we will get consistency/canonicitiy when
@@
   use the latter when constructing consistency/canonicity model.
+  -->
 
-* (Review A/B) __In the conclusion of rule Tyeq/Casety, R should be T?__
+* (A, B) __In the conclusion of rule Tyeq/Casety, R should be T?__
 

### 2023-02-08 pldi2023/review-reply-densed.md
Add a sentence about the relation between consistency and translation.

--- 2023-02-08 pldi2023/review-reply-densed.md
@@
 
-* (Review A) __"In terms of ergonomics, I wonder how powerful and usable" this is ...
-; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."__
+* (R-A, R-D) **Ergonomics is lacking because induction is allowed only at the top level.**
 
@@
 
-* (Review A) __Some comparison between this paper and [DCH 2003] [DR 2008]?__
+* (A) __Connection to work on ML-style modules [DCH 2003, DR 2008]?__
 
+  Thank you for the references.
+  Both ML-style modules and our families are modularity mechanisms.
+  ML-style modules are more about abstraction (caring about issues such as
+  signature ascription), while our families are more about extensibility.
+  Both use singletons to model and control the propagation of
+  definitions inside a module or family.
+  Work on mixin composition for ML-style modules focuses on making 
+  mixins play well with the peculiarities of ML-style modules,
+  while our work focuses on supporting mixins in the presence of extensible
+  inductive types.
+
+  <!--
   Generally speaking, compared to ML-style modules, families focus on overridability and code inheritance.
@@
 Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.
-
+  ...
   Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience.
@@
 Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phases is acceptable.
-
+  ...
   In fact, for plugin development, the family is compiled into modules/functors.
@@
 So our work is heavily influenced by ML-module.
-
+  ...
   Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008].
@@
 Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.
+  -->
 
-* (Review A/B/C/D) __Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition.
-"More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"__
+* (A, B, D) **The presentation of FMLTT is tough. The Agda-looking scripts are not readable.**
 
+  We sympathize with the reviewers. We will expand the section to clarify things
+  in response to the reviewers' questions. We will also use an appendix to
+  explain the formalization and proofs in greater detail.
+
+  <!--
   Thank you for your feedback. 
   We sympathize with this impression in Section 5, and we agree it is dense, hard to read and not giving too much inspiration to the readers.
+  ...
+  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater detail aiming for accessibility. 
+  -->
 
-  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater detail aiming for accessibility.
-
-* (Review-B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
+* (B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
 
@@
 
-* (Review C) __The utility of the FMLTT formalization seems unclear since consistency and canonicity can already be deduced by the translation to MLTT following the approach in FPOP__
+* (B, C) __Consistency is already implied for FMLTT by a translation to MLTT, and for FPOP by a translation to Coq__.
 
-  The technical detail here is that FMLTT also has this unconventional Wtype
-  formulation. While each classic Wtype is defined by a pair of types `A |- B`,
-  our unconventional Wtype is defined by a list of pairs of types `A_i ⊢ B_i`, and each
-  pair corresponds to one constructor.
+  We agree with the reviewers' insights.
 
+  We note that the effort required to prove consistency directly for FMLTT is
+  almost identical to that required to prove a translation to MLTT is
+  type-preserving: both proofs are structured as type-preserving, metacircular
+  interpretations.
+
+  <!--
+  Our main excuse for doing the proofs is that we found it more
+  educational for ourselves to directly prove FMLTT consistent than doing the
+  translation.
+  ...
+  We decided to prove consistency and canonicity directly for FMLTT, because our
+  formulation of W-types, which are part of both FMLTT and our target MLTT, is
+  slightly unconventional.
+  We define a W-type as given by a list of pairs of types $A_i ⊢ B_i$; each pair
+  corresponds to one constructor.
+  The convention is to define a W-type as given by a single pair of types $A ⊢ B$.
   The conventional and unconventional Wtype can be translated
   to each other because a list is also an inductive type.
-
+  So the conventional and unconventional Wtype can be
+  translated to each other. 
+  ...
   However, our translation doesn't translate this unconventional Wtype
@@
   consistency/canonicity.
-
+  ...
   We agree with Reviewer B's insight -- we will get consistency/canonicity when
@@
   use the latter when constructing the consistency/canonicity model.
+  -->
 
-* (Review A/B) __In the conclusion of rule Tyeq/Casety, R should be T?__
+* (A, B) __In the conclusion of rule Tyeq/Casety, R should be T?__
 

### 2023-02-08 pldi2023/review-reply-densed.md
move two minor questions to the end
edits

@@
 
-* (R-A, R-D) **Ergonomics is lacking because induction is allowed only at the top level.**
+* (A, D) **Ergonomics is lacking because induction is allowed only at the top level.**
 
@@
 Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".
-
+  ...
   Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A`
@@
   of type `tm : 𝕌` thus doing abstraction successfully.
-
+  ...
   Generally speaking, all the (extensible) inductive type will be simply
@@
   We agree with Reviewer B's insight -- we will get consistency/canonicity when
-  we translate the unconventional Wtype into the conventional one. The reason we
+  we translate the unconventional Wtype into the conventional one. In fact, our consistency 
+  proof is structurally identical to the  type-preservation proof for a translation to MLTT, 
+  since consistency is proved via a type-preserving, metacircular interpretation.
+  ...
+  The reason we
   didn't choose to do so is that we expect this translation a lot more verbose
@@
 
-* (A, B) __In the conclusion of rule Tyeq/Casety, R should be T?__
+* (B) __What happens if the programmer tries to mix contradictory language features?__
 
-  Yes. Thank you for noticing this typo.
+  We expect mixing two contradictory developments to cause the plugin to
+  generate unprovable proof obligations (cf. line 386). One example is mixing
+  HM-style polymorphism and references, which the reviewer mentioned. Another
+  example is mixing fixpoints into a family that proves strong normalization for
+  STLC.
 
-* (Review A) __"Why is W(t) a term and not a type? What are bold-W and bold-P needed for?"__
-
-  In a dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.
-
-  𝕎(t) is a typo and should be deleted. Thanks for pointing that out.
-
-  ℙ is another technical detail of the system. Its functionality is to transform
-  a linkage(overridable/extensible) into a module(sigma type). The reason it is
-  irreplaceable is that proper abstraction cannot happen on linkage (𝕃) but
-  only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
-  A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial
-  `A₆`.
-
-* (Review B) __How will our plugin react when trying to mixin the contradictory features?
-For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.__
-
-  <!-- I think we still need to clarify we doesn't support extending these -->
+  <!--
   We currently don't support extending STLC with polymorphism and reference, which requires extending the existent inductive family with new indices. 
-
-  Hypothetically speaking, We can expect our plugin will generate unprovable proof obligation under mixin, hindering qeding the proposition and thus closing the family.
-
+  ...
+  Hypothetically speaking, we can expect our plugin will generate unprovable proof obligation under mixin, hindering qeding the proposition and thus closing the family.
+  ...
   Another example would be extending *a family of STLC and its termination proof* with the general recursion feature.
 Our plugin will generate an unprovable proof obligation inside the reducibility argument for the fixpoint feature.
+  -->
 
@@
   Yes, modules/functors are generated in the background after each command is emitted.
+
+* (A, B) __In the conclusion of rule Tyeq/Casety, R should be T?__
+
+  Yes. Thank you for noticing this typo.
+
+* (A) __"Why is W(t) a term and not a type? What are bold-W and bold-P needed for?"__
+
+  In a dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.
+
+  𝕎(t) is a typo and should be deleted. Thanks for pointing that out.
+
+  ℙ is another technical detail of the system. Its functionality is to transform
+  a linkage(overridable/extensible) into a module(sigma type). The reason it is
+  irreplaceable is that proper abstraction cannot happen on linkage (𝕃) but
+  only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
+  A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial
+  `A₆`.

### 2023-02-08 pldi2023/review-reply-densed.md
shortening

@@
 
-* (Review D) __"What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching?
-Is it rejected?
-In the latter case, how does the plugin prevent other extensible definitions from referring to it?"__
+* (D) __What happens if a user tries to use `tm_rect` with standard pattern matching?__
 
-  Coq will reject the pattern matching on `tm`.
-For example, Figure 4, Line 494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the given inductive `tm`.
-The following field definition will be type-checked
-  based on this interface. As for Coq, it will only consider `tm` as an
-  arbitrary type and will fail to pattern-match it.
+  The programmer is prevented from using `tm_rect` within the family in which `tm` is defined,
+  because the plugin makes `tm` appear in the typing context as an axiomatized
+  type without exposing its concrete definition (line 494).
 

### 2023-02-08 pldi2023/review-reply-densed.md
edit

@@
 
-* (Review D) __""In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?"__
+* (D) __What does it mean to place a new field using annotations?__
 
-  For example
+  We clarify this using the following example.
+  The programmer uses the annotation `Inherit a2` to indicate that the new field
+  `b` should be placed immediately after the existing field `a2`.
+
   ```C
-Family A {Field a; Field b; Field c;}
+  Family A.
+    FDefinition a1 := ....
+    FDefinition a2 := ....
+    FDefinition c := ....
+  End A.
 
-Family B extends A {
-  (* Inherit a. *) 
-    (* this vernacular command line will make a difference once uncommented. 
-        Before uncommenting it, a2 will be the first field of Family B.
-        The fields are order sensitive, especially when the definition of a2 depends on a, placing a2 before a will fail the type-check. 
-      This ``annotation`` require the programmer`s effort. *) 
-  Field a2.
-}
+  Family B extends A.
+    Inherit a2.
+    FDefinition b := ...
+  End B
   ```

### 2023-02-08 pldi2023/review-reply-densed.md
shorten

@@
 
-* (Review D) __"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation?
-Is it that modules are being generated in the background, while commands are being processed?__
+* (D) __Line 487 (__*a module named STLC◦subst◦Cases is generated interactively ...*__) is confusing.__
 
-  We should be clearer that, the point we want to emphasize is, whenever one vernacular command is emitted, our plugin will translate and type-check.
-Basically, type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where type-checking only happens after a whole family is closed.
-
-  Yes, modules/functors are generated in the background after each command is emitted.
+  We wanted to emphasize that programming is interactive even within an `FRecursion` definition:
+  the programmer need not wait until the entire `FRecursion` definition is completed to have
+  their `Case` commands checked and translated by the plugin.
 

### 2023-02-08 pldi2023/review-reply-densed.md
shortening

@@
 We thank all the reviewers for their critical and encouraging feedback.
-Below we address the main concerns, which might be paraphrased for the space.
+Below we address the main concerns, which might be paraphrased for space.
 
+* (A, B, C, D) __There are limitations with the current design and implementation, such as
+  restricting induction to the top level and the awkwardness in handling terms like `denoteTy t`.__
+
+  We will clearly state the current limitations. We believe the limitations
+  identified by the reviewers represent interesting opportunities for future research.
+  For example, future work could explore a normalization algorithm for an
+  extended dependent type theory that admits more conversion equations on
+  elimination of extensible inductive types, as well as proof assistant
+  implementations that natively support extensible inductive types and thus
+  streamline normalization for open terms.
+
+  We will also add a comment about the trusted base of any developments using the
+  plugin, as the reviewers suggested.
+
+  <!--
   * (A, D) **Ergonomics is lacking because induction is allowed only at the top level.**
-
+  ...
   As Reviewer A/D highlighted, the current method for inductively eliminating data only involves using "FInduction" at the top level.
 We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.
-
-  <!-- The idea is 
-      fill in the new hole once pattern matching needs to extend the clauses
-      lifting nested induction to top level seems not necessary
-  -->
-
+  ...
+  The idea is 
+  - fill in the new hole once pattern matching needs to extend the clauses
+  - lifting nested induction to top level seems not necessary
+  ...
   We acknowledge this limitation of our current research and are trying to advance more in this direction.
@@
 It certainly requires a significant amount of engineering effort.
+  -->
+
+  <!--
+  * (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__
+  ...
+  Thank you for your feedback. We will clearly state the limitations of the
+  current implementation and meta-theory in the final revision. 
+  ...
+  do we want to mention overridability/pins?
+  ...
+  In particular, we will emphasize the issue raised by Review D regarding equality coercion
+  and the pervasive propositional equality on extensible inductive families.
+  This can result in unnecessary equality coercion when `denoteTy t` is not
+  definitionally equal to `nat`.
+  ...
+  We believe that this opens up new opportunities for future work in both theory
+  and practice, (a). such as a normalization algorithm for an extended calculus
+  that admits more conversion equations upon elimination of the extensible
+  inductive type; (b). and a proper proof assistant that supports native
+  extensible inductive types instead of the current encoding via our plugin,
+  thus allowing for proper reduction of open terms. This will change the kernel
+  of the proof assistant for sure. 
+  ...
+  In the final revision, we will also clarify that our plugin only translates
+  code into vanilla Coq terms without expanding the trusted codebase.
+  -->
 
@@
   We will ensure the release of our research artifacts for public access.
-
-* (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__
-
-  Thank you for your feedback. We will clearly state the limitations of the
-  current implementation and meta-theory in the final revision. 
-  <!-- do we want to mention overridability/pins? -->
-  In particular, we will emphasize the issue raised by Review D regarding equality coercion
-  and the pervasive propositional equality on extensible inductive families.
-  This can result in unnecessary equality coercion when `denoteTy t` is not
-  definitionally equal to `nat`.
-
-  We believe that this opens up new opportunities for future work in both theory
-  and practice, (a). such as a normalization algorithm for an extended calculus
-  that admits more conversion equations upon elimination of the extensible
-  inductive type; (b). and a proper proof assistant that supports native
-  extensible inductive types instead of the current encoding via our plugin,
-  thus allowing for proper reduction of open terms. This will change the kernel
-  of the proof assistant for sure.
-
-  In the final revision, we will also clarify that our plugin only translates
-  code into vanilla Coq terms without expanding the trusted codebase.
 
@@
 
-* (A, B) __In the conclusion of rule Tyeq/Casety, R should be T?__
+* (A, B) __In the conclusion of the rule TYEQ/CASETY, should $R$ be $T$?__
 

### 2023-02-08 pldi2023/review-reply-densed.md
reorder

@@
 
-* (A) __Connection to work on ML-style modules [DCH 2003, DR 2008]?__
-
-  Thank you for the references.
-  Both ML-style modules and our families are modularity mechanisms.
-  ML-style modules are more about abstraction (caring about issues such as
-  signature ascription), while our families are more about extensibility.
-  Both use singletons to model and control the propagation of
-  definitions inside a module or family.
-  Work on mixin composition for ML-style modules focuses on making 
-  mixins play well with the peculiarities of ML-style modules,
-  while our work focuses on supporting mixins in the presence of extensible
-  inductive types.
-
-  <!--
-  Generally speaking, compared to ML-style modules, families focus on overridability and code inheritance.
-The latter can be modelled by the functor in a verbose way.
-Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.
-  ...
-  Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience.
-For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation).
-Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phases is acceptable.
-  ...
-  In fact, for plugin development, the family is compiled into modules/functors.
-In meta-theory, we use sigma type as a conceptually simpler representation of modules.
-We also use singleton type to expose concrete type information in a family inspired by their work.
-So our work is heavily influenced by ML-module.
-  ...
-  Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008].
-However, compared to their work, we mainly focus on the prospect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
-Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.
-  -->
-
 * (A, B, D) **The presentation of FMLTT is tough. The Agda-looking scripts are not readable.**
@@
 
-* (B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
-
-  The FMLTT typing rules do not commit to a particular choice of the context type $A$,
-  though there exists a simple algorithm for picking the right $A$ (lines 785–787): a
-  field is checked in a context of type $A$ that hides W-type signatures behind
-  $\mathbb{U}$, unless that field invokes a constructor or eliminator of some
-  W-type. Intuitively, the type system is sound without this algorithm because
-  it restricts the ways a field can be used if a wrong $A$ is chosen.
-
-  <!--
-  In practice, this A is decided by implementation.
-Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".
-  ...
-  Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A`
-  has `tm : 𝕌` instead. `s` will make sure other fields stay the same type. To
-  show this explicitly at the plugin level, we look at Figure 4 (Line 494)
-  `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding
-  to `tm : 𝕌`). With this interface `STLC°tm` we cannot pattern match any term
-  of type `tm : 𝕌` thus doing abstraction successfully.
-  ...
-  Generally speaking, all the (extensible) inductive type will be simply
-  "wrapped" by a module type only exposing the
-  constructor (with no eliminators), just like how we generate `STLC°tm`.
-  -->
-
-* (B, D) __The plugin implementation is not included as a supplement.__
-
-  We will ensure the release of our research artifacts for public access.
-
 * (B, C) __Consistency is already implied for FMLTT by a translation to MLTT, and for FPOP by a translation to Coq__.
@@
   We note that the effort required to prove consistency directly for FMLTT is
-  almost identical to that required to prove a translation to MLTT is
+  similar to that required to prove a translation to MLTT
   type-preserving: both proofs are structured as type-preserving, metacircular
@@
   use the latter when constructing the consistency/canonicity model.
+  -->
+
+* (B, D) __The plugin implementation is not included as a supplement.__
+
+  We will ensure the release of our research artifacts for public access.
+
+* (A) __Connection to work on ML-style modules [DCH 2003, DR 2008]?__
+
+  Thank you for the references.
+  Both ML-style modules and our families are modularity mechanisms.
+  ML-style modules are more about abstraction (caring about issues such as
+  signature ascription), while our families are more about extensibility.
+  Both use singletons to model and control the propagation of
+  definitions inside a module or family.
+  Work on mixin composition for ML-style modules focuses on making 
+  mixins play well with the peculiarities of ML-style modules,
+  while our work focuses on supporting mixins in the presence of extensible
+  inductive types.
+
+  <!--
+  Generally speaking, compared to ML-style modules, families focus on overridability and code inheritance.
+The latter can be modelled by the functor in a verbose way.
+Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.
+  ...
+  Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience.
+For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation).
+Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phases is acceptable.
+  ...
+  In fact, for plugin development, the family is compiled into modules/functors.
+In meta-theory, we use sigma type as a conceptually simpler representation of modules.
+We also use singleton type to expose concrete type information in a family inspired by their work.
+So our work is heavily influenced by ML-module.
+  ...
+  Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008].
+However, compared to their work, we mainly focus on the prospect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
+Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.
+  -->
+
+* (B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
+
+  The FMLTT typing rules do not commit to a particular choice of the context type $A$,
+  though there exists a simple algorithm for picking the right $A$ (lines 785–787): a
+  field is checked in a context of type $A$ that hides W-type signatures behind
+  $\mathbb{U}$, unless that field invokes a constructor or eliminator of some
+  W-type. Intuitively, the type system is sound without this algorithm because
+  it restricts the ways a field can be used if a wrong $A$ is chosen.
+
+  <!--
+  In practice, this A is decided by implementation.
+Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".
+  ...
+  Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A`
+  has `tm : 𝕌` instead. `s` will make sure other fields stay the same type. To
+  show this explicitly at the plugin level, we look at Figure 4 (Line 494)
+  `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding
+  to `tm : 𝕌`). With this interface `STLC°tm` we cannot pattern match any term
+  of type `tm : 𝕌` thus doing abstraction successfully.
+  ...
+  Generally speaking, all the (extensible) inductive type will be simply
+  "wrapped" by a module type only exposing the
+  constructor (with no eliminators), just like how we generate `STLC°tm`.
   -->
@@
 
-* (A, B) __In the conclusion of the rule TYEQ/CASETY, should $R$ be $T$?__
-
-  Yes. Thank you for noticing this typo.
-
-* (A) __"Why is W(t) a term and not a type? What are bold-W and bold-P needed for?"__
-
-  In a dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.
-
-  𝕎(t) is a typo and should be deleted. Thanks for pointing that out.
-
-  ℙ is another technical detail of the system. Its functionality is to transform
-  a linkage(overridable/extensible) into a module(sigma type). The reason it is
-  irreplaceable is that proper abstraction cannot happen on linkage (𝕃) but
-  only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
-  A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial
-  `A₆`.
+We will address in our revision other reviewer comments not mentioned above.
+Again, we thank all the reviewers for their thoughtful and detailed comments, which will help improve the paper.

### 2023-02-08 pldi2023/review-reply-densed.md
satiate reviewer D's curiosity

@@
 
+* (D) __Out of curiosity, how does your plugin integrate with universes?__
+
+  We do not have intimate knowledge of Coq's universe polymorphism feature, but we expect
+  it to have the same level of compatibility with families as it does with Coq
+  modules—under the hood, upon closing a family, the plugin is just doing
+  repeated module inclusion and instantiation.
+
 We will address in our revision other reviewer comments not mentioned above.

### 2023-02-08 pldi2023/review-reply-final.md
Save Final/Clean Version for reply.

--- 2023-02-08 pldi2023/review-reply-densed.md
@@
 
-  <!--
-  * (A, D) **Ergonomics is lacking because induction is allowed only at the top level.**
-  ...
-  As Reviewer A/D highlighted, the current method for inductively eliminating data only involves using "FInduction" at the top level.
-We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.
-  ...
-  The idea is 
-  - fill in the new hole once pattern matching needs to extend the clauses
-  - lifting nested induction to top level seems not necessary
-  ...
-  We acknowledge this limitation of our current research and are trying to advance more in this direction.
-We hypothesize that we can reuse code involving nested induction/pattern matching, by filling in the new matching clauses after extension happens upon the inductive type.
-The plugin will generate an extra proof obligation to fill this hole.
-It certainly requires a significant amount of engineering effort.
-  -->
-
-  <!--
-  * (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__
-  ...
-  Thank you for your feedback. We will clearly state the limitations of the
-  current implementation and meta-theory in the final revision. 
-  ...
-  do we want to mention overridability/pins?
-  ...
-  In particular, we will emphasize the issue raised by Review D regarding equality coercion
-  and the pervasive propositional equality on extensible inductive families.
-  This can result in unnecessary equality coercion when `denoteTy t` is not
-  definitionally equal to `nat`.
-  ...
-  We believe that this opens up new opportunities for future work in both theory
-  and practice, (a). such as a normalization algorithm for an extended calculus
-  that admits more conversion equations upon elimination of the extensible
-  inductive type; (b). and a proper proof assistant that supports native
-  extensible inductive types instead of the current encoding via our plugin,
-  thus allowing for proper reduction of open terms. This will change the kernel
-  of the proof assistant for sure. 
-  ...
-  In the final revision, we will also clarify that our plugin only translates
-  code into vanilla Coq terms without expanding the trusted codebase.
-  -->
-
 * (A, B, D) **The presentation of FMLTT is tough. The Agda-looking scripts are not readable.**
@@
   explain the formalization and proofs in greater detail.
-
-  <!--
-  Thank you for your feedback. 
-  We sympathize with this impression in Section 5, and we agree it is dense, hard to read and not giving too much inspiration to the readers.
-  ...
-  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater detail aiming for accessibility. 
-  -->
 
@@
   interpretations.
-
-  <!--
-  Our main excuse for doing the proofs is that we found it more
-  educational for ourselves to directly prove FMLTT consistent than doing the
-  translation.
-  ...
-  We decided to prove consistency and canonicity directly for FMLTT, because our
-  formulation of W-types, which are part of both FMLTT and our target MLTT, is
-  slightly unconventional.
-  We define a W-type as given by a list of pairs of types $A_i ⊢ B_i$; each pair
-  corresponds to one constructor.
-  The convention is to define a W-type as given by a single pair of types $A ⊢ B$.
-  The conventional and unconventional Wtype can be translated
-  to each other because a list is also an inductive type.
-  So the conventional and unconventional Wtype can be
-  translated to each other. 
-  ...
-  However, our translation doesn't translate this unconventional Wtype
-  formulation back to the conventional one. So the target calculus after
-  translation is **MLTT + this unconventional Wtype**. In that case, translation
-  can only function as a guide to the plugin implementation, and
-  **pedantically** is not enough to show the consistency/canonicity for FMLTT
-  because, **pedantically**, target calculus is not proven to be
-  consistency/canonicity.
-  ...
-  We agree with Reviewer B's insight -- we will get consistency/canonicity when
-  we translate the unconventional Wtype into the conventional one. In fact, our consistency 
-  proof is structurally identical to the  type-preservation proof for a translation to MLTT, 
-  since consistency is proved via a type-preserving, metacircular interpretation.
-  ...
-  The reason we
-  didn't choose to do so is that we expect this translation a lot more verbose
-  than the current proof because of the simplicity of Wtype itself compared to
-  the rich functionality provided by (fake-)Agda's Inductive Facility. We only
-  use the latter when constructing the consistency/canonicity model.
-  -->
 
@@
 
-  <!--
-  Generally speaking, compared to ML-style modules, families focus on overridability and code inheritance.
-The latter can be modelled by the functor in a verbose way.
-Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.
-  ...
-  Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience.
-For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation).
-Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phases is acceptable.
-  ...
-  In fact, for plugin development, the family is compiled into modules/functors.
-In meta-theory, we use sigma type as a conceptually simpler representation of modules.
-We also use singleton type to expose concrete type information in a family inspired by their work.
-So our work is heavily influenced by ML-module.
-  ...
-  Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008].
-However, compared to their work, we mainly focus on the prospect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors.
-Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.
-  -->
-
 * (B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__
@@
 
-  <!--
-  In practice, this A is decided by implementation.
-Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".
-  ...
-  Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A`
-  has `tm : 𝕌` instead. `s` will make sure other fields stay the same type. To
-  show this explicitly at the plugin level, we look at Figure 4 (Line 494)
-  `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding
-  to `tm : 𝕌`). With this interface `STLC°tm` we cannot pattern match any term
-  of type `tm : 𝕌` thus doing abstraction successfully.
-  ...
-  Generally speaking, all the (extensible) inductive type will be simply
-  "wrapped" by a module type only exposing the
-  constructor (with no eliminators), just like how we generate `STLC°tm`.
-  -->
-
 * (B) __What happens if the programmer tries to mix contradictory language features?__
@@
   STLC.
-
-  <!--
-  We currently don't support extending STLC with polymorphism and reference, which requires extending the existent inductive family with new indices. 
-  ...
-  Hypothetically speaking, we can expect our plugin will generate unprovable proof obligation under mixin, hindering qeding the proposition and thus closing the family.
-  ...
-  Another example would be extending *a family of STLC and its termination proof* with the general recursion feature.
-Our plugin will generate an unprovable proof obligation inside the reducibility argument for the fixpoint feature.
-  -->
 

### 2023-02-08 pldi2023/review-reply-final.md
tweaks

@@
 We thank all the reviewers for their critical and encouraging feedback.
-Below we address the main concerns, which might be paraphrased for space.
+Below we address the main concerns, which we may have paraphrased for space.
 
@@
     FDefinition b := ...
-  End B
+  End B.
   ```
 
-* (D) __Line 487 (__*a module named STLC◦subst◦Cases is generated interactively ...*__) is confusing.__
+* (D) __Line 487__ ("a module named `STLC◦subst◦Cases` is generated interactively ...") __is confusing.__
 
@@
   modules—under the hood, upon closing a family, the plugin is just doing
-  repeated module inclusion and instantiation.
+  module inclusion and instantiation repeatedly.
 

### 2023-02-09 pldi2023/review-reply-final.md
tweaks
edit

@@
 We thank all the reviewers for their critical and encouraging feedback.
-Below we address the main concerns, which we may have paraphrased for space.
+Below we address the main concerns, which we may have paraphrased for brevity.
 
@@
 
-* (A, B, D) **The presentation of FMLTT is tough. The Agda-looking scripts are not readable.**
+* (A, B, D) **The presentation of FMLTT is tough.**
 
@@
   We note that the effort required to prove consistency directly for FMLTT is
-  similar to that required to prove a translation to MLTT
-  type-preserving: both proofs are structured as type-preserving, metacircular
-  interpretations.
+  similar to that required to prove type preservation for a translation to MLTT:
+  both proofs are structured as type-preserving interpretations.
 

### 2023-02-09 pldi2023/review-reply-final.md
Tweak.

@@
   We note that the effort required to prove consistency directly for FMLTT is
-  similar to that required to prove type preservation for a translation to MLTT:
-  both proofs are structured as type-preserving interpretations.
+  similar to that required to prove a translation to MLTT
+  type-preserving: both proofs are structured as type-preserving interpretations
+  in another dependent type theory.
 

### 2023-02-09 pldi2023/review-reply-final.md
into another TT

--- 2023-02-09 pldi2023/review-reply-final.md
@@
   similar to that required to prove type preservation for a translation to MLTT:
-  both proofs are structured as type-preserving interpretations.
+  both proofs are structured as type-preserving interpretations
+  into another dependent type theory.
 

### 2023-02-09 metatheory-condensed.tex
switch citations to author-year format

@@
 %\underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory (MLTT)~\cite{martin1982constructive}
-Martin-Löf dependent type theory (MLTT)~\cite{martin1982constructive}
+\citeauthor{martin1982constructive} dependent type theory (MLTT)
 with facilities to express family polymorphism
@@
 %
-As is the current trend~\cite{coquand2018canonicity,altkap2016,gratzer-sterling-birkedal-2019}
-with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
+Following the current trend~\cite{coquand2018canonicity,altkap2016,gratzer-sterling-birkedal-2019}
+in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
 substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part

### 2023-02-28 body.tex
move appendices to tr
ref to appendices
small edits
ack

@@
 
+\ifacknowledgments
+\section*{Acknowledgments}
+We thank the reviewers for their many useful comments.  We thank
+Anastasiya Kravchuk-Kirilyuk, Yao Li, and Andrew Myers % others?
+for valuable discussions.
+This work was supported in part by an Amazon research award on automated
+reasoning.
+The views and opinions expressed are those of the authors and do not necessarily
+reflect the position of any funding agency.
+\fi
+
 \setlength{\bibsep}{.8ex}
@@
 \bibliography{refs.bib}
+
+\ifreport
 
@@
 \section{More Typing Rules for \TT}
+\label{sec:more-rules}
 \input{all-typing-rule.tex}
@@
 \section{Using \TT's Linkage Transformers to Model a Derived Family}
+\label{sec:lt-in-action}
 \input{stlc-linkTran-typing.tex}
@@
 
-\section{The Complete Formalization and Proofs Written in Agda Syntax}
+\section{Definitive Version of the Formalization and Proofs}
 \label{sec:agda-files}
@@
 
-\if 0
-\section{Metatheory, in detail}
-\label{sec:metatheory}
-\input{metatheory.tex}
-
-\section{Example in Metatheory}
-\label{sec:coqexample}
-\input{coq-example.tex}
+%\section{Metatheory, in detail}
+%\label{sec:metatheory}
+%\input{metatheory.tex}
+%
+%\section{Example in Metatheory}
+%\label{sec:coqexample}
+%\input{coq-example.tex}
 

### 2023-02-28 introduction.tex
tweaks

@@
 
-In the Coq proof assistant~\cite{coq}, for instance, inductive types, as well as functions and
+In the \citeauthor{coq} proof assistant, for instance, inductive types, as well as functions and
 theorems over inductive types, are closed to extension.

### 2023-02-28 metatheory-condensed.tex
tweaks

@@
 We acknowledge that the presentation herein is dense for an audience
-without prior exposure to MLTT, so we summarize the salient points
+unacquainted with MLTT or explicit substitutions, so we summarize the salient points
 first before delving into the technical details.
@@
 %
-Following the current trend~\cite{coquand2018canonicity,altkap2016,gratzer-sterling-birkedal-2019}
+Following the current trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
 in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:

### 2023-03-02 introduction.tex
promise tool availability
CCS concepts and key words

@@
 in the new setting of proof engineering:
-the Expression Problem (EP) \cite{wadler-ep}.
+the \emph{expression problem} (EP) \cite{wadler-ep}.
 
@@
 \item We contribute a prototypical implementation of our language
-mechanism as a Coq plugin (\cref{sec:coqimpl}). The plugin works by
+mechanism as a Coq plugin (\cref{sec:coqimpl}),
+available at \url{https://anonymized-link-to-code}.
+The plugin works by
 compiling surface-language definitions into Coq modules parameterized by

### 2023-03-03 related-work.tex
add para on ML modules

@@
 Our approach is the first that applies family polymorphism~\cite{ernst2001family},
-an idea from the context of OO languages,
+an idea from OO research,
 to the context of mechanized proofs.
@@
 \citet{bac2006} address the expression problem for a core subset of Standard~ML
-by combining explicitly coded open recursion with a mechanism that
+by combining explicitly coded open recursion with a design that
 allows pattern-matching cases to be defined separately and combined
@@
 bestowing organizational power.
+
+\noindentparagraph{ML-style modules}\!\!, like families, are a modularity mechanism,
+but with a focus on abstraction rather than extensibility.
+Both \TT and the module system of \citet{stohar2000} use singletons to model and
+control the propagation of definitions.
+\mbox{MixML} \cite{rosdre2013} integrates mixins into ML and handles the
+idiosyncrasies of ML modules,
+while our work supports mixins in the presence of extensible inductive types.
 

### 2023-03-03 metatheory-condensed.tex
Make things more concise.

@@
 \item
-\TT also features \emph{linkage transformers} modeling the inductive
-construction of families from other families.
-They allow a linkage to be created by inheriting or
-overriding components of an existing linkage, or by extending an existing
-linkage with new components.
-
+\TT also features \emph{linkage transformers}, used to model the features like inheritance, extending and overriding components in \Lang .
 \item
@@
 
+% TODO: Remove mentioning conventional Wtype here
+
 W-types~\cite{martin1984intuitionistic} are a succinct way to model
@@
 $i$-th pair in~$\wsig$, and $\wcode{\wsig}$ giving the code of the W-type (\ruleref{tm/w}).
+
+% Add example for representing tm here
 

### 2023-03-03 metatheory-condensed.tex
wording

@@
 \item
-\TT also features \emph{linkage transformers}, used to model the features like inheritance, extending and overriding components in \Lang .
+\TT features \emph{linkage transformers}, which model how a family can be
+inductively constructed by inheriting fields from another family,
+adding new fields, or overriding existing fields.
 \item

### 2023-03-03 introduction.tex
kill two orphans

@@
 Language mechanisms including
+virtual classes~\cite{virtualclasses-1989,vc-calculus-2006},
 mixins~\cite{mixin-1990},
-virtual classes~\cite{virtualclasses-1989,vc-calculus-2006},
 virtual types~\cite{thorup97}, % and associated types~\cite{ckj05},
-extensible cases~\cite{bac2006}, etc.\ 
-are all based on this essential idea of inheritance.
+and extensible cases~\cite{bac2006}
+are based on this essential idea of inheritance.
 %

### 2023-03-03 metatheory-condensed.tex
kill two orphans

--- 2023-03-03 metatheory-condensed.tex
@@
 Most notably, it extends MLTT with what we call \emph{linkages}.
-(Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
+(Linkages are a namesake of the theoretical device through which \citet{zm2017} model
 family polymorphism in an OO setting, but the technical details differ significantly from the prior work.)
@@
 \item
-\TT also features \emph{linkage transformers}, used to model the features like inheritance, extending and overriding components in \Lang .
+\TT features \emph{linkage transformers}, which model how a family can be
+inductively constructed by inheriting fields from another family,
+adding new fields, and overriding existing fields.
 \item

### 2023-03-04 challenge.tex
tweak
shortening in §1–2

@@
 Integrating family polymorphism into a proof assistant presents challenges
-far beyond those found in an object-oriented setting~\cite{ncm2004,vc-calculus-2006,zm2017},
-because the underlying programming language is simultaneously
+far beyond those found in an object-oriented setting, %\cite{ncm2004,vc-calculus-2006,zm2017},
+as the underlying programming language is simultaneously
 functional, dependently typed, a logic, and an interactive tool.
@@
 The tension reflects a duality between variants and records.
-For record-like language constructs (e.g., classes and objects), it is
-safe for an extension to add more fields: existing fields can still be
-projected.
+With record-like language constructs (e.g., OO classes),
+it is safe for an extension to add new fields:
+existing fields can still be projected.
 But variant-like constructs (e.g., inductive types) do not automatically
 enjoy safe, modular addition of constructors:
-existing pattern matches could potentially become non-exhaustive.}
+existing pattern matches could become non-exhaustive.}
 In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
@@
 even in the middle of a family definition.
-Last but not least, in addition to producing theorems,
-it should be possible for terms defined in families to possess computational content.
+Last but not least, in addition to proving theorems,
+it should be possible for terms defined with families to possess computational content.
 

### 2023-03-04 casestudies.tex
Change according to Reviewer Please Review!!!

@@
 for simple imperative languages.
-In addition to the metatheories, this case study produces abstract
+In addition to the proof of its soundness, this case study produces abstract
 interpreters that are directly ready for program extraction.
@@
 ready for program extraction to OCaml.
-We put them to test over queries.
+We put the verified abstract interpreters to test over several simple queries. % Reviewer D : l899
 
@@
 inductive types for modeling extensible context-free grammars and derive
-decision procedures for them.
+decision procedures for language membership checking.
 %\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}

### 2023-03-04 challenge.tex
Change according to Reviewer Please Review!!!

@@
 \headinglabel{C1}{chg:extensible-v-exhaustive}.\,
-Extensible inductive types vs.\ exhaustive induction reasoning.%
+Extensible inductive types vs.\ exhaustive inductive reasoning.% Reviewer D
 }
@@
 However, there is a tension between extensibility of inductive types
-and exhaustivity of induction reasoning.\footnote{%
+and exhaustivity of inductive reasoning.\footnote{%
 The tension reflects a duality between variants and records.
@@
 Family polymorphism enables modular reuse via late binding:
-code of a base family can be reused by a derived family,
+the code of a base family can be reused by a derived family,
 because fields referenced by that code have meanings polymorphic to the
@@
 In \cref{fig:stlc-nonmechanized}, a proof assistant supporting family
-polymorphism cannot equate references to the substitution function and
-its definition (i.e., a pattern match against four cases),
+polymorphism cannot unfold references to the substitution function into its definition (i.e., a pattern match against four cases), %Review D : l135
 as a derived family may modify the definition of the function by adding

### 2023-03-04 coqimpl.tex
Change according to Reviewer Please Review!!!

@@
 First, a module named \lsti{STLC\_subst\_Cases} is generated interactively:
-every time the programmer completes a \lsti{Case}, a case handler (e.g., \lsti{Def subst\_tm_unit})
-is generated and added to the module.
+every time the programmer completes a \lsti{Case}, the case handler (e.g., \lsti{Def subst\_tm_unit})
+is type-checked and added to the module. % Reviewer D: l487
 

### 2023-03-04 introduction.tex
Change according to Reviewer Please Review!!!

@@
 supply of linguistic features that offer expressive power to solve the
-EP.
+EP.\EDJ{Reviewer D wants citation here}
 

### 2023-03-04 metatheory-condensed.tex
Change according to Reviewer Please Review!!!

@@
 
-We use Tarski-style universes~\cite{hofmann1997syntax}: a universe~$\cU$
-is inhabited by the \emph{codes} of types,
+In dependent type system, types are treated as terms, and should have a type. % Reviewer A : Why is W(t) a term and not a type?
+We use Tarski-style universes~\cite{hofmann1997syntax}: every type has a \emph{code}, inhabits in a universe~$\cU$,
 with $\El{t}$ decoding~$t$ (rule \ruleref{ty/el})
@@
 $\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
-$\Wsup{\wsig}{t_1}{t_2}$ constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
+we can constructs a term $\Wsup{\wsig}{t_1}{t_2}$ of type~$T$ (rule \ruleref{tm/wsup}). % Reviewer B: Line 666
 As the rules suggest, our formulation additionally supports

### 2023-03-04 related-work.tex
Change according to Reviewer Please Review!!!

@@
 from feature modules. Feature composition is manual, which seems to have
-motivated later approaches based on DTC.
-
-The design pattern \emph{data types à la carte} (DTC) introduced for Haskell by~\citet{dtc}
-is a source of inspiration for many approaches
+motivated later approaches based on \emph{data types à la carte} (DTC)~\cite{dtc}. % Reviewer B : Line 910
+DTC has in turn served as a source of inspiration for numerous work
 \cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.

### 2023-03-04 introduction.tex
Add a comment macro for each reviewer

--- 2023-03-03 introduction.tex
@@
 supply of linguistic features that offer expressive power to solve the
-EP.
+EP.\rD{There is no shortage of solutions to the EP in various
+languages, but there are no citations to those works.}
 

### 2023-03-04 casestudies.tex
All the stuff from reviewer.

@@
 
-Our implementation of family polymorphism retains the computation
-ability of the host proof assistant;
+Our implementation of family polymorphism retains the
+ability of program extraction in Coq;
 the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are

### 2023-03-04 coqimpl.tex
All the stuff from reviewer.

@@
 with the help of \lsti{tm_rect}, the recursor Coq generates for \lsti{tm}.
-Computational behaviors of \lsti{tm_prect_STLC} are immediate, by \lsti{reflexivity}.
+The computational behaviors of \lsti{tm_prect_STLC} are immediate, by \lsti{reflexivity}.
 
@@
 to the (already included) case handlers. %(\lsti{subst\_tm_unit}, etc.).
-Computational behaviors of \lsti{subst} are then immediate, by \lsti{reflexivity}.
+The computational behaviors of \lsti{subst} are then immediate, by \lsti{reflexivity}.
 

### 2023-03-04 lang-design.tex
Change according to Reviewer Please Review!!!
All the stuff from reviewer.

@@
 fixpoints can be mechanized using \Lang, in a style envisioned in
-\cref{fig:stlc-nonmechanized}.
+\cref{fig:stlc-nonmechanized}, directly adopted from \citet{sf-pl}. % Review B.244 - mention non-capture avoiding, SF is using non-capture avoiding.
 The base family "STLC" hosts the STLC metatheories, from abstract syntax
@@
 
-In family "STLCFix", inductive type "tm" \emph{further binds} the "tm" type
+In family "STLCFix", the inductive type "tm" \emph{further binds} the "tm" type
 in family "STLC". It has five constructors: four inherited from
@@
 With inductive types made extensible, exhaustivity of induction becomes
-in jeopardy, however.
-In particular, recursor "tm_rect" should no longer be allowed, because
+in jeopardy, however.\EDJ{``Becomes in jeopardy'' looks weird to me. how about changing to "is now in question"}
+In particular, the recursor "tm_rect" should no longer be allowed, because
 its type mentions "tm", whose meaning is late bound, yet
@@
 In the event that \Lang cannot infer where the programmer intends to place a new field,
-annotation is required.
+an annotation is required.
 

### 2023-03-04 metatheory-condensed.tex
All the stuff from reviewer.

@@
 \item 
-Linkages model families; they are like tuples composed of the fields,
+Linkages model families; they are like tuples composed of fields,
 with field names represented by variable bindings.
@@
 
-Well-formedness judgments of \emph{linkage signatures} have form
+Well-formedness judgments of \emph{linkage signatures} have the form
 $\goodSig{\Gm}{\lsig}{n}$.
@@
 %
-Well-formedness judgments of \emph{linkages} have form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
+Well-formedness judgments of \emph{linkages} have the form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
 where $\TyLkg{\lsig}$ is a type formed by~$\lsig$.

### 2023-03-04 lang-design.tex
Definite Article
make all comments into pdf comment from reviewer.

@@
 fixpoints can be mechanized using \Lang, in a style envisioned in
-\cref{fig:stlc-nonmechanized}, directly adopted from \citet{sf-pl}. % Review B.244 - mention non-capture avoiding, SF is using non-capture avoiding.
+\cref{fig:stlc-nonmechanized}, directly adopted from \citet{sf-pl}. \rB{mention non-capture avoiding}\EDJreply{SF is using non-capture avoiding. Is mentioning SF here enough?}
 The base family "STLC" hosts the STLC metatheories, from abstract syntax
@@
 
-In family "STLCFix", the inductive type "tm" \emph{further binds} the "tm" type
+In family "STLCFix", the\rD{l175}\EDJreply{Then do we add definite article in front of family "STLCFix"?} inductive type "tm" \emph{further binds} the "tm" type
 in family "STLC". It has five constructors: four inherited from

### 2023-03-04 introduction.tex
intro

--- 2023-03-03 introduction.tex
@@
 
-The EP is well studied in the conventional setting of functional
+The EP is well studied in the setting of ordinary functional
 programming and object-oriented (OO) programming.
@@
 supply of linguistic features that offer expressive power to solve the
-EP.
+EP. %\rD{There is no shortage of solutions to the EP in various
+%languages, but there are no citations to those works.}
 
-In contrast, proof assistants offer few linguistic solutions that
-address the EP.
-Yet, the challenge of writing extensible, type-safe code is
-as real, especially for metatheory mechanization.
-The programmer faces a tension between adding new constructors to an inductive data type
+In contrast, proof assistants offer few linguistic solutions to the EP.
+But the challenge of writing extensible, type-safe code is
+as real as in any other language.
+Metatheory mechanization epitomizes the difficulty:
+the programmer faces a tension between adding new constructors to an inductive type
 (e.g., new abstract syntax) and adding new functions and theorems over
-the data type (e.g., new meta\-theoretical results).
+the inductive type (e.g., new meta\-theoretical results).
 
@@
 %
-The programmer could also turn to design patterns and encodings~\cite{delaware2011,delaware2013,schwaab2013modular,keuchel2013generic,forsta2020}.
-But they tend to require heavy lifting from the programmer to make code
+The programmer could also turn to design patterns for structuring developments
+and tool support for cutting down on boilerplate
+\cite{delaware2011,delaware2013,schwaab2013modular,keuchel2013generic,forsta2020}.
+%\rD{The Coq à la Carte paper is arguably more than just a design pattern / encoding, given that it extends the surface syntax of the theorem prover.
+The same designation could ostensibly be applied to the plugin described in Sections 3-4.
+The difference, I suppose, is that the extended language in that work does not have a formal treatment.}
+But these solutions tend to require heavy lifting from the programmer to make code
 extensible and often lead to non-idiomatic programming styles.

### 2023-03-04 metatheory-condensed.tex

@@
 
-In dependent type system, types are treated as terms, and should have a type. % Reviewer A : Why is W(t) a term and not a type?
+In dependent type system, types are treated as terms, and should have a type. \rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}
 We use Tarski-style universes~\cite{hofmann1997syntax}: every type has a \emph{code}, inhabits in a universe~$\cU$,
@@
 $\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
-we can constructs a term $\Wsup{\wsig}{t_1}{t_2}$ of type~$T$ (rule \ruleref{tm/wsup}). % Reviewer B: Line 666
+we can constructs a term $\Wsup{\wsig}{t_1}{t_2}$ of type~$T$ (rule \ruleref{tm/wsup}). \rB{for readability please add text between "T," and "Wsup"}\EDJreply{changed, please review}
 As the rules suggest, our formulation additionally supports

### 2023-03-04 casestudies.tex
make all comments into pdf comment from reviewer.
resolve most comments

--- 2022-11-11 casestudies.tex
@@
 for simple imperative languages.
-In addition to the metatheories, this case study produces abstract
+In addition to a soundness proof, this case study produces abstract
 interpreters that are directly ready for program extraction.
@@
 
-Our implementation of family polymorphism retains the computation
-ability of the host proof assistant;
-the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
-ready for program extraction to OCaml.
-We put them to test over queries.
+Our implementation of family polymorphism is compatible with Coq's
+program extraction feature.
+We extract the two verified abstract interpreters to OCaml
+and put them to test over simple queries.
+%\rD{What were the results of testing the extracted abstract interpreters?
+I don't expect much in the way of performance, but did they return the correct results?
+What sorts of benchmarks did you use?}
 
@@
 inductive types for modeling extensible context-free grammars and derive
-decision procedures for them.
+decision procedures for language membership.
 %\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}

### 2023-03-04 challenge.tex
make all comments into pdf comment from reviewer.
resolve most comments

@@
 \headinglabel{C1}{chg:extensible-v-exhaustive}.\,
-Extensible inductive types vs.\ exhaustive inductive reasoning.% Reviewer D
+Extensible inductive types vs.\ exhaustive inductive reasoning.
 }
@@
 In \cref{fig:stlc-nonmechanized}, a proof assistant supporting family
-polymorphism cannot unfold references to the substitution function into its definition (i.e., a pattern match against four cases), %Review D : l135
+polymorphism cannot unfold references to the substitution function into a
+pattern match against four cases,
 as a derived family may modify the definition of the function by adding

### 2023-03-04 coqimpl.tex
make all comments into pdf comment from reviewer.
resolve most comments

@@
 Consider the translation of \lsti{subst} in family \lsti{STLC}.
-First, a module named \lsti{STLC\_subst\_Cases} is generated interactively:
-every time the programmer completes a \lsti{Case}, the case handler (e.g., \lsti{Def subst\_tm_unit})
-is type-checked and added to the module. % Reviewer D: l487
+First, a module named \lsti{STLC\_subst\_Cases} is generated on the fly.
+%every time the programmer completes a \lsti{Case}, the case handler (e.g., \lsti{Def subst\_tm_unit})
+%is type-checked and added to the module.
+%\rD{This is also confusing-- why is the programmer involved in the translation?
+Is it that modules are being generated in the background, while commands are being processed?}\EDJreply{changed to mention "type-checked" explicitly.
+please review}
+Importantly, programming remains interactive, as the programmer need not wait until the
+entire \lsti{FRecursion} definition is completed to have a \lsti{Case} command
+type-checked.
 

### 2023-03-04 lang-design.tex
resolve most comments

--- 2022-11-10 lang-design.tex
@@
 \cref{fig:stlc-nonmechanized}.
+%directly adopted from \citet{sf-pl}.
+\rB{mention non-capture avoiding}\EDJreply{SF is using non-capture avoiding.
+Is mentioning SF here enough?}\YZreply{It's mentioned in Case Studies.
+I don't think we need to mention it here.}
 The base family "STLC" hosts the STLC metatheories, from abstract syntax
@@
 
-In family "STLCFix", inductive type "tm" \emph{further binds} the "tm" type
+In family "STLCFix", %\rD{l175}\EDJreply{Then do we add definite article in front of family "STLCFix"?}
+the \lsti{FInductive} "tm" \emph{further binds} the "tm" type
 in family "STLC". It has five constructors: four inherited from
@@
 
-With inductive types made extensible, exhaustivity of induction becomes
-in jeopardy, however.
-In particular, recursor "tm_rect" should no longer be allowed, because
+With inductive types made extensible, exhaustivity of induction is now
+in question, however.
+In particular, the recursor "tm_rect" should no longer be allowed, because
 its type mentions "tm", whose meaning is late bound, yet

### 2023-03-04 metatheory-condensed.tex
resolve a comment

@@
 %
-A constructor is given by a pair of types
-$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
-given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
-$\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
-we can constructs a term $\Wsup{\wsig}{t_1}{t_2}$ of type~$T$ (rule \ruleref{tm/wsup}). \rB{for readability please add text between "T," and "Wsup"}\EDJreply{changed, please review}
-As the rules suggest, our formulation additionally supports
-\emph{W-type signatures}.
-A well formed W-type signature \goodWSig{\Gm}{\wsig}{n} is
-composed of $n$ pairs of types, each of which represents a constructor
-(rules \ruleref{wsig/empty} and \ruleref{wsig/add}),
-with $\wsigproj{i}{1}{\wsig}$ and $\wsigproj{i}{2}{\wsig}$ giving the
-$i$-th pair in~$\wsig$, and $\wcode{\wsig}$ giving the code of the W-type (\ruleref{tm/w}).
-
-% Add example for representing tm here
+In the literature,
+a W-type is given by a pair of types $\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$,
+but our formulation of W-types additionally
+allows a W-type signature \goodWSig{\Gm}{\wsig}{n} to
+specify $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
+Each pair models a constructor in an inductive type,
+with $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$
+and $\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$
+identifying the $i$-th constructor, and $\wcode{\wsig}$ giving the code of the
+W-type (\ruleref{tm/w}).
+%
+Rule~\ruleref{tm/wsup} states that 
+given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
+$\goodTerm{\Gm,\sub{\wsigproj{i}{2}{\wsig}}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{\El{\wcode{\wsig}}}$,
+we can construct a term $\Wsup{\wsig}{t_1}{t_2}$ of the W-type.
 

### 2023-03-04 metatheory-condensed.tex
reply to a comment

@@
 
-In dependent type system, types are treated as terms, and should have a type. \rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}
-We use Tarski-style universes~\cite{hofmann1997syntax}: every type has a \emph{code}, inhabits in a universe~$\cU$,
-with $\El{t}$ decoding~$t$ (rule \ruleref{ty/el})
-and $\codety{T}$ encoding type~$T$ (rule \ruleref{tm/c}).
+%In dependent type system, types are treated as terms, and should have a type.
+We use Tarski-style universes~\cite{hofmann1997syntax}:
+a universe~$\cU$ is not inhabited by types directly,
+but rather by the \emph{codes} of types:
+$\codety{T}$ encodes type~$T$ (rule \ruleref{tm/c}),
+and $\El{t}$ decodes~$t$ (rule \ruleref{ty/el}).
 MLTT has an infinite hierarchy of universes;
 we omit universe levels in the presentation for conciseness.
+\rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}
 

### 2023-03-04 metatheory-condensed.tex
Changing W-type section. Need to add an example now.

@@
 %
-Following the current trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
+In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}\YZ{Does coquand2018canonicity really use explicit subst?}
 in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
@@
 %
-In the literature,
-a W-type is given by a pair of types $\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$,
-but our formulation of W-types additionally
-allows a W-type signature \goodWSig{\Gm}{\wsig}{n} to
-specify $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
-Each pair models a constructor in an inductive type,
-with $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$
-and $\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$
-identifying the $i$-th constructor, and $\wcode{\wsig}$ giving the code of the
+Our enhanced W-type is decided by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
+composed of a list (\ruleref{wsig/empty} and \ruleref{wsig/add}) of $n$ pairs of 
+types $\{(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})\}_i$,
+modeling an inductive type of $n$ constructors. 
+%
+Rule~\ruleref{tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
+$\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$ models the $i$-th constructor---given 
+two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
+$\goodTerm{\Gm,\sub{\wsigproj{i}{2}{\wsig}}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{\El{\wcode{\wsig}}}$,
+we can construct a term $\Wsup{\wsig}{t_1}{t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
 W-type (\ruleref{tm/w}).
-%
-Rule~\ruleref{tm/wsup} states that 
-given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
-$\goodTerm{\Gm,\sub{\wsigproj{i}{2}{\wsig}}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{\El{\wcode{\wsig}}}$,
-we can construct a term $\Wsup{\wsig}{t_1}{t_2}$ of the W-type.
 
@@
 to avoid introducing non-dependent $n$-tuples, which linkages generalize.
+
+\EDJ{Add a concrete example of Inductive type ``tm'' here}
 

### 2023-03-04 metatheory-condensed.tex
minor

--- 2023-03-04 metatheory-condensed.tex
@@
 %
-Following the current trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
+In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
 in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:

### 2023-03-04 related-work.tex
make all comments into pdf comment from reviewer.
resolve most comments
minor

@@
 from feature modules. Feature composition is manual, which seems to have
-motivated later approaches based on \emph{data types à la carte} (DTC)~\cite{dtc}. % Reviewer B : Line 910
-DTC has in turn served as a source of inspiration for numerous work
-\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
-%\cite{vdr2022intrinsically}
+motivated later approaches based on \emph{data types à la carte} (DTC)~\cite{dtc}.
+%\rB{ define abbreviation DTC at first use?}
+
+%DTC is as a source of inspiration for may approaches
+%\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
 The original DTC encoding requires type-level general recursion that
@@
 for the underlying representation of type-level fixed points and
-avoid \lsti{Set} impredicativity.
+avoid \lsti{Set} impredicativity.\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}%\cite{vdr2022intrinsically}
 

### 2023-03-04 metatheory-condensed.tex
Reply the comments

--- 2023-03-04 metatheory-condensed.tex
@@
 %
-In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}\YZ{Does coquand2018canonicity really use explicit subst?}
+In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}\YZ{Does coquand2018canonicity really use explicit subst?}\EDJreply{Yes it is.
+Look at Section 2.1.
+When he starts reasoning, he actually deals with category with family (instead of the typing rules), where he consider morphism between contexts, which is substitution.
+But I agree it can be a bit ambiguous to see (because he literally show two version of MLTT without connecting them---one classical typing rule version, the other directly works on Category with Family (CwF)) but since his proof is mainly on Category with Family, so it counts as explicit substitution.
+But if you are concerned, we can change to Andras Kovacs "Generalized Universe Hierarchies and First-Class Universe Levels".
+or Andreas Abel's "Decidability of Conversion for Type Theory in Type Theory".
+The former one directly work on CwF, the latter one include substitution as part of the raw syntax.}
 in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:

### 2023-03-04 related-work.tex
Reply the comments

@@
 for the underlying representation of type-level fixed points and
-avoid \lsti{Set} impredicativity.\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}%\cite{vdr2022intrinsically}
+avoid \lsti{Set} impredicativity.\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}\EDJreply{My understanding is,
+ it is really an intrinsic typed version of MTC.
+It uses different set of tools (like containers (a la altenkirch, can be considered as an alternative of W-type) and stuff) but it is still a design pattern paper.
+ But good news is, container is closer to Wtype(inductive type) than the endofunctor encoding (classic MTC approach).}
+%\cite{vdr2022intrinsically}
 

### 2023-03-04 related-work.tex
One sentence about Intrinsically typed.

@@
 It uses different set of tools (like containers (a la altenkirch, can be considered as an alternative of W-type) and stuff) but it is still a design pattern paper.
- But good news is, container is closer to Wtype(inductive type) than the endofunctor encoding (classic MTC approach).}
+ But good news is, container is closer to Wtype(inductive type) than the endofunctor encoding (classic MTC approach)}\EDJreply{I also add one sentence after Coq a la carte.
+You can also check...
+I will add bib later.
+Interestingly, Intrinsic type doesn't refer Coq a la carte.}
 %\cite{vdr2022intrinsically}
@@
 individually developed feature extensions have to be composed separately
-by invoking the tool.
+by invoking the tool. Alternatively, \citet{vdr2022intrinsically} uses 
+containers and signature to mechanize intrinsically-typed target language. 
+Combined with Agda's advanced unification mechanism, the authors get rid of 
+a large number of boilerplate code. 
 %

### 2023-03-04 related-work.tex
Rephrasing, less confusing.

@@
 individually developed feature extensions have to be composed separately
-by invoking the tool. Alternatively, \citet{vdr2022intrinsically} uses 
-containers and signature to mechanize intrinsically-typed target language. 
-Combined with Agda's advanced unification mechanism, the authors get rid of 
-a large number of boilerplate code. 
+by invoking the tool. \citet{vdr2022intrinsically} mechanize intrinsically-typed
+target language based on container and signature\cite{morris2009indexed}. 
+Thanks to the nature of intrinsically-typed abstract syntax trees (ASTs) and Agda's advanced unification mechanism, the authors were able to eliminate a significant amount of code that dealt with "impossible" cases.
 %

### 2023-03-05 metatheory-condensed.tex
Adding WSup example Add backup for old tableSTLCbool (tablegenerators.com)

@@
 
-\EDJ{Add a concrete example of Inductive type ``tm'' here}
+We illustrate one concrete example of a "tm" with only two constructors, 
+also as a review for how W-type encodes a given inductive type. The key insight 
+is for a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive
+part of the given constructor stores in $\wsigproj{i}{1}{\wsig}$ and the ``arity'' of the 
+recursive part stores in $\wsigproj{i}{2}{\wsig}$. For example, $"app" : ("tm" \to "tm" \to "tm")$ 
+$\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
+(or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
+Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
+signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit".
+$"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context.
+The key expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict-positivity of Inductive type.
+
+\begin{figure}
+  \begin{tabular}{
+    >{\columncolor[HTML]{C0C0C0}}l lcl}
+    \hline
+    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
+      \multicolumn{1}{l|}{Equivalent Type} &
+      \multicolumn{1}{l}{} &
+      \multicolumn{1}{l|}{} \\
+    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| unit : tm!} &
+      \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
+      $\tau'$ &
+      \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
+    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| app  : tm -> tm -> tm!} &
+      \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
+      $\tau$ &
+      \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
+    \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}\lsti!unit!} &
+      $\coloneqq \Wsup{\tau}{\unit}{"elim-"\bbzero}$ &
+      \cellcolor[HTML]{C0C0C0}\lsti!app unit unit! &
+      $\coloneqq\Wsup{\tau}{\unit}{"unit"}$
+    \end{tabular}
+\end{figure}
 

### 2023-03-05 metatheory-condensed.tex
Fix Wsup - Wsup needs specify the index of the constructor.

@@
 $\goodTerm{\Gm,\sub{\wsigproj{i}{2}{\wsig}}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{\El{\wcode{\wsig}}}$,
-we can construct a term $\Wsup{\wsig}{t_1}{t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
+we can construct a term $\Wsup{i}{\wsig}{t_1}{t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
 W-type (\ruleref{tm/w}).
@@
     \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}\lsti!unit!} &
-      $\coloneqq \Wsup{\tau}{\unit}{"elim-"\bbzero}$ &
+      $\coloneqq \Wsup{1}{\tau}{\unit}{"elim-"\bbzero}$ &
       \cellcolor[HTML]{C0C0C0}\lsti!app unit unit! &
-      $\coloneqq\Wsup{\tau}{\unit}{"unit"}$
+      $\coloneqq\Wsup{0}{\tau}{\unit}{"unit"}$
     \end{tabular}
@@
 In contrast, $\wsig$ should be hidden from the typing context of any
-term that does not invoke $\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
+term that does not invoke $\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
 the term can be reused---without being rechecked---for a different
@@
 Later fields that invoke
-$\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
+$\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
 should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
@@
 The next four fields model the four constructors of~\lsti{tm}.
-Constructor \lsti{tm_unit} is modeled as $\Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}$
+Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}$
 and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
@@
   \\ \goodTerm{\Gm, \sub{\wsigproj{j}{2}{\wsig}}{(id, a)}}{b}{\El{(\wcode{\wsig})}}}
-{\goodTerm{\Gm}{\Wsup{T}{a}{b}}{\El{(\wcode{\wsig})}} }
+{\goodTerm{\Gm}{\Wsup{}{T}{a}{b}}{\El{(\wcode{\wsig})}} }
 $$
@@
 \\ \goodTerm{\Gamma}{T}{\TyWSingle{\tau}}}
-{\goodTerm{\Gamma}{\Wrec{\wsig}{h}{(\Wsup{\wsig}{a}{b})} \equiv 
+{\goodTerm{\Gamma}{\Wrec{\wsig}{h}{(\Wsup{}{\wsig}{a}{b})} \equiv 
 (\app{((\app{(\Recproj{j}{h})} )[("id", a)])} )[("id", \lam{(\Wrec{\wsig}{h[\pi_1]}{b})})]

### 2023-03-05 metatheory-condensed.tex
Add Caption.

@@
 
-We illustrate one concrete example of a "tm" with only two constructors, 
-also as a review for how W-type encodes a given inductive type. The key insight 
-is for a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive
-part of the given constructor stores in $\wsigproj{i}{1}{\wsig}$ and the ``arity'' of the 
-recursive part stores in $\wsigproj{i}{2}{\wsig}$. For example, $"app" : ("tm" \to "tm" \to "tm")$ 
-$\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
-(or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
-Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
-signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit".
-$"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context.
-The key expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict-positivity of Inductive type.
-
 \begin{figure}
@@
     \end{tabular}
+  \caption{W-type encoding of "tm \{unit, app\}"}
+  \label{fig:Wtype-example}
 \end{figure}
+
+We illustrate one concrete example in \cref{fig:Wtype-example} of a "tm" with only two constructors, 
+also as a review for how W-type encodes a given inductive type. The key insight 
+is for a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive
+part of the given constructor stores in $\wsigproj{i}{1}{\wsig}$ and the ``arity'' of the 
+recursive part stores in $\wsigproj{i}{2}{\wsig}$. For example, $"app" : ("tm" \to "tm" \to "tm")$ 
+$\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
+(or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
+Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
+signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit".
+$"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context.
+The expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict-positivity of Inductive type.
 

### 2023-03-05 metatheory-condensed.tex
Fix typo.

@@
 also as a review for how W-type encodes a given inductive type. The key insight 
-is for a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive
-part of the given constructor stores in $\wsigproj{i}{1}{\wsig}$ and the ``arity'' of the 
-recursive part stores in $\wsigproj{i}{2}{\wsig}$. For example, $"app" : ("tm" \to "tm" \to "tm")$ 
+is, to encode one constructor into a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive part will be stored in $\wsigproj{i}{1}{\wsig}$ 
+while the ``arity'' of the  recursive part stores in $\wsigproj{i}{2}{\wsig}$. 
+For example, $"app" : ("tm" \to "tm" \to "tm")$ 
 $\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
@@
 $"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context.
-The expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict-positivity of Inductive type.
+Generally speaking, the expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict positivity of Inductive type.
 

### 2023-03-05 metatheory-condensed.tex
Reply more. Add one bib.

@@
 We use Tarski-style universes~\cite{hofmann1997syntax}:
-a universe~$\cU$ is not inhabited by types directly,
+a universe~$\cU$ is not inhabited by types directly\footnote{On the contrary, 
+Russel-style Universe is inhabited by types directly. 
+Tarski-style universe is ``semantically more fundamental''\cite{luo2012notes} thus more friendly for meta-theoretical reasoning.},
 but rather by the \emph{codes} of types:
@@
 we omit universe levels in the presentation for conciseness.
-\rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}
+\rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}\EDJ{changed again, please review.}
 

### 2023-03-05 metatheory-condensed.tex
Start working on appendices.

@@
 Russel-style Universe is inhabited by types directly. 
-Tarski-style universe is ``semantically more fundamental''\cite{luo2012notes} thus more friendly for meta-theoretical reasoning.},
+Tarski-style universe is ``semantically more fundamental''\cite{luo2012notes} and more friendly for our meta-theoretical reasoning.},
 but rather by the \emph{codes} of types:
@@
 present it in
-\ifreport \cref{sec:more-rules}.
+\ifreport \cref{sec:complete-fmltt}.
 \else a supplemental appendix.

### 2023-03-05 metatheory2.tex
Start working on appendices.

In this section, we start by introducing the complete formulation of the syntax of \TT, 
and then we detail the proofs for syntactic translation (that translate away linkages), 
consistency and canonicity.

\subsection{Warm Up: Review Declarative MLTT}
We start with introduction to MLTT as \TT can be roughly considered as MLTT with linkages.

\input{syntax}

% Copy and Paste Intuition of MLTT from other places to here
\subsection{\TT : MLTT with Linkages}

% Copy and Paste Intuition of FMLTT from other places to here

% Deal with intuition

\subsection{Syntactic Translation get rid of Linkages}

\input{syn-translate.tex}

\subsection{Standard Model for Consistency}

\input{standardmodel.tex}

\subsection{Proof Relevant Logical Relation for Canonicity}
\input{canonicitymodel.tex}

### 2023-03-05 syn-translate.tex
Change makefile

@@
-The implementation described in \cref{sec:coqimpl} can be formalized in
-our metatheory. The described compilation strategy can be considered as
-a \textit{syntactic translation}, a function mapping from our syntactic
-(QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only
+The implementation described in \cref{sec:coqimpl} has hinted the fact that
+linkage can be translated into dependent tuple type.
+The described compilation strategy can be considered as
+a \textit{type preserving syntactic translation}, a function mapping from our syntactic
+model to a ``subpart''\footnote{Not exactly subpart---the only
 difference lies in "Wrec". In our metatheory, it is constructed
 using linkage; in ``subpart'', it will be using $\Sigma$-types (translation
-of the linkage).} of our syntactic (QIIT-)model \textit{without
-linkages}.\footnote{Here, syntactic QIIT-model can be roughly understood
-as just AST, with equational rules acting like quotient. So this mapping
-is mapping AST to AST, respecting quotient/judgemental equality.} This
-``subpart'' is the standard MLTT with inductive type signatures, an
+of the linkage).} of our syntactic model \textit{without
+linkages}.
+
+This ``subpart'' is the standard MLTT with an
 (unconventional) formulation of W-types,
@@
 
-We use $\denotesT{}$ as the translation.
-This translation behaves like identity\YZ{What does "invariant" mean?}\EDJreply{behaves like identity.
-Sorry I should have said "subpart is invariant under this translation" instead}
+We use $\denotesT{}$ as the translation. This translation behaves like identity
+\YZ{What does "invariant" mean?}\EDJreply{behaves like identity. Sorry I should have said "subpart is invariant under this translation" instead}
 under the subpart. Since we are in an intrinsic setting, everything here
@@
 (i.e. every type is tracking its context). In other words, every
-derivation of any judgement will track its ``context''. This will be
-clearer in the (pseudo-)Agda formalization.
+derivation of any judgement will track its ``context''.
 

### 2023-03-05 canonicitymodel.tex
Working on appendices.

@@
 Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
-We can even argue that, if this theorem is proven in a computable meta-logic\footnote{QIIT does have computational content~\cite{altkap2016}}, 
+We can even argue that, if this theorem is proven in a computable meta-logic, 
 then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.\YZ{Can we say something about strong/weak normalization?}\EDJreply{
@@
 
-The proof can be found in appendix.
+% The proof can be found in appendix. 
 The intuition of the model is similar to that of the consistency model---use the canonicity model of sigma type (from the earlier work) for linkage type and using Agda's inductive facility to justify our W-type.

### 2023-03-05 metatheory2.tex
Adding Info on appendices.
Change makefile
Working on appendices.

@@
 \subsection{Warm Up: Review Declarative MLTT}
-We start with introduction to MLTT as \TT can be roughly considered as MLTT with linkages.
+We begin with a review of MLTT as \TT can be roughly considered as MLTT with linkages.
+We work in an intrinsic-typed setting, so there is no un-typed preterms.
 
-\input{syntax}
+\input{all-mltt-typing}
 
 % Copy and Paste Intuition of MLTT from other places to here
-\subsection{\TT : MLTT with Linkages}
+\subsection{\TT : an MLTT variant with Linkages}
+Compared to MLTT, \TT has 3 more judgements about signatures and linkage transformers.
+
+\input{all-fmltt-typing}
 

### 2023-03-05 standardmodel.tex
Working on appendices.

--- 2022-11-01 standardmodel.tex
@@
-\newcommand{\goodCtxS}[2]{{ {#1} \ \models_S }}
-\newcommand{\goodTypeS}[3]{{ {#1} \models_S {#2} }}
-\newcommand{\goodTermS}[3]{{ {#1} \models_S {#2} : {#3} }}
-\newcommand{\goodSubS}[3]{{ {#1} \models_S {#2} : {#3} }}
-\newcommand{\goodSigS}[3]{{ {#1} \models_S {#2} \ \  \texttt{LSig}^{#3} }}
-\newcommand{\goodWSigS}[3]{{ {#1} \models_S {#2} \ \ \texttt{WSig}^{#3} }}
-\newcommand{\goodSealS}[4]{{ {#1} \models_S {#2} : {#3} \  |\  {#4} }}
-\newcommand{\goodInhS}[4]{{ {#1} \models_S {#2} : {#3} \twoheadrightarrow {#4}}}
-
 Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory.
-We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of QIIT.
-However, in this section, for readability, we only inductively interpret each syntax piece into naive set theory that is also respecting judgemental equalities.
-Please refer to our pseduo-Agda formulation for a more formal and complete version.
+We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of Quotient Inductive Inductive Type (QIIT).
+However, in our formulation, we only inductively interpret each syntax piece into naive set theory that is also respecting judgemental equalities.
 
@@
 
-The main intuition of the consistency model is that: (1) since we have used sigma type to translate the linkage, using the model for sigma type for linkage will work; (2) directly using the inductive facility of (pseudo)-Agda (or naive set theory) to justify the intro and elim rule of our W-type.
+The main intuition of the consistency model is that: (1) since we have used sigma type to translate the linkage, using the model for sigma type for linkage will work; (2) directly using the inductive facility of naive set theory to justify the intro and elim rule of our W-type.
 
-Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gm}{t}{T}}$ (i.e.
-every term is tracking its context and type).\footnote{And our (pseudo) Agda formulation is well-typed as well.
-Latex-formulation should be well-typed but we discarded all the type information in the paper.}
+Note that, since every piece is intrinsic well-typed, $\denotesS{t}$ means the same as $\denotesS{\goodTerm{\Gm}{t}{T}}$ (i.e. every term is tracking its context and type).
 
-Here we will only show some interpretation and also omit the universe level to sketch out the idea, please refer to appendix for the complete version.
+Here we will only show the interpretation for  omit the universe level to sketch out the idea. 
+% only show some interpretation and also
 
@@
   \denotesS{\goodCtx{\cdot}{i}} &= \{\star\} \quad \text{ a specific singleton set} \\
-  \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \Gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\Gm)\}
+  \denotesS{\goodCtx{\Gm, A}{i}} &= \{(\Gm, a) : \gm \in \denotesS{\goodCtx{\Gm}{}}, a \in \denotesS{\goodType{\Gm}{A}{}}(\gm)\}
   \\
@@
   \\
-  \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\Gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\Gm) \\
+  \denotesS{\goodTerm{\Gm}{\_}{T}} & = \prod_{\gm \in \denotesS{\goodCtx{\Gm}{i}}}\denotesS{T}(\gm) \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod4}(0.05,-0.15)(-0.10,0.30)
@@
   \\
-  \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\Gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\
+  \denotesS{\goodType{\Gm}{\Sigma A B}{}}(\gm) &:= \sum_{a \in \denotesS{A}\Gm} \denotesS{B}(\Gm, a) \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod5}(0.05,-0.15)(-0.10,0.30)
@@
   \denotesS{\goodType{\Gm}{\cL \sigma}{i}} & \text{ is a family of sets, defined inductively on } \sigma \tikzmarkend{stmod7}  \\
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\Gm) \land t \in \prod_{m' \in \denotesS{A}(\Gm)}\denotesS{T}(\Gm, m')  \}  \\
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\gm)&= \{(m, t) :  m \in \denotesS{\cL \tau}(\gm) \land t \in \prod_{m' \in \denotesS{A}(\gm)}\denotesS{T}(\Gm, m')  \}  \\
   & \quad \text{when } \denotesS{\sigma} = (\denotesS{\tau}, \denotesS{A}, \_, \denotesS{T}) \in \denotesS{\goodSig{\Gm}{\_}{n+1}} \\
-  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\Gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
+  \denotesS{\goodType{\Gm}{\cL \sigma}{i}} (\gm)&= \{\star\} \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
   \denotesS{\goodSig{\Gm}{\LSigEmp}{0}} &= \star \\
   \denotesS{\goodSig{\Gm}{\LSigAdd{\sigma}{s}{T}}{n+1}} &= (\sigma, A, f, T) \\
-  \denotesS{\goodTerm{\Gm}{\LkgEmp}{\cL \LSigEmp} } (\Gm) &= \star \\
-  \denotesS{\goodTerm{\Gm}{\LkgAdd{o}{t}}{\cL(\LSigAdd {\sigma}{s}  {T})}} \ (\Gm)&= (o(\Gm), (\lambda \ m' . t'(\Gm, m'))) \\
-  \denotesS{\goodTerm{\Gm}{\lkgproj{1}{t}}{\cL(\lsigproj{1}{\sigma})}}(\Gm) &=  "fst"~(t\Gm) \\ 
+  \denotesS{\goodTerm{\Gm}{\LkgEmp}{\cL \LSigEmp} } (\gm) &= \star \\
+  \denotesS{\goodTerm{\Gm}{\LkgAdd{o}{t}}{\cL(\LSigAdd {\sigma}{s}  {T})}} \ (\gm)&= (o(\gm), (\lambda \ m' . t'(\Gm, m'))) \\
+  \denotesS{\goodTerm{\Gm}{\lkgproj{1}{t}}{\cL(\lsigproj{1}{\sigma})}}(\gm) &=  "fst"~(t\Gm) \\ 
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod8}(0.05,-0.15)(-0.10,0.30)
@@
   &= \star \quad \text{ when } \denotesS{\sigma} \in \denotesS{\goodSig{\Gm}{\_}{0}} \\
-  \denotesS{\goodType{\Gm}{\bot}{i}} \ (\Gm) &= \emptyset \\
-  \denotesS{\goodType{\Gm}{\cB}{i}} \ (\Gm) &= \{0, 1\} \quad \text{a specific two element set} \\
+  \denotesS{\goodType{\Gm}{\bot}{i}} \ (\gm) &= \emptyset \\
+  \denotesS{\goodType{\Gm}{\cB}{i}} \ (\gm) &= \{0, 1\} \quad \text{a specific two element set} \\
   \tikzmarkin[disable rounded corners=true,set fill color=new,set border color=new]{stmod9}(0.05,-0.15)(-0.10,0.30)
@@
 
-We encourage the reader to read the (pseudo-)Agda-style proof with more type annotation so it is clearer about the mathematical object doing the interpretation.
+% We encourage the reader to read the (pseudo-)Agda-style proof with more type annotation so it is clearer about the mathematical object doing the interpretation. 
 Here, we need to clarify some ambiguity in our latex formulation.
@@
 The case is similar for $\denotesS{\cL \_}$, $\denotesS{\Tkg {\_}}$ and  $\denotesS{\Sigma \ \_ \ \_}$.
-\textbf{This is actually the formulation in Agda}---QIIT syntax consider $\TyTkg{} : "Sig"~\Gm~n \to "Ty"~\Gm$ as a constructor and our model needs to interpret this constructor to a function $\TyTkg{}^S : "Sig"^S \Gm^S~n \to "Ty"^S~\Gm^S$ with each judgement also interpreted.
+% \textbf{This is actually the formulation in Agda}---QIIT syntax consider $\TyTkg{} : "Sig"~\Gm~n \to "Ty"~\Gm$ as a constructor and our model needs to interpret this constructor to a function $\TyTkg{}^S : "Sig"^S \Gm^S~n \to "Ty"^S~\Gm^S$ with each judgement also interpreted.
 
-Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
+% Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps constructor $"Con"$ to $"Con"^S$, and constructor "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc.
 What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-definedness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.
@@
 We sketch the model for the inductive type and inheritance judgement here, and both of them require inductive facility in meta-logic.
-In set theory, that would be using \textit{Rule sets}~\cite{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead.
+In set theory, that would be using \textit{Rule sets}~\cite{timany2017consistency,aczel1998relating}, but here we informally use Agda's Inductive facility instead.
 Both of them are somewhat ``interpreted by themselves'' in Agda, because their syntactic definition is pretty ``well-founded''.
@@
 
-The verification of validity of this model is largely omitted here and we refer the reader to the appendix.
+The verification of validity of this model is largely omitted here.
 To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gm}{\lkgproj{1}{(\LkgAdd{o}{t})}}{\cL\sigma}} = \denotesS{\goodTerm{\Gm}{o}{\cL\sigma}}$ by function extensionality.
@@
 
-Since all the compatibility lemmas hold (by simply changing $\vdash$ into $\models$ in the syntactic rules then we would have all the compatibility lemmas\footnote{Apparently, in our (pseudo) Agda formulation, we don't need this extra step.}), we have constructed a model.
+Since all the compatibility lemmas hold (by simply changing $\vdash$ into $\models$ in the syntactic rules then we would have all the compatibility lemmas), we have constructed a model.
 Finally, notice that we interpret the bottom type using empty set, and thus we know it is not possible to derive $\cdot \vdash t : \bot$: otherwise we have $\cdot \models_S t : \bot$ which exactly means $\denotesS{t}(\star) \in \denotesS{\bot}(\star) = \emptyset$
 and that is a contradiction.\YZ{Explicitly state the theorem.}
-
-The full model formulated in (pseudo) Agda syntax can be found in the appendix.

### 2023-03-05 syn-translate.tex
Working on appendices.

@@
 
-We use the local morphism (the same one for encoding the Sealing
+We use the local morphism\cite{abbott2003category} (the same one for encoding the Sealing
 judgements) to translate inheritance judgement. This is mainly due to
-the existence of \ruleref{Inh} rule. Here we only typeset the overriding
-and inheritance rule, but all of the translation for inh judgment are
-induced by the $\beta$-rule of \ruleref{Inh}.
+the existence of \ruleref{Inh} rule. 
+% Here we only typeset the overriding
+% and inheritance rule, but all of the translation for inh judgment are
+% induced by the $\beta$-rule of \ruleref{Inh}.
 
@@
 $(\cC\sigma)[\tau]\equiv \cC(\sigma[\tau])$ and $\Sigr{\Delta}{n}[\tau]
-\equiv \Sigr{\Gamma}{n}$ together. The details (of all the equational
-rules) can be found in the appendices written in (pseudo-)Agda form. We
-encourage the reader to read that (pseudo-)Agda-style proof because it
-is more friendly to type theorists, easy to do type checking and
-validate the proof and it is clearer on the mathematical object each
-translation uses.
+\equiv \Sigr{\Gamma}{n}$ together. 
+% The details (of all the equational rules) can be found in the appendices written in (pseudo-)Agda form.
+We encourage the reader to read that (pseudo-)Agda-style proof because it is more friendly to type theorists, easy to do type checking and
+% validate the proof and it is clearer on the mathematical object each
+% translation uses.
 
-Ultimately, we have constructed a model for our QIIT-syntax using only a
-``subpart'' of our QIIT-syntax -- without these linkages. It is also
+Ultimately, we have constructed a model for our syntax using only a
+``subpart'' of our syntax -- without these linkages. It is also
 possible to compile our $\goodWSig{\Gamma}{w}{n}$ and W-type into a more
@@
 directly give a consistency and canonicity proof.
-
-% the link for syntactic translation
-% https://drive.google.com/file/d/1pMqn8DS4T4jiCubzk3HgMANjrfKpcGlI/view?usp=sharing

### 2023-03-05 canonicitymodel.tex
Working on appendices.

@@
-We follow the reducibility argument from~\citet{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model.
+We follow the reducibility argument from~\cite{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
+Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax.
+However, our proof will still work directly on the earlier introdcued syntax for readability.
 We will first specify the formulation of the most basic canonicity theorem.
@@
 
-Now we sketch out the big picture of the proof.
-The canonicity model $"Tm"^C~\Gamma^C~"T"^C$ will map each syntax piece $("t : Tm Γ T") \mapsto "(t, tₚ) :" "Tm"^C~\Gamma^C~"T"^C = \sum~("t : Tm Γ T")."Tm"_p~\Gamma^C~"T"^C~"t"$ to a dependent pair\footnote{This $"Tm"_p$ is usually called dependent model.}, of which the first part will be the same as the input syntax.
-Using this mapping on closed boolean term $"t : Tm ⋅ "\mathbb{B}$ will give us a proof of $"Tm"_p~\cdot^C~\mathbb{B}^C~"t"$, which will unfolds to our final goal $"t" \equiv "tt" + "t" \equiv "ff"$.
-Roughly, this can read as ``every closed boolean term is reducible''.
-Thus once we can ``propagate'' $(\cdot)^C$ into every syntactic piece, and construct this model, we are done.
-In our (pseudo) Agda formulation, we use $(\cdot)_2$ to replace the notation of $(\cdot)^C$ because we consider most parts of this model are actually a dependent pair.
-
-Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as ``a reducible set'': $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$.
-The reformulation can be understood as an unfolding of the \textit{type-theoretic logical relation proof}.\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory} If we use the terminology of the conventional formulation, this reducible set means if we apply its element $t$ with a substitution $\gamma$ from the set $\Gamma$ where we have $\Gamma_p$ witnesses the ``reducibility'' of $\Gamma$, then we will have $t[\gamma] \in T_p$ as a closed and \textbf{reducible} term.
-Now if $\Gamma$ is actually empty $\cdot$ and $T$ is $\cB$, we are done.
-
-% $\Gamma_p$ witnesses the set of substitution $ \gamma \in \Gamma$ makes applied term $t[\gamma]$ closed and \textbf{reducible} of types $T_p$.
-Thus, we proved $t[\gamma] \in T_p$---the set of closed reducible term of type $T$---when $\gamma \in \Gamma_p$.
+Like most logical relation proof,  we interpret each judgement and prove compatibility lemma.
 

### 2023-03-05 metatheory2.tex
Working on appendices.

@@
 In this section, we start by introducing the complete formulation of the syntax of \TT, 
-and then we detail the proofs for syntactic translation (that translate away linkages), 
+and then we detail the models and proofs for syntactic translation (that translate away linkages), 
 consistency and canonicity.
@@
 We begin with a review of MLTT as \TT can be roughly considered as MLTT with linkages.
-We work in an intrinsic-typed setting, so there is no un-typed preterms.
+For our syntax, we work in an intrinsic-typed setting, so there is no un-typed preterms 
+(i.e. every terms are well-typed by construction). We introduce judgemental equality $\equiv$
+for derivations (of each judgements). In other words, we will have judgemental equality between terms, 
+contexts and explicits substitutions. 
+We will make sure all the introduced models and proofs respecting judgemental equality.
 

### 2023-03-05 standardmodel.tex
Working on appendices.

@@
 The main goal is to interpret $\denotesS{\bot} = \emptyset$ which implies it is impossible $\goodTermS{\cdot}{t}{\bot}$ and thus $\bot$ is syntactically non-derivable.
-We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$ and that $\denotesS{}$ respecting judgemental equality.\footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo) Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}
+We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$ and that $\denotesS{}$ respecting judgemental equality.
+% \footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo) Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}
 

### 2023-03-05 canonicitymodel.tex
Save. Working on Appendices.

@@
 
-Like most logical relation proof,  we interpret each judgement and prove compatibility lemma.
+Like most logical relation proof,  we interpret each judgement, inductively interpret each syntax piece and prove compatibility lemma.
+
+\begin{align*}
+  \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
+  &\text{we define } \goodCtxS{\Gm}{i} :\iff \denotesS{\goodCtx{\Gm}{i}} \text{ is defined} \\
+  \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
+  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
+  &\text{we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesS{\goodType{\Gm}{T}{i}} \text{ is defined }
+\end{align*}
+
+Now that we construct the above model, we can map an arbitrary closed term $\goodTerm{\cdot}{t}{\cB}$ to get the result
+
+Then we extend the above model to \TT .
 

### 2023-03-05 metatheory-condensed.tex
edit

@@
 We use Tarski-style universes~\cite{hofmann1997syntax}:
-a universe~$\cU$ is not inhabited by types directly\footnote{On the contrary, 
-Russel-style Universe is inhabited by types directly. 
-Tarski-style universe is ``semantically more fundamental''\cite{luo2012notes} and more friendly for our meta-theoretical reasoning.},
+a universe~$\cU$ is not inhabited by types directly\footnote{By contrast, 
+Russel-style universes are inhabited by types directly. 
+Tarski-style universes are ``semantically more fundamental'' \cite{luo2012notes} and friendlier to our meta-theoretical reasoning.},
 but rather by the \emph{codes} of types:

### 2023-03-05 metatheory-condensed.tex
compress a little

@@
 %In dependent type system, types are treated as terms, and should have a type.
-We use Tarski-style universes~\cite{hofmann1997syntax}:
-a universe~$\cU$ is not inhabited by types directly\footnote{By contrast, 
-Russel-style universes are inhabited by types directly. 
-Tarski-style universes are ``semantically more fundamental'' \cite{luo2012notes} and friendlier to our meta-theoretical reasoning.},
-but rather by the \emph{codes} of types:
-$\codety{T}$ encodes type~$T$ (rule \ruleref{tm/c}),
-and $\El{t}$ decodes~$t$ (rule \ruleref{ty/el}).
+We use Tarski-style universes \cite{hofmann1997syntax}.
+Unlike Russell-style universes, they
+are not inhabited by types directly but rather by the \emph{codes} of types
+and arguably behave better \cite{luo2012notes}.
+$\codety{T}$ encodes type~$T$ (rule \ruleref{tm/c}) and thus inhabits~$\cU$.
+$\El{t}$ decodes~$t$ (rule \ruleref{ty/el}).
+%\footnote{By contrast,  Russel-style universes are inhabited by types directly.
+Tarski-style universes are ``semantically more fundamental'' \cite{luo2012notes} and friendlier to our meta-theoretical reasoning.}
 MLTT has an infinite hierarchy of universes;
 we omit universe levels in the presentation for conciseness.
-\rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}\EDJ{changed again, please review.}
+%\rA{Why is W(t) a term and not a type?}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}
 

### 2023-03-06 canonicitymodel.tex
Remove all mentioning about Agda Make sure proof stands alone in latex form.

--- 2023-03-05 canonicitymodel.tex
@@
 
-Like most logical relation proof,  we interpret each judgement and prove compatibility lemma.
+Like most logical relation proof,  we interpret each judgement, inductively interpret each syntax piece and prove compatibility lemma.
+We are working on an intrinsic setting thus even we omit context for brevity, the syntax piece is still well-typed.
+We only specify part of the logical relation proof for MLTT part.
+For the complete version, please refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic}.
 
-% The proof can be found in appendix. 
+\begin{align*}
+  \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
+  &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
+  \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
+  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
+  &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
+  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t.
+} \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
+  &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
+  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t.
+} \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
+  \denotesC{\goodType{\Gm}{\top}{}}(\gm) &= \{t : t \equiv \unit \} \\
+  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
+  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
+  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
+  &= \emptyset \text{ otherwise } \\ 
+  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
+  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
+  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
+  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
+  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
+  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
+  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
+  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &=
+\end{align*}
+
+Then we extend the above model to \TT , we still start with the judgement.
 The intuition of the model is similar to that of the consistency model---use the canonicity model of sigma type (from the earlier work) for linkage type and using Agda's inductive facility to justify our W-type.
+
+Now that we construct the above model, we can map an arbitrary closed term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$.
 

### 2023-03-06 metatheory2.tex
Remove all mentioning about Agda Make sure proof stands alone in latex form.

@@
 
-\subsection{Warm Up: Review Declarative MLTT}
+\subsection{Warm Up: Review Declarative Syntax for MLTT}
 We begin with a review of MLTT as \TT can be roughly considered as MLTT with linkages.
-For our syntax, we work in an intrinsic-typed setting, so there is no un-typed preterms 
-(i.e. every terms are well-typed by construction). We introduce judgemental equality $\equiv$
-for derivations (of each judgements). In other words, we will have judgemental equality between terms, 
+For our syntax, we work in an intrinsic-typed setting, so there is no un-typed preterms appearing anywhere. We introduce judgemental equality ($\equiv$)
+for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
 contexts and explicits substitutions. 
-We will make sure all the introduced models and proofs respecting judgemental equality.
+We will make sure all the introduced models and proofs respecting judgemental equality. In other words, we will consider our syntax as the result of taking quotient by these judgemental equality.
 
@@
 % Copy and Paste Intuition of MLTT from other places to here
-\subsection{\TT : an MLTT variant with Linkages}
+\subsection{Declarative Syntax for \TT : an MLTT variant with Linkages}
 Compared to MLTT, \TT has 3 more judgements about signatures and linkage transformers.

### 2023-03-06 canonicitymodel.tex
Working on canonicity model.

@@
+Now we prove the consistency and canonicity using canoncity model.
 We follow the reducibility argument from~\cite{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
@@
   \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
+  \denotesC{\goodTerm{\Gm}{\unit}{\top}} &= \\
+  \denotesC{\goodTerm{\Gm}{\true}{\cB}} &= \\
+  \denotesC{\goodTerm{\Gm}{\false}{\cB}} &= \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}} &= \\
+  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}} &= \\
   \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
@@
   \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
-  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &=
+  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}} &= \\
+  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
 \end{align*}
@@
 
-Now that we construct the above model, we can map an arbitrary closed term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$.
+The first observation is that, with this model, we know our system is consistent---we cannot derive $\goodTerm{\cdot}{t}{\bot}$.
+Otherwise, we have $\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \denotesS{\bot}(\star) = \emptyset$, that is an element in an empty set, a contradiction.
+
+\begin{theorem}[Consistency]
+  We cannot derive a closed term of bottom type, i.e. ,$\cdot \not \vdash t : \bot$ for arbitrary $t$.
+\end{theorem}
+
+What's more, with the above model, we can map an arbitrary closed term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
 

### 2023-03-06 canonicitymodel.tex
Working on the appendix.

@@
 
-Then we extend the above model to \TT , we still start with the judgement.
-The intuition of the model is similar to that of the consistency model---use the canonicity model of sigma type (from the earlier work) for linkage type and using Agda's inductive facility to justify our W-type.
+Then we extend the above model to \TT , we still start with the judgement. Since linkage transformers are just syntactic sugar, we will not interpret it.
+The intuition of the model is similar to that of the syntactic translation---use the canonicity model of sigma type (from the earlier work) for linkage type and using inductive facility in the meta-logic to justify our W-type.
+
+\begin{align*}
+  \denotesC{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigC{\Gm}{\sigma}{n}} :\iff \denotesC{\sigma} \in  \denotesC{\goodSig{\Gm}{\_}{n}} \\ 
+  \denotesC{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
+  & \text{ and we define } \goodWSigC{\Gm}{\tau}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \in \denotesC{\goodWSig{\Gm}{\_}{n}}
+\end{align*}
 
 The first observation is that, with this model, we know our system is consistent---we cannot derive $\goodTerm{\cdot}{t}{\bot}$.
-Otherwise, we have $\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \denotesS{\bot}(\star) = \emptyset$, that is an element in an empty set, a contradiction.
+Otherwise, we have $\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \denotesC{\bot}(\star) = \emptyset$, that is an element in an empty set, a contradiction.
 

### 2023-03-06 metatheory2.tex
Working on canonicity model.
Working on the appendix.

@@
 Compared to MLTT, \TT has 3 more judgements about signatures and linkage transformers.
+Moreover, the linkage transformer can be considered as a syntactic sugar (or library functions) and we will show how it can be translated away at the very end.
+Since it is a syntactic sugar, we will not consider about the linkage transformers during meta-theoretic reasoning.
 
@@
 
-\subsection{Syntactic Translation get rid of Linkages}
+\subsection{Syntactic Translation to get rid of Linkages}
 
@@
 \input{canonicitymodel.tex}
+
+\subsection{Syntactic Translation to get rid of Linkage Transformer}

### 2023-03-06 canonicitymodel.tex
Working on canonicity model.

@@
   \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
-  &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
+  % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
   \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
   & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
-  &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
+  % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
   \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t.
 } \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
-  &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
+  % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
   \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t.
@@
 
+\newcommand{\Glued}[1]{\ensuremath{{#1}^C}}
+\newcommand{\GSubstExt}[2]{\ensuremath{{#1},^c{#2}}}
+
+Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t.
+if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\Glued{\Pi}(A', B') = \denotesC{\TyPi{A}{B}}$.
+Using more type-theoretical language, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model.
+Similarly, we can have $\Glued{\Sigma}$ and $\GSubstExt{\gm}{t}$ for dependent tuple type and substitution extension.
+These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.
+
 Then we extend the above model to \TT , we still start with the judgement. Since linkage transformers are just syntactic sugar, we will not interpret it.
@@
 \begin{align*}
-  \denotesC{\goodSig{\Gm}{\_}{n}} & \text{ is a Set  and we define  } {\goodSigC{\Gm}{\sigma}{n}} :\iff \denotesC{\sigma} \in  \denotesC{\goodSig{\Gm}{\_}{n}} \\ 
-  \denotesC{\goodWSig{\Gm}{\_}{n}} &\text{ is a set of list of pairs of type interpretation of length } n \\
-  & \text{ and we define } \goodWSigC{\Gm}{\tau}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \in \denotesC{\goodWSig{\Gm}{\_}{n}}
+  \denotesC{\goodSig{\Gm}{\lsig}{n}} & \text{ is a list of 3-tuple of length $n$  } \\
+  % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
+  \denotesC{\goodWSig{\Gm}{\wsig}{n}} &\text{ is a list of 2-tuple of length $n$ } \\
+  % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
+  & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
+  \mathit{\cL}("nil") &= \denotesC{\top} \\
+  \mathit{\cL}((A,s,T) :: tl)(\gm) &= {\mathit{\cL}(tl)}(\gm) \times \mathit{\Pi}(A,T)(\gm) \\
+  \mathit{\cP}("nil") &= \denotesC{\top}\\
+  \mathit{\cP}((A,s,T) :: tl) &= \mathit{\Sigma}(\mathit{\cP}(tl), \sub{T}{(p_1, s)})\\
+  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{\cL}(\denotesC{\lsig}) \\
+  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{\cP}(\denotesC{\lsig}) \\
+  % 
+  \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
+  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesC{A}, \denotesC{s}, \denotesC{T})::\denotesC{\lsig}\\ 
+  \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
+  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0]\\ 
+  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= ("hd"~\denotesC{\lsig})[1]\\ 
+  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[2]\\
+  \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &=  \\
+  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}} &= \\
+  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \\
+  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \\
+  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= \\
+  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= \\
+  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= \\
+  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= \\
+  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{}} &= \\
+  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
+  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= \\
+  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\mathit{\Pi}{A}{\mathit{\Pi} {\mathit{\Pi}{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}}}\\
+  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \\
+  \denotesC{\Recproj{j}{\lkg}} &= \\
+  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\
 \end{align*}

### 2023-03-06 metatheory2.tex
Working on Canonicity Appendix.

--- 2023-03-06 metatheory2.tex
@@
 Compared to MLTT, \TT has 3 more judgements about signatures and linkage transformers.
+Moreover, the linkage transformer can be considered as a syntactic sugar (or library functions) (we will show why at the end).
+Since it is a syntactic sugar, we will not consider about the linkage transformers during meta-theoretic reasoning.
 
@@
 \input{canonicitymodel.tex}
+
+\subsection{Linkage Transformer as Syntax Sugar}
+Observe the syntax, linkage transformer and its judgement is defined above 
+other parts.
+
+So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$.
+Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality given in the syntax.
+In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.

### 2023-03-06 canonicitymodel.tex
Working on Canonicity Model.

@@
 
-Like most logical relation proof,  we interpret each judgement, inductively interpret each syntax piece and prove compatibility lemma.
+Like most logical relation proof,  we interpret each judgement, inductively interpret each syntax piece.
 We are working on an intrinsic setting thus even we omit context for brevity, the syntax piece is still well-typed.
-We only specify part of the logical relation proof for MLTT part.
-For the complete version, please refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic}.
+We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
+Interested audience can refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} for a complete technical treatements.
 
@@
   \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
-  \denotesC{\goodTerm{\Gm}{\unit}{\top}} &= \\
-  \denotesC{\goodTerm{\Gm}{\true}{\cB}} &= \\
-  \denotesC{\goodTerm{\Gm}{\false}{\cB}} &= \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}} &= \\
-  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}} &= \\
+  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm) &= \unit\\
+  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm) &= \true \\
+  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm) &= \false \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \\
+  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm) &= \eqrefl{\sub{t}{\gm}}\\
   \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
@@
 if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\Glued{\Pi}(A', B') = \denotesC{\TyPi{A}{B}}$.
-Using more type-theoretical language, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model.
+Type-theoretical speaking, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model.
 Similarly, we can have $\Glued{\Sigma}$ and $\GSubstExt{\gm}{t}$ for dependent tuple type and substitution extension.
@@
 
-Then we extend the above model to \TT , we still start with the judgement. Since linkage transformers are just syntactic sugar, we will not interpret it.
+Now we can extend the above model to \TT , we still start with the judgement. Since linkage transformers are just syntactic sugar, we will not interpret it.
 The intuition of the model is similar to that of the syntactic translation---use the canonicity model of sigma type (from the earlier work) for linkage type and using inductive facility in the meta-logic to justify our W-type.
@@
   \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
+  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm) &= (\denotesC{\ell}(\gm), \lambda u . t (\gm, u)) \\
   \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \\
-  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}} &= \\
   \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \\
   \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \\
-  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= \\
-  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= \\
-  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= \\
-  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= \\
-  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{}} &= \\
+  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
+  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesC{A} , \denotesC{B}) :: \denotesC{\wsig}\\
+  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[0]\\
+  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[1]\\
+  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
+  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}} &= \\
   \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
-  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= \\
   \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\mathit{\Pi}{A}{\mathit{\Pi} {\mathit{\Pi}{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}}}\\

### 2023-03-06 metatheory2.tex
Working on Canonicity Model.

@@
 
-\subsection{Standard Model for Consistency}
+% \subsection{Standard Model for Consistency}
 
-\input{standardmodel.tex}
+% \input{standardmodel.tex}
 
@@
 So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$.
-Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality given in the syntax.
+Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality\EDJ{Add substitution equalities for linkage transformers} given in the syntax.
 In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.

### 2023-03-07 canonicitymodel.tex
Save.

@@
   \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm) &= \false \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{a}(\gm) \text{ if } \denotes{c}(\gm) \equiv \true \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{b}(\gm) \text{ if } \denotes{c}(\gm) \equiv \false \\
   \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm) &= \eqrefl{\sub{t}{\gm}}\\
@@
 
-\newcommand{\Glued}[1]{\ensuremath{{#1}^C}}
+\newcommand{\Glued}[1]{\ensuremath{{#1}^c}}
+\newcommand{\GluedPi}[2]{\ensuremath{\Pi^c({#1},{#2})}}
 \newcommand{\GSubstExt}[2]{\ensuremath{{#1},^c{#2}}}
-
+\newcommand{\Gpair}[2]{\ensuremath{({#1},^c{#2})}}
+\newcommand{\Gfst}[1]{\ensuremath{\texttt{fst}^c~{#1}}}
+\newcommand{\Gsnd}[1]{\ensuremath{\texttt{snd}^c~{#1}}}
+\newcommand{\Gsub}[2]{\ensuremath{{#1}\!\left[{#2}\right]^c}}
+\newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^c({#1})}}
+\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^c)^{#1}}}
 Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t.
-if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\Glued{\Pi}(A', B') = \denotesC{\TyPi{A}{B}}$.
+if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\GluedPi{A'}{B'} = \denotesC{\TyPi{A}{B}}$.
 Type-theoretical speaking, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model.
-Similarly, we can have $\Glued{\Sigma}$ and $\GSubstExt{\gm}{t}$ for dependent tuple type and substitution extension.
+Similarly, we can have $\Glued{\Sigma}$, $\Gpair{a}{b}$  and $\GSubstExt{\gm}{t}$  for dependent tuple type, depende tuple construction and substitution extension, and also all other constructions.
 These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.
@@
   & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
-  \mathit{\cL}("nil") &= \denotesC{\top} \\
-  \mathit{\cL}((A,s,T) :: tl)(\gm) &= {\mathit{\cL}(tl)}(\gm) \times \mathit{\Pi}(A,T)(\gm) \\
-  \mathit{\cP}("nil") &= \denotesC{\top}\\
-  \mathit{\cP}((A,s,T) :: tl) &= \mathit{\Sigma}(\mathit{\cP}(tl), \sub{T}{(p_1, s)})\\
-  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{\cL}(\denotesC{\lsig}) \\
-  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{\cP}(\denotesC{\lsig}) \\
+  \mathit{L}("nil") &= \denotesC{\top} \\
+  \mathit{L}((A,s,T) :: tl)(\gm) &= {\mathit{L}(tl)}(\gm) \times \GluedPi{A}{T}(\gm) \quad \text{($\times$ is a special case for $\Glued{\Sigma}$)}  \\
+  \mathit{P}("nil") &= \denotesC{\top}\\
+  \mathit{P}((A,s,T) :: tl) &= \Glued{\Sigma}(\mathit{P}(tl), \Gsub{T}{\GSubstExt{\Glued{\pi_1}}{s}}) \quad \text{(doing substitution on $T$)} \\
+  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{L}(\denotesC{\lsig}) \\
+  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{P}(\denotesC{\lsig}) \\
   % 
@@
   \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
-  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0]\\ 
+  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0] \quad \text{ take the first element in the tuple}\\ 
   \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= ("hd"~\denotesC{\lsig})[1]\\ 
@@
   \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \\
-  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \\
-  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \\
+  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \Gfst{\denotes{\ell}} \\
+  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \Gapp{\Gsnd{\denotes{\ell}}} \\
   \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
@@
   \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
-  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\mathit{\Pi}{A}{\mathit{\Pi} {\mathit{\Pi}{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}}}\\
+  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}}\\
   \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \\

### 2023-03-07 canonicitymodel.tex
Save.
Working on appendices.

@@
 \newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^c)^{#1}}}
+\newcommand{\GLSigAdd}[3]{\ensuremath{\nu^{+c}({#1},{#2},{#3})}}
+\newcommand{\GCaseSig}[3]{\ensuremath{\texttt{CaseTy}^c({#1},{#2},{#3})}}
+
 Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t.
@@
 These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.
+What's more, during the construction of the model for \TT, we will use $\Glued{(\cdot)}$ when we can.
 
@@
   \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
-  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm) &= (\denotesC{\ell}(\gm), \lambda u . t (\gm, u)) \\
-  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \\
-  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \Gfst{\denotes{\ell}} \\
-  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \Gapp{\Gsnd{\denotes{\ell}}} \\
+  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm) &= (\denotesC{\lkg}(\gm), \lambda u . t (\gm, u)) \\
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
+  \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gpair{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{(\SubstId)}}{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}}}
+  \\ &\text{ when} \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\
+  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyTkg{\lsig}}} &= \Gfst{\denotes{\lkg}} \\
+  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}} &= \Gapp{\Gsnd{\denotes{\lkg}}} \\
   \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
@@
   \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
-  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}}\\
-  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \\
-  \denotesC{\Recproj{j}{\lkg}} &= \\
+  % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
+  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= 
+  {\GluedPi{\denotesC{A}}{\GluedPi {\GluedPi{\denotesC{B}}{\Gsub{\denotesC{R}}{\GSubstWeak{2}}}} {\Gsub{\denotesC{R}}{\GSubstWeak{2}}}}} \\
+  % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\
+  \text{we define } \RecSig{}{} &\text{ by induction on the signature} \\
+  RS~"nil"~R &= \denotesC{\LSigEmp} \\
+  RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\
+  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= RS~\denotesC{\tau}~\denotesC{R} \\
+  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg \\
   \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\

### 2023-03-07 syn-translate.tex
Working on appendices.
Save.

@@
   \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
-  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ are mutually recursive, } \\
-  & \text{defined by induction on the signature length} \\  
-  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
-  &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
+  \text{such that } &\goodTerm{\Gm}{\denotesT{\goodSig{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}} \\
+  \text{We make } & ({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ mutually recursive, } \\
+  & \text{defined by induction on $n$ and the signature length} \\  
+  % \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
+  % &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
   {\Sigr{\Gamma}{(n+1)}} &= 
@@
           ~(\Sigma~\cU
-                  ~(\Sigma~(\Pi(\TyTkg {("var"_0[p^1])})(El~"var"_0[p^1]))~\cU)) \\
+                  ~(\Sigma~(\Pi(\denotesT{\TyTkg {("var"_0[p^1])}})(El~"var"_0[p^1]))~\cU)) \\
   \Sigr{\Gamma}{0} &= \top \\
@@
   &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
+  \denotesT{\goodSig{\Gamma}{\LSigEmp}{0}} &= \unit \\ 
   \denotesT{\goodSig{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{(\codety{T})}) \\ 
+  \denotesT{\goodTerm{\Gamma}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \unit \\ 
   \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\TyLkg{(\LSigAdd{\sigma}{s}{T})}}} &= (\denotesT{m}, \lam{t}) \\ 
   \denotesT{\goodTerm{\Gamma}{\Tkg{m}}{\TyTkg {\sigma}}} & \text{ is defined upon } \denotesT{m} \text{ and inductively on the signature length} \\ 
+  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\
   \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= 
   (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{(\snd{\denotesT{m}})}, f = p_f\nu~\sigma \\ 
-  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{(\snd{\denotesT{m}})}, f = p_f\nu~\sigma 
 \end{align*}
-\begin{align*}
-  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \TyLkg{\sigma_1}}{\_}{\TyLkg{\sigma_2[p^1]}} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \TyLkg{\sigma_1}}{h}{\TyLkg{\sigma_2[p^1]}} \\
-  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\LSigAdd {\sigma_1} {s_1}{T})}{(\LSigAdd {\sigma_2} {s_2} {T[(p^1, \uparrow^s)]})}} &= \mu^+~(h[(p^1,\lkgproj{1}{"var"_0})])\\ &\quad \quad \{f_2[p^1]\}~(\lkgproj{2}{"var"_0})[(p^1, \uparrow^s[{p^1}^{\uparrow}])] \\
-\end{align*}
+% \begin{align*}
+%   \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \TyLkg{\sigma_1}}{\_}{\TyLkg{\sigma_2[p^1]}} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \TyLkg{\sigma_1}}{h}{\TyLkg{\sigma_2[p^1]}} \\
+%   \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\LSigAdd {\sigma_1} {s_1}{T})}{(\LSigAdd {\sigma_2} {s_2} {T[(p^1, \uparrow^s)]})}} &= \mu^+~(h[(p^1,\lkgproj{1}{"var"_0})])\\ &\quad \quad \{f_2[p^1]\}~(\lkgproj{2}{"var"_0})[(p^1, \uparrow^s[{p^1}^{\uparrow}])] \\
+% \end{align*}
 

### 2023-03-07 canonicitymodel.tex
Working on appendices.

@@
 Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax.
-However, our proof will still work directly on the earlier introdcued syntax for readability.
+Meanwhile, our proof will still work directly on the earlier introdcued syntax for readability.
 We will first specify the formulation of the most basic canonicity theorem.
@@
 
-Canonicity is one of the basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
+Canonicity is a basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
 We can even argue that, if this theorem is proven in a computable meta-logic, 

### 2023-03-07 syn-translate.tex
Working on appendices.

@@
 a \textit{type preserving syntactic translation}, a function mapping from our syntactic
-model to a ``subpart''\footnote{Not exactly subpart---the only
+model(typed Abstract Syntax Tree(AST)) to a ``subpart''\footnote{Not exactly subpart---the only
 difference lies in "Wrec". In our metatheory, it is constructed
 using linkage; in ``subpart'', it will be using $\Sigma$-types (translation
-of the linkage).} of our syntactic model \textit{without
+of the linkage).} of our syntactic model(typed AST) \textit{without
 linkages}.
@@
 syntactic translation for $\goodSig{\Gamma}{\_}{n}$ and $\cL$. The
-syntactic translation for $\cC$ and $\cCt$ are decided by their
+syntactic translation for $\cP$ are decided by their
 $\beta$-rules.
 
-We use the local morphism\cite{abbott2003category} (the same one for encoding the Sealing
-judgements) to translate inheritance judgement. This is mainly due to
-the existence of \ruleref{Inh} rule. 
+% We use the local morphism\cite{abbott2003category} (the same one for encoding the Sealing
+% judgements) to translate inheritance judgement. This is mainly due to
+% the existence of \ruleref{Inh} rule. 
 % Here we only typeset the overriding
@@
 $\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and
-$\denotesT{\cC}$ above, we actually have to prove the substitution laws
-$(\cC\sigma)[\tau]\equiv \cC(\sigma[\tau])$ and $\Sigr{\Delta}{n}[\tau]
+$\denotesT{\cP}$ above, we actually have to prove the substitution laws
+$\sub{(\TyTkg{\lsig})}{\gm}\equiv \TyTkg{\sub{\lsig}{\gm}}$ and $\sub{(\Sigr{\Delta}{n})}{\gm}
 \equiv \Sigr{\Gamma}{n}$ together. 

### 2023-03-07 metatheory2.tex
Working on examples.

@@
 So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$.
-Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality\EDJ{Add substitution equalities for linkage transformers} given in the syntax.
+Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality
+%\EDJ{Add substitution equalities for linkage transformers} 
+given in the syntax.
 In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.

### 2023-03-07 metatheory-condensed.tex
Working on examples.
Combine Wtype into the Figure.

@@
 
-\begin{figure}
-  \begin{tabular}{
-    >{\columncolor[HTML]{C0C0C0}}l lcl}
-    \hline
-    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
-      \multicolumn{1}{l|}{Equivalent Type} &
-      \multicolumn{1}{l}{} &
-      \multicolumn{1}{l|}{} \\
-    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| unit : tm!} &
-      \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
-      $\tau'$ &
-      \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
-    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| app  : tm -> tm -> tm!} &
-      \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
-      $\tau$ &
-      \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
-    \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}\lsti!unit!} &
-      $\coloneqq \Wsup{1}{\tau}{\unit}{"elim-"\bbzero}$ &
-      \cellcolor[HTML]{C0C0C0}\lsti!app unit unit! &
-      $\coloneqq\Wsup{0}{\tau}{\unit}{"unit"}$
-    \end{tabular}
-  \caption{W-type encoding of "tm \{unit, app\}"}
-  \label{fig:Wtype-example}
-\end{figure}
-
-We illustrate one concrete example in \cref{fig:Wtype-example} of a "tm" with only two constructors, 
-also as a review for how W-type encodes a given inductive type. The key insight 
-is, to encode one constructor into a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive part will be stored in $\wsigproj{i}{1}{\wsig}$ 
-while the ``arity'' of the  recursive part stores in $\wsigproj{i}{2}{\wsig}$. 
-For example, $"app" : ("tm" \to "tm" \to "tm")$ 
-$\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
-(or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
-Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
-signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit".
+% \begin{figure}
+%   \begin{tabular}{llll}
+%     \hline
+%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
+%       \multicolumn{1}{l|}{Equivalent Type} &
+%        &
+%       \multicolumn{1}{l|}{} \\
+%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}| unit : tm} &
+%       \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
+%       \multicolumn{1}{c}{$\tau'$} &
+%       \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
+%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}| app  : tm -\textgreater tm -\textgreater tm} &
+%       \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
+%       \multicolumn{1}{c}{$\tau$} &
+%       \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
+%     \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}unit} &
+%       $\coloneqq \Wsup{1}{\tau}{\unit}{"elim-"\bbzero}$ &
+%       \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}app unit unit} &
+%       $\coloneqq\Wsup{0}{\tau}{\unit}{"unit"}$ \\ \hline
+%     \cellcolor[HTML]{ECF4FF}$\lsig' = \LSigAdd{\LSigEmp}{\pi_2}{\TyS{\wcode{\tau'}}}$ &
+%       $o' = \LkgAdd{\LkgEmp}{\wcode{\tau'}}$ &
+%       \cellcolor[HTML]{ECF4FF}$\lsig = \LSigAdd{\LSigEmp}{\pi_2}{\TyS{\wcode{\tau}}}$ &
+%       $o = \LkgAdd{\LkgEmp}{\wcode{\tau}}$ \\
+%     \cellcolor[HTML]{ECF4FF}$\lsig'^+ = \LSigAdd{\lsig'}{s'}{..}$ &
+%       $o'^+ = \LkgAdd{o'}{t}$ &
+%       \cellcolor[HTML]{ECF4FF}$\lsig^+ = \LSigAdd{\lsig}{s}{..}$ &
+%       $o^+ = \LkgAdd{o}{\lkgproj{2}{o'^+}}$ \\
+%     \multicolumn{2}{l}{$A = \TySigma{\cU}{\El{\pi_2} \to \El{\pi_2}} $} &
+%       \multicolumn{2}{l}{$t = \goodTerm{\cdot , A}{\snd{\pi_2}}{\El{\fst{\pi_2}} \to \El{\fst{\pi_2}} }$} \\
+%     $\goodTerm{\cdot , \TyTkg{\lsig'}}{s'}{\sub{A}{\pi_1}}$ &
+%        &
+%       $\goodTerm{\cdot , \TyTkg{\lsig}}{s}{\sub{A}{\pi_1}}$ &
+
+%     \end{tabular}
+%   \caption{W-type encoding of "tm \{unit, app\}"}
+%   \label{fig:Wtype-example}
+% \end{figure}
+
+% We illustrate one concrete example in \cref{fig:Wtype-example} of a "tm" with only two constructors, 
+% also as a review for how W-type encodes a given inductive type. The key insight 
+% is, to encode one constructor into a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive part will be stored in $\wsigproj{i}{1}{\wsig}$ 
+% while the ``arity'' of the  recursive part stores in $\wsigproj{i}{2}{\wsig}$. 
+% For example, $"app" : ("tm" \to "tm" \to "tm")$ 
+% $\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
+% (or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
+% Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
+% signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit".
 $"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context.
@@
 
+\item We illustrate the structure of W-type signature (as a list) here.
+$\wsig_{\texttt{tm}}^0, \wsig_{\texttt{tm}}^1$ are (the subparts of) the signatures of constructors \{"tm\_unit"\} and \{"tm\_unit","tm\_app"\} respectively.
+Each item in this W-type signature, having two types, corresponds to one constructor, where the first type stores the non-recursive part of the constructor and the second type stores the ``arity'' of the  recursive part.
+For example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times (\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule \ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone, \wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \bbtwo$).
+This type means there is no information on non-recursive part (thus use unit type $\bbone$) and the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1 = \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\bbtwo}$.
+The story is similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and thus the ``arity'' is zero.
+
+Generally speaking, the expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict positivity of Inductive type.
 \item 

### 2023-03-10 metatheory-condensed.tex
A large change for named variable, Please review.

@@
 %
-We acknowledge that the presentation herein is dense for an audience
-unacquainted with MLTT or explicit substitutions, so we summarize the salient points
-first before delving into the technical details.
+
+For accessibility, we present MLTT using named variables and meta-level substitution,
+however we emphasize that in the meta-theoretic reasoning carried out in the appendix, 
+we will shift to debruijn indices and explicit substitution because it leads to clearer proof details.
+We summarize the salient points before delving into the technical details for audience
+unacquainted with MLTT.
 
@@
 
-\input{mltt-selected.tex}
+% \input{mltt-selected.tex}
+\input{mltt-selected-named.tex}
 \input{fmltt-selected.tex}
@@
 
-\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT
-(and \cref{fig:fmltt-selected} \TT).
+\cref{fig:mltt-selected} presents the syntax and selected typing rules of (predicative) MLTT
+(and \cref{fig:fmltt-selected} \TT). We formulate in an intrinsic style---every appearing syntax piece
+is well-typed in prior, and the premise judgement required for well-typed might be omitted for conciseness. 
+For example, the well-typed context $\goodCtx{\Gm}{}$ required by \ruleref{n/tm/lam} is omitted.  
+We also use round bracket to indicate substitution. 
+For each derivation (of each judgement) we introduce judgemental equality indicating convertibility 
+in MLTT.  
+We have dependent function and dependent tuple type equipped by standard dependent type theory.
 %We briefly review the base MLTT fragment first.
 %
-In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}\YZ{Does coquand2018canonicity really use explicit subst?}\EDJreply{Yes it is.
+% In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}\YZ{Does coquand2018canonicity really use explicit subst?}\EDJreply{Yes it is.
 Look at Section 2.1.
@@
 The former one directly work on CwF, the latter one include substitution as part of the raw syntax.}
-in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
-substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
-of the syntax rather than meta-operations.
-%
-Variables are in the form of \citeauthor{debruijn1964} indices:
-\var{n} is the variable bound by the $n$-th closest enclosing binder.
-For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
-%
-Substitutions are typed with the form 
+% in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
+% substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
+% of the syntax rather than meta-operations.
+%
+% Variables are in the form of \citeauthor{debruijn1964} indices:
+% \var{n} is the variable bound by the $n$-th closest enclosing binder.
+% For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
+%
+% Substitutions are typed with the form 
 %\EDJ{It seems to me G |- g : D is a judgement not a rule.
@@
 In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
-$\goodSub{\Gm}{\gm}{\Dl}$.
-The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
-yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
-%
-The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
-\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
-\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
-then applies~$\gm$.
-For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
-that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
-$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
-where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
+% $\goodSub{\Gm}{\gm}{\Dl}$.
+% The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
+% yields terms valid in~$\Gm$ (\ruleref{n/tm/sub} and \ruleref{n/ty/sub}).
+% %
+% The two main forms of substitutions are weakening (\ruleref{n/sub/wk}) and extension (\ruleref{n/sub/ext}):
+% \sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
+% \sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
+% then applies~$\gm$.
+% For example, rule \ruleref{n/tm/snd} states that if $t$ is a dependent pair
+% that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
+% $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
+% where $\SubstId$ is the identity substitution (\ruleref{n/sub/id}).
 
@@
 and arguably behave better \cite{luo2012notes}.
-$\codety{T}$ encodes type~$T$ (rule \ruleref{tm/c}) and thus inhabits~$\cU$.
-$\El{t}$ decodes~$t$ (rule \ruleref{ty/el}).
+$\codety{T}$ encodes type~$T$ (rule \ruleref{n/tm/c}) and thus inhabits~$\cU$.
+$\El{t}$ decodes~$t$ (rule \ruleref{n/ty/el}).
 %\footnote{By contrast,  Russel-style universes are inhabited by types directly.
 Tarski-style universes are ``semantically more fundamental'' \cite{luo2012notes} and friendlier to our meta-theoretical reasoning.}
-MLTT has an infinite hierarchy of universes;
+Our MLTT has an infinite hierarchy of universes;
 we omit universe levels in the presentation for conciseness.
@@
 A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
-in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
+in its type (rule \ruleref{n/tm/s})~\cite{aspinall1995singleton,stone2000}.
 
@@
 the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
-rather than \emph{existentially} quantified as is in \ruleref{tm/pair}---%
+rather than \emph{existentially} quantified as is in \ruleref{n/tm/pair}---%
 over how the fields in $t$'s context are defined.
@@
 
-The definitive version containing all the rules in \TT is stated in
+% The definitive version containing all the rules in \TT is stated in
+% terms of a meta\-language supporting quotient inductive-inductive types
+% (QIITs)~\cite{nordvall2010inductive,altkap2016}.
+% QIITs support equality constructors, which facilitate the expression of
+% conversion rules (e.g., \ruleref{tmeq/sub/id}).
+% The formalization is available as a supplemental file in Agda syntax,
+% though it is not checked by Agda because Agda does not support QIITs
+% natively.
+% An effort to mechanize the formalization is underway, using Agda's
+% \lsti{REWRITE} pragma to replace the limited uses of equality reflection.
+
+The definitive version containing all the rules in \TT is stated in the appendix 
+using debruijn indices and explicit substitution.
 terms of a meta\-language supporting quotient inductive-inductive types
@@
 natively.
-An effort to mechanize the formalization is underway, using Agda's
-\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
+An effort to mechanize the formalization is underway, by following  \citet{altkap2016} 
+to use quotient inductive-inductive types as 
+the logical framework to state the syntax and using Agda's
+\lsti{REWRITE} pragma to replace the limited uses of equality reflection appeared in 
+an extensional meta\-logic.
+
 %a \textbf{pen-and-paper} formalization

### 2023-03-10 metatheory-condensed.tex
Fix.

@@
 using debruijn indices and explicit substitution.
-terms of a meta\-language supporting quotient inductive-inductive types
-(QIITs)~\cite{nordvall2010inductive,altkap2016}.
-QIITs support equality constructors, which facilitate the expression of
-conversion rules (e.g., \ruleref{tmeq/sub/id}).
-The formalization is available as a supplemental file in Agda syntax,
-though it is not checked by Agda because Agda does not support QIITs
-natively.
 An effort to mechanize the formalization is underway, by following  \citet{altkap2016} 

### 2023-03-12 metatheory-condensed.tex
Partially Changing to named variables. For FMLTT. Please review.

@@
 \input{mltt-selected-named.tex}
-\input{fmltt-selected.tex}
+\input{fmltt-selected-named.tex}
 
@@
 For example, the well-typed context $\goodCtx{\Gm}{}$ required by \ruleref{n/tm/lam} is omitted.  
-We also use round bracket to indicate substitution. 
+We also use square bracket to indicate substitution. 
 For each derivation (of each judgement) we introduce judgemental equality indicating convertibility 
@@
 
-In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
+In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, a : A}{t}{T}$ is
 responsible for late binding.
@@
 Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
-the premise $\goodTerm{\Gm, A}{t}{T}$ makes clear that
+the premise $\goodTerm{\Gm, a : A}{t}{T}$ makes clear that
 the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
@@
 \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i}(-0.02,-0.10)(-0.02,0.28)
-A_i%
+a_i : A_i%
 \tikzmarkend{A-i}%
@@
 The third premise
-$\goodTerm{\Gm,\TyTkg{\lsig}}{\seal}{\sub{A}{\SubstWeak{1}}}$
+$\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
 in rules \ruleref{l/add} and \ruleref{lsig/add} is

### 2023-03-12 metatheory-condensed.tex
Save.

@@
 two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
-$\goodTerm{\Gm,\sub{\wsigproj{i}{2}{\wsig}}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{\El{\wcode{\wsig}}}$,
-we can construct a term $\Wsup{i}{\wsig}{t_1}{t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
+$\goodTerm{\Gm,{\wsigproj{i}{2}{\wsig}}[t_1/x]}{t_2}{\El{\wcode{\wsig}}}$,
+we can construct a term $\Wsup{i}{\wsig}{t_1}{a.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
 W-type (\ruleref{tm/w}).

### 2023-03-12 metatheory-condensed.tex
edit

--- 2023-03-12 metatheory-condensed.tex
@@
 
-For accessibility, we present MLTT using named variables and meta-level substitution,
-however we emphasize that in the meta-theoretic reasoning carried out in the appendix, 
-we will shift to debruijn indices and explicit substitution because it leads to clearer proof details.
-We summarize the salient points before delving into the technical details for audience
-unacquainted with MLTT.
+For accessibility of the main text, \cref{sec:metatheory2} presents \TT using
+named variables and meta-level substitution.
+For clarity of proof details, the technical appendix supplementing
+\cref{sec:metatheory2} instead uses de Bruijn indices and explicit substitution.
+%
+We acknowledge that the presentation is dense for an audience
+unacquainted with MLTT, so we summarize the salient points first.
 

### 2023-03-12 metatheory-condensed.tex
took a pass over MLTT review

@@
 Most notably, it extends MLTT with what we call \emph{linkages}.
-(Linkages are a namesake of the theoretical device through which \citet{zm2017} model
-family polymorphism in an OO setting, but the technical details differ significantly from the prior work.)
+Linkages are a namesake of the theoretical device through which \citet{zm2017} model
+family polymorphism in an OO setting, but the technical details differ significantly from the prior work.
 
@@
 
-\cref{fig:mltt-selected} presents the syntax and selected typing rules of (predicative) MLTT
-(and \cref{fig:fmltt-selected} \TT). We formulate in an intrinsic style---every appearing syntax piece
-is well-typed in prior, and the premise judgement required for well-typed might be omitted for conciseness. 
-For example, the well-typed context $\goodCtx{\Gm}{}$ required by \ruleref{n/tm/lam} is omitted.  
-We also use square bracket to indicate substitution. 
-For each derivation (of each judgement) we introduce judgemental equality indicating convertibility 
-in MLTT.  
-We have dependent function and dependent tuple type equipped by standard dependent type theory.
+\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT %(predicative) MLTT
+(and \cref{fig:fmltt-selected} \TT).
+Dependent function types \nTyPi{x}{A}{B},
+dependent product types \nTySigma{x}{A}{B}, and
+identity types \TyId{t_1}{t_2} are standard.
+So are their introduction and elimination forms.
+Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
+A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
+in its type (rule \ruleref{n/tm/s})~\cite{aspinall1995singleton,stone2000}.
+For each derivation (of each judgment) we introduce judgmental equality
+(of the form \goodTerm{\Gm}{t_1 \equiv t_2}{T})
+indicating convertibility.
+
+We use Tarski-style universes \cite{hofmann1997syntax},
+which, unlike Russell-style ones,
+are not inhabited by types directly but rather by the \emph{codes} of types
+and arguably behave better \cite{luo2012notes}.
+Term $\codety{T}$ encodes type~$T$ (\ruleref{n/tm/c}) and thus inhabits universe~$\cU$.
+Type $\El{t}$ decodes term~$t$ (\ruleref{n/ty/el}).
+%\footnote{By contrast,  Russel-style universes are inhabited by types directly.
+Tarski-style universes are ``semantically more fundamental'' \cite{luo2012notes} and friendlier to our meta-theoretical reasoning.}
+There is an infinite hierarchy of universes;
+we omit universe levels in the presentation for conciseness.
+%\rA{Why is W(t) a term and not a type?}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}
+
+We use intrinsically typed syntax \cite{church1940}---terms and types are well-formed by construction.
+For conciseness, typing rules omit obvious premises required for well-formedness.
+For example, \ruleref{n/tm/lam} implicitly requires that the context be well-typed
+($\goodCtx{\Gm}{}$).
+
 %We briefly review the base MLTT fragment first.
@@
 % where $\SubstId$ is the identity substitution (\ruleref{n/sub/id}).
-
-%In dependent type system, types are treated as terms, and should have a type.
-We use Tarski-style universes \cite{hofmann1997syntax}.
-Unlike Russell-style universes, they
-are not inhabited by types directly but rather by the \emph{codes} of types
-and arguably behave better \cite{luo2012notes}.
-$\codety{T}$ encodes type~$T$ (rule \ruleref{n/tm/c}) and thus inhabits~$\cU$.
-$\El{t}$ decodes~$t$ (rule \ruleref{n/ty/el}).
-%\footnote{By contrast,  Russel-style universes are inhabited by types directly.
-Tarski-style universes are ``semantically more fundamental'' \cite{luo2012notes} and friendlier to our meta-theoretical reasoning.}
-Our MLTT has an infinite hierarchy of universes;
-we omit universe levels in the presentation for conciseness.
-%\rA{Why is W(t) a term and not a type?}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}
-
-A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
-in its type (rule \ruleref{n/tm/s})~\cite{aspinall1995singleton,stone2000}.
 

### 2023-03-13 metatheory-condensed.tex
Fix macro usage.

@@
 two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
-$\goodTerm{\Gm,{\wsigproj{i}{2}{\wsig}}[t_1/x]}{t_2}{\El{\wcode{\wsig}}}$,
-we can construct a term $\Wsup{i}{\wsig}{t_1}{a.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
+$\goodTerm{\Gm, \mathit{self} : \nsub{\wsigproj{i}{2}{\wsig}}{x}{t_1}}{t_2}{\El{\wcode{\wsig}}}$,
+we can construct a term $\Wsup{i}{\wsig}{t_1}{\mathit{self}.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
 W-type (\ruleref{tm/w}).
@@
 
-In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, a : A}{t}{T}$ is
+In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ is
 responsible for late binding.
@@
 Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
-the premise $\goodTerm{\Gm, a : A}{t}{T}$ makes clear that
+the premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ makes clear that
 the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
@@
 \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i}(-0.02,-0.10)(-0.02,0.28)
-a_i : A_i%
+\mathit{self}_i : A_i%
 \tikzmarkend{A-i}%
@@
 constructed from the introduction forms
-$\small\InhId$, $\small\InhExt{h}{t}$, $\small\InhOv{h}{t}$, $\small\InhInh{h}$, etc.
+$\small\InhId$, $\small\InhExt{h}{x.t}$, $\small\InhOv{h}{x.t}$, $\small\InhInh{h}$, etc.
 \cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
 \ruleref{tmeq/ov/beta}.
-It states that applying the transformer $\small\InhOv{h}{t_2}$ to
-a linkage of form $\LkgAdd{\lkg}{t_1}$
+It states that applying the transformer $\small\InhOv{h}{\mathit{self}_2.t_2}$ to
+a linkage of form $\LkgAdd{\lkg}{\mathit{self}_1.t_1}$
 overrides the linkage's last field~$t_1$ with~$t_2$.
@@
 
-The definitive version containing all the rules in \TT is stated in the appendix 
+The definitive version containing all the rules in \TT is stated in
+\ifreport \cref{sec:complete-fmltt}.
+\else a supplemental appendix.
+\fi
 using debruijn indices and explicit substitution.
@@
 %
-We also typeset a more complete set of rules than the space allows and
-present it in
-\ifreport \cref{sec:complete-fmltt}.
-\else a supplemental appendix.
-\fi
+% We also typeset a more complete set of rules than the space allows and
+% present it in
+% \ifreport \cref{sec:complete-fmltt}.
+% \else a supplemental appendix.
+% \fi
 
@@
 However, the intro rules for the linkage types gives it the power of late-binding (also the fundation of inheritance and overriding).
-More concretely, the boxed premise in \ruleref{Lnkg Add} is making $t$ parametric on $A$, contrary to how in sigma type, the $t$ will be about a concrete $a : A$.
+More concretely, the boxed premise in \ruleref{Lnkg Add} is making $t$ parametric on $A$, contrary to how in sigma type, the $t$ will be about a concrete $\mathit{self} : A$.
 This parametricity makes $t$ inheritable and the earlier stuff about $A$ late-bound.

### 2023-03-13 metatheory-condensed.tex
Fix rule ref name

@@
 Our enhanced W-type is decided by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
-composed of a list (\ruleref{wsig/empty} and \ruleref{wsig/add}) of $n$ pairs of 
+composed of a list (\ruleref{n/wsig/empty} and \ruleref{n/wsig/add}) of $n$ pairs of 
 types $\{(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})\}_i$,
@@
 %
-Rule~\ruleref{tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
+Rule~\ruleref{n/tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
 $\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$ models the $i$-th constructor---given 
@@
 we can construct a term $\Wsup{i}{\wsig}{t_1}{\mathit{self}.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
-W-type (\ruleref{tm/w}).
+W-type (\ruleref{n/tm/w}).
 
@@
 W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
-case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
+case handlers for the $n$ constructors in~$\wsig$ (\ruleref{n/tm/wrec}).
 Each case handler has a type of form $\CaseSig{A}{B}{T}$, where
-$T$ is the motive of the recursion (\ruleref{tyeq/casety});
+$T$ is the motive of the recursion (\ruleref{n/tyeq/casety});
 for simplicity, we model non-dependent motives for recursion.
@@
 
-Rules \ruleref{tm/wsup} and \ruleref{tm/wrec} require access to~$\wsig$:
+Rules \ruleref{n/tm/wsup} and \ruleref{n/tm/wrec} require access to~$\wsig$:
 the W-type is exhaustively generated by its constructors, and
@@
 $\goodSig{\Gm}{\lsig}{n}$.
-As rules \ruleref{lsig/empty} and \ruleref{lsig/add} show, a linkage
+As rules \ruleref{n/lsig/empty} and \ruleref{n/lsig/add} show, a linkage
 signature is a list of $n$~types.
@@
 where $\TyLkg{\lsig}$ is a type formed by~$\lsig$.
-As rules \ruleref{l/empty} and \ruleref{l/add} show, a linkage is a list
+As rules \ruleref{n/l/empty} and \ruleref{n/l/add} show, a linkage is a list
 of $n$~terms, each representing a field of the family modeled by the linkage.
 
-In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ is
+In rule \ruleref{n/l/add}, the second premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ is
 responsible for late binding.
@@
 how the types of the fields prior to~$t$ are exposed to the typing of~$t$.
-Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
+Crucially, %unlike the second field in a dependent pair (\ruleref{n/tm/pair}),
 the premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ makes clear that
-the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
+the typing of~$t$ in \ruleref{n/l/add} is \emph{universally} quantified---%
 rather than \emph{existentially} quantified as is in \ruleref{n/tm/pair}---%
@@
 %recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$
-The type~$A$ in \ruleref{l/add} and \ruleref{lsig/add},
+The type~$A$ in \ruleref{n/l/add} and \ruleref{n/lsig/add},
 abstracting the types of the prior fields, does not necessarily contain the same types as those recorded by~$\TyLkg{\lsig}$,
@@
 $\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
-should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
-and \ruleref{tm/wrec} stipulate.
+should see the concrete signature~$\wsig$, as rules \ruleref{n/tm/wsup}
+and \ruleref{n/tm/wrec} stipulate.
 By contrast, $\wsig$ should be hidden
@@
 Each item in this W-type signature, having two types, corresponds to one constructor, where the first type stores the non-recursive part of the constructor and the second type stores the ``arity'' of the  recursive part.
-For example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times (\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule \ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone, \wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \bbtwo$).
+For example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times (\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule \ruleref{n/tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone, \wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \bbtwo$).
 This type means there is no information on non-recursive part (thus use unit type $\bbone$) and the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1 = \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\bbtwo}$.
@@
 So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
-as required by rule \ruleref{tm/wsup}.
+as required by rule \ruleref{n/tm/wsup}.
 
@@
 $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
-in rules \ruleref{l/add} and \ruleref{lsig/add} is
+in rules \ruleref{n/l/add} and \ruleref{n/lsig/add} is
 responsible for hiding W-type signatures.
-Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
+Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{n/tyeq/pk/add})
 that packages the types of all fields preceding the current field~$t$,
 and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
-$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new tuple of type~$A$
+$\TyTkg{\lsig}$ (rule \ruleref{n/tmeq/pk/add}) into a new tuple of type~$A$
 that hides W-type signatures behind $\cU$, if necessary.
@@
 applying~$h$ to a linkage of type~$\TyLkg{\lsig_1}$ yields a linkage of
-type~$\TyLkg{\lsig_2}$ (\ruleref{tm/inh}).
+type~$\TyLkg{\lsig_2}$ (\ruleref{n/tm/inh}).
 %Linkage transformers are second-class; the only way to use them is to
@@
 \cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
-\ruleref{tmeq/ov/beta}.
+\ruleref{n/tmeq/ov/beta}.
 It states that applying the transformer $\small\InhOv{h}{\mathit{self}_2.t_2}$ to
@@
 % QIITs support equality constructors, which facilitate the expression of
-% conversion rules (e.g., \ruleref{tmeq/sub/id}).
+% conversion rules (e.g., \ruleref{n/tmeq/sub/id}).
 % The formalization is available as a supplemental file in Agda syntax,

### 2023-03-13 metatheory2.tex
Start changing on appendices.

@@
-In this section, we start by introducing the complete formulation of the syntax of \TT, 
-and then we detail the models and proofs for syntactic translation (that translate away linkages), 
-consistency and canonicity.
+In this section, we lay out the complete version of the syntax of \TT but in the style of \citeauthor{debruijn1964} indices and \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
+substitutions~$\gm$ and their applications (e.g., in the form of \sub{T}{\gm}) are part
+of the syntax rather than meta-operations.
 
-\subsection{Warm Up: Review Declarative Syntax for MLTT}
-We begin with a review of MLTT as \TT can be roughly considered as MLTT with linkages.
-For our syntax, we work in an intrinsic-typed setting, so there is no un-typed preterms appearing anywhere. We introduce judgemental equality ($\equiv$)
+Other parts of the syntax remained the same as the main text.
+
+We also work in an intrinsic-typed setting, so there is no un-typed preterms appearing anywhere. We introduce judgemental equality ($\equiv$)
 for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
 contexts and explicits substitutions.
+
+This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT.
+
+After the syntax, we detail the models and proofs for syntactic translation (that translate away linkages),  consistency and canonicity.
+
 We will make sure all the introduced models and proofs respecting judgemental equality. In other words, we will consider our syntax as the result of taking quotient by these judgemental equality.
 
+\subsection{Review MLTT in Debruijn Indices and Explicit Subst}
+
+\TT can be roughly considered as MLTT with linkages so we still start with vanilla MLTT first.
+
 \input{all-mltt-typing}
+
+The change of the syntax formulation is important to the simplicity of our proof, as we want to handle the proof about substitutions and terms uniformly---thus making substitution as part of the syntax (explicit substitution) will be helpful.
+
+Variables are in the form of \citeauthor{debruijn1964} indices:
+\var{n} is the variable bound by the $n$-th closest enclosing binder.
+For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
+%
+Substitutions are typed with the form 
+%\EDJ{It seems to me G |- g : D is a judgement not a rule.
+Rules usually have two parts (premises + conclusion).
+Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here).
+I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this.
+In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
+$\goodSub{\Gm}{\gm}{\Dl}$.
+The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
+yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
+%
+The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
+\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
+\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
+then applies~$\gm$.
+For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
+that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
+$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
+where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
 
@@
 Compared to MLTT, \TT has 3 more judgements about signatures and linkage transformers.
-Moreover, the linkage transformer can be considered as a syntactic sugar (or library functions) (we will show why at the end).
+Moreover, the linkage transformer can be considered as a syntactic sugar (or ``library functions'' in the main text) (we will show why at the end).
 Since it is a syntactic sugar, we will not consider about the linkage transformers during meta-theoretic reasoning.

### 2023-03-13 syn-translate.tex
Start changing on appendices.

@@
 
-This ``subpart'' is the standard MLTT with an
+This ``subpart'' is the standard MLTT with our
 (unconventional) formulation of W-types,
@@
 Basically, we can compile away (the derivation of) judgement $\goodSig{\Gamma}{v}{n}$,
-the type $\cL$ and $\cC$, and related terms, by transforming them into
+the type $\cL$, $\cP$, and related terms, by transforming them into
 corresponding $\Sigma$-types.
@@
 
-We omit all the equational rules ($\beta,\eta$ and substitution) here.
-However, for coherence, when we mutual recursively define
+We omit all the validation of equational rules ($\beta,\eta$ and substitution) here.
+However, when we mutual recursively define
 $\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and

### 2023-03-13 metatheory-condensed.tex
Done Changing to named variables.
Remove mention of the universe level in main text
Change Back the Rule Name.

--- 2023-03-13 metatheory-condensed.tex
@@
 A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
-in its type (rule \ruleref{n/tm/s})~\cite{aspinall1995singleton,stone2000}.
+in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
 For each derivation (of each judgment) we introduce judgmental equality
@@
 and arguably behave better \cite{luo2012notes}.
-Term $\codety{T}$ encodes type~$T$ (\ruleref{n/tm/c}) and thus inhabits universe~$\cU$.
-Type $\El{t}$ decodes term~$t$ (\ruleref{n/ty/el}).
+Term $\codety{T}$ encodes type~$T$ (\ruleref{tm/c}) and thus inhabits universe~$\cU$.
+Type $\El{t}$ decodes term~$t$ (\ruleref{ty/el}).
 %\footnote{By contrast,  Russel-style universes are inhabited by types directly.
@@
 
-We use intrinsically typed syntax \cite{church1940}---terms and types are well-formed by construction.
+% We use intrinsically typed syntax \cite{church1940}---terms and types are well-formed by construction.
 For conciseness, typing rules omit obvious premises required for well-formedness.
-For example, \ruleref{n/tm/lam} implicitly requires that the context be well-typed
+For example, \ruleref{tm/lam} implicitly requires that the context be well-typed
 ($\goodCtx{\Gm}{}$).
@@
 % The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
-% yields terms valid in~$\Gm$ (\ruleref{n/tm/sub} and \ruleref{n/ty/sub}).
+% yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
 % %
-% The two main forms of substitutions are weakening (\ruleref{n/sub/wk}) and extension (\ruleref{n/sub/ext}):
+% The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
 % \sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
@@
 % then applies~$\gm$.
-% For example, rule \ruleref{n/tm/snd} states that if $t$ is a dependent pair
+% For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
 % that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
 % $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
-% where $\SubstId$ is the identity substitution (\ruleref{n/sub/id}).
+% where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
 
@@
 the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
-rather than \emph{existentially} quantified as is in \ruleref{n/tm/pair}---%
+rather than \emph{existentially} quantified as is in \ruleref{tm/pair}---%
 over how the fields in $t$'s context are defined.
@@
 
-\input{stlc-linkage-typing.tex}
+\input{stlc-linkage-typing-named.tex}
 
@@
 \Rule[name=Subst]
-{\goodType{\Delta}{T}{j} 
+{\goodType{\Delta}{T}{} 
   \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodType{\Gamma}{T[\gamma]}{j}},
+{\goodType{\Gamma}{T[\gamma]}{}},
 \Rule[]
@@
 {\goodWSig{\Gm}{\wsig}{n}
-  \quad \goodType{\Gm}{A}{i}
-  \quad \goodType{\Gm, A}{B}{i}}
+  \quad \goodType{\Gm}{A}{}
+  \quad \goodType{\Gm, A}{B}{}}
 {\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}}
@@
 {\goodWSig{\Gm}{\wsig}{n}}
-{\goodType{\Gm}{\TyWSingle {\wsig}}{i}
+{\goodType{\Gm}{\TyWSingle {\wsig}}{}
 \\ \goodTerm{\Gm}{\wsingle{\wsig}}{\TyWSingle{\wsig}}
@@
 {\goodSig{\Gm}{\lsig}{n}}
-{\goodType{\Gm}{\TyLkg{\lsig}}{i}
+{\goodType{\Gm}{\TyLkg{\lsig}}{}
 \\ \goodType{\Gm}{\TyTkg{\lsig}}{}}
@@
  \quad \goodSeal{\Gm}{\seal}{\lsig}{A}
- \quad \goodType{\Gm, A}{T}{i}}
+ \quad \goodType{\Gm, A}{T}{}}
 {\goodSig{\Gm}{(\LSigAdd {\lsig}{\seal}{T})}{n+1}}

### 2023-03-13 metatheory-condensed.tex
comments

@@
 Dependent function types \nTyPi{x}{A}{B},
-dependent product types \nTySigma{x}{A}{B}, and
+dependent pair types \nTySigma{x}{A}{B}, and
 identity types \TyId{t_1}{t_2} are standard.
@@
 in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
-For each derivation (of each judgment) we introduce judgmental equality
-(of the form \goodTerm{\Gm}{t_1 \equiv t_2}{T})
+For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}
+we introduce judgmental equality (of the form \goodTerm{\Gm}{t_1 \equiv t_2}{T})
 indicating convertibility.
@@
 There is an infinite hierarchy of universes;
-we omit universe levels in the presentation for conciseness.
+we omit universe levels in the presentation to avoid clutter.
 %\rA{Why is W(t) a term and not a type?}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}
@@
 % We use intrinsically typed syntax \cite{church1940}---terms and types are well-formed by construction.
-For conciseness, typing rules omit obvious premises required for well-formedness.
-For example, \ruleref{tm/lam} implicitly requires that the context be well-typed
-($\goodCtx{\Gm}{}$).
+For concision, typing rules omit obvious premises required for well-formedness.
+For example, \ruleref{tm/lam} implicitly requires $\goodCtx{\Gm}{}$
+(i.e., that the context be well-typed).
 
@@
 inductive types in MLTT.
-Together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
+Together with the identity type, they can express a
 whole host of inductive types~\cite{hugunin2021not}.
@@
 Rule~\ruleref{tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
-$\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$ models the $i$-th constructor---given 
+$\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$\YZ{this is still in de Brujin}
+models the $i$-th constructor---given
 two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
@@
 we can construct a term $\Wsup{i}{\wsig}{t_1}{\mathit{self}.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
-W-type (\ruleref{tm/w}).
+W-type (\ruleref{tm/w}).\YZ{What's the rationale behind naming the free variable 'self'?}
 

### 2023-03-13 metatheory2.tex
Fix issue on variables.
Save. Adding universe.

@@
 substitutions~$\gm$ and their applications (e.g., in the form of \sub{T}{\gm}) are part
-of the syntax rather than meta-operations.
+of the syntax rather than meta-operations. What's more, we also make universe level explicit for completeness.
 
@@
 
-\subsection{Review MLTT in Debruijn Indices and Explicit Subst}
+\subsection{Review MLTT in Debruijn Indices, Explicit Subst, and Explicit Universe Level}
 
@@
 
-The change of the syntax formulation is important to the simplicity of our proof, as we want to handle the proof about substitutions and terms uniformly---thus making substitution as part of the syntax (explicit substitution) will be helpful.
+\noindentparagraph{Debruijn Indices and Explicit Substitution.}  The change of the syntax formulation is important to the simplicity of our proof, as we want to handle the proof about substitutions and terms uniformly---thus making substitution as part of the syntax (explicit substitution) will be helpful.
 
@@
 Substitutions are typed with the form 
-%\EDJ{It seems to me G |- g : D is a judgement not a rule.
-Rules usually have two parts (premises + conclusion).
-Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here).
-I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this.
-In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
 $\goodSub{\Gm}{\gm}{\Dl}$.
@@
 where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
+
+Finally, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$. Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.
+
+\noindentparagraph{Explicit Universe Level.} Our treatment of universe level mainly follows that of \citet{altkap2016, kaposi2019gluing}.
+Universe level is mainly to get around the size issue---we cannot refer to(and thus quantifiy over) a set of all set (or a universe of all type).
+So everytime we want to quantify over a universe of types, we have to specify ``how large'' that universe is, that is basically what universe level is measuring.
+
+% We first need the mathmatical setup to talk about universe level, following \citet{sterling2019algebraic}:
+
+% \newcommand{\Set}[1]{\ensuremath{\texttt{Set}_{#1}}}
+
+% {\textbf{Set-theoretic Universe Assumption.}} As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our meta-logic.
+
+% For people familiar with Agda, we can consider $\Set{i}$ as the one in Agda.
+In other words, each $\Set{i}$ is (1).
+closed under dependent function type and dependent tuple type, (2).
+cumulative as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$
+
+Now we will need to attach syntax pieces with natural number indicating universe level.
+The explicit universe level $i$ now will attach to each context and type judgement $\goodCtx{\Gm}{i}$ and $\goodType{\Gm}{T}{i}$ basically indicating the context $\Gm$ and $T$ inhabits at what universe.
+Apparently, universe level will also attach to $\cU$.
+For two universe level $i,j$ (as natural number), we use $i \lcup j$ to denotes $\max(i,j)$.
 

### 2023-03-14 metatheory-condensed.tex
Reply the comments.

@@
 in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
-For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}
+For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement.
+It applies to each kind of judgement, for example, (G |- ), which looks like (G : Con), (G |-) is the derivation and (_ |- )is the judgement \\
+Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation.
+Similar for ( _ |- ) vs.
+(G |-) and (G |- _ ) vs (G |- T).
+\\
+I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e.
+there are no two different derivations lead to the same t; for each t, there is only one derivation for it).
+This is not true for extrinsic typing.
+\\
+If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
+I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
+}
 we introduce judgmental equality (of the form \goodTerm{\Gm}{t_1 \equiv t_2}{T})
@@
 Rule~\ruleref{tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
-$\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$\YZ{this is still in de Brujin}
+$\goodType{\Gm,x : \wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$\YZ{this is still in de Brujin}\EDJreply{Changed}
 models the $i$-th constructor---given
 two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
-$\goodTerm{\Gm, \mathit{self} : \nsub{\wsigproj{i}{2}{\wsig}}{x}{t_1}}{t_2}{\El{\wcode{\wsig}}}$,
-we can construct a term $\Wsup{i}{\wsig}{t_1}{\mathit{self}.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
-W-type (\ruleref{tm/w}).\YZ{What's the rationale behind naming the free variable 'self'?}
+$\goodTerm{\Gm, x : \nsub{\wsigproj{i}{2}{\wsig}}{x}{t_1}}{t_2}{\El{\wcode{\wsig}}}$,
+we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
+W-type (\ruleref{tm/w}).\YZ{What's the rationale behind naming the free variable 'self'?}\EDJreply{Self is not making sense, changed sorry.}
 

### 2023-03-14 metatheory2.tex
Attach universe level to MLTT.

@@
 
-Now we will need to attach syntax pieces with natural number indicating universe level.
+Now we will need to some attach syntax pieces with natural number indicating universe level.
 The explicit universe level $i$ now will attach to each context and type judgement $\goodCtx{\Gm}{i}$ and $\goodType{\Gm}{T}{i}$ basically indicating the context $\Gm$ and $T$ inhabits at what universe.
-Apparently, universe level will also attach to $\cU$.
+Apparently, universe level will also be attached to $\cU$.
 For two universe level $i,j$ (as natural number), we use $i \lcup j$ to denotes $\max(i,j)$.
+
+Since we work in an intrinsic-typed setting, some required premises can be omitted without ambiguity.
+For omitted context judgement $\goodCtx{\Gm}{k}$, the universe level is always $k$, hinted at the figure of judgement specification.
+Similar for type judgement $\goodType{\Gm}{T}{j}$.
 

### 2023-03-14 metatheory-condensed.tex
edits

@@
 %
-Our enhanced W-type is decided by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
-composed of a list (\ruleref{wsig/empty} and \ruleref{wsig/add}) of $n$ pairs of 
-types $\{(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})\}_i$,
-modeling an inductive type of $n$ constructors. 
-%
-Rule~\ruleref{tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
-$\goodType{\Gm,x : \wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$\YZ{this is still in de Brujin}\EDJreply{Changed}
-models the $i$-th constructor---given
-two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
+In our formulation, a W-type is given by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
+composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
+%
+Each pair models a constructor of the inductive type.
+The $i$-th pair, projected from the signature~$\wsig$ using the forms
+$\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and
+$\goodType{\Gm,x : \wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$,
+defines the $i$-th constructor.
+That is, given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
 $\goodTerm{\Gm, x : \nsub{\wsigproj{i}{2}{\wsig}}{x}{t_1}}{t_2}{\El{\wcode{\wsig}}}$,
-we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
-W-type (\ruleref{tm/w}).\YZ{What's the rationale behind naming the free variable 'self'?}\EDJreply{Self is not making sense, changed sorry.}
+we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type (\ruleref{tm/wsup}).
+$\wcode{\wsig}$ gives the code of the
+W-type (\ruleref{tm/w}).
 

### 2023-03-14 metatheory-condensed.tex
add comments from Slack

@@
 in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
-For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement.
+Definitional equalities have the form
+\goodTerm{\Gm}{t_1 \equiv t_2}{T} and \goodType{\Gm}{T_1 \equiv T_2}{}.\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.
+Maybe this is something require emphasis?
+Because it is usually handled by LF (like QIIT) but now we are using bare hands so....
+maybe we need to emphasize this}
+%For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement.
 It applies to each kind of judgement, for example, (G |- ), which looks like (G : Con), (G |-) is the derivation and (_ |- )is the judgement \\
-Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation.
+%Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation.
 Similar for ( _ |- ) vs.
@@
 \\
-I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e.
+%I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e.
 there are no two different derivations lead to the same t; for each t, there is only one derivation for it).
@@
 \\
-If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
-I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
-}
-we introduce judgmental equality (of the form \goodTerm{\Gm}{t_1 \equiv t_2}{T})
-indicating convertibility.
+%If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
+%I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
+%}
 

### 2023-03-14 metatheory-condensed.tex
Add comments on convertibility.

@@
 in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
-Definitional equalities have the form
-\goodTerm{\Gm}{t_1 \equiv t_2}{T} and \goodType{\Gm}{T_1 \equiv T_2}{}.\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.
+Definitional equalities (for each judgement) have the form
+\goodTerm{\Gm}{t_1 \equiv t_2}{T},\goodType{\Gm}{T_1 \equiv T_2}{},\goodCtx{\Gm_1 \equiv \Gm_2}{} ..., etc.
+We consider our syntax to be quotiented by these definition equalities, and thus each syntax piece is actually one (equivalent) class of syntax.
+Doing such quotient gives us convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient, the derivation of $t : A$ is the same as the derivation of $t : B$.
+\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.
 Maybe this is something require emphasis?

### 2023-03-14 metatheory2.tex
Add comments on convertibility.

@@
 
-We will make sure all the introduced models and proofs respecting judgemental equality. In other words, we will consider our syntax as the result of taking quotient by these judgemental equality.
+Again, following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}, we consider our syntax as the result of taking quotient by all of these judgemental equalities. This means 
+(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.
 

### 2023-03-14 metatheory2.tex
Attach universe level. And Fix.

@@
 
-Finally, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$. Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.
+To simplify, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$. Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.
+
+Finally, we have \ruleref{sub/dbj/shift} defined using \ruleref{sub/wk} and \ruleref{tm/var}.
+This rule applying substitution $\gm$ on the earlier part of the context.
+We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.
 

### 2023-03-14 canonicitymodel.tex
Add details about Grothendieck Universe

@@
 
-Like most logical relation proof,  we interpret each judgement, inductively interpret each syntax piece.
-We are working on an intrinsic setting thus even we omit context for brevity, the syntax piece is still well-typed.
-We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
+We first need the mathmatical setup to interpret universe level, following \citet{sterling2019algebraic}:
+
+\newcommand{\Set}[1]{\ensuremath{\texttt{Set}_{#1}}}
+
+{\textbf{Set-theoretic Universe Assumption.}} As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our meta-logic.
+
+For people familiar with Agda, we can roughly consider Grothendieck Universe $\Set{i}$ as the $\Set{i}$ in Agda. In other words, each $\Set{i}$ is 
+\begin{itemize}
+  \item Closed under dependent function type and dependent tuple type. For example, when we have $\denotesC{A}, \denotes{B} \in \Set{i}$, we will have $\denotes{\TyPi{A}{B}} \in \Set{i}$
+  \item the hierarchy is cumulative as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$
+\end{itemize}
+
+Like most logical relation proof,  we interpret each judgement and inductively interpret each syntax piece.
+We are working on an intrinsic setting thus even if we omit context for brevity, the syntax piece is still well-typed.
+% We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
 Interested audience can refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} for a complete technical treatements.

### 2023-03-14 syn-translate.tex
Fixing syn translation

@@
 \begin{align*}
-  \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
-  \text{such that } &\goodTerm{\Gm}{\denotesT{\goodSig{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}} \\
-  \text{We make } & ({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ mutually recursive, } \\
+  \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
+  \text{such that } &\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}} \\
+  \text{We make } & ({\Sigr{\Gamma}{n}{}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ mutually recursive, } \\
   & \text{defined by induction on $n$ and the signature length} \\  
@@
   % &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
-  {\Sigr{\Gamma}{(n+1)}} &= 
-    \Sigma~(\Sigr{\Gamma}{n})
-          ~(\Sigma~\cU
-                  ~(\Sigma~(\Pi(\denotesT{\TyTkg {("var"_0[p^1])}})(El~"var"_0[p^1]))~\cU)) \\
-  \Sigr{\Gamma}{0} &= \top \\
+  {\Sigr{\Gamma}{(n+1)}{j}} &= 
+    \Sigma~(\Sigr{\Gamma}{n}{})
+          ~(\Sigma~\cU_j \\
+          & \quad \quad ~(\Sigma~(\TyPi{\denotesT{\TyTkg {(\var{0}[\SubstWeak{1}])}}}{\El{\sub{\var{0}}{\SubstWeak{1}}}}) \\
+          & \quad \quad \quad \TyPi{\El{\sub{\var{0}}{\SubstWeak{1}}}}{\cU_j})) \\
+  \Sigr{\Gamma}{0}{j} &= \top \\
   \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}}, \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ defines upon } \denotesT{\sigma} 
@@
   \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= 
-    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(El \ (\app{("pjr"^3~\denotesT{\sigma})})[(p^1, \app{(\fst{("pjr"^2~{\denotesT{\sigma}})})})]) \\
+    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(\El{(\app{("pjr"^3~\denotesT{\sigma})})[(\SubstWeak{1}, \app{(\fst{("pjr"^2~{\denotesT{\sigma}})})})]}) \\
   \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &=
-  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(El~(\fst{(\snd{\denotesT{\sigma}})}))(El~(\app{("pjr"^3~\denotesT{\sigma})} )) \\
+  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(\El{(\fst{(\snd{\denotesT{\sigma}})})})(\El{(\app{("pjr"^3~\denotesT{\sigma})} )}) \\
   &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
-  \denotesT{\goodSig{\Gamma}{\LSigEmp}{0}} &= \unit \\ 
-  \denotesT{\goodSig{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{(\codety{T})}) \\ 
+  \denotesT{\goodSigU{j}{\Gamma}{\LSigEmp}{0}} &= \unit \\ 
+  \denotesT{\goodSigU{j}{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{(\codety{T})}) \\ 
   \denotesT{\goodTerm{\Gamma}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \unit \\ 
@@
 However, when we mutual recursively define
-$\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}} $, and
+$\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}{}} $, and
 $\denotesT{\cP}$ above, we actually have to prove the substitution laws
-$\sub{(\TyTkg{\lsig})}{\gm}\equiv \TyTkg{\sub{\lsig}{\gm}}$ and $\sub{(\Sigr{\Delta}{n})}{\gm}
-\equiv \Sigr{\Gamma}{n}$ together. 
+$\sub{(\TyTkg{\lsig})}{\gm}\equiv \TyTkg{\sub{\lsig}{\gm}}$ and $\sub{(\Sigr{\Delta}{n}{})}{\gm}
+\equiv \Sigr{\Gamma}{n}{}$ together. 
 % The details (of all the equational rules) can be found in the appendices written in (pseudo-)Agda form.

### 2023-03-14 canonicitymodel.tex
Fixing syn translation
Shift back to original version of the canonicity model.

@@
 \newcommand{\Set}[1]{\ensuremath{\texttt{Set}_{#1}}}
+\newtheorem{assumption}{Assumption}[section]
 
-{\textbf{Set-theoretic Universe Assumption.}} As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our meta-logic.
+\begin{assumption}[Set-theoretic Universe Assumption] As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our ambient meta-logic.
+\end{assumption}
 
-For people familiar with Agda, we can roughly consider Grothendieck Universe $\Set{i}$ as the $\Set{i}$ in Agda. In other words, each $\Set{i}$ is 
+We can roughly consider each Grothendieck Universe $\Set{i}$ as a $\Set{i}$ in Agda. In other words, each $\Set{i}$ is 
 \begin{itemize}
-  \item Closed under dependent function type and dependent tuple type. For example, when we have $\denotesC{A}, \denotes{B} \in \Set{i}$, we will have $\denotes{\TyPi{A}{B}} \in \Set{i}$
+  \item Closed under dependent function type and dependent tuple type.
+For example, later, for our interpretation of dependent function type, when we have $\denotesC{A}, \denotesC{B} \in \Set{i}$, we will have $\denotesC{\TyPi{A}{B}} \in \Set{i}$
   \item the hierarchy is cumulative as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$
@@
 \begin{align*}
-  \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
+  \denotesC{\goodCtx{\Gm}{k}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
   % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
-  \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
-  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
+  \denotesC{\goodType{\Gm}{T}{j}} & \text{ is a function maps to $\Set{j}$ s.t.} \\
+  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a an element of } \Set{j}  \\ 
   % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
@@
   \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
+  \denotesC{\goodType{\Gm}{\cU_j}{j}}(\gm) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\} \quad \text{     for some } \goodTerm{\cdot}{T}{\cU_j}  \\
+  \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm) &= \denotesC{T}(\gm) \\ 
+  \denotesC{\goodType{\Gm}{\El{T}}{j}}(\gm) &= \denotesC{T}(\gm) \\
 \end{align*}

### 2023-03-14 metatheory-condensed.tex
question

@@
 We consider our syntax to be quotiented by these definition equalities, and thus each syntax piece is actually one (equivalent) class of syntax.
-Doing such quotient gives us convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient, the derivation of $t : A$ is the same as the derivation of $t : B$.
-\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.
+Doing such quotient gives us convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient, the derivation of $t : A$ is the same as the derivation of $t : B$.\YZ{Since you are talking about quotienting, should you mention that the underlying logical framework is QIIT?}
+%\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.
 Maybe this is something require emphasis?

### 2023-03-14 canonicitymodel.tex
Working on canonicity model.

@@
 We are working on an intrinsic setting thus even if we omit context for brevity, the syntax piece is still well-typed.
+However, unlike most logical relation proof, we are having a proof-relevant logical realtion, essential for modeling universes properly~\cite{coquand2018canonicity}.
 % We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
@@
 \begin{align*}
-  \denotesC{\goodCtx{\Gm}{k}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
+  \denotesC{\goodCtx{\Gm}{k}} &= \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k} \quad \text{ i.e., sets indexed by closed substitution}\\
   % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
-  \denotesC{\goodType{\Gm}{T}{j}} & \text{ is a function maps to $\Set{j}$ s.t.} \\
-  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a an element of } \Set{j}  \\ 
+  \denotesC{\goodType{\Gm}{T}{j}} & \text{ is a function maps to Set, s.t. for arbitrary } \goodSub{\cdot}{\gm}{\Gm} , \\
+  & \text{ and arbitrary } \goodTerm{\cdot}{t}{\sub{T}{\gm}}, \quad \text{ we have } \denotesC{\goodType{\Gm}{T}{}}(\gm)(t) \in \Set{j}  \\ 
   % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
@@
 } \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
-  \denotesC{\goodType{\Gm}{\top}{}}(\gm) &= \{t : t \equiv \unit \} \\
+  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(t) &= \{\star \ | \text{  when } t \equiv \unit \} \\
   \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
@@
 \newcommand{\GCaseSig}[3]{\ensuremath{\texttt{CaseTy}^c({#1},{#2},{#3})}}
+
+The counter-intuitive part of the proof-relevant logical relation is the treatment of proposition---a proposition is encoded as a subsingleton data.
+For example, $\denotesC{\goodType{\Gm}{\top}{}} = $ $ \{\star \ | \text{  when } t \equiv \unit \}$ represents a singleton set when $t \equiv \unit$ or an empty set otherwise.
+However, not all the denotation of types are subsingleton---the salient counter-examples are $\denotes{\goodType{\Gm}{\cU}{}}$ and $\denotes{\goodType{\Gm}{\TyPi{A}{B}}{}}$.
+The non-trivial interpretation on these types is the main reason we use proof-relevant logical relations.
 

### 2023-03-15 canonicitymodel.tex
Save.

@@
 \begin{align*}
-  \denotesC{\goodCtx{\Gm}{k}} &= \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k} \quad \text{ i.e., sets indexed by closed substitution}\\
+  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k} \quad \text{ (also sets indexed by closed substitution)}\\
   % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
-  \denotesC{\goodType{\Gm}{T}{j}} & \text{ is a function maps to Set, s.t. for arbitrary } \goodSub{\cdot}{\gm}{\Gm} , \\
-  & \text{ and arbitrary } \goodTerm{\cdot}{t}{\sub{T}{\gm}}, \quad \text{ we have } \denotesC{\goodType{\Gm}{T}{}}(\gm)(t) \in \Set{j}  \\ 
+  \denotesC{\goodType{\Gm}{T}{j}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\{t : \goodTerm{\cdot}{t}{\sub{T}{\gm}}\}\to\Set{j} \\
   % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
-  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t.
-} \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
+  \denotesC{\goodSub{\Gm}{\delta}{\Delta}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)} \denotesC{\goodCtx{\Dl}{}}(\delta \circ \gm) \\
   % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
-  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t.
-} \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
-  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(t) &= \{\star \ | \text{  when } t \equiv \unit \} \\
-  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
-  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
-  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
-  &= \emptyset \text{ otherwise } \\ 
+  \denotesC{\goodTerm{\Gm}{t}{T}} \text{ is a dependent function}&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})\\
+  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \ | \text{  when } t \equiv \unit \} \\
+  \denotesC{\goodType{\Gm}{\bot}{}}(\gm)(\gm')(t) &= \emptyset \\
+  \denotesC{\goodType{\Gm}{\cB}{}}(\gm)(\gm')(t) &= \{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \ | \text{ when } t \equiv \false  \} \quad \text{disjoint union} \\
+  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \\
   \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
   \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
-  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
-  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
-  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm) &= \unit\\
-  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm) &= \true \\
-  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm) &= \false \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{a}(\gm) \text{ if } \denotes{c}(\gm) \equiv \true \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{b}(\gm) \text{ if } \denotes{c}(\gm) \equiv \false \\
-  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm) &= \eqrefl{\sub{t}{\gm}}\\
+  \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 
+  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_2 \gm_t)  \} \\
+  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm)(\gm') &= \star\\
+  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm)(\gm') &= \star^1 \\
+  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm)(\gm') &= \star^2 \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{a}(\gm) \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{b}(\gm)(\gm') \text{ if } \denotes{c}(\gm)(\gm') = \star^2 \\
+  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm)(\gm') &= \star\\
   \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
@@
   \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
-  \denotesC{\goodType{\Gm}{\cU_j}{j}}(\gm) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\} \quad \text{     for some } \goodTerm{\cdot}{T}{\cU_j}  \\
-  \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm) &= \denotesC{T}(\gm) \\ 
-  \denotesC{\goodType{\Gm}{\El{T}}{j}}(\gm) &= \denotesC{T}(\gm) \\
+  \denotesC{\goodType{\Gm}{\cU_j}{j+1}}(\gm)(\gm')(T) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\}  \to \Set{j} \\
+  \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm)(\gm') &= \denotesC{T}(\gm)(\gm') \\ 
+  \denotesC{\goodType{\Gm}{\El{T}}{j}}(\gm)(\gm')(t) &= \denotesC{T}(\gm)(\gm')(t) \\
 \end{align*}
@@
 The first observation is that, with this model, we know our system is consistent---we cannot derive $\goodTerm{\cdot}{t}{\bot}$.
-Otherwise, we have $\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \denotesC{\bot}(\star) = \emptyset$, that is an element in an empty set, a contradiction.
+Otherwise, we have $\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp)(\star) \in \denotesC{\bot}(\SubstEmp)(\star)(\sub{t}{\gm}) = \emptyset$, that is an element in an empty set, a contradiction.
 
@@
 
-What's more, with the above model, we can map an arbitrary closed term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
+What's more, with the above model, we can map an arbitrary closed boolean term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
 

### 2023-03-15 canonicitymodel.tex
Save.

@@
 
-What's more, with the above model, we can map an arbitrary closed boolean term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
+What's more, with the above model, we can map an arbitrary closed boolean term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp)(\star) = \star^1$ or $\star^2$, witnessing $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
 

### 2023-03-15 metatheory-condensed.tex
add a comment from Zhixuan

@@
 In our formulation, a W-type is given by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
-composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
+composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types".
+Since there are already boolean types and Σ types, Fin types can be defined.
+Using Fin, ordinary W types suffice to define trees with n constructors.
+So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}
 %

### 2023-03-15 metatheory-condensed.tex
incorporate some of Zhixuan's comment

@@
 
-Well-formedness judgments of \emph{linkage signatures} have the form
+Well-formed \emph{linkage signatures} have the judgment form
 $\goodSig{\Gm}{\lsig}{n}$.
@@
 %
-Well-formedness judgments of \emph{linkages} have the form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
+Well-formed \emph{linkages} have the judgment form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
 where $\TyLkg{\lsig}$ is a type formed by~$\lsig$.
@@
 \fi
-using debruijn indices and explicit substitution.
+using de Bruijn indices and explicit substitution.
 An effort to mechanize the formalization is underway, by following  \citet{altkap2016} 

### 2023-03-15 canonicitymodel.tex
Save.

@@
 Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax.
-Meanwhile, our proof will still work directly on the earlier introdcued syntax for readability.
+Meanwhile, our proof will still work directly on the earlier introduced syntax for readability\footnote{We still take quotient, so ultimately our syntax should be considered as an instance of any of the above logical framework, without exposing the reader too many technical details}.
+
 We will first specify the formulation of the most basic canonicity theorem.
@@
 However, unlike most logical relation proof, we are having a proof-relevant logical realtion, essential for modeling universes properly~\cite{coquand2018canonicity}.
+Our canonicity model at MLTT part is basically the same as shown in \citet{coquand2018canonicity}.
 % We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
@@
   \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \\
-  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
-  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
+  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
+  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}}(\gm)(\gm')(t) &= \sum_{u' \in \denotesC{A}(\gm)(\gm')} \\
   \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 

### 2023-03-15 metatheory-condensed.tex
Reply the comments

@@
 Using Fin, ordinary W types suffice to define trees with n constructors.
-So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}
+So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}\EDJreply{This following sentence has been replied to Zhixuan and he doesn't has further response: 1.
+We have to have enhanced W-type at FMLTT, that's for sure.
+Otherwise we (a).
+are too far away from the surface syntax because W-type itself only has one constructor all the time (b) this suggested encoding or syntactic sugar, since it interact with other connective like linkage (the Wrec is the interaction), so Wrec part is mostly unchanged  (c) since they will look the same, the encoding itself will look like an extra equality judgement in the syntax...
+which is not helpful for the complexity at all ...
+and even more complex  \\
+2.
+so to me the problem you guys really have is around the content in the appendix -- I think Zhixuan mean he doesn't think this is more expressive.
+But my point is, it is hard to really write out the above syntactic sugar, formally.
+I think the issue here is similar to other reviewer saying "syntax translation".
+It is very hard to write out the encoding of Fin already, and then encode it again into W-type.
+(Just ask my agda mechanization).
+Of course, some hand-waiving can be done.
+For example, the part Zhixuan says I think it okay to hand-wave.
+But the destruction part -- the part Wrec interact with this encoding, I don't believe handwaving is enough.
+\\
+3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
 %

### 2023-03-15 related-work.tex
incorporate some of Zhixuan's comment
took a pass over related work sec

--- 2023-03-04 related-work.tex
@@
 for the underlying representation of type-level fixed points and
-avoid \lsti{Set} impredicativity.\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}\EDJreply{My understanding is,
- it is really an intrinsic typed version of MTC.
+avoid \lsti{Set} impredicativity.
+%\Citet{vdr2022intrinsically} mechanize intrinsically-typed
+%target language based on container and signature \cite{altenkirch2015container}. 
+%Thanks to the nature of intrinsically-typed abstract syntax trees (ASTs) and Agda's advanced unification mechanism, the authors were able to eliminate a significant amount of code that dealt with "impossible" cases.
+%\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}\EDJreply{My understanding is,
+% it is really an intrinsic typed version of MTC.
 It uses different set of tools (like containers (a la altenkirch, can be considered as an alternative of W-type) and stuff) but it is still a design pattern paper.
- But good news is, container is closer to Wtype(inductive type) than the endofunctor encoding (classic MTC approach).}
-%\cite{vdr2022intrinsically}
+% But good news is, container is closer to Wtype(inductive type) than the endofunctor encoding (classic MTC approach)}\EDJreply{I also add one sentence after Coq a la carte.
+You can also check...
+I will add bib later.
+Interestingly, Intrinsic type doesn't refer Coq a la carte.}
 
@@
 For example, MTC~\cite{delaware2013} has been applied to implementing
-composable program adverbs~\cite{liwei2022} and been adapted to support
+composable program adverbs~\cite{liwei2022} and been adapted to allow
 unanticipated feature extensions that require type changes
-\cite{delaware2013b}.
+\cite{delaware2013b,vdr2022intrinsically}.
 
@@
 by invoking the tool.
-%
 %The approach does not seem to fully address the expression problem,
@@
 
-\noindentparagraph{Solutions to the expression problem} \!abound.
+\noindentparagraph{The expression problem.} Solutions abound.
 % \EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}\YZreply{Yes}
@@
 Our approach is the first that applies family polymorphism~\cite{ernst2001family},
-an idea from OO research,
+mostly seen in OO languages,
 to the context of mechanized proofs.

### 2023-03-15 related-work.tex
tweak

@@
 be more easily adapted for new purposes.
-For example, MTC~\cite{delaware2013} has been applied to implementing
+For example, MTC has been applied to implementing
 composable program adverbs~\cite{liwei2022} and been adapted to allow
-unanticipated feature extensions that require type changes
+feature extensions, such as reference cells and exceptions, that require type changes
 \cite{delaware2013b,vdr2022intrinsically}.

### 2023-03-15 metatheory-condensed.tex
More comments from Zhixuan

@@
 quantified over the earlier ones, a linkage component is \emph{universally}
-quantified over---and thus polymorphic to---the components preceding it.
+quantified over---and thus polymorphic to---the components preceding it.\Zhixuan{explain the intuition behind the (typing rule of) linkages, and try to connect it to the ``family'' in the earlier section/surface syntax}\EDJreply{I am not sure if more is needed because the section ``Family polymorphism via linkages'' is trying to do this.
+Maybe I should estabilish connection between Rule l/add with the compilation process of fam poly?}
 

### 2023-03-15 canonicitymodel.tex
Fix Logical Relation back to proof relevant one.

@@
   \denotesC{\goodTerm{\Gm}{t}{T}} \text{ is a dependent function}&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})\\
-  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \ | \text{  when } t \equiv \unit \} \\
+  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \} \\
   \denotesC{\goodType{\Gm}{\bot}{}}(\gm)(\gm')(t) &= \emptyset \\
@@
 
-\newcommand{\Glued}[1]{\ensuremath{{#1}^c}}
-\newcommand{\GluedPi}[2]{\ensuremath{\Pi^c({#1},{#2})}}
-\newcommand{\GSubstExt}[2]{\ensuremath{{#1},^c{#2}}}
-\newcommand{\Gpair}[2]{\ensuremath{({#1},^c{#2})}}
-\newcommand{\Gfst}[1]{\ensuremath{\texttt{fst}^c~{#1}}}
-\newcommand{\Gsnd}[1]{\ensuremath{\texttt{snd}^c~{#1}}}
-\newcommand{\Gsub}[2]{\ensuremath{{#1}\!\left[{#2}\right]^c}}
-\newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^c({#1})}}
-\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^c)^{#1}}}
-\newcommand{\GLSigAdd}[3]{\ensuremath{\nu^{+c}({#1},{#2},{#3})}}
-\newcommand{\GCaseSig}[3]{\ensuremath{\texttt{CaseTy}^c({#1},{#2},{#3})}}
+\newcommand{\Glued}[1]{\ensuremath{{#1}^\bullet}}
+\newcommand{\GluedPi}[2]{\ensuremath{\Pi^\bullet({#1},{#2})}}
+\newcommand{\GSubstExt}[2]{\ensuremath{{#1},^\bullet{#2}}}
+\newcommand{\Gpair}[2]{\ensuremath{({#1},^\bullet{#2})}}
+\newcommand{\Gfst}[1]{\ensuremath{\texttt{fst}^\bullet~{#1}}}
+\newcommand{\Gsnd}[1]{\ensuremath{\texttt{snd}^\bullet~{#1}}}
+\newcommand{\Gsub}[2]{\ensuremath{{#1}\!\left[{#2}\right]^\bullet}}
+\newcommand{\Glam}[1]{\ensuremath{\lambda^\bullet({#1})}}
+\newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^\bullet({#1})}}
+\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^\bullet)^{#1}}}
+\newcommand{\GLSigAdd}[3]{\ensuremath{\nu^{+\bullet}({#1},{#2},{#3})}}
+\newcommand{\GCaseSig}[3]{\ensuremath{\texttt{CaseTy}^\bullet({#1},{#2},{#3})}}
+\newcommand{\Gmodel}[1]{\ensuremath{{({#1})}^c}}
 
@@
 
-Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t.
-if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\GluedPi{A'}{B'} = \denotesC{\TyPi{A}{B}}$.
-Type-theoretical speaking, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model.
+Given above model for MLTT, we should have such a function ${\Pi}^c$ s.t.
+${\Pi}^c({\denotesC{A}},{\denotesC{B}}) = \denotesC{\TyPi{A}{B}}$.
+We want to reuse this function when defining the rest of the model of \TT, because as syntactic translation hinted, \TT reuses a lot from MLTT.
+Type-theoretical speaking, this ${\Pi}^c$ is using the \emph{internal dependent function type} of the above model.
+However this is not possible because the definition $\denotesC{\TyPi{A}{B}}$ is not solely based on the $\denotesC{A}$ and $\denotesC{B}$, but also on the syntax $\goodType{\Gm}{A}{}$ and $\goodType{\Gm}{B}{}$.
+
+Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$, that will return the syntax piece $S$ as well.
+Thus we can have a function $\Glued{\Pi}$ s.t.
+$\GluedPi{\denotesCC{A}}{\denotesCC{B}} = \denotesCC{\TyPi{A}{B}}$ now that syntax is also provided.
 Similarly, we can have $\Glued{\Sigma}$, $\Gpair{a}{b}$  and $\GSubstExt{\gm}{t}$  for dependent tuple type, depende tuple construction and substitution extension, and also all other constructions.
-These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.
+What's more, given $\Glued{S}$ a pair of syntax and model , we denote $\Gmodel{\Glued{S}}$ as getting the (canonicity) model part (the second element in that pair).
+% These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.
 What's more, during the construction of the model for \TT, we will use $\Glued{(\cdot)}$ when we can.
@@
   & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
-  \mathit{L}("nil") &= \denotesC{\top} \\
-  \mathit{L}((A,s,T) :: tl)(\gm) &= {\mathit{L}(tl)}(\gm) \times \GluedPi{A}{T}(\gm) \quad \text{($\times$ is a special case for $\Glued{\Sigma}$)}  \\
-  \mathit{P}("nil") &= \denotesC{\top}\\
-  \mathit{P}((A,s,T) :: tl) &= \Glued{\Sigma}(\mathit{P}(tl), \Gsub{T}{\GSubstExt{\Glued{\pi_1}}{s}}) \quad \text{(doing substitution on $T$)} \\
-  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{L}(\denotesC{\lsig}) \\
-  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{P}(\denotesC{\lsig}) \\
+  \Glued{\mathit{L}}("nil") &= ( \TyLkg{\LSigEmp} , \denotesC{\top} ) \\
+  \Glued{\mathit{L}}((A,s,T) :: tl) &= {\Glued{\mathit{L}}(tl)}(\gm) \times^\bullet \GluedPi{A}{T} \quad \text{($\times^\bullet$ is a special case for $\Glued{\Sigma}$)}  \\
+  \Glued{\mathit{P}}("nil") &= ( \TyTkg{\LSigEmp} , \denotesC{\top} )\\
+  \Glued{\mathit{P}}((A,s,T) :: tl) &= \Glued{\Sigma}(\Glued{\mathit{P}}(tl), \Gsub{T}{\GSubstExt{\Glued{\pi_1}}{s}}) \quad \text{(doing substitution on $T$)} \\
+  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \Gmodel{\Glued{\mathit{L}}(\denotesC{\lsig})} \quad \text{ discard syntax info} \\
+  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \Gmodel{\Glued{\mathit{P}}(\denotesC{\lsig})} \\
   % 
   \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
-  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesC{A}, \denotesC{s}, \denotesC{T})::\denotesC{\lsig}\\ 
+  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesCC{A}, \denotesCC{s}, \denotesCC{T})::\denotesC{\lsig}\\ 
   \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
-  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0] \quad \text{ take the first element in the tuple}\\ 
-  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= ("hd"~\denotesC{\lsig})[1]\\ 
-  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[2]\\
+  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= \Gmodel{("hd"~\denotesC{\lsig})[0]} \quad \text{ take the first element in the tuple,...}\\ 
+  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= \Gmodel{("hd"~\denotesC{\lsig})[1]} \quad \text{  ... and discard syntax}\\ 
+  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= \Gmodel{("hd"~\denotesC{\lsig})[2]}\\
   \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
-  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm) &= (\denotesC{\lkg}(\gm), \lambda u . t (\gm, u)) \\
+  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}} &= \Gmodel{\Gpair{\denotesCC{\lkg}}{\Glam{\denotesCC{t}}}} \\
   \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
   \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
-  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gpair{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{(\SubstId)}}{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}}}
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{(\SubstId)}}{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}}}}
   \\ &\text{ when} \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\
-  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyTkg{\lsig}}} &= \Gfst{\denotes{\lkg}} \\
-  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}} &= \Gapp{\Gsnd{\denotes{\lkg}}} \\
+  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gfst{\denotesCC{\lkg}}} \\
+  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}} &= \Gmodel{\Gapp{\Gsnd{\denotesCC{\lkg}}}} \\
   \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
-  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesC{A} , \denotesC{B}) :: \denotesC{\wsig}\\
-  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[0]\\
-  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[1]\\
+  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesCC{A} , \denotesCC{B}) :: \denotesC{\wsig}\\
+  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= \Gmodel{(\text{ $j$-th element of }~\denotesC{\wsig})[0]}\\
+  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= \Gmodel{(\text{ $j$-th element of }~\denotesC{\wsig})[1]}\\
   \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
@@
   \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= 
-  {\GluedPi{\denotesC{A}}{\GluedPi {\GluedPi{\denotesC{B}}{\Gsub{\denotesC{R}}{\GSubstWeak{2}}}} {\Gsub{\denotesC{R}}{\GSubstWeak{2}}}}} \\
+  \Gmodel{\GluedPi{\denotesC{A}}{\GluedPi {\GluedPi{\denotesC{B}}{\Gsub{\denotesC{R}}{\GSubstWeak{2}}}} {\Gsub{\denotesC{R}}{\GSubstWeak{2}}}}} \\
   % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\
   \text{we define } \RecSig{}{} &\text{ by induction on the signature} \\
-  RS~"nil"~R &= \denotesC{\LSigEmp} \\
+  RS~"nil"~R &= \denotesCC{\LSigEmp} \\
   RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\
-  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= RS~\denotesC{\tau}~\denotesC{R} \\
+  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{R}} \\
   \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg \\

### 2023-03-15 metatheory-condensed.tex
Add QIIT ref, Add Coquand Universe ref.

@@
 \goodTerm{\Gm}{t_1 \equiv t_2}{T},\goodType{\Gm}{T_1 \equiv T_2}{},\goodCtx{\Gm_1 \equiv \Gm_2}{} ..., etc.
-We consider our syntax to be quotiented by these definition equalities, and thus each syntax piece is actually one (equivalent) class of syntax.
+Following \citet{altkap2016}, we consider our syntax to be quotiented by these equalities, and thus each syntax piece is actually one (equivalent) class of syntax.
 Doing such quotient gives us convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient, the derivation of $t : A$ is the same as the derivation of $t : B$.\YZ{Since you are talking about quotienting, should you mention that the underlying logical framework is QIIT?}
@@
 
-We use Tarski-style universes \cite{hofmann1997syntax},
-which, unlike Russell-style ones,
-are not inhabited by types directly but rather by the \emph{codes} of types
-and arguably behave better \cite{luo2012notes}.
+Following \citet{kaposi2019gluing}, we use \emph{Universe à la Coquand} \cite{coquand2013presheaf}.
+Unlike Russell-style ones, Coquand universes are not inhabited by types directly but rather by the \emph{codes} of types
+and arguably behave better due to its closedness to Tarski-style ones~\cite{luo2012notes}. 
+% and compatibility with our algebraic formulation of MLTT~\cite{coquand2018canonicity}. % I am not sure if I want this sentence or not
 Term $\codety{T}$ encodes type~$T$ (\ruleref{tm/c}) and thus inhabits universe~$\cU$.

### 2023-03-15 coqimpl.tex
Add back limitation section and trusted code base section.

@@
 as the last line of \cref{fig:stlcfix-compiled} shows.
+
+\noindentparagraph{Trusted Code Base.} Finally we emphasize that our compilation can make sure the trusted code base is restricted to the Coq's kernel, as all the feature supported by our plugin has been compiled into Coq's surface syntax/Vernacular commands.
 
@@
 \begin{itemize}
-  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, notation for inductive type and parameters in the inductive type (we only support indexed inductive type right now)
+  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, syntax extensions for inductive type and parameters in the inductive type (we only support indexed inductive type right now)
   \item We don't support mutual fixpoint definition
@@
 \end{itemize}
-
-\fi

### 2023-03-15 metatheory-condensed.tex
Edit on QIIT para & a question

@@
 in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
-Definitional equalities (for each judgement) have the form
-\goodTerm{\Gm}{t_1 \equiv t_2}{T},\goodType{\Gm}{T_1 \equiv T_2}{},\goodCtx{\Gm_1 \equiv \Gm_2}{} ..., etc.
-Following \citet{altkap2016}, we consider our syntax to be quotiented by these equalities, and thus each syntax piece is actually one (equivalent) class of syntax.
-Doing such quotient gives us convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient, the derivation of $t : A$ is the same as the derivation of $t : B$.\YZ{Since you are talking about quotienting, should you mention that the underlying logical framework is QIIT?}
+Definitional equalities have the forms %(for each judgement)
+\goodCtx{\Gm_1 \equiv \Gm_2}{},
+\goodType{\Gm}{T_1 \equiv T_2}{},
+\goodTerm{\Gm}{t_1 \equiv t_2}{T},
+etc.
+Following \citet{altkap2016}, we regard our (intrinsically typed) syntax
+as being quotiented by these equalities.
+Quotienting facilitates coercion along equalities---given
+\goodCtx{\Gm_1 \equiv \Gm_2}{} and
+\goodType{\Gm_1}{T_1 \equiv T_2}{},
+the derivation of \goodTerm{\Gm_1}{t}{T_1} is considered definitionally equal to
+a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate?}
 %\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.

### 2023-03-15 metatheory-condensed.tex
Add mentioning traditional W-type.

@@
 Together with the identity type, they can express a
-whole host of inductive types~\cite{hugunin2021not}.
-%
-In our formulation, a W-type is given by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
-composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types".
+whole host of inductive types~\cite{hugunin2021not}. Specifically, with the traditional W-type and boolean types, we can encode an inductive type with multiple constructors.   
+%
+However, we formulate our W-type by an extra W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
+composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}) directly indicating each constructor and avoiding formentioned encoding.
+This is necessary to make our meta-theory syntax formulation closer to the surface syntax about inductive type and recursion handling supported by the plugin.\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types".
 Since there are already boolean types and Σ types, Fin types can be defined.

### 2023-03-15 metatheory-condensed.tex
tweak

--- 2023-03-15 metatheory-condensed.tex
@@
 
-Following \citet{kaposi2019gluing}, we use \emph{Universe à la Coquand} \cite{coquand2013presheaf}.
-Unlike Russell-style ones, Coquand universes are not inhabited by types directly but rather by the \emph{codes} of types
-and arguably behave better due to its closedness to Tarski-style ones~\cite{luo2012notes}. 
+We use universes à la \citet{coquand2013presheaf}, following \citet{kaposi2019gluing}.
+Unlike Russell-style ones, these universes are not inhabited by types directly but rather by the \emph{codes} of types
+and arguably behave better due to its closeness to Tarski-style universes~\cite{luo2012notes}. 
 % and compatibility with our algebraic formulation of MLTT~\cite{coquand2018canonicity}. % I am not sure if I want this sentence or not

### 2023-03-15 metatheory-condensed.tex
reply to a comment

@@
 \item 
-Linkages model families; they are like tuples composed of fields,
-with field names represented by variable bindings.
+Linkages model families, so they are like tuples composed of fields
+(with field names represented by variable bindings).
 But there is a twist: linkages support late binding.
@@
 quantified over---and thus polymorphic to---the components preceding it.\Zhixuan{explain the intuition behind the (typing rule of) linkages, and try to connect it to the ``family'' in the earlier section/surface syntax}\EDJreply{I am not sure if more is needed because the section ``Family polymorphism via linkages'' is trying to do this.
-Maybe I should estabilish connection between Rule l/add with the compilation process of fam poly?}
+Maybe I should estabilish connection between Rule l/add with the compilation process of fam poly?}\YZreply{I don't think you should mention anything about compilation process here.
+The surface syntax?
+Maybe.
+But I prefer not changing this sentence; it's supposed to be an executive summary rather than a detailed explanation.}
 

### 2023-03-15 metatheory-condensed.tex
typo?

@@
 We use universes à la \citet{coquand2013presheaf}, following \citet{kaposi2019gluing}.
-Unlike Russell-style ones, these universes are not inhabited by types directly but rather by the \emph{codes} of types
+Unlike Russell-style ones, these universes are not inhabited by types directly,
+but rather by the \emph{codes} of types,
 and arguably behave better due to its closeness to Tarski-style universes~\cite{luo2012notes}. 

### 2023-03-15 canonicitymodel.tex
Save Canonicity model.

@@
   \item the hierarchy is cumulative as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$
+  \item Thus if $\denotesC{A} \in \Set{i}, \denotesC{B} \in \Set{j}$, we will have $\denotesC{\TyPi{A}{B}} \in \Set{i} \cup \Set{j} = \Set{i \lcup j}$
 \end{itemize}
@@
   \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
-  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}}(\gm)(\gm')(t) &= \sum_{u' \in \denotesC{A}(\gm)(\gm')} \\
+  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}}(\gm)(\gm')(t) &= \sum_{u' \in \denotesC{A}(\gm)(\gm')(\fst{t})} \denotesC{B}(\SubstExt{\gm}{\fst{t}})((\gm', u'))(\snd{t}) \\
   \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 
@@
   \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm)(\gm') &= \star^2 \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{a}(\gm) \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{b}(\gm)(\gm') \text{ if } \denotes{c}(\gm)(\gm') = \star^2 \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{a}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{b}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^2 \\
   \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm)(\gm') &= \star\\
-  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
-  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
-  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
-  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}} &= \\
-  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
+  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}}(\gm)(\gm') &= \lambda u \lambda u' . \denotesC{t}(\SubstExt{\gm}{u}) ((\gm' , u')) \\
+  \denotesC{\goodTerm{\Gm}{\app{t}}{B}}(\gm)(\gm') &= \denotesC{t}(\pi_1 \gm)(\gm'[0])(\pi_2 \gm)(\gm'[1]) \\
+  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}}(\gm)(\gm') &= (\denotesC{a}(\gm)(\gm') , \denotesC{b}(\gm)(\gm') )\\
+  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[0] \\
+  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[1]\\
   \denotesC{\goodType{\Gm}{\cU_j}{j+1}}(\gm)(\gm')(T) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\}  \to \Set{j} \\
@@
 The counter-intuitive part of the proof-relevant logical relation is the treatment of proposition---a proposition is encoded as a subsingleton data.
-For example, $\denotesC{\goodType{\Gm}{\top}{}} = $ $ \{\star \ | \text{  when } t \equiv \unit \}$ represents a singleton set when $t \equiv \unit$ or an empty set otherwise.
-However, not all the denotation of types are subsingleton---the salient counter-examples are $\denotes{\goodType{\Gm}{\cU}{}}$ and $\denotes{\goodType{\Gm}{\TyPi{A}{B}}{}}$.
-The non-trivial interpretation on these types is the main reason we use proof-relevant logical relations.
+For example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \ | \text{ when } t \equiv \false  \}$ represents a disjoint union of two subsingleton set.
+The former one will be a singleton if we have a proof $t \equiv \true$ or an empty set otherwise, same as the later one.
+In other words, $\star^1$ witness a proof that $t \equiv true$.
+The same idea applies to the latter set.
 

### 2023-03-15 canonicitymodel.tex
Save Canonicity Model.

@@
   \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 
-  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_2 \gm_t)  \} \\
+  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \{ (\gamma' , t') : \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  \} \\
   \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm)(\gm') &= \star\\
@@
 The same idea applies to the latter set.
+Here $\star$, $\star^1$, $\star^2$ are just some arbitrary fixed elements.
 
@@
 Type-theoretical speaking, this ${\Pi}^c$ is using the \emph{internal dependent function type} of the above model.
-However this is not possible because the definition $\denotesC{\TyPi{A}{B}}$ is not solely based on the $\denotesC{A}$ and $\denotesC{B}$, but also on the syntax $\goodType{\Gm}{A}{}$ and $\goodType{\Gm}{B}{}$.
+However this is not possible because the definition $\denotesC{\TyPi{A}{B}}$ is not solely based on the $\denotesC{A}$ and $\denotesC{B}$, but also on the syntax $\goodType{\Gm}{A}{}$ and $\goodType{\Gm, A}{B}{}$.
 
@@
 
-What's more, with the above model, we can map an arbitrary closed boolean term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp)(\star) = \star^1$ or $\star^2$, witnessing $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
+What's more, with the above model, we can map an arbitrary closed boolean term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp)(\star) = \star^1$ or $\star^2$, witnessing the proof of $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
 

### 2023-03-15 coqimpl.tex
comments in limitations sec

--- 2023-03-04 coqimpl.tex
@@
 as the last line of \cref{fig:stlcfix-compiled} shows.
+
+\noindentparagraph{Trusted base.} Finally, we emphasize that our compilation can make sure the trusted code base is restricted to the Coq's kernel, as all the feature supported by our plugin has been compiled into Coq's surface syntax/Vernacular commands.
 
@@
 
-\subsection{Implementation Limitation}
+\fi
+
+\section{Limitations}
+\label{sec:limitations}
+
 We list some of the current limitation of our implementation. We believe some of these limitations can be remedied by more engineering efforts.
 \begin{itemize}
-  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, notation for inductive type and parameters in the inductive type (we only support indexed inductive type right now)
+  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, syntax extensions for inductive type and parameters in the inductive type (we only support indexed inductive type right now)\YZ{Can parameterized inductive types always be encoded as indexed inductive types?}
   \item We don't support mutual fixpoint definition
@@
   \item We only support automatic generation of partial recursor and the computation information for non-indexed inductive type 
+  \item ...\YZ{We should mention the limitations mentioned in Review D and in our response.
+All these limitations can be split into two categories: (1) those that we believe can be fixed with more engineering effort, and (2) those that are more research-y.}
 \end{itemize}
-
-\fi

### 2023-03-15 coqimpl.tex
Reply comments.

@@
 
-We list some of the current limitation of our implementation. We believe some of these limitations can be remedied by more engineering efforts.
+We list some of the current limitation of our implementation. We believe the following limitation can be rememdied by more engineering effort:
 \begin{itemize}
-  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, syntax extensions for inductive type and parameters in the inductive type (we only support indexed inductive type right now)\YZ{Can parameterized inductive types always be encoded as indexed inductive types?}
+  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, syntax extensions for inductive type and parameters in the inductive type (we only support indexed inductive type right now)\YZ{Can parameterized inductive types always be encoded as indexed inductive types?}\EDJreply{Yes.
+but practically the dependent elimination principle is different when using encoding.}
   \item We don't support mutual fixpoint definition
@@
   \item Unicode (i.e. "Require Import Coq.Unicode.Utf8.") is not compatible with our plugin 
-  \item We don't consider the compatibility of universe polymorphism mechanism with our plugin
   \item We only support automatic generation of partial recursor and the computation information for non-indexed inductive type 
@@
 \end{itemize}
+
+However, there are also limitations we believe require more research effort, including but not limited to: 
+\begin{itemize}
+  \item We didn't considered the compatibility of universe polymorphism mechanism of Coq with our plugin except we know our plugin elaborate families into modules.
+Thus our implementation is supposed to be as compatible with universe polymorphism as Coq's Module.
+  \item We don't support normalization of the open term using extensible inductive type.
+When applying recursor $Rec$ to open terms $t$ using extensible inductive type, the computation cannot be done by the Coq's kernel, thus the computation result of $Rec(t)$ is only propositionally equal (but not judgementally equal) to $Rec(t)$.
+It will complicate typechecking of the dependently typed terms.
+\end{itemize}

### 2023-03-15 coqimpl.tex
edit on "trusted base" para

--- 2023-03-15 coqimpl.tex
@@
 We implement our language design as a Coq plugin.
-Rather than modifying the Coq kernel, a prototypical implementation as a
-plugin allows a clearly defined trusted base and
-improved compatibility with different versions of Coq.
-
-The implementation works by translating programs in \Lang syntax into
+%
+It works by translating programs in \Lang syntax into
 programs that can be checked and evaluated by Coq.
@@
 
-\noindentparagraph{Trusted base.} Finally, we emphasize that our compilation can make sure the trusted code base is restricted to the Coq's kernel, as all the feature supported by our plugin has been compiled into Coq's surface syntax/Vernacular commands.
+\noindentparagraph{Trusted base.}
+
+Rather than modifying the Coq kernel to extend its core theory, a translation to
+Coq conveniently reduces the trusted base of any development using \Lang to Coq.
+%and improves compatibility with different versions of Coq.
+In particular, once a family is closed, \lsti{Print Assumptions} can be used
+to verify that there are no lingering \lsti{Axiom}s introduced by the translation.
+Ramifications of possible bugs in the \Lang implementation are limited to
+the usability of the plugin.
+%Finally, we emphasize that our compilation can make sure the trusted code base is restricted to the Coq's kernel, as all the feature supported by our plugin has been compiled into Coq's surface syntax/Vernacular commands.
 

### 2023-03-15 metatheory-condensed.tex
Fix B in Wsig

@@
 $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and
-$\goodType{\Gm,x : \wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$,
+$\goodTerm{\Gm}{\wsigproj{i}{2}{\wsig}}{\wsigproj{i}{1}{\wsig} \to \cU}$,
 defines the $i$-th constructor.
 That is, given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
-$\goodTerm{\Gm, x : \nsub{\wsigproj{i}{2}{\wsig}}{x}{t_1}}{t_2}{\El{\wcode{\wsig}}}$,
+$\goodTerm{\Gm, x : \El{\wsigproj{i}{2}{\wsig}(t_1)} }{t_2}{\El{\wcode{\wsig}}}$,
 we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type (\ruleref{tm/wsup}).
@@
 Each item in this W-type signature, having two types, corresponds to one constructor, where the first type stores the non-recursive part of the constructor and the second type stores the ``arity'' of the  recursive part.
-For example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times (\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule \ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone, \wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \bbtwo$).
-This type means there is no information on non-recursive part (thus use unit type $\bbone$) and the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1 = \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\bbtwo}$.
+For example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times (\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule \ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone, \wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \lambda \_ .
+\bbtwo$).
+This type means there is no information on non-recursive part (thus use unit type $\bbone$) and the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1 = \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\lambda \_ .\bbtwo}$.
 The story is similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and thus the ``arity'' is zero.

### 2023-03-16 canonicitymodel.tex
Save more on canonicity model.

@@
 
-Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$, that will return the syntax piece $S$ as well.
-Thus we can have a function $\Glued{\Pi}$ s.t.
+Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$, that will return the syntax piece $S$ as well.\footnote{This is also called glued interpretation in \citet{sterling2019algebraic}} Thus we can have a function $\Glued{\Pi}$ s.t.
 $\GluedPi{\denotesCC{A}}{\denotesCC{B}} = \denotesCC{\TyPi{A}{B}}$ now that syntax is also provided.
@@
 What's more, during the construction of the model for \TT, we will use $\Glued{(\cdot)}$ when we can.
+
+\newcommand{\GTy}[2]{\ensuremath{\Glued{\texttt{Ty}}_{#1}~{#2}}}
+\newcommand{\GCon}[1]{\ensuremath{\Glued{\texttt{Con}}_{#1}}}
+\newcommand{\GMWSig}[3]{\ensuremath{{\texttt{WSig}^C}_{#1}^{#3}~{#2}}}
+\newcommand{\GWSig}[3]{\ensuremath{{\Glued{\texttt{WSig}}}_{#1}^{#3}~{#2}}}
+
+There are actually more \emph{internal} type-theoretic construction to explore, and we need them to define the model for \TT as well.
+
+We denote $\GCon{k} \coloneq \sum_{\goodCtx{\Gm}{k}} \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k}$
+
+and $\GTy{j}{\Glued{\Gm}} \coloneq \sum_{\goodType{\Gm}{T}{j}} \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \Gmodel{\Glued{\Gm}}(\gm)}\{t : \goodTerm{\cdot}{t}{\sub{T}{\gm}}\}\to\Set{j} $ for $\Glued{\Gm} \in \GCon{k}$.
+
+Basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$.
+
+Then we define $\GMWSig{j}{\Glued{\Gm}}{n} \coloneq "Vector"^n~\sum_{\Glued{A} \in \GTy{j}{\Glued{\Gm}}} \GTy{j}{(\GSubstExt{\Glued{\Gm}}{\Glued{A}})}$ for $\Glued{\Gm} \in \GCon{k}$, which is basically a list of a pair of types of length $n$.
+As before, we can define glued interpretation $\GWSig{j}{\Glued{\Gm}}{n} \coloneq \{ \wsig : \goodWSig{\cdot}{\wsig}{n} \} \times \GMWSig{j}{\Glued{\Gm}}{n}$.
+This will be useful when we interpret judgement for W-type signature.
 
@@
   % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
-  \denotesC{\goodWSig{\Gm}{\wsig}{n}} &\text{ is a list of 2-tuple of length $n$ } \\
+  \denotesC{\goodWSigU{j}{\Gm}{\wsig}{n}} &: \GMWSig{j}{\denotesCC{\Gm}}{n} \\ 
+    &\text{ i.e., a list of 2-tuple of length $n$ } \\
   % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
@@
 \end{align*}
+together with the following inductively defined indexed set and its eliminator principle. These can be directly translated into Agda but we use a more conventional math notation
+\begin{align*}
+  "Inductive"~CW_p : (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
+    "CWsup"_p : n < N \to 
+\end{align*}
 

### 2023-03-16 metatheory-condensed.tex
Reply comments. Fix.

@@
 Maybe.
-But I prefer not changing this sentence; it's supposed to be an executive summary rather than a detailed explanation.}
+But I prefer not changing this sentence; it's supposed to be an executive summary rather than a detailed explanation.}\EDJreply{Actually I current am not sure where to add more sentences and how to elaborate more.
+Except for mentioning compilation process I don't think I have any good idea.
+So I will just stop here.}
 

### 2023-03-16 canonicitymodel.tex
Save Canonicity Model.

@@
 \newcommand{\GCon}[1]{\ensuremath{\Glued{\texttt{Con}}_{#1}}}
+\newcommand{\GTm}[2]{\ensuremath{\Glued{\texttt{Tm}}~{#1}~{#2}}}
 \newcommand{\GMWSig}[3]{\ensuremath{{\texttt{WSig}^C}_{#1}^{#3}~{#2}}}
@@
 
-There are actually more \emph{internal} type-theoretic construction to explore, and we need them to define the model for \TT as well.
+There are actually more \emph{internal} type-theoretic construction to we need:
 
@@
 
-and $\GTy{j}{\Glued{\Gm}} \coloneq \sum_{\goodType{\Gm}{T}{j}} \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \Gmodel{\Glued{\Gm}}(\gm)}\{t : \goodTerm{\cdot}{t}{\sub{T}{\gm}}\}\to\Set{j} $ for $\Glued{\Gm} \in \GCon{k}$.
+and $\GTy{j}{\Glued{\Gm}} \coloneq \sum_{\goodType{\Gm}{T}{j}} \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \Gmodel{\Glued{\Gm}}(\gm)}\{t : \goodTerm{\cdot}{t}{\sub{T}{\gm}}\}\to\Set{j} $ for $\Glued{\Gm} \in \GCon{k}$
 
-Basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$.
+and $\GTm{\Glued{\Gm}}{\Glued{T}} \coloneq \sum_{\goodTerm{\Gm}{t}{T}} \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in\Gmodel{\Glued{\Gm}}} \Gmodel{\Glued{T}}(\gm)(\gm')(\sub{t}{\gm})$
+
+Basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$.
+Bascially the four sets are collecting the glued interpretation.
 
@@
 The intuition of the model is similar to that of the syntactic translation---use the canonicity model of sigma type (from the earlier work) for linkage type and using inductive facility in the meta-logic to justify our W-type.
+
+\newcommand{\CWmodel}{\ensuremath{\mathit{W}^C}}
+\newcommand{\CWsup}{\ensuremath{\mathit{Wsup}^C}}
+\newcommand{\CWrec}{\ensuremath{\mathit{Wrec}^C}}
 
@@
 \end{align*}
-together with the following inductively defined indexed set and its eliminator principle. These can be directly translated into Agda but we use a more conventional math notation
+together and mutual recursively defined with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$) ....
+
 \begin{align*}
-  "Inductive"~CW_p : (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
-    "CWsup"_p : n < N \to 
+  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
+    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
+    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \sub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) }{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
 \end{align*}
+... and its eliminator principle $\CWrec$. 
+\begin{align*}
+  \CWrec &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
+    (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
+    & \to [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\ 
+    & \quad \quad \to 
+    b \in (\GTm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{\SubstId}{{a}}}) }{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}) \\ 
+    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
+    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \\
+    & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
+  \CWrec & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda )
+\end{align*}
+
+Note that in $\CWmodel$, the $\Glued{b}$ uses the definition of $\denotesC{\wcode{\wsig}}$, which after unfolding, will refers to $\CWmodel$ again and in strictly positive position.
+So it is a well-founded definition.
 

### 2023-03-16 canonicitymodel.tex
Save Canonicity Model.

@@
 However, unlike most logical relation proof, we are having a proof-relevant logical realtion, essential for modeling universes properly~\cite{coquand2018canonicity}.
-Our canonicity model at MLTT part is basically the same as shown in \citet{coquand2018canonicity}.
+Our canonicity model at MLTT part is very close to the one shown in \citet{coquand2018canonicity} and \citet{sterling2019algebraic}.
 % We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
@@
 \newcommand{\Gmodel}[1]{\ensuremath{{({#1})}^c}}
+\newcommand{\GEl}[1]{\ensuremath{\Glued{\texttt{El}}({#1})}}
+\newcommand{\Gwcode}[1]{\ensuremath{\Glued{\texttt{W}}({#1})}}
+\newcommand{\GSubstId}{\ensuremath{\GSubstWeak{0}}}
 
@@
 
-Basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$.
-Bascially the four sets are collecting the glued interpretation.
+These four sets are collecting the glued interpretation:
+basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$.
 
@@
 \newcommand{\CWmodel}{\ensuremath{\mathit{W}^C}}
-\newcommand{\CWsup}{\ensuremath{\mathit{Wsup}^C}}
-\newcommand{\CWrec}{\ensuremath{\mathit{Wrec}^C}}
+\newcommand{\CWsup}{\ensuremath{\mathit{W^Csup}}}
+\newcommand{\CWrec}{\ensuremath{\mathit{W^Crec}}}
 
@@
   RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\
-  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{R}} \\
+  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{"R"}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{"R"}} \\
   \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg \\
-  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\
+  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') &= \CWrec~(\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\wsig}{\sub{\lkg}{\gm'}}{w}))~f^r~\sub{t}{\gm'}~(\denotesC{t}(\gm)(\gm')) \\
+  \text{ where } \quad f^r~j~\Glued{a}~b~b^C &= 
+  \texttt{let}~\Glued{\rho}  \in \GTm{(\Glued{\cdot} \Glued{,} \Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}})}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})} 
+  \\
+  & \ \texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^C)~\texttt{ in } \\
+  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\Recproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
+  & \quad 
+   \text{ and also } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \quad ,\Glued{R} = \denotesCC{"R"}, \Glued{B} = \wsigproj{j}{2}{\Glued{\wsig}}
 \end{align*}
@@
     \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
-    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \sub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) }{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
+    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\wsig}}}{\GSubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
 \end{align*}
@@
 \begin{align*}
-  \CWrec &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
+  \CWrec \  &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
     (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
@@
     & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
-  \CWrec & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda )
+  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~P~f~(\sub{b}{\Gmodel{\Glued{\gm}}})~(b^C~\Glued{\gm}))
 \end{align*}
@@
 So it is a well-founded definition.
+We also don't distinguish $(b, b^C)$ and $\Glued{b}$ for simplicity.
 

### 2023-03-17 metatheory-condensed.tex
edit on what's new about our W-types

@@
 quantified over the earlier ones, a linkage component is \emph{universally}
-quantified over---and thus polymorphic to---the components preceding it.\Zhixuan{explain the intuition behind the (typing rule of) linkages, and try to connect it to the ``family'' in the earlier section/surface syntax}\EDJreply{I am not sure if more is needed because the section ``Family polymorphism via linkages'' is trying to do this.
+quantified over---and thus polymorphic to---the components preceding it.%\Zhixuan{explain the intuition behind the (typing rule of) linkages, and try to connect it to the ``family'' in the earlier section/surface syntax}\EDJreply{I am not sure if more is needed because the section ``Family polymorphism via linkages'' is trying to do this.
 Maybe I should estabilish connection between Rule l/add with the compilation process of fam poly?}\YZreply{I don't think you should mention anything about compilation process here.
@@
 inductive types in MLTT.
-Together with the identity type, they can express a
-whole host of inductive types~\cite{hugunin2021not}. Specifically, with the traditional W-type and boolean types, we can encode an inductive type with multiple constructors.   
-%
-However, we formulate our W-type by an extra W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
-composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}) directly indicating each constructor and avoiding formentioned encoding.
-This is necessary to make our meta-theory syntax formulation closer to the surface syntax about inductive type and recursion handling supported by the plugin.\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types".
+Together with the identity type, they can express a whole host of inductive
+types~\cite{hugunin2021not}, including those with multiple constructors.
+
+Our formulation of W-types differs from previous ones in that it makes
+individual constructors explicit and thus more closely matches the surface
+syntax of \Lang.
+%
+In \TT, a W-type is given by a \emph{W-type signature} $\goodWSig{\Gm}{\wsig}{n}$ 
+composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
+%\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types".
 Since there are already boolean types and Σ types, Fin types can be defined.
@@
 and even more complex  \\
-2.
+%2.
 so to me the problem you guys really have is around the content in the appendix -- I think Zhixuan mean he doesn't think this is more expressive.
@@
 \\
-3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
+%3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
 %

### 2023-03-17 canonicitymodel.tex
Save Canonicity Model.
Save Canonicity model.

@@
 \newcommand{\GSubstId}{\ensuremath{\GSubstWeak{0}}}
+\newcommand{\Gwsigproj}[3]{\ensuremath{\Glued{{\texttt{w}\pi^{#1}_{\texttt{#2}}}}({#3})}}
+\newcommand{\GRecproj}[2]{\ensuremath{\Glued{{\texttt{R}\pi^{#1}}}({#2})}}
 
@@
 \newcommand{\GTm}[2]{\ensuremath{\Glued{\texttt{Tm}}~{#1}~{#2}}}
+\newcommand{\GSub}[2]{\ensuremath{\Glued{\texttt{Sub}}~{#1}~{#2}}}
 \newcommand{\GMWSig}[3]{\ensuremath{{\texttt{WSig}^C}_{#1}^{#3}~{#2}}}
@@
 
-There are actually more \emph{internal} type-theoretic construction to we need:
+There are actually more \emph{internal} type-theoretic construction that we need:
 
@@
 
+and $\GSub{\Glued{\Gm}}{\Glued{\Dl}} \coloneq \sum_{\goodSub{\Gm}{\delta}{\Dl}} \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \Gmodel{\Glued{\Gm}}(\gm)} \Gmodel{\Glued{\Dl}}(\delta \circ \gm) $
+
 These four sets are collecting the glued interpretation:
-basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$.
+basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$, and etc.
+
+% Note that, for arbitrary $\Glued{\Gm} \in \GCon{k}$, we can consider it as $\Glued{\gm}$ as an element of $\GSub{\Glued{\cdot}}{}$
 
@@
   & \ \texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^C)~\texttt{ in } \\
-  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\Recproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
+  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
   & \quad 
-   \text{ and also } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \quad ,\Glued{R} = \denotesCC{"R"}, \Glued{B} = \wsigproj{j}{2}{\Glued{\wsig}}
+   \text{ and also } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \quad ,\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\Glued{\wsig}}
 \end{align*}
@@
     \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
-    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\wsig}}}{\GSubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
+    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
 \end{align*}
@@
     (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
-    & \to [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\ 
+    & \to [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
     & \quad \quad \to 
-    b \in (\GTm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{\SubstId}{{a}}}) }{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}) \\ 
-    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
+    \goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{\SubstId}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}} \\ 
+    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
     & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \\

### 2023-03-17 metatheory-condensed.tex
questions

@@
 the derivation of \goodTerm{\Gm_1}{t}{T_1} is considered definitionally equal to
-a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate?}
+a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate??}
 %\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.
@@
 $T$ is the motive of the recursion (\ruleref{tyeq/casety});
-for simplicity, we model non-dependent motives for recursion.
+for simplicity, we model non-dependent motives for recursion.\YZ{Does "for recursion" here convey relevant information?}
 The collection of case handlers~$\lkg$ encodes those defined and inherited

### 2023-03-17 abstract.tex
tweaks

@@
 The new design achieves reuse and extensibility via a form of family
-polymorphism, a seemingly object-oriented idea, that allows code and
+polymorphism, an object-oriented idea, that allows code and
 proofs to be polymorphic to their enclosing families.

### 2023-03-17 metatheory-condensed.tex
tweaks

@@
 
-Our formulation of W-types differs from previous ones in that it makes
-individual constructors explicit and thus more closely matches the surface
-syntax of \Lang.
-%
-In \TT, a W-type is given by a \emph{W-type signature} $\goodWSig{\Gm}{\wsig}{n}$ 
-composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
+Our formulation of W-types differs from previous ones in that it is
+straightforward to identify constructors from what we call \emph{W-type signatures}.
+%
+A signature $\goodWSig{\Gm}{\wsig}{n}$ is
+composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}),
 %\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types".
@@
 %
-Each pair models a constructor of the inductive type.
+each modeling a constructor of the inductive type.
 The $i$-th pair, projected from the signature~$\wsig$ using the forms

### 2023-03-17 metatheory-condensed.tex
Reply the comment.

--- 2023-03-17 metatheory-condensed.tex
@@
 the derivation of \goodTerm{\Gm_1}{t}{T_1} is considered definitionally equal to
-a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate??}
+a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate??}\EDJreply{Yes.
+In QIIT model, it looks like `t : Tm G1 T1' and  `t : Tm G2 T2'.
+In extenional Agda, because ``G1 = G2'', ``T1 = T2'', thus the two types `Tm G2 T2 = Tm G1 T1'.
+So they are the same derivation in QIIT.
+In our setup, we have the same idea, because after quotient, G1 refers to same object as G2, so the derivation is the same.}
 %\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.

### 2023-03-17 metatheory-condensed.tex
edit

--- 2023-03-17 metatheory-condensed.tex
@@
 %
-For readability, we use a notation with explicit names
-in this typing judgment:
-we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
-and write its type~$A_i$, which is a dependent tuple type, as a
-dependent record type with labels.
+The context types $A_i$ are a dependent tuple type, but for readability,
+we write them as dependent record types that give labels to the fields.
+%we use a notation with explicit names
+%in this typing judgment:
+%we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
+%and write its type~$A_i$, which is a dependent tuple type, as a
+%dependent record type with labels.
 %

### 2023-03-17 canonicitymodel.tex

@@
 
+Note that, we have $\prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \Gmodel{\Glued{\Gm}}(\gm)}$ prefix in $\GTy{}{}$, $\GTm{}{}$,$\GSub{}{}$.
+The interesting fact is, given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$ and $\gm' \in \Gmodel{\Glued{\Gm}}(\gm)$, we can consider $(\gm, \gm')$ as an element of $\GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
+Thus we consider $(\gm, \gm') ~ \Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
 % Note that, for arbitrary $\Glued{\Gm} \in \GCon{k}$, we can consider it as $\Glued{\gm}$ as an element of $\GSub{\Glued{\cdot}}{}$
@@
   \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
-  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}} &= \\
+  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}}(\gm)(\gm') &= \\
   \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
@@
     & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
-    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \\
+    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
     & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\

### 2023-03-17 metatheory-condensed.tex
resolve comment

@@
 the derivation of \goodTerm{\Gm_1}{t}{T_1} is considered definitionally equal to
-a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate??}\EDJreply{Yes.
+a derivation of \goodTerm{\Gm_2}{t}{T_2}.
+%\YZ{Is this accurate??}\EDJreply{Yes.
 In QIIT model, it looks like `t : Tm G1 T1' and  `t : Tm G2 T2'.

### 2023-03-17 metatheory-condensed.tex
Canonicity Model finished. OMG JS/Coquand really did their best to make things accessible.

@@
 $T$ is the motive of the recursion (\ruleref{tyeq/casety});
-for simplicity, we model non-dependent motives for recursion.\YZ{Does "for recursion" here convey relevant information?}
+for simplicity, we model non-dependent motives for recursion.\YZ{Does "for recursion" here convey relevant information?}\EDJreply{no.
+I actually wanted to say "non-dependent motives for non-dependent recursion".
+Maybe change to "only model non-dependent motives"?}
 The collection of case handlers~$\lkg$ encodes those defined and inherited

### 2023-03-17 metatheory-condensed.tex
Reply the comment.

@@
 I actually wanted to say "non-dependent motives for non-dependent recursion".
-Maybe change to "only model non-dependent motives"?}
+Maybe change to "only model non-dependent motives"??}
 The collection of case handlers~$\lkg$ encodes those defined and inherited

### 2023-03-17 metatheory-condensed.tex
comment

--- 2023-03-17 metatheory-condensed.tex
@@
 $\goodTerm{\Gm, x : \El{\wsigproj{i}{2}{\wsig}(t_1)} }{t_2}{\El{\wcode{\wsig}}}$,
-we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type (\ruleref{tm/wsup}).
+one can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type \El{\wcode{\wsig}} (\ruleref{tm/wsup}).
 $\wcode{\wsig}$ gives the code of the
@@
 This type means there is no information on non-recursive part (thus use unit type $\bbone$) and the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1 = \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\lambda \_ .\bbtwo}$.
-The story is similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and thus the ``arity'' is zero.
+The story is similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and thus the ``arity'' is zero.\YZ{Feels like it's more straightforward to just list the encodings of the four constructors of tm in a table, assuming an encoding of the identifier type.
+And this paragraph should be moved to the earlier paragraph on W-types.
+I'll take a stab at this.}
 

### 2023-03-17 metatheory-condensed.tex

@@
 $T$ is the motive of the recursion (\ruleref{tyeq/casety});
-for simplicity, we model non-dependent motives for recursion.\YZ{Does "for recursion" here convey relevant information?}
+for simplicity, we model non-dependent motives for recursion.\YZ{Does "for recursion" here convey relevant information?}\EDJreply{no.
+I actually wanted to say "non-dependent motives for non-dependent recursion".
+Maybe change to "only model non-dependent motives"??}
 The collection of case handlers~$\lkg$ encodes those defined and inherited

### 2023-03-17 metatheory-condensed.tex
resolve comment

--- 2023-03-17 metatheory-condensed.tex
@@
 $T$ is the motive of the recursion (\ruleref{tyeq/casety});
-for simplicity, we model non-dependent motives for recursion.\YZ{Does "for recursion" here convey relevant information?}
+for simplicity, we model only non-dependent motives.
 The collection of case handlers~$\lkg$ encodes those defined and inherited

### 2023-03-17 coqimpl.tex
comments in limitations section

@@
 
-We list some of the current limitation of our implementation. We believe the following limitation can be rememdied by more engineering effort:
-\begin{itemize}
-  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, syntax extensions for inductive type and parameters in the inductive type (we only support indexed inductive type right now)\YZ{Can parameterized inductive types always be encoded as indexed inductive types?}\EDJreply{Yes.
+The \Lang implementation currently does not support the full set of the Coq
+facility for inductive types and fixpoints;
+in particular, mutually inductive types and mutual fixpoints are not supported.
+The plugin supports indexed inductive types but not parameterized ones, though
+the latter can be encoded via the former.
+Partial recursors are automatically generated only for non-indexed inductive types.
+We believe that these limitations do not pose conceptual challenges and can be
+addressed with more engineering effort of the same kind that went into \Lang.
+
+%We list some of the current limitation of our implementation. We believe the following limitation can be rememdied by more engineering effort:
+%\begin{itemize}
+%  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, syntax extensions for inductive type and parameters in the inductive type (we only support indexed inductive type right now)\YZ{Can parameterized inductive types always be encoded as indexed inductive types?}\EDJreply{Yes.
 but practically the dependent elimination principle is different when using encoding.}
-  \item We don't support mutual fixpoint definition
-  \item We don't support incorporating "Hint" Vernacular commands, and thus the users cannot specify the database hints
-  \item Unicode (i.e. "Require Import Coq.Unicode.Utf8.") is not compatible with our plugin 
-  \item We only support automatic generation of partial recursor and the computation information for non-indexed inductive type 
-  \item ...\YZ{We should mention the limitations mentioned in Review D and in our response.
-All these limitations can be split into two categories: (1) those that we believe can be fixed with more engineering effort, and (2) those that are more research-y.}
-\end{itemize}
+%  \item We don't support mutual fixpoint definition
+%  \item We don't support incorporating "Hint" Vernacular commands, and thus the users cannot specify the database hints
+%  \item Unicode (i.e. "Require Import Coq.Unicode.Utf8.") is not compatible with our plugin 
+%  \item We only support automatic generation of partial recursor and the computation information for non-indexed inductive type 
+%\end{itemize}
 
@@
   \item We didn't considered the compatibility of universe polymorphism mechanism of Coq with our plugin except we know our plugin elaborate families into modules.
-Thus our implementation is supposed to be as compatible with universe polymorphism as Coq's Module.
+Thus our implementation is supposed to be as compatible with universe polymorphism as Coq's Module.\YZ{Is this really a limitation at all?
+Do we have to mention it?}
+  \item ... \YZ{mention restricting induction to the top level}
   \item We don't support normalization of the open term using extensible inductive type.

### 2023-03-17 coqimpl.tex
tweak

@@
 
-The \Lang implementation currently does not support the full set of the Coq
+The \Lang implementation currently does not support Coq's full
 facility for inductive types and fixpoints;
@@
 
-However, there are also limitations we believe require more research effort, including but not limited to: 
+However, there are also limitations we believe require more research effort or more thought from a language-design perspective, including but not limited to: 
 \begin{itemize}

### 2023-03-17 coqimpl.tex
Add top level recursion limitation.

@@
 Do we have to mention it?}
-  \item ... \YZ{mention restricting induction to the top level}
+  \item The current recursion mechanism supported by "FRecursion" and "FInduction" only works at top level.
+Further research efforts are necessary to enable the appearance of recursion handlers at any location while facilitating extensibility of the handlers at said locations.
+Additionally, a suitable Coq tactic is required to accompany this functionality.
+\YZ{mention restricting induction to the top level}
   \item We don't support normalization of the open term using extensible inductive type.

### 2023-03-17 canonicitymodel.tex
Canonicity Model finished. OMG JS/Coquand really did their best to make things accessible.
Add table for four signatures The table is inlined in the text, I believe you will reformat at the end, it is just help you to copy and paste the content.
Adding substitution info in Canonicity

@@
   \denotesC{\goodTerm{\Gm}{t}{T}} \text{ is a dependent function}&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})\\
+  \denotesC{\goodType{\Gm}{\sub{T}{\sigma}}{}}(\gm)(\gm')(t) &= \denotesC{T}(\sigma \circ \gm)(\denotes{\sigma}(\gm)(\gm'))(t)\\
   \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \} \\
@@
   \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \{ (\gamma' , t') : \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  \} \\
+  \denotesC{\goodTerm{\Gm}{\sub{t}{\sigma}}{\sub{T}{\sigma}}}(\gm)(\gm') &= \denotesC{t}(\sigma \circ \gm)(\denotesC{\sigma}(\gm)(\gm')) \\
   \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm)(\gm') &= \star\\
@@
 The interesting fact is, given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$ and $\gm' \in \Gmodel{\Glued{\Gm}}(\gm)$, we can consider $(\gm, \gm')$ as an element of $\GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
-Thus we consider $(\gm, \gm') ~ \Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
+Thus we consider the pair $(\gm, \gm')$ as $\Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
 % Note that, for arbitrary $\Glued{\Gm} \in \GCon{k}$, we can consider it as $\Glued{\gm}$ as an element of $\GSub{\Glued{\cdot}}{}$
@@
   \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesCC{A}, \denotesCC{s}, \denotesCC{T})::\denotesC{\lsig}\\ 
+  \denotesC{\goodSig{\Gm}{\sub{\lsig}{\gm}}{n}} & \text{ is done by point-wise/component-wise substitution} \\
   \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
@@
   \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesCC{A} , \denotesCC{B}) :: \denotesC{\wsig}\\
+  \denotesC{\goodWSig{\Gm}{\sub{\wsig}{\gm}}{n}} & \text{ is done by point-wise/component-wise substitution} \\
   \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= \Gmodel{(\text{ $j$-th element of }~\denotesC{\wsig})[0]}\\
@@
   \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
-  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}}(\gm)(\gm') &= \\
-  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
+  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}}(\gm)(\gm')(t) &= \CWmodel~(\Gsub{\denotesCC{\wsig}}{\Glued{\gm}})~t \\
+  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~(\Gsub{\denotesCC{\wsig}}{\Glued{\gm}})~(\Gsub{\denotesCC{a}}{\Glued{\gm}})~(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}}) \\
   % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\

### 2023-03-17 coqimpl.tex
limitations

--- 2023-03-17 coqimpl.tex
@@
 
-The \Lang implementation currently does not support the full set of the Coq
-facility for inductive types and fixpoints;
-in particular, mutually inductive types and mutual fixpoints are not supported.
-The plugin supports indexed inductive types but not parameterized ones, though
-the latter can be encoded via the former.
+The \Lang implementation currently does not yet bring extensibility to Coq's full
+facility for inductive types.
+Mutually inductive types and parameterized inductive types are not yet extensible,
+though indexed inductive types are supported and can be used to encode parameterized ones.
 Partial recursors are automatically generated only for non-indexed inductive types.
-We believe that these limitations do not pose conceptual challenges and can be
+The case studies in \cref{sec:casestudies} do not exercise these features.
+We believe that they do not pose conceptual challenges and can be
 addressed with more engineering effort of the same kind that went into \Lang.
@@
 
-However, there are also limitations we believe require more research effort, including but not limited to: 
-\begin{itemize}
-  \item We didn't considered the compatibility of universe polymorphism mechanism of Coq with our plugin except we know our plugin elaborate families into modules.
+What seems to require more thought from a language-design perspective is the
+restriction that recursion and induction (on extensible types) cannot be nested.
+%and must be at the top level through the \lsti{FRecursion} and \lsti{FInduction}
+%commands.
+A possible solution is to make the plugin generate proof obligations for
+nested pattern matches when the inductive types acquire new constructors.
+%
+Also limiting (and interesting for future research) is
+that generated propositional equalities cannot be used by Coq to automatically
+reduce terms to definitionally equal forms, which causes inconveniences for
+developments using intrinsically typed syntax.
+
+%\begin{itemize}
+%  \item We didn't considered the compatibility of universe polymorphism mechanism of Coq with our plugin except we know our plugin elaborate families into modules.
 Thus our implementation is supposed to be as compatible with universe polymorphism as Coq's Module.\YZ{Is this really a limitation at all?
 Do we have to mention it?}
-  \item ... \YZ{mention restricting induction to the top level}
-  \item We don't support normalization of the open term using extensible inductive type.
+%  \item The current recursion mechanism supported by "FRecursion" and "FInduction" only works at top level.
+Further research efforts are necessary to enable the appearance of recursion handlers at any location while facilitating extensibility of the handlers at said locations.
+Additionally, a suitable Coq tactic is required to accompany this functionality.
+%  \item We don't support normalization of the open term using extensible inductive type.
 When applying recursor $Rec$ to open terms $t$ using extensible inductive type, the computation cannot be done by the Coq's kernel, thus the computation result of $Rec(t)$ is only propositionally equal (but not judgementally equal) to $Rec(t)$.
 It will complicate typechecking of the dependently typed terms.
-\end{itemize}
+%\end{itemize}

### 2023-03-18 metatheory-condensed.tex
Add table for four signatures The table is inlined in the text, I believe you will reformat at the end, it is just help you to copy and paste the content.
move

@@
 
+\medskip
+
+\begin{tabular}{|
+>{\columncolor[HTML]{FFFFFF}}l |l|l|l|}
+\hline
+$\wsig_{\texttt{tm}}^{\texttt{unit}} \coloneq \WSigAdd{\WSigEmp}{\bbone}{\lambda \_ . \bbzero}$ 
+& $\wsig_{\texttt{tm}}^{\texttt{var}} \coloneq \WSigAdd{\wsig_{\texttt{tm}}^{\texttt{unit}}}{"id"}{\lambda \_ . \bbzero}$
+& $\wsig_{\texttt{tm}}^{\texttt{abs}} \coloneq \WSigAdd{\wsig_{\texttt{tm}}^{\texttt{var}}}{"id"}{\lambda \_ .\bbone} $ 
+& $\wsig_{\texttt{tm}}^{\texttt{app}} \coloneq \WSigAdd{\wsig_{\texttt{tm}}^{\texttt{abs}}}{\bbone}{\lambda \_ .\bbtwo}$  \\ \hline
+\end{tabular}
+
+We illustrate the structure of W-type signature (as a list) here.
+$\wsig_{\texttt{tm}}^0, \wsig_{\texttt{tm}}^1$ are (the subparts of) the
+signatures of constructors \{"tm\_unit"\} and \{"tm\_unit","tm\_app"\}
+respectively. Each item in this W-type signature, having two types, corresponds
+to one constructor, where the first type stores the non-recursive part of the
+constructor and the second type stores the ``arity'' of the  recursive part. For
+example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times
+(\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule
+\ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone,
+\wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \lambda \_ . \bbtwo$). This type means
+there is no information on non-recursive part (thus use unit type $\bbone$) and
+the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1
+= \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\lambda \_ .\bbtwo}$.  The story is
+similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and
+thus the ``arity'' is zero.\YZ{Feels like it's more straightforward to just list
+the encodings of the four constructors of tm in a table, assuming an encoding of
+the identifier type. And this paragraph should be moved to the earlier paragraph
+on W-types. I'll take a stab at this.}
+
+Generally speaking, the expressiveness of W-types comes from the ``arity'', as
+it encodes the case when recursive structure stores as the output of a function,
+mirroring the strict positivity of Inductive type.
+
+\medskip
+
 W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a
@@
 
-\item We illustrate the structure of W-type signature (as a list) here.
-$\wsig_{\texttt{tm}}^0, \wsig_{\texttt{tm}}^1$ are (the subparts of) the signatures of constructors \{"tm\_unit"\} and \{"tm\_unit","tm\_app"\} respectively.
-Each item in this W-type signature, having two types, corresponds to one constructor, where the first type stores the non-recursive part of the constructor and the second type stores the ``arity'' of the  recursive part.
-For example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times (\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule \ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone, \wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \lambda \_ .
-\bbtwo$).
-This type means there is no information on non-recursive part (thus use unit type $\bbone$) and the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1 = \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\lambda \_ .\bbtwo}$.
-The story is similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and thus the ``arity'' is zero.\YZ{Feels like it's more straightforward to just list the encodings of the four constructors of tm in a table, assuming an encoding of the identifier type.
-And this paragraph should be moved to the earlier paragraph on W-types.
-I'll take a stab at this.}
-
-Generally speaking, the expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict positivity of Inductive type.
 \item 

### 2023-03-18 metatheory-condensed.tex
W-types

--- 2023-03-17 metatheory-condensed.tex
@@
 
+For each pair of types identifying a constructor, the first type may be thought of
+as the non-inductive part of the constructor, and the second type as the arity
+of the inductive part.
+For example, the signature~$\wsig_{\texttt{tm}}$ of the W-type modeling
+the inductive type~\lsti{tm} (\cref{fig:stlc-mechanized})
+is constructed as follows,
+where $\bbzero$ is the bottom type~$\bot$, $\bbone$ the unit type~$\top$, $\bbtwo$ the boolean type,
+and $T_{\texttt{id}}$ a type encoding~\lsti{id}:
+%
+\begin{align*}
+\small
+\begin{tabular}{@{}llll@{}}
+\lsti!tm_unit : tm!
+&
+\lsti!tm_var : id -> tm!
+&
+\lsti!tm_abs : id -> tm -> tm!
+&
+\lsti!tm_app : tm -> tm -> tm!
+\smallskip\\
+$\wsig_{\texttt{tm}}^0 \coloneq \WSigAdd{\WSigEmp}{\bbone}{\lambda \_ . \bbzero}$ 
+& $\wsig_{\texttt{tm}}^1 \coloneq \WSigAdd{\wsig_{\texttt{tm}}^0}{T_{\texttt{id}}}{\lambda \_ . \bbzero}$
+& $\wsig_{\texttt{tm}}^2 \coloneq \WSigAdd{\wsig_{\texttt{tm}}^1}{T_{\texttt{id}}}{\lambda \_ .\bbone} $ 
+& $\wsig_{\texttt{tm}} \coloneq \WSigAdd{\wsig_{\texttt{tm}}^2}{\bbone}{\lambda \_ .\bbtwo}$
+\end{tabular}
+\end{align*}
+
+\noindent
+Thus, the encoding of \lsti{tm_abs} has type
+$T_{\texttt{id}} \to (\bbone \to \El{\wcode{\wsig_{\texttt{tm}}}}) \to \El{\wcode{\wsig_{\texttt{tm}}}}$,
+and that of \lsti{tm_app} has type
+$\bbone \to (\bbtwo \to \El{\wcode{\wsig_{\texttt{tm}}}}) \to \El{\wcode{\wsig_{\texttt{tm}}}}$.
+These types are strictly positive by construction.
+
+%We illustrate the structure of W-type signature (as a list) here.
+%$\wsig_{\texttt{tm}}^0, \wsig_{\texttt{tm}}^1$ are (the subparts of) the
+%signatures of constructors \{"tm\_unit"\} and \{"tm\_unit","tm\_app"\}
+%respectively. Each item in this W-type signature, having two types, corresponds
+%to one constructor, where the first type stores the non-recursive part of the
+%constructor and the second type stores the ``arity'' of the  recursive part. For
+%example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times
+%(\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule
+%\ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone,
+%\wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \lambda \_ . \bbtwo$). This type means
+%there is no information on non-recursive part (thus use unit type $\bbone$) and
+%the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1
+%= \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\lambda \_ .\bbtwo}$.  The story is
+%similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and
+%thus the ``arity'' is zero.\YZ{Feels like it's more straightforward to just list
+%the encodings of the four constructors of tm in a table, assuming an encoding of
+%the identifier type. And this paragraph should be moved to the earlier paragraph
+%on W-types. I'll take a stab at this.}
+%
+%Generally speaking, the expressiveness of W-types comes from the ``arity'', as
+%it encodes the case when recursive structure stores as the output of a function,
+%mirroring the strict positivity of Inductive type.
+
 W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a

### 2023-03-18 metatheory-condensed.tex
more details

@@
 
-For each pair of types identifying a constructor, the first type may be thought of
-as the non-inductive part of the constructor, and the second type as the arity
-of the inductive part.
+For each pair of types identifying a constructor, the first type models
+the non-inductive arguments of the constructor, and the second type models the \emph{arity}
+of the inductive arguments.
 For example, the signature~$\wsig_{\texttt{tm}}$ of the W-type modeling
@@
 \noindent
-Thus, the encoding of \lsti{tm_abs} has type
+While \lsti{tm_unit} and \lsti{tm_var} have no inductive arguments,
+\lsti{tm_abs} has one and \lsti{tm_app} has two.
+The encoding of \lsti{tm_abs} has type
 $T_{\texttt{id}} \to (\bbone \to \El{\wcode{\wsig_{\texttt{tm}}}}) \to \El{\wcode{\wsig_{\texttt{tm}}}}$,

### 2023-03-18 metatheory2.tex
Add details about Grothendieck Universe
edit in appendix

@@
-In this section, we lay out the complete version of the syntax of \TT but in the style of \citeauthor{debruijn1964} indices and \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
-substitutions~$\gm$ and their applications (e.g., in the form of \sub{T}{\gm}) are part
-of the syntax rather than meta-operations. What's more, we also make universe level explicit for completeness.
+This appendix supplements \cref{sec:metatheory2} with the details of the syntax
+and the proofs of the main results.
 
-Other parts of the syntax remained the same as the main text.
+Unlike in \cref{sec:metatheory2}, which uses named binders, in this appendix we
+use \citeauthor{debruijn1964} indices and \emph{explicit
+substitutions}~\cite{abadi1989subst,substcalculus}:
+substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}, which
+applies~$\gm$ to the type~$T$) are part of the syntax rather than
+meta-operations. 
+We work in an intrinsically typed setting: terms are well typed by construction.
+This style of syntax formulation follows a recent
+trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
+known as the ``algebraic presentation'' of MLTT.
+%We introduce judgmental equality ($\equiv$)
+%for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
+%contexts, explicits substitutions, and etc. 
+%This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
+Moreover, universe levels are explicit in this appendix.
 
-We also work in an intrinsic-typed setting, so there is no un-typed preterms appearing anywhere. We introduce judgemental equality ($\equiv$)
-for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
-contexts and explicits substitutions.
+Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
+In particular,
+%following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019},
+the syntax is still considered as being quotiented by the judgmental equalities.
+Quotienting facilitates coercion along equalities.
+Furthermore, the semantic model we develop in this appendix respects these
+judgmental equalities by construction.
+%This means 
+%(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.
 
-This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT.
-
-After the syntax, we detail the models and proofs for syntactic translation (that translate away linkages),  consistency and canonicity.
-
-Again, following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}, we consider our syntax as the result of taking quotient by all of these judgemental equalities. This means 
-(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.
+%After the syntax, we detail the models and proofs for syntactic translation (that translate away linkages),  consistency and canonicity.
 
@@
 So everytime we want to quantify over a universe of types, we have to specify ``how large'' that universe is, that is basically what universe level is measuring.
-
-% We first need the mathmatical setup to talk about universe level, following \citet{sterling2019algebraic}:
-
-% \newcommand{\Set}[1]{\ensuremath{\texttt{Set}_{#1}}}
-
-% {\textbf{Set-theoretic Universe Assumption.}} As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our meta-logic.
-
-% For people familiar with Agda, we can consider $\Set{i}$ as the one in Agda.
-In other words, each $\Set{i}$ is (1).
-closed under dependent function type and dependent tuple type, (2).
-cumulative as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$
 

### 2023-03-18 metatheory2.tex
Fixing typo. Adding ref.

@@
 \noindentparagraph{Explicit Universe Level.} Our treatment of universe level mainly follows that of \citet{altkap2016, kaposi2019gluing}.
-Universe level is mainly to get around the size issue---we cannot refer to(and thus quantifiy over) a set of all set (or a universe of all type).
+Universe level is mainly to get around the size issue---we cannot refer to (and thus quantifiy over) a set of all set (or a universe of all type)~\cite{hurkens1995simplification}.
 So everytime we want to quantify over a universe of types, we have to specify ``how large'' that universe is, that is basically what universe level is measuring.
 
-Now we will need to some attach syntax pieces with natural number indicating universe level.
+Now we will need to attach some syntax pieces with natural number indicating universe levels.
 The explicit universe level $i$ now will attach to each context and type judgement $\goodCtx{\Gm}{i}$ and $\goodType{\Gm}{T}{i}$ basically indicating the context $\Gm$ and $T$ inhabits at what universe.

### 2023-03-18 metatheory2.tex
pass over para on universe levels

--- 2023-03-18 metatheory2.tex
@@
 
-\subsection{Review MLTT in Debruijn Indices, Explicit Subst, and Explicit Universe Level}
+\subsection{MLTT with Explicit Substitutions and Universe Levels}
 
-\TT can be roughly considered as MLTT with linkages so we still start with vanilla MLTT first.
-
+We review the base MLTT fragment of \TT first.
 \input{all-mltt-typing}
@@
 
-\noindentparagraph{Explicit Universe Level.} Our treatment of universe level mainly follows that of \citet{altkap2016, kaposi2019gluing}.
-Universe level is mainly to get around the size issue---we cannot refer to(and thus quantifiy over) a set of all set (or a universe of all type).
-So everytime we want to quantify over a universe of types, we have to specify ``how large'' that universe is, that is basically what universe level is measuring.
+\noindentparagraph{Universe levels.}
 
-Now we will need to some attach syntax pieces with natural number indicating universe level.
-The explicit universe level $i$ now will attach to each context and type judgement $\goodCtx{\Gm}{i}$ and $\goodType{\Gm}{T}{i}$ basically indicating the context $\Gm$ and $T$ inhabits at what universe.
-Apparently, universe level will also be attached to $\cU$.
-For two universe level $i,j$ (as natural number), we use $i \lcup j$ to denotes $\max(i,j)$.
+Universe levels address the size issue---it is unsound to have a set of all sets
+(or a universe of all types).
+The level of a universe specifies ``how large'' that universe is.
+%that is basically what universe level is measuring.
 
-Since we work in an intrinsic-typed setting, some required premises can be omitted without ambiguity.
+%The treatment of universe levels follows \citet{altkap2016, kaposi2019gluing}.
+The judgment form $\goodType{\Gm}{T}{i}$ indicates that (the code of) the type $T$
+inhabits universe~$\cU_i$.
+Similarly, $\goodCtx{\Gm}{i}$ indicates that (the codes of) the types in $\Gm$
+inhabit universe~$\cU_i$.
+The notation $i \lcup j$ denotes $\max(i,j)$.
+
+%Since we work in an intrinsic-typed setting, some required premises can be omitted without ambiguity.
 For omitted context judgement $\goodCtx{\Gm}{k}$, the universe level is always $k$, hinted at the figure of judgement specification.

### 2023-03-18 metatheory2.tex
comments

@@
 
-\noindentparagraph{Debruijn Indices and Explicit Substitution.}  The change of the syntax formulation is important to the simplicity of our proof, as we want to handle the proof about substitutions and terms uniformly---thus making substitution as part of the syntax (explicit substitution) will be helpful.
+\noindentparagraph{De Bruijn Indices and Explicit Substitution.}
 
-Variables are in the form of \citeauthor{debruijn1964} indices:
-\var{n} is the variable bound by the $n$-th closest enclosing binder.
+\Citeauthor{debruijn1964} indices and explicit substitutions make details about
+binders and substitutions explicit.
+%The change of the syntax formulation is important to the simplicity of our
+%proof, as we want to handle the proof about substitutions and terms
+%uniformly---thus making substitution as part of the syntax (explicit
+%substitution) will be helpful.
+
+The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
 For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
@@
 
-To simplify, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$. Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.
+To simplify, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$.
+Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.\YZ{$\pi_1$ and $\pi_2$ should be added to the syntax figure}
 
@@
 This rule applying substitution $\gm$ on the earlier part of the context.
-We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.
+We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.\YZ{The shift rule has bugs: premise says n, conclusion says 1.}
 

### 2023-03-18 metatheory2.tex
question about linkage transformers

@@
 \subsection{MLTT with Explicit Substitutions and Universe Levels}
+\label{sec:mltt-full}
 
@@
 % Copy and Paste Intuition of MLTT from other places to here
-\subsection{Declarative Syntax for \TT : an MLTT variant with Linkages}
-Compared to MLTT, \TT has 3 more judgements about signatures and linkage transformers.
-Moreover, the linkage transformer can be considered as a syntactic sugar (or ``library functions'' in the main text) (we will show why at the end).
-Since it is a syntactic sugar, we will not consider about the linkage transformers during meta-theoretic reasoning.
+\subsection{\TT}
+\label{sec:fmltt-full}
+
+\TT extends the MLTT in \cref{sec:mltt-full} with W-type signatures and W-types,
+linkage signatures and linkages, and linkage transformers.
 
@@
 % Deal with intuition
+
+\noindentparagraph{Linkage transformers as syntactic sugar.}
+
+%Moreover, the linkage transformer can be considered as a syntactic sugar (or
+%``library functions'' in the main text) (we will show why at the end).
+%Since it is a syntactic sugar, we will not consider about the linkage
+%transformers during meta-theoretic reasoning.
+
+%Observe the syntax, linkage transformer and its judgement is defined above 
+%other parts. 
+%
+%So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$.
+Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality
+%%\EDJ{Add substitution equalities for linkage transformers} 
+%given in the syntax.
+In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.
+
+As mentioned in \cref{sec:metatheory2}, the five forms of linkage transformers can be
+thought of a library of functions that are used to construct linkages from other
+linkages.
+In particular, their typing rules are defined in terms of the rest of the typing
+rules in \TT.
+Thus, linkage transformers $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ can be defined
+as syntactic sugar via an inductive type with constructors\YZ{I reinterpreted what you wrote, but this sentence doesn't sound right to me.
+If linkage transformers are meta-level inductive types, aren't they just part of the syntax?}
+$\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$, and
+the syntax $\inh{h}{\ell}$ and $\sub{h}{\gm}$ can be defined by induction on
+this inductive type.
 
@@
 \input{canonicitymodel.tex}
-
-\subsection{Linkage Transformer as Syntax Sugar}
-Observe the syntax, linkage transformer and its judgement is defined above 
-other parts.
-
-So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$.
-Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality
-%\EDJ{Add substitution equalities for linkage transformers} 
-given in the syntax.
-In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.

### 2023-03-18 syn-translate.tex
Shift back to original version of the canonicity model.
edit in syn-translate.tex

@@
-The implementation described in \cref{sec:coqimpl} has hinted the fact that
-linkage can be translated into dependent tuple type.
-The described compilation strategy can be considered as
-a \textit{type preserving syntactic translation}, a function mapping from our syntactic
-model(typed Abstract Syntax Tree(AST)) to a ``subpart''\footnote{Not exactly subpart---the only
-difference lies in "Wrec". In our metatheory, it is constructed
-using linkage; in ``subpart'', it will be using $\Sigma$-types (translation
-of the linkage).} of our syntactic model(typed AST) \textit{without
-linkages}.
+We present a translation from the \TT syntax to the fragment of \TT without
+linkage signatures or linkages.
+The translation echoes how the \Lang plugin is implemented (\cref{sec:coqimpl})
+as a translation to Coq without facilities for family polymorphism.
+In \Lang, families are compiled to Coq modules; similarly, in this appendix
+we translate linkages in \TT to dependent tuples.
+The translation is type-preserving by construction, as we work in an
+intrinsically typed setting.
+%The described compilation strategy can be considered as
+%a \textit{type preserving syntactic translation}, a function mapping from our syntactic
+%model(typed Abstract Syntax Tree(AST)) to a ``subpart''\footnote{Not exactly subpart---the only
+%difference lies in "Wrec". In our metatheory, it is constructed
+%using linkage; in ``subpart'', it will be using $\Sigma$-types (translation
+%of the linkage).} of our syntactic model(typed AST) \textit{without
+%linkages}.
 
@@
 We use $\denotesT{}$ as the translation. This translation behaves like identity
-\YZ{What does "invariant" mean?}\EDJreply{behaves like identity. Sorry I should have said "subpart is invariant under this translation" instead}
 under the subpart. Since we are in an intrinsic setting, everything here
@@
   \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= 
-    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(\El{(\app{("pjr"^3~\denotesT{\sigma})})[(\SubstWeak{1}, \app{(\fst{("pjr"^2~{\denotesT{\sigma}})})})]}) \\
+    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(\El{(\app{("snd"^3~\denotesT{\sigma})})[(\SubstWeak{1}, \app{(\fst{("snd"^2~{\denotesT{\sigma}})})})]}) \\
   \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &=
-  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(\El{(\fst{(\snd{\denotesT{\sigma}})})})(\El{(\app{("pjr"^3~\denotesT{\sigma})} )}) \\
+  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(\El{(\fst{(\snd{\denotesT{\sigma}})})})(\El{(\app{("snd"^3~\denotesT{\sigma})} )}) \\
   &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\

### 2023-03-18 metatheory2.tex
Reply the comment.

@@
 To simplify, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$.
-Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.\YZ{$\pi_1$ and $\pi_2$ should be added to the syntax figure}
+Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.\YZ{$\pi_1$ and $\pi_2$ should be added to the syntax figure}\EDJreply{I added it, it is right after sub/dbj/shift.}
 
@@
 This rule applying substitution $\gm$ on the earlier part of the context.
-We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.\YZ{The shift rule has bugs: premise says n, conclusion says 1.}
+We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.\YZ{The shift rule has bugs: premise says n, conclusion says 1.}\EDJreply{Thanks, fixed.}
 
@@
 linkages.
-In particular, their typing rules are defined in terms of the rest of the typing
+In particular, their typing rules is defined in terms of the rest of the typing
 rules in \TT.
-Thus, linkage transformers $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ can be defined
-as syntactic sugar via an inductive type with constructors\YZ{I reinterpreted what you wrote, but this sentence doesn't sound right to me.
-If linkage transformers are meta-level inductive types, aren't they just part of the syntax?}
+Thus, linkage transformers $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ can be defined as syntactic sugar via an inductive type with constructors
+\YZ{I reinterpreted what you wrote, but this sentence doesn't sound right to me.
+If linkage transformers are meta-level inductive types, aren't they just part of the syntax?}\EDJreply{I think the reason that they are syntactic sugar is because they can be interpreted using FMLTT syntax(without them), it is fine either the syntactic sugar is defined via inductive type (and then interpreted), or directly interpreted.
+Maybe I should remove this misleading sentence?
+\\ What's more, the specification/definition of syntactic sugar itself, is inductive type.
+Because a linkage transformer looks inductive.
+Maybe I should directly separate the typing rules for linkage transformer as another figure to emphasize it as an inductive construction on the rest FMLTT syntax?}
 $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$, and

### 2023-03-18 syn-translate.tex
Add universe for Sigr

@@
 as a translation to Coq without facilities for family polymorphism.
-In \Lang, families are compiled to Coq modules; similarly, in this appendix
-we translate linkages in \TT to dependent tuples.
+%In \Lang, families are compiled to Coq modules; similarly, in this appendix
+%we translate linkages in \TT to dependent tuples.
 The translation is type-preserving by construction, as we work in an
@@
 
-This ``subpart'' is the standard MLTT with our
-(unconventional) formulation of W-types,
-and recursors that use $\Sigma$-types to aggregate handlers.
-
-Basically, we can compile away (the derivation of) judgement $\goodSig{\Gamma}{v}{n}$,
-the type $\cL$, $\cP$, and related terms, by transforming them into
-corresponding $\Sigma$-types.
+%This ``subpart'' is the standard MLTT with our
+%(unconventional) formulation of W-types,
+%and recursors that use $\Sigma$-types to aggregate handlers.
+%
+%Basically, we can compile away (the derivation of) judgement $\goodSig{\Gamma}{v}{n}$,
+%the type $\cL$, $\cP$, and related terms, by transforming them into
+%corresponding $\Sigma$-types.
 
@@
 
-We use $\denotesT{}$ as the translation. This translation behaves like identity
-under the subpart. Since we are in an intrinsic setting, everything here
-is well-typed (well-formed)including this translation. Thus
-$\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$
-(i.e. every type is tracking its context). In other words, every
-derivation of any judgement will track its ``context''.
+We define the translation $\denotesT{}$ below; the fragment of the syntax
+irrelevant to linkages is translated using the identity function and is thus elided.
+When the context~$\Gm$ is clear, we use $\denotesT{T}$ to mean $\denotesT{\goodType{\Gamma}{T}{}}$.
+
+We first define a new type $\goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}$  when
+$\goodCtx{\Gamma}{}$  is well-formed and $n \in \mathbb{N}$
+such that $\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}}$.
+We make $({\Sigr{\Gamma}{n}{}})$ and $\denotesT{\TyTkg {-}}$ mutually recursive,
+defined by induction on $n$ and the signature length.
+% \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
+% &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\
 
 \begin{align*}
-  \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
-  \text{such that } &\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}} \\
-  \text{We make } & ({\Sigr{\Gamma}{n}{}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ mutually recursive, } \\
-  & \text{defined by induction on $n$ and the signature length} \\  
-  % \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
-  % &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
   {\Sigr{\Gamma}{(n+1)}{j}} &= 
-    \Sigma~(\Sigr{\Gamma}{n}{})
+    \Sigma~(\Sigr{\Gamma}{n}{j})
           ~(\Sigma~\cU_j \\
-          & \quad \quad ~(\Sigma~(\TyPi{\denotesT{\TyTkg {(\var{0}[\SubstWeak{1}])}}}{\El{\sub{\var{0}}{\SubstWeak{1}}}}) \\
+          & \quad \quad ~(\Sigma~(\TyPi{\denotesT{\TyTkg {\sub{\var{0}}{\SubstWeak{1}}}}}{\El{\sub{\var{0}}{\SubstWeak{1}}}}) \\
           & \quad \quad \quad \TyPi{\El{\sub{\var{0}}{\SubstWeak{1}}}}{\cU_j})) \\
@@
   \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= 
-    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(\El{(\app{("snd"^3~\denotesT{\sigma})})[(\SubstWeak{1}, \app{(\fst{("snd"^2~{\denotesT{\sigma}})})})]}) \\
+    \Sigma~\denotesT{\TyTkg {\lsigproj{1}{\sigma}}}~(\El{\app{"snd"^3~\denotesT{\sigma}}[(\SubstWeak{1}, \app{\fst{("snd"^2~{\denotesT{\sigma}})}})]}) \\
   \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &=
-  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(\El{(\fst{(\snd{\denotesT{\sigma}})})})(\El{(\app{("snd"^3~\denotesT{\sigma})} )}) \\
+  \denotesT{\TyLkg{\lsigproj{1}{\sigma}}} \times \Pi(\El{\fst{(\snd{\denotesT{\sigma}})}})(\El{\app{"snd"^3~\denotesT{\sigma}}}) \\
   &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
   \denotesT{\goodSigU{j}{\Gamma}{\LSigEmp}{0}} &= \unit \\
-  \denotesT{\goodSigU{j}{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{(\codety{T})}) \\ 
+  \denotesT{\goodSigU{j}{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{\codety{T}}) \\
   \denotesT{\goodTerm{\Gamma}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \unit \\
-  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\TyLkg{(\LSigAdd{\sigma}{s}{T})}}} &= (\denotesT{m}, \lam{t}) \\ 
+  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\TyLkg{\LSigAdd{\sigma}{s}{T}}}} &= (\denotesT{m}, \lam{t}) \\
   \denotesT{\goodTerm{\Gamma}{\Tkg{m}}{\TyTkg {\sigma}}} & \text{ is defined upon } \denotesT{m} \text{ and inductively on the signature length} \\
@@
   (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{(\snd{\denotesT{m}})}, f = p_f\nu~\sigma 
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{\snd{\denotesT{m}}}, f = p_f\nu~\sigma 
 \end{align*}
@@
 
-The main idea of this syntactic translation is that, we make sure
-$\goodSig{\Gamma}{\nu^+ \_ }{n+1}$ and $\goodSig{\Gamma}{\mu^+ \_}{n+1}$
-wrapped into a dependent pair. These two definitions decide the
-syntactic translation for $\goodSig{\Gamma}{\_}{n}$ and $\cL$. The
-syntactic translation for $\cP$ are decided by their
-$\beta$-rules.
+The main idea is that
+%$\goodSig{\Gamma}{\nu^+ \_ }{n+1}$
+a linkage $\goodSig{\Gamma}{\mu^+ \_}{n+1}$
+is translated to a non-dependent tuple while introducing explicit universal
+quantification to the second component of the tuple;
+the universal quantification achieves late binding.
+%These two definitions decide the syntactic translation for
+%$\goodSig{\Gamma}{\_}{n}$ and $\cL$.
+The translation for $\cP$ is given by the relevant $\beta$-rules.
 
@@
 
-We omit all the validation of equational rules ($\beta,\eta$ and substitution) here.
-However, when we mutual recursively define
+We omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
+However, when we mutually recursively define
 $\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}{}} $, and
@@
 
-Ultimately, we have constructed a model for our syntax using only a
-``subpart'' of our syntax -- without these linkages. It is also
-possible to compile our $\goodWSig{\Gamma}{w}{n}$ and W-type into a more
-conventional formulation, but the recursor will always require a sigma
-type as handler and thus the elimination rule for our W-type is still
-non-conventional.
-
-This translation justifies the intuition -- a linkage is actually a
-module \textbf{with universal quantification wrapping its fields to
-achieve late binding}.
-
-After syntactic translation, we can directly have consistency and
-canonicity \textit{if we assume the consistency and canonicity of the
-``subpart''}. However, due to our non-conventional formulation of the
-inductive type, it is better we don't impose such assumption and
-directly give a consistency and canonicity proof.
+We have constructed a model for the \TT syntax using
+only the linkage-irrelevant fragment of the syntax.
+%It is also
+%possible to compile our $\goodWSig{\Gamma}{w}{n}$ and W-type into a more
+%conventional formulation, but the recursor will always require a sigma
+%type as handler and thus the elimination rule for our W-type is still
+%non-conventional.
+%
+%This translation justifies the intuition -- a linkage is actually a
+%module \textbf{with universal quantification wrapping its fields to
+%achieve late binding}. 
+%
+We would have consistency and canonicity for \TT immediately, if we could assume
+the consistency and canonicity of the linkage-irrelevant fragment.
+However, because our formulation of W-types is unconventional,
+we choose not to take such an assumption for granted and choose to directly prove
+consistency and canonicity for \TT.

### 2023-03-18 metatheory2.tex
resolve comment; please check for accuracy

--- 2023-03-18 metatheory2.tex
@@
 Thus, linkage transformers $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ can be defined
-as syntactic sugar via an inductive type with constructors\YZ{I reinterpreted what you wrote, but this sentence doesn't sound right to me.
-If linkage transformers are meta-level inductive types, aren't they just part of the syntax?}
+as syntactic sugar via an \TT inductive type with constructors
+%\YZ{I reinterpreted what you wrote, but this sentence doesn't sound right to me.
+If linkage transformers are meta-level inductive types, aren't they just part of the syntax?}\EDJreply{I think the reason that they are syntactic sugar is because they can be interpreted using FMLTT syntax(without them), it is fine either the syntactic sugar is defined via inductive type (and then interpreted), or directly interpreted.
+Maybe I should remove this misleading sentence?
+\\ What's more, the specification/definition of syntactic sugar itself, is inductive type.
+Because a linkage transformer looks inductive.
+Maybe I should directly separate the typing rules for linkage transformer as another figure to emphasize it as an inductive construction on the rest FMLTT syntax?}
 $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$, and
@@
 
-\subsection{Syntactic Translation to get rid of Linkages}
+\subsection{A Translation that Compiles Away Linkages}
 

### 2023-03-18 metatheory2.tex
edits

@@
 
-%Moreover, the linkage transformer can be considered as a syntactic sugar (or
-%``library functions'' in the main text) (we will show why at the end).
-%Since it is a syntactic sugar, we will not consider about the linkage
-%transformers during meta-theoretic reasoning.
-
 %Observe the syntax, linkage transformer and its judgement is defined above 
@@
 Thus, linkage transformers $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ can be defined
-as syntactic sugar via an \TT inductive type with constructors
+as syntactic sugar via an inductive type (in the meta\-logic) with four constructors
 %\YZ{I reinterpreted what you wrote, but this sentence doesn't sound right to me.
@@
 Maybe I should directly separate the typing rules for linkage transformer as another figure to emphasize it as an inductive construction on the rest FMLTT syntax?}
-$\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$, and
-the syntax $\inh{h}{\ell}$ and $\sub{h}{\gm}$ can be defined by induction on
-this inductive type.
+$\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, and $\InhNest{}{}$.
+Moreover, $\inh{h}{\ell}$ and $\sub{h}{\gm}$ can be defined
+as recursive functions (in the meta\-logic) by induction on this inductive type.
+We will treat linkage transformers as syntactic sugar in the meta\-theoretic development
+in the rest of \cref{sec:complete-fmltt}.
 

### 2023-03-18 canonicitymodel.tex
Save.
Took a pass over appendix. A question was added.

@@
-Now we prove the consistency and canonicity using canoncity model.
-We follow the reducibility argument from~\cite{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
+Now we prove consistency and canonicity using a canoncity model.
+We follow the reducibility argument of
+\citet{kaposi2019gluing,coquand2018canonicity,sterling2019algebraic}
+to construct this canonicity model.
+%
 Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax.
@@
 \begin{theorem}[Canonicity]
-  For a closed boolean term $\goodTerm{\cdot}{t}{\cB}$, we have $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$ hold.
+\label{thm:canonicity-appendix}
+  If $\goodTerm{\cdot}{t}{\cB}$, then either $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$.
 \end{theorem}
 
-Canonicity is a basic criteria to consider a dependent type theory to be ``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
-We can even argue that, if this theorem is proven in a computable meta-logic, 
-then by Curry-Howard Correspondence, this theorem provides a big-step interpreter for closed term of boolean type.\YZ{Can we say something about strong/weak normalization?}\EDJreply{
-  No. \\
-  1.
+Canonicity is a basic criterion that must be satisfied by a dependent type theory serving as
+``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
+We can even argue that, if this theorem is proven in a computable meta\-logic, 
+then by Curry–Howard Correspondence, the canonicity theorem provides a big-step
+interpreter for closed term of boolean type.%\YZ{Can we say something about
+%strong/weak normalization?}\EDJreply{
+%  No. \\
+%  1.
 In the declarative system, we cannot talk about strong/weak normalization (because we don't have operational semantic, if you recall the definition of strong/weak normalization, it based on reduction/path of steps/op semantic).
@@
 \\
-  2. In declarative system, we only say ``normalization'' directly, which means, we construct a function that \\
-      for arbitrary class of (judgemental-equivalent) terms, it will return one syntax representation (called normal-form) for each class \\
-    (apparently most people will make the normal form  without beta redex, however, interestingly, they may do eta expansion instead of eta reduction because theoretically easier) \\ 
-  3.
+%  2. In declarative system, we only say ``normalization'' directly, which means, we construct a function that \\
+%      for arbitrary class of (judgemental-equivalent) terms, it will return one syntax representation (called normal-form) for each class \\
+%    (apparently most people will make the normal form  without beta redex, however, interestingly, they may do eta expansion instead of eta reduction because theoretically easier) \\ 
+%  3.
 So translation won't work because it compiles away the normal form of linkage (the $\mu^+$).
 This part requires a direct proof which construct normal form and neutral form for our calculus, and a normalization model from which we extract a runnable interpreter for open terms \\
-  4. Of course we can conjecture we have that normalization function \\
-  5.
+%  4. Of course we can conjecture we have that normalization function \\
+%  5.
 Also we do have close term normalization, which is canonicity.
 Or more operationally speaking, we have termination (a runnable interpreter) for close term (since QIIT is runnable, the canoncity model itself is the runnable interpreter), \\
-  but we don't have termination (a runnable interpreter) for open terms.  \\ 
-  So earlier you say ``normalize'' a recursion on an inductive type, that is no doable -- if the programmer wants to program on meta-theory, they will have to manually write out the a chain of judgemental equality (with a bunch of beta rules inside) by themselves.
-}
+%  but we don't have termination (a runnable interpreter) for open terms.  \\ 
+%  So earlier you say ``normalize'' a recursion on an inductive type, that is no doable -- if the programmer wants to program on meta-theory, they will have to manually write out the a chain of judgemental equality (with a bunch of beta rules inside) by themselves.
+%}
 
-We first need the mathmatical setup to interpret universe level, following \citet{sterling2019algebraic}:
+We first need the mathematical setup to interpret universe levels, following \citet{sterling2019algebraic}:
 
@@
 
-\begin{assumption}[Set-theoretic Universe Assumption] As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our ambient meta-logic.
+\begin{assumption}[Set-theoretic Universe Assumption] As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our ambient meta\-logic.
 \end{assumption}
 
-We can roughly consider each Grothendieck Universe $\Set{i}$ as a $\Set{i}$ in Agda. In other words, each $\Set{i}$ is 
+We can roughly consider each Grothendieck universe $\Set{i}$ as the $\Set{i}$ in Agda. 
 \begin{itemize}
-  \item Closed under dependent function type and dependent tuple type.
-For example, later, for our interpretation of dependent function type, when we have $\denotesC{A}, \denotesC{B} \in \Set{i}$, we will have $\denotesC{\TyPi{A}{B}} \in \Set{i}$
-  \item the hierarchy is cumulative as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$
-  \item Thus if $\denotesC{A} \in \Set{i}, \denotesC{B} \in \Set{j}$, we will have $\denotesC{\TyPi{A}{B}} \in \Set{i} \cup \Set{j} = \Set{i \lcup j}$
+  \item Each $\Set{i}$ is closed under dependent function types and dependent
+  pair types. For example, later, for our interpretation of dependent function
+  types, when we have $\denotesC{A}, \denotesC{B} \in \Set{i}$, we will have
+  $\denotesC{\TyPi{A}{B}} \in \Set{i}$.
+  \item The universe hierarchy is cumulative, as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$.
+  \item Thus if $\denotesC{A} \in \Set{i}, \denotesC{B} \in \Set{j}$, we will
+  have $\denotesC{\TyPi{A}{B}} \in \Set{i} \cup \Set{j} = \Set{i \lcup j}$.
 \end{itemize}
 
-Like most logical relation proof,  we interpret each judgement and inductively interpret each syntax piece.
-We are working on an intrinsic setting thus even if we omit context for brevity, the syntax piece is still well-typed.
-However, unlike most logical relation proof, we are having a proof-relevant logical realtion, essential for modeling universes properly~\cite{coquand2018canonicity}.
-Our canonicity model at MLTT part is very close to the one shown in \citet{coquand2018canonicity} and \citet{sterling2019algebraic}.
+Like most logical-relations proofs, we interpret each judgment and inductively
+interpret each syntax piece. We are working in an intrinsic setting; thus, even if
+we omit contexts for brevity, the syntax piece is still well-typed. However,
+unlike most logical-relations proofs, our logical-relations model is proof-relevant,
+which is essential for modeling universes
+properly~\cite{coquand2018canonicity}.
+Our canonicity model for the base MLTT fragment is close to the ones
+in \citet{coquand2018canonicity} and \citet{sterling2019algebraic}.
 % We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
@@
 
-The counter-intuitive part of the proof-relevant logical relation is the treatment of proposition---a proposition is encoded as a subsingleton data.
-For example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \ | \text{ when } t \equiv \false  \}$ represents a disjoint union of two subsingleton set.
-The former one will be a singleton if we have a proof $t \equiv \true$ or an empty set otherwise, same as the later one.
-In other words, $\star^1$ witness a proof that $t \equiv true$.
-The same idea applies to the latter set.
-Here $\star$, $\star^1$, $\star^2$ are just some arbitrary fixed elements.
+%The difficult part of the proof-relevant logical-relations model is the
+%treatment of propositions---
+A proposition is encoded as a subsingleton data. For
+example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \
+| \text{ when } t \equiv \false  \}$ represents a disjoint union of two
+subsingleton sets, where the first one is either a singleton if we have a proof of
+$t \equiv \true$ or an empty set otherwise.
+In other words,
+$\star^1$ witnesses a proof that $t \equiv true$.
+Here, $\star$, $\star^1$, $\star^2$ are just some arbitrary fixed elements.
 
-Given above model for MLTT, we should have such a function ${\Pi}^c$ s.t.
-${\Pi}^c({\denotesC{A}},{\denotesC{B}}) = \denotesC{\TyPi{A}{B}}$.
-We want to reuse this function when defining the rest of the model of \TT, because as syntactic translation hinted, \TT reuses a lot from MLTT.
-Type-theoretical speaking, this ${\Pi}^c$ is using the \emph{internal dependent function type} of the above model.
-However this is not possible because the definition $\denotesC{\TyPi{A}{B}}$ is not solely based on the $\denotesC{A}$ and $\denotesC{B}$, but also on the syntax $\goodType{\Gm}{A}{}$ and $\goodType{\Gm, A}{B}{}$.
+Given the above model for MLTT, there should be a function ${\Pi}^c$ such
+that ${\Pi}^c({\denotesC{A}},{\denotesC{B}}) = \denotesC{\TyPi{A}{B}}$. We will
+use this function when defining the logical-relations model for the rest of \TT,
+Type-theoretically speaking, this ${\Pi}^c$ uses the \emph{internal dependent
+function type} of
+the above model.
+However, such a function $\Pi^c$ is not yet possible because the definition
+$\denotesC{\TyPi{A}{B}}$ is not based solely on $\denotesC{A}$ and
+$\denotesC{B}$,
+but also on the syntax $\goodType{\Gm}{A}{}$ and $\goodType{\Gm, A}{B}{}$.
 
-Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$, that will return the syntax piece $S$ as well.\footnote{This is also called glued interpretation in \citet{sterling2019algebraic}} Thus we can have a function $\Glued{\Pi}$ s.t.
-$\GluedPi{\denotesCC{A}}{\denotesCC{B}} = \denotesCC{\TyPi{A}{B}}$ now that syntax is also provided.
-Similarly, we can have $\Glued{\Sigma}$, $\Gpair{a}{b}$  and $\GSubstExt{\gm}{t}$  for dependent tuple type, depende tuple construction and substitution extension, and also all other constructions.
-What's more, given $\Glued{S}$ a pair of syntax and model , we denote $\Gmodel{\Glued{S}}$ as getting the (canonicity) model part (the second element in that pair).
+Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$
+that returns the syntax piece $S$ as well.\footnote{This is also called
+glued interpretation in \citet{sterling2019algebraic}.}
+Thus, we can have a
+function $\Glued{\Pi}$ such that $\GluedPi{\denotesCC{A}}{\denotesCC{B}} =
+\denotesCC{\TyPi{A}{B}}$ now that the syntax is available.
+Similarly, there are functions $\Glued{\Sigma}$, $\Gpair{a}{b}$  and
+$\GSubstExt{\gm}{t}$ for dependent pair types, dependent pairs, and
+substitution extension (and more for other constructions).
+Furthermore, given $\Glued{S}$ (i.e., the syntax and its semantic interpretation),
+we use $\Gmodel{\Glued{S}}$ to mean the latter of the two.\YZ{why not just $\denotesC{S}$?}
 % These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.
@@
 
-There are actually more \emph{internal} type-theoretic construction that we need:
+We need more \emph{internal} type-theoretic constructions:
 
-We denote $\GCon{k} \coloneq \sum_{\goodCtx{\Gm}{k}} \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k}$
+We define $\GCon{k} \coloneq \sum_{\goodCtx{\Gm}{k}} \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k}$
 
@@
 
-These four sets are collecting the glued interpretation:
-basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$, and etc.
+These four sets are collecting the glued interpretation.
+For each well-formed type $\goodType{\Gm}{T}{j}$, we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$;
+for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$;
+etc.
 
-Note that, we have $\prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \Gmodel{\Glued{\Gm}}(\gm)}$ prefix in $\GTy{}{}$, $\GTm{}{}$,$\GSub{}{}$.
-The interesting fact is, given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$ and $\gm' \in \Gmodel{\Glued{\Gm}}(\gm)$, we can consider $(\gm, \gm')$ as an element of $\GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
-Thus we consider the pair $(\gm, \gm')$ as $\Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
+Notice that
+$\prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \Gmodel{\Glued{\Gm}}(\gm)}$
+is part of $\GTy{\!}{\!}$, $\GTm{\!}{\!}$, and $\GSub{\!}{\!}$.
+The interesting fact is that, given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$
+and $\gm' \in \Gmodel{\Glued{\Gm}}(\gm)$, we can consider $(\gm, \gm')$ as an
+element of $\GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
+Thus, we consider the pair $(\gm,
+\gm')$ as $\Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
 % Note that, for arbitrary $\Glued{\Gm} \in \GCon{k}$, we can consider it as $\Glued{\gm}$ as an element of $\GSub{\Glued{\cdot}}{}$
 
-Then we define $\GMWSig{j}{\Glued{\Gm}}{n} \coloneq "Vector"^n~\sum_{\Glued{A} \in \GTy{j}{\Glued{\Gm}}} \GTy{j}{(\GSubstExt{\Glued{\Gm}}{\Glued{A}})}$ for $\Glued{\Gm} \in \GCon{k}$, which is basically a list of a pair of types of length $n$.
-As before, we can define glued interpretation $\GWSig{j}{\Glued{\Gm}}{n} \coloneq \{ \wsig : \goodWSig{\cdot}{\wsig}{n} \} \times \GMWSig{j}{\Glued{\Gm}}{n}$.
-This will be useful when we interpret judgement for W-type signature.
+Then we define $\GMWSig{j}{\Glued{\Gm}}{n} \coloneq "Vector"^n~\sum_{\Glued{A}
+\in \GTy{j}{\Glued{\Gm}}} \GTy{j}{(\GSubstExt{\Glued{\Gm}}{\Glued{A}})}$ for
+$\Glued{\Gm} \in \GCon{k}$, a length-$n$ list of pairs of types.
+As before, we can define glued interpretation $\GWSig{j}{\Glued{\Gm}}{n}
+\coloneq \{ \wsig : \goodWSig{\cdot}{\wsig}{n} \} \times
+\GMWSig{j}{\Glued{\Gm}}{n}$. This will be useful when we interpret judgments for
+W-type signatures.
 
-Now we can extend the above model to \TT , we still start with the judgement. Since linkage transformers are just syntactic sugar, we will not interpret it.
-The intuition of the model is similar to that of the syntactic translation---use the canonicity model of sigma type (from the earlier work) for linkage type and using inductive facility in the meta-logic to justify our W-type.
+Now we can extend the model above to include \TT constructs.
+%We still start with the judgments.
+%Since linkage transformers are just syntactic sugar, we will not interpret it.
+The key idea of this extension is similar to that of the syntactic translation:
+we interpret linkage types using the canonicity model of sigma types (developed in the prior work).
+We use the inductive facility of the meta\-logic to justify our W-types.
 
@@
 \end{align*}
-together and mutual recursively defined with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$) ....
 
+The semantic interpretations above are 
+mutually recursively defined with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)
+%
 \begin{align*}
@@
 \end{align*}
-... and its eliminator principle $\CWrec$. 
+and its eliminator $\CWrec$
 \begin{align*}
@@
 
-Note that in $\CWmodel$, the $\Glued{b}$ uses the definition of $\denotesC{\wcode{\wsig}}$, which after unfolding, will refers to $\CWmodel$ again and in strictly positive position.
-So it is a well-founded definition.
-We also don't distinguish $(b, b^C)$ and $\Glued{b}$ for simplicity.
+Note that in $\CWmodel$, the $\Glued{b}$ uses the definition of
+$\denotesC{\wcode{\wsig}}$, which after unfolding, recursively references $\CWmodel$
+in a strictly positive position. %So it is a well-founded definition.
+We do not distinguish $(b, b^C)$ and $\Glued{b}$ for simplicity.
 
-The first observation is that, with this model, we know our system is consistent---we cannot derive $\goodTerm{\cdot}{t}{\bot}$.
-Otherwise, we have $\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp)(\star) \in \denotesC{\bot}(\SubstEmp)(\star)(\sub{t}{\gm}) = \emptyset$, that is an element in an empty set, a contradiction.
+The first consequence of this model is the consistency of \TT---we cannot derive
+$\goodTerm{\cdot}{t}{\bot}$. Otherwise, we would have an element in the empty set,
+$\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp)(\star) \in
+\denotesC{\bot}(\SubstEmp)(\star)(\sub{t}{\gm}) = \emptyset$, a contradiction.
 
 \begin{theorem}[Consistency]
-  We cannot derive a closed term of bottom type, i.e. ,$\cdot \not \vdash t : \bot$ for arbitrary $t$.
+  The typing judgment $\goodTerm{\cdot}{t}{\bot}$ is not derivable for any term~$t$.
 \end{theorem}
 
-What's more, with the above model, we can map an arbitrary closed boolean term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp)(\star) = \star^1$ or $\star^2$, witnessing the proof of $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
+Next, with the logical-relations model, we can map an arbitrary closed boolean
+term $\goodTerm{\cdot}{t}{\cB}$ to get the result
+$\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp)(\star) = \star^1$ or $\star^2$,
+witnessing the proof of $t \equiv \true \text{ or } t \equiv\false$, arriving at
+\cref{thm:canonicity-appendix}.
 
-With the logical relation and the help of eta rule, we figure out the following canonical form
+Further, with the help of eta rules, we have the following canonical forms
 \begin{theorem}[Canonical Forms].
@@
     \item if $\goodTerm{\cdot}{t}{\Sigma A B}$ then $\goodTerm{\cdot}{t = (a, b)}{\Sigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[(id, a)]}$
-    \footnote{We emphasize the last one because $\cCt t$ is a dependent pair}
+    \footnote{We emphasize the last one because $\Tkg{\ell}$ is a dependent pair}
   \end{itemize}

### 2023-03-18 syn-translate.tex
edits
Took a pass over appendix. A question was added.

--- 2023-03-18 syn-translate.tex
@@
 
-We use $\denotesT{}$ as the translation. This translation behaves like identity
-under the subpart. Since we are in an intrinsic setting, everything here
-is well-typed (well-formed)including this translation. Thus
-$\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$
-(i.e. every type is tracking its context). In other words, every
-derivation of any judgement will track its ``context''.
+We define the translation $\denotesT{}$ below; the fragment of the syntax
+irrelevant to linkages is translated using the identity function and is thus elided.
+When the context~$\Gm$ is clear, we use $\denotesT{T}$ to mean $\denotesT{\goodType{\Gamma}{T}{}}$.
+
+We first define a new type $\goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}$  when
+$\goodCtx{\Gamma}{}$  is well-formed and $n \in \mathbb{N}$
+such that $\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}}$.
+We make $({\Sigr{\Gamma}{n}{}})$ and $\denotesT{\TyTkg {-}}$ mutually recursive,
+defined by induction on $n$ and the signature length.
+% \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
+% &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\
 
 \begin{align*}
-  \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
-  \text{such that } &\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}} \\
-  \text{We make } & ({\Sigr{\Gamma}{n}{}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ mutually recursive, } \\
-  & \text{defined by induction on $n$ and the signature length} \\  
-  % \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
-  % &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
   {\Sigr{\Gamma}{(n+1)}{j}} &= 

### 2023-03-18 canonicitymodel.tex
edits

@@
-Now we prove consistency and canonicity using a canoncity model.
+Now we prove canonicity (and consistency) for \TT using a logical-relations model.
 We follow the reducibility argument of
-\citet{kaposi2019gluing,coquand2018canonicity,sterling2019algebraic}
-to construct this canonicity model.
-%
-Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax.
-Meanwhile, our proof will still work directly on the earlier introduced syntax for readability\footnote{We still take quotient, so ultimately our syntax should be considered as an instance of any of the above logical framework, without exposing the reader too many technical details}.
+\citet{kaposi2019gluing}, \citet{coquand2018canonicity}, and \citet{sterling2019algebraic}
+to construct our model.
+The meta\-languages of these prior models are based on QIITs, categories with
+families \cite{dybjer1995internal}, and the generalized algebraic theory
+\cite{cartmell1986generalised}, respectively.
+Without exposing the reader to too many technical details, our meta\-language
+should be understood an instance of any of the above logical frameworks.
+%Our model works directly on the earlier introduced syntax for readability\footnote{We still
+%take quotient, so ultimately our syntax should be considered as an instance of
+%any of the above logical framework, without exposing the reader too many
+%technical details}.
 
-We will first specify the formulation of the most basic canonicity theorem.
+We state the canonicity theorem first:
 
@@
 
-Canonicity is a basic criterion that must be satisfied by a dependent type theory serving as
-``a programming language or as a computational foundation for mathematics''.\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
+\noindent
+Canonicity is a basic criterion that should be met by a dependent type theory serving as
+a programming language or as a computational foundation for mathematics.
+%\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
 We can even argue that, if this theorem is proven in a computable meta\-logic, 
-then by Curry–Howard Correspondence, the canonicity theorem provides a big-step
-interpreter for closed term of boolean type.%\YZ{Can we say something about
+then by the Curry–Howard correspondence, the canonicity theorem provides a big-step
+interpreter for closed terms of the boolean type.%\YZ{Can we say something about
 %strong/weak normalization?}\EDJreply{
@@
 Further, with the help of eta rules, we have the following canonical forms
-\begin{theorem}[Canonical Forms].
+\begin{theorem}[Canonical Forms]\hfill
   \begin{itemize}
-    \item if $\goodTerm{\cdot}{t}{El~(\wcode{\tau})}$ with $\goodWSig{\cdot}{\tau}{n}$, then $\goodTerm{\cdot}{t \equiv "Wsup"~j~a~b}{El~(\wcode{\tau})}$ for some $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot, B[(id, a)]}{b}{El~(\wcode{\tau})}$ and $j < n$
-    \item if $\goodTerm{\cdot}{t}{\mathbb{B}}$ then $\goodTerm{\cdot}{t \equiv "tt"}{\mathbb{B}}$ or $\goodTerm{\cdot}{t \equiv "ff"}{\mathbb{B}}$ 
-    \item if $\goodTerm{\cdot}{t}{\cL\sigma}$ with $\goodSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv \LkgAdd{o}{t}}{\cL\sigma}$
+    \item If $\goodTerm{\cdot}{t}{El~(\wcode{\tau})}$ with $\goodWSig{\cdot}{\tau}{n}$, then $\goodTerm{\cdot}{t \equiv "Wsup"~j~a~b}{El~(\wcode{\tau})}$ for some $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot, B[(id, a)]}{b}{El~(\wcode{\tau})}$ and $j < n$
+    \item If $\goodTerm{\cdot}{t}{\mathbb{B}}$ then $\goodTerm{\cdot}{t \equiv "tt"}{\mathbb{B}}$ or $\goodTerm{\cdot}{t \equiv "ff"}{\mathbb{B}}$ 
+    \item If $\goodTerm{\cdot}{t}{\cL\sigma}$ with $\goodSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv \LkgAdd{o}{t}}{\cL\sigma}$
 
       $\quad$ for some $\goodTerm{\cdot}{o}{\cL(\lkgproj{1}{\sigma})}$ and $\goodTerm{\cdot, \lsigproj{2}{\sigma}}{t}{\lkgproj{2}{\sigma}}$
-    \item if $\goodTerm{\cdot}{t}{\Sigma A B}$ then $\goodTerm{\cdot}{t = (a, b)}{\Sigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[(id, a)]}$
+    \item If $\goodTerm{\cdot}{t}{\Sigma A B}$ then $\goodTerm{\cdot}{t = (a, b)}{\Sigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[(id, a)]}$
     \footnote{We emphasize the last one because $\Tkg{\ell}$ is a dependent pair}

### 2023-03-18 syn-translate.tex
remove gratuitous parens
edits
Rewrite syn-translate. Please review

--- 2023-03-18 syn-translate.tex
@@
 
-We first define a new type $\goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}$  when
+We will first define two helpers
+\begin{itemize}
+  \item a new type $\goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}$  when
   $\goodCtx{\Gamma}{}$  is well-formed and $n \in \mathbb{N}$
-such that $\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}}$.
-We make $({\Sigr{\Gamma}{n}{}})$ and $\denotesT{\TyTkg {-}}$ mutually recursive,
-defined by induction on $n$ and the signature length.
-% \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
-% &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\
+  such that $\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}}$
+  \item and a function $\mathscr{P} : \{\lsig' : \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{n}{j}} \} \to \{ T : \goodType{\Gm}{T}{} \}$ map one syntax term of $\Sigr{\Gm}{n}{j}$ into a type.
+\end{itemize}
+mutually inductively on $n$, the signature length.
 
@@
           ~(\Sigma~\cU_j \\
-          & \quad \quad ~(\Sigma~(\TyPi{\denotesT{\TyTkg {\sub{\var{0}}{\SubstWeak{1}}}}}{\El{\sub{\var{0}}{\SubstWeak{1}}}}) \\
+          & \quad \quad ~(\Sigma~(\TyPi{{\mathscr{P}({\sub{\var{0}}{\SubstWeak{1}}})}}{\El{\sub{\var{0}}{\SubstWeak{1}}}}) \\
           & \quad \quad \quad \TyPi{\El{\sub{\var{0}}{\SubstWeak{1}}}}{\cU_j})) \\
   \Sigr{\Gamma}{0}{j} &= \top \\
-  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}}, \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ defines upon } \denotesT{\sigma} 
+  \mathscr{P}(\lsig') &=  \top \quad \text{ when } \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{0}{j}}  \\
+  \mathscr{P}(\lsig') &=  
+    \Sigma~(\fst{\lsig'})~(\El{\app{"snd"^3~\lsig'}[(\SubstWeak{1}, \app{\fst{("snd"^2~{\lsig'})}})]}) \\ & \text{ when } \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{n+1}{j}} \\
+\end{align*}
+
+% We make $({\Sigr{\Gamma}{n}{}})$ and
+
+% $\denotesT{\TyTkg {-}}$ mutually recursive,
+% defined by induction on $n$ and the signature length.
+% \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
+% &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\
+
+With the above two helpers, we can define denotation $\denotesT{}$.
+
+\begin{align*}
+  \denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}} &\in \{t : \goodTerm{\Gm}{t}{\Sigr{\Gm}{n}{j}} \} \\
+  {\goodType{\Gamma}{\TyTkg{\lsig}}{}} &= \mathscr{P}(\denotesT{\lsig}) \\
+  \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ defines upon } \denotesT{\lsig} 
   \text{ and inductively on the signature length} \\
-  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} = \top \quad
-      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\
-  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= 
-    \Sigma~\denotesT{\TyTkg {\lsigproj{1}{\sigma}}}~(\El{\app{"snd"^3~\denotesT{\sigma}}[(\SubstWeak{1}, \app{\fst{("snd"^2~{\denotesT{\sigma}})}})]}) \\
   \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &=
@@
   &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
+  \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &= \top \quad
+  \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\
   \denotesT{\goodSigU{j}{\Gamma}{\LSigEmp}{0}} &= \unit \\

### 2023-03-18 cc-condensed.tex
Fix typo.
Change Back the Rule Name.
a comment
Update cc-condense.

@@
 
-\noindent
+\noindent\YZ{These paragraphs here need to be updated}
 Consistency says that the type $\bot$ is not inhabited.
@@
 every proposition is trivially provable.
-
-\cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
-the well-formedness rules, which, in particular, interprets $\bot$ in
-\TT as an uninhabited type in the meta\-language.
-The construction of this interpretation follows the standard model of
-\citet{altkap2016}, extending it to handle linkages.
-A linkage~$\LkgAdd{\lkg}{t}$ can be interpreted into a non-dependent pair where the
-second component is a dependent function (modeling late binding),
-%\EDJ{Suggesting changing to ``can be interpreted into non-dependent pair with dependent function for late-binding''}
-as rule \ruleref{l/add} indicates.
-%
-We refer the reader to a supplemental file in Agda syntax for the complete proof.
 
@@
 
-We prove \cref{thm:canonicity} by constructing a logical-relations model
+We prove \cref{thm:consistency} and \cref{thm:canonicity} by constructing a logical-relations model
 for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
-In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
+
+In particular, (1) bottom type is interpreted as an empty set in the model, making sure no syntax derivation of bottom is possible and (2) a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
 into a logical predicate on closed terms: the predicate includes all
-closed, ``reducible'' terms of type~$T$.
+closed, ``reducible'' terms of type~$T$, deriving boolean canonicity.
 %\EDJ{I will add a quote around reducible here.
@@
 But this ``canonicity'' is not friendly to the audience not familar with declartive style.}
-The proof is available as a supplemental file in Agda syntax.
+% The proof is available as a supplemental file in Agda syntax.
+The proof is available in the appendix.

### 2023-03-18 cc-condensed.tex
edit

@@
 
-\noindent\YZ{These paragraphs here need to be updated}
+\noindent
 Consistency says that the type $\bot$ is not inhabited.
@@
 
-We prove \cref{thm:consistency} and \cref{thm:canonicity} by constructing a logical-relations model
-for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
-
-In particular, (1) bottom type is interpreted as an empty set in the model, making sure no syntax derivation of bottom is possible and (2) a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
-into a logical predicate on closed terms: the predicate includes all
-closed, ``reducible'' terms of type~$T$, deriving boolean canonicity.
+We prove \cref{thm:consistency,thm:canonicity} by constructing a logical-relations model
+for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,kaposi2019gluing,sterling2019algebraic}.
+%
+The model interprets the bottom type~$\bot$ as an empty set, from which
+\cref{thm:consistency} follows.
+A closed, well-formed type $\goodType{\cdot}{T}{}$ is interpreted as
+a logical predicate on closed terms: the predicate includes all
+closed, ``reducible'' terms of type~$T$. So \cref{thm:canonicity} follows.
 %\EDJ{I will add a quote around reducible here.
@@
 % The proof is available as a supplemental file in Agda syntax.
-The proof is available in the appendix.
+The construction of the logical-relations model is available in
+\ifreport
+\cref{sec:canonicity-model}.
+\else
+an appendix.
+\fi

### 2023-03-18 canonicitymodel.tex
Fix canonicity model, reply comments.

@@
   & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
-  \Glued{\mathit{L}}("nil") &= ( \TyLkg{\LSigEmp} , \denotesC{\top} ) \\
-  \Glued{\mathit{L}}((A,s,T) :: tl) &= {\Glued{\mathit{L}}(tl)}(\gm) \times^\bullet \GluedPi{A}{T} \quad \text{($\times^\bullet$ is a special case for $\Glued{\Sigma}$)}  \\
+  \mathit{L}("nil") &=  \denotesC{\top}  \\
+  \mathit{L}((A,s,T) :: tl)(\gm)(\gm')(t) &= {\mathit{L}(tl)}(\gm)(\gm')(\lkgproj{1}{t}) \times \Gmodel{\GluedPi{A}{T}}(\gm)(\gm')(\lam{\lkgproj{2}{t}})   \\
   \Glued{\mathit{P}}("nil") &= ( \TyTkg{\LSigEmp} , \denotesC{\top} )\\
   \Glued{\mathit{P}}((A,s,T) :: tl) &= \Glued{\Sigma}(\Glued{\mathit{P}}(tl), \Gsub{T}{\GSubstExt{\Glued{\pi_1}}{s}}) \quad \text{(doing substitution on $T$)} \\
-  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \Gmodel{\Glued{\mathit{L}}(\denotesC{\lsig})} \quad \text{ discard syntax info} \\
-  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \Gmodel{\Glued{\mathit{P}}(\denotesC{\lsig})} \\
+  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= {\mathit{L}(\denotesC{\lsig})} \\
+  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \Gmodel{\Glued{\mathit{P}}(\denotesC{\lsig})} \quad \text{ discard syntax info}  \\
   % 
@@
   \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
-  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}} &= \Gmodel{\Gpair{\denotesCC{\lkg}}{\Glam{\denotesCC{t}}}} \\
+  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm)(\gm') &= (\denotesC{\lkg}(\gm)(\gm') , \Gmodel{\Glam{\denotesCC{t}}}(\gm)(\gm')) \\
   \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
@@
   \\ &\text{ when} \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\
-  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gfst{\denotesCC{\lkg}}} \\
-  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}} &= \Gmodel{\Gapp{\Gsnd{\denotesCC{\lkg}}}} \\
+  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyLkg{\lsig}}}(\gm)(\gm') &= \denotesC{\ell}(\gm)(\gm')[0] \quad \text{ take the first element in the tuple} \\
+  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}}(\gm_+)(\gm_+') &= ((\denotesC{\ell}(\pi_1{\gm_+})(\gm_+'[0]))[1])(\pi_2{\gm_+})(\gm_+'[1]) \\
   \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\

### 2023-03-18 cc-condensed.tex
question
adding back the bit about how linkages
Fix canonicity model, reply comments.
edits

@@
-\noindentparagraph{Consistency.}
+\noindentparagraph{Consistency and canonicity.}
 One of the most fundamental properties of a dependent type theory is consistency.
@@
 every proposition is trivially provable.
-
-\noindentparagraph{Canonicity.}
 
@@
 %
+The model interprets a linkage~$\LkgAdd{\lkg}{t}$ as a pair where the
+second component is a function (modeling late binding),
+as rule \ruleref{l/add} indicates.
+%
 The model interprets the bottom type~$\bot$ as an empty set, from which
@@
 a logical predicate on closed terms: the predicate includes all
-closed, ``reducible'' terms of type~$T$. So \cref{thm:canonicity} follows.
+closed, ``reducible'' terms of type~$T$.
+\cref{thm:canonicity} follows from this interpretation.
 %\EDJ{I will add a quote around reducible here.
@@
 an appendix.
-\fi
+\fi\YZ{and as fake Agda sources?}

### 2023-03-18 metatheory-condensed.tex
edits

@@
 named variables and meta-level substitution.
-For clarity of proof details, the technical appendix supplementing
-\cref{sec:metatheory2} instead uses de Bruijn indices and explicit substitution.
+For clarity of proof details, a technical appendix supplementing
+\cref{sec:metatheory2} instead uses \citeauthor{debruijn1964} indices and explicit substitution \cite{abadi1989subst}.
 %
@@
 \ifreport \cref{sec:complete-fmltt}.
-\else a supplemental appendix.
+\else a technical appendix.
 \fi
-using de Bruijn indices and explicit substitution.
-An effort to mechanize the formalization is underway, by following  \citet{altkap2016} 
-to use quotient inductive-inductive types as 
-the logical framework to state the syntax and using Agda's
-\lsti{REWRITE} pragma to replace the limited uses of equality reflection appeared in 
-an extensional meta\-logic.
+This definitive version uses de Bruijn indices and explicit substitution.
+An effort to mechanize the formalization is underway. It follows \citet{altkap2016}
+in using quotient inductive-inductive types (QIITs) as 
+the logical framework to state intrinsically typed syntax (and also
+in using Agda's \lsti{REWRITE} pragma to replace the limited uses of equality
+reflection, as Agda does not support QIITs natively).
+%in the extensional meta\-logic.
 

### 2023-03-19 coqimpl-old.tex
relocate old writings

--- 2023-03-17 coqimpl.tex
@@
-We implement our language design as a Coq plugin.
-%
-It works by translating programs in \Lang syntax into
-programs that can be checked and evaluated by Coq.
-The translation is compatible with interactive theorem proving (\ref{chg:tooling}), in that
-a family is translated piece by piece, allowing each field to be defined
-and checked separately.
-The translation is modular and efficient\EDJ{I am very hesitant to say efficient :( }, in that
-code compiled for fields of a base family
-can be shared with derived families without having to be rechecked.
-
-\noindentparagraph{Explicit self parameterization.}
-
-The spirit of the translation is to take ``family polymorphism''
-literally: every field is translated into a Coq definition that is
-polymorphic to (i.e., universally quantified over) a representation of
-its enclosing family.
-While this universal quantification has been implicit with the \Lang
-syntax, it has to be made explicit in the translated Coq code.
-
-\cref{fig:stlc-compiled,fig:stlcfix-compiled} illustrate the translation of the
-\lsti{STLC} and \lsti{STLCFix} families from \cref{fig:stlc-mechanized}.
-Fields of a family are translated into parameterized Coq modules
-(or parameterized module types).
-
-As an example,
-consider field \lsti{env} in family \lsti{STLC}.
-It is translated into a top-level module named \lsti{STLC\_env}.
-This module has a parameter called \lsti{self} representing the enclosing
-family: fields of the current family in the context of \lsti{env} can be
-referenced through \lsti{self}.
-In particular, \lsti{env} is defined as \lsti{id}\,\lsti{->}\,\lsti{option ty},
-where \lsti{ty} is a late-bound reference to the \lsti{ty} field of the
-enclosing family.
-Hence, this reference to \lsti{ty} is translated to \lsti{self.ty},
-which is manifestly polymorphic to the enclosing family.
-This translation of the \lsti{env} field can be shared
-with a derived family even if it extends \lsti{ty} (e.g., \lsti{STLCProd})---%
-no recompilation is needed because \lsti{self.ty} is not tied to any
-concrete definition of \lsti{ty}.
-
-The type of \lsti{STLC\_env}'s \lsti{self} parameter is \lsti{STLC\_env\_Ctx},
-a module type constructed from \lsti{STLC\_ty}
-(i.e., the translation of the field before \lsti{env})
-and its context \lsti{STLC\_ty\_Ctx}.
-In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is
-constructed from \lsti{STLC\_subst},
-the translation of the field before \lsti{ty},
-and its context \lsti{STLC\_subst\_Ctx}.
-Thus, the \lsti{self} parameter can be used to reference those and only
-those fields in the current field's typing context, which echoes the
-discussion in \cref{sec:sound}.
-
-\input{stlc-compiled.tex}
-\input{stlcfix-compiled.tex}
-
-\noindentparagraph{Translating extensible inductive types.}
-
-An \lsti{FInductive} definition is translated to a parameterized module type.
-Consider the inductive type \lsti{tm}.
-In \cref{fig:stlc-compiled}, it is translated to a top-level module type
-\lsti{STLC\_tm} that declares a \lsti{tm} type,
-four functions standing for the constructors,
-%\EDJ{I am not sure if it is better we emphasize in the following way :`` that declares a tm type, its constructors (e.g.
-tmabs) as a normal function,..
-'' i.e.
-emphasize it is a normal function in the module type},
-a partial recursor (\lsti{tm_prect_STLC}),
-and the computational behaviors of the partial recursor
-(e.g., \lsti{tm_abs_eq_STLC}). %\cref{sec:prec}.
-%\EDJ{I propose to change this sentence to ``and the prop eq about computational behaviour on the partial recursor''.
-And thus in the other parts of the paper, we can use the word ``computational behaviour'' instead of prop eq.
-Because just saying prop eq seems a bit of lack of context/details.}
-
-Importantly,
-\lsti{STLC\_tm} merely declares the existence of these names and their types;
-it does not specify their definitions.
-Having these names and their types available through the context parameters (\lsti{self})
-suffices for the translations of the subsequent fields to be type-checked by Coq.
-Leaving the definitions unspecified enables \lsti{STLC} and \lsti{STLCFix}
-to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and upon \lsti{End}~\lsti{STLCFix}.
-
-The command \lsti{FInductive tm : Set += tm_fix : }\dadada~in family
-\lsti{STLCFix} is again translated to a module type \lsti{STLCFix\_tm} (\cref{fig:stlcfix-compiled}).
-It includes all the names declared by \lsti{STLC\_tm}
-via command \lsti!Include STLC\_tm(self)!,
-and additionally declares \lsti{tm_fix}, a partial recursor, and related equalities.
-
-\noindentparagraph{Translating recursion and induction.}
-
-An \lsti{FRecursion} definition is translated in two parts:
-first a module containing the definitions of all the case handlers,
-and then a module type declaring the existence of the recursive function as well as
-its computational behaviors.
-
-Consider the translation of \lsti{subst} in family \lsti{STLC}.
-First, a module named \lsti{STLC\_subst\_Cases} is generated on the fly.
-%every time the programmer completes a \lsti{Case}, the case handler (e.g., \lsti{Def subst\_tm_unit})
-%is type-checked and added to the module.
-%\rD{This is also confusing-- why is the programmer involved in the translation?
-Is it that modules are being generated in the background, while commands are being processed?}\EDJreply{changed to mention "type-checked" explicitly.
-please review}
-Importantly, programming remains interactive, as the programmer need not wait until the
-entire \lsti{FRecursion} definition is completed to have a \lsti{Case} command
-type-checked.
-
-Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
-%
-As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
-% \EDJ{If we don't allow further binding case handler(mentioned in your last reply), how do the user further bind the subst?
-I mean in the surface syntax}\YZreply{The position is that individual case handlers cannot be further bound, but the recursive function itself can be further bound by acquiring new case handlers.}
-so its definition is not exposed to the fields that come after it.
-Accordingly, the translation \lsti{STLC\_subst} merely declares the types of \lsti{subst}
-and the equalities about its computational behaviors, leaving \lsti{subst} undefined
-and the equalities unproven.
-%
-The equalities are stated in terms of the case handlers,
-whose definitions \emph{are} available through the \lsti{self} parameter.
-So Coq can simplify, for example, the type of \lsti{subst_tm_unit_eq} to
-\lsti[basicstyle=\fontsize{8.5}{9},basewidth=.95ex]{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
-These equalities about the computational behaviors of \lsti{subst}
-will be included and available for use in the translations of the
-subsequent fields through their \lsti{self} parameters.
-
-Importantly, code generated for the case handlers is shared with derived families.
-In \cref{fig:stlcfix-compiled}, module \lsti{STLCFix\_subst\_Cases}
-reuses---without rechecking---all the case handlers in
-\lsti{STLC\_subst\_Cases} via command \lsti{Include STLC\_subst\_Cases(self)}.
-
-The translation of \lsti{FInduction} is similar, except that there is no need to
-register computational behaviors, as \lsti{FInduction} proofs are considered opaque.
-
-\noindentparagraph{Translation of further-bindables vs.\ non-further-bindables.}
-
-In family \lsti{STLC},
-field \lsti{env} and the case handlers for \lsti{subst}
-are not further-bindable by derived families.
-In contrast, \lsti{tm}, \lsti{subst}, and the related equalities
-can be further bound.
-The distinction is reflected in the translations.
-The further-bindable fields are translated to module types that export only types of the fields.
-The non-further-bindable fields are translated to modules that export definitional equalities
-on the fields.
-Opaque fields in \Lang can be further bound (\cref{sec:override});
-they are translated to Coq modules that export opaque fields.
-
-%\footnotetext{%
-%Axioms in Coq module types are more like parameters than real axioms.
-%They become part of the signature described by the module type.
-%}
-
-\noindentparagraph{Eliminating \lsti{self} by aggregation.}
-
-Upon the conclusion of a family definition,
-a representation of the family is created.
-For example, module \lsti{STLC} in \cref{fig:stlc-compiled} is generated
-upon \lsti{End}~\lsti{STLC}.
-%
-This module can be viewed as the ``fixed point'' of the
-\lsti{self}-parameterized translations.
-The ``fixed point'' is taken step by step, by adding the translation of
-each field to this module in the same order as they appear in the family
-definition.
-
-For the non-further-bindables, the translated modules are directly included
-(e.g., \lsti{Include STLC\_env} and \lsti{Include STLC\_subst\_Cases} in \cref{fig:stlc-compiled}).
-%
-The instantiation of \lsti{self}s for these modules is implicit, thanks to a Coq nicety:
-when including a higher-order module, Coq automatically instantiates its
-parameter with the current interactive module environment.
-For instance, command \lsti{Include STLC\_subst\_Cases} is successfully executed,
-because Coq automatically instantiates the \lsti{self} parameter using the current
-module environment, which by construction contains all the fields required by \lsti{STLC\_subst\_Cases\_Ctx}.
-
-For the further-bindables, \lsti{Axiom}s declared in the module types must be instantiated.
-%
-\begin{itemize}
-  [align=left,itemsep=2pt,labelsep=*,leftmargin=*]
-
-\item
-In \cref{fig:stlc-compiled},
-an inductive type \lsti{tm} is generated, instantiating the axiomatized
-\lsti{tm} type and its constructors.
-%
-The partial recursor \lsti{tm_prect_STLC} is defined
-%to \lsti{\lambda}\,\lsti{P, tm_rect (\lambda}\,\lsti{t, option (P t))},
-with the help of \lsti{tm_rect}, the recursor Coq generates for \lsti{tm}.
-The computational behaviors of \lsti{tm_prect_STLC} are immediate, by \lsti{reflexivity}.
-
-\item
-Similarly, \lsti{subst} is instantiated by applying the recursor \lsti{tm_rect}
-to the (already included) case handlers. %(\lsti{subst\_tm_unit}, etc.).
-The computational behaviors of \lsti{subst} are then immediate, by \lsti{reflexivity}.
-
-\end{itemize}
-
-Module \lsti{STLCFix} in \cref{fig:stlcfix-compiled} is emitted upon
-\lsti{End}~\lsti{STLCFix}, in the same way as described above for \lsti{STLC}.
-The translation makes sharing evident.
-In particular, case handlers compiled for \lsti{STLC}
-are reused to instantiate \lsti{subst}, \lsti{substlem}, and alike.
-\lsti{STLC\_env} and \lsti{STLC\_typesafe}
-are also reused in the construction of module \lsti{STLCFix}.
-%
-One may argue that since the first four constructors of \lsti{tm} are
-repeated in \lsti{STLCFix}, the translation does not satisfy the
-modular compilation requirement.
-We could address this concern by using wrapper types,
-but we consider restating constructors a small price to pay in return
-for the clarity and concision of implementation.
-We emphasize that compiled case handlers, such as \lsti!subst\_tm_abs!,
-are entirely reusable without rechecking, even with restated constructors.
-%
-Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)
-can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a Coq module),
-as the last line of \cref{fig:stlcfix-compiled} shows.
-
-\noindentparagraph{Trusted base.}
-
-Rather than modifying the Coq kernel to extend its core theory, a translation to
-Coq conveniently reduces the trusted base of any development using \Lang to Coq.
-%and improves compatibility with different versions of Coq.
-In particular, once a family is closed, \lsti{Print Assumptions} can be used
-to verify that there are no lingering \lsti{Axiom}s introduced by the translation.
-Ramifications of possible bugs in the \Lang implementation are limited to
-the usability of the plugin.
-%Finally, we emphasize that our compilation can make sure the trusted code base is restricted to the Coq's kernel, as all the feature supported by our plugin has been compiled into Coq's surface syntax/Vernacular commands.
-
 \ifShowOldWriting
@@
 prec : forall (R : Set), ..,
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,@\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint.
-Let me fix other parts first}\YZreply{Is this constraint Detectable Partiality?}@
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,
       D -> option R
@@
   \begin{minted}[escapeinside=@@]{Coq}
-    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
-If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
-I mentioned left inverse 'linv' is the generalization}@
+    forall c : C, linv (inj c) = some c
   \end{minted}
@@
 C) → option C}, generally not provable, unless we have
-\mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.\YZ{Don't you really mean
-"unless we have the Detectable Partiality assumption instead"?}
+\mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
 This \ref{prop:detectable-partiality} is exactly transforming former to
@@
 complete extensional characterization of all the extension of a given (non-indexed)
-inductive type.\YZ{What about indexed inductive types?}\EDJreply{Done.
-I also have a formalization of this proof now, for indexed type and dependent eliminator.
-Though the formalization is about single constructor.}
+inductive type.
 
@@
 the family being defined so that they can also be used when defining
-the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}\EDJreply{Not sure.
-This auto-insertion will practically look like hijack the original Coq command.
-I am not sure this kind of hijacking can happen.}
+the following fields of the family.
 
@@
 % having boilerplate code, \textbf{semantically}.
-\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}\EDJreply{Done}
+\begin{figure}[!htb]
 \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
@@
 \fi
-
-\section{Limitations}
-\label{sec:limitations}
-
-The \Lang implementation currently does not support Coq's full
-facility for inductive types and fixpoints;
-in particular, mutually inductive types and mutual fixpoints are not supported.
-The plugin supports indexed inductive types but not parameterized ones, though
-the latter can be encoded via the former.
-Partial recursors are automatically generated only for non-indexed inductive types.
-We believe that these limitations do not pose conceptual challenges and can be
-addressed with more engineering effort of the same kind that went into \Lang.
-
-%We list some of the current limitation of our implementation. We believe the following limitation can be rememdied by more engineering effort:
-%\begin{itemize}
-%  \item We don't support the full power of inductive facility in Coq for extensible inductive type---concretely, we don't support mutual inductive type, syntax extensions for inductive type and parameters in the inductive type (we only support indexed inductive type right now)\YZ{Can parameterized inductive types always be encoded as indexed inductive types?}\EDJreply{Yes.
-but practically the dependent elimination principle is different when using encoding.}
-%  \item We don't support mutual fixpoint definition
-%  \item We don't support incorporating "Hint" Vernacular commands, and thus the users cannot specify the database hints
-%  \item Unicode (i.e. "Require Import Coq.Unicode.Utf8.") is not compatible with our plugin 
-%  \item We only support automatic generation of partial recursor and the computation information for non-indexed inductive type 
-%\end{itemize}
-
-However, there are also limitations we believe require more research effort or more thought from a language-design perspective, including but not limited to: 
-\begin{itemize}
-  \item We didn't considered the compatibility of universe polymorphism mechanism of Coq with our plugin except we know our plugin elaborate families into modules.
-Thus our implementation is supposed to be as compatible with universe polymorphism as Coq's Module.\YZ{Is this really a limitation at all?
-Do we have to mention it?}
-  \item ... \YZ{mention restricting induction to the top level}
-  \item We don't support normalization of the open term using extensible inductive type.
-When applying recursor $Rec$ to open terms $t$ using extensible inductive type, the computation cannot be done by the Coq's kernel, thus the computation result of $Rec(t)$ is only propositionally equal (but not judgementally equal) to $Rec(t)$.
-It will complicate typechecking of the dependently typed terms.
-\end{itemize}

### 2023-03-19 coqimpl.tex
relocate old writings

@@
 
-\ifShowOldWriting
-
-\newpage
-
-In this section, we describe how we can implement the proposed Vernacular commands.
-
-After considering the pros and cons, we decide to implement a Coq Plugin where we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the orignal surface Vernacular command) on the fly, instead of modifying the code-base of Coq.
-
-Despite possible difficulties for future maintenance, this approach has
-various advantages: 1. it is the easiest and the most accessible way to
-prototype as it relieves us the necessity of familarity of Coq base,
-especially for the implementation of the module and functor; 2. we have
-a clear definition of trusted-base---the whole Coq; 3. it is easy to
-debug---we just need to check the translated commands; 4. it can be
-well-incorporated with the existent tools like VSCoq; 5. it is more
-accessible for the interested audience who can then easily adapt our
-plugin and give a try---otherwise they have to download and re-build the
-whole customized Coq; 6. it is more stable because the surface syntax of
-Coq should be stable across different versions; 7. still, this plugin
-can capture the key ingredients of implementing family polymorphism
-inside Coq and act as a reference for guiding an appropriate
-implementation of Family Polymorphism inside all sorts of proof
-assistants.\YZ{Does the Coq implementation perform type checking to
-prevent illegal uses fam poly?}\EDJreply{Yes. Because I define feature of fampoly as a direct translation into Coq feature.  The fampoly is like a shorthand into multiple Coq commands. \\
-Do you find translation from my current metatheory into MLTT (without linkage) helpful? \\
-Actually I am also not sure what this question is aiming towards. \\
-Are you complaining there is no formal specification of my fampoly feature in paper and thus there is no proof my Coq surface syntax is related to metatheory?
-If so, I thought our Coq plugin is only advertising how powerful Fampoly is to solve expression problem in mechanized proving.
-I can decouple the relationship between our plugin and our metatheory if you find the gap is really there...
-And I can only claim metatheory is about a partial result on incorporating fampoly with dependent type, Coq plugin is using two example to illustrate how mechanized proving can get benefit and our Coq plugin is only inspired by our metatheory (and not claim any relationship).
-Doing so would require rearrange the metatheory section before this section.
-\\
-But I don't think the gap is there at all because if you want all possible gaps to disappear, then the action of using debruijn is wrong because we need to prove the one using debruijn and the surface syntax that is not using debruijn are ``equivalent''.
-What's more, module is not sigma type because Module is not a term in Coq and thus there is no formalism of Module at all.
-Let alone I don't see the plausibility of mechanizing dependent type without debruijn.
-\\
-Please comment back so that I know what corresponding modification is needed on the paper. 
-}\YZreply{My original question was more directed towards the fact
-that currently type checking is deferred to the Coq base. Generally,
-when one uses a statically typed language as a compilation target, they
-also want to have type checking at the source level, (1) because there
-could be type errors at the source level that could not otherwise be
-caught at the target level and (2) because it allows better error messages.
-So it needs to be addressed why deferring type checking to Coq is OK.}\EDJreply{Good point.
-I think (1) is largely mediated by the fact the semantic of family is very close to that of the module.
-And the supporting feature is simple enough that there won't appear some problems like type error at source-level but not target level(2) is happening for sure but some basic error message is almost "transparent" i.e.
-the error-messages from Coq can be directly interpreted for the user.
-But my justification is solely empirical and as you know the empirical study for this paper is not really strong enough to support anything.}
-
-\textbf{Family compiled into module.} The main idea of our plugin is to support family by compiling
-family components into Coq modules, family ``types'' into module types, and the context of a given term
-into parameters of the module.
-
-Since by default every field is inheritable, for each defining field (informally) $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
-``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
-\cL \sigma.
-T$, so that field $t$ can be inherited to all the future family with strengthen context $\sigma^+$ with the help of weakening from $\sigma^+$ to $\sigma$.
-However, instead of using Coq's universal quantifier, we choose module and module parameters to achieve this wrapping---each
-family type in the context will be one module parameter
-with the compiled module type, and thus each field of the family will
-compile into a functor parametrized by its context. Doing so we can also
-get rapid feedback from type checking when the users are defining each
-field interactively with Coq.
-Note that, when compiling the context during the  compilation of a default (non-overridable)  field, all the default (non-overridable) former fields will expose their definition in the context while overridable fields will only expose their types.
-% Insert one pseudo-code example for explanation of the mechanism
-\begin{figure}[!htb]
-  \begin{minipage}[t]{0.30\linewidth}
-\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
-Family A.
-Field a : nat := 1.
-Field b : a = 1 := eq_refl.
-Overridable Field a' : nat 
-                     := 1. 
-Overridable pins {a'}
-  Field b' : a' = 1 
-          := eq_refl.
-EndFamily.
-\end{lstlisting}
-  \end{minipage}%\YZ{Can the self's be omitted?}\EDJreply{Yes.
-But in our current plugin implementation it is not omitted.
-I think altering example is fine.
-If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}\EDJreply{Omitting self_ is easy and not much engineering effort is required.
-But omitting self_ will lead to `A.a' instead of `a' actually.
-So if we really want `a' as you last time mentioned, then a resolution is required and more engineering effort is required.
-But I don't see any technical difficulties.
-I have removed all the self, please check.}
-  \begin{minipage}[t]{0.35\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Module a_4 (self_A: EmptySig).
-Definition a : nat := 1. End a_4.
-Module Type a_6.
-Include a_4. End a_5.
-Module b_7 (self_A: a_6).
-Definition b 
-  : self_A.a = 1
-  := eq_refl. End b_7.
-Module Type b_9.
-Include a_6. Include b_7. End b_9.
-Module a'_10 (self_A: b_9).
-Definition a' : nat := 1. End a_4.
-\end{minted}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\linewidth}
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Module Type a'_12.
-Include b_9. Include a'_10. End a'_12.
-Module b'_16 (self_A: a'_12).
-Definition b' : self_A.a' = 1 := eq_refl. End b'_16.
-Module Type b'_19.
-Include b_9. Parameter a': nat.
-Parameter b': a' = 1. End b'_19.
-Module test_20 (self_A: b'_19).
-Definition test : nat := 1. End test_20.
-
-Module A. (* Final Aggregation *)
-Include a_4. Include b_7. Include a'_10.
-Include b'_16. Include test_20. End A.
-\end{minted}
-  \end{minipage}
-  \caption{Example Code and Plugin Translation}\label{fig:plugin-example1}
-\end{figure}
-
-% Concrete details of compilation of a family term data structure
-Taking \cref{fig:plugin-example1} as an example: immediately after the user inputs the line "Field a : nat := ..", our plugin will
-translate this statement into the "Module a_4", "Module Type a_6"
-two components.
-{The "EmptySig" is just an empty module type.} Then,
-these modules and module types will be generated and type-checked by Coq
-immediately. The compiled "Module a_6" is then part of the context of
-"Field b" (as the module type of "self_A") for "b" to refer to, so that "b" knows "a"'s concrete definition. The
-internal representation of family "A" is the list of these
-compiled modules and module types. We can achieve incremental
-checking as in \ref{chg:software-engineering}. This makes all kinds of inheritance plausible---for example, if we
-want to override "Field a", we simply replace that compiled module with
-the new module inside the list. 
-% This compiled module type serves as an abstraction that enables overriding.
-
-After closing the top family, the user can access family "A" like it is a module. This is achieved by a complete compilation from a family to a module,
-via aggregating the list of the modules in the internal representation:
-the plugin will repeatedly using Coq's "Include" command as the "Module
-A" as \cref{fig:plugin-example1} illustrates.
-References to family "A" outside "A" will be compiled to references to
-this Coq module "A".
-
-Note that, in our example, we include two
-functors without arguments---this is due to a peculiarity of Coq's
-"Include": uninstantiated parameters will be automatically
-instantiated by the appropriate fields in the surrounding defining
-module.
-For example, "b_7" requires a module of type "a_6". Inside module "A",
-when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat := 1}
-inside the current surrounding "A". Then when we
-\mintinline{Coq}{Include b_7} without specifying the module parameter,
-Coq will try to find fields in the surrounding module (i.e., module
-"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
-the \mintinline{Coq}{a : nat} that is included earlier.
-The resulting compiled module is just like mundane Coq's module and 
-thus maintains all the computational mechanism from Coq, fulfilling
-requirements in \ref{chg:software-engineering}.\YZ{
-  Do I understand correctly that an alternative to individually
-  compiling family members to modules is to compile a family as a whole,
-  but this alternative would not satisfy the "incremental type checking"
-  or "instant feedback" requirement?
-}\YZreply{
-  Looks like individually compiling family members addresses not only
-  this requirement, but also the "self vs. consistency" challenge.
-}\EDJreply{Yes. Indivially compiling family member can achieve both requirements}
-
-\textbf{Implementing "Overridable" and Overriding Chain.}
-We implement "Overridable" by simply annotating the corresponding field type with a mark and its dependency.
-Then when compute the context during the compilation of the term for that field, we make sure only all the dependee and the non-overridable field will expose its definition in the context.
-
-Most parts of the implementation is quite similar to the default fields, except that when inheriting an overridable field, we need to make sure all its dependee is inherited earlier.
-For example, in \cref{fig:plugin-example1}, the module "b'_16" will compile "b'" in the context of "a'_12" where the concrete definition of "a'" is exposed since "b'" depends on "a'"; while the module "test_20" is compiled in the context "b'_19" only knowing the type of "a'" and "b'".
-
-When aggregating the overridable fields into a module, since everything has concrete definition now, there won't be any problem---we just use the same way that aggregates the default fields.
-
-% explaining inheritance
-\textbf{Implementing Inheritance.}
-To simplify the implementation, we can consider all families as
-inheritance and the standalone families are extending the empty family.
-Thus, we need to deal with only inheritance during interactive theorem proving.
-
-Simiarly, we
-use a list to encode the three kinds of inheritance data in the
-implementation. The inheritance data will include the
-information of (1) the ``type'' of the parent family, (2) a list of
-operation indicating how each field of the parent will be dealt with
-(either inherited or overriden), (3) the operations indicating newly
-extended fields, and (4) the ``type'' of the children family.
-The surface syntax for the three kinds of inheritance
-are demonstrated in the exemplar \cref{fig:plugin-example3}. The
-implementation of overriding, inheritance, and extension is achieved by
-simply manipulating the internal representation of a family
-correspondingly---we can swap the module in the list for overriding,
-retain the module in the list for inheritance, and add new module into
-the list for extension. The compilation from a family (internal
-representation) to a module still acts the same.
-
-Though inheritance is solely a first-order data rather
-than a ``mapping'', it is still possible to \text{mixin} two inheritance
-data---by carefully ``mix'' the internal list, we can compose the
-inheritance. However, a good definition of mixin is still under
-investigation. We will show one practical example of \textit{mixin}
-later.\YZ{How about making mixin composition one of your Req's? Also
-does any example in the paper illustrate how to do this?}\EDJreply{mixin
-is not really formalized. There is no example illustrate it. The only
-example is the STLC example at the very end and I just kind of
-``mention'' it. I think here I just want to emphasize, inheritance as
-first order data is still possible to support composition.}
-
-Our implementation also makes sure the first attempt of defining "c"
-failed successfully in \cref{fig:plugin-example3}---proving "c"
-would require a concrete definition of "a", and our plugin rejects
-\mintinline{Coq}{c : a = 1 := eq_refl} in a context where we
-only know \mintinline{Coq}{a : nat}.
-
-% \textbf{Implementing "Final" and "Sealed Family".}
-% We implement "Final" by
-% exposing the whole definition directly into the compiled module type.
-% Doing so will prohibit overriding. Say, "bop", declared "Final" in
-% \cref{fig:plugin-example3}, is not overridable.
-% We require the users to manual decorate field as `Final' so that the
-% plugin will proceed with this special treatment.
-% %\YZ{This para is confusing: which kind of equality is used, propositional or judgmental?
-Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory.
-Now everything is about judgemental so I don't have to emphasize judgemental.
-I wanted to make the example small but I guess I can add one `Final' statement into our example.
-Don't resolve it until I add it into Figure 2.}
-
-% % \textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.}
-
-% The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type.
-The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
-% For example, the compiled module type of
-% "MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}".
-Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
-}
-
-% The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type.
-The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
-% For example, the compiled module type of
-% "MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}".
-Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin.
-}
-
-% Basically, {a : int -> int; property : a is monotonic}, in this case
-%   we may want to override "a" using different computation, but we still wnat
-%   this refinement/property
-
-\textbf{``Extensible'' Inductive Type and Defining Recursor/Eliminator.}
-We support extensible inductive types in Coq mentioned in
-\ref{chg:extensible-inductive-type}.
-% \footnote{Because to give a satisfactory
-% formal definition and semantic for extensible inductive type is not
-% something this paper aims at, and obviously requires more theoretical
-% research effort}
-%\YZ{I do consider that your plugin supports extensible inductive types.
-It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it.
-It is more like I don't think this is really the final form of extensible inductive type that everyone is happy about}
-
-% \begin{figure}[!htb]\YZ{Change this one into STLC example, to see the layout}
-%   \begin{minipage}[t]{0.32\linewidth}
-% \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-% Family B.
-%   FInductive b : Set 
-%     := tt : b | ff : b.
-%   Family neg_handler. 
-%     Field tt : self_B.b 
-%       := self_B.ff.
-%     Field ff : self_B.b 
-%       := self_B.tt.
-%   EndFamily.
-%   FRecursor neg 
-%     about self_B.b 
-%     motive (fun _ => self_B.b)
-%     using self_B.neg_handler
-%     by _rec.
-%   Field example := 
-%     self_B.neg self_B.tt. 
-% EndFamily.
-% Family B2 extends B.
-%   Extend FInductive b : Set 
-%     := uu : b.
-%   Extend Family neg_handler.
-%     Field uu : self_B2.b 
-%       := self_B2.uu.
-%   EndFamily. 
-% EndFamily.
-% \end{lstlisting}
-%   \end{minipage}
-%   \begin{minipage}[t]{0.65\linewidth}
-% \begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
-%   (* Abstraction for Inductive Type *)
-% Module Type b_3 (self_B: EmptySig_4).
-% Parameter (b : Set). Parameter (tt ff : b). End b_3.      
-% Module Type b_6 := b_3 EmptyMod.
-%   (* Recursor to Set for b *)
-% Module b_rec_12 (self_B: b_6).
-% Definition __recursor_type_b_rec :=
-%   forall P : self_B.b -> Set,
-%   P self_B.tt -> P self_B.ff 
-%   -> forall __i : self_B.b, P __i. End b_rec_12. 
-%   (* Field B.neg_handler.ff *)
-% Module ff_24 (self_B: b_6) 
-%              (self_neg_handler: tt_23 self_B).
-% Definition ff : self_B.b := self_B.tt. End ff_24.
-%   (* Intermediate Module solely for type checking *)
-% Module do_tc (self_B: neg_handler_26).
-% Include b_rec_12 self_B.
-% Parameter (recursor_for_type_checking : __recursor_type_b_rec).
-% Definition term_for_type_checking :=
-%   recursor_for_type_checking (fun _ : self_B.b => self_B.b)
-%     self_B.neg_handler.tt
-%     self_B.neg_handler.ff.  
-% End do_tc.
-% \end{minted}
-%   \end{minipage}
-% \caption{Example Code for Inductive Type}\label{fig:plugin-example2}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \begin{minipage}[t]{0.32\linewidth}
-% \begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-% Family STLC.
-%   FInductive tm : Set := tm_var : ... 
-%   FScheme tm_prec PRecT about self__STLC.tm.
-%   FScheme Comp about self__STLC.tm_prec.
-%   (* Generate Partial Recursor 
-%   and the computational axiom *)
-%   Family subst_internal. 
-%   Final Field tm_var : ...
-% := fun s x t => 
-%     if (eqb x s) 
-%     then t else (tm_var s).
-% (* ... other three cases *)
-%   EndFamily.
-%   FRecursor subst 
-%     about tm 
-%     motive (fun _ => id → tm → tm)
-%     using subst_internal
-%     by _rec.
-% (* Field test := (subst (tm_var 0) 0 (tm_var 0)).@\YZ{Why is "test" commented out?}\EDJreply{Because I want to show I can do it but I don't really want to d -- if I really do it the compilation result shown later will need to have this information.}@ *)
-% (* ... *)
-% EndFamily.
-% Family STLC_bool extends STLC.
-% (* ... *)
-% Extend FInductive tm : Set :=
-%   | tm_true : tm ...
-% Extend Family subst_internal.
-% Final Field tm_true : ...
-%   := fun x t => tm_true.
-% (* ... *)
-% EndFamily. 
-% Inherits subst.
-% (* ... *)
-% EndFamily.
-% \end{lstlisting}
-%   \end{minipage}
-%   \begin{minipage}[t]{0.65\linewidth}
-% \begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
-% Module Type STLC_Interface.
-%   Parameter (tm : Set).
-%   Parameter (tm_var : id → tm).
-%   Parameter (tm_abs : id → tm → tm).
-%   (* ... and all the fields defined before
-%   and the abstraction for inductive type ...*)
-% End STLC_Interface.
-% (* Collect eliminator to Set for tm *)
-% Module tm_rec_type (STLC: STLC_Interface).
-%   Definition __recursor_type_tm_rec :=
-%     forall P : STLC.tm -> Set,
-%     (forall n : id, P (STLC.tm_var n)) ->
-%     (forall (n : id) (i : STLC.tm),
-%     P i -> P (STLC.tm_abs n i)) ->
-%     ... -> forall i : STLC.tm, P i.
-% End tm_rec_type.
-% (* Compiled Field STLC.subst_internal.tm_var *)
-% Module tm_var_handle (self__STLC: STLC_Interface)
-%   (self__subst_internal: EmptySig_84).
-% Definition tm_var :
-%   forall (s : ident) (x : ident) (t : self__STLC.tm), self__STLC.tm :=
-%   fun s x t => if eqb x s then t else self__STLC.tm_var s.
-% End tm_var_handle. (* ... and more *)
-% (* Assembling STLC.subst_internal when compilation *)
-% Module subst_internal_410 (STLC: STLC_Interface). 
-%   Module subst_internal.
-%   Include tm_var_handle STLC. Include tm_abs_handle STLC.
-%   Include tm_app_handle STLC. Include tm_unit_handle STLC.
-%   End subst_internal.
-% End subst_internal_410.
-% Module do_tc (STLC: STLC_Interface).
-% (* Intermediate Module solely for type checking *)
-%   Include tm_rec_type STLC.
-%   Parameter (recursor_for_type_checking : __recursor_type_tm_rec).
-%   Definition term_for_type_checking :=
-%     recursor_for_type_checking (fun _  ⇒ id → STLC.tm → STLC.tm)
-%       subst_internal_410.tm_var ... .
-% End do_tc.
-% (* Ctx that expose concrete inductive type *)
-% Module Type ctx_subst_113.
-% Inductive __internal_tm : Set := ... 
-% Include subst_internal_410. (* ... and more inclusion *)
-% End ctx_subst_113.
-% (* Compilation of recursor, not reusable
-%     when inductive type extended due to the ctx *)
-% Module subst_128 (self__STLC: ctx_subst_113).
-% Include __motive_of_subst_107 self__STLC.
-% Definition subst :=
-%   self__STLC.__internal_tm_rec __motiveTsubst
-% 	self__STLC.subst_internal.tm_var ... .
-% End subst_128.
-% (* Assembling the whole STLC *)
-% Module STLC.
-% Include tm_5. Include tm_prec_72. Include tm_prectm_81.
-% Include subst_internal_410. Include subst_128.
-% End v_129.
-% (* Assembling STLC_bool.subst_internal via reusing 
-%     most of the former compilation results *)
-% Module subst_internal_248 (STLC_bool: STLC_Interface).
-% Include tm_true_197 STLC_bool. Include ...
-% (* The Reused Compilation result *)
-% Include tm_var_handle STLC_bool. Include ...
-% End v_248.
-% @\YZ{It'd be useful to (1) show how the parts of the STLC family are assembled and recursor inserted, and (2) show the code compiled for STLC\_bool and how code from STLC is shared with STLC\_bool.}@
-% \end{minted}
-%   \end{minipage}
-% \caption{Exemplar STLC Code, especially about Inductive Type}\label{fig:plugin-example2}
-% \end{figure}
-\include{ext-ind-type-compilation-example.tex}
-
-% We start with the programming interface.
-We use "FInductive" to define an extensible inductive type and in the children family we use "Extend Finductive ..." to extend the type with new constructors.
-In \cref{fig:plugin-example2}, we make the example of boolean and three-valued boolean.
-
-% Recall, to construct a recursor, we need to create for each
-% constructor a handler (function) and aggregate them into a family (e.g.
-"subst_internal" in \cref{fig:plugin-example2}).
-We use "FRecursor" to declare a recursor (e.g.
-"subst") by specifying the motive, the aiming inductive type and the handler family.
-Once recursor is created it will look like any other function fields (e.g.
-"subst").
-Here in \cref{fig:plugin-example2} we can apply the function "subst" to "tm_var 0" as a new field.
-The inheritance of recursor is mainly delegated by the inheritance of the handler family---if the user extend with new constructor (e.g.
-"tm_true"), then the inherited family (e.g.
-"subst_handler") will need to extend correspondingly (e.g.
-"subst_internal.tm_true").
-If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "subst".\EDJ{This paragraph is duplicating but I am not sure how to make it concise.}
-
-We illustrate the usage and plugin implementation of inductive type and reasoning in the exemplar \cref{fig:plugin-example2}, the example of
-"subst" on STLC. When we define the extensible inductive type "tm", all
-the following fields (e.g., "subst_internal.tm_var") can only know "tm"
-is a type (of "Set") and there are at least two constructors for it,
-just like what we declare in ``the abstracted interface''---the compiled
-module type "STLC_Interface". However, we will not export
-%\YZ{export or put?}\EDJreply{Ok.
-%I just figure out how my wording is confusing. put is absolutely correct
-%but by put I want to mean ``export''. I think I use ``export'' in
-%several places...}
-the eliminator of the inductive type "tm" to "STLC_Interface", \textbf{since
-the mere existence of the eliminator will assert that there are only two
-constructors} in "tm", lead to an \textbf{non-extensible} "tm"!
-Thus the "STLC_Interface" will only include the ``constructor'' themselves in the form of parameter function.
-Doing so, all the subsequent fields (e.g.
-"subst_internal.tm_var" and the corresponding compiled "tm_var_handle") can be inherited to the
-context where "tm" is extended with new constructors without re-typechecking (re-compilation).
-
-To construct recursive function on a given inductive type, we use "FRecursion".
-"FRecursion" is a syntax sugar with (1) a normal family that includes all the case handling as field members, and (2) a step of doing type-checking and exhaustiveness checking (as in \ref{chg:extensible-inductive-type}).
-By considering all the case handling gathered in a family, the extension of the case handling can be simplified as the extension of the family.
-In our example, the resulting handler family is "subst_internal_410".
-While to achieve type-checking, we require the signature of eliminator about "tm" to be exposed.
-Here we show one
-exemplar extraction of the eliminator to \mintinline{Coq}{Set} as module
-"tm_rec_type". This concrete eliminator will be used by our "FRecursion".
-For example, in \cref{fig:plugin-example2}, after closing an "FRecursion" block, our plugin will feed Coq the module "do_tc" for doing type-checking with specified
-handlers (e.g. "subst_internal_410") during compilation---what it does is
-basically
-assert the existence of the recursor and apply to see if handlers can
-pass the type checking by Coq.\YZ{
-  Sounds like we can still claim the plugin is doing type checking.
-  It's just that the checking is done by generating some Coq code and
-  then having Coq check it.
-}\EDJreply{Yes! If my phrasing doesn't make things clear (i.e. our plugin is doing typechecking),
-feel free to change the phrasing for me}
-
-% Note that, at this point of the example,
-% "FRecursion" will only do type check via "do_tc". 
-% The compilation of "FRecursor"  only happens when closing the whole
-% family, by including "subst_18". 
-% However, notice the context (parameter) of the "subst_18": it is "ctx_subst_113", where the concrete inductive type is exposed in the context, instead of "STLC_Interface".
-This concrete exposition makes "__internal_tm_rec" accessible, but also makes "subst_18" non-inheritable when "tm" is extended.
-Thus in every children family, modules like "subst_18" needs to be regenerated.
-% Thus a "FRecursion" field is unlike normal fields, 
-% even though it is exposed as a plain
-% function for the succeeding fields in the programming interface.
-
-Note that, the "FRecursion" field (e.g. "subst") is not actually
-inheritable, because it is not compiled like any other fields.
-Recall how other fields are compiled into reusable pieces of
-(parametrized) modules; a recursor field is not compiled like that.
-
-In our example "STLC", when the "FRecursion" block is closed, it will only do type check via "do_tc".  The compilation of "FRecursion"  
-only happens when closing the top-level
-family "STLC", by generating "subst_18" on-the-fly and assembly it into "module STLC". .
-However, notice the context (parameter) of the "subst_18": it is "ctx_subst_113", where the concrete inductive type is exposed in the context, instead of "STLC_Interface".
-This concrete exposition makes "__internal_tm_rec" accessible, but also makes "subst_18" non-inheritable when "tm" is extended.
-Thus, for all the children family, we have to regenerate module and replace the
-module with the ``new'' recursor using ``the same'' handler
-family (e.g. "Module subst_242", whose context "self__STLC_bool" is also different from that of the parent).
-
-This is affordable since each aggregated recursor handler
-doesn't need to be rechecked: look at surface syntax, the user only need to handle the new constructors like "tm_true" in "STLC_bool.subst"; during compilation,"Module subst_internal_248" will directly include the old handlers
-from the parent (e.g. "STLC.subst_internal.tm_var") that have already been type-checked (since they only rely on ``the abstracted interface''
-module "STLC_Interface" and thus well-typed). 
-Thanks to family inheritance, their corresponding compiled module from the parents---for example, 
-the module "tm_var_handler"---doesn't require another type-check and is
-re-used during the type-checking and compilation of "STLC_bool.subst".
-
-Of course, in the implementation, these details are all hidden and when the user ``inherit'' the recursor "subst" in the surface syntax,
-our plugin actually will actually do a second exhaustiveness
-checking in children family mentioned above, with the ``same'' motive and the ``same'' handler family.
-%\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea.
-Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''}
-
-The plugin realizes the "Extend FInductive" via overriding---%
-when extending an inductive type $A$ with new constructor $c$, our
-plugin will generate a new (syntactic) inductive definition $A'$ with
-this new $c$ on the fly and feed it to Coq. But we must be careful on
-the exposing data in the family type (the compiled module type) for the
-succeeding field. In other words, we must be careful on the interface
-(context) based on which the succeeding fields are defined.
-
-% Insert one pseudo-code example for explanation of the mechanism
-% Use the natural number example
-
-% Explain why the module type need extra care
-
-% explain how recursor is constructed
-% explain we have the incremental-checking for any recursor
-
-% explain ftheorem wrapping this complicated recursor
-% This design of decoupling of the case handlers and exhaustiveness
-% checking handles the semantic well, however, it brings a lot of
-% boilerplate code---we need to create an extra (handler) family, manually
-% specifying the type of each case handlers, and then using "FRecursor" to
-% ``tie the knot'' and construct a recursor field from the handlers.
-
-For theorem proving, we provide a command "FInductive" that is doing similarly as "FRecursion" in programming interface but using tactic programming.
-Like "FRecursion", "FInduction" can handle newly added constructor in the children family as well.
-
-% This command (1) can avoid specifying the handlers family and the type of each case handlers when writing induction like above, (2) invoke
-% proof interaction mode and thus allow tactic programming, (3) and is also open to extension like "FRecursor". We expect this "FTheorem" to be used in theorem proving but not general programming.
-
-To summarize, (1) to make each recursor
-handler inheritable, we need to seal an abstraction around the inductive
-type (e.g., module "STLC_Interface"); but to construct a recursor (and carry out the
-exhaustiveness checking), we need to break this abstraction and see the
-concrete definition of the inductive type (e.g., module "tm_rec_type"). Our
-meta-theory need to handle these two seemingly contradicting ideas
-simultaneously when solving \ref{chg:extensible-inductive-type}.
-(2) \ref{chg:extensible-inductive-type} is also resolved by this
-\textit{decoupling} of the syntax of implementation of recursor handlers
-and exhaustiveness checking under the syntax sugar. 
-The former is handled by family inheritance
-and thus avoiding the boilerplate code; and the non-inheritance of the
-compiled recursor ensures the exhaustiveness checking happens for every
-recursor.
-
-% Attempt to rewrite the section for Overridable pins inductive/Closing Fact , and also introduce the concept of mixin
-
-\paragraph{Tactics and propositional recursors.}
-
-However, "FRecursor" provides barely enough functionality for
-manipulating our(extensible) inductive type, and it is far from
-satisfactory to be a legitimate programming interface. (1) We do not yet
-have any computational information using this "FRecursion" until its
-compilation into module, i.e. we cannot do computation with "subst" during
-the construction of the family using the current compilation strategy.
-(2) The necessary tactic like "discriminate" and "injection" is not
-applicable on (extensible) inductive type.
-
-\textit{Generating Scheme for Propositional Computation Axiom for FRecursor}.
-Thus we support "FScheme" to remedy problem (1): our plugin can automatically generate(postulate) the computational ``axiom'' for the specified  into the family and prove them during compilation into modules.
-For example, "FScheme Comp about subst" in \cref{fig:plugin-example2} will compile into "Module Type substtm_139", filled with all the computational axiom for "subst".
-There module type when included in the context can be used to reduce "subst"; they will also be proved after compilation.
-
-However, it is only introducing computational information \textbf{propositionally} but not judgementally, and we might need the help of rewriting tactic on these computational ``axiom'' to justify the computation.
-
-% However, this method, though remedies problem (1) to some extent, is only introducing computational information \textbf{propositionally} but not judgementally.
-Thus, for example, during theorem proving, we might need the help of rewriting tactic on these computational ``axiom'' to justify the computation.
-And in the full dependent type theory, propositional equality is sometimes not enough and bringing a lot of coding challenges\footnote{For example, for "u : F a"， "v : F b", "u = v" is not even well-typed when "a = b" only propositionally hold.
-Transport is required during reasoning.}.
-
-% To resolve (2), we can simply  Our following three sections will try to resolve these problems. 
-% we notice that, the power of "FRecursor" comes from its distinction from the mundane fields and its ability to stab through the abstraction of "FInductive".
-However, the latter is quite similar on how "Overridable ..
-pins {..}" can see through the abstraction of former pinned overrdiable fields.
-This inspires us to generalize pins to not only apply to overrdiable fields, but also "FInductive" and "FRecursor" fields.
-
-\textit{Overridable pins Inductive Type and FRecursor}. We notice that,
-the power of "FRecursor" comes from its distinction from the mundane
-fields and its ability to stab through the abstraction of "FInductive".
-Interestingly, the latter is quite similar to how "Overridable .. pins
-{..}" can see through the abstraction of former pinned overrdiable
-fields. This inspires us to generalize "pins"---we support pinning
-inductive type and "FRecursor". Once "FInductive" and "FRecursor" fields
-are pinned, we are directly exposed with the primitive coq-providing
-recursor and \textbf{judgemental equality} of the computational
-information for "FRecursor".
-
-The compilation mechanism is to simply make the underlying, compiled
-definition of inductive type and recursor included as part of the
-parameters of the module, just like how the concrete definition of
-overridable fields are included inside the parameters when pinned. This
-way, the computational information from vanilla Coq can be transmitted
-into the context.
-
-% However, only overridable fields can pin the (extensible) inductive type and "FRecursor".
-
-% Upon this design, it is natural to make "Overridable" fields to provide ``automatic overriding'' as it is likely the inherited tactic is able to provide a proof even in the context with the extended inductive type.
-
-% But a question arises: do we consider an ``automatic overridden'' field \textit{inheriting} the parent field or \textit{overriding} the parent field? 
-% TODO: Mention still needing total recursor, for exhaustiveness checking
-
-% To fulfill \ref{chg:software-engineering}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of
-
-% For example,
-
-%   For example, currently it is not possible to define a
-% vanilla inductive type dependent on our ``extensible'' inductive type
-% due to the invisibility, unless we compile the whole family. Similarly,
-% customized tactic expressions as required by \ref{chg:software-engineering} cannot
-% use lemmas or theorems in the fields of the current defining family as
-% well.
-
-% To remedy this, we support a "MetaData" command. This command
-% bundles any original Coq primitive (in the current family
-% context) into the compiled module type, and makes them visible to the
-% following defining field. Doing so, we can define customized tactic as
-% in \ref{chg:software-engineering} that uses lemmas proved in the family.
-
-% Following needs a rewrite
-To achieve "injection" in problem (2), we can easily generate/postulate
-$n$ ``injection'' propositions for $n$ constructors, just like how we
-did for computational information. But for "discriminate" tactic, this
-way will generate $O(n^2)$ propositions in a verbose way.
-
-What's more, to manipulate inductive type, it is not necessarily enough
-to only support "injection" and "discriminate" tactic (and their
-corresponding axioms). Recall in vanilla inductive type, we believe that
-eliminator is the strong-enough ``toolkit'' because the eliminator (with
-its computational information) is the \textit{universal property} of the
-inductive type---it extensionally describes the inductive type up to
-(certain) isomorphism. This universal property is much more powerful
-than "injection" and "discriminate" and maybe the most powerful tool we
-can have to reason about inductive type.
-
-For example, given a new inductive type, to prove "discriminate" between two constructors, i.e., "c$_i$ ..
-$\neq$ c$_j$ .." for two different constructors, we simply use the eliminator to map each constructor to a different natural number, assuming we can "discriminate" different natural numbers.\footnote{Rigorously, to avoid any suspicious circular reasoning, we can use large elimination to map them to $\top$ and $\bot$, and because $\exists a : \_, \top$ holds for the former not the latter, thus $\top \neq \bot$.} "injection" can be supported in a similar way---we can simply let eliminator map each inductive type to the arguments of the constructors.
-
-Thus it is natural to think if (an altered form of) eliminator can be
-postulated into our context, since apparently we cannot directly
-postulate the original eliminator here---because the sole existence of
-the eliminator prevents any possible future extension of the inductive
-type.
-
-Thus we alter eliminators into \text{partial recursor}, such that (1) it
-will not hinder the future extension, and (2) it is still powerful
-enough to derive injectivity ("inject") and disjointness ("discriminate") of constructors.
-What's more, (3) we can prove this \text{partial recursor} is
-actually good enough because it can act as an \textit{extensional
-characterization} for extensible inductive types \textit{to some extent}.
-We will call it \textit{propositional} because we only propositionally
-postulate its computational information for Coq.
-
-\textbf{Propositional Partial Recursor.} Partial recursor is simply adding partiality into the signature of the eliminator, by wrapping each motive with "option". 
-% This section is introducing the concept of \textit{partial recursor}.
-The motivations are various: (a) we want to have some semantic characterization of extensible inductive type.
-(b) We want to show that, some tactics reasoning like "inversion" is still valid even in extensible inductive type.
-(c) More generally speaking, notice that, for both "FRecursor" and "Closing Fact", they are not necessarily ``\textit{conservative extension}''---they will possibly disrupt arbitrary extension of inductive type.
-For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)} in the "Family STLC" either using "FRecursor" or "Closing", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
-We want to see what leads to a conservative extension---and the partial recursor is helpful for this question.
-
-% This raise an interesting question: what propositions are conservative extension? 
-% For example, apparently we know injectivity of constructor, (i.e.
-\mintinline{Coq}{STLC.tt ≠ STLC.ff}) must be conservative because it holds for all extension of the inductive type "STLC.tm".
-But is there a \textbf{``best'' proposition} (i.e.
-the proposition that everything the proposition derives are exactly conservative extension)?
-
-% Since injectivity of the constructors are derived by recursor in vanilla inductive type, an educated guess would be related to recursor.
-But we know the original recursor won't work because the mere existence of the original recursor will break extensibility.
-Thus we need to modify it.
-% The simplest idea is to allow partiality---i.e., returning
-% \mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
-
-% We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors?
-I suppose it has to do with the computational axioms?}\EDJreply{Yes.
-I want to emphasize the equality used by computational axiom is not judgemental equality.
-So Coq cannot automatically compute/reduce when a partial recursor applied to a term.
-This reduction can only be semi-automatically done by using [rewrite tactic and the computational axiom].}
-For example, the propositional partial recursor for \mintinline{Coq}{B.b} is
-
-\begin{minted}[escapeinside=@@]{Coq}
-(* Inside Family B *)
-FInductive b : Set := tt : b | ff : b.
-b_prec : forall R, option R -> option R -> b -> option R.
-\end{minted}
-with appropriate computational axioms.  This partial
-recursor differs from original recursor only by decorating the return
-type with an "option" and can thus be automatically generated (and
-the case is similar for those computational axiom). Thanks to the
-partiality, partial recursor doesn't break
-extensibility (i.e., all future extensions of "b" can support this
-"b_prec") because for example, "b_prec" can just return "None" when bump into extended constructor in the future.
-
-The second thing to check is that, if partial recursor is theoretically enough for injectivity of the constructor. The rough idea to achieve this
-is to use "b_prec" to reflect our special inductive type into vanilla
-inductive type, and use the original "injection" and "discriminate"
-tactics. For example, we want to prove \mintinline{Coq}{B.tt ≠
-B.ff}, we simply 
-\begin{minted}{Coq}
-Definition reflect : B.b → option bool 
-                   := b_prec (fun _ => bool) (Some true) (Some false).
-Definition tt_neq_ff : B.tt = B.ff → False.
-(* Because B.tt = B.ff → reflect B.tt = reflect B.ff 
-      → Some true = Some false    (by computational axiom)
-      → False           (by discriminate of bool and injection of Some) *)
-\end{minted}
-
-Thirdly, even though we cannot prove partial recursors are \textit{the
-best} reasoning tool for extensible inductive types,
-the above "reflect" function implies that propositional partial
-recursors are a good \textit{extensional characterization} of
-(non-indexed) extensible inductive type (under certain
-constraints)---because we can ``embed'' vanilla inductive types into
-types supporting a partial recursor: "reflect" is actually a left
-inverse of an injection "bool → self_B.b". A bit more formally, and
-restricting our focus to non-indexed inductive types:
-
-\begin{theorem}\label{thm:prec-complete} Given a list of $n$ pairs of types $\{ x : "A"_i \vdash "B"_i(x) \}_{i}$ s.t.
-
-  \begin{itemize}
-    \item \textlabel{(Detectable Partiality)}{prop:detectable-partiality} for each pair $x : A_i \vdash B_i(x)$,
-    \begin{minted}{Coq}
-      Axiom detectable:
-      forall {T} {a} (f: Bᵢ a -> option T),
-        {forall x, f x <> None} + {exists x, f x = None}.
-    \end{minted} 
-    \item Define $"C"$ as the inductive type using this list, i.e.,
-    \begin{minted}{Coq}
-Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
-    \end{minted}
-    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
-    \begin{itemize}
-      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
-      \item a partial recursor
-      \begin{minted}[escapeinside=@@]{Coq}
-prec : forall (R : Set), ..,
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,@\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint.
-Let me fix other parts first}\YZreply{Is this constraint Detectable Partiality?}@
-      D -> option R
-      \end{minted}
-      \item and $n$ (propositional) computational axioms: for all $i$, 
-      \begin{minted}{Coq}
-prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
-    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
-      \end{minted}
-    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
-    \end{itemize}
-  \end{itemize}
-  Then there exists an \textbf{embedding} from "C" to "D".
-  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of\/
-  "C" and "prec" of\/ "D" (both acting like identity) such that
-  \begin{minted}[escapeinside=@@]{Coq}
-    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
-If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
-I mentioned left inverse 'linv' is the generalization}@
-  \end{minted}
-\end{theorem}
-The reason we have this weird looking \ref{prop:detectable-partiality}
-is that the construction of "linv"---at one point, using "prec" to
-construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
-C) → option C}, generally not provable, unless we have
-\mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.\YZ{Don't you really mean
-"unless we have the Detectable Partiality assumption instead"?}
-This \ref{prop:detectable-partiality} is exactly transforming former to
-the latter:
-recall that, according to the semantic of inductive type,
-\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
-"linv" on the substructure of "D". This function points out each of the
-translation result from "D" to "C" and some of them are "None" because
-of failing. For us, we only care about if \textbf{any} of the
-substructure of "D" fails to translate---if any substructure fails,
-then the whole translation fails and thus we should have "None" as
-return.
-
-\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that includes all of the examples shown in this paper.
-
-\ref{prop:detectable-partiality} is not generally supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
-However, in those cases, we still have partial recursor provable and useful---we just cannot see the partial recursor as a strong toolkit anymore.
-
-We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
-actually be considered as a constructor, because they can be reflected
-to real constructors of "C" using the left inverse "linv" (directly generalization of "reflect : self_B.b → option bool" of the above). Thus, \cref{thm:prec-complete}
-implies that (1) every future extension of the inductive type can
-support this partial recursion (trivially); (2) every type supporting
-this partial recursor with its computational axioms at least supports
-these constructors because of the embedding. In other words, \textit{a
-type (at least) supports these constructors if and only if this type
-supports the corresponding partial recursor (with the computational
-axioms)}.  Thus, we can argue that the partial recursor gives a sound and
-complete extensional characterization of all the extension of a given (non-indexed)
-inductive type.\YZ{What about indexed inductive types?}\EDJreply{Done.
-I also have a formalization of this proof now, for indexed type and dependent eliminator.
-Though the formalization is about single constructor.}
-
-We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
-\begin{minted}{Coq}
-  pdelim : ∀ (P : D -> Set), .., 
-  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
-    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
-\end{minted}
-To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable.
-Please refer to the supplementary material and appendix for a formalization of the left inverse in the context of indexed inductive type.
-
-Though partial recursor is provable by simply using "FRecursor", for convenience, we support using "FScheme" to automatic generate it.
-For example, "FScheme tm_prec PRecT" in \cref{fig:plugin-example2} can lead to compilation of "Module tm_prec_72".
-At the meantime, we implement a tactic "prec_discriminate" that when fed with the appropriate partial recursor (with computational axiom generated), it can discriminate an absurd equation between two different constructors of extensible inductive type.
-\begin{minted}{Coq}
-  FInductive tm : Set := tunit : tm | tvar : ident -> tm | ..
-  FScheme tm_prec PRecT about self__STLC.tm.
-  FScheme Rec2D about self__STLC.tm_prec.
-  (* ... *)
-  Goal ∀ i, tvar i = tunit -> False.
-  intros i H. prec_discriminate self__STLC.tm_prec H. Qed.
-\end{minted}
-
-Finally, we should point out that, the mere postulation of partial recursor is \textit{conservative}.
-In other word, generally speaking, using "FRecursor" to prove theorems is not necessarily a ``\textit{conservative extension}''---the proved theorems might hinder future extensions of the inductive type.
-For example, for \cref{fig:STLC-example}, once we proved \mintinline{Coq}{var_not_value : ∀ i, ~ value (tm_var i)}\footnote{Achieved by three steps: first proving partial recursor using "FRecursor", then use partial recursor to prove "tm_var i $\neq$ tm_lam v body" and finally using "FRecursor" prove "var_not_value" by induction on "value"} in the "Family STLC" either using "FRecursor", then any children family of "STLC" cannot extend a new constructor \mintinline{Coq}{vvar: ∀ i, value (tm_var i)}.
-But the postulation of partial recursor will not cause this problem.
-
-\textbf{Escape Hatches: "Meta-data" and "Closing Fact".} There are cases the developers want to refer to global reasoning or just escape from the programming paradigm of family polymorphism.
-Thus we provide the following mechanism.
-
-\textit{"MetaData".} Due to our implementation of the family,
-the family data is invisible to Coq internal and thus causes some
-inconvenience. For example, in
-\cref{fig:plugin-example-global-reasoning-meta-data} the first attempt
-of defining tactic "inv" fails. Because the earlier defining "tm",
-"value", and "value_not_app" are currently just data structure in the
-plugin and not visible for Coq internal (only visible after the family
-is compiled into module).
-
-Thus to refer to them, we have to scope the defining tactic with
-"MetaData" and our plugin will set up the appropriate environment. After
-a "MetaData" block is closed, the plugin will aggregate the content into
-the family being defined so that they can also be used when defining
-the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}\EDJreply{Not sure.
-This auto-insertion will practically look like hijack the original Coq command.
-I am not sure this kind of hijacking can happen.}
-
-This can not only fulfill \ref{chg:software-engineering} and define new customized tactic but also integrate any vanilla Coq's feature into family.
-For example, if the user don't want to use "FInductive" but only the non-extensible "Inductive" type, the user can simply declare the inductive type inside "MetaData" to be part of the family.
-
-\textit{"Closing Fact".} We have described how to use partial recursor to "discriminate" different constructors of an extensible inductive type.
-Practically, "discriminate" is used when proving inversion lemma, like forementioned "var_not_value".
-
-However, it seems verbose to prove this simple inversion lemma using "FRecursor", let alone the fact this "FRecursor" will need to be extended every time the inductive type (e.g.
-"value") is extended.
-Thus we provide a command "Closing Fact".
-"Closing Fact" will prove the given lemma in the context where all the definitions \textbf{are exposed}.
-Thus the above "var_not_value" can be simply achieved by
-\begin{minted}[fontsize=\footnotesize]{Coq}
-Closing Fact var_not_value : 
-  forall i, ~ self__STLC.value (self__STLC.tm_var i) 
-    by { intros i H; inversion H; eauto }.
-\end{minted}
-We can use "inversion" tactic here thanks to the exposure of the vanilla inductive type in the context.
-
-We simply support inheritance for "Closing Fact".
-However, we fix the semantic of the "Closing Fact" by this one-line \textbf{untyped} proof script---and thus we restrict "Closing Fact" to be used in a proof-irrelevant way.
-
-\textit{"Overridable ..
-pins {..}" vs.
-"Closing Fact".} A keen reader will notice a great deal of similarity between "Overridable ..
-pins {..}" vs.
-"Closing Fact" as they both expose the concrete definition of the former fields (including extensible inductive type).
-In fact, using "Overridable" can prove this "var_not_value" in an almost identical way.
-The only difference seems to be the ``auto-inheritance'' of "Closing Fact" with the help of the proof script where even if the extensible inductive type is extended---but this difference seems superficial as we can implement something similar for "Overridable".
-
-The main difference is that, the semantic of a "Closing Fact" field is the \textbf{untyped} proof script but the semantic of a "Overridable" field is always a \textbf{typed} proof term.
-Thus when a pinned extensible inductive type/former field is extended/overridden, we can still \textbf{inherit} the "Closing Fact" but we can only \textbf{override} the overridable field.
-
-This difference might seem small and irrelevant to the user from the first sight but this becomes quite relevant when we introduce \textit{mixin} of two inheritances.
-% \textbf{Overridable Field and Closing Fact}
-% It is natural to consider the ``automatic overridden'' resulting inheritance, as we are using the same tactic (or the same untyped Gallina term to provide the proof).
-But for type theorists, these are definitely overriding because these are definitely different typed Gallina term!
-
-% Instead of making a decision here, we introduce "Closing Fact":  "Closing Fact" will make a field tied with the untyped proof script and only allows inheritance.
-Of course, during inheritance, the untyped proof script will always be re-run and introduce (potentially) \textbf{different} typed terms\footnote{Thus called "Closing Fact" as its semantic depends on the closed  surrounding family in a global sense}.
-Thus, we encourage the user to use it in a proof-irrelevant way, where the difference between typed terms can be ignored.
-Then, the ``automatic overridden'' will always lead to overriding.
-
-% This is similar to "FRecursor"---we only allow inheritance for "FRecursor" as its semantic can be considered as a tie between a handler family and an extensible inductive type, where during each inheritance, the tie will be rechecked.
-
-% Just like "Overridable ..
-pins {..}", "Closing Fact" provides the ability to stab through all the abstraction of the former fields, and thus has the similar functionality  as "Overridable ..
-pins {..}".
-
-% \textbf{Mixin}.
-The reason we introduce two different semantics for two similar functionality for "Overridable ...
-pins {...}" and "Closing Fact" is due to the fact that, we want to introduce mixin in our facility.
-
-% \textit{Mixin} is implemented by making inheritance judgement/data supporting \textit{weakening} and composition.  
-\textbf{Mixin} is implemented by making inheritance judgement/data supporting \textit{weakening} and composition.More concretely,  for an inheritance from parent family type $\sigma_1$ to children family type ${\sigma_2}$ and another inheritance from ${\sigma_1}$ to ${\sigma_3}$, we can weaken(expand the input of) the second inheritance to ${\sigma_2}$ then the output children family type becomes ${\sigma_3'}$.
-Then the ``composition'' of the very first inheritance and the altered second inheritance is the resulting mixin.
-
-However, this implementation breaks commutativity: for example, when the two inheritances are overriding a given field in different ways, the order of composition becomes important.
-Thus to avoid this ambiguity, we simply \textit{don't allow overriding fields} inside any inheritance when we do mixin.
-In this case, a proof irrelevant "Closing Fact" is a useful alternative for overriding.
-
-Note that, every top level family can be considered as an inheritance because (1) our implementation is simplified by considering every family as an inheritance, and (2) we want to align with the complete family polymorphism advocated by \citet{zm2017} where the user can override the parent of the inheriting family (inside a family due to late-binding), and in these case the nested inheriting family is more a inheritance data than a ``standalone'' linkage.
-Thus we will store for every top level family in our plugin the corresponding inheritance data together with a ``reference'' to the parent family.
-
-% Mixin example:
-
-% Then our mixin command can single out the inheritance data from "STLC_bool" and "STLC_prod" and ``mixin'' the inheritance to reach "STLC_bool_prod".
-
-% However, we don't allow any overriding happening in either mix-edin inheritance, because it will break the commutativity and complicate the semantic without much benefit.
-Overriding is not allowed in this case, but "FRecursor" and "Closing Fact" is still allowed since they only allow to be inherited.
-
-% \textbf{Hack: Global Reasoning.}\YZ{This is more like an escape hatch. Escape from family poly into vanilla Coq. Maybe these two hacks should be grouped into one section}
-% % explain the places using ``Closing Fact''
-% % 1. No need to export partial recursor
-% % 2. Computational rule for the total recursor
-% Unfortunately, there are cases an extensible proof brings more boilerplate code. For example, in STLC, we expect
-% ``\mintinline{Coq}{~ value (tm_app x y)}'' to hold in all future
-% extensions of STLC---applications will never be considered as values
-% by any extension.
-% We know the correct way to prove the proposition
-% \mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use
-% "FRecursor" to inductively reason "value (tm_app x y)". However, this
-% means that every time we extend "value" with new value forms, we will
-% need to extend corresponding inductive cases for "app_not_value". We can
-% imagine the newly extended proof to be boring case analysis that the
-% newly added value forms are not "tm_app". What's worse, similar
-% scenario can happen on other statements like \mintinline{Coq}{~ value
-% (tm_if cond x y)}. Everything here is extended correctly but we end up
-% having boilerplate code, \textbf{semantically}.
-\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}\EDJreply{Done}
-\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
-Family STLC.
-  FInductive tm : Set := ... 
-  FInductive value : tm -> Prop := ... 
-  Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
-      by { intros x y H; inversion H; eauto }.
-  Fail Ltac inv := 
-    match goal with 
-    | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
-    (* Fail: value, tm_app, value_not_app unfound *)
-    end. 
-  MetaData tactic1.
-    Ltac inv := (* ... the same definition as above ... *)
-  EndMetaData.
-  (* ... STLC example continue ... *)
-EndFamily.
-\end{lstlisting}  
-\caption{Example for Global Reasoning and MetaData, using in STLC example}\label{fig:plugin-example-global-reasoning-meta-data}
-\end{figure}
-
-% Our proposal is to allow global reasoning by providing "Closing Fact" command. Ultimately, "Closing Fact" is asserting a constraint upon all the
-% possible extension of the surrounding family and claiming that a proof script
-% can solve the constraints.\YZ{This is a good one-liner that summarizes the intended usage scenarios of "Closing Fact".
-Should probably bring it out earlier.}\EDJreply{I put it here as you suggested but I think this one-liner at here will cause more confusion because it is too abstract.}
-% The assertion will only be verified when the whole family is closed and compiled into a module.
-
-% % We provide "Closing Fact" as a command to
-% % carry out this global reasoning, attached with a proof script and proof
-% % term. This "Closing Fact" will look like an assertion during the
-% % construction of the enclosing family. This assertion will only be
-% % verified when the whole family is closed and compiled into a module,
-% % with the help of the attached proof script and proof terms.
-
-% Here we show the example to prove "value_not_app" in
-% \cref{fig:plugin-example-global-reasoning-meta-data}, the script inside
-% curly brackets after "by" is the attached proof script that will be
-% executed upon "EndFamily", which triggers the compilation of the family
-% being defined.
-% Since we are dealing with a concrete (vanilla) inductive type during compilation,
-% we can use Coq's "inversion" tactic to complete this proof.
-% This "Closing Fact", as an assertion, will have no problem being
-% inherited, but the proof script will be executed every time a derived
-% family is compiled.
-
-% The downside is that "Closing Fact" cannot immediately check the correctness of the
-% attached proof term/proof script because this verification only happens when we
-% finished defining a family, and thus a bit harder to handle.\YZ{'harder' in what way?
-I suppose you mean the "Fail Ltac" in Figure 5?}\EDJreply{The proof script cannot be run immediately to reply the user whether the proof script is even correct.
-This is bad engineering to some extent.
-In those bad cases, because error only happens at the the very end ``EndFamily'' , the user will need to go back to the place of Closing Fact to adjust the proof script, or to adjust other parts.
-I am not sure if early check is possible to be done (engineering-wise) I think it is possible (imagine I just do an early compilation, i.e.
-only compile part of the family to the point closing fact is used) but it is a big engineering difficulty for me.}
-% We expect it to be used only sporadically for maintainability, and as an alternative to using "FRecursor".
-This is called global reasoning because this is a post-hoc reasoning when the family as a \textit{whole} is compiled into a module.
-
-% However, to verify and prove the recursor and partial recursor and their
-% computational rules (as in \ref{chg:definition-relevant-reasoning}), our plugin has to
-% compile the extensible family into a non-extensible Coq Module and define
-% the recursors by instantiating the content of Coq module with concrete
-% inductive types. We provide "Closing Fact" as a command to carry out this global reasoning,
-
-% This style of ``Global Reasoning'' says that we can prove the
-% corresponding theorem only when we know the ``full picture'', in particular,
-% all the constructors of an inductive type. 
-% We provide "Closing Fact" as
-% a command to carry out this global reasoning
-
-%\YZ{Are recursors and partial recursors automatically generated and proved by the plugin, or must they be asserted and proved by the programmer?}\EDJreply{(1) Closing Fact is not only used to generate recursor and partial recursor.
-I think the intro of this section is just giving people wrong impression.
-So I rewrite the whole section now.
-(2)  Recursors and Partial recursors are possible to be automatically generated but my plugin didn't implement it.
-Current stage is that user only need to assert them.
-Nobody has to prove it (plugin will prove it).
-Here when I say the user has to provide proof script, it means for arbitrary proposition if the user want to carry out global reasoning.}
-
-% ---we can attach to "Closing
-% Fact" either (proof) terms or Coq's proof script, then only when
-% compilation to module happens, is the term/proof script
-% type-checked to see if a "Closing Fact" statement can be proved. Note
-% that, "Closing Fact" cannot immediately check the correctness of the
-% term/proof script because this verification only happens when we
-% finished defining a family, and thus a bit harder to handle.
-
-% This "Closing Fact" also can ease us from the necessity to let plugin
-% generate proof for recursors and partial recursors and their
-% computational rules (as in \ref{chg:definition-relevant-reasoning}), and thus making
-% plugin development easier.
-%\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy.
-Question is if it makes the programmer's life easier.}\EDJreply{You are right.
-It does sound like a bad excuse.
-I rewrite the whole section now.
-Closing Fact is initially used when the user wants global reasoning instead of extensible proof by induction, for example, the inversion lemma.
-\\ It just happens that they can be used by me so that I can postpone generation of these rule plugin (which is hard for me to do).
-This paragraph is just me being honest that my plugin is not mature enough to generate everything.}
-% For example, the computation rule for a
-% concrete recursor like "B.neg" and partial recursors in
-% \cref{fig:plugin-example2} can be directly asserted by using "Closing
-% Fact" and their proofs verified later during compilation.
-%\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know.
-There must be some other ways in the literature.
-}\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}\EDJreply{Please check.}
-% This ``Global Reasoning'' will be again useful in later examples.
-
-\fi
-
 \section{Limitations}

### 2023-03-19 metatheory-condensed-old.tex
relocate old writings

--- 2023-03-18 metatheory-condensed.tex
@@
-We contribute \TT, a core type theory that extends
-%\underline{f}a\underline{m}i\underline{l}y polymorphism
-%\underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory (MLTT)~\cite{martin1982constructive}
-\citeauthor{martin1982constructive} dependent type theory (MLTT)
-with facilities to express family polymorphism
-while maintaining consistency and canonicity.
-%\NA{maybe you can give an overview of the salient points of the design and what they're trying to achieve? What is the core aspect modeled by the theory?}
-%
-
-For accessibility of the main text, \cref{sec:metatheory2} presents \TT using
-named variables and meta-level substitution.
-For clarity of proof details, the technical appendix supplementing
-\cref{sec:metatheory2} instead uses de Bruijn indices and explicit substitution.
-%
-We acknowledge that the presentation is dense for an audience
-unacquainted with MLTT, so we summarize the salient points first.
-
-\noindentparagraph{Summary.}
-
-\TT is intended as a foundational model. So unlike our programmer-facing plugin,
-%typing is structural rather than nominal\EDJ{I don't actually know what do you mean by structural here. I think surface syntax (Coq) is pretty structual as well.}, and
-fields automatically axiomatized by the plugin require explicit definitions in \TT.
-%
-\TT provides MLTT-style constructs that can be used to express families and
-family polymorphism. %rather than building it in directly.
-%
-Most notably, it extends MLTT with what we call \emph{linkages}.
-Linkages are a namesake of the theoretical device through which \citet{zm2017} model
-family polymorphism in an OO setting, but the technical details differ significantly from the prior work.
-
-\begin{itemize}
-[labelsep=*,leftmargin=1pc,itemsep=3pt]
-\item 
-Linkages model families, so they are like tuples composed of fields
-(with field names represented by variable bindings).
-But there is a twist: linkages support late binding.
-Unlike dependent tuples where a later component is \emph{existentially}
-quantified over the earlier ones, a linkage component is \emph{universally}
-quantified over---and thus polymorphic to---the components preceding it.%\Zhixuan{explain the intuition behind the (typing rule of) linkages, and try to connect it to the ``family'' in the earlier section/surface syntax}\EDJreply{I am not sure if more is needed because the section ``Family polymorphism via linkages'' is trying to do this.
-Maybe I should estabilish connection between Rule l/add with the compilation process of fam poly?}\YZreply{I don't think you should mention anything about compilation process here.
-The surface syntax?
-Maybe.
-But I prefer not changing this sentence; it's supposed to be an executive summary rather than a detailed explanation.}\EDJreply{Actually I current am not sure where to add more sentences and how to elaborate more.
-Except for mentioning compilation process I don't think I have any good idea.
-So I will just stop here.}
-
-\item
-\TT features \emph{linkage transformers}, which model how a family can be
-inductively constructed by inheriting fields from another family,
-adding new fields, and overriding existing fields.
-\item
-Inductive types are modeled as W-types~\cite{martin1984intuitionistic} and their
-extension as overriding.
-
-\item
-Consistency and canonicity of \TT are proved by giving semantic interpretations to the syntactic typing judgments.
-\end{itemize}
-
-%\TT supports extending inductive types through
-%formation rules that enable the encoding of the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
-
-%Also unlike our plugin, \TT does not feature automations such as
-%the generation of propositional equalities and partial recursors,
-%but it is expressive enough to encode their definitions.
-
-% \input{mltt-selected.tex}
-\input{mltt-selected-named.tex}
-\input{fmltt-selected-named.tex}
-
-\noindentparagraph{Brief review of MLTT.}
-
-\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT %(predicative) MLTT
-(and \cref{fig:fmltt-selected} \TT).
-Dependent function types \nTyPi{x}{A}{B},
-dependent pair types \nTySigma{x}{A}{B}, and
-identity types \TyId{t_1}{t_2} are standard.
-So are their introduction and elimination forms.
-Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
-A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
-in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
-Definitional equalities have the forms %(for each judgement)
-\goodCtx{\Gm_1 \equiv \Gm_2}{},
-\goodType{\Gm}{T_1 \equiv T_2}{},
-\goodTerm{\Gm}{t_1 \equiv t_2}{T},
-etc.
-Following \citet{altkap2016}, we regard our (intrinsically typed) syntax
-as being quotiented by these equalities.
-Quotienting facilitates coercion along equalities---given
-\goodCtx{\Gm_1 \equiv \Gm_2}{} and
-\goodType{\Gm_1}{T_1 \equiv T_2}{},
-the derivation of \goodTerm{\Gm_1}{t}{T_1} is considered definitionally equal to
-a derivation of \goodTerm{\Gm_2}{t}{T_2}.
-%\YZ{Is this accurate??}\EDJreply{Yes.
-In QIIT model, it looks like `t : Tm G1 T1' and  `t : Tm G2 T2'.
-In extenional Agda, because ``G1 = G2'', ``T1 = T2'', thus the two types `Tm G2 T2 = Tm G1 T1'.
-So they are the same derivation in QIIT.
-In our setup, we have the same idea, because after quotient, G1 refers to same object as G2, so the derivation is the same.}
-%\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B.
-Maybe this is something require emphasis?
-Because it is usually handled by LF (like QIIT) but now we are using bare hands so....
-maybe we need to emphasize this}
-%For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement.
-It applies to each kind of judgement, for example, (G |- ), which looks like (G : Con), (G |-) is the derivation and (_ |- )is the judgement \\
-%Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation.
-Similar for ( _ |- ) vs.
-(G |-) and (G |- _ ) vs (G |- T).
-\\
-%I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e.
-there are no two different derivations lead to the same t; for each t, there is only one derivation for it).
-This is not true for extrinsic typing.
-\\
-%If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
-%I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
-%}
-
-We use universes à la \citet{coquand2013presheaf}, following \citet{kaposi2019gluing}.
-Unlike Russell-style ones, these universes are not inhabited by types directly,
-but rather by the \emph{codes} of types,
-and arguably behave better due to its closeness to Tarski-style universes~\cite{luo2012notes}. 
-% and compatibility with our algebraic formulation of MLTT~\cite{coquand2018canonicity}. % I am not sure if I want this sentence or not
-Term $\codety{T}$ encodes type~$T$ (\ruleref{tm/c}) and thus inhabits universe~$\cU$.
-Type $\El{t}$ decodes term~$t$ (\ruleref{ty/el}).
-%\footnote{By contrast,  Russel-style universes are inhabited by types directly.
-Tarski-style universes are ``semantically more fundamental'' \cite{luo2012notes} and friendlier to our meta-theoretical reasoning.}
-There is an infinite hierarchy of universes;
-we omit universe levels in the presentation to avoid clutter.
-%\rA{Why is W(t) a term and not a type?}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}
-
-% We use intrinsically typed syntax \cite{church1940}---terms and types are well-formed by construction.
-For concision, typing rules omit obvious premises required for well-formedness.
-For example, \ruleref{tm/lam} implicitly requires $\goodCtx{\Gm}{}$
-(i.e., that the context be well-typed).
-
-%We briefly review the base MLTT fragment first.
-%
-% In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}\YZ{Does coquand2018canonicity really use explicit subst?}\EDJreply{Yes it is.
-Look at Section 2.1.
-When he starts reasoning, he actually deals with category with family (instead of the typing rules), where he consider morphism between contexts, which is substitution.
-But I agree it can be a bit ambiguous to see (because he literally show two version of MLTT without connecting them---one classical typing rule version, the other directly works on Category with Family (CwF)) but since his proof is mainly on Category with Family, so it counts as explicit substitution.
-But if you are concerned, we can change to Andras Kovacs "Generalized Universe Hierarchies and First-Class Universe Levels".
-or Andreas Abel's "Decidability of Conversion for Type Theory in Type Theory".
-The former one directly work on CwF, the latter one include substitution as part of the raw syntax.}
-% in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
-% substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
-% of the syntax rather than meta-operations.
-%
-% Variables are in the form of \citeauthor{debruijn1964} indices:
-% \var{n} is the variable bound by the $n$-th closest enclosing binder.
-% For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
-%
-% Substitutions are typed with the form 
-%\EDJ{It seems to me G |- g : D is a judgement not a rule.
-Rules usually have two parts (premises + conclusion).
-Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here).
-I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this.
-In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
-% $\goodSub{\Gm}{\gm}{\Dl}$.
-% The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
-% yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
-% %
-% The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
-% \sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
-% \sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
-% then applies~$\gm$.
-% For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
-% that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
-% $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
-% where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
-
-\noindentparagraph{Introducing and eliminating inductive types.}
-
-% TODO: Remove mentioning conventional Wtype here
-
-W-types~\cite{martin1984intuitionistic} are a succinct way to model
-inductive types in MLTT.
-Together with the identity type, they can express a whole host of inductive
-types~\cite{hugunin2021not}, including those with multiple constructors.
-
-Our formulation of W-types differs from previous ones in that it is
-straightforward to identify constructors from what we call \emph{W-type signatures}.
-%
-A signature $\goodWSig{\Gm}{\wsig}{n}$ is
-composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}),
-%\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types".
-Since there are already boolean types and Σ types, Fin types can be defined.
-Using Fin, ordinary W types suffice to define trees with n constructors.
-So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}\EDJreply{This following sentence has been replied to Zhixuan and he doesn't has further response: 1.
-We have to have enhanced W-type at FMLTT, that's for sure.
-Otherwise we (a).
-are too far away from the surface syntax because W-type itself only has one constructor all the time (b) this suggested encoding or syntactic sugar, since it interact with other connective like linkage (the Wrec is the interaction), so Wrec part is mostly unchanged  (c) since they will look the same, the encoding itself will look like an extra equality judgement in the syntax...
-which is not helpful for the complexity at all ...
-and even more complex  \\
-%2.
-so to me the problem you guys really have is around the content in the appendix -- I think Zhixuan mean he doesn't think this is more expressive.
-But my point is, it is hard to really write out the above syntactic sugar, formally.
-I think the issue here is similar to other reviewer saying "syntax translation".
-It is very hard to write out the encoding of Fin already, and then encode it again into W-type.
-(Just ask my agda mechanization).
-Of course, some hand-waiving can be done.
-For example, the part Zhixuan says I think it okay to hand-wave.
-But the destruction part -- the part Wrec interact with this encoding, I don't believe handwaving is enough.
-\\
-%3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
-%
-each modeling a constructor of the inductive type.
-The $i$-th pair, projected from the signature~$\wsig$ using the forms
-$\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and
-$\goodTerm{\Gm}{\wsigproj{i}{2}{\wsig}}{\wsigproj{i}{1}{\wsig} \to \cU}$,
-defines the $i$-th constructor.
-That is, given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
-$\goodTerm{\Gm, x : \El{\wsigproj{i}{2}{\wsig}(t_1)} }{t_2}{\El{\wcode{\wsig}}}$,
-one can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type \El{\wcode{\wsig}} (\ruleref{tm/wsup}).
-$\wcode{\wsig}$ gives the code of the
-W-type (\ruleref{tm/w}).
-
-For each pair of types identifying a constructor, the first type may be thought of
-as the non-inductive part of the constructor, and the second type as the arity
-of the inductive part.
-For example, the signature~$\wsig_{\texttt{tm}}$ of the W-type modeling
-the inductive type~\lsti{tm} (\cref{fig:stlc-mechanized})
-is constructed as follows,
-where $\bbzero$ is the bottom type~$\bot$, $\bbone$ the unit type~$\top$, $\bbtwo$ the boolean type,
-and $T_{\texttt{id}}$ a type encoding~\lsti{id}:
-%
-\begin{align*}
-\small
-\begin{tabular}{@{}llll@{}}
-\lsti!tm_unit : tm!
-&
-\lsti!tm_var : id -> tm!
-&
-\lsti!tm_abs : id -> tm -> tm!
-&
-\lsti!tm_app : tm -> tm -> tm!
-\smallskip\\
-$\wsig_{\texttt{tm}}^0 \coloneq \WSigAdd{\WSigEmp}{\bbone}{\lambda \_ . \bbzero}$ 
-& $\wsig_{\texttt{tm}}^1 \coloneq \WSigAdd{\wsig_{\texttt{tm}}^0}{T_{\texttt{id}}}{\lambda \_ . \bbzero}$
-& $\wsig_{\texttt{tm}}^2 \coloneq \WSigAdd{\wsig_{\texttt{tm}}^1}{T_{\texttt{id}}}{\lambda \_ .\bbone} $ 
-& $\wsig_{\texttt{tm}} \coloneq \WSigAdd{\wsig_{\texttt{tm}}^2}{\bbone}{\lambda \_ .\bbtwo}$
-\end{tabular}
-\end{align*}
-
-\noindent
-Thus, the encoding of \lsti{tm_abs} has type
-$T_{\texttt{id}} \to (\bbone \to \El{\wcode{\wsig_{\texttt{tm}}}}) \to \El{\wcode{\wsig_{\texttt{tm}}}}$,
-and that of \lsti{tm_app} has type
-$\bbone \to (\bbtwo \to \El{\wcode{\wsig_{\texttt{tm}}}}) \to \El{\wcode{\wsig_{\texttt{tm}}}}$.
-These types are strictly positive by construction.
-
-%We illustrate the structure of W-type signature (as a list) here.
-%$\wsig_{\texttt{tm}}^0, \wsig_{\texttt{tm}}^1$ are (the subparts of) the
-%signatures of constructors \{"tm\_unit"\} and \{"tm\_unit","tm\_app"\}
-%respectively. Each item in this W-type signature, having two types, corresponds
-%to one constructor, where the first type stores the non-recursive part of the
-%constructor and the second type stores the ``arity'' of the  recursive part. For
-%example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times
-%(\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule
-%\ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone,
-%\wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \lambda \_ . \bbtwo$). This type means
-%there is no information on non-recursive part (thus use unit type $\bbone$) and
-%the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1
-%= \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\lambda \_ .\bbtwo}$.  The story is
-%similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and
-%thus the ``arity'' is zero.\YZ{Feels like it's more straightforward to just list
-%the encodings of the four constructors of tm in a table, assuming an encoding of
-%the identifier type. And this paragraph should be moved to the earlier paragraph
-%on W-types. I'll take a stab at this.}
-%
-%Generally speaking, the expressiveness of W-types comes from the ``arity'', as
-%it encodes the case when recursive structure stores as the output of a function,
-%mirroring the strict positivity of Inductive type.
-
-W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a
-W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
-case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
-Each case handler has a type of form $\CaseSig{A}{B}{T}$, where
-$T$ is the motive of the recursion (\ruleref{tyeq/casety});
-for simplicity, we model only non-dependent motives.
-The collection of case handlers~$\lkg$ encodes those defined and inherited
-by an \lsti{FRecursion} command in our plugin.
-We choose to type it with a linkage type $\TyLkg{\RecSig{\wsig}{T}}$
-to avoid introducing non-dependent $n$-tuples, which linkages generalize.
-
-% \begin{figure}
-%   \begin{tabular}{llll}
-%     \hline
-%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
-%       \multicolumn{1}{l|}{Equivalent Type} &
-%        &
-%       \multicolumn{1}{l|}{} \\
-%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}| unit : tm} &
-%       \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
-%       \multicolumn{1}{c}{$\tau'$} &
-%       \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
-%     \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}| app  : tm -\textgreater tm -\textgreater tm} &
-%       \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
-%       \multicolumn{1}{c}{$\tau$} &
-%       \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
-%     \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}unit} &
-%       $\coloneqq \Wsup{1}{\tau}{\unit}{"elim-"\bbzero}$ &
-%       \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}app unit unit} &
-%       $\coloneqq\Wsup{0}{\tau}{\unit}{"unit"}$ \\ \hline
-%     \cellcolor[HTML]{ECF4FF}$\lsig' = \LSigAdd{\LSigEmp}{\pi_2}{\TyS{\wcode{\tau'}}}$ &
-%       $o' = \LkgAdd{\LkgEmp}{\wcode{\tau'}}$ &
-%       \cellcolor[HTML]{ECF4FF}$\lsig = \LSigAdd{\LSigEmp}{\pi_2}{\TyS{\wcode{\tau}}}$ &
-%       $o = \LkgAdd{\LkgEmp}{\wcode{\tau}}$ \\
-%     \cellcolor[HTML]{ECF4FF}$\lsig'^+ = \LSigAdd{\lsig'}{s'}{..}$ &
-%       $o'^+ = \LkgAdd{o'}{t}$ &
-%       \cellcolor[HTML]{ECF4FF}$\lsig^+ = \LSigAdd{\lsig}{s}{..}$ &
-%       $o^+ = \LkgAdd{o}{\lkgproj{2}{o'^+}}$ \\
-%     \multicolumn{2}{l}{$A = \TySigma{\cU}{\El{\pi_2} \to \El{\pi_2}} $} &
-%       \multicolumn{2}{l}{$t = \goodTerm{\cdot , A}{\snd{\pi_2}}{\El{\fst{\pi_2}} \to \El{\fst{\pi_2}} }$} \\
-%     $\goodTerm{\cdot , \TyTkg{\lsig'}}{s'}{\sub{A}{\pi_1}}$ &
-%        &
-%       $\goodTerm{\cdot , \TyTkg{\lsig}}{s}{\sub{A}{\pi_1}}$ &
-
-%     \end{tabular}
-%   \caption{W-type encoding of "tm \{unit, app\}"}
-%   \label{fig:Wtype-example}
-% \end{figure}
-
-% We illustrate one concrete example in \cref{fig:Wtype-example} of a "tm" with only two constructors, 
-% also as a review for how W-type encodes a given inductive type. The key insight 
-% is, to encode one constructor into a pair $(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})$, the non-recursive part will be stored in $\wsigproj{i}{1}{\wsig}$ 
-% while the ``arity'' of the  recursive part stores in $\wsigproj{i}{2}{\wsig}$. 
-% For example, $"app" : ("tm" \to "tm" \to "tm")$ 
-% $\cong ((\bbtwo \to "tm") \to "tm")$, saying the ``arity'' of the recursive part is $\bbtwo$ 
-% (or boolean type, as a type with two elements). Similarly, the ``arity'' of recursive part of "unit" is 0. 
-% Both constructors don't store other information thus $\bbone$. We hightlight $\tau$ as the 
-% signature of our "tm" and showcase the meta-theoretic encoding of "unit" and "app~unit~unit".
-$"elim-"\bbzero$ indicates the elimination of the bottom type $\bbzero$ in the context.
-Generally speaking, the expressiveness of W-types comes from the ``arity'', as it encodes the case when recursive structure stores as the output of a function, mirroring the strict positivity of Inductive type.
-
-Rules \ruleref{tm/wsup} and \ruleref{tm/wrec} require access to~$\wsig$:
-the W-type is exhaustively generated by its constructors, and
-its elimination must exhaustively handle all the constructors in
-its signature.
-
-In contrast, $\wsig$ should be hidden from the typing context of any
-term that does not invoke $\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
-the term can be reused---without being rechecked---for a different
-W-type signature~$\wsig'$ that extends~$\wsig$ with additional
-constructors.
-Moreover, the typing of the term should be made parametric to
-the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
-so that the term can be reused---without being rechecked---when those fields
-are overridden to support the extended signature~$\wsig'$.
-%
-Such abstraction required by family polymorphism is supported via linkages,
-which we discuss next.
-
-\noindentparagraph{Family polymorphism via linkages.}
-
-Family polymorphism requires late binding.
-In \TT, families are expressed through linkages, and late binding of
-field references is achieved by requiring that typing be polymorphic to
-the definition of that field.
-
-Well-formed \emph{linkage signatures} have the judgment form
-$\goodSig{\Gm}{\lsig}{n}$.
-As rules \ruleref{lsig/empty} and \ruleref{lsig/add} show, a linkage
-signature is a list of $n$~types.
-%
-Well-formed \emph{linkages} have the judgment form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
-where $\TyLkg{\lsig}$ is a type formed by~$\lsig$.
-As rules \ruleref{l/empty} and \ruleref{l/add} show, a linkage is a list
-of $n$~terms, each representing a field of the family modeled by the linkage.
-
-In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ is
-responsible for late binding.
-Here, $A$ abstracts the context of the current field~$t$, controlling
-how the types of the fields prior to~$t$ are exposed to the typing of~$t$.
-Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
-the premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ makes clear that
-the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
-rather than \emph{existentially} quantified as is in \ruleref{tm/pair}---%
-over how the fields in $t$'s context are defined.
-Late binding enables reuse.
-A different linkage~$\ell'$ that overrides fields in $t$'s context (and thus models a derived family)
-can reuse~$t$---without retyping it---by first
-projecting~$t$ from $\LkgAdd{\lkg}{t}$ and then appending it to~$\lkg'$.
-%\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.
-
-%recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$
-The type~$A$ in \ruleref{l/add} and \ruleref{lsig/add},
-abstracting the types of the prior fields, does not necessarily contain the same types as those recorded by~$\TyLkg{\lsig}$,
-%\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
-because a field defined as the code $\wcode{\wsig}$ of a W-type has to
-expose different types to different fields that come after it.
-Later fields that invoke
-$\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
-should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
-and \ruleref{tm/wrec} stipulate.
-By contrast, $\wsig$ should be hidden
-from all other fields, so that they can be reused
-in a different context where the W-type signature~$\wsig$ is replaced by an extended one~$\wsig'$.
-
-\definecolor{A-color}{HTML}{E7EFC5}
-\definecolor{T-color}{HTML}{FFF1D0}
-\definecolor{t-color}{HTML}{FAE0E4}
-
-\input{stlc-linkage-typing-named.tex}
-
-We use \cref{fig:stlc-linkage-typing} to illustrate.
-On the left is code excerpted from \cref{fig:stlc-mechanized}, and on the right
-is how the corresponding fields are modeled and typed in \TT.
-%
-$\lkg_i$ is the linkage that adds the $i$-th field~$t_i$ with the typing
-$\goodTerm{%
-\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i}(-0.02,-0.10)(-0.02,0.28)
-\mathit{self}_i : A_i%
-\tikzmarkend{A-i}%
-}{%
-\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i}(-0.02,-0.10)(-0.02,0.28)%
-t_i%
-\tikzmarkend{t-i}%
-}{%
-\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i}(-0.02,-0.10)(-0.02,0.28)%
-T_i%
-\tikzmarkend{T-i}%
-}$.
-%
-The context types $A_i$ are a dependent tuple type, but for readability,
-we write them as dependent record types that give labels to the fields.
-%we use a notation with explicit names
-%in this typing judgment:
-%we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
-%and write its type~$A_i$, which is a dependent tuple type, as a
-%dependent record type with labels.
-%
-\begin{itemize}
-% [leftmargin=4ex,itemsep=.5ex]
-  [labelsep=*,leftmargin=1pc,itemsep=3pt]
-
-\item 
-The first field $t_1$ is defined as $\wcode{\wsig_{\texttt{tm}}}$
-and given the singleton type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$, where $\wsig_{\texttt{tm}}$
-is the W-type signature for \lsti{tm}.
-This typing is recorded by $\lsig_1$ and is thus available in all $\lsig_i$'s.
-
-\item 
-The next four fields model the four constructors of~\lsti{tm}.
-Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}$
-and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
-the typing context containing the first field~\lsti{tm}.
-The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
-this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
-So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
-as required by rule \ruleref{tm/wsup}.
-
-\item
-Similarly, $\wsig_{\texttt{tm}}$ is exposed in the typing context of
-$t_{10}$, which models the recursive function \lsti{subst} by invoking
-the recursor $\Wrec{\wsig_{\texttt{tm}}}{\cDot}{\cDot}$.
-%
-Like \lsti{subst}, partial recursors in \Lang are axiomatized by the plugin
-(\cref{sec:latebind,sec:prec}), and they can similarly be defined in \TT
-using $\texttt{Wrec}$.
-
-\item
-By contrast, $\wsig_{\texttt{tm}}$ is hidden from the typing of all other fields.
-Their typing should depend on the knowledge that \lsti!tm! has type~$\cU$,
-rather than \TyS{\wcode{\wsig_{\texttt{tm}}}},
-so that they can be reused in a context where
-\lsti!tm! is defined as $\wcode{\wsig'_{\texttt{tm}}}$,
-where $\wsig'_{\texttt{tm}}$ extends~$\wsig_{\texttt{tm}}$ with additional constructors.
-%
-For example, in \cref{fig:stlc-linkage-typing}, the typing of the case handlers of
-\lsti{subst} (e.g., $t_6$) is oblivious to the definition of \lsti!tm!---it sees only $\text{\lsti!tm!}:\cU$---so
-the case handlers can be reused by a linkage modeling \lsti{STLCFix}.
-
-\end{itemize}
-
-The third premise
-$\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
-in rules \ruleref{l/add} and \ruleref{lsig/add} is
-responsible for hiding W-type signatures.
-Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
-that packages the types of all fields preceding the current field~$t$,
-and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
-$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new tuple of type~$A$
-that hides W-type signatures behind $\cU$, if necessary.
-%
-It is straightforward to find the~$\seal$ that fits the bill, though this process
-is not automated in \TT.
-In particular, when no hiding is needed, $\seal$ is $\var{0}$
-%\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right?
-Or you meant to say it is identity ``sealing''?}\YZreply{Yes, I meant identity sealing.
-I cannot say 'sealing' because the paper currently does not use the word sealing anywhere.}\EDJreply{What about just remove the parenthesis and ``the identity''}
-and $A$ is $\TyTkg{\lsig}$.
-In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are $\var{0}$,
-and $\seal_6$ is from
-\begingroup
-\fontsize{9}{9}\selectfont%
-$\big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$
-\endgroup
-to
-\begingroup
-\fontsize{9}{9}\selectfont%
-$\big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$%
-\endgroup
-.
-
-When a family is concluded (e.g., \lsti!End!~\lsti!STLC!),
-a linkage~$\lkg$ containing all the fields is available
-(e.g., $\lkg_{53}$ in \cref{fig:stlc-linkage-typing}).
-Fields of the family can then be accessed by projecting them out of the tuple
-$\Tkg{\lkg}$.
-
-\noindentparagraph{Linkage transformers.}
-
-Inheritance and code reuse can already be expressed through the projection
-of fields out of linkages and their inclusion into new linkages.
-To make common patterns of linkage manipulations more convenient,
-\TT provides a ``library'' of \emph{linkage transformers}, whose well-formedness judgments
-have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$. The idea is that
-applying~$h$ to a linkage of type~$\TyLkg{\lsig_1}$ yields a linkage of
-type~$\TyLkg{\lsig_2}$ (\ruleref{tm/inh}).
-%Linkage transformers are second-class; the only way to use them is to
-%apply them to linkages to form new linkages.
-
-Derived families can be modeled as linkage transformers inductively
-constructed from the introduction forms
-$\small\InhId$, $\small\InhExt{h}{x.t}$, $\small\InhOv{h}{x.t}$, $\small\InhInh{h}$, etc.
-\cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
-\ruleref{tmeq/ov/beta}.
-It states that applying the transformer $\small\InhOv{h}{\mathit{self}_2.t_2}$ to
-a linkage of form $\LkgAdd{\lkg}{\mathit{self}_1.t_1}$
-overrides the linkage's last field~$t_1$ with~$t_2$.
-For instance, the construction below shows that 
-$\small\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$
-is used as the first step in creating a linkage transformer modeling a derived family that
-overrides $\wsig_{\texttt{tm}}$ with an extended
-signature~$\wsig'_{\texttt{tm}}$:
-\vspace{.9ex}
-
-\noindent
-\input{stlcfix-linkage-typing.tex}
-
-\noindent
-\ifreport \cref{sec:lt-in-action}
-\else A supplemental appendix
-\fi
-sketches how the other introduction forms of linkage transformers
-can be used to model the construction of a derived family as a linkage transformer.
-
-\noindentparagraph{The complete formalization.}
-
-% The definitive version containing all the rules in \TT is stated in
-% terms of a meta\-language supporting quotient inductive-inductive types
-% (QIITs)~\cite{nordvall2010inductive,altkap2016}.
-% QIITs support equality constructors, which facilitate the expression of
-% conversion rules (e.g., \ruleref{tmeq/sub/id}).
-% The formalization is available as a supplemental file in Agda syntax,
-% though it is not checked by Agda because Agda does not support QIITs
-% natively.
-% An effort to mechanize the formalization is underway, using Agda's
-% \lsti{REWRITE} pragma to replace the limited uses of equality reflection.
-
-The definitive version containing all the rules in \TT is stated in
-\ifreport \cref{sec:complete-fmltt}.
-\else a supplemental appendix.
-\fi
-using de Bruijn indices and explicit substitution.
-An effort to mechanize the formalization is underway, by following  \citet{altkap2016} 
-to use quotient inductive-inductive types as 
-the logical framework to state the syntax and using Agda's
-\lsti{REWRITE} pragma to replace the limited uses of equality reflection appeared in 
-an extensional meta\-logic.
-
-%a \textbf{pen-and-paper} formalization
-%assuming the facility of quotient inductive-inductive
-%types~\cite{nordvall2010inductive,altkap2016} and extensional equality
-%(equality reflection), available as a supplemental file in a
-%\textbf{(pseudo-)Agda syntax}. Our formalization is directly following
-%Agda syntax and type-checking (equality-checking) is done manually,
-%leaving traces in the comments. We are currently working on mechanizing
-%it and relying on Agda's rewriting mechanism to replace our limited
-%usage of equality reflection. 
-%
-% We also typeset a more complete set of rules than the space allows and
-% present it in
-% \ifreport \cref{sec:complete-fmltt}.
-% \else a supplemental appendix.
-% \fi
-
-\input{cc-condensed.tex}
-
 \ifShowOldWriting

### 2023-03-19 metatheory-condensed.tex
relocate old writings

@@
 \input{cc-condensed.tex}
-
-\ifShowOldWriting
-\newpage
-
-To formalize our metatheory, we follow the formulation of~\citet{altkap2016}, which provides a declarative and intrinsic style of the predicative \textit{Martin-Lof Type Theory} (MLTT)~\cite{martin1982constructive},
-using \textbf{Quotient Inductive Inductive Type}(QIIT)\footnote{Inductive Inductive Type(IIT) is a generalization of mutual inductive type where (indexed-)types \mintinline{Coq}{A : Type, B : A -> Type} can be defined mutually.
-QIIT further enhances IIT by allowing equality constructor (i.e.
-mathematical quotient).}, \textbf{explicit substitution} and \textbf{debruijn indices} as the main tool.
-Generally speaking, we \textbf{use each QIIT type to represent each kind of judgements, and thus a term will represent a derivation of a judgement}.
-The classical formulation of dependent type theory requires a lot of \textit{presupposition} and quotient afterwards, causing a lot of duplicacy.
-Both issues can be concisely solved by using QIIT expressing intrinsically typed syntax instead.
-What's more, working \textbf{in a type-theoretical setting} using QIIT makes checking of the pen-and-paper proof easier.
-However, we will follow conventional notation and thus the reader might not notice the usage of QIIT explicitly.\EDJ{We might be able to remove QIIT in the main text totally.
-Let's try to do it later.}
-
-$$
-\Rule[name=Subst]
-{\goodType{\Delta}{T}{} 
-  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodType{\Gamma}{T[\gamma]}{}},
-\Rule[]
-{\goodTerm{\Delta}{t}{T}
-  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
-{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
-\quad
-\Rule[name=Dbj Indices]
-{}
-{\goodSub{\Gamma, A}{p^1 \gamma}{\Gamma}},
-\Rule[]
-{}
-{\goodSub{\Gamma, A}{"var"_0}{A[p^1]}}
-$$
-
-% Explicit substitution
-Instead of meta-level substitution, we use \textbf{explicit substitution}, i.e., substitution itself is a expression in the program and will be done in the runtime.
-Thus substitution has its own judgement $\goodSub{\Gamma}{\gamma}{\Delta}$.
-We can apply substitution to types and terms\footnote{The type judgement is the form $\goodType{\Gamma}{T}{}$.
-The term judgement is the form $\goodTerm{\Gamma}{t}{T}$} in MLTT via \ruleref{Subst}.
-Note how substitution will change the context---we suggest intuitvely understanding context as holes of the term/types and substitution as filling those holes, where the substitution itself has holes.
-After filling the holes, then the new context is about the holes of the substitution.
-We have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
-% Debruijn Indices
-Contrary to named variables, \textbf{debruijn indices} refer to the variable in the context like operating a stack.
-We will use "var"$_i$ indicating the $i$-th variable (from right to left, $i$ starts with 0) in the context.
-On the other hand, we can also get the rest part of the stack, or ``subpart of "id"'' using $\goodSub{\Gamma, A}{p^1 \gamma}{\Gamma}$.
-For example, we can get the very first variable in the context $\Gamma, x : T \vdash x : T$ by $\goodTerm{\Gamma, T}{var_0}{T[p^1]}$.
-Note here $T$ is a type in $\Gamma$, but now shifted into $\goodType{\Gamma, T}{T[p^1]}{}$ using $p^1$.
-Similarly, to get the second and prior variables, we use "var"$_i$ and $p^{i+1}$ correspondingly.
-
-Now, we will mainly focus on the newly introduced typing rules, and illustrate their connection to the plugin implementation.
-For those curious about the complete meta-theory formulation, we refer to our appendices or \citet{altkap2016, kaposi2017type} or maybe those not using QIIT \cite{coquand2018canonicity, sterling2019algebraic}.
-Our metatheory is formulated using QIIT in an extensional setting.
-
-$$
-\Rule[name=wsig/ept]
-{""}
-{\goodWSig{\Gm}{\WSigEmp}{0}}
-\quad 
-\Rule[name=wsigeq/ept/eta]
-{\goodWSig{\Gm}{\wsig}{0}}
-{\goodWSig{\Gm}{\wsig \equiv \WSigEmp}{0}}
-\quad
-\Rule[name=wsig/add]
-{\goodWSig{\Gm}{\wsig}{n}
-  \quad \goodType{\Gm}{A}{}
-  \quad \goodType{\Gm, A}{B}{}}
-{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}}
-\quad
-\Rule[name=wsig/universe]
-{\goodWSig{\Gm}{\wsig}{n}}
-{\goodType{\Gm}{\TyWSingle {\wsig}}{}
-\\ \goodTerm{\Gm}{\wsingle{\wsig}}{\TyWSingle{\wsig}}
-}
-$$
-$$
-\Rule[name=ind/cstr]
-{{\goodWSig{\Gm}{\wsig}{n}}
-  \\ \goodTerm{\Gm}{a}{\wsigproj{j}{1}{\wsig}}
-  \\ \goodTerm{\Gm, \sub{\wsigproj{j}{2}{\wsig}}{(id, a)}}{b}{\El{(\wcode{\wsig})}}}
-{\goodTerm{\Gm}{\Wsup{}{T}{a}{b}}{\El{(\wcode{\wsig})}} }
-$$
-$$
-\Rule[name=ty/casesig]
-{\goodType{\Gm}{A}{}
-\quad \goodType{\Gm, A}{B}{}
-\quad \goodType{\Gm}{R}{}
-}
-{\goodType{\Gm}{\CaseSig{A}{B}{R} \equiv (\TyPi {A} {(\TyPi {(\TyPi {B} {(\sub{R}{ \SubstComp {\pi_1}{\pi_1}})})} {\sub{R}{\SubstComp{\pi_1}{\pi_1}}})} )}{}}
-\quad 
-\Rule[name=Hdlers]
-{\goodWSig{\Gamma}{\tau}{n}
-\quad \goodType{\Gamma}{R}{}
-}
-{\goodSig{\Gamma}{"RecSig"~\tau~R}{n}}
-$$
-$$
-\Rule[name=reclkg/proj]
-{ \goodWSig{\Gm}{\wsig}{N}
-\\ \goodTerm{\Gm}{o}{\TyLkg{(\RecSig{\wsig}{R})}}
-\\ j < N
-}
-{\goodTerm{\Gm}{\Recproj{j}{o}}{\sub{(\CaseSig{(\pi_1^j \wsig)}{(\pi_2^j \wsig)}{R})}{\pi_1}}}
-$$
-$$
-\Rule[name=tm/wrec]
-{ \goodWSig{\Gm}{\wsig}{N}
-\\ \goodType{\Gm}{R}{}
-\\ \goodTerm{\Gm}{r}{\TyLkg{(\RecSig{\wsig}{R})}}
-\\ \goodTerm{\Gm}{w}{\El{(\wcode{\wsig})}}
-}
-{\goodTerm{\Gm}{\Wrec{\wsig}{w}{r}}{R}}
-\YZ{I suppose Rec-Constr is only for exhaustivity checking?
-If so, then would it be more intuitive to not include the w in the term?
-And I suppose the programmer will want to use sealing to hide its type, or otherwise it would fail to normalize for lack of a beta rule?}\EDJreply{ 1.
-Yes.
-But not include w will cause $"Wrec"~r$ be a function type.
-we are not supposed to add a new intro-rule for function type(or for any other type), this is bad meta-theory design.
-\\
-  2.
-It is actually the other way around -- if things are abstracted, then no beta rule can apply.
-More concretely, $T : \bW\tau$ is abstracted into $T' : \cU$, then this later one we cannot use $"Wrec"$ on it for sure.
-But once we have $\{T : \bW\tau, "rec" : \cL(RecSig~\tau)\}$ we are okay, and this, after wrap "rec" with an lambda abstraction, we will have $\cC\{T : \bW\tau, "rec" : \cL(RecSig~\tau~R), "recf" : T -> R\}$ as compilation type.
-Then with sealing abstraction, we will have $\Sigma(T : U, \Sigma ("rec" : ?, \Sigma("ref" : T -> R)))$
-}\YZreply{
-  What happens if the programmer forgets to seal "rec"? I suppose rec
-  could still be inherited into a derived family. But it would not be
-  useful if the inductive type is extended there, because rec is
-  applicable only to the inductive type defined in the base family.
-}\EDJreply{
-  You are right.
-Since "rec" has type "\cL(RecSig~\tau~R)", after inheritance, it is basically the same type "rec : \cL(RecSig~\tau[k]~R[k])" for some "k" stands for the upcasting.
-but the most important $\tau[k]$ has an  unchanged shape (like the number of constructors) so it is still a recursor for inductive type of signature $\tau$.
-Once the parent inductive type $\bW\tau$ is extended $\bW\tau_2$ in the children family, with enriched signature $\tau_2$, then because of the signature inconsistency with $\tau[k]$, this old recursor is not useful for this new inductive type.
-\\
-  In other words, this recursor linkage "rec" can be used for any inductive type with signature $\tau$.
-}
-%\YZreply{
-%  So there IS a beta rule for recursor (but you don't show it), right?
-%}\EDJreply{
-%  Yes. 
-%  % https://github.com/DKXXXL/Redstone/blob/46c2420982e1fac50138924cdd7e152786cd38c8/Sketch/Extensible-Module-1/Syntax.agda#L589 \\ 
-%  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$. \\
-%  We also have canonicity (i.e. a closed term of boolean reduces to true or false), so there is definitely a beta rule for inductive type.
-%}\YZreply{I'd add typeset this beta rule}\EDJreply{I just realize you use the words normalize.
-I will say currently this system cannot ``reduce'' any open term.
-But can only ``reduce'' close term.
-\\
-%When the programmer program on this meta-theory, they don't need canonicity or normalization at all (they write out a chain of  judgemental equality by themselves).
-If they program on a proof assistant, then our proof assistant will need normalization to do type checking.
-%
-$$
-$$
-\Rule[name=tmeq/rec/beta]
-{\goodTerm{\Gamma}{h}{\cL(\RecSig{\tau}{R})} 
-\\ \goodTerm{\Gamma}{T}{\TyWSingle{\tau}}}
-{\goodTerm{\Gamma}{\Wrec{\wsig}{h}{(\Wsup{}{\wsig}{a}{b})} \equiv 
-(\app{((\app{(\Recproj{j}{h})} )[("id", a)])} )[("id", \lam{(\Wrec{\wsig}{h[\pi_1]}{b})})]
-}{R}}\EDJ{"app" and "lam" are lambda application and abstraction}
-$$
-
-We start with the formulation of (extensible) inductive type, which is close to W-type (\ruleref{WSig-Add} requires a pair of types $A \vdash B$ as well), but with mulitple constructors.
-Thus we have a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$ indicating signature of inductive type by its constructors.
-Unconventionally, a signature can induce a "universe" of types $\bW\tau$---in other words, $T : \bW\tau$ will be the inductive type instead of $\bW\tau$ itself.
-"Wsup" is used to construct a specific term of a given inductive type.
-
-% When programming via meta-theory, we usually seal the inductive type, only expose its constructors (as a mundane function member) and hide the eliminator by abstracting $T : \bW\tau$ into $"W"~T : \cU$.
-Then the following fields---those only rely on this interface with $\cU$---can be inherited to other contexts (e.g.
-using enriched inductive type), as long as the targeting contexts can be abstracted into this {interface} again.
-
-Similar to plugin implementation, we delegate handlers (and their reuse) for recursion to linkages (and their inheritance).
-We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers for all constructors.
-\ruleref{Rec-Constr} uses "RecSig" to achieve exhaustiveness checking on elimination.
-The partial recursors and the computational axioms are treated as normal fields as expected, since we know they are provable in the plugin implementation.
-
-To introduce the concrete programming on inductive type in metatheory, we need to first introduce the formulation of families.
-
-%\YZ{
-%    Is extensionality in conflict with the claim in §3 that "subst tm_unit x t" and "tm_unit" are
-%    not definitionally equal but propositionally equal in family STLC?
-%}\EDJreply{No, TLDR: it is a different level. \\ 
-%There are two type theories, one is Agda-like (= Latex = Conventional math), the other is the one formulated in QIIT.
-This sentence is saying that the Agda-like one is extensional MLTT (i.e.
-we have function extensionality and etc here in Latex writing directly).
-\\ But the one you are asking about is in the metatheory in QIIT: if we have an identity type $Id x y$ (propositionally equal) can we have a judgemetnal equality for the two terms in QIIT $x \equiv y$.
-\\
-%Apparently, we don't setup our QIIT formulation in this way (I thought we need, but now I don't think we need it. Our QIIT formulation is purely in intensional MLTT now).\\ 
-%But even if we do setup like that, I will not say a conflict is happening (because it is to our advantage if we have judgemental equal in metatheory).
-I will say it is a bit off as to how Coq is formulated (Coq is closer to an intensional MLTT, so no judgemental equal only propositional one).
-%}
-
-% linkage, signature and compilation and seal
-$$
-\Rule[name=lsig/ept]
-{""}
-{\goodSig{\Gm}{\LSigEmp}{0}}
-\quad
-\Rule[name=ty/lkg]
-{\goodSig{\Gm}{\lsig}{n}}
-{\goodType{\Gm}{\TyLkg{\lsig}}{}
-\\ \goodType{\Gm}{\TyTkg{\lsig}}{}}
-\quad
-\Rule[name=lnkg/pkg]
-{ \goodTerm{\Gm}{o}{\TyLkg {\lsig}} 
-}
-{
-  \goodTerm{\Gm}{\Tkg {o}}{\TyTkg{\lsig}}
-}
-$$
-$$
-\Rule[name=lsig/add]
-{\goodSig{\Gm}{\lsig}{n} 
- \quad \goodType{\Gm}{A}{}
- \quad \goodSeal{\Gm}{\seal}{\lsig}{A}
- \quad \goodType{\Gm, A}{T}{}}
-{\goodSig{\Gm}{(\LSigAdd {\lsig}{\seal}{T})}{n+1}}
-\quad
-\Biggl( \boxed{\goodSeal{\Gamma}{s}{\sigma}{A}} 
-:= \boxed{\goodTerm{\Gamma, \cC\sigma}{s}{A[p^1]}} \Biggr)
-$$
-$$
-\Rule[name=lnkg/ept]
-{""}
-{\goodTerm{\Gm}{\LkgEmp}{\TyLkg {\LSigEmp}}  
-}
-\quad
-\Rule[name=lkg/add]
-{ \goodTerm{\Gm}{o}{\TyLkg {\lsig}} 
-\\ \goodType{\Gm}{A}{}
-\\  \goodSeal{\Gm}{\seal}{\lsig}{A} 
-\\  \boxed{\goodTerm{\Gm, A}{t}{T}}
-}
-{\goodTerm{\Gm}{(\LkgAdd {o} {t})}{\TyLkg{\LSigAdd {\lsig} {\seal} {T}}}
-}
-$$
-$$
-\Rule[name=lkg/proj]
-{\goodTerm{\Gm}{o}{\TyLkg{\lsig}}}
-{\goodTerm{\Gm}{\lkgproj{1} {o}}{\TyLkg {(\lsigproj{1} {\lsig})}}
-\\ \boxed{\goodTerm{\Gm, \lsigprojT {\lsig}}{\lkgproj{2}{o}}{\lsigproj{2}{\lsig}}}
-}
-$$
-
-We start with signature (in a brand new judgement $\goodSig{\Gamma}{\sigma}{n}$).
-A signature is a list of types contained by a \textit{linkage}\footnote{Following the convention by \citet{zm2017}, we name the meta-theoretical correspondents of \textit{family} as \textit{linkage}.}, tracking the length as superscript.
-We will use $\{a : T_1, b : T_2, ..\}$ as a shorthand for signature with type $T_1, T_2, ..$ for readability.
-
-\ruleref{lkg/proj} indicates how linkage captures the idea of \textit{code reuse}, as $\lkgproj{2}{-}$ will lead to the boxed term that is parametric on the abstracted interface of linkage $\lsigprojT{\lsig}$ .
-To code reuse, we simply project out a field from the parent and add back to the children family.
-\newpage 
-Adding new types into signature \ruleref{Sig Add} requires a \textit{sealing judgement} $\goodSeal{\Gamma}{f}{\sigma}{A}$, which is an alias of a term judgement.
-Sealing is corresponding \textit{to indicating whether each field is overridable(extensible) or non-overridable (thus definitionally exposed) in the Coq Plugin}.
-Concretely, we encode our STLC example to show the motivation of the sealing.
-%\YZ{Is sealing used only for late-binding the signatures of inductive types?
-Do I understand correctly that late binding of other kinds of fields (like subst) is achieved by the boxed premise in rule "LNKG ADD"?}\EDJreply{All late-bound are achieved by ``selective-abstraction'' imposed by sealing.
-For example, recursion will be consequently abstracted because of the abstraction of the inductive type it recurse on.
-%}\EDJreply{Sorry, I see what you mean.
-Yes.
-the box part is important for late-binding.
-But the functionality of sealing is always there.
-They are both required to achieve overriding.
-A illustrative example is, if the type is $S(\bot)$, then even with the boxed premise, it is still not overridable because the type itself restricts all the possibilities.}\YZreply{
-%  By "selective abstraction", do you mean in the metatheory the programmer chooses what and how to abstract?
-%}\EDJreply{Yes.}
-
-% For example, if a field is overridable, we abstract that to its type; if a field has exposed its definition, we abstract it into the singleton type containing it instead.
-Sealing $\goodSeal{\Gamma}{f}{\sigma}{A}$ is useful when the surrounding interface is $\sigma$ and we want to construct the new field based on $A$.
-
-% We can now point out how "Overridable" and "pins" is modeled.
-Despite the fact we have a complicated dynamic of invisibility between overridable and normal fields in the plugin, every field can be elaborated into the (pseudo)-syntax form \\ \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...}  where every former field x,y,z,...
-hide its definition when type-check t, and remaining former fields have their definition exposed in t.
-% This above elaborated "hiding" syntax directly corresponding to the $f$ in $(\mu^+ ~o~\{f\}~t)$, where each fields decides the abstraction of the prior.
-
-\begin{figure}[!htb]
-  \lstset{
-      basicstyle=\fontsize{8}{8.5}\ttfamily,
-  % numbers=left,
-  }
-
-  \begin{minipage}{\textwidth}
-  \begin{multicols}{3}
-
-  \definecolor{codecomment-color}{HTML}{0DA3FF}
-
-  \begin{lstlisting}
-  Family STLC. 
-   FInductive ty (* no hiding *)
-
-   FInductive tm (* hiding ty *)
-
-   FRecursion subst
-     Case ... (* hiding ty tm *)
-     ...
-   End subst (* hiding ty *)
-
-   Field t1 := (tvar 0) 
-   (* hiding ty tm subst *)
-
-   Field t2 := (tapp t1 t1)
-   (* hiding ty tm subst *)
-
-  ...
-  \end{lstlisting}
-
-  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
-
-  \columnbreak
-  % \definecolor{codecomment-color}{HTML}{5D030F}
-
-  \begin{lstlisting}
-
-    (* s[@₁@]  : {} | {} *)
-    - o[@₁@]  = μ+ {} s[@₁@] [@$W\tau_1$@] : 𝕃{ty : [@$\TyS{W\tau_1}$@]}
-    (* s[@₂@]  : {ty : [@$\TyS{W\tau_1}$@]} | [ty : 𝕌; bool : ty; ...] *) 
-    - o[@₂@]  = μ+ o[@₁@] s[@₂@] [@$W\tau_2$@] : 𝕃{ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]}
-    (* s[@₃@][@₁@]  : {ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]}
-            | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
-    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] ([@μ@]+ ..) : 𝕃{ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig)}
-    (* s[@₃@] : {ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig)} 
-          | [ty : 𝕌; ... ; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig)] *)
-    - o[@₃@] = μ+ o[@₃@][@₁@] s[@₃@] (λ (Wrec ..)) : 𝕃{ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}
-
-    (* s[@₄@] : {ty : [@$\TyS{W\tau_1}$@]; ... subst : tm → nat → tm → tm } 
-          | [.. ; tm : 𝕌; tvar : nat → tm; .. subst : tm → nat → tm → tm] *)
-    - o[@₄@] = μ+ o[@₃@] s[@₄@] [@("tvar"~0)@] : 𝕃{ty : [@$\TyS{W\tau_1}$@]; ... t[@₁@] : [@\TyS{"tvar"~0}@]}
-
-    (* s[@₅@] : {ty : [@$\TyS{W\tau_1}$@]; tm : [@$\TyS{W\tau_2}$@]; subst : 𝕃(RecSig); t1 : [@\TyS{"tvar"~0}@]} 
-          | [ty : 𝕌; tm : 𝕌; subst : 𝕃(RecSig); t1 : [@\TyS{"tvar"~0}@]] *)
-    - o[@₅@] = μ+ o[@₄@] s[@₅@] .. : 𝕃{ ... }
-  \end{lstlisting}
-
-  \columnbreak
-
-  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
-
-  \end{multicols}
-  \end{minipage}
-  \caption{Sealing and Linkage}\label{fig:sealing+linkage}
-  \end{figure}
-
-In \cref{fig:sealing+linkage}, the "hiding" comment illustrates the late-bound prior fields for each fields.
-For "tm", "ty" is late-bound and thus extending "ty" won't cause disruption on the inheritance of "tm".
-For "subst" itself, the definition of "tm"  cannot hidden for the sake of exhaustiveness checking.
-However, for each case handler inside "subst", "tm" is late-bound and thus each handler is inheritable.
-
-Each hiding comment corresponds to the textual-aligned sealing derivation on RHS.
-For readability, RHS is informal---curly brackets indicate a signature (omitting sealing at type level) and square bracket indicate a nested $\Sigma$-type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.
-
-For "tm", "ty" is hidden and thus the sealing "s₂" will make "ty" into an arbitrary type "𝕌" and expose its constructor "bool", "arrow", and etc.
-Then when constructing "o₂", the newly added field "tm" as "⋆ : 𝕎τ₂" is constructed in the context where "ty : 𝕌", ignoring its original eliminator.
-(Because eliminator is only possible when "ty : 𝕎τ₁").
-On the other hand, for "subst", only "ty" is hidden and thus the sealing "s₃" leave "tm : 𝕎τ₂" untouched and "Wrec" becomes possible when constructing "o₃".
-
-Basically, the hidden prior fields---inductive type are sealed into arbitrary type with ``constructors only'', where the recursion is constructed without sealing the inductive type.
-Generally speaking, each extensible inductive type are hidden for all the other fields except for the recursion on itself.
-
-We choose to expose the definition of $t_1$ to all the following fields including $t_2$ by using singleton type $\TyS{-}$.
-If other fields, say $t_3$, want to hide it, just seal $t_1$ into type "tm"; and then overriding $t_1$ will not disrupt the inheritablity of $t_3$.
-But such overriding will definitely disrupt the inheritablity of $t_2$ as $s_5$ doesn't hide $t_1$.
-Generally speaking, any field that has a clear "hiding" semantic on the prior fields can be understood as sealing.
-All of our plugin commands have clear "hiding" semantics.
-
-% Intuitively, sealing corresponds to the following surface syntax\\
-% \mintinline{Coq}{Field hiding {x, y, z ...} newfield := t...} where every former field "x","y","z",...
-hide its definition when type-check "t", and remaining former fields have their definition exposed for "t".
-Thus, "x","y","z", ...
-can be overridden but other fields have to stay the same if we want to inherit "t".
-In meta-theory, we control the sealing $f$ so that "x","y","z",...
-only has type exposed and remaining fields are definitionally exposed using singleton type.
-
-% The reader should feel this subtle difference compared to our current plugin implementation---$(\mu^+ ~o~\{f\}~t)$ reads as \textit{``$t$ is a new well-typed field when we abstract the \textbf{prior} fields using $f$''}.
-In other words, every field in meta-theory has the right to choose how they perceive and abstract the prior fields.
-While in the plugin, "Overridable" keyword will decide how the current defining field is percived by all the following fields.
-
-% Given a signature $\sigma$, we have the linkage type $\cL\sigma$ inhabited by linkages; and the \textit{compilation type} $\cC\sigma$ of the linkage type.
-The compilation type corresponds to how we compile a context into a \mintinline{Coq}{Module Type} in the plugin implementation---in metatheory, $\cC\sigma$ will be a nested sigma type.
-For example, $\cC\{a : \cU, b : a \to \bot \} = \Sigma(a : \cU).\Sigma (b : a \to \bot).
-\top$.
-
-Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) $\Sigma$-type instead of a linkage type.
-For example, $\TyLkg{\{a : S(\bot), b : a \to \bot\}}$ can only be sealed into $\TyTkg{\{a : \cU, b : a \to \bot\}}$ but never $\TyLkg{\{a : \cU, b : a \to \bot\}}$.
-Because the latter cannot be inhabitted---by projecting the second field of the $ t : \TyLkg{\{a : \cU, b : a \to \bot\}}$, we can get a proof of bottom type.
-For simplicity, we make sealing a term in the context of compilation type instead of linkage type.
-
-% The difference between compilation type and linkage types lies in their introduction rules.
-The compilation types are just syntactic sugar for sigma types.
-However, the intro rules for the linkage types gives it the power of late-binding (also the fundation of inheritance and overriding).
-More concretely, the boxed premise in \ruleref{Lnkg Add} is making $t$ parametric on $A$, contrary to how in sigma type, the $t$ will be about a concrete $\mathit{self} : A$.
-This parametricity makes $t$ inheritable and the earlier stuff about $A$ late-bound.
-
-% The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$.
-For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators).
-Then any later fields $\goodTerm{\Gamma, A}{t}{T}$ relies only on the \textit{interface} $A$ can be inherited to other contexts (e.g.
-using extended inductive type), as long as the targeting contexts can be abstracted into the surrounding \textit{interface} for $t$ (i.e.
-$\Gamma, A$)  as well.
-
-% The importance of the compilation type lies in its difference from the linkage type.
-First we have a singleton type $"s A" : S(T) \iff "A" \equiv T$, i.e.
-a type of one single term.
-We can see that we have a term $x : \cC\{A : S(\bot), B : A \to \bot \} \vdash h_1 : \cC\{A : \cU, B:A \to \bot \}$ which is equivalent to $x : \Sigma (A: S(\bot)).
-A \to \bot \vdash h_1 : \Sigma (A: \cU).
-A \to \bot$.
-This can be considered as using seal to do an abstraction on $A:S(\bot)$.
-
-% But we can never have a term $x : \cL\{a : S(\bot), b : a \to \bot \} \vdash h_2 : \cL\{a : \cU, b : a \to \bot \}$, as we know $\cL\{a : S(\bot), b : a \to \bot \}$ is inhabited
-% and that leads to a close term of $\cL\{a : \cU, b : a \to \bot \}$ due to the nature of late-binding of $\cL\sigma$---once we override $a$ with $\top$ we will have inconsistency using $b$.
-This example illustrates the programming difference between the sigma type and linkage type.
-
-% However, this form of abstracting a type is very important when programming in our metatheory---as later we will show inductive type of different constructors locates in different type (universe), and to make later fields inheritable we need to work on an interface of abstracted type.
-Thus the compilation type is indispensable.
-
-% Looking closely, the sealing judgement helps to model most machinaries about abstraction in plugin implementation.
-
-% Inductive type and recursor and handlers
-
-$$
-\Rule[name=Inh]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-\quad \goodTerm{\Gamma}{l}{\cL \sigma_1}
-}
-{\goodTerm{\Gamma}{("inh" \ h \ l)}{\cL \sigma_2}},
-\Rule[name=Inh-Id]
-{}
-{\goodInh{\Gamma}{"inhid"}{\sigma}{\sigma}},
-\Rule[name=Inh-Ext]
-{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
-  \quad \goodTerm{\Gamma, A_2}{t}{T}}
-{\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}
-$$
-
-% Inhertance judgement
-Currently we can achieve inheritance by projecting out the field, and then add back the projection into the children family.
-That means the formulation of linkage already captures the concept of \textit{code reuse}.
-
-However, in the plugin implementation, we use "Family"s to organize the code reuse.
-They are second class objects like Coq's "Module".
-What's more, a "Family" (with or without "extends" clause) can be considered as a standalone piece of code trying to \textit{inherit} an existent (but maybe empty) family.
-Thus we model them using \textit{inheritance judgement} $\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}$ instead of considering them as first class terms.
-Inheritance judgement is like ``library function'' upon linkage for achieving inheritance and overriding.
-The only way to use $h$ is by \ruleref{Inh} and transforming a linkage of signature $\sigma_1$ to another linkage of $\sigma_2$.
-Inheritance judgement can be inductively constructed via inheriting operations, overriding operations, extending operations and etc, corresponding to "Inherit", "Override" and "Extend" command  in our plugin implementation.
-
-In this way, we will consider the program in our metatheory similar to those vanilla MLTT program, except there are meta variables denoting inheirtance judgement (i.e.
-"Family"), of which the only usage is to apply \ruleref{Inh}.
-We don't yet consider non-trivial equality between derivations of inheritance judgements.
-
-\subsection{Syntactic Translation that removes Linkages}
-To further provide the intuition why our core calculus is sound, we provide a syntactic translation from our core calculus to the subpart where linkages are absent.
-This can justify the \textit{consistency} and \textit{canonicity} modulo the quirky (but apparent) formulation of our inductive type.
-To be complete, we also provide a consistency and canonicity proof for our complete core calculus in the appendix.
-
-Thanks to QIIT formulation, syntactic translation becomes a function between QIIT type, which by definition respects all the judgemental equalities.
-We will now sketch this translation $\denotesT{\_}$, which keeps most parts untouched except for the linkages related.
-We omit most of the inheritance translation.
-
-\begin{align*}
-  \text{We first define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{}  \text{ when } \goodCtx{\Gamma}{}  \text{ is well-formed and } n \in \mathbb{N} \\
-  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\TyTkg {-}} \text{ are mutually recursive, } \\
-  & \text{defined by induction on the signature length} \\  
-  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \\ 
-  &\text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \text{ defined} \iff \goodTerm{\Gamma}{\denotesT{\sigma}}{\Sigr{\Gamma}{n}} \\ 
-  {\Sigr{\Gamma}{(n+1)}} &= 
-    \Sigma~(\Sigr{\Gamma}{n})
-          ~(\Sigma~\cU
-                  ~(\Sigma~(\Pi(\TyTkg {("var"_0[p^1])})(El~"var"_0[p^1]))~\cU)) \\
-  \Sigr{\Gamma}{0} &= \top \\
-  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}}, \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ defines upon } \denotesT{\sigma} 
-  \text{ and inductively on the signature length} \\
-  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} = \top \quad
-      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
-  \denotesT{\goodType{\Gamma}{\TyTkg {\sigma}}{}} &= 
-    \Sigma~\denotesT{\TyTkg {(\lsigproj{1}{\sigma})}}~(El \ (\app{("pjr"^3~\denotesT{\sigma})})[(p^1, \app{(\fst{("pjr"^2~{\denotesT{\sigma}})})})]) \\
-  \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &=
-  \denotesT{\TyLkg{(\lsigproj{1}{\sigma})}} \times \Pi(El~(\fst{(\snd{\denotesT{\sigma}})}))(El~(\app{("pjr"^3~\denotesT{\sigma})} )) \\
-  &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
-  \denotesT{\goodSig{\Gamma}{\LSigAdd{\sigma}{f}{T}}{n+1}} &= (\denotesT{\sigma}, \codety{A}, \lam{f}, \lam{(\codety{T})}) \\ 
-  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\TyLkg{(\LSigAdd{\sigma}{s}{T})}}} &= (\denotesT{m}, \lam{t}) \\ 
-  \denotesT{\goodTerm{\Gamma}{\Tkg{m}}{\TyTkg {\sigma}}} & \text{ is defined upon } \denotesT{m} \text{ and inductively on the signature length} \\ 
-  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= 
-  (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{(\snd{\denotesT{m}})}, f = p_f\nu~\sigma \\ 
-  \denotesT{\goodTerm{\Gamma}{\Tkg {m}}{\TyTkg {\sigma}}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
-\end{align*}
-\begin{align*}
-  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \TyLkg{\sigma_1}}{\_}{\TyLkg{\sigma_2[p^1]}} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \TyLkg{\sigma_1}}{h}{\TyLkg{\sigma_2[p^1]}} \\
-  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\LSigAdd {\sigma_1} {s_1}{T})}{(\LSigAdd {\sigma_2} {s_2} {T[(p^1, \uparrow^s)]})}} &= \mu^+~(h[(p^1,\lkgproj{1}{"var"_0})])\\ &\quad \quad \{f_2[p^1]\}~(\lkgproj{2}{"var"_0})[(p^1, \uparrow^s[{p^1}^{\uparrow}])] \\
-\end{align*}
-
-\begin{figure}[!htb]
-  \lstset{
-      basicstyle=\fontsize{8}{8.5}\ttfamily,
-  % numbers=left,
-  }
-
-  \begin{minipage}{\textwidth}
-  \begin{multicols}{3}
-
-  \definecolor{codecomment-color}{HTML}{0DA3FF}
-
-  \begin{lstlisting}
-  Family STLC_bool extends STLC.
-
-   FInductive ty += ty_bool
-   FInductive tm += tm_true ..
-
-   FRecursion subst
-     (* Inherits all others *)
-
-     Case tm_true ...
-
-   End subst
-
-   Inherit t1
-
-  ...
-  \end{lstlisting}
-
-  % \makeline[0pt]{Parser-exmp-before-start}{Parser-exmp-before-end}[codecomment-color!50]
-
-  \columnbreak
-  % \definecolor{codecomment-color}{HTML}{5D030F}
-
-  \begin{lstlisting}
-
-  i[@₀@] = inh-id : Inh ⋅ {} {} 
-  i[@₁@] = inh-ov i[@₀@] (W [@τ₁@]') : Inh ⋅ {ty : 𝕊(W τ₁)} {ty : 𝕊(W τ₁')}
-  s[@₂@]' : {ty : 𝕊(W τ₁')} | [ty : 𝕌] 
-  i[@₂@] = inh-ov i₁ (W τ₂') 
-     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂') }
-  (* recall s[@₃₁@] : {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂)} 
-                | [ty : 𝕌;  ..] *)
-  s[@₃₁@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂')} | [ty : 𝕌; ..; tm_true : tm; ..]
-  ↑s[@₃@] : Tm ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..]) [ty : 𝕌; ..] 
-  (* ↑s[@₃@] map (the type of) s[@₃₁@]' to that of the sealing s[@₃₁@] before inheritance *)
-  ii[@₃₀₀@] = inh-id 
-           ⫶ Inh ( ⋅, [ty : 𝕌; ..; tm_true : tm; ..])
-              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
-              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]]
-  ii[@₃₀₁@] = inh-extend (inh-extend ii[@₃₀₀@] {p[@₁@]} .. ) {p[@₁@]} ..
-        : Inh ( ⋅, [ty : 𝕌; tm :𝕌; ..; tm_true : tm; ..]) 
-              (RecSig τ₂  tm)[p[@₀@], ↑s[@₃@]] (RecSig τ₂'  tm)
-  i[@₃₁@] = inh+inh i[@₂@] ↑s[@₃@] ii[@₃₀₁@] : 
-      Inh ⋅ {ty : 𝕊(W τ₁); ..; subst' : 𝕃(RecSig τ₂ tm)} 
-            {ty : 𝕊(W τ₁'); ..; subst' : 𝕃(RecSig τ₂' tm)}
-  s[@₃@]' : {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm) }
-      | [ty : 𝕌; .. ; tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm)] 
-  i[@₃@] = inh-ov i[@₃₁@] ..
-     : Inh ⋅ {ty : 𝕊(W τ₁); tm : 𝕊(W τ₂); subst' : 𝕃(RecSig τ₂ tm); subst : tm → nat → tm → tm}  
-             {ty : 𝕊(W τ₁'); tm : 𝕊(W τ₂'); subst' : 𝕃(RecSig τ₂' tm); subst : subst : tm → nat → tm → tm}
-  (* recall s₄ : {ty : 𝕊(W τ₁); .. subst : .. } 
-      | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..] *)
-  s[@₄@]' : {ty : 𝕊(W τ₁'); .. subst : .. } 
-        | [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
-  ↑s[@₄@] : Tm ( ⋅, [ty : 𝕌; .. ; tm : 𝕌; .. ; subst' : 𝕃(RecSig τ₂ tm);subst : ..]) 
-          [ty : 𝕌; .. ; tm : 𝕌; .. ;subst' : 𝕃(RecSig τ₂ tm); subst : ..]
-  i[@₄@] = inh-inh i[@₃@] ↑s[@₄@] 
-         : Inh ⋅ {ty : 𝕊(W τ₁); .. subst : ..; t1 : 𝕊(tm_var 0)} 
-                 {ty : 𝕊(W τ₁'); .. subst : ..; t1 : 𝕊(tm_var 0)} 
-  \end{lstlisting}
-
-  \columnbreak
-
-  % \makeline[.5\textwidth+9pt]{Parser-exmp-after-start}{Parser-exmp-after-end}[codecomment-color!50]
-
-  \end{multicols}
-  \end{minipage}
-  \caption{Inheritance Metatheory}\label{fig:inheritance}
-  \end{figure}
-
-\fi

### 2023-03-19 canonicitymodel.tex
Reply and add new comments.

@@
 Without exposing the reader to too many technical details, our meta\-language
-should be understood an instance of any of the above logical frameworks.
+should be understood an instance of any of the above logical frameworks---the difference is our formulation did quotient manually but if we use their framework, quotient is done automatically.
 %Our model works directly on the earlier introduced syntax for readability\footnote{We still
@@
 %\footnote{See nlab explanation: \href{https://ncatlab.org/nlab/show/canonical+form}{Canonical Form}}
-We can even argue that, if this theorem is proven in a computable meta\-logic, 
-then by the Curry–Howard correspondence, the canonicity theorem provides a big-step
-interpreter for closed terms of the boolean type.%\YZ{Can we say something about
+% We can even argue that, if this theorem is proven in a computable meta\-logic, 
+% then by the Curry–Howard correspondence, the canonicity theorem provides a big-step
+% interpreter for closed terms of the boolean type.
+%\YZ{Can we say something about
 %strong/weak normalization?}\EDJreply{
@@
 
-\begin{assumption}[Set-theoretic Universe Assumption] As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our ambient meta\-logic.
+\begin{assumption}[Set-theoretic Universe Assumption] We assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our ambient meta\-logic.
 \end{assumption}
 
-We can roughly consider each Grothendieck universe $\Set{i}$ as the $\Set{i}$ in Agda. 
+We can roughly consider each Grothendieck universe $\Set{i}$ as the $\Set{i}$ in Agda: 
 \begin{itemize}
@@
 %treatment of propositions---
-A proposition is encoded as a subsingleton data. For
+Proof-relevant model treats proposition in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}
+a proposition is encoded as a subsingleton data. For
 example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \
@@
 subsingleton sets, where the first one is either a singleton if we have a proof of
-$t \equiv \true$ or an empty set otherwise.
+$t \equiv \true $ or an empty set otherwise.
 In other words,
-$\star^1$ witnesses a proof that $t \equiv true$.
+$\star^1$ witnesses a proof that $t \equiv \true$.
 Here, $\star$, $\star^1$, $\star^2$ are just some arbitrary fixed elements.
@@
 Given the above model for MLTT, there should be a function ${\Pi}^c$ such
-that ${\Pi}^c({\denotesC{A}},{\denotesC{B}}) = \denotesC{\TyPi{A}{B}}$. We will
-use this function when defining the logical-relations model for the rest of \TT,
+that ${\Pi}^c({\denotesC{A}},{\denotesC{B}}) = \denotesC{\TyPi{A}{B}}$. 
 Type-theoretically speaking, this ${\Pi}^c$ uses the \emph{internal dependent
 function type} of
-the above model.
+the above model. We hope to
+use this function when defining the logical-relations model for the rest of \TT.
 However, such a function $\Pi^c$ is not yet possible because the definition
@@
 Furthermore, given $\Glued{S}$ (i.e., the syntax and its semantic interpretation),
-we use $\Gmodel{\Glued{S}}$ to mean the latter of the two.\YZ{why not just $\denotesC{S}$?}
+we use $\Gmodel{\Glued{S}}$ to mean the latter of the two.\YZ{why not just $\denotesC{S}$?}\EDJreply{Using superscript C is shorter and more precise -- we are not "reconstructing the C model from the syntax, but taken out the C model directly from the pair".
+Look at the canonicity model where superscript c is used, if change to $\denotesC{-}$ on them, it is not really well-founded recursion, technically speaking (only equivalent to one).}
 % These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.

### 2023-03-19 metatheory2.tex
Took a pass over appendix. A question was added.
edit
Reply and add new comments.

@@
 %substitution) will be helpful. 
-
+\EDJ{Why comment out this sentence?}
 The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
@@
 Universe levels address the size issue---it is unsound to have a set of all sets
-(or a universe of all types).
+(or a universe of all types)~\cite{hurkens1995simplification}\EDJ{Why remove this citation?}.
 The level of a universe specifies ``how large'' that universe is.
@@
 
-\subsection{Proof Relevant Logical Relation for Canonicity}
+\subsection{A Proof Relevant Logical-Relations Model for Canonicity}
+\label{sec:canonicity-model}
 \input{canonicitymodel.tex}

### 2023-03-19 syn-translate.tex
Reply and add new comments.

@@
   \Sigr{\Gamma}{0}{j} &= \top \\
-  \mathscr{P}(\lsig') &=  \top \quad \text{ when } \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{0}{j}}  \\
   \mathscr{P}(\lsig') &=  
     \Sigma~(\fst{\lsig'})~(\El{\app{"snd"^3~\lsig'}[(\SubstWeak{1}, \app{\fst{("snd"^2~{\lsig'})}})]}) \\ & \text{ when } \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{n+1}{j}} \\
+  \mathscr{P}(\lsig') &=  \top \quad \text{ when } \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{0}{j}}  \\
 \end{align*}
@@
 We omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
-However, when we mutually recursively define
-$\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}{}} $, and
-$\denotesT{\cP}$ above, we actually have to prove the substitution laws
+Note that, when we mutually recursively define the type
+${\Sigr{\Gamma}{n}{}}$ (of signatures $\denotesT{\goodSig{\Gamma}{t}{n}}$), and
+$\denotesT{\cP} = \mathscr{P}$ above, we actually have to prove the two substitution laws
 $\sub{(\TyTkg{\lsig})}{\gm}\equiv \TyTkg{\sub{\lsig}{\gm}}$ and $\sub{(\Sigr{\Delta}{n}{})}{\gm}

### 2023-03-19 canonicitymodel.tex
Phrasing.

@@
 is part of $\GTy{\!}{\!}$, $\GTm{\!}{\!}$, and $\GSub{\!}{\!}$.
-The interesting fact is that, given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$
+An interesting fact about them is : given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$
 and $\gm' \in \Gmodel{\Glued{\Gm}}(\gm)$, we can consider $(\gm, \gm')$ as an
 element of $\GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
-Thus, we consider the pair $(\gm,
-\gm')$ as $\Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
-% Note that, for arbitrary $\Glued{\Gm} \in \GCon{k}$, we can consider it as $\Glued{\gm}$ as an element of $\GSub{\Glued{\cdot}}{}$
+Thus, we can consider the pair $(\gm,
+\gm')$ as an element $\Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
 
@@
 we interpret linkage types using the canonicity model of sigma types (developed in the prior work).
-We use the inductive facility of the meta\-logic to justify our W-types.
+We use the inductive facility of the ambient meta\-logic to justify our W-types.
 

### 2023-03-19 syn-translate.tex
Phrasing.

--- 2023-03-18 syn-translate.tex
@@
 We omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
-However, when we mutually recursively define
-$\denotesT{\goodSig{\Gamma}{\_}{n}}$, ${\Sigr{\Gamma}{n}{}} $, and
-$\denotesT{\cP}$ above, we actually have to prove the substitution laws
+Note that, when we mutually recursively define the type
+${\Sigr{\Gamma}{n}{}}$ (of signatures $\denotesT{\goodSig{\Gamma}{t}{n}}$), and the function
+$\denotesT{\cP} = \mathscr{P}$ above, we actually have to prove the two substitution laws
 $\sub{(\TyTkg{\lsig})}{\gm}\equiv \TyTkg{\sub{\lsig}{\gm}}$ and $\sub{(\Sigr{\Delta}{n}{})}{\gm}

### 2023-03-21 metatheory2.tex
Working on cover letter.

@@
 binders and substitutions explicit.
+%
 %The change of the syntax formulation is important to the simplicity of our
@@
 \EDJ{Why comment out this sentence?}
+%
 The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
@@
 Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.\YZ{$\pi_1$ and $\pi_2$ should be added to the syntax figure}\EDJreply{I added it, it is right after sub/dbj/shift.}
+
+Consequently, function application changes to a equivalent formulation, and becomes a ``direct inverse'' of typing rule for function abstraction.
+For example, the named notation $\napp{f}{t}$ can be equivalently represented by $\sub{\app{f}}{\SubstExt{\SubstId}{t}}$.
 

### 2023-03-21 coqimpl.tex
move a redundant mention of mutually inductive types

@@
 Partial recursors are automatically generated only for non-indexed inductive types.
-The case studies in \cref{sec:casestudies} do not exercise these features.
+These features are not exercised by our case studies (\cref{sec:casestudies})
+but may be useful for modeling other languages.
 We believe that they do not pose conceptual challenges and can be
@@
 %
-Also limiting (and interesting for future research) is
+Also limiting---and interesting for future research---is
 that generated propositional equalities cannot be used by Coq to automatically

### 2023-03-21 coqimpl.tex
tweak in §5

@@
 %
-Also limiting---and interesting for future research---is
-that generated propositional equalities cannot be used by Coq to automatically
-reduce terms to definitionally equal forms, which causes inconveniences for
+Also interesting for future research is the support for automatically converting
+terms to propositionally equally typed forms using generated propositional equalities.
+The lack of this automation currently may cause inconveniences for
 developments using intrinsically typed syntax.

### 2023-03-21 metatheory-condensed.tex
comment
add more details about bold-P

@@
 Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
+We use $x$ and $\mathit{self}$ to denote variables.
 A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
@@
 
-Well-formed \emph{linkage signatures} have the judgment form
-$\goodSig{\Gm}{\lsig}{n}$.
+\emph{Linkage signatures} have the form $\goodSig{\Gm}{\lsig}{n}$.
 As rules \ruleref{lsig/empty} and \ruleref{lsig/add} show, a linkage
@@
 %
-Well-formed \emph{linkages} have the judgment form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
+\emph{Linkages} have the form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
 where $\TyLkg{\lsig}$ is a type formed by~$\lsig$.
@@
 
-In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ is
+In rules \ruleref{l/add} and \ruleref{lsig/add},
+the second premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ is
 responsible for late binding.
@@
 \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i}(-0.02,-0.10)(-0.02,0.28)
-\mathit{self}_i : A_i%
+\mathit{self} : A_i%
 \tikzmarkend{A-i}%
@@
 and given the singleton type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$, where $\wsig_{\texttt{tm}}$
-is the W-type signature for \lsti{tm}.
+is the W-type signature constructed earlier for the extensible inductive type~\lsti{tm}.
 This typing is recorded by $\lsig_1$ and is thus available in all $\lsig_i$'s.
@@
 \item
-Similarly, $\wsig_{\texttt{tm}}$ is exposed in the typing context of
+Likewise, $\wsig_{\texttt{tm}}$ is exposed in the typing context of
 $t_{10}$, which models the recursive function \lsti{subst} by invoking
@@
 
-The third premise
-$\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
-in rules \ruleref{l/add} and \ruleref{lsig/add} is
+In rules \ruleref{l/add} and \ruleref{lsig/add},
+the third premise $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$ is
 responsible for hiding W-type signatures.
-Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
-that packages the types of all fields preceding the current field~$t$,
-and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
-$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new tuple of type~$A$
+Here, $\TyTkg{\lsig}$ packages $\lsig$---recall that $\lsig$ contains the
+(self-parameterized) types of all the fields preceding the current
+field~$t$---into a dependent tuple type (\ruleref{tyeq/pk/add}).
+The term $\seal$ turns a tuple of type $\TyTkg{\lsig}$ into a new tuple of type~$A$
 that hides W-type signatures behind $\cU$, if necessary.
@@
 is not automated in \TT.
-In particular, when no hiding is needed, $\seal$ is $\var{0}$
+In particular, when no hiding is needed, $\seal$ is simply~$x$,
 %\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right?
@@
 and $A$ is $\TyTkg{\lsig}$.
-In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are $\var{0}$,
-and $\seal_6$ is from
+In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are simply $x$,
+and $\seal_6$ has the typing
 \begingroup
 \fontsize{9}{9}\selectfont%
-$\big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$
+\goodTerm{
+  \mathit{self} :
+  \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+}{
+  \seal_6
+}{
+  \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+}.
 \endgroup
-to
-\begingroup
-\fontsize{9}{9}\selectfont%
-$\big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]$%
-\endgroup
-.
 
@@
 $\Tkg{\lkg}$.
+As \ruleref{tmeq/pk/add} indicates, $\Tkg{\lkg}$ ties the recursive knot:
+it packages the linkage~$\lkg$ into a dependent tuple of type~$\TyTkg{\lsig}$,
+by instantiating the $\mathit{self}$ parameters.
 

### 2023-03-21 stlc-linkage-typing-named.tex
Done Changing to named variables.
Fix B in Wsig
W-types
bugfix in STLC encoding figure
tweaks to avoid margin overrun

--- 2022-11-11 stlc-linkage-typing.tex
@@
 &
-\goodTerm{\TyTkg{\lsig_0}}{\seal_1}{\sub{
+\goodTerm{x : \TyTkg{\lsig_0}}{\seal_1}{{
     \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-a}(-0.02,-0.10)(-0.05,0.28)
@@
     \tikzmarkend{A1-a}
-}{\SubstWeak{1}}}
+}{}}
 \\
@@
 &
-\goodTerm{\TyTkg{\lsig_1}}{\seal_2}{\sub{
+\goodTerm{x : \TyTkg{\lsig_1}}{\seal_2}{{
     \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-a}(-0.02,-0.10)(-0.05,0.28)
@@
     \tikzmarkend{A2-a}
-}{\SubstWeak{1}}}
+}{}}
 \\
@@
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-b}(-0.02,-0.10)(-0.02,0.28)
-    \Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}
+    \Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}
     \tikzmarkend{t2-b}
@@
 &
-\goodTerm{\TyTkg{\lsig_5}}{\seal_6}{\sub{
+\goodTerm{x : \TyTkg{\lsig_5}}{\seal_6}{{
     \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-a}(-0.02,-0.10)(-0.05,0.28)
@@
     \tikzmarkend{A6-a}
-}{\SubstWeak{1}}}
+}}
 \\
@@
 }{
-    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-b}(-0.02,-0.10)(-0.02,0.28)
-    \CaseSig{\top}{\bot}{
-        \fontsize{8.5}{10.2}\selectfont
-        \text{$\El{\mathit{self}.\texttt{tm}}\to\mathit{id}\to\El{\mathit{self}.\texttt{tm}}$}
-    }
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-b}(-0.05,-0.30)(-0.02,0.46)
+%   \CaseSig{\top}{\bot}{
+    \texttt{CaseTy}\left(\top,\bot,
+        \begin{minipage}{.15\textwidth}%
+        \vspace{-7.5pt}%
+        \fontsize{8.0}{9.2}\selectfont%
+        \begin{gather*}
+        \begin{array}{@{}l@{}}
+            \El{\mathit{self}.\texttt{tm}} \to
+            \\
+            \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
+        \end{array}
+        \end{gather*}
+        \end{minipage}
+    \right)
+%   }
     \tikzmarkend{T6-b}
@@
 &
-\goodTerm{\TyTkg{\lsig_9}}{\seal_{10}}{\sub{
+\goodTerm{x : \TyTkg{\lsig_9}}{\seal_{10}}{{
     \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-a}(-0.02,-0.10)(-0.05,0.28)%
@@
     \tikzmarkend{A10-a}
-}{\SubstWeak{1}}}
+}{}}
 \\
@@
 }{
-    \begin{minipage}{.20\textwidth}%
+    \begin{minipage}{.15\textwidth}%
     \vspace{-7.5pt}%
@@
     \begin{gather*}
-    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.26)(-0.02,0.40)
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.45)(-0.02,0.60)
     \begin{array}{@{}l@{}}
-        \El{\mathit{self}.\texttt{tm}} \to \El{\mathit{self}.\texttt{tm}}
+        \El{\mathit{self}.\texttt{tm}} \to
         \\
-        \to \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
+        \El{\mathit{self}.\texttt{tm}} \to
+        \\
+        \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
     \end{array}

### 2023-03-21 casestudies.tex
move a redundant mention of mutually inductive types
composing contradictory features

@@
 For example, composing \lsti{STLCIsorec} and \lsti{STLCProd}
-(\cref{fig:stlc-isorec-prod}) creates a need to extend \lsti{tysubst} to
+(\cref{fig:stlc-isorec-prod}) creates an obligation to extend \lsti{tysubst} to
 handle \lsti{ty_prod}, which the type-checker enforces.
+%
+Composing contradictory features (e.g., a fixpoints construct and a strong
+normalization theorem) would lead to unprovable proof obligations.
 
@@
 
-The current plugin implementation does not yet support
-mutually inductive types or mutual induction, which are
-useful for modeling languages much more complex than STLCs.
-However,
-we believe accommodating mutually inductive types does
-not pose theoretical challenges.
+%The current plugin implementation does not yet support
+%mutually inductive types or mutual induction, which are
+%useful for modeling languages much more complex than STLCs.
+%However,
+%we believe accommodating mutually inductive types does
+%not pose theoretical challenges.
 

### 2023-03-21 metatheory2.tex
add π1 π2 to MLTT syntax display
resolve some comments

@@
 %
+Using explicit substitutions obviates the need for special treatment of
+substitutions in the proofs, as substitutions are part of the syntax.
 %The change of the syntax formulation is important to the simplicity of our
@@
 %substitution) will be helpful. 
-\EDJ{Why comment out this sentence?}
 %
@@
 
-To simplify, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$.
-Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.\YZ{$\pi_1$ and $\pi_2$ should be added to the syntax figure}\EDJreply{I added it, it is right after sub/dbj/shift.}
+To simplify, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$. Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.
 
@@
 This rule applying substitution $\gm$ on the earlier part of the context.
-We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.\YZ{The shift rule has bugs: premise says n, conclusion says 1.}\EDJreply{Thanks, fixed.}
+We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.
 
@@
 Universe levels address the size issue---it is unsound to have a set of all sets
-(or a universe of all types)~\cite{hurkens1995simplification}\EDJ{Why remove this citation?}.
+(or a universe of all types)~\cite{hurkens1995simplification}.
 The level of a universe specifies ``how large'' that universe is.

### 2023-03-21 syn-translate.tex
is this right?
\setc

@@
   such that $\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}}$
-  \item and a function $\mathscr{P} : \{\lsig' : \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{n}{j}} \} \to \{ T : \goodType{\Gm}{T}{} \}$ map one syntax term of $\Sigr{\Gm}{n}{j}$ into a type.
+  \item and a function $\mathscr{P} : \setc{\lsig'}{ \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{n}{j}} } \to \setc{T}{ \goodType{\Gm}{T}{} }$ mapping a term of type $\Sigr{\Gm}{n}{j}$ into a type.
 \end{itemize}
@@
 \begin{align*}
-  \denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}} &\in \{t : \goodTerm{\Gm}{t}{\Sigr{\Gm}{n}{j}} \} \\
+  \denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}} &\in \setc{t}{ \goodTerm{\Gm}{t}{\Sigr{\Gm}{n}{j}} } \\
   {\goodType{\Gamma}{\TyTkg{\lsig}}{}} &= \mathscr{P}(\denotesT{\lsig}) \\

### 2023-03-21 syn-translate.tex
comment

@@
 \end{itemize}
-mutually inductively on $n$, the signature length.
+They are defined mutually inductively on $n$, the signature length.
 
@@
   (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{\snd{\denotesT{m}}}, f = p_f\nu~\sigma 
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{\snd{\denotesT{m}}}, f = p_f\nu~\sigma\YZ{syntax pf nu undefined?}
 \end{align*}

### 2023-03-21 metatheory2.tex
small improvements

@@
 As mentioned in \cref{sec:metatheory2}, the five forms of linkage transformers can be
-thought of a library of functions that are used to construct linkages from other
+thought of as a library of functions that are used to construct linkages from other
 linkages.

### 2023-03-21 syn-translate.tex
Fix and Reply comments

@@
   (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{\snd{\denotesT{m}}}, f = p_f\nu~\sigma\YZ{syntax pf nu undefined?}
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{\snd{\denotesT{m}}}, f = \lsigproj{s}{\lsig} \YZ{syntax pf nu undefined?}\EDJreply{Fixed.}
 \end{align*}

### 2023-03-22 casestudies.tex
Update statistics about FPOP for STLC.

@@
 %not pose theoretical challenges.
+When programming in a direct style, before using FPOP, the STLC base module 
+will take around 350LOCs, and STLC with one feature is around 400 LOCs, because
+it requires to repeat the definition of the STLC base definition. It will take 
+more space when we add more features into STLC.
+
+After using FPOP, the 
+STLC base family also take about 350 LOCs. However, one feature as children family takes only about 100~250 LOCs. What's more, the mixin of two features usually only take 1~20 LOCs.
 

### 2023-03-22 casestudies.tex
Rephrase in a more qualitative style.

--- 2023-03-21 casestudies.tex
@@
 %not pose theoretical challenges.
+
+% I think the following is a bit weird ... not enough quantative and not enough qualititive
+% When programming in a direct style, before using FPOP, the STLC base module 
+% will take around 350LOCs, and STLC with one feature is around 400 LOCs, because
+% it requires to repeat the definition of the STLC base definition. It will take 
+% more space when we add more features into STLC.
+
+% After using FPOP, the 
+% STLC base family also take about 350 LOCs. However, one feature as children family takes only about 100~250 LOCs. What's more, the mixin of two features usually only take 1~20 LOCs.
+When programming in a direct style, before using FPOP, the STLC base module will take around 350 LOCs while using FPOP, the STLC base family takes about same amount of the code.
+
+However, without FPOP, STLC with one feature will require around 400 LOCs as the definition of STLC base is repeated.
+On the other hand, using FPOP, each feature as extended family only requires ranging from 100~250 LOCs.
+What's more, without FPOP, STLC with multiple features will require more and more while using FPOP's mixin feature, tens of LOCs is enough.
 

### 2023-03-22 casestudies.tex
Update qualitative statistic for analysis case.

--- 2023-03-21 casestudies.tex
@@
 %not pose theoretical challenges.
+
+% I think the following is a bit weird ... not enough quantative and not enough qualititive
+% When programming in a direct style, before using \Lang, the STLC base module 
+% will take around 350LOCs, and STLC with one feature is around 400 LOCs, because
+% it requires to repeat the definition of the STLC base definition. It will take 
+% more space when we add more features into STLC.
+
+% After using \Lang, the 
+% STLC base family also take about 350 LOCs. However, one feature as children family takes only about 100~250 LOCs. What's more, the mixin of two features usually only take 1~20 LOCs.
+When programming in a direct style, before using \Lang, the STLC base module will take around 350 LOCs while using \Lang, the STLC base family takes about same amount of the code.
+
+However, without \Lang, STLC with one feature will require around 400 LOCs as the definition of STLC base is repeated.
+On the other hand, using \Lang, each feature as extended family only requires ranging from 100~250 LOCs.
+What's more, without \Lang, STLC with multiple features will require more and more while using \Lang's mixin feature, tens of LOCs is enough.
 
@@
 
+If not using \Lang, we can construct \lsti{Imp}, \lsti{ImpGAI} and \lsti{ImpTI} in a similar way---but \lsti{Imp} is now a module instead of a family, \lsti{ImpGAI} is now a functor instead of a family with overridable fields, and \lsti{ImpTI} is a concrete instantiation of the functor \lsti{ImpGAI}.
+However, \lsti{ImpCP} cannot reuse the above framework as it extends syntax with natural numbers.
+Thus \lsti{ImpCP} will require repeating a large amount of definitions, and it alone will needs about 900~LOCs.
+
 In addition to the two case studies above, we also use extensible

### 2023-03-22 casestudies.tex
edits

@@
 %
-The base STLC family consists of about 400~LOC.
+The linguistic nature of our approach allows us to retain a programming
+style similar to the original proofs in Software Foundations.
+
+The base STLC family consists of \textasciitilde360~LOC,
+about the same as an STLC development not using \Lang.
+%
 Lines of code in each of the four derived families
 ($\mathrm{Y}$, $\times$, $+$, and $\mu$ in the Venn diagram)
-vary from 100 to 250, largely depending on how many constructors they
-add to the inductive types.
-%
-The linguistic nature of our approach allows us to retain a programming
-style similar to the original proofs in Software Foundations.
+vary from 100 to 250,
+largely depending on how many constructors they add to the inductive types.
+Without \Lang, the same STLC code would have to be duplicated for each feature.
 
@@
 Individually developed features can be easily composed (as mixins) to
-form new STLC variants.
+form new STLC variants (e.g., $\mathrm{Y}$$+$$\mu$).
+%
+Such a feature composition often requires only a few lines of code.
 
@@
 features working correctly in isolation may require coordination when composed.
-For example, composing \lsti{STLCIsorec} and \lsti{STLCProd}
+For example, composing $\times$ and $\mu$  %\lsti{STLCIsorec} and \lsti{STLCProd}
 (\cref{fig:stlc-isorec-prod}) creates an obligation to extend \lsti{tysubst} to
@@
 % STLC base family also take about 350 LOCs. However, one feature as children family takes only about 100~250 LOCs. What's more, the mixin of two features usually only take 1~20 LOCs.
-When programming in a direct style, before using \Lang, the STLC base module will take around 350 LOCs while using \Lang, the STLC base family takes about same amount of the code.
+%When programming in a direct style, before using \Lang, the STLC base module will take around 350 LOCs while using \Lang, the STLC base family takes about same amount of the code.
 
-However, without \Lang, STLC with one feature will require around 400 LOCs as the definition of STLC base is repeated.
+%However, without \Lang, STLC with one feature will require around 400 LOCs as the definition of STLC base is repeated.
 On the other hand, using \Lang, each feature as extended family only requires ranging from 100~250 LOCs.

### 2023-03-22 syn-translate.tex
edits

--- 2023-03-21 syn-translate.tex
@@
   (\denotesT{\Tkg {o}}, t[(\SubstExt{p^1}{f})][(\SubstExt{"id"}{\denotesT{\Tkg {o}}})]) \\
-  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{\snd{\denotesT{m}}}, f = p_f\nu~\sigma\YZ{syntax pf nu undefined?}
+  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = \lsigproj{1}{m}, t = \app{\snd{\denotesT{m}}}, f = \lsigproj{s}{\lsig}
 \end{align*}

### 2023-03-22 casestudies.tex
ImpCP

@@
 The code is organized into four families.
-A base family, \lsti{Imp}, defines via \lsti{FInductive} the abstract
+A base family \lsti{Imp} (\textasciitilde 200~LOC) defines via \lsti{FInductive} the abstract
 syntax of a while-language with pure expressions and impure statements.
@@
 Family \lsti{Imp} defines the interpreter via \lsti{FRecursion}.
-It contains \textasciitilde 200~LOC.
 
-A second family, \lsti{ImpGAI}, extends \lsti{Imp} and uses
-\textasciitilde 550~LOC.
+A second family \lsti{ImpGAI} (\textasciitilde 550~LOC) extends \lsti{Imp}.
 It exports a generic framework for deriving
@@
 
-The other two families then extend \lsti{ImpGAI}.
+The next two families then extend \lsti{ImpGAI}.
 Family \lsti{ImpTI} (\textasciitilde200~LOC) is an abstract
@@
 Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
-with natural number arithmetic,
+with natural-number arithmetic,
 and further binds the generic abstract interpreter to perform constant propagation.
+%
+An alternative to \Lang would be to define \lsti{ImpGAI} as a functor parameterized
+by the abstract domain and such. But it would prevent \lsti{ImpCP} from reusing
+the code and proofs in \lsti{ImpGAI}, because \lsti{ImpCP} involves extending
+inductive types with new constructors.
 
@@
 program extraction feature.
-We extract the two verified abstract interpreters to OCaml
-and put them to test over simple queries.
+We extract the two verified abstract interpreters to OCaml.
+Testing the extracted program over simple queries returns expected results.
 %\rD{What were the results of testing the extracted abstract interpreters?
@@
 
-We note that our plugin cannot yet allow the expression of two derived abstract
+We note that \Lang cannot yet allow the expression of two derived abstract
 interpreters nested within the same family.
@@
 
-If not using \Lang, we can construct \lsti{Imp}, \lsti{ImpGAI} and \lsti{ImpTI} in a similar way---but \lsti{Imp} is now a module instead of a family, \lsti{ImpGAI} is now a functor instead of a family with overridable fields, and \lsti{ImpTI} is a concrete instantiation of the functor \lsti{ImpGAI}.
+%If not using \Lang, we can construct \lsti{Imp}, \lsti{ImpGAI} and \lsti{ImpTI} in a similar way---but \lsti{Imp} is now a module instead of a family, \lsti{ImpGAI} is now a functor instead of a family with overridable fields, and \lsti{ImpTI} is a concrete instantiation of the functor \lsti{ImpGAI}.
 However, \lsti{ImpCP} cannot reuse the above framework as it extends syntax with natural numbers.

### 2023-03-22 canonicitymodel.tex
comments

@@
   \denotesC{\goodType{\Gm}{\bot}{}}(\gm)(\gm')(t) &= \emptyset \\
-  \denotesC{\goodType{\Gm}{\cB}{}}(\gm)(\gm')(t) &= \{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \ | \text{ when } t \equiv \false  \} \quad \text{disjoint union} \\
-  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \\
+  \denotesC{\goodType{\Gm}{\cB}{}}(\gm)(\gm')(t) &= 
+    \begin{cases}
+      \{\star^1\} & \text{if } t \equiv \true \\
+      \{\star^2\} & \text{if } t \equiv \false
+    \end{cases}\YZ{Is this more accurate than 'disjoint union'?}
+    \\
+%   \setc{\star^1}{ \text{ when } t \equiv \true } \dot{\cup} \setc{\star^2}{\text{ when } t \equiv \false } \quad \text{disjoint union} \\
+  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \YZ{What if the conditions do not hold?
+Empty set?}\\
   \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
@@
   \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 
-  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \{ (\gamma' , t') : \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  \} \\
+  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \setc{ (\gamma' , t') }{ \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  } \\
   \denotesC{\goodTerm{\Gm}{\sub{t}{\sigma}}{\sub{T}{\sigma}}}(\gm)(\gm') &= \denotesC{t}(\sigma \circ \gm)(\denotesC{\sigma}(\gm)(\gm')) \\
@@
   \denotesC{\goodTerm{\Gm}{\snd{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[1]\\
-  \denotesC{\goodType{\Gm}{\cU_j}{j+1}}(\gm)(\gm')(T) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\}  \to \Set{j} \\
+  \denotesC{\goodType{\Gm}{\cU_j}{j+1}}(\gm)(\gm')(T) &= \setc{t} { \goodTerm{\cdot}{t}{\El{T}} }  \to \Set{j} \\
   \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm)(\gm') &= \denotesC{T}(\gm)(\gm') \\ 
@@
 
-%The difficult part of the proof-relevant logical-relations model is the
-%treatment of propositions---
-Proof-relevant model treats proposition in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}
-a proposition is encoded as a subsingleton data. For
+The difficult part of the proof-relevant logical-relations model is the
+treatment of propositions.
+The model treats proposition in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}\YZreply{Why is it 'difficult'?
+Is there a more appropriate adjective than 'difficult'?}
+a proposition is encoded as a sub\-singleton. For
 example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \
 | \text{ when } t \equiv \false  \}$ represents a disjoint union of two
-subsingleton sets, where the first one is either a singleton if we have a proof of
+sub\-singleton sets, where the first one is either a singleton if we have a proof of
 $t \equiv \true $ or an empty set otherwise.
@@
 Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$
-that returns the syntax piece $S$ as well.\footnote{This is also called
+that also returns the syntax piece $S$.\footnote{This is also called
 glued interpretation in \citet{sterling2019algebraic}.}
@@
 \denotesCC{\TyPi{A}{B}}$ now that the syntax is available.
-Similarly, there are functions $\Glued{\Sigma}$, $\Gpair{a}{b}$  and
+Similarly, there are functions $\Glued{\Sigma}$, $\Gpair{a}{b}$, and
 $\GSubstExt{\gm}{t}$ for dependent pair types, dependent pairs, and
@@
 Furthermore, given $\Glued{S}$ (i.e., the syntax and its semantic interpretation),
-we use $\Gmodel{\Glued{S}}$ to mean the latter of the two.\YZ{why not just $\denotesC{S}$?}\EDJreply{Using superscript C is shorter and more precise -- we are not "reconstructing the C model from the syntax, but taken out the C model directly from the pair".
+we use $\Gmodel{\Glued{S}}$ to mean the latter of the two.%\YZ{why not just $\denotesC{S}$?}\EDJreply{Using superscript C is shorter and more precise -- we are not "reconstructing the C model from the syntax, but taken out the C model directly from the pair".
 Look at the canonicity model where superscript c is used, if change to $\denotesC{-}$ on them, it is not really well-founded recursion, technically speaking (only equivalent to one).}

### 2023-03-22 canonicitymodel.tex
Reply the comments.

@@
       \{\star^2\} & \text{if } t \equiv \false
-    \end{cases}\YZ{Is this more accurate than 'disjoint union'?}
+    \end{cases}\YZ{Is this more accurate than 'disjoint union'?}\EDJreply{It is accurate.
+It is actually the way Coquand write it.
+But I find this way weird type-theoretically---at the very end, when we want to prove a closed term t is either true or false, we will use this cB.
+But defining in this Coquand way, type-theoretically, we cannot extract a proof term t = true.
+(i.e.
+there is no proof term stored here!).
+I know my above sentence is a bit awkward but I just find this Coquand way of saying very...counter-intuitive.}
     \\
@@
   \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \YZ{What if the conditions do not hold?
-Empty set?}\\
+Empty set?}\EDJreply{Yes.
+It is a family of set indexed by t (or just a function maps t to a set), and if t is neither of the above, then the function will map to empty set.}\\
   \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
@@
 
-The difficult part of the proof-relevant logical-relations model is the
+The eccentric part of the proof-relevant logical-relations model is the
 treatment of propositions.
 The model treats proposition in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}\YZreply{Why is it 'difficult'?
-Is there a more appropriate adjective than 'difficult'?}
+Is there a more appropriate adjective than 'difficult'?}\EDJreply{Don't you find difficult?
+At first sight, I cannot understand why this sentence, proposed by Sterling in two of his paper, can be valid mathematically.
+It only makes sense to me until I am in a type-theoretic world (by Kaposi, since Eq is an inductive type now) (or maybe topos-theoretically where categorical logic will encode equality using data like equalizer).
+What about changing the phrase to "eccentric" or "weird"?}
 a proposition is encoded as a sub\-singleton. For

### 2023-03-22 canonicitymodel.tex
fixes

@@
 Without exposing the reader to too many technical details, our meta\-language
-should be understood an instance of any of the above logical frameworks---the difference is our formulation did quotient manually but if we use their framework, quotient is done automatically.
+should be understood as an instance of any of the above logical frameworks---the
+difference is that quotienting is manual in our formulation, whereas it is
+automatic with the logical frameworks.
 %Our model works directly on the earlier introduced syntax for readability\footnote{We still
@@
 \noindent
-Canonicity is a basic criterion that should be met by a dependent type theory serving as
+Canonicity is a fundamental property that should be satisfied by a dependent type theory serving as
 a programming language or as a computational foundation for mathematics.
@@
 
-We first need the mathematical setup to interpret universe levels, following \citet{sterling2019algebraic}:
+First, we need the mathematical setup to interpret universe levels, following \citet{sterling2019algebraic}:
 
@@
   \item The universe hierarchy is cumulative, as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$.
-  \item Thus if $\denotesC{A} \in \Set{i}, \denotesC{B} \in \Set{j}$, we will
+  \item Thus, if $\denotesC{A} \in \Set{i}, \denotesC{B} \in \Set{j}$, we will
   have $\denotesC{\TyPi{A}{B}} \in \Set{i} \cup \Set{j} = \Set{i \lcup j}$.
@@
 \begin{align*}
-  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k} \quad \text{ (also sets indexed by closed substitution)}\\
+  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \setc{\gm}{\goodSub{\cdot}{\gm}{\Gm}} \to \Set{k} \quad \text{ (i.e., sets indexed by closed substitution)}\\
   % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
-  \denotesC{\goodType{\Gm}{T}{j}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\{t : \goodTerm{\cdot}{t}{\sub{T}{\gm}}\}\to\Set{j} \\
+  \denotesC{\goodType{\Gm}{T}{j}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\setc{t}{\goodTerm{\cdot}{t}{\sub{T}{\gm}}}\to\Set{j} \\
   % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
@@
   % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
-  \denotesC{\goodTerm{\Gm}{t}{T}} \text{ is a dependent function}&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})\\
+  \denotesC{\goodTerm{\Gm}{t}{T}} \text{ is a dependent function}&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})\\
   \denotesC{\goodType{\Gm}{\sub{T}{\sigma}}{}}(\gm)(\gm')(t) &= \denotesC{T}(\sigma \circ \gm)(\denotes{\sigma}(\gm)(\gm'))(t)\\
@@
 It is a family of set indexed by t (or just a function maps t to a set), and if t is neither of the above, then the function will map to empty set.}\\
-  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
+  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})(\gm', u')(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
   \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}}(\gm)(\gm')(t) &= \sum_{u' \in \denotesC{A}(\gm)(\gm')(\fst{t})} \denotesC{B}(\SubstExt{\gm}{\fst{t}})((\gm', u'))(\snd{t}) \\
@@
   \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm)(\gm') &= \star\\
-  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}}(\gm)(\gm') &= \lambda u \lambda u' . \denotesC{t}(\SubstExt{\gm}{u}) ((\gm' , u')) \\
+  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}}(\gm)(\gm') &= \lambda u \lambda u' . \denotesC{t}(\SubstExt{\gm}{u}) (\gm' , u') \\
   \denotesC{\goodTerm{\Gm}{\app{t}}{B}}(\gm)(\gm') &= \denotesC{t}(\pi_1 \gm)(\gm'[0])(\pi_2 \gm)(\gm'[1]) \\

### 2023-03-22 syn-translate.tex
fixes

@@
 \begin{itemize}
-  \item a new type $\goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}$  when
-  $\goodCtx{\Gamma}{}$  is well-formed and $n \in \mathbb{N}$
-  such that $\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}}$
-  \item and a function $\mathscr{P} : \setc{\lsig'}{ \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{n}{j}} } \to \setc{T}{ \goodType{\Gm}{T}{} }$ mapping a term of type $\Sigr{\Gm}{n}{j}$ into a type.
+  \item A type $\goodType{\Gamma}{\Sigr{\Gamma}{n}{j}}{}$ where
+  $\goodCtx{\Gamma}{}$ and $n \in \mathbb{N}$.
+  The idea is that $\goodTerm{\Gm}{\denotesT{\goodSigU{j}{\Gamma}{\lsig}{n}}}{\Sigr{\Gm}{n}{j}}$
+  \item A function $\mathscr{P} : \setc{\lsig'}{ \goodTerm{\Gm}{\lsig'}{\Sigr{\Gm}{n}{j}} } \to \setc{T}{ \goodType{\Gm}{T}{} }$ mapping a term of type $\Sigr{\Gm}{n}{j}$ to a type.
 \end{itemize}
 They are defined mutually inductively on $n$, the signature length.
-
 \begin{align*}
@@
 With the above two helpers, we can define denotation $\denotesT{}$.
-
+%
 \begin{align*}
@@
   {\goodType{\Gamma}{\TyTkg{\lsig}}{}} &= \mathscr{P}(\denotesT{\lsig}) \\
-  \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ defines upon } \denotesT{\lsig} 
+  \denotesT{\goodType{\Gamma}{\TyLkg{\sigma}}{}} &\text{ is defined upon } \denotesT{\lsig}
   \text{ and inductively on the signature length} \\

### 2023-03-22 canonicitymodel.tex
resolve comment

@@
   \denotesC{\goodType{\Gm}{\cB}{}}(\gm)(\gm')(t) &= 
-    \begin{cases}
-      \{\star^1\} & \text{if } t \equiv \true \\
-      \{\star^2\} & \text{if } t \equiv \false
-    \end{cases}\YZ{Is this more accurate than 'disjoint union'?}\EDJreply{It is accurate.
+    %\YZ{Is this more accurate than 'disjoint union'?}\EDJreply{It is accurate.
 It is actually the way Coquand write it.
@@
 I know my above sentence is a bit awkward but I just find this Coquand way of saying very...counter-intuitive.}
-    \\
-%   \setc{\star^1}{ \text{ when } t \equiv \true } \dot{\cup} \setc{\star^2}{\text{ when } t \equiv \false } \quad \text{disjoint union} \\
+  \setc{\star^1}{ \text{ when } t \equiv \true } \uplus \setc{\star^2}{\text{ when } t \equiv \false } \quad \text{disjoint union} \\
   \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \YZ{What if the conditions do not hold?
@@
 a proposition is encoded as a sub\-singleton. For
-example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \
+example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \uplus \{\star^2 \
 | \text{ when } t \equiv \false  \}$ represents a disjoint union of two

### 2023-03-22 canonicitymodel.tex
resolve comments

@@
   \setc{\star^1}{ \text{ when } t \equiv \true } \uplus \setc{\star^2}{\text{ when } t \equiv \false } \quad \text{disjoint union} \\
-  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \YZ{What if the conditions do not hold?
+  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \setc{\star}{\text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm}} %\YZ{What if the conditions do not hold?
 Empty set?}\EDJreply{Yes.
-It is a family of set indexed by t (or just a function maps t to a set), and if t is neither of the above, then the function will map to empty set.}\\
+It is a family of set indexed by t (or just a function maps t to a set), and if t is neither of the above, then the function will map to empty set.}
+  \\
   \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})(\gm', u')(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
@@
 
-The eccentric part of the proof-relevant logical-relations model is the
+What is perhaps unintuitive about the proof-relevant logical-relations model is the
 treatment of propositions.
-The model treats proposition in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}\YZreply{Why is it 'difficult'?
+%The model treats propositions in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}\YZreply{Why is it 'difficult'?
 Is there a more appropriate adjective than 'difficult'?}\EDJreply{Don't you find difficult?
@@
 What about changing the phrase to "eccentric" or "weird"?}
-a proposition is encoded as a sub\-singleton. For
-example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \uplus \{\star^2 \
-| \text{ when } t \equiv \false  \}$ represents a disjoint union of two
+A proposition is encoded as a sub\-singleton. For
+example, $\setc{\star^1}{\text{ when } t \equiv \true} \uplus \setc{\star^2}{\text{ when } t \equiv \false}$ represents a disjoint union of two
 sub\-singleton sets, where the first one is either a singleton if we have a proof of

### 2023-03-22 canonicitymodel.tex
fixes

@@
   \begin{itemize}
-    \item If $\goodTerm{\cdot}{t}{El~(\wcode{\tau})}$ with $\goodWSig{\cdot}{\tau}{n}$, then $\goodTerm{\cdot}{t \equiv "Wsup"~j~a~b}{El~(\wcode{\tau})}$ for some $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot, B[(id, a)]}{b}{El~(\wcode{\tau})}$ and $j < n$
+    \item If $\goodTerm{\cdot}{t}{\El{\wcode{\tau}}}$ and $\goodWSig{\cdot}{\tau}{n}$, then $\goodTerm{\cdot}{t \equiv \Wsup{j}{\tau}{a}{b}}{\El{\wcode{\tau}}}$ for some $\goodTerm{\cdot}{a}{A}$, $\goodTerm{B[("id", a)]}{b}{\El{\wcode{\tau}}}$, and $j < n$
     \item If $\goodTerm{\cdot}{t}{\mathbb{B}}$ then $\goodTerm{\cdot}{t \equiv "tt"}{\mathbb{B}}$ or $\goodTerm{\cdot}{t \equiv "ff"}{\mathbb{B}}$ 
-    \item If $\goodTerm{\cdot}{t}{\cL\sigma}$ with $\goodSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv \LkgAdd{o}{t}}{\cL\sigma}$
-
-      $\quad$ for some $\goodTerm{\cdot}{o}{\cL(\lkgproj{1}{\sigma})}$ and $\goodTerm{\cdot, \lsigproj{2}{\sigma}}{t}{\lkgproj{2}{\sigma}}$
-    \item If $\goodTerm{\cdot}{t}{\Sigma A B}$ then $\goodTerm{\cdot}{t = (a, b)}{\Sigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[(id, a)]}$
+    \item If $\goodTerm{\cdot}{t}{\TyLkg{\lsig}}$ with $\goodSig{\cdot}{\sigma}{n}$, then $\goodTerm{\cdot}{t \equiv \LkgAdd{o}{t}}{\TyLkg{\lsig}}$ 
+      for some $\goodTerm{\cdot}{o}{\TyLkg{\lkgproj{1}{\sigma}}}$ and $\goodTerm{\lsigproj{2}{\sigma}}{t}{\lkgproj{2}{\sigma}}$
+    \item If $\goodTerm{\cdot}{t}{\TySigma A B}$ then $\goodTerm{\cdot}{t \equiv (a, b)}{\TySigma A B}$ with $\goodTerm{\cdot}{a}{A}$ and $\goodTerm{\cdot}{b}{B[("id", a)]}$
     \footnote{We emphasize the last one because $\Tkg{\ell}$ is a dependent pair}

### 2023-03-23 canonicitymodel.tex
question
Remove the intricate proposition paragraph Now that we don't encode prop using data. State FTLR.

@@
 I know my above sentence is a bit awkward but I just find this Coquand way of saying very...counter-intuitive.}
-  \setc{\star^1}{ \text{ when } t \equiv \true } \uplus \setc{\star^2}{\text{ when } t \equiv \false } \quad \text{disjoint union} \\
-  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \setc{\star}{\text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm}} %\YZ{What if the conditions do not hold?
+    \begin{cases}
+      \{\star^1\} & \text{if } t \equiv \true \\
+      \{\star^2\} & \text{if } t \equiv \false \\
+      \emptyset & \text{otherwise }
+    \end{cases}\\
+  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= 
+  \begin{cases}
+    \{\star\} & \text{if } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm}\\
+    \emptyset & \text{otherwise }
+  \end{cases}\\
+  % \setc{\star}{\text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm}} %\YZ{What if the conditions do not hold?
 Empty set?}\EDJreply{Yes.
 It is a family of set indexed by t (or just a function maps t to a set), and if t is neither of the above, then the function will map to empty set.}
-  \\
   \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})(\gm', u')(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
@@
   \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm)(\gm') &= \star^2 \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{a}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
-  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{b}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^2 \\
+  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= 
+  \begin{cases}
+    \denotesC{a}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
+    \denotesC{b}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^2
+  \end{cases}\\
+  % \denotesC{a}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
+  % \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{b}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^2 \\
   \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm)(\gm') &= \star\\
@@
 
-What is perhaps unintuitive about the proof-relevant logical-relations model is the
-treatment of propositions.
+% What is perhaps unintuitive about the proof-relevant logical-relations model is the
+% treatment of propositions.
 %The model treats propositions in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}\YZreply{Why is it 'difficult'?
@@
 What about changing the phrase to "eccentric" or "weird"?}
-A proposition is encoded as a sub\-singleton. For
-example, $\setc{\star^1}{\text{ when } t \equiv \true} \uplus \setc{\star^2}{\text{ when } t \equiv \false}$ represents a disjoint union of two
-sub\-singleton sets, where the first one is either a singleton if we have a proof of
-$t \equiv \true $ or an empty set otherwise.
-In other words,
-$\star^1$ witnesses a proof that $t \equiv \true$.
+% A proposition is encoded as a sub\-singleton. For
+% example, $\setc{\star^1}{\text{ when } t \equiv \true} \uplus \setc{\star^2}{\text{ when } t \equiv \false}$ represents a disjoint union of two
+% sub\-singleton sets, where the first one is either a singleton if we have a proof of
+% $t \equiv \true $ or an empty set otherwise.
+% In other words,
+% $\star^1$ witnesses a proof that $t \equiv \true$.
 Here, $\star$, $\star^1$, $\star^2$ are just some arbitrary fixed elements.
@@
 in a strictly positive position. %So it is a well-founded definition.
-We do not distinguish $(b, b^C)$ and $\Glued{b}$ for simplicity.
+We do not distinguish $(b, b^C)$ and $\Glued{b}$ for simplicity. Again, we omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
+
+We state the fundamental property of logical relation here---in our proof relevant model, it is admitting that our model is really a model, and our syntax is the initial model so there is map from our syntax to our canonicity model:
+\begin{lemma}[Fundamental Theorem of Logical Relation]
+  If we have a (intrinsically well-typed) term $\goodTerm{\Gm}{t}{T}$, then we have its interpretation as a dependent function $\denotesC{t} : \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})$
+\end{lemma}
 
@@
 \denotesC{\bot}(\SubstEmp)(\star)(\sub{t}{\gm}) = \emptyset$, a contradiction.
+\YZ{Have you stated the fundamental property of the logical-relations model?}\EDJreply{No. Now stated.}
 
@@
 $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp)(\star) = \star^1$ or $\star^2$,
-witnessing the proof of $t \equiv \true \text{ or } t \equiv\false$, arriving at
+witnessing the proof of $t \equiv \true \text{ or } t \equiv\false$ by the definition of our model, arriving at
 \cref{thm:canonicity-appendix}.

### 2023-03-23 canonicitymodel.tex
Fix.

@@
 \begin{lemma}[Fundamental Theorem of Logical Relation]
-  If we have a (intrinsically well-typed) term $\goodTerm{\Gm}{t}{T}$, then we have its interpretation as a dependent function $\denotesC{t} : \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})$
+  If we have a (intrinsically well-typed) term $\goodTerm{\Gm}{t}{T}$, then we have its interpretation as a dependent function \\ $\denotesC{t} : \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})$
 \end{lemma}

### 2023-03-23 metatheory2.tex
tweaks
Working on adding explanation of boldP

@@
 % Deal with intuition
+\noindentparagraph{The necessity of $\TyTkg{\cdot}$.} 
+According to the equality rules of $\ruleref{}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is very similar to $\TyLkg{\lsig}$ but we are using dependent pair $\pair{}{}$ instead of $ $ to package the fields of a linkage.
+We can compute a concrete example to persuade the audience.
+
+Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
+$\TyTkg{\cdot}$ is indispensible because we cannot do \emph{abstraction} with $\TyLkg{\cdot}$ alone, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{}$ is usually a dependent tuple type.
+We make a small example here.
 
@@
 
-\subsection{A Translation that Compiles Away Linkages}
+\subsection{A Translation that Compiles Linkages Away}
 

### 2023-03-23 metatheory-condensed.tex
bugfixes

@@
 \goodTerm{
-  \mathit{self} :
+  x :
   \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
@@
 constructed from the introduction forms
-$\small\InhId$, $\small\InhExt{h}{x.t}$, $\small\InhOv{h}{x.t}$, $\small\InhInh{h}$, etc.
+$\small\InhId$, $\small\InhExt{h}{\mathit{self}.t}$, $\small\InhOv{h}{\mathit{self}.t}$, %$\small\InhInh{h}$,
+etc.
 \cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
 \ruleref{tmeq/ov/beta}.
-It states that applying the transformer $\small\InhOv{h}{\mathit{self}_2.t_2}$ to
-a linkage of form $\LkgAdd{\lkg}{\mathit{self}_1.t_1}$
+It states that applying the transformer $\small\InhOv{h}{\mathit{self}_{\!2}.t_2}$ to
+a linkage of form $\LkgAdd{\lkg}{\mathit{self}_{\!1}.t_1}$
 overrides the linkage's last field~$t_1$ with~$t_2$.
 For instance, the construction below shows that 
-$\small\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$
+$\small\InhOv{\InhId}{\mathit{self}.\wcode{\wsig'_{\texttt{tm}}}}$
 is used as the first step in creating a linkage transformer modeling a derived family that

### 2023-03-23 metatheory-condensed.tex
fix margin overrun

@@
 For instance, the construction below shows that 
-$\small\InhOv{\InhId}{\mathit{self}.\wcode{\wsig'_{\texttt{tm}}}}$
+$\small\InhOv{\InhId}{\mathit{self}'\!.\wcode{\wsig'_{\texttt{tm}}}}$
 is used as the first step in creating a linkage transformer modeling a derived family that

### 2023-03-23 metatheory-condensed.tex
fix

@@
 can reuse~$t$---without retyping it---by first
-projecting~$t$ from $\LkgAdd{\lkg}{t}$ and then appending it to~$\lkg'$.
+projecting~$t$ from $\LkgAdd{\lkg}{\mathit{self}.t}$ and then appending it to~$\lkg'$.
 %\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.
@@
 The next four fields model the four constructors of~\lsti{tm}.
-Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}$
+Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{x.\bot}$
 and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for

### 2023-03-23 canonicitymodel.tex
edits

--- 2023-03-23 canonicitymodel.tex
@@
 \noindent
-Canonicity is a fundamental property that should be satisfied by a dependent type theory serving as
+Canonicity is a key criterion for a dependent type theory to be considered as
 a programming language or as a computational foundation for mathematics.
@@
 
-We state the fundamental property of logical relation here---in our proof relevant model, it is admitting that our model is really a model, and our syntax is the initial model so there is map from our syntax to our canonicity model:
-\begin{lemma}[Fundamental Theorem of Logical Relation]
-  If we have a (intrinsically well-typed) term $\goodTerm{\Gm}{t}{T}$, then we have its interpretation as a dependent function $\denotesC{t} : \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})$
-\end{lemma}
+We state the fundamental property of the logical-relations model.
+%in our proof relevant model, it is admitting that our model is really a model,
+%and our syntax is the initial model so there is map from our syntax to our
+%canonicity model:
+\begin{theorem}[Fundamental Property]
+  If\/ $\goodTerm{\Gm}{t}{T}$, then its semantic interpretation is a dependent function such that
+  $\denotesC{t} : \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})$.
+\end{theorem}
 
@@
 \denotesC{\bot}(\SubstEmp)(\star)(\sub{t}{\gm}) = \emptyset$, a contradiction.
-\YZ{Have you stated the fundamental property of the logical-relations model?}\EDJreply{No. Now stated.}
 
@@
 
-Further, with the help of eta rules, we have the following canonical forms
+Further, with the help of eta rules, we have the following canonical forms.
 \begin{theorem}[Canonical Forms]\hfill

### 2023-03-23 metatheory-condensed.tex
edits

@@
 Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{x.\bot}$
-and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
+and typed with $\El{\mathit{self}."tm"}$, where $\mathit{self}$ stands for
 the typing context containing the first field~\lsti{tm}.
 The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
-this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
-So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
+this typing context; that is, $\mathit{self}."tm"$ has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
+So $\El{\mathit{self}."tm"}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
 as required by rule \ruleref{tm/wsup}.

### 2023-03-23 stlc-linkage-typing-named.tex
more bugfixes
fix
Adding boldP computation.

@@
 &
-\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{
+\lsig_1 \coloneqq \LSigAdd{\lsig_0}{x.\seal_1}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-a}(-0.02,-0.10)(-0.02,0.28)
@@
 \lkg_1 \coloneqq \LkgAdd{\lkg_0}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-a}(-0.02,-0.10)(-0.02,0.28)
@@
 &
-\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{
+\lsig_2 \coloneqq \LSigAdd{\lsig_1}{x.\seal_2}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-a}(-0.02,-0.10)(-0.02,0.28)
@@
 \lkg_2 \coloneqq \LkgAdd{\lkg_1}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-a}(-0.02,-0.10)(-0.02,0.28)
@@
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-b}(-0.02,-0.10)(-0.02,0.28)
-    \Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}
+    \Wsup{}{\wsig_{\texttt{tm}}}{\top}{x.\bot}
     \tikzmarkend{t2-b}
@@
 &
-\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{
+\lsig_6 \coloneqq \LSigAdd{\lsig_5}{x.\seal_6}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-a}(-0.02,-0.10)(-0.02,0.28)
@@
 \lkg_6 \coloneqq \LkgAdd{\lkg_5}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-a}(-0.02,-0.10)(-0.02,0.28)
@@
 &
-\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{
+\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{x.\seal_{10}}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-a}(-0.02,-0.10)(-0.02,0.28)
@@
 \lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
+    \mathit{self}\!.\,
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-a}(-0.02,-0.10)(-0.02,0.28)
@@
 \end{tabular}
-
+\begin{tabular}{ll}
+  \hline
+  $\TyTkg{\lsig_1} = \nTySigma{\_}{\top}{\TyS{\wcode{\wsig_{\texttt{tm}}}}} $ &  
+  $\TyTkg{\lsig_2} = \nTySigma{\_}{\TyTkg{\lsig_1}}{\El{\wcode{\wsig_{\texttt{tm}}}}}$ \\  
+  $\TyTkg{\lsig_6} = \nTySigma{P_5}{\TyTkg{\lsig_5}}{\CaseSig{\top}{\bot}{\El{P_5.tm} \to \mathit{id} \to \El{P_5.tm}}}$\EDJ{It strike me that, we don't have field projection for dependent tuple...
+so it is a bit ambiguous.
+I am not sure if I should continue; but some explanation of this field projection is required I believe} & \\
+  $\TyTkg{\lsig_{10}} = \nTySigma{\_}{\TyTkg{\lsig_9}}{\El{\wcode{\wsig_{\texttt{tm}}}} \to \El{\wcode{\wsig_{\texttt{tm}}}} \to \mathit{id} \to \El{\wcode{\wsig_{\texttt{tm}}}}}$ & \\ 
+\end{tabular}
 \caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.
@@
 \label{fig:stlc-linkage-typing}
+
 \end{figure}

### 2023-03-23 metatheory2.tex
Working on the example.

--- 2023-03-21 metatheory2.tex
@@
 % Deal with intuition
+\noindentparagraph{The necessity of $\TyTkg{\cdot}$.} 
+According to the equality rules of $\ruleref{tyeq/pk/add}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is very similar to $\TyLkg{\lsig}$ but we are using dependent tuple $\pair{\cdot}{\cdot}$ instead of $\LkgAdd{\cdot}{\cdot}$ to package the fields of a linkage.
+We have computed several examples in \cref{fig:stlc-linkage-typing}.
+
+Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
+$\TyTkg{\cdot}$ is indispensible because we cannot do \emph{abstraction} with $\TyLkg{\cdot}$ alone, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{lsig/add}$ and $\ruleref{l/add}$ is usually a dependent tuple type.
+
+We illustrate with a small example, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function maps from the first to the second type.
+Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{}}$.
+Say we want to add a new field into this linkage and abstracting both extensible inductive types, then in $\ruleref{l/add}$, the chosen $A$ must abstract both inductive type to $\cU$.
+The  question now is whether $A$ is a linkage type or a dependent tuple type; and for each case we will intuitively choose the following approach to do abstraction.
+
+However, the first approach won't work because such linkage cannot exist.
+
+To summarize, the reason we need dependent tuple type is because of the overridability nature of our linkages, together with the frequent usage of universe type in the context of dependent type programming.
+Once we want to abstract a type in a signature, we may have to use dependent tuple type as the abstraction.
 

### 2023-03-23 metatheory2.tex
Finish Explaning the necessity of boldP.

--- 2023-03-21 metatheory2.tex
@@
 % Deal with intuition
+\noindentparagraph{The necessity of $\TyTkg{\cdot}$.} 
+According to the equality rules of $\ruleref{tyeq/pk/add}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is structurally similar to $\TyLkg{\lsig}$ but we are using dependent tuple $\pair{\cdot}{\cdot}$ instead of $\LkgAdd{\cdot}{\cdot}$ to package the fields of a linkage.
+We have computed several examples in \cref{fig:stlc-linkage-typing}.
+
+Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
+$\TyTkg{\cdot}$ is indispensible because \emph{we cannot do abstraction with $\TyLkg{\cdot}$ alone}, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{lsig/add}$ and $\ruleref{l/add}$ is \emph{usually a dependent tuple type}.
+
+\begin{align*}
+  \lsig &\coloneq \{ T_A : \TyS{\wcode{\wsig_1}}; &&T_B : \TyS{\wcode{\wsig_2}}; &&f : \El{T_A} \to \El{T_B} \} \\
+  \lsig^A &\coloneq \{T_A : \cU; &&T_B : \cU; &&f : \El{T_A} \to \El{T_B} \}
+\end{align*}
+
+We illustrate with a small example using named convention, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function $f$ maps from the first to the second type.
+Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{\lsig}}$.
+Say we want to add a new field into this linkage and abstracting both extensible inductive types, then in $\ruleref{l/add}$, the chosen $A$ must abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
+The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction.
+
+\begin{align*}
+  & \text{doing abstraction} \\ 
+  \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
+  \text{Approach } \beta \quad&\quad  \TyLkg{\lsig} \longrightarrow \TyTkg{\lsig} \longrightarrow \TyTkg{\lsig^A}\\
+\end{align*}
+
+However, the seemingly easier approach $\alpha$ is impossible because linkage $\TyLkg{\lsig^A}$ is uninhabited---say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
+To make the inconsistency clearer, by double abstraction we will have an impossible function $g \coloneq \goodTerm{\cdot}{\nlam{T_A}{\nlam{T_B}{\ell^A.f}}}{\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A} \to \El{T_B}}}} $, i.e.
+a function maps from arbitrary type to another arbitrary type .
+We can easily derive inconsistency setting applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
+
+Thus we can only choose the approach $\beta$ instead, which justifies why we have $\TyTkg{\cdot}$ in our typing rule, as $\TyTkg{\lsig}$ is usually the first step to do abstraction.
+
+To summarize, the reason we need dependent tuple type is because of the overridability nature of our linkages, together with the frequent usage of universe type in the context of dependent type programming.
+Once we want to abstract a type in a signature, we may have to use dependent tuple type as the abstraction instead of linkages.
+Since abstraction is usually dependent tuple type, we introduce $\TyTkg{\cdot}$ and its computation rules to make the process of doing abstraction in \TT easier.
 

### 2023-03-23 metatheory2.tex
boldP explanation added. Phrasing.

@@
 \noindentparagraph{The necessity of $\TyTkg{\cdot}$.} 
-According to the equality rules of $\ruleref{tyeq/pk/add}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is structurally similar to $\TyLkg{\lsig}$ but we are using dependent tuple $\pair{\cdot}{\cdot}$ instead of $\LkgAdd{\cdot}{\cdot}$ to package the fields of a linkage.
-We have computed several examples in \cref{fig:stlc-linkage-typing}.
+According to the equality rules $\ruleref{tyeq/pk/add}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is structurally similar to $\TyLkg{\lsig}$ but we are using dependent tuple $\pair{\cdot}{\cdot}$ instead of $\LkgAdd{\cdot}{\cdot}$ to package the fields of a linkage.
+We have computed several examples in \cref{fig:stlc-linkage-typing}, where the reader should also notice the high similarity between $\TyTkg{\lsig}$ and $\TyLkg{\lsig}$.
 
@@
 Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{\lsig}}$.
-Say we want to add a new field into this linkage and abstracting both extensible inductive types, then in $\ruleref{l/add}$, the chosen $A$ must abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
-The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction.
+Say we want to add a new field into this linkage, and in $\ruleref{l/add}$, we chose $A$ to abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
+The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction (choose $s$).
 
 \begin{align*}
-  & \text{doing abstraction} \\ 
+  & \text{doing abstraction (choosing appropriate $s$ in \ruleref{l/add})} \\ 
   \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
@@
 a function maps from arbitrary type to another arbitrary type .
-We can easily derive inconsistency setting applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
+We can easily derive inconsistency by applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
 

### 2023-03-23 metatheory2.tex
edit

@@
 \noindentparagraph{The necessity of $\TyTkg{\cdot}$.}
-According to the equality rules $\ruleref{tyeq/pk/add}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is structurally similar to $\TyLkg{\lsig}$ but we are using dependent tuple $\pair{\cdot}{\cdot}$ instead of $\LkgAdd{\cdot}{\cdot}$ to package the fields of a linkage.
-We have computed several examples in \cref{fig:stlc-linkage-typing}, where the reader should also notice the high similarity between $\TyTkg{\lsig}$ and $\TyLkg{\lsig}$.
+
+As shown by the equality rule \ruleref{tyeq/pk/add},
+for a given linkage signature $\lsig \coloneq \LSigAdd{\cdot}{\cdot}{\cdot}$,
+the type $\TyTkg{\lsig}$ is structurally similar to the linkage type $\TyLkg{\lsig}$
+except that $\TyTkg{\lsig}$ is a dependent tuple type---%
+it packages the fields of $\lsig$ into a dependent tuple type.
+%
+We have computed several examples in
+\cref{fig:stlc-linkage-typing}, where the reader should also notice the high
+similarity between $\TyTkg{\lsig}$ and $\TyLkg{\lsig}$.
 
@@
 
-However, the seemingly easier approach $\alpha$ is impossible because linkage $\TyLkg{\lsig^A}$ is uninhabited---say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
+However, the seemingly easier approach $\alpha$ is impossible because linkage
+$\TyLkg{\lsig^A}$ is uninhabited---%
+say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
 To make the inconsistency clearer, by double abstraction we will have an impossible function $g \coloneq \goodTerm{\cdot}{\nlam{T_A}{\nlam{T_B}{\ell^A.f}}}{\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A} \to \El{T_B}}}} $, i.e.

### 2023-03-23 metatheory-condensed.tex
notation for field projection

@@
 we write them as dependent record types that give labels to the fields.
+Field accesses are notated, for example, $\getfield{\mathit{self}}{"tm"}$.
 %we use a notation with explicit names
@@
 Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{x.\bot}$
-and typed with $\El{\mathit{self}."tm"}$, where $\mathit{self}$ stands for
+and has type $\El{\getfield{\mathit{self}}{"tm"}}$, where $\mathit{self}$ stands for
 the typing context containing the first field~\lsti{tm}.
 The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
-this typing context; that is, $\mathit{self}."tm"$ has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
-So $\El{\mathit{self}."tm"}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
+this typing context; that is, $\getfield{\mathit{self}}{"tm"}$ has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
+So $\El{\getfield{\mathit{self}}{"tm"}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
 as required by rule \ruleref{tm/wsup}.

### 2023-03-23 stlc-linkage-typing-named.tex
notation for field projection

@@
     \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-b}(-0.02,-0.10)(-0.02,0.28)
-    \El{\mathit{self.}\texttt{tm}}
+    \El{\getfield{\mathit{self}}{\texttt{tm}}}
     \tikzmarkend{T2-b}
@@
         \begin{array}{@{}l@{}}
-            \El{\mathit{self}.\texttt{tm}} \to
+            \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
             \\
-            \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
+            \mathit{id} \to \El{\getfield{\mathit{self}}{\texttt{tm}}}
         \end{array}
@@
     \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-b}(-0.02,-0.10)(-0.02,0.28)
-    \lambda t.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{t}
+    \lambda x.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{x}
     \tikzmarkend{t10-b}
@@
     \begin{array}{@{}l@{}}
-        \El{\mathit{self}.\texttt{tm}} \to
+        \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
         \\
-        \El{\mathit{self}.\texttt{tm}} \to
+        \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
         \\
-        \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
+        \mathit{id} \to \El{\getfield{\mathit{self}}{\texttt{tm}}}
     \end{array}
@@
   $\TyTkg{\lsig_2} = \nTySigma{\_}{\TyTkg{\lsig_1}}{\El{\wcode{\wsig_{\texttt{tm}}}}}$ \\  
-  $\TyTkg{\lsig_6} = \nTySigma{P_5}{\TyTkg{\lsig_5}}{\CaseSig{\top}{\bot}{\El{P_5.tm} \to \mathit{id} \to \El{P_5.tm}}}$\EDJ{It strike me that, we don't have field projection for dependent tuple...
+  $\TyTkg{\lsig_6} = \nTySigma{P_5}{\TyTkg{\lsig_5}}{\CaseSig{\top}{\bot}{\El{\getfield{P_5}{"tm"}} \to \mathit{id} \to \El{\getfield{P_5}{"tm"}}}}$\EDJ{It strike me that, we don't have field projection for dependent tuple...
 so it is a bit ambiguous.

### 2023-03-23 stlc-linkage-typing-named.tex
revamp figure 8

--- 2023-03-21 stlc-linkage-typing-named.tex
@@
 %
+
+\begingroup
+\renewcommand*{\arraystretch}{1.55}
+
+\fcolorbox{black}{faint-gray}{
+\begin{minipage}{25ex}%
+\vspace{-1.2ex}%
+\begin{gather*}%
+\begin{array}{@{}l@{}}
+\goodTerm{x :
+    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{Pi-c}(0.00,-0.15)(0.00,0.30)%
+    \TyTkg{\lsig_{i-1}}
+    \tikzmarkend{Pi-c}
+}{\seal_i}{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{Ai-c}(0.00,-0.15)(0.00,0.30)%
+    A_i
+    \tikzmarkend{Ai-c}
+}
+\\
+\goodTerm{\mathit{self} :
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{Ai-b}(0.00,-0.15)(0.00,0.30)%
+    A_i
+    \tikzmarkend{Ai-b}
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{ti-b}(-0.02,-0.10)(-0.02,0.28)
+    t_i
+    \tikzmarkend{ti-b}
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{Ti-b}(0.00,-0.15)(0.00,0.30)
+    T_i
+    \tikzmarkend{Ti-b}
+}
+\\
+\text{\scriptsize shown below for $i\in\{1,2,6,10\}$}
+\end{array} 
+\end{gather*}
+\end{minipage}
+}%
+\hfill
+\fcolorbox{black}{faint-gray}{
+\begin{minipage}{25ex}%
+\vspace{-1.2ex}%
+\begin{gather*}%
+\begin{array}{@{}l@{}}
+\lsig_i \coloneq \LSigAdd{\lsig_{i-1}}{x.\seal_i}{\mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{Ti-a}(0.00,-0.15)(0.00,0.30)
+    T_i
+    \tikzmarkend{Ti-a}
+}
+\\
+\lkg_i \coloneq \LkgAdd{\lkg_{i-1}}{\mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{ti-a}(-0.02,-0.10)(-0.02,0.28)
+    t_i
+    \tikzmarkend{ti-a}
+}
+\\
+\goodTerm{\cdot}{\lkg_i}{\TyLkg{\lsig_i}}
+\end{array} 
+\end{gather*}
+\end{minipage}
+}%
+\hfill
+\fcolorbox{black}{faint-gray}{
+\begin{minipage}{38.5ex}%
+\vspace{-1.2ex}%
+\begin{gather*}%
+\begin{array}{@{}l@{}}
+\TyTkg{\lsig_i} \coloneq \nTySigma{x}{\TyTkg{\lsig_{i-1}}}{\nsub{T_i}{\mathit{self}}{\seal_i}}
+\\
+\Tkg{\lkg_i} \coloneq \big(\Tkg{\lkg_{i-1}},
+    \nsub{t_i}{\mathit{self}}{\nsub{\seal_i}{x}{\Tkg{\lkg_{i-1}}}}\!
+\big)
+\\
+\goodTerm{\cdot}{\Tkg{\lkg_i}}{\TyTkg{\lsig_i}}
+\end{array} 
+\end{gather*}
+\end{minipage}
+}
+\endgroup
+
+\vspace{2ex}
+
 \begin{tabular}{@{}l@{\ \ \ }L@{\quad}L@{\quad}L@{\quad}L@{}}
@@
 \lkg_0 \coloneqq \LkgEmp
-&
-\goodTerm{\cdot}{\lkg_0}{\TyLkg{\lsig_0}}
+%&
+%\goodTerm{\cdot}{\lkg_0}{\TyLkg{\lsig_0}}
 &
@@
 &
-\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{
-    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-a}(-0.02,-0.10)(-0.02,0.28)
-    T_1
-    \tikzmarkend{T1-a}
+\multicolumn{4}{@{}l@{}}{
+\goodTerm{
+    \mathit{x} : 
+    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P0-c}(0.00,-0.15)(0.00,0.30)%
+    \big[\big]
+    \tikzmarkend{P0-c}%
+}{
+    \seal_1
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-c}(0.00,-0.15)(0.00,0.30)%
+    \big[\big]
+    \tikzmarkend{A1-c}%
 } 
-&
-\lkg_1 \coloneqq \LkgAdd{\lkg_0}{
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-a}(-0.02,-0.10)(-0.02,0.28)
-    t_1
-    \tikzmarkend{t1-a}
 }
-&
-\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
-&
-\goodTerm{x : \TyTkg{\lsig_0}}{\seal_1}{{
-    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-a}(-0.02,-0.10)(-0.05,0.28)
-    A_1
-    \tikzmarkend{A1-a}
-}{}}
-\\
+\\[2pt]
 \cellcolor{Bkg-color}
-\multirow{-2}{*}{%
-\lsti!FInductive!\ \ 
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.03,-0.10)(-0.02,0.28)%
-    \lsti!tm!%
-    \tikzmarkend{t1-c}%
-\ \;\lsti!:=!}
 &
@@
 }
+\\[2pt]
+\cellcolor{Bkg-color}
+\multirow{-3}{*}{%
+\lsti!FInductive!\ \ 
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.03,-0.10)(-0.02,0.28)%
+    \lsti!tm!%
+    \tikzmarkend{t1-c}%
+\ \;\lsti!:=!}
+&
+\lsig_1 \coloneqq \LSigAdd{\lsig_0}{x.\seal_1}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-a}(-0.02,-0.10)(-0.02,0.28)
+    T_1
+    \tikzmarkend{T1-a}
+}
+&
+\lkg_1 \coloneqq \LkgAdd{\lkg_0}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-a}(-0.02,-0.10)(-0.02,0.28)
+    t_1
+    \tikzmarkend{t1-a}
+}
+%&
+%\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
+%&
+%\goodTerm{x : \TyTkg{\lsig_0}}{\seal_1}{{
+%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-a}(-0.02,-0.10)(-0.05,0.28)
+%    A_1
+%    \tikzmarkend{A1-a}
+%}{}}
 \\
@@
 &
-\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{
-    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-a}(-0.02,-0.10)(-0.02,0.28)
-    T_2
-    \tikzmarkend{T2-a}
+\multicolumn{4}{@{}l@{}}{
+\goodTerm{
+    \mathit{x} :
+    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P2-c}(0.00,-0.15)(0.00,0.30)%
+    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
+    \tikzmarkend{P2-c}%
+}{
+    \seal_2
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-c}(0.00,-0.15)(0.00,0.30)
+    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
+    \tikzmarkend{A2-c}
 }
-&
-\lkg_2 \coloneqq \LkgAdd{\lkg_1}{
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-a}(-0.02,-0.10)(-0.02,0.28)
-    t_2
-    \tikzmarkend{t2-a}
 }
-&
-\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
-&
-\goodTerm{x : \TyTkg{\lsig_1}}{\seal_2}{{
-    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-a}(-0.02,-0.10)(-0.05,0.28)
-    A_2
-    \tikzmarkend{A2-a}
-}{}}
-\\
+\\[2pt]
 \cellcolor{Bkg-color}
-\multirow{-2}{*}{%
-    \lsti!|!\ %
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-c}(0.01,-0.10)(-0.02,0.28)%
-    \lsti!tm_unit!%
-    \tikzmarkend{t2-c}%
-    \ \ \lsti!: tm!
-}
 &
@@
 }
-\\
+\\[2pt]
+\cellcolor{Bkg-color}
+\multirow{-3}{*}{%
+    \lsti!|!\ %
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-c}(0.01,-0.10)(-0.02,0.28)%
+    \lsti!tm_unit!%
+    \tikzmarkend{t2-c}%
+    \ \ \lsti!: tm!
+}
+&
+\lsig_2 \coloneqq \LSigAdd{\lsig_1}{x.\seal_2}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-a}(-0.02,-0.10)(-0.02,0.28)
+    T_2
+    \tikzmarkend{T2-a}
+}
+&
+\lkg_2 \coloneqq \LkgAdd{\lkg_1}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-a}(-0.02,-0.10)(-0.02,0.28)
+    t_2
+    \tikzmarkend{t2-a}
+}
+%&
+%\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
+%&
+%\goodTerm{x : \TyTkg{\lsig_1}}{\seal_2}{{
+%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-a}(-0.02,-0.10)(-0.05,0.28)
+%    A_2
+%    \tikzmarkend{A2-a}
+%}{}}
+\\[2pt]
 \cellcolor{Bkg-color-2}
@@
 &
-\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{
-    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-a}(-0.02,-0.10)(-0.02,0.28)
-    T_6
-    \tikzmarkend{T6-a}
+\multicolumn{4}{@{}l@{}}{
+\goodTerm{
+    \mathit{x} :
+    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P6-c}(0.00,-0.15)(0.00,0.30)%
+    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+    \tikzmarkend{P6-c}%
+}{
+    \seal_6
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-c}(0.01,-0.10)(-0.02,0.28)
+    \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+    \tikzmarkend{A6-c}
 }
-&
-\lkg_6 \coloneqq \LkgAdd{\lkg_5}{
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-a}(-0.02,-0.10)(-0.02,0.28)
-    t_6
-    \tikzmarkend{t6-a}
 }
-&
-\goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
-&
-\goodTerm{x : \TyTkg{\lsig_5}}{\seal_6}{{
-    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-a}(-0.02,-0.10)(-0.05,0.28)
-    A_6
-    \tikzmarkend{A6-a}
-}}
-\\
+\\[2pt]
 \cellcolor{Bkg-color}
-\multirow{-2}{*}{%
-    \lsti!Case tm_unit :=!\ \;%
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-c}(0.01,-0.10)(-0.02,0.28)%
-    \dadada%
-    \tikzmarkend{t6-c}%
-    \lsti!.!%
-}
 &
@@
 }
-\\
+\\[2pt]
+\cellcolor{Bkg-color}
+\multirow{-4}{*}{%
+    \lsti!Case tm_unit :=!\ \;%
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-c}(0.01,-0.10)(-0.02,0.28)%
+    \dadada%
+    \tikzmarkend{t6-c}%
+    \lsti!.!%
+}
+&
+\lsig_6 \coloneqq \LSigAdd{\lsig_5}{x.\seal_6}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-a}(-0.02,-0.10)(-0.02,0.28)
+    T_6
+    \tikzmarkend{T6-a}
+}
+&
+\lkg_6 \coloneqq \LkgAdd{\lkg_5}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-a}(-0.02,-0.10)(-0.02,0.25)
+    t_6
+    \tikzmarkend{t6-a}
+}
+%&
+%\goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
+%&
+%\goodTerm{x : \TyTkg{\lsig_5}}{\seal_6}{{
+%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-a}(-0.02,-0.10)(-0.05,0.28)
+%    A_6
+%    \tikzmarkend{A6-a}
+%}}
+\\[2pt]
 \cellcolor{Bkg-color-2}
@@
 &
-\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{
-    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-a}(-0.02,-0.10)(-0.02,0.28)
-    T_{10}
-    \tikzmarkend{T10-a}
+\multicolumn{4}{@{}l@{}}{
+\goodTerm{
+    \mathit{x} :
+    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P10-c}(0.00,-0.32)(0.00,0.47)%
+    \!\left[{%
+    \begin{minipage}{.31\textwidth}%
+    \vspace{-7.5pt}%
+    \fontsize{8.0}{9.2}\selectfont%
+    \begin{gather*}
+    \begin{array}{@{}l@{}}
+        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
+        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
+        \dadada,
+        \smallskip\\
+        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
+        \dadada
+    \end{array}
+    \end{gather*}
+    \end{minipage}
+    }\right]\!
+    \tikzmarkend{P10-c}%
+}{
+    \seal_{10}
+}{
+    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-c}(0.00,-0.32)(0.00,0.47)%
+    \!\left[{%
+    \begin{minipage}{.31\textwidth}%
+    \vspace{-7.5pt}%
+    \fontsize{8.0}{9.2}\selectfont%
+    \begin{gather*}
+    \begin{array}{@{}l@{}}
+        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
+        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
+        \dadada,
+        \smallskip\\
+        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
+        \dadada
+    \end{array}
+    \end{gather*}
+    \end{minipage}
+    }\right]\!
+    \tikzmarkend{A10-c}%
 }
-&
-\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-a}(-0.02,-0.10)(-0.02,0.28)
-    t_{10}
-    \tikzmarkend{t10-a}
 }
-&
-\goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
-&
-\goodTerm{x : \TyTkg{\lsig_9}}{\seal_{10}}{{
-    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-a}(-0.02,-0.10)(-0.05,0.28)%
-    A_{10}
-    \tikzmarkend{A10-a}
-}{}}
-\\
+\\[2pt]
 \cellcolor{Bkg-color}
-\multirow{-2}{*}{%
-    \lsti!End!\ \ %
-    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-c}(0.01,-0.10)(-0.02,0.28)%
-    \lsti!subst!%
-    \tikzmarkend{t10-c}%
-    \lsti!.!%
-}
 &
@@
     \begin{gather*}
-    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.45)(-0.02,0.60)
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.45)(-0.02,0.57)
     \begin{array}{@{}l@{}}
-        \El{\mathit{self}.\texttt{tm}} \to
+        \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
         \\
-        \El{\mathit{self}.\texttt{tm}} \to
+        \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
         \\
-        \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
+        \mathit{id} \to \El{\getfield{\mathit{self}}{\texttt{tm}}}
     \end{array}
@@
 }
-\\
+\\[2pt]
+\cellcolor{Bkg-color}
+\multirow{-5}{*}{%
+    \lsti!End!\ \ %
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-c}(0.01,-0.10)(-0.02,0.28)%
+    \lsti!subst!%
+    \tikzmarkend{t10-c}%
+    \lsti!.!%
+}
+&
+\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{x.\seal_{10}}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-a}(-0.02,-0.10)(-0.02,0.28)
+    T_{10}
+    \tikzmarkend{T10-a}
+}
+&
+\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
+    \mathit{self}.\,
+    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-a}(-0.02,-0.10)(-0.02,0.28)
+    t_{10}
+    \tikzmarkend{t10-a}
+}
+%&
+%\goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
+%&
+%\goodTerm{x : \TyTkg{\lsig_9}}{\seal_{10}}{{
+%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-a}(-0.02,-0.10)(-0.05,0.28)%
+%    A_{10}
+%    \tikzmarkend{A10-a}
+%}{}}
+\\[2pt]
 \cellcolor{Bkg-color-2}
@@
 
+%\begin{tabular}{ll}
+%  \hline
+%  $\TyTkg{\lsig_1} = \nTySigma{\_}{\top}{\TyS{\wcode{\wsig_{\texttt{tm}}}}} $ &  
+%  $\TyTkg{\lsig_2} = \nTySigma{\_}{\TyTkg{\lsig_1}}{\El{\wcode{\wsig_{\texttt{tm}}}}}$ \\  
+%  $\TyTkg{\lsig_6} = \nTySigma{P_5}{\TyTkg{\lsig_5}}{\CaseSig{\top}{\bot}{\El{\getfield{P_5}{"tm"}} \to \mathit{id} \to \El{\getfield{P_5}{"tm"}}}}$ & \\ 
+%  $\TyTkg{\lsig_{10}} = \nTySigma{\_}{\TyTkg{\lsig_9}}{\El{\wcode{\wsig_{\texttt{tm}}}} \to \El{\wcode{\wsig_{\texttt{tm}}}} \to \mathit{id} \to \El{\wcode{\wsig_{\texttt{tm}}}}}$ & \\ 
+%\end{tabular}
+
 \caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.

### 2023-03-23 metatheory2.tex
edit

--- 2023-03-23 metatheory2.tex
@@
 % Deal with intuition
-\noindentparagraph{The necessity of $\TyTkg{\cdot}$.} 
-According to the equality rules of $\ruleref{tyeq/pk/add}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is structurally similar to $\TyLkg{\lsig}$ but we are using dependent tuple $\pair{\cdot}{\cdot}$ instead of $\LkgAdd{\cdot}{\cdot}$ to package the fields of a linkage.
-We have computed several examples in \cref{fig:stlc-linkage-typing}.
+\noindentparagraph{Why $\TyTkg{\cdot}$?}
 
-Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
+%As shown by the equality rule \ruleref{tyeq/pk/add},
+%for a given linkage signature $\lsig \coloneq \LSigAdd{\cdot}{\cdot}{\cdot}$,
+%the type $\TyTkg{\lsig}$ is structurally similar to the linkage type $\TyLkg{\lsig}$
+%except that $\TyTkg{\lsig}$ is a dependent tuple type---%
+%it packages the fields of $\lsig$ into a dependent tuple type.
+%
+%We have computed several examples in
+%\cref{fig:stlc-linkage-typing}, where the reader should also notice the high
+%similarity between $\TyTkg{\lsig}$ and $\TyLkg{\lsig}$.
+
+$\TyTkg{\lsig}$ is needed in rules $\ruleref{lsig/add}$ and $\ruleref{l/add}$:
+these rules have a premise  $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
+responsible for possibly hiding W-type signatures (in $\TyTkg{\lsig}$) behind $\cU$ (in $A$);
+see $\seal_6$ in \cref{fig:stlc-linkage-typing} for an example.
+Here, both $\TyTkg{\lsig}$ and $A$ are dependent tuple types rather than linkage types.
+
+%Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
 $\TyTkg{\cdot}$ is indispensible because \emph{we cannot do abstraction with $\TyLkg{\cdot}$ alone}, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{lsig/add}$ and $\ruleref{l/add}$ is \emph{usually a dependent tuple type}.
 
+An alternative to $\TyTkg{\lsig}$ would be to use linkage types $\TyLkg{\cdot}$ in this premise.
+However, this alternative is infeasible.
+We illustrate using a simple example.
+Consider the linkage signature $\lsig$ below (it is written using the record
+syntax for illustration but should not be confused with records, as a later field
+is universally quantified over a $\mathit{self}$ variable).
+The job is to hide the W-types $\TyS{\wcode{\wsig_1}}$ and
+$\TyS{\wcode{\wsig_2}}$ in $\lsig$ behind $\cU$ in $\lsig^A$:
+%
 \begin{align*}
-  \lsig &\coloneq \{ T_A : \TyS{\wcode{\wsig_1}}; &&T_B : \TyS{\wcode{\wsig_2}}; &&f : \El{T_A} \to \El{T_B} \} \\
-  \lsig^A &\coloneq \{T_A : \cU; &&T_B : \cU; &&f : \El{T_A} \to \El{T_B} \}
+  \lsig &\coloneq \{ T_A : \TyS{\wcode{\wsig_1}}; &&T_B : \TyS{\wcode{\wsig_2}}; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
+  \lsig^A &\coloneq \{T_A : \cU; &&T_B : \cU; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
+  & && \goodTerm{x : \TyLkg{\lsig}}{\seal}{\TyLkg{\lsig^A}}
 \end{align*}
 
-We illustrate with a small example using named convention, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function $f$ maps from the first to the second type.
+%We illustrate with a small example using named convention, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function $f$ maps from the first to the second type.
 Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{\lsig}}$.
-Say we want to add a new field into this linkage and abstracting both extensible inductive types, then in $\ruleref{l/add}$, the chosen $A$ must abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
-The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction.
+Say we want to add a new field into this linkage, and in $\ruleref{l/add}$, we chose $A$ to abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
+The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction (choose $s$).
 
-\begin{align*}
-  & \text{doing abstraction} \\ 
-  \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
-  \text{Approach } \beta \quad&\quad  \TyLkg{\lsig} \longrightarrow \TyTkg{\lsig} \longrightarrow \TyTkg{\lsig^A}\\
-\end{align*}
+The problem lies in that the type $\TyLkg{\lsig^A}$ above cannot uninhabited.
+Suppose there is some $\lkg^A$ such that $\cdot \vdash \lkg^A : \TyLkg{\lsig^A}$.
+Then it follows from the self-parameterization that projecting out the field~$f$
+has the typing
+$\goodTerm{T_A : \cU, T_B: \cU}{\getfield{\lkg^A}{f}}{\El{T_A} \to \El{T_B}}$.
+This typing means that there is a function of type
+$\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A}
+\to \El{T_B}}}$, i.e., a function from an arbitrary type to another arbitrary
+type. We can easily derive inconsistency from this function:
+we can apply it to $\top$ and $\bot$ and get $\top \to \bot$.
 
-However, the seemingly easier approach $\alpha$ is impossible because linkage $\TyLkg{\lsig^A}$ is uninhabited---say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
+%\begin{align*}
+%  & \text{doing abstraction (choosing appropriate $s$ in \ruleref{l/add})} \\ 
+%  \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
+%  \text{Approach } \beta \quad&\quad  \TyLkg{\lsig} \longrightarrow \TyTkg{\lsig} \longrightarrow \TyTkg{\lsig^A}\\
+%\end{align*}
+%
+%However, the seemingly easier approach $\alpha$ is impossible because linkage
+%$\TyLkg{\lsig^A}$ is uninhabited---%
+%say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
 To make the inconsistency clearer, by double abstraction we will have an impossible function $g \coloneq \goodTerm{\cdot}{\nlam{T_A}{\nlam{T_B}{\ell^A.f}}}{\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A} \to \El{T_B}}}} $, i.e.
 a function maps from arbitrary type to another arbitrary type .
-We can easily derive inconsistency setting applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
+We can easily derive inconsistency by applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
+%
+%Thus we can only choose the approach $\beta$ instead, which justifies why we have $\TyTkg{\cdot}$ in our typing rule, as $\TyTkg{\lsig}$ is usually the first step to do abstraction.
 
-Thus we can only choose the approach $\beta$ instead, which justifies why we have $\TyTkg{\cdot}$ in our typing rule, as $\TyTkg{\lsig}$ is usually the first step to do abstraction.
+Thus, it is critical that the hiding of W-type signatures operate on dependent
+tuple types rather than linkage types.
+In \TT, we introduce the syntax $\TyTkg{\lsig}$, $\Tkg{\lkg}$, and their computation rules to make
+this hiding process more succinct.
 
-To summarize, the reason we need dependent tuple type is because of the overridability nature of our linkages, together with the frequent usage of universe type in the context of dependent type programming.
+%To summarize, the reason we need dependent tuple type is because of the overridability nature of our linkages, together with the frequent usage of universe type in the context of dependent type programming.
 Once we want to abstract a type in a signature, we may have to use dependent tuple type as the abstraction instead of linkages.

### 2023-03-23 lang-design.tex
edits

@@
 the four cases from "STLC" are automatically inherited and reused,
-with "STLCFix" retroactively adding a fifth case to form the new "subst" function.
+with "STLCFix" adding a fifth case retroactively to form a new "subst" function.
 %

### 2023-03-23 metatheory-condensed.tex
revamp figure 8
edits

@@
 \definecolor{t-color}{HTML}{FAE0E4}
+\definecolor{P-color}{HTML}{BFEFFF}
 
@@
 we write them as dependent record types that give labels to the fields.
-Field accesses are notated, for example, $\getfield{\mathit{self}}{"tm"}$.
+Field accesses are notated, for example, as $\getfield{\mathit{self}}{"tm"}$.
 %we use a notation with explicit names
@@
 The term $\seal$ turns a tuple of type $\TyTkg{\lsig}$ into a new tuple of type~$A$
-that hides W-type signatures behind $\cU$, if necessary.
+that hides W-type signatures behind $\cU$, if necessary.\footnote{%
+In the presence of hiding, it is important that the context type~$A$ be a
+dependent tuple type rather than a linkage type; we expand on this issue in an
+appendix.}
 %
@@
 In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are simply $x$,
-and $\seal_6$ has the typing
-\begingroup
-\fontsize{9}{9}\selectfont%
-\goodTerm{
-  x :
-  \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
-}{
-  \seal_6
-}{
-  \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
-}.
-\endgroup
+while $\seal_6$ hides $"tm" : \TyS{\wcode{\wsig_{\texttt{tm}}}}$
+as $"tm" : \cU$ in $A_6$
+so that the typing of~$t_6$ is oblivious to the concrete signature
+$\wsig_{\texttt{tm}}$ and therefore can be reused.
+%\begingroup
+%\fontsize{9}{9}\selectfont%
+%\goodTerm{
+%  x :
+%  \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+%}{
+%  \seal_6
+%}{
+%  \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
+%}.
+%\endgroup
 

### 2023-03-23 metatheory2.tex
Question, for double checking.

@@
 
-The problem lies in that the type $\TyLkg{\lsig^A}$ above cannot uninhabited.
+The problem lies in that the type $\TyLkg{\lsig^A}$ above cannot be inhabited.
 Suppose there is some $\lkg^A$ such that $\cdot \vdash \lkg^A : \TyLkg{\lsig^A}$.
@@
 has the typing
-$\goodTerm{T_A : \cU, T_B: \cU}{\getfield{\lkg^A}{f}}{\El{T_A} \to \El{T_B}}$.
+$\goodTerm{T_A : \cU, T_B: \cU}{\getfield{\lkg^A}{f}}{\El{T_A} \to \El{T_B}}$\EDJ{This field projection should be the usual one right? because ellA here is a linkage}.
 This typing means that there is a function of type

### 2023-03-23 metatheory-condensed.tex
address comment

@@
 In the presence of hiding, it is important that the context type~$A$ be a
-dependent tuple type rather than a linkage type; we expand on this issue in an
+dependent tuple type rather than a linkage type; we expand on this point in an
 appendix.}

### 2023-03-23 metatheory2.tex
address comment

@@
 Suppose there is some $\lkg^A$ such that $\cdot \vdash \lkg^A : \TyLkg{\lsig^A}$.
-Then it follows from the self-parameterization that projecting out the field~$f$
+Then it follows from the self-parameterization that projecting out the last field~$f$
 has the typing
-$\goodTerm{T_A : \cU, T_B: \cU}{\getfield{\lkg^A}{f}}{\El{T_A} \to \El{T_B}}$\EDJ{This field projection should be the usual one right? because ellA here is a linkage}.
+$\goodTerm{T_A : \cU, T_B: \cU}{\lkgproj{2}{\lkg^A}}{\El{T_A} \to \El{T_B}}$.
 This typing means that there is a function of type
@@
 tuple types rather than linkage types.
-In \TT, we introduce the syntax $\TyTkg{\lsig}$, $\Tkg{\lkg}$, and their computation rules to make
-this hiding process more succinct.
+To this end, we introduce the syntax $\TyTkg{\lsig}$ and its computation rules
+(e.g., \ruleref{tyeq/pk/add}) to make it convenient to package linkage signatures
+into dependent tuple types.
 

### 2023-03-24 metatheory2.tex
cover letter
tweak

@@
 % Deal with intuition
-\noindentparagraph{Why $\TyTkg{\cdot}$?}
+\noindentparagraph{Why $\TyTkg{\lsig}$?}
 
@@
 Then it follows from the self-parameterization that projecting out the last field~$f$
-has the typing
-$\goodTerm{T_A : \cU, T_B: \cU}{\lkgproj{2}{\lkg^A}}{\El{T_A} \to \El{T_B}}$.
+from $\lkg^A$ (notated as $\lkg^A.f$ for illustration) has the typing
+$\goodTerm{T_A : \cU, T_B: \cU}{\lkg^A.f}{\El{T_A} \to \El{T_B}}$.
 This typing means that there is a function of type
@@
 type. We can easily derive inconsistency from this function:
-we can apply it to $\top$ and $\bot$ and get $\top \to \bot$.
+applying it to $\top$ and $\bot$ yields a function of type $\top\to\bot$.
 

### 2023-03-24 metatheory-condensed.tex
edit

@@
 appendix.}
-%
+
 It is straightforward to find the~$\seal$ that fits the bill, though this process
 is not automated in \TT.
-In particular, when no hiding is needed, $\seal$ is simply~$x$,
+In particular, when no hiding is needed (that is, when the field being checked
+invokes a W-type constructor or eliminator), $\seal$ is simply~$x$,
+and $A$ is $\TyTkg{\lsig}$.
+In \cref{fig:stlc-linkage-typing}, $\seal_2$ and $\seal_{10}$ are~$x$.
 %\EDJ{$\var{0}$ is not identity (substitution), but the first variable, right?
@@
 I cannot say 'sealing' because the paper currently does not use the word sealing anywhere.}\EDJreply{What about just remove the parenthesis and ``the identity''}
-and $A$ is $\TyTkg{\lsig}$.
-In \cref{fig:stlc-linkage-typing}, $\seal_1$, $\seal_2$, and $\seal_{10}$ are simply $x$,
-while $\seal_6$ hides $"tm" : \TyS{\wcode{\wsig_{\texttt{tm}}}}$
-as $"tm" : \cU$ in $A_6$
-so that the typing of~$t_6$ is oblivious to the concrete signature
+Otherwise, it is needed to hide W-type signatures. In \cref{fig:stlc-linkage-typing},
+$\seal_6$ hides $"tm" : \TyS{\wcode{\wsig_{\texttt{tm}}}}$
+as $"tm" : \cU$ in~$A_6$,
+so that~$t_6$, typed under $\mathit{self}:A_6$, is oblivious to the concrete signature
 $\wsig_{\texttt{tm}}$ and therefore can be reused.

### 2023-03-24 metatheory2.tex
Questions/Comments. add back mentioning of based path induction.

@@
 applies~$\gm$ to the type~$T$) are part of the syntax rather than
-meta-operations. 
-We work in an intrinsically typed setting: terms are well typed by construction.
+meta-operations. We specifically use based path induction
+as eliminator for identity type~\cite{program2013homotopy}.
+We work in an intrinsically typed setting: terms are well typed by construction, 
+%
+and consequently some premises are omitted without ambiguity and can be inferred from the context. \EDJ{I added this sentence back here. Why removed this sentence earlier?}
+% 
 This style of syntax formulation follows a recent
@@
 Similar for type judgement $\goodType{\Gm}{T}{j}$.
+\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}
 

### 2023-03-24 metatheory2.tex
replies

@@
 %
-and consequently some premises are omitted without ambiguity and can be inferred from the context. \EDJ{I added this sentence back here. Why removed this sentence earlier?}
+and consequently some premises are omitted without ambiguity and can be inferred from the context.
+\EDJ{I added this sentence back here.
+Why removed this sentence earlier?}\YZreply{I don't find a strong causal connection between the two parts of the sentence}
 % 
@@
 Similar for type judgement $\goodType{\Gm}{T}{j}$.
-\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}
+\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}
 

### 2023-03-24 metatheory2.tex
Phrasing.

@@
 applies~$\gm$ to the type~$T$) are part of the syntax rather than
-meta-operations. We specifically use based path induction
-as eliminator for identity type~\cite{program2013homotopy}.
+meta-operations.
 We work in an intrinsically typed setting: terms are well typed by construction, 
@@
 %(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.
+We use based path induction
+as the elimination principle for identity types~\cite{program2013homotopy}.
 
@@
 
-Consequently, function application changes to a equivalent formulation, and becomes a ``direct inverse'' of typing rule for function abstraction.
+Consequently, function application changes to an equivalent formulation, and becomes a ``direct inverse'' of typing rule for function abstraction.
 For example, the named notation $\napp{f}{t}$ can be equivalently represented by $\sub{\app{f}}{\SubstExt{\SubstId}{t}}$.
@@
 Finally, we have \ruleref{sub/dbj/shift} defined using \ruleref{sub/wk} and \ruleref{tm/var}.
-This rule applying substitution $\gm$ on the earlier part of the context.
+This rule applying substitution $\gm$ to the earlier portion of the context.
 We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.

### 2023-03-24 metatheory2.tex
Does Section 6 not use based path induction?
Reply comments.

@@
 \EDJ{I added this sentence back here.
-Why removed this sentence earlier?}\YZreply{I don't find a strong causal connection between the two parts of the sentence}
+Why removed this sentence earlier?}\YZreply{I don't find a strong causal connection between the two parts of the sentence}\EDJreply{If it is not intrinsically typed (everything is raw), then no inference can be done.}
 % 
@@
 Similar for type judgement $\goodType{\Gm}{T}{j}$.
-\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}
+\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}\EDJreply{Do you think ``omitting premises'' should be mentioned here or earlier?}
 

### 2023-03-24 canonicitymodel.tex
Fix typo.
Fix typo on identity type.
BugFix.

@@
 Our canonicity model for the base MLTT fragment is close to the ones
-in \citet{coquand2018canonicity} and \citet{sterling2019algebraic}.
+in \citet{coquand2018canonicity} and \citet{sterling2019algebraic}, where we utilize the machinary from ambient meta-logic (e.g.
+use dependent function and dependent tuple in ambient meta-logic to model the one in \TT).
 % We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
@@
   \denotesC{\goodType{\Gm}{\sub{T}{\sigma}}{}}(\gm)(\gm')(t) &= \denotesC{T}(\sigma \circ \gm)(\denotes{\sigma}(\gm)(\gm'))(t)\\
-  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \} \\
+  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \} \quad \text{  a singleton set } \\
   \denotesC{\goodType{\Gm}{\bot}{}}(\gm)(\gm')(t) &= \emptyset \\
@@
   \end{cases}\\
-  % \setc{\star}{\text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm}} %\YZ{What if the conditions do not hold?
-Empty set?}\EDJreply{Yes.
-It is a family of set indexed by t (or just a function maps t to a set), and if t is neither of the above, then the function will map to empty set.}
-  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})(\gm', u')(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
+  \denotesC{\goodTerm{\Gm}{\Jeq{w}{t}}{\sub{C}{\SubstExt{\SubstExt{id}{v}}{t} }}}(\gm)(\gm') &= \denotesC{w}(\gm)(\gm') \quad \quad \text{ given } \goodTerm{\Gm}{t}{\TyId{u}{v}} \\ 
+  & \text{Since $\denotesC{t}(\gm)(\gm')$ witnesses } \sub{t}{\gm} \equiv \eqrefl{\sub{u}{\gm}} \text{ and } \sub{u}{\gm} \equiv \sub{v}{\gm} \\
+  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
   \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}}(\gm)(\gm')(t) &= \sum_{u' \in \denotesC{A}(\gm)(\gm')(\fst{t})} \denotesC{B}(\SubstExt{\gm}{\fst{t}})((\gm', u'))(\snd{t}) \\
-  \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 
+  \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\}  \\ 
   \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \setc{ (\gamma' , t') }{ \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  } \\
+  \denotesC{\goodSub{\Gm}{\pi_1 \delta}{\Dl}}(\gm)(\gm') &= \denotesC{\delta}(\gm)(\gm')[0] \quad \text{ get the first element of the tuple} \\
+  \denotesC{\goodTerm{\Gm}{\pi_2 \delta}{\Dl}}(\gm)(\gm') &= \denotesC{\delta}(\gm)(\gm')[1] \quad \text{ get the second element of the tuple} \\
+  \denotesC{\goodSub{\Gm}{\SubstExt{\delta}{t}}{\Dl}}(\gm)(\gm') &= (\denotesC{\delta}(\gm)(\gm'),\denotesC{t}(\gm)(\gm')) \\ 
   \denotesC{\goodTerm{\Gm}{\sub{t}{\sigma}}{\sub{T}{\sigma}}}(\gm)(\gm') &= \denotesC{t}(\sigma \circ \gm)(\denotesC{\sigma}(\gm)(\gm')) \\
@@
   \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}}(\gm)(\gm') &= (\denotesC{a}(\gm)(\gm') , \denotesC{b}(\gm)(\gm') )\\
-  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[0] \\
+  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[0]  \quad \text{ extract the first element in the tuple }\\
   \denotesC{\goodTerm{\Gm}{\snd{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[1]\\
@@
 \newcommand{\Gwcode}[1]{\ensuremath{\Glued{\texttt{W}}({#1})}}
-\newcommand{\GSubstId}{\ensuremath{\GSubstWeak{0}}}
+\newcommand{\GSubstId}{\ensuremath{\Glued{"id"}}}
 \newcommand{\Gwsigproj}[3]{\ensuremath{\Glued{{\texttt{w}\pi^{#1}_{\texttt{#2}}}}({#3})}}
@@
   % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
-  & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
+  \text{we define }\TyLkg{}, \TyTkg{} & \text{ inductively on its input signature, by defining } \mathit{L},\Glued{\mathit{P}} \text{ first}  \\ 
   \mathit{L}("nil") &=  \denotesC{\top}  \\
@@
   \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
-  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesCC{A}, \denotesCC{s}, \denotesCC{T})::\denotesC{\lsig}\\ 
+  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesCC{A}, \denotesCC{s}, \denotesCC{T})::\denotesC{\lsig} \quad \text{ given } \goodTerm{\Gm, \TyTkg{\lsig}}{s}{A} \\ 
   \denotesC{\goodSig{\Gm}{\sub{\lsig}{\gm}}{n}} & \text{ is done by point-wise/component-wise substitution} \\
@@
   \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
-  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{(\SubstId)}}{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}}}}
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}}}}
   \\ &\text{ when} \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\
@@
 
-The semantic interpretations above are 
-mutually recursively defined with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)
+The semantic interpretations above are defined together with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)
 %
@@
     & \quad \quad \to 
-    \goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{\SubstId}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}} \\ 
-    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
+    \goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}} \\ 
+    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
     & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
@@
 
-Note that in $\CWmodel$, the $\Glued{b}$ uses the definition of
+Note that in $\CWsup$, the $\Glued{b}$ uses the definition of
 $\denotesC{\wcode{\wsig}}$, which after unfolding, recursively references $\CWmodel$

### 2023-03-24 canonicitymodel.tex
Bug fix. Fix alignment.
Fix alignment
Bugfix.

@@
   \denotesC{\goodSub{\Gm}{\SubstExt{\delta}{t}}{\Dl}}(\gm)(\gm') &= (\denotesC{\delta}(\gm)(\gm'),\denotesC{t}(\gm)(\gm')) \\ 
+  \denotesC{\goodSub{\Gm}{"id"}{\Gm}}(\gm)(\gm') &= \gm' \\
+  \denotesC{\goodSub{\Gm}{\delta_1 \circ \delta_2}{\Dl}}(\gm)(\gm') &= \denotesC{\delta_1}(\SubstComp{\delta_2}{\gm})(\denotesC{\delta_2}(\gm)(\gm')) \\
   \denotesC{\goodTerm{\Gm}{\sub{t}{\sigma}}{\sub{T}{\sigma}}}(\gm)(\gm') &= \denotesC{t}(\sigma \circ \gm)(\denotesC{\sigma}(\gm)(\gm')) \\
@@
 is part of $\GTy{\!}{\!}$, $\GTm{\!}{\!}$, and $\GSub{\!}{\!}$.
-An interesting fact about them is : given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$
+A useful fact about them is : given a pair of arbitrary $\goodSub{\cdot}{\gm}{\Gm}$
 and $\gm' \in \Gmodel{\Glued{\Gm}}(\gm)$, we can consider $(\gm, \gm')$ as an
-element of $\GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
-Thus, we can consider the pair $(\gm,
-\gm')$ as an element $\Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
+element of $\GSub{\Glued{\cdot}}{\Glued{\Gm}}$, and vice versa.
+Thus, we consider the pair $(\gm,
+\gm')$ the equivalent form of an element $\Glued{\gm} \in \GSub{\Glued{\cdot}}{\Glued{\Gm}}$.
 
@@
   \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
-  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm)(\gm') &= (\denotesC{\lkg}(\gm)(\gm') , \Gmodel{\Glam{\denotesCC{t}}}(\gm)(\gm')) \\
+  \denotesC{{\LkgAdd{\lkg}{t}}}(\gm)(\gm') &= (\denotesC{\lkg}(\gm)(\gm') , \Gmodel{\Glam{\denotesCC{t}}}(\gm)(\gm')) \\
   \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
   \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
-  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}}}}
-  \\ &\text{ when} \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\Glued{\ell'}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}}}
+  \\ &\text{ where } \Glued{\ell'} = \denotesCC{\Tkg{\lkgproj{1}{\lkg}}} 
+  \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\
   \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyLkg{\lsig}}}(\gm)(\gm') &= \denotesC{\ell}(\gm)(\gm')[0] \quad \text{ take the first element in the tuple} \\
@@
   \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}}(\gm)(\gm')(t) &= \CWmodel~(\Gsub{\denotesCC{\wsig}}{\Glued{\gm}})~t \\
-  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~(\Gsub{\denotesCC{\wsig}}{\Glued{\gm}})~(\Gsub{\denotesCC{a}}{\Glued{\gm}})~(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}}) \\
+  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~(\Gsub{\denotesCC{a}}{\Glued{\gm}})~(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}}) \\
   % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
@@
   \denotesC{\goodType{\Gm}{\RecSig{\wsig}{"R"}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{"R"}} \\
-  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg \\
-  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') &= \CWrec~(\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\wsig}{\sub{\lkg}{\gm'}}{w}))~f^r~\sub{t}{\gm'}~(\denotesC{t}(\gm)(\gm')) \\
+  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg 
+\end{align*}
+\begin{align*}
+  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') = \CWrec \quad &\Gsub{\denotesCC{\wsig}}{\Glued{\gm}} \\ 
+  & (\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\sub{\wsig}{\gm}}{\sub{\lkg}{\gm}}{w}))\\
+  &f^r\\ &\sub{t}{\gm}\\ 
+  &(\denotesC{t}(\gm)(\gm')) 
+\end{align*}
+\begin{align*}  
   \text{ where } \quad f^r~j~\Glued{a}~b~b^C &= 
@@
   \\
-  & \ \texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^C)~\texttt{ in } \\
+  &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^C)~\texttt{ in } \\
   & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
   & \quad 
-   \text{ and also } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \quad ,\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\Glued{\wsig}}
+   \text{ given } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \\
+   & \quad \text{ where }\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\Glued{\wsig}}
 \end{align*}
@@
     (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
-    & \to [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
+    & \to 
+    [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
     & \quad \quad \to 
     \goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}} \\ 
-    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
+    & \quad \quad \to b^C : (\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}})} \to P~(\sub{b}{ \gm })) \\
     & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\

### 2023-03-24 canonicitymodel.tex
Bugfix.

@@
 Our canonicity model for the base MLTT fragment is close to the ones
-in \citet{coquand2018canonicity} and \citet{sterling2019algebraic}, where we utilize the machinary from ambient meta-logic (e.g.
-use dependent function and dependent tuple in ambient meta-logic to model the one in \TT).
+in \citet{coquand2018canonicity} and \citet{sterling2019algebraic} in that it
+utilizes the facilitates of the ambient meta\-logic; for example, we use dependent functions and
+dependent tuples in the ambient meta\-logic to model dependent functions and tuples in \TT.
 % We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model).
@@
   \denotesC{\goodSub{\Gm}{"id"}{\Gm}}(\gm)(\gm') &= \gm' \\
+  \denotesC{\goodSub{\Gm}{\SubstEmp}{\cdot}}(\gm)(\gm') &= \star \\
   \denotesC{\goodSub{\Gm}{\delta_1 \circ \delta_2}{\Dl}}(\gm)(\gm') &= \denotesC{\delta_1}(\SubstComp{\delta_2}{\gm})(\denotesC{\delta_2}(\gm)(\gm')) \\
@@
 As before, we can define glued interpretation $\GWSig{j}{\Glued{\Gm}}{n}
-\coloneq \{ \wsig : \goodWSig{\cdot}{\wsig}{n} \} \times
+\coloneq 
+% \{ \wsig : \goodWSig{\cdot}{\wsig}{n} \} 
+\setc{\wsig}{\goodWSig{\cdot}{\wsig}{n}}
+\times
 \GMWSig{j}{\Glued{\Gm}}{n}$. This will be useful when we interpret judgments for
@@
   % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
-  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= 
-  \Gmodel{\GluedPi{\denotesC{A}}{\GluedPi {\GluedPi{\denotesC{B}}{\Gsub{\denotesC{R}}{\GSubstWeak{2}}}} {\Gsub{\denotesC{R}}{\GSubstWeak{2}}}}} \\
+  \denotesC{\goodType{\Gm}{\CaseSig{"A"}{"B"}{"R"}}{}} &= 
+  \Gmodel{\GluedPi{\denotesCC{"A"}}{\GluedPi {\GluedPi{\denotesCC{"B"}}{\Gsub{\denotesCC{"R"}}{\GSubstWeak{2}}}} {\Gsub{\denotesCC{"R"}}{\GSubstWeak{2}}}}} \\
   % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\

### 2023-03-24 metatheory2.tex
resolve a comment

@@
 meta-operations.
-We work in an intrinsically typed setting: terms are well typed by construction, 
+We work in an intrinsically typed setting: terms are well typed by construction.
 %
-and consequently some premises are omitted without ambiguity and can be inferred from the context.
-\EDJ{I added this sentence back here.
-Why removed this sentence earlier?}\YZreply{I don't find a strong causal connection between the two parts of the sentence}\EDJreply{If it is not intrinsically typed (everything is raw), then no inference can be done.}
+Consequently, we omit without ambiguity some obvious premises needed for well-formedness.
 % 

### 2023-03-24 metatheory2.tex
reply

@@
 Similar for type judgement $\goodType{\Gm}{T}{j}$.
-\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}\EDJreply{Do you think ``omitting premises'' should be mentioned here or earlier?}
+\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}\EDJreply{Do you think ``omitting premises'' should be mentioned here or earlier?}\YZreply{It's not just about omitting premises, right?
+You also omit universal levels in some typing rules.}
 

### 2023-03-24 canonicitymodel.tex
edit
Bugfix.

@@
 properly~\cite{coquand2018canonicity}.
-Our canonicity model for the base MLTT fragment is close to the ones
+Our canonicity model for the base MLTT fragment is following the ones
 in \citet{coquand2018canonicity} and \citet{sterling2019algebraic} in that it
@@
 \begin{align*}  
-  \text{ where } \quad f^r~j~\Glued{a}~b~b^C &= 
+  \text{ where } \quad f^r~j~\Glued{a}~b~b^c &= 
   \texttt{let}~\Glued{\rho}  \in \GTm{(\Glued{\cdot} \Glued{,} \Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}})}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})} 
   \\
-  &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^C)~\texttt{ in } \\
+  &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^c)~\texttt{ in } \\
   & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
@@
     \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
-    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
+    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}) \\
+    & \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
 \end{align*}
@@
     & \quad \quad \to 
-    \goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}} \\ 
-    & \quad \quad \to b^C : (\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}})} \to P~(\sub{b}{ \gm })) \\
+    \setc{b}{\goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
+    & \quad \quad \to  \prod_{\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}})}} P~(\sub{b}{ \gm }) \\
     & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
     & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
-  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~P~f~(\sub{b}{\Gmodel{\Glued{\gm}}})~(b^C~\Glued{\gm}))
+  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~P~f~(\sub{b}{\gm})~(\Gmodel{\Glued{b}}~\Glued{\gm}))
 \end{align*}
@@
 in a strictly positive position. %So it is a well-founded definition.
-We do not distinguish $(b, b^C)$ and $\Glued{b}$ for simplicity. Again, we omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
+We do not distinguish $(b, b^c)$ and $\Glued{b}$ for simplicity. Again, we omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
 

### 2023-03-24 metatheory2.tex
Resolve Comment.

--- 2023-03-24 metatheory2.tex
@@
 Similar for type judgement $\goodType{\Gm}{T}{j}$.
-\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}\EDJreply{Do you think ``omitting premises'' should be mentioned here or earlier?}
+% \EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}\EDJreply{Do you think ``omitting premises'' should be mentioned here or earlier?}\YZreply{It's not just about omitting premises, right?
+You also omit universal levels in some typing rules.}\EDJreply{Actually now I am not sure in our formulation, the omitted universe level $k$ is really used anywhere.
+I think I am okay with commenting this sentence.}
 

### 2023-03-24 canonicitymodel.tex
tweak

@@
 properly~\cite{coquand2018canonicity}.
-Our canonicity model for the base MLTT fragment is following the ones
+Our canonicity model for the base MLTT fragment follows the constructions
 in \citet{coquand2018canonicity} and \citet{sterling2019algebraic} in that it

### 2023-03-25 canonicitymodel.tex
Fix position.
Fix canonicity model.
bug fix.

@@
   \end{cases}\\
-  \denotesC{\goodTerm{\Gm}{\Jeq{w}{t}}{\sub{C}{\SubstExt{\SubstExt{id}{v}}{t} }}}(\gm)(\gm') &= \denotesC{w}(\gm)(\gm') \quad \quad \text{ given } \goodTerm{\Gm}{t}{\TyId{u}{v}} \\ 
-  & \text{Since $\denotesC{t}(\gm)(\gm')$ witnesses } \sub{t}{\gm} \equiv \eqrefl{\sub{u}{\gm}} \text{ and } \sub{u}{\gm} \equiv \sub{v}{\gm} \\
   \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
@@
   \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm)(\gm') &= \star\\
+  \denotesC{\goodTerm{\Gm}{\Jeq{w}{t}}{\sub{C}{\SubstExt{\SubstExt{id}{v}}{t} }}}(\gm)(\gm') &= \denotesC{w}(\gm)(\gm') \quad \quad \text{ given } \goodTerm{\Gm}{t}{\TyId{u}{v}} \\ 
+  & \text{Since $\denotesC{t}(\gm)(\gm')$ witnesses } \sub{t}{\gm} \equiv \eqrefl{\sub{u}{\gm}} \text{ and } \sub{u}{\gm} \equiv \sub{v}{\gm} \\
   \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}}(\gm)(\gm') &= \lambda u \lambda u' . \denotesC{t}(\SubstExt{\gm}{u}) (\gm' , u') \\
@@
 glued interpretation in \citet{sterling2019algebraic}.}
-Thus, we can have a
+Then, we can have a
 function $\Glued{\Pi}$ such that $\GluedPi{\denotesCC{A}}{\denotesCC{B}} =
@@
   % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
-  \text{we define }\TyLkg{}, \TyTkg{} & \text{ inductively on its input signature, by defining } \mathit{L},\Glued{\mathit{P}} \text{ first}  \\ 
+  \text{we define }\TyLkg{}, \TyTkg{} & \text{ inductively on its input signature, via } \mathit{L},\Glued{\mathit{P}}  \\ 
   \mathit{L}("nil") &=  \denotesC{\top}  \\
@@
   \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
-  \denotesC{{\LkgAdd{\lkg}{t}}}(\gm)(\gm') &= (\denotesC{\lkg}(\gm)(\gm') , \Gmodel{\Glam{\denotesCC{t}}}(\gm)(\gm')) \\
+  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm)(\gm') &= (\denotesC{\lkg}(\gm)(\gm') , \Gmodel{\Glam{\denotesCC{t}}}(\gm)(\gm')) \\
   \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
@@
   % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\
-  \text{we define } \RecSig{}{} &\text{ by induction on the signature} \\
+  \text{we define } \RecSig{}{} &\text{ by induction on the signature, via } RS \\
   RS~"nil"~R &= \denotesCC{\LSigEmp} \\

### 2023-03-25 canonicitymodel.tex
Phrasing.
Elaborate glued interpretation, first attempt

@@
 \newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^\bullet({#1})}}
-\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^\bullet)^{#1}}}
+\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^{#1})^{\bullet}}}
 \newcommand{\GLSigAdd}[3]{\ensuremath{\nu^{+\bullet}({#1},{#2},{#3})}}
@@
 for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$;
-etc.
+etc. The audience should reminisce these structure with the one appearing at the beginning of the MLTT canonicity model.
 
@@
 W-type signatures.
+
+\textbf{Unfortunately}, at this point, our notation is already highly densed because of frequent appearances of $\bullet$. These burden
 
@@
   \Glued{\mathit{P}}("nil") &= ( \TyTkg{\LSigEmp} , \denotesC{\top} )\\
-  \Glued{\mathit{P}}((A,s,T) :: tl) &= \Glued{\Sigma}(\Glued{\mathit{P}}(tl), \Gsub{T}{\GSubstExt{\Glued{\pi_1}}{s}}) \quad \text{(doing substitution on $T$)} \\
+  \Glued{\mathit{P}}((A,s,T) :: tl) &= \Glued{\Sigma}(\Glued{\mathit{P}}(tl), \Gsub{T}{\GSubstExt{{\GSubstWeak{1}}}{s}}) \quad \text{(doing substitution on $T$)} \\
   \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= {\mathit{L}(\denotesC{\lsig})} \\
@@
   \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
-  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\Glued{\ell'}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}}}
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\Glued{\ell'}}{\Gsub{\Gsub{\denotesCC{\lkgproj{2}{\lkg}}}{\GSubstExt{{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}}}
   \\ &\text{ where } \Glued{\ell'} = \denotesCC{\Tkg{\lkgproj{1}{\lkg}}} 
   \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\ 
+\end{align*}
+This is the first point our proof becomes cryptic because of this bullet notation.
+
+The first complexity comes from the three ``bullets after comma'' notation.
+The way to interpret this notation is that, this ``$\GSubstExt{\_}{\_}$'' should be considered as a binary function requiring two input of glued interpretation.
+
+However, a careful reader will notice that, because of the overloading of the notation of dependent tuple construction and substitution extension, our glued notation is also overloading.
+The first one ``$\Gpair{\Glued{\ell'}}{...}$'' should be the glued interpretation for dependent tuple, but the other two should be the one for substitution extension.
+
+Similarly, the second complexity comes from the bullets after the square bracket, which as expected, is the glued interpretation for substitution.
+Again, ``$\Gsub{\_}{\_}$'' should be considered as a binary function consuming two glued interpretation, corresponding to the substitution operation ``$\sub{\_}{\_}$''.
+
+In all, the whole interpretation should be considered as:
+\begin{enumerate}
+  \item We first get $\denotesCC{\lkgproj{2}{\lkg}}$, the glued interpretation of the projection of the linkage 
+  \item Then we apply it with two substitution,  $\GSubstExt{{\GSubstWeak{1}}}{s}$ and $\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }$
+  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{...}$
+\end{enumerate}
+
+\begin{align*}
   \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyLkg{\lsig}}}(\gm)(\gm') &= \denotesC{\ell}(\gm)(\gm')[0] \quad \text{ take the first element in the tuple} \\
@@
    \text{ given } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \\
-   & \quad \text{ where }\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\Glued{\wsig}}
+   & \quad \text{ where }\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\denotesCC{\wsig}}
 \end{align*}

### 2023-03-25 canonicitymodel.tex
Help the audience parse by using background colour

@@
 
+Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.
+
 In all, the whole interpretation should be considered as:
@@
   \item We first get $\denotesCC{\lkgproj{2}{\lkg}}$, the glued interpretation of the projection of the linkage 
-  \item Then we apply it with two substitution,  $\GSubstExt{{\GSubstWeak{1}}}{s}$ and $\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }$
-  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{...}$
+  \item Then we apply it with two substitution,  $\Gsub{\_}{\GSubstExt{{\GSubstWeak{1}}}{s}}$ and $\Gsub{\_}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}$
+  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{\_}$
 \end{enumerate}
@@
   % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\
+\end{align*}
+We note that, the interpretation of $\Wsup{i}{\wsig}{a}{b}$ is defined by $\CWsup$ (defined below) will take three elements, $i$, a glued $(\Gsub{\denotesCC{a}}{\Glued{\gm}})$ and a glued $(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})$.
+As we mentioned earlier, $\Glued{\gm}$ is the equivalent form of $(\gm, \gm')$.
+
+The interpretation for $\CaseSig{"A"}{"B"}{"R"}$ is also bullet-drowned.
+However, it should be read as the glued interpretation of $ \TyPi{A}{\TyPi {\TyPi{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}} $, which is exactly from the debruinj indxed-version of the typing rule \ruleref{tyeq/casety}.
+
+In the following interpretation, parsing might be hard, so we might use background colour to make things easier to read, and there is no meaning for the choice of colour.
+
+\begin{align*}
   \text{we define } \RecSig{}{} &\text{ by induction on the signature, via } RS \\
@@
   "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
-    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
-    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}) \\
+    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\
+    & \to \Glued{b} \in (\GTm{
+      \colorbox{blue!10}{(\Glued{\cdot} \Glued{,} \Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}})} 
+      }{
+        \colorbox{green!10}{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}
+        }) \\
     & \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
@@
     & \quad \quad \to 
-    \setc{b}{\goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
-    & \quad \quad \to  \prod_{\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}})}} P~(\sub{b}{ \gm }) \\
+    \setc{b}{\goodTerm{({\cdot} , 
+    \colorbox{blue!10}{\sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}}
+    ) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
+    & \quad \quad \to  \prod_{\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} 
+    \colorbox{blue!10}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}} P~(\sub{b}{ \gm }) \\
     & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
     & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
-  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~P~f~(\sub{b}{\gm})~(\Gmodel{\Glued{b}}~\Glued{\gm}))
+  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~\Glued{\wsig}~P~f~(\sub{b}{\gm})~(\Gmodel{\Glued{b}}~\Glued{\gm}))
 \end{align*}

### 2023-03-25 canonicitymodel.tex
Adding several color box to help parsing

@@
   \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
-  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\Glued{\ell'}}{\Gsub{\Gsub{\denotesCC{\lkgproj{2}{\lkg}}}{\GSubstExt{{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}}}
+  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\Glued{\ell'}}{
+    \colorbox{blue!17}{\Gsub{
+    \colorbox{blue!9}{\Gsub{
+    \colorbox{blue!1}{\denotesCC{\lkgproj{2}{\lkg}}}
+    }{\GSubstExt{{\GSubstWeak{1}}}{s}}}
+    }{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}}
+    }}
   \\ &\text{ where } \Glued{\ell'} = \denotesCC{\Tkg{\lkgproj{1}{\lkg}}} 
@@
 
+To reflect that, we use background colour to help parsing for the proof, and there is no meaning for the choice of colour.
+
 \begin{align*}
@@
   \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}}(\gm)(\gm')(t) &= \CWmodel~(\Gsub{\denotesCC{\wsig}}{\Glued{\gm}})~t \\
-  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~(\Gsub{\denotesCC{a}}{\Glued{\gm}})~(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}}) \\
+  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~
+  \colorbox{blue!10}{(\Gsub{\denotesCC{a}}{\Glued{\gm}})}~
+  \colorbox{green!10}{(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})} \\
   % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
@@
 However, it should be read as the glued interpretation of $ \TyPi{A}{\TyPi {\TyPi{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}} $, which is exactly from the debruinj indxed-version of the typing rule \ruleref{tyeq/casety}.
-
-In the following interpretation, parsing might be hard, so we might use background colour to make things easier to read, and there is no meaning for the choice of colour.
 
@@
   \text{ where } \quad f^r~j~\Glued{a}~b~b^c &= 
-  \texttt{let}~\Glued{\rho}  \in \GTm{(\Glued{\cdot} \Glued{,} \Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}})}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})} 
+  \texttt{let}~\Glued{\rho}  \in \GTm{
+    \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} \colorbox{blue!5}{\Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}}})}
+    }{
+      \colorbox{green!10}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})}
+    } 
   \\
   &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^c)~\texttt{ in } \\
-  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
+  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\Gsub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
   & \quad 
@@
 in a strictly positive position. %So it is a well-founded definition.
-We do not distinguish $(b, b^c)$ and $\Glued{b}$ for simplicity. Again, we omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
+We do not distinguish $(b, b^c)$ and $\Glued{b}$ for simplicity.
+
+The idea of the proof of W-type is, as mentioned, mirroring the facility of the inductive type in the ambient logic into \TT.
+The main difference between $\CWmodel$ in the ambient logic and $\wcode{\cdot}$ in the \TT, is that $\CWmodel$ is only witnessing those reducible \emph{closed terms}.
+Thus when using $\CWmodel$ to model $\wcode{\cdot}$, we need to do closed substitution properly.
+
+Again, we omit validating the equational rules ($\beta$, $\eta$, and substitution) here.
 

### 2023-03-25 canonicitymodel.tex
Done adding color for parsing.
cut clutter

--- 2023-03-25 canonicitymodel.tex
@@
 
-\textbf{Unfortunately}, at this point, our notation is already highly densed because of frequent appearances of $\bullet$. These burden
+%\textbf{Unfortunately}, at this point, our notation is already highly densed because of frequent appearances of $\bullet$. These burden
 
@@
 The key idea of this extension is similar to that of the syntactic translation:
-we interpret linkage types using the canonicity model of sigma types (developed in the prior work).
+we interpret linkage types using the canonicity model of $\Sigma$-types (developed in the prior work).
 We use the inductive facility of the ambient meta\-logic to justify our W-types.
@@
   % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
-  \text{we define }\TyLkg{}, \TyTkg{} & \text{ inductively on its input signature, via } \mathit{L},\Glued{\mathit{P}}  \\ 
+  \text{we interpret }\TyLkg{}, \TyTkg{} & \text{ by induction on the input linkage signature, via } \mathit{L} \text{ and } \Glued{\mathit{P}}  \\ 
   \mathit{L}("nil") &=  \denotesC{\top}  \\
@@
   \\ &\text{ where } \Glued{\ell'} = \denotesCC{\Tkg{\lkgproj{1}{\lkg}}} 
-  \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\ 
+  \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_
+  \\ &{~\begin{minipage}{.58\textwidth}\small
+    In this definition, $\Gpair{\_}{\_}$ notates the glued interpretation of
+    dependent tuple construction, $[\GSubstExt{\_}{\_}]$ notates the glued
+    interpretation of substitution extension, and $\Gsub{\_}{\_}$ notates the
+    glued interpretation of applying a substitution.
+  \end{minipage}}
 \end{align*}
-This is the first point our proof becomes cryptic because of this bullet notation.
 
-The first complexity comes from the three ``bullets after comma'' notation.
+%The first complexity comes from the three ``bullets after comma'' notation.
 The way to interpret this notation is that, this ``$\GSubstExt{\_}{\_}$'' should be considered as a binary function requiring two input of glued interpretation.
 
-However, a careful reader will notice that, because of the overloading of the notation of dependent tuple construction and substitution extension, our glued notation is also overloading.
+%However, a careful reader will notice that, because of the overloading of the notation of dependent tuple construction and substitution extension, our glued notation is also overloaded.
 The first one ``$\Gpair{\Glued{\ell'}}{...}$'' should be the glued interpretation for dependent tuple, but the other two should be the one for substitution extension.
 
-Similarly, the second complexity comes from the bullets after the square bracket, which as expected, is the glued interpretation for substitution.
+%Similarly, the second complexity comes from the bullets after the square bracket, which as expected, is the glued interpretation for substitution.
 Again, ``$\Gsub{\_}{\_}$'' should be considered as a binary function consuming two glued interpretation, corresponding to the substitution operation ``$\sub{\_}{\_}$''.
 
-Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.
+%Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.
 
-In all, the whole interpretation should be considered as:
-\begin{enumerate}
-  \item We first get $\denotesCC{\lkgproj{2}{\lkg}}$, the glued interpretation of the projection of the linkage 
-  \item Then we apply it with two substitution,  $\Gsub{\_}{\GSubstExt{{\GSubstWeak{1}}}{s}}$ and $\Gsub{\_}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}$
-  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{\_}$
-\end{enumerate}
+%The right-hand side can be read as follows:
+%\begin{enumerate}
+%  \item $\denotesCC{\lkgproj{2}{\lkg}}$ is the glued interpretation of a linkage projection
+%  \item Then we apply it with two substitution,  $\Gsub{\_}{\GSubstExt{{\GSubstWeak{1}}}{s}}$ and $\Gsub{\_}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}$
+%  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{\_}$
+%\end{enumerate}
 

### 2023-03-25 canonicitymodel.tex
One more colorbox for help parsing.

--- 2023-03-25 canonicitymodel.tex
@@
   \\ &\text{ where } \Glued{\ell'} = \denotesCC{\Tkg{\lkgproj{1}{\lkg}}} 
-  \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\ 
-\end{align*}
-This is the first point our proof becomes cryptic because of this bullet notation.
-
-The first complexity comes from the three ``bullets after comma'' notation.
+  \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_
+  \\ &{~\begin{minipage}{.58\textwidth}\small
+    In this definition, $\Gpair{\_}{\_}$ notates the glued interpretation of
+    dependent tuple construction, $[\GSubstExt{\_}{\_}]$ notates the glued
+    interpretation of substitution extension, and $\Gsub{\_}{\_}$ notates the
+    glued interpretation of applying a substitution.
+    \\
+    The color boxes indicate the scopes of the two instances of $\Gsub{\_}{\_}$.
+  \end{minipage}} \\[5pt]
+%
+%%The first complexity comes from the three ``bullets after comma'' notation.
 The way to interpret this notation is that, this ``$\GSubstExt{\_}{\_}$'' should be considered as a binary function requiring two input of glued interpretation.
-
-However, a careful reader will notice that, because of the overloading of the notation of dependent tuple construction and substitution extension, our glued notation is also overloading.
+%
+%%However, a careful reader will notice that, because of the overloading of the notation of dependent tuple construction and substitution extension, our glued notation is also overloaded.
 The first one ``$\Gpair{\Glued{\ell'}}{...}$'' should be the glued interpretation for dependent tuple, but the other two should be the one for substitution extension.
-
-Similarly, the second complexity comes from the bullets after the square bracket, which as expected, is the glued interpretation for substitution.
+%
+%%Similarly, the second complexity comes from the bullets after the square bracket, which as expected, is the glued interpretation for substitution.
 Again, ``$\Gsub{\_}{\_}$'' should be considered as a binary function consuming two glued interpretation, corresponding to the substitution operation ``$\sub{\_}{\_}$''.
-
-Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.
-
-In all, the whole interpretation should be considered as:
-\begin{enumerate}
-  \item We first get $\denotesCC{\lkgproj{2}{\lkg}}$, the glued interpretation of the projection of the linkage 
-  \item Then we apply it with two substitution,  $\Gsub{\_}{\GSubstExt{{\GSubstWeak{1}}}{s}}$ and $\Gsub{\_}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}$
-  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{\_}$
-\end{enumerate}
-
-To reflect that, we use background colour to help parsing for the proof, and there is no meaning for the choice of colour.
-
-\begin{align*}
+%
+%%Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.
+%
   \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyLkg{\lsig}}}(\gm)(\gm') &= \denotesC{\ell}(\gm)(\gm')[0] \quad \text{ take the first element in the tuple} \\
@@
   &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^c)~\texttt{ in } \\
-  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\Gsub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
+  & \Gmodel{
+    \Gsub{
+      \colorbox{blue!27}{\Gapp{
+        \colorbox{blue!18}{\Gsub{
+          \colorbox{blue!9}{\Gapp{
+            \colorbox{blue!0}{\GRecproj{j}{\Gsub{\denotesCC{\lkg}}{\Glued{\gm}}}}
+            }}
+          }{
+            \GSubstExt{\GSubstId}{\Glued{a}}
+          }}}}
+      }{
+        \GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}
+      }
+    }(\epsilon)(\star) \\
   & \quad 
@@
     & \to \Glued{b} \in (\GTm{
-      \colorbox{blue!10}{(\Glued{\cdot} \Glued{,} \Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}})} 
+      \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} 
+      \colorbox{blue!5}{\Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}} )} 
       }{

### 2023-03-25 canonicitymodel.tex
parse
Typo.

@@
     \\
-    The color boxes indicate the scopes of the two instances of $\Gsub{\_}{\_}$.
+    The colored boxes indicate the scopes of the two instances of $\Gsub{\_}{\_}$.
+    We use colored boxes here and below to help the reader parse nested glued interpretations.
   \end{minipage}} \\[5pt]
@@
 \end{align*}
-We note that, the interpretation of $\Wsup{i}{\wsig}{a}{b}$ is defined by $\CWsup$ (defined below) will take three elements, $i$, a glued $(\Gsub{\denotesCC{a}}{\Glued{\gm}})$ and a glued $(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})$.
+We note that, the interpretation of $\Wsup{i}{\wsig}{a}{b}$ is defined by $\CWsup$ (defined below).
+It will take three elements: $i$, a glued $(\Gsub{\denotesCC{a}}{\Glued{\gm}})$ and a glued $(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})$.
 As we mentioned earlier, $\Glued{\gm}$ is the equivalent form of $(\gm, \gm')$.
@@
       }{
-        \GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}
+        \GSubstExt{\GSubstId}{
+          \colorbox{green!10}{\Glam{\Glued{\rho}}}
+          }
       }

### 2023-03-25 canonicitymodel.tex
Fail to fix font size on the subscript of the dependent product

@@
     ) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
-    & \quad \quad \to  \prod_{\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} 
-    \colorbox{blue!10}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}} P~(\sub{b}{ \gm }) \\
+    & \quad \quad \to  \prod_{
+      {\Glued{\gm} \in \colorbox{blue!1}{\GSub{\Glued{\cdot}}{
+        \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} 
+    \colorbox{blue!5}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}}}
+    }
+    } P~(\sub{b}{ \gm }) \\
     & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\

### 2023-03-25 canonicitymodel.tex
edits

--- 2023-03-25 canonicitymodel.tex
@@
 \end{align*}
-We note that, the interpretation of $\Wsup{i}{\wsig}{a}{b}$ is defined by $\CWsup$ (defined below).
-It will take three elements: $i$, a glued $(\Gsub{\denotesCC{a}}{\Glued{\gm}})$ and a glued $(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})$.
-As we mentioned earlier, $\Glued{\gm}$ is the equivalent form of $(\gm, \gm')$.
+The interpretation of $\Wsup{i}{\wsig}{a}{b}$ is defined via $\CWsup$, which is
+in turn defined below. This construction takes three arguments as input: $i$, a glued
+$(\Gsub{\denotesCC{a}}{\Glued{\gm}})$ and a glued
+$(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})$.  As mentioned earlier,
+$\Glued{\gm}$ is an equivalent form of $(\gm, \gm')$.
 
-The interpretation for $\CaseSig{"A"}{"B"}{"R"}$ is also bullet-drowned.
+%The interpretation for $\CaseSig{"A"}{"B"}{"R"}$ is also bullet-drowned.
 However, it should be read as the glued interpretation of $ \TyPi{A}{\TyPi {\TyPi{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}} $, which is exactly from the debruinj indxed-version of the typing rule \ruleref{tyeq/casety}.

### 2023-03-25 canonicitymodel.tex
Fix.

@@
     & \to 
-    [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
+    [ j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
     & \quad \quad \to 
@@
     ) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
-    & \quad \quad \to  \prod_{
-      {\Glued{\gm} \in \colorbox{blue!1}{\GSub{\Glued{\cdot}}{
-        \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} 
-    \colorbox{blue!5}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}}}
-    }
-    } P~(\sub{b}{ \gm }) \\
+    &  \quad \quad \to  
+      \left( {\Glued{\gm} \in \GSub{\Glued{\cdot}}{
+        \colorbox{blue!11}{(\Glued{\cdot} \Glued{,} 
+    \colorbox{blue!1}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}}
+    }\to P~(\sub{b}{ \gm }) \right) \\
     & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\

### 2023-03-25 canonicitymodel.tex
\Big

--- 2023-03-25 canonicitymodel.tex
@@
     & \to 
-    [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
+    \Big( j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\
     & \quad \quad \to 
@@
     } P~(\sub{b}{ \gm }) \\
-    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
+    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) \Big) \\
     & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\

### 2023-03-25 canonicitymodel.tex
\Bigg

--- 2023-03-25 canonicitymodel.tex
@@
     & \to 
-    [ j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
+    \Bigg( j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
     & \quad \quad \to 
@@
     }\to P~(\sub{b}{ \gm }) \right) \\
-    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
+    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) \Bigg) \\
     & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\

### 2023-03-25 metatheory-condensed.tex
Move based path induction from appendix to Section 6.
tweak
shorter citation
revive metadata

@@
 We acknowledge that the presentation is dense for an audience
-unacquainted with MLTT, so we summarize the salient points first.
+without intimate knowledge of MLTT, so we summarize the salient points first.
 
@@
 So are their introduction and elimination forms.
+We use based path induction%
+\defcitealias{program2013homotopy}{UFP~2013}%
+as the elimination principle for identity types~\citepalias{program2013homotopy}.
 Capture-avoiding substitution is notated $\bullet[\bullet/x]$.

### 2023-03-31 lang-design.tex
Add back partial recursor

@@
 
-%\begin{centered}
-%\begin{minipage}{.97\textwidth}
-%\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
-%\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
-%tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
-%                [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
-%                [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] ->
-%                \forall t, option (P t)
-%\end{lstlisting}
-%\end{minipage}
-%\end{centered}
+\noindentparagraph{Property about Partial Recursors.}
+
+Partial recursor is simply adding partiality into the signature of the eliminator, by wrapping each motive with "option".
+This partial recursor seems non-canonical compared to the (dependent) eliminator for Inductive type as the latter can be considered as universal property.
+This extensionality makes dependent eliminator one of \textit{the most powerful} reasoning tools around inductive type.
+
+Even though we cannot prove partial recursors are \textit{the
+most powerful} reasoning tool for extensible inductive types,
+the fact that propositional partial
+recursors can derive injectivity alreadt demonstrate itself as a good \textit{extensional characterization} of
+(non-indexed) extensible inductive type (under certain
+constraints)---because we can ``embed'' vanilla inductive types into
+types supporting a partial recursor: a bit more formally, and
+restricting our focus to non-indexed inductive types:
+
+\begin{theorem}\label{thm:prec-complete} Given a list of $n$ pairs of types $\{ x : "A"_i \vdash "B"_i(x) \}_{i}$ s.t.
+
+  \begin{itemize}
+    \item \textlabel{(Detectable Partiality)}{prop:detectable-partiality} for each pair $x : A_i \vdash B_i(x)$,
+    \begin{minted}{Coq}
+      Axiom detectable:
+      forall {T} {a} (f: Bᵢ a -> option T),
+        {forall x, f x <> None} + {exists x, f x = None}.
+    \end{minted} 
+    \item Define $"C"$ as the inductive type using this list, i.e.,
+    \begin{minted}{Coq}
+Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
+    \end{minted}
+    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
+    \begin{itemize}
+      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
+      \item a partial recursor
+      \begin{minted}[escapeinside=@@]{Coq}
+prec : forall (R : Set), ..,
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,@\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint.
+Let me fix other parts first}\YZreply{Is this constraint Detectable Partiality?}@
+      D -> option R
+      \end{minted}
+      \item and $n$ (propositional) computational axioms: for all $i$, 
+      \begin{minted}{Coq}
+prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
+    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
+      \end{minted}
+    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
+    \end{itemize}
+  \end{itemize}
+  Then there exists an \textbf{embedding} from "C" to "D".
+  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
+  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of\/
+  "C" and "prec" of\/ "D" (both acting like identity) such that
+  \begin{minted}[escapeinside=@@]{Coq}
+    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
+If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
+I mentioned left inverse 'linv' is the generalization}@
+  \end{minted}
+\end{theorem}
+The reason we have this weird looking \ref{prop:detectable-partiality}
+is due to the construction of "linv"---at one point, using "prec" to
+construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
+C) → option C}, generally not provable, unless we have
+\mintinline{Coq}{option (Bᵢ a → C)} instead.\YZ{Don't you really mean
+"unless we have the Detectable Partiality assumption instead"?}
+This \ref{prop:detectable-partiality} is exactly transforming the former to
+the latter:
+recall that, according to the semantic of inductive type,
+\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
+"linv" on the substructure of "D". This function points out each of the
+translation result from "D" to "C" and some of them are "None" because
+of failing. For us, we only care about if \textbf{any} of the
+substructure of "D" fails to translate---if any substructure fails,
+then the whole translation fails and thus we should have "None" as
+return.
+
+\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that includes all of the examples shown in this paper.
+
+\ref{prop:detectable-partiality} is not generally supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
+However, in those cases, we still have partial recursor provable and useful---we just cannot see the partial recursor as a strong toolkit anymore.
+
+We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
+actually be considered as a constructor, because they can be reflected
+to real constructors of "C" using the left inverse "linv". Thus, \cref{thm:prec-complete}
+implies that (1) every future extension of the inductive type can
+support this partial recursion (trivially); (2) every type supporting
+this partial recursor with its computational axioms at least supports
+these constructors because of the embedding. In other words, \textit{a
+type (at least) supports these constructors if and only if this type
+supports the corresponding partial recursor (with the computational
+axioms)}.  Thus, we can argue that the partial recursor gives a sound and
+complete extensional characterization of all the extension of a given (non-indexed)
+inductive type.\YZ{What about indexed inductive types?}\EDJreply{Done.
+I also have a formalization of this proof now, for indexed type and dependent eliminator.
+Though the formalization is about single constructor.}
+
+We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
+\begin{minted}{Coq}
+  pdelim : ∀ (P : D -> Set), .., 
+  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
+    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
+\end{minted}
+To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable. 
+% Please refer to the supplementary material and appendix for a formalization of the left inverse in the context of indexed inductive type.
 

### 2023-03-31 body.tex
Start working on appendices.
Save Canonicity Model.
ack Zhixuan
tweaks
Set out the frame

@@
-\begin{abstract}
-\input{abstract}
-\end{abstract}
+% \begin{abstract}
+% \input{abstract}
+% \end{abstract}
 
-\maketitle
+% \maketitle
+
+\section*{Family Polymorphism by Coq Plugin}
 
@@
 
-\section{\TT: A Core Dependent Type Theory}
+\section{A Preview of \TT: A Core Dependent Type Theory}
 \label{sec:metatheory2}
@@
 We thank the reviewers for their many useful comments.  We thank
-Anastasiya Kravchuk-Kirilyuk, Yao Li, and Andrew Myers % others?
-for valuable discussions.
+Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
+for valuable discussions and feedback.
 This work was supported in part by an Amazon research award on automated
@@
 
-\ifreport
+% \ifreport
 
-\appendix
+% \appendix
 
@@
 
-\section{More Typing Rules for \TT}
-\label{sec:more-rules}
-\input{all-typing-rule.tex}
+\section*{Technicality of \TT}
+
+\section{Syntax and Semantic Models of \TT}
+\label{sec:complete-fmltt}
+% \input{all-typing-rule.tex} -- we make all-typing rules into metatheory2.tex
+\input{metatheory2.tex}
 
@@
 
-\newpage
+% \newpage
 
-\section{Definitive Version of the Formalization and Proofs}
-\label{sec:agda-files}
-\input{agda-files.tex}
+% \section{Definitive Version of the Formalization and Proofs}
+% \label{sec:agda-files}
+% \input{agda-files.tex}
 
@@
 
-\fi
+% \fi

### 2023-03-31 metatheory2.tex
Set out the frame.

@@
-This appendix supplements \cref{sec:metatheory2} with the details of the syntax
+This section supplements \cref{sec:metatheory2} with the details of the syntax
 and the proofs of the main results.
@@
 %This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
-Moreover, universe levels are explicit in this appendix.
+Moreover, universe levels are explicit in this appendix. Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
 
-Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
-In particular,
-%following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019},
-the syntax is still considered as being quotiented by the judgmental equalities.
-Quotienting facilitates coercion along equalities.
-Furthermore, the semantic model we develop in this appendix respects these
-judgmental equalities by construction.
 %This means 
 %(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.
-We use based path induction
+We still use based path induction
 as the elimination principle for identity types~\cite{program2013homotopy}.
@@
 \input{all-mltt-typing}
+
+\noindentparagraph{MLTT as a programming language.}  Conventionally, the operational semantic sketches \emph{the interpreter} of a given formalization of a programming language.
+However, compared to this traditional setup, a careful reader will notice the absence of operational semantic in our MLTT.
+It is no doubt that MLTT will be a logical system we can write proofs, but it is unclear if MLTT can really be considered as a programming language.
+
+Indeed, for MLTT, we only have syntactic term and their judgemental equality (conversion rules) between them.
+However, we still can have \emph{an interpreter} about our MLTT in our ambient meta-logic.
+For a programmer, this is just like writting a Haskell meta-circular interpreter~\cite{abelson1996structure} that interpret an abstract syntax tree into Haskell's data structure.
+For a logician, this is just interpreting the syntactic $\goodTerm{\cdot}{t}{\cB}$ to an element $t \in \{\true, \false \}$.
+For a formal semanticist, this can be considered as a naive form of denotational semantic.
+
+This interpreter is in the form of our canonicity theorem~\cref{thm:canonicity} in the light of Curry-Howard Correspondence, as any closed program of type $\cB$ is evaluated to either $\true$ or $\false$ under this theorem.
+What's more, our canonicity theorem not only return the evaluation result, but also \emph{reflects} this computational traces into MLTT as a proof term of the derivation $t \equiv \true$ or $t \equiv \false$.
+
+\noindentparagraph{Quotient upon the syntax.} As we mentioned in \cref{sec:metatheory2},
+%following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019},
+the syntax is considered as being quotiented by the judgmental equalities, and 
+quotienting facilitates type coercion along equalities.
+
+By looking at our syntax as some sorts of abstract syntax tree (data), after quotient,
+each syntax piece is actually an equivalent class of unquotiented syntax, where 
+the equivalence is exactly the judgemental equalities.  A proper introduction 
+of this quotient concept 
+can be found in \citet{altkap2016,kaposi2017type}, where a type-theoretic quotient formulation 
+of MLTT is given.
+
+Once it is clear that the syntax of our MLTT as a logic system is a quotiented 
+data, then it is natural to consider model of this logic system 
+as a function mapping from those quotiented data (syntax). The (semantic) model introduced below 
+will follow this idea. More specifically, the semantic model as a function will have to respects 
+judgemental equalities to be a well-defined function from quotiented data. In other words,
+for two judgementally equal syntax pieces, the model needs to map them into the same data.
+
+% Furthermore, the semantic model we develop in this appendix respects these
+% judgmental equalities by construction.
 

### 2023-03-31 ttintro.tex
Set out the frame.

--- 2023-03-24 metatheory2.tex
@@
-This appendix supplements \cref{sec:metatheory2} with the details of the syntax
-and the proofs of the main results.
+\noindentparagraph{\TT as a programming language.}  Conventionally, the operational semantic sketches \emph{the interpreter} of a given formalization of a programming language.
+However, compared to this traditional setup, a careful reader will notice the absence of operational semantic in our \TT and doubt if \TT can really be considered as a programming language.
 
-Unlike in \cref{sec:metatheory2}, which uses named binders, in this appendix we
-use \citeauthor{debruijn1964} indices and \emph{explicit
-substitutions}~\cite{abadi1989subst,substcalculus}:
-substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}, which
-applies~$\gm$ to the type~$T$) are part of the syntax rather than
-meta-operations.
-We work in an intrinsically typed setting: terms are well typed by construction.
-%
-Consequently, we omit without ambiguity some obvious premises needed for well-formedness.
-% 
-This style of syntax formulation follows a recent
-trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
-known as the ``algebraic presentation'' of MLTT.
-%We introduce judgmental equality ($\equiv$)
-%for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
-%contexts, explicits substitutions, and etc. 
-%This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
-Moreover, universe levels are explicit in this appendix.
+Indeed, for \TT, we only have syntactic term and their judgemental equality (conversion rules) between them.
+However, we still can have \emph{an interpreter} about our \TT in our ambient meta-logic.
+For a programmer, this is just like writting a Haskell meta-circular interpreter~\cite{abelson1996structure} that interpret an abstract syntax tree into Haskell's data structure.
+For a logician, this is just interpreting the syntactic $\goodTerm{\cdot}{t}{\cB}$ to an element $t \in \{\true, \false \}$.
+For a formal semanticist, this can be considered as a naive form of denotational semantic.
 
-Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
-In particular,
-%following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019},
-the syntax is still considered as being quotiented by the judgmental equalities.
-Quotienting facilitates coercion along equalities.
-Furthermore, the semantic model we develop in this appendix respects these
-judgmental equalities by construction.
-%This means 
-%(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.
-We use based path induction
-as the elimination principle for identity types~\cite{program2013homotopy}.
-
-%After the syntax, we detail the models and proofs for syntactic translation (that translate away linkages),  consistency and canonicity.
-
-\subsection{MLTT with Explicit Substitutions and Universe Levels}
-\label{sec:mltt-full}
-
-We review the base MLTT fragment of \TT first.
-\input{all-mltt-typing}
-
-\noindentparagraph{De Bruijn Indices and Explicit Substitution.}
-
-\Citeauthor{debruijn1964} indices and explicit substitutions make details about
-binders and substitutions explicit.
-%
-Using explicit substitutions obviates the need for special treatment of
-substitutions in the proofs, as substitutions are part of the syntax.
-%The change of the syntax formulation is important to the simplicity of our
-%proof, as we want to handle the proof about substitutions and terms
-%uniformly---thus making substitution as part of the syntax (explicit
-%substitution) will be helpful. 
-%
-The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
-For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
-%
-Substitutions are typed with the form 
-$\goodSub{\Gm}{\gm}{\Dl}$.
-The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
-yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
-%
-The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
-\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
-\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
-then applies~$\gm$.
-For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
-that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
-$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
-where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
-
-To simplify, $\SubstWeak{n}$ is a short hand for $\pi_1^n~"id"$ and $\var{n}$ is a short hand for $\pi_2~\pi_1^{n}$. Thus during meta-theoretic reasoning, we will only deal with $\pi_1$ and $\pi_2$.
-
-Consequently, function application changes to an equivalent formulation, and becomes a ``direct inverse'' of typing rule for function abstraction.
-For example, the named notation $\napp{f}{t}$ can be equivalently represented by $\sub{\app{f}}{\SubstExt{\SubstId}{t}}$.
-
-Finally, we have \ruleref{sub/dbj/shift} defined using \ruleref{sub/wk} and \ruleref{tm/var}.
-This rule applying substitution $\gm$ to the earlier portion of the context.
-We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.
-
-\noindentparagraph{Universe levels.}
-
-Universe levels address the size issue---it is unsound to have a set of all sets
-(or a universe of all types)~\cite{hurkens1995simplification}.
-The level of a universe specifies ``how large'' that universe is.
-%that is basically what universe level is measuring.
-
-%The treatment of universe levels follows \citet{altkap2016, kaposi2019gluing}.
-The judgment form $\goodType{\Gm}{T}{i}$ indicates that (the code of) the type $T$
-inhabits universe~$\cU_i$.
-Similarly, $\goodCtx{\Gm}{i}$ indicates that (the codes of) the types in $\Gm$
-inhabit universe~$\cU_i$.
-The notation $i \lcup j$ denotes $\max(i,j)$.
-
-%Since we work in an intrinsic-typed setting, some required premises can be omitted without ambiguity.
-For omitted context judgement $\goodCtx{\Gm}{k}$, the universe level is always $k$, hinted at the figure of judgement specification.
-Similar for type judgement $\goodType{\Gm}{T}{j}$.
-\EDJ{Why removing the mentioning of ``omitted/inferrable k here?'' I know it is confusing but I think it is important to make things rigorous.}\YZreply{I agree it's important, and I'm happy to include it if you make it less confusing.}\EDJreply{Do you think ``omitting premises'' should be mentioned here or earlier?}\YZreply{It's not just about omitting premises, right?
-You also omit universal levels in some typing rules.}
-
-% Copy and Paste Intuition of MLTT from other places to here
-\subsection{\TT}
-\label{sec:fmltt-full}
-
-\TT extends the MLTT in \cref{sec:mltt-full} with W-type signatures and W-types,
-linkage signatures and linkages, and linkage transformers.
-
-\input{all-fmltt-typing}
-
-% Copy and Paste Intuition of FMLTT from other places to here
-
-% Deal with intuition
-\noindentparagraph{Why $\TyTkg{\lsig}$?}
-
-%As shown by the equality rule \ruleref{tyeq/pk/add},
-%for a given linkage signature $\lsig \coloneq \LSigAdd{\cdot}{\cdot}{\cdot}$,
-%the type $\TyTkg{\lsig}$ is structurally similar to the linkage type $\TyLkg{\lsig}$
-%except that $\TyTkg{\lsig}$ is a dependent tuple type---%
-%it packages the fields of $\lsig$ into a dependent tuple type.
-%
-%We have computed several examples in
-%\cref{fig:stlc-linkage-typing}, where the reader should also notice the high
-%similarity between $\TyTkg{\lsig}$ and $\TyLkg{\lsig}$.
-
-$\TyTkg{\lsig}$ is needed in rules $\ruleref{lsig/add}$ and $\ruleref{l/add}$:
-these rules have a premise  $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
-responsible for possibly hiding W-type signatures (in $\TyTkg{\lsig}$) behind $\cU$ (in $A$);
-see $\seal_6$ in \cref{fig:stlc-linkage-typing} for an example.
-Here, both $\TyTkg{\lsig}$ and $A$ are dependent tuple types rather than linkage types.
-
-%Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
-$\TyTkg{\cdot}$ is indispensible because \emph{we cannot do abstraction with $\TyLkg{\cdot}$ alone}, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{lsig/add}$ and $\ruleref{l/add}$ is \emph{usually a dependent tuple type}.
-
-An alternative to $\TyTkg{\lsig}$ would be to use linkage types $\TyLkg{\cdot}$ in this premise.
-However, this alternative is infeasible.
-We illustrate using a simple example.
-Consider the linkage signature $\lsig$ below (it is written using the record
-syntax for illustration but should not be confused with records, as a later field
-is universally quantified over a $\mathit{self}$ variable).
-The job is to hide the W-types $\TyS{\wcode{\wsig_1}}$ and
-$\TyS{\wcode{\wsig_2}}$ in $\lsig$ behind $\cU$ in $\lsig^A$:
-%
-\begin{align*}
-  \lsig &\coloneq \{ T_A : \TyS{\wcode{\wsig_1}}; &&T_B : \TyS{\wcode{\wsig_2}}; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
-  \lsig^A &\coloneq \{T_A : \cU; &&T_B : \cU; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
-  & && \goodTerm{x : \TyLkg{\lsig}}{\seal}{\TyLkg{\lsig^A}}
-\end{align*}
-
-%We illustrate with a small example using named convention, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function $f$ maps from the first to the second type.
-Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{\lsig}}$.
-Say we want to add a new field into this linkage, and in $\ruleref{l/add}$, we chose $A$ to abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
-The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction (choose $s$).
-
-The problem lies in that the type $\TyLkg{\lsig^A}$ above cannot be inhabited.
-Suppose there is some $\lkg^A$ such that $\cdot \vdash \lkg^A : \TyLkg{\lsig^A}$.
-Then it follows from the self-parameterization that projecting out the last field~$f$
-from $\lkg^A$ (notated as $\lkg^A.f$ for illustration) has the typing
-$\goodTerm{T_A : \cU, T_B: \cU}{\lkg^A.f}{\El{T_A} \to \El{T_B}}$.
-This typing means that there is a function of type
-$\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A}
-\to \El{T_B}}}$, i.e., a function from an arbitrary type to another arbitrary
-type. We can easily derive inconsistency from this function:
-applying it to $\top$ and $\bot$ yields a function of type $\top\to\bot$.
-
-%\begin{align*}
-%  & \text{doing abstraction (choosing appropriate $s$ in \ruleref{l/add})} \\ 
-%  \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
-%  \text{Approach } \beta \quad&\quad  \TyLkg{\lsig} \longrightarrow \TyTkg{\lsig} \longrightarrow \TyTkg{\lsig^A}\\
-%\end{align*}
-%
-%However, the seemingly easier approach $\alpha$ is impossible because linkage
-%$\TyLkg{\lsig^A}$ is uninhabited---%
-%say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
-To make the inconsistency clearer, by double abstraction we will have an impossible function $g \coloneq \goodTerm{\cdot}{\nlam{T_A}{\nlam{T_B}{\ell^A.f}}}{\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A} \to \El{T_B}}}} $, i.e.
-a function maps from arbitrary type to another arbitrary type .
-We can easily derive inconsistency by applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
-%
-%Thus we can only choose the approach $\beta$ instead, which justifies why we have $\TyTkg{\cdot}$ in our typing rule, as $\TyTkg{\lsig}$ is usually the first step to do abstraction.
-
-Thus, it is critical that the hiding of W-type signatures operate on dependent
-tuple types rather than linkage types.
-To this end, we introduce the syntax $\TyTkg{\lsig}$ and its computation rules
-(e.g., \ruleref{tyeq/pk/add}) to make it convenient to package linkage signatures
-into dependent tuple types.
-
-%To summarize, the reason we need dependent tuple type is because of the overridability nature of our linkages, together with the frequent usage of universe type in the context of dependent type programming.
-Once we want to abstract a type in a signature, we may have to use dependent tuple type as the abstraction instead of linkages.
-Since abstraction is usually dependent tuple type, we introduce $\TyTkg{\cdot}$ and its computation rules to make the process of doing abstraction in \TT easier.
-
-\noindentparagraph{Linkage transformers as syntactic sugar.}
-
-%Observe the syntax, linkage transformer and its judgement is defined above 
-%other parts. 
-%
-%So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$.
-Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality
-%%\EDJ{Add substitution equalities for linkage transformers} 
-%given in the syntax.
-In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.
-
-As mentioned in \cref{sec:metatheory2}, the five forms of linkage transformers can be
-thought of as a library of functions that are used to construct linkages from other
-linkages.
-In particular, their typing rules are defined in terms of the rest of the typing
-rules in \TT.
-Thus, linkage transformers $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ can be defined
-as syntactic sugar via an inductive type (in the meta\-logic) with four constructors
-%\YZ{I reinterpreted what you wrote, but this sentence doesn't sound right to me.
-If linkage transformers are meta-level inductive types, aren't they just part of the syntax?}\EDJreply{I think the reason that they are syntactic sugar is because they can be interpreted using FMLTT syntax(without them), it is fine either the syntactic sugar is defined via inductive type (and then interpreted), or directly interpreted.
-Maybe I should remove this misleading sentence?
-\\ What's more, the specification/definition of syntactic sugar itself, is inductive type.
-Because a linkage transformer looks inductive.
-Maybe I should directly separate the typing rules for linkage transformer as another figure to emphasize it as an inductive construction on the rest FMLTT syntax?}
-$\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, and $\InhNest{}{}$.
-Moreover, $\inh{h}{\ell}$ and $\sub{h}{\gm}$ can be defined
-as recursive functions (in the meta\-logic) by induction on this inductive type.
-We will treat linkage transformers as syntactic sugar in the meta\-theoretic development
-in the rest of \cref{sec:complete-fmltt}.
-
-\subsection{A Translation that Compiles Linkages Away}
-
-\input{syn-translate.tex}
-
-% \subsection{Standard Model for Consistency}
-
-% \input{standardmodel.tex}
-
-\subsection{A Proof Relevant Logical-Relations Model for Canonicity}
-\label{sec:canonicity-model}
-\input{canonicitymodel.tex}
+This interpreter is in the form of our canonicity theorem~\cref{thm:canonicity} in the light of Curry-Howard Correspondence, as any closed program of type $\cB$ is evaluated to either $\true$ or $\false$ under this theorem.
+What's more, our canonicity theorem not only return the evaluation result, but also \emph{reflects} this computational traces into \TT as a proof term of the derivation $t \equiv \true$ or $t \equiv \false$.

### 2023-03-31 metatheory2.tex
Typo.

@@
 
-Indeed, for MLTT, we only have syntactic term and their judgemental equality (conversion rules) between them.
+Indeed, for MLTT, we only have syntactic term and their judgemental equality (conversion rules) among them.
 However, we still can have \emph{an interpreter} about our MLTT in our ambient meta-logic.
@@
 This interpreter is in the form of our canonicity theorem~\cref{thm:canonicity} in the light of Curry-Howard Correspondence, as any closed program of type $\cB$ is evaluated to either $\true$ or $\false$ under this theorem.
-What's more, our canonicity theorem not only return the evaluation result, but also \emph{reflects} this computational traces into MLTT as a proof term of the derivation $t \equiv \true$ or $t \equiv \false$.
+What's more, our canonicity theorem also \emph{reflects} this computational traces into MLTT as a proof term of the derivation $t \equiv \true$ or $t \equiv \false$.
+From this perspective, we can see the similarity between the unidirectional small-step relation and the bidirectional conversion rules.
 

### 2023-03-31 metatheory2.tex
Adding TOC.

@@
 However, compared to this traditional setup, a careful reader will notice the absence of operational semantic in our MLTT.
-It is no doubt that MLTT will be a logical system we can write proofs, but it is unclear if MLTT can really be considered as a programming language.
+It is no doubt that MLTT is a logical system where we can write proofs, but it is unclear if MLTT can really be considered as a programming language.
 
@@
 However, we still can have \emph{an interpreter} about our MLTT in our ambient meta-logic.
-For a programmer, this is just like writting a Haskell meta-circular interpreter~\cite{abelson1996structure} that interpret an abstract syntax tree into Haskell's data structure.
-For a logician, this is just interpreting the syntactic $\goodTerm{\cdot}{t}{\cB}$ to an element $t \in \{\true, \false \}$.
+For a programmer, this is like writting a Haskell meta-circular interpreter~\cite{abelson1996structure} that interprets an abstract syntax tree into Haskell's data structure.
+For a logician, this is interpreting the syntactic $\goodTerm{\cdot}{t}{\cB}$ to an element $\llbracket {t} \rrbracket \in \{\true, \false \}$.
 For a formal semanticist, this can be considered as a naive form of denotational semantic.

### 2023-03-31 lang-design.tex
Added TOC.

@@
 "unless we have the Detectable Partiality assumption instead"?}
+\EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.}
 This \ref{prop:detectable-partiality} is exactly transforming the former to

### 2023-03-31 frontmatters.tex
Adding front matters.

% \begin{abstract}
% \input{abstract}
% \end{abstract}

% \maketitle
% adding chapter here

\tableofcontents

### 2023-03-31 body.tex
Set out the frame.
Adding TOC.
Added TOC.
Adding front matters.
Add Front Matter

@@
-% \begin{abstract}
-% \input{abstract}
-% \end{abstract}
+\input{frontmatters.tex}
 
-% \maketitle
+\newpage
 
-\section*{Family Polymorphism by Coq Plugin}
+\chapter{Family Polymorphism as Coq Plugin}
 
@@
 
-\ifacknowledgments
-\section*{Acknowledgments}
-We thank the reviewers for their many useful comments.  We thank
-Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
-for valuable discussions and feedback.
-This work was supported in part by an Amazon research award on automated
-reasoning.
-The views and opinions expressed are those of the authors and do not necessarily
-reflect the position of any funding agency.
-\fi
-
-\setlength{\bibsep}{.8ex}
-\bibliographystyle{tex-macros/ACM-Reference-Format.bst}
-\bibliography{refs.bib}
-
 % \ifreport
@@
 
-\section*{Technicality of \TT}
+\chapter{Technicality of \TT}
 
@@
 
-\newpage
+% \newpage
 
@@
 % \fi
+
+% \ifacknowledgments
+% \section*{Acknowledgments}
+% We thank the reviewers for their many useful comments.  We thank
+% Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
+% for valuable discussions and feedback.
+% This work was supported in part by an Amazon research award on automated
+% reasoning.
+% The views and opinions expressed are those of the authors and do not necessarily
+% reflect the position of any funding agency.
+% \fi
+
+\setlength{\bibsep}{.8ex}
+\bibliographystyle{tex-macros/ACM-Reference-Format.bst}
+\bibliography{refs.bib}

### 2023-03-31 frontmatters.tex
Adding frontmatters.
Add Front Matter

@@
-% \begin{abstract}
-% \input{abstract}
-% \end{abstract}
+% Title Page
 
-% \maketitle
-% adding chapter here
+\vspace*{\fill}
+\begin{center}
+  \textsc{\Huge \thetitle}
+\end{center}
 
+\vspace*{\fill}
+
+\newpage
+
+% Author's Declaration
+\vspace*{\fill}
+
+\chapterN{Author's Declaration}
+
+\vspace*{\fill}
+
+\newpage
+
+% Statement of contributions
+
+\vspace*{\fill}
+
+\chapterN{Statement of contributions}
+
+\vspace*{\fill}
+
+\newpage
+
+% Abstract
+\vspace*{\fill}
+\chapterN{Abstract}
+
+\input{abstract}
+\vspace*{\fill}
+
+\newpage
+
+% Acknowledgements
+
+\vspace*{\fill}
+\chapterN{Acknowledgments}
+
+We thank Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
+for valuable discussions and feedback.
+This work was supported in part by an Amazon research award on automated
+reasoning.
+The views and opinions expressed are those of the authors and do not necessarily
+reflect the position of any funding agency.
+\vspace*{\fill}
+
+\newpage
 \tableofcontents

### 2023-04-01 frontmatters.tex
Fix Front matter
Fix front.
Fix.

@@
 % Title Page
-
+\pagenumbering{roman}
 \vspace*{\fill}
@@
 
+\setcounter{page}{1}
+
 % Author's Declaration
@@
 \chapterN{Author's Declaration}
+
+This thesis consists of material all of which I authored or co-authored: see Statement of Contributions included in the thesis.
+This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.
 
@@
 \chapterN{Statement of contributions}
+
+Ende Jin and Prof. Yizhou Zhang co-authored Chapter 1.
+
+Ende Jin is the sole author of Chapter 2, which were written under the  under the supervision of Prof. Yizhou Zhang.
+
+Both Chapter 1 and Chapter 2 are already published result.
+
+\noindentparagraph{Research presented in Chapter 1 and 2:} Ende Jin is the primary investigator of both theoretical development and practical implementation of the prototype.
+This work was supported in part by an Amazon research award on automated
+reasoning.
 

### 2023-04-01 lang-design.tex
address an AE reviewer's comment

@@
 though we believe the design could be adapted to other proof assistants such as Lean.
-We call our design and implementation \Lang (\underline{f}amily
-\underline{\smash{po}}lymorphism for a \underline{\smash{p}}roof assistant).
+We call our design and implementation \ac{FPOP}.
 In this section, we focus on the language design of \Lang.
@@
 power the \lsti{finjection} and \lsti{fdiscriminate} tactics.
-Partial recursors are automatically generated for all inductive types
+Partial recursors can be automatically generated for inductive types
 defined with \lsti{FInductive}.

### 2023-04-01 introduction.tex
hyphen

@@
 in the new setting of proof engineering:
-the \emph{expression problem} (EP) \cite{wadler-ep}.
+the \ac{EP} \cite{wadler-ep}.
 
@@
 The EP is well studied in the setting of ordinary functional
-programming and object-oriented (OO) programming.
+programming and \ac{OO} programming.
 Modern languages, such as Scala \cite{scala-oopsla05}, have a good
@@
 As an example, the diagram to the right depicts an extensible
-mechanization of the simply typed λ-calculus (STLC), using family
+mechanization of the \ac{STLC}, using family
 polymorphism.
@@
 \item We capture the essence of the new language mechanism formally by extending
-Martin–Löf type theory with facilities to express family polymorphism (\cref{sec:metatheory2}).
+Martin-Löf type theory with facilities to express family polymorphism (\cref{sec:metatheory2}).
 We prove foundational meta\-theoretical results including consistency

### 2023-04-01 metatheory-condensed.tex
Adding list of figures.
Adding List of Abbrev.
List of abbreviation added

@@
-We contribute \TT, a core type theory that extends
+We contribute \ac{FMLTT}, a core type theory that extends
 %\underline{f}a\underline{m}i\underline{l}y polymorphism
@@
 So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}\EDJreply{This following sentence has been replied to Zhixuan and he doesn't has further response: 1.
-We have to have enhanced W-type at FMLTT, that's for sure.
+We have to have enhanced W-type at \TT, that's for sure.
 Otherwise we (a).
@@
 
-\definecolor{A-color}{HTML}{E7EFC5}
-\definecolor{T-color}{HTML}{FFF1D0}
-\definecolor{t-color}{HTML}{FAE0E4}
-\definecolor{P-color}{HTML}{BFEFFF}
-
 \input{stlc-linkage-typing-named.tex}
@@
 An effort to mechanize the formalization is underway. It follows \citet{altkap2016}
-in using quotient inductive-inductive types (QIITs) as 
+in using \ac{QIIT} as 
 the logical framework to state intrinsically typed syntax (and also

### 2023-04-01 related-work.tex
fix
List of abbreviation added

@@
 from feature modules. Feature composition is manual, which seems to have
-motivated later approaches based on \emph{data types à la carte} (DTC)~\cite{dtc}.
+motivated later approaches based on \ac{DTC}~\cite{dtc}.
 %\rB{ define abbreviation DTC at first use?}
@@
 fails the strict positivity check imposed by proof assistants including Coq and Agda.
-\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
+\citet{delaware2013} introduce \ac{MTC}:
 it overcomes the problem by using Church encodings for
@@
 
-Pumpkin Pi~\cite{ringer2021pumpkin} is a Coq plug-in that helps
+Pumpkin Pi~\cite{ringer2021pumpkin} is a Coq plugin that helps
 repair proofs broken by changes in type definitions.

### 2023-04-01 casestudies.tex
say less
seems more appropriate to say RState is a concretization relation?

@@
 is stated with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
-The theorem says that the abstraction relation \lsti{RState} over a
+The theorem says that the concretization relation \lsti{RState} over a
 concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
@@
 This family leaves fields representing the abstract domain, the
-abstraction relation, monotonicity of transfer functions, etc.\ 
+concretization relation, monotonicity of transfer functions, etc.\ 
 largely unspecified or unproven---a derived family can further bind
@@
 %
-An alternative to \Lang would be to define \lsti{ImpGAI} as a functor parameterized
-by the abstract domain and such. But it would prevent \lsti{ImpCP} from reusing
-the code and proofs in \lsti{ImpGAI}, because \lsti{ImpCP} involves extending
-inductive types with new constructors.
+%An alternative to \Lang would be to define \lsti{ImpGAI} as a functor parameterized
+%by the abstract domain and such. But it would prevent \lsti{ImpCP} from reusing
+%the code and proofs in \lsti{ImpGAI}, because \lsti{ImpCP} involves extending
+%inductive types with new constructors.
 

### 2023-04-02 coqimpl.tex
add a bit from rebuttal
address reviewer comments

@@
 to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and upon \lsti{End}~\lsti{STLCFix}.
+In particular, non-exhaustive pattern matching is prevented because
+an ordinary recursor like \lsti{tm_rect} is not available.
 
@@
 to verify that there are no lingering \lsti{Axiom}s introduced by the translation.
-Ramifications of possible bugs in the \Lang implementation are limited to
+The ramifications of possible bugs in the \Lang implementation are limited to
 the usability of the plugin.
@@
 We believe that they do not pose conceptual challenges and can be
-addressed with more engineering effort of the same kind that went into \Lang.
+addressed with more engineering effort on the same level as the current \Lang implementation.
 

### 2023-04-02 metatheory-condensed.tex
address reviewer comments

@@
 
+\defcitealias{program2013homotopy}{UFP~2013}
+
 \cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT %(predicative) MLTT
@@
 So are their introduction and elimination forms.
-We use based path induction%
-\defcitealias{program2013homotopy}{UFP~2013}%
+We use based path induction
 as the elimination principle for identity types~\citepalias{program2013homotopy}.
@@
 % and compatibility with our algebraic formulation of MLTT~\cite{coquand2018canonicity}. % I am not sure if I want this sentence or not
-Term $\codety{T}$ encodes type~$T$ (\ruleref{tm/c}) and thus inhabits universe~$\cU$.
-Type $\El{t}$ decodes term~$t$ (\ruleref{ty/el}).
+The term $\codety{T}$ encodes type~$T$ (\ruleref{tm/c}) and thus inhabits universe~$\cU$.
+The type $\El{t}$ decodes term~$t$ (\ruleref{ty/el}).
 %\footnote{By contrast,  Russel-style universes are inhabited by types directly.
@@
 case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
-Each case handler has a type of form $\CaseSig{A}{B}{T}$, where
+Each case handler has a type of the form $\CaseSig{A}{B}{T}$, where
 $T$ is the motive of the recursion (\ruleref{tyeq/casety});
@@
 
-Rules \ruleref{tm/wsup} and \ruleref{tm/wrec} require access to~$\wsig$:
+The rules \ruleref{tm/wsup} and \ruleref{tm/wrec} require access to~$\wsig$:
 the W-type is exhaustively generated by its constructors, and
@@
 etc.
-\cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
+\cref{fig:fmltt-selected} shows the $\beta$-rule of an example transformer,
 \ruleref{tmeq/ov/beta}.

### 2023-04-02 metatheory-condensed.tex
Mention J in the text.

@@
 So are their introduction and elimination forms.
-We use based path induction
-as the elimination principle for identity types~\citepalias{program2013homotopy}.
+We use based path induction $\Jeq{\cdot}{r}$
+as the elimination principle for 
+a term $r$ of identity types~\citepalias{program2013homotopy}. 
 Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
@@
 
-\emph{Linkage signatures} have the form $\goodSig{\Gm}{\lsig}{n}$.
-As rules \ruleref{lsig/empty} and \ruleref{lsig/add} show, a linkage
-signature is a list of $n$~types.
+\emph{Linkage signatures} have the judgment form $\goodSig{\Gm}{\lsig}{n}$.
+A linkage signature is a list of $n$~types
+(\ruleref{lsig/empty} and \ruleref{lsig/add}).
 %
-\emph{Linkages} have the form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
+\emph{Linkages} have the judgment form $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$,
 where $\TyLkg{\lsig}$ is a type formed by~$\lsig$.
-As rules \ruleref{l/empty} and \ruleref{l/add} show, a linkage is a list
-of $n$~terms, each representing a field of the family modeled by the linkage.
+A linkage is a list of $n$~terms, each representing a field of the family
+modeled by the linkage (\ruleref{l/empty} and \ruleref{l/add}).
 
-In rules \ruleref{l/add} and \ruleref{lsig/add},
+In the rules \ruleref{l/add} and \ruleref{lsig/add},
 the second premise $\goodTerm{\Gm, \mathit{self} : A}{t}{T}$ is
@@
 how the types of the fields prior to~$t$ are exposed to the typing of~$t$.
+%
+As discussed later, the third premise $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
+is responsible for creating the context type~$A$ that possibly hides W-type
+signatures in $\lsig$, which records the types of the prior fields.
+
 Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
@@
 $\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
-should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
+should see the concrete signature~$\wsig$, as the rules \ruleref{tm/wsup}
 and \ruleref{tm/wrec} stipulate.
@@
 
-In rules \ruleref{l/add} and \ruleref{lsig/add},
+In \ruleref{l/add} and \ruleref{lsig/add},
 the third premise $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$ is

### 2023-04-02 introduction.tex
Adding List of Abbrev.
TODO

--- 2023-04-01 introduction.tex
@@
 mechanism as a Coq plugin (\cref{sec:coqimpl}),
-available at \url{https://anonymized-link-to-code}.
+available at \url{https://anonymized-link-to-code}.\YZ{TODO}
 The plugin works by

### 2023-04-02 metatheory-condensed.tex
add a forward reference as suggested by shepherd
tweak
TR

@@
 named variables and meta-level substitution.
-For clarity of proof details, a technical appendix supplementing
-\cref{sec:metatheory2} instead uses \citeauthor{debruijn1964} indices and explicit substitution \cite{abadi1989subst}.
+For clarity of proof details,
+\ifreport
+\cref{sec:complete-fmltt}
+\else
+a technical appendix \cite{fpop-tr}
+\fi
+supplementing \cref{sec:metatheory2}
+instead uses \citeauthor{debruijn1964} indices and explicit substitution \cite{abadi1989subst}.
 %
@@
 So are their introduction and elimination forms.
-We use based path induction $\Jeq{\cdot}{r}$
+We use based path induction $\Jeq{\cdot}{t}$
 as the elimination principle for 
-a term $r$ of identity types~\citepalias{program2013homotopy}. 
+a term~$t$ of an identity type~\citepalias{program2013homotopy}. 
 Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
@@
 In the presence of hiding, it is important that the context type~$A$ be a
-dependent tuple type rather than a linkage type; we expand on this point in an
-appendix.}
+dependent tuple type rather than a linkage type; we expand on this point in
+\ifreport
+\cref{sec:complete-fmltt}.
+\else
+an appendix.
+\fi
+}
 
@@
 \ifreport \cref{sec:complete-fmltt}.
-\else a technical appendix.
+\else a technical appendix \cite{fpop-tr}.
 \fi

### 2023-04-05 casestudies.tex
cut to save space
back on 22 pages

@@
 
+\begin{wrapfigure}[13]{r}{0.132\textwidth}
+\includegraphics[scale=.66]{graphics/ai-casestudy.pdf}
+\end{wrapfigure}
+
 Our second case study is a mechanization of abstract interpreters
@@
 
-\begin{wrapfigure}[13]{r}{0.138\textwidth}
-\includegraphics[scale=.68]{graphics/ai-casestudy.pdf}
-\end{wrapfigure}
-
 \noindent
@@
 
-The next two families then extend \lsti{ImpGAI}.
+The next two families both extend \lsti{ImpGAI}.
 Family \lsti{ImpTI} (\textasciitilde200~LOC) is an abstract
-interpreter that does type inference~\cite{cousot1997types}.
+interpreter doing type inference~\cite{cousot1997types}.
 Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
@@
 
-We note that \Lang cannot yet allow the expression of two derived abstract
-interpreters nested within the same family.
-Work on nested inheritance~\cite{ncm2004,zm2017} points to a direction
-to further increase the expressive power of our language design.
+%We note that \Lang cannot yet allow the expression of two derived abstract
+%interpreters nested within the same family.
+%Work on nested inheritance~\cite{ncm2004,zm2017} points to a direction
+%to further increase the expressive power of our language design.
 

### 2023-04-05 related-work.tex
back on 22 pages

@@
 The design requires distinct names for a base inductive type and its
-extensions (including distinct names for constructors), while family
-polymorphism in our approach allows names to be late bound.
+extensions (including distinct names for constructors), while \Lang allows names to be late bound.
 

### 2023-04-05 frontmatters.tex
Adding List of Figures.
Adding list of figures.
List of Figures now working. But might fail sometime. Just be aware.
Fix TOC
Adding List of Abbrev.
Add backmatters
Adding Front Page
Update Acknowledgement

@@
 \pagenumbering{roman}
-\vspace*{\fill}
-\begin{center}
-  \textsc{\Huge \thetitle}
-\end{center}
 
-\vspace*{\fill}
+\input{titlepage.tex}
 
@@
 
-We thank Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
+We are grateful to the reviewers for their many useful comments.  We thank
+Benjamin Delaware,
+Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
 for valuable discussions and feedback.
@@
 reflect the position of any funding agency.
+
 \vspace*{\fill}
@@
 \tableofcontents
+
+\newpage
+
+\vspace*{\fill}
+
+\listoffigures
+
+\vspace*{\fill}
+
+\newpage
+
+\vspace*{\fill}
+\chapterN{List of Abbreviations}
+
+\input{listofabbrev.tex}
+
+\vspace*{\fill}
+
+% we omit list of symbols first
+% \newpage
+
+% \vspace*{\fill}
+% \chapterN{List of Symbols}
+
+\vspace*{\fill}

### 2023-04-05 metatheory2.tex
id
Update Acknowledgement
Removing mentioning of the appendix.

@@
 
-Unlike in \cref{sec:metatheory2}, which uses named binders, in this appendix we
+Unlike in \cref{sec:metatheory2}, which uses named binders, here we
 use \citeauthor{debruijn1964} indices and \emph{explicit
@@
 %This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
-Moreover, universe levels are explicit in this appendix. Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
+Moreover, universe levels are explicit here. Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
 
@@
 
-% Furthermore, the semantic model we develop in this appendix respects these
+% Furthermore, the semantic model we develop here respects these
 % judgmental equalities by construction.
@@
 where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
+We occasionally use the notation "id" for $\SubstId$.
 

### 2023-04-05 stlc-linkTran-typing.tex
Fix Wsup - Wsup needs specify the index of the constructor.
Removing mentioning of the appendix.

@@
-In this appendix, we sketch how to use the ``library'' of linkage
+Here, we sketch how to use the ``library'' of linkage
 transformers in \TT to inductively construct a linkage transformer that models a
@@
 \lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_{\texttt{tm}}'}}$                      \\ \cline{2-3} 
-\codecomment{existing constructors}   & 2        & $\InhOv{h_1}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$           \\ \cline{2-3} 
-\lsti!| tm_true |!\ \;\dadada    & 3        & $\InhExt{h_2}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$          \\ \cline{2-3} 
+\codecomment{existing constructors}   & 2        & $\InhOv{h_1}{\Wsup{}{\wsig_{\texttt{tm}}'}{\top}{\bot}}$           \\ \cline{2-3} 
+\lsti!| tm_true |!\ \;\dadada    & 3        & $\InhExt{h_2}{\Wsup{}{\wsig_{\texttt{tm}}'}{\top}{\bot}}$          \\ \cline{2-3} 
 \rowcolor[HTML]{CDCDCD} 

### 2023-04-17 lang-design.tex
Adding List of Abbrev.
Removing mentioning of the appendix.
Phrasing.

@@
 The reason we have this weird looking \ref{prop:detectable-partiality}
-is due to the construction of "linv"---at one point, using "prec" to
+is due to a technical detail of the construction of "linv".
+
+At one point, using "prec" to
 construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
-C) → option C}, generally not provable, unless we have
+C) → option C}. This is generally not provable, unless we have
 \mintinline{Coq}{option (Bᵢ a → C)} instead.\YZ{Don't you really mean
@@
 \EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.}
-This \ref{prop:detectable-partiality} is exactly transforming the former to
+
+This \ref{prop:detectable-partiality} can be used to transform the former to
 the latter:
@@
 To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable.
-% Please refer to the supplementary material and appendix for a formalization of the left inverse in the context of indexed inductive type.
 

### 2023-04-17 metatheory-condensed.tex
Removing mentioning of the appendix.
Move Why TyTkg into main text.

@@
 
+% Deal with intuition
+\noindentparagraph{Why $\TyTkg{\lsig}$?}
+
+%As shown by the equality rule \ruleref{tyeq/pk/add},
+%for a given linkage signature $\lsig \coloneq \LSigAdd{\cdot}{\cdot}{\cdot}$,
+%the type $\TyTkg{\lsig}$ is structurally similar to the linkage type $\TyLkg{\lsig}$
+%except that $\TyTkg{\lsig}$ is a dependent tuple type---%
+%it packages the fields of $\lsig$ into a dependent tuple type.
+%
+%We have computed several examples in
+%\cref{fig:stlc-linkage-typing}, where the reader should also notice the high
+%similarity between $\TyTkg{\lsig}$ and $\TyLkg{\lsig}$.
+
+$\TyTkg{\lsig}$ is needed in rules $\ruleref{lsig/add}$ and $\ruleref{l/add}$:
+these rules have a premise  $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
+responsible for possibly hiding W-type signatures (in $\TyTkg{\lsig}$) behind $\cU$ (in $A$);
+see $\seal_6$ in \cref{fig:stlc-linkage-typing} for an example.
+Here, both $\TyTkg{\lsig}$ and $A$ are dependent tuple types rather than linkage types.
+
+%Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
+$\TyTkg{\cdot}$ is indispensible because \emph{we cannot do abstraction with $\TyLkg{\cdot}$ alone}, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{lsig/add}$ and $\ruleref{l/add}$ is \emph{usually a dependent tuple type}.
+
+An alternative to $\TyTkg{\lsig}$ would be to use linkage types $\TyLkg{\cdot}$ in this premise.
+However, this alternative is infeasible.
+We illustrate using a simple example.
+Consider the linkage signature $\lsig$ below (it is written using the record
+syntax for illustration but should not be confused with records, as a later field
+is universally quantified over a $\mathit{self}$ variable).
+The job is to hide the W-types $\TyS{\wcode{\wsig_1}}$ and
+$\TyS{\wcode{\wsig_2}}$ in $\lsig$ behind $\cU$ in $\lsig^A$:
+%
+\begin{align*}
+  \lsig &\coloneq \{ T_A : \TyS{\wcode{\wsig_1}}; &&T_B : \TyS{\wcode{\wsig_2}}; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
+  \lsig^A &\coloneq \{T_A : \cU; &&T_B : \cU; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
+  & && \goodTerm{x : \TyLkg{\lsig}}{\seal}{\TyLkg{\lsig^A}}
+\end{align*}
+
+%We illustrate with a small example using named convention, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function $f$ maps from the first to the second type.
+Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{\lsig}}$.
+Say we want to add a new field into this linkage, and in $\ruleref{l/add}$, we chose $A$ to abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
+The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction (choose $s$).
+
+The problem lies in that the type $\TyLkg{\lsig^A}$ above cannot be inhabited.
+Suppose there is some $\lkg^A$ such that $\cdot \vdash \lkg^A : \TyLkg{\lsig^A}$.
+Then it follows from the self-parameterization that projecting out the last field~$f$
+from $\lkg^A$ (notated as $\lkg^A.f$ for illustration) has the typing
+$\goodTerm{T_A : \cU, T_B: \cU}{\lkg^A.f}{\El{T_A} \to \El{T_B}}$.
+This typing means that there is a function of type
+$\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A}
+\to \El{T_B}}}$, i.e., a function from an arbitrary type to another arbitrary
+type. We can easily derive inconsistency from this function:
+applying it to $\top$ and $\bot$ yields a function of type $\top\to\bot$.
+
+%\begin{align*}
+%  & \text{doing abstraction (choosing appropriate $s$ in \ruleref{l/add})} \\ 
+%  \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
+%  \text{Approach } \beta \quad&\quad  \TyLkg{\lsig} \longrightarrow \TyTkg{\lsig} \longrightarrow \TyTkg{\lsig^A}\\
+%\end{align*}
+%
+%However, the seemingly easier approach $\alpha$ is impossible because linkage
+%$\TyLkg{\lsig^A}$ is uninhabited---%
+%say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
+To make the inconsistency clearer, by double abstraction we will have an impossible function $g \coloneq \goodTerm{\cdot}{\nlam{T_A}{\nlam{T_B}{\ell^A.f}}}{\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A} \to \El{T_B}}}} $, i.e.
+a function maps from arbitrary type to another arbitrary type .
+We can easily derive inconsistency by applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
+%
+%Thus we can only choose the approach $\beta$ instead, which justifies why we have $\TyTkg{\cdot}$ in our typing rule, as $\TyTkg{\lsig}$ is usually the first step to do abstraction.
+
+Thus, it is critical that the hiding of W-type signatures operate on dependent
+tuple types rather than linkage types.
+To this end, we introduce the syntax $\TyTkg{\lsig}$ and its computation rules
+(e.g., \ruleref{tyeq/pk/add}) to make it convenient to package linkage signatures
+into dependent tuple types.
+
+%To summarize, the reason we need dependent tuple type is because of the overridability nature of
+
 \noindentparagraph{The complete formalization.}
@@
 % \ifreport \cref{sec:complete-fmltt}.
-% \else a supplemental appendix.
+% \else a 
 % \fi

### 2023-04-17 metatheory2.tex
Move Why TyTkg into main text.

@@
 
-% Deal with intuition
-\noindentparagraph{Why $\TyTkg{\lsig}$?}
-
-%As shown by the equality rule \ruleref{tyeq/pk/add},
-%for a given linkage signature $\lsig \coloneq \LSigAdd{\cdot}{\cdot}{\cdot}$,
-%the type $\TyTkg{\lsig}$ is structurally similar to the linkage type $\TyLkg{\lsig}$
-%except that $\TyTkg{\lsig}$ is a dependent tuple type---%
-%it packages the fields of $\lsig$ into a dependent tuple type.
-%
-%We have computed several examples in
-%\cref{fig:stlc-linkage-typing}, where the reader should also notice the high
-%similarity between $\TyTkg{\lsig}$ and $\TyLkg{\lsig}$.
-
-$\TyTkg{\lsig}$ is needed in rules $\ruleref{lsig/add}$ and $\ruleref{l/add}$:
-these rules have a premise  $\goodTerm{\Gm,x : \TyTkg{\lsig}}{\seal}{A}$
-responsible for possibly hiding W-type signatures (in $\TyTkg{\lsig}$) behind $\cU$ (in $A$);
-see $\seal_6$ in \cref{fig:stlc-linkage-typing} for an example.
-Here, both $\TyTkg{\lsig}$ and $A$ are dependent tuple types rather than linkage types.
-
-%Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity.
-$\TyTkg{\cdot}$ is indispensible because \emph{we cannot do abstraction with $\TyLkg{\cdot}$ alone}, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{lsig/add}$ and $\ruleref{l/add}$ is \emph{usually a dependent tuple type}.
-
-An alternative to $\TyTkg{\lsig}$ would be to use linkage types $\TyLkg{\cdot}$ in this premise.
-However, this alternative is infeasible.
-We illustrate using a simple example.
-Consider the linkage signature $\lsig$ below (it is written using the record
-syntax for illustration but should not be confused with records, as a later field
-is universally quantified over a $\mathit{self}$ variable).
-The job is to hide the W-types $\TyS{\wcode{\wsig_1}}$ and
-$\TyS{\wcode{\wsig_2}}$ in $\lsig$ behind $\cU$ in $\lsig^A$:
-%
-\begin{align*}
-  \lsig &\coloneq \{ T_A : \TyS{\wcode{\wsig_1}}; &&T_B : \TyS{\wcode{\wsig_2}}; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
-  \lsig^A &\coloneq \{T_A : \cU; &&T_B : \cU; &&f : \El{\getfield{\mathit{self}}{T_A}} \to \El{\getfield{\mathit{self}}{T_B}} \} \\
-  & && \goodTerm{x : \TyLkg{\lsig}}{\seal}{\TyLkg{\lsig^A}}
-\end{align*}
-
-%We illustrate with a small example using named convention, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function $f$ maps from the first to the second type.
-Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{\lsig}}$.
-Say we want to add a new field into this linkage, and in $\ruleref{l/add}$, we chose $A$ to abstract both inductive type to $\cU$, resulting something like $\lsig^A$.
-The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction (choose $s$).
-
-The problem lies in that the type $\TyLkg{\lsig^A}$ above cannot be inhabited.
-Suppose there is some $\lkg^A$ such that $\cdot \vdash \lkg^A : \TyLkg{\lsig^A}$.
-Then it follows from the self-parameterization that projecting out the last field~$f$
-from $\lkg^A$ (notated as $\lkg^A.f$ for illustration) has the typing
-$\goodTerm{T_A : \cU, T_B: \cU}{\lkg^A.f}{\El{T_A} \to \El{T_B}}$.
-This typing means that there is a function of type
-$\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A}
-\to \El{T_B}}}$, i.e., a function from an arbitrary type to another arbitrary
-type. We can easily derive inconsistency from this function:
-applying it to $\top$ and $\bot$ yields a function of type $\top\to\bot$.
-
-%\begin{align*}
-%  & \text{doing abstraction (choosing appropriate $s$ in \ruleref{l/add})} \\ 
-%  \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
-%  \text{Approach } \beta \quad&\quad  \TyLkg{\lsig} \longrightarrow \TyTkg{\lsig} \longrightarrow \TyTkg{\lsig^A}\\
-%\end{align*}
-%
-%However, the seemingly easier approach $\alpha$ is impossible because linkage
-%$\TyLkg{\lsig^A}$ is uninhabited---%
-%say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage.
-To make the inconsistency clearer, by double abstraction we will have an impossible function $g \coloneq \goodTerm{\cdot}{\nlam{T_A}{\nlam{T_B}{\ell^A.f}}}{\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A} \to \El{T_B}}}} $, i.e.
-a function maps from arbitrary type to another arbitrary type .
-We can easily derive inconsistency by applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.
-%
-%Thus we can only choose the approach $\beta$ instead, which justifies why we have $\TyTkg{\cdot}$ in our typing rule, as $\TyTkg{\lsig}$ is usually the first step to do abstraction.
-
-Thus, it is critical that the hiding of W-type signatures operate on dependent
-tuple types rather than linkage types.
-To this end, we introduce the syntax $\TyTkg{\lsig}$ and its computation rules
-(e.g., \ruleref{tyeq/pk/add}) to make it convenient to package linkage signatures
-into dependent tuple types.
-
-%To summarize, the reason we need dependent tuple type is because of the overridability nature of our linkages, together with the frequent usage of universe type in the context of dependent type programming.
-Once we want to abstract a type in a signature, we may have to use dependent tuple type as the abstraction instead of linkages.
-Since abstraction is usually dependent tuple type, we introduce $\TyTkg{\cdot}$ and its computation rules to make the process of doing abstraction in \TT easier.
-
 \noindentparagraph{Linkage transformers as syntactic sugar.}

### 2023-04-19 frontmatters.tex
Update statement of contributions.

@@
 
-Ende Jin and Prof. Yizhou Zhang co-authored Chapter 1.
+Ende Jin is the sole author of \cref{sec:metatheory2}, which were written under the under the supervision of Prof. Yizhou Zhang.
 
-Ende Jin is the sole author of Chapter 2, which were written under the  under the supervision of Prof. Yizhou Zhang.
+Ende Jin and Prof. Yizhou Zhang co-authored the rest of the thesis.
 
-Both Chapter 1 and Chapter 2 are already published result.
+This thesis is a published result.
 
-\noindentparagraph{Research presented in Chapter 1 and 2:} Ende Jin is the primary investigator of both theoretical development and practical implementation of the prototype.
+\noindentparagraph{Research presented in the thesis:} Ende Jin is the primary investigator of both theoretical development and practical implementation of the prototype.
 This work was supported in part by an Amazon research award on automated

### 2023-04-26 coqimpl-old.tex
Formatting.

@@
 is that the construction of "linv"---at one point, using "prec" to
-construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
+construct "linv", we need to prove \\ \mintinline{Coq}{∀ a, (Bᵢ a → option
 C) → option C}, generally not provable, unless we have
 \mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
-This \ref{prop:detectable-partiality} is exactly transforming former to
+To resolve that, the \ref{prop:detectable-partiality} can help to transform the former to
 the latter:

### 2023-04-26 lang-design.tex
Formatting.

@@
 At one point, using "prec" to
-construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
-C) → option C}. This is generally not provable, unless we have
+construct "linv", we need to prove \\ \mintinline{Coq}{∀ a, (Bᵢ a → option
+C) → option C}. This is generally not provable, unless we have \\
 \mintinline{Coq}{option (Bᵢ a → C)} instead.\YZ{Don't you really mean
 "unless we have the Detectable Partiality assumption instead"?}
-\EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.}
-
-This \ref{prop:detectable-partiality} can be used to transform the former to
+\EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.} To resolve that, the \ref{prop:detectable-partiality} can be used to transform the former to
 the latter:
@@
 \end{minted}
-To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable.
+To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}.
 
-\noindent
 As expected, "tm_prect_STLC" is axiomatized along with

### 2023-05-23 stlc-compiled.tex
fix typos and awkward wordings
Save. Fixing overflow.

@@
 [@\outdent@]Module [@\textls[-30]{STLC}@]\_[@\textls[-30]{subst}@]\_[@\textls[-30]{Cases}@] (self[@\,@]:[@\,@][@\textls[-30]{STLC}@]\_[@\textls[-30]{subst}@]\_[@\textls[-30]{Cases}@]\_[@\textls[-30]{Ctx}@]).
-Def subst\_tm_unit :=[@\codecomment{\textls[-20]{emmited upon definition of case}}@]
+Def subst\_tm_unit :=[@\codecomment{\textls[-20]{emitted upon definition of case}}@]
   \lambda (x : id) (t : self.tm), self.tm_unit.
@@
 Inductive tm : Set :=
-| tm_unit : tm | tm_var : id -> tm
-| tm_abs : id -> tm -> tm | tm_app : tm -> tm -> tm.
-[@\codecomment{Instantiate \texttt{tm} partial recursor \& its computational behaviors}@]
-Def tm_prect_STLC := \lambda P, tm_rect (\lambda[@\,@]t, option (P t)).
+| tm_unit | tm_var (v : id) [@\dadada@]
+[@\codecomment{Inst. partial recursor \& computational behaviors}@]
+Def tm_prect_STLC P := 
+      tm_rect (\lambda[@\,@]t, option (P t)).
 Fact [@\textls[-0]{tm\_unit\_eq\_STLC}@] : [@\dadada@]. reflexivity. Qed.

### 2023-05-23 stlc-mechanized.tex
trying out a name for the language design and impl
Save. Fixing overflow.

@@
 \begin{figure}
+\vspace{-10pt}
 \begin{minipage}{\textwidth}
@@
   motive \lambda G t T (_ : hasty G t T),
-  G = empty -> \forall t', step t t' -> hasty empty t' T.
+  G = empty -> \forall t', step t t' -> 
+  hasty empty t' T.
 Case ht_unit. [@\dadada@] Qed.[@\hfill@]Case ht_var. [@\dadada@] Qed.
@@
 | tm_fix : id -> tm -> tm.
-
-FRecursion subst on tm motive \lambda _, id -> tm -> tm.
+FRecursion subst on tm 
+      motive (\lambda _, id -> tm -> tm).
 Case tm_fix := \lambda y t' IHt' x t, [@\dadada@].
@@
 FInductive hasty : env -> tm -> ty -> Prop +=
-| ht_fix : \forall G x t T, hasty (extend G x T) t T ->
-  hasty G (tm_fix x t) T.
+| ht_fix : \forall G x t T, hasty (extend G x T) t T 
+-> hasty G (tm_fix x t) T.
 
@@
 \end{minipage}
-\caption{Using our plugin to mechanize STLC and the fixpoints extension,
+\caption{Using \Lang to mechanize STLC and the fixpoints extension,
 as envisioned in \cref{fig:stlc-nonmechanized}.}

### 2023-05-23 stlcfix-compiled.tex
minor
Save. Fixing overflow.

@@
 Inductive tm : Set :=
-| tm_unit : tm | tm_var : id -> tm
-| tm_abs : id -> tm -> tm | tm_app : tm -> tm -> tm
-| tm_fix : id -> tm -> tm.
-[@\codecomment{Instantiate \texttt{tm} partial recursors \& their comp.\ behaviors}@] [@\dadada@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
+| tm_unit | tm_var (v : id)
+| tm_abs (v : id) (b : tm) | tm_app (a b : tm)
+| tm_fix (v : id) (b : tm).
+[@\codecomment{Inst. \texttt{tm} partial recursors \& their comp.\ behaviors}@] [@\dadada@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
 Include STLCFix\_subst\_Cases. [@\tanc{c-STLCFix-d-subst-start}@]
@@
 subst\_tm_fix.
-Fact [@\textls[-40]{subst\_tm\_unit\_eq}@] : [@\dadada@]. [@\textls[-40]{{\color{dkblue}reflexivity}.}@] Qed.
+Fact [@\textls[-0]{subst\_tm\_unit\_eq}@] : [@\dadada@]. reflexivity. Qed.
 Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]

### 2023-06-03 lang-design.tex
Save.

@@
 a constructor "ht_fix" to "hasty", the programmer is required in family "STLCFix" to
-extend the proof of "substlem" to handle this extra case.
+extend the proof of "weakenlem" to handle this extra case.
 
@@
 For example, within family "STLC", the partial recursor for
-"tm" has the following type:
+"tm" has the following type, with extra equations describing
+its behvaiour on each constuctors:
 
@@
   [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
+(* For arbitrary [@P@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@], and we denote tmPR := tm_prect_STLC [@P@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@] *)
+(* We will have following equations *)
+tm_prect_STLC_unit : tmPR tm_unit      = [@$r_1$@]
+tm_prect_STLC_var  : tmPR (tm_var x)   = [@$r_2$@] x
+tm_prect_STLC_abs  : tmPR (tm_abs x b) = [@$r_3$@] a b (tmPR b)
+tm_prect_STLC_app  : tmPR (tm_app a b) = [@$r_4$@] a (tmPR a) b (tmPR b)
 \end{lstlisting}
@@
 
+This corresponds to the dependent eliminator for the non-extensible version of "tm".
+
 \noindentparagraph{Property about Partial Recursors.}
 
-Partial recursor is simply adding partiality into the signature of the eliminator, by wrapping each motive with "option".
+A partial recursor is simply adding partiality into the signature of the ordinary eliminator---by wrapping each motive with "option"---together 
+with the equations describing its computational behaviors inherited from the ordinary eliminators.
+
 This partial recursor seems non-canonical compared to the (dependent) eliminator for Inductive type as the latter can be considered as universal property.
@@
 the fact that propositional partial
-recursors can derive injectivity alreadt demonstrate itself as a good \textit{extensional characterization} of
+recursors can derive injectivity already demonstrate itself as a good \textit{extensional characterization} of
 (non-indexed) extensible inductive type (under certain

### 2023-06-03 lang-design.tex
Fixing prec.

@@
   [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
-(* For arbitrary [@P@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@], and we denote tmPR := tm_prect_STLC [@P@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@] *)
+(* For arbitrary [@$P$@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@], and we denote [@"tmPR"@] := [@"tm_prect_STLC"@] [@$P$@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@] *)
 (* We will have following equations *)
@@
 \end{theorem}
-The reason we have this weird looking \ref{prop:detectable-partiality}
-is due to a technical detail of the construction of "linv".
+\begin{proof} 
+First, for a fixed \mintinline{Coq}{x : Aᵢ}, given \mintinline{Coq}{fᵢ :  Bᵢ x -> T},
+\\ we have \mintinline{Coq}{lift fᵢ : Bᵢ x -> option T}, simply wrapping option around the returning result.
 
-At one point, using "prec" to
-construct "linv", we need to prove \\ \mintinline{Coq}{∀ a, (Bᵢ a → option
-C) → option C}. This is generally not provable, unless we have \\
-\mintinline{Coq}{option (Bᵢ a → C)} instead.\YZ{Don't you really mean
-"unless we have the Detectable Partiality assumption instead"?}
-\EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.} To resolve that, the \ref{prop:detectable-partiality} can be used to transform the former to
-the latter:
-recall that, according to the semantic of inductive type,
-\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
-"linv" on the substructure of "D". This function points out each of the
-translation result from "D" to "C" and some of them are "None" because
-of failing. For us, we only care about if \textbf{any} of the
-substructure of "D" fails to translate---if any substructure fails,
-then the whole translation fails and thus we should have "None" as
-return.
+Second, with \ref{prop:detectable-partiality} we can construct
+\begin{minted}{Coq}
+Definition unlift:
+  forall {a} {T : Set} (f: (forall c: B a, option T)) 
+                       (h : forall x, f x <> None),
+    {g : (forall c: B a, T) | f = lift g}.
+
+Theorem detect_lifted:
+  forall {a} {T : B a -> Type} (f: (forall c: B a, option (T c))),
+    { g | f = lift g} + {exists x, f x = None}.
+\end{minted}
+The idea of "unlift" is that if "f" never returns "None", then it is always a result of a "lift". "detect_lifted" is just another way of saying  \ref{prop:detectable-partiality}.
+
+Then we can construct "inj" and "linv" as follows, using eliminator and partial recursor.
+\begin{minted}{Coq}
+Definition inj : C -> D :=
+  C_rect (fun (_ : C) ⇒ D) 
+    ...
+    (fun (xᵢ : Aᵢ) (_ : Bᵢ x -> Cᵢ) (Xᵢ : Bᵢ x -> D) ⇒ dᵢ xᵢ Xᵢ)
+    ... .
+
+Definition linv : D -> option C :=
+  prec C 
+    ...
+    (fun (aᵢ : Aᵢ) (bᵢ' : Bᵢ aᵢ → option C) ⇒
+       match (detectable bᵢ') with
+       | left h =>
+           let (bᵢ, _) := unlift bᵢ' h in Some (cᵢ aᵢ bᵢ)
+       | right _ => None
+       end)
+    ... .
+\end{minted}
+Since each ``constructor'' in "C" and "D" are well-correpsonded, when invoking "C_rect" and "prec", each case is handled similarly, and they only vary along the $\{ "A"_i \vdash "B"_i \}_{i}$ types.
+
+The definition of "inj" and "linv" is well-formed but their behaviour is still burried in the definition. The following two equations can clarify them.
+
+\begin{minted}{Coq}
+(* inj_compute can be proved thanks to 
+    the computational behaviour about C_rect built-in Coq *)
+Lemma inj_compute  : forall a b,
+  inj (cᵢ a b)  = dᵢ a (fun x => inj (b x)).
+
+(* linv_compute can be proved thanks to 
+    the n (propositional) computational axioms about prec *)
+Lemma linv_compute : forall a b,
+  linv (dᵢ a b) = 
+    let b' := fun x => linv (b x) in 
+    match detect_lifted b' with
+    | inleft (exist _ x _) => Some (cᵢ a x)
+    | inright _ => None
+    end.
+\end{minted}
+
+Finally,  these two equations is enough to prove that "linv" is really a left inverse, via induction
+\begin{minted}{Coq}
+linv (inj (cᵢ a b)) 
+= linv (dᵢ a (fun x ⇒ inj (b x)))         (* by inj_compute *)
+= Some (cᵢ a (fun x ⇒ linv (inj (b x))))  (* by linv_compute *)
+= Some (cᵢ a (fun x => b x))              (* by I.H. *)
+= Some (cᵢ a b)                           (* by eta rule *)
+\end{minted}
+\end{proof}
+
+% The reason we have this weird looking \ref{prop:detectable-partiality}
+% is due to a technical detail of the construction of "linv".
+
+% At one point, using "prec" to
+% construct "linv", we need to prove \\ \mintinline{Coq}{∀ a, (Bᵢ a → option
+% C) → option C}. This is generally not provable, unless we have \\
+% \mintinline{Coq}{option (Bᵢ a → C)} instead.\YZ{Don't you really mean
+% "unless we have the Detectable Partiality assumption instead"?}
+% \EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.} To resolve that, the \ref{prop:detectable-partiality} can be used to transform the former to
+% the latter:
+% recall that, according to the semantic of inductive type,
+% \mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
+% "linv" on the substructure of "D". This function points out each of the
+% translation result from "D" to "C" and some of them are "None" because
+% of failing. For us, we only care about if \textbf{any} of the
+% substructure of "D" fails to translate---if any substructure fails,
+% then the whole translation fails and thus we should have "None" as
+% return.
 
@@
 To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}.
+We have mechanized the argument for indexed inductive type and dependent eliminator, but for the case we only have one constructor in "C" and "D".
+This mechanization is a direct generalization of the proof given in
 

### 2023-06-03 lang-design.tex
Fix typo for Prec.

@@
 (* We will have following equations *)
-tm_prect_STLC_unit : tmPR tm_unit      = [@$r_1$@]
-tm_prect_STLC_var  : tmPR (tm_var x)   = [@$r_2$@] x
-tm_prect_STLC_abs  : tmPR (tm_abs x b) = [@$r_3$@] a b (tmPR b)
-tm_prect_STLC_app  : tmPR (tm_app a b) = [@$r_4$@] a (tmPR a) b (tmPR b)
+tm_unit_eq_STLC : tmPR tm_unit      = [@$r_1$@]
+tm_var_eq_STLC  : tmPR (tm_var x)   = [@$r_2$@] x
+tm_abs_eq_STLC  : tmPR (tm_abs x b) = [@$r_3$@] a b (tmPR b)
+tm_app_eq_STLC  : tmPR (tm_app a b) = [@$r_4$@] a (tmPR a) b (tmPR b)
 \end{lstlisting}
@@
 
-This corresponds to the dependent eliminator for the non-extensible version of "tm".
+These four equations uniquely describe the "tm_prect_STLC" on the four constructors and thus "tm_prect_STLC" cannot be a trivial function.
+This partial recursor also corresponds to the dependent eliminator for the non-extensible version of "tm".
 
@@
       Axiom detectable:
-      forall {T} {a} (f: Bᵢ a -> option T),
+      forall {T} {a} (f: Bᵢ a → option T),
         {forall x, f x <> None} + {exists x, f x = None}.
@@
     \begin{minted}{Coq}
-Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
+Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x → C) → C | ...
     \end{minted}
@@
     \begin{itemize}
-      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
+      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x → D) → D}  
       \item a partial recursor
@@
 prec : forall (R : Set), ..,
-  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,@\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint.
+  (rᵢ : (forall (x : Aᵢ), (Bᵢ x → option R) → option R)), ..,@\YZ{Can R be indexed?}\EDJreply{This part actually currently incorrect and needs an extra constraint.
 Let me fix other parts first}\YZreply{Is this constraint Detectable Partiality?}@
-      D -> option R
+      D → option R
       \end{minted}
@@
 prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
-    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
+    = rᵢ aᵢ (fun x ⇒ prec {R} r₁ r₂ .. rₙ (bᵢ x))
       \end{minted}
@@
   Then there exists an \textbf{embedding} from "C" to "D".
-  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
-  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of\/
+  More concretely, we can have \mintinline{Coq}{inj : C → D} and 
+  \mintinline{Coq}{linv : D → option C} defined using the eliminator of\/
   "C" and "prec" of\/ "D" (both acting like identity) such that
@@
 \begin{proof} 
-First, for a fixed \mintinline{Coq}{x : Aᵢ}, given \mintinline{Coq}{fᵢ :  Bᵢ x -> T},
-\\ we have \mintinline{Coq}{lift fᵢ : Bᵢ x -> option T}, simply wrapping option around the returning result.
+First, for a fixed \mintinline{Coq}{x : Aᵢ}, given \mintinline{Coq}{fᵢ :  Bᵢ x → T},
+\\ we have \mintinline{Coq}{lift fᵢ : Bᵢ x → option T}, simply wrapping option around the returning result.
 
@@
 Theorem detect_lifted:
-  forall {a} {T : B a -> Type} (f: (forall c: B a, option (T c))),
+  forall {a} {T : Set} (f: (forall c: B a, option T)),
     { g | f = lift g} + {exists x, f x = None}.
@@
 \begin{minted}{Coq}
-Definition inj : C -> D :=
+Definition inj : C → D :=
   C_rect (fun (_ : C) ⇒ D) 
     ...
-    (fun (xᵢ : Aᵢ) (_ : Bᵢ x -> Cᵢ) (Xᵢ : Bᵢ x -> D) ⇒ dᵢ xᵢ Xᵢ)
+    (fun (xᵢ : Aᵢ) (_ : Bᵢ x → Cᵢ) (Xᵢ : Bᵢ x → D) ⇒ dᵢ xᵢ Xᵢ)
     ... .
 
-Definition linv : D -> option C :=
+Definition linv : D → option C :=
   prec C 
@@
        match (detectable bᵢ') with
-       | left h =>
+       | left h ⇒
            let (bᵢ, _) := unlift bᵢ' h in Some (cᵢ aᵢ bᵢ)
-       | right _ => None
+       | right _ ⇒ None
        end)
@@
 Lemma inj_compute  : forall a b,
-  inj (cᵢ a b)  = dᵢ a (fun x => inj (b x)).
+  inj (cᵢ a b)  = dᵢ a (fun x ⇒ inj (b x)).
 
@@
   linv (dᵢ a b) = 
-    let b' := fun x => linv (b x) in 
+    let b' := fun x ⇒ linv (b x) in 
     match detect_lifted b' with
-    | inleft (exist _ x _) => Some (cᵢ a x)
-    | inright _ => None
+    | inleft (exist _ x _) ⇒ Some (cᵢ a x)
+    | inright _ ⇒ None
     end.
@@
 = Some (cᵢ a (fun x ⇒ linv (inj (b x))))  (* by linv_compute *)
-= Some (cᵢ a (fun x => b x))              (* by I.H. *)
+= Some (cᵢ a (fun x ⇒ b x))              (* by I.H. *)
 = Some (cᵢ a b)                           (* by eta rule *)
@@
 \begin{minted}{Coq}
-  pdelim : ∀ (P : D -> Set), .., 
-  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
-    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
+  pdelim : ∀ (P : D → Set), .., 
+  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x → D), (∀ (b : Bᵢ x), option (P (w b))) 
+    → option (P (dᵢ x w)))),  .., ∀ (d : D) → option (P d)
 \end{minted}
 To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}.
-We have mechanized the argument for indexed inductive type and dependent eliminator, but for the case we only have one constructor in "C" and "D".
-This mechanization is a direct generalization of the proof given in
-
-As expected, "tm_prect_STLC" is axiomatized along with
+We have mechanized the argument for indexed W type and dependent eliminator, but for the case with only one constructor in "C" and "D".
+This mechanization\footnote{\url{https://gist.github.com/DKXXXL/9b4e48350b966fa3739cfb71f1a9a852}} is a direct generalization of the proof for \cref{thm:prec-complete}.
+
+As mentioned, "tm_prect_STLC" is axiomatized along with
 four equalities describing its computational behaviors, one for each constructor.
-For instance, the equality for constructor "tm_abs" is as follows:
+After unfold the definition of "tmPR", the equality for constructor "tm_abs" is as follows:
 

### 2023-06-03 lang-design.tex
Fix typo.

@@
 
-These four equations uniquely describe the "tm_prect_STLC" on the four constructors and thus "tm_prect_STLC" cannot be a trivial function.
-This partial recursor also corresponds to the dependent eliminator for the non-extensible version of "tm".
+"tm_prect_STLC" cannot be a trivial function even though its return type has "option".
+Because these four equations uniquely describe the "tm_prect_STLC" on the four constructors and make "tm_prect_STLC" behave similar to the ordinary dependent eliminator of the non-extensible version of "tm".
 
@@
 axioms)}.  Thus, we can argue that the partial recursor gives a sound and
-complete extensional characterization of all the extension of a given (non-indexed)
+complete extensional characterization of all the extensions of a given (non-indexed)
 inductive type.\YZ{What about indexed inductive types?}\EDJreply{Done.

### 2023-06-03 metatheory2.tex
Fix typo.

@@
 
-\noindentparagraph{MLTT as a programming language.}  Conventionally, the operational semantic sketches \emph{the interpreter} of a given formalization of a programming language.
-However, compared to this traditional setup, a careful reader will notice the absence of operational semantic in our MLTT.
+\noindentparagraph{MLTT as a programming language.}  Conventionally, the operational semantics sketches \emph{the interpreter} of a given formalization of a programming language.
+However, compared to this traditional setup, a careful reader will notice the absence of operational semantics in our MLTT.
 It is no doubt that MLTT is a logical system where we can write proofs, but it is unclear if MLTT can really be considered as a programming language.
@@
 as a function mapping from those quotiented data (syntax). The (semantic) model introduced below 
-will follow this idea. More specifically, the semantic model as a function will have to respects 
+will follow this idea. More specifically, the semantic model as a function will have to respect 
 judgemental equalities to be a well-defined function from quotiented data. In other words,

### 2023-06-03 ttintro.tex
Fix typo.

@@
-\noindentparagraph{\TT as a programming language.}  Conventionally, the operational semantic sketches \emph{the interpreter} of a given formalization of a programming language.
-However, compared to this traditional setup, a careful reader will notice the absence of operational semantic in our \TT and doubt if \TT can really be considered as a programming language.
+\noindentparagraph{\TT as a programming language.}  Conventionally, the operational semantics sketches \emph{the interpreter} of a given formalization of a programming language.
+However, compared to this traditional setup, a careful reader will notice the absence of operational semantics in our \TT and doubt if \TT can really be considered as a programming language.
 

### 2023-06-03 lang-design.tex
Fix Grammar.

@@
 "tm" has the following type, with extra equations describing
-its behvaiour on each constuctors:
+its behavior on each constuctor:
 
@@
 (* For arbitrary [@$P$@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@], and we denote [@"tmPR"@] := [@"tm_prect_STLC"@] [@$P$@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@] *)
-(* We will have following equations *)
+(* We will have the following equations *)
 tm_unit_eq_STLC : tmPR tm_unit      = [@$r_1$@]
@@
 "tm_prect_STLC" cannot be a trivial function even though its return type has "option".
-Because these four equations uniquely describe the "tm_prect_STLC" on the four constructors and make "tm_prect_STLC" behave similar to the ordinary dependent eliminator of the non-extensible version of "tm".
+Because these four equations uniquely describe the "tm_prect_STLC" on the four constructors and make "tm_prect_STLC" behave similarly to the ordinary dependent eliminator of the non-extensible version of "tm".
 
@@
 
-This partial recursor seems non-canonical compared to the (dependent) eliminator for Inductive type as the latter can be considered as universal property.
+This partial recursor seems non-canonical compared to the (dependent) eliminator for inductive types as the latter can be considered a universal property.
 This extensionality makes dependent eliminator one of \textit{the most powerful} reasoning tools around inductive type.
@@
 constraints)---because we can ``embed'' vanilla inductive types into
-types supporting a partial recursor: a bit more formally, and
+types supporting a partial recursor: a bit more formally speaking, and
 restricting our focus to non-indexed inductive types:
@@
   \begin{minted}[escapeinside=@@]{Coq}
-    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correpsond to 'reflect' above?
+    forall c : C, linv (inj c) = some c@\YZ{Does 'linv' correspond to 'reflect' above?
 If so, then the reader should be explicitly reminded of this fact rather than having to guess it.}\EDJreply{Added below.
@@
 \end{minted}
-Since each ``constructor'' in "C" and "D" are well-correpsonded, when invoking "C_rect" and "prec", each case is handled similarly, and they only vary along the $\{ "A"_i \vdash "B"_i \}_{i}$ types.
-
-The definition of "inj" and "linv" is well-formed but their behaviour is still burried in the definition. The following two equations can clarify them.
+Since each ``constructor'' in "C" and "D" are well-corresponded, when invoking "C_rect" and "prec", each case is handled similarly, and they only vary along the $\{ "A"_i \vdash "B"_i \}_{i}$ types.
+
+The definition of "inj" and "linv" is well-formed but their behavior is still buried in this complicated formulation. The following two equations can clarify them.
 
@@
 
-Finally,  these two equations is enough to prove that "linv" is really a left inverse, via induction
+Finally,  these two equations are enough to prove that "linv" is really a left inverse, via induction
 \begin{minted}{Coq}
@@
 
-\ref{prop:detectable-partiality} is not generally supported when using function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
-However, in those cases, we still have partial recursor provable and useful---we just cannot see the partial recursor as a strong toolkit anymore.
+\ref{prop:detectable-partiality} is not generally supported when using a function (infinitary branching) in the constructors, for example, the case of using \textit{higher order abstract syntax} to represent functions in "tm".
+However, in those cases, we still have the partial recursor provable and useful---we just cannot see the partial recursor as a strong toolkit anymore.
 
 We emphasize that in \cref{thm:prec-complete}, each function $"d"_i$ can
-actually be considered as a constructor, because they can be reflected
+actually be considered a constructor because they can be reflected
 to real constructors of "C" using the left inverse "linv". Thus, \cref{thm:prec-complete}
@@
 
-We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
+We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to construct this left inverse for (non-indexed) inductive type. 
 \begin{minted}{Coq}
@@
 
-As mentioned, "tm_prect_STLC" is axiomatized along with
+As mentioned earlier, "tm_prect_STLC" is axiomatized along with
 four equalities describing its computational behaviors, one for each constructor.
-After unfold the definition of "tmPR", the equality for constructor "tm_abs" is as follows:
+After unfolding the definition of "tmPR", the equality for constructor "tm_abs" is as follows:
 

### 2023-06-03 lang-design.tex
Fix Grammar

@@
 
-We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to construct this left inverse for (non-indexed) inductive type. 
+We can support a partial dependent eliminator as well. Though the partial dependent eliminator can deduce "prec", our "prec" is enough to construct this left inverse for (non-indexed) inductive type. 
 \begin{minted}{Coq}

### 2023-06-03 lang-design.tex
Fix typo.

@@
 the fact that propositional partial
-recursors can derive injectivity already demonstrate itself as a good \textit{extensional characterization} of
+recursors can derive injectivity already demonstrates itself as a good \textit{extensional characterization} of
 (non-indexed) extensible inductive type (under certain

### 2023-06-05 introduction.tex
Fix Typo.

--- 2023-04-01 introduction.tex
@@
 mechanism as a Coq plugin (\cref{sec:coqimpl}),
-available at \url{https://anonymized-link-to-code}.
+available at \url{https://github.com/DKXXXL/FPOP}.\YZ{TODO}
 The plugin works by

### 2023-06-05 lang-design.tex
Fix Typo.

@@
 
-\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a"}---that includes all of the examples shown in this paper.
+\ref{prop:detectable-partiality} is derivable for all the finitary branching inductive type\footnote{By simply enumerating through the finite domain "Bᵢ a".}---that includes all of the examples shown in this paper.
 

### 2023-06-05 metatheory2.tex
Fix Typo.

@@
 We review the base MLTT fragment of \TT first.
+The following rules are the fully expanded version of \cref{fig:mltt-selected}, where each type (except universes) has four kinds of rules: introduction, elimination, computation and uniqueness rules.
+Computation rules indicate the judgemental equality when elimination rules apply to the introduction form.
+Uniqueness rules indicate how some terms are possibly uniquely constructed by introduction rules.
+The substitution rule is in the form of de Bruijn indices.
+
+The first four rules are following the style of \citet{program2013homotopy, pfenning2009lecture}, while the substitution rules are naturally required for any language with first-class functions.
 \input{all-mltt-typing}
@@
 linkage signatures and linkages, and linkage transformers.
+The following rules are the fully expanded version of \cref{fig:fmltt-selected}, where each type has five kinds of rules: introduction, elimination, computation, uniqueness rules and substitution lemma.
+Computation rules indicate the judgemental equality when elimination rules apply to the introduction form.
+Uniqueness rules indicates how some terms are possibly uniquely constructed by introduction rules.
+The substitution rule is in the form of de Bruijn indices.
+
+The first four rules are following the style of \citet{program2013homotopy, pfenning2009lecture}, while the substitution rules are naturally required for any language with first-class functions.
+
+Apart from types, we also lay out the complete version of judgements for signatures and linkage transformers.
+The signature judgements are also following these four parts---introduction, elimination, computation and uniqueness.
 

### 2023-06-05 metatheory2.tex
Fix typo.
Fix typo

@@
 The following rules are the fully expanded version of \cref{fig:mltt-selected}, where each type (except universes) has four kinds of rules: introduction, elimination, computation and uniqueness rules.
-Computation rules indicate the judgemental equality when elimination rules apply to the introduction form.
-Uniqueness rules indicate how some terms are possibly uniquely constructed by introduction rules.
+Computation rules ($\beta$-rules) indicate the judgemental equality for expressing reduction when elimination rules apply to the introduction form.
+Uniqueness rules ($\eta$-rules) indicate how some terms are possibly uniquely constructed by introduction rules.
 The substitution rule is in the form of de Bruijn indices.
@@
 The following rules are the fully expanded version of \cref{fig:fmltt-selected}, where each type has five kinds of rules: introduction, elimination, computation, uniqueness rules and substitution lemma.
-Computation rules indicate the judgemental equality when elimination rules apply to the introduction form.
-Uniqueness rules indicates how some terms are possibly uniquely constructed by introduction rules.
+Computation rules ($\beta$-rules) indicate the judgemental equality for expressing reduction when elimination rules apply to the introduction form.
+Uniqueness rules ($\eta$-rules) indicates how some terms are possibly uniquely constructed by introduction rules.
 The substitution rule is in the form of de Bruijn indices.
@@
 Apart from types, we also lay out the complete version of judgements for signatures and linkage transformers.
-The signature judgements are also following these four parts---introduction, elimination, computation and uniqueness.
+The signature judgements are also following these five parts---introduction, elimination, computation, uniqueness and substitution rules.
 

### 2023-06-05 metatheory2.tex
Fix typo.

@@
 The following rules are the fully expanded version of \cref{fig:mltt-selected}, where each type (except universes) has four kinds of rules: introduction, elimination, computation and uniqueness rules.
-Computation rules ($\beta$-rules) indicate the judgemental equality for expressing reduction when elimination rules apply to the introduction form.
-Uniqueness rules ($\eta$-rules) indicate how some terms are possibly uniquely constructed by introduction rules.
+Computation rules ($\beta$-equalities) indicate the judgemental equality for expressing reduction when elimination rules apply to the introduction form.
+Uniqueness rules ($\eta$-equalities) indicate how some terms are possibly uniquely constructed by introduction rules.
 The substitution rule is in the form of de Bruijn indices.
@@
 The following rules are the fully expanded version of \cref{fig:fmltt-selected}, where each type has five kinds of rules: introduction, elimination, computation, uniqueness rules and substitution lemma.
-Computation rules ($\beta$-rules) indicate the judgemental equality for expressing reduction when elimination rules apply to the introduction form.
-Uniqueness rules ($\eta$-rules) indicates how some terms are possibly uniquely constructed by introduction rules.
+Computation rules ($\beta$-equalities) indicate the judgemental equality for expressing reduction when elimination rules apply to the introduction form.
+Uniqueness rules ($\eta$-equalities) indicate how some terms are possibly uniquely constructed by introduction rules.
 The substitution rule is in the form of de Bruijn indices.

### 2023-06-07 canonicitymodel.tex
\setc
Explain another notation for dependent function.
Make colorbox consistent.
Save. Fixing overflow.
Fix overflow.
Save.

@@
 \begin{align*}
-  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \setc{\gm}{\goodSub{\cdot}{\gm}{\Gm}} \to \Set{k} \quad \text{ (i.e., sets indexed by closed substitution)}\\
+  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \setc{\gm}{\goodSub{\cdot}{\gm}{\Gm}} \to \Set{k} \quad \\
+   & \text{ (i.e., sets indexed by closed substitution)}\\
   % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
@@
   \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\}  \\ 
-  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \setc{ (\gamma' , t') }{ \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  } \\
+  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \setc{ (\gamma' , t') }{ \gm' \in \denotesC{{\Gm}}(\pi_1 \gm_t), t' \in \denotesC{{\sub{T}{\gm}}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  } \\
   \denotesC{\goodSub{\Gm}{\pi_1 \delta}{\Dl}}(\gm)(\gm') &= \denotesC{\delta}(\gm)(\gm')[0] \quad \text{ get the first element of the tuple} \\
@@
   \denotesC{\goodTerm{\Gm}{\Jeq{w}{t}}{\sub{C}{\SubstExt{\SubstExt{id}{v}}{t} }}}(\gm)(\gm') &= \denotesC{w}(\gm)(\gm') \quad \quad \text{ given } \goodTerm{\Gm}{t}{\TyId{u}{v}} \\ 
-  & \text{Since $\denotesC{t}(\gm)(\gm')$ witnesses } \sub{t}{\gm} \equiv \eqrefl{\sub{u}{\gm}} \text{ and } \sub{u}{\gm} \equiv \sub{v}{\gm} \\
+   \text{Since $\denotesC{t}(\gm)(\gm')$ witnesses } &\sub{t}{\gm} \equiv \eqrefl{\sub{u}{\gm}} \text{ and } \sub{u}{\gm} \equiv \sub{v}{\gm} \\
   \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}}(\gm)(\gm') &= \lambda u \lambda u' . \denotesC{t}(\SubstExt{\gm}{u}) (\gm' , u') \\
@@
 for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$;
-etc. The audience should reminisce these structure with the one appearing at the beginning of the MLTT canonicity model.
+etc. These structures should remind the reader of the ones appearing at the beginning of the MLTT canonicity model.
 
@@
   \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~
-  \colorbox{blue!10}{(\Gsub{\denotesCC{a}}{\Glued{\gm}})}~
-  \colorbox{green!10}{(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})} \\
+  \colorbox{blue!9}{(\Gsub{\denotesCC{a}}{\Glued{\gm}})}~
+  \colorbox{green!9}{(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})} \\
   % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
@@
   \texttt{let}~\Glued{\rho}  \in \GTm{
-    \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} \colorbox{blue!5}{\Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}}})}
+    \colorbox{blue!17}{(\Glued{\cdot} \Glued{,} \colorbox{blue!9}{\Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}}})}
     }{
-      \colorbox{green!10}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})}
+      \colorbox{green!9}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})}
     } 
@@
       \colorbox{blue!27}{\Gapp{
-        \colorbox{blue!18}{\Gsub{
+        \colorbox{blue!17}{\Gsub{
           \colorbox{blue!9}{\Gapp{
-            \colorbox{blue!0}{\GRecproj{j}{\Gsub{\denotesCC{\lkg}}{\Glued{\gm}}}}
+            \colorbox{blue!1}{\GRecproj{j}{\Gsub{\denotesCC{\lkg}}{\Glued{\gm}}}}
             }}
@@
         \GSubstExt{\GSubstId}{
-          \colorbox{green!10}{\Glam{\Glued{\rho}}}
+          \colorbox{green!9}{\Glam{\Glued{\rho}}}
           }
@@
 The semantic interpretations above are defined together with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)
+
+(Note: we also use $(a \in A \to B(a))$ as another notation for dependent function $\prod_{a \in A} B(a)$).
 %
 \begin{align*}
-  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
+  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \setc{t}{\goodTerm{\cdot}{t}{\El{\wcode{\wsig}}}} \to \Set{i + 1} \quad "where" \\
     \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\
     & \to \Glued{b} \in (\GTm{
-      \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} 
-      \colorbox{blue!5}{\Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}} )} 
+      \colorbox{blue!17}{(\Glued{\cdot} \Glued{,} 
+      \colorbox{blue!9}{\Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}} )} 
       }{
-        \colorbox{green!10}{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}
+        \colorbox{green!9}{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}
         }) \\
@@
   \CWrec \  &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
-    (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
+    (P : \setc{t}{\goodTerm{\cdot}{t}{\El{\wcode{\wsig}}}} \to \Set{k})  \\
     & \to 
@@
     \setc{b}{\goodTerm{({\cdot} , 
-    \colorbox{blue!10}{\sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}}
+    \colorbox{blue!9}{\sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}}
     ) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
@@
       \left( {\Glued{\gm} \in \GSub{\Glued{\cdot}}{
-        \colorbox{blue!11}{(\Glued{\cdot} \Glued{,} 
-    \colorbox{blue!1}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}}
+        \colorbox{blue!17}{(\Glued{\cdot} \Glued{,} 
+    \colorbox{blue!9}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}}
     }\to P~(\sub{b}{ \gm }) \right) \\

### 2023-06-07 introduction.tex
Save.

@@
 logical reasoning, however, poses significant technical challenges.
-As we analyze, pillars of dependent type theories—including
+As our analysis
+will show, pillars of dependent type theories—including
 inductive types, definitional equality, and logical consistency—are

### 2023-06-08 frontmatters.tex
Update contribution section.
Fix Typo.
Fix frontmatters.

@@
 
+I understand that my thesis may be made electronically available to the public.
+
 \vspace*{\fill}
@@
 
-Ende Jin is the sole author of \cref{sec:metatheory2}, which were written under the under the supervision of Prof. Yizhou Zhang.
+Ende Jin was the sole author of \cref{sec:complete-fmltt}, which were written under the supervision of Prof. Yizhou Zhang.
 
-Ende Jin and Prof. Yizhou Zhang co-authored the rest of the thesis.
+Jin E, Amin N, Zhang Y.
+Extensible Metatheory Mechanization via Family Polymorphism: Technical Report.
+Technical Report CS-2023-01.
+School of Computer Science, University of Waterloo.
+https://cs.
+uwaterloo.
+ca/sites/ca.
+computer-science/files/uploads/files/cs-2023-01.
+pdf; 2023.
+\\
 
-This thesis is a published result.
+This thesis consists in part of a manuscript written for publication. The exception to sole authorship of material is as follows:
 
-\noindentparagraph{Research presented in the thesis:} Ende Jin is the primary investigator of both theoretical development and practical implementation of the prototype.
-This work was supported in part by an Amazon research award on automated
-reasoning.
+\noindentparagraph{Research presented in rest of the thesis:}
+The research was conducted at the University of Waterloo by Ende Jin under the supervision of Dr.
+Yizhou Zhang.
+Ende Jin proposed the primary design of the language, prototyped the Coq plugin for realizing the language, implemented and evaluated the case studies, and formalized and proved meta-theoretic arguments for the core type theory.
+Ende Jin estabilished the complete language design with consultations from Dr.
+Yizhou Zhang.
+
+Jin E, Amin N, Zhang Y. Extensible Metatheory Mechanization via Family Polymorphism. Proc. of the ACM on Programming Languages (PACMPL). 2023;7. DOI: https://doi.org/10.1145/3591286
 
@@
 
-We are grateful to the reviewers for their many useful comments.  We thank
+I am grateful to the reviewers for their many useful comments.  I thank
 Benjamin Delaware,
-Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
-for valuable discussions and feedback.
+Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, Zhixuan Yang, Prabhakar Ragde, Werner Dietl
+for valuable discussions and feedback. I also thank my advisors: Yizhou Zhang and Ondřej Lhoták for their patience and guidance.
 This work was supported in part by an Amazon research award on automated
 reasoning.
-The views and opinions expressed are those of the authors and do not necessarily
+The views and opinions expressed are those of the author and do not necessarily
 reflect the position of any funding agency.

### 2023-06-08 frontmatters.tex
Fixing page number
Fixing format.
Fix typo.

@@
+\pagenumbering{gobble}
+
 % Title Page
-\pagenumbering{roman}
 
@@
 \newpage
-
-\setcounter{page}{1}
+\pagenumbering{roman}
+\setcounter{page}{2}
 
@@
 
-Ende Jin was the sole author of \cref{sec:complete-fmltt}, which were written under the supervision of Prof. Yizhou Zhang.
+Ende Jin was the sole author for \cref{sec:complete-fmltt}, which were written under the supervision of Prof. Yizhou Zhang.
 
@@
 
-I am grateful to the reviewers for their many useful comments.  I thank
+I am grateful for many useful comments and suggestions.  I thank
 Benjamin Delaware,
 Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, Zhixuan Yang, Prabhakar Ragde, Werner Dietl
-for valuable discussions and feedback. I also thank my advisors: Yizhou Zhang and Ondřej Lhoták for their patience and guidance.
+for valuable discussions and feedback.
+
+I also thank my advisors: Yizhou Zhang and Ondřej Lhoták for their patience and guidance.
+
 This work was supported in part by an Amazon research award on automated

### 2023-06-11 frontmatters.tex
Fix typo.

@@
 
-Ende Jin was the sole author for \cref{sec:complete-fmltt}, which were written under the supervision of Prof. Yizhou Zhang.
+Ende Jin was the sole author for \cref{sec:complete-fmltt}, which were written under the supervision of Dr. Yizhou Zhang.
 
@@
 
-\noindentparagraph{Research presented in rest of the thesis:}
+\noindentparagraph{Research presented in the rest of the thesis:}
 The research was conducted at the University of Waterloo by Ende Jin under the supervision of Dr.
